### 2020-01-21 ResultNotebooks/results_socbudget.Rmd
Initial commit, 1st draft

---
title: "First SOC budget analysis"
output: html_notebook
---

```{r load data}
library(moinput)
library(mrvalidation)
library(socbudget)

setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")

SoilCarbon <- calcOutput("CarbonBudget", aggregate = FALSE)
Landuse    <- calcOutput("Landuse",      aggregate=FALSE)

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

SOCStock     <- dimSums(dimSums(SoilCarbon, dim=3.2)* Landuse, dim=3)
glo_SOCStock <- dimSums(SOCStock, dim=1) 
cell2iso     <- toolGetMapping("CountryToCellMapping.csv", type = "cell")
iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
iso2reg      <- toolMappingFile("regional",getConfig("regionmapping"),readcsv = TRUE)
reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")

```

Load validation data:

```{r Validation Stocks}

  LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
  LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
  LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
  WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
  GSOC       <- calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate="REG+GLO", try=TRUE)
  SoilGrids  <- calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate="REG+GLO", try=TRUE)

```

Plot global stocks

```{r plot glo stocks}

plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red", xlim=c(1951,2017),  ylim=c(400000,1300000))
lines(getYears(LPJ_IPCC, as.integer = TRUE), LPJ_IPCC["GLO",,1], type="l", col="green")
lines(getYears(LPJ_car, as.integer = TRUE), LPJ_car["GLO",,], type="l", col="darkgreen")
lines(getYears(LPJ_rev21, as.integer = TRUE), LPJ_rev21["GLO",,2], type="l", col="lightgreen")
lines(getYears(WISE, as.integer = TRUE), WISE["GLO",,], type="l", col="brown")
lines(getYears(SoilGrids, as.integer = TRUE), SoilGrids["GLO",,], type="l", col="blue")
lines(getYears(GSOC, as.integer = TRUE), GSOC["GLO",,], type="l", col="pink")
plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red")
```

Plot regional stocks

```{r plot reg stocks}

library(mip)

getNames(reg_SOCStock) <- paste0("modeloutput.SOCbudget.", getNames(WISE,dim=3))
getSets(reg_SOCStock)  <- getSets(WISE)

hist <- rbind(as.quitte(WISE["GLO",,,invert=TRUE]),
              as.quitte(GSOC["GLO",,,invert=TRUE]),
              as.quitte(SoilGrids["GLO",,,invert=TRUE]),
              as.quitte(LPJ_IPCC["GLO",,,invert=TRUE][,,1]),
              as.quitte(LPJ_car["GLO",,,invert=TRUE]),
              as.quitte(LPJ_rev21["GLO",,,invert=TRUE][,,2]))

mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y")

```

```{r Validation Densities}
    cell_GSOC      <- readSource("GSOC",  convert="onlycorrect")
    cell_WISE      <- readSource("WISE",  convert="onlycorrect")
    cell_SoilGrids <- readSource("SoilGrids",  convert="onlycorrect")

```
```{r plot density maps}

```

### 2020-01-21 SOCbudgetPaper/SOCbudget.Rmd
Initial commit, 1st draft

---
title: Historical Soil Organic Carbon Budget
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Kristine
    surname: Karstens
    affiliation: 1
    email: kristine.karstenst@pik-potsdam.de
    corresponding: true
  - given_name: Benjamin Leon
    surname: Bodirsky
    affiliation: 1
  - given_name: Alexander
    surname: Popp
    affiliation: 1

affiliation:
  - code: 1
    address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany

abstract: |

 SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
 this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
 models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these 
 biophysical processes are still little investigated.
 Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and 
 land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
 We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
 Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from … . We also find that SOC is very 
 sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.

bibliography: SOCbudget.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:

algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.

copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company.

availability:
  codedata: |
    use this to add a statement when having data sets and software code available

authorcontribution: |
  Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.

disclaimer: |
  We like Copernicus.

acknowledgements: |
  Thanks to the rticles contributors!

appendix: |
  \section{Figures and tables in appendices}
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  \subsection{Option 1}
  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
  They will be correctly named automatically.
  \subsection{Option 2}
  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
  `\appendixfigures` needs to be added in front of appendix figures
  `\appendixtables` needs to be added in front of appendix tables

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.

output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

```{r child = 'introduction.Rmd'}
```

```{r child = 'methods.Rmd'}
```

```{r child = 'results.Rmd'}
```

```{r child = 'discussion.Rmd'}
```

```{r child = 'conclusion.Rmd'}
```

### 2020-01-21 SOCbudgetPaper/SOCbudget.tex
Initial commit, 1st draft

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

%% copernicus_rticles_template (flag for rticles template detection - do not remove!)

%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.

%% 2-column papers and discussion papers
\documentclass[gc, manuscript]{copernicus}

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Web Ecology (we)
% Wind Energy Science (wes)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
%
\usepackage{algorithmic}
\usepackage{algorithm}

\begin{document}

\title{Historical Soil Organic Carbon Budget}

\Author[1]{Kristine}{Karstens}
\Author[1]{Benjamin Leon}{Bodirsky}
\Author[1]{Alexander}{Popp}

\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

\runningtitle{R Markdown Template for Copernicus}

\runningauthor{Nüst et al.}

\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
SOC one of larges c sinks on earth (3 times larger biospehre pool).
Agricultural management leads to a depletion of soil organic crabon.
However this depletion of soil organic carbon (SOC) pools are so far not
well represented in global assessments of historic carbon emissions.
While SOC models often represent well the biochemical processes that
lead to the accumulation and decay of SOC, the management decisions
driving these biophysical processes are still little investigated. Here
we create a spatial explicit data set for crop residue and manure
management on cropland based on global historic production (FAOSTAT) and
land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
create a half-degree resolution soil organic carbon budget on mineral
soils. We estimate that due to arable farming soils have lost over (?)
GtOC of which (??) GtOC have been released within the period 1990-2010.
Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than
Tier 1 methods, which may origin from \ldots{} . We also find that SOC
is very sensity to management decision such as residue recycling
indicating the nessessity to incorporated better management data in soil
models.
\end{abstract}

\copyrightstatement{The author's copyright for this publication is transferred to
institution/company.}

\introduction

Introduction text goes here. You can change the name of the section if
neccessary using
\texttt{\textbackslash{}introduction{[}modified\ heading{]}}.

The following settings can or must be configured in the header of this
file and are bespoke for Copernicus manuscripts:

\begin{itemize}
\item
  The \texttt{journal} you are submitting to using the official
  abbreviation.
\item
  You can use the function
  \texttt{rticles::copernicus\_journal\_abbreviations(name\ =\ \textquotesingle{}...\textquotesingle{})}
  to search the existing journals.
\end{itemize}

\section{Method (50)}

\subsection{Carbon Budget (50)}

Based on the IPCC guidelines \citep[\citet{IPCC2019}]{IPCC2006} for soil
organic carbon stocks, we combining approaches to estimate SOC stocks by
weighting inflows via dead plant material (see @ref(sec:carboninputs))
against outflows through SOC decay (see @ref(sec:tier1),
@ref(sec:tier2)). Carbon displacement via leaching and erosion is
neglected for this budget.

Connecting carbon inputs with soil carbon dynamics will create a budget
of in- and outflows.

\subsubsection{Soil Carbon turnover following (new) Tier 2 method
(300)}\label{sec:tier2}

Yearly turnover between three different SOC pools for the topsoil is
model using global parameters.

\subsubsection{Soil Carbon turnover following Tier 1
(150)}\label{sec:tier1}

Stock changes are calculated using simple change factors on the SOC
pools for the topsoil.

\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}

Carbon inputs are estimated based on the land use. Whereas cropland
inputs are mainly formed by residue, dead below ground and cover crop
biomass alterate by the corresponding management practice, pasture and
natural vegetation inputs are estimate via modelled annual litterfall
rates. For the former we use various data sets and rule based
approaches, which are specified in @ref(sec:agrimanagement)

\subsection{Agricultural management (50)}\label{sec:agrimanagement}

We combine data sets to estimate agricultural flows and management
decisions on cropland.

\subsubsection{Landuse and Landuse Change (150)}

We use LUH2v2 data for major Landuse types and their transition and fit
cropspecific areas to country scale FAO data.

\subsubsection{Crop Production and Residues (300)}

FAO Production values are combined with Feed estimations from
{[}Isabelles Paper{]} and rule based demand shares. LPJmL yield and LUH
landuse patterns are used to scale down to half-degree.

\subsubsection{Livestock Distribution and Manure Excretion (300)}

Based on {[}Gridded Livestock of the world{]} we use rule based
asumption to estimate livestock and manure distribution on the globe.
Animal waste system shares are used as is {[}Bodirsky{]}.

\subsubsection{Irrigation (100)}

Simple growing period calculations together with irrigation shares of
LUH2v2 are use to estimate water effects on decay rates.

\subsubsection{Tillage (100)}

Tillage data sets of {[}Vera, others{]} together with rules are used to
drive tillage effect on decay rates.

\section{Content section with citations}

See the
\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{R
Markdown docs for bibliographies and citations}.

Copernicus supports biblatex and a sample bibliography is in file
\texttt{sample.bib}. Read {[}{]}, and {[}see {]}.

\section{Content section with R code chunks}

You should always use \texttt{echo\ =\ FALSE} on R Markdown code blocks
as they add formatting and styling not desired by Copernicus. The hidden
workflow results in 42.

You can add verbatim code snippets without extra styles by using
\texttt{\textasciigrave{}\textasciigrave{}\textasciigrave{}} without
additional instructions.

\begin{verbatim}
sum <- 1 + 41
\end{verbatim}

\section{Content section with list}

If you want to insert a list, you must

\begin{itemize}
\item
  leave
\item
  empty lines
\item
  between each list item
\end{itemize}

\section{Examples from the official template}

\subsection{FIGURES}

When figures and tables are placed at the end of the MS (article in
one-column style), please add \clearpage between bibliography and first
table and/or figure as well as between each table and/or figure.

\subsection{FIGURES}

When figures and tables are placed at the end of the MS (article in
one-column style), please add \clearpage between bibliography and first
table and/or figure as well as between each table and/or figure.

\subsubsection{ONE-COLUMN FIGURES}

Include a 12cm width figure of Nikolaus Copernicus from
\href{https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg}{Wikipedia}
with caption using R Markdown.

\subsubsection{TWO-COLUMN FIGURES}

You can also include a larger figure.

\subsection{TABLES}

You can ad \LaTeX table in an R Markdown document to meet the template
requirements.

\subsubsection{ONE-COLUMN TABLE}

\begin{table}[t]
\caption{TEXT}
\begin{tabular}{l c r}
\tophline

a & b & c \\
\middlehline
1 & 2 & 3 \\

\bottomhline
\end{tabular}
\belowtable{Table Footnotes}
\end{table}

\subsubsection{TWO-COLUMN TABLE}

\begin{table*}[t]
\caption{TEXT}
\begin{tabular}{l c r}
\tophline

a & b & c \\
\middlehline
1 & 2 & 3 \\

\bottomhline
\end{tabular}
\belowtable{Table footnotes}
\end{table*}

\subsection{MATHEMATICAL EXPRESSIONS}

All papers typeset by Copernicus Publications follow the math
typesetting regulations given by the IUPAC Green Book (IUPAC:
Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell
Science, available at:
http://old.iupac.org/publications/books/gbook/green\_book\_2ed.pdf,
1993).

Physical quantities/variables are typeset in italic font (t for time, T
for Temperature)

Indices which are not defined are typeset in italic font (x, y, z, a, b,
c)

Items/objects which are defined are typeset in roman font (Car A, Car B)

Descriptions/specifications which are defined by itself are typeset in
roman font (abs, rel, ref, tot, net, ice)

Abbreviations from 2 letters are typeset in roman font (RH, LAI)

Vectors are identified in bold italic font using \vec{x}

Matrices are identified in bold roman font

Multiplication signs are typeset using the LaTeX commands
\texttt{\textbackslash{}times} (for vector products, grids, and
exponential notations) or \texttt{\textbackslash{}cdot}

The character * should not be applied as mutliplication sign

\subsection{ALGORITHM}

If you want to use algorithms, you can either enable the required
packages in the header (the default, see \texttt{algorithms:\ true}), or
make sure yourself that the \LaTeX packages \texttt{algorithms} and
\texttt{algorithmicx} are installed so that \texttt{algorithm.sty}
respectively \texttt{algorithmic.sty} can be loaded by the Copernicus
template. Copernicus staff will remove all undesirable packages from
your LaTeX source code, so please stick to using the header option,
which only adds the two acceptable packages.

\begin{algorithm}
\caption{Algorithm Caption}
\label{a1}
\begin{algorithmic}
\STATE $i\gets 10$
  \IF {$i\geq 5$} 
\STATE $i\gets i-1$
  \ELSE
\IF {$i\leq 3$}
\STATE $i\gets i+2$
  \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

\conclusions

The conclusion goes here. You can modify the section name with
\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.

\codedataavailability{use this to add a statement when having data sets and software code
available} %% use this section when having data sets and software code available

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Figures and tables in appendices}

Regarding figures and tables in appendices, the following two options
are possible depending on your general handling of figures and tables in
the manuscript environment: \subsection{Option 1} If you sorted all
figures and tables into the sections of the text, please also sort the
appendix figures and appendix tables into the respective appendix
sections. They will be correctly named automatically.
\subsection{Option 2} If you put all figures after the reference list,
please insert appendix tables and figures after the normal tables and
figures.

To rename them correctly to A1, A2, etc., please add the following
commands in front of them: \texttt{\textbackslash{}appendixfigures}
needs to be added in front of appendix figures
\texttt{\textbackslash{}appendixtables} needs to be added in front of
appendix tables

Please add \texttt{\textbackslash{}clearpage} between each table and/or
figure. Further guidelines on figures and tables can be found below.
\noappendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky and
Popp revised paper.} %% optional section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\disclaimer{We like Copernicus.} %% optional section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
Thanks to the rticles contributors!
\end{acknowledgements}

%% REFERENCES
%% DN: pre-configured to BibTeX for rticles

%% The reference list is compiled as follows:
%%
%% \begin{thebibliography}{}
%%
%% \bibitem[AUTHOR(YEAR)]{LABEL1}
%% REFERENCE 1
%%
%% \bibitem[AUTHOR(YEAR)]{LABEL2}
%% REFERENCE 2
%%
%% \end{thebibliography}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
\bibliographystyle{copernicus}
\bibliography{SOCbudget.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

\end{document}

### 2020-01-21 SOCbudgetPaper/conclusion.Rmd
Initial commit, 1st draft

\conclusions

The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.

### 2020-01-21 SOCbudgetPaper/discussion.Rmd
Initial commit, 1st draft

--- 2020-01-21 SOCbudgetPaper/SOCbudget.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
+# Examples from the official template
 
-%% copernicus_rticles_template (flag for rticles template detection - do not remove!)
+## FIGURES
 
-%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
+When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage between bibliography and first table and/or figure as well as between each table and/or figure.
 
-%% 2-column papers and discussion papers
-\documentclass[gc, manuscript]{copernicus}
+## FIGURES
 
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
+When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage between bibliography and first table and/or figure as well as between each table and/or figure.
 
-% Advances in Geosciences (adgeo)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
-% Atmospheric Chemistry and Physics (acp)
-% Atmospheric Measurement Techniques (amt)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% DEUQUA Special Publications (deuquasp)
-% Drinking Water Engineering and Science (dwes)
-% Earth Surface Dynamics (esurf)
-% Earth System Dynamics (esd)
-% Earth System Science Data (essd)
-% E&G Quaternary Science Journal (egqsj)
-% Fossil Record (fr)
-% Geochronology (gchron)
-% Geographica Helvetica (gh)
-% Geoscience Communication (gc)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% History of Geo- and Space Sciences (hgss)
-% Hydrology and Earth System Sciences (hess)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Primate Biology (pb)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Web Ecology (we)
-% Wind Energy Science (wes)
+### ONE-COLUMN FIGURES
 
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
+Include a 12cm width figure of Nikolaus Copernicus from [Wikipedia](https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg) with caption using R Markdown.
 
-% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
-%
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-\begin{document}
-
-\title{Historical Soil Organic Carbon Budget}
-
-\Author[1]{Kristine}{Karstens}
-\Author[1]{Benjamin Leon}{Bodirsky}
-\Author[1]{Alexander}{Popp}
-
-\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
-
-%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
-
-\runningtitle{R Markdown Template for Copernicus}
-
-\runningauthor{Nüst et al.}
-
-\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
-
-\maketitle
-
-\begin{abstract}
-SOC one of larges c sinks on earth (3 times larger biospehre pool).
-Agricultural management leads to a depletion of soil organic crabon.
-However this depletion of soil organic carbon (SOC) pools are so far not
-well represented in global assessments of historic carbon emissions.
-While SOC models often represent well the biochemical processes that
-lead to the accumulation and decay of SOC, the management decisions
-driving these biophysical processes are still little investigated. Here
-we create a spatial explicit data set for crop residue and manure
-management on cropland based on global historic production (FAOSTAT) and
-land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
-create a half-degree resolution soil organic carbon budget on mineral
-soils. We estimate that due to arable farming soils have lost over (?)
-GtOC of which (??) GtOC have been released within the period 1990-2010.
-Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than
-Tier 1 methods, which may origin from \ldots{} . We also find that SOC
-is very sensity to management decision such as residue recycling
-indicating the nessessity to incorporated better management data in soil
-models.
-\end{abstract}
-
-\copyrightstatement{The author's copyright for this publication is transferred to
-institution/company.}
-
-\introduction
-
-Introduction text goes here. You can change the name of the section if
-neccessary using
-\texttt{\textbackslash{}introduction{[}modified\ heading{]}}.
-
-The following settings can or must be configured in the header of this
-file and are bespoke for Copernicus manuscripts:
-
-\begin{itemize}
-\item
-  The \texttt{journal} you are submitting to using the official
-  abbreviation.
-\item
-  You can use the function
-  \texttt{rticles::copernicus\_journal\_abbreviations(name\ =\ \textquotesingle{}...\textquotesingle{})}
-  to search the existing journals.
-\end{itemize}
-
-\section{Method (50)}
-
-\subsection{Carbon Budget (50)}
-
-Based on the IPCC guidelines \citep[\citet{IPCC2019}]{IPCC2006} for soil
-organic carbon stocks, we combining approaches to estimate SOC stocks by
-weighting inflows via dead plant material (see @ref(sec:carboninputs))
-against outflows through SOC decay (see @ref(sec:tier1),
-@ref(sec:tier2)). Carbon displacement via leaching and erosion is
-neglected for this budget.
-
-Connecting carbon inputs with soil carbon dynamics will create a budget
-of in- and outflows.
-
-\subsubsection{Soil Carbon turnover following (new) Tier 2 method
-(300)}\label{sec:tier2}
-
-Yearly turnover between three different SOC pools for the topsoil is
-model using global parameters.
-
-\subsubsection{Soil Carbon turnover following Tier 1
-(150)}\label{sec:tier1}
-
-Stock changes are calculated using simple change factors on the SOC
-pools for the topsoil.
-
-\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}
-
-Carbon inputs are estimated based on the land use. Whereas cropland
-inputs are mainly formed by residue, dead below ground and cover crop
-biomass alterate by the corresponding management practice, pasture and
-natural vegetation inputs are estimate via modelled annual litterfall
-rates. For the former we use various data sets and rule based
-approaches, which are specified in @ref(sec:agrimanagement)
-
-\subsection{Agricultural management (50)}\label{sec:agrimanagement}
-
-We combine data sets to estimate agricultural flows and management
-decisions on cropland.
-
-\subsubsection{Landuse and Landuse Change (150)}
-
-We use LUH2v2 data for major Landuse types and their transition and fit
-cropspecific areas to country scale FAO data.
-
-\subsubsection{Crop Production and Residues (300)}
-
-FAO Production values are combined with Feed estimations from
-{[}Isabelles Paper{]} and rule based demand shares. LPJmL yield and LUH
-landuse patterns are used to scale down to half-degree.
-
-\subsubsection{Livestock Distribution and Manure Excretion (300)}
-
-Based on {[}Gridded Livestock of the world{]} we use rule based
-asumption to estimate livestock and manure distribution on the globe.
-Animal waste system shares are used as is {[}Bodirsky{]}.
-
-\subsubsection{Irrigation (100)}
-
-Simple growing period calculations together with irrigation shares of
-LUH2v2 are use to estimate water effects on decay rates.
-
-\subsubsection{Tillage (100)}
-
-Tillage data sets of {[}Vera, others{]} together with rules are used to
-drive tillage effect on decay rates.
-
-\section{Content section with citations}
-
-See the
-\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{R
-Markdown docs for bibliographies and citations}.
-
-Copernicus supports biblatex and a sample bibliography is in file
-\texttt{sample.bib}. Read {[}{]}, and {[}see {]}.
-
-\section{Content section with R code chunks}
-
-You should always use \texttt{echo\ =\ FALSE} on R Markdown code blocks
-as they add formatting and styling not desired by Copernicus. The hidden
-workflow results in 42.
-
-You can add verbatim code snippets without extra styles by using
-\texttt{\textasciigrave{}\textasciigrave{}\textasciigrave{}} without
-additional instructions.
-
-\begin{verbatim}
-sum <- 1 + 41
-\end{verbatim}
-
-\section{Content section with list}
-
-If you want to insert a list, you must
-
-\begin{itemize}
-\item
-  leave
-\item
-  empty lines
-\item
-  between each list item
-\end{itemize}
-
-\section{Examples from the official template}
-
-\subsection{FIGURES}
-
-When figures and tables are placed at the end of the MS (article in
-one-column style), please add \clearpage between bibliography and first
-table and/or figure as well as between each table and/or figure.
-
-\subsection{FIGURES}
-
-When figures and tables are placed at the end of the MS (article in
-one-column style), please add \clearpage between bibliography and first
-table and/or figure as well as between each table and/or figure.
-
-\subsubsection{ONE-COLUMN FIGURES}
-
-Include a 12cm width figure of Nikolaus Copernicus from
-\href{https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg}{Wikipedia}
-with caption using R Markdown.
-
-\subsubsection{TWO-COLUMN FIGURES}
+### TWO-COLUMN FIGURES
 
@@
 
-\subsection{TABLES}
+## TABLES
 
-You can ad \LaTeX table in an R Markdown document to meet the template
-requirements.
+You can ad \LaTeX table in an R Markdown document to meet the template requirements.
 
-\subsubsection{ONE-COLUMN TABLE}
+### ONE-COLUMN TABLE
 
@@
 
-\subsubsection{TWO-COLUMN TABLE}
+### TWO-COLUMN TABLE
 
@@
 
-\subsection{MATHEMATICAL EXPRESSIONS}
+## MATHEMATICAL EXPRESSIONS
 
-All papers typeset by Copernicus Publications follow the math
-typesetting regulations given by the IUPAC Green Book (IUPAC:
-Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell
-Science, available at:
-http://old.iupac.org/publications/books/gbook/green\_book\_2ed.pdf,
-1993).
+All papers typeset by Copernicus Publications follow the math typesetting regulations given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
 
-Physical quantities/variables are typeset in italic font (t for time, T
-for Temperature)
+Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
 
-Indices which are not defined are typeset in italic font (x, y, z, a, b,
-c)
+Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
 
@@
 
-Descriptions/specifications which are defined by itself are typeset in
-roman font (abs, rel, ref, tot, net, ice)
+Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
 
@@
 
-Multiplication signs are typeset using the LaTeX commands
-\texttt{\textbackslash{}times} (for vector products, grids, and
-exponential notations) or \texttt{\textbackslash{}cdot}
+Multiplication signs are typeset using the LaTeX commands `\times` (for vector products, grids, and exponential notations) or `\cdot`
 
@@
 
-\subsection{ALGORITHM}
+## ALGORITHM
 
-If you want to use algorithms, you can either enable the required
-packages in the header (the default, see \texttt{algorithms:\ true}), or
-make sure yourself that the \LaTeX packages \texttt{algorithms} and
-\texttt{algorithmicx} are installed so that \texttt{algorithm.sty}
-respectively \texttt{algorithmic.sty} can be loaded by the Copernicus
-template. Copernicus staff will remove all undesirable packages from
-your LaTeX source code, so please stick to using the header option,
-which only adds the two acceptable packages.
+  If you want to use algorithms, you can either enable the required packages in the header (the default, see `algorithms: true`), or make sure yourself that the \LaTeX packages `algorithms` and `algorithmicx` are installed so that `algorithm.sty` respectively `algorithmic.sty` can be loaded by the Copernicus template.
+Copernicus staff will remove all undesirable packages from your LaTeX source code, so please stick to using the header option, which only adds the two acceptable packages.
 
@@
 \end{algorithm}
-
-\conclusions
-
-The conclusion goes here. You can modify the section name with
-\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
-
-\codedataavailability{use this to add a statement when having data sets and software code
-available} %% use this section when having data sets and software code available
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Figures and tables in appendices}
-
-Regarding figures and tables in appendices, the following two options
-are possible depending on your general handling of figures and tables in
-the manuscript environment: \subsection{Option 1} If you sorted all
-figures and tables into the sections of the text, please also sort the
-appendix figures and appendix tables into the respective appendix
-sections. They will be correctly named automatically.
-\subsection{Option 2} If you put all figures after the reference list,
-please insert appendix tables and figures after the normal tables and
-figures.
-
-To rename them correctly to A1, A2, etc., please add the following
-commands in front of them: \texttt{\textbackslash{}appendixfigures}
-needs to be added in front of appendix figures
-\texttt{\textbackslash{}appendixtables} needs to be added in front of
-appendix tables
-
-Please add \texttt{\textbackslash{}clearpage} between each table and/or
-figure. Further guidelines on figures and tables can be found below.
-\noappendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky and
-Popp revised paper.} %% optional section
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\disclaimer{We like Copernicus.} %% optional section
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgements}
-Thanks to the rticles contributors!
-\end{acknowledgements}
-
-%% REFERENCES
-%% DN: pre-configured to BibTeX for rticles
-
-%% The reference list is compiled as follows:
-%%
-%% \begin{thebibliography}{}
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL1}
-%% REFERENCE 1
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL2}
-%% REFERENCE 2
-%%
-%% \end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{SOCbudget.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-\end{document}

### 2020-01-21 SOCbudgetPaper/introduction.Rmd
Initial commit, 1st draft

\introduction

Introduction text goes here.
You can change the name of the section if neccessary using `\introduction[modified heading]`.

The following settings can or must be configured in the header of this file and are bespoke for Copernicus manuscripts:

  - The `journal` you are submitting to using the official abbreviation.

  - You can use the function `rticles::copernicus_journal_abbreviations(name = '...')` to search the existing journals.

### 2020-01-21 SOCbudgetPaper/methods.Rmd
Initial commit, 1st draft

# Method (50)

<!-- Soil carbon dynamics are modeled using a yearly based SOC model and carbon input and management data. -->

## Carbon Budget (50)

Based on the IPCC guidelines [@IPCC2019] for soil organic carbon stocks, we combining approaches to estimate SOC stocks by weighting inflows via dead plant material (see \@ref(sec:carboninputs)) against outflows through SOC decay (see \@ref(sec:tier2)). Carbon displacement via leaching and erosion is neglected for this budget. The simple Tier 1 method based on stock change factors is used to cross validated results \@ref(sec:tier1)

### Carbon Inputs to the Soil {#sec:carboninputs}

Carbon inputs estimations are based on the land use type. Whereas cropland inputs are mainly formed by disaggregated country statistics on residue, dead below ground and cover crop biomass, pasture and natural vegetation inputs are estimate via modelled annual litterfall rates. For the former we use several data sets and rule based approaches to reflect management decisions, which are specified in \@ref(sec:agrimanagement). The later are based on LPJmL4 estimations for natural vegetation (see Sybilles paper) and grassland (see Susannes Paper).

### Soil Carbon turnover following (new) Tier 2 method (300) {#sec:tier2}

Yearly turnover between three different SOC pools for the topsoil is model using global parameters.

### Soil Carbon turnover following Tier 1 (150) {#sec:tier1}

Following the tier 1 approach of the IPCC guidelines 2006, stocks are estimated via stock change factors reflecting , in- and outflows are not calculated explicitly

Stock changes are calculated using simple change factors on the SOC pools for the topsoil.

## Agricultural management	(50) {#sec:agrimanagement}
We combine data sets to estimate agricultural flows and management decisions on cropland.

### Landuse and Landuse Change (150)
We use LUH2v2 data for major Landuse types and their transition and fit cropspecific areas to country scale FAO data.

### Crop Production and Residues	(300)
FAO Production values are combined with Feed estimations from [Isabelles Paper] and rule based demand shares. LPJmL yield and LUH landuse patterns are used to scale down to half-degree.

### Livestock Distribution and Manure Excretion	(300)
Based on [Gridded Livestock of the world] we use rule based asumption to estimate livestock and manure distribution on the globe. Animal waste system shares are used as is [Bodirsky].

### Irrigation (100)
Simple growing period calculations together with irrigation shares of LUH2v2 are use to estimate water effects on decay rates.

### Tillage (100)
Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.

### 2020-01-21 SOCbudgetPaper/results.Rmd
Initial commit, 1st draft

# Content section with citations

See the [R Markdown docs for bibliographies and citations](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html).

Copernicus supports biblatex and a sample bibliography is in file `sample.bib`.
Read [], and [see ].

# Content section with R code chunks

```{r, echo = FALSE}
sum <- 1 + 41
```

You should always use `echo = FALSE` on R Markdown code blocks as they add formatting and styling not desired by Copernicus.
The hidden workflow results in `r sum`.

You can add verbatim code snippets without extra styles by using ` ``` ` without additional instructions.

```
sum <- 1 + 41
```

# Content section with list

If you want to insert a list, you must

- leave

- empty lines

- between each list item

### 2020-01-21 ResultNotebooks/results_socbudget.R

@@ -1,0 +1,72 @@
+---
+title: "First SOC budget analysis"
+output: html_notebook
+---
+
+  
+
+```{r load data}
+library(moinput)
+library(mrvalidation)
+library(socbudget)
+
+setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
+setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+
+SoilCarbon <- calcOutput("CarbonBudget", aggregate = FALSE)
+Landuse    <- calcOutput("Landuse",      aggregate=FALSE)
+
+```
+
+Calculate global carbon stocks
+
+```{r Carbon Stock (cell iso reg glo)}
+
+SOCStock     <- dimSums(dimSums(SoilCarbon, dim=3.2)* Landuse, dim=3)
+glo_SOCStock <- dimSums(SOCStock, dim=1) 
+cell2iso     <- toolGetMapping("CountryToCellMapping.csv", type = "cell")
+iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
+iso2reg      <- toolMappingFile("regional",getConfig("regionmapping"),readcsv = TRUE)
+reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
+
+```
+
+Load validation data:
+
+```{r Validation Stocks}
+
+  LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
+  LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
+  LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
+  WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
+  GSOC       <- calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate="REG+GLO", try=TRUE)
[…]

### 2020-01-22 SOCbudgetPaper/SOCbudget.Rmd
several Updates + first part methods

@@
     affiliation: 1
-
 affiliation:
@@
     address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany
-
 abstract: | 
-
  SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
@@
  sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models. 
-
 bibliography: SOCbudget.bib
@@
   author: Nüst et al.
-# This section is mandatory even if you declare that no competing interests are present.
 competinginterests: |
@@
 # OPTIONAL:
-
 algorithms: true
 # See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
-
 copyrightstatement: |
   The author's copyright for this publication is transferred to institution/company.
-
 availability:
@@
     use this to add a statement when having data sets and software code available
-
 authorcontribution: |
   Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.
-
 disclaimer: |
   We like Copernicus.
-
 acknowledgements: |
   Thanks to the rticles contributors!
-
 appendix: |
   \section{Figures and tables in appendices}
-  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
   \subsection{Option 1}
@@
 
-  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
   `\appendixfigures` needs to be added in front of appendix figures
@@
   Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
-
+  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
+  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
 output:
-  rticles::copernicus_article: default
   bookdown::pdf_book:
     base_format: rticles::copernicus_article # for using bookdown features like \@ref()
+  rticles::copernicus_article: default
+  md_document:
+    variant: markdown_github
 ---

### 2020-01-22 SOCbudgetPaper/SOCbudget.tex
several Updates + first part methods

@@
 
-The following settings can or must be configured in the header of this
-file and are bespoke for Copernicus manuscripts:
-
-\begin{itemize}
-\item
-  The \texttt{journal} you are submitting to using the official
-  abbreviation.
-\item
-  You can use the function
-  \texttt{rticles::copernicus\_journal\_abbreviations(name\ =\ \textquotesingle{}...\textquotesingle{})}
-  to search the existing journals.
-\end{itemize}
-
 \section{Method (50)}
@@
 
-Based on the IPCC guidelines \citep[\citet{IPCC2019}]{IPCC2006} for soil
-organic carbon stocks, we combining approaches to estimate SOC stocks by
-weighting inflows via dead plant material (see @ref(sec:carboninputs))
-against outflows through SOC decay (see @ref(sec:tier1),
-@ref(sec:tier2)). Carbon displacement via leaching and erosion is
-neglected for this budget.
+Based on the IPCC guidelines vol.~4 and their refinement
+(\citep{ipcc_2006_2006},\citep{ipcc_2019_2019}) for soil organic carbon
+stocks, we combining approaches to estimate SOC stocks by weighting
+inflows via dead plant material (see \ref{sec:carboninputs}) against
+outflows through SOC decay (see \ref{sec:tier2}). Carbon displacement
+via leaching and erosion is neglected in this study. We calculate annual
+land use type specific budgets for cropland, pastures and natural
+vegetation, also representin land conversion as transfer between landuse
+type budgets. A simple approach based o the tier 1 method of the older
+IPCC guidelines vol 4. (\citet{ipcc_2006_2006}) using stock change
+factors, is applied to cross validated results \ref{sec:tier1}
 
-Connecting carbon inputs with soil carbon dynamics will create a budget
-of in- and outflows.
+\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}
+
+Carbon input estimations are based on the land use type. Whereas
+cropland inputs are mainly formed by disaggregated country statistics on
+residue, dead below ground and cover crop biomass, pasture and natural
+vegetation inputs are estimate via modelled annual litterfall rates.
+Using the steady-state method of the IPCC guidelines
+(\citep{ipcc_2019_2019}) carbon inputs have to be accompined by data on
+lignin and nitrogen content to allocate dead plant biomass to the
+corrosponding soil pools based on the chemical texture. Sources for all
+use data can be found in table \ref{tab:datasourceinputs}
+
+\begin{table*}[h]
+\caption{Sources for carbon input data}
+\begin{tabular}{l l l}
+\tophline
+ Land use types   & carbon inputs & nitrogen and lignin content \\
+\middlehline
+ Cropland         & FAO statistics, AQUASTAT, LPJmL4 [1] & default values [\cite{ipcc_2006_2006}] \\
+ Pasture          & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4 - manage grassland [3] & default values [2] \\
+ Natural vegetation & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4  & \begin{minipage}[t]{0.37\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY \strut \end{minipage}\tabularnewline
+\bottomhline
+\end{tabular}
+\label{tab:datasourceinputs}
+\belowtable{}
+\end{table*}
 
@@
 
-Yearly turnover between three different SOC pools for the topsoil is
-model using global parameters.
+We are following the steady-state method of the refinement of the IPCC
+guidelines vol.~4 (\citet{ipcc_2019_2019}) by calculating yearly
+turnover and transfer rates between three different SOC pools for the
+topsoil (0-30 cm). The approach is based on global parameters
+({[}@ref(ipcc\_2019\_2019){]}) as well as half-degree data on sand
+fraction (SoilGrids), temperature, preciptation and potential
+evapotranspiration (CRU). Next to the given climatic and natural
+biophysical properties irrigation regime (rainfed vs.~irrigated) as well
+as tillage (as soil disturbance indicator) modificate processes. For
+cropland an assessment of tillage types and irrigation conditions has
+been made, whereas on pastures and natural vegetation, we assume rainfed
+and non-tilled conditions.
 
@@
 
-Stock changes are calculated using simple change factors on the SOC
-pools for the topsoil.
-
-\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}
-
-Carbon inputs are estimated based on the land use. Whereas cropland
-inputs are mainly formed by residue, dead below ground and cover crop
-biomass alterate by the corresponding management practice, pasture and
-natural vegetation inputs are estimate via modelled annual litterfall
-rates. For the former we use various data sets and rule based
-approaches, which are specified in @ref(sec:agrimanagement)
+Following the tier 1 approach of the IPCC guidelines vol.~4
+(\citet{ipcc_2006_2006}), stocks are estimated via stock change factors
+given by the IPCC for the topsoil (0-30 cm) and based on measurements.
+The factors are differentiate between different crop and management
+systems reflecting different dynamics under changed in- and outflows
+without explicitly tracking these. They can be seen as conservative
+guesses and will be used to evaluate our modelling based results.
 
@@
 
-\section{Content section with citations}
+\section{Results}
 
-See the
-\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{R
-Markdown docs for bibliographies and citations}.
-
-Copernicus supports biblatex and a sample bibliography is in file
-\texttt{sample.bib}. Read {[}{]}, and {[}see {]}.
-
-\section{Content section with R code chunks}
-
-You should always use \texttt{echo\ =\ FALSE} on R Markdown code blocks
-as they add formatting and styling not desired by Copernicus. The hidden
-workflow results in 42.
-
-You can add verbatim code snippets without extra styles by using
-\texttt{\textasciigrave{}\textasciigrave{}\textasciigrave{}} without
-additional instructions.
-
-\begin{verbatim}
-sum <- 1 + 41
-\end{verbatim}
-
-\section{Content section with list}
-
-If you want to insert a list, you must
-
-\begin{itemize}
-\item
-  leave
-\item
-  empty lines
-\item
-  between each list item
-\end{itemize}
-
-\section{Examples from the official template}
-
-\subsection{FIGURES}
-
-When figures and tables are placed at the end of the MS (article in
-one-column style), please add \clearpage between bibliography and first
-table and/or figure as well as between each table and/or figure.
-
-\subsection{FIGURES}
-
-When figures and tables are placed at the end of the MS (article in
-one-column style), please add \clearpage between bibliography and first
-table and/or figure as well as between each table and/or figure.
-
-\subsubsection{ONE-COLUMN FIGURES}
-
-Include a 12cm width figure of Nikolaus Copernicus from
-\href{https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg}{Wikipedia}
-with caption using R Markdown.
-
-\subsubsection{TWO-COLUMN FIGURES}
-
-You can also include a larger figure.
-
-\subsection{TABLES}
-
-You can ad \LaTeX table in an R Markdown document to meet the template
-requirements.
-
-\subsubsection{ONE-COLUMN TABLE}
-
-\begin{table}[t]
-\caption{TEXT}
-\begin{tabular}{l c r}
-\tophline
-
-a & b & c \\
-\middlehline
-1 & 2 & 3 \\
-
-\bottomhline
-\end{tabular}
-\belowtable{Table Footnotes}
-\end{table}
-
-\subsubsection{TWO-COLUMN TABLE}
-
-\begin{table*}[t]
-\caption{TEXT}
-\begin{tabular}{l c r}
-\tophline
-
-a & b & c \\
-\middlehline
-1 & 2 & 3 \\
-
-\bottomhline
-\end{tabular}
-\belowtable{Table footnotes}
-\end{table*}
-
-\subsection{MATHEMATICAL EXPRESSIONS}
-
-All papers typeset by Copernicus Publications follow the math
-typesetting regulations given by the IUPAC Green Book (IUPAC:
-Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell
-Science, available at:
-http://old.iupac.org/publications/books/gbook/green\_book\_2ed.pdf,
-1993).
-
-Physical quantities/variables are typeset in italic font (t for time, T
-for Temperature)
-
-Indices which are not defined are typeset in italic font (x, y, z, a, b,
-c)
-
-Items/objects which are defined are typeset in roman font (Car A, Car B)
-
-Descriptions/specifications which are defined by itself are typeset in
-roman font (abs, rel, ref, tot, net, ice)
-
-Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-
-Vectors are identified in bold italic font using \vec{x}
-
-Matrices are identified in bold roman font
-
-Multiplication signs are typeset using the LaTeX commands
-\texttt{\textbackslash{}times} (for vector products, grids, and
-exponential notations) or \texttt{\textbackslash{}cdot}
-
-The character * should not be applied as mutliplication sign
-
-\subsection{ALGORITHM}
-
-If you want to use algorithms, you can either enable the required
-packages in the header (the default, see \texttt{algorithms:\ true}), or
-make sure yourself that the \LaTeX packages \texttt{algorithms} and
-\texttt{algorithmicx} are installed so that \texttt{algorithm.sty}
-respectively \texttt{algorithmic.sty} can be loaded by the Copernicus
-template. Copernicus staff will remove all undesirable packages from
-your LaTeX source code, so please stick to using the header option,
-which only adds the two acceptable packages.
-
-\begin{algorithm}
-\caption{Algorithm Caption}
-\label{a1}
-\begin{algorithmic}
-\STATE $i\gets 10$
-  \IF {$i\geq 5$} 
-\STATE $i\gets i-1$
-  \ELSE
-\IF {$i\leq 3$}
-\STATE $i\gets i+2$
-  \ENDIF
-\ENDIF
-\end{algorithmic}
-\end{algorithm}
+\section{Discussion}
 
@@
 \appendix
-\section{Figures and tables in appendices}
+\section{Figures and tables in appendices}\subsection{Option 1}
 
-Regarding figures and tables in appendices, the following two options
-are possible depending on your general handling of figures and tables in
-the manuscript environment: \subsection{Option 1} If you sorted all
-figures and tables into the sections of the text, please also sort the
-appendix figures and appendix tables into the respective appendix
-sections. They will be correctly named automatically.
-\subsection{Option 2} If you put all figures after the reference list,
-please insert appendix tables and figures after the normal tables and
-figures.
+If you sorted all figures and tables into the sections of the text,
+please also sort the appendix figures and appendix tables into the
+respective appendix sections. They will be correctly named
+automatically. \subsection{Option 2} If you put all figures after the
+reference list, please insert appendix tables and figures after the
+normal tables and figures.
 
-To rename them correctly to A1, A2, etc., please add the following
-commands in front of them: \texttt{\textbackslash{}appendixfigures}
-needs to be added in front of appendix figures
-\texttt{\textbackslash{}appendixtables} needs to be added in front of
-appendix tables
+\texttt{\textbackslash{}appendixfigures} needs to be added in front of
+appendix figures \texttt{\textbackslash{}appendixtables} needs to be
+added in front of appendix tables
 
@@
 figure. Further guidelines on figures and tables can be found below.
+Regarding figures and tables in appendices, the following two options
+are possible depending on your general handling of figures and tables in
+the manuscript environment: To rename them correctly to A1, A2, etc.,
+please add the following commands in front of them:
 \noappendix

### 2020-01-22 SOCbudgetPaper/introduction.Rmd
several Updates + first part methods

@@
 You can change the name of the section if neccessary using `\introduction[modified heading]`.
-
-The following settings can or must be configured in the header of this file and are bespoke for Copernicus manuscripts:
-
-  - The `journal` you are submitting to using the official abbreviation.
-
-  - You can use the function `rticles::copernicus_journal_abbreviations(name = '...')` to search the existing journals.

### 2020-01-28 SOCbudgetPaper/SOCbudget.Rmd
Rewrite of methods part 1 on Calculating carbon stocks

@@
     base_format: rticles::copernicus_article # for using bookdown features like \@ref()
+    md_extensions: +raw_attribute
   rticles::copernicus_article: default
@@
 ---
+\newpage
 
@@
 ```
+\newpage
 
@@
 ```
+\newpage

### 2020-01-28 SOCbudgetPaper/SOCbudget.tex
Rewrite of methods part 1 on Calculating carbon stocks

@@
 
-\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}
+\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstenst@pik-potsdam.de}{\nolinkurl{kristine.karstenst@pik-potsdam.de}})}
 
@@
 \begin{abstract}
-SOC one of larges c sinks on earth (3 times larger biospehre pool).
-Agricultural management leads to a depletion of soil organic crabon.
-However this depletion of soil organic carbon (SOC) pools are so far not
-well represented in global assessments of historic carbon emissions.
-While SOC models often represent well the biochemical processes that
-lead to the accumulation and decay of SOC, the management decisions
-driving these biophysical processes are still little investigated. Here
-we create a spatial explicit data set for crop residue and manure
-management on cropland based on global historic production (FAOSTAT) and
-land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
-create a half-degree resolution soil organic carbon budget on mineral
-soils. We estimate that due to arable farming soils have lost over (?)
-GtOC of which (??) GtOC have been released within the period 1990-2010.
-Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than
-Tier 1 methods, which may origin from \ldots{} . We also find that SOC
-is very sensity to management decision such as residue recycling
-indicating the nessessity to incorporated better management data in soil
-models.
+SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
+this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
+models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these
+biophysical processes are still little investigated.
+Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and
+land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
+We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
+Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from \ldots{} . We also find that SOC is very
+sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
 \end{abstract}
 
-\copyrightstatement{The author's copyright for this publication is transferred to
-institution/company.}
+\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
+
+\newpage
 
@@
 
-Introduction text goes here. You can change the name of the section if
-neccessary using
-\texttt{\textbackslash{}introduction{[}modified\ heading{]}}.
+Introduction text goes here.
+You can change the name of the section if neccessary using \texttt{\textbackslash{}introduction{[}modified\ heading{]}}.
+\newpage
 
-\section{Method (50)}
+\hypertarget{method-50}{%
+\section{Method (50)}\label{method-50}}
 
-\subsection{Carbon Budget (50)}
+\hypertarget{sec:carbonbudget}{%
+\subsection{Carbon Stocks following (new) Tier 2 method (50)}\label{sec:carbonbudget}}
 
-Based on the IPCC guidelines vol.~4 and their refinement
-(\citep{ipcc_2006_2006},\citep{ipcc_2019_2019}) for soil organic carbon
-stocks, we combining approaches to estimate SOC stocks by weighting
-inflows via dead plant material (see \ref{sec:carboninputs}) against
-outflows through SOC decay (see \ref{sec:tier2}). Carbon displacement
-via leaching and erosion is neglected in this study. We calculate annual
-land use type specific budgets for cropland, pastures and natural
-vegetation, also representin land conversion as transfer between landuse
-type budgets. A simple approach based o the tier 1 method of the older
-IPCC guidelines vol 4. (\citet{ipcc_2006_2006}) using stock change
-factors, is applied to cross validated results \ref{sec:tier1}
+We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
 
-\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}
+\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
+\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}
 
-Carbon input estimations are based on the land use type. Whereas
-cropland inputs are mainly formed by disaggregated country statistics on
-residue, dead below ground and cover crop biomass, pasture and natural
-vegetation inputs are estimate via modelled annual litterfall rates.
-Using the steady-state method of the IPCC guidelines
-(\citep{ipcc_2019_2019}) carbon inputs have to be accompined by data on
-lignin and nitrogen content to allocate dead plant biomass to the
-corrosponding soil pools based on the chemical texture. Sources for all
-use data can be found in table \ref{tab:datasourceinputs}
+In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
+\begin{equation}
+SOC^{eq} =\frac{C^{\textrm{in}}}{k}
+\label{eq:inoutflow}
+\end{equation}
+with \(C^{\textrm{in}}\) being carbon inputs to the soil and \(k\) denoting the soil organic carbon decay rate.
+We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon pools and entangled dynamics between them.
+Carbon inflow to each subpool (see \ref{sec:carboninputs}) and decay rates (see \ref{sec:tier2}) of each subpool are still the key components to determining steady-state SOC stocks.
+
+\hypertarget{sec:carboninputs}{%
+\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}}
+
+We account for different carbon inputs sources depending on the land use type (see table \ref{tab:datasourceinputs}).
+Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see @ref(ipcc\_2019\_2019)).
+For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub pools.
+This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
 
  \begin{table*}[h]
-\caption{Sources for carbon input data}
-\begin{tabular}{l l l}
+ \caption{Type and data sources for carbon inputs to different land use types}
+ \begin{tabular}{l l l l}
  \tophline
- Land use types   & carbon inputs & nitrogen and lignin content \\
+  Land use types   & source of carbon inputs & data source & nitrogen and lignin content \\
  \middlehline
- Cropland         & FAO statistics, AQUASTAT, LPJmL4 [1] & default values [\cite{ipcc_2006_2006}] \\
- Pasture          & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4 - manage grassland [3] & default values [2] \\
- Natural vegetation & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4  & \begin{minipage}[t]{0.37\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY \strut \end{minipage}\tabularnewline
+ \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
+                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
+                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
+                            \hline
+ \multirow{2}{*}{Pasture}  & annual litterfall & LPJmL4 [3] & default values given by [2] \\ 
+                            & manure  & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
+                            \hline
+  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
  \bottomhline
@@
 
-\subsubsection{Soil Carbon turnover following (new) Tier 2 method
-(300)}\label{sec:tier2}
+\hypertarget{sec:tier2}{%
+\subsubsection{Soil Organic Carbon decay (300)}\label{sec:tier2}}
 
-We are following the steady-state method of the refinement of the IPCC
-guidelines vol.~4 (\citet{ipcc_2019_2019}) by calculating yearly
-turnover and transfer rates between three different SOC pools for the
-topsoil (0-30 cm). The approach is based on global parameters
-({[}@ref(ipcc\_2019\_2019){]}) as well as half-degree data on sand
-fraction (SoilGrids), temperature, preciptation and potential
-evapotranspiration (CRU). Next to the given climatic and natural
-biophysical properties irrigation regime (rainfed vs.~irrigated) as well
-as tillage (as soil disturbance indicator) modificate processes. For
-cropland an assessment of tillage types and irrigation conditions has
-been made, whereas on pastures and natural vegetation, we assume rainfed
-and non-tilled conditions.
+Decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
+Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates \(k_{sub}\) given by the IPCC via
 
-\subsubsection{Soil Carbon turnover following Tier 1
-(150)}\label{sec:tier1}
+\begin{equation}
+k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \underbrace{\cdot sf_{t,i} \quad \cdot till_{t,i}.}_{\text{only included for some subpools}}
+\label{eq:decayrates}
+\end{equation}
 
-Following the tier 1 approach of the IPCC guidelines vol.~4
-(\citet{ipcc_2006_2006}), stocks are estimated via stock change factors
-given by the IPCC for the topsoil (0-30 cm) and based on measurements.
-The factors are differentiate between different crop and management
-systems reflecting different dynamics under changed in- and outflows
-without explicitly tracking these. They can be seen as conservative
-guesses and will be used to evaluate our modelling based results.
+For cropland we performed an assessment of tillage types and irrigation conditions, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as considered effects for each land use types are shown in table \ref{tab:datasourcedecay}.
+To account for variations of decay rates within grid cells based on different tillage and irrigation regimes, average rates based on area shares are calculated.
 
-\subsection{Agricultural management (50)}\label{sec:agrimanagement}
+ \begin{table*}[h]
+ \caption{Type and data sources for carbon inputs to different land use types}
+ \begin{tabular}{l l l l}
+ \tophline
+  Land use types   & type of decay driver & parameter use to represent driver & data source \\
+ \middlehline
+ \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm &  [SoilGrids]  \\
+                      \cline{2-4}
 
-We combine data sets to estimate agricultural flows and management
-decisions on cropland.
+                      & Mircobial activity & air temperature & [CRUp4.0] \\
+                      \cline{2-4}
 
-\subsubsection{Landuse and Landuse Change (150)}
+                      & Water restriction & precipitation \& potential evapotranspiration & [CRUp4.0] \\
+                      \cline{1-4}
+\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & [sec:irrigation] \\ 
+                      \cline{2-4}
 
-We use LUH2v2 data for major Landuse types and their transition and fit
-cropspecific areas to country scale FAO data.
+                      & Soil disturbance & tillage & [sec:tillage] \\
+ \bottomhline
+ \end{tabular}
+ \label{tab:datasourcedecay}
+ \belowtable{}
+ \end{table*}
 
-\subsubsection{Crop Production and Residues (300)}
+\hypertarget{soc-transfer-between-land-use-types}{%
+\subsubsection{SOC transfer between land use types}\label{soc-transfer-between-land-use-types}}
 
-FAO Production values are combined with Feed estimations from
-{[}Isabelles Paper{]} and rule based demand shares. LPJmL yield and LUH
-landuse patterns are used to scale down to half-degree.
+We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
+If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well.
+We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
 
-\subsubsection{Livestock Distribution and Manure Excretion (300)}
+\begin{equation}
+SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall sub, i  
+\label{eq:ctransfer}
+\end{equation}
 
-Based on {[}Gridded Livestock of the world{]} we use rule based
-asumption to estimate livestock and manure distribution on the globe.
-Animal waste system shares are used as is {[}Bodirsky{]}.
+with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land use type \(lut\).
+Note that \(!lut\) denotes the (sum over) land use type(s), which decrease(s) in the specific time step \(t\).
+Data sources and methodology on land use states and changes are descripted in \ref{sec:landuse}.
 
-\subsubsection{Irrigation (100)}
+\hypertarget{total-soc-stocks}{%
+\subsubsection{Total SOC stocks}\label{total-soc-stocks}}
 
-Simple growing period calculations together with irrigation shares of
-LUH2v2 are use to estimate water effects on decay rates.
+Carbon stocks \(SOC\) for each subpool (sub) converge towards the calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut), each subpool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.
 
-\subsubsection{Tillage (100)}
+\begin{equation}
+SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
+\label{eq:steadystate}
+\end{equation}
 
-Tillage data sets of {[}Vera, others{]} together with rules are used to
-drive tillage effect on decay rates.
+The global SOC stock for each time step can than be calculated via
 
-\section{Results}
+\begin{equation}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{land use type specific SOC stock within a grid cell}}}_{\text{total SOC stock within a grid cell}}
+\label{eq:totalstock}
+\end{equation}
 
-\section{Discussion}
+\newpage
+
+\hypertarget{sec:tier1}{%
+\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}
+
+Following the tier 1 approach of the IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurements.
+The factors are differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+They can be seen as conservative guesses and will be used to evaluate our modelling based results.
+
+\hypertarget{sec:agrimanagement}{%
+\subsection{Agricultural management (50)}\label{sec:agrimanagement}}
+
+We combine data sets to estimate agricultural flows and management decisions on cropland.
+
+\hypertarget{sec:landuse}{%
+\subsubsection{Landuse and Landuse Change (150)}\label{sec:landuse}}
+
+We use the Land-Use Harmonization 2 (LUH2, \citep{LUH2}) half-degree data for major land-use states.
+Cropland is subdivided using country scale FAO data (\citep{FAOSTAT}) as well as crop type specific half-degree data from LUH2v2 to disaggregate into our 19 crop types on half degree resolution (see appendix for table on crop types and their mapping to FAOSTAT and LUH types as well as for technical discription of the disaggregation method).
+Landuse transitions are calculate as area differences, allowing any land use type to either expand or shrink in each time step.
+Within cropland we do not track area transitions, following the averaging approach as descripted in \ref{sec:tier2}/\ref{sec:carbonbudget}
+
+\hypertarget{crop-production-and-residues-300}{%
+\subsubsection{Crop Production and Residues (300)}\label{crop-production-and-residues-300}}
+
+FAOSTAT production (\citep{FAOSTAT}) values are combined with feed estimations including residue intake (\citep{weindl}) and rule based demand shares.
+LPJmL yield and calculate land use pattern (see @ref(\#sec:landuse)) are used to scale down to half-degree.
+
+\hypertarget{livestock-distribution-and-manure-excretion-300}{%
+\subsubsection{Livestock Distribution and Manure Excretion (300)}\label{livestock-distribution-and-manure-excretion-300}}
+
+Based on {[}Gridded Livestock of the world{]} we use rule based asumption to estimate livestock and manure distribution on the globe. Animal waste system shares are used as is {[}Bodirsky{]}.
+
+\hypertarget{irrigation-100}{%
+\subsubsection{Irrigation (100)}\label{irrigation-100}}
+
+Simple growing period calculations together with irrigation shares of LUH2v2 are use to estimate water effects on decay rates.
+
+\hypertarget{tillage-100}{%
+\subsubsection{Tillage (100)}\label{tillage-100}}
+
+Tillage data sets of {[}Vera, others{]} together with rules are used to drive tillage effect on decay rates.
+
+\newpage
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\newpage
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Shortcommings:
+
+\begin{itemize}
+\item
+  Carbon displacement via leaching and erosion is neglected in this study.
+\item
+  Non-net/Gross land use transitions are not tracked in this study.
+  \newpage
+\end{itemize}
 
@@
 
-The conclusion goes here. You can modify the section name with
-\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
+The conclusion goes here.
+You can modify the section name with \texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
+\newpage
 
-\codedataavailability{use this to add a statement when having data sets and software code
-available} %% use this section when having data sets and software code available
+\codedataavailability{use this to add a statement when having data sets and software code available} %% use this section when having data sets and software code available
 
@@
 \appendix
-\section{Figures and tables in appendices}\subsection{Option 1}
+\section{Figures and tables in appendices}
+\subsection{Option 1}
 
-If you sorted all figures and tables into the sections of the text,
-please also sort the appendix figures and appendix tables into the
-respective appendix sections. They will be correctly named
-automatically. \subsection{Option 2} If you put all figures after the
-reference list, please insert appendix tables and figures after the
-normal tables and figures.
+If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
+They will be correctly named automatically.
 
-\texttt{\textbackslash{}appendixfigures} needs to be added in front of
-appendix figures \texttt{\textbackslash{}appendixtables} needs to be
-added in front of appendix tables
+\subsection{Option 2}
 
-Please add \texttt{\textbackslash{}clearpage} between each table and/or
-figure. Further guidelines on figures and tables can be found below.
-Regarding figures and tables in appendices, the following two options
-are possible depending on your general handling of figures and tables in
-the manuscript environment: To rename them correctly to A1, A2, etc.,
-please add the following commands in front of them:
+If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
+
+\texttt{\textbackslash{}appendixfigures} needs to be added in front of appendix figures
+\texttt{\textbackslash{}appendixtables} needs to be added in front of appendix tables
+
+Please add \texttt{\textbackslash{}clearpage} between each table and/or figure. Further guidelines on figures and tables can be found below.
+Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
+To rename them correctly to A1, A2, etc., please add the following commands in front of them:
 \noappendix
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky and
-Popp revised paper.} %% optional section
+\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.} %% optional section
 

### 2020-01-28 SOCbudgetPaper/discussion.Rmd
Rewrite of methods part 1 on Calculating carbon stocks

@@
-# Examples from the official template
+# Discussion
 
-## FIGURES
+Shortcommings:
 
-When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage between bibliography and first table and/or figure as well as between each table and/or figure.
+- Carbon displacement via leaching and erosion is neglected in this study.
 
-## FIGURES
-
-When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage between bibliography and first table and/or figure as well as between each table and/or figure.
-
-### ONE-COLUMN FIGURES
-
-Include a 12cm width figure of Nikolaus Copernicus from [Wikipedia](https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg) with caption using R Markdown.
-
-### TWO-COLUMN FIGURES
-
-You can also include a larger figure.
-
-## TABLES
-
-You can ad \LaTeX table in an R Markdown document to meet the template requirements.
-
-### ONE-COLUMN TABLE
-
-\begin{table}[t]
-\caption{TEXT}
-\begin{tabular}{l c r}
-\tophline
-
-a & b & c \\
-\middlehline
-1 & 2 & 3 \\
-
-\bottomhline
-\end{tabular}
-\belowtable{Table Footnotes}
-\end{table}
-
-### TWO-COLUMN TABLE
-
-\begin{table*}[t]
-\caption{TEXT}
-\begin{tabular}{l c r}
-\tophline
-
-a & b & c \\
-\middlehline
-1 & 2 & 3 \\
-
-\bottomhline
-\end{tabular}
-\belowtable{Table footnotes}
-\end{table*}
-
-## MATHEMATICAL EXPRESSIONS
-
-All papers typeset by Copernicus Publications follow the math typesetting regulations given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-
-Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-
-Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-
-Items/objects which are defined are typeset in roman font (Car A, Car B)
-
-Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-
-Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-
-Vectors are identified in bold italic font using \vec{x}
-
-Matrices are identified in bold roman font
-
-Multiplication signs are typeset using the LaTeX commands `\times` (for vector products, grids, and exponential notations) or `\cdot`
-
-The character * should not be applied as mutliplication sign
-
-## ALGORITHM
-
-  If you want to use algorithms, you can either enable the required packages in the header (the default, see `algorithms: true`), or make sure yourself that the \LaTeX packages `algorithms` and `algorithmicx` are installed so that `algorithm.sty` respectively `algorithmic.sty` can be loaded by the Copernicus template.
-Copernicus staff will remove all undesirable packages from your LaTeX source code, so please stick to using the header option, which only adds the two acceptable packages.
-
-\begin{algorithm}
-\caption{Algorithm Caption}
-\label{a1}
-\begin{algorithmic}
-\STATE $i\gets 10$
-  \IF {$i\geq 5$} 
-\STATE $i\gets i-1$
-  \ELSE
-\IF {$i\leq 3$}
-\STATE $i\gets i+2$
-  \ENDIF
-\ENDIF
-\end{algorithmic}
-\end{algorithm}
+- Non-net/Gross land use transitions are not tracked in this study.

### 2020-01-28 SOCbudgetPaper/methods.Rmd
Update SOCbudgetPaper/methods.Rmd

@@
 
-## Carbon Budget (50)
+## Carbon Stocks following (new) Tier 2 method (50) {#sec:carbonbudget}
 
-Based on the IPCC guidelines [@IPCC2019] for soil organic carbon stocks, we combining approaches to estimate SOC stocks by weighting inflows via dead plant material (see \@ref(sec:carboninputs)) against outflows through SOC decay (see \@ref(sec:tier2)).
-Carbon displacement via leaching and erosion is neglected for this budget.
-The simple Tier 1 method based on stock change factors is used to cross validated results \@ref(sec:tier1)
+We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
+
+### Steady-state SOC stocks and decay rates
+
+In a simple first order kinetic approach the steady-state soil organic carbon stocks $SOC^{eq}$ are given by 
+\begin{equation}
+SOC^{eq} =\frac{C^{\textrm{in}}}{k}
+\label{eq:inoutflow}
+\end{equation}
+with $C^{\textrm{in}}$ being carbon inputs to the soil and $k$ denoting the soil organic carbon decay rate.
+We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.
+4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon pools and entangled dynamics between them.
+Carbon inflow to each subpool (see \@ref(sec:carboninputs)) and decay rates (see \@ref(sec:tier2)) of each subpool are still the key components to determining steady-state SOC stocks.
 
 ### Carbon Inputs to the Soil {#sec:carboninputs}
+BB: maybe consider to include as text instead of table. The table takes it a bit out of focus. Maybe you want to have the focus on this rather than on other parts.
 
-Carbon inputs estimations are based on the land use type.
-Whereas cropland inputs are mainly formed by disaggregated country statistics on residue, dead below ground and cover crop biomass, pasture and natural vegetation inputs are estimate via modelled annual litterfall rates.
-For the former we use several data sets and rule based approaches to reflect management decisions, which are specified in \@ref(sec:agrimanagement).
-The later are based on LPJmL4 estimations for natural vegetation (see Sybilles paper) and grassland (see Susannes Paper).
+We account for different carbon inputs sources depending on the land use type (see table \@ref(tab:datasourceinputs)).
+Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
+For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub pools.
+This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \@ref(tab:datasourceinputs).
 
-### Soil Carbon turnover following (new) Tier 2 method (300) {#sec:tier2}
+```{=latex}
 
-Yearly turnover between three different SOC pools for the topsoil is model using global parameters.
+ \begin{table*}[h]
+ \caption{Type and data sources for carbon inputs to different land use types}
+ \begin{tabular}{l l l l}
+ \tophline
+  Land use types   & source of carbon inputs & data source & nitrogen and lignin content \\
+ \middlehline
+ \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
+                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
+                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
+                            \hline
+ \multirow{2}{*}{Pasture}  & annual litterfall & LPJmL4 [3] & default values given by [2] \\ 
+                            & manure  & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
+                            \hline
+  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
+ \bottomhline
+ \end{tabular}
+ \label{tab:datasourceinputs}
+ \belowtable{}
+ \end{table*}
+```
 
-### Soil Carbon turnover following Tier 1 (150) {#sec:tier1}
+### Soil Organic Carbon decay (300) {#sec:tier2}
 
-Following the tier 1 approach of the IPCC guidelines 2006, stocks are estimated via stock change factors reflecting , in- and outflows are not calculated explicitly
+Decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
+Following the steady-state method of the refinement of the IPCC guidelines vol.
+4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates $k_{sub}$ given by the IPCC via
 
-Stock changes are calculated using simple change factors on the SOC pools for the topsoil.
+\begin{equation}
+k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \underbrace{\cdot sf_{t,i} \quad \cdot till_{t,i}.}_{\text{only included for some subpools}}
+\label{eq:decayrates}
+\end{equation}
+
+BB: braket includes sandyness instead of water.
+I would leave out the braket, and just state that non-crop areas are neither tilled nor irrigated. then the formulas are still correct.
+
+For cropland we performed an assessment of tillage types and irrigation conditions, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as considered effects for each land use types are shown in table \@ref(tab:datasourcedecay).
+To account for variations of decay rates within grid cells based on different tillage and irrigation regimes, average rates based on area shares are calculated.
+
+```{=latex}
+
+ \begin{table*}[h]
+ \caption{Type and data sources for carbon inputs to different land use types}
+ \begin{tabular}{l l l l}
+ \tophline
+  Land use types   & type of decay driver & parameter use to represent driver & data source \\
+ \middlehline
+ \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm &  [SoilGrids]  \\
+                      \cline{2-4}
+
+                      & Mircobial activity & air temperature & [CRUp4.0] \\
+                      \cline{2-4}
+
+                      & Water restriction & precipitation \& potential evapotranspiration & [CRUp4.0] \\
+                      \cline{1-4}
+\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & [sec:irrigation] \\ 
+                      \cline{2-4}
+
+                      & Soil disturbance & tillage & [sec:tillage] \\
+ \bottomhline
+ \end{tabular}
+ \label{tab:datasourcedecay}
+ \belowtable{}
+ \end{table*}
+```
+
+### SOC transfer between land use types
+
+We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
+If land is converted from one land use type (lut) into another(!lut) (BB: like this?
+then you need to change the order of the !lut in the equation; lut is reduced, !lut is expanded), a respective share of the SOC stocks have to be reallocated as well.
+We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via
+
+\begin{equation}
+SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall sub, i  
+\label{eq:ctransfer}
+\end{equation}
+
+with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land use type $lut$.
+Note that $!lut$ denotes the (sum over) (BB: why braket?) land use type(s), which decrease(s) in the specific time step $t$.
+Data sources and methodology on land use states and changes are descripted in \@ref(sec:landuse).
+
+### Total SOC stocks
+
+Carbon stocks $SOC$ for each subpool (sub) converge towards the calculate steady-state stock $SOC^{eq}$ for each land-use types (lut), each subpool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
+
+\begin{equation}
+SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
+\label{eq:steadystate}
+\end{equation}
+
+BB: What is delta t doing in this equation? The value woudl be 1, no?
+
+The global SOC stock for each time step can than be calculated via
+
+\begin{equation}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{land use type specific SOC stock within a grid cell}}}_{\text{total SOC stock within a grid cell}}
+\label{eq:totalstock}
+\end{equation}
+
+<!-- Rewriting was till here!!!!!!!!!!! -->
+
+\newpage
+
+## Carbon Budget following Tier 1 (150) {#sec:tier1}
+
+Following the tier 1 approach of the IPCC guidelines vol.
+4 (@ipcc_2006_2006), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurements.
+The factors  differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+This method is used additionally to the Tier-2 approach to evaluate model accuracy.
 
 ## Agricultural management	(50) {#sec:agrimanagement}
-We combine data sets to estimate agricultural flows and management decisions on cropland.
+We combine data sets to estimate agricultural flows and management decisions on cropland. BB: wow. definetly shorter than 50 words! but do you need text hear at all?
 
-### Landuse and Landuse Change (150)
-We use LUH2v2 data for major Landuse types and their transition and fit cropspecific areas to country scale FAO data.
+### Landuse and Landuse Change (150) {#sec:landuse}
+
+We use the Land-Use Harmonization 2 (LUH2, [@LUH2]) half-degree data for major land-use states.
+Cropland is subdivided using country scale FAO data ([@FAOSTAT]) as well as crop type (BB: crop functional type?
+ots not crop-specific) specific half-degree data from LUH2v2 to disaggregate into our 19 crop types on half degree resolution (see appendix for table on crop types and their mapping to FAOSTAT and LUH types as well as for technical discription of the disaggregation method).
+Landuse transitions are calculated as area differences, allowing any land use type to either expand or shrink in each time step.
+Within cropland we do not track area transitions, following the averaging approach as descripted in \@ref(sec:tier2)/\@ref(sec:carbonbudget)
+<!-- How to explain why we do not track inside-cropland transitions? --> BB: just dont explain it.
 
 ### Crop Production and Residues	(300)
-FAO Production values are combined with Feed estimations from [Isabelles Paper] and rule based demand shares. LPJmL yield and LUH landuse patterns are used to scale down to half-degree.
+FAOSTAT production ([@FAOSTAT]) values are combined with feed basket estimates including residue and pasture intake ([@weindl]) and rule based demand shares(BB:?).
+LPJmL yields and calculated land use patterns (see \@ref(#sec:landuse)) are used to scale down to half-degree.
 
 ### Livestock Distribution and Manure Excretion	(300)
-Based on [Gridded Livestock of the world] we use rule based asumption to estimate livestock and manure distribution on the globe. Animal waste system shares are used as is [Bodirsky].
+Based on [Gridded Livestock of the world] we use rule based assumptions(BB?) to estimate livestock and manure distribution on global scale(BB: more detail).
+Animal waste system shares are used as in [Bodirsky].
 
 ### Irrigation (100)
-Simple growing period calculations together with irrigation shares of LUH2v2 are use to estimate water effects on decay rates.
+Simple growing period calculations together with irrigation shares of LUH2v2 are used (BB: you dont like d's, on't you?) to estimate water effects on decay rates.
 

### 2020-02-10 SOCbudgetPaper/SOCbudget.Rmd
rewising bennis comments, including docx, including newest resultNotebooks (with first plot)

--- 2020-01-22 SOCbudgetPaper/SOCbudget.Rmd
@@
 output:
-  bookdown::pdf_book:
-    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
+  bookdown::word_document2: default
   rticles::copernicus_article: default
@@
     variant: markdown_github
+  bookdown::pdf_book:
+    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
+    md_extensions: +raw_attribute
 ---
+\newpage
 

### 2020-02-10 SOCbudgetPaper/SOCbudget.tex
rewising bennis comments, including docx, including newest resultNotebooks (with first plot)

@@
 with \(C^{\textrm{in}}\) being carbon inputs to the soil and \(k\) denoting the soil organic carbon decay rate.
-We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon pools and entangled dynamics between them.
-Carbon inflow to each subpool (see \ref{sec:carboninputs}) and decay rates (see \ref{sec:tier2}) of each subpool are still the key components to determining steady-state SOC stocks.
+We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
+Carbon inflow to each sub-pool (see \ref{sec:carboninputs}) and decay rates (see \ref{sec:tier2}) of each sub-pool are still the key components to determining steady-state SOC stocks.
 
@@
 Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see @ref(ipcc\_2019\_2019)).
-For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub pools.
+For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub-pools.
 This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
@@
 Decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
-Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates \(k_{sub}\) given by the IPCC via
+Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates.
+Since the three different SOC sub-pools represent different SOC characteristics, the global decay rates and the associated drivers differ.
+Thus \(k_{sub}\) is given by
 
 \begin{equation}
-k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \underbrace{\cdot sf_{t,i} \quad \cdot till_{t,i}.}_{\text{only included for some subpools}}
+k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \quad \underbrace{\cdot till_{t,i} \quad \overbrace{\cdot sf_{t,i}.}^\text{only for sub = active}}_{\text{for sub = (active,slow)}}
 \label{eq:decayrates}
@@
 with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land use type \(lut\).
-Note that \(!lut\) denotes the (sum over) land use type(s), which decrease(s) in the specific time step \(t\).
+Note that \(!lut\) denotes the sum over all other land use types, which decreases in the specific time step \(t\).
 Data sources and methodology on land use states and changes are descripted in \ref{sec:landuse}.
@@
 
-Carbon stocks \(SOC\) for each subpool (sub) converge towards the calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut), each subpool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.
+Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.
 
@@
 \newpage
-
-\hypertarget{sec:tier1}{%
-\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}
-
-Following the tier 1 approach of the IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurements.
-The factors are differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-They can be seen as conservative guesses and will be used to evaluate our modelling based results.
-
-\hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management (50)}\label{sec:agrimanagement}}
-
-We combine data sets to estimate agricultural flows and management decisions on cropland.
-
-\hypertarget{sec:landuse}{%
-\subsubsection{Landuse and Landuse Change (150)}\label{sec:landuse}}
-
-We use the Land-Use Harmonization 2 (LUH2, \citep{LUH2}) half-degree data for major land-use states.
-Cropland is subdivided using country scale FAO data (\citep{FAOSTAT}) as well as crop type specific half-degree data from LUH2v2 to disaggregate into our 19 crop types on half degree resolution (see appendix for table on crop types and their mapping to FAOSTAT and LUH types as well as for technical discription of the disaggregation method).
-Landuse transitions are calculate as area differences, allowing any land use type to either expand or shrink in each time step.
-Within cropland we do not track area transitions, following the averaging approach as descripted in \ref{sec:tier2}/\ref{sec:carbonbudget}
-
-\hypertarget{crop-production-and-residues-300}{%
-\subsubsection{Crop Production and Residues (300)}\label{crop-production-and-residues-300}}
-
-FAOSTAT production (\citep{FAOSTAT}) values are combined with feed estimations including residue intake (\citep{weindl}) and rule based demand shares.
-LPJmL yield and calculate land use pattern (see @ref(\#sec:landuse)) are used to scale down to half-degree.
-
-\hypertarget{livestock-distribution-and-manure-excretion-300}{%
-\subsubsection{Livestock Distribution and Manure Excretion (300)}\label{livestock-distribution-and-manure-excretion-300}}
-
-Based on {[}Gridded Livestock of the world{]} we use rule based asumption to estimate livestock and manure distribution on the globe. Animal waste system shares are used as is {[}Bodirsky{]}.
-
-\hypertarget{irrigation-100}{%
-\subsubsection{Irrigation (100)}\label{irrigation-100}}
-
-Simple growing period calculations together with irrigation shares of LUH2v2 are use to estimate water effects on decay rates.
-
-\hypertarget{tillage-100}{%
-\subsubsection{Tillage (100)}\label{tillage-100}}
-
-Tillage data sets of {[}Vera, others{]} together with rules are used to drive tillage effect on decay rates.
 

### 2020-02-10 SOCbudgetPaper/methods.Rmd
rewising bennis comments, including docx, including newest resultNotebooks (with first plot)

@@
 We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon pools and entangled dynamics between them.
-Carbon inflow to each subpool (see \@ref(sec:carboninputs)) and decay rates (see \@ref(sec:tier2)) of each subpool are still the key components to determining steady-state SOC stocks.
+4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
+Carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and decay rates (see \@ref(sec:tier2)) of each sub-pool are still the key components to determining steady-state SOC stocks.
 
 ### Carbon Inputs to the Soil {#sec:carboninputs}
-BB: maybe consider to include as text instead of table. The table takes it a bit out of focus. Maybe you want to have the focus on this rather than on other parts.
+<!-- BB: maybe consider to include as text instead of table. The table takes it a bit out of focus. Maybe you want to have the focus on this rather than on other parts. -->
 
@@
 Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
-For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub pools.
+For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub-pools.
 This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
@@
 Following the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates $k_{sub}$ given by the IPCC via
+4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates.
+Since the three different SOC sub-pools represent different SOC characteristics, the global decay rates and the associated drivers differ.
+Thus $k_{sub}$ is given by
 
 \begin{equation}
-k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \underbrace{\cdot sf_{t,i} \quad \cdot till_{t,i}.}_{\text{only included for some subpools}}
+k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \quad \underbrace{\cdot till_{t,i} \quad \overbrace{\cdot sf_{t,i}.}^\text{only for sub = active}}_{\text{for sub = (active,slow)}}
 \label{eq:decayrates}
 \end{equation}
-
-BB: braket includes sandyness instead of water.
-I would leave out the braket, and just state that non-crop areas are neither tilled nor irrigated. then the formulas are still correct.
 
@@
 We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
-If land is converted from one land use type (lut) into another(!lut) (BB: like this?
-then you need to change the order of the !lut in the equation; lut is reduced, !lut is expanded), a respective share of the SOC stocks have to be reallocated as well.
+If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well.
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via
+
+<!--
+If land is converted from one land use type (lut) into another(!lut)
+(BB: like this? then you need to change the order of the !lut in the equation; lut is reduced, !lut is expanded)
+KK: Not sure, since it might be three different land use types there is no one into 'the' other.
+-->
 
@@
 with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land use type $lut$.
-Note that $!lut$ denotes the (sum over) (BB: why braket?) land use type(s), which decrease(s) in the specific time step $t$.
+Note that $!lut$ denotes the sum over all other land use types, which decreases in the specific time step $t$.
 Data sources and methodology on land use states and changes are descripted in \@ref(sec:landuse).
@@
 
-Carbon stocks $SOC$ for each subpool (sub) converge towards the calculate steady-state stock $SOC^{eq}$ for each land-use types (lut), each subpool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
+Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculate steady-state stock $SOC^{eq}$ for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
 
@@
 
+<!-- 
 BB: What is delta t doing in this equation? The value woudl be 1, no? 
+KK: Yes, but the unit is important, since k is 1/a, we need something with a as unit to counteract it.
+We could also just write \cdot 1\unit{a}. I had the feeling the Detla is more elegant.
+-->
 
@@
 
-<!-- Rewriting was till here!!!!!!!!!!! -->
-
 \newpage
 
+<!--
 ## Carbon Budget following Tier 1 (150) {#sec:tier1}
@@
 Within cropland we do not track area transitions, following the averaging approach as descripted in \@ref(sec:tier2)/\@ref(sec:carbonbudget)
-<!-- How to explain why we do not track inside-cropland transitions? --> BB: just dont explain it.
+-->
+<!-- How to explain why we do not track inside-cropland transitions? --> <!-- BB: just dont explain it. -->
 
+<!--
 ### Crop Production and Residues	(300)
@@
 Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.
+-->

### 2020-03-19 ResultNotebooks/results_socbudget.Rmd
rewising bennis comments, including docx, including newest resultNotebooks (with first plot)
update method part

@@
 title: "First SOC budget analysis"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
 ---
@@
 setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+#setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+setConfig(sourcefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/sources/")
 
@@
 
-```{r plot reg stocks}
+```{r plot reg stock, out.width="100%", out.height=1000}
 
 library(mip)
+library(quitte)
+library(ggplot2)
 
@@
 
-```{r Validation Densities}
+```{r table dens climatezones}
     cell_GSOC      <- readSource("GSOC",  convert="onlycorrect")
@@
 
+    koeppen     <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
+    getNames(koeppen) <- tolower(getNames(koeppen))
+    koepp2IPCC  <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
+    koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
+    climateIPCC <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
+    climateIPCC <- climateIPCC[,getYears(SoilCarbon),]
+
+    SOCDens_lut <- dimSums(SoilCarbon, dim=3.2) 
+    SoilCshare  <- SOCDens_lut[,,"crop"]/collapseNames(SOCDens_lut[,,"natveg"])
+    SoilCshare[is.na(SoilCshare)]       <- 1
+    SoilCshare[is.infinite(SoilCshare)] <- 1
+
+    SoilCshare_ipcczone <- dimSums(climateIPCC * SoilCshare * Landuse[,,"crop"], dim=1)/dimSums(climateIPCC * Landuse[,,"crop"], dim=1)
+    print(SoilCshare_ipcczone)
+
 ```
+
 ```{r plot density maps}
 
+cropland_share <- dimSums(SoilCarbon[,,"crop"], dim=3)/dimSums(SoilCarbon[,,"natveg"], dim=3)
+
+luplot::plotmap(cropland_share[,"y2000",])
+
 ```

### 2020-03-19 SOCbudgetPaper/SOCbudget.Rmd
update method part

--- 2020-01-22 SOCbudgetPaper/SOCbudget.Rmd
@@
 output:
+  rticles::copernicus_article: default
+  bookdown::word_document2: default
+  md_document:
+    variant: markdown_github
   bookdown::pdf_book:
     base_format: rticles::copernicus_article # for using bookdown features like \@ref()
-  rticles::copernicus_article: default
-  md_document:
-    variant: markdown_github
+    md_extensions: +raw_attribute
 ---
+\newpage
 

### 2020-03-19 SOCbudgetPaper/SOCbudget.tex
update method part

@@
 
-\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
+\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam,
+Germany}
 
@@
 
-\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstenst@pik-potsdam.de}{\nolinkurl{kristine.karstenst@pik-potsdam.de}})}
+\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}
 
@@
 \begin{abstract}
-SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
-this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
-models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these
-biophysical processes are still little investigated.
-Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and
-land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
-We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
-Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from \ldots{} . We also find that SOC is very
-sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
+SOC one of larges c sinks on earth (3 times larger biospehre pool).
+Agricultural management leads to a depletion of soil organic crabon.
+However this depletion of soil organic carbon (SOC) pools are so far not
+well represented in global assessments of historic carbon emissions.
+While SOC models often represent well the biochemical processes that
+lead to the accumulation and decay of SOC, the management decisions
+driving these biophysical processes are still little investigated. Here
+we create a spatial explicit data set for crop residue and manure
+management on cropland based on global historic production (FAOSTAT) and
+land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
+create a half-degree resolution soil organic carbon budget on mineral
+soils. We estimate that due to arable farming soils have lost over (?)
+GtOC of which (??) GtOC have been released within the period 1990-2010.
+Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than
+Tier 1 methods, which may origin from \ldots{} . We also find that SOC
+is very sensity to management decision such as residue recycling
+indicating the nessessity to incorporated better management data in soil
+models.
 \end{abstract}
 
-\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
+\copyrightstatement{The author's copyright for this publication is
+transferred to institution/company.}
 
@@
 
-Introduction text goes here.
-You can change the name of the section if neccessary using \texttt{\textbackslash{}introduction{[}modified\ heading{]}}.
-\newpage
+Introduction text goes here. You can change the name of the section if
+neccessary using
+\texttt{\textbackslash{}introduction{[}modified\ heading{]}}. \newpage
 
-\hypertarget{method-50}{%
-\section{Method (50)}\label{method-50}}
+\section{Method (50)}
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{Carbon Stocks following (new) Tier 2 method (50)}\label{sec:carbonbudget}}
+\subsection{Carbon Stocks following (new) Tier 2 method
+(50)}\label{sec:carbonbudget}}
 
-We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
+We calculate annual land use type specific soil organic carbon stocks
+for cropland, pastures and natural vegetation on half-degree resolution
+for the period of 1965 to 2010 based on the following three steps: (1)
+Calculating the land use (sub-)type specific steady-states and decay
+rates for SOC stocks given the current biophysical, climatic and
+agronomic conditions, (2) accounting for land conversation effects by
+transferring SOC between land use types and (3) updating SOC stocks
+based on the previous stock, the steady-state and the decay rate.
 
-\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
-\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}
+\subsubsection{Steady-state SOC stocks and decay rates}
 
-In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
-\begin{equation}
+In a simple first order kinetic approach the steady-state soil organic
+carbon stocks \(SOC^{eq}\) are given by \begin{equation}
 SOC^{eq} =\frac{C^{\textrm{in}}}{k}
 \label{eq:inoutflow}
-\end{equation}
-with \(C^{\textrm{in}}\) being carbon inputs to the soil and \(k\) denoting the soil organic carbon decay rate.
-We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
-Carbon inflow to each sub-pool (see \ref{sec:carboninputs}) and decay rates (see \ref{sec:tier2}) of each sub-pool are still the key components to determining steady-state SOC stocks.
+\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil
+and \(k\) denoting the soil organic carbon decay rate. We use for our
+calculations the steady-state method of the refinement of the IPCC
+guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which
+assume three soil carbon sub-pools (active, slow and passive) and
+entangled dynamics between them. Carbon inflow to each sub-pool (see
+@ref(sec:carboninputs)) and decay rates (see @ref(sec:tier2)) of each
+sub-pool are still the key components to determining steady-state SOC
+stocks.
 
@@
 
-We account for different carbon inputs sources depending on the land use type (see table \ref{tab:datasourceinputs}).
-Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see @ref(ipcc\_2019\_2019)).
-For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub-pools.
-This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
+We account for different carbon inputs sources depending on the land use
+type (see table @ref(tab:datasourceinputs)). Following the IPCC
+methodology carbon inputs are disaggregated into different structural
+components depending on their lignin and nitrogen content (see
+@ref(ipcc\_2019\_2019)). For each structural components the sum over all
+carbon inputs sources is allocated to the respective SOC sub-pools. This
+implies that not only the amount of carbon, but also their structural
+composition is determining the effective inflow. Data sources for all
+considered carbon inputs as well as for lignin and nitrogen content can
+be found in table @ref(tab:datasourceinputs).
 
@@
 
-Decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
-Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates.
-Since the three different SOC sub-pools represent different SOC characteristics, the global decay rates and the associated drivers differ.
-Thus \(k_{sub}\) is given by
+Decay rates are influenced by climatic conditions, biophysical and
+biochemical soil properties as well as management factors that vary over
+time (t) and space (i). Following the steady-state method of the
+refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for
+mineral soils we consider temperature (temp), water (wat), sand fraction
+(sf) and tillage (till) effects to spatially disaggregate default global
+decay rates. Since the three different SOC sub-pools represent different
+SOC characteristics, the global decay rates and the associated drivers
+differ. Thus \(k_{sub}\) is given by
 
 \begin{equation}
-k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \quad \underbrace{\cdot till_{t,i} \quad \overbrace{\cdot sf_{t,i}.}^\text{only for sub = active}}_{\text{for sub = (active,slow)}}
+\begin{aligned}
+& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
+& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
+& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
 \label{eq:decayrates}
+\end{aligned}
 \end{equation}
 
-For cropland we performed an assessment of tillage types and irrigation conditions, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as considered effects for each land use types are shown in table \ref{tab:datasourcedecay}.
-To account for variations of decay rates within grid cells based on different tillage and irrigation regimes, average rates based on area shares are calculated.
+For cropland we performed an assessment of tillage types and irrigation
+conditions, whereas on pastures and natural vegetation, we assume
+rainfed and non-tilled conditions. Data sources as well as considered
+effects for each land use types are shown in table
+@ref(tab:datasourcedecay). To account for variations of decay rates
+within grid cells based on different tillage and irrigation regimes,
+average rates based on area shares are calculated.
 
@@
 
-\hypertarget{soc-transfer-between-land-use-types}{%
-\subsubsection{SOC transfer between land use types}\label{soc-transfer-between-land-use-types}}
+\subsubsection{SOC transfer between land use types}
 
-We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
-If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well.
-We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
+We calculate SOC stocks based on the area shares of land use types (lut)
+within the half-degree grid cells (i). If land is converted from one
+land use type into others (!lut), a respective share of the SOC stocks
+have to be reallocated as well. We account for land conversion at the
+beginning of each time step \(t\) by calculating a preliminary stock
+\(SOC_{lut,t*}\) via
 
@@
 
-with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land use type \(lut\).
-Note that \(!lut\) denotes the sum over all other land use types, which decreases in the specific time step \(t\).
-Data sources and methodology on land use states and changes are descripted in \ref{sec:landuse}.
+with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
+expansion for a given land use type \(lut\). Note that \(!lut\) denotes
+the sum over all other land use types, which decreases in the specific
+time step \(t\). Data sources and methodology on land use states and
+changes are described in @ref(sec:landuse).
 
-\hypertarget{total-soc-stocks}{%
-\subsubsection{Total SOC stocks}\label{total-soc-stocks}}
+\subsubsection{Total SOC stocks}
 
-Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.
+Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
+calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut),
+each sub-pool (sub) and each annual time step (t) as represented in
+equation @ref(eq:steadystate).
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall lut, sub, i.
 \label{eq:steadystate}
@@
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{land use type specific SOC stock within a grid cell}}}_{\text{total SOC stock within a grid cell}}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ - land use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ - total SOC stock within cell}}
 \label{eq:totalstock}
@@
 
-\newpage
+\hypertarget{sec:tier1}{%
+\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+Following the tier 1 approach of the IPCC guidelines vol.~4
+(\citet{ipcc_2006_2006}), stocks are estimated via stock change factors
+given by the IPCC for the topsoil (0-30 cm) and based on measurement
+data and expert choice. The factors differentiate between different crop
+and management systems reflecting different dynamics under changed in-
+and outflows without explicitly tracking these. This method is used
+additionally to the Tier-2 approach to evaluate model accuracy.
+
+\hypertarget{sec:agrimanagement}{%
+\subsection{Agricultural management (50)}\label{sec:agrimanagement}}
+
+\hypertarget{sec:landuse}{%
+\subsubsection{Landuse and Landuse Change (150)}\label{sec:landuse}}
+
+Our land use patterns are based on the Land-Use Harmonization 2 (LUH2,
+\citep{LUH2}) data set, which we aggregate from quarter degree to half
+degree resolution. To transform the five different cropland
+subcategories of LUH2 into our crop functional types we define country
+and year specific area shares based on FAOSTAT data (\citep{FAOSTAT})
+(see @ref(appendix:fao2luh2mag). Where no data is available we fill data
+gaps by using similar countries as proxies or global averages. Land use
+transitions are calculated as area differences of the land use data,
+allowing any land use type to either expand or shrink in each time step.
+
+\subsubsection{Crop Production and Residues (300)}
+
+Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
+half-degree cropland pattern (see @ref(\#sec:landuse)) we compile crop
+functional type specific half-degree production patterns. We calibrate
+them to country level FAOSTAT production (\citep{FAOSTAT}) values to be
+in line with historical observations.
+
+Starting from this harvested fraction we calculate above ground and
+below ground residual biomass using harvest indices and shoot:root
+ratios as described in (\citep{bodirsky2012}). We assume that all below
+ground biomass is left on the fields, whereas above ground biosmass can
+be drawn away for other purposes such as feeding animals
+(\citep{weindl}), fuel or for material uses. We calculate recycle shares
+by subtracting all residue usages (see \citep{bodirsky1012}) from the
+total above ground biomass.
+
+To transform dry matter values into carbon quantities used as soil
+input, we compiled crop functional type specific carbon to dry matter
+(c:dm) ratios for harvested organs and applied a universal c:dm ratio of
+0.44 for all above and below ground residue biomass.
+
+\subsubsection{Pasture production}
+
+Using livestock production statistics as well as feed mix assumptions as
+describted in (\citep{weindl}) we estimating country specific pasture
+production.
+
+Using the same method as for crop production we disaggregate and
+calibrate half-degree pasture production pattern.
+
+\subsubsection{Livestock Distribution and Manure Excretion (300)}
+
+Since their is no native, biophysical indicator to disaggregate country
+level FAOSTAT livestock production values to half-degree pattern, we use
+the following rule based assumptions:
+
+\begin{itemize}
+\item
+  Based on the development state of a country, we devide poultry, egg
+  and monogastric meat production into intensive and extensive systems.
+\item
+  Extensive production is allocated to urban area share within a country
+  based on the assumption that these animals are held in households,
+  subsistence or small-holder farming systems that are labour rich.
+\item
+  Intensive production is distributed within a country using the crop
+  production (excluding 2nd generation bioenergy crops) share, assuming
+  that feed availability is the most driving factor for livestock
+  location.
+\item
+  We split dairy and ruminated meat production into pasture fed and
+  non-pasture fed fractions based on the aggregated feed mix within a
+  country
+\end{itemize}
+
+\begin{itemize}
+\item
+  Pasture fed animals are apportioned to half-degree cells according to
+  the pasture production shares of that cell within a country.
+\item
+  Non-pasture fed animals are distributed again using crop production
+  (excluding 2nd generation bioenergy crops) shares.
+\end{itemize}
+
+We calculate excretions using the nitrogen balance of the livestock
+system (\citep{bodirsky1012}) by assuming that every feed intake not
+incoorpared in the slaugher mass, has to excreted. Using C:N ratios
+(given by \citep[(][]{ipcc_2019_2019}) we estimate carbon content
+
+\subsubsection{Irrigation (100)}
+
+Simple growing period calculations together with irrigation shares of
+LUH2v2 are used (BB: you dont like d's, on't you?) to estimate water
+effects on decay rates.
+
+\subsubsection{Tillage (100)}
+
+Tillage data sets of {[}Vera, others{]} together with rules are used to
+drive tillage effect on decay rates. \newpage
+
+\section{Results}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}
 
@@
 \item
-  Carbon displacement via leaching and erosion is neglected in this study.
+  Carbon displacement via leaching and erosion is neglected in this
+  study.
 \item
   Non-net/Gross land use transitions are not tracked in this study.
-  \newpage
+\item
+  Within cropland we do not track area transitions, but rather look at
+  statistical distributions of the crop functional types. Due to crop
+  rotations and missing data on crop specific distributions, these
+  transitions would be any way rather uncertain. \newpage
 \end{itemize}
@@
 
-The conclusion goes here.
-You can modify the section name with \texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
+The conclusion goes here. You can modify the section name with
+\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
 \newpage
 
-\codedataavailability{use this to add a statement when having data sets and software code available} %% use this section when having data sets and software code available
+\codedataavailability{use this to add a statement when having data sets
+and software code
+available} %% use this section when having data sets and software code available
 
@@
 
-If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-They will be correctly named automatically.
+If you sorted all figures and tables into the sections of the text,
+please also sort the appendix figures and appendix tables into the
+respective appendix sections. They will be correctly named
+automatically.
 
@@
 
-If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
+If you put all figures after the reference list, please insert appendix
+tables and figures after the normal tables and figures.
 
-\texttt{\textbackslash{}appendixfigures} needs to be added in front of appendix figures
-\texttt{\textbackslash{}appendixtables} needs to be added in front of appendix tables
+\texttt{\textbackslash{}appendixfigures} needs to be added in front of
+appendix figures \texttt{\textbackslash{}appendixtables} needs to be
+added in front of appendix tables
 
-Please add \texttt{\textbackslash{}clearpage} between each table and/or figure. Further guidelines on figures and tables can be found below.
-Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-To rename them correctly to A1, A2, etc., please add the following commands in front of them:
+Please add \texttt{\textbackslash{}clearpage} between each table and/or
+figure. Further guidelines on figures and tables can be found below.
+Regarding figures and tables in appendices, the following two options
+are possible depending on your general handling of figures and tables in
+the manuscript environment: To rename them correctly to A1, A2, etc.,
+please add the following commands in front of them:
 \noappendix
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.} %% optional section
+\authorcontribution{Karstens wrote code and paper build on work of
+Bodirsky. Bodirsky and Popp revised paper.} %% optional section
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
+\competinginterests{The authors declare no competing
+interests.} %% this section is mandatory even if you declare that no competing interests are present
 

### 2020-03-19 SOCbudgetPaper/methods.Rmd
update method part

@@
 ### Carbon Inputs to the Soil {#sec:carboninputs}
-<!-- BB: maybe consider to include as text instead of table. The table takes it a bit out of focus. Maybe you want to have the focus on this rather than on other parts. -->
 
@@
 \begin{equation}
-k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \quad \underbrace{\cdot till_{t,i} \quad \overbrace{\cdot sf_{t,i}.}^\text{only for sub = active}}_{\text{for sub = (active,slow)}}
+\begin{aligned}
+& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
+& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
+& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
 \label{eq:decayrates}
+\end{aligned}
 \end{equation}
@@
 We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
-If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well.
+If land is converted from one land use type into others (!lut), a respective share of the SOC stocks have to be reallocated as well.
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via
-
-<!--
-If land is converted from one land use type (lut) into another(!lut)
-(BB: like this? then you need to change the order of the !lut in the equation; lut is reduced, !lut is expanded)
-KK: Not sure, since it might be three different land use types there is no one into 'the' other.
--->
 
@@
 Note that $!lut$ denotes the sum over all other land use types, which decreases in the specific time step $t$.
-Data sources and methodology on land use states and changes are descripted in \@ref(sec:landuse).
+Data sources and methodology on land use states and changes are described in \@ref(sec:landuse).
 
@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall lut, sub, i.
 \label{eq:steadystate}
 \end{equation}
-
-<!-- 
-BB: What is delta t doing in this equation? The value woudl be 1, no? 
-KK: Yes, but the unit is important, since k is 1/a, we need something with a as unit to counteract it.
-We could also just write \cdot 1\unit{a}. I had the feeling the Detla is more elegant.
--->
 
@@
 
-<!--
 ## Carbon Budget following Tier 1 (150) {#sec:tier1}
@@
 Following the tier 1 approach of the IPCC guidelines vol.
-4 (@ipcc_2006_2006), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurements.
+4 (@ipcc_2006_2006), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurement data and expert choice.
 The factors  differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
@@
 ## Agricultural management	(50) {#sec:agrimanagement}
-We combine data sets to estimate agricultural flows and management decisions on cropland. BB: wow. definetly shorter than 50 words! but do you need text hear at all?
 
@@
 
-We use the Land-Use Harmonization 2 (LUH2, [@LUH2]) half-degree data for major land-use states.
-Cropland is subdivided using country scale FAO data ([@FAOSTAT]) as well as crop type (BB: crop functional type?
-ots not crop-specific) specific half-degree data from LUH2v2 to disaggregate into our 19 crop types on half degree resolution (see appendix for table on crop types and their mapping to FAOSTAT and LUH types as well as for technical discription of the disaggregation method).
-Landuse transitions are calculated as area differences, allowing any land use type to either expand or shrink in each time step.
-Within cropland we do not track area transitions, following the averaging approach as descripted in \@ref(sec:tier2)/\@ref(sec:carbonbudget)
--->
-<!-- How to explain why we do not track inside-cropland transitions? --> <!-- BB: just dont explain it. -->
+Our land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
+To transform the five different cropland subcategories of LUH2 into our crop functional types we define country and year specific area shares based on FAOSTAT data ([@FAOSTAT]) (see \@ref(appendix:fao2luh2mag).
+Where no data is available we fill data gaps by using similar countries as proxies or global averages.
+Land use transitions are calculated as area differences of the land use data, allowing any land use type to either expand or shrink in each time step.
 
-<!--
 ### Crop Production and Residues	(300)
-FAOSTAT production ([@FAOSTAT]) values are combined with feed basket estimates including residue and pasture intake ([@weindl]) and rule based demand shares(BB:?).
-LPJmL yields and calculated land use patterns (see \@ref(#sec:landuse)) are used to scale down to half-degree.
+Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland pattern (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns.
+We calibrate them to country level FAOSTAT production ([@FAOSTAT]) values to be in line with historical observations.
+
+Starting from this harvested fraction we calculate above ground and below ground residual biomass using harvest indices and shoot:root ratios as described in ([@bodirsky2012]).
+We assume that all below ground biomass is left on the fields, whereas above ground biosmass can be drawn away for other purposes such as feeding animals ([@weindl]), fuel or for material uses.
+We calculate recycle shares by subtracting all residue usages (see [@bodirsky1012]) from the total above ground biomass.
+
+<!-- Add formular or drawings here? -->
+
+<!-- Ask benni if it is ok to just cide his paper here and not go into detail, how we distribute the residues -->
+
+To transform dry matter values into carbon quantities used as soil input, we compiled crop functional type specific carbon to dry matter (c:dm) ratios for harvested organs and applied a universal c:dm ratio of 0.44 for all above and below ground residue biomass.
+
+<!-- Do we have to explain our choice for the crop functional type specific data. Anyway we have to discuss here a bit, because the overall balance is not clear to me. -->
+
+### Pasture production
+Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production. 
+<!-- Is this true? Ask Isabelle again. -->
+
+Using the same method as for crop production we disaggregate and calibrate half-degree pasture production pattern.
+<!--As for cropland production we used pasture yields from LPJmL and pasture and rangeland patterns to derive half-degree pasture production patterns.
+Again country specific pasture production is used to calibrated the pattern.-->
 
 ### Livestock Distribution and Manure Excretion	(300)
-Based on [Gridded Livestock of the world] we use rule based assumptions(BB?) to estimate livestock and manure distribution on global scale(BB: more detail).
-Animal waste system shares are used as in [Bodirsky].
+Since their is no native, biophysical indicator to disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions:
+
+* Based on the development state of a country, we devide poultry, egg and monogastric meat production into intensive and extensive systems.
+
+* Extensive production is allocated to urban area share within a country based on the assumption that these animals are held in households, subsistence or small-holder farming systems that are labour rich.
+
+* Intensive production is distributed within a country using the crop production (excluding 2nd generation bioenergy crops) share, assuming that feed availability is the most driving factor for livestock location.
+
+* We split dairy and ruminated meat production into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country
+
+<!-- how to spell out that crop is everything else then pasture? -->
+
+* Pasture fed animals are apportioned to half-degree cells according to the pasture production shares of that cell within a country.
+
+* Non-pasture fed animals are distributed again using crop production (excluding 2nd generation bioenergy crops) shares.
+
+<!-- Add formular, drawings or visualisation here? -->
+
+<!-- TILL HERE -->
+
+We calculate excretions using the nitrogen balance of the livestock system ([@bodirsky1012]) by assuming that every feed intake not incoorpared in the slaugher mass, has to excreted.
+Using C:N ratios (given by [(@ipcc_2019_2019)]) we estimate carbon content
 
@@
 Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.
--->

### 2020-03-23 SOCbudgetPaper/methods.Rmd
Update methods.Rmd

@@
 
-## Agricultural management	(50) {#sec:agrimanagement}
+## Agricultural management data on 0.5 degrees	(50) {#sec:agrimanagement}
 
@@
 
-Our land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-To transform the five different cropland subcategories of LUH2 into our crop functional types we define country and year specific area shares based on FAOSTAT data ([@FAOSTAT]) (see \@ref(appendix:fao2luh2mag).
-Where no data is available we fill data gaps by using similar countries as proxies or global averages.
-Land use transitions are calculated as area differences of the land use data, allowing any land use type to either expand or shrink in each time step.
+Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
+We disaggregate the five different cropland subcategories of LUH2 into our X crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(appendix:fao2luh2mag)<!--glaube nicht dass du da noch n appendix brauchst-->.
+For X countries without FAOSTAT data, we fill data gaps by using similar countries as proxies or global averages<!--heisst das die haben HURTT data aber keine FAO daten?
+woher kriegt dann hurtt die daten?-->.
+Land use transitions are calculated as area differences of the land use data.
 
 ### Crop Production and Residues	(300)
-Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland pattern (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns.
-We calibrate them to country level FAOSTAT production ([@FAOSTAT]) values to be in line with historical observations.
+Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns.
+We calibrate cellular yields with one country-level calibration factor for each crop-group to meet historical FAOSTAT production ([@FAOSTAT]).
 
-Starting from this harvested fraction we calculate above ground and below ground residual biomass using harvest indices and shoot:root ratios as described in ([@bodirsky2012]).
-We assume that all below ground biomass is left on the fields, whereas above ground biosmass can be drawn away for other purposes such as feeding animals ([@weindl]), fuel or for material uses.
-We calculate recycle shares by subtracting all residue usages (see [@bodirsky1012]) from the total above ground biomass.
+Starting from these production estimates of the harvested organs and their respective crop area, we estimate above-ground and below-ground residual biomass using yield-dependent harvest indices and shoot:root ratios as described in ([@bodirsky2012]).
+We assume that all below ground biomass is left on the fields.
+A fixed share of the above-ground residues is assumed to be burnt on field, which depends on the per-capita income of the country.
+For low-income countries according to worldbank definitions (<1000 USD/yr), we assume a ...
+etc.
+ground biomass can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material uses.
+We calculate recycle shares by subtracting all residue usages (see [@bodirsky1012, @weindl2019]) from the total above ground biomass.
 
 <!-- Add formular or drawings here? --> 
+<!-- if you want, but first finish the rest -->
+<!-- Ask benni if it is ok to just cite his paper here and not go into detail, how we distribute the residues --> <!-- Benni: No, explain in detail as it is important for your manuscript and adds credibility to your detailed calcualtions.
+Also, some assumptions may have changed (e.g.
+DevelopmentState as function of gdp) -->
 
-<!-- Ask benni if it is ok to just cide his paper here and not go into detail, how we distribute the residues -->
-
-To transform dry matter values into carbon quantities used as soil input, we compiled crop functional type specific carbon to dry matter (c:dm) ratios for harvested organs and applied a universal c:dm ratio of 0.44 for all above and below ground residue biomass.
+To transform dry matter estimates into carbon, we compiled crop functional type specific carbon to dry matter (c:dm) ratios (see appendix table X) [citations] for harvested organs and applied a universal c:dm ratio of 0.44 [citation] for all above and below ground residue biomass.
 

### 2020-03-23 SOCbudgetPaper/methods.Rmd
Update methods.Rmd

@@
 ### Livestock Distribution and Manure Excretion	(300)
-Since their is no native, biophysical indicator to disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions:
 
-* Based on the development state of a country, we devide poultry, egg and monogastric meat production into intensive and extensive systems.
+To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which was inspired by the approach of [@gilbert].
 
-* Extensive production is allocated to urban area share within a country based on the assumption that these animals are held in households, subsistence or small-holder farming systems that are labour rich.
+For poultry, egg and monogastric meat production we 
+* divide into intensive and extensiv eproduction systems based on a per-capita (explain rules in more detail)
+
+* Based on the development state (thats not a fixed term) of a country, we devide poultry, egg and monogastric meat production into intensive and extensive systems.
+
+* Extensive production is allocated to built-up area share <!-- built-up is the more correct terminology for our urban are --> within a country based on the assumption that these animals are held in households, subsistence or small-holder farming systems.
 
@@
 
-* We split dairy and ruminated meat production into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country
+For dairy and ruminated meat production we
+
+* splot into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
 
@@
 
-* Pasture fed animals are apportioned to half-degree cells according to the pasture production shares of that cell within a country.
+* Pasture fed animals are distributed to half-degree cells according to the pasture production shares of that cell within a country (not methodologically clear description).
 
-* Non-pasture fed animals are distributed again using crop production (excluding 2nd generation bioenergy crops) shares.
+* Non-pasture fed animals are distributed again using crop production (excluding 2nd generation bioenergy crops) shares (not clear, either).
 
-<!-- Add formular, drawings or visualisation here? -->
+<!-- Add formular, drawings or visualisation here? --> <!-- No formulas needed, but precise wording -->
 
@@
 
-We calculate excretions using the nitrogen balance of the livestock system ([@bodirsky1012]) by assuming that every feed intake not incoorpared in the slaugher mass, has to excreted.
-Using C:N ratios (given by [(@ipcc_2019_2019)]) we estimate carbon content
+We calculate excretions by first estimating the nitrogen balance of the livestock system ([@bodirsky1012]) on the basis of comprehensive livestock feed baskets ([weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaugher mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]).
+What about animal waste management losses?
 
 ### Irrigation (100)
-Simple growing period calculations together with irrigation shares of LUH2v2 are used (BB: you dont like d's, on't you?) to estimate water effects on decay rates.
+Simple growing period calculations together with irrigation shares of LUH2v2 are used (BB: you dont like d's, on't you) to estimate water effects on decay rates.
 

### 2020-03-26 SOCbudgetPaper/SOCbudget.tex
update residue part (and before)

@@
 \hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management (50)}\label{sec:agrimanagement}}
+\subsection{Agricultural management data on 0.5 degrees
+(50)}\label{sec:agrimanagement}}
 
@@
 
-Our land use patterns are based on the Land-Use Harmonization 2 (LUH2,
+Land use patterns are based on the Land-Use Harmonization 2 (LUH2,
 \citep{LUH2}) data set, which we aggregate from quarter degree to half
-degree resolution. To transform the five different cropland
-subcategories of LUH2 into our crop functional types we define country
-and year specific area shares based on FAOSTAT data (\citep{FAOSTAT})
-(see @ref(appendix:fao2luh2mag). Where no data is available we fill data
-gaps by using similar countries as proxies or global averages. Land use
-transitions are calculated as area differences of the land use data,
-allowing any land use type to either expand or shrink in each time step.
+degree resolution. We disaggregate the five different cropland
+subcategories of LUH2 into our 17 crop groups, assuming relative shares
+for each gridcell based on the country and year specific area shares of
+FAOSTAT data (\citep{FAOSTAT}). Land use transitions are calculated as
+net area differences of the land use data on half-degree.
 
-\subsubsection{Crop Production and Residues (300)}
+\subsubsection{Crop, Crop Residues and Pasture Production (300)}
 
 Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
-half-degree cropland pattern (see @ref(\#sec:landuse)) we compile crop
+half-degree cropland patterns (see @ref(\#sec:landuse)) we compile crop
 functional type specific half-degree production patterns. We calibrate
-them to country level FAOSTAT production (\citep{FAOSTAT}) values to be
-in line with historical observations.
+cellular yields with one country-level calibration factor for each
+crop-group to meet historical FAOSTAT production (\citep{FAOSTAT}).
 
-Starting from this harvested fraction we calculate above ground and
-below ground residual biomass using harvest indices and shoot:root
-ratios as described in (\citep{bodirsky2012}). We assume that all below
-ground biomass is left on the fields, whereas above ground biosmass can
-be drawn away for other purposes such as feeding animals
-(\citep{weindl}), fuel or for material uses. We calculate recycle shares
-by subtracting all residue usages (see \citep{bodirsky1012}) from the
-total above ground biomass.
+Crop residues production is based on a revised methodology as described
+in (\citep{bodirsky2012}) and will be explained in key aspects again due
+to its central role for soil carbon modelling. Starting from crop
+production estimates of the harvested organs and their respective crop
+area, we estimate above-ground (ag) and below-ground (bg) residual
+biomass using yield-dependent harvest indices and shoot:root ratios. We
+assume that all bg residues are recycled to the soil, whereas ag
+residues can be burned or harvested for other purposes such as feeding
+animals (\citep{weindl}), fuel or for material use.
 
-To transform dry matter values into carbon quantities used as soil
-input, we compiled crop functional type specific carbon to dry matter
-(c:dm) ratios for harvested organs and applied a universal c:dm ratio of
-0.44 for all above and below ground residue biomass.
+A fixed share of the ag residues is assumed to be burned on field, which
+depends on the per-capita income of the country. Following
+\citep{smil1999}) we assume 25\% burn share for low-income countries
+according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\%
+for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolated
+shares for all middle-income countries.
 
-\subsubsection{Pasture production}
+Residue demand for feed is based on regional livestock specific feed
+basekts (see \citep{weindl}) taking available ag residual biomass as
+well as livestock production systems and thus again development state
+into account. Here we aggregate crop groups to residue groups based on
+their suitability as feed (see @ref(append:Table\_kcr2kres)).
+
+We estimated a material use share for the straw residues group of 5\%
+and a fuel demand share demand of 10\% in low income countries.
+
+The remaining ag residues are assumend to be recycled to the soil.
 
@@
 describted in (\citep{weindl}) we estimating country specific pasture
-production.
+production. Following the same approach as for crop production we
+disaggregate and calibrate half-degree pasture production pattern from
+grass yields from LPJmL and pasture area and rangeland patterns ( (see
+@ref(\#sec:landuse))) to derive half-degree pasture production patterns.
 
-Using the same method as for crop production we disaggregate and
-calibrate half-degree pasture production pattern.
+To transform dry matter estimates into carbon, we compiled crop and
+residue group specific carbon to dry matter (c:dm) ratios (see appendix
+table X) {[}citations{]} for harvested organs and applied a universal
+c:dm ratio of 0.44 {[}LPJml???{]} for all above and below ground residue
+biomass as well as for grass biomass.
 
@@
 
-Since their is no native, biophysical indicator to disaggregate country
-level FAOSTAT livestock production values to half-degree pattern, we use
-the following rule based assumptions:
+To disaggregate country level FAOSTAT livestock production values to
+half-degree pattern, we use the following rule based assumptions which
+was inspired by the approach of \citep{gilbert}.
+
+For poultry, egg and monogastric meat production we
 
@@
 \item
-  Based on the development state of a country, we devide poultry, egg
-  and monogastric meat production into intensive and extensive systems.
+  divide into intensive and extensiv eproduction systems based on a
+  per-capita (explain rules in more detail)
 \item
-  Extensive production is allocated to urban area share within a country
-  based on the assumption that these animals are held in households,
-  subsistence or small-holder farming systems that are labour rich.
+  Based on the development state (thats not a fixed term) of a country,
+  we devide poultry, egg and monogastric meat production into intensive
+  and extensive systems.
+\item
+  Extensive production is allocated to built-up area share within a
+  country based on the assumption that these animals are held in
+  households, subsistence or small-holder farming systems.
 \item
@@
   location.
-\item
-  We split dairy and ruminated meat production into pasture fed and
+\end{itemize}
+
+For dairy and ruminated meat production we split into pasture fed and
 non-pasture fed fractions based on the aggregated feed mix within a
-  country
-\end{itemize}
+country (explain rules)
 
@@
 \item
-  Pasture fed animals are apportioned to half-degree cells according to
-  the pasture production shares of that cell within a country.
+  Pasture fed animals are distributed to half-degree cells according to
+  the pasture production shares of that cell within a country (not
+  methodologically clear description).
 \item
   Non-pasture fed animals are distributed again using crop production
-  (excluding 2nd generation bioenergy crops) shares.
+  (excluding 2nd generation bioenergy crops) shares (not clear, either).
 \end{itemize}
 
-We calculate excretions using the nitrogen balance of the livestock
-system (\citep{bodirsky1012}) by assuming that every feed intake not
-incoorpared in the slaugher mass, has to excreted. Using C:N ratios
-(given by \citep[(][]{ipcc_2019_2019}) we estimate carbon content
+We calculate excretions by first estimating the nitrogen balance of the
+livestock system (\citep{bodirsky1012}) on the basis of comprehensive
+livestock feed baskets ({[}weindl{]}), assuming that all nitrogen in
+protein feed intake, minus the nitrogen in the slaugher mass, is
+excreted. Carbon in excreted manure is estimated by applying fixed C:N
+ratios (given by \citep[(][]{ipcc_2019_2019}). What about animal waste
+management losses?
 
@@
 Simple growing period calculations together with irrigation shares of
-LUH2v2 are used (BB: you dont like d's, on't you?) to estimate water
+LUH2v2 are used (BB: you dont like d's, on't you) to estimate water
 effects on decay rates.

### 2020-03-26 SOCbudgetPaper/methods.Rmd
update residue part (and before)

@@
 Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories of LUH2 into our X crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(appendix:fao2luh2mag)<!--glaube nicht dass du da noch n appendix brauchst-->.
-For X countries without FAOSTAT data, we fill data gaps by using similar countries as proxies or global averages<!--heisst das die haben HURTT data aber keine FAO daten?
-woher kriegt dann hurtt die daten?-->.
-Land use transitions are calculated as area differences of the land use data.
+We disaggregate the five different cropland subcategories of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]).
+Land use transitions are calculated as net area differences of the land use data on half-degree.
 
-### Crop Production and Residues	(300)
+### Crop, Crop Residues and Pasture Production	(300)
+
 Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns.
@@
 
-Starting from these production estimates of the harvested organs and their respective crop area, we estimate above-ground and below-ground residual biomass using yield-dependent harvest indices and shoot:root ratios as described in ([@bodirsky2012]).
-We assume that all below ground biomass is left on the fields.
-A fixed share of the above-ground residues is assumed to be burnt on field, which depends on the per-capita income of the country.
-For low-income countries according to worldbank definitions (<1000 USD/yr), we assume a ...
-etc.
-ground biomass can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material uses.
-We calculate recycle shares by subtracting all residue usages (see [@bodirsky1012, @weindl2019]) from the total above ground biomass.
+Crop residues production is based on a revised methodology as described in ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Starting from crop production estimates of the harvested organs and their respective crop area, we estimate above-ground (ag) and below-ground (bg) residual biomass using yield-dependent harvest indices and shoot:root ratios.
+We assume that all bg residues are recycled to the soil, whereas ag residues can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
+
+A fixed share of the ag residues is assumed to be burned on field, which depends on the per-capita income of the country.
+Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolated shares for all middle-income countries.
+
+Residue demand for feed is based on regional livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock production systems and thus again development state into account.
+Here we aggregate crop groups to residue groups based on their suitability as feed (see \@ref(append:Table_kcr2kres)).
+
+We estimated a material use share for the straw residues group of 5% and a fuel demand share demand of 10% in low income countries.
+
+The remaining ag residues are assumend to be recycled to the soil.
 
@@
 <!-- if you want, but first finish the rest -->
-<!-- Ask benni if it is ok to just cite his paper here and not go into detail, how we distribute the residues --> <!-- Benni: No, explain in detail as it is important for your manuscript and adds credibility to your detailed calcualtions.
-Also, some assumptions may have changed (e.g.
-DevelopmentState as function of gdp) -->
 
-To transform dry matter estimates into carbon, we compiled crop functional type specific carbon to dry matter (c:dm) ratios (see appendix table X) [citations] for harvested organs and applied a universal c:dm ratio of 0.44 [citation] for all above and below ground residue biomass.
+Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production.
+Following the same approach as for crop production we disaggregate and calibrate half-degree pasture production pattern from grass yields from LPJmL and pasture area and rangeland patterns ( (see \@ref(#sec:landuse))) to derive half-degree pasture production patterns.
 
-<!-- Do we have to explain our choice for the crop functional type specific data. Anyway we have to discuss here a bit, because the overall balance is not clear to me. -->
-
-### Pasture production
-Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production. 
-<!-- Is this true? Ask Isabelle again. -->
-
-Using the same method as for crop production we disaggregate and calibrate half-degree pasture production pattern.
-<!--As for cropland production we used pasture yields from LPJmL and pasture and rangeland patterns to derive half-degree pasture production patterns.
-Again country specific pasture production is used to calibrated the pattern.-->
+To transform dry matter estimates into carbon, we compiled crop and residue group specific carbon to dry matter (c:dm) ratios (see appendix table X) [citations] for harvested organs and applied a universal c:dm ratio of 0.44 [LPJml???] for all above and below ground residue biomass as well as for grass biomass.
 
@@
 For poultry, egg and monogastric meat production we
+
 * divide into intensive and extensiv eproduction systems based on a per-capita (explain rules in more detail)
@@
 
-For dairy and ruminated meat production we
-
-* splot into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
+For dairy and ruminated meat production we split into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
 

### 2020-03-28 SOCbudgetPaper/methods.Rmd
Update methods.Rmd

@@
 
-We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
+We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states <!-- BB: Do we calculate the steady states?
+I would say we are calculating budgets with in- and outflows, so its a dynamic state.
+Steady state we only use to initialize in 1960, no?
+This is not only for here but throughout the text.
+Steady state is misleading in my opinion, we are doing better :) --> and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
 
@@
 
-We account for different carbon inputs sources depending on the land use type (see table \@ref(tab:datasourceinputs)).
+We account for different carbon input sources depending on the land use type (see table \@ref(tab:datasourceinputs)).
 Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
@@
 
+<-- you also need a section on the initialisation, assuming a (now really) steady state for 1960 -->
+
 ## Carbon Budget following Tier 1 (150) {#sec:tier1}

### 2020-03-28 SOCbudgetPaper/methods.Rmd
Update methods.Rmd

@@
 
-We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states <!-- BB: Do we calculate the steady states?
+Following the tier 2 approach of the IPCC guidelines vol.
+4 (@ipcc_2006_2006), we calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states <!-- BB: Do we calculate the steady states?
 I would say we are calculating budgets with in- and outflows, so its a dynamic state.
 Steady state we only use to initialize in 1960, no?
+Maybe there is a steady-state subcalculation in the soil pool module, but the general approach is a budgeting approach, new=old+in-out, no?
 This is not only for here but throughout the text.
@@
 We account for different carbon input sources depending on the land use type (see table \@ref(tab:datasourceinputs)).
-Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
-For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub-pools.
+Following the IPCC methodology carbon inputs are disaggregated into different structural components <-- right wording?
+--> depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
+For each structural component the sum over all carbon input sources is allocated to the respective SOC sub-pools.
 This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
@@
 Following the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates.
-Since the three different SOC sub-pools represent different SOC characteristics, the global decay rates and the associated drivers differ.
+4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates <-- you dont disaggregate a global decay rate, do you?-->.
+Since the three different SOC sub-pools represent different SOC characteristics, the decay rates across the world differ <-- this sentence is rendundant with the previous.
+Remove completely?
+-->.
 Thus $k_{sub}$ is given by
@@
 
-For cropland we performed an assessment of tillage types and irrigation conditions, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
+For cropland we distinguish the effect of different tillage (section X) and irrigation (section Y) practices on decay rates <-- what do you want to say be assessment?
+-->, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
 Data sources as well as considered effects for each land use types are shown in table \@ref(tab:datasourcedecay).
-To account for variations of decay rates within grid cells based on different tillage and irrigation regimes, average rates based on area shares are calculated.
+To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
 
@@
 We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
-If land is converted from one land use type into others (!lut), a respective share of the SOC stocks have to be reallocated as well.
+If land is converted from one land use type into others (!lut), the respective share of the SOC stocks is reallocated.
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via
@@
 Data sources and methodology on land use states and changes are described in \@ref(sec:landuse).
+<-- if you want to talk about net-expansion, this is probably the better space than below -->
 
@@
 
-Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculate steady-state stock $SOC^{eq}$ for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
+Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
 
@@
 
+<-- you also need a section on the initialisation, assuming a (now really) steady state for 1960 -->
+
 \newpage
-
-<-- you also need a section on the initialisation, assuming a (now really) steady state for 1960 -->
 
@@
 
-Following the tier 1 approach of the IPCC guidelines vol.
-4 (@ipcc_2006_2006), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurement data and expert choice.
-The factors  differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-This method is used additionally to the Tier-2 approach to evaluate model accuracy.
+Additionally to the tier 2 approach of the IPCC guidelines vol.
+4 (@ipcc_2006_2006), we also estimate SOC pools using the IPCC tier 1 approach for comparison.
+Here, stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on a review of measurement data (and expert choice) <-- how important is the expert choice here?
+otherwise leave out, i guess all approaches require some expert info, and in the IPCC guidelines i think most parameters are rather based on literatre -->.
+The factors differentiate  different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+
+<-- also include an equation here -->
 
@@
 Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]).
+We disaggregate the five different cropland subcategories (name them) of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]).
 Land use transitions are calculated as net area differences of the land use data on half-degree.
@@
 We calibrate cellular yields with one country-level calibration factor for each crop-group to meet historical FAOSTAT production ([@FAOSTAT]).
+Crop yields are therefore also expressed as production by physical area, not by harvested area <-- or something similar expressing the problem of multiple harvests and fallow -->.
 
-Crop residues production is based on a revised methodology as described in ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
 Starting from crop production estimates of the harvested organs and their respective crop area, we estimate above-ground (ag) and below-ground (bg) residual biomass using yield-dependent harvest indices and shoot:root ratios.
@@
 A fixed share of the ag residues is assumed to be burned on field, which depends on the per-capita income of the country.
-Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolated shares for all middle-income countries.
+Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolate shares for all middle-income countries depending on their per-capita income.
 
-Residue demand for feed is based on regional livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock production systems and thus again development state into account.
-Here we aggregate crop groups to residue groups based on their suitability as feed (see \@ref(append:Table_kcr2kres)).
+Residue demand for feed is based on country and livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock production systems and thus again development state <-- are livestock systems dependent on development state?
+i dont think so--> into account.
+The three residue groups (straw, high-lignin and low-lignin residues) of the feed baskets are disaggregated to the crop groups, using....
+(see \@ref(append:Table_kcr2kres)).
+<-- its a disggregation, no -->
 
-We estimated a material use share for the straw residues group of 5% and a fuel demand share demand of 10% in low income countries.
+We estimated a material use share for the straw residues group of 5% and a fuel demand share demand of 10% <-- only for straw, or for all groups? --> in low income countries.
 
-The remaining ag residues are assumend to be recycled to the soil.
+The remaining ag residues as well as all bg residues are assumend to be recycled to the soil.
 
@@
 
-To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which was inspired by the approach of [@gilbert].
+To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert].
 

### 2020-04-02 SOCbudgetPaper/methods.Rmd
Update

@@
 
-Following the tier 2 approach of the IPCC guidelines vol.
-4 (@ipcc_2006_2006), we calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states <!-- BB: Do we calculate the steady states?
-I would say we are calculating budgets with in- and outflows, so its a dynamic state.
-Steady state we only use to initialize in 1960, no?
-Maybe there is a steady-state subcalculation in the soil pool module, but the general approach is a budgeting approach, new=old+in-out, no?
-This is not only for here but throughout the text.
-Steady state is misleading in my opinion, we are doing better :) --> and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
+Following the tier 2 approach of the refinement of IPCC guidelines vol.
+4 (@ipcc_2019_2019), we calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
 
@@
 We account for different carbon input sources depending on the land use type (see table \@ref(tab:datasourceinputs)).
-Following the IPCC methodology carbon inputs are disaggregated into different structural components <-- right wording?
---> depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
-For each structural component the sum over all carbon input sources is allocated to the respective SOC sub-pools.
+Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
+For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
 This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
@@
 
-Decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
+The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
 Following the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates <-- you dont disaggregate a global decay rate, do you?-->.
-Since the three different SOC sub-pools represent different SOC characteristics, the decay rates across the world differ <-- this sentence is rendundant with the previous.
-Remove completely?
--->.
+4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
 Thus $k_{sub}$ is given by
@@
 
-For cropland we distinguish the effect of different tillage (section X) and irrigation (section Y) practices on decay rates <-- what do you want to say be assessment?
--->, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
+For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
 Data sources as well as considered effects for each land use types are shown in table \@ref(tab:datasourcedecay).
@@
 Data sources and methodology on land use states and changes are described in \@ref(sec:landuse).
-<-- if you want to talk about net-expansion, this is probably the better space than below -->
 
@@
 
-<-- you also need a section on the initialisation, assuming a (now really) steady state for 1960 -->
+### Initialisation of SOC pools
+
+To initialize all SOC sub-pools we assume that
 
@@
 
-Additionally to the tier 2 approach of the IPCC guidelines vol.
-4 (@ipcc_2006_2006), we also estimate SOC pools using the IPCC tier 1 approach for comparison.
-Here, stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on a review of measurement data (and expert choice) <-- how important is the expert choice here?
-otherwise leave out, i guess all approaches require some expert info, and in the IPCC guidelines i think most parameters are rather based on literatre -->.
+Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.
+4 (@ipcc_2019_2019), we also estimate SOC pools using the IPCC tier 1 approach of IPCC guidelines vol.
+4 (@ipcc_2006_2006) for comparison.
+Here, stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on a review of measurement data.
 The factors differentiate  different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+The SOC stocks as thus calulated
 
-<-- also include an equation here -->
+\begin{equation}
+SOC_{t} = SOC_{t-1} - \frac{SOC_{t-1}-SOC_{\text{target}}}{D}
+\intertext{with}
+SOC_{\text{target}} = \sum_{c,s,i} SOC_{\text{ref}_{c,s,i}} \cdot F_{\text{LU}_{c,s,i}} \cdot F_{\text{MG}_{c,s,i}} \cdot F_{\text{I}_{c,s,i}} \cdot A_{c,s,i}}
+\label{eq:tier1}
+\end{equation}
+
+<!- also include an equation here -->
+<!- even if there are just "copied" out of te guidelines so to say? -->
+<!- more details will follow - how deep to go? -->
 
@@
 Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (name them) of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]).
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop type mapping).
 Land use transitions are calculated as net area differences of the land use data on half-degree.
@@
 
-Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns.
-We calibrate cellular yields with one country-level calibration factor for each crop-group to meet historical FAOSTAT production ([@FAOSTAT]).
-Crop yields are therefore also expressed as production by physical area, not by harvested area <-- or something similar expressing the problem of multiple harvests and fallow -->.
+Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop group specific half-degree production patterns.
+We calibrate cellular yields with one country-level calibration factor for each crop group to meet historical FAOSTAT production ([@FAOSTAT]).
+Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
 
@@
 
-Residue demand for feed is based on country and livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock production systems and thus again development state <-- are livestock systems dependent on development state?
-i dont think so--> into account.
+Residue demand for feed is based on country and livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock productivity into account.
 The three residue groups (straw, high-lignin and low-lignin residues) of the feed baskets are disaggregated to the crop groups, using....
@@
 
-To transform dry matter estimates into carbon, we compiled crop and residue group specific carbon to dry matter (c:dm) ratios (see appendix table X) [citations] for harvested organs and applied a universal c:dm ratio of 0.44 [LPJml???] for all above and below ground residue biomass as well as for grass biomass.
+To transform dry matter estimates into carbon, we compiled crop group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
+
+<!-- Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation -->
 
@@
 
-For poultry, egg and monogastric meat production we
-
-* divide into intensive and extensiv eproduction systems based on a per-capita (explain rules in more detail)
-
-* Based on the development state (thats not a fixed term) of a country, we devide poultry, egg and monogastric meat production into intensive and extensive systems.
-
-* Extensive production is allocated to built-up area share <!-- built-up is the more correct terminology for our urban are --> within a country based on the assumption that these animals are held in households, subsistence or small-holder farming systems.
+For poultry, egg and monogastric meat production we divide into intensive and extensive production systems based on a per-capita income of the country.
+For low-income countries according to worldbank definitions (<1000 USD/yr), we assume extensive production systems.
+We located them according to (built-up areas shares| population shares) based on the idea that these animals are held in households, subsistence or small-holder farming systems.
 
@@
 
-For dairy and ruminated meat production we split into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
+For dairy and ruminated meat production we
+
+* splot into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
 

### 2020-04-02 ResultNotebooks/results_socbudget.tex
Update pdf and results

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={First SOC budget analysis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{First SOC budget analysis}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(moinput)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: magclass
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'magclass'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:grid':
## 
##     getNames
\end{verbatim}

\begin{verbatim}
## Loading required package: madrat
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(mrvalidation)}
\KeywordTok{library}\NormalTok{(socbudget)}

\KeywordTok{setConfig}\NormalTok{(}\DataTypeTok{cachefolder =} \StringTok{"/home/kristine/WORK/R libraries/rd3mod\_inputdata/cache/rev4.18.cellpanda/"}\NormalTok{, }\DataTypeTok{forcecache =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Configuration update:
##   cachefolder: NA -> /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda
##   forcecache: FALSE -> TRUE
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")}
\KeywordTok{setConfig}\NormalTok{(}\DataTypeTok{sourcefolder =} \StringTok{"/home/kristine/WORK/R libraries/rd3mod\_inputdata/sources/"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Configuration update:
##   sourcefolder: NA -> /home/kristine/WORK/R libraries/rd3mod_inputdata/sources
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SoilCarbon <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"CarbonBudget"}\NormalTok{, }\DataTypeTok{aggregate =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("CarbonBudget", aggregate = FALSE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcCarbonBudget.Rda
## Exit calcOutput("CarbonBudget", aggregate = FALSE) in 10.32 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Landuse    <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"Landuse"}\NormalTok{,      }\DataTypeTok{aggregate=}\OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("Landuse", aggregate = FALSE)
\end{verbatim}

\begin{verbatim}
## Warning: More than one function found for type "Landuse" and prefix "calc". Use
## last occurrence (package "socbudget")
\end{verbatim}

\begin{verbatim}
## > WARNING: More than one function found for type " Landuse " and prefix " calc 
## > ". Use last occurrence (package " socbudget ")
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcLanduse.Rda
## Exit calcOutput("Landuse", aggregate = FALSE) in 1.87 seconds
\end{verbatim}

Calculate global carbon stocks

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SOCStock     <{-}}\StringTok{ }\KeywordTok{dimSums}\NormalTok{(}\KeywordTok{dimSums}\NormalTok{(SoilCarbon, }\DataTypeTok{dim=}\FloatTok{3.2}\NormalTok{)}\OperatorTok{*}\StringTok{ }\NormalTok{Landuse, }\DataTypeTok{dim=}\DecValTok{3}\NormalTok{)}
\NormalTok{glo\_SOCStock <{-}}\StringTok{ }\KeywordTok{dimSums}\NormalTok{(SOCStock, }\DataTypeTok{dim=}\DecValTok{1}\NormalTok{) }
\NormalTok{cell2iso     <{-}}\StringTok{ }\KeywordTok{toolGetMapping}\NormalTok{(}\StringTok{"CountryToCellMapping.csv"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"cell"}\NormalTok{)}
\NormalTok{iso\_SOCStock <{-}}\StringTok{ }\KeywordTok{toolCountryFill}\NormalTok{(}\KeywordTok{toolAggregate}\NormalTok{(SOCStock, }\DataTypeTok{rel=}\NormalTok{cell2iso, }\DataTypeTok{from=}\StringTok{"celliso"}\NormalTok{, }\DataTypeTok{to=}\StringTok{"iso"}\NormalTok{, }\DataTypeTok{dim=}\DecValTok{1}\NormalTok{), }\DataTypeTok{fill=}\DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## NOTE:  - toolCountryFill set missing values for IMPORTANT countries to  0 :
## NOTE:  --- Bahrain(BHR)  --- Hong Kong(HKG)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{iso2reg      <{-}}\StringTok{ }\KeywordTok{toolMappingFile}\NormalTok{(}\StringTok{"regional"}\NormalTok{,}\KeywordTok{getConfig}\NormalTok{(}\StringTok{"regionmapping"}\NormalTok{),}\DataTypeTok{readcsv =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{reg\_SOCStock <{-}}\StringTok{ }\KeywordTok{toolAggregate}\NormalTok{(iso\_SOCStock, }\DataTypeTok{rel=}\NormalTok{iso2reg, }\DataTypeTok{from=}\StringTok{"CountryCode"}\NormalTok{, }\DataTypeTok{to=}\StringTok{"RegionCode"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Load validation data:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  LPJ\_IPCC   <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"LPJ\_IPCC2006"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006",     aggregate = "REG+GLO", try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_LPJ_IPCC2006.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006",     aggregate = "REG+GLO", try = TRUE) in 0.34 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  LPJ\_rev21  <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"LPJmL\_rev21"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_LPJmL_rev21.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate = "REG+GLO",     try = TRUE) in 0.14 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  LPJ\_car    <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"LPJmLCarbon"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_LPJmLCarbon.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate = "REG+GLO",     try = TRUE) in 0.08 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  WISE       <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"WISE"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_WISE.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate = "REG+GLO",     try = TRUE) in 0.09 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  GSOC       <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"GSOC"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_GSOC.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate = "REG+GLO",     try = TRUE) in 0.09 seconds
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  SoilGrids  <{-}}\StringTok{ }\KeywordTok{calcOutput}\NormalTok{(}\StringTok{"ValidCarbonStocks"}\NormalTok{, }\DataTypeTok{datasource =} \StringTok{"SoilGrids"}\NormalTok{, }\DataTypeTok{aggregate=}\StringTok{"REG+GLO"}\NormalTok{, }\DataTypeTok{try=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Run calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate = "REG+GLO",     try = TRUE)
## >  - force cache 
## > /home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/calcValidCarbonStocks-datasource_SoilGrids.Rda
## Exit calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate = "REG+GLO",     try = TRUE) in 0.09 seconds
\end{verbatim}

Plot global stocks

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(glo\_SOCStock, }\DataTypeTok{as.integer=}\OtherTok{TRUE}\NormalTok{)), glo\_SOCStock, }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{, }\DataTypeTok{xlim=}\KeywordTok{c}\NormalTok{(}\DecValTok{1951}\NormalTok{,}\DecValTok{2017}\NormalTok{),  }\DataTypeTok{ylim=}\KeywordTok{c}\NormalTok{(}\DecValTok{400000}\NormalTok{,}\DecValTok{1300000}\NormalTok{))}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(LPJ\_IPCC, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), LPJ\_IPCC[}\StringTok{"GLO"}\NormalTok{,,}\DecValTok{1}\NormalTok{], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"green"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(LPJ\_car, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), LPJ\_car[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"darkgreen"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(LPJ\_rev21, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), LPJ\_rev21[}\StringTok{"GLO"}\NormalTok{,,}\DecValTok{2}\NormalTok{], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"lightgreen"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(WISE, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), WISE[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"brown"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(SoilGrids, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), SoilGrids[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"blue"}\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(GSOC, }\DataTypeTok{as.integer =} \OtherTok{TRUE}\NormalTok{), GSOC[}\StringTok{"GLO"}\NormalTok{,,], }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"pink"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{results_socbudget_files/figure-latex/plot glo stocks-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\KeywordTok{getYears}\NormalTok{(glo\_SOCStock, }\DataTypeTok{as.integer=}\OtherTok{TRUE}\NormalTok{)), glo\_SOCStock, }\DataTypeTok{type=}\StringTok{"l"}\NormalTok{, }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{results_socbudget_files/figure-latex/plot glo stocks-2.pdf}

Plot regional stocks

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(mip)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: quitte
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(quitte)}
\KeywordTok{library}\NormalTok{(ggplot2)}

\KeywordTok{getNames}\NormalTok{(reg\_SOCStock) <{-}}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\StringTok{"modeloutput.SOCbudget."}\NormalTok{, }\KeywordTok{getNames}\NormalTok{(WISE,}\DataTypeTok{dim=}\DecValTok{3}\NormalTok{))}
\KeywordTok{getSets}\NormalTok{(reg\_SOCStock)  <{-}}\StringTok{ }\KeywordTok{getSets}\NormalTok{(WISE) }

\NormalTok{hist <{-}}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{as.quitte}\NormalTok{(WISE[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(GSOC[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(SoilGrids[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(LPJ\_IPCC[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{][,,}\DecValTok{1}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(LPJ\_car[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{]),}
              \KeywordTok{as.quitte}\NormalTok{(LPJ\_rev21[}\StringTok{"GLO"}\NormalTok{,,,}\DataTypeTok{invert=}\OtherTok{TRUE}\NormalTok{][,,}\DecValTok{2}\NormalTok{]))}

\KeywordTok{mipLineHistorical}\NormalTok{(reg\_SOCStock, hist, }\DataTypeTok{facet.ncol=}\DecValTok{4}\NormalTok{, }\DataTypeTok{scales=}\StringTok{"free\_y"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=1\linewidth]{results_socbudget_files/figure-latex/reg c stock-1}

\begin{Shaded}
\begin{Highlighting}[]
    \CommentTok{\#cell\_GSOC      <{-} readSource("GSOC",  convert="onlycorrect")}
    \CommentTok{\#cell\_WISE      <{-} readSource("WISE",  convert="onlycorrect")}
    \CommentTok{\#cell\_SoilGrids <{-} readSource("SoilGrids",  convert="onlycorrect")}

    \CommentTok{\#koeppen     <{-} readSource("LPJml\_rev21", subtype ="koeppen", convert="onlycorrect" )}
    \CommentTok{\#koepp2IPCC  <{-} toolGetMapping("mapping\_koeppen\_ipcc.csv", type = "sectoral")}
    \CommentTok{\#climateIPCC <{-} toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen\_geiger", to="ipcc\_reduced")}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cropland\_share <{-}}\StringTok{ }\KeywordTok{dimSums}\NormalTok{(SoilCarbon[,,}\StringTok{"crop"}\NormalTok{], }\DataTypeTok{dim=}\DecValTok{3}\NormalTok{)}\OperatorTok{/}\KeywordTok{dimSums}\NormalTok{(SoilCarbon[,,}\StringTok{"natveg"}\NormalTok{], }\DataTypeTok{dim=}\DecValTok{3}\NormalTok{)}

\NormalTok{luplot}\OperatorTok{::}\KeywordTok{plotmap}\NormalTok{(cropland\_share[,}\StringTok{"y2000"}\NormalTok{,])}
\end{Highlighting}
\end{Shaded}

\includegraphics{results_socbudget_files/figure-latex/plot density maps-1.pdf}

\end{document}

### 2020-04-02 ResultNotebooks/results_socbudget_old.Rmd
Update pdf and results

--- 2020-03-19 ResultNotebooks/results_socbudget.Rmd
@@
 setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
-#setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(sourcefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/sources/")
+#setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43/", forcecache = TRUE)
+setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+#setConfig(sourcefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/sources/")
 
@@
 cell_WISE      <- readSource("WISE",  convert="onlycorrect")
-    cell_SoilGrids <- readSource("SoilGrids",  convert="onlycorrect")
+cell_SoilGrids <- readSource("SoilGrids", subtype="cstock_0_30", convert="onlycorrect")
 

### 2020-04-02 SOCbudgetPaper/SOCbudget.tex
Update pdf and results

@@
 
-We calculate annual land use type specific soil organic carbon stocks
-for cropland, pastures and natural vegetation on half-degree resolution
-for the period of 1965 to 2010 based on the following three steps: (1)
-Calculating the land use (sub-)type specific steady-states and decay
-rates for SOC stocks given the current biophysical, climatic and
-agronomic conditions, (2) accounting for land conversation effects by
-transferring SOC between land use types and (3) updating SOC stocks
-based on the previous stock, the steady-state and the decay rate.
+Following the tier 2 approach of the refinement of IPCC guidelines
+vol.~4 (\citet{ipcc_2019_2019}), we calculate annual land use type
+specific soil organic carbon stocks for cropland, pastures and natural
+vegetation on half-degree resolution for the period of 1965 to 2010
+based on the following three steps: (1) Calculating the land use
+(sub-)type specific steady-states and decay rates for SOC stocks given
+the current biophysical, climatic and agronomic conditions, (2)
+accounting for land conversation effects by transferring SOC between
+land use types and (3) updating SOC stocks based on the previous stock,
+the steady-state and the decay rate.
 
@@
 
-We account for different carbon inputs sources depending on the land use
+We account for different carbon input sources depending on the land use
 type (see table @ref(tab:datasourceinputs)). Following the IPCC
-methodology carbon inputs are disaggregated into different structural
-components depending on their lignin and nitrogen content (see
-@ref(ipcc\_2019\_2019)). For each structural components the sum over all
-carbon inputs sources is allocated to the respective SOC sub-pools. This
-implies that not only the amount of carbon, but also their structural
-composition is determining the effective inflow. Data sources for all
-considered carbon inputs as well as for lignin and nitrogen content can
-be found in table @ref(tab:datasourceinputs).
+methodology carbon inputs are disaggregated into metabolic and
+structural components depending on their lignin and nitrogen content
+(see @ref(ipcc\_2019\_2019)). For each component the sum over all carbon
+input sources is allocated to the respective SOC sub-pools via transfer
+coefficients. This implies that not only the amount of carbon, but also
+their structural composition is determining the effective inflow. Data
+sources for all considered carbon inputs as well as for lignin and
+nitrogen content can be found in table @ref(tab:datasourceinputs).
 
@@
 
-Decay rates are influenced by climatic conditions, biophysical and
-biochemical soil properties as well as management factors that vary over
-time (t) and space (i). Following the steady-state method of the
-refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for
-mineral soils we consider temperature (temp), water (wat), sand fraction
-(sf) and tillage (till) effects to spatially disaggregate default global
-decay rates. Since the three different SOC sub-pools represent different
-SOC characteristics, the global decay rates and the associated drivers
-differ. Thus \(k_{sub}\) is given by
+The sub-pool specific decay rates are influenced by climatic conditions,
+biophysical and biochemical soil properties as well as management
+factors that vary over time (t) and space (i). Following the
+steady-state method of the refinement of the IPCC guidelines vol.~4
+(\citet{ipcc_2019_2019}) for mineral soils we consider temperature
+(temp), water (wat), sand fraction (sf) and tillage (till) effects to
+account for spatial variation of decay rates. Thus \(k_{sub}\) is given
+by
 
@@
 
-For cropland we performed an assessment of tillage types and irrigation
-conditions, whereas on pastures and natural vegetation, we assume
-rainfed and non-tilled conditions. Data sources as well as considered
-effects for each land use types are shown in table
+For cropland we distinguish the effect of different tillage (see
+@ref(\#sec:tillage)) and irrigation (see @ref(\#sec:irrigation))
+practices on decay rates, whereas on pastures and natural vegetation, we
+assume rainfed and non-tilled conditions. Data sources as well as
+considered effects for each land use types are shown in table
 @ref(tab:datasourcedecay). To account for variations of decay rates
-within grid cells based on different tillage and irrigation regimes,
+within each grid cell due to different tillage and irrigation regimes,
 average rates based on area shares are calculated.
@@
 within the half-degree grid cells (i). If land is converted from one
-land use type into others (!lut), a respective share of the SOC stocks
-have to be reallocated as well. We account for land conversion at the
-beginning of each time step \(t\) by calculating a preliminary stock
-\(SOC_{lut,t*}\) via
+land use type into others (!lut), the respective share of the SOC stocks
+is reallocated. We account for land conversion at the beginning of each
+time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
 
@@
 Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
-calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut),
-each sub-pool (sub) and each annual time step (t) as represented in
-equation @ref(eq:steadystate).
+calculated steady-state stock \(SOC^{eq}\) for each land-use types
+(lut), each sub-pool (sub) and each annual time step (t) as represented
+in equation @ref(eq:steadystate).
 
@@
 
+\subsubsection{Initialisation of SOC pools}
+
+To initialize all SOC sub-pools we assume that
+
 \newpage
@@
 
-Following the tier 1 approach of the IPCC guidelines vol.~4
-(\citet{ipcc_2006_2006}), stocks are estimated via stock change factors
-given by the IPCC for the topsoil (0-30 cm) and based on measurement
-data and expert choice. The factors differentiate between different crop
-and management systems reflecting different dynamics under changed in-
-and outflows without explicitly tracking these. This method is used
-additionally to the Tier-2 approach to evaluate model accuracy.
+Additionally to the tier 2 approach of the refinement of IPCC guidelines
+vol.~4 (\citet{ipcc_2019_2019}), we also estimate SOC pools using the
+IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006})
+for comparison. Here, stocks are estimated via stock change factors
+given by the IPCC for the topsoil (0-30 cm) and based on a review of
+measurement data. The factors differentiate different crop and
+management systems reflecting different dynamics under changed in- and
+outflows without explicitly tracking these. The SOC stocks as thus
+calulated
+
+\begin{equation}
+SOC_{\text{target}} = \sum_{c,s,i} SOC_{\text{ref}_{c,s,i}} \cdot F_{\text{LU}_{c,s,i}} \cdot F_{\text{MG}_{c,s,i}} \cdot F_{\text{I}_{c,s,i}} \cdot A_{c,s,i}
+\label{eq:tier1}
+\end{equation}
+
+\textless!- also include an equation here --\textgreater{} \textless!-
+even if there are just ``copied'' out of te guidelines so to say?
+--\textgreater{} \textless!- more details will follow - how deep to go?
+--\textgreater{}
 
@@
 degree resolution. We disaggregate the five different cropland
-subcategories of LUH2 into our 17 crop groups, assuming relative shares
-for each gridcell based on the country and year specific area shares of
-FAOSTAT data (\citep{FAOSTAT}). Land use transitions are calculated as
-net area differences of the land use data on half-degree.
+subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17
+crop groups, assuming relative shares for each gridcell based on the
+country and year specific area shares of FAOSTAT data (\citep{FAOSTAT})
+(see @ref(append:Table\_luh2fao2mag) for more details on the crop type
+mapping). Land use transitions are calculated as net area differences of
+the land use data on half-degree.
 
@@
 half-degree cropland patterns (see @ref(\#sec:landuse)) we compile crop
-functional type specific half-degree production patterns. We calibrate
-cellular yields with one country-level calibration factor for each
-crop-group to meet historical FAOSTAT production (\citep{FAOSTAT}).
+group specific half-degree production patterns. We calibrate cellular
+yields with one country-level calibration factor for each crop group to
+meet historical FAOSTAT production (\citep{FAOSTAT}). Note that by using
+physical cropland areas we account for multiple crop harvest events as
+well as for fallows.
 
-Crop residues production is based on a revised methodology as described
-in (\citep{bodirsky2012}) and will be explained in key aspects again due
+Crop residue production and management is based on a revised methodology
+of (\citep{bodirsky2012}) and will be explained in key aspects again due
 to its central role for soil carbon modelling. Starting from crop
@@
 according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\%
-for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolated
-shares for all middle-income countries.
+for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate
+shares for all middle-income countries depending on their per-capita
+income.
 
-Residue demand for feed is based on regional livestock specific feed
+Residue demand for feed is based on country and livestock specific feed
 basekts (see \citep{weindl}) taking available ag residual biomass as
-well as livestock production systems and thus again development state
-into account. Here we aggregate crop groups to residue groups based on
-their suitability as feed (see @ref(append:Table\_kcr2kres)).
+well as livestock productivity into account. The three residue groups
+(straw, high-lignin and low-lignin residues) of the feed baskets are
+disaggregated to the crop groups, using\ldots. (see
+@ref(append:Table\_kcr2kres)). \textless-- its a disggregation, no
+--\textgreater{}
 
 We estimated a material use share for the straw residues group of 5\%
-and a fuel demand share demand of 10\% in low income countries.
+and a fuel demand share demand of 10\% \textless-- only for straw, or
+for all groups? --\textgreater{} in low income countries.
 
-The remaining ag residues are assumend to be recycled to the soil.
+The remaining ag residues as well as all bg residues are assumend to be
+recycled to the soil.
 
@@
 
-To transform dry matter estimates into carbon, we compiled crop and
-residue group specific carbon to dry matter (c:dm) ratios (see appendix
-table X) {[}citations{]} for harvested organs and applied a universal
-c:dm ratio of 0.44 {[}LPJml???{]} for all above and below ground residue
-biomass as well as for grass biomass.
+To transform dry matter estimates into carbon, we compiled crop group
+and plant part specific carbon to dry matter (c:dm) ratios (see
+@ref(append:Table\_c2dm)) (\citet{Ma2018}).
 
@@
 half-degree pattern, we use the following rule based assumptions which
-was inspired by the approach of \citep{gilbert}.
+were inspired by the approach of \citep{gilbert}.
 
-For poultry, egg and monogastric meat production we
+For poultry, egg and monogastric meat production we divide into
+intensive and extensive production systems based on a per-capita income
+of the country. For low-income countries according to worldbank
+definitions (\textless1000 USD/yr), we assume extensive production
+systems. We located them according to (built-up areas shares\textbar{}
+population shares) based on the idea that these animals are held in
+households, subsistence or small-holder farming systems.
 
 \begin{itemize}
-\item
-  divide into intensive and extensiv eproduction systems based on a
-  per-capita (explain rules in more detail)
-\item
-  Based on the development state (thats not a fixed term) of a country,
-  we devide poultry, egg and monogastric meat production into intensive
-  and extensive systems.
-\item
-  Extensive production is allocated to built-up area share within a
-  country based on the assumption that these animals are held in
-  households, subsistence or small-holder farming systems.
 \item
@@
   location.
+\item
+  For dairy and ruminated meat production we
+\item
+  splot into pasture fed and non-pasture fed fractions based on the
+  aggregated feed mix within a country (explain rules)
 \end{itemize}
-
-For dairy and ruminated meat production we split into pasture fed and
-non-pasture fed fractions based on the aggregated feed mix within a
-country (explain rules)
 

### 2020-04-02 SOCbudgetPaper/methods.Rmd
Update pdf and results

@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \frac{SOC_{t-1}-SOC_{\text{target}}}{D}
-\intertext{with}
-SOC_{\text{target}} = \sum_{c,s,i} SOC_{\text{ref}_{c,s,i}} \cdot F_{\text{LU}_{c,s,i}} \cdot F_{\text{MG}_{c,s,i}} \cdot F_{\text{I}_{c,s,i}} \cdot A_{c,s,i}}
+SOC_{\text{target}} = \sum_{c,s,i} SOC_{\text{ref}_{c,s,i}} \cdot F_{\text{LU}_{c,s,i}} \cdot F_{\text{MG}_{c,s,i}} \cdot F_{\text{I}_{c,s,i}} \cdot A_{c,s,i}
 \label{eq:tier1}
@@
 
-For dairy and ruminated meat production we
+* For dairy and ruminated meat production we
 

### 2020-04-03 SOCbudgetPaper/SOCbudget.tex
include figures in pdf and docx

@@
 
+\begin{figure}
+\includegraphics[width=12cm]{images/figs_draft-1} \caption{two column figure}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=12cm]{images/figs_draft-2} \caption{two column figure}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=12cm]{images/figs_draft-3} \caption{two column figure}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=12cm]{images/RegionPlot+Valid} \caption{two column figure}\label{fig:unnamed-chunk-9}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=12cm]{images/maps/CShare_1965} \includegraphics[width=12cm]{images/maps/CShare_2010} \caption{two column figure}\label{fig:unnamed-chunk-10}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=12cm]{images/maps/CIncrease_1965} \includegraphics[width=12cm]{images/maps/CIncrease_2010} \caption{two column figure}\label{fig:unnamed-chunk-11}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=12cm]{images/maps/CShare_4ColorClimates} \caption{one column figure}\label{fig:unnamed-chunk-12}
+\end{figure}
 \newpage

### 2020-05-03 ResultNotebooks/SOCflows_results.Rmd
methods update, add more result rmds

---
  title: "First SOC budget analysis"
output:
  html_document:
  df_print: paged
pdf_document: default
---

```{r load}
library(moinput)
library(mrvalidation)
library(socbudget)
library(magpiesets)

#input       <- "/home/kristine/WORK/R libraries/rd3mod_inputdata/output/rev1_690d3718e151be1b450b394c1064b1c5_validation.tgz"
#inputfolder <- gsub("[^/]*$","", input)
#untar(input, inputfolder)

setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")

```

```{r residues}

### Residues
  # originally from moinput
  ResidueDemand       <- dimSums(collapseNames(calcOutput("ResFieldBalancePast", cellular=TRUE, products = "kres", aggregate = FALSE)[,,"c"]), dim=1)
  ResidueRecyclingAg  <- ResidueDemand[,,"recycle"]
  ResidueRecyclingBg  <- dimSums(calcOutput("ResBiomass", cellular=TRUE, aggregate = FALSE)[,,"bg"][,,c("c")], dim=c(1,3))
  ResiduesMoinput     <- dimSums(ResidueRecyclingAg, dim=3) + ResidueRecyclingBg

  # Residues with cut of overflow
  ResiduesDensities   <- calcOutput("CarbonResidues", aggregate=FALSE)[,,"c"]
  Landuse             <- calcOutput("Landuse", aggregate=FALSE)
  ResiduesSOCbudget   <- dimSums(Landuse[,getYears(ResiduesDensities),"crop"]*ResiduesDensities, dim=c(1,3))

  print(mbind(ResiduesMoinput, ResiduesSOCbudget, ResiduesSOCbudget/ResiduesMoinput))

  # Corrected Residues shares for different plant parts  
  RatioPlantParts     <- mbind(ResidueRecyclingAg, setNames(ResidueRecyclingBg, "res_roots"))/(dimSums(ResidueRecyclingAg, dim=3)+ResidueRecyclingBg)

  ResiduesCorrected   <- collapseNames(RatioPlantParts * ResiduesSOCbudget)

  # total residues and residue demands

  ResidueBg     <- ResidueRecyclingBg
  ResidueAg     <- dimSums(calcOutput("ResBiomass", cellular=TRUE, aggregate = FALSE)[,,"ag"][,,c("c")], dim=c(1,3))
  ResidueTotal  <- ResidueAg + ResidueBg

```

Quite a lot of residues get lost, due to cutting extreme values in the later years.
Cutting algorithm, cuts everything bigger the 95%-percentile.
How we deal with that? 
* Should we do a yearly extrem value cutting?
* Should we investigate that deeper?
* Should we leave it like this?

```{r manure}

### Manure
  # originally from moinput
  ManureApplication   <- calcOutput("ManureRecyclingCroplandPast", products = "kli", cellular = TRUE, aggregate = FALSE)[,,"c"]
  ManureGrazing       <- calcOutput("Excretion", cellular = TRUE, attributes = "npkc", aggregate = FALSE)[,,"stubble_grazing"][,,"c"]
  ManureMoinput       <- collapseNames(dimSums(ManureApplication, dim=1) + dimSums(ManureGrazing, dim=1))

  # Manure with cut of overflow
  ManureDensities     <- calcOutput("CarbonManure", aggregate=FALSE)[,,"c"]
  Landuse             <- calcOutput("Landuse", aggregate=FALSE)
  ManureSOCbudget     <- dimSums(Landuse[,getYears(ManureDensities),"crop"]*ManureDensities, dim=1)

  print(mbind(dimSums(ManureMoinput, dim=3), dimSums(ManureSOCbudget, dim=3), dimSums(ManureSOCbudget, dim=3)/dimSums(ManureMoinput, dim=3)))

  ManureCorrected     <- ManureSOCbudget

  ManureExcreted      <- dimSums(calcOutput("Excretion", cellular = TRUE, attributes = "npkc", aggregate = FALSE)[,,"c"][,,c("confinement","stubble_grazing")], dim=c(1,3.3))

  print(mbind(dimSums(ManureExcreted, dim=3), dimSums(ManureMoinput, dim=3), dimSums(ManureCorrected, dim=3)))
  print(quantile(ManureApplication+ManureGrazing, probs=seq(0,1,0.02)))
  range(ManureApplication+ManureGrazing)
```

Same problem here. Manure share is going down.

```{r production}

#Production     <- collapseNames(calcOutput("Production", products="kcr", cellular=TRUE, attributes="dm", irrigation=FALSE, aggregate = FALSE))
Balance        <- calcOutput("FAOmassbalance", aggregate=FALSE)

CropProduction <- dimSums(Balance[,,findset("kcr")][,,"production"][,,"c"], dim=c(1,3))
FeedProduction <- dimSums(Balance[,,c("pasture",findset("kap"),findset("kres"),"scp"), invert=TRUE][,,"feed"][,,"c"], dim=c(1,3))
ResidueFeed    <- dimSums(Balance[,,findset("kres")][,,"feed"][,,"c"], dim=c(1,3))

print(mbind(CropProduction,FeedProduction,ResidueFeed))
```

Biomass excreted ist bigger than feed intake via plant harvest and residues taken for feed.

```{r soc}

  Litter             <- dimSums(calcOutput("CarbonLitter", aggregate=FALSE)[,,"c"]*Landuse[,,"natveg"], dim=1)

  # Calculate effective input
  param <- readSource("IPCCSoil", convert=FALSE)
  param.f1_metab2a          <- setYears(param[,,"f1"], NULL) 
  # stabilization efficiencies for metabolic decay products entering the active pool
  param.f3_struc2s          <- setYears(param[,,"f3"], NULL) 
  # stabilization efficiencies for structural decay products entering the slow pool
  param.f2_struc2a <- mbind(setNames(setYears(param[,,"f2_ft"], NULL), "crop"),
                            setNames(setYears(param[,,"f2_nt"], NULL), "natveg"))
  # stabilization efficiencies for structural decay products entering the active pool (tillage and no tillage specific)

  cell.input       <- mbind(add_dimension(dimSums(calcOutput("CarbonInput", landtype="crop", aggregate = FALSE),dim=3.2),
                                          dim=3.1, nm = "crop",   add="landuse"),
                            add_dimension(collapseNames(calcOutput("CarbonInput", landtype="natveg", aggregate = FALSE)),
                                          dim=3.1, nm = "natveg", add="landuse"))
  CarbonInput      <- dimSums(cell.input[,,"totalC"]*Landuse, dim=c(1))

  CarbonInputEff   <- collapseNames(dimSums((cell.input[,,"metabDOC"] * param.f1_metab2a +
                                             cell.input[,,"strucDOC"] * param.f2_struc2a +
                                             cell.input[,,"ligninC"]  * param.f3_struc2s) * Landuse, dim=c(1)))

  CarbonInputTot   <- collapseNames(dimSums(SOC[,,"steadystate"]*SOC[,,"decayrate"]*Landuse, dim=c(1,3.3)))

  # Load inputs, active steady state stock and decay rate

  SOC                <- calcOutput("SoilCarbon", output="full", aggregate=FALSE)
  Landuse            <- calcOutput("Landuse",       aggregate=FALSE)
  LanduseChange      <- calcOutput("LanduseChange", aggregate=FALSE)

  CarbonStocks       <- collapseNames(dimSums(SOC[,,c("steadystate","actualstate")]*Landuse, dim=1))
  CarbonTransfer     <- dimSums(toolConditionalReplace(SOC[,,"carbontransfer"], "<0", 0),dim=c(1,3.3))
  CarbonRelease      <- collapseNames(dimSums(SOC[,,"interstate"]*SOC[,,"decayrate"]*Landuse, dim=c(1, 3.3)))

print(mbind(CarbonTransfer, CarbonRelease))
```

### 2020-05-03 ResultNotebooks/results_socbudget.Rmd
methods update, add more result rmds

@@
 
-setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
-#setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(sourcefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/sources/")
+setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
+#setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
+#setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43/", forcecache = TRUE)
+#setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43/", forcecache = TRUE)
+setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+#setConfig(sourcefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/sources/")
 
-SoilCarbon <- calcOutput("CarbonBudget", aggregate = FALSE)
+#SoilCarbon <- readRDS("/home/kristine/WORK/Paper/historicalsocbudget/ResultNotebooks/soilcarbon.rds")
+SoilCarbon <- calcOutput("SoilCarbon", output="full", aggregate = FALSE)
 Landuse    <- calcOutput("Landuse",      aggregate=FALSE)
@@
 
-SOCStock     <- dimSums(dimSums(SoilCarbon, dim=3.2)* Landuse, dim=3)
+SOCStock     <- dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=3)
 glo_SOCStock <- dimSums(SOCStock, dim=1) 
@@
 reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
+
+```
+
+```{r Carbon loss}
+SOCStock     <- dimSums(dimSums(SoilCarbon, dim=3.2)* Landuse[,,"crop"], dim=c(1,3.3))
+
+SOCDiff      <- SOCStock[,,"crop"] - SOCStock[,,"natveg"]
+
+plot(SOCDiff)
 
@@
               as.quitte(SoilGrids["GLO",,,invert=TRUE]),
-              as.quitte(LPJ_IPCC["GLO",,,invert=TRUE][,,1]),
-              as.quitte(LPJ_car["GLO",,,invert=TRUE]),
               as.quitte(LPJ_rev21["GLO",,,invert=TRUE][,,2]))
 
-mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y")
+mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y", ylab="SOC (0-30cm) in MtC")
 
+print(dimSums(reg_SOCStock, dim=1))
+```
+```{r barplot reg stock, out.width="100%", out.height=1000}
+
+library(mip)
+library(quitte)
+library(ggplot2)
+
+getNames(reg_SOCStock) <- paste0("modeloutput.SOCbudget.", getNames(WISE,dim=3))
+getSets(reg_SOCStock)  <- getSets(WISE) 
+year <- "y2010"
+
+all <- rbind( as.quitte(reg_SOCStock[,year,]),
+              as.quitte(WISE["GLO",,,invert=TRUE][,1,]),
+              as.quitte(GSOC["GLO",,,invert=TRUE][,1,]),
+              as.quitte(SoilGrids["GLO",,,invert=TRUE][,1,]),
+              as.quitte(LPJ_IPCC["GLO",,,invert=TRUE][,,1][,year,]),
+              as.quitte(LPJ_car["GLO",,,invert=TRUE][,year,]),
+              as.quitte(LPJ_rev21["GLO",,,invert=TRUE][,,2][,year,]))
+all$scenario <- all$model
+all$model    <- "HistSOC"
+
+mipBarYearData(reg_SOCStock)
 ```
@@
 ```{r table dens climatezones}
-    cell_GSOC      <- readSource("GSOC",  convert="onlycorrect")
-    cell_WISE      <- readSource("WISE",  convert="onlycorrect")
-    cell_SoilGrids <- readSource("SoilGrids",  convert="onlycorrect")
 
@@
 
-    SOCDens_lut <- dimSums(SoilCarbon, dim=3.2) 
-    SoilCshare  <- SOCDens_lut[,,"crop"]/collapseNames(SOCDens_lut[,,"natveg"])
-    SoilCshare[is.na(SoilCshare)]       <- 1
-    SoilCshare[is.infinite(SoilCshare)] <- 1
+    SOCDens_lut <- dimSums(SoilCarbon[,,c("steadystate","actualstate")], dim=3.3)
 
-    SoilCshare_ipcczone <- dimSums(climateIPCC * SoilCshare * Landuse[,,"crop"], dim=1)/dimSums(climateIPCC * Landuse[,,"crop"], dim=1)
-    print(SoilCshare_ipcczone)
+    SoilCshare_ipcczone2 <- dimSums(climateIPCC * SOCDens_lut[,,"crop"] * Landuse[,,"crop"], dim=1)/
+                            dimSums(climateIPCC * SOCDens_lut[,,"natveg"] * Landuse[,,"crop"], dim=1)
+
+    print(collapseNames(SoilCshare_ipcczone2[,findset("past"),]))
 
@@
 
-cropland_share <- dimSums(SoilCarbon[,,"crop"], dim=3)/dimSums(SoilCarbon[,,"natveg"], dim=3)
+cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3)
+cropland_share[Landuse[,,"crop"]==0] <- NA
 
-luplot::plotmap(cropland_share[,"y2000",])
+cropland_increase <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)-dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3) 
+cropland_increase[Landuse[,,"crop"]==0] <- NA
+
+library(RColorBrewer)
+cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(1000)
+outputdir <- "/home/kristine/WORK/Paper/historicalsocbudget/SOCbudgetPaper/images/maps/"
+
+png(paste0(outputdir,"CShare_1965.png"),width = 1280, height = 720, res = 108)
+luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
+dev.off()
+png(paste0(outputdir,"CShare_2010.png"),width = 1280, height = 720, res = 108)
+luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
+dev.off()
+
+png(paste0(outputdir,"CIncrease_1965.png"),width = 1280, height = 720, res = 108)
+luplot::plotmap(cropland_increase[,"y1965",], legend_range = c(-80,80), legend_colours = cols, main="Absoulte C change on Cropland 1965")
+dev.off()
+
+png(paste0(outputdir,"CIncrease_2010.png"),width = 1280, height = 720, res = 108)
+luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols, main="Absoulte C change on Cropland 2010")
+dev.off()
+
+library(RColorBrewer)
+pal_red    <- c("#110044","#774499","#FF0088","#FFAADD")
+pal_green  <- c("#004400","#00FF00","#FFFF00","#FFFFCC")
+pal_blue   <- c("#000044","#0000FF","#00FFFF","#CCFFFF")
+pal_yellow <- c("#FFDD99","#feb24c","#FF0000","#440000")
+
+pal_red    <- c("#FFAAAA","#AAFFFF")
+pal_green  <- c("#AAFFAA","#FFAAFF")
+pal_blue   <- c("#AAAAFF","#FFFFAA")
+pal_grey   <- c("#AAAAAA","#222222")
+
+cols <- c(colorRampPalette(pal_red)(500)[500:1],
+          colorRampPalette(pal_yellow)(500),
+          colorRampPalette(pal_blue)(500)[500:1],
+          colorRampPalette(pal_green)(500)[500:1]
+          )
+
+test <- cropland_share
+test[Landuse[,,"crop"]==0] <- NA
+test[test>2]  <- 1.9999
+test[test==0] <- 0.0001
+
+koeppen     <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
+getNames(koeppen) <- tolower(getNames(koeppen))
+koepp2IPCC  <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
+koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
+climateIPCC <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
+climateIPCC <- climateIPCC[,getYears(SoilCarbon),]
+
+test <- test*climateIPCC[,,1]+(test+2)*climateIPCC[,,2]+(test+4)*climateIPCC[,,3]+(test+6)*climateIPCC[,,4]
+luplot::plotmap(test[,"y2010",], legend_range = c(0,8), legend_colours = cols)
+
+outputdir <- "/home/kristine/WORK/Paper/historicalsocbudget/SOCbudgetPaper/images/maps/"
+png(paste0(outputdir,"CShare_4ColorClimates.png"),width = 1280, height = 720, res = 108)
+luplot::plotmap(test[,"y2010",], legend_range = c(0,8), legend_colours = cols, main="Stock Change Factor on Cropland 2010 (climate zones color coded)")
+dev.off()
 
 ```
+
+```{r plot density maps}
+
+cols <- colorRampPalette(brewer.pal(9,"Oranges"))(1000)
+cropland <- dimSums(SoilCarbon[,,"crop"], dim=3)
+cropland[Landuse[,,"crop"]==0] <- NA
+
+luplot::plotmap(cropland[,"y2010",], legend_colours = cols, legend_range=range(cropland[,"y2010",]/5, na.rm=TRUE))
+
+natland <- dimSums(SoilCarbon[,,"natveg"], dim=3)
+
+luplot::plotmap(natland[,"y2010",], legend_colours = cols, legend_range=range(cropland[,"y2010",]/5, na.rm=TRUE))
+
+```
+
+```{r output carbon flows}
+
+```

### 2020-05-03 ResultNotebooks/scf_colormaps.Rmd
methods update, add more result rmds

---
  title: "Stock Change Factor maps (coloured)"
output:
  html_document:
  df_print: paged
pdf_document: default
---

```{r load data}
library(moinput)

SoilCarbon <- readRDS("/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_SOC_0.5.0/SoilCarbon.rds")
Landuse    <- readRDS("/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_SOC_0.5.0/Landuse.rds")

setConfig(cachefolder  = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43/", forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(mappingfolder = "/home/kristine/mnt/rd3mod/inputdata/mappings/")

koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
getNames(koeppen) <- tolower(getNames(koeppen))
koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
climateIPCC       <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
climateIPCC       <- climateIPCC[,getYears(SoilCarbon),]

```

```{r calculate SCF}

# one colour scheme
cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/
                  dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3)
cropland_share[Landuse[,,"crop"]==0] <- NA

# 4 colour scheme
test          <- cropland_share
test[Landuse[,,"crop"]==0] <- NA
test[test>2]  <- 1.9999
test[test==0] <- 0.0001
test          <- test*climateIPCC[,,1]+(test+2)*climateIPCC[,,2]+(test+4)*climateIPCC[,,3]+(test+6)*climateIPCC[,,4]

```

```{r plot density maps}

# No climate zone specification
library(RColorBrewer)
cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(1000)

luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
png("CShare_1965.png",width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
dev.off()

luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
png("CShare_2010.png",width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
dev.off()

```

```{r plot density maps 4 colours}

# With climate zone specification
library(RColorBrewer)
pal_pink   <- c("#330022","#aa99aa","#ccbbcc","#FFBBEE")
pal_green  <- c("#003300","#99aa99","#bbccbb","#CCFFCC")
pal_blue   <- c("#000044","#9999aa","#bbbbcc","#CCFFFF")
pal_red    <- c("#FFEECC","#ccc2bb","#aaa299","#440000")

pal_pink   <- c("#110044","#774499","#aaaaaa","#FF0088","#FFAADD")
pal_green  <- c("#004400","#00FF00","#aaaaaa","#FFFF00","#FFFFCC")
pal_blue   <- c("#000044","#0000FF","#aaaaaa","#00FFFF","#CCFFFF")
pal_red    <- c("#FFDD99","#feb24c","#aaaaaa","#FF0000","#440000")

cols <- c(colorRampPalette(pal_pink)(500)[500:1],
          colorRampPalette(pal_red)(500),
          colorRampPalette(pal_blue)(500)[500:1],
          colorRampPalette(pal_green)(500)[500:1]
)

luplot::plotmap(test[,"y2010",], legend_range = c(0,8), legend_colours = cols, main="Stock Change Factor on Cropland 2010 (climate zones color coded)")
png("CShare_4ColorClimates.png", width = 1280, height = 720, res = 108)

luplot::plotmap(test[,"y2010",], legend_range = c(0,8), legend_colours = cols, main="Stock Change Factor on Cropland 2010 (climate zones color coded)")
dev.off()
```

### 2020-05-03 SOCbudgetPaper/SOCbudget.Rmd
methods update, add more result rmds

--- 2020-01-28 SOCbudgetPaper/SOCbudget.Rmd
@@
 abstract: | 
- SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
- this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
- models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these 
- biophysical processes are still little investigated.
- Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and 
- land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
+ SOC one of larges c sinks on earth (3 times larger biosphere pool). 
+ Agricultural management leads to a depletion of soil organic carbon.
+However this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions.
+ While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated.
+ Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half degree resolution soil organic carbon budget on mineral soils.
 We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
- Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from … . We also find that SOC is very 
- sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models. 
+ We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
+We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
 bibliography: SOCbudget.bib
@@
 output:
+  rticles::copernicus_article: default
+  md_document:
+    variant: markdown_github
+  bookdown::word_document2: default
   bookdown::pdf_book:
@@
     md_extensions: +raw_attribute
-  rticles::copernicus_article: default
-  md_document:
-    variant: markdown_github
 ---

### 2020-05-03 SOCbudgetPaper/SOCbudget.tex
methods update, add more result rmds

@@
 \begin{abstract}
-SOC one of larges c sinks on earth (3 times larger biospehre pool).
-Agricultural management leads to a depletion of soil organic crabon.
+SOC one of larges c sinks on earth (3 times larger biosphere pool).
+Agricultural management leads to a depletion of soil organic carbon.
 However this depletion of soil organic carbon (SOC) pools are so far not
@@
 land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
-create a half-degree resolution soil organic carbon budget on mineral
+create a half degree resolution soil organic carbon budget on mineral
 soils. We estimate that due to arable farming soils have lost over (?)
 GtOC of which (??) GtOC have been released within the period 1990-2010.
-Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than
-Tier 1 methods, which may origin from \ldots{} . We also find that SOC
-is very sensity to management decision such as residue recycling
-indicating the nessessity to incorporated better management data in soil
-models.
+We show that, our results on global scale based on Tier 2 IPCC
+methodolgy are in good agreement with Tier 1 default assumptions. We
+also find that SOC is very sensitive to management decision such as
+residue recycling indicating the nessessity to incorporated better
+management data in soil models.
 \end{abstract}
@@
 Following the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 (\citet{ipcc_2019_2019}), we calculate annual land use type
-specific soil organic carbon stocks for cropland, pastures and natural
-vegetation on half-degree resolution for the period of 1965 to 2010
-based on the following three steps: (1) Calculating the land use
-(sub-)type specific steady-states and decay rates for SOC stocks given
-the current biophysical, climatic and agronomic conditions, (2)
-accounting for land conversation effects by transferring SOC between
-land use types and (3) updating SOC stocks based on the previous stock,
-the steady-state and the decay rate.
+vol.~4 (\citet{ipcc_2019_2019}), we estimate global land-use type
+specific soil organic carbon (SOC) stocks for cropland and natural
+vegetation on half-degree resolution from 1965 to 2010. We assume the
+actual SOC state converges towards a stable steady state, that itself is
+changing over time and space depending on biophysical, climatic and
+agronomic conditions. Therefor we conduct the following three steps: (1)
+Calculating annual land-use (sub-)type specific steady states and decay
+rates for SOC stocks, (2) accounting for land conversion by transferring
+SOC between land-use types and (3) updating SOC stocks based on the
+previous stock, the steady state and the decay rate.
 
@@
 assume three soil carbon sub-pools (active, slow and passive) and
-entangled dynamics between them. Carbon inflow to each sub-pool (see
-@ref(sec:carboninputs)) and decay rates (see @ref(sec:tier2)) of each
-sub-pool are still the key components to determining steady-state SOC
-stocks.
+entangled dynamics between them. Annual carbon inflow to each sub-pool
+(see @ref(sec:carboninputs)) and annual decay rates (see
+@ref(sec:tier2)) of each sub-pool are still the key components to
+determining steady-state SOC stocks.
 
 \hypertarget{sec:carboninputs}{%
-\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}}
+\paragraph{Carbon Inputs to the Soil}\label{sec:carboninputs}}
 
-We account for different carbon input sources depending on the land use
+We account for different carbon input sources depending on the land-use
 type (see table @ref(tab:datasourceinputs)). Following the IPCC
@@
  \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land use types}
+ \caption{Type and data sources for carbon inputs to different land-use types}
  \begin{tabular}{l l l l}
  \tophline
-  Land use types   & source of carbon inputs & data source & nitrogen and lignin content \\
+  land-use types   & source of carbon inputs & data source & nitrogen and lignin content \\
  \middlehline
@@
                             & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
-                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
-                            \hline
- \multirow{2}{*}{Pasture}  & annual litterfall & LPJmL4 [3] & default values given by [2] \\ 
                             & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
@@
 \hypertarget{sec:tier2}{%
-\subsubsection{Soil Organic Carbon decay (300)}\label{sec:tier2}}
+\paragraph{Soil Organic Carbon decay (300)}\label{sec:tier2}}
 
@@
 biophysical and biochemical soil properties as well as management
-factors that vary over time (t) and space (i). Following the
+factors that all vary over time (t) and space (i). Following the
 steady-state method of the refinement of the IPCC guidelines vol.~4
@@
 @ref(\#sec:tillage)) and irrigation (see @ref(\#sec:irrigation))
-practices on decay rates, whereas on pastures and natural vegetation, we
-assume rainfed and non-tilled conditions. Data sources as well as
-considered effects for each land use types are shown in table
+practices on decay rates, whereas on natural vegetation, we assume
+rainfed and non-tilled conditions. Data sources as well as considered
+effects for each land-use types are shown in table
 @ref(tab:datasourcedecay). To account for variations of decay rates
@@
  \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land use types}
+ \caption{Type and data sources for carbon inputs to different land-use types}
  \begin{tabular}{l l l l}
  \tophline
-  Land use types   & type of decay driver & parameter use to represent driver & data source \\
+  land-use types   & type of decay driver & parameter use to represent driver & data source \\
  \middlehline
@@
 
-\subsubsection{SOC transfer between land use types}
+\subsubsection{SOC transfer between land-use types}
 
-We calculate SOC stocks based on the area shares of land use types (lut)
+We calculate SOC stocks based on the area shares of land-use types (lut)
 within the half-degree grid cells (i). If land is converted from one
-land use type into others (!lut), the respective share of the SOC stocks
+land-use type into others (!lut), the respective share of the SOC stocks
 is reallocated. We account for land conversion at the beginning of each
@@
 with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
-expansion for a given land use type \(lut\). Note that \(!lut\) denotes
-the sum over all other land use types, which decreases in the specific
-time step \(t\). Data sources and methodology on land use states and
+expansion for a given land-use type \(lut\). Note that \(!lut\) denotes
+the sum over all other land-use types, which decreases in the specific
+time step \(t\). Data sources and methodology on land-use states and
 changes are described in @ref(sec:landuse).
@@
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ - land use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ - total SOC stock within cell}}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ - land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ - total SOC stock within cell}}
 \label{eq:totalstock}
@@
 
-To initialize all SOC sub-pools we assume that
+To initialize all SOC sub-pools we assume that cropped land natural
+vegetation * steady-states or * spin up
 
@@
 
+Agricultural management data is based on R library package moinput
+(\citep{bodirsky_in_prep}), which compiles in its core country-specific
+FAO production and cropland statistics (\citep{FAOSTAT}) to a
+compherensive and constistent data suite. The data is prepared in 5 year
+time steps from 1965 to 2010, which also restricts our analysis to this
+time span. For all the following data, if not declared differently, we
+interpolate values linearly between the time steps and hold it constant
+before the first time step for a short spin-up phase from 1961 to 1965.
+
 \hypertarget{sec:landuse}{%
@@
 
-Land use patterns are based on the Land-Use Harmonization 2 (LUH2,
+Land-use patterns are based on the Land-Use Harmonization 2 (LUH2,
 \citep{LUH2}) data set, which we aggregate from quarter degree to half
@@
 (see @ref(append:Table\_luh2fao2mag) for more details on the crop type
-mapping). Land use transitions are calculated as net area differences of
-the land use data on half-degree.
+mapping). land-use transitions are calculated as net area differences of
+the land-use data on half-degree.
 
 \subsubsection{Crop, Crop Residues and Pasture Production (300)}
+
+\paragraph{Crop Production}
 
@@
 well as for fallows.
+
+\paragraph{Crop Residue Production}
 
@@
 
-A fixed share of the ag residues is assumed to be burned on field, which
-depends on the per-capita income of the country. Following
+\subparagraph{Burned Residues}
+
+A fixed share of the ag residues is assumed to be burned on field
+depending on the per-capita income of the country. Following
 \citep{smil1999}) we assume 25\% burn share for low-income countries
@@
 shares for all middle-income countries depending on their per-capita
-income.
+income. Depending on the crop type 80--90\% of the residue carbon burned
+on the fields are lost within the combustion process
+(\citep{ipcc_2006_2006}).
 
-Residue demand for feed is based on country and livestock specific feed
-basekts (see \citep{weindl}) taking available ag residual biomass as
-well as livestock productivity into account. The three residue groups
-(straw, high-lignin and low-lignin residues) of the feed baskets are
-disaggregated to the crop groups, using\ldots. (see
-@ref(append:Table\_kcr2kres)). \textless-- its a disggregation, no
---\textgreater{}
+\subparagraph{Residue Usage}
 
-We estimated a material use share for the straw residues group of 5\%
-and a fuel demand share demand of 10\% \textless-- only for straw, or
-for all groups? --\textgreater{} in low income countries.
+We compile out of our 17 crop groups, three used residue groups (straw,
+high-lignin and low-lignin residues) with additional demand for other
+purposes and one residues with no double use (see
+@ref(append:Table\_kcr2kres)).
+
+\subparagraph{Feed Demand}
+
+Residue feed demand for five different livestock groups is based on
+country- and residue-group-specific feed basekts (see \citep{weindl})
+taking available ag residual biomass as well as livestock productivity
+into account.
+
+\subparagraph{Material and Fuel demand}
+
+We estimate a material-use share for the straw residues group of 5\% and
+a fuel-share of 10\% for all used residues groups in low income
+countries according to worldbank definitions
+(\(<\,1000\,\tfrac{USD}{yr}\)). For high-income
+(\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is
+assumend, leaving middle-income countries with linearly interpolate
+shares depending on their per-capita income.
+
+\subparagraph{Recycled Residues}
 
 The remaining ag residues as well as all bg residues are assumend to be
-recycled to the soil.
+recycled to the soil. We cut high recycling shares per hectar at the
+95\%-percentile to corrected for outliers.
+
+\paragraph{Pasture Production}
 
@@
 
+\paragraph{Dry Matter to Carbon Transformation}
+
 To transform dry matter estimates into carbon, we compiled crop group
@@
 
+\paragraph{Livestock Distribution}
+
 To disaggregate country level FAOSTAT livestock production values to
 half-degree pattern, we use the following rule based assumptions which
-were inspired by the approach of \citep{gilbert}.
+were inspired by the approach of \citep{gilbert} and uses feed basket
+assumptions based on a revised methodology of \citet{weindl}.
 
-For poultry, egg and monogastric meat production we divide into
-intensive and extensive production systems based on a per-capita income
-of the country. For low-income countries according to worldbank
-definitions (\textless1000 USD/yr), we assume extensive production
-systems. We located them according to (built-up areas shares\textbar{}
-population shares) based on the idea that these animals are held in
-households, subsistence or small-holder farming systems.
+\subparagraph{Livestock Managament Intensities}
 
-\begin{itemize}
-\item
-  Intensive production is distributed within a country using the crop
-  production (excluding 2nd generation bioenergy crops) share, assuming
-  that feed availability is the most driving factor for livestock
-  location.
-\item
-  For dairy and ruminated meat production we
-\item
-  splot into pasture fed and non-pasture fed fractions based on the
-  aggregated feed mix within a country (explain rules)
-\end{itemize}
+To account for different livestock management intensities, we
+differentiate between extensive and an intensive systems assuming that
+they have different requirements on production factors such as feed
+availabiliy, access to markets and infrastructure (e.g.~roads, water).
 
-\begin{itemize}
-\item
-  Pasture fed animals are distributed to half-degree cells according to
-  the pasture production shares of that cell within a country (not
-  methodologically clear description).
-\item
-  Non-pasture fed animals are distributed again using crop production
-  (excluding 2nd generation bioenergy crops) shares (not clear, either).
-\end{itemize}
+\subparagraph{Dairy and ruminant Meat Production}
 
-We calculate excretions by first estimating the nitrogen balance of the
-livestock system (\citep{bodirsky1012}) on the basis of comprehensive
-livestock feed baskets ({[}weindl{]}), assuming that all nitrogen in
-protein feed intake, minus the nitrogen in the slaugher mass, is
+For extensive dairy and ruminant meat production we estimate, that
+livestock is located rather immovable on pastures and rangeland. We use
+the country-level pasture feed stuff shares within the feed baskets, to
+split up pasture-fed, extensive from the rather intensive, crop-fed
+livestock. Extensive production of dairy and ruminant meat is allocated
+weighted by the pasture production values to all half-degree gridcell of
+a country. On the other hand intensive dairy and ruminant meat
+production is assumend to be located proportinal to crop production to
+have short transport distances for feed stuff.
+
+\subparagraph{Poultry, Egg and monogastric Meat Production}
+
+For poultry, egg and monogastric meat production we use the per-capita
+income of the country to divide into intensive and extensive production
+systems. For low-income countries according to worldbank definitions
+(\textless1000 USD/yr), we assume extensive production systems. We
+located them according to built-up areas shares based on the idea that
+these animals are held in households, subsistence or small-holder
+farming systems with a high labour per animal ratio. Intensive
+production is distributed within a country using the crop production
+share, assuming that feed availability is the most driving factor for
+livestock location.
+
+\paragraph{Manure Excretion}
+
+Manure production and management is based on a revised methodology of
+(\citep{bodirsky2012}) and will be explained in key aspects again due to
+its central role for soil carbon modelling. Based on the gridded
+livestock distribution we calculate excretions by estimating the
+nitrogen balance of the livestock system on the basis of comprehensive
+livestock feed baskets (\citep{weindl}), assuming that all nitrogen in
+protein feed intake, minus the nitrogen in the slaughter mass, is
 excreted. Carbon in excreted manure is estimated by applying fixed C:N
-ratios (given by \citep[(][]{ipcc_2019_2019}). What about animal waste
-management losses?
+ratios (given by \citep[(][]{ipcc_2019_2019}).
+
+\paragraph{Manure Storage}
+
+Depending on the feed system we assume manure to be handled in four
+different ways: All manure orginated from pasture feed intake is
+excreted directly to pastures and rangelands (pasture grazing) deducting
+manure collected as fuel. Manure fuel shares are estimated using IPCC
+default values (\citet{ippc_2006_2006}). Whereas for low-income
+countries according to worldbank definitions (\textless1000 USD/yr), we
+adopt a share of 25\% of crop residues in feed intake directly consumend
+and excreted on crop fields (stubble grazing), we do not consider any
+stubble grazing in high-income countries (\(>\,10000\,\tfrac{USD}{yr}\),
+leaving middle-income countries with linearly interpolate shares
+depending on their per-capita income. For all other feed items we assume
+the manure to be stored in animal waste management systems associated to
+animal houses.
+
+\paragraph{Manure Recycling}
+
+To estimate the carbon actually recycled to the soil, we account for
+carbon losses during storage and recycling shares in different animal
+waste management and grazing systems. Whereas we assume no losses for
+pasture and stubble grazing, we consider all carbon lost for manure
+collected as fuel. For manure stored in different animal waste
+management system we compiled carbon loss rates partly depending on
+nitrogen loss rates as specified in \citep{bodirsky2012} (see
+@ref(append:Table\_clossAWMS))).
 

### 2020-05-03 SOCbudgetPaper/methods.Rmd
methods update, add more result rmds

@@
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019), we calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.
+4 (@ipcc_2019_2019), we estimate global land-use type specific soil organic carbon (SOC) stocks for cropland and natural vegetation on half-degree resolution from 1965 to 2010.
+We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
+Therefor we conduct the following three steps:
+(1) Calculating annual land-use (sub-)type specific steady states and decay rates for SOC stocks, 
+(2) accounting for land conversion by transferring SOC between land-use types and 
+(3) updating SOC stocks based on the previous stock, the steady state and the decay rate.
 
@@
 4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
-Carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and decay rates (see \@ref(sec:tier2)) of each sub-pool are still the key components to determining steady-state SOC stocks.
+Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are still the key components to determining steady-state SOC stocks.
 
-### Carbon Inputs to the Soil {#sec:carboninputs}
+#### Carbon Inputs to the Soil {#sec:carboninputs}
 
-We account for different carbon input sources depending on the land use type (see table \@ref(tab:datasourceinputs)).
+We account for different carbon input sources depending on the land-use type (see table \@ref(tab:datasourceinputs)).
 Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
@@
 
-### Soil Organic Carbon decay (300) {#sec:tier2}
+#### Soil Organic Carbon decay (300) {#sec:tier2}
 
-The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i).
+The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
 Following the steady-state method of the refinement of the IPCC guidelines vol.
@@
 
-For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as considered effects for each land use types are shown in table \@ref(tab:datasourcedecay).
+For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas on natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as considered effects for each land-use types are shown in table \@ref(tab:datasourcedecay).
 To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
@@
 
-### SOC transfer between land use types
+### SOC transfer between land-use types
 
-We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i).
-If land is converted from one land use type into others (!lut), the respective share of the SOC stocks is reallocated.
+We calculate SOC stocks based on the area shares of land-use types (lut) within the half-degree grid cells (i).
+If land is converted from one land-use type into others (!lut), the respective share of the SOC stocks is reallocated.
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via
@@
 
-with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land use type $lut$.
-Note that $!lut$ denotes the sum over all other land use types, which decreases in the specific time step $t$.
-Data sources and methodology on land use states and changes are described in \@ref(sec:landuse).
+with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land-use type $lut$.
+Note that $!lut$ denotes the sum over all other land-use types, which decreases in the specific time step $t$.
+Data sources and methodology on land-use states and changes are described in \@ref(sec:landuse).
 
@@
 
-To initialize all SOC sub-pools we assume that
+To initialize all SOC sub-pools we assume that cropped land 
+natural vegetation 
+* steady-states or 
+* spin up
 
@@
 
+Agricultural management data is based on R library package moinput ([@bodirsky_in_prep]), which compiles in its core country-specific FAO production and cropland statistics ([@FAOSTAT]) to a compherensive and constistent data suite.
+The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis to this time span.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold it constant before the first time step for a short spin-up phase from 1961 to 1965.
+
 ### Landuse and Landuse Change (150) {#sec:landuse}
 
-Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
+Land-use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
 We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop type mapping).
-Land use transitions are calculated as net area differences of the land use data on half-degree.
+land-use transitions are calculated as net area differences of the land-use data on half-degree.
 
@@
 
+#### Crop Production
 Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop group specific half-degree production patterns.
@@
 
+#### Crop Residue Production
 Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
-Starting from crop production estimates of the harvested organs and their respective crop area, we estimate above-ground (ag) and below-ground (bg) residual biomass using yield-dependent harvest indices and shoot:root ratios.
+Starting from crop production estimates of the harvested organs and their respective crop area, we estimate above-ground (ag) and below-ground (bg) residual biomass using yield-dependent <!-- do not know what was meant with this anymore (why "yield-dependent"?).
+I guess that has to be explained again than -.- --> harvest indices and shoot:root ratios.
 We assume that all bg residues are recycled to the soil, whereas ag residues can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
 
-A fixed share of the ag residues is assumed to be burned on field, which depends on the per-capita income of the country.
+##### Burned Residues
+A fixed share of the ag residues is assumed to be burned on field depending on the per-capita income of the country.
 Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop type 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).
 
-Residue demand for feed is based on country and livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock productivity into account.
-The three residue groups (straw, high-lignin and low-lignin residues) of the feed baskets are disaggregated to the crop groups, using....
-(see \@ref(append:Table_kcr2kres)).
-<-- its a disggregation, no -->
+##### Residue Usage
+We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see \@ref(append:Table_kcr2kres)).
 
-We estimated a material use share for the straw residues group of 5% and a fuel demand share demand of 10% <-- only for straw, or for all groups? --> in low income countries.
+##### Feed Demand
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock productivity into account.
+<!-- more details needed here? -->
 
+##### Material and Fuel demand
+We estimate a material-use share for the straw residues group of 5% and a fuel-share of 10% for all used residues groups in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
+For high-income  ($>\,10000\,\tfrac{USD}{yr}$ no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+
+##### Recycled Residues
 The remaining ag residues as well as all bg residues are assumend to be recycled to the soil. 
+We cut high recycling shares per hectar at the 95%-percentile to corrected for outliers.
 
-<!-- Add formular or drawings here? --> 
-<!-- if you want, but first finish the rest -->
-
+#### Pasture Production
 Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production.
@@
 
+#### Dry Matter to Carbon Transformation
 To transform dry matter estimates into carbon, we compiled crop group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
@@
 
-To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert].
+#### Livestock Distribution
+To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert] and uses feed basket assumptions based on a revised methodology of @weindl.
 
-For poultry, egg and monogastric meat production we divide into intensive and extensive production systems based on a per-capita income of the country.
+##### Livestock Managament Intensities
+To account for different livestock management intensities, we differentiate between extensive and an intensive systems assuming that they have different requirements on production factors such as feed availabiliy, access to markets and infrastructure (e.g.
+roads, water).
+
+##### Dairy and ruminant Meat Production
+For extensive dairy and ruminant meat production we estimate, that livestock is located rather immovable on pastures and rangeland.
+We use the country-level pasture feed stuff shares within the feed baskets, to split up pasture-fed, extensive from the rather intensive, crop-fed livestock.
+Extensive production of dairy and ruminant meat is allocated weighted by the pasture production values to all half-degree gridcell of a country.
+On the other hand intensive dairy and ruminant meat production is assumend to be located proportinal to crop production to have short transport distances for feed stuff.
+<!-- yeah, actually was that the full idea here?
+-->
+
+##### Poultry, Egg and monogastric Meat Production
+For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
 For low-income countries according to worldbank definitions (<1000 USD/yr), we assume extensive production systems.
-We located them according to (built-up areas shares| population shares) based on the idea that these animals are held in households, subsistence or small-holder farming systems.
+We located them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
+Intensive production is distributed within a country using the crop production share, assuming that feed availability is the most driving factor for livestock location.
 
-* Intensive production is distributed within a country using the crop production (excluding 2nd generation bioenergy crops) share, assuming that feed availability is the most driving factor for livestock location.
+#### Manure Excretion
+Manure production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets ([@weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]).
 
-* For dairy and ruminated meat production we
+#### Manure Storage
+Depending on the feed system we assume manure to be handled in four different ways:
+All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing) deducting manure collected as fuel.
+Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
+Whereas for low-income countries according to worldbank definitions (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses.
 
-* splot into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
-
-<!-- how to spell out that crop is everything else then pasture? -->
-
-* Pasture fed animals are distributed to half-degree cells according to the pasture production shares of that cell within a country (not methodologically clear description).
-
-* Non-pasture fed animals are distributed again using crop production (excluding 2nd generation bioenergy crops) shares (not clear, either).
-
-<!-- Add formular, drawings or visualisation here? --> <!-- No formulas needed, but precise wording -->
-
-<!-- TILL HERE -->
-
-We calculate excretions by first estimating the nitrogen balance of the livestock system ([@bodirsky1012]) on the basis of comprehensive livestock feed baskets ([weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaugher mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]).
-What about animal waste management losses?
+#### Manure Recycling
+To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
+Whereas we assume no losses for pasture and stubble grazing, we consider all carbon lost for manure collected as fuel.
+For manure stored in different animal waste management system we compiled carbon loss rates partly depending on nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
 

### 2020-05-03 SOCbudgetPaper/results.Rmd
methods update, add more result rmds

@@
-# Content section with citations
+# Results
 
-See the [R Markdown docs for bibliographies and citations](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html).
-
-Copernicus supports biblatex and a sample bibliography is in file `sample.bib`.
-Read [], and [see ].
-
-# Content section with R code chunks
-
-```{r, echo = FALSE}
-sum <- 1 + 41
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "two column figure", fig.show='hold'}
+knitr::include_graphics("images/figs_draft-1.png")
 ```
 
-You should always use `echo = FALSE` on R Markdown code blocks as they add formatting and styling not desired by Copernicus.
-The hidden workflow results in `r sum`.
-
-You can add verbatim code snippets without extra styles by using ` ``` ` without additional instructions.
-
-```
-sum <- 1 + 41
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "two column figure", fig.show='hold'}
+knitr::include_graphics("images/figs_draft-2.png")
 ```
 
-# Content section with list
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "two column figure", fig.show='hold'}
+knitr::include_graphics("images/figs_draft-3.png")
+```
 
-If you want to insert a list, you must
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "two column figure", fig.show='hold'}
+knitr::include_graphics("images/RegionPlot+Valid.png")
+```
 
-- leave
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "two column figure", fig.show='hold'}
+knitr::include_graphics(c("images/maps/CShare_1965.png","images/maps/CShare_2010.png"))
+```
 
-- empty lines
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "two column figure", fig.show='hold'}
+knitr::include_graphics(c("images/maps/CIncrease_1965.png","images/maps/CIncrease_2010.png"))
+```
 
-- between each list item
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "one column figure"}
+knitr::include_graphics("images/maps/CShare_4ColorClimates.png")
+```

### 2020-05-03 FlowFigure/manipulate_sankey.R

@@ -1,0 +1,256 @@
+
+nitrogen_sankey<-function(x,svg_in,svg_out="sankey.svg",scaling=1){
+  require("xml2")
+  svg<-read_xml(svg_in)
+  svg2<-xml_find_all(svg, "//svg:g")
+  paths<-xml_find_all(svg2, ".//svg:path")
+  texts<-xml_find_all(svg2, ".//svg:text")
+  
+  mapping<-toolMappingFile(type = "sectoral",name = "svg_nitrogen.csv",readcsv = TRUE)
+  
+
+  
+  settext<-function(id,value){
+    text=texts[which(xml_attr(texts,attr="id")==id)]
+    if(length(text)!=1){warning(paste0("text id ",id," does not exist or exists multiple"))}
+    xml_text(text)<-value
+  }
+  
+  setstyle<-function(id,value,unit="px",style="stroke-width"){
+    path=paths[which(xml_attr(paths,attr="id")==id)]
+    if(length(path)!=1){stop(paste0("id ",id," does not exist or exists multiple"))}
+    text = xml_attr(path,attr="style")
+    text=strsplit(text,split = ";")[[1]]
+    pos<-which(substring(text,first = 0,last=nchar(style))==style)
+    text[pos]<-paste0(style,":",value,unit)
+    text<-paste(text,collapse=";")
+    xml_attr(path,"style")<-text
+  }
+  
+  numformat <- function(val) {
+    if((val<0.95)&(val>0.05)) {
+      return(sub("^(-?)0.", "\\1.", sprintf("%.1f", val)) )
+    } else if(val<=0.05) {
+      return(" ")
+    } else{
+      return(sprintf("%.0f", val) )  
+    }
+  }
+  
+  
+  
+  ## add year and region to header
+  
+  region<-getRegions(x)
+  if (length(region)==249){
+    region="World"
+    settext("region",paste0("of the world"))
+  } else if (length(region)==1){
+    mapping2<-toolMappingFile(type = "regional",name = "regionmappingMAgPIE.csv",readcsv = TRUE)
+    region<-mapping2[which(mapping2[,2]==region),1]
+    settext("region",paste0("of ",region))
+  } else {region <- "aggregate world region"}
+  
+
[…]

### 2020-05-03 FlowFigure/toolFlowFigure.R

@@ -1,0 +1,106 @@
+#' toolFlowFigure
+#'
+#' Create a flow figure out of data, mapping and proxy pattern
+#'
+#' @param data
+#' @param proxy
+#' @param mapping
+#' @param scaling
+#'
+#' @return file path to svg
+#' @author Benjamin Leon Bodirsky, Kristine Karstens
+#'
+#' @importFrom xml2 read_xml xml_find_all
+#' @export
+
+toolFlowFigure <- function(data, proxy, mapping, scaling=1){
+
+  proxy <- "/home/kristine/WORK/Paper/historicalsocbudget/FlowFigure/sankey_nitrogen_CHN_2010.svg"
+
+  svg      <- read_xml(proxy)
+  groups   <- xml_find_all(svg, "//svg:g")
+  paths    <- xml_find_all(groups, ".//svg:path")
+  texts    <- xml_find_all(groups, ".//svg:text")
+
+  mapping <- toolGetMapping(where = "socbudet", name = mapping)
+
+
+
+  settext<-function(id,value){
+    text=texts[which(xml_attr(texts,attr="id")==id)]
+    if(length(text)!=1){warning(paste0("text id ",id," does not exist or exists multiple"))}
+    xml_text(text)<-value
+  }
+
+  setstyle<-function(id,value,unit="px",style="stroke-width"){
+    path=paths[which(xml_attr(paths,attr="id")==id)]
+    if(length(path)!=1){stop(paste0("id ",id," does not exist or exists multiple"))}
+    text = xml_attr(path,attr="style")
+    text=strsplit(text,split = ";")[[1]]
+    pos<-which(substring(text,first = 0,last=nchar(style))==style)
+    text[pos]<-paste0(style,":",value,unit)
+    text<-paste(text,collapse=";")
+    xml_attr(path,"style")<-text
+  }
+
+  numformat <- function(val) {
+    if((val<0.95)&(val>0.05)) {
+      return(sub("^(-?)0.", "\\1.", sprintf("%.1f", val)) )
+    } else if(val<=0.05) {
+      return(" ")
+    } else{
+      return(sprintf("%.0f", val) )
+    }
+  }
+
+
+
+  ## add year and region to header
+
+  region<-getRegions(x)
+  if (length(region)==249){
+    region="World"
+    settext("region",paste0("of the world"))
+  } else if (length(region)==1){
+    mapping2<-toolMappingFile(type = "regional",name = "regionmappingMAgPIE.csv",readcsv = TRUE)
[…]

### 2020-05-19 SOCbudgetPaper/methods.Rmd
Update methods.Rmd

@@
 We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
-Therefor we conduct the following three steps:
-(1) Calculating annual land-use (sub-)type specific steady states and decay rates for SOC stocks, 
-(2) accounting for land conversion by transferring SOC between land-use types and 
-(3) updating SOC stocks based on the previous stock, the steady state and the decay rate.
+Therefore we conduct the following three steps within each yearly timestep:
+(1) We calculate annual land-use (sub-)type specific steady states and decay rates for SOC stocks, 
+(2) We account for land conversion by transferring SOC between land-use types and 
+(3) We estimate SOC stocks based on the stocks of the previous time period, the steady state stocks and the decay rate.
 
@@
 SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall lut, sub, i.
+\label{eq:steadystate}
+\end{equation}
+
+This equation can also be reformulated to a massbalance equation as follows:
+
+\begin{equation}
+SOC_{t} = SOC_{t-1} (1 - k_{t} \cdot 1\unit{a}) + C^{\textrm{in}}  \qquad \forall lut, sub, i.
 \label{eq:steadystate}

### 2020-05-19 SOCbudgetPaper/methods.Rmd
Update methods.Rmd

@@
 \begin{equation}
-SOC^{eq} =\frac{C^{\textrm{in}}}{k}
+SOC^{eq}_t =\frac{C^{\textrm{in}}_t}{k_t}
 \label{eq:inoutflow}
@@
 
+BB: I would include the t, also to show that the steady-state is time-dependent.
+
 #### Carbon Inputs to the Soil {#sec:carboninputs}
@@
 We account for different carbon input sources depending on the land-use type (see table \@ref(tab:datasourceinputs)).
+Name here the inputs, and refer to the respective sections, dont just refer to the table.
 Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
@@
  \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
+ \caption{Type and data sources for carbon inputs to different land-use types }
  \begin{tabular}{l l l l}
  \tophline
-  land-use types   & source of carbon inputs & data source & nitrogen and lignin content \\
+  land-use types BB: table header should be bold   & source of carbon inputs & data source & nitrogen and lignin content \\
  \middlehline
@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} (1 - k_{t} \cdot 1\unit{a}) + C^{\textrm{in}}  \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} - k_{t} \cdot 1\unit{a} + C^{\textrm{in}}  \qquad \forall lut, sub, i.
 \label{eq:steadystate}
@@
 
-Agricultural management data is based on R library package moinput ([@bodirsky_in_prep]), which compiles in its core country-specific FAO production and cropland statistics ([@FAOSTAT]) to a compherensive and constistent data suite.
+Agricultural management data is estimated using the R library moinput ([cite git link]), which compiles country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
+BB: Isnt the whole calculation based on R libraries?
 The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis to this time span.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold it constant before the first time step for a short spin-up phase from 1961 to 1965.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold it constant before the first time step from 1961 to 1965.
+BB: too short for a spin-up, so dont use this word.
+Maybe also just ignore this detail.
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop type mapping).
-land-use transitions are calculated as net area differences of the land-use data on half-degree.
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each gridcell based on the country and year specific area harvested shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop type mapping).
+Land-use transitions are calculated as net area differences of the land-use data on half-degree.
 
@@
 We assume that all bg residues are recycled to the soil, whereas ag residues can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
+    BB: One new thing is here that we use a doublecropping factor to account for lower harvest indices of multicropping vs a very high single-cropped yield
 
@@
 
-##### Feed Demand
 Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock productivity into account.
-<!-- more details needed here? -->
 
-##### Material and Fuel demand
 We estimate a material-use share for the straw residues group of 5% and a fuel-share of 10% for all used residues groups in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
@@
 
-##### Recycled Residues
 The remaining ag residues as well as all bg residues are assumend to be recycled to the soil. 
@@
 Following the same approach as for crop production we disaggregate and calibrate half-degree pasture production pattern from grass yields from LPJmL and pasture area and rangeland patterns ( (see \@ref(#sec:landuse))) to derive half-degree pasture production patterns.
+BB: Oh, does that mean we assume that only the pasture grows that is being eaten?
+Thats not too much, is it?
+Maybe ask Isabelle for a better solution (e.g.
+direct pasture yields from LPJmL).
+And then the "pasture production" is just the part being removed by grazing.
+Does the IPCC specifiy something in respect to pastures?
 
 #### Dry Matter to Carbon Transformation
-To transform dry matter estimates into carbon, we compiled crop group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
+To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
 
@@
 To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert] and uses feed basket assumptions based on a revised methodology of @weindl.
-
-##### Livestock Managament Intensities
-To account for different livestock management intensities, we differentiate between extensive and an intensive systems assuming that they have different requirements on production factors such as feed availabiliy, access to markets and infrastructure (e.g.
-roads, water).
-
-##### Dairy and ruminant Meat Production
-For extensive dairy and ruminant meat production we estimate, that livestock is located rather immovable on pastures and rangeland.
-We use the country-level pasture feed stuff shares within the feed baskets, to split up pasture-fed, extensive from the rather intensive, crop-fed livestock.
-Extensive production of dairy and ruminant meat is allocated weighted by the pasture production values to all half-degree gridcell of a country.
-On the other hand intensive dairy and ruminant meat production is assumend to be located proportinal to crop production to have short transport distances for feed stuff.
+We differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
+For ruminants, we use the country-level shares of gras within the feed baskets to split up pasture-fed, extensive systems from the rather intensive crop-fed livestock.
+For extensive dairy and ruminant meat production we estimate, that livestock is located in proximity to pastures and rangelands and distribute manure excretion proportional to pasture production of all half-degree gridcell of a country.
+Intensive dairy and ruminant meat production is assumend to be located proportinal to crop production to have short transport distances for feed stuff.
 <!-- yeah, actually was that the full idea here?
@@
 
-##### Poultry, Egg and monogastric Meat Production
+BB: What does not become clear to me: Do the extensive cows only eat pasture, and the intensive one only crops?
+Does that in the end mean we just redistirbute grazed pasture to pastures and crop-based feed to croplands proportional to production?
+Or is there more to it?
+
 For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
 For low-income countries according to worldbank definitions (<1000 USD/yr), we assume extensive production systems.
-We located them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
+We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
 Intensive production is distributed within a country using the crop production share, assuming that feed availability is the most driving factor for livestock location.
 
-#### Manure Excretion
+#### Manure Excretion, Storage and Recycling
 Manure production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
@@
 Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]).
-
-#### Manure Storage
 Depending on the feed system we assume manure to be handled in four different ways:
-All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing) deducting manure collected as fuel.
+All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
 Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
@@
 For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses. 
-
-#### Manure Recycling
 To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider all carbon lost for manure collected as fuel.
-For manure stored in different animal waste management system we compiled carbon loss rates partly depending on nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
+Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
+For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
 
 ### Irrigation (100)
-Simple growing period calculations together with irrigation shares of LUH2v2 are used (BB: you dont like d's, on't you) to estimate water effects on decay rates.
+BB: Where does the irrigated area come from?
+Simple growing period calculations together with irrigation shares of LUH2v2 are used to estimate water effects on decay rates.
 

### 2020-06-16 ResultNotebooks/results_SOCcompare.Rmd
Update result scripts to new structure

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrSOCbudget)
library(magpiesets)

rev <- "rev0.6_mrSOC"

setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")

SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
Landuse    <- read.magpie("Data/Landuse.rds")

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

SOCStock     <- dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=3)
glo_SOCStock <- dimSums(SOCStock, dim=1) 
cell2iso     <- toolGetMapping("CountryToCellMapping.csv", type = "cell")
iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
iso2reg      <- toolMappingFile("regional",getConfig("regionmapping"),readcsv = TRUE)
reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")

print(mbind(dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=c(1,3)),
            dimSums(dimSums(SoilCarbon[,,"steadystate"], dim=3.3)* Landuse, dim=c(1,3))))

```

Load validation data:

```{r Validation Stocks}

  LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
  LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
  LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
  WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
  GSOC       <- calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate="REG+GLO", try=TRUE)
  SoilGrids  <- calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate="REG+GLO", try=TRUE)

```

Plot global stocks

```{r plot glo stocks}

plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red", xlim=c(1951,2017),  ylim=c(400000,1300000))
lines(getYears(LPJ_IPCC, as.integer = TRUE), LPJ_IPCC["GLO",,1], type="l", col="green")
lines(getYears(LPJ_car, as.integer = TRUE), LPJ_car["GLO",,], type="l", col="darkgreen")
lines(getYears(LPJ_rev21, as.integer = TRUE), LPJ_rev21["GLO",,2], type="l", col="lightgreen")
lines(getYears(WISE, as.integer = TRUE), WISE["GLO",,], type="l", col="brown")
lines(getYears(SoilGrids, as.integer = TRUE), SoilGrids["GLO",,], type="l", col="blue")
lines(getYears(GSOC, as.integer = TRUE), GSOC["GLO",,], type="l", col="pink")
plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red")
```

Plot regional stocks

```{r plot reg stock, out.width="100%", out.height=1000}

library(mip)
library(quitte)
library(ggplot2)

getNames(reg_SOCStock) <- paste0("modeloutput.SOCbudget.", getNames(WISE,dim=3))
getSets(reg_SOCStock)  <- getSets(WISE)

hist <- rbind(as.quitte(WISE["GLO",,,invert=TRUE]),
              as.quitte(GSOC["GLO",,,invert=TRUE]),
              as.quitte(SoilGrids["GLO",,,invert=TRUE]),
              as.quitte(LPJ_rev21["GLO",,,invert=TRUE][,,2]))

mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y", ylab="SOC (0-30cm) in MtC")
```

```{r bar plot}

### FROM ISABELLE

# p.tc <- ggplot() 
# # agass Hauptszenarien --> Balken
# p.tc <-p.tc + geom_bar(data = dat1a, aes(x = Region, y = Value, fill=Mainscen,group=factor(Mainscen)),stat = "identity",width = 0.7,position=position_dodge(0.7),alpha=0.8, show_guide=FALSE) 
# p.tc <- p.tc + scale_fill_manual(values = c("#999999","#FF3300","#66CC00","#3366FF"),breaks=c("baseline","S1a","S3a","S3c"),
#                                  labels=c("Baseline     ","Climate_impacts     ","Shift_to_rangeland     ","Shift_to_mixed     "))
# # agass mit co2 --> graue K?stchen
# p.tc <- p.tc + geom_point(data=dat2a,aes(x = Region, y = Value, group=factor(Mainscen)),size=4,position=position_dodge(0.7),colour="black",shape=22,fill="grey")
# 
# # Spread GCMs ohne co2
# p.tc <- p.tc + geom_errorbar(data=dat4a,aes(x = Region, ymin = min,ymax=max, group=Mainscen),position=position_dodge(0.7),width=.2,colour="dark red", linetype="dashed",size=0.6,show_guide=FALSE)
# #p.tc <- p.tc + geom_point(data=dat3a,aes(x = Region, y = Value, group=factor(Mainscen)),size=1.5,position=position_dodge(0.7),colour="red", show_guide=FALSE)
# p.tc <- p.tc + geom_point(data=dat4a,aes(x = Region, y = min, group=factor(Mainscen)),size=3.8,position=position_dodge(0.7),colour="dark red",shape=21,fill="white", show_guide=FALSE)
# p.tc <- p.tc + geom_point(data=dat4a,aes(x = Region, y = max, group=factor(Mainscen)),size=2.5,position=position_dodge(0.7),colour="dark red",shape=19, show_guide=FALSE)
# 
# # Spread GCMs mit co2
# p.tc <- p.tc + geom_errorbar(data=dat4b,aes(x = Region, ymin = min,ymax=max, group=Mainscen),position=position_dodge(0.7),width=.2,colour="#003300",size=0.6, show_guide=FALSE)
# #p.tc <- p.tc + geom_point(data=dat3b,aes(x = Region, y = Value, group=factor(Mainscen)),size=1.5,position=position_dodge(0.7),colour="#003300", show_guide=FALSE)
# p.tc <- p.tc + geom_point(data=dat4b,aes(x = Region, y = min, group=factor(Mainscen)),size=3.5,position=position_dodge(0.7),colour="#003300",shape=23,fill="white", show_guide=FALSE)
# p.tc <- p.tc + geom_point(data=dat4b,aes(x = Region, y = max, group=factor(Mainscen)),size=3.8,position=position_dodge(0.7),colour="#003300",shape=18, show_guide=FALSE)
# 
# #Layout
# p.tc <- p.tc + labs(y="Average annual TC rates 2005-2045 [%] \n",x="")
# guides(colour = guide_legend(override.aes = list(shape = NA)))
# p.tc <- p.tc + geom_hline(yintercept=0)
# 
# #p.tc <- p.tc + ylim(0, 2.4)
# p.tc <- p.tc + theme_bw() +
#   theme(axis.text = element_text(size = 16,face="bold"),
#         axis.title = element_text(size = 18,face="bold"),
#                 
#         legend.background = element_rect(fill = "white",colour="black",size=0.7),
#         legend.text =       element_text(size = 14, face="bold"),
#         legend.title =      element_blank(),
#         legend.position = c(0.25, 1.01),
#         legend.key = element_rect(colour = "white",size=2),
#         legend.key.height = unit(0.7,"cm"),
#         legend.key.width = unit(1.8,"cm"),
#         legend.margin = unit(100,"lines"),
#         
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         
#         panel.border = element_rect(size = 0.8, colour = "black"),
#         plot.margin = unit(c(0.5,0.5,0.5,0.5),"cm")  # top, right, bottom, and left margins
#   )
# 
# #Legend 1
# p.tc <- p.tc + annotate("rect", xmin = 0.6, xmax = 4.5, ymin = 2.3, ymax = 3.3,fill = "white",colour="black",size=0.7)
# 
# p.tc <- p.tc + annotate("text", x = 0.8, y = 3.1, label = "IAASTD~climate~scenario~w/o~CO[2]~effect",parse = TRUE,size = 4.7, hjust=0)
# 
# p.tc <- p.tc + annotate("text", x = 1.8, y = 2.9, label = "climate_impacts",size = 4.7, hjust=0)
# p.tc <- p.tc + annotate("rect", xmin = 0.8, xmax = 1.6, ymin = 2.85, ymax = 2.95,fill = "#FF3300",size=0.7)
# 
# p.tc <- p.tc + annotate("text", x = 1.8, y = 2.7, label = "shift_to_rangeland",size = 4.7, hjust=0)
# p.tc <- p.tc + annotate("rect", xmin = 0.8, xmax = 1.6, ymin = 2.65, ymax = 2.75,fill = "#66CC00",size=0.7)
# 
# p.tc <- p.tc + annotate("text", x = 1.8, y = 2.5, label = "shift_to_mixed",size = 4.7, hjust=0)
# p.tc <- p.tc + annotate("rect", xmin = 0.8, xmax = 1.6, ymin = 2.45, ymax = 2.55,fill = "#3366FF",size=0.7)
# 
# 
# #Legend 2
# p.tc <- p.tc + annotate("rect", xmin = 7.6, xmax = 11.4,  ymin = 2.3, ymax = 3.3,fill = "white",colour="black",size=0.7)
# 
# p.tc <- p.tc + annotate("text", x = 9, y = 3.0, label = "full~CO[2]~effect~(IAASTD)",parse = TRUE, size = 4.7, hjust=0)
# p.tc <- p.tc + annotate("point", x = 8.4, y = 3.0, colour="black",shape=22,fill="grey",size=4.5)
# 
# p.tc <- p.tc + annotate("text", x = 9, y = 2.8, label = "full~CO[2]~effect~(all~GCMs)",parse = TRUE, size = 4.7, hjust=0)
# p.tc <- p.tc + annotate("segment", x = 8, xend=8.8, y = 2.8, yend=2.8,colour="#003300",size=0.7)
# p.tc <- p.tc + annotate("point", x = 8, y = 2.8, colour="#003300",shape=23,fill="white",size=3.8)
# p.tc <- p.tc + annotate("point", x = 8.8, y = 2.8, colour="#003300",shape=18,size=5)
# 
# p.tc <- p.tc + annotate("text", x = 9, y = 2.6, label = "w/o~CO[2]~effect~(all~GCMs)",parse = TRUE,size = 4.7, hjust=0)
# p.tc <- p.tc + annotate("segment", x = 8, xend=8.8, y = 2.6, yend=2.6,colour="dark red",linetype="dashed",size=0.7)
# p.tc <- p.tc + annotate("point", x = 8, y = 2.6, colour="dark red",shape=21,fill="white",size=4)
# p.tc <- p.tc + annotate("point", x = 8.8, y = 2.6, colour="dark red",shape=19,size=4)
# 
# 
# print(p.tc)

```

### 2020-06-16 ResultNotebooks/results_SOCflows.Rmd
Update result scripts to new structure

--- 2020-05-03 ResultNotebooks/SOCflows_results.Rmd
@@
 ---
-  title: "First SOC budget analysis"
+  title: "First SOC budget flow analysis"
 output:
@@
 ```{r load}
-library(moinput)
+library(mrcommons)
 library(mrvalidation)
-library(socbudget)
+library(mrSOCbudget)
 library(magpiesets)
 
-#input       <- "/home/kristine/WORK/R libraries/rd3mod_inputdata/output/rev1_690d3718e151be1b450b394c1064b1c5_validation.tgz"
-#inputfolder <- gsub("[^/]*$","", input)
-#untar(input, inputfolder)
+rev <- "rev0.6_mrSOC"
 
-setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
+setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
 setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
+untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+
+SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
+Landuse    <- read.magpie("Data/Landuse.rds")
+
+```
+
+```{r load data}
+
+files <- list.files("Data")
+
+CarbonInput      <- readRDS("Data/CarbonInput.rds")
+CarbonLitter     <- readRDS("Data/CarbonLitter.rds")
+CarbonManure     <- readRDS("Data/CarbonManure.rds")           
+CarbonResidues   <- readRDS("Data/CarbonResidues.rds")
+CarbonHarvest    <- readRDS("Data/Crop_Harvest.rds")
+SoilCarbon       <- readRDS("Data/SoilCarbon.rds")
+
+Landuse          <- readRDS("Data/Landuse.rds")
+LanduseChange    <- readRDS("Data/LanduseChange.rds")
+
+FAOmassbalance   <- readRDS("Data/FAOmassbalance_ISO.rds" )
+ManureExcreted   <- readRDS("Data/Manure_excreted.rds")
+ManureRecycled   <- readRDS("Data/Manure_recycled.rds")
+ResidueBiomass   <- readRDS("Data/Residue_Biomass.rds")
+ResiduesAgDemand <- readRDS("Data/ResiduesAg_FieldBalance.rds")
 

### 2020-06-16 ResultNotebooks/results_SOCmaps.Rmd
Update result scripts to new structure

---
title: "Global SOC maps and emissions"
output:
  md_document:
    variant: markdown_github
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrSOCbudget)
library(magpiesets)

rev <- "rev0.6_mrSOC"

setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")

SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
Landuse    <- read.magpie("Data/Landuse.rds")

```

Calc global numbers

```{r Carbon Emissions (glo)}

SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
SOCStock        <- dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2)
glo_SOCStock    <- dimSums(SOCStock, dim=1) 
glo_SOCEmission <- glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"]  
plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission/1000, type="l", xlab="Time", ylab="Global SOC Debt in GtC")
plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCStock[,,"actualstate"])
lines(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCStock[,,"naturalstate"])
```

Plot cellular maps

```{r plot density maps}

cropland_dens    <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)
cropland_dens[cropland_dens==0] <- NA

cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/dimSums(SoilCarbon[,,"naturalstate"][,,"natveg"], dim=3)
cropland_share[Landuse[,,"crop"]==0] <- NA

cropland_increase <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)-dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3) 
cropland_increase[Landuse[,,"crop"]==0] <- NA

library(RColorBrewer)
cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(7)
cols_dens <- colorRampPalette(brewer.pal(11, "RdYlGn"))(30)[c(1:20, rep(c(21:30), each=10))]
outputdir <- "Output/Images"

png(paste0(outputdir,"CDens_2010.png"), width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_dens[,"y2010",], legend_range = c(0,300), legend_colours = cols_dens , main="Soil Organic Carbon Stocks under Cropland")
dev.off()

# png(paste0(outputdir,"CShare_1965.png"),width = 1280, height = 720, res = 108)
# luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
# dev.off()

png(paste0(outputdir,"CShare_2010.png"), width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
dev.off()

# png(paste0(outputdir,"CIncrease_1965.png"),width = 1280, height = 720, res = 108)
# luplot::plotmap(cropland_increase[,"y1965",], legend_range = c(-80,80), legend_colours = cols, main="Absoulte C change on Cropland 1965")
# dev.off()

png(paste0(outputdir,"CIncrease_2010.png"),width = 1280, height = 720, res = 108)
luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols, legend_title = "tC/ha", main="Absoulte C change on Cropland 2010")
dev.off()

```

```{r grid plot}
library(gridExtra)

par(mfrow = c(2, 2))

luplot::plotmap(cropland_dens[,"y2010",], legend_range = c(0,300), legend_colours = cols_dens , 
                main="SOC Stocks under Cropland")
plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission/1000, type="l", xlab="Time",
     ylab="Global SOC Debt in GtC")
luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols)
luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols,
                legend_title = "tC/ha")

```

```{r plot diff maps, out.width="100%", out.height=1000}

library(luplot)
library(magpiesets)

cellSOCdens   <- (dimSums(dimSums(SoilCarbon[,,"naturalstate"], dim=3.2) * Landuse, dim=c(3))/dimSums(Landuse, dim=c(3)))[,findset("past"),]
cellSOCdens[is.na(cellSOCdens)] <- 0

cellSoilGrids <- readSource("SoilGrids", subtype="cstock_0_30", convert="onlycorrect")
cellLPJml     <- calcOutput("LPJmL", version="LPJmL4", climatetype="CRU_4", subtype="soilc_layer", time="raw", selectyears="past", aggregate=FALSE)
cellLPJml     <- collapseNames(cellLPJml[,,"layer1"]+1/3*cellLPJml[,,"layer2"], collapsedim = c(2,3))

diffSoilGrids <- cellSOCdens - cellSoilGrids
diffLPJml     <- cellSOCdens - cellLPJml

luplot::plotmap(diffLPJml[,10,], legend_range = c(-80,80))
```

```{r plot density maps 2}

cols <- colorRampPalette(brewer.pal(9,"Oranges"))(1000)
cropland <- dimSums(SoilCarbon[,,"crop"], dim=3)
cropland[Landuse[,,"crop"]==0] <- NA

luplot::plotmap(cropland[,"y2010",], legend_colours = cols, legend_range=range(cropland[,"y2010",]/5, na.rm=TRUE))

natland <- dimSums(SoilCarbon[,,"natveg"], dim=3)

luplot::plotmap(natland[,"y2010",], legend_colours = cols, legend_range=range(cropland[,"y2010",]/5, na.rm=TRUE))

```

### 2020-06-16 ResultNotebooks/results_SOCtable.Rmd
Update result scripts to new structure

---
title: "Table on stock change factors per clinate zone"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrSOCbudget)
library(magpiesets)

rev <- "rev0.6_mrSOC"

setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")

SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
Landuse    <- read.magpie("Data/Landuse.rds")

```

```{r table scf climatezones}

library(knitr)
library(stats)
library(luplot)
library(reshape2)

koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
getNames(koeppen) <- tolower(getNames(koeppen))
koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
climateIPCC       <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
climateIPCC       <- climateIPCC[,getYears(SoilCarbon),]

rm(koeppen)

SOCDens           <- dimSums(SoilCarbon[,,c("steadystate","actualstate","naturalstate")], dim=3.3) 
SOCDens[,,"naturalstate.crop"] <- SOCDens[,,"actualstate.crop"]

CShare_ipcczone   <- dimSums(climateIPCC * SOCDens[,,"crop"]   * Landuse[,,"crop"], dim=1)/
                     dimSums(climateIPCC * SOCDens[,,"natveg"] * Landuse[,,"crop"], dim=1)

print(collapseNames(CShare_ipcczone[,findset("past"),]))

actualstate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"actualstate"])), 
                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")

naturalstate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"naturalstate"])), 
                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")

steadystate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"steadystate"])), 
                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")

saveRDS(list(naturalstate_table = naturalstate_table,
             actualstate_table  = actualstate_table, 
             steadystate_table  = steadystate_table), "Output/SFC_climatezone.rds")

#knitr::kable(naturalstate_table, "latex",digits=2, label="naturalstate_table")
#knitr::kable(actualstate_table, "markdown", digits=2)
```

```{r table ipcc scf}

library(knitr)
library(stats)
library(luplot)
library(reshape2)

CShare_ipcc <- new.magpie("GLO","y2010",outer(c("IPCC2006","IPCC2019"),unique(koepp2IPCC$ipcc_reduced), paste, sep="."))

CShare_ipcc[,,"IPCC2006.tropical_moist"]  <- "0.48"
CShare_ipcc[,,"IPCC2006.tropical_dry"]    <- "0.58-0.64" 
CShare_ipcc[,,"IPCC2006.temperate_moist"] <- "0.69"
CShare_ipcc[,,"IPCC2006.temperate_dry"]   <- "0.80"

CShare_ipcc[,,"IPCC2019.tropical_moist"]  <- "0.83"
CShare_ipcc[,,"IPCC2019.tropical_dry"]    <- "0.92" 
CShare_ipcc[,,"IPCC2019.temperate_moist"] <- "0.69-0.70"
CShare_ipcc[,,"IPCC2019.temperate_dry"]   <- "0.76-0.77"

comparison <- mbind(CShare_ipcc, 
                    round(add_dimension(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),
                                  add="SOCbudget",nm="model"),2))

print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))

```

### 2020-06-16 ResultNotebooks/results_SOCtesting.Rmd
Update result scripts to new structure

--- 2020-05-03 ResultNotebooks/results_socbudget.Rmd
@@
 ```{r load data}
-library(moinput)
+library(mrcommons)
 library(mrvalidation)
-library(socbudget)
+library(mrSOCbudget)
 
-setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
-#setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
+#setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.43_SOC_0.5.0/", forcecache = TRUE)
+setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43_mrSOC_0.4.1/", forcecache = TRUE)
 #setConfig(cachefolder = "/home/kristine/mnt/rd3mod/inputdata/cache/rev4.43/", forcecache = TRUE)
@@
 reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
+
+print(mbind(dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=c(1,3)),
+            dimSums(dimSums(SoilCarbon[,,"steadystate"], dim=3.3)* Landuse, dim=c(1,3))))
 
@@
 ```
+
 ```{r barplot reg stock, out.width="100%", out.height=1000}
@@
 ```
+Plot global stocks
+
+```{r plot glo stocks}
+
+plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red", xlim=c(1951,2017),  ylim=c(400000,1300000))
+lines(getYears(LPJ_IPCC, as.integer = TRUE), LPJ_IPCC["GLO",,1], type="l", col="green")
+lines(getYears(LPJ_car, as.integer = TRUE), LPJ_car["GLO",,], type="l", col="darkgreen")
+lines(getYears(LPJ_rev21, as.integer = TRUE), LPJ_rev21["GLO",,2], type="l", col="lightgreen")
+lines(getYears(WISE, as.integer = TRUE), WISE["GLO",,], type="l", col="brown")
+lines(getYears(SoilGrids, as.integer = TRUE), SoilGrids["GLO",,], type="l", col="blue")
+lines(getYears(GSOC, as.integer = TRUE), GSOC["GLO",,], type="l", col="pink")
+plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red")
+```
+
+Plot cellular stocks
+
+```{r plot reg stock, out.width="100%", out.height=1000}
+
+library(luplot)
+library(magpiesets)
+
+cellSOCdens   <- (dimSums(dimSums(SoilCarbon[,,"naturalstate"], dim=3.2) * Landuse, dim=c(3))/dimSums(Landuse, dim=c(3)))[,findset("past"),]
+cellSOCdens[is.na(cellSOCdens)] <- 0
+
+cellSoilGrids <- readSource("SoilGrids", subtype="cstock_0_30", convert="onlycorrect")
+cellLPJml     <- calcOutput("LPJmL", version="LPJmL4", climatetype="CRU_4", subtype="soilc_layer", time="raw", selectyears="past", aggregate=FALSE)
+cellLPJml     <- collapseNames(cellLPJml[,,"layer1"]+1/3*cellLPJml[,,"layer2"], collapsedim = c(2,3))
+
+diffSoilGrids <- cellSOCdens - cellSoilGrids
+diffLPJml     <- cellSOCdens - cellLPJml
+
+luplot::plotmap(diffLPJml[,10,], legend_range = c(-80,80))
+```
 
 ```{r table dens climatezones}
+
+    library(magpiesets)
 
@@
 
-    SOCDens_lut <- dimSums(SoilCarbon[,,c("steadystate","actualstate")], dim=3.3)
+    SOCDens_lut <- dimSums(SoilCarbon[,,c("steadystate","actualstate","naturalstate")], dim=3.3) 
+    SOCDens_lut[,,"naturalstate.crop"] <- SOCDens_lut[,,"actualstate.crop"]
 
@@
 
-cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3)
+cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/dimSums(SoilCarbon[,,"naturalstate"][,,"natveg"], dim=3)
 cropland_share[Landuse[,,"crop"]==0] <- NA
@@
 library(RColorBrewer)
-cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(1000)
+cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(7)[]
 outputdir <- "/home/kristine/WORK/Paper/historicalsocbudget/SOCbudgetPaper/images/maps/"

### 2020-06-16 ResultNotebooks/results_socbudget.R

@@ -1,72 +1,0 @@
----
-title: "First SOC budget analysis"
-output: html_notebook
----
-
-  
-
-```{r load data}
-library(moinput)
-library(mrvalidation)
-library(socbudget)
-
-setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-
-SoilCarbon <- calcOutput("CarbonBudget", aggregate = FALSE)
-Landuse    <- calcOutput("Landuse",      aggregate=FALSE)
-
-```
-
-Calculate global carbon stocks
-
-```{r Carbon Stock (cell iso reg glo)}
-
-SOCStock     <- dimSums(dimSums(SoilCarbon, dim=3.2)* Landuse, dim=3)
-glo_SOCStock <- dimSums(SOCStock, dim=1) 
-cell2iso     <- toolGetMapping("CountryToCellMapping.csv", type = "cell")
-iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
-iso2reg      <- toolMappingFile("regional",getConfig("regionmapping"),readcsv = TRUE)
-reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
-
-```
-
-Load validation data:
-
-```{r Validation Stocks}
-
-  LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
-  LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
-  LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
-  WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
-  GSOC       <- calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate="REG+GLO", try=TRUE)
[…]

### 2020-06-16 SOCbudgetPaper/discussion.Rmd
update method part
Update Paper part to new structure + small draft for intro, some ideas for discussion

@@
 # Discussion
+
+Big points:
+
+- SOC initialization is bad (Task: find a better representation)
+
+- natural land representation is lacking proper  parametrization of natural input (n, lg content of litterfall) (--> ask Christop again)
+
+- boreal zone and dry regions are not well represented (maybe because of the bad parameterization of the soil model for natural soils)
+
+- As pointed out by Keel (2017), Smith (2019) the results might be highly sensitive to carbon input calculations more precisly to below and above ground residue carbon estimates derived from harvested quantities.
+It has been questioned, that below ground residues might increase with a fixed root:shot ratio (maybe specifically in high end farming systems (?)).
+Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
+
+- fertilizer interaction (--> ask LPJmLer again)
+
+- diaggregation of manure with urban area is leeding to a lot of displaced manure (?) that is cut off
+
+- mismatches between FAO/LUH
 
@@
 - Non-net/Gross land use transitions are not tracked in this study.
+
+- Within cropland we do not track area transitions, but rather look at statistical distributions of the crop functional types.
+Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.

### 2020-06-16 SOCbudgetPaper/introduction.Rmd
Update Paper part to new structure + small draft for intro, some ideas for discussion

@@
 
-Introduction text goes here.
-You can change the name of the section if neccessary using `\introduction[modified heading]`.
+<!-- Relevance of the general topic
+- why important (rescue the world)
+Stand der Forschung, sher spezifisch auf deine fragestellung
+- Gibt viel biophysikalische Modellierung
+Forschungslücke (fängt mit "However" an.)
+- Zu wenig Fokus auf management, obwohl sehr einflussreich
+Fragestellung, Subfragestellungen
+Struktur der Arbeit
+-->
+
+<!-- You can change the name of the section if neccessary using `\introduction[modified heading]`. -->
+
+Within the last centuries the recognition of soil organic carbon (SOC) has developed from being a major source of agricultural carbon emissions towards a potentially large additional sink to offset other green house gas emissions and thus mitigate climate change.
+Whereas the extent of sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, over topping the atmospheric and even more the biospheric carbon pool multiple times.
+Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+permafrost melting).
+The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+
+Mapping the worlds SOC has been of great interest over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
+It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard.
+These factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for SOC dynamics till dawn of civilization <!--to heroic find another word-->.
+
+Recent studies identified the SOC debt of anthropogenic source at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
+Other studies have focus more closely on spatially disaggregation SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
+Most of these studies doing cutting edge research in their field, but seem to lack proper representation of available information on land management.
+
+On the other hand small-scale models (ref.
+Daycent, RothC, Ecosse, C-Tool) are able to represent fairly well SOC dynamics on field-scale driven by high quality management data on yield levels, fertilizer inputs and various other on farming activities.
+Due to the lack of proper, concise management data to feed these input demanding models on larger extends, it seems still very complex to scale them up to a global extent.
+
+This study however wants to contribute filling the gap on impacts of agricultural management like yield levels, residue recycling, manure amendments, irrigation and tillage on SOC stocks of the worlds croplands.
+It is doing so by combining work on agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as OC flow dynamics within the agricultural system.
+
+This is critically important since managing agricultural soil might be one of the few setscrew for humans to _naturally_ combat climate change (ref.
+to natural climate solutions).
+Other _natural_ options, which might be summarized by restoring potential natural stocks via renaturation of managed land, seem to be often limited due to increased pressure on the land systems for feed, food and energy demand of a growing and developing population on earth.
+
+<!--
+
+Within the terrestrical carbon cycle, soil organic carbon (SOC) accounts for over 1500 GtC/PgC.
+It is b far bigger than the biosperic pool and almost twice as big as the athmosperic pool.
+Small changes in SOC stocks migh lead to
+
+ Assessing the size of SOC pool as been a subject to multiple analysis, using variuos kinds of tools and methods to address getting a grip on SOC states.
+
+Not only the state but also the development over time as been lately researched more and more (ref: sanderman, Pugh, LPJmL), looking on the key aspects that change over time: climatic conditions as well as human interventions like land use change.
+Whereas it is commonly agreed upon that SOC stocks decline for most of the cropping activities for humid and wet climates (dry cimlates might be an exception to this general statement), the level of degradation is an ongoing topic of research.
+It has been access heavly from the experiemntal site as well as been researched by the theoretical models.
+
+Soil organic carbon (SOC) enhancement is important for climate change mitigation and, if practiced large scale a negative emissions technique as well.
+This is why, we need a good understanding of the current state of SOC and its most important drivers.
+
+CLAIM
+
+Soil organic carbon (SOC) enhancement is important for climate change mitigation and, if practiced large scale a negative emissions technique as well.
+This is why, we need a good understanding of the current state of SOC and its most important drivers.
+
+EVIDENCE
+
+The SOC carbon pool up to 1 meter depth contains about 1500 GtC and is the largest terrestrial carbon pool on earth.
+It is twice as big as the atmospheric and almost three times bigger than the vegetation carbon pool.
+Even small changes might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration.
+
+REBUTTAL
+
+However most of the SOC is stored in boreal climate zones as well as in peat and wetlands around the world, which might be more exposed to carbon degradation, than actually being able to store additional carbon.
+Although this holds true, the total amount of SOC under cropped and grazed land that might be altered by improved agricultural techniques is still substantial and might lead according to Minansy et al.
+(2018) to a additive carbon uptake of around 5-10 GtC per year.
+
+SOC enhancement will also lead to co-benefits for soil quality and therefor also for resilience, ecosystem services and in the end crop yields.
+
+Their are widely known and tested SOC enhancement techniques that could used large scale, if C storage would be rewarded (carbon price).
+
+WARRANTS
+
+The bigger the pool the bigger even small changes will be.
+
+CC mitigation needs co-benifits
+
+* sanderman
+* WISE
+* SoilGrids
+* Stoorvogel
+* HANPP
+
+* Smith/Powlson
+* Lal
+* Bathjes
+-->

### 2020-06-16 SOCbudgetPaper/results.Rmd
Update Paper part to new structure + small draft for intro, some ideas for discussion

@@
 # Results
+
+```{r, results = 'asis', echo=FALSE, fig.show='hold', fig.width=2, fig.height=2, fig.cap="Test", fig.subcap = c("1","2")}
+knitr::include_graphics(c("images/maps/CShare_1965.png"))
+knitr::include_graphics(c("images/maps/CIncrease_1965.png"))
+```
+
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "one column figure"}
+knitr::include_graphics("images/maps/CShare_4ColorClimates.png")
+```
+
+plot(1:10)
+plot(rnorm(10))
+```
 
@@
 ```
+
+```{r 4planel plot, out.width="50%", fig.hold="hold"}
+par(mar = c(4, 4, 0.1, 0.1))
+
+```
+
+```{r table, echo = FALSE}
+
+table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
+knitr::kable(table$naturalstate_table, "latex", digit=2, caption="SCF compared to potential natural state")
+knitr::kable(table$actualstate_table,  "latex", digit=2, caption="SCF compared to actual natural state")
+knitr::kable(table$steadystate_table,  "latex", digit=2, caption="Steady state SCF compared to actual natural state")
+
+```

### 2020-06-16 SOCbudgetPaper/SOCbudget.tex
clean up file list for Paper Part

@@
 
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
 \begin{document}
@@
 
-Introduction text goes here. You can change the name of the section if
-neccessary using
-\texttt{\textbackslash{}introduction{[}modified\ heading{]}}. \newpage
+Within the last centuries the recognition of soil organic carbon (SOC)
+has developed from being a major source of agricultural carbon emissions
+towards a potentially large additional sink to offset other green house
+gas emissions and thus mitigate climate change. Whereas the extent of
+sink potential by SOC enhancement is highly debated (add refs), it is
+largely agreed upon that the SOC pool itself is the biggest terrestrial
+carbon pool, over topping the atmospheric and even more the biospheric
+carbon pool multiple times. Even small changes in SOC drivers might lead
+to substantial shifts in earth carbon cycled and influence the
+atmospheric CO2 concentration (ref. permafrost melting). The specific
+amount of carbon stored in the soil is uncertain though and estimates
+ranging from 1500 to 2400 GtC for the first meter of the soil profile
+(Bathjes, 1996).
+
+Mapping the worlds SOC has been of great interest over the last decades
+and led to an increased quality of SOC maps as well as a even better
+understanding of factors driving magnitude, distribution and dynamics of
+SOC pools. It is undisputed that natural properties like climatic,
+biophysical and landscape characteristics play the most important role
+in this regard. These factors can not be easily altered by human
+intervention, leaving land cover, land use and land management changes
+the most important factors for SOC dynamics till dawn of civilization .
+
+Recent studies identified the SOC debt of anthropogenic source at around
+116 GtC (Sanderman et al.), which compares to older estimates of around
+60-130 GtC (Lal, 2006). Other studies have focus more closely on
+spatially disaggregation SOC changes via advanced digital soil mapping
+techniques (S-World; Stoorvogel 2, 2017) or better representation of
+biogeochemcial processes within SOC dynamics (). Most of these studies
+doing cutting edge research in their field, but seem to lack proper
+representation of available information on land management.
+
+On the other hand small-scale models (ref. Daycent, RothC, Ecosse,
+C-Tool) are able to represent fairly well SOC dynamics on field-scale
+driven by high quality management data on yield levels, fertilizer
+inputs and various other on farming activities. Due to the lack of
+proper, concise management data to feed these input demanding models on
+larger extends, it seems still very complex to scale them up to a global
+extent.
+
+This study however wants to contribute filling the gap on impacts of
+agricultural management like yield levels, residue recycling, manure
+amendments, irrigation and tillage on SOC stocks of the worlds
+croplands. It is doing so by combining work on agricultural management
+data on global level with a robust but light weighted SOC model to
+estimate SOC stocks and stock change factors as well as OC flow dynamics
+within the agricultural system.
+
+This is critically important since managing agricultural soil might be
+one of the few setscrew for humans to \emph{naturally} combat climate
+change (ref. to natural climate solutions). Other \emph{natural}
+options, which might be summarized by restoring potential natural stocks
+via renaturation of managed land, seem to be often limited due to
+increased pressure on the land systems for feed, food and energy demand
+of a growing and developing population on earth.
+
+\newpage
 
@@
 changing over time and space depending on biophysical, climatic and
-agronomic conditions. Therefor we conduct the following three steps: (1)
-Calculating annual land-use (sub-)type specific steady states and decay
-rates for SOC stocks, (2) accounting for land conversion by transferring
-SOC between land-use types and (3) updating SOC stocks based on the
-previous stock, the steady state and the decay rate.
+agronomic conditions. Therefore we conduct the following three steps
+within each yearly timestep: (1) We calculate annual land-use (sub-)type
+specific steady states and decay rates for SOC stocks, (2) We account
+for land conversion by transferring SOC between land-use types and (3)
+We estimate SOC stocks based on the stocks of the previous time period,
+the steady state stocks and the decay rate.
 
@@
 carbon stocks \(SOC^{eq}\) are given by \begin{equation}
-SOC^{eq} =\frac{C^{\textrm{in}}}{k}
+SOC^{eq}_t =\frac{C^{\textrm{in}}_t}{k_t}
 \label{eq:inoutflow}
@@
 
+BB: I would include the t, also to show that the steady-state is
+time-dependent.
+
 \hypertarget{sec:carboninputs}{%
@@
 We account for different carbon input sources depending on the land-use
-type (see table @ref(tab:datasourceinputs)). Following the IPCC
-methodology carbon inputs are disaggregated into metabolic and
-structural components depending on their lignin and nitrogen content
-(see @ref(ipcc\_2019\_2019)). For each component the sum over all carbon
-input sources is allocated to the respective SOC sub-pools via transfer
-coefficients. This implies that not only the amount of carbon, but also
-their structural composition is determining the effective inflow. Data
-sources for all considered carbon inputs as well as for lignin and
-nitrogen content can be found in table @ref(tab:datasourceinputs).
+type (see table @ref(tab:datasourceinputs)). Name here the inputs, and
+refer to the respective sections, dont just refer to the table.
+Following the IPCC methodology carbon inputs are disaggregated into
+metabolic and structural components depending on their lignin and
+nitrogen content (see @ref(ipcc\_2019\_2019)). For each component the
+sum over all carbon input sources is allocated to the respective SOC
+sub-pools via transfer coefficients. This implies that not only the
+amount of carbon, but also their structural composition is determining
+the effective inflow. Data sources for all considered carbon inputs as
+well as for lignin and nitrogen content can be found in table
+@ref(tab:datasourceinputs).
 
  \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
+ \caption{Type and data sources for carbon inputs to different land-use types }
  \begin{tabular}{l l l l}
  \tophline
-  land-use types   & source of carbon inputs & data source & nitrogen and lignin content \\
+  land-use types BB: table header should be bold   & source of carbon inputs & data source & nitrogen and lignin content \\
  \middlehline
@@
 
+This equation can also be reformulated to a massbalance equation as
+follows:
+
+\begin{equation}
+SOC_{t} = SOC_{t-1} - k_{t} \cdot 1\unit{a} + C^{\textrm{in}}  \qquad \forall lut, sub, i.
+\label{eq:steadystate}
+\end{equation}
+
 The global SOC stock for each time step can than be calculated via
@@
 
-Agricultural management data is based on R library package moinput
-(\citep{bodirsky_in_prep}), which compiles in its core country-specific
-FAO production and cropland statistics (\citep{FAOSTAT}) to a
-compherensive and constistent data suite. The data is prepared in 5 year
-time steps from 1965 to 2010, which also restricts our analysis to this
-time span. For all the following data, if not declared differently, we
-interpolate values linearly between the time steps and hold it constant
-before the first time step for a short spin-up phase from 1961 to 1965.
+Agricultural management data is estimated using the R library moinput
+({[}cite git link{]}), which compiles country-specific FAO production
+and cropland statistics (\citep{FAOSTAT}) to a comprehensive and
+constistent data suite. BB: Isnt the whole calculation based on R
+libraries? The data is prepared in 5 year time steps from 1965 to 2010,
+which also restricts our analysis to this time span. For all the
+following data, if not declared differently, we interpolate values
+linearly between the time steps and hold it constant before the first
+time step from 1961 to 1965. BB: too short for a spin-up, so dont use
+this word. Maybe also just ignore this detail.
 
@@
 subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17
-crop groups, assuming relative shares for each gridcell based on the
-country and year specific area shares of FAOSTAT data (\citep{FAOSTAT})
-(see @ref(append:Table\_luh2fao2mag) for more details on the crop type
-mapping). land-use transitions are calculated as net area differences of
-the land-use data on half-degree.
+crop groups, applying the relative shares for each gridcell based on the
+country and year specific area harvested shares of FAOSTAT data
+(\citep{FAOSTAT}) (see @ref(append:Table\_luh2fao2mag) for more details
+on the crop type mapping). Land-use transitions are calculated as net
+area differences of the land-use data on half-degree.
 
@@
 residues can be burned or harvested for other purposes such as feeding
-animals (\citep{weindl}), fuel or for material use.
+animals (\citep{weindl}), fuel or for material use. BB: One new thing is
+here that we use a doublecropping factor to account for lower harvest
+indices of multicropping vs a very high single-cropped yield
 
@@
 
-\subparagraph{Feed Demand}
-
 Residue feed demand for five different livestock groups is based on
@@
 into account.
-
-\subparagraph{Material and Fuel demand}
 
@@
 shares depending on their per-capita income.
-
-\subparagraph{Recycled Residues}
 
@@
 @ref(\#sec:landuse))) to derive half-degree pasture production patterns.
+BB: Oh, does that mean we assume that only the pasture grows that is
+being eaten? Thats not too much, is it? Maybe ask Isabelle for a better
+solution (e.g.~direct pasture yields from LPJmL). And then the ``pasture
+production'' is just the part being removed by grazing. Does the IPCC
+specifiy something in respect to pastures?
 
@@
 
-To transform dry matter estimates into carbon, we compiled crop group
+To transform dry matter estimates into carbon, we compiled crop-group
 and plant part specific carbon to dry matter (c:dm) ratios (see
@@
 were inspired by the approach of \citep{gilbert} and uses feed basket
-assumptions based on a revised methodology of \citet{weindl}.
+assumptions based on a revised methodology of \citet{weindl}. We
+differentiate ruminant and monogastric systems, as well as extensive and
+an intensive systems. For ruminants, we use the country-level shares of
+gras within the feed baskets to split up pasture-fed, extensive systems
+from the rather intensive crop-fed livestock. For extensive dairy and
+ruminant meat production we estimate, that livestock is located in
+proximity to pastures and rangelands and distribute manure excretion
+proportional to pasture production of all half-degree gridcell of a
+country. Intensive dairy and ruminant meat production is assumend to be
+located proportinal to crop production to have short transport distances
+for feed stuff.
 
-\subparagraph{Livestock Managament Intensities}
-
-To account for different livestock management intensities, we
-differentiate between extensive and an intensive systems assuming that
-they have different requirements on production factors such as feed
-availabiliy, access to markets and infrastructure (e.g.~roads, water).
-
-\subparagraph{Dairy and ruminant Meat Production}
-
-For extensive dairy and ruminant meat production we estimate, that
-livestock is located rather immovable on pastures and rangeland. We use
-the country-level pasture feed stuff shares within the feed baskets, to
-split up pasture-fed, extensive from the rather intensive, crop-fed
-livestock. Extensive production of dairy and ruminant meat is allocated
-weighted by the pasture production values to all half-degree gridcell of
-a country. On the other hand intensive dairy and ruminant meat
-production is assumend to be located proportinal to crop production to
-have short transport distances for feed stuff.
-
-\subparagraph{Poultry, Egg and monogastric Meat Production}
+BB: What does not become clear to me: Do the extensive cows only eat
+pasture, and the intensive one only crops? Does that in the end mean we
+just redistirbute grazed pasture to pastures and crop-based feed to
+croplands proportional to production? Or is there more to it?
 
@@
 (\textless1000 USD/yr), we assume extensive production systems. We
-located them according to built-up areas shares based on the idea that
+locate them according to built-up areas shares based on the idea that
 these animals are held in households, subsistence or small-holder
@@
 
-\paragraph{Manure Excretion}
+\paragraph{Manure Excretion, Storage and Recycling}
 
@@
 excreted. Carbon in excreted manure is estimated by applying fixed C:N
-ratios (given by \citep[(][]{ipcc_2019_2019}).
-
-\paragraph{Manure Storage}
-
-Depending on the feed system we assume manure to be handled in four
-different ways: All manure orginated from pasture feed intake is
-excreted directly to pastures and rangelands (pasture grazing) deducting
-manure collected as fuel. Manure fuel shares are estimated using IPCC
-default values (\citet{ippc_2006_2006}). Whereas for low-income
-countries according to worldbank definitions (\textless1000 USD/yr), we
-adopt a share of 25\% of crop residues in feed intake directly consumend
-and excreted on crop fields (stubble grazing), we do not consider any
-stubble grazing in high-income countries (\(>\,10000\,\tfrac{USD}{yr}\),
-leaving middle-income countries with linearly interpolate shares
-depending on their per-capita income. For all other feed items we assume
-the manure to be stored in animal waste management systems associated to
-animal houses.
-
-\paragraph{Manure Recycling}
-
-To estimate the carbon actually recycled to the soil, we account for
-carbon losses during storage and recycling shares in different animal
-waste management and grazing systems. Whereas we assume no losses for
-pasture and stubble grazing, we consider all carbon lost for manure
-collected as fuel. For manure stored in different animal waste
-management system we compiled carbon loss rates partly depending on
+ratios (given by \citep[(][]{ipcc_2019_2019}). Depending on the feed
+system we assume manure to be handled in four different ways: All manure
+orginated from pasture feed intake is excreted directly to pastures and
+rangelands (pasture grazing), deducting manure collected as fuel. Manure
+fuel shares are estimated using IPCC default values
+(\citet{ippc_2006_2006}). Whereas for low-income countries according to
+worldbank definitions (\textless1000 USD/yr), we adopt a share of 25\%
+of crop residues in feed intake directly consumend and excreted on crop
+fields (stubble grazing), we do not consider any stubble grazing in
+high-income countries (\(>\,10000\,\tfrac{USD}{yr}\), leaving
+middle-income countries with linearly interpolate shares depending on
+their per-capita income. For all other feed items we assume the manure
+to be stored in animal waste management systems associated to animal
+houses. To estimate the carbon actually recycled to the soil, we account
+for carbon losses during storage and recycling shares in different
+animal waste management and grazing systems. Whereas we assume no losses
+for pasture and stubble grazing, we consider that the manure collected
+as fuel is not recycled. For manure stored in different animal waste
+management system we compiled carbon loss rates partly depending on the
 nitrogen loss rates as specified in \citep{bodirsky2012} (see
@@
 
-Simple growing period calculations together with irrigation shares of
-LUH2v2 are used (BB: you dont like d's, on't you) to estimate water
-effects on decay rates.
+BB: Where does the irrigated area come from? Simple growing period
+calculations together with irrigation shares of LUH2v2 are used to
+estimate water effects on decay rates.
 
@@
 \begin{figure}
-\includegraphics[width=12cm]{images/figs_draft-1} \caption{two column figure}\label{fig:unnamed-chunk-6}
+\includegraphics[width=17.78in]{../OldDataNotTracked/images/maps/CShare_1965} \includegraphics[width=17.78in]{../OldDataNotTracked/images/maps/CIncrease_1965} \caption{Test}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=12cm]{images/figs_draft-2} \caption{two column figure}\label{fig:unnamed-chunk-7}
+\includegraphics[width=12cm]{../OldDataNotTracked/images/figs_draft-1} \caption{two column figure}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=12cm]{images/figs_draft-3} \caption{two column figure}\label{fig:unnamed-chunk-8}
+\includegraphics[width=12cm]{../OldDataNotTracked/images/figs_draft-2} \caption{two column figure}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=12cm]{images/RegionPlot+Valid} \caption{two column figure}\label{fig:unnamed-chunk-9}
+\includegraphics[width=12cm]{../OldDataNotTracked/images/figs_draft-3} \caption{two column figure}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=12cm]{images/maps/CShare_1965} \includegraphics[width=12cm]{images/maps/CShare_2010} \caption{two column figure}\label{fig:unnamed-chunk-10}
+\includegraphics[width=12cm]{../OldDataNotTracked/images/RegionPlot+Valid} \caption{two column figure}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=12cm]{images/maps/CIncrease_1965} \includegraphics[width=12cm]{images/maps/CIncrease_2010} \caption{two column figure}\label{fig:unnamed-chunk-11}
+\includegraphics[width=12cm]{../OldDataNotTracked/images/maps/CShare_4ColorClimates} \caption{one column figure}\label{fig:unnamed-chunk-11}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=12cm]{images/maps/CShare_4ColorClimates} \caption{one column figure}\label{fig:unnamed-chunk-12}
-\end{figure}
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{par}\NormalTok{(}\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}
+
+\caption{\label{tab:table}SCF compared to potential natural state}
+\centering
+\begin{tabular}[t]{l|r|r|r|r}
+\hline
+  & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
+\hline
+1965 & 0.25 & 0.33 & 0.55 & 0.48\\
+\hline
+1970 & 0.29 & 0.36 & 0.56 & 0.50\\
+\hline
+1975 & 0.32 & 0.38 & 0.57 & 0.51\\
+\hline
+1980 & 0.34 & 0.38 & 0.56 & 0.53\\
+\hline
+1985 & 0.36 & 0.40 & 0.60 & 0.56\\
+\hline
+1990 & 0.38 & 0.42 & 0.63 & 0.58\\
+\hline
+1995 & 0.40 & 0.45 & 0.63 & 0.59\\
+\hline
+2000 & 0.42 & 0.47 & 0.64 & 0.60\\
+\hline
+2005 & 0.45 & 0.54 & 0.69 & 0.65\\
+\hline
+2010 & 0.50 & 0.58 & 0.73 & 0.67\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:table}SCF compared to actual natural state}
+\centering
+\begin{tabular}[t]{l|r|r|r|r}
+\hline
+  & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
+\hline
+1965 & 0.25 & 0.33 & 0.55 & 0.48\\
+\hline
+1970 & 0.29 & 0.36 & 0.56 & 0.50\\
+\hline
+1975 & 0.32 & 0.38 & 0.57 & 0.51\\
+\hline
+1980 & 0.34 & 0.38 & 0.56 & 0.53\\
+\hline
+1985 & 0.36 & 0.40 & 0.60 & 0.56\\
+\hline
+1990 & 0.38 & 0.42 & 0.63 & 0.58\\
+\hline
+1995 & 0.40 & 0.45 & 0.63 & 0.59\\
+\hline
+2000 & 0.42 & 0.47 & 0.64 & 0.60\\
+\hline
+2005 & 0.45 & 0.54 & 0.69 & 0.65\\
+\hline
+2010 & 0.50 & 0.58 & 0.74 & 0.68\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:table}Steady state SCF compared to actual natural state}
+\centering
+\begin{tabular}[t]{l|r|r|r|r}
+\hline
+  & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
+\hline
+1965 & 0.22 & 0.32 & 0.53 & 0.47\\
+\hline
+1970 & 0.24 & 0.46 & 0.52 & 0.53\\
+\hline
+1975 & 0.25 & 0.51 & 0.69 & 0.55\\
+\hline
+1980 & 0.26 & 0.45 & 0.68 & 0.63\\
+\hline
+1985 & 0.29 & 0.57 & 0.90 & 0.87\\
+\hline
+1990 & 0.31 & 0.60 & 0.90 & 0.87\\
+\hline
+1995 & 0.35 & 0.62 & 0.85 & 0.72\\
+\hline
+2000 & 0.37 & 0.67 & 0.85 & 0.78\\
+\hline
+2005 & 0.41 & 0.78 & 1.05 & 0.91\\
+\hline
+2010 & 0.47 & 0.83 & 1.30 & 0.95\\
+\hline
+\end{tabular}
+\end{table}
 \newpage
@@
 \section{Discussion}
+
+Big points:
+
+\begin{itemize}
+\item
+  SOC initialization is bad (Task: find a better representation)
+\item
+  natural land representation is lacking proper parametrization of
+  natural input (n, lg content of litterfall) (--\textgreater{} ask
+  Christop again)
+\item
+  boreal zone and dry regions are not well represented (maybe because of
+  the bad parameterization of the soil model for natural soils)
+\item
+  As pointed out by Keel (2017), Smith (2019) the results might be
+  highly sensitive to carbon input calculations more precisly to below
+  and above ground residue carbon estimates derived from harvested
+  quantities. It has been questioned, that below ground residues might
+  increase with a fixed root:shot ratio (maybe specifically in high end
+  farming systems (?)). Following this argumentation SOC results shown
+  in this study might especially in high-yielding farming system (europe
+  etc.) overestimate actual SOC stocks.
+\item
+  fertilizer interaction (--\textgreater{} ask LPJmLer again)
+\item
+  diaggregation of manure with urban area is leeding to a lot of
+  displaced manure (?) that is cut off
+\item
+  mismatches between FAO/LUH
+\end{itemize}
 

### 2020-06-16 SOCbudgetPaper/results.Rmd
clean up file list for Paper Part

--- 2020-01-21 SOCbudgetPaper/results.Rmd
@@
-# Content section with citations
+# Results
 
-See the [R Markdown docs for bibliographies and citations](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html).
+```{r table, echo = FALSE}
 
-Copernicus supports biblatex and a sample bibliography is in file `sample.bib`.
-Read [], and [see ].
-
-# Content section with R code chunks
-
-```{r, echo = FALSE}
-sum <- 1 + 41
-```
-
-You should always use `echo = FALSE` on R Markdown code blocks as they add formatting and styling not desired by Copernicus.
-The hidden workflow results in `r sum`.
-
-You can add verbatim code snippets without extra styles by using ` ``` ` without additional instructions.
+table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
+knitr::kable(table$naturalstate_table, "latex", digit=2, caption="SCF compared to potential natural state")
+knitr::kable(table$actualstate_table,  "latex", digit=2, caption="SCF compared to actual natural state")
+knitr::kable(table$steadystate_table,  "latex", digit=2, caption="Steady state SCF compared to actual natural state")
 
 ```
-sum <- 1 + 41
-```
-
-# Content section with list
-
-If you want to insert a list, you must
-
-- leave
-
-- empty lines
-
-- between each list item

### 2020-06-18 SOCbudgetPaper/discussion.Rmd
Update on method (tillage, irrigatoin, initialization, some smaller comments) + Update intro

@@
 
+- pasture dynamics are neglected and treated as natveg
+
+- no tillage adaptaion is neglected on cropland. Pastures are assumed to be not tilled at all (propably at least heavy managed pastures are tilled with some rotation)
+
+- irrigated areas are no crop specific and growing periods are not considered (since this is very complex to calculate average growing periods).
+Crop specific growing periods might be possible using LPJmL data.
+
 - Carbon displacement via leaching and erosion is neglected in this study.

### 2020-06-18 SOCbudgetPaper/introduction.Rmd
Update on method (tillage, irrigatoin, initialization, some smaller comments) + Update intro

--- 2020-01-22 SOCbudgetPaper/introduction.Rmd
@@
 
-Introduction text goes here.
-You can change the name of the section if neccessary using `\introduction[modified heading]`.
+<!-- Relevance of the general topic
+- why important (rescue the world)
+Stand der Forschung, sher spezifisch auf deine fragestellung
+- Gibt viel biophysikalische Modellierung
+Forschungslücke (fängt mit "However" an.)
+- Zu wenig Fokus auf management, obwohl sehr einflussreich
+Fragestellung, Subfragestellungen
+Struktur der Arbeit
+-->
+
+<!-- You can change the name of the section if neccessary using `\introduction[modified heading]`. -->
+
+Within the last centuries the recognition of soil organic carbon (SOC) has developed from being a major source of agricultural carbon emissions towards a potentially large additional sink to offset other green house gas emissions and thus mitigate climate change.
+Whereas the extent of sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, over topping the atmospheric and even more the biospheric carbon pool multiple times.
+Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+permafrost melting).
+The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+
+Mapping the worlds SOC has researches intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
+It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard.
+Nevertheless,
+these factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for SOC dynamics till dawn of civilization <!--to heroic find another word-->.
+
+Recent studies identified the SOC debt of anthropogenic source at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
+Other studies have focus more closely on spatially disaggregation SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
+Most of these studies doing cutting edge research in their field, but seem to lack proper data on land management.
+
+On the other hand small-scale models (ref.
+Daycent, RothC, Ecosse, C-Tool) are able to represent well SOC dynamics on field-scale driven by detailed management data on yield levels, fertilizer inputs and various other on farming activities.
+Due to the lack of proper, concise management data to feed these input demanding models on larger extends, it seems still very complex to scale them up to a global extent.
+
+This study however wants to contribute filling the gap on impacts of agricultural management like yield levels, residue recycling, manure amendments, irrigation and tillage on SOC stocks of the worlds croplands.
+It is doing so by combining work on agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
+
+This is critically important since managing agricultural soil might be one of the few setscrew for humans to _naturally_ combat climate change (ref.
+to natural climate solutions).
+Other _natural_ options, which might be summarized by restoring potential natural stocks via renaturation of managed land, seem to be often limited due to increased pressure on the land systems for feed, food and energy demand of a growing and developing population on earth.
+
+This paper will introduce in the the first part of the method section the basic concept of SOC dynamics.
+We continue in the second part with a detailed description of the used global gridded management data on production levels, residue recycling rates, manure amendments, irrigation and tillage adoption.
+In a third, small part, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+In our result section we will focus on the SOC dynamics on global croplands by analysing (1) the spatial explicit SOC distribution and depletion over the last decades, (2) compare the climate zone specfic stock changes with default stock changes factors from the IPCC Tier 1 approach and (3) compare global agricultrual carbon flows and stocks to contextualize the importants of the different management aspects.
+
+Finally, we will discuss our findings and their implications on SOC model development and conlude with an analysis and outlook on the the opportunities for SOC management for climate change mitigation and negative emissions.
+
+<!--
+* sanderman
+* WISE
+* SoilGrids
+* Stoorvogel
+* HANPP
+
+* Smith/Powlson
+* Lal
+* Bathjes
+-->

### 2020-06-18 SOCbudgetPaper/methods.Rmd
Update on method (tillage, irrigatoin, initialization, some smaller comments) + Update intro

@@
 
-BB: I would include the t, also to show that the steady-state is time-dependent.
+<!-- BB: I would include the t, also to show that the steady-state is time-dependent. -->
 
@@
  \tophline
-  land-use types BB: table header should be bold   & source of carbon inputs & data source & nitrogen and lignin content \\
+  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
@@
 
-This equation can also be reformulated to a massbalance equation as follows:
+Reformulated equation \ref(eq:steadystate), we obtain a massbalance equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - k_{t} \cdot 1\unit{a} + C^{\textrm{in}}  \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall lut, sub, i.
 \label{eq:steadystate}
@@
 
-To initialize all SOC sub-pools we assume that cropped land 
-natural vegetation 
-* steady-states or 
-* spin up
+To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initilzation year 1961.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency for the initial conditions will decrease.
 
@@
 
-Agricultural management data is estimated using the R library moinput ([cite git link]), which compiles country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
-BB: Isnt the whole calculation based on R libraries?
+Agricultural management data is estimated using the R library mrcommons ([cite git link]), which compiles country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
+<!-- BB: Isnt the whole calculation based on R libraries?
+- KK: Sure but I thought makig the relevenat packages visible here makes sense?
+-->
 The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis to this time span.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold it constant before the first time step from 1961 to 1965.
-BB: too short for a spin-up, so dont use this word.
-Maybe also just ignore this detail.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps.
 
@@
 We assume that all bg residues are recycled to the soil, whereas ag residues can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
-    BB: One new thing is here that we use a doublecropping factor to account for lower harvest indices of multicropping vs a very high single-cropped yield
+    <!-- BB: One new thing is here that we use a doublecropping factor to account for lower harvest indices of multicropping vs a very high single-cropped yield - KK: ASk Benni again about this -->
 
@@
 Following the same approach as for crop production we disaggregate and calibrate half-degree pasture production pattern from grass yields from LPJmL and pasture area and rangeland patterns ( (see \@ref(#sec:landuse))) to derive half-degree pasture production patterns.
-BB: Oh, does that mean we assume that only the pasture grows that is being eaten?
+<!-- BB: Oh, does that mean we assume that only the pasture grows that is being eaten?
 Thats not too much, is it?
@@
 Does the IPCC specifiy something in respect to pastures?
+- KK: Clarify with Benni.But actually pastrue production does not matter since just the patterns are used to distribute livestock.
+Anyway are word on the neglected SOC cycling on pasture compared to natveg has to come somewhere.
+-->
 
@@
 
-BB: What does not become clear to me: Do the extensive cows only eat pasture, and the intensive one only crops?
+<!-- BB: What does not become clear to me: Do the extensive cows only eat pasture, and the intensive one only crops?
 Does that in the end mean we just redistirbute grazed pasture to pastures and crop-based feed to croplands proportional to production?
 Or is there more to it?
+- KK: Now acctually not.-->
 
@@
 ### Irrigation (100)
-BB: Where does the irrigated area come from?
-Simple growing period calculations together with irrigation shares of LUH2v2 are used to estimate water effects on decay rates.
+
+We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
+The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
+We apply aggregated area shares, leading to the same impact of irrigation of all our crop groups.
+Moreover we assume the irrigation effect to te present for all month of a year, that has been marked as irrigated.
 
 ### Tillage (100)
+
+To account for tillage distribution between the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we use the simple assumptions that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
+This rules are inspired by the approach of [@porvolik], but neglects the less common adoption of no tillage farming on agricultural land.
+
 Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.

### 2020-06-18 SOCbudgetPaper/introduction.Rmd
Update introduction.Rmd

@@
 
-Within the last centuries the recognition of soil organic carbon (SOC) has developed from being a major source of agricultural carbon emissions towards a potentially large additional sink to offset other green house gas emissions and thus mitigate climate change.
-Whereas the extent of sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, over topping the atmospheric and even more the biospheric carbon pool multiple times.
+<!-- this is not a mitigation paper. Your first paragraph is however on mitigation. I would frame it closer to your research. This is a better intro for your next paper -->
+
+The depletion of Soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
+Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
 Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
@@
 
-Mapping the worlds SOC has researches intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
-It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard.
-Nevertheless,
-these factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for SOC dynamics till dawn of civilization <!--to heroic find another word-->.
-
-Recent studies identified the SOC debt of anthropogenic source at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
-Other studies have focus more closely on spatially disaggregation SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
-Most of these studies doing cutting edge research in their field, but seem to lack proper data on land management.
-
-On the other hand small-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to represent well SOC dynamics on field-scale driven by detailed management data on yield levels, fertilizer inputs and various other on farming activities.
-Due to the lack of proper, concise management data to feed these input demanding models on larger extends, it seems still very complex to scale them up to a global extent.
-
-This study however wants to contribute filling the gap on impacts of agricultural management like yield levels, residue recycling, manure amendments, irrigation and tillage on SOC stocks of the worlds croplands.
-It is doing so by combining work on agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
-
 This is critically important since managing agricultural soil might be one of the few setscrew for humans to _naturally_ combat climate change (ref.
@@
 Other _natural_ options, which might be summarized by restoring potential natural stocks via renaturation of managed land, seem to be often limited due to increased pressure on the land systems for feed, food and energy demand of a growing and developing population on earth.
+
+Mapping the worlds SOC has been researched intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
+It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard.
+Nevertheless,
+these factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for current SOC dynamics.
+<!--to heroic find another word-->.
+
+Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
+Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
+Despite their importance in their field, these studies lack a detailed consideration of land management.
+
+Also small-scale models (ref.
+Daycent, RothC, Ecosse, C-Tool) are able to represent well SOC dynamics on field-scale driven by detailed management data on yield levels, fertilizer inputs and various other on farming activities.
+Due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
+
+Our study therefore estimates the change of global SOC stocks under specific consideration of agricultural mangement, including the impacts of different yield levels, residue recycling, manure amendments, irrigation and tillage.
+Our approach combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
 

### 2020-06-18 SOCbudgetPaper/introduction.Rmd
Update introduction.Rmd

@@
 
-The depletion of Soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
-Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
-Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+Soil organic carbon (SOC) is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
+Estimates of this carbon pool are still connected to large uncertainties, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+Even small changes in SOC can therefore lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
 permafrost melting).
-The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+Mapping the worlds SOC has been investigated intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
+Natural properties like climatic, biophysical and landscape characteristics clearly play the most important role in this regard.
 
-This is critically important since managing agricultural soil might be one of the few setscrew for humans to _naturally_ combat climate change (ref.
-to natural climate solutions).
-Other _natural_ options, which might be summarized by restoring potential natural stocks via renaturation of managed land, seem to be often limited due to increased pressure on the land systems for feed, food and energy demand of a growing and developing population on earth.
-
-Mapping the worlds SOC has been researched intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
-It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard.
-Nevertheless,
-these factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for current SOC dynamics.
-<!--to heroic find another word-->.
-
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change today.
+These  The depletion of Soil organic carbon (SOC) due to agricultural management has been a major source of carbon emissions.
 Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
 Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
-Despite their importance in their field, these studies lack a detailed consideration of land management.
+Despite providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management.
 
-Also small-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to represent well SOC dynamics on field-scale driven by detailed management data on yield levels, fertilizer inputs and various other on farming activities.
-Due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
+Field-scale models (ref.
+Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on yield levels, fertilizer inputs and various other on farming activities.
+Yet, due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
 
-Our study therefore estimates the change of global SOC stocks under specific consideration of agricultural mangement, including the impacts of different yield levels, residue recycling, manure amendments, irrigation and tillage.
-Our approach combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
+Our study combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
+We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation and tillage.
+We thereby provide the first global, spatial explicit SOC maps that consider agricultural management.
 

### 2020-06-26 SOCbudgetPaper/introduction.Rmd
update introduction (reinstall first paragraph as before)

@@
 
-<!-- Relevance of the general topic
-- why important (rescue the world)
-Stand der Forschung, sher spezifisch auf deine fragestellung
-- Gibt viel biophysikalische Modellierung
-Forschungslücke (fängt mit "However" an.)
-- Zu wenig Fokus auf management, obwohl sehr einflussreich
-Fragestellung, Subfragestellungen
-Struktur der Arbeit
--->
+The depletion of soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
+Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
+Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+permafrost melting).
+The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
 
-<!-- You can change the name of the section if neccessary using `\introduction[modified heading]`. -->
-
-<!-- this is not a mitigation paper. Your first paragraph is however on mitigation. I would frame it closer to your research. This is a better intro for your next paper -->
-
-Soil organic carbon (SOC) is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
-Estimates of this carbon pool are still connected to large uncertainties, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
-Even small changes in SOC can therefore lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
-permafrost melting).
 Mapping the worlds SOC has been investigated intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
@@
 
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change today.
-These  The depletion of Soil organic carbon (SOC) due to agricultural management has been a major source of carbon emissions.
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change on managed land today.
 Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
@@
 In our result section we will focus on the SOC dynamics on global croplands by analysing (1) the spatial explicit SOC distribution and depletion over the last decades, (2) compare the climate zone specfic stock changes with default stock changes factors from the IPCC Tier 1 approach and (3) compare global agricultrual carbon flows and stocks to contextualize the importants of the different management aspects.
-
 Finally, we will discuss our findings and their implications on SOC model development and conlude with an analysis and outlook on the the opportunities for SOC management for climate change mitigation and negative emissions.

### 2020-06-26 SOCbudgetPaper/SOCbudget.tex
update method part + smaller stuff in other section and result figure

@@
 
-Within the last centuries the recognition of soil organic carbon (SOC)
-has developed from being a major source of agricultural carbon emissions
-towards a potentially large additional sink to offset other green house
-gas emissions and thus mitigate climate change. Whereas the extent of
-sink potential by SOC enhancement is highly debated (add refs), it is
-largely agreed upon that the SOC pool itself is the biggest terrestrial
-carbon pool, over topping the atmospheric and even more the biospheric
-carbon pool multiple times. Even small changes in SOC drivers might lead
-to substantial shifts in earth carbon cycled and influence the
-atmospheric CO2 concentration (ref. permafrost melting). The specific
-amount of carbon stored in the soil is uncertain though and estimates
-ranging from 1500 to 2400 GtC for the first meter of the soil profile
-(Bathjes, 1996).
+The depletion of soil organic carbon (SOC) due agricultural management
+has been a major source of carbon emissions, but also offers the
+potential to turn into a large additional carbon sink and thus mitigate
+climate change. Whereas the extent of the sink potential by SOC
+enhancement is highly debated (add refs), it is largely agreed upon that
+the SOC pool itself is the biggest terrestrial carbon pool, exceeding
+the atmospheric and even more the biospheric carbon pool multiple times.
+Even small changes in SOC drivers might lead to substantial shifts in
+earth carbon cycled and influence the atmospheric CO2 concentration
+(ref. permafrost melting). The specific amount of carbon stored in the
+soil is uncertain though and estimates ranging from 1500 to 2400 GtC for
+the first meter of the soil profile (Bathjes, 1996).
 
-Mapping the worlds SOC has been of great interest over the last decades
-and led to an increased quality of SOC maps as well as a even better
-understanding of factors driving magnitude, distribution and dynamics of
-SOC pools. It is undisputed that natural properties like climatic,
-biophysical and landscape characteristics play the most important role
-in this regard. These factors can not be easily altered by human
-intervention, leaving land cover, land use and land management changes
-the most important factors for SOC dynamics till dawn of civilization .
+Mapping the worlds SOC has been investigated intensivly over the last
+decades and led to an increased quality of SOC maps as well as a even
+better understanding of factors driving magnitude, distribution and
+dynamics of SOC pools. Natural properties like climatic, biophysical and
+landscape characteristics clearly play the most important role in this
+regard.
 
-Recent studies identified the SOC debt of anthropogenic source at around
-116 GtC (Sanderman et al.), which compares to older estimates of around
-60-130 GtC (Lal, 2006). Other studies have focus more closely on
-spatially disaggregation SOC changes via advanced digital soil mapping
-techniques (S-World; Stoorvogel 2, 2017) or better representation of
-biogeochemcial processes within SOC dynamics (). Most of these studies
-doing cutting edge research in their field, but seem to lack proper
-representation of available information on land management.
+Human intervention, including land cover change and land management, has
+however added a further driver to SOC change, which alters terrestrial
+carbon pools in much shorter time scales and is likely the dominant
+driver of SOC change on managed land today. Recent studies identified
+the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which
+compares to older estimates of around 60-130 GtC (Lal, 2006). Other
+studies have focused more closely on spatial disaggregated SOC changes
+via advanced digital soil mapping techniques (S-World; Stoorvogel 2,
+2017) or better representation of biogeochemcial processes within SOC
+dynamics (). Despite providing a rough estimate of the order of
+magnitude of change, these studies lack a detailed consideration of land
+management.
 
-On the other hand small-scale models (ref. Daycent, RothC, Ecosse,
-C-Tool) are able to represent fairly well SOC dynamics on field-scale
-driven by high quality management data on yield levels, fertilizer
-inputs and various other on farming activities. Due to the lack of
-proper, concise management data to feed these input demanding models on
-larger extends, it seems still very complex to scale them up to a global
-extent.
+Field-scale models (ref. Daycent, RothC, Ecosse, C-Tool) are able to
+capture these land mangement impacts by using detailed information on
+yield levels, fertilizer inputs and various other on farming activities.
+Yet, due to the lack of comprehensive global management data to feed
+these input demanding models, it seems still very complex to scale them
+up to a global extent.
 
-This study however wants to contribute filling the gap on impacts of
-agricultural management like yield levels, residue recycling, manure
-amendments, irrigation and tillage on SOC stocks of the worlds
-croplands. It is doing so by combining work on agricultural management
-data on global level with a robust but light weighted SOC model to
-estimate SOC stocks and stock change factors as well as OC flow dynamics
-within the agricultural system.
+Our study combines an spatial-explicit estimate of agricultural
+management data on global level with a robust but light weighted SOC
+model to estimate SOC stocks and stock change factors as well as organic
+carbon flow dynamics within the agricultural system. We thereby consider
+change in SOC caused by historical land cover change as well as of
+different agricultural mangement practices, including residue recycling,
+manure amendments, irrigation and tillage. We thereby provide the first
+global, spatial explicit SOC maps that consider agricultural management.
 
-This is critically important since managing agricultural soil might be
-one of the few setscrew for humans to \emph{naturally} combat climate
-change (ref. to natural climate solutions). Other \emph{natural}
-options, which might be summarized by restoring potential natural stocks
-via renaturation of managed land, seem to be often limited due to
-increased pressure on the land systems for feed, food and energy demand
-of a growing and developing population on earth.
+This paper will introduce in the the first part of the method section
+the basic concept of SOC dynamics. We continue in the second part with a
+detailed description of the used global gridded management data on
+production levels, residue recycling rates, manure amendments,
+irrigation and tillage adoption. In a third, small part, we shortly
+refer to the concept of stock change factors as outlined in the Tier 1
+approach of the IPCC guidelines. In our result section we will focus on
+the SOC dynamics on global croplands by analysing (1) the spatial
+explicit SOC distribution and depletion over the last decades, (2)
+compare the climate zone specfic stock changes with default stock
+changes factors from the IPCC Tier 1 approach and (3) compare global
+agricultrual carbon flows and stocks to contextualize the importants of
+the different management aspects. Finally, we will discuss our findings
+and their implications on SOC model development and conlude with an
+analysis and outlook on the the opportunities for SOC management for
+climate change mitigation and negative emissions.
 
@@
 \section{Method (50)}
+
+We compiled calculations as open-source R packages available under
+github.com/pik-piam/mrcommons (management related functions) and
+github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which
+are both based on github.com/pik-piam/madrat, a framework improving
+reproducibility and transparency in data processing. In the following
+chapter we outine the most important relationships and assumptions. See
+table @ref(append:subsection2mrfunctions) for further information on
+corresponding code within the R packages.
 
@@
 Following the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 (\citet{ipcc_2019_2019}), we estimate global land-use type
-specific soil organic carbon (SOC) stocks for cropland and natural
-vegetation on half-degree resolution from 1965 to 2010. We assume the
-actual SOC state converges towards a stable steady state, that itself is
-changing over time and space depending on biophysical, climatic and
-agronomic conditions. Therefore we conduct the following three steps
-within each yearly timestep: (1) We calculate annual land-use (sub-)type
-specific steady states and decay rates for SOC stocks, (2) We account
-for land conversion by transferring SOC between land-use types and (3)
-We estimate SOC stocks based on the stocks of the previous time period,
-the steady state stocks and the decay rate.
+vol.~4 (\citet{ipcc_2019_2019}), we estimate soil organic carbon (SOC)
+stocks for cropland and all other land represented by natural vegetation
+on half-degree resolution from 1975 to 2010. We assume the actual SOC
+state converges towards a stable steady state, that itself is changing
+over time and space depending on biophysical, climatic and agronomic
+conditions. Therefore we conduct the following three steps within each
+yearly timestep: (1) We calculate annual land-use (sub-)type specific
+steady states and decay rates for SOC stocks, (2) We account for land
+conversion by transferring SOC between land-use types and (3) We
+estimate SOC stocks based on the stocks of the previous time period, the
+steady state stocks and the decay rate.
 
@@
 carbon stocks \(SOC^{eq}\) are given by \begin{equation}
-SOC^{eq}_t =\frac{C^{\textrm{in}}_t}{k_t}
+SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
 \label{eq:inoutflow}
-\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil
-and \(k\) denoting the soil organic carbon decay rate. We use for our
+\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil,
+\(k\) denoting the soil organic carbon decay rate; as well as \(i\)
+representing grid cells indecies and \(t\) years. We use for our
 calculations the steady-state method of the refinement of the IPCC
@@
 
-BB: I would include the t, also to show that the steady-state is
-time-dependent.
-
 \hypertarget{sec:carboninputs}{%
@@
 
-We account for different carbon input sources depending on the land-use
-type (see table @ref(tab:datasourceinputs)). Name here the inputs, and
-refer to the respective sections, dont just refer to the table.
+We account for different carbon input sources depending on our two
+distinguished land-use types (see table @ref(tab:datasourceinputs)); in
+particular for recycled crop residues, below ground biomass of crops
+(for both see @ref(sec:residues)) and recycled manure (see
+\ref(sec:livst\_manure)) on croplands; and litterfall
+(\ref(\citet{LPJmL4_1})) on natural vegetation.
+
 Following the IPCC methodology carbon inputs are disaggregated into
@@
 \begin{equation}
-SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall sub, i  
+SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
 \label{eq:ctransfer}
@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
 \label{eq:steadystate}
@@
 
-This equation can also be reformulated to a massbalance equation as
-follows:
+Reformulated equation \ref(eq:steadystate), we obtain a massbalance
+equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - k_{t} \cdot 1\unit{a} + C^{\textrm{in}}  \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
 \label{eq:steadystate}
@@
 
-To initialize all SOC sub-pools we assume that cropped land natural
-vegetation * steady-states or * spin up
+To initialize all SOC sub-pools we assume that cropped land as well as
+natural vegetation are in a steady state for the specific configuration
+present within the initilzation year 1961. We assume after a spin-up
+period of 15 years reliable results start from 1975 and improve over
+time, since dependency for the initial conditions will decrease.
 
 \newpage
-
-\hypertarget{sec:tier1}{%
-\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}
-
-Additionally to the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 (\citet{ipcc_2019_2019}), we also estimate SOC pools using the
-IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006})
-for comparison. Here, stocks are estimated via stock change factors
-given by the IPCC for the topsoil (0-30 cm) and based on a review of
-measurement data. The factors differentiate different crop and
-management systems reflecting different dynamics under changed in- and
-outflows without explicitly tracking these. The SOC stocks as thus
-calulated
-
-\begin{equation}
-SOC_{\text{target}} = \sum_{c,s,i} SOC_{\text{ref}_{c,s,i}} \cdot F_{\text{LU}_{c,s,i}} \cdot F_{\text{MG}_{c,s,i}} \cdot F_{\text{I}_{c,s,i}} \cdot A_{c,s,i}
-\label{eq:tier1}
-\end{equation}
-
-\textless!- also include an equation here --\textgreater{} \textless!-
-even if there are just ``copied'' out of te guidelines so to say?
---\textgreater{} \textless!- more details will follow - how deep to go?
---\textgreater{}
 
@@
 
-Agricultural management data is estimated using the R library moinput
-({[}cite git link{]}), which compiles country-specific FAO production
-and cropland statistics (\citep{FAOSTAT}) to a comprehensive and
-constistent data suite. BB: Isnt the whole calculation based on R
-libraries? The data is prepared in 5 year time steps from 1965 to 2010,
-which also restricts our analysis to this time span. For all the
-following data, if not declared differently, we interpolate values
-linearly between the time steps and hold it constant before the first
-time step from 1961 to 1965. BB: too short for a spin-up, so dont use
-this word. Maybe also just ignore this detail.
+We compile country-specific FAO production and cropland statistics
+(\citep{FAOSTAT}) to a comprehensive and constistent data suite. The
+data is prepared in 5 year time steps from 1965 to 2010, which also
+restricts our analysis together with the accounting for a spin up phase
+to time span from 1975 to 2010. For all the following data, if not
+declared differently, we interpolate values linearly between the time
+steps.
 
@@
 (\citep{FAOSTAT}) (see @ref(append:Table\_luh2fao2mag) for more details
-on the crop type mapping). Land-use transitions are calculated as net
+on the crop group mapping). Land-use transitions are calculated as net
 area differences of the land-use data on half-degree.
 
-\subsubsection{Crop, Crop Residues and Pasture Production (300)}
+\hypertarget{sec:residues}{%
+\subsubsection{Crop and Crop Residues (300)}\label{sec:residues}}
 
@@
 to its central role for soil carbon modelling. Starting from crop
-production estimates of the harvested organs and their respective crop
-area, we estimate above-ground (ag) and below-ground (bg) residual
-biomass using yield-dependent harvest indices and shoot:root ratios. We
-assume that all bg residues are recycled to the soil, whereas ag
-residues can be burned or harvested for other purposes such as feeding
-animals (\citep{weindl}), fuel or for material use. BB: One new thing is
-here that we use a doublecropping factor to account for lower harvest
-indices of multicropping vs a very high single-cropped yield
+production (\(CP\)) estimates of the harvested organs and their
+respective harvested crop area (\(CA\)), we estimate above-ground
+(\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group
+(\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios
+(\(RS\)) as follows
+
+\begin{equation}
+\begin{aligned}
+AGR & = CP \cdot HI_{\textrm{slope}} + CA \cdot HI_{\textrm{intercept}}\qquad & \textrm{and} \\
+BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
+\label{eq:resbiomass}
+\end{aligned}
+\end{equation}
+
+Following the IPCC guidelines, we split the harvest index into a yield
+and an area dependend fraction (\citep{ipcc_2006_2006}). Note that
+deviating from \citep{bodirsky2012} we use harvested instead of physical
+crop area to account for increased residue biomass due to multiple
+cropping and decreased amounts on fallow land. We assume that all BGR
+are recycled to the soil, whereas AGR can be burned or harvested for
+other purposes such as feeding animals (\citep{weindl}), fuel or for
+material use.
 
@@
 
-A fixed share of the ag residues is assumed to be burned on field
-depending on the per-capita income of the country. Following
-\citep{smil1999}) we assume 25\% burn share for low-income countries
-according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\%
-for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate
-shares for all middle-income countries depending on their per-capita
-income. Depending on the crop type 80--90\% of the residue carbon burned
-on the fields are lost within the combustion process
-(\citep{ipcc_2006_2006}).
+A fixed share of the AGR is assumed to be burned on field depending on
+the per-capita income of the country. Following \citep{smil1999}) we
+assume 25\% burn share for low-income countries according to worldbank
+definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\% for high-income
+(\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate shares for all
+middle-income countries depending on their per-capita income. Depending
+on the crop group 80--90\% of the residue carbon burned on the fields
+are lost within the combustion process (\citep{ipcc_2006_2006}).
 
@@
 purposes and one residues with no double use (see
-@ref(append:Table\_kcr2kres)).
-
-Residue feed demand for five different livestock groups is based on
-country- and residue-group-specific feed basekts (see \citep{weindl})
-taking available ag residual biomass as well as livestock productivity
-into account.
-
-We estimate a material-use share for the straw residues group of 5\% and
-a fuel-share of 10\% for all used residues groups in low income
-countries according to worldbank definitions
-(\(<\,1000\,\tfrac{USD}{yr}\)). For high-income
+@ref(append:Table\_kcr2kres)). Residue feed demand for five different
+livestock groups is based on country- and residue-group-specific feed
+basekts (see \citep{weindl}) taking available AGR biomass as well as
+livestock productivity into account. We estimate a material-use share
+for the straw residues group of 5\% and a fuel-share of 10\% for all
+used residues groups in low income countries according to worldbank
+definitions (\(<\,1000\,\tfrac{USD}{yr}\)). For high-income
 (\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is
 assumend, leaving middle-income countries with linearly interpolate
-shares depending on their per-capita income.
-
-The remaining ag residues as well as all bg residues are assumend to be
-recycled to the soil. We cut high recycling shares per hectar at the
-95\%-percentile to corrected for outliers.
-
-\paragraph{Pasture Production}
-
-Using livestock production statistics as well as feed mix assumptions as
-describted in (\citep{weindl}) we estimating country specific pasture
-production. Following the same approach as for crop production we
-disaggregate and calibrate half-degree pasture production pattern from
-grass yields from LPJmL and pasture area and rangeland patterns ( (see
-@ref(\#sec:landuse))) to derive half-degree pasture production patterns.
-BB: Oh, does that mean we assume that only the pasture grows that is
-being eaten? Thats not too much, is it? Maybe ask Isabelle for a better
-solution (e.g.~direct pasture yields from LPJmL). And then the ``pasture
-production'' is just the part being removed by grazing. Does the IPCC
-specifiy something in respect to pastures?
+shares depending on their per-capita income. The remaining AGR as well
+as all BGR are assumend to be recycled to the soil. We cut high
+recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for
+outliers.
 
@@
 
-\subsubsection{Livestock Distribution and Manure Excretion (300)}
+\hypertarget{sec:livst_manure}{%
+\subsubsection{Livestock Distribution and Manure Excretion
+(300)}\label{sec:livst_manure}}
+
+We assume that manure is applied at its excretion place, leaving the
+livestock distribution the driving factor for the spatial pattern of
+manuring.
 
@@
 were inspired by the approach of \citep{gilbert} and uses feed basket
-assumptions based on a revised methodology of \citet{weindl}. We
+assumptions based on a revised methodology of \citet{weindl.We}
 differentiate ruminant and monogastric systems, as well as extensive and
-an intensive systems. For ruminants, we use the country-level shares of
-gras within the feed baskets to split up pasture-fed, extensive systems
-from the rather intensive crop-fed livestock. For extensive dairy and
-ruminant meat production we estimate, that livestock is located in
-proximity to pastures and rangelands and distribute manure excretion
-proportional to pasture production of all half-degree gridcell of a
-country. Intensive dairy and ruminant meat production is assumend to be
-located proportinal to crop production to have short transport distances
-for feed stuff.
-
-BB: What does not become clear to me: Do the extensive cows only eat
-pasture, and the intensive one only crops? Does that in the end mean we
-just redistirbute grazed pasture to pastures and crop-based feed to
-croplands proportional to production? Or is there more to it?
-
-For poultry, egg and monogastric meat production we use the per-capita
-income of the country to divide into intensive and extensive production
-systems. For low-income countries according to worldbank definitions
-(\textless1000 USD/yr), we assume extensive production systems. We
-locate them according to built-up areas shares based on the idea that
-these animals are held in households, subsistence or small-holder
-farming systems with a high labour per animal ratio. Intensive
-production is distributed within a country using the crop production
-share, assuming that feed availability is the most driving factor for
-livestock location.
+an intensive systems. For ruminants, we assume that livestock is located
+at the production spot sof the feed intake. We distingush between grazed
+pasture, which is converted into livestock products from extensive
+systems; and all other (crop-based) feed stuff, which we consider to be
+consumend in intensive systems. For poultry, egg and monogastric meat
+production we use the per-capita income of the country to divide into
+intensive and extensive production systems. For low-income countries
+according to worldbank definitions (\textless1000 USD/yr), we assume
+extensive production systems. We locate them according to built-up areas
+shares based on the idea that these animals are held in households,
+subsistence or small-holder farming systems with a high labour per
+animal ratio. Intensive production is distributed within a country using
+again crop-based feed production share, assuming that feed availability
+is the most driving factor for livestock location.
 
@@
 nitrogen loss rates as specified in \citep{bodirsky2012} (see
-@ref(append:Table\_clossAWMS))).
+@ref(append:Table\_clossAWMS))). We cut high application shares at
+\(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.
 
@@
 
-BB: Where does the irrigated area come from? Simple growing period
-calculations together with irrigation shares of LUH2v2 are used to
-estimate water effects on decay rates.
+We use irrigation area shares to modify the water effect on
+decomposition by weighting the irrigated and rainfed water factors based
+on these shares. The LUH2v2 data set provides irrigated fractions for
+their cropland subcategories. We apply aggregated area shares, leading
+to the same impact of irrigation of all our crop groups. Moreover we
+assume the irrigation effect to te present for all month of a year, that
+has been marked as irrigated.
 
@@
 
-Tillage data sets of {[}Vera, others{]} together with rules are used to
-drive tillage effect on decay rates. \newpage
+To account for tillage distribution between the three tillage types
+specfied by the IPCC - full tillage, reduced tillage and no tillage -,
+we use the simple assumptions that all natural land and pastures are not
+tilled, whereas annual crops are under full and perennials under reduced
+tillage. This rules are inspired by the approach of \citep{porvolik},
+but neglects the less common adoption of no tillage farming on
+agricultural land.
+
+\newpage
+
+\hypertarget{sec:tier1}{%
+\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}
+
+Additionally to the tier 2 approach of the refinement of IPCC guidelines
+vol.~4 (\citet{ipcc_2019_2019}) and the detailed analysis of management
+data coming with it, SOC changes can be estimated using the IPCC tier 1
+approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006},
+\citet{ipcc_2019_2019}). Here, stocks are calculated via stock change
+factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on
+observational data. SCF are differentiated by different crop and
+management systems (\(m\)) reflecting different dynamics under changed
+in- and outflows without explicitly tracking these. Moreover SCF vary
+for different climatic zones (\(c\)) specified by the IPCC (see
+@ref(append:climatemap)). The actual SOC stocks as thus calulated based
+on a given reference SOC stock by
+
+\begin{equation}
+SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F_{\text{LU}_{c,m}} \cdot F_{\text{MG}_{c,m}} \cdot F_{\text{I}_{c,m}} \cdot A_{m,i} \qquad\forall\quad t,
+\label{eq:tier1}
+\end{equation}
+
+with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into
+corresponding climate zones \(c\). For this analysis we us the default
+SCF as comparison and constistency check for our more detailed budget
+approach.
+
+\newpage
 
 \section{Results}
-
-\begin{figure}
-\includegraphics[width=17.78in]{../OldDataNotTracked/images/maps/CShare_1965} \includegraphics[width=17.78in]{../OldDataNotTracked/images/maps/CIncrease_1965} \caption{Test}\label{fig:unnamed-chunk-6}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=12cm]{../OldDataNotTracked/images/figs_draft-1} \caption{two column figure}\label{fig:unnamed-chunk-7}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=12cm]{../OldDataNotTracked/images/figs_draft-2} \caption{two column figure}\label{fig:unnamed-chunk-8}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=12cm]{../OldDataNotTracked/images/figs_draft-3} \caption{two column figure}\label{fig:unnamed-chunk-9}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=12cm]{../OldDataNotTracked/images/RegionPlot+Valid} \caption{two column figure}\label{fig:unnamed-chunk-10}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=12cm]{../OldDataNotTracked/images/maps/CShare_4ColorClimates} \caption{one column figure}\label{fig:unnamed-chunk-11}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-\caption{\label{tab:table}SCF compared to actual natural state}
+\caption{\label{tab:table2}SCF compared to IPCC Tier 1}
 \centering
-\begin{tabular}[t]{l|r|r|r|r}
+\begin{tabular}[t]{l|l|l|l|l}
 \hline
-  & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
+Data1 & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
 \hline
-1965 & 0.25 & 0.33 & 0.55 & 0.48\\
+IPCC2006 & 0.48 & 0.58-0.64 & 0.80 & 0.69\\
 \hline
-1970 & 0.29 & 0.36 & 0.56 & 0.50\\
+IPCC2019 & 0.83 & 0.92 & 0.76-0.77 & 0.69-0.70\\
 \hline
-1975 & 0.32 & 0.38 & 0.57 & 0.51\\
-\hline
-1980 & 0.34 & 0.38 & 0.56 & 0.53\\
-\hline
-1985 & 0.36 & 0.40 & 0.60 & 0.56\\
-\hline
-1990 & 0.38 & 0.42 & 0.63 & 0.58\\
-\hline
-1995 & 0.40 & 0.45 & 0.63 & 0.59\\
-\hline
-2000 & 0.42 & 0.47 & 0.64 & 0.60\\
-\hline
-2005 & 0.45 & 0.54 & 0.69 & 0.65\\
-\hline
-2010 & 0.50 & 0.58 & 0.74 & 0.68\\
+model & 0.5 & 0.58 & 0.73 & 0.67\\
 \hline
@@
 
-\begin{table}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# All defaults}
+\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include\_graphics}\NormalTok{(}\StringTok{"../ResultNotebooks/Output/Images/4panelfigure.png"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-\caption{\label{tab:table}Steady state SCF compared to actual natural state}
-\centering
-\begin{tabular}[t]{l|r|r|r|r}
-\hline
-  & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
-\hline
-1965 & 0.22 & 0.32 & 0.53 & 0.47\\
-\hline
-1970 & 0.24 & 0.46 & 0.52 & 0.53\\
-\hline
-1975 & 0.25 & 0.51 & 0.69 & 0.55\\
-\hline
-1980 & 0.26 & 0.45 & 0.68 & 0.63\\
-\hline
-1985 & 0.29 & 0.57 & 0.90 & 0.87\\
-\hline
-1990 & 0.31 & 0.60 & 0.90 & 0.87\\
-\hline
-1995 & 0.35 & 0.62 & 0.85 & 0.72\\
-\hline
-2000 & 0.37 & 0.67 & 0.85 & 0.78\\
-\hline
-2005 & 0.41 & 0.78 & 1.05 & 0.91\\
-\hline
-2010 & 0.47 & 0.83 & 1.30 & 0.95\\
-\hline
-\end{tabular}
-\end{table}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/4panelfigure}
 \newpage
@@
 \begin{itemize}
+\item
+  pasture dynamics are neglected and treated as natveg / Note that
+  manure excreted to pastures is neglected within these analysis, since
+  we focus purely on cropland dynamics.
+\item
+  no tillage adaptaion is neglected on cropland. Pastures are assumed to
+  be not tilled at all (propably at least heavy managed pastures are
+  tilled with some rotation)
+\item
+  irrigated areas are no crop specific and growing periods are not
+  considered (since this is very complex to calculate average growing
+  periods). Crop specific growing periods might be possible using LPJmL
+  data.
 \item

### 2020-06-26 SOCbudgetPaper/appendix.Rmd
update method part + smaller stuff in other section and result figure

* table on method subsections to functions within R packages

* table on mapping LUH2FAO2MAG

* kcr2kres mapping

* carbon 2 dry matter @Ma2018 (??) -> Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation

* closs in AWMS - Table

* map on climate zone used for SCF

### 2020-06-26 SOCbudgetPaper/methods.Rmd
update method part + smaller stuff in other section and result figure

@@
 
-<!-- Soil carbon dynamics are modeled using a yearly based SOC model and carbon input and management data. -->
+We compiled calculations as open-source R packages available under github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on github.com/pik-piam/madrat, a framework improving reproducibility and transparency in data processing.
+In the following chapter we outine the most important relationships and assumptions. See table \@ref(append:subsection2mrfunctions) for further information on corresponding code within the R packages.
 
@@
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019), we estimate global land-use type specific soil organic carbon (SOC) stocks for cropland and natural vegetation on half-degree resolution from 1965 to 2010.
+4 (@ipcc_2019_2019), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
 We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
@@
 \begin{equation}
-SOC^{eq}_t =\frac{C^{\textrm{in}}_t}{k_t}
+SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
 \label{eq:inoutflow}
 \end{equation}
-with $C^{\textrm{in}}$ being carbon inputs to the soil and $k$ denoting the soil organic carbon decay rate.
+with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cells indecies and $t$ years.
 We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.
@@
 
-<!-- BB: I would include the t, also to show that the steady-state is time-dependent. -->
-
 #### Carbon Inputs to the Soil {#sec:carboninputs}
 
-We account for different carbon input sources depending on the land-use type (see table \@ref(tab:datasourceinputs)).
-Name here the inputs, and refer to the respective sections, dont just refer to the table.
+We account for different carbon input sources depending on our two distinguished land-use types (see table \@ref(tab:datasourceinputs)); in particular for recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)) on croplands; and litterfall (\ref(@LPJmL4_1)) on natural vegetation.
+
 Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
@@
 ```
+
+<!-- add references here for the table-->
 
@@
 
+<!-- add references here for the table-->
+
 ### SOC transfer between land-use types
@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall lut, sub, i.
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
 \label{eq:steadystate}
@@
 
-## Carbon Budget following Tier 1 (150) {#sec:tier1}
-
-Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019), we also estimate SOC pools using the IPCC tier 1 approach of IPCC guidelines vol.
-4 (@ipcc_2006_2006) for comparison.
-Here, stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on a review of measurement data.
-The factors differentiate  different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-The SOC stocks as thus calulated
-
-\begin{equation}
-SOC_{\text{target}} = \sum_{c,s,i} SOC_{\text{ref}_{c,s,i}} \cdot F_{\text{LU}_{c,s,i}} \cdot F_{\text{MG}_{c,s,i}} \cdot F_{\text{I}_{c,s,i}} \cdot A_{c,s,i}
-\label{eq:tier1}
-\end{equation}
-
-<!- also include an equation here -->
-<!- even if there are just "copied" out of te guidelines so to say? -->
-<!- more details will follow - how deep to go? -->
-
 ## Agricultural management data on 0.5 degrees	(50) {#sec:agrimanagement}
 
-Agricultural management data is estimated using the R library mrcommons ([cite git link]), which compiles country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
-<!-- BB: Isnt the whole calculation based on R libraries?
-- KK: Sure but I thought makig the relevenat packages visible here makes sense?
--->
-The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis to this time span.
+We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
+The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis together with the accounting for a spin up phase to time span from 1975 to 2010.
 For all the following data, if not declared differently, we interpolate values linearly between the time steps.
@@
 Land-use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each gridcell based on the country and year specific area harvested shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop type mapping).
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each gridcell based on the country and year specific area harvested shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop group mapping).
 Land-use transitions are calculated as net area differences of the land-use data on half-degree.
 
-### Crop, Crop Residues and Pasture Production	(300)
+### Crop and Crop Residues (300) {#sec:residues}
 
@@
 Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
-Starting from crop production estimates of the harvested organs and their respective crop area, we estimate above-ground (ag) and below-ground (bg) residual biomass using yield-dependent <!-- do not know what was meant with this anymore (why "yield-dependent"?).
-I guess that has to be explained again than -.- --> harvest indices and shoot:root ratios.
-We assume that all bg residues are recycled to the soil, whereas ag residues can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
-    <!-- BB: One new thing is here that we use a doublecropping factor to account for lower harvest indices of multicropping vs a very high single-cropped yield - KK: ASk Benni again about this -->
+Starting from crop production ($CP$) estimates of the harvested organs and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
+
+\begin{equation}
+\begin{aligned}
+AGR & = CP \cdot HI_{\textrm{slope}} + CA \cdot HI_{\textrm{intercept}}\qquad & \textrm{and} \\
+BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
+\label{eq:resbiomass}
+\end{aligned}
+\end{equation}
+
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction ([@ipcc_2006_2006]).
+Note that deviating from [@bodirsky2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
 
 ##### Burned Residues
-A fixed share of the ag residues is assumed to be burned on field depending on the per-capita income of the country.
+A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
 Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop type 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).
+Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).
 
@@
 We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see \@ref(append:Table_kcr2kres)).
-
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock productivity into account.
-
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see [@weindl]) taking available AGR biomass as well as livestock productivity into account.
 We estimate a material-use share for the straw residues group of 5% and a fuel-share of 10% for all used residues groups in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
 For high-income  ($>\,10000\,\tfrac{USD}{yr}$ no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
-
-The remaining ag residues as well as all bg residues are assumend to be recycled to the soil. 
-We cut high recycling shares per hectar at the 95%-percentile to corrected for outliers.
-
-#### Pasture Production
-Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production.
-Following the same approach as for crop production we disaggregate and calibrate half-degree pasture production pattern from grass yields from LPJmL and pasture area and rangeland patterns ( (see \@ref(#sec:landuse))) to derive half-degree pasture production patterns.
-<!-- BB: Oh, does that mean we assume that only the pasture grows that is being eaten?
-Thats not too much, is it?
-Maybe ask Isabelle for a better solution (e.g.
-direct pasture yields from LPJmL).
-And then the "pasture production" is just the part being removed by grazing.
-Does the IPCC specifiy something in respect to pastures?
-- KK: Clarify with Benni.But actually pastrue production does not matter since just the patterns are used to distribute livestock.
-Anyway are word on the neglected SOC cycling on pasture compared to natveg has to come somewhere.
--->
+The remaining AGR as well as all BGR are assumend to be recycled to the soil.
+We cut high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to corrected for outliers.
 
@@
 
-<!-- Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation -->
+### Livestock Distribution and Manure Excretion	(300) {#sec:livst_manure}
 
-### Livestock Distribution and Manure Excretion	(300)
+We assume that manure is applied at its excretion place, leaving the livestock distribution the driving factor for the spatial pattern of manuring.
 
 #### Livestock Distribution
-To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert] and uses feed basket assumptions based on a revised methodology of @weindl.
-We differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
-For ruminants, we use the country-level shares of gras within the feed baskets to split up pasture-fed, extensive systems from the rather intensive crop-fed livestock.
-For extensive dairy and ruminant meat production we estimate, that livestock is located in proximity to pastures and rangelands and distribute manure excretion proportional to pasture production of all half-degree gridcell of a country.
-Intensive dairy and ruminant meat production is assumend to be located proportinal to crop production to have short transport distances for feed stuff.
-<!-- yeah, actually was that the full idea here?
--->
-
-<!-- BB: What does not become clear to me: Do the extensive cows only eat pasture, and the intensive one only crops?
-Does that in the end mean we just redistirbute grazed pasture to pastures and crop-based feed to croplands proportional to production?
-Or is there more to it?
-- KK: Now acctually not.-->
-
+To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert] and uses feed basket assumptions based on a revised methodology of @weindl.We differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
+For ruminants, we assume that livestock is located at the production spot sof the feed intake.
+We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feed stuff, which we consider to be consumend in intensive systems.
 For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
@@
 We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
-Intensive production is distributed within a country using the crop production share, assuming that feed availability is the most driving factor for livestock location.
+Intensive production is distributed within a country using again crop-based feed production share, assuming that feed availability is the most driving factor for livestock location.
 
@@
 For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
+We cut high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to corrected for outliers.
 
@@
 
-Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.
+\newpage
+
+## Carbon Budget following Tier 1 (150) {#sec:tier1}
+
+Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.
+4 (@ipcc_2019_2019) and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.
+4 (@ipcc_2006_2006, @ipcc_2019_2019).
+Here, stocks are calculated via stock change factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+SCF are differentiated by different crop and management systems ($m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover SCF vary for different climatic zones ($c$) specified by the IPCC (see \@ref(append:climatemap)).
+The actual SOC stocks as thus calulated based on a given reference SOC stock by
+
+\begin{equation}
+SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F_{\text{LU}_{c,m}} \cdot F_{\text{MG}_{c,m}} \cdot F_{\text{I}_{c,m}} \cdot A_{m,i} \qquad\forall\quad t,
+\label{eq:tier1}
+\end{equation}
+
+with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
+For this analysis we us the default SCF as comparison and constistency check for our more detailed budget approach.
+
+<!-- KK: Representation of climate zones and grid cells is not consistent and has to be reformulated. -->

### 2020-06-26 SOCbudgetPaper/results.Rmd
update method part + smaller stuff in other section and result figure

@@
 ```{r table, echo = FALSE}
-
 table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
 knitr::kable(table$naturalstate_table, "latex", digit=2, caption="SCF compared to potential natural state")
-knitr::kable(table$actualstate_table,  "latex", digit=2, caption="SCF compared to actual natural state")
-knitr::kable(table$steadystate_table,  "latex", digit=2, caption="Steady state SCF compared to actual natural state")
+```
 
+```{r table2, echo = FALSE}
+table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
+knitr::kable(table$comparison_table, "latex", digit=2, caption="SCF compared to IPCC Tier 1")
 ```
+
+```{r out.width = "16cm"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
+```

### 2020-06-26 ResultNotebooks/results_SOCmaps.Rmd
update results

@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
 
@@
 glo_SOCStock    <- dimSums(SOCStock, dim=1) 
-glo_SOCEmission <- glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"]  
+glo_SOCEmission <- (glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
 plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission/1000, type="l", xlab="Time", ylab="Global SOC Debt in GtC")
@@
 lines(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCStock[,,"naturalstate"])
+
+SOCemis <- ggplot() + geom_line(data= as.data.frame(glo_SOCEmission)[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
+  labs( x = "Year", 
+        y = "Global cummulative SOC emission in GtC") +
+  scale_x_discrete(breaks=seq(1965,2010,5)) +
+  scale_y_continuous(limits=c(25,40), breaks=seq(25,40,5)) +
+  theme(rect=element_blank())
+
 ```
@@
 
+library("ggplot2")
+theme_set(theme_bw())
+library("sf")
+library("rnaturalearth")
+library("rnaturalearthdata")
 library(RColorBrewer)
-cols      <- colorRampPalette(brewer.pal(11, "RdYlGn"))(7)
-cols_dens <- colorRampPalette(brewer.pal(11, "RdYlGn"))(30)[c(1:20, rep(c(21:30), each=10))]
+
+cols_dens <- colorRampPalette(brewer.pal(11, "RdYlBu"))(30)[c(1:20, rep(c(21:30), each=10))]
+
 outputdir <- "Output/Images"
 
-png(paste0(outputdir,"CDens_2010.png"), width = 1280, height = 720, res = 108)
-luplot::plotmap(cropland_dens[,"y2010",], legend_range = c(0,300), legend_colours = cols_dens , main="Soil Organic Carbon Stocks under Cropland")
-dev.off()
+world         <- ne_countries(scale = "medium", returnclass = "sf")
+magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
+share2plot    <- merge(as.data.frame(cropland_share), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+increase2plot <- merge(as.data.frame(cropland_increase), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+density2plot  <- merge(as.data.frame(cropland_dens), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)
+world.robin   <- st_transform(world,"+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
 
-# png(paste0(outputdir,"CShare_1965.png"),width = 1280, height = 720, res = 108)
-# luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
+SOCshare <- ggplot(data = world) + geom_sf() +
+    geom_tile(data = subset(share2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits=c(0,2), colours = cols, 
+                      name = expression("ratio of cropland SOC to potential natural SOC"), 
+                      breaks=c(0.3,0.6,0.9,1.1,1.4,1.7,2),
+                      labels=c("0     0.3","  0.6", "  0.9","  1.1","  1.4","  1.7","> 2"), na.value = "white",
+                      guide = guide_legend(
+                        title.theme=element_text(size=10),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = T,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          )
+
+SOCincrease <- ggplot(data = world) + geom_sf() +
+    geom_tile(data = subset(increase2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
+                      name = expression("cropland SOC change in tons of carbon per hectar"), 
+                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
+                      breaks=c(-50,-20,-10,10,20,50,100),
+                      labels=c("> -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
+                      guide = guide_legend(
+                        title.theme=element_text(size=10),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = T,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          )
+
+cols <- colorRampPalette(brewer.pal(9, "YlGn"))(8)[2:8]
+display.brewer.pal(9, "YlGn")
+display.brewer.pal(9, "RdYlGn")
+
+SOCdensity <- ggplot(data = world) + geom_sf() +
+    geom_tile(data = subset(density2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits=c(0,150), colours = cols, 
+                      name = expression("cropland SOC in tons of carbon per hectar"), 
+                      values=c(0,5,10,20,40,70,100,150)/150,
+                      breaks=c(5,10,20,40,70,100,150),
+                      labels=c("0     5","  10","  20","  40","  70","  100"," >150"), na.value = "white",
+                      guide = guide_legend(
+                        title.theme=element_text(size=10),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = T,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          )
+
+# png(paste0(outputdir,"CDens_2010.png"), width = 1280, height = 720, res = 108)
+# luplot::plotmap2(cropland_dens[,"y2010",], legend_range = c(0,300), lowcol = cols[1], midcol = cols[4], highcol = cols[7], ncol=7, legend_discrete = TRUE)
 # dev.off()
-
-png(paste0(outputdir,"CShare_2010.png"), width = 1280, height = 720, res = 108)
-luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
-dev.off()
-
-# png(paste0(outputdir,"CIncrease_1965.png"),width = 1280, height = 720, res = 108)
-# luplot::plotmap(cropland_increase[,"y1965",], legend_range = c(-80,80), legend_colours = cols, main="Absoulte C change on Cropland 1965")
+# 
+# # png(paste0(outputdir,"CShare_1965.png"),width = 1280, height = 720, res = 108)
+# # luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
+# # dev.off()
+# 
+# png(paste0(outputdir,"CShare_2010.png"), width = 1280, height = 720, res = 108)
+# luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
 # dev.off()
-
-png(paste0(outputdir,"CIncrease_2010.png"),width = 1280, height = 720, res = 108)
-luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols, legend_title = "tC/ha", main="Absoulte C change on Cropland 2010")
-dev.off()
+# 
+# # png(paste0(outputdir,"CIncrease_1965.png"),width = 1280, height = 720, res = 108)
+# # luplot::plotmap(cropland_increase[,"y1965",], legend_range = c(-80,80), legend_colours = cols, main="Absoulte C change on Cropland 1965")
+# # dev.off()
+# 
+# png(paste0(outputdir,"CIncrease_2010.png"),width = 1280, height = 720, res = 108)
+# luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols, legend_title = "tC/ha", main="Absoulte C change on Cropland 2010")
+# dev.off()
 
@@
 
-par(mfrow = c(2, 2))
-
-luplot::plotmap(cropland_dens[,"y2010",], legend_range = c(0,300), legend_colours = cols_dens , 
-                main="SOC Stocks under Cropland")
-plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission/1000, type="l", xlab="Time",
-     ylab="Global SOC Debt in GtC")
-luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols)
-luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols,
-                legend_title = "tC/ha")
+grid <- grid.arrange(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2)
+grid
+ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid)
 

### 2020-06-29 SOCbudgetPaper/introduction.Rmd
Update SOCbudgetPaper/introduction.Rmd

@@
 
-The depletion of soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
-Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
-Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+The depletion of soil organic carbon (SOC) due to agricultural management has been <!--still is, no?
+try to avoid passive tense and if possible just use "is a major source" here, if possible--> a major source of carbon emissions, however, SOC has <!-- may have?
+in the next paragraph you say the extent is debated - only the total value?--> the potential, given XXX, to instead act as a large carbon sink and thus mitigate climate change <!-- maybe citation needed here --> .
+
+<!-- the way this sentence was written originally, "the depletion..., but also offers..." makes it look like it is the depletion which offers the potential.
+You also had the word additional,but it is a bit unclear because we don't know additional to what?
+for the XXX, i am thinking to put "given alternative/better/reduced management (regimes)", to -->
+
+Whereas the extent of the sink potential by SOC enhancement is debated (add refs), <!-- what is SOC enhancement?
+unclear to me if the potential is the SOC sink, or the sink potential is the enhancement, which sounds like particular techniques or policies?
+if former, i would write  "...the extent of the sink potential of SOC (no enhancement)" --> it is well established that the SOC pool itself is the biggest terrestrial carbon pool, exceeding that of the atmospheric and biospheric carbon pools multiple times.
+Even small changes in drivers of SOC may thus lead to substantial shifts in the amount of earth carbon cycled and influence the amount of CO2 in the atmosphere  (ref.
 permafrost melting).
-The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
 
-Mapping the worlds SOC has been investigated intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
-Natural properties like climatic, biophysical and landscape characteristics clearly play the most important role in this regard.
+<!-- Reading this first paragraph, the first sentence is introducing ag management and more "flows" of SOC, whereas the 2nd sentence is talking about the "stock".
+After reading this paragraph, I am not clear if intention is to measure SOC flows - and especially measuring the climate impacts/mitigation of ag practices - especially since you say "enhancement", or rather measuring the pool in total, which the last sentence is more saying...
+and which I understand to be true in the next parts (along with flows).
+I would put the second part first...and this way the more certain thing is also stated first.
+something like,,, :
+
+"Soil Organic Carbon, the amount of carbon stored (and cycled?) through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and biospheric pools multiple times (cite).
+As such, even small changes in drivers of SOC may thus lead to substantial shifts in the amount of earth carbon cycled and influence the amount of CO2 in the atmosphere  (ref.
+permafrost melting).
+The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+The depletion of soil organic carbon (SOC) due to agricultural management has been (is/has been) a major source of carbon emissions, however, SOC has/may also have the potential, given XXX, to instead act as a large carbon sink and thus mitigate climate change <!-- maybe citation needed here  -->
+
+The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
+Natural properties like climatic, biophysical and landscape characteristics clearly play some of the most important roles in this regard.
+
+<!-- multiple properties play multiple roleS, and i would switch it to "some of the most important roles" so that it flows into the importance of Human intervention in next line, esp.
+when its the dominant driver on mnged land" -->
+
+<!-- Oxford comma is illegal in german and optional in english but i personally really like it for clarity when separating lists -->
 
 Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change on managed land today.
-Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
-Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
-Despite providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management.
+Recent studies identify the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC <!--look up --> (Lal, 2006).
+Other studies have focused more closely on spatially disaggregated changes SOC via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics  (<!-- find microbial study -->).
+While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, (including agricultural activities?).
 
 Field-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on yield levels, fertilizer inputs and various other on farming activities.
-Yet, due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
+Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm activities.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
 
-Our study combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
+Our study combines an spatially explicit estimate of agricultural management data on the global level with a robust but light-weight SOC model in order to estimate SOC stocks and stock change factors, as well as organic carbon flow dynamics within the agricultural system.
 We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation and tillage.
@@
 
-This paper will introduce in the the first part of the method section the basic concept of SOC dynamics.
-We continue in the second part with a detailed description of the used global gridded management data on production levels, residue recycling rates, manure amendments, irrigation and tillage adoption.
-In a third, small part, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-In our result section we will focus on the SOC dynamics on global croplands by analysing (1) the spatial explicit SOC distribution and depletion over the last decades, (2) compare the climate zone specfic stock changes with default stock changes factors from the IPCC Tier 1 approach and (3) compare global agricultrual carbon flows and stocks to contextualize the importants of the different management aspects.
-Finally, we will discuss our findings and their implications on SOC model development and conlude with an analysis and outlook on the the opportunities for SOC management for climate change mitigation and negative emissions.
+<!--  To me, spatially explicit is much more natural than spatial explicit, when its used as an adjective. Google scholar shows me way more results for "spatially explicit" also. -->)
+
+<!-- "robust but light-weight" why makes it robust?
+lightweight?
+the "heavyness" of the model to me seems not so relevant here, maybe an advantage to be discussed in discussion.
+I would almost definitely take it out here in intro, and maybe "robust" as well (to me it makes sense to say the model is robust because we did so and so tests, in the discussion) -->
+
+This paper will first introduce in the methods section a presentation of SOC dynamics.
+We continue with a detailed description of the global gridded management data used here, including (crop??) production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
+Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC over the last decades <!-- maybe put explicitly which years -->, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) comparing global agricultural carbon flows and stocks to contextualize the importance of various management aspects.
+
+<!-- "basic concept of SOC dynamics" already exists right?
+So i would say it just presents them, otherwise sounds like you've invented the concept.
+Otherwise would write something like, "the concept of SOC dynamics as used in this paper" or something like that...
+-->
+
+<!-- for 3  not clear what is being compared.
+right now it sounds like comparing flows vs.
+stocks (within global value) since there is only one global value.
+Or are you comparing across regions/management practices ?
+-->
+
+Finally, we will discuss our findings - including their implications for SOC model development - and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+
+<!-- I would not use the term analyse when already in conclusion -->
 

### 2020-06-29 SOCbudgetPaper/methods.Rmd
Update SOCbudgetPaper/methods.Rmd

@@
 
-We compiled calculations as open-source R packages available under github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on github.com/pik-piam/madrat, a framework improving reproducibility and transparency in data processing.
+We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
 In the following chapter we outine the most important relationships and assumptions. See table \@ref(append:subsection2mrfunctions) for further information on corresponding code within the R packages.
@@
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
-We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
+4 (@ipcc_2019_2019), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree <!-- (55km at the equator) put this in for clarification?
+--> resolution from 1975 to 2010.
+We assume the actual SOC state converges towards a stable steady state, which itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
 Therefore we conduct the following three steps within each yearly timestep:
-(1) We calculate annual land-use (sub-)type specific steady states and decay rates for SOC stocks, 
+(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks, 
 (2) We account for land conversion by transferring SOC between land-use types and 
 (3) We estimate SOC stocks based on the stocks of the previous time period, the steady state stocks and the decay rate.
+
+<!-- maybe make it more clear what the steady state is ?
+what does actual (actual SOC) mean here ?
+what does actual mean?
+quantity of total carbon?
+so far not so clear to me, and its a bit confusing when you say its still changing over time.
+i think these lines need to be more clear, something like "which still includes changing internal dynamics over space and time depending on biophysical..." -->
 
@@
 \end{equation}
-with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cells indecies and $t$ years.
+with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cells indices and $t$ years.
 We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
-Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are still the key components to determining steady-state SOC stocks.
+4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and interactions between them.
+Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are the key components to determining steady-state SOC stocks.
 
@@
 
-We account for different carbon input sources depending on our two distinguished land-use types (see table \@ref(tab:datasourceinputs)); in particular for recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)) on croplands; and litterfall (\ref(@LPJmL4_1)) on natural vegetation.
+We account for different carbon input sources depending on the two land-use types (croplands and natural vegetation) we distinguish (see table \@ref(tab:datasourceinputs)).
+Sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); For natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the (main/only?) source.
 
-Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
+<!-- i took out in particular because this means there are more than the 3 you list?
+didn't look at the list though.
+If the 3 you listed here for croplands are not exhaustive, I would rather write "Important sources for cropland are ..., while litterfall dominates for nat veg") -->
+
+Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
 For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
+This implies that both the amount of carbon as well as its structural composition determine the effective inflow.
 Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \@ref(tab:datasourceinputs).

### 2020-06-30 SOCbudgetPaper/methods.Rmd
Update SOCbudgetPaper/methods.Rmd

@@
 \end{equation}
-with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cells indices and $t$ years.
-We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and interactions between them.
+with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cell indices and $t$ years.
+We use for our calculations the steady-state method from the refinement of the IPCC guidelines vol.
+4 (@ipcc_2019_2019) for mineral soils, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
 Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are the key components to determining steady-state SOC stocks.
@@
 
-We account for different carbon input sources depending on the two land-use types (croplands and natural vegetation) we distinguish (see table \@ref(tab:datasourceinputs)).
-Sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); For natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the (main/only?) source.
+We account for different carbon input sources depending on the two land-use types we distinguish, croplands and natural vegetation (see table \@ref(tab:datasourceinputs)).
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); For natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the (main/only) source of carbon.
 
-<!-- i took out in particular because this means there are more than the 3 you list?
-didn't look at the list though.
-If the 3 you listed here for croplands are not exhaustive, I would rather write "Important sources for cropland are ..., while litterfall dominates for nat veg") -->
+<!-- i took out in particular because this means there are more than the 3 you list? you could say "C sources for cropland are namely..." instead of in particular, but its a bit redundant here  -->
 
 Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
-For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
+For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
 This implies that both the amount of carbon as well as its structural composition determine the effective inflow.
@@
 Following the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
+4 (@ipcc_2019_2019) for mineral soils <!-- Can you shorten this because you already said it word-for-word in previous paragraph?
+--> we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
 Thus $k_{sub}$ is given by
@@
 
-For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas on natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as considered effects for each land-use types are shown in table \@ref(tab:datasourcedecay).
+For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as drivers considered for each land-use types are shown in table \@ref(tab:datasourcedecay).
 To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
@@
 
-Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
+Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use type (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
 
@@
 
-Reformulated equation \ref(eq:steadystate), we obtain a massbalance equation as follows
+Reformulating equation \ref(eq:steadystate), we obtain a massbalance equation as follows
 
@@
 
-To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initilzation year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency for the initial conditions will decrease.
+To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions will decrease.
+
+<!-- do u test the assumption of 15 year spin up not influencing results? I feel like this si something people may critique -->
 
@@
 
-## Agricultural management data on 0.5 degrees	(50) {#sec:agrimanagement}
+## Agricultural management data on 0.5 degree resolution	(50) {#sec:agrimanagement}
 
-We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
-The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis together with the accounting for a spin up phase to time span from 1975 to 2010.
+We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and consistent data suite.
+<!-- I don't know what you are trying to say with "...to a comprehensive and consistent data suite"    into a harmonized and consistent data set?
+comprehensive in what way?
+the term data suite sounds very odd to me --> The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
 For all the following data, if not declared differently, we interpolate values linearly between the time steps.
@@
 Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop group specific half-degree production patterns.
-We calibrate cellular yields with one country-level calibration factor for each crop group to meet historical FAOSTAT production ([@FAOSTAT]).
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production ([@FAOSTAT]).
 Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
@@
 #### Crop Residue Production
-Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
-Starting from crop production ($CP$) estimates of the harvested organs and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
+Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling.
+Starting from crop production ($CP$) estimates of the harvested organs <!-- organs very odd here, and I'm not sure what the unit being measured/estimated is.
+Is it just crop yields?
+"Starting from respective crop yields ($CP$) and harvested area ($CA$)..." maybe simplify something like that  --> and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
 
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction ([@ipcc_2006_2006]).
-Note that deviating from [@bodirsky2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction ([@ipcc_2006_2006]).
+Note that, deviating from [@bodirsky2012], we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
 We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
@@
 ##### Residue Usage
-We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see \@ref(append:Table_kcr2kres)).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see [@weindl]) taking available AGR biomass as well as livestock productivity into account.
-We estimate a material-use share for the straw residues group of 5% and a fuel-share of 10% for all used residues groups in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
-For high-income  ($>\,10000\,\tfrac{USD}{yr}$ no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Table_kcr2kres)).
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account.
+We estimate, for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr}$) <!-- specify per capita -->,  a material-use share for the straw residues group <!-- is straw residues a group?
+--> of 5% and a fuel-share of 10% for all used residues groups <!-- does this mean "10%  of all residues used as fuel in low income.." ?--> in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
+For high-income countries, ($>\,10000\,\tfrac{USD}{yr}$ no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
 The remaining AGR as well as all BGR are assumend to be recycled to the soil.
-We cut high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to corrected for outliers.
+We limit high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
 
@@
 #### Livestock Distribution
-To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert] and uses feed basket assumptions based on a revised methodology of @weindl.We differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
-For ruminants, we assume that livestock is located at the production spot sof the feed intake.
-We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feed stuff, which we consider to be consumend in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
+To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, inspired by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl.
+We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
+<!-- "inspired by" sounds vague to me and personally I would never write it.
+"loosely based on" or "drawing from" is what I would say...
+-->
+For ruminants, we assume that livestock is located where the production of feed occurs.
+We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
 For low-income countries according to worldbank definitions (<1000 USD/yr), we assume extensive production systems.
-We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
-Intensive production is distributed within a country using again crop-based feed production share, assuming that feed availability is the most driving factor for livestock location.
+We locate them according to the share of built-up areas based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio <!-- after "based on" it is not clear to me.
+THe animals are held in EITHER households, OR subsistence and small holders?
+which of these has the high labour per animal ratio?
+--> Intensive production is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.
+<!-- what happens with extensive/intensive in higher income countries is missing or unclear to me -->
 
 #### Manure Excretion, Storage and Recycling
-Manure production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Manure production and management is based on a revised methodology of ([@bodirsky2012]) and is presented here due to its central role in soil carbon modelling.
 Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets ([@weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
@@
 Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
-Whereas for low-income countries according to worldbank definitions (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
-For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses. 
+Whereas for low-income countries according to worldbank definitions (<1000 USD/yr) <!-- I feel like you don't ahve to say "according to world bank...
+" after the first time -->, we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$; middle-income countries see linearly interpolated shares depending on their per-capita income.
+For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
 To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
-We cut high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to corrected for outliers.
+For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
+We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
 
@@
 The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
-We apply aggregated area shares, leading to the same impact of irrigation of all our crop groups.
-Moreover we assume the irrigation effect to te present for all month of a year, that has been marked as irrigated.
+We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
+Moreover we assume the irrigation effect to be present for all months of a year, that has been marked as irrigated.
+<!-- assuming that irrigation effect happens 12 months?
+or just the months "marked as irrigated" ?
+this term confusing -->
 
@@
 
-To account for tillage distribution between the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we use the simple assumptions that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rules are inspired by the approach of [@porvolik], but neglects the less common adoption of no tillage farming on agricultural land.
+To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
+This rule is based on the approach of [@porvolik], but neglects the (less common) adoption of no-tillage farming on agricultural land.
 
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis we us the default SCF as comparison and constistency check for our more detailed budget approach.
+For this analysis we us the default SCF as a comparison and constistency check for our more detailed budget approach.
 

### 2020-07-01 SOCbudgetPaper/introduction.Rmd
Update introduction.Rmd by kristine as merge of thoughts

--- 2020-06-26 SOCbudgetPaper/introduction.Rmd
@@
 
-The depletion of soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
-Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
-Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and biospheric pools multiple times (Bathjes).
+As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere  (ref.
 permafrost melting).
-The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
+Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles in this regard.
 
-Mapping the worlds SOC has been investigated intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
-Natural properties like climatic, biophysical and landscape characteristics clearly play the most important role in this regard.
-
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change on managed land today.
-Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006).
-Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->).
-Despite providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management.
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today.
+Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC <!--look up --> (Lal, 2006).
+Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics  (Hararuk, 20XX).
+While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especial agricultural activities.
 
 Field-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on yield levels, fertilizer inputs and various other on farming activities.
-Yet, due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
+Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm activities.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
 
-Our study combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
-We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation and tillage.
-We thereby provide the first global, spatial explicit SOC maps that consider agricultural management.
+Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
+This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
+We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation, and tillage.
+We thereby provide the first global, spatially explicit SOC loss maps that consider detailed agricultural management.
 
-This paper will introduce in the the first part of the method section the basic concept of SOC dynamics.
-We continue in the second part with a detailed description of the used global gridded management data on production levels, residue recycling rates, manure amendments, irrigation and tillage adoption.
-In a third, small part, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-In our result section we will focus on the SOC dynamics on global croplands by analysing (1) the spatial explicit SOC distribution and depletion over the last decades, (2) compare the climate zone specfic stock changes with default stock changes factors from the IPCC Tier 1 approach and (3) compare global agricultrual carbon flows and stocks to contextualize the importants of the different management aspects.
-Finally, we will discuss our findings and their implications on SOC model development and conlude with an analysis and outlook on the the opportunities for SOC management for climate change mitigation and negative emissions.
+This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
+We continue with a detailed description of the global gridded management data used here, including (crop??) production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
+Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
 
-<!--
-* sanderman
-* WISE
-* SoilGrids
-* Stoorvogel
-* HANPP
+In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
 
-* Smith/Powlson
-* Lal
-* Bathjes
--->
+Finally, we will discuss our findings - including their implications for SOC model development - and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.

### 2020-07-01 SOCbudgetPaper/methods.Rmd
Update methods.Rmd from kristine first try of merging in davids changes

@@
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree <!-- (55km at the equator) put this in for clarification?
---> resolution from 1975 to 2010.
-We assume the actual SOC state converges towards a stable steady state, which itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
+4 (@ipcc_2019_2019; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
+We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
 Therefore we conduct the following three steps within each yearly timestep:
@@
 (3) We estimate SOC stocks based on the stocks of the previous time period, the steady state stocks and the decay rate.
-
-<!-- maybe make it more clear what the steady state is ?
-what does actual (actual SOC) mean here ?
-what does actual mean?
-quantity of total carbon?
-so far not so clear to me, and its a bit confusing when you say its still changing over time.
-i think these lines need to be more clear, something like "which still includes changing internal dynamics over space and time depending on biophysical..." -->
 
@@
 with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cell indices and $t$ years.
-We use for our calculations the steady-state method from the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
+We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
 Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are the key components to determining steady-state SOC stocks.
@@
 We account for different carbon input sources depending on the two land-use types we distinguish, croplands and natural vegetation (see table \@ref(tab:datasourceinputs)).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); For natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the (main/only) source of carbon.
-
-<!-- i took out in particular because this means there are more than the 3 you list? you could say "C sources for cropland are namely..." instead of in particular, but its a bit redundant here  -->
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); for natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the only source of carbon.
 
@@
 The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
-Following the steady-state method of the refinement of the IPCC guidelines vol.
-4 (@ipcc_2019_2019) for mineral soils <!-- Can you shorten this because you already said it word-for-word in previous paragraph?
---> we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
+Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
 Thus $k_{sub}$ is given by
@@
 
-<!-- do u test the assumption of 15 year spin up not influencing results? I feel like this si something people may critique -->
-
 \newpage
@@
 
-We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and consistent data suite.
-<!-- I don't know what you are trying to say with "...to a comprehensive and consistent data suite"    into a harmonized and consistent data set?
-comprehensive in what way?
-the term data suite sounds very odd to me --> The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps.
+We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a harmonized and consistent data set.
+The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
 
@@
 Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account.
-We estimate, for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr}$) <!-- specify per capita -->,  a material-use share for the straw residues group <!-- is straw residues a group?
---> of 5% and a fuel-share of 10% for all used residues groups <!-- does this mean "10%  of all residues used as fuel in low income.." ?--> in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
+We estimate, for low-income countries ($<\,1000\,\tfrac{USD}{yr}$) per capita,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low income countries ($<\,1000\,\tfrac{USD}{yr}$).
 For high-income countries, ($>\,10000\,\tfrac{USD}{yr}$ no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
@@
 #### Livestock Distribution
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, inspired by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl.
+To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl.
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-<!-- "inspired by" sounds vague to me and personally I would never write it.
-"loosely based on" or "drawing from" is what I would say...
--->
 For ruminants, we assume that livestock is located where the production of feed occurs.
@@
 For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
-For low-income countries according to worldbank definitions (<1000 USD/yr), we assume extensive production systems.
-We locate them according to the share of built-up areas based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio <!-- after "based on" it is not clear to me.
-THe animals are held in EITHER households, OR subsistence and small holders?
-which of these has the high labour per animal ratio?
---> Intensive production is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.
-<!-- what happens with extensive/intensive in higher income countries is missing or unclear to me -->
+For low-income countries (<1000 USD/yr), we assume extensive production systems.
+We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio.
+Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.
 
@@
 Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
-Whereas for low-income countries according to worldbank definitions (<1000 USD/yr) <!-- I feel like you don't ahve to say "according to world bank...
-" after the first time -->, we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$; middle-income countries see linearly interpolated shares depending on their per-capita income.
+Whereas for low-income countries (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$; middle-income countries see linearly interpolated shares depending on their per-capita income.
 For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 

### 2020-07-14 SOCbudgetPaper/methods.Rmd

@@
 Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling.
-Starting from crop production ($CP$) estimates of the harvested organs <!-- organs very odd here, and I'm not sure what the unit being measured/estimated is.
-Is it just crop yields?
-"Starting from respective crop yields ($CP$) and harvested area ($CA$)..." maybe simplify something like that  --> and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
+Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
 
@@
 A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
-Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
 Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).
@@
 Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account.
-We estimate, for low-income countries ($<\,1000\,\tfrac{USD}{yr}$) per capita,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low income countries ($<\,1000\,\tfrac{USD}{yr}$).
-For high-income countries, ($>\,10000\,\tfrac{USD}{yr}$ no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
+We estimate, for low-income countries,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low-income countries.
+For high-income countries, no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
 The remaining AGR as well as all BGR are assumend to be recycled to the soil.
@@
 For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
-For low-income countries (<1000 USD/yr), we assume extensive production systems.
+For low-income countries, we assume extensive production systems.
 We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio.
@@
 Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
-Whereas for low-income countries (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$; middle-income countries see linearly interpolated shares depending on their per-capita income.
+Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
 For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
@@
 We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
-Moreover we assume the irrigation effect to be present for all months of a year, that has been marked as irrigated.
-<!-- assuming that irrigation effect happens 12 months?
-or just the months "marked as irrigated" ?
-this term confusing -->
+Moreover we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
 

### 2020-07-14 SOCbudgetPaper/SOCbudget.Rmd
small update on written paper part

--- 2020-03-19 SOCbudgetPaper/SOCbudget.Rmd
@@
 ---
-title: Historical Soil Organic Carbon Budget
+title: Management induced changes of soil organic carbon on global croplands 
+       \newline or \newline
+       How management changed soil organic carbon on global croplands
+       \newline or \newline
+       Agricultural soil have lost 13 GtC topsoil carbon
+       \newline or \newline
+       Linking agricultural management data to soil modeleling – a global approach
+
 journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
@@
     affiliation: 1
-    email: kristine.karstenst@pik-potsdam.de
+    email: kristine.karstens@pik-potsdam.de
     corresponding: true
@@
     affiliation: 1
+  - given_name: Your
+    surname: Name
+    affiliation: 2
   - given_name: Alexander
@@
     address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany
+  - code: 2
+    address: Your affiliation 
 abstract: | 
- SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
- this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
- models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these 
- biophysical processes are still little investigated.
- Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and 
- land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
- We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
- Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from … . We also find that SOC is very 
- sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models. 
+  Soil organic carbon (SOC) is one of larges c stocks on earth (3 times larger biosphere pool). 
 bibliography: SOCbudget.bib
@@
 authorcontribution: |
-  Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.
+  Karstens wrote code and paper build on work of Bodirsky. Bodirsky, <here_could_be_your_names> and Popp revised paper.
 disclaimer: |
@@
 output:
-  rticles::copernicus_article: default
-  bookdown::word_document2: default
-  md_document:
-    variant: markdown_github
   bookdown::pdf_book:
@@
     md_extensions: +raw_attribute
+  md_document:
+    variant: markdown_github
+  bookdown::word_document2: default
+  bookdown::pdf_document2: default
+  rticles::copernicus_article: default
 ---

### 2020-07-14 SOCbudgetPaper/SOCbudget.tex
small update on written paper part

@@
 
-\title{Historical Soil Organic Carbon Budget}
+\title{Management induced changes of soil organic carbon on global croplands \newline or \newline How management changed soil organic carbon on global croplands \newline or \newline Agricultural soil have lost 13 GtC topsoil carbon \newline or \newline Linking agricultural management data to soil modeleling -- a global approach}
 
@@
 \Author[1]{Benjamin Leon}{Bodirsky}
+\Author[2]{Your}{Name}
 \Author[1]{Alexander}{Popp}
 
-\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam,
-Germany}
+\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
+\affil[2]{Your affiliation}
 
@@
 
-\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}
+\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}
 
@@
 \begin{abstract}
-SOC one of larges c sinks on earth (3 times larger biosphere pool).
-Agricultural management leads to a depletion of soil organic carbon.
-However this depletion of soil organic carbon (SOC) pools are so far not
-well represented in global assessments of historic carbon emissions.
-While SOC models often represent well the biochemical processes that
-lead to the accumulation and decay of SOC, the management decisions
-driving these biophysical processes are still little investigated. Here
-we create a spatial explicit data set for crop residue and manure
-management on cropland based on global historic production (FAOSTAT) and
-land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
-create a half degree resolution soil organic carbon budget on mineral
-soils. We estimate that due to arable farming soils have lost over (?)
-GtOC of which (??) GtOC have been released within the period 1990-2010.
-We show that, our results on global scale based on Tier 2 IPCC
-methodolgy are in good agreement with Tier 1 default assumptions. We
-also find that SOC is very sensitive to management decision such as
-residue recycling indicating the nessessity to incorporated better
-management data in soil models.
+Soil organic carbon (SOC) is one of larges c stocks on earth (3 times larger biosphere pool).
 \end{abstract}
 
-\copyrightstatement{The author's copyright for this publication is
-transferred to institution/company.}
+\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
 
@@
 
-The depletion of soil organic carbon (SOC) due agricultural management
-has been a major source of carbon emissions, but also offers the
-potential to turn into a large additional carbon sink and thus mitigate
-climate change. Whereas the extent of the sink potential by SOC
-enhancement is highly debated (add refs), it is largely agreed upon that
-the SOC pool itself is the biggest terrestrial carbon pool, exceeding
-the atmospheric and even more the biospheric carbon pool multiple times.
-Even small changes in SOC drivers might lead to substantial shifts in
-earth carbon cycled and influence the atmospheric CO2 concentration
-(ref. permafrost melting). The specific amount of carbon stored in the
-soil is uncertain though and estimates ranging from 1500 to 2400 GtC for
-the first meter of the soil profile (Bathjes, 1996).
+The depletion of soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
+Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
+Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+permafrost melting).
+The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
 
-Mapping the worlds SOC has been investigated intensivly over the last
-decades and led to an increased quality of SOC maps as well as a even
-better understanding of factors driving magnitude, distribution and
-dynamics of SOC pools. Natural properties like climatic, biophysical and
-landscape characteristics clearly play the most important role in this
-regard.
+Mapping the worlds SOC has been investigated intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
+Natural properties like climatic, biophysical and landscape characteristics clearly play the most important role in this regard.
 
-Human intervention, including land cover change and land management, has
-however added a further driver to SOC change, which alters terrestrial
-carbon pools in much shorter time scales and is likely the dominant
-driver of SOC change on managed land today. Recent studies identified
-the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which
-compares to older estimates of around 60-130 GtC (Lal, 2006). Other
-studies have focused more closely on spatial disaggregated SOC changes
-via advanced digital soil mapping techniques (S-World; Stoorvogel 2,
-2017) or better representation of biogeochemcial processes within SOC
-dynamics (). Despite providing a rough estimate of the order of
-magnitude of change, these studies lack a detailed consideration of land
-management.
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change on managed land today.
+Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC (Lal, 2006).
+Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics ().
+Despite providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management.
 
-Field-scale models (ref. Daycent, RothC, Ecosse, C-Tool) are able to
-capture these land mangement impacts by using detailed information on
-yield levels, fertilizer inputs and various other on farming activities.
-Yet, due to the lack of comprehensive global management data to feed
-these input demanding models, it seems still very complex to scale them
-up to a global extent.
+Field-scale models (ref.
+Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on yield levels, fertilizer inputs and various other on farming activities.
+Yet, due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
 
-Our study combines an spatial-explicit estimate of agricultural
-management data on global level with a robust but light weighted SOC
-model to estimate SOC stocks and stock change factors as well as organic
-carbon flow dynamics within the agricultural system. We thereby consider
-change in SOC caused by historical land cover change as well as of
-different agricultural mangement practices, including residue recycling,
-manure amendments, irrigation and tillage. We thereby provide the first
-global, spatial explicit SOC maps that consider agricultural management.
+Our study combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
+We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation and tillage.
+We thereby provide the first global, spatial explicit SOC maps that consider agricultural management.
 
-This paper will introduce in the the first part of the method section
-the basic concept of SOC dynamics. We continue in the second part with a
-detailed description of the used global gridded management data on
-production levels, residue recycling rates, manure amendments,
-irrigation and tillage adoption. In a third, small part, we shortly
-refer to the concept of stock change factors as outlined in the Tier 1
-approach of the IPCC guidelines. In our result section we will focus on
-the SOC dynamics on global croplands by analysing (1) the spatial
-explicit SOC distribution and depletion over the last decades, (2)
-compare the climate zone specfic stock changes with default stock
-changes factors from the IPCC Tier 1 approach and (3) compare global
-agricultrual carbon flows and stocks to contextualize the importants of
-the different management aspects. Finally, we will discuss our findings
-and their implications on SOC model development and conlude with an
-analysis and outlook on the the opportunities for SOC management for
-climate change mitigation and negative emissions.
+This paper will introduce in the the first part of the method section the basic concept of SOC dynamics.
+We continue in the second part with a detailed description of the used global gridded management data on production levels, residue recycling rates, manure amendments, irrigation and tillage adoption.
+In a third, small part, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+In our result section we will focus on the SOC dynamics on global croplands by analysing (1) the spatial explicit SOC distribution and depletion over the last decades, (2) compare the climate zone specfic stock changes with default stock changes factors from the IPCC Tier 1 approach and (3) compare global agricultrual carbon flows and stocks to contextualize the importants of the different management aspects.
+Finally, we will discuss our findings and their implications on SOC model development and conlude with an analysis and outlook on the the opportunities for SOC management for climate change mitigation and negative emissions.
 
@@
 
-\section{Method (50)}
+\hypertarget{method-50}{%
+\section{Method (50)}\label{method-50}}
 
-We compiled calculations as open-source R packages available under
-github.com/pik-piam/mrcommons (management related functions) and
-github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which
-are both based on github.com/pik-piam/madrat, a framework improving
-reproducibility and transparency in data processing. In the following
-chapter we outine the most important relationships and assumptions. See
-table @ref(append:subsection2mrfunctions) for further information on
-corresponding code within the R packages.
+We compiled calculations as open-source R packages available under github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on github.com/pik-piam/madrat, a framework improving reproducibility and transparency in data processing.
+In the following chapter we outine the most important relationships and assumptions. See table \ref{append:subsection2mrfunctions} for further information on corresponding code within the R packages.
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{Carbon Stocks following (new) Tier 2 method
-(50)}\label{sec:carbonbudget}}
+\subsection{Carbon Stocks following (new) Tier 2 method (50)}\label{sec:carbonbudget}}
 
-Following the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 (\citet{ipcc_2019_2019}), we estimate soil organic carbon (SOC)
-stocks for cropland and all other land represented by natural vegetation
-on half-degree resolution from 1975 to 2010. We assume the actual SOC
-state converges towards a stable steady state, that itself is changing
-over time and space depending on biophysical, climatic and agronomic
-conditions. Therefore we conduct the following three steps within each
-yearly timestep: (1) We calculate annual land-use (sub-)type specific
-steady states and decay rates for SOC stocks, (2) We account for land
-conversion by transferring SOC between land-use types and (3) We
-estimate SOC stocks based on the stocks of the previous time period, the
-steady state stocks and the decay rate.
+Following the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
+We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
+Therefore we conduct the following three steps within each yearly timestep:
+(1) We calculate annual land-use (sub-)type specific steady states and decay rates for SOC stocks,
+(2) We account for land conversion by transferring SOC between land-use types and
+(3) We estimate SOC stocks based on the stocks of the previous time period, the steady state stocks and the decay rate.
 
-\subsubsection{Steady-state SOC stocks and decay rates}
+\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
+\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}
 
-In a simple first order kinetic approach the steady-state soil organic
-carbon stocks \(SOC^{eq}\) are given by \begin{equation}
+In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
+\begin{equation}
 SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
 \label{eq:inoutflow}
-\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil,
-\(k\) denoting the soil organic carbon decay rate; as well as \(i\)
-representing grid cells indecies and \(t\) years. We use for our
-calculations the steady-state method of the refinement of the IPCC
-guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which
-assume three soil carbon sub-pools (active, slow and passive) and
-entangled dynamics between them. Annual carbon inflow to each sub-pool
-(see @ref(sec:carboninputs)) and annual decay rates (see
-@ref(sec:tier2)) of each sub-pool are still the key components to
-determining steady-state SOC stocks.
+\end{equation}
+with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cells indecies and \(t\) years.
+We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
+Annual carbon inflow to each sub-pool (see \ref{sec:carboninputs}) and annual decay rates (see \ref{sec:tier2}) of each sub-pool are still the key components to determining steady-state SOC stocks.
 
-\hypertarget{sec:carboninputs}{%
-\paragraph{Carbon Inputs to the Soil}\label{sec:carboninputs}}
+\hypertarget{carbon-inputs-to-the-soil-seccarboninputs}{%
+\paragraph*{Carbon Inputs to the Soil \{\#sec:carboninputs\}}\label{carbon-inputs-to-the-soil-seccarboninputs}}
+\addcontentsline{toc}{paragraph}{Carbon Inputs to the Soil \{\#sec:carboninputs\}}
 
-We account for different carbon input sources depending on our two
-distinguished land-use types (see table @ref(tab:datasourceinputs)); in
-particular for recycled crop residues, below ground biomass of crops
-(for both see @ref(sec:residues)) and recycled manure (see
-\ref(sec:livst\_manure)) on croplands; and litterfall
-(\ref(\citet{LPJmL4_1})) on natural vegetation.
+We account for different carbon input sources depending on our two distinguished land-use types (see table \ref{tab:datasourceinputs}); in particular for recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see @ref(sec:livst\_manure)) on croplands; and litterfall (\citep{LPJmL4_1}) on natural vegetation.
 
-Following the IPCC methodology carbon inputs are disaggregated into
-metabolic and structural components depending on their lignin and
-nitrogen content (see @ref(ipcc\_2019\_2019)). For each component the
-sum over all carbon input sources is allocated to the respective SOC
-sub-pools via transfer coefficients. This implies that not only the
-amount of carbon, but also their structural composition is determining
-the effective inflow. Data sources for all considered carbon inputs as
-well as for lignin and nitrogen content can be found in table
-@ref(tab:datasourceinputs).
+Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citep{ipcc_2019_2019}).
+For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
+This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
 
@@
  \middlehline
- \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
-                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
+ \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, \\ref{sec:residues}] & default values given by [2]  \\
+                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, \\ref{sec:residues}] & default values given by [2] \\
                             & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
@@
  \end{tabular}
+ \belowtable{}
  \label{tab:datasourceinputs}
- \belowtable{}
  \end{table*}
 
-\hypertarget{sec:tier2}{%
-\paragraph{Soil Organic Carbon decay (300)}\label{sec:tier2}}
+\hypertarget{soil-organic-carbon-decay-300-sectier2}{%
+\paragraph*{Soil Organic Carbon decay (300) \{\#sec:tier2\}}\label{soil-organic-carbon-decay-300-sectier2}}
+\addcontentsline{toc}{paragraph}{Soil Organic Carbon decay (300) \{\#sec:tier2\}}
 
-The sub-pool specific decay rates are influenced by climatic conditions,
-biophysical and biochemical soil properties as well as management
-factors that all vary over time (t) and space (i). Following the
-steady-state method of the refinement of the IPCC guidelines vol.~4
-(\citet{ipcc_2019_2019}) for mineral soils we consider temperature
-(temp), water (wat), sand fraction (sf) and tillage (till) effects to
-account for spatial variation of decay rates. Thus \(k_{sub}\) is given
-by
+The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
+Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citep{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
+Thus \(k_{sub}\) is given by
 
@@
 
-For cropland we distinguish the effect of different tillage (see
-@ref(\#sec:tillage)) and irrigation (see @ref(\#sec:irrigation))
-practices on decay rates, whereas on natural vegetation, we assume
-rainfed and non-tilled conditions. Data sources as well as considered
-effects for each land-use types are shown in table
-@ref(tab:datasourcedecay). To account for variations of decay rates
-within each grid cell due to different tillage and irrigation regimes,
-average rates based on area shares are calculated.
+For cropland we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas on natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as considered effects for each land-use types are shown in table \ref{tab:datasourcedecay}.
+To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
 
@@
  \end{tabular}
+ \belowtable{}
  \label{tab:datasourcedecay}
- \belowtable{}
  \end{table*}
 
-\subsubsection{SOC transfer between land-use types}
+\hypertarget{soc-transfer-between-land-use-types}{%
+\subsubsection{SOC transfer between land-use types}\label{soc-transfer-between-land-use-types}}
 
-We calculate SOC stocks based on the area shares of land-use types (lut)
-within the half-degree grid cells (i). If land is converted from one
-land-use type into others (!lut), the respective share of the SOC stocks
-is reallocated. We account for land conversion at the beginning of each
-time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
+We calculate SOC stocks based on the area shares of land-use types (lut) within the half-degree grid cells (i).
+If land is converted from one land-use type into others (!lut), the respective share of the SOC stocks is reallocated.
+We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
 
@@
 
-with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
-expansion for a given land-use type \(lut\). Note that \(!lut\) denotes
-the sum over all other land-use types, which decreases in the specific
-time step \(t\). Data sources and methodology on land-use states and
-changes are described in @ref(sec:landuse).
+with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land-use type \(lut\).
+Note that \(!lut\) denotes the sum over all other land-use types, which decreases in the specific time step \(t\).
+Data sources and methodology on land-use states and changes are described in \ref{sec:landuse}.
 
-\subsubsection{Total SOC stocks}
+\hypertarget{total-soc-stocks}{%
+\subsubsection{Total SOC stocks}\label{total-soc-stocks}}
 
-Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
-calculated steady-state stock \(SOC^{eq}\) for each land-use types
-(lut), each sub-pool (sub) and each annual time step (t) as represented
-in equation @ref(eq:steadystate).
+Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.
 
@@
 SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
-\label{eq:steadystate}
+\label{eq:SOCstate}
 \end{equation}
 
-Reformulated equation \ref(eq:steadystate), we obtain a massbalance
-equation as follows
+Reformulated equation \eqref{eq:SOCstate}, we obtain a massbalance equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
-\label{eq:steadystate}
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \\eqref{eq:inoutflow})}}  \qquad \forall\quad lut, sub, i.
+\label{eq:steadystate2budget}
 \end{equation}
@@
 
-\subsubsection{Initialisation of SOC pools}
+\hypertarget{initialisation-of-soc-pools}{%
+\subsubsection{Initialisation of SOC pools}\label{initialisation-of-soc-pools}}
 
-To initialize all SOC sub-pools we assume that cropped land as well as
-natural vegetation are in a steady state for the specific configuration
-present within the initilzation year 1961. We assume after a spin-up
-period of 15 years reliable results start from 1975 and improve over
-time, since dependency for the initial conditions will decrease.
+To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initilzation year 1961.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency for the initial conditions will decrease.
 
@@
 \hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data on 0.5 degrees
-(50)}\label{sec:agrimanagement}}
+\subsection{Agricultural management data on 0.5 degrees (50)}\label{sec:agrimanagement}}
 
-We compile country-specific FAO production and cropland statistics
-(\citep{FAOSTAT}) to a comprehensive and constistent data suite. The
-data is prepared in 5 year time steps from 1965 to 2010, which also
-restricts our analysis together with the accounting for a spin up phase
-to time span from 1975 to 2010. For all the following data, if not
-declared differently, we interpolate values linearly between the time
-steps.
+We compile country-specific FAO production and cropland statistics (\citep{FAOSTAT}) to a comprehensive and constistent data suite.
+The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis together with the accounting for a spin up phase to time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps.
 
@@
 
-Land-use patterns are based on the Land-Use Harmonization 2 (LUH2,
-\citep{LUH2}) data set, which we aggregate from quarter degree to half
-degree resolution. We disaggregate the five different cropland
-subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17
-crop groups, applying the relative shares for each gridcell based on the
-country and year specific area harvested shares of FAOSTAT data
-(\citep{FAOSTAT}) (see @ref(append:Table\_luh2fao2mag) for more details
-on the crop group mapping). Land-use transitions are calculated as net
-area differences of the land-use data on half-degree.
+Land-use patterns are based on the Land-Use Harmonization 2 (LUH2, \citep{LUH2}) data set, which we aggregate from quarter degree to half degree resolution.
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each gridcell based on the country and year specific area harvested shares of FAOSTAT data (\citep{FAOSTAT}) (see @ref(append:Table\_luh2fao2mag) for more details on the crop group mapping).
+Land-use transitions are calculated as net area differences of the land-use data on half-degree.
 
@@
 
-\paragraph{Crop Production}
+\textbf{Crop Production}
+Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns.
+We calibrate cellular yields with one country-level calibration factor for each crop group to meet historical FAOSTAT production (\citep{FAOSTAT}).
+Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
 
-Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
-half-degree cropland patterns (see @ref(\#sec:landuse)) we compile crop
-group specific half-degree production patterns. We calibrate cellular
-yields with one country-level calibration factor for each crop group to
-meet historical FAOSTAT production (\citep{FAOSTAT}). Note that by using
-physical cropland areas we account for multiple crop harvest events as
-well as for fallows.
-
-\paragraph{Crop Residue Production}
-
-Crop residue production and management is based on a revised methodology
-of (\citep{bodirsky2012}) and will be explained in key aspects again due
-to its central role for soil carbon modelling. Starting from crop
-production (\(CP\)) estimates of the harvested organs and their
-respective harvested crop area (\(CA\)), we estimate above-ground
-(\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group
-(\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios
-(\(RS\)) as follows
+\textbf{Crop Residue Production}
+Crop residue production and management is based on a revised methodology of (\citep{bodirsky2012}) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Starting from crop production (\(CP\)) estimates of the harvested organs and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios (\(RS\)) as follows
 
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield
-and an area dependend fraction (\citep{ipcc_2006_2006}). Note that
-deviating from \citep{bodirsky2012} we use harvested instead of physical
-crop area to account for increased residue biomass due to multiple
-cropping and decreased amounts on fallow land. We assume that all BGR
-are recycled to the soil, whereas AGR can be burned or harvested for
-other purposes such as feeding animals (\citep{weindl}), fuel or for
-material use.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction (\citep{ipcc_2006_2006}).
+Note that deviating from \citep{bodirsky2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals (\citep{weindl}), fuel or for material use.
 
-\subparagraph{Burned Residues}
+\textbf{Burned Residues}
+A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
+Following \citep{smil1999}) we assume 25\% burn share for low-income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process (\citep{ipcc_2006_2006}).
 
-A fixed share of the AGR is assumed to be burned on field depending on
-the per-capita income of the country. Following \citep{smil1999}) we
-assume 25\% burn share for low-income countries according to worldbank
-definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\% for high-income
-(\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate shares for all
-middle-income countries depending on their per-capita income. Depending
-on the crop group 80--90\% of the residue carbon burned on the fields
-are lost within the combustion process (\citep{ipcc_2006_2006}).
+\textbf{Residue Usage}
+We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see @ref(append:Table\_kcr2kres)).
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see \citep{weindl}) taking available AGR biomass as well as livestock productivity into account.
+We estimate a material-use share for the straw residues group of 5\% and a fuel-share of 10\% for all used residues groups in low income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)).
+For high-income (\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+The remaining AGR as well as all BGR are assumend to be recycled to the soil.
+We cut high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.
 
-\subparagraph{Residue Usage}
-
-We compile out of our 17 crop groups, three used residue groups (straw,
-high-lignin and low-lignin residues) with additional demand for other
-purposes and one residues with no double use (see
-@ref(append:Table\_kcr2kres)). Residue feed demand for five different
-livestock groups is based on country- and residue-group-specific feed
-basekts (see \citep{weindl}) taking available AGR biomass as well as
-livestock productivity into account. We estimate a material-use share
-for the straw residues group of 5\% and a fuel-share of 10\% for all
-used residues groups in low income countries according to worldbank
-definitions (\(<\,1000\,\tfrac{USD}{yr}\)). For high-income
-(\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is
-assumend, leaving middle-income countries with linearly interpolate
-shares depending on their per-capita income. The remaining AGR as well
-as all BGR are assumend to be recycled to the soil. We cut high
-recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for
-outliers.
-
-\paragraph{Dry Matter to Carbon Transformation}
-
-To transform dry matter estimates into carbon, we compiled crop-group
-and plant part specific carbon to dry matter (c:dm) ratios (see
-@ref(append:Table\_c2dm)) (\citet{Ma2018}).
+\textbf{Dry Matter to Carbon Transformation}
+To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see @ref(append:Table\_c2dm)) (\citet{Ma2018}).
 
 \hypertarget{sec:livst_manure}{%
-\subsubsection{Livestock Distribution and Manure Excretion
-(300)}\label{sec:livst_manure}}
+\subsubsection{Livestock Distribution and Manure Excretion (300)}\label{sec:livst_manure}}
 
-We assume that manure is applied at its excretion place, leaving the
-livestock distribution the driving factor for the spatial pattern of
-manuring.
+We assume that manure is applied at its excretion place, leaving the livestock distribution the driving factor for the spatial pattern of manuring.
 
-\paragraph{Livestock Distribution}
+\textbf{Livestock Distribution}
+To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of \citep{gilbert} and uses feed basket assumptions based on a revised methodology of \citet{weindl.We} differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
+For ruminants, we assume that livestock is located at the production spot sof the feed intake.
+We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feed stuff, which we consider to be consumend in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
+For low-income countries according to worldbank definitions (\textless1000 USD/yr), we assume extensive production systems.
+We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
+Intensive production is distributed within a country using again crop-based feed production share, assuming that feed availability is the most driving factor for livestock location.
 
-To disaggregate country level FAOSTAT livestock production values to
-half-degree pattern, we use the following rule based assumptions which
-were inspired by the approach of \citep{gilbert} and uses feed basket
-assumptions based on a revised methodology of \citet{weindl.We}
-differentiate ruminant and monogastric systems, as well as extensive and
-an intensive systems. For ruminants, we assume that livestock is located
-at the production spot sof the feed intake. We distingush between grazed
-pasture, which is converted into livestock products from extensive
-systems; and all other (crop-based) feed stuff, which we consider to be
-consumend in intensive systems. For poultry, egg and monogastric meat
-production we use the per-capita income of the country to divide into
-intensive and extensive production systems. For low-income countries
-according to worldbank definitions (\textless1000 USD/yr), we assume
-extensive production systems. We locate them according to built-up areas
-shares based on the idea that these animals are held in households,
-subsistence or small-holder farming systems with a high labour per
-animal ratio. Intensive production is distributed within a country using
-again crop-based feed production share, assuming that feed availability
-is the most driving factor for livestock location.
+\textbf{Manure Excretion, Storage and Recycling}
+Manure production and management is based on a revised methodology of (\citep{bodirsky2012}) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets (\citep{weindl}), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep[(][]{ipcc_2019_2019}).
+Depending on the feed system we assume manure to be handled in four different ways:
+All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
+Manure fuel shares are estimated using IPCC default values (\citet{ippc_2006_2006}).
+Whereas for low-income countries according to worldbank definitions (\textless1000 USD/yr), we adopt a share of 25\% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries (\(>\,10000\,\tfrac{USD}{yr}\), leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses.
+To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
+Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
+For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky2012} (see @ref(append:Table\_clossAWMS))).
+We cut high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.
 
-\paragraph{Manure Excretion, Storage and Recycling}
+\hypertarget{irrigation-100}{%
+\subsubsection{Irrigation (100)}\label{irrigation-100}}
 
-Manure production and management is based on a revised methodology of
-(\citep{bodirsky2012}) and will be explained in key aspects again due to
-its central role for soil carbon modelling. Based on the gridded
-livestock distribution we calculate excretions by estimating the
-nitrogen balance of the livestock system on the basis of comprehensive
-livestock feed baskets (\citep{weindl}), assuming that all nitrogen in
-protein feed intake, minus the nitrogen in the slaughter mass, is
-excreted. Carbon in excreted manure is estimated by applying fixed C:N
-ratios (given by \citep[(][]{ipcc_2019_2019}). Depending on the feed
-system we assume manure to be handled in four different ways: All manure
-orginated from pasture feed intake is excreted directly to pastures and
-rangelands (pasture grazing), deducting manure collected as fuel. Manure
-fuel shares are estimated using IPCC default values
-(\citet{ippc_2006_2006}). Whereas for low-income countries according to
-worldbank definitions (\textless1000 USD/yr), we adopt a share of 25\%
-of crop residues in feed intake directly consumend and excreted on crop
-fields (stubble grazing), we do not consider any stubble grazing in
-high-income countries (\(>\,10000\,\tfrac{USD}{yr}\), leaving
-middle-income countries with linearly interpolate shares depending on
-their per-capita income. For all other feed items we assume the manure
-to be stored in animal waste management systems associated to animal
-houses. To estimate the carbon actually recycled to the soil, we account
-for carbon losses during storage and recycling shares in different
-animal waste management and grazing systems. Whereas we assume no losses
-for pasture and stubble grazing, we consider that the manure collected
-as fuel is not recycled. For manure stored in different animal waste
-management system we compiled carbon loss rates partly depending on the
-nitrogen loss rates as specified in \citep{bodirsky2012} (see
-@ref(append:Table\_clossAWMS))). We cut high application shares at
-\(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.
+We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
+The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
+We apply aggregated area shares, leading to the same impact of irrigation of all our crop groups.
+Moreover we assume the irrigation effect to te present for all month of a year, that has been marked as irrigated.
 
-\subsubsection{Irrigation (100)}
+\hypertarget{tillage-100}{%
+\subsubsection{Tillage (100)}\label{tillage-100}}
 
-We use irrigation area shares to modify the water effect on
-decomposition by weighting the irrigated and rainfed water factors based
-on these shares. The LUH2v2 data set provides irrigated fractions for
-their cropland subcategories. We apply aggregated area shares, leading
-to the same impact of irrigation of all our crop groups. Moreover we
-assume the irrigation effect to te present for all month of a year, that
-has been marked as irrigated.
-
-\subsubsection{Tillage (100)}
-
-To account for tillage distribution between the three tillage types
-specfied by the IPCC - full tillage, reduced tillage and no tillage -,
-we use the simple assumptions that all natural land and pastures are not
-tilled, whereas annual crops are under full and perennials under reduced
-tillage. This rules are inspired by the approach of \citep{porvolik},
-but neglects the less common adoption of no tillage farming on
-agricultural land.
+To account for tillage distribution between the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we use the simple assumptions that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
+This rules are inspired by the approach of \citep{porvolik}, but neglects the less common adoption of no tillage farming on agricultural land.
 
@@
 
-Additionally to the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 (\citet{ipcc_2019_2019}) and the detailed analysis of management
-data coming with it, SOC changes can be estimated using the IPCC tier 1
-approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006},
-\citet{ipcc_2019_2019}). Here, stocks are calculated via stock change
-factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on
-observational data. SCF are differentiated by different crop and
-management systems (\(m\)) reflecting different dynamics under changed
-in- and outflows without explicitly tracking these. Moreover SCF vary
-for different climatic zones (\(c\)) specified by the IPCC (see
-@ref(append:climatemap)). The actual SOC stocks as thus calulated based
-on a given reference SOC stock by
+Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}).
+Here, stocks are calculated via stock change factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+SCF are differentiated by different crop and management systems (\(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover SCF vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}).
+The actual SOC stocks as thus calulated based on a given reference SOC stock by
 
@@
 
-with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into
-corresponding climate zones \(c\). For this analysis we us the default
-SCF as comparison and constistency check for our more detailed budget
-approach.
+with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
+For this analysis we us the default SCF as comparison and constistency check for our more detailed budget approach.
+
+According to \citep{ipcc_2019_2019} the formulation of absolute changes like
+might be neccessary as well.
 
@@
 
-\section{Results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+We present simulation results of the SOC budget of the period from 1975 to 2010 mainly focusing on cropland areas.
+Detailed total global and regional results of the state of the worlds SOC in comparison to other estimates can be found in the supplement.
+
+\hypertarget{soc-distribution-and-depletion}{%
+\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}
+
+Figure
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# All defaults}
+\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include\_graphics}\NormalTok{(}\StringTok{"../ResultNotebooks/Output/Images/4panelfigure.png"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a) ... , (b) ... , (c) ... , (d) ... .}\label{fig:SOCmaps}
+\end{figure}
+
+\newpage
+
+\hypertarget{comparison-of-stock-changes-factors}{%
+\subsection{comparison of stock changes factors}\label{comparison-of-stock-changes-factors}}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# All defaults}
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include\_graphics}\NormalTok{(}\StringTok{"../ResultNotebooks/Output/Images/4panelfigure.png"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/4panelfigure}
 \newpage
 
-\section{Discussion}
+\hypertarget{global-agricultrual-carbon-flows-and-stocks}{%
+\subsection{global agricultrual carbon flows and stocks}\label{global-agricultrual-carbon-flows-and-stocks}}
+
+\newpage
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 \item
-  natural land representation is lacking proper parametrization of
-  natural input (n, lg content of litterfall) (--\textgreater{} ask
-  Christop again)
+  natural land representation is lacking proper parametrization of natural input (n, lg content of litterfall) (--\textgreater{} ask Christop again)
 \item
-  boreal zone and dry regions are not well represented (maybe because of
-  the bad parameterization of the soil model for natural soils)
+  boreal zone and dry regions are not well represented (maybe because of the bad parameterization of the soil model for natural soils)
 \item
-  As pointed out by Keel (2017), Smith (2019) the results might be
-  highly sensitive to carbon input calculations more precisly to below
-  and above ground residue carbon estimates derived from harvested
-  quantities. It has been questioned, that below ground residues might
-  increase with a fixed root:shot ratio (maybe specifically in high end
-  farming systems (?)). Following this argumentation SOC results shown
-  in this study might especially in high-yielding farming system (europe
-  etc.) overestimate actual SOC stocks.
+  As pointed out by Keel (2017), Smith (2019) the results might be highly sensitive to carbon input calculations more precisly to below and above ground residue carbon estimates derived from harvested quantities.
+It has been questioned, that below ground residues might increase with a fixed root:shot ratio (maybe specifically in high end farming systems (?)).
+Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
 \item
@@
 \item
-  diaggregation of manure with urban area is leeding to a lot of
-  displaced manure (?) that is cut off
+  ``The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
+  current C stock to the future steady-state soil C stock calculated based on current conditions.'' Leading to the fact that our total stock results are highly uncertained.
+\item
+  diaggregation of manure with urban area is leeding to a lot of displaced manure (?) that is cut off
 \item
@@
 \item
-  pasture dynamics are neglected and treated as natveg / Note that
-  manure excreted to pastures is neglected within these analysis, since
-  we focus purely on cropland dynamics.
+  pasture dynamics are neglected and treated as natveg / Note that manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.
 \item
-  no tillage adaptaion is neglected on cropland. Pastures are assumed to
-  be not tilled at all (propably at least heavy managed pastures are
-  tilled with some rotation)
+  no tillage adaptaion is neglected on cropland. Pastures are assumed to be not tilled at all (propably at least heavy managed pastures are tilled with some rotation)
 \item
-  irrigated areas are no crop specific and growing periods are not
-  considered (since this is very complex to calculate average growing
-  periods). Crop specific growing periods might be possible using LPJmL
-  data.
+  irrigated areas are no crop specific and growing periods are not considered (since this is very complex to calculate average growing periods).
+Crop specific growing periods might be possible using LPJmL data.
 \item
-  Carbon displacement via leaching and erosion is neglected in this
-  study.
+  Carbon displacement via leaching and erosion is neglected in this study.
 \item
@@
 \item
-  Within cropland we do not track area transitions, but rather look at
-  statistical distributions of the crop functional types. Due to crop
-  rotations and missing data on crop specific distributions, these
-  transitions would be any way rather uncertain. \newpage
+  Within cropland we do not track area transitions, but rather look at statistical distributions of the crop functional types.
+Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
+  \newpage
 \end{itemize}
@@
 
-The conclusion goes here. You can modify the section name with
-\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
+The conclusion goes here.
+You can modify the section name with \texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
 \newpage
 
-\codedataavailability{use this to add a statement when having data sets
-and software code
-available} %% use this section when having data sets and software code available
+\codedataavailability{use this to add a statement when having data sets and software code available} %% use this section when having data sets and software code available
 
@@
 
-If you sorted all figures and tables into the sections of the text,
-please also sort the appendix figures and appendix tables into the
-respective appendix sections. They will be correctly named
-automatically.
+If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
+They will be correctly named automatically.
 
@@
 
-If you put all figures after the reference list, please insert appendix
-tables and figures after the normal tables and figures.
+If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
 
-\texttt{\textbackslash{}appendixfigures} needs to be added in front of
-appendix figures \texttt{\textbackslash{}appendixtables} needs to be
-added in front of appendix tables
+\texttt{\textbackslash{}appendixfigures} needs to be added in front of appendix figures
+\texttt{\textbackslash{}appendixtables} needs to be added in front of appendix tables
 
-Please add \texttt{\textbackslash{}clearpage} between each table and/or
-figure. Further guidelines on figures and tables can be found below.
-Regarding figures and tables in appendices, the following two options
-are possible depending on your general handling of figures and tables in
-the manuscript environment: To rename them correctly to A1, A2, etc.,
-please add the following commands in front of them:
+Please add \texttt{\textbackslash{}clearpage} between each table and/or figure. Further guidelines on figures and tables can be found below.
+Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
+To rename them correctly to A1, A2, etc., please add the following commands in front of them:
 \noappendix
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of
-Bodirsky. Bodirsky and Popp revised paper.} %% optional section
+\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky, and Popp revised paper.} %% optional section
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing
-interests.} %% this section is mandatory even if you declare that no competing interests are present
+\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
 

### 2020-07-14 SOCbudgetPaper/discussion.Rmd
update method part + smaller stuff in other section and result figure
small update on written paper part

@@
 
+- "The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
+current C stock to the future steady-state soil C stock calculated based on current conditions." Leading to the fact that our total stock results are highly uncertained.
+
 - diaggregation of manure with urban area is leeding to a lot of displaced manure (?) that is cut off
@@
 
-- pasture dynamics are neglected and treated as natveg
+- pasture dynamics are neglected and treated as natveg  / Note that manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.
 

### 2020-07-14 SOCbudgetPaper/methods.Rmd
small update on written paper part

--- 2020-07-01 SOCbudgetPaper/methods.Rmd
@@
 
-We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
+We compiled calculations as open-source R packages available under github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on github.com/pik-piam/madrat, a framework improving reproducibility and transparency in data processing.
 In the following chapter we outine the most important relationships and assumptions. See table \@ref(append:subsection2mrfunctions) for further information on corresponding code within the R packages.
 
-## Carbon Stocks following (new) Tier 2 method (50) {#sec:carbonbudget}
+## Carbon Stocks following (new) Tier 2 method {#sec:carbonbudget}
 
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
-We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
+4 (@ipcc_2019_2019), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
+We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
 Therefore we conduct the following three steps within each yearly timestep:
-(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks, 
+(1) We calculate annual land-use (sub-)type specific steady states and decay rates for SOC stocks, 
 (2) We account for land conversion by transferring SOC between land-use types and 
@@
 SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
-\label{eq:inoutflow}
+(\#eq:inoutflow)
 \end{equation}
-with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cell indices and $t$ years.
-We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
-Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are the key components to determining steady-state SOC stocks.
+with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cells indecies and $t$ years.
+We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.
+4 (@ipcc_2019_2019) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
+Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are still the key components to determining steady-state SOC stocks and will be discussed in the following two paragraphs.
 
-#### Carbon Inputs to the Soil {#sec:carboninputs}
+#### Carbon Inputs to the Soil {-}
 
-We account for different carbon input sources depending on the two land-use types we distinguish, croplands and natural vegetation (see table \@ref(tab:datasourceinputs)).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); for natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the only source of carbon.
+We account for different carbon input sources depending on our two distinguished land-use types (see table \@ref(tab:datasourceinputs)); in particular for recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livst_manure)) on croplands; and litterfall ([@LPJmL4_1]) on natural vegetation.
 
-Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
-For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that both the amount of carbon as well as its structural composition determine the effective inflow.
+Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see [@ipcc_2019_2019]).
+For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
+This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
 Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \@ref(tab:datasourceinputs).
@@
  \middlehline
- \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
-                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
+ \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, \@ref(sec:residues)] & default values given by [2]  \\
+                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, \@ref(sec:residues)] & default values given by [2] \\
                             & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
@@
  \end{tabular}
- \label{tab:datasourceinputs}
  \belowtable{}
+ (\#tab:datasourceinputs)
  \end{table*}
@@
 
-#### Soil Organic Carbon decay (300) {#sec:tier2}
+#### Soil Organic Carbon decay {-}
 
 The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
-Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
+Following the steady-state method of the refinement of the IPCC guidelines vol.
+4 ([@ipcc_2019_2019]) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
 Thus $k_{sub}$ is given by
@@
 & k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
-\label{eq:decayrates}
+(\#eq:decayrates)
 \end{aligned}
@@
 
-For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as drivers considered for each land-use types are shown in table \@ref(tab:datasourcedecay).
+For cropland we distinguish the effect of different tillage (see \@ref(sec:tillage)) and irrigation (see \@ref(sec:irrigation)) practices on decay rates, whereas on natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as considered effects for each land-use types are shown in table \@ref(tab:datasourcedecay).
 To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
@@
  \end{tabular}
- \label{tab:datasourcedecay}
  \belowtable{}
+ (\#tab:datasourcedecay)
  \end{table*}
@@
 SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
-\label{eq:ctransfer}
+(\#eq:ctransfer)
 \end{equation}
@@
 
-Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use type (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
+Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
 
@@
 SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
-\label{eq:steadystate}
+(\#eq:SOCstate)
 \end{equation}
 
-Reformulating equation \ref(eq:steadystate), we obtain a massbalance equation as follows
+Reformulated equation \@ref(eq:SOCstate), we obtain a massbalance equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
-\label{eq:steadystate}
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \@ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
+(\#eq:steadystate2budget)
 \end{equation}
@@
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ - land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ - total SOC stock within cell}}
-\label{eq:totalstock}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
+(\#eq:totalstock)
 \end{equation}
@@
 
-To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions will decrease.
+To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initilzation year 1961.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency for the initial conditions will decrease.
 
@@
 
-## Agricultural management data on 0.5 degree resolution	(50) {#sec:agrimanagement}
+## Agricultural management data on 0.5 degrees	(50) {#sec:agrimanagement}
 
-We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a harmonized and consistent data set.
-The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
+We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a comprehensive and constistent data suite.
+The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis together with the accounting for a spin up phase to time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps.
 
@@
 
-#### Crop Production
-Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop group specific half-degree production patterns.
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production ([@FAOSTAT]).
+#### Crop Production {-}
+Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(sec:landuse)) we compile crop group specific half-degree production patterns.
+We calibrate cellular yields with one country-level calibration factor for each crop group to meet historical FAOSTAT production ([@FAOSTAT]).
 Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
 
-#### Crop Residue Production
-Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling.
-Starting from crop production ($CP$) estimates of the harvested organs <!-- organs very odd here, and I'm not sure what the unit being measured/estimated is.
-Is it just crop yields?
-"Starting from respective crop yields ($CP$) and harvested area ($CA$)..." maybe simplify something like that  --> and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
+#### Crop Residue Production {-}
+Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
+Starting from crop production ($CP$) estimates of the harvested organs and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
 
@@
 BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
-\label{eq:resbiomass}
+(\#eq:resbiomass)
 \end{aligned}
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction ([@ipcc_2006_2006]).
-Note that, deviating from [@bodirsky2012], we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction ([@ipcc_2006_2006]).
+Note that deviating from [@bodirsky2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
 We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
 
-##### Burned Residues
+#### Burned Residues {-}
 A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
@@
 
-##### Residue Usage
-We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Table_kcr2kres)).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account.
-We estimate, for low-income countries ($<\,1000\,\tfrac{USD}{yr}$) per capita,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low income countries ($<\,1000\,\tfrac{USD}{yr}$).
-For high-income countries, ($>\,10000\,\tfrac{USD}{yr}$ no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
+#### Residue Usage {-}
+We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see \@ref(append:Table_kcr2kres)).
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see [@weindl]) taking available AGR biomass as well as livestock productivity into account.
+We estimate a material-use share for the straw residues group of 5% and a fuel-share of 10% for all used residues groups in low income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$).
+For high-income  ($>\,10000\,\tfrac{USD}{yr}$ no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
 The remaining AGR as well as all BGR are assumend to be recycled to the soil.
-We limit high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
+We cut high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to corrected for outliers.
 
-#### Dry Matter to Carbon Transformation
+#### Dry Matter to Carbon Transformation {-}
 To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
@@
 
-#### Livestock Distribution
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl.
-We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-For ruminants, we assume that livestock is located where the production of feed occurs.
-We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
-For low-income countries (<1000 USD/yr), we assume extensive production systems.
-We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio.
-Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.
+#### Livestock Distribution {-}
+To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of [@gilbert] and uses feed basket assumptions based on a revised methodology of @weindl.We differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
+For ruminants, we assume that livestock is located at the production spot sof the feed intake.
+We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feed stuff, which we consider to be consumend in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
+For low-income countries according to worldbank definitions (<1000 USD/yr), we assume extensive production systems.
+We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
+Intensive production is distributed within a country using again crop-based feed production share, assuming that feed availability is the most driving factor for livestock location.
 
-#### Manure Excretion, Storage and Recycling
-Manure production and management is based on a revised methodology of ([@bodirsky2012]) and is presented here due to its central role in soil carbon modelling.
+#### Manure Excretion, Storage and Recycling {-}
+Manure production and management is based on a revised methodology of ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling.
 Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets ([@weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
@@
 Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
-Whereas for low-income countries (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
+Whereas for low-income countries according to worldbank definitions (<1000 USD/yr), we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries ($>\,10000\,\tfrac{USD}{yr}$, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses. 
 To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
-We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
+For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
+We cut high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to corrected for outliers.
 
@@
 The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
-We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
-Moreover we assume the irrigation effect to be present for all months of a year, that has been marked as irrigated.
-<!-- assuming that irrigation effect happens 12 months?
-or just the months "marked as irrigated" ?
-this term confusing -->
+We apply aggregated area shares, leading to the same impact of irrigation of all our crop groups.
+Moreover we assume the irrigation effect to te present for all month of a year, that has been marked as irrigated.
 
@@
 
-To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rule is based on the approach of [@porvolik], but neglects the (less common) adoption of no-tillage farming on agricultural land.
+To account for tillage distribution between the three tillage types specfied by the IPCC -- full tillage, reduced tillage and no tillage --, we use the simple assumptions that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
+This rules are inspired by the approach of [@porvolik], but neglects the less common adoption of no tillage farming on agricultural land.
 
@@
 SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F_{\text{LU}_{c,m}} \cdot F_{\text{MG}_{c,m}} \cdot F_{\text{I}_{c,m}} \cdot A_{m,i} \qquad\forall\quad t,
-\label{eq:tier1}
+(\#eq:tier1)
 \end{equation}
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis we us the default SCF as a comparison and constistency check for our more detailed budget approach.
+For this analysis we us the default SCF as comparison and constistency check for our more detailed budget approach.
 
-<!-- KK: Representation of climate zones and grid cells is not consistent and has to be reformulated. -->
+According to [@ipcc_2019_2019] the formulation of absolute changes like
+<!-- plugin absolute chnage equation of refinement cropland chapter -->
+might be neccessary as well.

### 2020-07-14 SOCbudgetPaper/results.Rmd
small update on written paper part

@@
 # Results
+
+We present simulation results of the SOC budget of the period from 1975 to 2010 mainly focusing on cropland areas.
+Detailed total global and regional results of the state of the worlds SOC in comparison to other estimates can be found in the supplement.
+
+## SOC distribution and depletion
+Figure
+
+```{r caption_SOCmaps, echo=FALSE}
+caption <- "(a) ... , (b) ... , (c) ... , (d) ... ."
+```
+
+```{r SOCmaps, out.width = "18cm", fig.cap = caption}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
+```
+
+\newpage
+
+## comparison of stock changes factors
 
@@
 
-```{r out.width = "16cm"}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
-```
+\newpage
+
+## global agricultrual carbon flows and stocks

### 2020-07-14 SOCbudgetPaper/methods.Rmd

--- 2020-07-14 SOCbudgetPaper/methods.Rmd
@@
 
-## Carbon Stocks following (new) Tier 2 method (50) {#sec:carbonbudget}
+## Carbon Stocks following (new) Tier 2 method {#sec:carbonbudget}
 
@@
 SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
-\label{eq:inoutflow}
+(\#eq:inoutflow)
 \end{equation}
@@
 We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
-Annual carbon inflow to each sub-pool (see \@ref(sec:carboninputs)) and annual decay rates (see \@ref(sec:tier2)) of each sub-pool are the key components to determining steady-state SOC stocks.
+Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
 
-#### Carbon Inputs to the Soil {#sec:carboninputs}
+#### Carbon Inputs to the Soil {-}
 
 We account for different carbon input sources depending on the two land-use types we distinguish, croplands and natural vegetation (see table \@ref(tab:datasourceinputs)).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \ref(sec:livst_manure)); for natural vegetation litterfall (\ref(@LPJmL4_1)) litterfall is the only source of carbon.
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livst_manure)); for natural vegetation litterfall ([@LPJmL4_1]) litterfall is the only source of carbon.
 
@@
 
-#### Soil Organic Carbon decay (300) {#sec:tier2}
+#### Soil Organic Carbon decay {-}
 
@@
 SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
-\label{eq:steadystate}
+(\#eq:SOCstate)
 \end{equation}
 
-Reformulating equation \ref(eq:steadystate), we obtain a massbalance equation as follows
+Reformulating equation \@ref(eq:steadystate), we obtain a massbalance equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
-\label{eq:steadystate}
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \@ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
+(\#eq:steadystate2budget)
 \end{equation}
@@
 
-## Agricultural management data on 0.5 degree resolution	(50) {#sec:agrimanagement}
+## Agricultural management data on 0.5 degree resolution {#sec:agrimanagement}
 
@@
 
-#### Crop Production
+#### Crop Production {-}
 Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop group specific half-degree production patterns.
@@
 
-#### Crop Residue Production
+#### Crop Residue Production {-}
 Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling.
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction ([@ipcc_2006_2006]).
-Note that, deviating from [@bodirsky2012], we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction ([@ipcc_2006_2006]).
+Note that deviating from [@bodirsky2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
 We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
 
-##### Burned Residues
+##### Burned Residues {-}
 A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
@@
 
-##### Residue Usage
+##### Residue Usage {-}
 We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Table_kcr2kres)).
@@
 
-#### Dry Matter to Carbon Transformation
+#### Dry Matter to Carbon Transformation {-}
 To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
@@
 
-#### Livestock Distribution
+#### Livestock Distribution {-}
 To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl.
@@
 
-#### Manure Excretion, Storage and Recycling
+#### Manure Excretion, Storage and Recycling {-}
 Manure production and management is based on a revised methodology of ([@bodirsky2012]) and is presented here due to its central role in soil carbon modelling.
@@
 
-### Irrigation (100)
+### Irrigation
 
@@
 
-### Tillage (100)
+### Tillage
 
@@
 
-<!-- KK: Representation of climate zones and grid cells is not consistent and has to be reformulated. -->
+According to [@ipcc_2019_2019] the formulation of absolute changes like
+<!-- plugin absolute chnage equation of refinement cropland chapter -->
+might be neccessary as well.

### 2020-08-14 ResultNotebooks/results_SOCcompare.Rmd
refine results scripts (specially flow figure)

@@
 rev       <- "rev0.6_mrSOC"
+outputdir <- "Output/Images"
 
@@
 
-mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y", ylab="SOC (0-30cm) in MtC")
+regSOCcompare <- mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y", ylab="SOC (0-30cm) in MtC")
+
+ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid)
 ```
@@
 ```
+
+```{r plot diff maps, out.width="100%", out.height=1000}
+
+library(luplot)
+library(magpiesets)
+
+cellSOCdens   <- (dimSums(dimSums(SoilCarbon[,,"naturalstate"], dim=3.2) * Landuse, dim=c(3))/dimSums(Landuse, dim=c(3)))[,findset("past"),]
+cellSOCdens[is.na(cellSOCdens)] <- 0
+
+cellSoilGrids <- readSource("SoilGrids", subtype="cstock_0_30", convert="onlycorrect")
+cellLPJml     <- calcOutput("LPJmL", version="LPJmL4", climatetype="CRU_4", subtype="soilc_layer", time="raw", selectyears="past", aggregate=FALSE)
+cellLPJml     <- collapseNames(cellLPJml[,,"layer1"]+1/3*cellLPJml[,,"layer2"], collapsedim = c(2,3))
+
+diffSoilGrids <- cellSOCdens - cellSoilGrids
+diffLPJml     <- cellSOCdens - cellLPJml
+
+luplot::plotmap(diffLPJml[,10,], legend_range = c(-80,80))
+```

### 2020-08-14 ResultNotebooks/results_SOCflows.Rmd
refine results scripts (specially flow figure)

@@
 
-rev <- "rev0.6_mrSOC"
+rev       <- "rev0.7_mrSOC"
+outputdir <- "Output/Images"
 
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
-
-SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Landuse.rds")
+#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
 
@@
 CarbonHarvest    <- readRDS("Data/Crop_Harvest.rds")
-SoilCarbon       <- readRDS("Data/SoilCarbon.rds")
+SoilCarbonOut    <- readRDS("Data/SoilCarbon.rds")
+SoilCarbonDecay  <- readRDS("Data/SoilDecay.rds")
+SoilCarbonStSt   <- readRDS("Data/SoilSteadyState.rds")
 
@@
 ResiduesAgDemand <- readRDS("Data/ResiduesAg_FieldBalance.rds")
+
+```
+
+```{r globals}
+grepSpaceTime     <- function(x){return(dimSums(x[,"y2010","c"],dim=1))}
+
+#Agriculture
+kcr        <- findset("kcr")
+kli        <- findset("kli")
+kres       <- findset("kres")
+kExResPast <- setdiff(findset("kall"),c(kres,kli,"pasture","wood","woodfuel","scp","fish"))
+kPrim      <- setdiff(findset("kall"),c(kli,"scp","fish"))
+
+CropHANPP         <- dimSums(grepSpaceTime(FAOmassbalance[,,kcr][,,c("food","waste","other_util","bioenergy","feed",
+                                                                     "milling","refining","extracting","fermentation","distilling")]), dim=3)
+CropFEED          <- grepSpaceTime(dimSums(FAOmassbalance[,,kExResPast][,,"feed"], dim=3.1))
+CropPRIM          <- CropHANPP-CropFEED
+ResFEED           <- grepSpaceTime(dimSums(FAOmassbalance[,,kres][,,"feed"], dim=3.1))
+ResPRIM           <- dimSums(grepSpaceTime(FAOmassbalance[,,kres][,,c("other_util","bioenergy")]), dim=3)
+
+FEED              <- grepSpaceTime(dimSums(FAOmassbalance[,,kPrim][,,"feed"], dim=3.1))
+LivstHANPP        <- dimSums(grepSpaceTime(FAOmassbalance[,,kli][,,c("food","waste","other_util")]), dim=3)
+LivstCropHANPP    <- (CropFEED+ResFEED)/FEED * LivstHANPP
+
+ResidueBiomassTot <- grepSpaceTime(dimSums(ResidueBiomass, dim=3.2))
+ResidueRecycledTot<- grepSpaceTime(dimSums(ResiduesAgDemand[,,"recycle"], dim=3.2))
+
+ManureExcretedTot <- grepSpaceTime(dimSums(ManureExcreted[,,c("stubble_grazing","confinement")], dim=c(3.1,3.2)))
+ManureRecycledTot <- grepSpaceTime(dimSums(ManureRecycled, dim=3.1))
+
+#Soil
+
+### Carbon To
+CarbonTo            <- dimSums(toolConditionalReplace(SoilCarbonOut[,"y2010","carbontransfer"], "<0", 0),dim=c(1,3.3))
+CarbonToEff         <- dimSums(SoilCarbonOut[,"y2010","carbontransfer"], dim=c(1,3.3))
+CarbonToPools       <- dimSums(toolConditionalReplace(SoilCarbonOut[,"y2010","carbontransfer"], "<0", 0),dim=c(1))
+
+####### delete when new output on decay is ready
+# Decay
+
+ActiveDecay        <- mbind(setNames(calcOutput("ActiveDecay",  landtype="crop",   aggregate = FALSE),"crop"),
+                              setNames(calcOutput("ActiveDecay",  landtype="natveg", aggregate = FALSE),"natveg"))
+SlowDecay          <- mbind(setNames(calcOutput("SlowDecay",  landtype="crop",   aggregate = FALSE),"crop"),
+                              setNames(calcOutput("SlowDecay",  landtype="natveg", aggregate = FALSE),"natveg"))
+PassiveDecay       <- mbind(setNames(calcOutput("PassiveDecay", landtype="crop",   aggregate = FALSE),"crop"),
+                              setNames(calcOutput("PassiveDecay", landtype="natveg", aggregate = FALSE),"natveg"))
+
+subpools     <- c("active","slow","passive")
+names        <- as.vector(outer(getNames(Landuse), subpools, paste, sep="."))
+Decay <- new.magpie(getCells(Landuse), getYears(Landuse), names, fill = 0)
+
+Decay[,,"active"]                  <- ActiveDecay
+Decay[,,"slow"]                    <- SlowDecay
+Decay[,,"passive"]                 <- PassiveDecay
+
+rm(ActiveDecay, SlowDecay, PassiveDecay)
+
+#
+####### end delete
+
+DecayCut          <- Decay
+DecayCut[Decay>1] <- 1
+
+### Carbon In
+f1        <- param[,,"f1"]
+f2_crop   <- param[,,"f2_ft"]
+f2_natveg <- param[,,"f2_nt"]
+f3        <- param[,,"f3"]
+
+# f2 has to be updated with new test runs
+
+MineralizeFracInput <- setCells(mbind(add_dimension(collapseNames(CarbonInput[,,"totalC",invert=TRUE][1,,"res"]), dim=3.1, add = "lu", nm = "crop"),
+                                      add_dimension(collapseNames(CarbonInput[,,"totalC",invert=TRUE][1,,"res"]), dim=3.1, add = "lu", nm = "natveg")), "GLO")
+MineralizeFracInput[,,"metabDOC"]        <- f1
+MineralizeFracInput[,,"crop.strucDOC"]   <- f2_crop
+MineralizeFracInput[,,"natveg.strucDOC"] <- f2_natveg
+MineralizeFracInput[,,"ligninC"]         <- f3
+
+input2lu <- as.data.frame(list(c("res", findset("kli"), "litfall"),
+                            c(rep("crop",6), "natveg")), col.names=c("input","lu"))
+
+CarbonInEff       <- collapseNames(dimSums(toolAggregate(CarbonInput[,"y2010",][,,"totalC",invert=TRUE], rel=input2lu, dim=3.1) *
+                                            MineralizeFracInput[,"y2010",] * Landuse[,"y2010",], dim=c(1,3.2)))
+
+CarbonInSource   <- collapseNames(dimSums(CarbonInput[,"y2010",][,,"totalC",invert=TRUE]  * 
+                                            toolAggregate(MineralizeFracInput[,"y2010",] * Landuse[,"y2010",], rel=input2lu, from="lu", to="input", dim=3.1), dim=c(1,3.2)))
+
+CarbonInRaw       <- collapseNames(dimSums(SoilCarbonOut[,"y2010","steadystate"]*DecayCut[,"y2010",]*Landuse[,"y2010",], dim=c(1,3.3)))
+
+### Carbon Out
+
+CarbonOutRaw      <- collapseNames(dimSums(SoilCarbonOut[,"y2010","interstate"]*DecayCut[,"y2010",]*Landuse[,"y2010",], dim=c(1,3.3)))
+CarbonOutEff      <- CarbonOutRaw + CarbonInEff - CarbonInRaw
+
+### SOC State
+
+SOCstate <- dimSums(SoilCarbonOut[,c("y2009","y2010"),c("actualstate","interstate")]*Landuse[,c("y2009","y2010"),], dim=c(1,3.3))
 

### 2020-08-14 ResultNotebooks/results_SOCmaps.Rmd
refine results scripts (specially flow figure)

@@
 
+library(ggplot2)
+
 SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
@@
 lines(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCStock[,,"naturalstate"])
-
-SOCemis <- ggplot() + geom_line(data= as.data.frame(glo_SOCEmission)[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
-  labs( x = "Year", 
-        y = "Global cummulative SOC emission in GtC") +
-  scale_x_discrete(breaks=seq(1965,2010,5)) +
-  scale_y_continuous(limits=c(25,40), breaks=seq(25,40,5)) +
-  theme(rect=element_blank())
 
@@
 
-library("ggplot2")
-theme_set(theme_bw())
-library("sf")
-library("rnaturalearth")
-library("rnaturalearthdata")
-library(RColorBrewer)
-
-cols_dens <- colorRampPalette(brewer.pal(11, "RdYlBu"))(30)[c(1:20, rep(c(21:30), each=10))]
-
-outputdir <- "Output/Images"
-
-world         <- ne_countries(scale = "medium", returnclass = "sf")
 magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
@@
 density2plot  <- merge(as.data.frame(cropland_dens), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)
-world.robin   <- st_transform(world,"+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
-
-SOCshare <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(share2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(0,2), colours = cols, 
-                      name = expression("ratio of cropland SOC to potential natural SOC"), 
-                      breaks=c(0.3,0.6,0.9,1.1,1.4,1.7,2),
-                      labels=c("0     0.3","  0.6", "  0.9","  1.1","  1.4","  1.7","> 2"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=10),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = T,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
-          )
-
-SOCincrease <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(increase2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression("cropland SOC change in tons of carbon per hectar"), 
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("> -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=10),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = T,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
-          )
-
-cols <- colorRampPalette(brewer.pal(9, "YlGn"))(8)[2:8]
-display.brewer.pal(9, "YlGn")
-display.brewer.pal(9, "RdYlGn")
-
-SOCdensity <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(density2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(0,150), colours = cols, 
-                      name = expression("cropland SOC in tons of carbon per hectar"), 
-                      values=c(0,5,10,20,40,70,100,150)/150,
-                      breaks=c(5,10,20,40,70,100,150),
-                      labels=c("0     5","  10","  20","  40","  70","  100"," >150"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=10),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = T,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
-          )
 
@@
 ```{r grid plot}
+
+library("ggplot2")
+theme_set(theme_bw())
+library("sf")
+library("rnaturalearth")
+library("rnaturalearthdata")
+library(RColorBrewer)
+library(sp)
+library(rworldmap)
+library(oce)
+
+legend_size <- 10
+
+map.world <- map_data("world")
+world         <- ne_countries(scale = "medium", returnclass = "sf")
+data(coastlineWorld)  
+coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+
+SOCemis <- ggplot() + geom_line(data= as.data.frame(glo_SOCEmission[,seq(1975,2010,5),])[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
+  labs( x = "Year", 
+        y = "Global cummulative \nSOC emission in GtC") +
+  scale_x_discrete(breaks=seq(1965,2010,5)) +
+  scale_y_continuous(limits=c(30,40), breaks=seq(30,40,5)) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8))
+
+cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)
+
+SOCshare <- ggplot(data = world) + geom_sf() + 
+    geom_tile(data = subset(share2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits=c(0,2), colours = cols, 
+                      name = expression("ratio of cropland SOC to potential natural SOC"), 
+                      breaks=c(0.3,0.6,0.9,1.1,1.4,1.7,2),
+                      labels=c("0     0.3","  0.6", "  0.9","  1.1","  1.4","  1.7","> 2"), na.value = "white",
+                      guide = guide_legend(
+                        title.theme=element_text(size=legend_size),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = F,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          legend.text=element_text(size=legend_size),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          )
+
+#geom_map(dat=map.world, map = map.world, aes(map_id=region), fill="NA", color="black") +
+
+SOCincrease <- ggplot(data = world) + geom_sf() +
+    geom_tile(data = subset(increase2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
+                      name = expression("cropland SOC change in tons of carbon per hectar"), 
+                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
+                      breaks=c(-50,-20,-10,10,20,50,100),
+                      labels=c("> -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
+                      guide = guide_legend(
+                        title.theme=element_text(size=legend_size),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = F,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          legend.text=element_text(size=legend_size),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          )
+
+cols <- colorRampPalette(brewer.pal(9, "YlGn"))(8)[2:8]
+
+SOCdensity <- ggplot(data = world) + geom_sf() +
+    geom_tile(data = subset(density2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits=c(0,150), colours = cols, 
+                      name = expression("cropland SOC in tons of carbon per hectar"), 
+                      values=c(0,5,10,20,40,70,100,150)/150,
+                      breaks=c(5,10,20,40,70,100,150),
+                      labels=c("0     5","  10","  20","  40","  70","  100"," >150"), na.value = "white",
+                      guide = guide_legend(
+                        title.theme=element_text(size=legend_size),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = F,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          legend.text=element_text(size=legend_size),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          )
+
 library(gridExtra)
-
-grid <- grid.arrange(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2)
+library(cowplot)
+grid <- plot_grid(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2, labels=c("a","b", "c", "d"),
+                  label_fontface = "plain", label_size = 10)
 grid
@@
 ```
-
-```{r plot diff maps, out.width="100%", out.height=1000}
-
-library(luplot)
-library(magpiesets)
-
-cellSOCdens   <- (dimSums(dimSums(SoilCarbon[,,"naturalstate"], dim=3.2) * Landuse, dim=c(3))/dimSums(Landuse, dim=c(3)))[,findset("past"),]
-cellSOCdens[is.na(cellSOCdens)] <- 0
-
-cellSoilGrids <- readSource("SoilGrids", subtype="cstock_0_30", convert="onlycorrect")
-cellLPJml     <- calcOutput("LPJmL", version="LPJmL4", climatetype="CRU_4", subtype="soilc_layer", time="raw", selectyears="past", aggregate=FALSE)
-cellLPJml     <- collapseNames(cellLPJml[,,"layer1"]+1/3*cellLPJml[,,"layer2"], collapsedim = c(2,3))
-
-diffSoilGrids <- cellSOCdens - cellSoilGrids
-diffLPJml     <- cellSOCdens - cellLPJml
-
-luplot::plotmap(diffLPJml[,10,], legend_range = c(-80,80))
-```
-
-```{r plot density maps 2}
-
-cols <- colorRampPalette(brewer.pal(9,"Oranges"))(1000)
-cropland <- dimSums(SoilCarbon[,,"crop"], dim=3)
-cropland[Landuse[,,"crop"]==0] <- NA
-
-luplot::plotmap(cropland[,"y2010",], legend_colours = cols, legend_range=range(cropland[,"y2010",]/5, na.rm=TRUE))
-
-natland <- dimSums(SoilCarbon[,,"natveg"], dim=3)
-
-luplot::plotmap(natland[,"y2010",], legend_colours = cols, legend_range=range(cropland[,"y2010",]/5, na.rm=TRUE))
-
-```

### 2020-08-14 ResultNotebooks/results_SOCtable.Rmd
update results
refine results scripts (specially flow figure)

@@
 
-rev <- "rev0.6_mrSOC"
+rev       <- "rev0.7_mrSOC"
+outputdir <- "Output/"
 
@@
 
+getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))
+
 print(collapseNames(CShare_ipcczone[,findset("past"),]))
@@
              actualstate_table  = actualstate_table, 
-             steadystate_table  = steadystate_table), "Output/SFC_climatezone.rds")
+             steadystate_table  = steadystate_table), paste0(outputdir, "SFC_climatezone.rds"))
 
@@
 
+getNames(CShare_ipcc) <- gsub("_", " ", getNames(CShare_ipcc))
+
 comparison <- mbind(CShare_ipcc, 
+                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),"y2010"),
+                                  nm="SOCbudget 1990",add="model"),2),
                     round(add_dimension(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),
-                                  add="SOCbudget",nm="model"),2))
+                                  nm="SOCbudget 2010",add="model"),2))
 
@@
 
+comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(comparison)), 
+                            Data1 ~ Data2, value.var="Value", 
+                            row.names=levels(as.data.frame(comparison)$Data1)))
+
+saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"SFC_comparison.rds"))
+
 ```
+
+```{r kable2image}
+library(kableExtra)
+
+table <- readRDS("Output/SFC_comparison.rds")
+colnames(table[[1]]) <- gsub("_", " ", colnames(table[[1]]))
+colnames(table[[1]])[1] <- "Source"
+save_kable(kable_styling(kable(table$comparison_table, "latex", digit=2), latex_options = c("striped", "scale_down")), 
+         file="Output/Images/TableSCF_comparison.png")
+
+table <- readRDS("Output/SFC_climatezone.rds")
+colnames(table[[1]]) <- gsub("_", " ", colnames(table[[1]]))
+colnames(table[[1]][1]) <- ""
+as_image(kable_styling(kable(table$naturalstate_table[3:10,], "latex", digit=2), latex_options = c("striped", "scale_down")), 
+         width=4, file="Output/Images/TableSCF_timeline.png")
+```

### 2020-08-14 SOCbudgetPaper/SOCbudget.Rmd
refine results scripts (specially flow figure)

@@
 abstract: |
-  Soil organic carbon (SOC) is one of larges c stocks on earth (3 times larger biosphere pool). 
+  Soil organic carbon (SOC) is one of larges carbon stocks on earth.
+It is three times larger than the biospheric pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
+Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
+While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
+ Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
+We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been regained within the period 1975-2010.
+ We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
+We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
 bibliography: SOCbudget.bib
@@
   codedata: |
-    use this to add a statement when having data sets and software code available
+    Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/. Data used for the output can be found under <something-with-zenodo>.  
 authorcontribution: |
-  Karstens wrote code and paper build on work of Bodirsky. Bodirsky, <here_could_be_your_names> and Popp revised paper.
+  Karstens wrote code and paper build on work of Bodirsky (and <here_could_be_your_names>). 
+  Bodirsky, <here_could_be_your_names> and Popp revised paper.
 disclaimer: |
@@
   Thanks to the rticles contributors!
-appendix: |
-  \section{Figures and tables in appendices}
-  \subsection{Option 1}
-  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-  They will be correctly named automatically.
-  \subsection{Option 2}
-  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-
-  `\appendixfigures` needs to be added in front of appendix figures
-  `\appendixtables` needs to be added in front of appendix tables
-
-  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
-  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
 output:
@@
 \newpage
+
+```{r child = 'appendix.Rmd'}
+```
+\newpage

### 2020-08-14 SOCbudgetPaper/SOCbudget.Rmd
update written part to first official version (woohoo)

@@
 title: Management induced changes of soil organic carbon on global croplands 
-       \newline or \newline
-       How management changed soil organic carbon on global croplands
-       \newline or \newline
-       Agricultural soil have lost 13 GtC topsoil carbon
-       \newline or \newline
-       Linking agricultural management data to soil modeleling – a global approach
-
 journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
@@
 Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
-We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been regained within the period 1975-2010.
+We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
 We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
@@
     variant: markdown_github
+  rticles::copernicus_article: default
   bookdown::word_document2: default
   bookdown::pdf_document2: default
-  rticles::copernicus_article: default
 ---

### 2020-08-14 SOCbudgetPaper/SOCbudget.tex
update written part to first official version (woohoo)

@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
 \begin{document}
 
-\title{Management induced changes of soil organic carbon on global croplands \newline or \newline How management changed soil organic carbon on global croplands \newline or \newline Agricultural soil have lost 13 GtC topsoil carbon \newline or \newline Linking agricultural management data to soil modeleling -- a global approach}
+\title{Management induced changes of soil organic carbon on global croplands}
 
@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of larges c stocks on earth (3 times larger biosphere pool).
+Soil organic carbon (SOC) is one of larges carbon stocks on earth.
+It is three times larger than the biospheric pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
+Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
+While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
+Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
+We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
+We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
+We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
 \end{abstract}
@@
 
-The depletion of soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change.
-Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times.
-Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref.
+Possible titles:
+
+\begin{itemize}
+\item
+  How management changed soil organic carbon on global croplands
+\item
+  Agricultural soil have lost 37 GtC topsoil carbon
+\item
+  Linking agricultural management data to soil modeleling -- a global approach
+\end{itemize}
+
+Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and biospheric pools multiple times (Bathjes).
+As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref.
 permafrost melting).
-The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
-
-Mapping the worlds SOC has been investigated intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools.
-Natural properties like climatic, biophysical and landscape characteristics clearly play the most important role in this regard.
-
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely the dominant driver of SOC change on managed land today.
-Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC (Lal, 2006).
-Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics ().
-Despite providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management.
-
+The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
+Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles in this regard.\\
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today.
+Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC (Lal, 2006).
+Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX).
+While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especial agricultural activities.\\
 Field-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on yield levels, fertilizer inputs and various other on farming activities.
-Yet, due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.
-
-Our study combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
-We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation and tillage.
-We thereby provide the first global, spatial explicit SOC maps that consider agricultural management.
-
-This paper will introduce in the the first part of the method section the basic concept of SOC dynamics.
-We continue in the second part with a detailed description of the used global gridded management data on production levels, residue recycling rates, manure amendments, irrigation and tillage adoption.
-In a third, small part, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-In our result section we will focus on the SOC dynamics on global croplands by analysing (1) the spatial explicit SOC distribution and depletion over the last decades, (2) compare the climate zone specfic stock changes with default stock changes factors from the IPCC Tier 1 approach and (3) compare global agricultrual carbon flows and stocks to contextualize the importants of the different management aspects.
-Finally, we will discuss our findings and their implications on SOC model development and conlude with an analysis and outlook on the the opportunities for SOC management for climate change mitigation and negative emissions.
-
+Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm activities.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.\\
+Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
+This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
+We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation, and tillage.
+We thereby provide the first global, spatially explicit SOC loss maps that consider detailed agricultural management.\\
+This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
+We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
+Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
+Finally, we will discuss our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.\\
 \newpage
 
-\hypertarget{method-50}{%
-\section{Method (50)}\label{method-50}}
+\hypertarget{method}{%
+\section{Method}\label{method}}
 
-We compiled calculations as open-source R packages available under github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on github.com/pik-piam/madrat, a framework improving reproducibility and transparency in data processing.
-In the following chapter we outine the most important relationships and assumptions. See table \ref{append:subsection2mrfunctions} for further information on corresponding code within the R packages.
+We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
+In the following chapter we outine the most important relationships and assumptions. Table \ref{append:subsection2mrfunctions} provides further information on corresponding code within the R packages.
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{Carbon Stocks following (new) Tier 2 method (50)}\label{sec:carbonbudget}}
+\subsection{SOC Stocks and Stock changes following the Tier 2 steady-state method}\label{sec:carbonbudget}}
 
-Following the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}), we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
-We assume the actual SOC state converges towards a stable steady state, that itself is changing over time and space depending on biophysical, climatic and agronomic conditions.
+Following the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
+We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
 Therefore we conduct the following three steps within each yearly timestep:
-(1) We calculate annual land-use (sub-)type specific steady states and decay rates for SOC stocks,
-(2) We account for land conversion by transferring SOC between land-use types and
-(3) We estimate SOC stocks based on the stocks of the previous time period, the steady state stocks and the decay rate.
+(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks,
+(2) we account for land conversion by transferring SOC from and to an other land-use type representad as natural vegetation,
+(3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
 
@@
 \end{equation}
-with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cells indecies and \(t\) years.
-We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them.
-Annual carbon inflow to each sub-pool (see \ref{sec:carboninputs}) and annual decay rates (see \ref{sec:tier2}) of each sub-pool are still the key components to determining steady-state SOC stocks.
+with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cell indices and \(t\) years.
+We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
+Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
 
-\hypertarget{carbon-inputs-to-the-soil-seccarboninputs}{%
-\paragraph*{Carbon Inputs to the Soil \{\#sec:carboninputs\}}\label{carbon-inputs-to-the-soil-seccarboninputs}}
-\addcontentsline{toc}{paragraph}{Carbon Inputs to the Soil \{\#sec:carboninputs\}}
+\textbf{Carbon Inputs to the Soil}
 
-We account for different carbon input sources depending on our two distinguished land-use types (see table \ref{tab:datasourceinputs}); in particular for recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see @ref(sec:livst\_manure)) on croplands; and litterfall (\citep{LPJmL4_1}) on natural vegetation.
+We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \ref{tab:datasourceinputs}).
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see \ref{sec:livstmanure}); for natural vegetation litterfall \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
 
-Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citep{ipcc_2019_2019}).
-For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow.
+Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citet{ipcc_2019_2019}).
+For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
+This implies that both the amount of carbon as well as its structural composition determine the effective inflow.
 Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
@@
  \middlehline
- \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, \\ref{sec:residues}] & default values given by [2]  \\
-                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, \\ref{sec:residues}] & default values given by [2] \\
+ \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
+                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
                             & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
@@
  \end{tabular}
+ \label{tab:datasourceinputs}
  \belowtable{}
- \label{tab:datasourceinputs}
  \end{table*}
 
-\hypertarget{soil-organic-carbon-decay-300-sectier2}{%
-\paragraph*{Soil Organic Carbon decay (300) \{\#sec:tier2\}}\label{soil-organic-carbon-decay-300-sectier2}}
-\addcontentsline{toc}{paragraph}{Soil Organic Carbon decay (300) \{\#sec:tier2\}}
+\textbf{SOC decay}
 
 The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
-Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citep{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
+Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
 Thus \(k_{sub}\) is given by
@@
 
-For cropland we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas on natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as considered effects for each land-use types are shown in table \ref{tab:datasourcedecay}.
+For cropland we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as drivers considered for each land-use types are shown in table \ref{tab:datasourcedecay}.
 To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
@@
 
-\hypertarget{total-soc-stocks}{%
-\subsubsection{Total SOC stocks}\label{total-soc-stocks}}
+\hypertarget{total-soc-stocks-and-stock-changes}{%
+\subsubsection{Total SOC stocks and stock changes}\label{total-soc-stocks-and-stock-changes}}
 
-Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.
+Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use type (lut), each sub-pool (sub) and each annual time step (t) like
 
@@
 
-Reformulated equation \eqref{eq:SOCstate}, we obtain a massbalance equation as follows
+Reformulating this equation, we obtain a massbalance equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \\eqref{eq:inoutflow})}}  \qquad \forall\quad lut, sub, i.
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
 \label{eq:steadystate2budget}
@@
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ - land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ - total SOC stock within cell}}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
 \label{eq:totalstock}
 \end{equation}
+
+According to the IPCC guidelines SOC changes can be calculates as the difference of two (aufeinander folgenden) years (see Eq.
+5.0A in \citep{ipcc_2019_2019}).
+This however will also include naturally occured changes due to climatic variance over time.
+For our study we will define the absolute and relative SOC changes in relation to a potential natural state \(SOC_{pnv}\) under the same climatic conditions at time \(t\) in grid cell \(i\), that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
+The absolute changes \(\Delta SOC\) and relative changes \(F^{SCF}\) a thus given by
+
+\begin{equation}
+\Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
+\label{eq:totalstock}
+\end{equation}
+
+Note that the absolution changes \(\Delta SOC\) can be also interpreted as cummulative SOC emissions or as SOC depth of mankind; whereas relative changes \(F^{SCF}\) can be denoted as stock change factors as defined within the IPCC guidelines of 2006 \citep{ipcc_2006_2006}.
 
@@
 
-To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initilzation year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency for the initial conditions will decrease.
+To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions will decrease.
 
@@
 \hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data on 0.5 degrees (50)}\label{sec:agrimanagement}}
+\subsection{Agricultural management data on 0.5 degree resolution}\label{sec:agrimanagement}}
 
-We compile country-specific FAO production and cropland statistics (\citep{FAOSTAT}) to a comprehensive and constistent data suite.
-The data is prepared in 5 year time steps from 1965 to 2010, which also restricts our analysis together with the accounting for a spin up phase to time span from 1975 to 2010.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps.
+We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
+The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
 
 \hypertarget{sec:landuse}{%
-\subsubsection{Landuse and Landuse Change (150)}\label{sec:landuse}}
+\subsubsection{Landuse and Landuse Change}\label{sec:landuse}}
 
-Land-use patterns are based on the Land-Use Harmonization 2 (LUH2, \citep{LUH2}) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each gridcell based on the country and year specific area harvested shares of FAOSTAT data (\citep{FAOSTAT}) (see @ref(append:Table\_luh2fao2mag) for more details on the crop group mapping).
+Land-use patterns are based on the Land-Use Harmonization 2 \citep{Hurtt2020} data set, which we aggregate from quarter degree to half degree resolution.
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see \ref{append:Tableluh2fao2mag} for more details on the crop group mapping).
 Land-use transitions are calculated as net area differences of the land-use data on half-degree.
@@
 \hypertarget{sec:residues}{%
-\subsubsection{Crop and Crop Residues (300)}\label{sec:residues}}
+\subsubsection{Crop and Crop Residues Production}\label{sec:residues}}
 
 \textbf{Crop Production}
-Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns.
-We calibrate cellular yields with one country-level calibration factor for each crop group to meet historical FAOSTAT production (\citep{FAOSTAT}).
+
+Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns.
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
 Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
@@
 \textbf{Crop Residue Production}
-Crop residue production and management is based on a revised methodology of (\citep{bodirsky2012}) and will be explained in key aspects again due to its central role for soil carbon modelling.
-Starting from crop production (\(CP\)) estimates of the harvested organs and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios (\(RS\)) as follows
+
+Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modelling.
+Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios (\(RS\)) as follows
 
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction (\citep{ipcc_2006_2006}).
-Note that deviating from \citep{bodirsky2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
-We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals (\citep{weindl}), fuel or for material use.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction \citep{ipcc_2006_2006}.
+Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals \citep{weindl}, fuel or for material use.
 
 \textbf{Burned Residues}
+
 A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
-Following \citep{smil1999}) we assume 25\% burn share for low-income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process (\citep{ipcc_2006_2006}).
+Following \citep{smil1999}) we assume 25\% burn share for low-income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process \citep{ipcc_2006_2006}.
 
 \textbf{Residue Usage}
-We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see @ref(append:Table\_kcr2kres)).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see \citep{weindl}) taking available AGR biomass as well as livestock productivity into account.
-We estimate a material-use share for the straw residues group of 5\% and a fuel-share of 10\% for all used residues groups in low income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)).
-For high-income (\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
+
+We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \ref{append:Tablekcr2kres}).
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see \citep{weindl}), taking available AGR biomass as well as livestock productivity into account.
+We estimate, for low-income countries, a material-use share for straw residues of 5\% and a fuel-share of 10\% for all used residues groups in low-income countries.
+For high-income countries, no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
 The remaining AGR as well as all BGR are assumend to be recycled to the soil.
-We cut high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.
+We limit high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
 
 \textbf{Dry Matter to Carbon Transformation}
-To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see @ref(append:Table\_c2dm)) (\citet{Ma2018}).
 
-\hypertarget{sec:livst_manure}{%
-\subsubsection{Livestock Distribution and Manure Excretion (300)}\label{sec:livst_manure}}
+To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \ref{append:Tablec2dm}).
+
+\hypertarget{sec:livstmanure}{%
+\subsubsection{Livestock Distribution and Manure Excretion}\label{sec:livstmanure}}
 
@@
 \textbf{Livestock Distribution}
-To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of \citep{gilbert} and uses feed basket assumptions based on a revised methodology of \citet{weindl.We} differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
-For ruminants, we assume that livestock is located at the production spot sof the feed intake.
-We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feed stuff, which we consider to be consumend in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems.
-For low-income countries according to worldbank definitions (\textless1000 USD/yr), we assume extensive production systems.
-We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio.
-Intensive production is distributed within a country using again crop-based feed production share, assuming that feed availability is the most driving factor for livestock location.
+To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014}, and uses feed basket assumptions based on a revised methodology from \citep{weindl}.
+We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
+For ruminants, we assume that livestock is located where the production of feed occurs.
+We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
+For low-income countries, we assume extensive production systems.
+We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio.
+Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production, assuming that feed availability is the most driving factor for livestock location.
 
 \textbf{Manure Excretion, Storage and Recycling}
-Manure production and management is based on a revised methodology of (\citep{bodirsky2012}) and will be explained in key aspects again due to its central role for soil carbon modelling.
-Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets (\citep{weindl}), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep[(][]{ipcc_2019_2019}).
+Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modelling.
+Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets \citep{weindl}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep{ipcc_2019_2019}).
 Depending on the feed system we assume manure to be handled in four different ways:
 All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Manure fuel shares are estimated using IPCC default values (\citet{ippc_2006_2006}).
-Whereas for low-income countries according to worldbank definitions (\textless1000 USD/yr), we adopt a share of 25\% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries (\(>\,10000\,\tfrac{USD}{yr}\), leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
-For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses.
+Manure fuel shares are estimated using IPCC default values \citep{ippc_2006_2006}.
+Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
+For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing.
 To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky2012} (see @ref(append:Table\_clossAWMS))).
-We cut high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.
+For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see \ref{append:TableclossAWMS}).
+We limit high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
 
-\hypertarget{irrigation-100}{%
-\subsubsection{Irrigation (100)}\label{irrigation-100}}
+\hypertarget{sec:irrigation}{%
+\subsubsection{Irrigation}\label{sec:irrigation}}
 
@@
 The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
-We apply aggregated area shares, leading to the same impact of irrigation of all our crop groups.
-Moreover we assume the irrigation effect to te present for all month of a year, that has been marked as irrigated.
+We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
+Moreover we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
 
-\hypertarget{tillage-100}{%
-\subsubsection{Tillage (100)}\label{tillage-100}}
+\hypertarget{sec:tillage}{%
+\subsubsection{Tillage}\label{sec:tillage}}
 
-To account for tillage distribution between the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we use the simple assumptions that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rules are inspired by the approach of \citep{porvolik}, but neglects the less common adoption of no tillage farming on agricultural land.
+To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
+This rule is based on the approach of \citep{porwollik_generating_2018}, but neglects the (less common) adoption of no-tillage farming on agricultural land.
 
@@
 \hypertarget{sec:tier1}{%
-\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}
+\subsection{SOC Stocks and Stock changes following Tier 1}\label{sec:tier1}}
 
-Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}).
-Here, stocks are calculated via stock change factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-SCF are differentiated by different crop and management systems (\(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover SCF vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}).
+Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}).
+Here, stocks are calculated via stock change factors (F\^{}\{SCF\}) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+\(F^{SCF}\) are differentiated by different crop, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover \(F^{SCF}\) vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}).
 The actual SOC stocks as thus calulated based on a given reference SOC stock by
@@
 \begin{equation}
-SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F_{\text{LU}_{c,m}} \cdot F_{\text{MG}_{c,m}} \cdot F_{\text{I}_{c,m}} \cdot A_{m,i} \qquad\forall\quad t,
+SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
 \label{eq:tier1}
@@
 with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis we us the default SCF as comparison and constistency check for our more detailed budget approach.
-
-According to \citep{ipcc_2019_2019} the formulation of absolute changes like
-might be neccessary as well.
-
+For this analysis we use the default \(F^{SCF}\) from the Tier 1 method of \citep{ipcc_2006_2006}, and \citep{ipcc_2019_2019} as a comparison and constistency check for our more detailed budget approach.
 \newpage
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# All defaults}
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include\_graphics}\NormalTok{(}\StringTok{"../ResultNotebooks/Output/Images/4panelfigure.png"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a) ... , (b) ... , (c) ... , (d) ... .}\label{fig:SOCmaps}
+\begin{figure}[H]
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas.
+(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades (despite ongoing cropland expansion).
+(c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
+Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.}\label{fig:SOCmaps}
 \end{figure}
 
-\newpage
+\hypertarget{global-agricultural-carbon-flows-and-stocks}{%
+\subsection{global agricultural carbon flows and stocks}\label{global-agricultural-carbon-flows-and-stocks}}
+
+\begin{figure}[H]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/OuFlowFig} \caption{Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
+Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil).}\label{fig:FlowFig}
+\end{figure}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:table}SCF compared to potential natural state}
-\centering
-\begin{tabular}[t]{l|r|r|r|r}
-\hline
-  & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
-\hline
-1965 & 0.25 & 0.33 & 0.55 & 0.48\\
-\hline
-1970 & 0.29 & 0.36 & 0.56 & 0.50\\
-\hline
-1975 & 0.32 & 0.38 & 0.57 & 0.51\\
-\hline
-1980 & 0.34 & 0.38 & 0.56 & 0.53\\
-\hline
-1985 & 0.36 & 0.40 & 0.60 & 0.56\\
-\hline
-1990 & 0.38 & 0.42 & 0.63 & 0.58\\
-\hline
-1995 & 0.40 & 0.45 & 0.63 & 0.59\\
-\hline
-2000 & 0.42 & 0.47 & 0.64 & 0.60\\
-\hline
-2005 & 0.45 & 0.54 & 0.69 & 0.65\\
-\hline
-2010 & 0.50 & 0.58 & 0.73 & 0.67\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:table2}SCF compared to IPCC Tier 1}
-\centering
-\begin{tabular}[t]{l|l|l|l|l}
-\hline
-Data1 & tropical\_moist & tropical\_dry & temperate\_dry & temperate\_moist\\
-\hline
-IPCC2006 & 0.48 & 0.58-0.64 & 0.80 & 0.69\\
-\hline
-IPCC2019 & 0.83 & 0.92 & 0.76-0.77 & 0.69-0.70\\
-\hline
-model & 0.5 & 0.58 & 0.73 & 0.67\\
-\hline
-\end{tabular}
-\end{table}
+\begin{figure}[H]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{$F^{SOC}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{SOC}$ by additional 5-8 percent.
+Also: modelled $F^{SOC}$ have increased for all climates over time .}\label{fig:SCFtable}
+\end{figure}
 
 \newpage
-
-\hypertarget{global-agricultrual-carbon-flows-and-stocks}{%
-\subsection{global agricultrual carbon flows and stocks}\label{global-agricultrual-carbon-flows-and-stocks}}
 
@@
 
-Big points:
-
 \begin{itemize}
 \item
-  SOC initialization is bad (Task: find a better representation)
+  The natural land representation and therefor also SOC initialization is lacking proper parametrization of both decay behaviour and natural input (e.g.~nitrogen and lignin content of litterfall).
+As all results are valued against these potential natural SOC the absolute values and also the changes over time have to be used with caution.
+Especially in boreal zone and in very dry regions the natural land representation might be far off, due to the focus for the parametrization for cropped areas.
+  The good agreement of the stock change factors \(F^{SCF}\) with the Tier 1 default values might serve as a hint that the natural land representation is not totally off.
+Boreal and very dry zone might be uncropped, leaving the results on SOC changes on cropland untouch from the bad borreal land representation.
 \item
-  natural land representation is lacking proper parametrization of natural input (n, lg content of litterfall) (--\textgreater{} ask Christop again)
-\item
-  boreal zone and dry regions are not well represented (maybe because of the bad parameterization of the soil model for natural soils)
+  This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
+Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+It has been argued that for intensivly tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
+Following these argumentation SOC stocks in croplands might even be underestimated.
 \item
@@
 Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
-\item
-  fertilizer interaction (--\textgreater{} ask LPJmLer again)
-\item
-  ``The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
-  current C stock to the future steady-state soil C stock calculated based on current conditions.'' Leading to the fact that our total stock results are highly uncertained.
-\item
-  diaggregation of manure with urban area is leeding to a lot of displaced manure (?) that is cut off
-\item
-  mismatches between FAO/LUH
+(In general the sensitivity to and the uncertainty of certain parameters might be very high.)
 \end{itemize}
 
-Shortcommings:
+\newpage
+
+Smaller points and shortcommings:
 
@@
 \item
-  pasture dynamics are neglected and treated as natveg / Note that manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.
+  Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
+Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer ammounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
 \item
-  no tillage adaptaion is neglected on cropland. Pastures are assumed to be not tilled at all (propably at least heavy managed pastures are tilled with some rotation)
+  Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
+(Note that also manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.)
 \item
-  irrigated areas are no crop specific and growing periods are not considered (since this is very complex to calculate average growing periods).
+  No tillage adaptaion is neglected on cropland due to less common adoption of no tillage and conservation agriculture.
+Pastures are assumed to be not tilled at all (propably only heavy managed pastures are tilled with some rotation)
+\item
+  Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
 Crop specific growing periods might be possible using LPJmL data.
+\item
+  flooded rice area are not represented correctly as parameterization does not hold true for flooded conditions.
 \item
@@
 \item
-  Within cropland we do not track area transitions, but rather look at statistical distributions of the crop functional types.
+  Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
 Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
+\item
+  The disaggregation of manure to uild-up areas (in the case of extensive monograstrics) is leeding to a lot of displaced manure (?) that is cut off
+\item
+  It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
+\item
+  ``The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
+  current C stock to the future steady-state soil C stock calculated based on current conditions.'' Leading to the fact that our total stock results are highly uncertained.
   \newpage
@@
 
-The conclusion goes here.
-You can modify the section name with \texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
+Outlook including perspective on mitigation (and soc enhancement in the future).
 \newpage
 
-\codedataavailability{use this to add a statement when having data sets and software code available} %% use this section when having data sets and software code available
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\hypertarget{append:subsection2mrfunctions}{%
+\subsection{table on method subsections to functions within R packages}\label{append:subsection2mrfunctions}}
+
+\hypertarget{append:Tableluh2fao2mag}{%
+\subsection{table on mapping LUH2FAO2MAG}\label{append:Tableluh2fao2mag}}
+
+\hypertarget{append:Tablekcr2kres}{%
+\subsection{kcr2kres mapping}\label{append:Tablekcr2kres}}
+
+\hypertarget{append:Tablec2dm}{%
+\subsection{carbon 2 dry matter}\label{append:Tablec2dm}}
+
+Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
+
+\hypertarget{append:TableclossAWMS}{%
+\subsection{closs in AWMS - Table}\label{append:TableclossAWMS}}
+
+\hypertarget{append:climatemap}{%
+\subsection{map on climate zone used for SCF}\label{append:climatemap}}
+
+\newpage
+
+\codedataavailability{Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/.
+Data used for the output can be found under .} %% use this section when having data sets and software code available
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Figures and tables in appendices}
-\subsection{Option 1}
-
-If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-They will be correctly named automatically.
-
-\subsection{Option 2}
-
-If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-
-\texttt{\textbackslash{}appendixfigures} needs to be added in front of appendix figures
-\texttt{\textbackslash{}appendixtables} needs to be added in front of appendix tables
-
-Please add \texttt{\textbackslash{}clearpage} between each table and/or figure. Further guidelines on figures and tables can be found below.
-Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-To rename them correctly to A1, A2, etc., please add the following commands in front of them:
-\noappendix
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky, and Popp revised paper.} %% optional section
+\authorcontribution{Karstens wrote code and paper build on work of Bodirsky (and ).
+Bodirsky, and Popp revised paper.} %% optional section
 

### 2020-08-14 SOCbudgetPaper/appendix.Rmd
update written part to first official version (woohoo)

@@
-* table on method subsections to functions within R packages
+# Appendix
 
-* table on mapping LUH2FAO2MAG
+## table on method subsections to functions within R packages {#append:subsection2mrfunctions}
 
-* kcr2kres mapping
+## table on mapping LUH2FAO2MAG {#append:Tableluh2fao2mag}
 
-* carbon 2 dry matter @Ma2018 (??) -> Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
+## kcr2kres mapping {#append:Tablekcr2kres}
 
-* closs in AWMS - Table
+## carbon 2 dry matter {#append:Tablec2dm}
+Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
 
-* map on climate zone used for SCF
+## closs in AWMS - Table {#append:TableclossAWMS}
+
+## map on climate zone used for SCF {#append:climatemap}

### 2020-08-14 SOCbudgetPaper/discussion.Rmd
update written part to first official version (woohoo)

--- 2020-06-16 SOCbudgetPaper/discussion.Rmd
@@
 
-Big points:
+- The natural land representation and therefor also SOC initialization is lacking proper parametrization of both decay behaviour and natural input (e.g.
+nitrogen and lignin content of litterfall).
+As all results are valued against these potential natural SOC the absolute values and also the changes over time have to be used with caution.
+Especially in boreal zone and in very dry regions the natural land representation might be far off, due to the focus for the parametrization for cropped areas.
+The good agreement of the stock change factors $F^{SCF}$ with the Tier 1 default values might serve as a hint that the natural land representation is not totally off.
+Boreal and very dry zone might be uncropped, leaving the results on SOC changes on cropland untouch from the bad borreal land representation.
 
-- SOC initialization is bad (Task: find a better representation)
-
-- natural land representation is lacking proper  parametrization of natural input (n, lg content of litterfall) (--> ask Christop again)
-
-- boreal zone and dry regions are not well represented (maybe because of the bad parameterization of the soil model for natural soils)
+- This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
+Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+It has been argued that for intensivly tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
+Following these argumentation SOC stocks in croplands might even be underestimated.
 
@@
 Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
+(In general the sensitivity to and the uncertainty of certain parameters might be very high.)
 
-- fertilizer interaction (--> ask LPJmLer again)
+<!--Tasks: 
+Find better natural land representation and initalization.
+Using LPJmL values might be complex since the split up into SOC subpools is needed.
+Getting better parametrization for the CENTURY based approach for decay behaviour and better values for nitrogen and lignin of natural litter fall.
+-->
 
-- diaggregation of manure with urban area is leeding to a lot of displaced manure (?) that is cut off
+\newpage
 
-- mismatches between FAO/LUH
+Smaller points and shortcommings:
 
-Shortcommings:
+- Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
+Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer ammounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
+
+- Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
+(Note that also manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.)
+
+- No tillage adaptaion is neglected on cropland due to less common adoption of no tillage and conservation agriculture.
+Pastures are assumed to be not tilled at all (propably only heavy managed pastures are tilled with some rotation)
+
+- Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
+Crop specific growing periods might be possible using LPJmL data.
+
+- flooded rice area are not represented correctly as parameterization does not hold true for flooded conditions.
 
@@
 
-- Within cropland we do not track area transitions, but rather look at statistical distributions of the crop functional types.
+- Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
 Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
+
+- The disaggregation of manure to uild-up areas (in the case of extensive monograstrics) is leeding to a lot of displaced manure (?) that is cut off
+
+- It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
+
+- "The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
+current C stock to the future steady-state soil C stock calculated based on current conditions." Leading to the fact that our total stock results are highly uncertained.

### 2020-08-14 SOCbudgetPaper/introduction.Rmd
update written part to first official version (woohoo)

@@
 \introduction
+
+Possible titles:
+
+* How management changed soil organic carbon on global croplands
+
+* Agricultural soil have lost 37 GtC topsoil carbon
+
+* Linking agricultural management data to soil modeleling – a global approach
 
@@
 Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles in this regard.
-
 Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today.
@@
 While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especial agricultural activities.
-
 Field-scale models (ref.
@@
 However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
-
 Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
@@
 We thereby provide the first global, spatially explicit SOC loss maps that consider detailed agricultural management.
-
 This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
-We continue with a detailed description of the global gridded management data used here, including (crop??) production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
+We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
 Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-
 In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
-
-Finally, we will discuss our findings - including their implications for SOC model development - and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+Finally, we will discuss our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.

### 2020-08-14 SOCbudgetPaper/methods.Rmd
update written part to first official version (woohoo)

@@
-# Method (50)
+# Method
 
 We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
-In the following chapter we outine the most important relationships and assumptions. See table \@ref(append:subsection2mrfunctions) for further information on corresponding code within the R packages.
+In the following chapter we outine the most important relationships and assumptions.
+Table \@ref(append:subsection2mrfunctions) provides further information on corresponding code within the R packages.
 
-## Carbon Stocks following (new) Tier 2 method {#sec:carbonbudget}
+## SOC Stocks and Stock changes following the Tier 2 steady-state method {#sec:carbonbudget}
 
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks for cropland and all other land represented by natural vegetation on half-degree resolution from 1975 to 2010.
+4 (@ipcc_2019_2019; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
 We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
@@
 (1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks, 
-(2) We account for land conversion by transferring SOC between land-use types and 
-(3) We estimate SOC stocks based on the stocks of the previous time period, the steady state stocks and the decay rate.
+(2) we account for land conversion by transferring SOC from and to an other land-use type representad as natural vegetation,
+(3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
 
@@
 
-#### Carbon Inputs to the Soil {-}
+**Carbon Inputs to the Soil**
 
-We account for different carbon input sources depending on the two land-use types we distinguish, croplands and natural vegetation (see table \@ref(tab:datasourceinputs)).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livst_manure)); for natural vegetation litterfall ([@LPJmL4_1]) litterfall is the only source of carbon.
+We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \@ref(tab:datasourceinputs)).
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livstmanure)); for natural vegetation litterfall [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
 
-Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \@ref(ipcc_2019_2019)).
+Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see @ipcc_2019_2019).
 For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
@@
 
-#### Soil Organic Carbon decay {-}
+**SOC decay**
 
@@
 
-For cropland we distinguish the effect of different tillage (see \@ref(#sec:tillage)) and irrigation (see \@ref(#sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
+For cropland we distinguish the effect of different tillage (see \@ref(sec:tillage)) and irrigation (see \@ref(sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
 Data sources as well as drivers considered for each land-use types are shown in table \@ref(tab:datasourcedecay).
@@
 
-### Total SOC stocks
+### Total SOC stocks and stock changes
 
-Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use type (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \@ref(eq:steadystate).
+Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use type (lut), each sub-pool (sub) and each annual time step (t) like
 
@@
 
-Reformulating equation \@ref(eq:steadystate), we obtain a massbalance equation as follows
+Reformulating this equation, we obtain a massbalance equation as follows
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \@ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
+SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
 (\#eq:steadystate2budget)
@@
 
+According to the IPCC guidelines SOC changes can be calculates as the difference of two (aufeinander folgenden) years (see Eq.
+5.0A in [@ipcc_2019_2019]).
+This however will also include naturally occured changes due to climatic variance over time.
+For our study we will define the absolute and relative SOC changes in relation to a potential natural state $SOC_{pnv}$ under the same climatic conditions at time $t$ in grid cell $i$, that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
+The absolute changes $\Delta SOC$ and relative changes $F^{SCF}$ a thus given by
+
+\begin{equation}
+\Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
+(\#eq:totalstock)
+\end{equation}
+
+Note that the absolution changes $\Delta SOC$ can be also interpreted as cummulative SOC emissions or as SOC depth of mankind; whereas relative changes $F^{SCF}$ can be denoted as stock change factors as defined within the IPCC guidelines of 2006 [@ipcc_2006_2006].
+
 ### Initialisation of SOC pools
@@
 
-We compile country-specific FAO production and cropland statistics ([@FAOSTAT]) to a harmonized and consistent data set.
+We compile country-specific FAO production and cropland statistics [@faostat_faostat_2016] to a harmonized and consistent data set.
 The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
@@
 
-### Landuse and Landuse Change (150) {#sec:landuse}
+### Landuse and Landuse Change {#sec:landuse}
 
-Land-use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each gridcell based on the country and year specific area harvested shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(append:Table_luh2fao2mag) for more  details on the crop group mapping).
+Land-use patterns are based on the Land-Use Harmonization 2 [@Hurtt2020] data set, which we aggregate from quarter degree to half degree resolution.
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see \@ref(append:Tableluh2fao2mag) for more  details on the crop group mapping).
 Land-use transitions are calculated as net area differences of the land-use data on half-degree.
 
-### Crop and Crop Residues (300) {#sec:residues}
+### Crop and Crop Residues Production {#sec:residues}
 
-#### Crop Production {-}
-Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop group specific half-degree production patterns.
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production ([@FAOSTAT]).
+**Crop Production**
+
+Using half-degree yield data from LPJmL [@schaphoff_lpjml4_2018] as well as half-degree cropland patterns (see \@ref(sec:landuse)) we compile crop group specific half-degree production patterns.
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production [@faostat_faostat_2016].
 Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
 
-#### Crop Residue Production {-}
-Crop residue production and management is based on a revised methodology of ([@bodirsky2012]) and key aspects are explained here given its central role for soil carbon modelling.
+**Crop Residue Production**
+
+Crop residue production and management is based on a revised methodology of [@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modelling.
 Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction ([@ipcc_2006_2006]).
-Note that deviating from [@bodirsky2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
-We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction [@ipcc_2006_2006].
+Note that deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals [@weindl], fuel or for material use.
 
-##### Burned Residues {-}
+**Burned Residues**
+
 A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
 Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process ([@ipcc_2006_2006]).
+Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process [@ipcc_2006_2006].
 
-##### Residue Usage {-}
-We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Table_kcr2kres)).
+**Residue Usage**
+
+We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Tablekcr2kres)).
 Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account.
@@
 
-#### Dry Matter to Carbon Transformation {-}
-To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Table_c2dm)) (@Ma2018).
+**Dry Matter to Carbon Transformation**
 
-### Livestock Distribution and Manure Excretion	(300) {#sec:livst_manure}
+To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Tablec2dm)).
+
+### Livestock Distribution and Manure Excretion {#sec:livstmanure}
 
@@
 
-#### Livestock Distribution {-}
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from by the approach of [@gilbert], and uses feed basket assumptions based on a revised methodology from @weindl.
+**Livestock Distribution**
+To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of [@robinson_mapping_2014], and uses feed basket assumptions based on a revised methodology from [@weindl].
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
@@
 
-#### Manure Excretion, Storage and Recycling {-}
-Manure production and management is based on a revised methodology of ([@bodirsky2012]) and is presented here due to its central role in soil carbon modelling.
-Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets ([@weindl]), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [(@ipcc_2019_2019)]).
+**Manure Excretion, Storage and Recycling**
+Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modelling.
+Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets [@weindl], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [@ipcc_2019_2019]).
 Depending on the feed system we assume manure to be handled in four different ways:
 All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Manure fuel shares are estimated using IPCC default values (@ippc_2006_2006).
+Manure fuel shares are estimated using IPCC default values [@ippc_2006_2006].
 Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
@@
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky2012] (see \@ref(append:Table_clossAWMS))).
+For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky_n2o_2012] (see \@ref(append:TableclossAWMS)).
 We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
 
-### Irrigation
+### Irrigation {#sec:irrigation}
 
@@
 
-### Tillage
+### Tillage {#sec:tillage}
 
 To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rule is based on the approach of [@porvolik], but neglects the (less common) adoption of no-tillage farming on agricultural land.
+This rule is based on the approach of [@porwollik_generating_2018], but neglects the (less common) adoption of no-tillage farming on agricultural land.
 
@@
 
-## Carbon Budget following Tier 1 (150) {#sec:tier1}
+## SOC Stocks and Stock changes following Tier 1 {#sec:tier1}
 
 Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019) and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.
+4 [@ipcc_2019_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.
 4 (@ipcc_2006_2006, @ipcc_2019_2019).
-Here, stocks are calculated via stock change factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-SCF are differentiated by different crop and management systems ($m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover SCF vary for different climatic zones ($c$) specified by the IPCC (see \@ref(append:climatemap)).
+Here, stocks are calculated via stock change factors (F^{SCF}) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+$F^{SCF}$ are differentiated by different crop, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover $F^{SCF}$ vary for different climatic zones ($c$) specified by the IPCC (see \@ref(append:climatemap)).
 The actual SOC stocks as thus calulated based on a given reference SOC stock by
@@
 \begin{equation}
-SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F_{\text{LU}_{c,m}} \cdot F_{\text{MG}_{c,m}} \cdot F_{\text{I}_{c,m}} \cdot A_{m,i} \qquad\forall\quad t,
+SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
 (\#eq:tier1)
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis we us the default SCF as a comparison and constistency check for our more detailed budget approach.
-
-According to [@ipcc_2019_2019] the formulation of absolute changes like
-<!-- plugin absolute chnage equation of refinement cropland chapter -->
-might be neccessary as well.
+For this analysis we use the default $F^{SCF}$ from the Tier 1 method of [@ipcc_2006_2006], and [@ipcc_2019_2019] as a comparison and constistency check for our more detailed budget approach.

### 2020-08-14 SOCbudgetPaper/results.Rmd
update written part to first official version (woohoo)

@@
 ```{r caption_SOCmaps, echo=FALSE}
-caption <- "(a) ... , (b) ... , (c) ... , (d) ... ."
+caption_SOCmaps <- "(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas.
+(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades (despite ongoing cropland expansion).
+(c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
+Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas."
 ```
 
-```{r SOCmaps, out.width = "18cm", fig.cap = caption}
+```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="H"}
 # All defaults
@@
 
-\newpage
+## global agricultural carbon flows and stocks
+
+```{r caption_FlowFig, echo=FALSE}
+caption_FlowFig <- "Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
+Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil)."
+```
+
+```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/OuFlowFig.png")
+```
 
@@
 
-```{r table, echo = FALSE}
-table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
-knitr::kable(table$naturalstate_table, "latex", digit=2, caption="SCF compared to potential natural state")
+```{r caption_SCFtable, echo=FALSE}
+caption_SCFtable <- "$F^{SOC}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{SOC}$ by additional 5-8 percent.
+Also: modelled $F^{SOC}$ have increased for all climates over time ."
 ```
 
-```{r table2, echo = FALSE}
-table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
-knitr::kable(table$comparison_table, "latex", digit=2, caption="SCF compared to IPCC Tier 1")
+```{r SCFtable, out.width = "16cm", fig.cap = caption_SCFtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
 ```
 
+<!-- ```{r table2, echo = FALSE}
+# table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
+# knitr::kable(table$comparison_table, "latex", digit=2, 
+#              caption="$F^{SOC}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{SOC}$ by additional 5-8 percent.
+Also: modelled $F^{SOC}$ have increased for all climates over time .")
+# ```
+# ```{r table, echo=FALSE}
+# table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
+# knitr::kable(table$naturalstate_table[3:10,], "latex", digit=2, 
+#              caption="S)
+# ``` -->
+
 \newpage
-
-## global agricultrual carbon flows and stocks

### 2020-10-06 ResultNotebooks/results_SOCcompare.Rmd
Mayor update of everything

@@
 
-rev       <- "rev0.6_mrSOC"
-outputdir <- "Output/Images"
+rev       <- "rev0.76baseline"
+outputdir <- "Output/"
 
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
 
-SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Landuse.rds")
+SoilCarbon <- read.magpie("Data/Baseline/SoilCarbon.rds")
+Landuse    <- read.magpie("Data/Baseline/Landuse.rds")
 
@@
 
-print(mbind(dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=c(1,3)),
-            dimSums(dimSums(SoilCarbon[,,"steadystate"], dim=3.3)* Landuse, dim=c(1,3))))
+print(dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=c(1,3)))
 
@@
   LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
-  LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
-  LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
+  #LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
+  #LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
   WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
@@
   SoilGrids  <- calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate="REG+GLO", try=TRUE)
+  LPJmL4  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL4Paper", aggregate="REG+GLO", try=TRUE)
 
@@
 
-plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red", xlim=c(1951,2017),  ylim=c(400000,1300000))
+plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red", xlim=c(1951,2017),  ylim=c(350000,1300000))
 lines(getYears(LPJ_IPCC, as.integer = TRUE), LPJ_IPCC["GLO",,1], type="l", col="green")
-lines(getYears(LPJ_car, as.integer = TRUE), LPJ_car["GLO",,], type="l", col="darkgreen")
-lines(getYears(LPJ_rev21, as.integer = TRUE), LPJ_rev21["GLO",,2], type="l", col="lightgreen")
+#lines(getYears(LPJ_car, as.integer = TRUE), LPJ_car["GLO",,], type="l", col="darkgreen")
+lines(getYears(LPJmL4, as.integer = TRUE), LPJmL4["GLO",,], type="l", col="lightgreen")
 lines(getYears(WISE, as.integer = TRUE), WISE["GLO",,], type="l", col="brown")
@@
 plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red")
+```
+
+```{r table ipcc scf}
+
+library(knitr)
+library(stats)
+library(luplot)
+library(reshape2)
+
+SOCStock <- mbind(glo_SOCStock, reg_SOCStock)
+glo_SOC  <- mbind(setNames(SOCStock[,"y2010",],"SOC Budget"), 
+                  setNames(LPJmL4[,"y2010",], "LPJmL4Paper"),
+                  setYears(setNames(WISE[,1,],"WISE"), "y2010"),
+                  setYears(setNames(SoilGrids[,1,],"SoilGrids"),"y2010"),
+                  setYears(setNames(GSOC[,1,],"GSOC"),"y2010"))
+
+#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+
+comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(glo_SOC/1000)), 
+                           Data1 ~ Region, value.var="Value", 
+                            row.names=levels(as.data.frame(glo_SOC)$Data1)))
+
+saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"gloSOC_comparison.rds"))
+
+```
+
+```{r kable2image}
+library(kableExtra)
+
+table <- readRDS("Output/gloSOC_comparison.rds")
+colnames(table[[1]])[1] <- "Model/Database"
+colnames(table[[1]])[5] <- "Global SOC in GtC"
+save_kable(kable_styling(kable(table$comparison_table[c(1,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
+         file="Output/Images/TableSOC_comparison.png")
+
+save_kable(kable_styling(kable(table$comparison_table[setdiff(1:13,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
+         file="Output/Images/TableRegSOC_comparison.png")
+
 ```

### 2020-10-06 ResultNotebooks/results_SOClitter.Rmd
Mayor update of everything

---
title: "Global SOC maps and emissions"
output:
  md_document:
    variant: markdown_github
  html_document:  
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrSOCbudget)
library(magpiesets)

rev       <- c(Default="rev0.71default", Till="rev0.72default", HiHi="rev0.71LGhigh_Nhigh", HiLo="rev0.71LGhigh_Nlow", LoHi="rev0.71LGlow_Nhigh", LoLo="rev0.71LGlow_Nlow")
outputdir <- "Output/Images"

```

Calc global numbers

```{r Carbon Emissions (glo)}

library(ggplot2)

col <- c("black","red","green","blue","brown","pink")

range        <- NULL
litter_param <- NULL

for(r in 1:length(rev)){
  SoilCarbon <- read.magpie(paste0("Data/",names(rev[r]),"/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/",names(rev[r]),"/Landuse.rds"))
  SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
  SOCStock        <- dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2)
  glo_SOCStock    <- dimSums(SOCStock, dim=1) 
  glo_SOCEmission <- (glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
  if(r==1) plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission, type="l", xlab="Time", ylab="Global SOC Debt in GtC", ylim=c(30,60))
  else     lines(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission, type="l", xlab="Time", ylab="Global SOC Debt in GtC", col=col[r])
  range <- cbind(range,range(glo_SOCEmission))
}

litter_param <- collapseNames(read.magpie("/home/kristine/mnt/rd3mod/inputdata/sources/SOCbudgetParam/all.csv"))
litter_param <- mbind(litter_param, setNames(litter_param[,,"Default"],gsub("Default","Till",getNames(litter_param[,,"Default"]))))

print(cbind(rev, max=round(t(range)[,2],1), min=round(t(range)[,1],1), 
            lgc=as.array(litter_param[1,1,"lgc_ratio"][,,names(rev)]), 
            nc =as.array(litter_param[1,1,"nc_ratio"][,,names(rev)]), 
            colour=col))
```

### 2020-10-06 ResultNotebooks/results_SOCmaps.Rmd
Mayor update of everything

@@
 
-rev       <- "rev0.7_mrSOC"
+rev       <- "rev0.76baseline"
 outputdir <- "Output/Images"
@@
 
-SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Landuse.rds")
+SoilCarbon <- read.magpie("Data/Baseline/SoilCarbon.rds")
+Landuse    <- read.magpie("Data/Baseline/Landuse.rds")
 
@@
 cropland_dens    <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)
-cropland_dens[cropland_dens==0] <- NA
+cropland_dens[Landuse[,,"crop"]==0] <- NA
 
@@
 library(oce) 
+library(scales)
 
@@
 
-SOCemis <- ggplot() + geom_line(data= as.data.frame(glo_SOCEmission[,seq(1975,2010,5),])[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
+squish_trans <- function(from, to, factor) {
+
+  trans <- function(x) {
+    if (any(is.na(x))) return(x)
+    # get indices for the relevant regions
+    isq <- x > from & x < to
+    ito <- x >= to
+    # apply transformation
+    x[isq] <- from + (x[isq] - from)/factor
+    x[ito] <- from + (to - from)/factor + (x[ito] - to)
+    return(x)
+  }
+
+  inv <- function(x) {
+    if (any(is.na(x))) return(x)
+    # get indices for the relevant regions
+    isq <- x > from & x < from + (to - from)/factor
+    ito <- x >= from + (to - from)/factor
+    # apply transformation
+    x[isq] <- from + (x[isq] - from) * factor
+    x[ito] <- to + (x[ito] - (from + (to - from)/factor))
+    return(x)
+  }
+  # return the transformation
+  return(trans_new("squished", trans, inv))
+}
+
+SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission[,seq(1975,2010,5),])[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
   labs( x = "Year", 
-        y = "Global cummulative \nSOC emission in GtC") +
-  scale_x_discrete(breaks=seq(1965,2010,5)) +
-  scale_y_continuous(limits=c(30,40), breaks=seq(30,40,5)) +
+        y = "Global SOC gap\n in GtC") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous(trans = squish_trans(-35,0,15),
+                     limits = c(-45,0),
+                     breaks = c(0, -15, -35, -40, -45)) +
   theme(rect=element_blank(),
         axis.title.y= element_text(size=legend_size),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8))
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8)) +
+  annotate(geom="text", label="SOC debt", x=factor(1995), y=-35, size = 3) +
+  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -39, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
 
@@
     scale_fill_stepsn(limits=c(0,2), colours = cols, 
-                      name = expression("ratio of cropland SOC to potential natural SOC"), 
+                      name = expression(" cropland SOC stock change factor (unitless)"), 
                       breaks=c(0.3,0.6,0.9,1.1,1.4,1.7,2),
@@
     scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression("cropland SOC change in tons of carbon per hectar"), 
+                      name = expression("cropland SOC gap in tons of carbon per hectar"), 
                       values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,

### 2020-10-06 ResultNotebooks/results_SOCscenario.Rmd
Mayor update of everything

---
title: "Global SOC scenarios"
output:
  md_document:
    variant: markdown_github
  html_document:  
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrSOCbudget)
library(magpiesets)

rev <- "rev0.76baseline"

setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")

#revisions   <- c(Baseline="rev0.74baseline", Till="rev0.74tillage_mixedtill", ResRe="rev0.74residue_freeze1975", ManRe="rev0.74manure_freeze1975", Yield="rev0.74yield_freeze1975")
filesNeeded <- c("SoilCarbon.rds","Landuse.rds")

revisions   <- c(Baseline="rev0.76baseline", Allon="rev0.76allon_freeze1975", Allon2="rev0.76allon2_freeze1975",        ResRe="rev0.76rrecycle_freeze1975",
                Till="rev0.76tillage_mixedtill", ResSta="rev0.76residue_freeze1975", Yield="rev0.76yield_freeze1975", ManRe="rev0.76manure_freeze1975")

# for(item in revisions){
#    dir.create(path=paste0("Data/",names(which(revisions[]==item))))
#    untar(paste0(getConfig("outputfolder"),"/",item, "_h12_carbonbudget.tgz"), files=filesNeeded, exdir=paste0("Data/", names(which(revisions[]==item))))#, extras="--keep-newer-files ")
# }

outputdir <- "Output/Images"

```

Calc global numbers

```{r Carbon Emissions (glo)}

library(ggplot2)

glo_SOCStock <- NULL

for(r in 1:length(revisions)){
  SoilCarbon <- read.magpie(paste0("Data/",names(revisions[r]),"/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/",names(revisions[r]),"/Landuse.rds"))
  SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
  SOCStock        <- toolConditionalReplace(dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2), "is.nan()",0)
  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), dim=3.1, nm=names(revisions[r]), add="scenario")) 
}

# 
```

```{r plot Carbon Emissions (glo)}

plot_rev <- c(3,5,6,8)

names(revisions) <- c("Historical Management", 
                      "Allon",    
                      "Constant management from 1975",   
                      "ResRe", 
                      "Constant tillage area from 1975",
                      "Constant residue input from 1975",
                      "Yield", 
                      "Constant manure application from 1975")

colr <- c("black","green","red","brown","pink","darkgreen","yellow","blue")
glo_SOCEmission <- collapseNames(glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
glo_SOCShare    <- collapseNames( toolConditionalReplace(glo_SOCStock[,,"actualstate"]/glo_SOCStock[,,"naturalstate"], "is.na()",0) )

getNames(glo_SOCEmission) <- names(revisions)
plot(getYears( glo_SOCEmission, as.integer = TRUE), -glo_SOCEmission[,,1], type="l", xlab="Time", ylab="Global SOC gap in GtC", ylim=c(-45,-35))
legend("topleft", legend=names(revisions[c(1,plot_rev)]), col=colr[c(1,plot_rev)], lty=1, cex=0.8)
for(r in plot_rev) lines(getYears(glo_SOCEmission, as.integer = TRUE), -glo_SOCEmission[,,names(revisions[r])], type="l", col=colr[r])
print(cbind(colr[c(1,plot_rev)], names(revisions[c(1,plot_rev)])))
print(glo_SOCStock[,,"actualstate"])

#plot(getYears( glo_SOCEmission, as.integer = TRUE), dimSums(Landuse[,,"crop"], dim=1), type="l")
#plot(getYears( glo_SOCEmission, as.integer = TRUE), dimSums(glo_SOCStock[,,"Baseline.actualstate"], dim=1), type="l")

```

```{r ggplot}

legend_size <- 10

SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission[,seq(1975,2010,5),c(1,plot_rev)]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
  labs( x = "Year", 
        y = "Global SOC gap in GtC",
        color="") +
  scale_x_discrete(breaks=seq(1975,2010,5)) +
  scale_y_continuous(limits = c(-45,-35),
                     breaks = c(-36, -38, -40,- 42, -44)) +
  theme(rect=element_blank(),
        axis.title.y= element_text(size=legend_size),
        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8))

SOCemis

ggsave(paste0(outputdir,"/scenario.png"), plot=SOCemis)

```

### 2020-10-06 ResultNotebooks/results_SOCtable.Rmd
Mayor update of everything

--- 2020-06-16 ResultNotebooks/results_SOCtable.Rmd
@@
 
-rev <- "rev0.6_mrSOC"
+rev       <- "rev0.76baseline"
+outputdir <- "Output/"
 
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/Baselin", extras="--keep-newer-files ")
 
-SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Landuse.rds")
+SoilCarbon <- read.magpie("Data/Baseline/SoilCarbon.rds")
+Landuse    <- read.magpie("Data/Baseline/Landuse.rds")
 
@@
 
-SOCDens           <- dimSums(SoilCarbon[,,c("steadystate","actualstate","naturalstate")], dim=3.3) 
+SOCDens           <- dimSums(SoilCarbon[,,c("actualstate","naturalstate")], dim=3.3) 
 SOCDens[,,"naturalstate.crop"] <- SOCDens[,,"actualstate.crop"]
@@
                      dimSums(climateIPCC * SOCDens[,,"natveg"] * Landuse[,,"crop"], dim=1)
+
+getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))
 
@@
 
-steadystate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"steadystate"])), 
-                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")
-
 saveRDS(list(naturalstate_table = naturalstate_table,
-             actualstate_table  = actualstate_table, 
-             steadystate_table  = steadystate_table), "Output/SFC_climatezone.rds")
+             actualstate_table  = actualstate_table), paste0(outputdir, "SFC_climatezone.rds"))
 
@@
 
-CShare_ipcc <- new.magpie("GLO","y2010",outer(c("IPCC2006","IPCC2019"),unique(koepp2IPCC$ipcc_reduced), paste, sep="."))
+CShare_ipcc <- IShare_ipcc <- new.magpie("GLO","y2010",outer(c("IPCC2006","IPCC2019"),unique(koepp2IPCC$ipcc_reduced), paste, sep="."))
 
-CShare_ipcc[,,"IPCC2006.tropical_moist"]  <- "0.48"
-CShare_ipcc[,,"IPCC2006.tropical_dry"]    <- "0.58-0.64" 
-CShare_ipcc[,,"IPCC2006.temperate_moist"] <- "0.69"
-CShare_ipcc[,,"IPCC2006.temperate_dry"]   <- "0.80"
+CShare_ipcc <- IShare_ipcc <- new.magpie("GLO","y2010",outer(c("IPCC2006","IPCC2019"),c("tropical_moist","tropical_dry","temperate_dry","temperate_moist"), paste, sep="."))
 
-CShare_ipcc[,,"IPCC2019.tropical_moist"]  <- "0.83"
-CShare_ipcc[,,"IPCC2019.tropical_dry"]    <- "0.92" 
-CShare_ipcc[,,"IPCC2019.temperate_moist"] <- "0.69-0.70"
-CShare_ipcc[,,"IPCC2019.temperate_dry"]   <- "0.76-0.77"
+CShare_ipcc <- add_dimension(CShare_ipcc, dim=3.2, nm="default", add="subsystem")
+IShare_ipcc <- add_dimension(IShare_ipcc, dim=3.2, nm="low_input", add="subsystem")
 
-comparison <- mbind(CShare_ipcc, 
+CShare_ipcc[,,"IPCC2006.default.tropical_moist"]  <- "0.48"
+CShare_ipcc[,,"IPCC2006.default.tropical_dry"]    <- "0.58-0.64" 
+CShare_ipcc[,,"IPCC2006.default.temperate_moist"] <- "0.69"
+CShare_ipcc[,,"IPCC2006.default.temperate_dry"]   <- "0.80"
+
+CShare_ipcc[,,"IPCC2019.default.tropical_moist"]  <- "0.83"
+CShare_ipcc[,,"IPCC2019.default.tropical_dry"]    <- "0.92" 
+CShare_ipcc[,,"IPCC2019.default.temperate_moist"] <- "0.69-0.70"
+CShare_ipcc[,,"IPCC2019.default.temperate_dry"]   <- "0.76-0.77"
+
+#Low input regime factor
+IShare_ipcc[,,"IPCC2006.low_input.tropical_moist"]  <- "0.44"
+IShare_ipcc[,,"IPCC2006.low_input.tropical_dry"]    <- "0.55-0.61" 
+IShare_ipcc[,,"IPCC2006.low_input.temperate_moist"] <- "0.66"
+IShare_ipcc[,,"IPCC2006.low_input.temperate_dry"]   <- "0.74"
+
+IShare_ipcc[,,"IPCC2019.low_input.tropical_moist"]  <- "0.76"
+IShare_ipcc[,,"IPCC2019.low_input.tropical_dry"]    <- "0.87" 
+IShare_ipcc[,,"IPCC2019.low_input.temperate_moist"] <- "0.66-0.67"
+IShare_ipcc[,,"IPCC2019.low_input.temperate_dry"]   <- "0.70-0.71"
+
+getNames(CShare_ipcc) <- gsub("_", " ", getNames(CShare_ipcc))
+getNames(IShare_ipcc) <- gsub("_", " ", getNames(IShare_ipcc))
+
+comparison <- mbind(CShare_ipcc, IShare_ipcc,
+                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),"y2010"),
+                                  nm="SOC budget.1990",add="model.year"),2),
                     round(add_dimension(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),
-                                  add="SOCbudget",nm="model"),2))
+                                  nm="SOC budget.2010",add="model.year"),2))
 
-print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+
+comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(comparison)), 
+                           Data1 + Data2 ~ Data3, value.var="Value", 
+                            row.names=levels(as.data.frame(comparison)$Data1)))
+
+saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"SFC_comparison.rds"))
 
 ```
+
+```{r kable2image}
+library(kableExtra)
+
+table <- readRDS("Output/SFC_comparison.rds")
+colnames(table[[1]]) <- gsub("\\.", " ", colnames(table[[1]]))
+colnames(table[[1]])[1] <- "Source"
+colnames(table[[1]])[2] <- "Subsystem/Year"
+save_kable(kable_styling(kable(table$comparison_table, "latex", digit=2), latex_options = c("striped", "scale_down")), 
+         file="Output/Images/TableSCF_comparison.png")
+
+table <- readRDS("Output/SFC_climatezone.rds")
+colnames(table[[1]]) <- gsub("_", " ", colnames(table[[1]]))
+colnames(table[[1]][1]) <- ""
+as_image(kable_styling(kable(table$naturalstate_table[3:10,], "latex", digit=2), latex_options = c("striped", "scale_down")), 
+         width=4, file="Output/Images/TableSCF_timeline.png")
+```

### 2020-10-06 SOCbudgetPaper/SOCbudget.Rmd
Mayor update of everything

@@
 output:
+  rticles::copernicus_article: default
+  md_document:
+    variant: markdown_github
   bookdown::pdf_book:
@@
     md_extensions: +raw_attribute
-  md_document:
-    variant: markdown_github
-  rticles::copernicus_article: default
   bookdown::word_document2: default

### 2020-10-06 SOCbudgetPaper/SOCbudget.tex
Mayor update of everything

@@
 % E&G Quaternary Science Journal (egqsj)
+% European Journal of Mineralogy (ejm)
 % Fossil Record (fr)
@@
 % Journal of Sensors and Sensor Systems (jsss)
+% Magnetic Resonance (mr)
 % Mechanical Sciences (ms)
@@
 % The Cryosphere (tc)
+% Weather and Climate Dynamics (wcd)
 % Web Ecology (we)
@@
 
-\title{Management induced changes of soil organic carbon on global croplands}
+\title{Management induced changes of soil organic carbon on global
+croplands}
 
@@
 
-\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
+\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam,
+Germany}
 \affil[2]{Your affiliation}
-
-%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
 
@@
 
-\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}
+\correspondence{Kristine\ Karstens\ (kristine.karstens@pik-potsdam.de)}
 
@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of larges carbon stocks on earth.
-It is three times larger than the biospheric pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
-Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
-While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
-Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
-We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
-We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
-We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
+Soil organic carbon (SOC) is one of larges carbon stocks on earth. It is
+three times larger than the biospheric pool and more than twice as big
+as the athmospheric pool, when looking into the first meter of the earth
+soil profile. Human cropping activties led and still lead to a depletion
+of SOC though, which are so far not well represented in global
+assessments of historic carbon emissions. While SOC models often
+represent well the biochemical processes that lead to the accumulation
+and decay of SOC, the management decisions driving these biophysical
+processes are still little investigated on the global scale. Here we
+create a spatial explicit data set for agricutural management on
+cropland, especially for crop residue and manure management, based on
+global historic production (FAOSTAT), and land-use (LUH2) data; and
+combine it with the IPCC Tier 2 approach to create a half degree
+resolution SOC stock changes on mineral soils. We estimate that due to
+arable farming soils have lost over 37 GtOC of which 4 GtOC have been
+regained within the period 1975-2010. We show that, our results on
+global scale based on Tier 2 IPCC methodolgy are in good agreement with
+Tier 1 default assumptions. We also find that SOC is very sensitive to
+management decision such as residue recycling indicating the nessessity
+to incorporated better management data in soil models.
 \end{abstract}
 
-\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
+\copyrightstatement{The author's copyright for this publication is
+transferred to institution/company.}
 
@@
 
-Possible titles:
+Soil Organic Carbon (SOC), the amount of organic carbon stored through
+the earth's soil, is the largest terrestrial carbon pool, exceeding the
+carbon in the atmospheric and vegetation pools multiple times (Batjes).
+As such, even small changes in drivers of SOC may thus lead to
+substantial shifts in earth carbon cycle and influence the amount of CO2
+in the atmosphere (ref. permafrost melting). The specific amount of
+carbon stored in the soil is, however, uncertain, with estimates ranging
+from 1500 to 2400 GtC for the first meter of the soil profile (Batjes,
+1996; more). The quality of SOC maps has markedly improved in recent
+decades, along with better understanding of the factors driving the
+global magnitude, distribution, and dynamics of SOC pools. Natural
+properties like climatic, biophysical, and landscape characteristics
+clearly play the most important roles to determine SOC variations over
+space and time. Human intervention, including land cover change and land
+management, has however added a further driver to SOC change, which
+alters terrestrial carbon pools in much shorter time scales and is
+likely one of the most dominant driver of SOC changes on managed land
+today (cite). Recent studies identify the anthropogenic SOC debt for the
+first meter of the soil profile at around 116 GtC (Sanderman et al.),
+compared to previous estimates between 60-130 GtC (Lal, 2006). Other
+studies have focused more closely on spatially disaggregation of SOC
+changes via advanced digital soil mapping techniques (S-World;
+Stoorvogel 2, 2017) or better representation of biogeochemical processes
+within SOC dynamics (Hararuk, 20XX). While providing a rough estimate of
+the order of magnitude of change, these studies lack a detailed
+consideration of land management, especially of agricultural activities.
+Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to
+capture these land management impacts by using detailed information on
+crop yield levels, fertilizer inputs and various other on-farm measures.
+However, due to the lack of comprehensive global management data as
+input to these models, scaling up to the global extent remains a complex
+challenge. Our study combines an spatially explicit estimate of
+agricultural management data on the global level with a 3-pool SOC model
+parametrized for global croplands. This allows us to estimate SOC stock
+change factors, as well as organic carbon flow dynamics within the
+agricultural system. We thereby consider change in SOC caused by
+historical land cover change as well as of different agricultural
+management practices, including residue recycling, manure amendments,
+irrigation, and tillage. We provide the first global, spatially explicit
+SOC loss maps that considers detailed agricultural management. This
+paper will first introduce in the methods section the basic concept of
+SOC dynamics as applied in this study. We continue with a detailed
+description of the global gridded management data used here, including
+crop production levels, residue recycling rates, manure amendments, and
+the adoption of irrigation and tillage practices. Lastly, we shortly
+refer to the concept of stock change factors as outlined in the Tier 1
+approach of the IPCC guidelines. In our results section we focus on the
+SOC dynamics of global croplands by (1) analysing the spatially explicit
+distribution and depletion of SOC from 1975 to 2010, (2) comparing
+impacts of different management effects on SOC emissions and (3)
+quantify global agricultural carbon flows and stocks to compare the
+importance of various management aspects. Finally, we discuss and
+compare our findings -- including their implications for SOC model
+development -- and conclude with an outlook on the ability of SOC
+management to mitigate climate change and contribute to negative
+greenhouse gas emissions. \newpage
 
-\begin{itemize}
-\item
-  How management changed soil organic carbon on global croplands
-\item
-  Agricultural soil have lost 37 GtC topsoil carbon
-\item
-  Linking agricultural management data to soil modeleling -- a global approach
-\end{itemize}
+\section{Method}
 
-Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and biospheric pools multiple times (Bathjes).
-As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref.
-permafrost melting).
-The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
-The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
-Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles in this regard.\\
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today.
-Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC (Lal, 2006).
-Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX).
-While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especial agricultural activities.\\
-Field-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm activities.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.\\
-Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
-This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
-We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation, and tillage.
-We thereby provide the first global, spatially explicit SOC loss maps that consider detailed agricultural management.\\
-This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
-We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
-Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
-Finally, we will discuss our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.\\
-\newpage
-
-\hypertarget{method}{%
-\section{Method}\label{method}}
-
-We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
-In the following chapter we outine the most important relationships and assumptions. Table \ref{append:subsection2mrfunctions} provides further information on corresponding code within the R packages.
+We compile calculations as open-source R packages available at
+github.com/pik-piam/mrcommons (management related functions) and
+github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which
+are both based on the MADRaT package (package citation?), a framework
+which aims to improve reproducibility and transparency in data
+processing. In the following chapter we outine the most important
+relationships and assumptions. Table @ref(append:subsection2mrfunctions)
+provides further information on corresponding code within the R
+packages.
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{SOC Stocks and Stock changes following the Tier 2 steady-state method}\label{sec:carbonbudget}}
+\subsection{SOC Stocks and Stock changes following the Tier 2
+steady-state method}\label{sec:carbonbudget}}
 
-Following the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
-We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
-Therefore we conduct the following three steps within each yearly timestep:
-(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks,
-(2) we account for land conversion by transferring SOC from and to an other land-use type representad as natural vegetation,
-(3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
+Following the tier 2 approach of the refinement of IPCC guidelines
+vol.~4 (\citet{ipcc_2019_2019}; short Tier 2 steady-state method) , we
+estimate soil organic carbon (SOC) stocks and their change over time for
+cropland on half-degree resolution from 1975 to 2010. We assume the
+current SOC state converges towards a steady state, which itself is
+depending on biophysical, climatic and agronomic conditions. Therefore
+we conduct the following three steps within each yearly timestep: (1) We
+calculate annual land-use (sub-)type-specific steady states and decay
+rates for SOC stocks, (2) we account for land conversion by transferring
+SOC from and to an other land-use type representad as natural
+vegetation, (3) we estimate SOC stocks and changes based on the stocks
+of the previous time period, the steady state stocks and the decay rate.
 
-\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
-\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}
+\subsubsection{Steady-state SOC stocks and decay rates}
 
-In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
-\begin{equation}
+In a simple first order kinetic approach the steady-state soil organic
+carbon stocks \(SOC^{eq}\) are given by \begin{equation}
 SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
-\label{eq:inoutflow}
-\end{equation}
-with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cell indices and \(t\) years.
-We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
-Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
+(\#eq:inoutflow)
+\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil,
+\(k\) denoting the soil organic carbon decay rate; as well as \(i\)
+representing grid cell indices and \(t\) years. We use for our
+calculations the Tier 2 steady-state method, which assumes three soil
+carbon sub-pools (active, slow and passive) and interactions between
+them. Annual carbon inflow to each sub-pool and annual decay rates of
+each sub-pool are the key components to determining steady-state SOC
+stocks.
 
@@
 
-We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \ref{tab:datasourceinputs}).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see \ref{sec:livstmanure}); for natural vegetation litterfall \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
+We account for different carbon input sources depending on the two
+land-use types we distinguish: croplands and natural vegetated land as
+representative for all other land use (see table
+@ref(tab:datasourceinputs)). Carbon sources for cropland are recycled
+crop residues, below ground biomass of crops (for both see
+@ref(sec:residues)) and recycled manure (see @ref(sec:livstmanure)); for
+natural vegetation litterfall \citep{schaphoff_lpjml4_2018} is the only
+source of carbon inflow to the soil.
 
-Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citet{ipcc_2019_2019}).
-For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that both the amount of carbon as well as its structural composition determine the effective inflow.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
+Following the IPCC carbon accounting methodology, carbon inputs are
+disaggregated into metabolic and structural components depending on
+their lignin and nitrogen content (see \citet{ipcc_2019_2019}). For each
+component the sum of all carbon input sources is allocated to the
+respective SOC sub-pools via transfer coefficients. This implies that
+both the amount of carbon as well as its structural composition
+determine the effective inflow. Data sources for all considered carbon
+inputs as well as for lignin and nitrogen content can be found in table
+@ref(tab:datasourceinputs).
 
@@
 
-The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
-Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
-Thus \(k_{sub}\) is given by
+The sub-pool specific decay rates are influenced by climatic conditions,
+biophysical and biochemical soil properties as well as management
+factors that all vary over time (t) and space (i). Following the Tier 2
+steady-state method we consider temperature (temp), water (wat), sand
+fraction (sf) and tillage (till) effects to account for spatial
+variation of decay rates. Thus \(k_{sub}\) is given by
 
@@
 
-For cropland we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as drivers considered for each land-use types are shown in table \ref{tab:datasourcedecay}.
-To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
+For cropland we distinguish the effect of different tillage (see
+@ref(sec:tillage)) and irrigation (see @ref(sec:irrigation)) practices
+on decay rates, whereas for natural vegetation, we assume rainfed and
+non-tilled conditions. Data sources as well as drivers considered for
+each land-use types are shown in table @ref(tab:datasourcedecay). To
+account for variations of decay rates within each grid cell due to
+different tillage and irrigation regimes, average rates based on area
+shares are calculated.
 
@@
  \belowtable{}
- \label{tab:datasourcedecay}
+ (\#tab:datasourcedecay)
  \end{table*}
 
-\hypertarget{soc-transfer-between-land-use-types}{%
-\subsubsection{SOC transfer between land-use types}\label{soc-transfer-between-land-use-types}}
+\subsubsection{SOC transfer between land-use types}
 
-We calculate SOC stocks based on the area shares of land-use types (lut) within the half-degree grid cells (i).
-If land is converted from one land-use type into others (!lut), the respective share of the SOC stocks is reallocated.
-We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
+We calculate SOC stocks based on the area shares of land-use types (lut)
+within the half-degree grid cells (i). If land is converted from one
+land-use type into others (!lut), the respective share of the SOC stocks
+is reallocated. We account for land conversion at the beginning of each
+time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
 
@@
 SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
-\label{eq:ctransfer}
+(\#eq:ctransfer)
 \end{equation}
 
-with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land-use type \(lut\).
-Note that \(!lut\) denotes the sum over all other land-use types, which decreases in the specific time step \(t\).
-Data sources and methodology on land-use states and changes are described in \ref{sec:landuse}.
+with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
+expansion for a given land-use type \(lut\). Note that \(!lut\) denotes
+the sum over all other land-use types, which decreases in the specific
+time step \(t\). Data sources and methodology on land-use states and
+changes are described in @ref(sec:landuse).
 
-\hypertarget{total-soc-stocks-and-stock-changes}{%
-\subsubsection{Total SOC stocks and stock changes}\label{total-soc-stocks-and-stock-changes}}
+\subsubsection{Total SOC stocks and stock changes}
 
-Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use type (lut), each sub-pool (sub) and each annual time step (t) like
+Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
+calculated steady-state stock \(SOC^{eq}\) for each land-use type (lut),
+each sub-pool (sub) and each annual time step (t) like
 
@@
 SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
-\label{eq:SOCstate}
+(\#eq:SOCstate)
 \end{equation}
@@
 SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
-\label{eq:steadystate2budget}
+(\#eq:steadystate2budget)
 \end{equation}
@@
 SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
-\label{eq:totalstock}
+(\#eq:totalstock)
 \end{equation}
 
-According to the IPCC guidelines SOC changes can be calculates as the difference of two (aufeinander folgenden) years (see Eq.
-5.0A in \citep{ipcc_2019_2019}).
-This however will also include naturally occured changes due to climatic variance over time.
-For our study we will define the absolute and relative SOC changes in relation to a potential natural state \(SOC_{pnv}\) under the same climatic conditions at time \(t\) in grid cell \(i\), that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
-The absolute changes \(\Delta SOC\) and relative changes \(F^{SCF}\) a thus given by
+According to the IPCC guidelines SOC changes can be calculates as the
+difference of two (aufeinander folgenden) years (see Eq. 5.0A in
+\citep{ipcc_2019_2019}). This however will also include naturally
+occured changes due to climatic variance over time. For our study we
+will define the absolute and relative SOC changes in relation to a
+potential natural state \(SOC_{pnv}\) under the same climatic conditions
+at time \(t\) in grid cell \(i\), that is based on the natural
+vegetation SOC calculations as defined above without accounting for land
+conversion from cropland at any time. The absolute changes
+\(\Delta SOC\) and relative changes \(F^{SCF}\) a thus given by
 
@@
 \Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
-\label{eq:totalstock}
+(\#eq:totalstock)
 \end{equation}
 
-Note that the absolution changes \(\Delta SOC\) can be also interpreted as cummulative SOC emissions or as SOC depth of mankind; whereas relative changes \(F^{SCF}\) can be denoted as stock change factors as defined within the IPCC guidelines of 2006 \citep{ipcc_2006_2006}.
+Note that the absolution changes \(\Delta SOC\) can be also interpreted
+as the SOC gap due to human cropping activits; whereas relative changes
+\(F^{SCF}\) can be denoted as stock change factors as defined within the
+IPCC guidelines of 2006 \citep{ipcc_2006_2006}. Note that the SOC gap is
+to the negated cummulative SOC emission also refered to as SOC debt
+(cite Sanderman) of mankind.
 
-\hypertarget{initialisation-of-soc-pools}{%
-\subsubsection{Initialisation of SOC pools}\label{initialisation-of-soc-pools}}
+\subsubsection{Initialisation of SOC pools}
 
-To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions will decrease.
+To initialize all SOC sub-pools we assume that cropped land as well as
+natural vegetation are in a steady state for the specific configuration
+present within the initialization year 1961. We assume after a spin-up
+period of 15 years reliable results start from 1975 and improve over
+time, since dependency on the initial conditions will decrease.
 
@@
 \hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data on 0.5 degree resolution}\label{sec:agrimanagement}}
+\subsection{Agricultural management data on 0.5 degree
+resolution}\label{sec:agrimanagement}}
 
-We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
+We compile country-specific FAO production and cropland statistics
+\citep{faostat_faostat_2016} to a harmonized and consistent data set.
+The data is prepared in 5 year time steps from 1965 to 2010, which,
+together with the spin up phase, restricts our analysis to time span
+from 1975 to 2010. For all the following data, if not declared
+differently, we interpolate values linearly between the time steps and
+hold them constant before 1965.
 
@@
 
-Land-use patterns are based on the Land-Use Harmonization 2 \citep{Hurtt2020} data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see \ref{append:Tableluh2fao2mag} for more details on the crop group mapping).
-Land-use transitions are calculated as net area differences of the land-use data on half-degree.
+Land-use patterns are based on the Land-Use Harmonization 2
+\citep{Hurtt2020} data set, which we aggregate from quarter degree to
+half degree resolution. We disaggregate the five different cropland
+subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17
+crop groups, applying the relative shares for each grid cell based on
+the country and year specific area harvested shares of FAOSTAT data
+\citep{faostat_faostat_2016} (see @ref(append:Tableluh2fao2mag) for more
+details on the crop group mapping). Land-use transitions are calculated
+as net area differences of the land-use data on half-degree.
 
@@
 
-Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns.
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
-Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
+Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as
+well as half-degree cropland patterns (see @ref(sec:landuse)) we compile
+crop group specific half-degree production patterns. We calibrate
+cellular yields with a country-level calibration factor for each crop
+group to meet historical FAOSTAT production
+\citep{faostat_faostat_2016}. Note that by using physical cropland areas
+we account for multiple crop harvest events as well as for fallows.
 
@@
 
-Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modelling.
-Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios (\(RS\)) as follows
+Crop residue production and management is based on a revised methodology
+of \citep{bodirsky_n2o_2012} and key aspects are explained here given
+its central role for soil carbon modelling. Starting from harvested crop
+production (\(CP\)) estimates and their respective harvested crop area
+(\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\))
+residual biomass using crop group (\(cg\)) specific harvest indices
+(\(HI\)) and root:shoot ratios (\(RS\)) as follows
 
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction \citep{ipcc_2006_2006}.
-Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
-We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals \citep{weindl}, fuel or for material use.
+Following the IPCC guidelines, we split the harvest index into a yield
+and an area dependend fraction \citep{ipcc_2006_2006}. Note that
+deviating from \citep{bodirsky_n2o_2012} we use harvested instead of
+physical crop area to account for increased residue biomass due to
+multiple cropping and decreased amounts on fallow land. We assume that
+all BGR are recycled to the soil, whereas AGR can be burned or harvested
+for other purposes such as feeding animals \citep{weindl}, fuel or for
+material use.
 
@@
 
-A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
-Following \citep{smil1999}) we assume 25\% burn share for low-income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process \citep{ipcc_2006_2006}.
+A fixed share of the AGR is assumed to be burned on field depending on
+the per-capita income of the country. Following \citep{smil1999}) we
+assume 25\% burn share for low-income countries according to worldbank
+definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income
+(\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for
+all middle-income countries depending on their per-capita income.
+Depending on the crop group 80--90\% of the residue carbon burned on the
+fields are lost within the combustion process \citep{ipcc_2006_2006}.
 
@@
 
-We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \ref{append:Tablekcr2kres}).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see \citep{weindl}), taking available AGR biomass as well as livestock productivity into account.
-We estimate, for low-income countries, a material-use share for straw residues of 5\% and a fuel-share of 10\% for all used residues groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
-The remaining AGR as well as all BGR are assumend to be recycled to the soil.
-We limit high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
+We compile out of our 17 crop groups, three residue groups (straw,
+high-lignin and low-lignin residues) with additional demand for other
+purposes and one residue with no double use (see
+@ref(append:Tablekcr2kres)). Residue feed demand for five different
+livestock groups is based on country- and residue-group-specific feed
+baskets (see \citep{weindl}), taking available AGR biomass as well as
+livestock productivity into account. We estimate, for low-income
+countries, a material-use share for straw residues of 5\% and a
+fuel-share of 10\% for all used residues groups in low-income countries.
+For high-income countries, no withdrawal for material or fuel use is
+assumend, and middle-income countries use shares are linearly
+interpolated based on per-capita income. The remaining AGR as well as
+all BGR are assumend to be recycled to the soil. We limit high recycling
+shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
 
@@
 
-To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \ref{append:Tablec2dm}).
+To transform dry matter estimates into carbon, we compiled crop-group
+and plant part specific carbon to dry matter (c:dm) ratios (see
+@ref(append:Tablec2dm)).
 
 \hypertarget{sec:livstmanure}{%
-\subsubsection{Livestock Distribution and Manure Excretion}\label{sec:livstmanure}}
+\subsubsection{Livestock Distribution and Manure
+Excretion}\label{sec:livstmanure}}
 
-We assume that manure is applied at its excretion place, leaving the livestock distribution the driving factor for the spatial pattern of manuring.
+We assume that manure is applied at its excretion place, leaving the
+livestock distribution the driving factor for the spatial pattern of
+manuring.
 
-\textbf{Livestock Distribution}
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014}, and uses feed basket assumptions based on a revised methodology from \citep{weindl}.
-We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-For ruminants, we assume that livestock is located where the production of feed occurs.
-We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
-For low-income countries, we assume extensive production systems.
-We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio.
-Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production, assuming that feed availability is the most driving factor for livestock location.
+\textbf{Livestock Distribution} To disaggregate country level FAOSTAT
+livestock production values to the half-degree scale, we use the
+following rule-based assumptions, drawing from the approach of
+\citep{robinson_mapping_2014}, and uses feed basket assumptions based on
+a revised methodology from \citep{weindl}. We differentiate between
+ruminant and monogastric systems, as well as extensive and intensive
+systems. For ruminants, we assume that livestock is located where the
+production of feed occurs. We distingush between grazed pasture, which
+is converted into livestock products from extensive systems; and all
+other (crop-based) feeds, which we consider to be consumend in intensive
+systems. For poultry, egg and monogastric meat production we use the
+per-capita income of the country to divide between intensive and
+extensive production systems. For low-income countries, we assume
+extensive production systems. We locate them according to the share of
+built-up areas based on the idea that these animals are held in
+subsistence or small-holder farming systems with a high labour per
+animal ratio. Intensive production associated with higher income
+countries, is distributed within a country using again share of
+crop-based feed production, assuming that feed availability is the most
+driving factor for livestock location.
 
-\textbf{Manure Excretion, Storage and Recycling}
-Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modelling.
-Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets \citep{weindl}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep{ipcc_2019_2019}).
-Depending on the feed system we assume manure to be handled in four different ways:
-All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Manure fuel shares are estimated using IPCC default values \citep{ippc_2006_2006}.
-Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing.
-To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see \ref{append:TableclossAWMS}).
-We limit high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
+\textbf{Manure Excretion, Storage and Recycling} Manure production and
+management is based on a revised methodology of
+\citep{bodirsky_n2o_2012} and is presented here due to its central role
+in soil carbon modelling. Based on the gridded livestock distribution we
+calculate excretions by estimating the nitrogen balance of the livestock
+system on the basis of comprehensive livestock feed baskets
+\citep{weindl}, assuming that all nitrogen in protein feed intake, minus
+the nitrogen in the slaughter mass, is excreted. Carbon in excreted
+manure is estimated by applying fixed C:N ratios (given by
+\citep{ipcc_2019_2019}). Depending on the feed system we assume manure
+to be handled in four different ways: All manure orginated from pasture
+feed intake is excreted directly to pastures and rangelands (pasture
+grazing), deducting manure collected as fuel. Manure fuel shares are
+estimated using IPCC default values \citep{ippc_2006_2006}. Whereas for
+low-income countries, we adopt a share of 25\% of crop residues in feed
+intake directly consumend and excreted on crop fields (stubble grazing),
+we do not consider any stubble grazing in high-income countries;
+middle-income countries see linearly interpolated shares depending on
+their per-capita income. For all other feed items we assume the manure
+to be stored in animal waste management systems associated with
+livestock housing. To estimate the carbon actually recycled to the soil,
+we account for carbon losses during storage and recycling shares in
+different animal waste management and grazing systems. Whereas we assume
+no losses for pasture and stubble grazing, we consider that the manure
+collected as fuel is not recycled. For manure stored in different animal
+waste management systems we compiled carbon loss rates partly depending
+on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}
+(see @ref(append:TableclossAWMS)). We limit high application shares at
+\(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
 
@@
 
-We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
-The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
-We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
-Moreover we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
+We use irrigation area shares to modify the water effect on
+decomposition by weighting the irrigated and rainfed water factors based
+on these shares. The LUH2v2 data set provides irrigated fractions for
+their cropland subcategories. We apply aggregated area shares, leading
+to the same impact of irrigation on all of our crop groups. Moreover we
+assume the irrigation effect to be present for all 12 months of a year
+in grid cells, that has been marked as irrigated.
 
@@
 
-To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rule is based on the approach of \citep{porwollik_generating_2018}, but neglects the (less common) adoption of no-tillage farming on agricultural land.
+To account for the distribution of tillage to the three different
+tillage types specfied by the IPCC - full tillage, reduced tillage and
+no tillage -, we assume that all natural land and pastures are not
+tilled, whereas as default annual crops are under full and perennials
+under reduced tillage. Furthermore we assume no tillage in cropland
+cells specified as no tillage cell based on the historic global gridded
+tillage data set from Prowollik \citep{porwollik_generating_2018}.
+Porwollik data set is extended by to the period of 1974-2010 by
+combining country-level data on conservation agriculture (ca) area
+values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the
+methodolgy of Porwollik to identify potential no tillage cells.
 
@@
 \hypertarget{sec:tier1}{%
-\subsection{SOC Stocks and Stock changes following Tier 1}\label{sec:tier1}}
+\subsection{SOC Stocks and Stock changes following Tier
+1}\label{sec:tier1}}
 
-Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}).
-Here, stocks are calculated via stock change factors (F\^{}\{SCF\}) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-\(F^{SCF}\) are differentiated by different crop, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover \(F^{SCF}\) vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}).
-The actual SOC stocks as thus calulated based on a given reference SOC stock by
+Additionally to the tier 2 approach of the refinement of IPCC guidelines
+vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management
+data coming with it, SOC changes can be estimated using the IPCC tier 1
+approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006},
+\citet{ipcc_2019_2019}). Here, stocks are calculated via stock change
+factors (F\^{}\{SCF\}) given by the IPCC for the topsoil (0-30 cm) and
+based on observational data. \(F^{SCF}\) are differentiated by different
+crop, management and input systems (here summarized under \(m\))
+reflecting different dynamics under changed in- and outflows without
+explicitly tracking these. Moreover \(F^{SCF}\) vary for different
+climatic zones (\(c\)) specified by the IPCC (see
+@ref(append:climatemap)). The actual SOC stocks as thus calulated based
+on a given reference SOC stock by
 
@@
 SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
-\label{eq:tier1}
+(\#eq:tier1)
 \end{equation}
 
-with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis we use the default \(F^{SCF}\) from the Tier 1 method of \citep{ipcc_2006_2006}, and \citep{ipcc_2019_2019} as a comparison and constistency check for our more detailed budget approach.
-\newpage
+with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into
+corresponding climate zones \(c\). For this analysis we use the default
+\(F^{SCF}\) from the Tier 1 method of \citep{ipcc_2006_2006}, and
+\citep{ipcc_2019_2019} as a comparison and constistency check for our
+more detailed budget approach. \newpage
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\section{Results}
 
-We present simulation results of the SOC budget of the period from 1975 to 2010 mainly focusing on cropland areas.
-Detailed total global and regional results of the state of the worlds SOC in comparison to other estimates can be found in the supplement.
+We present simulation results of our SOC budget focusing on cropland
+areas for the year 2010 as well as global trends of SOC stock changes
+for the period from 1975 to 2010.
 
-\hypertarget{soc-distribution-and-depletion}{%
-\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}
-
-Figure
+\subsection{SOC distribution and depletion}
 
@@
 \includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas.
-(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades (despite ongoing cropland expansion).
+(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades.
 (c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
@@
 
-\hypertarget{global-agricultural-carbon-flows-and-stocks}{%
-\subsection{global agricultural carbon flows and stocks}\label{global-agricultural-carbon-flows-and-stocks}}
+In fig.~1(a) we provide the first world map of SOC on croplands
+considering real world management data on the global scale. Our
+spatially explicit results moreover show hot spots of SOC losses as well
+as gains in two different ways: 1. Absolute SOC changes (see fig.~1(c))
+indicate areas with high importance for the global SOC emissions. The
+might be driven by huge relative losses or a high natural stock, from
+which even small deviations could lead to substantial losses. 2. To
+attribute SOC losses to insufficient agricultural management relative
+SOC changes (\(F^{SCF}\), see fig.~1(d)) are a helpful tool. They
+indicate areas with huge difference in carbon inflows or SOC decay
+compared to natural vegetation, that might be overcome due to improved
+agricultural pratices.
+
+\subsection{Agricultural management effects on SOC emissions and
+cycling}
+
+Global cummulative SOC emissions are decreasing (see fig.~1(c)). Fig. 2
+reveals the relative impact of management effects by freezing tillage
+areas as well as carbon inflows from residues or manure at the level of
+1975. Our counterfactual scenarios show that the increasing residue
+carbon input had the biggest overall effect on SOC stocks. Without
+changes in management regimes especially in residue inflows to the soil,
+global cummulative SOC emissions would still grow. The strong effect of
+carbon residue amounts are also visible in the carbon flow diagram for
+agricultural production for the year 2010 (see fig.~3)
+
+\begin{figure}[H]
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario} \caption{Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline.}\label{fig:SOCscen}
+\end{figure}
 
@@
 
-\hypertarget{comparison-of-stock-changes-factors}{%
-\subsection{comparison of stock changes factors}\label{comparison-of-stock-changes-factors}}
-
-\begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{$F^{SOC}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{SOC}$ by additional 5-8 percent.
-Also: modelled $F^{SOC}$ have increased for all climates over time .}\label{fig:SCFtable}
-\end{figure}
-
 \newpage
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}
+
+\subsection{Including agricultural management data changes the sign of
+the trend}
+
+This study provides an analysis on historic SOC stock changes on
+cropland. We determine the SOC trends on cropland compared to a
+potential natural vegeted world under the same historic climatic
+conditions (\(SOC_{natveg}\)) leading to a cummulative SOC emission
+since dawn of mankind of around 37 GtC in 2010. Whereas recent modelling
+estimates on global SOC emissions indicate an on going increase of SOC
+emissions (Sanderman et al, Pugh et al.), our study indicates that the
+global SOC gap might be closing slowly.
+
+According to Sanderman et al.~cummulative SOC emissions since beginning
+of human cropping activities have been at around 37 GtC for the first 30
+cm of the soil with half of it attributed to grazing. It was also
+pointed out, that this results might be conservative estimates and low
+compared to experimental results, leaving our estimate of 37 GtC in 2010
+for cropland emissions only in alignment with Sanderman et al.~estimates
+considering the hugh uncertainties in modelling SOC on the global scale.
+Furtherore the results of Sanderman et al.~calculated historical trends
+based on agricultural land expansion without considering SOC variations
+due to different management systems at all. In Pugh et al.~management
+effects like tillage and residue recycling have been considered, but
+neither changes over time nor alignment to oberserved historical data
+like yields levels or no tillage areas were taken into account. Under
+this assumptions this study only found marginal effect of crop
+productivity and other management effects on cummulative SOC emissions.
+
+Nevertheless our results show, when purely focusing on the past decades,
+the moderate global cropland expansion of around 11\% between 1974 and
+2010, is outweighted by improved agricultural yields ad practises.
+Moreover our sensitivity analysis indicades that 1. yield increases and
+with that the increase in residue biomass might play the most important
+role, followed by 2. enhanced residue recycling rates, 3. improved
+manure recycling (e.g.~due to improved animal waste management system)
+and 4. the adoption of no tillage practises.
+
+Modelling management effects on the global scale comes however with huge
+parametic and systematic uncertainties. One aspect pointed out by Keel
+(2017) as well as by Smith (2019) might be that carbon input
+calculations are highly sensitive to the choice of allometric functions
+determining below and above ground residue estimates from harvested
+quantities. Keel et al.~questioned, that below ground residues might
+increase with a fixed root:shot ratio rather than being independend from
+productivity gains. Following this argumentation SOC results shown in
+this study might especially in high-yielding farming system overestimate
+actual SOC stocks. However according to our study above ground residue
+biomass recycling seem to contribute even more to overall SOC stocks due
+to higher input rates.
+
+\subsection{Modeled management effect inline with default IPCC
+assumptions}
+
+To validated the effect of our modelled SOC stocks and stock changes
+under management, we compare our results to default IPCC stock changes
+factors (cite), which are based on measurement data for croplands (see
+@ref(tab:SCFtable)).
+
+\begin{figure}[H]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows IPCC 2006 as well as 2019 default (medium input) and low input regimes stock change factors $F_{SCF}$ without other subsystem consideration compared results of this study (SOC budget) for 1990 and 2010.}\label{fig:SCFtable}
+\end{figure}
+
+Our climate zone specific, aggregated \(F_{SCF}\) correspond very well
+to the tier 1 estimation of IPCC, 2006 of \(F_{SOC}\). For the tropical
+regions the assumptions changed notablly from the guidelines in 2006 to
+the update in 2019, leaving our results too low in comparison with IPCC,
+2019. Considering yield gaps in mainly developing regions in the tropics
+the default assumption of medium input systems, might be an
+overestimation of actual SOC state. The additional effect of considering
+low inpout regimes in tropical regions can however not explain the full
+mismatch to IPCC 2019 values but account for at least 5-7\% of it.
+
+\subsection{SOC stocks inline with literature}
+
+The worlds SOC stock and its changes are highly uncertain (cite), which
+is visible by the wide range of global SOC stock estimates (see
+@ref(tab:SOCtable)).
+
+\begin{figure}[H]
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/TableSOC_comparison} \caption{Modelled as well as data based estimation for global SOC stock (for 2010) in GtC for the first 30 cm of soil.}\label{fig:SOCtable}
+\end{figure}
+
+The global esimates of SOC stock by this study are on the lower edge
+compared to other modelled results or more data driven estimates.
+Looking on regional results (fig.~SX in supplement), our estimates turn
+out to be in good agreement for most regions with the largest deviations
+for boreal areas. Considering that the model was parametrized for
+croplands, these mismatches are not superising, since the temperature
+effects on decomposition is most likly fundamentally different for
+permaforst soils than assumend in this study. To avoid that this bias
+influences our results, our study focusses exclusively on cropland
+soils, excluding most of the boreal zone. Morover, when focussing on SOC
+changes, pristine natural vegetated areas without human land management
+under the same historic climatic conditions cancel out in the
+calculation of SOC emissions.
+
+However the natural land representation that is dominating the total SOC
+stocks of the world and with it also the SOC initialization is lacking
+proper parametrization of nitrogen and lignin content of litterfall.
+This leaves carbon inputs and decay behaviour for natural land rather
+uncertained. As all results are valued against these potential natural
+SOC the absolute values and also the emissions over time have to be used
+with caution. Especially in less forested areas the natural land
+representation might be off, due to parameterization assumptions of the
+natural litterfall. Nevertheless total SOC stocks are in a reasonable
+range and the stock change factors \(F^{SCF}\) are in good agreement
+with the Tier 1 default values of 2006. In addition we preformed a
+sensitivity analysis for lignin and nitrogen parameterization of natural
+litterfall (see appendix). It shows that the general trend of decreasing
+SOC emissions is not altered even by rather unlikely parameter choices,
+that would overestimate natural SOC stocks strongly and therefor lead to
+high emission in the case of land conversion.
+
+\subsection{Important short commings --\textgreater{} move to appendix}
+
+Smaller points and shortcommings:
+
+-This study as the IPCC guidelines suggested has limited her focus to
+the first 30 cm of the soil profile, leaving changes in the subsoil
+unnoticed. Nevertheless studies (see Don on tillage) have shown, the
+subsoil to be a game changer in evaluating total SOC losses or gains for
+no-tillage systems. It has been argued that for intensivly tilled soils,
+subsoil SOC is increasing due to the import of carbon rich topsoil to
+deeper soil layers. Following these argumentation SOC stocks in
+croplands might even be underestimated.
 
@@
 \item
-  The natural land representation and therefor also SOC initialization is lacking proper parametrization of both decay behaviour and natural input (e.g.~nitrogen and lignin content of litterfall).
-As all results are valued against these potential natural SOC the absolute values and also the changes over time have to be used with caution.
-Especially in boreal zone and in very dry regions the natural land representation might be far off, due to the focus for the parametrization for cropped areas.
-  The good agreement of the stock change factors \(F^{SCF}\) with the Tier 1 default values might serve as a hint that the natural land representation is not totally off.
-Boreal and very dry zone might be uncropped, leaving the results on SOC changes on cropland untouch from the bad borreal land representation.
+  Fertilizer interaction is not included here by accounting for
+  additional N supply that would alter C:N ratio of the carbon inputs.
+  Tier 2 steady-state method is neglecting fertilizer application,
+  however we would have fertilizer ammounts at hand to include them, if
+  proper representation of fertilizer within the method would be
+  possible to add.
 \item
-  This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
-Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-It has been argued that for intensivly tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
-Following these argumentation SOC stocks in croplands might even be underestimated.
+  Pasture dynamics are neglected and treated as natural vegetation,
+  which might be -- looking on pasture degradation due to overgrazing --
+  oversimplified for some spots, but is inline with assumption on
+  pasture SOC stocks done before (see Tier 1 IPCC). (Note that also
+  manure excreted to pastures is neglected within these analysis, since
+  we focus purely on cropland dynamics.)
 \item
-  As pointed out by Keel (2017), Smith (2019) the results might be highly sensitive to carbon input calculations more precisly to below and above ground residue carbon estimates derived from harvested quantities.
-It has been questioned, that below ground residues might increase with a fixed root:shot ratio (maybe specifically in high end farming systems (?)).
-Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
-(In general the sensitivity to and the uncertainty of certain parameters might be very high.)
-\end{itemize}
-
-\newpage
-
-Smaller points and shortcommings:
-
-\begin{itemize}
+  No tillage adaptaion is neglected on cropland due to less common
+  adoption of no tillage and conservation agriculture. Pastures are
+  assumed to be not tilled at all (propably only heavy managed pastures
+  are tilled with some rotation)
 \item
-  Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
-Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer ammounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
+  Irrigated areas are not crop specific and irrigation is not restricted
+  to growing periods (since it is very complex to calculate average
+  growing periods). Crop specific growing periods might be possible
+  using LPJmL data.
 \item
-  Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
-(Note that also manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.)
+  flooded rice area are not represented correctly as parameterization
+  does not hold true for flooded conditions.
 \item
-  No tillage adaptaion is neglected on cropland due to less common adoption of no tillage and conservation agriculture.
-Pastures are assumed to be not tilled at all (propably only heavy managed pastures are tilled with some rotation)
-\item
-  Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
-Crop specific growing periods might be possible using LPJmL data.
-\item
-  flooded rice area are not represented correctly as parameterization does not hold true for flooded conditions.
-\item
-  Carbon displacement via leaching and erosion is neglected in this study.
+  Carbon displacement via leaching and erosion is neglected in this
+  study.
 \item
@@
 \item
-  Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
-Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
+  Within cropland we do not track croparea transitions, but rather look
+  at statistical distributions of the crop functional types. Due to crop
+  rotations and missing data on crop specific distributions, these
+  transitions would be any way rather uncertain.
 \item
-  The disaggregation of manure to uild-up areas (in the case of extensive monograstrics) is leeding to a lot of displaced manure (?) that is cut off
+  The disaggregation of manure to uild-up areas (in the case of
+  extensive monograstrics) is leeding to a lot of displaced manure (?)
+  that is cut off
 \item
-  It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
+  It is known that there are mismatches between FAO statistics and LUH
+  areas. As far as possibles there were harmonized within this study.
 \item
-  ``The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
-  current C stock to the future steady-state soil C stock calculated based on current conditions.'' Leading to the fact that our total stock results are highly uncertained.
-  \newpage
+  ``The Tier 2 method does not simulate C change but simply calculates
+  an annual C stock change from the current C stock to the future
+  steady-state soil C stock calculated based on current conditions.''
+  Leading to the fact that our total stock results are highly
+  uncertained. \newpage
 \end{itemize}
@@
 
-Outlook including perspective on mitigation (and soc enhancement in the future).
-\newpage
+Outlook including perspective on mitigation (and soc enhancement in the
+future). \newpage
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+\section{Appendix}
 
 \hypertarget{append:subsection2mrfunctions}{%
-\subsection{table on method subsections to functions within R packages}\label{append:subsection2mrfunctions}}
+\subsection{table on method subsections to functions within R
+packages}\label{append:subsection2mrfunctions}}
 
 \hypertarget{append:Tableluh2fao2mag}{%
-\subsection{table on mapping LUH2FAO2MAG}\label{append:Tableluh2fao2mag}}
+\subsection{table on mapping
+LUH2FAO2MAG}\label{append:Tableluh2fao2mag}}
 
@@
 
-Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
+Litter is coming from LPJmL in carbon units - transformation with 0.44
+is done twice reverting the effect of the transformation
 
@@
 
-\codedataavailability{Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/.
-Data used for the output can be found under .} %% use this section when having data sets and software code available
+\codedataavailability{Software code for paper and result prepartion can
+be found under www.github.com/k4rst3ns/. Data used for the output can be
+found under
+.} %% use this section when having data sets and software code available
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of Bodirsky (and ).
-Bodirsky, and Popp revised paper.} %% optional section
+\authorcontribution{Karstens wrote code and paper build on work of
+Bodirsky (and ). Bodirsky, and Popp revised paper.} %% optional section
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
+\competinginterests{The authors declare no competing
+interests.} %% this section is mandatory even if you declare that no competing interests are present
 

### 2020-10-06 SOCbudgetPaper/discussion.Rmd
Mayor update of everything

@@
 
-- The natural land representation and therefor also SOC initialization is lacking proper parametrization of both decay behaviour and natural input (e.g.
-nitrogen and lignin content of litterfall).
-As all results are valued against these potential natural SOC the absolute values and also the changes over time have to be used with caution.
-Especially in boreal zone and in very dry regions the natural land representation might be far off, due to the focus for the parametrization for cropped areas.
-The good agreement of the stock change factors $F^{SCF}$ with the Tier 1 default values might serve as a hint that the natural land representation is not totally off.
-Boreal and very dry zone might be uncropped, leaving the results on SOC changes on cropland untouch from the bad borreal land representation.
+## Including agricultural management data changes the sign of the trend
 
-- This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
+This study provides an analysis on historic SOC stock changes on cropland.
+We determine the SOC trends on cropland compared to a potential natural vegeted world under the same historic climatic conditions ($SOC_{natveg}$) leading to a cummulative SOC emission since dawn of mankind of around 37 GtC in 2010.
+Whereas recent modelling estimates on global SOC emissions indicate an on going increase of SOC emissions (Sanderman et al, Pugh et al.), our study indicates that the global SOC gap might be closing slowly.
+
+According to Sanderman et al.
+cummulative SOC emissions since beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+It was also pointed out, that this results might be conservative estimates and low compared to experimental results, leaving our estimate of 37 GtC in 2010 for cropland emissions only in alignment with Sanderman et al.
+estimates considering the hugh uncertainties in modelling SOC on the global scale.
+Furtherore the results of Sanderman et al.
+calculated historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+In Pugh et al.
+management effects like tillage and residue recycling have been considered, but neither changes over time nor alignment to oberserved historical data like yields levels or no tillage areas were taken into account.
+Under this assumptions this study only found marginal effect of crop productivity and other management effects on cummulative SOC emissions.
+
+Nevertheless our results show, when purely focusing on the past decades, the moderate global cropland expansion of around 11% between 1974 and 2010, is outweighted by improved agricultural yields ad practises.
+Moreover our sensitivity analysis indicades that 1.
+yield increases and with that the increase in residue biomass might play the most important role, followed by 2.
+enhanced residue recycling rates, 3.
+improved manure recycling (e.g.
+due to improved animal waste management system) and 4.
+the adoption of no tillage practises.
+
+Modelling management effects on the global scale comes however with huge parametic and systematic uncertainties.
+One aspect pointed out by Keel (2017) as well as by Smith (2019) might be that carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
+Keel et al.
+questioned, that below ground residues might increase with a fixed root:shot ratio rather than being independend from productivity gains.
+Following this argumentation SOC results shown in this study might especially in high-yielding farming system overestimate actual SOC stocks.
+However according to our study above ground residue biomass recycling seem to contribute even more to overall SOC stocks due to higher input rates.
+
+## Modeled management effect inline with default IPCC assumptions
+
+To validated the effect of our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (cite), which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
+
+```{r caption_SCFtable, echo=FALSE}
+caption_SCFtable <- "This table shows IPCC 2006 as well as 2019 default (medium input) and low input regimes stock change factors $F_{SCF}$ without other subsystem consideration compared results of this study (SOC budget) for 1990 and 2010."
+```
+
+```{r SCFtable, out.width = "16cm", fig.cap = caption_SCFtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
+```
+
+Our climate zone specific, aggregated $F_{SCF}$ correspond very well to the tier 1 estimation of IPCC, 2006 of $F_{SOC}$.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
+The additional effect of considering low inpout regimes in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.
+
+## SOC stocks inline with literature
+
+The worlds SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+
+```{r caption_SOCtable, echo=FALSE}
+caption_SOCtable <- "Modelled as well as data based estimation for global SOC stock (for 2010) in GtC for the first 30 cm of soil."
+```
+
+```{r SOCtable, out.width = "8cm", fig.cap = caption_SOCtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSOC_comparison.png")
+```
+
+The global esimates of SOC stock by this study are on the lower edge compared to other modelled results or more data driven estimates.
+Looking on regional results (fig.
+SX in supplement), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not superising, since the temperature effects on decomposition is most likly fundamentally different for permaforst soils than assumend in this study.
+To avoid that this bias influences our results, our study focusses exclusively on cropland soils, excluding most of the boreal zone.
+Morover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+
+However the natural land representation that is dominating the total SOC stocks of the world and with it also the SOC initialization is lacking proper parametrization of nitrogen and lignin content of litterfall.
+This leaves carbon inputs and decay behaviour for natural land rather uncertained.
+As all results are valued against these potential natural SOC the absolute values and also the emissions over time have to be used with caution.
+Especially in less forested areas the natural land representation might be off, due to parameterization assumptions of the natural litterfall.
+Nevertheless total SOC stocks are in a reasonable range and the stock change factors $F^{SCF}$ are in good agreement with the Tier 1 default values of 2006.
+In addition we preformed a sensitivity analysis for lignin and nitrogen parameterization of natural litterfall (see appendix).
+It shows that the general trend of decreasing SOC emissions is not altered even by rather unlikely parameter choices, that would overestimate natural SOC stocks strongly and therefor lead to high emission in the case of land conversion.
+
+## Important short commings --> move to appendix
+
+Smaller points and shortcommings:
+
+-This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
 Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
@@
 Following these argumentation SOC stocks in croplands might even be underestimated.
-
-- As pointed out by Keel (2017), Smith (2019) the results might be highly sensitive to carbon input calculations more precisly to below and above ground residue carbon estimates derived from harvested quantities.
-It has been questioned, that below ground residues might increase with a fixed root:shot ratio (maybe specifically in high end farming systems (?)).
-Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
-(In general the sensitivity to and the uncertainty of certain parameters might be very high.)
-
-<!--Tasks: 
-Find better natural land representation and initalization.
-Using LPJmL values might be complex since the split up into SOC subpools is needed.
-Getting better parametrization for the CENTURY based approach for decay behaviour and better values for nitrogen and lignin of natural litter fall.
--->
-
-\newpage
-
-Smaller points and shortcommings:
 

### 2020-10-06 SOCbudgetPaper/introduction.Rmd
Mayor update of everything

@@
 
-Possible titles:
-
-* How management changed soil organic carbon on global croplands
-
-* Agricultural soil have lost 37 GtC topsoil carbon
-
-* Linking agricultural management data to soil modeleling – a global approach
-
-Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and biospheric pools multiple times (Bathjes).
+Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth’s soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and vegetation pools multiple times (Batjes).
 As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref.
 permafrost melting).
-The specific amount of carbon stored in the soil is however uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).
+The specific amount of carbon stored in the soil is, however, uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Batjes, 1996; more).
 The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
-Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles in this regard.
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today.
-Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC <!--look up --> (Lal, 2006).
+Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite).
+Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC  (Lal, 2006).
 Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX).
-While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especial agricultural activities.
-Field-scale models (ref.
-Daycent, RothC, Ecosse, C-Tool) are able to capture these land mangement impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm activities.
+While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especially of agricultural activities.
+Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to capture these land management impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm measures.
 However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
@@
 This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
-We thereby consider change in SOC caused by historical land cover change as well as of different agricultural mangement practices, including residue recycling, manure amendments, irrigation, and tillage.
-We thereby provide the first global, spatially explicit SOC loss maps that consider detailed agricultural management.
+We thereby consider change in SOC caused by historical land cover change as well as of different agricultural management practices, including residue recycling, manure amendments, irrigation, and tillage.
+We provide the first global, spatially explicit SOC loss maps that considers detailed agricultural management.
 This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
@@
 Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing the climate zone specfic stock changes with default stock change factors from the IPCC Tier 1 approach and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
-Finally, we will discuss our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
+Finally, we discuss and compare our findings – including their implications for SOC model development – and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.

### 2020-10-06 SOCbudgetPaper/methods.Rmd
Mayor update of everything

@@
 
-Note that the absolution changes $\Delta SOC$ can be also interpreted as cummulative SOC emissions or as SOC depth of mankind; whereas relative changes $F^{SCF}$ can be denoted as stock change factors as defined within the IPCC guidelines of 2006 [@ipcc_2006_2006].
+Note that the absolution changes $\Delta SOC$ can be also interpreted as the SOC gap due to human cropping activits; whereas relative changes $F^{SCF}$ can be denoted as stock change factors as defined within the IPCC guidelines of 2006 [@ipcc_2006_2006].
+Note that the SOC gap is to the negated cummulative SOC emission also refered to as SOC debt (cite Sanderman) of mankind.
 
@@
 
-To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rule is based on the approach of [@porwollik_generating_2018], but neglects the (less common) adoption of no-tillage farming on agricultural land.
+To account for the distribution of tillage to the three different tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas as default annual crops are under full and perennials under reduced tillage.
+Furthermore we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from Prowollik  [@porwollik_generating_2018].
+Porwollik data set is extended by to the period of 1974-2010 by combining country-level data on conservation agriculture (ca) area values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the methodolgy of Porwollik to identify potential no tillage cells.
 

### 2020-10-06 SOCbudgetPaper/results.Rmd
Mayor update of everything

--- 2020-06-26 SOCbudgetPaper/results.Rmd
@@
 
-```{r table, echo = FALSE}
-table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
-knitr::kable(table$naturalstate_table, "latex", digit=2, caption="SCF compared to potential natural state")
+We present simulation results of our SOC budget focusing on cropland areas for the year 2010 as well as global trends of SOC stock changes for the period from 1975 to 2010.
+
+## SOC distribution and depletion
+
+```{r caption_SOCmaps, echo=FALSE}
+caption_SOCmaps <- "(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas.
+(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+(c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
+Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas."
 ```
 
-```{r table2, echo = FALSE}
-table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
-knitr::kable(table$comparison_table, "latex", digit=2, caption="SCF compared to IPCC Tier 1")
-```
-
-```{r out.width = "16cm"}
+```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="H"}
 # All defaults
@@
 ```
+
+In fig.
+1(a) we provide the first world map of SOC on croplands considering real world management data on the global scale.
+Our spatially explicit results moreover show hot spots of SOC losses as well as gains in two different ways:
+1.
+Absolute SOC changes (see fig.
+1(c)) indicate areas with high importance for the global SOC emissions.
+The might be driven by huge relative losses or a high natural stock, from which even small deviations could lead to substantial losses.
+2.
+To attribute SOC losses to insufficient agricultural management relative SOC changes ($F^{SCF}$, see fig.
+1(d)) are a helpful tool.
+They indicate areas with huge difference in carbon inflows or SOC decay compared to natural vegetation, that might be overcome due to improved agricultural pratices.
+
+## Agricultural management effects on SOC emissions and cycling
+
+Global cummulative SOC emissions are decreasing (see fig.
+1(c)).
+Fig.
+2 reveals the relative impact of management effects by freezing tillage areas as well as carbon inflows from residues or manure at the level of 1975.
+Our counterfactual scenarios show that the increasing residue carbon input had the biggest overall effect on SOC stocks.
+Without changes in management regimes especially in residue inflows to the soil, global cummulative SOC emissions would still grow.
+The strong effect of carbon residue amounts are also visible in the carbon flow diagram for agricultural production for the year 2010 (see fig.
+3)
+
+```{r caption_SOCscen, echo=FALSE}
+caption_SOCscen <- "Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline."
+```
+
+```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
+```
+
+```{r caption_FlowFig, echo=FALSE}
+caption_FlowFig <- "Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
+Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil)."
+```
+
+```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/OuFlowFig.png")
+```
+
+\newpage

### 2020-10-06 SOCbudgetPaper/SOCbudget.tex
Update SOCbudget.tex

@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of larges carbon stocks on earth. It is
+Soil organic carbon (SOC) is one of larges carbon stocks on Earth. It is
 three times larger than the biospheric pool and more than twice as big
@@
 (c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
-Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.}\label{fig:SOCmaps}
+Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.
+(BB: Increase map size, legend is larger than map, no need for that)}\label{fig:SOCmaps}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario} \caption{Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline.}\label{fig:SOCscen}
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario} \caption{Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline.
+(BB: order of legend should be equal to order of lines.
+Management=tillage+reside+manure?)}\label{fig:SOCscen}
 \end{figure}
@@
 \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/OuFlowFig} \caption{Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
-Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil).}\label{fig:FlowFig}
+Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil).(BB: adjuste flow size to numbers, include arrows for losses, e.g.
+--> x ,otherwise its confusing, trees need black outer line, roots have no clean upper end)}\label{fig:FlowFig}
 \end{figure}
@@
 
-This study provides an analysis on historic SOC stock changes on
-cropland. We determine the SOC trends on cropland compared to a
-potential natural vegeted world under the same historic climatic
-conditions (\(SOC_{natveg}\)) leading to a cummulative SOC emission
-since dawn of mankind of around 37 GtC in 2010. Whereas recent modelling
-estimates on global SOC emissions indicate an on going increase of SOC
+This study provides an analysis of historic SOC stock changes on
+cropland. We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic
+conditions (\(SOC_{natveg}\)). Our results show, that human activity lead to cummulative SOC emissions
+of around 37 GtC in 2010. Whereas recent modelling
+estimates of global SOC emissions indicate an ongoing increase of SOC
 emissions (Sanderman et al, Pugh et al.), our study indicates that the
-global SOC gap might be closing slowly.
+global SOC gap is slowly closing due to improved management.
 
@@
 cm of the soil with half of it attributed to grazing. It was also
-pointed out, that this results might be conservative estimates and low
-compared to experimental results, leaving our estimate of 37 GtC in 2010
-for cropland emissions only in alignment with Sanderman et al.~estimates
-considering the hugh uncertainties in modelling SOC on the global scale.
-Furtherore the results of Sanderman et al.~calculated historical trends
+pointed out that these results might be conservative estimates and low
+compared to experimental results. Considering the huge uncertainties in modelling SOC on the global scale, 
+our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with Sanderman et al.~estimates.
+
+Furthermore (BB: consider using a spelling checker) the results of Sanderman et al.~calculated (BB: the results calculated?) historical trends
 based on agricultural land expansion without considering SOC variations
-due to different management systems at all. In Pugh et al.~management
-effects like tillage and residue recycling have been considered, but
+due to different management systems at all. Pugh et al.~ considered management
+effects like tillage and residue recycling in a static way, but
 neither changes over time nor alignment to oberserved historical data
-like yields levels or no tillage areas were taken into account. Under
-this assumptions this study only found marginal effect of crop
+like yields-levels or no-tillage areas were taken into account. Under
+these assumptions this study only found marginal effects (BB: be more precise) of crop
 productivity and other management effects on cummulative SOC emissions.
 
-Nevertheless our results show, when purely focusing on the past decades,
-the moderate global cropland expansion of around 11\% between 1974 and
-2010, is outweighted by improved agricultural yields ad practises.
-Moreover our sensitivity analysis indicades that 1. yield increases and
+Our study for the first time uses a dynanmic management dataset as driver for SOC dynamics. We
+show that the moderate global cropland expansion of around 11\% between 1974 and
+2010 and the resulting depletion of SOC stocks in converted cropland has been outweighted by improved agricultural yields ad practises.
+Moreover our sensitivity analysis indicades that (1) yield increases and
 with that the increase in residue biomass might play the most important
-role, followed by 2. enhanced residue recycling rates, 3. improved
+role, followed by (2) enhanced residue recycling rates, (3) improved
 manure recycling (e.g.~due to improved animal waste management system)
-and 4. the adoption of no tillage practises.
+and (4) the adoption of no tillage practises.
 
-Modelling management effects on the global scale comes however with huge
-parametic and systematic uncertainties. One aspect pointed out by Keel
-(2017) as well as by Smith (2019) might be that carbon input
+Modelling management effects on the global scale comes however with 
+parametric and structural uncertainties. As pointed out by Keel
+(2017) and Smith (2019), carbon input
 calculations are highly sensitive to the choice of allometric functions
 determining below and above ground residue estimates from harvested
-quantities. Keel et al.~questioned, that below ground residues might
-increase with a fixed root:shot ratio rather than being independend from
+quantities. Keel et al.~questioned whether below ground residues might
+increase with a fixed root:shoot ratio rather than being independend from
 productivity gains. Following this argumentation SOC results shown in
 this study might especially in high-yielding farming system overestimate
-actual SOC stocks. However according to our study above ground residue
-biomass recycling seem to contribute even more to overall SOC stocks due
+actual SOC stocks (BB: No.
+Very likely, root:shoot ratios come from developed countries.
+They rather underestimate SOC in LICs).
+However (?), according to our study above ground residue biomass recycling seems to contribute even more to overall SOC stocks due
 to higher input rates.
 
-\subsection{Modeled management effect inline with default IPCC
+\subsection{Modeled management effect in line with default IPCC
 assumptions}
@@
 to the tier 1 estimation of IPCC, 2006 of \(F_{SOC}\). For the tropical
-regions the assumptions changed notablly from the guidelines in 2006 to
+regions the assumptions changed notably from the guidelines in 2006 to
 the update in 2019, leaving our results too low in comparison with IPCC,

### 2020-10-07 SOCbudgetPaper/SOCbudget.tex
Update SOCbudget.tex

@@
 cm of the soil with half of it attributed to grazing. It was also
-pointed out that these results might be conservative estimates and low
+pointed out that these results might be conservatively low
 compared to experimental results. Considering the huge uncertainties in modelling SOC on the global scale, 
@@
 
-To validated the effect of our modelled SOC stocks and stock changes
+To validate the effect of our modelled SOC stocks and stock changes
 under management, we compare our results to default IPCC stock changes
-factors (cite), which are based on measurement data for croplands (see
+factors (cite) which are based on measurement data for croplands (see
 @ref(tab:SCFtable)).
@@
 \begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows IPCC 2006 as well as 2019 default (medium input) and low input regimes stock change factors $F_{SCF}$ without other subsystem consideration compared results of this study (SOC budget) for 1990 and 2010.}\label{fig:SCFtable}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows different estimates of stock change factors $F_{SCF}$, defined as the soil carbon stock of managed croplands relative to the counterfactual soil carbon stock of undisturbed natural vegetation.
+We compare average values for the four IPCC climate zone classifications ...
+(BB: Your caption has to be self-explaining for people who do not know the rest of your article.
+All elements have to be defined and explained.)IPCC 2006 as well as 2019 default (medium input) and low input regimes(BB:regimes?) stock change factors  without other subsystem consideration compared results of this study (SOC budget) for 1990 and 2010.}\label{fig:SCFtable}
 \end{figure}
 
-Our climate zone specific, aggregated \(F_{SCF}\) correspond very well
-to the tier 1 estimation of IPCC, 2006 of \(F_{SOC}\). For the tropical
+We also intercompared our stock-change factors \(F_{SCF}\), which describe the ratio between ... . 
+To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones (see table X).
+Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006. For the tropical
 regions the assumptions changed notably from the guidelines in 2006 to
-the update in 2019, leaving our results too low in comparison with IPCC,
-2019. Considering yield gaps in mainly developing regions in the tropics
+the update in 2019 (BB: do you have any indication why they changed so dramatically?), leaving our results too low in comparison with IPCC,
+2019. The development of \(F_{SCF}\) over time in our study shows the substantial impact of changed management.
+Considering yield gaps in mainly developing regions in the tropics
 the default assumption of medium input systems, might be an
@@
 \begin{figure}[H]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/TableSOC_comparison} \caption{Modelled as well as data based estimation for global SOC stock (for 2010) in GtC for the first 30 cm of soil.}\label{fig:SOCtable}
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/TableSOC_comparison} \caption{Modelled as well as data based estimation for global SOC stock (for 2010) in GtC for the first 30 cm of soil.(BB: total land or just croplands?
+All other studies are for 2010?)}\label{fig:SOCtable}
 \end{figure}
 
-The global esimates of SOC stock by this study are on the lower edge
+(BB: do these studies assume landuse? what do they do for urban land?)
+
+The global esimates of SOC stock by this study are on the lower end
 compared to other modelled results or more data driven estimates.
@@
 for boreal areas. Considering that the model was parametrized for
-croplands, these mismatches are not superising, since the temperature
-effects on decomposition is most likly fundamentally different for
-permaforst soils than assumend in this study. To avoid that this bias
+croplands, these mismatches are not superising since the temperature
+effects on decomposition are fundamentally different for
+permaforst soils. To avoid that this bias
 influences our results, our study focusses exclusively on cropland
@@
 
-However the natural land representation that is dominating the total SOC
-stocks of the world and with it also the SOC initialization is lacking
-proper parametrization of nitrogen and lignin content of litterfall.
+Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands (BB: and pastures?), which are however only estimated in a basic manner.
+For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall.
 This leaves carbon inputs and decay behaviour for natural land rather
-uncertained. As all results are valued against these potential natural
-SOC the absolute values and also the emissions over time have to be used
+uncertain. (delete: As all results are valued against these potential natural
+SOC - BB: the valuation against the natural state is actually making the results more robust.
+We are only looking at the differences caused by management, not on the absolute states.
+Make this clear here again (even though you already mention it in the previous paragraph in a differnet context)
+The absolute values of stocks and emissions from land-use change therefore have to be used
 with caution. Especially in less forested areas the natural land
@@
 range and the stock change factors \(F^{SCF}\) are in good agreement
-with the Tier 1 default values of 2006. In addition we preformed a
-sensitivity analysis for lignin and nitrogen parameterization of natural
-litterfall (see appendix). It shows that the general trend of decreasing
-SOC emissions is not altered even by rather unlikely parameter choices,
-that would overestimate natural SOC stocks strongly and therefor lead to
-high emission in the case of land conversion.
+with the Tier 1 default values of 2006. Moreover, emissions or carbon sinks
+from changed management are not affected from the natural vegetation representation.
+To identify the importance of lignin and nitrogen parameterization of natural
+litterfall, we conducted a sensitivity analysis (see appendix). It shows that the general trend of decreasing
+SOC emissions is not altered even by rather unlikely parameter choices (BB: Why do you look at unlikely parameter choices???).
 
@@
 
+(BB: Or include into outlook)
 Smaller points and shortcommings:

### 2020-11-04 ResultNotebooks/results_SOCcompare.Rmd
update discussion part after bennis suggestions - tex is not compiling

@@
 
+```{r point plot reg stock, out.width=800, out.height=600 }
+library(ggplot2)
+
+leg_name <- "Data Source"
+comp <- ggplot(data = as.data.frame(glo_SOC["GLO",,invert=TRUE][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
+  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
+  facet_wrap(~ Region, scales="free", drop=TRUE) + 
+  scale_size_manual(values=c(4,2,2,2,2)) + 
+  scale_shape_manual(values=c(19,6,2,5,0)) +
+  labs(color  = leg_name, size = leg_name, shape = leg_name) +
+  xlab("") + ylab("SOC (0-30cm) in GtC")
+
+comp
+
+ggsave(paste0(outputdir,"/comparisonfigure.png"), plot=comp)
+
+```
+
 ```{r plot reg stock, out.width="100%", out.height=1000}
@@
               as.quitte(SoilGrids["GLO",,,invert=TRUE]),
-              as.quitte(LPJ_rev21["GLO",,,invert=TRUE][,,2]))
+              as.quitte(LPJmL4["GLO",,,invert=TRUE]))
 
@@
 ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid)
+
 ```

### 2020-11-04 ResultNotebooks/results_SOCmaps.Rmd
update discussion part after bennis suggestions - tex is not compiling

@@
 
-```{r grid plot}
+```{r grid plot, out.height=800 , out.width=1200}
 
@@
         axis.title.y= element_text(size=legend_size),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8)) +
+        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 10, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size)#,
+        #plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
+        ) +
   annotate(geom="text", label="SOC debt", x=factor(1995), y=-35, size = 3) +
@@
           rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7))
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
+          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
           )
@@
 library(cowplot)
+SOCdensity
+SOCincrease
+SOCemis
+SOCshare
+
 grid <- plot_grid(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2, labels=c("a","b", "c", "d"),
@@
 grid
-ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid)
+ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid, width = 20, height = 14, dpi = 300, units = "cm")
 

### 2020-11-04 ResultNotebooks/results_SOCscenario.Rmd
update discussion part after bennis suggestions - tex is not compiling

@@
 
-plot_rev <- c(3,5,6,8)
+plot_rev <- c(5,8,6,3)
 
@@
 
+theme_set(theme_bw())
 legend_size <- 10
@@
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits = c(-45,-35),
+  scale_y_continuous(limits = c(-44,-36),
                      breaks = c(-36, -38, -40,- 42, -44)) +
   theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8))
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size))
 

### 2020-11-04 SOCbudgetPaper/SOCbudget.Rmd
update discussion part after bennis suggestions - tex is not compiling

@@
 abstract: |
-  Soil organic carbon (SOC) is one of larges carbon stocks on earth.
+  Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
 It is three times larger than the biospheric pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.

### 2020-11-04 SOCbudgetPaper/discussion.Rmd
update discussion part after bennis suggestions - tex is not compiling

--- 2020-08-14 SOCbudgetPaper/discussion.Rmd
@@
 
-- The natural land representation and therefor also SOC initialization is lacking proper parametrization of both decay behaviour and natural input (e.g.
-nitrogen and lignin content of litterfall).
-As all results are valued against these potential natural SOC the absolute values and also the changes over time have to be used with caution.
-Especially in boreal zone and in very dry regions the natural land representation might be far off, due to the focus for the parametrization for cropped areas.
-The good agreement of the stock change factors $F^{SCF}$ with the Tier 1 default values might serve as a hint that the natural land representation is not totally off.
-Boreal and very dry zone might be uncropped, leaving the results on SOC changes on cropland untouch from the bad borreal land representation.
+## Including agricultural management data changes the sign of the trend
 
-- This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
+This study provides an analysis of historic SOC stock changes on cropland.
+We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions ($SOC_{natveg}$).
+Our results show, that human activity lead to cummulative SOC emission of around 37 GtC in 2010.
+Whereas recent modelling estimates of global SOC emissions indicate an on going increase of SOC emissions (Sanderman et al, Pugh et al.), our study indicates that the global SOC gap is slowly closing due to improved management.
+
+According to Sanderman et al.
+cummulative SOC emissions since beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+It was also pointed out that these results might be conservatively low compared to experimental results.
+Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with Sanderman et al.
+estimates.
+
+Furthermore the results of Sanderman et al.
+modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Pugh et al.
+considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to oberserved historical data like yields-levels or no-tillage areas were taken into account.
+The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substainal decrease in SOC emissions, but only account for less than 1% change in SOC emissions.
+
+Our study for the first time uses a dynanmic management dataset as driver for SOC dynamics.
+We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighted by improved agricultural yields and practises.
+This is challenging Pugh et al.
+findings of only small effects due to improved pratices.
+Moreover our sensitivity analysis indicades that (1) recycled residue biomass increases driven by higher yields as well as higher higher recycling rates, (2) improved manure recycling (e.g.
+due to improved animal waste management system) and (3) the adoption of no tillage practises.
+
+Modelling management effects on the global scale comes however with parametric and structural uncertainties.
+As pointed out by Keel (2017) and Smith (2019), carbon input calculations are highly sensitive to the choice of allometric functions
+determining below and above ground residue estimates from harvested quantities.
+Keel et al.~questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independend from productivity gains.
+Moreover the study pointed out that plant breeding shift allometries and might been not reflected in outdated data sources.
+Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding.
+However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.
+
+## Modeled management effect in line with default IPCC assumptions
+
+To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (cite) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
+To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.
+
+```{r captionSCFtable, echo=FALSE}
+captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+We compare average values for four IPCC climate zone classifications.
+IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget).}"
+```
+
+```{r SCFtable, out.width = "16cm", fig.cap = captionSCFtable, echo=FALSE}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
+```
+
+Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
+The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.
+
+Looking again on the time trend our study shows the substantial impact of changed management to the development of $F_{SCF}$.
+
+## SOC stocks inline with literature
+
+The worlds SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+
+```{r captionSOCtable, echo=FALSE}
+captionSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
+```
+
+```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSOC_comparison.png")
+```
+
+The global esimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking on regional results (fig.
+SX in supplement), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not superising since the temperature effects on decomposition are fundamentally different for permaforst soils.
+To avoid that this bias influences our results, our study focusses exclusively on cropland soils, excluding most of the boreal zone.
+Morover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+
+Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
+For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
+This leaves carbon inputs and decay behaviour for natural land and pastures rather uncertain.
+The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
+Especially in less forested areas the natural land representation might be off, due to parameterization assumptions of the natural litterfall.
+
+Nevertheless total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetaded areas at all.
+Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions.
+Than former natural land SOC stocks are declining to lower levels.
+However as pointed out before, due to low cropland expansion rates, there SOC declines are outweighted by the general trend of increasing SOC stocks on existing cropland due to management improvements.
+We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parameterization of natural litterfall, which tend to change the global SOC stock substantially for different choices
+(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
+
+## Important short commings --> move to appendix
+
+Smaller points and shortcommings:
+
+-This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
 Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.

### 2020-11-05 SOCbudgetPaper/discussion.Rmd
after spell check

--- 2020-08-14 SOCbudgetPaper/discussion.Rmd
@@
 
-- The natural land representation and therefor also SOC initialization is lacking proper parametrization of both decay behaviour and natural input (e.g.
-nitrogen and lignin content of litterfall).
-As all results are valued against these potential natural SOC the absolute values and also the changes over time have to be used with caution.
-Especially in boreal zone and in very dry regions the natural land representation might be far off, due to the focus for the parametrization for cropped areas.
-The good agreement of the stock change factors $F^{SCF}$ with the Tier 1 default values might serve as a hint that the natural land representation is not totally off.
-Boreal and very dry zone might be uncropped, leaving the results on SOC changes on cropland untouch from the bad borreal land representation.
+## Including agricultural management data changes the sign of the trend
 
-- This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
-Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-It has been argued that for intensivly tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
-Following these argumentation SOC stocks in croplands might even be underestimated.
+This study provides an analysis of historic SOC stock changes on cropland.
+We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions ($SOC_{natveg}$).
+Our results show that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
+Whereas recent modelling estimates of global SOC emissions indicate an ongoing increase of SOC emissions (Sanderman et al, Pugh et al.), our study indicates that the global SOC gap is slowly closing due to improved management.
 
-- As pointed out by Keel (2017), Smith (2019) the results might be highly sensitive to carbon input calculations more precisly to below and above ground residue carbon estimates derived from harvested quantities.
-It has been questioned, that below ground residues might increase with a fixed root:shot ratio (maybe specifically in high end farming systems (?)).
-Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
-(In general the sensitivity to and the uncertainty of certain parameters might be very high.)
+According to Sanderman et al.
+cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil, with half of this value attributed to grazing.
+Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as Sanderman et al.
+estimations.
+However, there are large uncertainties in modelling SOC at the global scale, and the Sanderman et al.
+results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
 
-<!--Tasks: 
-Find better natural land representation and initalization.
-Using LPJmL values might be complex since the split up into SOC subpools is needed.
-Getting better parametrization for the CENTURY based approach for decay behaviour and better values for nitrogen and lignin of natural litter fall.
--->
+Furthermore, Sanderman et al.
+modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Pugh et al.
+considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account.
+The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decrease in SOC emissions, accounting for only less than 1% change in SOC emissions.
 
-\newpage
+Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
+We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improved agricultural yields and practices.
+This challenges Pugh et al.'s findings of only small effects due to improved practices.
+Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g.
+due to improved animal waste management systems) and (3) the adoption of no tillage practices.
 
-Smaller points and shortcommings:
+Modelling management effects on the global scale comes however with parametric and structural uncertainties.
+As pointed out by Keel (2017) and Smith (2019), carbon input calculations are highly sensitive to the choice of allometric functions
+determining below and above ground residue estimates from harvested quantities.
+Keel et al.
+question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources.
+Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
+However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass.
+Rather, it is more likely that we are still missing carbon inputs to the soil.
 
-- Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
-Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer ammounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
+## Modeled management effect in line with default IPCC assumptions
 
-- Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
-(Note that also manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.)
+To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (cite) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
+To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
 
-- No tillage adaptaion is neglected on cropland due to less common adoption of no tillage and conservation agriculture.
-Pastures are assumed to be not tilled at all (propably only heavy managed pastures are tilled with some rotation)
+```{r captionSCFtable, echo=FALSE}
+captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+We compare average values for four IPCC climate zone classifications.
+IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget).}"
+```
+
+```{r SCFtable, out.width = "16cm", fig.cap = captionSCFtable, echo=FALSE}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
+```
+
+Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006.
+For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state.
+The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21% of it.
+
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$.
+
+## SOC stocks inline with literature
+
+The worlds SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+
+```{r captionSOCtable, echo=FALSE}
+captionSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
+```
+
+```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSOC_comparison.png")
+```
+
+The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking at regional results (fig.
+SX in supplement), our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not suprising as the temperature effects on decomposition are fundamentally different for permafrost soils.
+To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
+Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+
+Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
+For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
+This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain.
+The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
+Especially in less forested areas the natural land representation higher uncertainty may arise from the parametrization assumptions of natural litterfall.
+
+Nevertheless, total SOC stocks are in a reasonable range and results for SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
+The SOC gap from changed management is only affected by the natural vegetation representation in the case of land use transitions.
+In this case, former natural land SOC stocks decline.
+However, as pointed out before, due to low cropland expansion rates, land use change-related SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
+We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
+(see appendix). This shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
+
+## Important shortcomings --> move to appendix
+
+Smaller points and shortcomings:
+
+-This study as the IPCC guidelines suggested has limited its focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
+Nevertheless, studies (see Don on tillage) have shown the subsoil to be very influential in evaluating total SOC losses or gains for no-tillage systems.
+It has been argued that for intensively tilled soils, subsoil SOC increases due to the import of carbon rich topsoil to deeper soil layers.
+Following this argumentation, SOC stocks in croplands might even be underestimated.
+
+- Fertilizer interaction is not included here by accounting for additional N supply that would alter the C:N ratio of the carbon inputs.
+For the IPCC Tier 2 steady-state method, fertilizer application is neglected; However we would have fertilizer amounts available to include, if proper representation of fertilizer within the method would be possible to add.
+
+- Pasture dynamics are neglected and treated as natural vegetation, which may be -- especially with regard to pasture degradation due to overgrazing -- oversimplified for some areas, but is in line with assumptions on previous pasture SOC stocks (see Tier 1 IPCC).
+(Note that also manure excreted to pastures is neglected within this analysis, since we focus purely on cropland dynamics.)
+
+- No-tillage practices are neglected on cropland due to the less common adoption of no tillage and conservation agriculture.
+Pastures are assumed to be not tilled at all (probably only heavily managed pastures are tilled with some rotation)
 
 - Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
-Crop specific growing periods might be possible using LPJmL data.
+Crop specific growing periods may be possible using LPJmL data.
 
-- flooded rice area are not represented correctly as parameterization does not hold true for flooded conditions.
+- flooded rice area are not represented correctly as the parametrization does not hold true for flooded conditions.
 
@@
 
-- The disaggregation of manure to uild-up areas (in the case of extensive monograstrics) is leeding to a lot of displaced manure (?) that is cut off
+- The disaggregation of manure to build-up areas (in the case of extensive monogastric) is leading to a lot of displaced manure (?) that is cut off
 
-- It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
+- It is known that there are mismatches between FAO statistics and LUH areas. As far as possible these were harmonized within this study.
 
 - "The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
-current C stock to the future steady-state soil C stock calculated based on current conditions." Leading to the fact that our total stock results are highly uncertained.
+current C stock to the future steady-state soil C stock calculated based on current conditions." This leads to the fact that our total stock results are highly uncertain.

### 2020-11-09 ResultNotebooks/results_SOCcompare.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

--- 2020-10-06 ResultNotebooks/results_SOCcompare.Rmd
@@
 rev       <- "rev0.76baseline"
-outputdir <- "Output/"
+outputdir <- "Output/Images/"
 
@@
 
-```{r plot reg stock, out.width="100%", out.height=1000}
-
-library(mip)
-library(quitte)
+```{r point plot reg stock, out.width=800, out.height=600}
 library(ggplot2)
 
-getNames(reg_SOCStock) <- paste0("modeloutput.SOCbudget.", getNames(WISE,dim=3))
-getSets(reg_SOCStock)  <- getSets(WISE)
+leg_name <- "Data Source"
 
-hist <- rbind(as.quitte(WISE["GLO",,,invert=TRUE]),
-              as.quitte(GSOC["GLO",,,invert=TRUE]),
-              as.quitte(SoilGrids["GLO",,,invert=TRUE]),
-              as.quitte(LPJ_rev21["GLO",,,invert=TRUE][,,2]))
+glob <- ggplot(data = as.data.frame(glo_SOC["GLO",,][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
+  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
+  scale_size_manual(values=c(4,2,2,2,2)) + 
+  scale_shape_manual(values=c(19,6,2,5,0)) +
+  labs(color  = leg_name, size = leg_name, shape = leg_name) +
+  xlab("") + ylab("SOC (0-30cm) in GtC")
 
-regSOCcompare <- mipLineHistorical(reg_SOCStock, hist, facet.ncol=4, scales="free_y", ylab="SOC (0-30cm) in MtC")
+comp <- ggplot(data = as.data.frame(glo_SOC["GLO",,invert=TRUE][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
+  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
+  facet_wrap(~ Region, scales="free", drop=TRUE) + 
+  scale_size_manual(values=c(4,2,2,2,2)) + 
+  scale_shape_manual(values=c(19,6,2,5,0)) +
+  labs(color  = leg_name, size = leg_name, shape = leg_name) +
+  xlab("") + ylab("SOC (0-30cm) in GtC")
 
-ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid)
+glob
+comp
+
+ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob, height=6, width=3, units="cm")
+ggsave(paste0(outputdir,"/reg_comparisonfigure.png"), plot=comp)
+
 ```
 
-```{r bar plot}
+```{r point plot glo stock, fig.width=3, fig.height=4}
+library(ggplot2)
 
-### FROM ISABELLE
-
-# p.tc <- ggplot() 
-# # agass Hauptszenarien --> Balken
-# p.tc <-p.tc + geom_bar(data = dat1a, aes(x = Region, y = Value, fill=Mainscen,group=factor(Mainscen)),stat = "identity",width = 0.7,position=position_dodge(0.7),alpha=0.8, show_guide=FALSE) 
-# p.tc <- p.tc + scale_fill_manual(values = c("#999999","#FF3300","#66CC00","#3366FF"),breaks=c("baseline","S1a","S3a","S3c"),
-#                                  labels=c("Baseline     ","Climate_impacts     ","Shift_to_rangeland     ","Shift_to_mixed     "))
-# # agass mit co2 --> graue K?stchen
-# p.tc <- p.tc + geom_point(data=dat2a,aes(x = Region, y = Value, group=factor(Mainscen)),size=4,position=position_dodge(0.7),colour="black",shape=22,fill="grey")
-# 
-# # Spread GCMs ohne co2
-# p.tc <- p.tc + geom_errorbar(data=dat4a,aes(x = Region, ymin = min,ymax=max, group=Mainscen),position=position_dodge(0.7),width=.2,colour="dark red", linetype="dashed",size=0.6,show_guide=FALSE)
-# #p.tc <- p.tc + geom_point(data=dat3a,aes(x = Region, y = Value, group=factor(Mainscen)),size=1.5,position=position_dodge(0.7),colour="red", show_guide=FALSE)
-# p.tc <- p.tc + geom_point(data=dat4a,aes(x = Region, y = min, group=factor(Mainscen)),size=3.8,position=position_dodge(0.7),colour="dark red",shape=21,fill="white", show_guide=FALSE)
-# p.tc <- p.tc + geom_point(data=dat4a,aes(x = Region, y = max, group=factor(Mainscen)),size=2.5,position=position_dodge(0.7),colour="dark red",shape=19, show_guide=FALSE)
-# 
-# # Spread GCMs mit co2
-# p.tc <- p.tc + geom_errorbar(data=dat4b,aes(x = Region, ymin = min,ymax=max, group=Mainscen),position=position_dodge(0.7),width=.2,colour="#003300",size=0.6, show_guide=FALSE)
-# #p.tc <- p.tc + geom_point(data=dat3b,aes(x = Region, y = Value, group=factor(Mainscen)),size=1.5,position=position_dodge(0.7),colour="#003300", show_guide=FALSE)
-# p.tc <- p.tc + geom_point(data=dat4b,aes(x = Region, y = min, group=factor(Mainscen)),size=3.5,position=position_dodge(0.7),colour="#003300",shape=23,fill="white", show_guide=FALSE)
-# p.tc <- p.tc + geom_point(data=dat4b,aes(x = Region, y = max, group=factor(Mainscen)),size=3.8,position=position_dodge(0.7),colour="#003300",shape=18, show_guide=FALSE)
-# 
-# #Layout
-# p.tc <- p.tc + labs(y="Average annual TC rates 2005-2045 [%] \n",x="")
-# guides(colour = guide_legend(override.aes = list(shape = NA)))
-# p.tc <- p.tc + geom_hline(yintercept=0)
-# 
-# #p.tc <- p.tc + ylim(0, 2.4)
-# p.tc <- p.tc + theme_bw() +
-#   theme(axis.text = element_text(size = 16,face="bold"),
-#         axis.title = element_text(size = 18,face="bold"),
-#                 
-#         legend.background = element_rect(fill = "white",colour="black",size=0.7),
-#         legend.text =       element_text(size = 14, face="bold"),
-#         legend.title =      element_blank(),
-#         legend.position = c(0.25, 1.01),
-#         legend.key = element_rect(colour = "white",size=2),
-#         legend.key.height = unit(0.7,"cm"),
-#         legend.key.width = unit(1.8,"cm"),
-#         legend.margin = unit(100,"lines"),
-#         
-#         panel.grid.major = element_blank(),
-#         panel.grid.minor = element_blank(),
-#         
-#         panel.border = element_rect(size = 0.8, colour = "black"),
-#         plot.margin = unit(c(0.5,0.5,0.5,0.5),"cm")  # top, right, bottom, and left margins
-#   )
-# 
-# #Legend 1
-# p.tc <- p.tc + annotate("rect", xmin = 0.6, xmax = 4.5, ymin = 2.3, ymax = 3.3,fill = "white",colour="black",size=0.7)
-# 
-# p.tc <- p.tc + annotate("text", x = 0.8, y = 3.1, label = "IAASTD~climate~scenario~w/o~CO[2]~effect",parse = TRUE,size = 4.7, hjust=0)
-# 
-# p.tc <- p.tc + annotate("text", x = 1.8, y = 2.9, label = "climate_impacts",size = 4.7, hjust=0)
-# p.tc <- p.tc + annotate("rect", xmin = 0.8, xmax = 1.6, ymin = 2.85, ymax = 2.95,fill = "#FF3300",size=0.7)
-# 
-# p.tc <- p.tc + annotate("text", x = 1.8, y = 2.7, label = "shift_to_rangeland",size = 4.7, hjust=0)
-# p.tc <- p.tc + annotate("rect", xmin = 0.8, xmax = 1.6, ymin = 2.65, ymax = 2.75,fill = "#66CC00",size=0.7)
-# 
-# p.tc <- p.tc + annotate("text", x = 1.8, y = 2.5, label = "shift_to_mixed",size = 4.7, hjust=0)
-# p.tc <- p.tc + annotate("rect", xmin = 0.8, xmax = 1.6, ymin = 2.45, ymax = 2.55,fill = "#3366FF",size=0.7)
-# 
-# 
-# #Legend 2
-# p.tc <- p.tc + annotate("rect", xmin = 7.6, xmax = 11.4,  ymin = 2.3, ymax = 3.3,fill = "white",colour="black",size=0.7)
-# 
-# p.tc <- p.tc + annotate("text", x = 9, y = 3.0, label = "full~CO[2]~effect~(IAASTD)",parse = TRUE, size = 4.7, hjust=0)
-# p.tc <- p.tc + annotate("point", x = 8.4, y = 3.0, colour="black",shape=22,fill="grey",size=4.5)
-# 
-# p.tc <- p.tc + annotate("text", x = 9, y = 2.8, label = "full~CO[2]~effect~(all~GCMs)",parse = TRUE, size = 4.7, hjust=0)
-# p.tc <- p.tc + annotate("segment", x = 8, xend=8.8, y = 2.8, yend=2.8,colour="#003300",size=0.7)
-# p.tc <- p.tc + annotate("point", x = 8, y = 2.8, colour="#003300",shape=23,fill="white",size=3.8)
-# p.tc <- p.tc + annotate("point", x = 8.8, y = 2.8, colour="#003300",shape=18,size=5)
-# 
-# p.tc <- p.tc + annotate("text", x = 9, y = 2.6, label = "w/o~CO[2]~effect~(all~GCMs)",parse = TRUE,size = 4.7, hjust=0)
-# p.tc <- p.tc + annotate("segment", x = 8, xend=8.8, y = 2.6, yend=2.6,colour="dark red",linetype="dashed",size=0.7)
-# p.tc <- p.tc + annotate("point", x = 8, y = 2.6, colour="dark red",shape=21,fill="white",size=4)
-# p.tc <- p.tc + annotate("point", x = 8.8, y = 2.6, colour="dark red",shape=19,size=4)
-# 
-# 
-# print(p.tc)
-
+glob
+ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob)
 ```

### 2020-11-09 ResultNotebooks/results_SOCsceninit.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

--- 2020-10-06 ResultNotebooks/results_SOCscenario.Rmd
@@
 
-rev <- "rev0.76baseline"
+rev <- "rev0.77init_lu"
 
@@
 
-#revisions   <- c(Baseline="rev0.74baseline", Till="rev0.74tillage_mixedtill", ResRe="rev0.74residue_freeze1975", ManRe="rev0.74manure_freeze1975", Yield="rev0.74yield_freeze1975")
+revisions   <- c(InitLu="rev0.77init_lu", InitNv="rev0.77init_natveg", InitMix="rev0.77init_mixed")
 filesNeeded <- c("SoilCarbon.rds","Landuse.rds")
-
-revisions   <- c(Baseline="rev0.76baseline", Allon="rev0.76allon_freeze1975", Allon2="rev0.76allon2_freeze1975",        ResRe="rev0.76rrecycle_freeze1975",
-                Till="rev0.76tillage_mixedtill", ResSta="rev0.76residue_freeze1975", Yield="rev0.76yield_freeze1975", ManRe="rev0.76manure_freeze1975")
 
@@
 
-plot_rev <- c(3,5,6,8)
+plot_rev <- c(2,3)
 
-names(revisions) <- c("Historical Management", 
-                      "Allon",    
-                      "Constant management from 1975",   
-                      "ResRe", 
-                      "Constant tillage area from 1975",
-                      "Constant residue input from 1975",
-                      "Yield", 
-                      "Constant manure application from 1975")
+names(revisions) <- c("steady state of land use","steady state of potential natural vegetation","mean of steady states above")
 
@@
 
+theme_set(theme_bw())
 legend_size <- 10
 
-SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission[,seq(1975,2010,5),c(1,plot_rev)]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
+SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission[,seq(1975,2010,1),c(1,plot_rev)]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
   labs( x = "Year", 
         y = "Global SOC gap in GtC",
-        color="") +
+        color="Initilization of SOC stocks in 1960 by") +
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits = c(-45,-35),
-                     breaks = c(-36, -38, -40,- 42, -44)) +
+  scale_y_continuous(limits = c(-45,-0),
+                     breaks = seq(-45,0,5)) +
   theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8))
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size))
 
@@
 
-ggsave(paste0(outputdir,"/scenario.png"), plot=SOCemis)
+ggsave(paste0(outputdir,"/scenario_lu.png"), plot=SOCemis)
 

### 2020-11-09 SOCbudgetPaper/SOCbudget.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

@@
   Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
-It is three times larger than the biospheric pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
+It is three times larger than the vegetation pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
 Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
@@
 We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
-bibliography: SOCbudget.bib
+bibliography: SOCbudget2.bib
 running:
@@
 output:
-  rticles::copernicus_article: default
-  md_document:
-    variant: markdown_github
   bookdown::pdf_book:
@@
     md_extensions: +raw_attribute
+  md_document:
+    variant: markdown_github
   bookdown::word_document2: default
+  rticles::copernicus_article: default
   bookdown::pdf_document2: default

### 2020-11-09 SOCbudgetPaper/SOCbudget.tex
update with spell check, references for discussion part, additional analysis on initialization, update of figures

--- 2020-10-06 SOCbudgetPaper/SOCbudget.tex
@@
 %% 2-column papers and discussion papers
+\RequirePackage[2020/02/02]{latexrelease}
 \documentclass[gc, manuscript]{copernicus}
@@
 % Hydrology and Earth System Sciences (hess)
+% Journal of Bone and Joint Infection (jbji)
 % Journal of Micropalaeontology (jm)
@@
 
+% Pandoc citation processing
+
 % The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
@@
 
-\title{Management induced changes of soil organic carbon on global
-croplands}
+\title{Management induced changes of soil organic carbon on global croplands}
 
@@
 
-\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam,
-Germany}
+\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
 \affil[2]{Your affiliation}
@@
 
-\correspondence{Kristine\ Karstens\ (kristine.karstens@pik-potsdam.de)}
+\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}
 
@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of larges carbon stocks on earth. It is
-three times larger than the biospheric pool and more than twice as big
-as the athmospheric pool, when looking into the first meter of the earth
-soil profile. Human cropping activties led and still lead to a depletion
-of SOC though, which are so far not well represented in global
-assessments of historic carbon emissions. While SOC models often
-represent well the biochemical processes that lead to the accumulation
-and decay of SOC, the management decisions driving these biophysical
-processes are still little investigated on the global scale. Here we
-create a spatial explicit data set for agricutural management on
-cropland, especially for crop residue and manure management, based on
-global historic production (FAOSTAT), and land-use (LUH2) data; and
-combine it with the IPCC Tier 2 approach to create a half degree
-resolution SOC stock changes on mineral soils. We estimate that due to
-arable farming soils have lost over 37 GtOC of which 4 GtOC have been
-regained within the period 1975-2010. We show that, our results on
-global scale based on Tier 2 IPCC methodolgy are in good agreement with
-Tier 1 default assumptions. We also find that SOC is very sensitive to
-management decision such as residue recycling indicating the nessessity
-to incorporated better management data in soil models.
+Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
+It is three times larger than the vegetation pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
+Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
+While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
+Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
+We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
+We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
+We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
 \end{abstract}
 
-\copyrightstatement{The author's copyright for this publication is
-transferred to institution/company.}
+\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
 
@@
 
-Soil Organic Carbon (SOC), the amount of organic carbon stored through
-the earth's soil, is the largest terrestrial carbon pool, exceeding the
-carbon in the atmospheric and vegetation pools multiple times (Batjes).
-As such, even small changes in drivers of SOC may thus lead to
-substantial shifts in earth carbon cycle and influence the amount of CO2
-in the atmosphere (ref. permafrost melting). The specific amount of
-carbon stored in the soil is, however, uncertain, with estimates ranging
-from 1500 to 2400 GtC for the first meter of the soil profile (Batjes,
-1996; more). The quality of SOC maps has markedly improved in recent
-decades, along with better understanding of the factors driving the
-global magnitude, distribution, and dynamics of SOC pools. Natural
-properties like climatic, biophysical, and landscape characteristics
-clearly play the most important roles to determine SOC variations over
-space and time. Human intervention, including land cover change and land
-management, has however added a further driver to SOC change, which
-alters terrestrial carbon pools in much shorter time scales and is
-likely one of the most dominant driver of SOC changes on managed land
-today (cite). Recent studies identify the anthropogenic SOC debt for the
-first meter of the soil profile at around 116 GtC (Sanderman et al.),
-compared to previous estimates between 60-130 GtC (Lal, 2006). Other
-studies have focused more closely on spatially disaggregation of SOC
-changes via advanced digital soil mapping techniques (S-World;
-Stoorvogel 2, 2017) or better representation of biogeochemical processes
-within SOC dynamics (Hararuk, 20XX). While providing a rough estimate of
-the order of magnitude of change, these studies lack a detailed
-consideration of land management, especially of agricultural activities.
-Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to
-capture these land management impacts by using detailed information on
-crop yield levels, fertilizer inputs and various other on-farm measures.
-However, due to the lack of comprehensive global management data as
-input to these models, scaling up to the global extent remains a complex
-challenge. Our study combines an spatially explicit estimate of
-agricultural management data on the global level with a 3-pool SOC model
-parametrized for global croplands. This allows us to estimate SOC stock
-change factors, as well as organic carbon flow dynamics within the
-agricultural system. We thereby consider change in SOC caused by
-historical land cover change as well as of different agricultural
-management practices, including residue recycling, manure amendments,
-irrigation, and tillage. We provide the first global, spatially explicit
-SOC loss maps that considers detailed agricultural management. This
-paper will first introduce in the methods section the basic concept of
-SOC dynamics as applied in this study. We continue with a detailed
-description of the global gridded management data used here, including
-crop production levels, residue recycling rates, manure amendments, and
-the adoption of irrigation and tillage practices. Lastly, we shortly
-refer to the concept of stock change factors as outlined in the Tier 1
-approach of the IPCC guidelines. In our results section we focus on the
-SOC dynamics of global croplands by (1) analysing the spatially explicit
-distribution and depletion of SOC from 1975 to 2010, (2) comparing
-impacts of different management effects on SOC emissions and (3)
-quantify global agricultural carbon flows and stocks to compare the
-importance of various management aspects. Finally, we discuss and
-compare our findings -- including their implications for SOC model
-development -- and conclude with an outlook on the ability of SOC
-management to mitigate climate change and contribute to negative
-greenhouse gas emissions. \newpage
+Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and vegetation pools multiple times (Batjes).
+As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref.
+permafrost melting).
+The specific amount of carbon stored in the soil is, however, uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Batjes, 1996; more).
+The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
+Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools over much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite).
+Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC (Lal, 2006).
+Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX).
+While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especially of agricultural activities.
+Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to capture these land management impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm measures.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
+Our study combines a spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
+This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
+We thereby consider change in SOC caused by historical land cover change as well as of different agricultural management practices, including residue recycling, manure amendments, irrigation, and tillage.
+We provide the first global, spatially explicit SOC loss maps that considers detailed agricultural management.
+This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
+Additionally, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+Lastly, we continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
+In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
+Finally, we discuss and compare our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+\newpage
 
-\section{Method}
+\hypertarget{method}{%
+\section{Method}\label{method}}
 
-We compile calculations as open-source R packages available at
-github.com/pik-piam/mrcommons (management related functions) and
-github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which
-are both based on the MADRaT package (package citation?), a framework
-which aims to improve reproducibility and transparency in data
-processing. In the following chapter we outine the most important
-relationships and assumptions. Table @ref(append:subsection2mrfunctions)
-provides further information on corresponding code within the R
-packages.
+We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
+In the following section we outline the most important relationships and assumptions.
+Table \ref{append:subsection2mrfunctions} provides further information on corresponding code within the R packages.
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{SOC Stocks and Stock changes following the Tier 2
-steady-state method}\label{sec:carbonbudget}}
+\subsection{SOC Stocks and Stock changes following the Tier 2 steady-state method}\label{sec:carbonbudget}}
 
-Following the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 (\citet{ipcc_2019_2019}; short Tier 2 steady-state method) , we
-estimate soil organic carbon (SOC) stocks and their change over time for
-cropland on half-degree resolution from 1975 to 2010. We assume the
-current SOC state converges towards a steady state, which itself is
-depending on biophysical, climatic and agronomic conditions. Therefore
-we conduct the following three steps within each yearly timestep: (1) We
-calculate annual land-use (sub-)type-specific steady states and decay
-rates for SOC stocks, (2) we account for land conversion by transferring
-SOC from and to an other land-use type representad as natural
-vegetation, (3) we estimate SOC stocks and changes based on the stocks
-of the previous time period, the steady state stocks and the decay rate.
+Following the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}; short Tier 2 steady-state method), we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
+We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
+Therefore, we conduct the following three steps within each yearly time-step:
+(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks,
+(2) we account for land conversion by transferring SOC from and to another land-use type represented as natural vegetation,
+(3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
 
-\subsubsection{Steady-state SOC stocks and decay rates}
+\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
+\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}
 
-In a simple first order kinetic approach the steady-state soil organic
-carbon stocks \(SOC^{eq}\) are given by \begin{equation}
+In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
+\begin{equation}
 SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
-(\#eq:inoutflow)
-\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil,
-\(k\) denoting the soil organic carbon decay rate; as well as \(i\)
-representing grid cell indices and \(t\) years. We use for our
-calculations the Tier 2 steady-state method, which assumes three soil
-carbon sub-pools (active, slow and passive) and interactions between
-them. Annual carbon inflow to each sub-pool and annual decay rates of
-each sub-pool are the key components to determining steady-state SOC
-stocks.
+\label{eq:inoutflow}
+\end{equation}
+with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cell identifier and \(t\) years.
+We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them, following the approach in the Century model (Parton et al, 1988).
+Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
 
@@
 
-We account for different carbon input sources depending on the two
-land-use types we distinguish: croplands and natural vegetated land as
-representative for all other land use (see table
-@ref(tab:datasourceinputs)). Carbon sources for cropland are recycled
-crop residues, below ground biomass of crops (for both see
-@ref(sec:residues)) and recycled manure (see @ref(sec:livstmanure)); for
-natural vegetation litterfall \citep{schaphoff_lpjml4_2018} is the only
-source of carbon inflow to the soil.
+We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \ref{tab:datasourceinputs}).
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see \ref{sec:livstmanure}); for natural vegetation litterfall (including fine root turnover) \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
 
-Following the IPCC carbon accounting methodology, carbon inputs are
-disaggregated into metabolic and structural components depending on
-their lignin and nitrogen content (see \citet{ipcc_2019_2019}). For each
-component the sum of all carbon input sources is allocated to the
-respective SOC sub-pools via transfer coefficients. This implies that
-both the amount of carbon as well as its structural composition
-determine the effective inflow. Data sources for all considered carbon
-inputs as well as for lignin and nitrogen content can be found in table
-@ref(tab:datasourceinputs).
+Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citet{ipcc_2019_2019}).
+For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
+This implies that both the amount of carbon and its structural composition determine the effective inflow.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
 
@@
 
-The sub-pool specific decay rates are influenced by climatic conditions,
-biophysical and biochemical soil properties as well as management
-factors that all vary over time (t) and space (i). Following the Tier 2
-steady-state method we consider temperature (temp), water (wat), sand
-fraction (sf) and tillage (till) effects to account for spatial
-variation of decay rates. Thus \(k_{sub}\) is given by
+The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
+Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
+Thus, \(k_{sub}\) is given by
 
@@
 & k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
-(\#eq:decayrates)
+\label{eq:decayrates}
 \end{aligned}
@@
 
-For cropland we distinguish the effect of different tillage (see
-@ref(sec:tillage)) and irrigation (see @ref(sec:irrigation)) practices
-on decay rates, whereas for natural vegetation, we assume rainfed and
-non-tilled conditions. Data sources as well as drivers considered for
-each land-use types are shown in table @ref(tab:datasourcedecay). To
-account for variations of decay rates within each grid cell due to
-different tillage and irrigation regimes, average rates based on area
-shares are calculated.
+For cropland, we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
+Data sources as well as drivers considered for each land-use types are shown in table \ref{tab:datasourcedecay}.
+To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
 
@@
  \belowtable{}
- (\#tab:datasourcedecay)
+ \label{tab:datasourcedecay}
  \end{table*}
 
-\subsubsection{SOC transfer between land-use types}
+\hypertarget{soc-transfer-between-land-use-types}{%
+\subsubsection{SOC transfer between land-use types}\label{soc-transfer-between-land-use-types}}
 
-We calculate SOC stocks based on the area shares of land-use types (lut)
-within the half-degree grid cells (i). If land is converted from one
-land-use type into others (!lut), the respective share of the SOC stocks
-is reallocated. We account for land conversion at the beginning of each
-time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
+We calculate SOC stocks based on the area shares of land-use types (lut) within the half-degree grid cells (i).
+If land is converted from one land-use type into others (!lut), the respective share of the SOC stocks is reallocated.
+We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
 
@@
 SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
-(\#eq:ctransfer)
+\label{eq:ctransfer}
 \end{equation}
 
-with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
-expansion for a given land-use type \(lut\). Note that \(!lut\) denotes
-the sum over all other land-use types, which decreases in the specific
-time step \(t\). Data sources and methodology on land-use states and
-changes are described in @ref(sec:landuse).
+with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land-use type \(lut\).
+Note that \(!lut\) denotes the sum over all other land-use types, which decreases in the specific time step \(t\).
+Data sources and methodology on land-use states and changes are described in \ref{sec:landuse}.
 
-\subsubsection{Total SOC stocks and stock changes}
+\hypertarget{total-soc-stocks-and-stock-changes}{%
+\subsubsection{Total SOC stocks and stock changes}\label{total-soc-stocks-and-stock-changes}}
 
-Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
-calculated steady-state stock \(SOC^{eq}\) for each land-use type (lut),
-each sub-pool (sub) and each annual time step (t) like
+Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use type (lut), each sub-pool (sub) and each annual time step (t) like
 
@@
 SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
-(\#eq:SOCstate)
+\label{eq:SOCstate}
 \end{equation}
 
-Reformulating this equation, we obtain a massbalance equation as follows
+Reformulating this equation, we obtain a mass balance equation as follows
 
@@
 SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
-(\#eq:steadystate2budget)
+\label{eq:steadystate2budget}
 \end{equation}
 
-The global SOC stock for each time step can than be calculated via
+The global SOC stock for each time step can then be calculated via
 
@@
 SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
-(\#eq:totalstock)
+\label{eq:totalstock}
 \end{equation}
 
-According to the IPCC guidelines SOC changes can be calculates as the
-difference of two (aufeinander folgenden) years (see Eq. 5.0A in
-\citep{ipcc_2019_2019}). This however will also include naturally
-occured changes due to climatic variance over time. For our study we
-will define the absolute and relative SOC changes in relation to a
-potential natural state \(SOC_{pnv}\) under the same climatic conditions
-at time \(t\) in grid cell \(i\), that is based on the natural
-vegetation SOC calculations as defined above without accounting for land
-conversion from cropland at any time. The absolute changes
-\(\Delta SOC\) and relative changes \(F^{SCF}\) a thus given by
+According to the IPCC guidelines SOC changes can be expressed as the difference of two consecutive years (see Eq.
+5.0A in \citep{ipcc_2019_2019}).
+This however, will also include naturally occurring changes due to climatic variance over time.
+For our study we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC_{pnv}\) under the same climatic conditions at time \(t\) in grid cell \(i\), that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
+The absolute changes \(\Delta SOC\) and relative changes \(F^{SCF}\) a thus given by
 
@@
 \Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
-(\#eq:totalstock)
+\label{eq:totalstock}
 \end{equation}
 
-Note that the absolution changes \(\Delta SOC\) can be also interpreted
-as the SOC gap due to human cropping activits; whereas relative changes
-\(F^{SCF}\) can be denoted as stock change factors as defined within the
-IPCC guidelines of 2006 \citep{ipcc_2006_2006}. Note that the SOC gap is
-to the negated cummulative SOC emission also refered to as SOC debt
-(cite Sanderman) of mankind.
+Note that the absolution changes \(\Delta SOC\) can be also interpreted as the SOC gap due to human cropping activities; whereas relative changes \(F^{SCF}\) can be denoted as stock change factors as defined within the IPCC guidelines of 2006 \citep{ipcc_2006_2006}.
+Note that the SOC gap is equivalent to the negated cumulative SOC emission (\citep{pugh_simulated_2015}) also referred to as SOC debt (\citep{sanderman_soil_2017}) of human land use.
 
-\subsubsection{Initialisation of SOC pools}
+\hypertarget{initialization-of-soc-pools}{%
+\subsubsection{Initialization of SOC pools}\label{initialization-of-soc-pools}}
 
-To initialize all SOC sub-pools we assume that cropped land as well as
-natural vegetation are in a steady state for the specific configuration
-present within the initialization year 1961. We assume after a spin-up
-period of 15 years reliable results start from 1975 and improve over
-time, since dependency on the initial conditions will decrease.
+To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions decreases.
+
+\hypertarget{sec:tier1}{%
+\subsection{SOC Stocks and Stock changes following Tier 1}\label{sec:tier1}}
+
+Additionally, to the tier 2 approach of the refinement of IPCC guidelines vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}).
+Here, stocks are calculated via stock change factors (\(F^{SCF}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+\(F^{SCF}\) are differentiated by different crop, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover, \(F^{SCF}\) vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}).
+The actual SOC stocks as thus calculated based on a given reference SOC stock by
+
+\begin{equation}
+SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
+\label{eq:tier1}
+\end{equation}
+
+with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
+For this analysis we use the default \(F^{SCF}\) from the Tier 1 method of \citep{ipcc_2006_2006}, and \citep{ipcc_2019_2019} as a comparison and consistency check for our more detailed budget approach.
 
@@
 \hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data on 0.5 degree
-resolution}\label{sec:agrimanagement}}
+\subsection{Agricultural management data on 0.5 degree resolution}\label{sec:agrimanagement}}
 
-We compile country-specific FAO production and cropland statistics
-\citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5 year time steps from 1965 to 2010, which,
-together with the spin up phase, restricts our analysis to time span
-from 1975 to 2010. For all the following data, if not declared
-differently, we interpolate values linearly between the time steps and
-hold them constant before 1965.
+We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
+The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
 
 \hypertarget{sec:landuse}{%
-\subsubsection{Landuse and Landuse Change}\label{sec:landuse}}
+\subsubsection{Land use and Land-use change}\label{sec:landuse}}
 
-Land-use patterns are based on the Land-Use Harmonization 2
-\citep{Hurtt2020} data set, which we aggregate from quarter degree to
-half degree resolution. We disaggregate the five different cropland
-subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17
-crop groups, applying the relative shares for each grid cell based on
-the country and year specific area harvested shares of FAOSTAT data
-\citep{faostat_faostat_2016} (see @ref(append:Tableluh2fao2mag) for more
-details on the crop group mapping). Land-use transitions are calculated
-as net area differences of the land-use data on half-degree.
+Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we aggregate from quarter degree to half degree resolution.
+We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see \ref{append:Tableluh2fao2mag} for more details on the crop group mapping).
+Land-use transitions are calculated as net area differences of the land-use data on half-degree.
 
@@
 
-Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as
-well as half-degree cropland patterns (see @ref(sec:landuse)) we compile
-crop group specific half-degree production patterns. We calibrate
-cellular yields with a country-level calibration factor for each crop
-group to meet historical FAOSTAT production
-\citep{faostat_faostat_2016}. Note that by using physical cropland areas
-we account for multiple crop harvest events as well as for fallows.
+Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns.
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
+Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
 
@@
 
-Crop residue production and management is based on a revised methodology
-of \citep{bodirsky_n2o_2012} and key aspects are explained here given
-its central role for soil carbon modelling. Starting from harvested crop
-production (\(CP\)) estimates and their respective harvested crop area
-(\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\))
-residual biomass using crop group (\(cg\)) specific harvest indices
-(\(HI\)) and root:shoot ratios (\(RS\)) as follows
+Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modelling.
+Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
 
@@
 BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
-(\#eq:resbiomass)
+\label{eq:resbiomass}
 \end{aligned}
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield
-and an area dependend fraction \citep{ipcc_2006_2006}. Note that
-deviating from \citep{bodirsky_n2o_2012} we use harvested instead of
-physical crop area to account for increased residue biomass due to
-multiple cropping and decreased amounts on fallow land. We assume that
-all BGR are recycled to the soil, whereas AGR can be burned or harvested
-for other purposes such as feeding animals \citep{weindl}, fuel or for
-material use.
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction \citep{ipcc_2006_2006}.
+Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
+We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
 
@@
 
-A fixed share of the AGR is assumed to be burned on field depending on
-the per-capita income of the country. Following \citep{smil1999}) we
-assume 25\% burn share for low-income countries according to worldbank
-definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income
-(\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for
-all middle-income countries depending on their per-capita income.
-Depending on the crop group 80--90\% of the residue carbon burned on the
-fields are lost within the combustion process \citep{ipcc_2006_2006}.
+A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
+Following \citep{smil_nitrogen_1999}) we assume 25\% burn share for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process \citep{ipcc_2006_2006}.
 
@@
 
-We compile out of our 17 crop groups, three residue groups (straw,
-high-lignin and low-lignin residues) with additional demand for other
-purposes and one residue with no double use (see
-@ref(append:Tablekcr2kres)). Residue feed demand for five different
-livestock groups is based on country- and residue-group-specific feed
-baskets (see \citep{weindl}), taking available AGR biomass as well as
-livestock productivity into account. We estimate, for low-income
-countries, a material-use share for straw residues of 5\% and a
-fuel-share of 10\% for all used residues groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is
-assumend, and middle-income countries use shares are linearly
-interpolated based on per-capita income. The remaining AGR as well as
-all BGR are assumend to be recycled to the soil. We limit high recycling
-shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
+We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \ref{append:Tablekcr2kres}).
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see \citep{weindl_livestock_2017}), taking available AGR biomass as well as livestock productivity into account.
+We estimate, for low-income countries, a material-use share for straw residues of 5\% and a fuel-share of 10\% for all used residues groups in low-income countries.
+For high-income countries, no withdrawal for material or fuel use is assumed, and middle-income countries use shares are linearly interpolated based on per-capita income.
+The remaining AGR as well as all BGR are expected to be recycled to the soil.
+We limit high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
 
@@
 
-To transform dry matter estimates into carbon, we compiled crop-group
-and plant part specific carbon to dry matter (c:dm) ratios (see
-@ref(append:Tablec2dm)).
+To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \ref{append:Tablec2dm}).
 
 \hypertarget{sec:livstmanure}{%
-\subsubsection{Livestock Distribution and Manure
-Excretion}\label{sec:livstmanure}}
+\subsubsection{Livestock Distribution and Manure Excretion}\label{sec:livstmanure}}
 
-We assume that manure is applied at its excretion place, leaving the
-livestock distribution the driving factor for the spatial pattern of
-manuring.
+We assume that manure is applied at its excretion place, leaving the livestock distribution the driving factor for the spatial pattern of manuring.
 
-\textbf{Livestock Distribution} To disaggregate country level FAOSTAT
-livestock production values to the half-degree scale, we use the
-following rule-based assumptions, drawing from the approach of
-\citep{robinson_mapping_2014}, and uses feed basket assumptions based on
-a revised methodology from \citep{weindl}. We differentiate between
-ruminant and monogastric systems, as well as extensive and intensive
-systems. For ruminants, we assume that livestock is located where the
-production of feed occurs. We distingush between grazed pasture, which
-is converted into livestock products from extensive systems; and all
-other (crop-based) feeds, which we consider to be consumend in intensive
-systems. For poultry, egg and monogastric meat production we use the
-per-capita income of the country to divide between intensive and
-extensive production systems. For low-income countries, we assume
-extensive production systems. We locate them according to the share of
-built-up areas based on the idea that these animals are held in
-subsistence or small-holder farming systems with a high labour per
-animal ratio. Intensive production associated with higher income
-countries, is distributed within a country using again share of
-crop-based feed production, assuming that feed availability is the most
-driving factor for livestock location.
+\textbf{Livestock Distribution}
+To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014}, and uses feed basket assumptions based on a revised methodology from \citep{weindl_livestock_2017}.
+We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
+For ruminants, we assume that livestock is located where the production of feed occurs.
+We distinguish between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumed in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
+For low-income countries, we assume extensive production systems.
+We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labor per animal ratio.
+Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production, assuming that feed availability is the most driving factor for livestock location.
 
-\textbf{Manure Excretion, Storage and Recycling} Manure production and
-management is based on a revised methodology of
-\citep{bodirsky_n2o_2012} and is presented here due to its central role
-in soil carbon modelling. Based on the gridded livestock distribution we
-calculate excretions by estimating the nitrogen balance of the livestock
-system on the basis of comprehensive livestock feed baskets
-\citep{weindl}, assuming that all nitrogen in protein feed intake, minus
-the nitrogen in the slaughter mass, is excreted. Carbon in excreted
-manure is estimated by applying fixed C:N ratios (given by
-\citep{ipcc_2019_2019}). Depending on the feed system we assume manure
-to be handled in four different ways: All manure orginated from pasture
-feed intake is excreted directly to pastures and rangelands (pasture
-grazing), deducting manure collected as fuel. Manure fuel shares are
-estimated using IPCC default values \citep{ippc_2006_2006}. Whereas for
-low-income countries, we adopt a share of 25\% of crop residues in feed
-intake directly consumend and excreted on crop fields (stubble grazing),
-we do not consider any stubble grazing in high-income countries;
-middle-income countries see linearly interpolated shares depending on
-their per-capita income. For all other feed items we assume the manure
-to be stored in animal waste management systems associated with
-livestock housing. To estimate the carbon actually recycled to the soil,
-we account for carbon losses during storage and recycling shares in
-different animal waste management and grazing systems. Whereas we assume
-no losses for pasture and stubble grazing, we consider that the manure
-collected as fuel is not recycled. For manure stored in different animal
-waste management systems we compiled carbon loss rates partly depending
-on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}
-(see @ref(append:TableclossAWMS)). We limit high application shares at
-\(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
+\textbf{Manure Excretion, Storage and Recycling}
+Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modelling.
+Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep{ipcc_2019_2019}).
+Depending on the feed system we assume manure to be handled in four different ways:
+All manure originated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
+Manure fuel shares are estimated using IPCC default values \citep{ipcc_2006_2006}.
+Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
+For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing.
+To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
+Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
+For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see \ref{append:TableclossAWMS}).
+We limit high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
 
@@
 
-We use irrigation area shares to modify the water effect on
-decomposition by weighting the irrigated and rainfed water factors based
-on these shares. The LUH2v2 data set provides irrigated fractions for
-their cropland subcategories. We apply aggregated area shares, leading
-to the same impact of irrigation on all of our crop groups. Moreover we
-assume the irrigation effect to be present for all 12 months of a year
-in grid cells, that has been marked as irrigated.
+We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
+The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for their cropland subcategories.
+We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
+Moreover, we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
 
@@
 
-To account for the distribution of tillage to the three different
-tillage types specfied by the IPCC - full tillage, reduced tillage and
-no tillage -, we assume that all natural land and pastures are not
-tilled, whereas as default annual crops are under full and perennials
-under reduced tillage. Furthermore we assume no tillage in cropland
-cells specified as no tillage cell based on the historic global gridded
-tillage data set from Prowollik \citep{porwollik_generating_2018}.
-Porwollik data set is extended by to the period of 1974-2010 by
-combining country-level data on conservation agriculture (ca) area
-values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the
-methodolgy of Porwollik to identify potential no tillage cells.
-
+To account for the distribution of tillage to the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas as default annual crops are under full and perennials under reduced tillage.
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from \citep{porwollik_generating_2018}.
+This data set is extended by to the period of 1974-2010 by combining country-level data on conservation agriculture area values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the methodology of \citep{porwollik_generating_2018} to identify potential no tillage cells.
 \newpage
 
-\hypertarget{sec:tier1}{%
-\subsection{SOC Stocks and Stock changes following Tier
-1}\label{sec:tier1}}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-Additionally to the tier 2 approach of the refinement of IPCC guidelines
-vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management
-data coming with it, SOC changes can be estimated using the IPCC tier 1
-approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006},
-\citet{ipcc_2019_2019}). Here, stocks are calculated via stock change
-factors (F\^{}\{SCF\}) given by the IPCC for the topsoil (0-30 cm) and
-based on observational data. \(F^{SCF}\) are differentiated by different
-crop, management and input systems (here summarized under \(m\))
-reflecting different dynamics under changed in- and outflows without
-explicitly tracking these. Moreover \(F^{SCF}\) vary for different
-climatic zones (\(c\)) specified by the IPCC (see
-@ref(append:climatemap)). The actual SOC stocks as thus calulated based
-on a given reference SOC stock by
+We present simulation results of our SOC budget focusing on cropland areas for the year 2010 as well as global trends of SOC stock changes for the period from 1975 to 2010.
 
-\begin{equation}
-SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
-(\#eq:tier1)
-\end{equation}
-
-with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into
-corresponding climate zones \(c\). For this analysis we use the default
-\(F^{SCF}\) from the Tier 1 method of \citep{ipcc_2006_2006}, and
-\citep{ipcc_2019_2019} as a comparison and constistency check for our
-more detailed budget approach. \newpage
-
-\section{Results}
-
-We present simulation results of our SOC budget focusing on cropland
-areas for the year 2010 as well as global trends of SOC stock changes
-for the period from 1975 to 2010.
-
-\subsection{SOC distribution and depletion}
+\hypertarget{soc-distribution-and-depletion}{%
+\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}
 
@@
 
-In fig.~1(a) we provide the first world map of SOC on croplands
-considering real world management data on the global scale. Our
-spatially explicit results moreover show hot spots of SOC losses as well
-as gains in two different ways: 1. Absolute SOC changes (see fig.~1(c))
-indicate areas with high importance for the global SOC emissions. The
-might be driven by huge relative losses or a high natural stock, from
-which even small deviations could lead to substantial losses. 2. To
-attribute SOC losses to insufficient agricultural management relative
-SOC changes (\(F^{SCF}\), see fig.~1(d)) are a helpful tool. They
-indicate areas with huge difference in carbon inflows or SOC decay
-compared to natural vegetation, that might be overcome due to improved
-agricultural pratices.
+In fig.~1(a) we provide the first world map of SOC on croplands considering real world management data on the global scale.
+Our spatially explicit results moreover show hot spots of SOC losses as well as gains in two different ways:
+1.
+Absolute SOC changes (see fig.~1(c)) indicate areas with high importance for the global SOC emissions.
+The might be driven by huge relative losses or a high natural stock, from which even small deviations could lead to substantial losses.
+2.
+To attribute SOC losses to insufficient agricultural management relative SOC changes (\(F^{SCF}\), see fig.~1(d)) are a helpful tool.
+They indicate areas with huge difference in carbon inflows or SOC decay compared to natural vegetation, that might be overcome due to improved agricultural practices.
 
-\subsection{Agricultural management effects on SOC emissions and
-cycling}
+\hypertarget{agricultural-management-effects-on-soc-emissions-and-cycling}{%
+\subsection{Agricultural management effects on SOC emissions and cycling}\label{agricultural-management-effects-on-soc-emissions-and-cycling}}
 
-Global cummulative SOC emissions are decreasing (see fig.~1(c)). Fig. 2
-reveals the relative impact of management effects by freezing tillage
-areas as well as carbon inflows from residues or manure at the level of
-1975. Our counterfactual scenarios show that the increasing residue
-carbon input had the biggest overall effect on SOC stocks. Without
-changes in management regimes especially in residue inflows to the soil,
-global cummulative SOC emissions would still grow. The strong effect of
-carbon residue amounts are also visible in the carbon flow diagram for
-agricultural production for the year 2010 (see fig.~3)
+Global cumulative SOC emissions are decreasing (see fig.~1(c)).
+Fig.
+2 reveals the relative impact of management effects by freezing tillage areas as well as carbon inflows from residues or manure at the level of 1975.
+Our counterfactual scenarios show that the increasing residue carbon input had the biggest overall effect on SOC stocks.
+Without changes in management regimes especially in residue inflows to the soil, global cumulative SOC emissions would still grow.
+The strong effect of carbon residue amounts are also visible in the carbon flow diagram for agricultural production for the year 2010 (see fig.~3)
 
@@
 \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/OuFlowFig} \caption{Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
-Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil).}\label{fig:FlowFig}
+Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposed before official counting as soil).}\label{fig:FlowFig}
 \end{figure}
@@
 
-\section{Discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-\subsection{Including agricultural management data changes the sign of
-the trend}
+\hypertarget{including-agricultural-management-data-changes-the-sign-of-the-trend}{%
+\subsection{Including agricultural management data changes the sign of the trend}\label{including-agricultural-management-data-changes-the-sign-of-the-trend}}
 
-This study provides an analysis on historic SOC stock changes on
-cropland. We determine the SOC trends on cropland compared to a
-potential natural vegeted world under the same historic climatic
-conditions (\(SOC_{natveg}\)) leading to a cummulative SOC emission
-since dawn of mankind of around 37 GtC in 2010. Whereas recent modelling
-estimates on global SOC emissions indicate an on going increase of SOC
-emissions (Sanderman et al, Pugh et al.), our study indicates that the
-global SOC gap might be closing slowly.
+This study provides an analysis of historic SOC stock changes on cropland.
+We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions (\(SOC_{natveg}\)).
+Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
+Whereas recent modelling estimates of global SOC emissions indicate an ongoing increase of SOC emissions (\citep{pugh_simulated_2015}, \citep{sanderman_soil_2017}), our study indicates that the global SOC gap is slowly closing due to improved management.
 
-According to Sanderman et al.~cummulative SOC emissions since beginning
-of human cropping activities have been at around 37 GtC for the first 30
-cm of the soil with half of it attributed to grazing. It was also
-pointed out, that this results might be conservative estimates and low
-compared to experimental results, leaving our estimate of 37 GtC in 2010
-for cropland emissions only in alignment with Sanderman et al.~estimates
-considering the hugh uncertainties in modelling SOC on the global scale.
-Furtherore the results of Sanderman et al.~calculated historical trends
-based on agricultural land expansion without considering SOC variations
-due to different management systems at all. In Pugh et al.~management
-effects like tillage and residue recycling have been considered, but
-neither changes over time nor alignment to oberserved historical data
-like yields levels or no tillage areas were taken into account. Under
-this assumptions this study only found marginal effect of crop
-productivity and other management effects on cummulative SOC emissions.
+According to \citep{sanderman_soil_2017} cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+It was also pointed out that these results might be conservatively low compared to experimental results.
+Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with \citep{sanderman_soil_2017} estimates.
 
-Nevertheless our results show, when purely focusing on the past decades,
-the moderate global cropland expansion of around 11\% between 1974 and
-2010, is outweighted by improved agricultural yields ad practises.
-Moreover our sensitivity analysis indicades that 1. yield increases and
-with that the increase in residue biomass might play the most important
-role, followed by 2. enhanced residue recycling rates, 3. improved
-manure recycling (e.g.~due to improved animal waste management system)
-and 4. the adoption of no tillage practises.
+Furthermore, the results of \citep{sanderman_soil_2017} modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+\citep{pugh_simulated_2015} considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
+The study moreover concludes that productivity gains for crops while increasing yield levels by 18\%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1\% change in SOC emissions.
 
-Modelling management effects on the global scale comes however with huge
-parametic and systematic uncertainties. One aspect pointed out by Keel
-(2017) as well as by Smith (2019) might be that carbon input
-calculations are highly sensitive to the choice of allometric functions
-determining below and above ground residue estimates from harvested
-quantities. Keel et al.~questioned, that below ground residues might
-increase with a fixed root:shot ratio rather than being independend from
-productivity gains. Following this argumentation SOC results shown in
-this study might especially in high-yielding farming system overestimate
-actual SOC stocks. However according to our study above ground residue
-biomass recycling seem to contribute even more to overall SOC stocks due
-to higher input rates.
+Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
+We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
+This is challenging \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
+Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g.~due to improved animal waste management system) and (3) the adoption of no tillage practices.
 
-\subsection{Modeled management effect inline with default IPCC
-assumptions}
+Modelling management effects on the global scale comes however with parametric and structural uncertainties.
+As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
+\citep{keel_large_2017} questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources.
+Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding.
+However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.
 
-To validated the effect of our modelled SOC stocks and stock changes
-under management, we compare our results to default IPCC stock changes
-factors (cite), which are based on measurement data for croplands (see
-@ref(tab:SCFtable)).
+\hypertarget{modeled-management-effect-in-line-with-default-ipcc-assumptions}{%
+\subsection{Modeled management effect in line with default IPCC assumptions}\label{modeled-management-effect-in-line-with-default-ipcc-assumptions}}
+
+To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}) which are based on measurement data for croplands (see \ref{tab:SCFtable}).
+To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.
+
+\begin{figure}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+We compare average values for four IPCC climate zone classifications.
+IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget).}\label{fig:SCFtable}
+\end{figure}
+
+Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006.
+For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
+The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7\% of it.
+
+Looking again on the time trend our study shows the substantial impact of changed management on the development of \(F_{SCF}\).
+
+\hypertarget{soc-stocks-inline-with-literature}{%
+\subsection{SOC stocks inline with literature}\label{soc-stocks-inline-with-literature}}
+
+The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \ref{tab:SOCtable}).
 
 \begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows IPCC 2006 as well as 2019 default (medium input) and low input regimes stock change factors $F_{SCF}$ without other subsystem consideration compared results of this study (SOC budget) for 1990 and 2010.}\label{fig:SCFtable}
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.}\label{fig:SOCtable}
 \end{figure}
 
-Our climate zone specific, aggregated \(F_{SCF}\) correspond very well
-to the tier 1 estimation of IPCC, 2006 of \(F_{SOC}\). For the tropical
-regions the assumptions changed notablly from the guidelines in 2006 to
-the update in 2019, leaving our results too low in comparison with IPCC,
-2019. Considering yield gaps in mainly developing regions in the tropics
-the default assumption of medium input systems, might be an
-overestimation of actual SOC state. The additional effect of considering
-low inpout regimes in tropical regions can however not explain the full
-mismatch to IPCC 2019 values but account for at least 5-7\% of it.
+The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking on regional results \ref{append:regcompare}, our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils.
+To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
+Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
 
-\subsection{SOC stocks inline with literature}
+Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
+For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
+This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain.
+The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
+Especially in less forested areas the natural land representation might be off, due to parametrization assumptions of the natural litterfall.
 
-The worlds SOC stock and its changes are highly uncertain (cite), which
-is visible by the wide range of global SOC stock estimates (see
-@ref(tab:SOCtable)).
+Nevertheless, total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
+Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions.
+Then former natural land SOC stocks are declining to lower levels.
+However, as pointed out before, due to low cropland expansion rates, their SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
+We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
+(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
 
-\begin{figure}[H]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/TableSOC_comparison} \caption{Modelled as well as data based estimation for global SOC stock (for 2010) in GtC for the first 30 cm of soil.}\label{fig:SOCtable}
-\end{figure}
+\hypertarget{important-shortcomings-might-be-added-as-well}{%
+\subsection{Important shortcomings --\textgreater{} might be added as well}\label{important-shortcomings-might-be-added-as-well}}
 
-The global esimates of SOC stock by this study are on the lower edge
-compared to other modelled results or more data driven estimates.
-Looking on regional results (fig.~SX in supplement), our estimates turn
-out to be in good agreement for most regions with the largest deviations
-for boreal areas. Considering that the model was parametrized for
-croplands, these mismatches are not superising, since the temperature
-effects on decomposition is most likly fundamentally different for
-permaforst soils than assumend in this study. To avoid that this bias
-influences our results, our study focusses exclusively on cropland
-soils, excluding most of the boreal zone. Morover, when focussing on SOC
-changes, pristine natural vegetated areas without human land management
-under the same historic climatic conditions cancel out in the
-calculation of SOC emissions.
+The \textbf{initialization of SOC stocks} in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start.
+That is inline with IPCC guidelines calling for a run-in period of 5-20 years.
+However, the results are heavily influenced by it as \ref{append:initcompare} suggested.
+Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore.
+For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010.
+For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010.
+The relative declining trend for the SOC gap between 1975 and 2010 (-12\%) is however almost the same as for our historic estimation (-14\%).
+According to LUH \citep{hurtt_harmonization_2020} cropland area has increased over 62\% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries.
+Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.
 
-However the natural land representation that is dominating the total SOC
-stocks of the world and with it also the SOC initialization is lacking
-proper parametrization of nitrogen and lignin content of litterfall.
-This leaves carbon inputs and decay behaviour for natural land rather
-uncertained. As all results are valued against these potential natural
-SOC the absolute values and also the emissions over time have to be used
-with caution. Especially in less forested areas the natural land
-representation might be off, due to parameterization assumptions of the
-natural litterfall. Nevertheless total SOC stocks are in a reasonable
-range and the stock change factors \(F^{SCF}\) are in good agreement
-with the Tier 1 default values of 2006. In addition we preformed a
-sensitivity analysis for lignin and nitrogen parameterization of natural
-litterfall (see appendix). It shows that the general trend of decreasing
-SOC emissions is not altered even by rather unlikely parameter choices,
-that would overestimate natural SOC stocks strongly and therefor lead to
-high emission in the case of land conversion.
+Following the IPCC guidelines this study has limited its focus to the \textbf{first 30 cm of the soil profile}.
+In this regard several aspects are over simplified within our approach.
+Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
+This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
+Consequently, our SOC loss estimates are too high.
+Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
+Following this arguments SOC stocks in croplands might even be underestimated.
 
-\subsection{Important short commings --\textgreater{} move to appendix}
-
-Smaller points and shortcommings:
-
--This study as the IPCC guidelines suggested has limited her focus to
-the first 30 cm of the soil profile, leaving changes in the subsoil
-unnoticed. Nevertheless studies (see Don on tillage) have shown, the
-subsoil to be a game changer in evaluating total SOC losses or gains for
-no-tillage systems. It has been argued that for intensivly tilled soils,
-subsoil SOC is increasing due to the import of carbon rich topsoil to
-deeper soil layers. Following these argumentation SOC stocks in
-croplands might even be underestimated.
-
-\begin{itemize}
-\item
-  Fertilizer interaction is not included here by accounting for
-  additional N supply that would alter C:N ratio of the carbon inputs.
-  Tier 2 steady-state method is neglecting fertilizer application,
-  however we would have fertilizer ammounts at hand to include them, if
-  proper representation of fertilizer within the method would be
-  possible to add.
-\item
-  Pasture dynamics are neglected and treated as natural vegetation,
-  which might be -- looking on pasture degradation due to overgrazing --
-  oversimplified for some spots, but is inline with assumption on
-  pasture SOC stocks done before (see Tier 1 IPCC). (Note that also
-  manure excreted to pastures is neglected within these analysis, since
-  we focus purely on cropland dynamics.)
-\item
-  No tillage adaptaion is neglected on cropland due to less common
-  adoption of no tillage and conservation agriculture. Pastures are
-  assumed to be not tilled at all (propably only heavy managed pastures
-  are tilled with some rotation)
-\item
-  Irrigated areas are not crop specific and irrigation is not restricted
-  to growing periods (since it is very complex to calculate average
-  growing periods). Crop specific growing periods might be possible
-  using LPJmL data.
-\item
-  flooded rice area are not represented correctly as parameterization
-  does not hold true for flooded conditions.
-\item
-  Carbon displacement via leaching and erosion is neglected in this
-  study.
-\item
-  Non-net/Gross land use transitions are not tracked in this study.
-\item
-  Within cropland we do not track croparea transitions, but rather look
-  at statistical distributions of the crop functional types. Due to crop
-  rotations and missing data on crop specific distributions, these
-  transitions would be any way rather uncertain.
-\item
-  The disaggregation of manure to uild-up areas (in the case of
-  extensive monograstrics) is leeding to a lot of displaced manure (?)
-  that is cut off
-\item
-  It is known that there are mismatches between FAO statistics and LUH
-  areas. As far as possibles there were harmonized within this study.
-\item
-  ``The Tier 2 method does not simulate C change but simply calculates
-  an annual C stock change from the current C stock to the future
-  steady-state soil C stock calculated based on current conditions.''
-  Leading to the fact that our total stock results are highly
-  uncertained. \newpage
-\end{itemize}
+Carbon displacement via \textbf{leaching and erosion} is neglected in this study.
+As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain.
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+\newpage
 
@@
 
-Outlook including perspective on mitigation (and soc enhancement in the
-future). \newpage
+Outlook including perspective on mitigation (and soc enhancement in the future).
+\newpage
 
-\section{Appendix}
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
 
 \hypertarget{append:subsection2mrfunctions}{%
-\subsection{table on method subsections to functions within R
-packages}\label{append:subsection2mrfunctions}}
+\subsection{table on method subsections to functions within R packages}\label{append:subsection2mrfunctions}}
 
 \hypertarget{append:Tableluh2fao2mag}{%
-\subsection{table on mapping
-LUH2FAO2MAG}\label{append:Tableluh2fao2mag}}
+\subsection{table on mapping LUH2FAO2MAG}\label{append:Tableluh2fao2mag}}
 
@@
 
-Litter is coming from LPJmL in carbon units - transformation with 0.44
-is done twice reverting the effect of the transformation
+Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
 
@@
 
+\hypertarget{append:regcompare}{%
+\subsection{regional SOC stock in GtC from different sources}\label{append:regcompare}}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{../ResultNotebooks/Output/Images/reg_comparisonfigure} \caption{Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.}\label{fig:regSOCtable}
+\end{figure}
+
+\hypertarget{append:initcompare}{%
+\subsection{global SOC gap in GtC for different initialization choices}\label{append:initcompare}}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{../ResultNotebooks/Output/Images/scenario_lu} \caption{Global SOC gap in GtC for the first 30 cm of soil for different initialization choices.}\label{fig:initSOC}
+\end{figure}
+
+\hypertarget{more-short-comings}{%
+\subsection{more short comings}\label{more-short-comings}}
+
+\begin{itemize}
+\item
+  Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
+Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer amounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
+\item
+  Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
+(Note that also manure excreted to as well as tillage on pastures are neglected within this analysis, since we focus purely on cropland dynamics.)
+\item
+  Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
+Crop specific growing periods might be possible using LPJmL data.
+\item
+  Flooded rice area are not represented correctly as parametrization does not hold true for flooded conditions.
+\item
+  Non-net/Gross land use transitions are not tracked in this study.
+\item
+  Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
+Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
+\item
+  The disaggregation of manure to build-up areas (in the case of extensive monogastrics) is leading to a lot of displaced manure (?) that is cut off
+\item
+  It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
   \newpage
+\end{itemize}
 
-\codedataavailability{Software code for paper and result prepartion can
-be found under www.github.com/k4rst3ns/. Data used for the output can be
-found under
-.} %% use this section when having data sets and software code available
+\codedataavailability{Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/.
+Data used for the output can be found under .} %% use this section when having data sets and software code available
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of
-Bodirsky (and ). Bodirsky, and Popp revised paper.} %% optional section
+\authorcontribution{Karstens wrote code and paper build on work of Bodirsky (and ).
+Bodirsky, and Popp revised paper.} %% optional section
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing
-interests.} %% this section is mandatory even if you declare that no competing interests are present
+\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
 
@@
 \bibliographystyle{copernicus}
-\bibliography{SOCbudget.bib}
+\bibliography{SOCbudget2.bib}
 %%

### 2020-11-09 SOCbudgetPaper/appendix.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

@@
 ## map on climate zone used for SCF {#append:climatemap}
+
+## regional SOC stock in GtC from different sources {#append:regcompare}
+
+```{r captionregSOCtable, echo=FALSE}
+captionregSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
+```
+
+```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
+```
+
+## global SOC gap in GtC for different initialization choices {#append:initcompare}
+
+```{r captioninitSOC, echo=FALSE}
+captioninitSOC <- "Global SOC gap in GtC for the first 30 cm of soil for different initialization choices."
+```
+
+```{r initSOC, out.width = "100%", fig.cap = captioninitSOC, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_lu.png")
+```
+
+## more short comings
+
+- Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
+Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer amounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
+
+- Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
+(Note that also manure excreted to as well as tillage on pastures are neglected within this analysis, since we focus purely on cropland dynamics.)
+
+- Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
+Crop specific growing periods might be possible using LPJmL data.
+
+- Flooded rice area are not represented correctly as parametrization does not hold true for flooded conditions.
+
+- Non-net/Gross land use transitions are not tracked in this study.
+
+- Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
+Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
+
+- The disaggregation of manure to build-up areas (in the case of extensive monogastrics) is leading to a lot of displaced manure (?) that is cut off
+
+- It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.

### 2020-11-09 SOCbudgetPaper/discussion.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

--- 2020-11-04 SOCbudgetPaper/discussion.Rmd
@@
 We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions ($SOC_{natveg}$).
-Our results show, that human activity lead to cummulative SOC emission of around 37 GtC in 2010.
-Whereas recent modelling estimates of global SOC emissions indicate an on going increase of SOC emissions (Sanderman et al, Pugh et al.), our study indicates that the global SOC gap is slowly closing due to improved management.
+Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
+Whereas recent modelling estimates of global SOC emissions indicate an ongoing increase of SOC emissions ([@pugh_simulated_2015], [@sanderman_soil_2017]), our study indicates that the global SOC gap is slowly closing due to improved management.
 
-According to Sanderman et al.
-cummulative SOC emissions since beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
 It was also pointed out that these results might be conservatively low compared to experimental results.
-Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with Sanderman et al.
-estimates.
+Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with [@sanderman_soil_2017] estimates.
 
-Furthermore the results of Sanderman et al.
-modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
-Pugh et al.
-considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to oberserved historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substainal decrease in SOC emissions, but only account for less than 1% change in SOC emissions.
+Furthermore, the results of [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+[@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
+The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1% change in SOC emissions.
 
-Our study for the first time uses a dynanmic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighted by improved agricultural yields and practises.
-This is challenging Pugh et al.
-findings of only small effects due to improved pratices.
-Moreover our sensitivity analysis indicades that (1) recycled residue biomass increases driven by higher yields as well as higher higher recycling rates, (2) improved manure recycling (e.g.
-due to improved animal waste management system) and (3) the adoption of no tillage practises.
+Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
+We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
+This is challenging [@pugh_simulated_2015] findings of only small effects due to improved practices.
+Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g.
+due to improved animal waste management system) and (3) the adoption of no tillage practices.
 
 Modelling management effects on the global scale comes however with parametric and structural uncertainties.
-As pointed out by Keel (2017) and Smith (2019), carbon input calculations are highly sensitive to the choice of allometric functions
-determining below and above ground residue estimates from harvested quantities.
-Keel et al.~questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independend from productivity gains.
-Moreover the study pointed out that plant breeding shift allometries and might been not reflected in outdated data sources.
+As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
+[@keel_large_2017] questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources.
 Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding.
@@
 
-To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (cite) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
+To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
 To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.
@@
 ```{r captionSCFtable, echo=FALSE}
-captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+captionSCFtable <- "This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
 We compare average values for four IPCC climate zone classifications.
-IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget).}"
+IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget)."
 ```
@@
 Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
+For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
 Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
@@
 
-Looking again on the time trend our study shows the substantial impact of changed management to the development of $F_{SCF}$.
+Looking again on the time trend our study shows the substantial impact of changed management on the development of $F_{SCF}$.
 
@@
 
-The worlds SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
 
@@
 captionSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
 In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
@@
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/TableSOC_comparison.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
 ```
 
-The global esimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
-Looking on regional results (fig.
-SX in supplement), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
-Considering that the model was parametrized for croplands, these mismatches are not superising since the temperature effects on decomposition are fundamentally different for permaforst soils.
-To avoid that this bias influences our results, our study focusses exclusively on cropland soils, excluding most of the boreal zone.
-Morover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking on regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils.
+To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
+Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
 
@@
 For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
-This leaves carbon inputs and decay behaviour for natural land and pastures rather uncertain.
+This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain.
 The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
-Especially in less forested areas the natural land representation might be off, due to parameterization assumptions of the natural litterfall.
+Especially in less forested areas the natural land representation might be off, due to parametrization assumptions of the natural litterfall.
 
-Nevertheless total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetaded areas at all.
+Nevertheless, total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
 Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions.
-Than former natural land SOC stocks are declining to lower levels.
-However as pointed out before, due to low cropland expansion rates, there SOC declines are outweighted by the general trend of increasing SOC stocks on existing cropland due to management improvements.
-We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parameterization of natural litterfall, which tend to change the global SOC stock substantially for different choices
+Then former natural land SOC stocks are declining to lower levels.
+However, as pointed out before, due to low cropland expansion rates, their SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
+We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
 (see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
 
-## Important short commings --> move to appendix
+## Important shortcomings --> might be added as well
 
-Smaller points and shortcommings:
+The **initialization of SOC stocks** in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start.
+That is inline with IPCC guidelines calling for a run-in period of 5-20 years.
+However, the results are heavily influenced by it as \@ref(append:initcompare) suggested.
+Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore.
+For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010.
+For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010.
+The relative declining trend for the SOC gap between 1975 and 2010 (-12%) is however almost the same as for our historic estimation (-14%).
+According to LUH [@hurtt_harmonization_2020] cropland area has increased over 62% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries.
+Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.
 
--This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed.
-Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-It has been argued that for intensivly tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
-Following these argumentation SOC stocks in croplands might even be underestimated.
+Following the IPCC guidelines this study has limited its focus to the **first 30 cm of the soil profile**.
+In this regard several aspects are over simplified within our approach.
+Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
+This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
+Consequently, our SOC loss estimates are too high.
+Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
+Following this arguments SOC stocks in croplands might even be underestimated.
 
-- Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
-Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer ammounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
-
-- Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
-(Note that also manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.)
-
-- No tillage adaptaion is neglected on cropland due to less common adoption of no tillage and conservation agriculture.
-Pastures are assumed to be not tilled at all (propably only heavy managed pastures are tilled with some rotation)
-
-- Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
-Crop specific growing periods might be possible using LPJmL data.
-
-- flooded rice area are not represented correctly as parameterization does not hold true for flooded conditions.
-
-- Carbon displacement via leaching and erosion is neglected in this study.
-
-- Non-net/Gross land use transitions are not tracked in this study.
-
-- Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
-Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
-
-- The disaggregation of manure to uild-up areas (in the case of extensive monograstrics) is leeding to a lot of displaced manure (?) that is cut off
-
-- It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
-
-- "The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
-current C stock to the future steady-state soil C stock calculated based on current conditions." Leading to the fact that our total stock results are highly uncertained.
+Carbon displacement via **leaching and erosion** is neglected in this study.
+As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain.
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.

### 2020-11-09 SOCbudgetPaper/introduction.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

@@
 Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite).
+Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools over much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite).
 Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC  (Lal, 2006).
@@
 However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
-Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
+Our study combines a spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
 This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
@@
 This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
-We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
-Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+Additionally, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
+Lastly, we continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
 In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
-Finally, we discuss and compare our findings – including their implications for SOC model development – and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+Finally, we discuss and compare our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.

### 2020-11-09 SOCbudgetPaper/methods.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

--- 2020-08-14 SOCbudgetPaper/methods.Rmd
@@
 We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
-In the following chapter we outine the most important relationships and assumptions.
+In the following section we outline the most important relationships and assumptions.
 Table \@ref(append:subsection2mrfunctions) provides further information on corresponding code within the R packages.
@@
 Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019; short Tier 2 steady-state method) , we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
+4 (@ipcc_2019_2019; short Tier 2 steady-state method), we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
 We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
-Therefore we conduct the following three steps within each yearly timestep:
+Therefore, we conduct the following three steps within each yearly time-step:
 (1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks, 
-(2) we account for land conversion by transferring SOC from and to an other land-use type representad as natural vegetation,
+(2) we account for land conversion by transferring SOC from and to another land-use type represented as natural vegetation,
 (3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
@@
 \end{equation}
-with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cell indices and $t$ years.
-We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them.
+with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cell identifier and $t$ years.
+We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them, following the approach in the Century model (Parton et al, 1988).
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
@@
 We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \@ref(tab:datasourceinputs)).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livstmanure)); for natural vegetation litterfall [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
+Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livstmanure)); for natural vegetation litterfall (including fine root turnover) [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
 
@@
 For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that both the amount of carbon as well as its structural composition determine the effective inflow.
+This implies that both the amount of carbon and its structural composition determine the effective inflow.
 Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \@ref(tab:datasourceinputs).
@@
 Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
-Thus $k_{sub}$ is given by
+Thus, $k_{sub}$ is given by
 
@@
 
-For cropland we distinguish the effect of different tillage (see \@ref(sec:tillage)) and irrigation (see \@ref(sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
+For cropland, we distinguish the effect of different tillage (see \@ref(sec:tillage)) and irrigation (see \@ref(sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
 Data sources as well as drivers considered for each land-use types are shown in table \@ref(tab:datasourcedecay).
@@
 
-Reformulating this equation, we obtain a massbalance equation as follows
+Reformulating this equation, we obtain a mass balance equation as follows
 
@@
 
-The global SOC stock for each time step can than be calculated via
+The global SOC stock for each time step can then be calculated via
 
@@
 
-According to the IPCC guidelines SOC changes can be calculates as the difference of two (aufeinander folgenden) years (see Eq.
+According to the IPCC guidelines SOC changes can be expressed as the difference of two consecutive years (see Eq.
 5.0A in [@ipcc_2019_2019]).
-This however will also include naturally occured changes due to climatic variance over time.
-For our study we will define the absolute and relative SOC changes in relation to a potential natural state $SOC_{pnv}$ under the same climatic conditions at time $t$ in grid cell $i$, that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
+This however, will also include naturally occurring changes due to climatic variance over time.
+For our study we defined the absolute and relative SOC changes in relation to a potential natural state $SOC_{pnv}$ under the same climatic conditions at time $t$ in grid cell $i$, that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
 The absolute changes $\Delta SOC$ and relative changes $F^{SCF}$ a thus given by
@@
 
-Note that the absolution changes $\Delta SOC$ can be also interpreted as cummulative SOC emissions or as SOC depth of mankind; whereas relative changes $F^{SCF}$ can be denoted as stock change factors as defined within the IPCC guidelines of 2006 [@ipcc_2006_2006].
+Note that the absolution changes $\Delta SOC$ can be also interpreted as the SOC gap due to human cropping activities; whereas relative changes $F^{SCF}$ can be denoted as stock change factors as defined within the IPCC guidelines of 2006 [@ipcc_2006_2006].
+Note that the SOC gap is equivalent to the negated cumulative SOC emission ([@pugh_simulated_2015]) also referred to as SOC debt ([@sanderman_soil_2017]) of human land use.
 
-### Initialisation of SOC pools
+### Initialization of SOC pools
 
 To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions will decrease.
+We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions decreases.
+
+## SOC Stocks and Stock changes following Tier 1 {#sec:tier1}
+
+Additionally, to the tier 2 approach of the refinement of IPCC guidelines vol.
+4 [@ipcc_2019_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.
+4 (@ipcc_2006_2006, @ipcc_2019_2019).
+Here, stocks are calculated via stock change factors ($F^{SCF}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+$F^{SCF}$ are differentiated by different crop, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover, $F^{SCF}$ vary for different climatic zones ($c$) specified by the IPCC (see \@ref(append:climatemap)).
+The actual SOC stocks as thus calculated based on a given reference SOC stock by
+
+\begin{equation}
+SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
+(\#eq:tier1)
+\end{equation}
+
+with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
+For this analysis we use the default $F^{SCF}$ from the Tier 1 method of [@ipcc_2006_2006], and [@ipcc_2019_2019] as a comparison and consistency check for our more detailed budget approach.
 
@@
 
-### Landuse and Landuse Change {#sec:landuse}
+### Land use and Land-use change {#sec:landuse}
 
-Land-use patterns are based on the Land-Use Harmonization 2 [@Hurtt2020] data set, which we aggregate from quarter degree to half degree resolution.
+Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we aggregate from quarter degree to half degree resolution.
 We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see \@ref(append:Tableluh2fao2mag) for more  details on the crop group mapping).
@@
 Crop residue production and management is based on a revised methodology of [@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modelling.
-Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest indices ($HI$) and root:shoot ratios ($RS$) as follows
+Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest index values ($HI$) and root:shoot ratios ($RS$) as follows
 
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependend fraction [@ipcc_2006_2006].
+Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction [@ipcc_2006_2006].
 Note that deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
-We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals [@weindl], fuel or for material use.
+We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals [@weindl_livestock_2017], fuel or for material use.
 
@@
 A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
-Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following [@smil_nitrogen_1999]) we assume 25% burn share for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
 Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process [@ipcc_2006_2006].
@@
 We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Tablekcr2kres)).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl]), taking available AGR biomass as well as livestock productivity into account.
+Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl_livestock_2017]), taking available AGR biomass as well as livestock productivity into account.
 We estimate, for low-income countries,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumend, and middle-income countries use shares are linearly interpolated based on per-capita income.
-The remaining AGR as well as all BGR are assumend to be recycled to the soil.
+For high-income countries, no withdrawal for material or fuel use is assumed, and middle-income countries use shares are linearly interpolated based on per-capita income.
+The remaining AGR as well as all BGR are expected to be recycled to the soil.
 We limit high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
@@
 **Livestock Distribution**
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of [@robinson_mapping_2014], and uses feed basket assumptions based on a revised methodology from [@weindl].
+To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of [@robinson_mapping_2014], and uses feed basket assumptions based on a revised methodology from [@weindl_livestock_2017].
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
 For ruminants, we assume that livestock is located where the production of feed occurs.
-We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumend in intensive systems.
+We distinguish between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumed in intensive systems.
 For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
 For low-income countries, we assume extensive production systems.
-We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labour per animal ratio.
+We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labor per animal ratio.
 Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.
@@
 Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modelling.
-Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets [@weindl], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
 Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [@ipcc_2019_2019]).
 Depending on the feed system we assume manure to be handled in four different ways:
-All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Manure fuel shares are estimated using IPCC default values [@ippc_2006_2006].
-Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
+All manure originated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
+Manure fuel shares are estimated using IPCC default values [@ipcc_2006_2006].
+Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
 For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
@@
 We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
-The LUH2v2 data set provides irrigated fractions for their cropland subcategories.
+The LUH2v2 [@hurtt_harmonization_2020] data set provides irrigated fractions for their cropland subcategories.
 We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
-Moreover we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
+Moreover, we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
 
@@
 
-To account for the distribution of tillage of the three tillage types specfied by the IPCC - full tillage, reduced tillage and no tillage -, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage.
-This rule is based on the approach of [@porwollik_generating_2018], but neglects the (less common) adoption of no-tillage farming on agricultural land.
-
-\newpage
-
-## SOC Stocks and Stock changes following Tier 1 {#sec:tier1}
-
-Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.
-4 [@ipcc_2019_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.
-4 (@ipcc_2006_2006, @ipcc_2019_2019).
-Here, stocks are calculated via stock change factors (F^{SCF}) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-$F^{SCF}$ are differentiated by different crop, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover $F^{SCF}$ vary for different climatic zones ($c$) specified by the IPCC (see \@ref(append:climatemap)).
-The actual SOC stocks as thus calulated based on a given reference SOC stock by
-
-\begin{equation}
-SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
-(\#eq:tier1)
-\end{equation}
-
-with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis we use the default $F^{SCF}$ from the Tier 1 method of [@ipcc_2006_2006], and [@ipcc_2019_2019] as a comparison and constistency check for our more detailed budget approach.
+To account for the distribution of tillage to the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas as default annual crops are under full and perennials under reduced tillage.
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from  [@porwollik_generating_2018].
+This data set is extended by to the period of 1974-2010 by combining country-level data on conservation agriculture area values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the methodology of [@porwollik_generating_2018] to identify potential no tillage cells.

### 2020-11-09 SOCbudgetPaper/results.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

@@
 1(d)) are a helpful tool.
-They indicate areas with huge difference in carbon inflows or SOC decay compared to natural vegetation, that might be overcome due to improved agricultural pratices.
+They indicate areas with huge difference in carbon inflows or SOC decay compared to natural vegetation, that might be overcome due to improved agricultural practices.
 
@@
 
-Global cummulative SOC emissions are decreasing (see fig.
+Global cumulative SOC emissions are decreasing (see fig.
 1(c)).
@@
 Our counterfactual scenarios show that the increasing residue carbon input had the biggest overall effect on SOC stocks.
-Without changes in management regimes especially in residue inflows to the soil, global cummulative SOC emissions would still grow.
+Without changes in management regimes especially in residue inflows to the soil, global cumulative SOC emissions would still grow.
 The strong effect of carbon residue amounts are also visible in the carbon flow diagram for agricultural production for the year 2010 (see fig.
@@
 caption_FlowFig <- "Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
-Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil)."
+Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposed before official counting as soil)."
 ```

### 2020-11-09 SOCbudgetPaper/discussion.Rmd

@@
 According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
-It was also pointed out that these results might be conservatively low compared to experimental results.
-Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with [@sanderman_soil_2017] estimates.
+Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as [@sanderman_soil_2017] estimations.
+However, there are large uncertainties in modelling SOC at the global scale, and the [@sanderman_soil_2017] results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
 
-Furthermore, the results of [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Furthermore, [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
 [@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1% change in SOC emissions.
+The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decrease in SOC emissions, accounting for only less than 1% change in SOC emissions.
 
@@
 We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
-This is challenging [@pugh_simulated_2015] findings of only small effects due to improved practices.
-Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g.
-due to improved animal waste management system) and (3) the adoption of no tillage practices.
+This challenges [@pugh_simulated_2015] findings of only small effects due to improved practices.
+
+Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g.
+due to improved animal waste management systems) and (3) the adoption of no tillage practices.
 
@@
 As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
-[@keel_large_2017] questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources.
-Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding.
-However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.
+[@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources.
+Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
+However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass.
+Rather, it is more likely that we are still missing carbon inputs to the soil.
 
@@
 To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
-To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.
+To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
 
 ```{r captionSCFtable, echo=FALSE}
-captionSCFtable <- "This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
 We compare average values for four IPCC climate zone classifications.
@@
 For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
-Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
-The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state.
+The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21% of it.
 
-Looking again on the time trend our study shows the substantial impact of changed management on the development of $F_{SCF}$.
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$.
 
@@
 
-The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+The world's SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
 
@@
 
-The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
-Looking on regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
-Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils.
+The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking at regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not suprising as the temperature effects on decomposition are fundamentally different for permafrost soils.
 To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
-Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
 
-Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
+Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
 For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
@@
 The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
-Especially in less forested areas the natural land representation might be off, due to parametrization assumptions of the natural litterfall.
+Especially in less forested areas the natural land representation higher uncertainty may arise from the parametrization assumptions of natural litterfall.
 
-Nevertheless, total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
-Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions.
-Then former natural land SOC stocks are declining to lower levels.
-However, as pointed out before, due to low cropland expansion rates, their SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
+Nevertheless, total SOC stocks are in a reasonable range and results for SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
+The SOC gap from changed management is only affected by the natural vegetation representation in the case of land use transitions.
+In this case, former natural land SOC stocks decline.
+However, as pointed out before, due to low cropland expansion rates, land use change-related SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
 We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
-(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
+(see appendix). This shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
 

### 2020-11-09 SOCbudgetPaper/SOCbudget.Rmd
fix merging issues

@@
 output:
+  bookdown::word_document2: default
+  md_document:
+    variant: markdown_github
   bookdown::pdf_book:
@@
     md_extensions: +raw_attribute
-  md_document:
-    variant: markdown_github
-  bookdown::word_document2: default
   rticles::copernicus_article: default

### 2020-11-09 SOCbudgetPaper/SOCbudget.tex
fix merging issues

@@
 \Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
-\label{eq:totalstock}
+\label{eq:stockdiff}
 \end{equation}
@@
 According to \citep{sanderman_soil_2017} cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
-It was also pointed out that these results might be conservatively low compared to experimental results.
-Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with \citep{sanderman_soil_2017} estimates.
+Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as \citep{sanderman_soil_2017} estimations.
+However, there are large uncertainties in modelling SOC at the global scale, and the \citep{sanderman_soil_2017} results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
 
-Furthermore, the results of \citep{sanderman_soil_2017} modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Furthermore, \citep{sanderman_soil_2017} modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
 \citep{pugh_simulated_2015} considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that productivity gains for crops while increasing yield levels by 18\%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1\% change in SOC emissions.
+The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a~substantial decrease in SOC emissions, accounting for only less than 1\% change in SOC emissions.
 
@@
 We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
-This is challenging \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
-Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g.~due to improved animal waste management system) and (3) the adoption of no tillage practices.
+This challenges \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
+
+Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g.~due to improved animal waste management systems) and (3) the adoption of no tillage practices.
 
@@
 As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
-\citep{keel_large_2017} questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources.
-Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding.
-However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.
+\citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources.
+Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
+However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass.
+Rather, it is more likely that we are still missing carbon inputs to the soil.
 
@@
 To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}) which are based on measurement data for croplands (see \ref{tab:SCFtable}).
-To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.
+To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
 
 \begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
 We compare average values for four IPCC climate zone classifications.
@@
 For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
-Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
-The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7\% of it.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state.
+The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21\% of it.
 
-Looking again on the time trend our study shows the substantial impact of changed management on the development of \(F_{SCF}\).
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of \(F_{SCF}\).
 
@@
 
-The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \ref{tab:SOCtable}).
+The world's SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \ref{tab:SOCtable}).
 
@@
 
-The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
-Looking on regional results \ref{append:regcompare}, our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
-Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils.
+The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking at regional results \ref{append:regcompare}, our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not~suprising as the temperature effects on decomposition are fundamentally different for permafrost soils.
 To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
-Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
 
-Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
+Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
 For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.

### 2020-11-10 SOCbudgetPaper/SOCbudget.Rmd
correct mispelling

--- 2020-11-09 SOCbudgetPaper/SOCbudget.Rmd
@@
   Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
-It is three times larger than the vegetation pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
-Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
+It is three times larger than the vegetation pool and more than twice as big as the atmospheric pool, when looking into the first meter of the earth soil profile.
+Human cropping activities led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
 While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
-Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
-We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
-We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
-We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
+Here we create a spatial explicit data set for agricultural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
+We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975--2010.
+We show that, our results on global scale based on Tier 2 IPCC methodology are in good agreement with Tier 1 default assumptions.
+We also find that SOC is very sensitive to management decision such as residue recycling indicating the necessity to incorporated better management data in soil models.
 bibliography: SOCbudget2.bib

### 2020-11-10 SOCbudgetPaper/SOCbudget.tex
correct mispelling

@@
 Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
-It is three times larger than the vegetation pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
-Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
+It is three times larger than the vegetation pool and more than twice as big as the atmospheric pool, when looking into the first meter of the earth soil profile.
+Human cropping activities led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
 While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
-Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
-We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
-We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
-We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
+Here we create a spatial explicit data set for agricultural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
+We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975--2010.
+We show that, our results on global scale based on Tier 2 IPCC methodology are in good agreement with Tier 1 default assumptions.
+We also find that SOC is very sensitive to management decision such as residue recycling indicating the necessity to incorporated better management data in soil models.
 \end{abstract}
@@
 
-Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modelling.
+Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
 Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
@@
 \textbf{Manure Excretion, Storage and Recycling}
-Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modelling.
+Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
 Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
@@
 Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
-Whereas recent modelling estimates of global SOC emissions indicate an ongoing increase of SOC emissions (\citep{pugh_simulated_2015}, \citep{sanderman_soil_2017}), our study indicates that the global SOC gap is slowly closing due to improved management.
+Whereas recent modeling estimates of global SOC emissions indicate an ongoing increase of SOC emissions (\citep{pugh_simulated_2015}, \citep{sanderman_soil_2017}), our study indicates that the global SOC gap is slowly closing due to improved management.
 
@@
 Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as \citep{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modelling SOC at the global scale, and the \citep{sanderman_soil_2017} results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
+However, there are large uncertainties in modeling SOC at the global scale, and the \citep{sanderman_soil_2017} results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
 
-Furthermore, \citep{sanderman_soil_2017} modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
 \citep{pugh_simulated_2015} considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
@@
 
-Modelling management effects on the global scale comes however with parametric and structural uncertainties.
+modeling management effects on the global scale comes however with parametric and structural uncertainties.
 As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
@@
 
-To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}) which are based on measurement data for croplands (see \ref{tab:SCFtable}).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}) which are based on measurement data for croplands (see \ref{tab:SCFtable}).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
@@
 \begin{figure}[H]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
 Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.}\label{fig:SOCtable}
+In this context they will be compared to modeled data (LPJmL4, this study) for the year 2010.}\label{fig:SOCtable}
 \end{figure}
 
-The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates.
+The global estimates of SOC stock from this study are on the lower end compared to other modeled results or more data driven estimates.
 Looking at regional results \ref{append:regcompare}, our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.

### 2020-11-10 SOCbudgetPaper/discussion.Rmd
correct mispelling

@@
 Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
-Whereas recent modelling estimates of global SOC emissions indicate an ongoing increase of SOC emissions ([@pugh_simulated_2015], [@sanderman_soil_2017]), our study indicates that the global SOC gap is slowly closing due to improved management.
+Whereas recent modeling estimates of global SOC emissions indicate an ongoing increase of SOC emissions ([@pugh_simulated_2015], [@sanderman_soil_2017]), our study indicates that the global SOC gap is slowly closing due to improved management.
 
@@
 Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as [@sanderman_soil_2017] estimations.
-However, there are large uncertainties in modelling SOC at the global scale, and the [@sanderman_soil_2017] results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
+However, there are large uncertainties in modeling SOC at the global scale, and the [@sanderman_soil_2017] results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
 
-Furthermore, [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
 [@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
@@
 
-Modelling management effects on the global scale comes however with parametric and structural uncertainties.
+modeling management effects on the global scale comes however with parametric and structural uncertainties.
 As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
@@
 
-To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
@@
 ```{r captionSOCtable, echo=FALSE}
-captionSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+captionSOCtable <- "modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
 Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
+In this context they will be compared to modeled data (LPJmL4, this study) for the year 2010."
 ```
@@
 
-The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates.
+The global estimates of SOC stock from this study are on the lower end compared to other modeled results or more data driven estimates.
 Looking at regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.

### 2020-11-10 SOCbudgetPaper/methods.Rmd
fix merging issues
correct mispelling

@@
 \Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
-(\#eq:totalstock)
+(\#eq:stockdiff)
 \end{equation}
@@
 
-Crop residue production and management is based on a revised methodology of [@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modelling.
+Crop residue production and management is based on a revised methodology of [@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modeling.
 Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest index values ($HI$) and root:shoot ratios ($RS$) as follows
@@
 **Manure Excretion, Storage and Recycling**
-Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modelling.
+Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modeling.
 Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.

### 2020-12-07 SOCbudgetPaper/methods.Rmd
full update
spell check
add appendix

@@
-# Method
+# Methods
 
-We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
-In the following section we outline the most important relationships and assumptions.
-Table \@ref(append:subsection2mrfunctions) provides further information on corresponding code within the R packages.
+In Sect.
+\@ref(sec:carbonbudget) we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.
+4 ([@calvo_buendia_ipcc_2019]).
+We additionally describe how we configured and extended the steady-state method (for model code see @mrsoil_2020).
+In Sect.
+\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see @mrcommons_2020).
+In Sect.
+\@ref(sec:scenarios) we define the scenarios used to complement our historic model results.
 
-## SOC Stocks and Stock changes following the Tier 2 steady-state method {#sec:carbonbudget}
+## SOC stocks and stock changes following the steady-state method {#sec:carbonbudget}
 
-Following the tier 2 approach of the refinement of IPCC guidelines vol.
-4 (@ipcc_2019_2019; short Tier 2 steady-state method), we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
-We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
-Therefore, we conduct the following three steps within each yearly time-step:
-(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks, 
-(2) we account for land conversion by transferring SOC from and to another land-use type represented as natural vegetation,
-(3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
+Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.
+4 ([@calvo_buendia_ipcc_2019]; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon ($SOC$) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+We restrict our analysis to the first 0-30 cm of the soil profile.
+Moreover, we assume the current $SOC$ state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
+Therefore, we take the following three steps for each year of our simulation period: 
+(1) We calculate annual land-use type-specific steady states and decay rates for $SOC$ stocks (Sect. \@ref(#sec:steadystates)); 
+(2) we account for land conversion by transferring $SOC$ from and natural vegetation (Sect. \@ref(#sec:carbontransfer)),
+(3) we estimate $SOC$ stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \@ref(#sec:totalsoc)). 
+To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \@ref(#sec.initsoc)).
 
-### Steady-state SOC stocks and decay rates
+### Steady-state SOC stocks and decay rates {#sec:steadystates}
 
-In a simple first order kinetic approach the steady-state soil organic carbon stocks $SOC^{eq}$ are given by 
+In a simple first order kinetic approach the steady-state soil organic carbon stocks $SOC^{\mathrm{eq}}$ are given by 
 \begin{equation}
-SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
+SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
 (\#eq:inoutflow)
 \end{equation}
-with $C^{\textrm{in}}$ being carbon inputs to the soil, $k$ denoting the soil organic carbon decay rate; as well as $i$ representing grid cell identifier and $t$ years.
-We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them, following the approach in the Century model (Parton et al, 1988).
-Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
+with $C^{\textrm{in}}$ being the carbon inputs to the soil, $k$ denotes the soil organic carbon decay rate.
+This equation is valid for all grid cells $i$ and all years $t$.
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model (@parton_analysis_1987).
+Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
+We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
 
-**Carbon Inputs to the Soil**
-
-We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \@ref(tab:datasourceinputs)).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \@ref(sec:residues)) and recycled manure (see \@ref(sec:livstmanure)); for natural vegetation litterfall (including fine root turnover) [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
-
-Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see @ipcc_2019_2019).
-For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that both the amount of carbon and its structural composition determine the effective inflow.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \@ref(tab:datasourceinputs).
+Carbon inputs for cropland are below and above ground residues left or returned to the field (see Sect.
+\@ref(sec:residues)) and manure inputs (see Sect.
+\@ref(sec:livstmanure)); for natural vegetation litterfall (including fine root turnover) ([@schaphoff_lpjml4_2018]) is the only source of carbon inflow to the soil.
+Following the IPCC guidelines (@calvo_buendia_ipcc_2019), carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
+For each component the sum of all carbon input sources is allocated to the respective $SOC$ sub-pools via transfer coefficients.
+This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \@ref(tab:datasourceinputs).
 
   ```{=latex}
-
     \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types }
- \begin{tabular}{l l l l}
+    \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
+    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.25\textwidth}}
     \tophline
@@
     \middlehline
- \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
-                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
-                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
+    \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+                                                                        \cite{faostat_faostat_2016}, \\
+                                                                        \cite{schaphoff_lpjml4_2018} (LPJmL),\\
+                                                                        \cite{weindl_livestock_2017} \end{minipage}} & 
+                                                        \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+                                                                        lg/c generic values according to Table 5.5B\\
+                                                                        in \cite{calvo_buendia_ipcc_2019},\\
+                                                                        crop-type specific n/c-ratios from
+                                                                        \cite{dietrich_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
+                              & below ground residues &  &  \\ \cline{2-2}
+                              & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
                               \hline
-  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
+    Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018} (LPJmL) 
+                              & \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut 
+                                  \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
    \bottomhline
@@
 
-<!-- add references here for the table-->
-
-**SOC decay**
-
-The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
-Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
-Thus, $k_{sub}$ is given by
+The sub-pool specific decay rates $k_{sub}$ are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space ($i$) and time ($t$).
+Following the steady-state method [@calvo_buendia_ipcc_2019], we consider temperature ($temp$), water ($wat$), sand fraction ($sf$) and tillage ($till$) effects to account for spatial variation of decay rates.
+Thus, $k_{sub}$ are given by
 
@@
 \begin{aligned}
-& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
-& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
-& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
+& k_{i,t,\mathrm{active},lu}  & = &~ k_{\mathrm{active}}  ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ & \cdot~ sf_{i}\\
+& k_{i,t,\mathrm{slow},lu}    & = &~ k_{\mathrm{slow}}    ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ &\\
+& k_{i,t,\mathrm{passive},lu} & = &~ k_{\mathrm{passive}} ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ & ~ & 
 (\#eq:decayrates)
-\end{aligned}
+\end{aligned}.
 \end{equation}
 
-For cropland, we distinguish the effect of different tillage (see \@ref(sec:tillage)) and irrigation (see \@ref(sec:irrigation)) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as drivers considered for each land-use types are shown in table \@ref(tab:datasourcedecay).
-To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
+For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect.
+\@ref(sec:tillage)) and irrigation (see Sect.
+\@ref(sec:irrigation)) practices on decay rates.
+We calculated area weighted means for $till$ and $wat$ on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \@ref(tab:datasourcedecay); equations are displayed by equation 5.0B--5.0F in [@calvo_buendia_ipcc_2019].
 
@@
  \middlehline
- \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm &  [SoilGrids]  \\
+ \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
+                                     & \cite{hengl_soilgrids250m:_2017} \\
                       \cline{2-4}
 
-                      & Mircobial activity & air temperature & [CRUp4.0] \\
+                      & Mircobial activity & air temperature & \cite{harris_version_2020} \\
                       \cline{2-4}
 
-                      & Water restriction & precipitation \& potential evapotranspiration & [CRUp4.0] \\
+                      & Water restriction & precipitation \& potential evapotranspiration & \cite{harris_version_2020} \\
                       \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & [sec:irrigation] \\ 
+\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. @ref(sec:irrigation) \\ 
                       \cline{2-4}
 
-                      & Soil disturbance & tillage & [sec:tillage] \\
+                      & Soil disturbance & tillage & Sect. @ref(sec:irrigation) \\
  \bottomhline
@@
 
-<!-- add references here for the table-->
+### SOC transfer between land-use types {#sec:carbontransfer}
 
-### SOC transfer between land-use types
-
-We calculate SOC stocks based on the area shares of land-use types (lut) within the half-degree grid cells (i).
-If land is converted from one land-use type into others (!lut), the respective share of the SOC stocks is reallocated.
-We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via
+We calculate $SOC$ stocks based on the area shares of land-use types ($lu$) within our grid cells ($i$).
+If land is converted from one land-use type $lu=\{crop,natveg\}$ into the other $!lu=\{natveg,crop\}$, a respective share of the $SOC$ is reallocated.
+We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{t^*}$ via
 
 \begin{equation}
-SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
+SOC_{i,t^*,sub,lu} = SOC_{i,t-1,sub,lu} - \frac{SOC_{i,t-1,sub,lu}}{A_{i,t-1,lu}} \cdot  AR_{i,t,lu} + \frac{SOC_{i,t-1,sub,!lu}}{A_{i,t-1,!lu}} \cdot  AE_{i,t,lu}
 (\#eq:ctransfer)
@@
 
-with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land-use type $lut$.
-Note that $!lut$ denotes the sum over all other land-use types, which decreases in the specific time step $t$.
-Data sources and methodology on land-use states and changes are described in \@ref(sec:landuse).
+with $A_{lu}$ being the land-use type specific areas, $AR_{lu}$ and $AE_{lu}$ the area reduction resp.
+area expansion of the two land-use types.
+Data sources and methodology on land-use states and changes are described in Sect.
+\@ref(sec:landuse).
 
-### Total SOC stocks and stock changes
+### Total SOC stocks and stock changes {#sec:totalsoc}
 
-Carbon stocks $SOC$ for each sub-pool (sub) converge towards the calculated steady-state stock $SOC^{eq}$ for each land-use type (lut), each sub-pool (sub) and each annual time step (t) like
+$SOC$ converge towards the calculated steady-state stock $SOC^{\mathrm{eq}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
+SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t-1,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
 (\#eq:SOCstate)
@@
 
+Note that the decay rates have to be multiplied by one year ($1a$) to form a dimensionless factor.
 Reformulating this equation, we obtain a mass balance equation as follows
@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
+SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} - \underbrace{SOC_{i,t-1,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
 (\#eq:steadystate2budget)
@@
 
-The global SOC stock for each time step can then be calculated via
+The global $SOC$ stock for each time step $t$ can then be calculated via
 
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
 (\#eq:totalstock)
@@
 
-According to the IPCC guidelines SOC changes can be expressed as the difference of two consecutive years (see Eq.
-5.0A in [@ipcc_2019_2019]).
-This however, will also include naturally occurring changes due to climatic variance over time.
-For our study we defined the absolute and relative SOC changes in relation to a potential natural state $SOC_{pnv}$ under the same climatic conditions at time $t$ in grid cell $i$, that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
-The absolute changes $\Delta SOC$ and relative changes $F^{SCF}$ a thus given by
+According to the IPCC guidelines $SOC$ changes can be expressed as the difference of two consecutive years (see Eq.
+5.0A in [@calvo_buendia_ipcc_2019]).
+This, however, will also include naturally occurring changes due to climatic varition over time.
+For our study, we defined the absolute and relative SOC changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
+The absolute changes $\Delta SOC$ and relative changes $F^{\mathrm{SCF}}$ are thus given by
 
 \begin{equation}
-\Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
+\Delta SOC_{i,t} = SOC_{i,t} - SOC^{\mathrm{pnv}}_{i,t}\qquad \text{and} \qquad  F^{\mathrm{SCF}}_{i,t} = \frac{SOC_{i,t}}{SOC^{\mathrm{pnv}}_{i,t}} .
 (\#eq:stockdiff)
@@
 
-Note that the absolution changes $\Delta SOC$ can be also interpreted as the SOC gap due to human cropping activities; whereas relative changes $F^{SCF}$ can be denoted as stock change factors as defined within the IPCC guidelines of 2006 [@ipcc_2006_2006].
-Note that the SOC gap is equivalent to the negated cumulative SOC emission ([@pugh_simulated_2015]) also referred to as SOC debt ([@sanderman_soil_2017]) of human land use.
+Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_hs_ipcc_2006].
+Note that the SOC debt is equivalent to the negated cumulative SOC emission ([@pugh_simulated_2015]) of human land use.
 
-### Initialization of SOC pools
+### Initialization of SOC pools {#sec:initsoc}
 
-To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions decreases.
+To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
+While land-use, climate and litter input information are available till 1901, agricultural management data on residue and manure inputs are hold constant at the level of 1965.
+After a spin-up period of 64 years with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975.
+Note that irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
 
-## SOC Stocks and Stock changes following Tier 1 {#sec:tier1}
+<!-- backup
+```{=latex}
+ \begin{table*}[h]
+ \caption{Setup of spin-up and simulation phase}
+ \begin{tabular}{l l l}
+ \tophline
+ \multirow{3}{*}{1901--1964} 
+   & Constant management spin-up phase with
+   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
+   & constant management data from 1965 (manure and residue inputs, tillage and irrigation areas)\\
+                      \cline{1-4}
+ \multirow{3}{*}{1965--1974}
+   & Dynamic management spin-up phase with
+   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
+   & historic management data (manure and residue inputs, tillage and irrigation areas)
+                      \cline{1-4}
+ \multirow{3 }{*}{1975--2010}                     
+   & Simulation phase 
+   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
+   & historic management data (manure and residue inputs, tillage and irrigation areas)              
+ \bottomhline
+ \end{tabular}
+ \belowtable{}
+ (\#tab:datasourcedecay)
+ \end{table*}
+``` -->
+## SOC stocks and stock changes following Tier 1 {#sec:tier1}
 
-Additionally, to the tier 2 approach of the refinement of IPCC guidelines vol.
-4 [@ipcc_2019_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.
-4 (@ipcc_2006_2006, @ipcc_2019_2019).
-Here, stocks are calculated via stock change factors ($F^{SCF}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-$F^{SCF}$ are differentiated by different crop, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover, $F^{SCF}$ vary for different climatic zones ($c$) specified by the IPCC (see \@ref(append:climatemap)).
-The actual SOC stocks as thus calculated based on a given reference SOC stock by
+Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+Estimates of $F^{\mathrm{SCF}}$ are differentiated by different crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig.
+\@ref(append:climatemap)).
+The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}$ by
 
 \begin{equation}
-SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
+SOC_{i,t} = \sum_{c,m} T_{c,i} \cdot SOC^{\mathrm{ref}}_{i,t} \cdot F^{\mathrm{SCF}}_{c,m}
 (\#eq:tier1)
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis we use the default $F^{SCF}$ from the Tier 1 method of [@ipcc_2006_2006], and [@ipcc_2019_2019] as a comparison and consistency check for our more detailed budget approach.
+For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of [@eggleston_hs_ipcc_2006] and [@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
-\newpage
-
-## Agricultural management data on 0.5 degree resolution {#sec:agrimanagement}
+## Agricultural management data at 0.5 degree resolution {#sec:agrimanagement}
 
 We compile country-specific FAO production and cropland statistics [@faostat_faostat_2016] to a harmonized and consistent data set.
-The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
+The data is prepared in 5-year time steps from 1965 to 2010, which, together with the spin-up phase, restricts our analysis to the time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
 
-### Land use and Land-use change {#sec:landuse}
+### Land use and land-use change {#sec:landuse}
 
-Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see \@ref(append:Tableluh2fao2mag) for more  details on the crop group mapping).
-Land-use transitions are calculated as net area differences of the land-use data on half-degree.
+Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see Table \@ref(append:Tableluh2fao2mag) for more details on the crop group mapping).
+By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
+Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
 
-### Crop and Crop Residues Production {#sec:residues}
+### Crop and crop residues production {#sec:residues}
 
-**Crop Production**
-
-Using half-degree yield data from LPJmL [@schaphoff_lpjml4_2018] as well as half-degree cropland patterns (see \@ref(sec:landuse)) we compile crop group specific half-degree production patterns.
+Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL [@schaphoff_lpjml4_2018] as well as half-degree cropland patterns (see Sect.
+\@ref(sec:landuse)).
 We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production [@faostat_faostat_2016].
-Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
-
-**Crop Residue Production**
+Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallow land.
 
@@
 \begin{aligned}
-AGR & = CP \cdot HI_{\textrm{slope}} + CA \cdot HI_{\textrm{intercept}}\qquad & \textrm{and} \\
-BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
+AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
+\qquad & \textrm{and} \\
+BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
 (\#eq:resbiomass)
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction [@ipcc_2006_2006].
-Note that deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
-We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals [@weindl_livestock_2017], fuel or for material use.
+Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_hs_ipcc_2006].
+Note that deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+We assume that all $BGR$ are left in the soil, whereas $AGR$ can be burned or harvested for other purposes such as feeding animals [@weindl_livestock_2017], fuel or for material use.
 
-**Burned Residues**
+A fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
+Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_hs_ipcc_2006].
 
-A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
-Following [@smil_nitrogen_1999]) we assume 25% burn share for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group 80–90% of the residue carbon burned on the fields are lost within the combustion process [@ipcc_2006_2006].
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see Table \@ref(append:Tablekcr2kres)).
+Residue feed demand for five different livestock groups is based on country-specific feed baskets (see [@weindl_livestock_2017]), that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
+We estimate a material-use share for the straw-residue group of 5% and a fuel-share of 10% for all used residue groups in low-income countries.
+For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above.
+The remaining $AGR$ as well as all $BGR$ are expected to be left on the field.
+We limit high returning rates at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
 
-**Residue Usage**
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(append:Tablec2dm)).
 
-We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \@ref(append:Tablekcr2kres)).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see [@weindl_livestock_2017]), taking available AGR biomass as well as livestock productivity into account.
-We estimate, for low-income countries,  a material-use share for straw residues of 5% and a fuel-share of 10% for all used residues groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumed, and middle-income countries use shares are linearly interpolated based on per-capita income.
-The remaining AGR as well as all BGR are expected to be recycled to the soil.
-We limit high recycling shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
+### Livestock distribution and manure excretion {#sec:livstmanure}
 
-**Dry Matter to Carbon Transformation**
+Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems 	at the present day a lot of the manure has to be stored and can be applied on croplands.
+We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
 
-To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \@ref(append:Tablec2dm)).
+To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of [@robinson_mapping_2014] and applying feed basket assumptions based on a revised methodology from [@weindl_livestock_2017].
+We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
+Due to great feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
+We distinguish between grazed pasture, which is converted into livestock products in extensive systems, and primary-crop feed stuff, which we consider to be consumed in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to distinguish between intensive and extensive production systems.
+For low-income countries, we assume only extensive production systems.
+We locate them according to the share of built-up areas based on the asumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
+Intensive production associated with high-income countries, is distributed within a country using the share of primary-crop production, assuming that feed availability is the most determining factor for livestock location.
+For middle-income countries we split the livestock production into extensive and intensive systems based on the per-capita income.
 
-### Livestock Distribution and Manure Excretion {#sec:livstmanure}
-
-We assume that manure is applied at its excretion place, leaving the livestock distribution the driving factor for the spatial pattern of manuring.
-
-**Livestock Distribution**
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of [@robinson_mapping_2014], and uses feed basket assumptions based on a revised methodology from [@weindl_livestock_2017].
-We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-For ruminants, we assume that livestock is located where the production of feed occurs.
-We distinguish between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumed in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
-For low-income countries, we assume extensive production systems.
-We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labor per animal ratio.
-Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production,  assuming that feed availability is the most driving factor for livestock location.
-
-**Manure Excretion, Storage and Recycling**
 Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modeling.
-Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios (given by [@ipcc_2019_2019]).
+Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed c/n ratios (given by [@calvo_buendia_ipcc_2019]).
 Depending on the feed system we assume manure to be handled in four different ways:
-All manure originated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Manure fuel shares are estimated using IPCC default values [@ipcc_2006_2006].
+All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
 Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing. 
-To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in [@bodirsky_n2o_2012] (see \@ref(append:TableclossAWMS)).
+For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing. 
+To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems.
+Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
+For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in [@bodirsky_n2o_2012] (see Sect.
+\@ref(append:TableclossAWMS) for more details).
 We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
@@
 
-We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
 The LUH2v2 [@hurtt_harmonization_2020] data set provides irrigated fractions for their cropland subcategories.
-We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
-Moreover, we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
+We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient $wat$ on decay rates using these shares to compute the weighted mean between rainfed and irrigated $wat$ factors.
+As a result $wat$ is the same for all crop groups within a grid cell.
+Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a (partly) irrigated grid cells, since we do not have crop group specific growing periods at hand.
+In addition, we expect, either water limitations to be no problem during the off-season (causing our assumption to be a good approximation), or irrigated growing periods to span over the whole year in water-restricted areas.
 
@@
 
-To account for the distribution of tillage to the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas as default annual crops are under full and perennials under reduced tillage.
+In order to derive a spatial distribution of the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
 Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from  [@porwollik_generating_2018].
-This data set is extended by to the period of 1974-2010 by combining country-level data on conservation agriculture area values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the methodology of [@porwollik_generating_2018] to identify potential no tillage cells.
+This data set is extended to the period of 1975-2010 by combining country-level data on areas under conservation agriculture from [@fao_aquastat_2016] available from 1975 on with LUH2 physical crop areas, applying the methodology of [@porwollik_generating_2018] to identify potential no-tillage grid-cells.
+
+## Scenario definitions {#sec:scenarios}
+
+To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
+In the following sections we outline name and idea of these scenarios (for technical implementation see [@mrsoil_karstens_2020]).
+
+### Management scenarios {#sec:scen_management}
+
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practises are neglected (adoption starts in 1975 according to the available data set); (2) the \textit{constResidues} and the \textif{constManure} scenario assume constant input rates from residues resp.
+manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards.
+Within the \textit{constResidue} scenario at least three effects overlay each other: (1) yields and with them residue biomass increase due to productivity gains, (2) rates of residue left or returned to fields are raising, and (3) shifts of crop groups change the amount of residue biomass due to crop-group specific harvest index values.
+The \textit{constManagement} combines all three scenarios.
+
+### Initialization choice {#sec:scen_initial}
+
+As outlined in Sect.
+@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
+Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
+We additionally combined the \textit{Initial-natveg} scenario with the (\textit{constManagement} scenario.
+
+### Parameterization of natural litter fall {#sec:scen_litterpnv}
+
+Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short lg/c) and nitrogen-to-carbon (short n/c) ratios for all land under natural vegetation.
+This is however a strong simplification.
+Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
+Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
+Whereas lg/c ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, n/c ratios range between $0.001-0.05$ with an average of $0.015$ (default value).
+We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
+We test however a range of different plant types as specified in Table @ref(tab:scen_litterpnv).
+
+```{=latex}
+
+ \begin{table*}[h]
+ \caption{Scenarios for sensitivity analysis on litter parameterization}
+ \begin{tabular}{l l l}
+ \tophline
+  Scenario name & plant type & n/c & lg/c \\
+  \middlehline
+  \textit{LitterPNV-CenturyAverage} & mix over all tree types & 0.015  & 0.25 \\
+  \textit{LitterPNV-PerennialGrasses} & perennial grasses     & 0.029  & 0.11 \\
+  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
+  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
+  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
+  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
+  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
+ \bottomhline
+ \end{tabular}
+ \belowtable{}
+ (\#tab:scen_litterpnv)
+ \end{table*}
+```

### 2020-12-10 ResultNotebooks/results_SOCcompare.Rmd
mayor update on figures and flowfig

@@
 library(mrcommons)
-library(mrvalidation)
-library(mrSOCbudget)
+library(mrsoil)
 library(magpiesets)
 
-rev       <- "rev0.76baseline"
-outputdir <- "Output/Images/"
+rev       <- "rev0.11"
+outputdir <- "Output/"
 
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+#untar(paste0(getConfig("outputfolder"),"/",rev,"histManagement_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
 
-SoilCarbon <- read.magpie("Data/Baseline/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Baseline/Landuse.rds")
+SoilCarbon <- read.magpie("Data/histManagement/SoilCarbon.rds")
+Landuse    <- read.magpie("Data/histManagement/Landuse.rds")
 
@@
 iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
-iso2reg      <- toolMappingFile("regional",getConfig("regionmapping"),readcsv = TRUE)
+iso2reg      <- read.csv2("Data/histManagement/regionmappingH12.csv")
 reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
-
-print(dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=c(1,3)))
-
 ```
@@
 
+library(mrvalidation)
+
   LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
-  #LPJ_rev21  <- calcOutput("ValidCarbonStocks", datasource = "LPJmL_rev21", aggregate="REG+GLO", try=TRUE)
-  #LPJ_car    <- calcOutput("ValidCarbonStocks", datasource = "LPJmLCarbon", aggregate="REG+GLO", try=TRUE)
   WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
@@
 
-Plot global stocks
-
-```{r plot glo stocks}
-
-plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red", xlim=c(1951,2017),  ylim=c(350000,1300000))
-lines(getYears(LPJ_IPCC, as.integer = TRUE), LPJ_IPCC["GLO",,1], type="l", col="green")
-#lines(getYears(LPJ_car, as.integer = TRUE), LPJ_car["GLO",,], type="l", col="darkgreen")
-lines(getYears(LPJmL4, as.integer = TRUE), LPJmL4["GLO",,], type="l", col="lightgreen")
-lines(getYears(WISE, as.integer = TRUE), WISE["GLO",,], type="l", col="brown")
-lines(getYears(SoilGrids, as.integer = TRUE), SoilGrids["GLO",,], type="l", col="blue")
-lines(getYears(GSOC, as.integer = TRUE), GSOC["GLO",,], type="l", col="pink")
-plot(c(getYears(glo_SOCStock, as.integer=TRUE)), glo_SOCStock, type="l", col="red")
-```
-
-```{r table ipcc scf}
+```{r table glo soc stocks}
 
@@
 
+supp.labs <- getNames(glo_SOC["GLO",,invert=TRUE])
+names(supp.labs) <- c("OJ", "VC")
+
 leg_name <- "Data Source"
@@
 ```
-
-```{r plot diff maps, out.width="100%", out.height=1000}
-
-library(luplot)
-library(magpiesets)
-
-cellSOCdens   <- (dimSums(dimSums(SoilCarbon[,,"naturalstate"], dim=3.2) * Landuse, dim=c(3))/dimSums(Landuse, dim=c(3)))[,findset("past"),]
-cellSOCdens[is.na(cellSOCdens)] <- 0
-
-cellSoilGrids <- readSource("SoilGrids", subtype="cstock_0_30", convert="onlycorrect")
-cellLPJml     <- calcOutput("LPJmL", version="LPJmL4", climatetype="CRU_4", subtype="soilc_layer", time="raw", selectyears="past", aggregate=FALSE)
-cellLPJml     <- collapseNames(cellLPJml[,,"layer1"]+1/3*cellLPJml[,,"layer2"], collapsedim = c(2,3))
-
-diffSoilGrids <- cellSOCdens - cellSoilGrids
-diffLPJml     <- cellSOCdens - cellLPJml
-
-luplot::plotmap(diffLPJml[,10,], legend_range = c(-80,80))
-```
+`

### 2020-12-10 ResultNotebooks/results_SOCflows.Rmd
Mayor update of everything
mayor update on figures and flowfig

@@
 library(mrvalidation)
-library(mrSOCbudget)
+library(mrsoil)
 library(magpiesets)
 
-rev       <- "rev0.7_mrSOC"
+rev <- "rev0.78baseline"
+
 outputdir <- "Output/Images"
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/Baseline", extras="--keep-newer-files ")
+
+revfolder <- "Data/Baseline/"
 
@@
 
-CarbonInput      <- readRDS("Data/CarbonInput.rds")
-CarbonLitter     <- readRDS("Data/CarbonLitter.rds")
-CarbonManure     <- readRDS("Data/CarbonManure.rds")           
-CarbonResidues   <- readRDS("Data/CarbonResidues.rds")
-CarbonHarvest    <- readRDS("Data/Crop_Harvest.rds")
-SoilCarbonOut    <- readRDS("Data/SoilCarbon.rds")
-SoilCarbonDecay  <- readRDS("Data/SoilDecay.rds")
-SoilCarbonStSt   <- readRDS("Data/SoilSteadyState.rds")
+CarbonInput      <- readRDS(paste0(revfolder,"CarbonInput.rds"))
+CarbonLitter     <- readRDS(paste0(revfolder,"CarbonLitter.rds"))
+CarbonManure     <- readRDS(paste0(revfolder,"CarbonManure.rds"))          
+CarbonResidues   <- readRDS(paste0(revfolder,"CarbonResidues.rds"))
+CarbonHarvest    <- readRDS(paste0(revfolder,"Crop_Harvest.rds"))
+SoilCarbonOut    <- readRDS(paste0(revfolder,"SoilCarbon.rds"))
 
-Landuse          <- readRDS("Data/Landuse.rds")
-LanduseChange    <- readRDS("Data/LanduseChange.rds")
+Landuse          <- readRDS(paste0(revfolder,"Landuse.rds"))
+LanduseChange    <- readRDS(paste0(revfolder,"LanduseChange.rds"))
 
-FAOmassbalance   <- readRDS("Data/FAOmassbalance_ISO.rds" )
-ManureExcreted   <- readRDS("Data/Manure_excreted.rds")
-ManureRecycled   <- readRDS("Data/Manure_recycled.rds")
-ResidueBiomass   <- readRDS("Data/Residue_Biomass.rds")
-ResiduesAgDemand <- readRDS("Data/ResiduesAg_FieldBalance.rds")
+FAOmassbalance   <- readRDS(paste0(revfolder,"FAOmassbalance_ISO.rds"))
+ManureExcreted   <- readRDS(paste0(revfolder,"Manure_excreted.rds"))
+ManureRecycled   <- readRDS(paste0(revfolder,"Manure_recycled.rds"))
+ResidueBiomass   <- readRDS(paste0(revfolder,"Residue_Biomass.rds"))
+ResiduesAgDemand <- readRDS(paste0(revfolder,"ResiduesAg_FieldBalance.rds"))
 
@@
                               setNames(calcOutput("PassiveDecay", landtype="natveg", aggregate = FALSE),"natveg"))
-
+decay <- calcOutput("Decay", aggregate = FALSE)
 subpools     <- c("active","slow","passive")
@@
 ```
-
-Biomass excreted ist bigger than feed intake via plant harvest and residues taken for feed.
-
-```{r soc}
-
-  Litter             <- dimSums(calcOutput("CarbonLitter", aggregate=FALSE)[,,"c"]*Landuse[,,"natveg"], dim=1)
-
-  # Calculate effective input
-  param <- readSource("IPCCSoil", convert=FALSE)
-  param.f1_metab2a          <- setYears(param[,,"f1"], NULL) 
-  # stabilization efficiencies for metabolic decay products entering the active pool
-  param.f3_struc2s          <- setYears(param[,,"f3"], NULL) 
-  # stabilization efficiencies for structural decay products entering the slow pool
-  param.f2_struc2a <- mbind(setNames(setYears(param[,,"f2_ft"], NULL), "crop"),
-                            setNames(setYears(param[,,"f2_nt"], NULL), "natveg"))
-  # stabilization efficiencies for structural decay products entering the active pool (tillage and no tillage specific)
-
-  cell.input       <- mbind(add_dimension(dimSums(calcOutput("CarbonInput", landtype="crop", aggregate = FALSE),dim=3.2),
-                                          dim=3.1, nm = "crop",   add="landuse"),
-                            add_dimension(collapseNames(calcOutput("CarbonInput", landtype="natveg", aggregate = FALSE)),
-                                          dim=3.1, nm = "natveg", add="landuse"))
-  CarbonInput      <- dimSums(cell.input[,,"totalC"]*Landuse, dim=c(1))
-
-  CarbonInputEff   <- collapseNames(dimSums((cell.input[,,"metabDOC"] * param.f1_metab2a +
-                                             cell.input[,,"strucDOC"] * param.f2_struc2a +
-                                             cell.input[,,"ligninC"]  * param.f3_struc2s) * Landuse, dim=c(1)))
-
-  CarbonInputTot   <- collapseNames(dimSums(SOC[,,"steadystate"]*SOC[,,"decayrate"]*Landuse, dim=c(1,3.3)))
-
-  # Load inputs, active steady state stock and decay rate
-
-  SOC                <- calcOutput("SoilCarbon", output="full", aggregate=FALSE)
-  Landuse            <- calcOutput("Landuse",       aggregate=FALSE)
-  LanduseChange      <- calcOutput("LanduseChange", aggregate=FALSE)
-
-  CarbonStocks       <- collapseNames(dimSums(SOC[,,c("steadystate","actualstate")]*Landuse, dim=1))
-  CarbonTransfer     <- dimSums(toolConditionalReplace(SOC[,,"carbontransfer"], "<0", 0),dim=c(1,3.3))
-  CarbonRelease      <- collapseNames(dimSums(SOC[,,"interstate"]*SOC[,,"decayrate"]*Landuse, dim=c(1, 3.3)))
-
-print(mbind(CarbonTransfer, CarbonRelease))
-```

### 2020-12-10 ResultNotebooks/results_SOCmaps.Rmd
mayor update on figures and flowfig

@@
 library(mrvalidation)
-library(mrSOCbudget)
+library(mrsoil)
 library(magpiesets)
 
-rev       <- "rev0.76baseline"
-outputdir <- "Output/Images"
+rev <- "rev0.11"
 
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
 
-SoilCarbon <- read.magpie("Data/Baseline/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Baseline/Landuse.rds")
+filesNeeded <- c("SoilCarbon.rds","Landuse.rds")
+
+scenarios    <- c("histManagement","constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
+                  "constManagement2-1975", "constResrate-1975", "constYield-1975",
+                 "Initial-natveg", "Initial-natveg_constManangement-1975", "Initial-natveg_constManangement2-1975", 
+                 "LitterPNV-PerennialGrasses", "LitterPNV-AASS", "LitterPNV-BRCF", 
+                 "LitterPNV-MTCF", "LitterPNV-BLEAF",  "LitterPNV-TREF")
+
+# for(item in scenarios){
+#    dir.create(path=paste0("Data/",paste0(rev,item)))
+#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+#    files=filesNeeded, exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ")
+# }
+
+outputdir <- "Output/Images"
+
+SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
 
@@
 glo_SOCEmission <- (glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
-plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCEmission/1000, type="l", xlab="Time", ylab="Global SOC Debt in GtC")
-plot(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCStock[,,"actualstate"])
-lines(getYears(glo_SOCEmission, as.integer = TRUE),glo_SOCStock[,,"naturalstate"])
+
+SoilCarbon <- SoilCarbon[,findset("past"),]
+Landuse    <- Landuse[,findset("past"),]
 
@@
   labs( x = "Year", 
-        y = "Global SOC gap\n in GtC") +
+        y = "Global SOC loss\n in GtC") +
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(trans = squish_trans(-35,0,15),
-                     limits = c(-45,0),
-                     breaks = c(0, -15, -35, -40, -45)) +
+  scale_y_continuous(trans = squish_trans(-18,0,10),
+                     limits = c(-28,0),
+                     breaks = c(0, -20, -22, -24, -26, -28)) +
   theme(rect=element_blank(),
@@
         ) +
-  annotate(geom="text", label="SOC debt", x=factor(1995), y=-35, size = 3) +
-  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -39, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
+  annotate(geom="text", label="SOC debt", size=14, x=factor(1995), y=-20, size = 3) +
+  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -24, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
 
@@
     scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression("cropland SOC gap in tons of carbon per hectar"), 
+                      name = expression("cropland SOC gap in tons of carbon per hectare"), 
                       values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
                       breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("> -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
+                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
                       guide = guide_legend(
@@
     scale_fill_stepsn(limits=c(0,150), colours = cols, 
-                      name = expression("cropland SOC in tons of carbon per hectar"), 
+                      name = expression("cropland SOC in tons of carbon per hectare"), 
                       values=c(0,5,10,20,40,70,100,150)/150,
@@
 
-grid <- plot_grid(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2, labels=c("a","b", "c", "d"),
+grid <- plot_grid(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2, labels=c("(a)","(b)", "(c)", "(d)"),
                   label_fontface = "plain", label_size = 10)

### 2020-12-10 ResultNotebooks/results_SOCscenario.Rmd
mayor update on figures and flowfig

--- 2020-10-06 ResultNotebooks/results_SOCscenario.Rmd
@@
 library(mrvalidation)
-library(mrSOCbudget)
+library(mrsoil)
 library(magpiesets)
 
-rev <- "rev0.76baseline"
+rev <- "rev0.11"
 
@@
 
-#revisions   <- c(Baseline="rev0.74baseline", Till="rev0.74tillage_mixedtill", ResRe="rev0.74residue_freeze1975", ManRe="rev0.74manure_freeze1975", Yield="rev0.74yield_freeze1975")
 filesNeeded <- c("SoilCarbon.rds","Landuse.rds")
 
-revisions   <- c(Baseline="rev0.76baseline", Allon="rev0.76allon_freeze1975", Allon2="rev0.76allon2_freeze1975",        ResRe="rev0.76rrecycle_freeze1975",
-                Till="rev0.76tillage_mixedtill", ResSta="rev0.76residue_freeze1975", Yield="rev0.76yield_freeze1975", ManRe="rev0.76manure_freeze1975")
+scenarios    <- c("histManagement","constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
+                  "constManagement2-1975", "constResrate-1975", "constYield-1975",
+                 "Initial-natveg", "Initial-natveg_constManagement-1975", "Initial-natveg_constManagement2-1975", 
+                 "LitterPNV-PerennialGrasses", "LitterPNV-AASS", "LitterPNV-BRCF", 
+                 "LitterPNV-MTCF", "LitterPNV-BLEAF",  "LitterPNV-TREF")
 
-# for(item in revisions){
-#    dir.create(path=paste0("Data/",names(which(revisions[]==item))))
-#    untar(paste0(getConfig("outputfolder"),"/",item, "_h12_carbonbudget.tgz"), files=filesNeeded, exdir=paste0("Data/", names(which(revisions[]==item))))#, extras="--keep-newer-files ")
-# }
+# untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
+# 
+for(item in scenarios[10:11]){
+   untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+   files=filesNeeded, exdir=paste0("Data/", item), extras="--keep-newer-files ")
+}
 
-outputdir <- "Output/Images"
+outputdir <- "Output/"
 
@@
 
-```{r Carbon Emissions (glo)}
-
-library(ggplot2)
+```{r Load data and calc globals}
 
@@
 
-for(r in 1:length(revisions)){
-  SoilCarbon <- read.magpie(paste0("Data/",names(revisions[r]),"/SoilCarbon.rds"))
-  Landuse    <- read.magpie(paste0("Data/",names(revisions[r]),"/Landuse.rds"))
+for(scen in scenarios){
+  SoilCarbon <- read.magpie(paste0("Data/", scen, "/SoilCarbon.rds"))
+  Landuse    <- read.magpie(paste0("Data/", scen, "/Landuse.rds"))
   SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
   SOCStock        <- toolConditionalReplace(dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2), "is.nan()",0)
-  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), dim=3.1, nm=names(revisions[r]), add="scenario")) 
+  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), 
+                                                       dim=3.1, nm=scen, add="scenario")) 
 }
 
-# 
+glo_SOCDebt  <- collapseNames(glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
+glo_SOCShare <- collapseNames(toolConditionalReplace(glo_SOCStock[,,"actualstate"]/glo_SOCStock[,,"naturalstate"], "is.na()",0) )
+
+saveRDS(mbind(add_dimension(collapseNames(glo_SOCStock[,,"actualstate"]), dim=3.1, add="type", nm="stock"),
+              add_dimension(              glo_SOCDebt,                    dim=3.1, add="type", nm="debt"),
+              add_dimension(              glo_SOCShare,                   dim=3.1, add="type", nm="share")), 
+        paste0(outputdir, "glo_SOCscenarions.rds"))
 ```
 
-```{r plot Carbon Emissions (glo)}
+```{r Carbon inputs (glo)}
 
-plot_rev <- c(3,5,6,8)
+CarbonInput  <- read.magpie(paste0("Data/histManagement/CarbonInput.rds"))
+glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))
 
-names(revisions) <- c("Historical Management", 
-                      "Allon",    
-                      "Constant management from 1975",   
-                      "ResRe", 
-                      "Constant tillage area from 1975",
-                      "Constant residue input from 1975",
-                      "Yield", 
-                      "Constant manure application from 1975")
+ci2plot <- c('Above Ground Residues'="ag.crop",          
+             'Below Ground Residues'="bg.crop",          
+             'Manure from Monogastrics'="livst_chick.crop",
+             'Manure from Monogastrics'="livst_egg.crop",
+             'Manure from Monogastrics'="livst_pig.crop",
+             'Manure from Ruminants'="livst_milk.crop",  
+             'Manure from Ruminants'="livst_rum.crop",   
+             'Natural Litterfall and Fine Roots'="litfall.natveg")
 
-colr <- c("black","green","red","brown","pink","darkgreen","yellow","blue")
-glo_SOCEmission <- collapseNames(glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
-glo_SOCShare    <- collapseNames( toolConditionalReplace(glo_SOCStock[,,"actualstate"]/glo_SOCStock[,,"naturalstate"], "is.na()",0) )
-
-getNames(glo_SOCEmission) <- names(revisions)
-plot(getYears( glo_SOCEmission, as.integer = TRUE), -glo_SOCEmission[,,1], type="l", xlab="Time", ylab="Global SOC gap in GtC", ylim=c(-45,-35))
-legend("topleft", legend=names(revisions[c(1,plot_rev)]), col=colr[c(1,plot_rev)], lty=1, cex=0.8)
-for(r in plot_rev) lines(getYears(glo_SOCEmission, as.integer = TRUE), -glo_SOCEmission[,,names(revisions[r])], type="l", col=colr[r])
-print(cbind(colr[c(1,plot_rev)], names(revisions[c(1,plot_rev)])))
-print(glo_SOCStock[,,"actualstate"])
-
-#plot(getYears( glo_SOCEmission, as.integer = TRUE), dimSums(Landuse[,,"crop"], dim=1), type="l")
-#plot(getYears( glo_SOCEmission, as.integer = TRUE), dimSums(glo_SOCStock[,,"Baseline.actualstate"], dim=1), type="l")
+glo_SOCInput <- toolAggregate(glo_SOCInput, rel=data.frame(from=ci2plot,to=names(ci2plot)), from="from", to="to", dim=3)
 
@@
 
-```{r ggplot}
+```{r Plot Management Scenarios}
 
+readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))
+
+manage_scen <-  c(histManagement   = "histManagement",
+                  constTillage     = "constTillage-mixedtill", 
+                  constManure      = "constManure-1975",
+                  constResidues    = "constResidues-1975", 
+                  constManagement  = "constManagement-1975")
+
+library(ggplot2)
+library(scales)
+library(gridExtra)
+library(cowplot)
+
+theme_set(theme_bw())
 legend_size <- 10
+scen_cols   <- c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000")[5:1]
+input_cols  <- scen_cols[1:4]
 
-SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission[,seq(1975,2010,5),c(1,plot_rev)]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
+limits <- c(-30,-21)
+breaks <- c(-22, -24,- 26, -28, -30)
+
+hSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),manage_scen]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
   labs( x = "Year", 
-        y = "Global SOC gap in GtC",
+        y = "Global SOC debt\n in GtC",
         color="") +
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits = c(-45,-35),
-                     breaks = c(-36, -38, -40,- 42, -44)) +
+  scale_y_continuous(limits=limits, breaks=breaks) +
+  scale_color_manual(values=scen_cols, labels = names(manage_scen)) +
   theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8))
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+        legend.position="bottom", legend.direction="vertical",
+        legend.margin=margin(-10,-10,0,0),
+        legend.box.margin=margin(-10,-10,0,0))
 
-SOCemis
+hSOCinput <- ggplot(data= as.data.frame(glo_SOCInput[,seq(1975,2010,1),c(3:4,2,1)]),  aes(x = Year, y = Value, group=Data1, fill=Data1)) + geom_area() +
+  labs( x = "Year", 
+        y = "Global SOC input\n in GtC",
+        fill="") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous() +
+  scale_fill_manual(values=input_cols) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+        legend.position="bottom", 
+        legend.direction="vertical",
+        legend.margin=margin(-10,-10,0,0),
+        legend.box.margin=margin(-10,-10,0,0))
 
-ggsave(paste0(outputdir,"/scenario.png"), plot=SOCemis)
+vSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1974,2010,1),manage_scen]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
+  labs( x = "Year", 
+        y = "Global SOC debt in GtC",
+        color="") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous(limits = limits,
+                     breaks = breaks) +
+  scale_color_manual(values=scen_cols, labels = names(manage_scen)) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+        legend.position="right", legend.direction="vertical")
+
+vSOCinput <- ggplot(data= as.data.frame(glo_SOCInput[,seq(1975,2010,1),c(3:4,2,1)]),  aes(x = Year, y = Value, group=Data1, fill=Data1)) + geom_area() +
+  labs( x = "Year", 
+        y = "Global SOC inputs in GtC",
+        fill="") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous() +
+  scale_fill_manual(values=input_cols) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size), legend.direction="vertical")
+
+hSOCemis
+hSOCinput
+vSOCemis
+vSOCinput
+
+hgrid <- plot_grid(hSOCemis, hSOCinput, ncol=2, labels=c("(a)","(b)"),
+                  label_fontface = "plain", label_size = 10, align="hv")
+
+vgrid <- plot_grid(plot_grid(vSOCemis  + theme(legend.position = "none") , 
+                             vSOCinput + theme(legend.position = "none"), ncol=1, labels=c("(a)","(b)"),
+                             label_fontface = "plain", label_size = 10, align="v", axis="tblr"),
+                   plot_grid(get_legend(vSOCemis), get_legend(vSOCinput), ncol=1,
+                             label_fontface = "plain", label_size = 10, align="v"),
+                   align="h")
+
+hgrid
+vgrid
+
+ggsave(paste0(outputdir,"Images/scenario_horiz.png"), plot=hgrid, width = 15, height = 10, dpi = 300, units = "cm")
+ggsave(paste0(outputdir,"Images/scenario_verti.png"), plot=vgrid, width = 12, height = 15, dpi = 300, units = "cm")
 
 ```
+
+```{r Plot Initialization Scenarios}
+
+glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
+
+init_scen   <-  c(`histManagement.Initial-natveg`  = "Initial-natveg", 
+                  `constManagement.Initial-natveg` = "Initial-natveg_constManagement-1975",
+                  `histManagement.Initial-lu`      = "histManagement", 
+                  `constManagement.Initial-lu`     = "constManagement-1975")
+
+glo_SOCDebt <- glo_SOCDebt[,,init_scen]
+getNames(glo_SOCDebt) <- names(init_scen)
+
+library(ggplot2)
+library(scales)
+library(gridExtra)
+library(cowplot)
+
+theme_set(theme_bw())
+legend_size <- 10
+scen_cols   <- c("#785EF0","#FE6100","#FFB000")
+input_cols  <- scen_cols[1:4]
+
+limits <- c(-30,-10)
+breaks <- c(-10, -15,- 20, -25, -30)
+
+initSOCemis <- ggplot() + geom_path(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),]), 
+                                    aes(x = Year, y = Value, group=interaction(Data1,Data2), linetype=Data1, color=Data2)) +
+  labs( x = "Year", 
+        y = "Global SOC debt\n in GtC",
+        color    ="Initalitzation",
+        linetype ="Management") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous(limits=limits, breaks=breaks) +
+  scale_color_manual(values=c(scen_cols)[1:2], labels = unique(gsub("(.*)\\.(.*)","\\2", names(init_scen)))) +
+  scale_linetype_manual(values=c(1,2), labels=unique(gsub("(.*)\\.(.*)","\\1", names(init_scen)))) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+        legend.position="right", legend.direction="vertical")
+
+initSOCemis
+
+ggsave(paste0(outputdir,"Images/scenario_init.png"), plot= initSOCemis, width = 10, height = 6, dpi = 300, units = "cm")
+```
+
+```{r Plot Litter param Scenarios, out.height="6cm", out.width="18cm"}
+
+glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
+glo_SOCStock <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"stock"])
+
+litter_scen <-  c(`LitterPNV-Mean` = "histManagement",
+                  `LitterPNV-grass`= "LitterPNV-PerennialGrasses", 
+                  `LitterPNV-AASS` = "LitterPNV-AASS", 
+                  `LitterPNV-BRCF` = "LitterPNV-BRCF", 
+                  `LitterPNV-MTCF` = "LitterPNV-MTCF", 
+                  `LitterPNV-BLEAF`= "LitterPNV-BLEAF",  
+                  `LitterPNV-TREF` = "LitterPNV-TREF")
+
+library(ggplot2)
+library(scales)
+library(gridExtra)
+library(cowplot)
+
+theme_set(theme_bw())
+legend_size <- 10
+scen_cols   <- c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000")[5:1]
+input_cols  <- scen_cols[1:4]
+
+limits <- c(-30,-18)
+breaks <- c(-18, -22, -26, -30)
+
+litterSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),litter_scen]), 
+                                    aes(x = Year, y = Value, group=Data1, color=Data1, size=Data1)) +
+  labs( x = "Year", 
+        y = "Global SOC debt\n in GtC",
+        color="") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous(limits=limits, breaks=breaks) +
+  scale_color_manual(values=c("red", scen_cols, "blue"), labels = names(litter_scen)) +
+  scale_size_manual(values=c(1,rep(0.5, length(litter_scen)-1)), labels = names(litter_scen)) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+        legend.position = "none")
+
+litterSOCemis
+
+limits <- c(625,705)
+breaks <- c(635, 655, 675, 695)
+
+litterSOCstock <- ggplot() + geom_line(data= as.data.frame(glo_SOCStock[,seq(1975,2010,1),litter_scen]/1000), 
+                                    aes(x = Year, y = Value, group=Data1, color=Data1, size=Data1)) +
+  labs( x = "Year", 
+        y = "Global SOC stock\n in GtC",
+        color="Litter parameterization",
+        size="Litter parameterization") +
+  scale_x_discrete(breaks=seq(1975,2010,5)) +
+  scale_y_continuous(limits=limits, breaks=breaks) +
+  scale_color_manual(values=c("red", scen_cols, "blue"), labels = names(litter_scen)) +
+  scale_size_manual(values=c(1,rep(0.5, length(litter_scen)-1)), labels = names(litter_scen)) +
+  theme(rect=element_blank(),
+        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
+        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
+        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+        legend.direction="vertical",
+        legend.margin=margin(-10,-10,0,0),
+        legend.box.margin=margin(-10,-10,0,0)) +
+  guides(color=guide_legend(), size = guide_legend())
+
+litterSOCstock
+
+littergrid <- plot_grid(litterSOCemis, litterSOCstock + theme(legend.position = "none"), get_legend(litterSOCstock),
+                                  ncol=3, labels=c("(a)","(b)"), axis = "l",
+                                  label_fontface = "plain", label_size = 10, align="v")
+littergrid
+
+ggsave(paste0(outputdir,"Images/scenario_litter.png"), plot=littergrid, width = 18, height = 6, dpi = 300, units = "cm")
+
+```

### 2020-12-10 ResultNotebooks/results_SOCtable.Rmd
mayor update on figures and flowfig

--- 2020-06-16 ResultNotebooks/results_SOCtable.Rmd
@@
 library(mrvalidation)
-library(mrSOCbudget)
+library(mrsoil)
 library(magpiesets)
 
-rev <- "rev0.6_mrSOC"
+rev       <- "rev0.11"
+outputdir <- "Output/"
 
@@
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data", extras="--keep-newer-files ")
+#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
 
-SoilCarbon <- read.magpie("Data/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/Landuse.rds")
+SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))[,1975:2010,]
+Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))[,1975:2010,]
 
@@
 library(stats)
-library(luplot)
 library(reshape2)
 
-koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
+years             <- intersect(findset("past"), getYears(SoilCarbon))
+
+### calculating climate mask
+koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )[,years,]
+koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
 getNames(koeppen)         <- tolower(getNames(koeppen))
-koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
 koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
 climateIPCC       <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
-climateIPCC       <- climateIPCC[,getYears(SoilCarbon),]
-
 rm(koeppen)
 
-SOCDens           <- dimSums(SoilCarbon[,,c("steadystate","actualstate","naturalstate")], dim=3.3) 
+### calculating stock change factors
+SOCDens           <- dimSums(SoilCarbon[,years,c("actualstate","naturalstate")], dim=3.3) 
 SOCDens[,,"naturalstate.crop"] <- SOCDens[,,"actualstate.crop"]
 
-CShare_ipcczone   <- dimSums(climateIPCC * SOCDens[,,"crop"]   * Landuse[,,"crop"], dim=1)/
-                     dimSums(climateIPCC * SOCDens[,,"natveg"] * Landuse[,,"crop"], dim=1)
+CShare_ipcczone   <- dimSums(climateIPCC * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
+                     dimSums(climateIPCC * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)
 
-print(collapseNames(CShare_ipcczone[,findset("past"),]))
+getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))
 
-actualstate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"actualstate"])), 
-                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")
+SCF_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,years,"naturalstate"])), 
+                             Year ~ Data1, value.var="Value", row.names=years), row.names = "Year")
+saveRDS(list(SCF_table = SCF_table), paste0(outputdir, "SFC_climatezone.rds"))
 
-naturalstate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"naturalstate"])), 
-                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")
+knitr::kable(SCF_table, "latex",digits=2, label="naturalstate_table")
 
-steadystate_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,findset("past"),"steadystate"])), 
-                            Year ~ Data1, value.var="Value", row.names=findset("past")), row.names = "Year")
+```
 
-saveRDS(list(naturalstate_table = naturalstate_table,
-             actualstate_table  = actualstate_table, 
-             steadystate_table  = steadystate_table), "Output/SFC_climatezone.rds")
+```{r plot climte map}
 
-#knitr::kable(naturalstate_table, "latex",digits=2, label="naturalstate_table")
-#knitr::kable(actualstate_table, "markdown", digits=2)
+magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
+climate2plot <- dimSums(climateIPCC*new.magpie("GLO",NULL,getNames(climateIPCC),c(1:4)), dim=3)
+climate2plot <- merge(as.data.frame(climate2plot), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+
+library("ggplot2")
+theme_set(theme_bw())
+library("sf")
+library("rnaturalearth")
+library("rnaturalearthdata")
+library(RColorBrewer)
+library(sp)
+library(rworldmap)
+library(oce) 
+library(scales)
+
+legend_size <- 10
+
+map.world <- map_data("world")
+world         <- ne_countries(scale = "medium", returnclass = "sf")
+data(coastlineWorld)  
+coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+
+cols          <- colorRampPalette(brewer.pal(4, "RdYlBu"))(4)
+
+ClimateZones <- ggplot(data = world) + geom_sf() +
+    geom_tile(data = subset(climate2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
+    scale_fill_manual(name = expression("climate zones"), 
+                      values = cols, na.value = "white",
+                      breaks= c(1,2,3,4),
+                      labels = gsub("_", " ", getNames(climateIPCC)),
+                      guide = guide_legend(
+                        title.theme=element_text(size=legend_size),
+                        direction = "horizontal",
+                        title.position = 'top',
+                        # I shift the labels around, the should be placed 
+                        # exactly at the right end of each legend key
+                        title.hjust = 0.5,
+                        label.hjust = 1,
+                        nrow = 1,
+                        byrow = F,
+                        label.position = "bottom",
+                        keyheight = 0.5,
+                        keywidth  = 2
+                      )) +
+    theme(legend.position="bottom",
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          legend.text=element_text(size=legend_size),
+          rect=element_blank(),
+          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
+          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
+          )
+ClimateZones 
+ggsave(paste0(outputdir,"/climatezones.png"), plot=ClimateZones , width = 10, height = 7, dpi = 300, units = "cm")
 ```
@@
 
-CShare_ipcc <- new.magpie("GLO","y2010",outer(c("IPCC2006","IPCC2019"),unique(koepp2IPCC$ipcc_reduced), paste, sep="."))
+MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
+                                                                          unique(koepp2IPCC$ipcc_reduced), paste, sep="."))
 
-CShare_ipcc[,,"IPCC2006.tropical_moist"]  <- "0.48"
-CShare_ipcc[,,"IPCC2006.tropical_dry"]    <- "0.58-0.64" 
-CShare_ipcc[,,"IPCC2006.temperate_moist"] <- "0.69"
-CShare_ipcc[,,"IPCC2006.temperate_dry"]   <- "0.80"
+MShare_ipcc <- add_dimension(MShare_ipcc, dim=3.2, nm="medium.invariant", add="input.year")
+LShare_ipcc <- add_dimension(LShare_ipcc, dim=3.2, nm="low.invariant",    add="input.year")
+HShare_ipcc <- add_dimension(HShare_ipcc, dim=3.2, nm="high.invariant",   add="input.year")
 
-CShare_ipcc[,,"IPCC2019.tropical_moist"]  <- "0.83"
-CShare_ipcc[,,"IPCC2019.tropical_dry"]    <- "0.92" 
-CShare_ipcc[,,"IPCC2019.temperate_moist"] <- "0.69-0.70"
-CShare_ipcc[,,"IPCC2019.temperate_dry"]   <- "0.76-0.77"
+MShare_ipcc[,,"IPCC2006.medium"][,,"tropical_moist"]  <- "0.48"
+MShare_ipcc[,,"IPCC2006.medium"][,,"tropical_dry"]    <- "0.58--0.64" 
+MShare_ipcc[,,"IPCC2006.medium"][,,"temperate_moist"] <- "0.69"
+MShare_ipcc[,,"IPCC2006.medium"][,,"temperate_dry"]   <- "0.80"
+
+MShare_ipcc[,,"IPCC2019.medium"][,,"tropical_moist"]  <- "0.83"
+MShare_ipcc[,,"IPCC2019.medium"][,,"tropical_dry"]    <- "0.92" 
+MShare_ipcc[,,"IPCC2019.medium"][,,"temperate_moist"] <- "0.69--0.70"
+MShare_ipcc[,,"IPCC2019.medium"][,,"temperate_dry"]   <- "0.76--0.77"
+
+#Low input regime factor
+LShare_ipcc[,,"IPCC2006.low"][,,"tropical_moist"]  <- "0.44"
+LShare_ipcc[,,"IPCC2006.low"][,,"tropical_dry"]    <- "0.55--0.61" 
+LShare_ipcc[,,"IPCC2006.low"][,,"temperate_moist"] <- "0.66"
+LShare_ipcc[,,"IPCC2006.low"][,,"temperate_dry"]   <- "0.74"
+
+LShare_ipcc[,,"IPCC2019.low"][,,"tropical_moist"]  <- "0.76"
+LShare_ipcc[,,"IPCC2019.low"][,,"tropical_dry"]    <- "0.87" 
+LShare_ipcc[,,"IPCC2019.low"][,,"temperate_moist"] <- "0.66--0.67"
+LShare_ipcc[,,"IPCC2019.low"][,,"temperate_dry"]   <- "0.70--0.71"
+
+#High input regime factor
+HShare_ipcc[,,"IPCC2006.high"][,,"tropical_moist"]  <- "0.53"
+HShare_ipcc[,,"IPCC2006.high"][,,"tropical_dry"]    <- "0.60--0.67" 
+HShare_ipcc[,,"IPCC2006.high"][,,"temperate_moist"] <- "0.77"
+HShare_ipcc[,,"IPCC2006.high"][,,"temperate_dry"]   <- "0.83"
+
+HShare_ipcc[,,"IPCC2019.high"][,,"tropical_moist"]  <- "0.92"
+HShare_ipcc[,,"IPCC2019.high"][,,"tropical_dry"]    <- "0.96" 
+HShare_ipcc[,,"IPCC2019.high"][,,"temperate_moist"] <- "0.77--0.78"
+HShare_ipcc[,,"IPCC2019.high"][,,"temperate_dry"]   <- "0.79--0.80"
+
+CShare_ipcc <- mbind(LShare_ipcc,MShare_ipcc,HShare_ipcc)
+getNames(CShare_ipcc) <- gsub("_", " ", getNames(CShare_ipcc))
 
 comparison <- mbind(CShare_ipcc, 
-                    round(add_dimension(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),
-                                  add="SOCbudget",nm="model"),2))
+                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),NULL),
+                                  nm="This Study.hist.1990",add="model.input.year"),2),
+                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),NULL),
+                                  nm="This Study.hist.2010",add="model.input.year"),2))
 
-print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+
+comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(comparison)), 
+                           Data1 + Data2 + Data3 ~ Data4, value.var="Value", 
+                            row.names=levels(as.data.frame(comparison)$Data1)))
+
+colnames(comparison_table) <- gsub("\\.", " ", colnames(comparison_table))
+colnames(comparison_table)[1:3] <- c("Source", "Input", "Year")
+
+saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"SFC_comparison.rds"))
 
 ```
+
+```{r kable2image}
+library(kableExtra)
+
+getwd()
+table <- readRDS(paste0(outputdir,"/SFC_comparison.rds"))
+
+save  <- kbl(table$comparison_table, latex = T, digit=2)%>%
+            kable_styling(latex_options = c("scale_down","striped"), full_width = F)%>%
+              row_spec(7:8, bold=TRUE)
+
+save_kable(save, file="Output/Images/TableSCF_comparison.png")
+```

### 2020-12-11 SOCbudgetPaper/appendix.Rmd
shortened verion of appendix

--- 2020-08-14 SOCbudgetPaper/appendix.Rmd
@@
 
-## table on method subsections to functions within R packages {#append:subsection2mrfunctions}
-
-## table on mapping LUH2FAO2MAG {#append:Tableluh2fao2mag}
-
-## kcr2kres mapping {#append:Tablekcr2kres}
+## map on climate zone used for SCF {#append:climatemap}
 
 ## carbon 2 dry matter {#append:Tablec2dm}
-Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
 
@@
 
-## map on climate zone used for SCF {#append:climatemap}
+## regional SOC stock in GtC from different sources {#append:regcompare}
+
+```{r captionregSOCtable, echo=FALSE}
+captionregSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
+```
+
+```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
+```
+
+with 
+Canada, Australia and New Zealand: CAZ; 
+China: CHA; 
+European Union: EUR; 
+India: IND; 
+Japan: JPN; 
+Latin America: LAM;
+Middle East and north Africa: MEA; 
+non-EU member states: NEU;
+other Asia: OAS; reforming countries: REF; Sub-Saharan Africa:
+SSA; United States: USA.
+
+## global SOC debt in GtC for various litter parameterizations {#append:senslitterpnv}

### 2020-12-11 SOCbudgetPaper/introduction.Rmd
mayor update

--- 2020-10-06 SOCbudgetPaper/introduction.Rmd
@@
 
-Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth’s soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and vegetation pools multiple times (Batjes).
-As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref.
-permafrost melting).
-The specific amount of carbon stored in the soil is, however, uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Batjes, 1996; more).
-The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
+Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth’s soil, is the largest terrestrial organic carbon pool.
+It exceeds the carbon in the atmospheric and vegetation pools multiple times ([@batjes_total_1996]).
+Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere ([@friedlingstein_global_2019], [@minasny_soil_2017]).
+The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile ([@batjes_total_1996], [@sanderman_soil_2017]).
+
 Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite).
-Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC  (Lal, 2006).
-Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX).
-While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especially of agricultural activities.
-Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to capture these land management impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm measures.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
-Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
-This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
-We thereby consider change in SOC caused by historical land cover change as well as of different agricultural management practices, including residue recycling, manure amendments, irrigation, and tillage.
-We provide the first global, spatially explicit SOC loss maps that considers detailed agricultural management.
-This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
-We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
-Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
-Finally, we discuss and compare our findings – including their implications for SOC model development – and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018]).
+However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
+Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
+
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of LUC emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
+[@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter ([@lal_world_2001]).
+
+Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions ([@friedlingstein_global_2019]).
+While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
+Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest ([@strassmann_simulating_2008], [@betts_climate_2015]).
+BKMs are designed to estimate LUC realted emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management ([@friedlingstein_global_2019], [@houghton_carbon_2012], [@hansis_relevance_2015]).
+
+Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g.
+[@bondeau_modelling_2007], [@lindeskog_implications_2013]).
+[@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
+They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
+
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.
+growing seasons ([portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g.
+[@pugh_simulated_2015], [@lutz_simulating_2019]).
+
+More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.
+on tillage systems, see [@porwollik_generating_2018], [@prestele_spatially_2018]) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.
+[@iizumi_modeling_2019], [@minoli_modelling_2019]), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+
+<!--Recent studies on soil sequestration for net greenhouse gas removal ([@paustian_quantifying_2019], [@smith_how_2020]) therefore called for 'well-integrated data-model frameworks' including high quality management data to quantify and monitor SOC stock changes.-->
+
+Field-scale models (@del_grosso_simulated_2001], [@coleman_simulating_1997],[@smith_estimating_2010], [@taghizadeh-toosi_c-tool_2014]) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
+
+The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historical agricultural management.
+To achieve this objective we create and provide a comprehensive dataset of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
+Using a scenario analysis, we decompose the contribution of different management activities.
+Our analysis highlights the importance of considering agricultural management for improving estimates of SOC stocks.
+We conclude, that historic agricultural management let to increasing SOC stocks in the most recent decades.

### 2020-12-11 SOCbudgetPaper/discussion.Rmd
small add on

@@
 
-## Including agricultural management data changes the sign of the trend
+## Agricultural management matters
 
-This study provides an analysis of historic SOC stock changes on cropland.
-We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions ($SOC_{natveg}$).
-Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
-Whereas recent modeling estimates of global SOC emissions indicate an ongoing increase of SOC emissions ([@pugh_simulated_2015], [@sanderman_soil_2017]), our study indicates that the global SOC gap is slowly closing due to improved management.
+This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to constant management assumptions.
+This result remains robust to variations of central model parameters and initialization stocks.
+While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to net reduction of the SOC debt.
 
-According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
-Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as [@sanderman_soil_2017] estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and the [@sanderman_soil_2017] results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
+The evaluation of SOC stock gains and losses, however, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+If defining the SOC debt of 1975 the baseline and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink according to our study.
+Annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 
-Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
-[@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decrease in SOC emissions, accounting for only less than 1% change in SOC emissions.
+## SOC debt and SOC drivers in literature
+
+According to [@sanderman_soil_2017] SOC debt since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+Our estimate of 22 GtC in 2010 for cropland debt is higher as [@sanderman_soil_2017] estimations.
+However, there are large uncertainties in modeling SOC at the global scale, and the [@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results, suggesting that our results are in an acceptable range.
+
+Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+[@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
+The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decline in SOC debt (less than 1% change).
 
 Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
-This challenges [@pugh_simulated_2015] findings of only small effects due to improved practices.
+We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
+This is in contrast to [@pugh_simulated_2015] findings of only small effects due to improved practices.
 
-Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g.
-due to improved animal waste management systems) and (3) the adoption of no tillage practices.
+Moreover, our sensitivity analysis indicates that (1) higher returned residue biomass driver by yield increases play the most important role, followed by (2) greater manure applications due to an increasing livestock sector as well as improved animal waste management systems, and (3) the adoption of no tillage practices.
+Unfortunately, disentangling drivers of higher residue and manure inputs remains unsolved within this study.
 
-modeling management effects on the global scale comes however with parametric and structural uncertainties.
+## Limitations and uncertainties
+
+Modeling management effects at the global scale, however, comes with parametric and structural uncertainties.
+High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.
+the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs.
 As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
 [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources.
-Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
-However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass.
-Rather, it is more likely that we are still missing carbon inputs to the soil.
+Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
+Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
+However, with an evaluation of management effects (see Sect.
+\@ref(sec:ipcccompare)), there is no indication for a systematic overrating of residue biomass.
+Rather, it is also likely that we are still missing carbon inputs e.g.
+from cover- and inter-cropping to the soil.
 
-## Modeled management effect in line with default IPCC assumptions
+Another aspect concerns the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
+As shown in Fig.
+\@ref(fig:SOCinit) the SOC debt estimation halves (from ~26 GtC to ~14 GtC) as well as the decrease in SOC debt (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
+[@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
+Our results show that these legacy fluxes are compensated by agricultural management changes even under the strong assumption of the Initial-natveg scenario.
+However, it remains open, if our simple, depth-restricted SOC model itself can represent these fluxes correctly.
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes, that most of the SOC dynamic happens in the topsoil.
+In this regard several aspects are over simplified within our approach.
+Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
+This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
+Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies ([@powlson_limited_2014]) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.
+
+This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
+However, as pointed out by [@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions ([@wang_human-induced_2017]).
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
+Human interventions have led to large changes in SOC stocks there as well ([@sanderman_soil_2017],[@friedlingstein_global_2019]).
+This study does not try to be an comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
+
+## Modeled management effect in line with default IPCC assumptions {#sec:ipcccompare}
+
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) which are based on measurement data for croplands (see Table \@ref(tab:SCFtable)).
+To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
+\@ref(append:climatemap)).
 
 ```{r captionSCFtable, echo=FALSE}
-captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation.
 We compare average values for four IPCC climate zone classifications.
-IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget)."
+IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in [@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019] are compared to 1990 and 2010 results for historic input levels of this study."
 ```
@@
 
-Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006.
-For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
-Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state.
-The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21% of it.
+Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines ([@calvo_buendia_ipcc_2019]).
+Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), leaving our tropical results in between.
+Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially.
+Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
+Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
+\@ref(fig:SOCmaps)(c)) might be not surprising.
 
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$.
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$ as also indicated by the time trend of the SOC debt.
 
-## SOC stocks inline with literature
+## SOC stocks in line with literature
 
-The world's SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(tab:SOCtable)).
 
 ```{r captionSOCtable, echo=FALSE}
-captionSOCtable <- "modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data (LPJmL4, this study) for the year 2010."
+captionSOCtable <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
+This makes it hard to pinpoint a specific year for these SOC estimations.
+In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.
+LPJmL4 "
 ```
@@
 
-The global estimates of SOC stock from this study are on the lower end compared to other modeled results or more data driven estimates.
-Looking at regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.
-Considering that the model was parametrized for croplands, these mismatches are not suprising as the temperature effects on decomposition are fundamentally different for permafrost soils.
-To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
-Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+The global estimates of SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+Looking at regional results in Fig.
+\@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
+To avoid that this bias influences our results, our study focuses on SOC changes on cropland.
+Pristine natural vegetated areas without human land management cancel out in calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.
 
-Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
-For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
-This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain.
-The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
-Especially in less forested areas the natural land representation higher uncertainty may arise from the parametrization assumptions of natural litterfall.
+Additionally, our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural vegetation, which are, however, only estimated in a basic manner.
+For example, we do not differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass plant types.
+This leaves carbon inputs and decay behavior for natural litterfall rather uncertain.
+The absolute values of SOC stocks and debt from land-use change have to be used with caution.
+As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land may arise especially in less forested areas.
 
-Nevertheless, total SOC stocks are in a reasonable range and results for SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
-The SOC gap from changed management is only affected by the natural vegetation representation in the case of land use transitions.
-In this case, former natural land SOC stocks decline.
-However, as pointed out before, due to low cropland expansion rates, land use change-related SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
-We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
-(see appendix). This shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
-
-## Important shortcomings --> might be added as well
-
-The **initialization of SOC stocks** in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start.
-That is inline with IPCC guidelines calling for a run-in period of 5-20 years.
-However, the results are heavily influenced by it as \@ref(append:initcompare) suggested.
-Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore.
-For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010.
-For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010.
-The relative declining trend for the SOC gap between 1975 and 2010 (-12%) is however almost the same as for our historic estimation (-14%).
-According to LUH [@hurtt_harmonization_2020] cropland area has increased over 62% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries.
-Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.
-
-Following the IPCC guidelines this study has limited its focus to the **first 30 cm of the soil profile**.
-In this regard several aspects are over simplified within our approach.
-Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
-This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Consequently, our SOC loss estimates are too high.
-Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
-Following this arguments SOC stocks in croplands might even be underestimated.
-
-Carbon displacement via **leaching and erosion** is neglected in this study.
-As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain.
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+We conducted a sensitivity analysis (Fig.
+\@ref(append:senslitterpnv) based on various plant parameterizations from the Century model (see Sect.
+\@ref(sec:scen_litterpnv)).
+This shows that the general trend of decreasing SOC debt of ~4 GtC is not altered under various estimates for natural SOC stocks.

### 2020-12-11 SOCbudgetPaper/methods.Rmd
some smaller updates in method

@@
 This equation is valid for all grid cells $i$ and all years $t$.
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model (@parton_analysis_1987).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [(@parton_analysis_1987]).
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
@@
 \@ref(sec:livstmanure)); for natural vegetation litterfall (including fine root turnover) ([@schaphoff_lpjml4_2018]) is the only source of carbon inflow to the soil.
-Following the IPCC guidelines (@calvo_buendia_ipcc_2019), carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
+Following the IPCC guidelines [@calvo_buendia_ipcc_2019], carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
 For each component the sum of all carbon input sources is allocated to the respective $SOC$ sub-pools via transfer coefficients.
@@
 5.0A in [@calvo_buendia_ipcc_2019]).
-This, however, will also include naturally occurring changes due to climatic varition over time.
+This, however, will also include naturally occurring changes due to climatic variation over time.
 For our study, we defined the absolute and relative SOC changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
@@
 Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see Table \@ref(append:Tableluh2fao2mag) for more details on the crop group mapping).
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see FAO2LUH2MAG_croptypes.csv in [@mrcommons]).
 By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see Table \@ref(append:Tablekcr2kres)).
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in [@mrcommons]).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets (see [@weindl_livestock_2017]), that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
@@
 The remaining $AGR$ as well as all $BGR$ are expected to be left on the field.
-We limit high returning rates at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
+We limit high returning rates at $10\unit{tC}\unit{ha}^{-1}}$ to correct for outliers.
 
@@
 \@ref(append:TableclossAWMS) for more details).
-We limit high application shares at $10\tfrac{\unit{tC}}{\unit{ha}}$ to correct for outliers.
+We limit high application shares at $10\unit{tC}\unit{ha}^{-1}}$ to correct for outliers.
 
@@
 Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from  [@porwollik_generating_2018].
-This data set is extended to the period of 1975-2010 by combining country-level data on areas under conservation agriculture from [@fao_aquastat_2016] available from 1975 on with LUH2 physical crop areas, applying the methodology of [@porwollik_generating_2018] to identify potential no-tillage grid-cells.
+This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from [@fao_aquastat_2016] and half-degree resolution physical crop areas from [@hurtt_harmonization_2020], applying the methodology of [@porwollik_generating_2018] to identify potential no-tillage grid cells.
 
@@
 Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
+Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines ([@calvo_buendia_ipcc_2019]), but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
 Whereas lg/c ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, n/c ratios range between $0.001-0.05$ with an average of $0.015$ (default value).
@@
   \middlehline
-  \textit{LitterPNV-CenturyAverage} & mix over all tree types & 0.015  & 0.25 \\
-  \textit{LitterPNV-PerennialGrasses} & perennial grasses     & 0.029  & 0.11 \\
+  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
+  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
   \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\

### 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
added all  authors

--- 2020-11-09 SOCbudgetPaper/SOCbudget.Rmd
@@
     affiliation: 1
-  - given_name: Your
-    surname: Name
+  - given_name: Jan Philipp
+    surname: Dietrich
+    affiliation: 1 
+  - given_name: Marta
+    surname: Dondini
     affiliation: 2
+  - given_name: Jens
+    surname: Heinke
+    affiliation: 1
+  - given_name: Matthias
+    surname: Kuhnert
+    affiliation: 2
+  - given_name: Christoph
+    surname: Müller
+    affiliation: 1 
+  - given_name: Susanne
+    surname: Rolinski
+    affiliation: 1
+  - given_name: Pete
+    surname: Smith
+    affiliation: 2
+  - given_name: Isabelle
+    surname: Weindl
+    affiliation: 1    
+  - given_name: Hermann
+    surname: Lotze-Campen
+    affiliation: 1
   - given_name: Alexander
@@
   - code: 1
-    address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany
+    address: Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany
   - code: 2
-    address: Your affiliation 
+    address: Institute of Biological & Environmental Sciences, University of Aberdeen, Aberdeen, UK
+  - code: 3
+    address: Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany
+
 abstract: |
-  Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
-It is three times larger than the vegetation pool and more than twice as big as the athmospheric pool, when looking into the first meter of the earth soil profile.
-Human cropping activties led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
-While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
-Here we create a spatial explicit data set for agricutural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
-We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975-2010.
-We show that, our results on global scale based on Tier 2 IPCC methodolgy are in good agreement with Tier 1 default assumptions.
-We also find that SOC is very sensitive to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.
+  Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
+It is three times larger than the vegetation carbon pool and more than twice as big as the atmospheric carbon pool, when looking into the first meter of the Earth's soil profile.
+SOC has been depleted by anthropogenic land cover change and agricultural management.
+However, the latter has so far not been well represented in global carbon stock assessments.
+While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
+Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
+We estimate that due to arable farming, soils have around 26 GtC relative to a counterfactual natural state in 1975.
+Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
+We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
+
 bibliography: SOCbudget2.bib
@@
   codedata: |
-    Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/. Data used for the output can be found under <something-with-zenodo>.  
+  We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions),  github.com/pik-piam/mrsoil (soil dynamic related functions) and  github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing.
+Model results are accessable under (zenode_link).
+Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
 authorcontribution: |
-  Karstens wrote code and paper build on work of Bodirsky (and <here_could_be_your_names>). 
-  Bodirsky, <here_could_be_your_names> and Popp revised paper.
+  KK wrote code and paper build on work of BLB, IW. 
+  JPD revised and improved the model code. 
+  BB, CM, MK, IW, JH, SR, MD, SM, HLC, AP revised paper.
 disclaimer: |
@@
 acknowledgements: |
-  Thanks to the rticles contributors!
+  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors. 
+appendix: |
+
 output:

### 2020-12-11 SOCbudgetPaper/methods.Rmd
typo

@@
 
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practises are neglected (adoption starts in 1975 according to the available data set); (2) the \textit{constResidues} and the \textif{constManure} scenario assume constant input rates from residues resp.
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set); (2) the \textit{constResidues} and the \textif{constManure} scenario assume constant input rates from residues resp.
 manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards.

### 2020-12-11 SOCbudgetPaper/results.Rmd
update of results

--- 2020-08-14 SOCbudgetPaper/results.Rmd
@@
 
-We present simulation results of the SOC budget of the period from 1975 to 2010 mainly focusing on cropland areas.
-Detailed total global and regional results of the state of the worlds SOC in comparison to other estimates can be found in the supplement.
-
 ## SOC distribution and depletion
-Figure
 
 ```{r caption_SOCmaps, echo=FALSE}
-caption_SOCmaps <- "(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas.
-(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades (despite ongoing cropland expansion).
-(c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
-Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas."
+caption_SOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas.
+(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics.
+Whereas absolute losses ($\DeltaSOC$) are often highest in temperate dry regions, relative losses ($F^\mathrm{SCF}$) are often larger in tropical moist areas.
+(d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+"
 ```
@@
 
-## global agricultural carbon flows and stocks
+In Fig.
+\@ref(fig:SOCmaps)(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
+Values ranging between well over $100\unit{t ha^{-1}}$ in northern temperate croplands, to less than $5\unit{t ha^{-1}}$ for arid and semiarid croplands.
+Our spatially explicit results show hot spots of SOC losses and gains compared to SOC under potential natural vegetation in two different ways: 
+1.
+Absolute SOC changes $\Delta SOC$ (see Fig.
+\@ref(fig:SOCmaps)(b)) indicate areas with high importance for the global SOC losses.
+The might be driven by large relative changes (e.g.
+in Central Africa) or a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.
+North-east Asia).
+In contrast, large net gains of SOC happen primarily in developed countries of the North according to our results.
+2.
+Relative SOC changes measured as stock changes factors $F^{SCF}$ (see Fig.
+\@ref(fig:SOCmaps)(c)) are a helpful tool, to measure the impact of human cropping activities.
+They indicate areas with large difference in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
+Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
+Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have improved, but presumably irrigated croplands at the border to dry, unsuitable areas worldwide as well.
+
+Focusing of the temporal dimension of SOC stock changes (see Fig.
+\@ref(fig:SOCmaps)(c)), global SOC debt has decreased around 15% over the last couple of decades to $22\unit{GtC}$.
+This leads to a sequestration rate of $0.11\unit{GtC yr^{-1}}$ and considering our estimate of the global SOC stock of $675\unit{GtC}$ in 2010 to $0.2$ per mille per year.
+
+## Agricultural management effects on SOC debt
+
+Figure \@ref(fig:SOCscen)(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices.
+Without changes in management regimes, especially in C inflows from crop residues to the soil, global SOC debt on cropland would be increase at a rate of $0.1\unit{GtC yr^{-1}}$.
+The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of $0.06\unit{GtC yr^{-1}}$.
+Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of $0.09\unit{GtC yr^{-1}}$ for both), the effect of no-tillage gets stronger till 2010.
+The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \@ref(fig:SOCscen)(b) as well.
+
+```{r caption_SOCscen, echo=FALSE}
+caption_SOCscen <- "(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+ConstManagement combines all three modifications.
+Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
+Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
+```
+
+```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
+```
+
+The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig.
+\@ref(fig:SOcinit).
+The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
+The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC) as well as the decrease in of $\Delta SOC$ for over the last 35 years (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.
+
+```{r caption_SOCinit, echo=FALSE}
+caption_SOCinit <- "Global $\Delta SOC$ for different SOC initialization choices in the start year 1901."
+```
+
+```{r SOCinit, out.width = "18cm", fig.cap = caption_SOCinit, echo=FALSE, fig.pos="H"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
+```
+ Figure \@ref(fig:)
+
+## Carbon flows in the agricultural system
+
+Figure \@ref(fig:FlowFig) illustrates the carbon flows within the agricultural production system on cropland for the year 2010.
 
 ```{r caption_FlowFig, echo=FALSE}
-caption_FlowFig <- "Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
-Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil)."
+caption_FlowFig <- "Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues."
 ```
@@
 knitr::include_graphics("../ResultNotebooks/Output/Images/OuFlowFig.png")
-```
-
-## comparison of stock changes factors
-
-```{r caption_SCFtable, echo=FALSE}
-caption_SCFtable <- "$F^{SOC}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{SOC}$ by additional 5-8 percent.
-Also: modelled $F^{SOC}$ have increased for all climates over time ."
-```
-
-```{r SCFtable, out.width = "16cm", fig.cap = caption_SCFtable, echo=FALSE, fig.pos="H"}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
 ```
@@
 # ``` -->
-
-\newpage

### 2020-12-11 ResultNotebooks/results_SOCcompare.Rmd
update of results images and tables

@@
 
-SOCStock     <- dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=3)
+SoilCarbon[,,c("naturalstate.crop")] <- SoilCarbon[,,c("naturalstate.natveg")]
+SOCStock     <- dimSums(dimSums(SoilCarbon[,,c("naturalstate","actualstate")], dim=3.3)* Landuse, dim=3.2)
 glo_SOCStock <- dimSums(SOCStock, dim=1) 
@@
 SOCStock <- mbind(glo_SOCStock, reg_SOCStock)
-glo_SOC  <- mbind(setNames(SOCStock[,"y2010",],"SOC Budget"), 
+glo_SOC  <- mbind(setNames(SOCStock[,"y2010","actualstate"],"This Study"),
+                  setNames(SOCStock[,"y2010","naturalstate"],"This Study (PNV)"),
                   setNames(LPJmL4[,"y2010",], "LPJmL4Paper"),

### 2020-12-11 ResultNotebooks/results_SOCmaps.Rmd
update of results images and tables

@@
 
-library("ggplot2")
+library(ggplot2)
 theme_set(theme_bw())
-library("sf")
-library("rnaturalearth")
-library("rnaturalearthdata")
+library(sf)
+library(rnaturalearth)
+library(rnaturalearthdata)
 library(RColorBrewer)
@@
   labs( x = "Year", 
-        y = "Global SOC loss\n in GtC") +
+        y = expression(paste('Global \u0394SOC in GtC'))) +
   scale_x_discrete(breaks=seq(1975,2010,5)) +
@@
         ) +
-  annotate(geom="text", label="SOC debt", size=14, x=factor(1995), y=-20, size = 3) +
+  annotate(geom="text", label="SOC debt", x=factor(1995), y=-20, size = 3) +
   annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -24, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
@@
     scale_fill_stepsn(limits=c(0,2), colours = cols, 
-                      name = expression(" cropland SOC stock change factor (unitless)"), 
+                      name = expression("cropland SOC stock change factor (unitless)"), 
                       breaks=c(0.3,0.6,0.9,1.1,1.4,1.7,2),
@@
     scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression("cropland SOC gap in tons of carbon per hectare"), 
+                      name = expression(paste('croplands \u0394SOC in tons of carbon per hectare')), 
                       values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
@@
 
-grid <- plot_grid(SOCdensity, SOCemis, SOCincrease, SOCshare, ncol=2, labels=c("(a)","(b)", "(c)", "(d)"),
+grid <- plot_grid(SOCdensity,SOCincrease, SOCshare, SOCemis,  ncol=2, labels=c("(a)","(b)", "(c)", "(d)"),
                   label_fontface = "plain", label_size = 10)

### 2020-12-11 ResultNotebooks/results_SOCscenario.Rmd
update of results images and tables

@@
 
-# untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
+# untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"), 
+#       exdir="Data/histManagement", extras="--keep-newer-files ")
 # 
-for(item in scenarios[10:11]){
-   untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-   files=filesNeeded, exdir=paste0("Data/", item), extras="--keep-newer-files ")
-}
+# for(item in scenarios){
+#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+#    files=filesNeeded, exdir=paste0("Data/", item), extras="--keep-newer-files ")
+# }
 
@@
 CarbonInput  <- read.magpie(paste0("Data/histManagement/CarbonInput.rds"))
+Landuse      <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
 glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))
@@
 
-readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))
+glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
 
@@
   labs( x = "Year", 
-        y = "Global SOC debt\n in GtC",
+        y = expression(paste('Global \u0394SOC in GtC')),
         color="") +
@@
   labs( x = "Year", 
-        y = "Global SOC debt in GtC",
+        y = expression(paste('Global \u0394SOC in GtC')),
         color="") +
@@
   labs( x = "Year", 
-        y = "Global SOC debt\n in GtC",
+        y = expression(paste('Global \u0394SOC in GtC')),
         color    ="Initalitzation",

### 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
update text and style files

--- 2020-11-10 SOCbudgetPaper/SOCbudget.Rmd
@@
   Thanks to the rticles contributors!
+appendix:
 output:
+  rticles::copernicus_article: default
+  md_document:
+    variant: markdown_github
+  bookdown::word_document2: default
   bookdown::pdf_book:
@@
     md_extensions: +raw_attribute
-  md_document:
-    variant: markdown_github
-  bookdown::word_document2: default
-  rticles::copernicus_article: default
   bookdown::pdf_document2: default
@@
 ```
-\newpage

### 2020-12-11 SOCbudgetPaper/appendix.Rmd
update text and style files

--- 2020-11-09 SOCbudgetPaper/appendix.Rmd
@@
 
-```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
+```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 
-```{r initSOC, out.width = "100%", fig.cap = captioninitSOC, echo=FALSE, fig.pos="H"}
+```{r initSOC, out.width = "100%", fig.cap = captioninitSOC, echo=FALSE, fig.pos="h"}
 # All defaults

### 2020-12-11 SOCbudgetPaper/discussion.Rmd
update text and style files

--- 2020-11-09 SOCbudgetPaper/discussion.Rmd
@@
 According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
-Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as [@sanderman_soil_2017] estimations.
-However, there are large uncertainties in modelling SOC at the global scale, and the [@sanderman_soil_2017] results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
+It was also pointed out that these results might be conservatively low compared to experimental results.
+Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with [@sanderman_soil_2017] estimates.
 
-Furthermore, [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
+Furthermore, the results of [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
 [@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decrease in SOC emissions, accounting for only less than 1% change in SOC emissions.
+The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1% change in SOC emissions.
 
@@
 We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
-This challenges [@pugh_simulated_2015] findings of only small effects due to improved practices.
-
-Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g.
-due to improved animal waste management systems) and (3) the adoption of no tillage practices.
+This is challenging [@pugh_simulated_2015] findings of only small effects due to improved practices.
+Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g.
+due to improved animal waste management system) and (3) the adoption of no tillage practices.
 
@@
 As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
-[@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources.
-Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
-However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass.
-Rather, it is more likely that we are still missing carbon inputs to the soil.
+[@keel_large_2017] questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources.
+Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding.
+However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.
 
@@
 To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
+To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.
 
 ```{r captionSCFtable, echo=FALSE}
-captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
+captionSCFtable <- "This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
 We compare average values for four IPCC climate zone classifications.
@@
 For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
-Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state.
-The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21% of it.
+Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state.
+The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.
 
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$.
+Looking again on the time trend our study shows the substantial impact of changed management on the development of $F_{SCF}$.
 
@@
 
-The world's SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
+The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).
 
@@
 
-```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
+```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 
-The global estimates of SOC stock from this study are on the lower end compared to other modelled results or more data driven estimates.
-Looking at regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.
-Considering that the model was parametrized for croplands, these mismatches are not suprising as the temperature effects on decomposition are fundamentally different for permafrost soils.
+The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates.
+Looking on regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas.
+Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils.
 To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
-Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
+Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
 
-Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
+Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
 For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
@@
 The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
-Especially in less forested areas the natural land representation higher uncertainty may arise from the parametrization assumptions of natural litterfall.
+Especially in less forested areas the natural land representation might be off, due to parametrization assumptions of the natural litterfall.
 
-Nevertheless, total SOC stocks are in a reasonable range and results for SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
-The SOC gap from changed management is only affected by the natural vegetation representation in the case of land use transitions.
-In this case, former natural land SOC stocks decline.
-However, as pointed out before, due to low cropland expansion rates, land use change-related SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
+Nevertheless, total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
+Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions.
+Then former natural land SOC stocks are declining to lower levels.
+However, as pointed out before, due to low cropland expansion rates, their SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
 We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
-(see appendix). This shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
+(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
 
-## Important shortcomings --> might be added as well
+## Important short comings --> might be added as well
 
-The **initialization of SOC stocks** in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start.
+The initialization of SOC stock in 1960 is assumed to be in steady state followed by a run-in period of 15 years, till the analysis of the results start.
 That is inline with IPCC guidelines calling for a run-in period of 5-20 years.
-However, the results are heavily influenced by it as \@ref(append:initcompare) suggested.
-Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore.
-For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010.
-For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010.
-The relative declining trend for the SOC gap between 1975 and 2010 (-12%) is however almost the same as for our historic estimation (-14%).
-According to LUH [@hurtt_harmonization_2020] cropland area has increased over 62% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries.
-Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.
+However, the results in the early years might be still distorted by a longer living memory effects than this 15 years of spin-up.
 
-Following the IPCC guidelines this study has limited its focus to the **first 30 cm of the soil profile**.
+Following the IPCC guidelines this study has limited its focus to the first 30 cm of the soil profile.
 In this regard several aspects are over simplified within our approach.
@@
 It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
-Following this arguments SOC stocks in croplands might even be underestimated.
+Following this argumentation SOC stocks in croplands might even be underestimated.
 
-Carbon displacement via **leaching and erosion** is neglected in this study.
-As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain.
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+Carbon displacement via leaching and erosion is neglected in this study.
+As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertian.
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach on SOC focusing especialy on emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.

### 2020-12-11 SOCbudgetPaper/methods.Rmd
update text and style files

--- 2020-12-07 SOCbudgetPaper/methods.Rmd
@@
 
-$SOC$ converge towards the calculated steady-state stock $SOC^{\mathrm{eq}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like
+$SOC$ converge towards the calculated steady-state stock $SOC^{\mathrm{eq}}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like
 
@@
 \@ref(append:climatemap)).
-The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}$ by
+The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}}$ by
 
@@
 
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practises are neglected (adoption starts in 1975 according to the available data set); (2) the \textit{constResidues} and the \textif{constManure} scenario assume constant input rates from residues resp.
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practises are neglected (adoption starts in 1975 according to the available data set); (2) the \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
 manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards.
@@
  \caption{Scenarios for sensitivity analysis on litter parameterization}
- \begin{tabular}{l l l}
+ \begin{tabular}{l l l l}
  \tophline
@@
  \belowtable{}
- (\#tab:scen_litterpnv)
+ (\#tab:scenlitterpnv)
  \end{table*}

### 2020-12-11 SOCbudgetPaper/results.Rmd
update text and style files

--- 2020-11-09 SOCbudgetPaper/results.Rmd
@@
 
-```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="H"}
+```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 
-```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="H"}
+```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 
-```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="H"}
+```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
 # All defaults

### 2020-12-11 SOCbudgetPaper/introduction.Rmd

@@
 
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of LUC emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
 [@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter ([@lal_world_2001]).

### 2020-12-11 SOCbudgetPaper/methods.Rmd

--- 2020-12-11 SOCbudgetPaper/methods.Rmd
@@
 
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set); (2) the \textit{constResidues} and the \textif{constManure} scenario assume constant input rates from residues resp.
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set); (2) the \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
 manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards.

### 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
compliling version

--- 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
@@
     address: Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany
-
 abstract: |
@@
 We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
-
-bibliography: SOCbudget2.bib
+bibliography: SOCbudget.bib
 running:
@@
   Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors. 
-appendix: |
-
 output:

### 2020-12-11 SOCbudgetPaper/SOCbudget.tex
compliling version

@@
 %% 2-column papers and discussion papers
-\RequirePackage[2020/02/02]{latexrelease}
 \documentclass[gc, manuscript]{copernicus}
@@
 \Author[1]{Benjamin Leon}{Bodirsky}
-\Author[2]{Your}{Name}
+\Author[1]{Jan Philipp}{Dietrich}
+\Author[2]{Marta}{Dondini}
+\Author[1]{Jens}{Heinke}
+\Author[2]{Matthias}{Kuhnert}
+\Author[1]{Christoph}{Müller}
+\Author[1]{Susanne}{Rolinski}
+\Author[2]{Pete}{Smith}
+\Author[1]{Isabelle}{Weindl}
+\Author[1]{Hermann}{Lotze-Campen}
 \Author[1]{Alexander}{Popp}
 
-\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
-\affil[2]{Your affiliation}
+\affil[1]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany}
+\affil[2]{Institute of Biological \& Environmental Sciences, University of Aberdeen, Aberdeen, UK}
+\affil[3]{Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany}
 
@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of larges carbon stocks on Earth.
-It is three times larger than the vegetation pool and more than twice as big as the atmospheric pool, when looking into the first meter of the earth soil profile.
-Human cropping activities led and still lead to a depletion of SOC though, which are so far not well represented in global assessments of historic carbon emissions.
-While SOC models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated on the global scale.
-Here we create a spatial explicit data set for agricultural management on cropland, especially for crop residue and manure management, based on global historic production (FAOSTAT), and land-use (LUH2) data; and combine it with the IPCC Tier 2 approach to create a half degree resolution SOC stock changes on mineral soils.
-We estimate that due to arable farming soils have lost over 37 GtOC of which 4 GtOC have been regained within the period 1975--2010.
-We show that, our results on global scale based on Tier 2 IPCC methodology are in good agreement with Tier 1 default assumptions.
-We also find that SOC is very sensitive to management decision such as residue recycling indicating the necessity to incorporated better management data in soil models.
+Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
+It is three times larger than the vegetation carbon pool and more than twice as big as the atmospheric carbon pool, when looking into the first meter of the Earth's soil profile.
+SOC has been depleted by anthropogenic land cover change and agricultural management.
+However, the latter has so far not been well represented in global carbon stock assessments.
+While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
+Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
+We estimate that due to arable farming, soils have around 26 GtC relative to a counterfactual natural state in 1975.
+Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
+We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
 \end{abstract}
@@
 
-Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth's soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and vegetation pools multiple times (Batjes).
-As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref.
-permafrost melting).
-The specific amount of carbon stored in the soil is, however, uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Batjes, 1996; more).
-The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools.
+Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool.
+It exceeds the carbon in the atmospheric and vegetation pools multiple times (\citep{batjes_total_1996}).
+Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere (\citep{friedlingstein_global_2019}, \citep{minasny_soil_2017}).
+The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (\citep{batjes_total_1996}, \citep{sanderman_soil_2017}).
+
 Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools over much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite).
-Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC (Lal, 2006).
-Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX).
-While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especially of agricultural activities.
-Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to capture these land management impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm measures.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge.
-Our study combines a spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands.
-This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system.
-We thereby consider change in SOC caused by historical land cover change as well as of different agricultural management practices, including residue recycling, manure amendments, irrigation, and tillage.
-We provide the first global, spatially explicit SOC loss maps that considers detailed agricultural management.
-This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study.
-Additionally, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines.
-Lastly, we continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.
-In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects.
-Finally, we discuss and compare our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
+Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}).
+However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
+Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
+
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
+\citep{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter (\citep{lal_world_2001}).
+
+Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions (\citep{friedlingstein_global_2019}).
+While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
+Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest (\citep{strassmann_simulating_2008}, \citep{betts_climate_2015}).
+BKMs are designed to estimate LUC realted emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management (\citep{friedlingstein_global_2019}, \citep{houghton_carbon_2012}, \citep{hansis_relevance_2015}).
+
+Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g.~\citep{bondeau_modelling_2007}, \citep{lindeskog_implications_2013}).
+\citep{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
+They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
+
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons ({[}portmann\_mirca2000global\_2010{]}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).
+
+More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+
+Field-scale models (\citet{del_grosso_simulated_2001}{]}, \citep{coleman_simulating_1997},\citep{smith_estimating_2010}, \citep{taghizadeh-toosi_c-tool_2014}) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
+
+The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historical agricultural management.
+To achieve this objective we create and provide a comprehensive dataset of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
+Using a scenario analysis, we decompose the contribution of different management activities.
+Our analysis highlights the importance of considering agricultural management for improving estimates of SOC stocks.
+We conclude, that historic agricultural management let to increasing SOC stocks in the most recent decades.
 \newpage
 
-\hypertarget{method}{%
-\section{Method}\label{method}}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
-We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on the MADRaT package (package citation?), a framework which aims to improve reproducibility and transparency in data processing.
-In the following section we outline the most important relationships and assumptions.
-Table \ref{append:subsection2mrfunctions} provides further information on corresponding code within the R packages.
+In Sect.
+\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}).
+We additionally describe how we configured and extended the steady-state method (for model code see \citet{mrsoil_2020}).
+In Sect.
+\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
+Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citet{mrcommons_2020}).
+In Sect.
+\ref{sec:scenarios} we define the scenarios used to complement our historic model results.
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{SOC Stocks and Stock changes following the Tier 2 steady-state method}\label{sec:carbonbudget}}
+\subsection{SOC stocks and stock changes following the steady-state method}\label{sec:carbonbudget}}
 
-Following the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}; short Tier 2 steady-state method), we estimate soil organic carbon (SOC) stocks and their change over time for cropland on half-degree resolution from 1975 to 2010.
-We assume the current SOC state converges towards a steady state, which itself is depending on biophysical, climatic and agronomic conditions.
-Therefore, we conduct the following three steps within each yearly time-step:
-(1) We calculate annual land-use (sub-)type-specific steady states and decay rates for SOC stocks,
-(2) we account for land conversion by transferring SOC from and to another land-use type represented as natural vegetation,
-(3) we estimate SOC stocks and changes based on the stocks of the previous time period, the steady state stocks and the decay rate.
+Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+We restrict our analysis to the first 0-30 cm of the soil profile.
+Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
+Therefore, we take the following three steps for each year of our simulation period:
+(1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
+(2) we account for land conversion by transferring \(SOC\) from and natural vegetation (Sect. \ref{sec:carbontransfer}),
+(3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
+To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \ref{sec:initsoc}).
 
-\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
-\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}
+\hypertarget{sec:steadystates}{%
+\subsubsection{Steady-state SOC stocks and decay rates}\label{sec:steadystates}}
 
-In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
+In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{\mathrm{eq}}\) are given by
 \begin{equation}
-SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
+SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
 \label{eq:inoutflow}
 \end{equation}
-with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cell identifier and \(t\) years.
-We use for our calculations the Tier 2 steady-state method, which assumes three soil carbon sub-pools (active, slow and passive) and interactions between them, following the approach in the Century model (Parton et al, 1988).
-Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are the key components to determining steady-state SOC stocks.
+with \(C^{\textrm{in}}\) being the carbon inputs to the soil, \(k\) denotes the soil organic carbon decay rate.
+This equation is valid for all grid cells \(i\) and all years \(t\).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep[(][]{parton_analysis_1987}).
+Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
+We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
 
-\textbf{Carbon Inputs to the Soil}
-
-We account for different carbon input sources depending on the two land-use types we distinguish: croplands and natural vegetated land as representative for all other land use (see table \ref{tab:datasourceinputs}).
-Carbon sources for cropland are recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see \ref{sec:livstmanure}); for natural vegetation litterfall (including fine root turnover) \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
-
-Following the IPCC carbon accounting methodology, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citet{ipcc_2019_2019}).
-For each component the sum of all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients.
-This implies that both the amount of carbon and its structural composition determine the effective inflow.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.
+Carbon inputs for cropland are below and above ground residues left or returned to the field (see Sect.
+\ref{sec:residues}) and manure inputs (see Sect.
+\ref{sec:livstmanure}); for natural vegetation litterfall (including fine root turnover) (\citep{schaphoff_lpjml4_2018}) is the only source of carbon inflow to the soil.
+Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
+For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients.
+This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \ref{tab:datasourceinputs}.
 
   \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types }
- \begin{tabular}{l l l l}
+  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
+  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.25\textwidth}}
   \tophline
@@
   \middlehline
- \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
-                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
-                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
+  \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+                                                                      \cite{faostat_faostat_2016}, \\
+                                                                      \cite{schaphoff_lpjml4_2018} (LPJmL),\\
+                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
+                                                      \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+                                                                      lg/c generic values according to Table 5.5B\\
+                                                                      in \cite{calvo_buendia_ipcc_2019},\\
+                                                                      crop-type specific n/c-ratios from
+                                                                      \cite{dietrich_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
+                            & below ground residues &  &  \\ \cline{2-2}
+                            & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
                             \hline
-  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
+  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018} (LPJmL) 
+                            & \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut 
+                                \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
  \bottomhline
@@
 
-\textbf{SOC decay}
-
-The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i).
-Following the Tier 2 steady-state method we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates.
-Thus, \(k_{sub}\) is given by
+The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space (\(i\)) and time (\(t\)).
+Following the steady-state method \citep{calvo_buendia_ipcc_2019}, we consider temperature (\(temp\)), water (\(wat\)), sand fraction (\(sf\)) and tillage (\(till\)) effects to account for spatial variation of decay rates.
+Thus, \(k_{sub}\) are given by
 
@@
 \begin{aligned}
-& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
-& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
-& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
+& k_{i,t,\mathrm{active},lu}  & = &~ k_{\mathrm{active}}  ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ & \cdot~ sf_{i}\\
+& k_{i,t,\mathrm{slow},lu}    & = &~ k_{\mathrm{slow}}    ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ &\\
+& k_{i,t,\mathrm{passive},lu} & = &~ k_{\mathrm{passive}} ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ & ~ & 
 \label{eq:decayrates}
-\end{aligned}
+\end{aligned}.
 \end{equation}
 
-For cropland, we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas for natural vegetation, we assume rainfed and non-tilled conditions.
-Data sources as well as drivers considered for each land-use types are shown in table \ref{tab:datasourcedecay}.
-To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.
+For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect.
+\ref{sec:tillage}) and irrigation (see Sect.
+\ref{sec:irrigation}) practices on decay rates.
+We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in \citep{calvo_buendia_ipcc_2019}.
 
@@
  \middlehline
- \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm &  [SoilGrids]  \\
+ \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
+                                     & \cite{hengl_soilgrids250m:_2017} \\
                       \cline{2-4}
 
-                      & Mircobial activity & air temperature & [CRUp4.0] \\
+                      & Mircobial activity & air temperature & \cite{harris_version_2020} \\
                       \cline{2-4}
 
-                      & Water restriction & precipitation \& potential evapotranspiration & [CRUp4.0] \\
+                      & Water restriction & precipitation \& potential evapotranspiration & \cite{harris_version_2020} \\
                       \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & [sec:irrigation] \\ 
+\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
                       \cline{2-4}
 
-                      & Soil disturbance & tillage & [sec:tillage] \\
+                      & Soil disturbance & tillage & Sect. \ref{sec:irrigation} \\
  \bottomhline
@@
 
-\hypertarget{soc-transfer-between-land-use-types}{%
-\subsubsection{SOC transfer between land-use types}\label{soc-transfer-between-land-use-types}}
+\hypertarget{sec:carbontransfer}{%
+\subsubsection{SOC transfer between land-use types}\label{sec:carbontransfer}}
 
-We calculate SOC stocks based on the area shares of land-use types (lut) within the half-degree grid cells (i).
-If land is converted from one land-use type into others (!lut), the respective share of the SOC stocks is reallocated.
-We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via
+We calculate \(SOC\) stocks based on the area shares of land-use types (\(lu\)) within our grid cells (\(i\)).
+If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the \(SOC\) is reallocated.
+We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{t^*}\) via
 
 \begin{equation}
-SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall\quad sub, i  
+SOC_{i,t^*,sub,lu} = SOC_{i,t-1,sub,lu} - \frac{SOC_{i,t-1,sub,lu}}{A_{i,t-1,lu}} \cdot  AR_{i,t,lu} + \frac{SOC_{i,t-1,sub,!lu}}{A_{i,t-1,!lu}} \cdot  AE_{i,t,lu}
 \label{eq:ctransfer}
@@
 
-with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land-use type \(lut\).
-Note that \(!lut\) denotes the sum over all other land-use types, which decreases in the specific time step \(t\).
-Data sources and methodology on land-use states and changes are described in \ref{sec:landuse}.
+with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) and \(AE_{lu}\) the area reduction resp.
+area expansion of the two land-use types.
+Data sources and methodology on land-use states and changes are described in Sect.
+\ref{sec:landuse}.
 
-\hypertarget{total-soc-stocks-and-stock-changes}{%
-\subsubsection{Total SOC stocks and stock changes}\label{total-soc-stocks-and-stock-changes}}
+\hypertarget{sec:totalsoc}{%
+\subsubsection{Total SOC stocks and stock changes}\label{sec:totalsoc}}
 
-Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use type (lut), each sub-pool (sub) and each annual time step (t) like
+\(SOC\) converge towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
 
 \begin{equation}
-SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
+SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t-1,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
 \label{eq:SOCstate}
@@
 
+Note that the decay rates have to be multiplied by one year (\(1a\)) to form a dimensionless factor.
 Reformulating this equation, we obtain a mass balance equation as follows
@@
 \begin{equation}
-SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation (1))}}  \qquad \forall\quad lut, sub, i.
+SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} - \underbrace{SOC_{i,t-1,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
 \label{eq:steadystate2budget}
@@
 
-The global SOC stock for each time step can then be calculated via
+The global \(SOC\) stock for each time step \(t\) can then be calculated via
 
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ -- total SOC stock within cell}}
+SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
 \label{eq:totalstock}
@@
 
-According to the IPCC guidelines SOC changes can be expressed as the difference of two consecutive years (see Eq.
-5.0A in \citep{ipcc_2019_2019}).
-This however, will also include naturally occurring changes due to climatic variance over time.
-For our study we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC_{pnv}\) under the same climatic conditions at time \(t\) in grid cell \(i\), that is based on the natural vegetation SOC calculations as defined above without accounting for land conversion from cropland at any time.
-The absolute changes \(\Delta SOC\) and relative changes \(F^{SCF}\) a thus given by
+According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years (see Eq.
+5.0A in \citep{calvo_buendia_ipcc_2019}).
+This, however, will also include naturally occurring changes due to climatic variation over time.
+For our study, we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
+The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by
 
 \begin{equation}
-\Delta SOC_{t, i} = SOC_{t, i} - SOC_{pnv, t, i}\quad, \qquad \quad  F^{SCF}_{t, i} = \frac{SOC_{t, i}}{SOC_{pnv, t, i}}.
+\Delta SOC_{i,t} = SOC_{i,t} - SOC^{\mathrm{pnv}}_{i,t}\qquad \text{and} \qquad  F^{\mathrm{SCF}}_{i,t} = \frac{SOC_{i,t}}{SOC^{\mathrm{pnv}}_{i,t}} .
 \label{eq:stockdiff}
@@
 
-Note that the absolution changes \(\Delta SOC\) can be also interpreted as the SOC gap due to human cropping activities; whereas relative changes \(F^{SCF}\) can be denoted as stock change factors as defined within the IPCC guidelines of 2006 \citep{ipcc_2006_2006}.
-Note that the SOC gap is equivalent to the negated cumulative SOC emission (\citep{pugh_simulated_2015}) also referred to as SOC debt (\citep{sanderman_soil_2017}) of human land use.
+Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_hs_ipcc_2006}.
+Note that the SOC debt is equivalent to the negated cumulative SOC emission (\citep{pugh_simulated_2015}) of human land use.
 
-\hypertarget{initialization-of-soc-pools}{%
-\subsubsection{Initialization of SOC pools}\label{initialization-of-soc-pools}}
+\hypertarget{sec:initsoc}{%
+\subsubsection{Initialization of SOC pools}\label{sec:initsoc}}
 
-To initialize all SOC sub-pools we assume that cropped land as well as natural vegetation are in a steady state for the specific configuration present within the initialization year 1961.
-We assume after a spin-up period of 15 years reliable results start from 1975 and improve over time, since dependency on the initial conditions decreases.
+To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
+While land-use, climate and litter input information are available till 1901, agricultural management data on residue and manure inputs are hold constant at the level of 1965.
+After a spin-up period of 64 years with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975.
+Note that irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
 
 \hypertarget{sec:tier1}{%
-\subsection{SOC Stocks and Stock changes following Tier 1}\label{sec:tier1}}
+\subsection{SOC stocks and stock changes following Tier 1}\label{sec:tier1}}
 
-Additionally, to the tier 2 approach of the refinement of IPCC guidelines vol.~4 \citep{ipcc_2019_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}).
-Here, stocks are calculated via stock change factors (\(F^{SCF}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-\(F^{SCF}\) are differentiated by different crop, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover, \(F^{SCF}\) vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}).
-The actual SOC stocks as thus calculated based on a given reference SOC stock by
+Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
+Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig.
+\ref{append:climatemap}).
+The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by
 
 \begin{equation}
-SOC^{\text{target}}_{i} = \sum_{c,m} T_{c,i} \cdot SOC^{\text{ref}} \cdot F^{SCF}_{c,m} \qquad\forall\quad t,
+SOC_{i,t} = \sum_{c,m} T_{c,i} \cdot SOC^{\mathrm{ref}}_{i,t} \cdot F^{\mathrm{SCF}}_{c,m}
 \label{eq:tier1}
@@
 with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis we use the default \(F^{SCF}\) from the Tier 1 method of \citep{ipcc_2006_2006}, and \citep{ipcc_2019_2019} as a comparison and consistency check for our more detailed budget approach.
-
-\newpage
+For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_hs_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
 \hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data on 0.5 degree resolution}\label{sec:agrimanagement}}
+\subsection{Agricultural management data at 0.5 degree resolution}\label{sec:agrimanagement}}
 
 We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5 year time steps from 1965 to 2010, which, together with the spin up phase, restricts our analysis to time span from 1975 to 2010.
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and hold them constant before 1965.
+The data is prepared in 5-year time steps from 1965 to 2010, which, together with the spin-up phase, restricts our analysis to the time span from 1975 to 2010.
+For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
 
 \hypertarget{sec:landuse}{%
-\subsubsection{Land use and Land-use change}\label{sec:landuse}}
+\subsubsection{Land use and land-use change}\label{sec:landuse}}
 
-Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we aggregate from quarter degree to half degree resolution.
-We disaggregate the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country and year specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see \ref{append:Tableluh2fao2mag} for more details on the crop group mapping).
-Land-use transitions are calculated as net area differences of the land-use data on half-degree.
+Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{mrcommons}).
+By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
+Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
 
 \hypertarget{sec:residues}{%
-\subsubsection{Crop and Crop Residues Production}\label{sec:residues}}
+\subsubsection{Crop and crop residues production}\label{sec:residues}}
 
-\textbf{Crop Production}
-
-Using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns.
+Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see Sect.
+\ref{sec:landuse}).
 We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
-Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.
-
-\textbf{Crop Residue Production}
+Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallow land.
 
@@
 \begin{aligned}
-AGR & = CP \cdot HI_{\textrm{slope}} + CA \cdot HI_{\textrm{intercept}}\qquad & \textrm{and} \\
-BGR & = (CP + AGR) \cdot RS \qquad                                            & \forall\quad cg, i, t.
+AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
+\qquad & \textrm{and} \\
+BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
 \label{eq:resbiomass}
@@
 
-Following the IPCC guidelines, we split the harvest index into a yield and an area dependent fraction \citep{ipcc_2006_2006}.
-Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased amounts on fallow land.
-We assume that all BGR are recycled to the soil, whereas AGR can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
+Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_hs_ipcc_2006}.
+Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
 
-\textbf{Burned Residues}
+A fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
+Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_hs_ipcc_2006}.
 
-A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country.
-Following \citep{smil_nitrogen_1999}) we assume 25\% burn share for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process \citep{ipcc_2006_2006}.
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{mrcommons}).
+Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
+We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
+For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above.
+The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field.
+We limit high returning rates at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
 
-\textbf{Residue Usage}
-
-We compile out of our 17 crop groups, three residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residue with no double use (see \ref{append:Tablekcr2kres}).
-Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed baskets (see \citep{weindl_livestock_2017}), taking available AGR biomass as well as livestock productivity into account.
-We estimate, for low-income countries, a material-use share for straw residues of 5\% and a fuel-share of 10\% for all used residues groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumed, and middle-income countries use shares are linearly interpolated based on per-capita income.
-The remaining AGR as well as all BGR are expected to be recycled to the soil.
-We limit high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
-
-\textbf{Dry Matter to Carbon Transformation}
-
-To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see \ref{append:Tablec2dm}).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{append:Tablec2dm}).
 
 \hypertarget{sec:livstmanure}{%
-\subsubsection{Livestock Distribution and Manure Excretion}\label{sec:livstmanure}}
+\subsubsection{Livestock distribution and manure excretion}\label{sec:livstmanure}}
 
-We assume that manure is applied at its excretion place, leaving the livestock distribution the driving factor for the spatial pattern of manuring.
+Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems at the present day a lot of the manure has to be stored and can be applied on croplands.
+We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
 
-\textbf{Livestock Distribution}
-To disaggregate country level FAOSTAT livestock production values to the half-degree scale, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014}, and uses feed basket assumptions based on a revised methodology from \citep{weindl_livestock_2017}.
+To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from \citep{weindl_livestock_2017}.
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-For ruminants, we assume that livestock is located where the production of feed occurs.
-We distinguish between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feeds, which we consider to be consumed in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to divide between intensive and extensive production systems.
-For low-income countries, we assume extensive production systems.
-We locate them according to the share of built-up areas based on the idea that these animals are held in subsistence or small-holder farming systems with a high labor per animal ratio.
-Intensive production associated with higher income countries, is distributed within a country using again share of crop-based feed production, assuming that feed availability is the most driving factor for livestock location.
+Due to great feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
+We distinguish between grazed pasture, which is converted into livestock products in extensive systems, and primary-crop feed stuff, which we consider to be consumed in intensive systems.
+For poultry, egg and monogastric meat production we use the per-capita income of the country to distinguish between intensive and extensive production systems.
+For low-income countries, we assume only extensive production systems.
+We locate them according to the share of built-up areas based on the asumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
+Intensive production associated with high-income countries, is distributed within a country using the share of primary-crop production, assuming that feed availability is the most determining factor for livestock location.
+For middle-income countries we split the livestock production into extensive and intensive systems based on the per-capita income.
 
-\textbf{Manure Excretion, Storage and Recycling}
 Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
-Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep{ipcc_2019_2019}).
+Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
+Carbon in excreted manure is estimated by applying fixed c/n ratios (given by \citep{calvo_buendia_ipcc_2019}).
 Depending on the feed system we assume manure to be handled in four different ways:
-All manure originated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Manure fuel shares are estimated using IPCC default values \citep{ipcc_2006_2006}.
+All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
 Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items we assume the manure to be stored in animal waste management systems associated with livestock housing.
-To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled.
-For manure stored in different animal waste management systems we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see \ref{append:TableclossAWMS}).
-We limit high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to correct for outliers.
+For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
+To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems.
+Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
+For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see Sect.
+\ref{append:TableclossAWMS} for more details).
+We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
 
@@
 
-We use irrigation area shares to modify the water effect on decomposition by weighting the irrigated and rainfed water factors based on these shares.
 The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for their cropland subcategories.
-We apply aggregated area shares, leading to the same impact of irrigation on all of our crop groups.
-Moreover, we assume the irrigation effect to be present for all 12 months of a year in grid cells, that has been marked as irrigated.
+We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient \(wat\) on decay rates using these shares to compute the weighted mean between rainfed and irrigated \(wat\) factors.
+As a result \(wat\) is the same for all crop groups within a grid cell.
+Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a (partly) irrigated grid cells, since we do not have crop group specific growing periods at hand.
+In addition, we expect, either water limitations to be no problem during the off-season (causing our assumption to be a good approximation), or irrigated growing periods to span over the whole year in water-restricted areas.
 
@@
 
-To account for the distribution of tillage to the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas as default annual crops are under full and perennials under reduced tillage.
+In order to derive a spatial distribution of the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
 Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from \citep{porwollik_generating_2018}.
-This data set is extended by to the period of 1974-2010 by combining country-level data on conservation agriculture area values from AQUASTAT (FAO, 2016) and LUH2 crop areas together with the methodology of \citep{porwollik_generating_2018} to identify potential no tillage cells.
+This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from \citep{fao_aquastat_2016} and half-degree resolution physical crop areas from \citep{hurtt_harmonization_2020}, applying the methodology of \citep{porwollik_generating_2018} to identify potential no-tillage grid cells.
+
+\hypertarget{sec:scenarios}{%
+\subsection{Scenario definitions}\label{sec:scenarios}}
+
+To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
+In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{mrsoil_karstens_2020}).
+
+\hypertarget{sec:scen_management}{%
+\subsubsection{Management scenarios}\label{sec:scen_management}}
+
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set); (2) the \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
+manure (in \(\tfrac{t}{ha}\)) at the level of 1975 onwards.
+Within the \textit{constResidue} scenario at least three effects overlay each other: (1) yields and with them residue biomass increase due to productivity gains, (2) rates of residue left or returned to fields are raising, and (3) shifts of crop groups change the amount of residue biomass due to crop-group specific harvest index values.
+The \textit{constManagement} combines all three scenarios.
+
+\hypertarget{sec:scen_initial}{%
+\subsubsection{Initialization choice}\label{sec:scen_initial}}
+
+As outlined in Sect.
+\citet{ref}(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
+Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
+We additionally combined the \textit{Initial-natveg} scenario with the (\textit{constManagement} scenario.
+
+\hypertarget{sec:scenlitterpnv}{%
+\subsubsection{Parameterization of natural litter fall}\label{sec:scenlitterpnv}}
+
+Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short lg/c) and nitrogen-to-carbon (short n/c) ratios for all land under natural vegetation.
+This is however a strong simplification.
+Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
+Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
+Whereas lg/c ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, n/c ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
+We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
+We test however a range of different plant types as specified in Table \citet{ref}(tab:scenlitterpnv).
+
+ \begin{table*}[h]
+ \caption{Scenarios for sensitivity analysis on litter parameterization}
+ \begin{tabular}{l l l l}
+ \tophline
+  Scenario name & plant type & n/c & lg/c \\
+  \middlehline
+  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
+  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
+  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
+  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
+  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
+  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
+  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
+ \bottomhline
+ \end{tabular}
+ \belowtable{}
+ \label{tab:scenlitterpnv}
+ \end{table*}
+
 \newpage
@@
 
-We present simulation results of our SOC budget focusing on cropland areas for the year 2010 as well as global trends of SOC stock changes for the period from 1975 to 2010.
-
 \hypertarget{soc-distribution-and-depletion}{%
@@
 
-\begin{figure}[H]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas.
-(b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades.
-(c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics.
-Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.}\label{fig:SOCmaps}
+\begin{figure}[h]
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas.
+(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics.
+Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
+(d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+}\label{fig:SOCmaps}
 \end{figure}
 
-In fig.~1(a) we provide the first world map of SOC on croplands considering real world management data on the global scale.
-Our spatially explicit results moreover show hot spots of SOC losses as well as gains in two different ways:
+In Fig.
+\ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
+Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands, to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
+Our spatially explicit results show hot spots of SOC losses and gains compared to SOC under potential natural vegetation in two different ways:
 1.
-Absolute SOC changes (see fig.~1(c)) indicate areas with high importance for the global SOC emissions.
-The might be driven by huge relative losses or a high natural stock, from which even small deviations could lead to substantial losses.
+Absolute SOC changes \(\Delta SOC\) (see Fig.
+\ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses.
+The might be driven by large relative changes (e.g.~in Central Africa) or a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-east Asia).
+In contrast, large net gains of SOC happen primarily in developed countries of the North according to our results.
 2.
-To attribute SOC losses to insufficient agricultural management relative SOC changes (\(F^{SCF}\), see fig.~1(d)) are a helpful tool.
-They indicate areas with huge difference in carbon inflows or SOC decay compared to natural vegetation, that might be overcome due to improved agricultural practices.
+Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig.
+\ref{fig:SOCmaps}(c)) are a helpful tool, to measure the impact of human cropping activities.
+They indicate areas with large difference in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
+Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
+Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have improved, but presumably irrigated croplands at the border to dry, unsuitable areas worldwide as well.
 
-\hypertarget{agricultural-management-effects-on-soc-emissions-and-cycling}{%
-\subsection{Agricultural management effects on SOC emissions and cycling}\label{agricultural-management-effects-on-soc-emissions-and-cycling}}
+Focusing of the temporal dimension of SOC stock changes (see Fig.
+\ref{fig:SOCmaps}(c)), global SOC debt has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\).
+This leads to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\) and considering our estimate of the global SOC stock of \(675\unit{GtC}\) in 2010 to \(0.2\) per mille per year.
 
-Global cumulative SOC emissions are decreasing (see fig.~1(c)).
-Fig.
-2 reveals the relative impact of management effects by freezing tillage areas as well as carbon inflows from residues or manure at the level of 1975.
-Our counterfactual scenarios show that the increasing residue carbon input had the biggest overall effect on SOC stocks.
-Without changes in management regimes especially in residue inflows to the soil, global cumulative SOC emissions would still grow.
-The strong effect of carbon residue amounts are also visible in the carbon flow diagram for agricultural production for the year 2010 (see fig.~3)
+\hypertarget{agricultural-management-effects-on-soc-debt}{%
+\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
+
+Figure \ref{fig:SOCscen}(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices.
+Without changes in management regimes, especially in C inflows from crop residues to the soil, global SOC debt on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\).
+The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of \(0.06\unit{GtC yr^{-1}}\).
+Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both), the effect of no-tillage gets stronger till 2010.
+The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \ref{fig:SOCscen}(b) as well.
+
+\begin{figure}[h]
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+ConstManagement combines all three modifications.
+Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
+Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
+\end{figure}
+
+The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig.
+\ref{fig:SOcinit}.
+The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
+The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
 
 \begin{figure}[H]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario} \caption{Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline.}\label{fig:SOCscen}
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
 \end{figure}
 
-\begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/OuFlowFig} \caption{Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.
-Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposed before official counting as soil).}\label{fig:FlowFig}
+Figure \ref{fig:}
+
+\hypertarget{carbon-flows-in-the-agricultural-system}{%
+\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
+
+Figure \ref{fig:FlowFig} illustrates the carbon flows within the agricultural production system on cropland for the year 2010.
+
+\begin{figure}[h]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.}\label{fig:FlowFig}
 \end{figure}
-
-\newpage
 
@@
 
-\hypertarget{including-agricultural-management-data-changes-the-sign-of-the-trend}{%
-\subsection{Including agricultural management data changes the sign of the trend}\label{including-agricultural-management-data-changes-the-sign-of-the-trend}}
+\hypertarget{agricultural-management-matters}{%
+\subsection{Agricultural management matters}\label{agricultural-management-matters}}
 
-This study provides an analysis of historic SOC stock changes on cropland.
-We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions (\(SOC_{natveg}\)).
-Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010.
-Whereas recent modeling estimates of global SOC emissions indicate an ongoing increase of SOC emissions (\citep{pugh_simulated_2015}, \citep{sanderman_soil_2017}), our study indicates that the global SOC gap is slowly closing due to improved management.
+This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to constant management assumptions.
+This result remains robust to variations of central model parameters and initialization stocks.
+While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to net reduction of the SOC debt.
 
-According to \citep{sanderman_soil_2017} cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
-Our estimate of 37 GtC in 2010 for cropland emissions only is twice as high as \citep{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and the \citep{sanderman_soil_2017} results have been pointed out as potentially conservatively low compared with experimental results, leaving our results in an aceptable range.
+The evaluation of SOC stock gains and losses, however, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+If defining the SOC debt of 1975 the baseline and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink according to our study.
+Annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 
-Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all.
-\citep{pugh_simulated_2015} considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a~substantial decrease in SOC emissions, accounting for only less than 1\% change in SOC emissions.
+\hypertarget{soc-debt-and-soc-drivers-in-literature}{%
+\subsection{SOC debt and SOC drivers in literature}\label{soc-debt-and-soc-drivers-in-literature}}
+
+According to \citep{sanderman_soil_2017} SOC debt since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+Our estimate of 22 GtC in 2010 for cropland debt is higher as \citep{sanderman_soil_2017} estimations.
+However, there are large uncertainties in modeling SOC at the global scale, and the \citep{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results, suggesting that our results are in an acceptable range.
+
+Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
+The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a~substantial decline in SOC debt (less than 1\% change).
 
 Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices.
-This challenges \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
+We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
+This is in contrast to \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
 
-Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates play the most important role, followed by (2) improved manure recycling (e.g.~due to improved animal waste management systems) and (3) the adoption of no tillage practices.
+Moreover, our sensitivity analysis indicates that (1) higher returned residue biomass driver by yield increases play the most important role, followed by (2) greater manure applications due to an increasing livestock sector as well as improved animal waste management systems, and (3) the adoption of no tillage practices.
+Unfortunately, disentangling drivers of higher residue and manure inputs remains unsolved within this study.
 
-modeling management effects on the global scale comes however with parametric and structural uncertainties.
+\hypertarget{limitations-and-uncertainties}{%
+\subsection{Limitations and uncertainties}\label{limitations-and-uncertainties}}
+
+Modeling management effects at the global scale, however, comes with parametric and structural uncertainties.
+High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs.
 As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
 \citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries which might not be reflected in outdated data sources.
-Following this line of argumentation, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
-However, with an evaluation of management effects (see below), there is no indication for a systematic overrating of residue biomass.
-Rather, it is more likely that we are still missing carbon inputs to the soil.
+Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
+Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
+However, with an evaluation of management effects (see Sect.
+\ref{sec:ipcccompare}), there is no indication for a systematic overrating of residue biomass.
+Rather, it is also likely that we are still missing carbon inputs e.g.~from cover- and inter-cropping to the soil.
 
-\hypertarget{modeled-management-effect-in-line-with-default-ipcc-assumptions}{%
-\subsection{Modeled management effect in line with default IPCC assumptions}\label{modeled-management-effect-in-line-with-default-ipcc-assumptions}}
+Another aspect concerns the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
+As shown in Fig.
+\ref{fig:SOCinit} the SOC debt estimation halves (from \textasciitilde26 GtC to \textasciitilde14 GtC) as well as the decrease in SOC debt (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
+\citep{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
+Our results show that these legacy fluxes are compensated by agricultural management changes even under the strong assumption of the Initial-natveg scenario.
+However, it remains open, if our simple, depth-restricted SOC model itself can represent these fluxes correctly.
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}) which are based on measurement data for croplands (see \ref{tab:SCFtable}).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones.
-
-\begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \caption{This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation.
-We compare average values for four IPCC climate zone classifications.
-IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget).}\label{fig:SCFtable}
-\end{figure}
-
-Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006.
-For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019.
-Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems might be an overestimation of actual SOC state.
-The additional effect of considering low inputs of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for 14-21\% of it.
-
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of \(F_{SCF}\).
-
-\hypertarget{soc-stocks-inline-with-literature}{%
-\subsection{SOC stocks inline with literature}\label{soc-stocks-inline-with-literature}}
-
-The world's SOC stock and its changes are highly uncertain (cite), which is seen in the wide range of global SOC stock estimates (see \ref{tab:SOCtable}).
-
-\begin{figure}[H]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data (LPJmL4, this study) for the year 2010.}\label{fig:SOCtable}
-\end{figure}
-
-The global estimates of SOC stock from this study are on the lower end compared to other modeled results or more data driven estimates.
-Looking at regional results \ref{append:regcompare}, our estimates turn out to be in good agreement for most regions, with the largest deviations for boreal areas.
-Considering that the model was parametrized for croplands, these mismatches are not~suprising as the temperature effects on decomposition are fundamentally different for permafrost soils.
-To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone.
-Moreover, when focussing on SOC changes on cropland, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.
-
-Our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner.
-For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes.
-This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain.
-The absolute values of stocks and emissions from land-use change therefore have to be used with caution.
-Especially in less forested areas the natural land representation higher uncertainty may arise from the parametrization assumptions of natural litterfall.
-
-Nevertheless, total SOC stocks are in a reasonable range and results for SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all.
-The SOC gap from changed management is only affected by the natural vegetation representation in the case of land use transitions.
-In this case, former natural land SOC stocks decline.
-However, as pointed out before, due to low cropland expansion rates, land use change-related SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
-We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
-(see appendix). This shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.
-
-\hypertarget{important-shortcomings-might-be-added-as-well}{%
-\subsection{Important shortcomings --\textgreater{} might be added as well}\label{important-shortcomings-might-be-added-as-well}}
-
-The \textbf{initialization of SOC stocks} in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start.
-That is inline with IPCC guidelines calling for a run-in period of 5-20 years.
-However, the results are heavily influenced by it as \ref{append:initcompare} suggested.
-Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore.
-For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010.
-For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010.
-The relative declining trend for the SOC gap between 1975 and 2010 (-12\%) is however almost the same as for our historic estimation (-14\%).
-According to LUH \citep{hurtt_harmonization_2020} cropland area has increased over 62\% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries.
-Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.
-
-Following the IPCC guidelines this study has limited its focus to the \textbf{first 30 cm of the soil profile}.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes, that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are over simplified within our approach.
@@
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Consequently, our SOC loss estimates are too high.
-Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers.
-Following this arguments SOC stocks in croplands might even be underestimated.
+Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (\citep{powlson_limited_2014}) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.
 
-Carbon displacement via \textbf{leaching and erosion} is neglected in this study.
-As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain.
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
+However, as pointed out by \citep{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions (\citep{wang_human-induced_2017}).
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
+Human interventions have led to large changes in SOC stocks there as well (\citep{sanderman_soil_2017},\citep{friedlingstein_global_2019}).
+This study does not try to be an comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
+
+\hypertarget{sec:ipcccompare}{%
+\subsection{Modeled management effect in line with default IPCC assumptions}\label{sec:ipcccompare}}
+
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) which are based on measurement data for croplands (see Table \ref{tab:SCFtable}).
+To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
+\ref{append:climatemap}).
+
+\textbackslash begin\{figure\}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \textbackslash caption\{This table shows different estimates of stock change factors \(F_{SCF}\), defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation.
+We compare average values for four IPCC climate zone classifications.
+IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in \citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019} are compared to 1990 and 2010 results for historic input levels of this study.\}\label{fig:SCFtable}
+\textbackslash end\{figure\}
+
+Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}).
+Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_hs_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), leaving our tropical results in between.
+Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially.
+Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
+Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
+\ref{fig:SOCmaps}(c)) might be not surprising.
+
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of \(F_{SCF}\) as also indicated by the time trend of the SOC debt.
+
+\hypertarget{soc-stocks-in-line-with-literature}{%
+\subsection{SOC stocks in line with literature}\label{soc-stocks-in-line-with-literature}}
+
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}, \citep{schaphoff_lpjml4_2018-1}) in Table \ref{tab:SOCtable}).
+
+\begin{figure}[h]
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
+This makes it hard to pinpoint a specific year for these SOC estimations.
+In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCtable}
+\end{figure}
+
+The global estimates of SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+Looking at regional results in Fig.
+\ref{append:regcompare}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
+To avoid that this bias influences our results, our study focuses on SOC changes on cropland.
+Pristine natural vegetated areas without human land management cancel out in calculation of \(\Delta SOC\) and \(F^\mathrm{SCF}\).
+
+Additionally, our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural vegetation, which are, however, only estimated in a basic manner.
+For example, we do not differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass plant types.
+This leaves carbon inputs and decay behavior for natural litterfall rather uncertain.
+The absolute values of SOC stocks and debt from land-use change have to be used with caution.
+As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land may arise especially in less forested areas.
+
+We conducted a sensitivity analysis (Fig.
+\ref{append:senslitterpnv} based on various plant parameterizations from the Century model (see Sect.
+@ref(sec:scen\_litterpnv)).
+This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC is not altered under various estimates for natural SOC stocks.
 \newpage
@@
 
-\hypertarget{append:subsection2mrfunctions}{%
-\subsection{table on method subsections to functions within R packages}\label{append:subsection2mrfunctions}}
-
-\hypertarget{append:Tableluh2fao2mag}{%
-\subsection{table on mapping LUH2FAO2MAG}\label{append:Tableluh2fao2mag}}
-
-\hypertarget{append:Tablekcr2kres}{%
-\subsection{kcr2kres mapping}\label{append:Tablekcr2kres}}
+\hypertarget{append:climatemap}{%
+\subsection{map on climate zone used for SCF}\label{append:climatemap}}
 
@@
 
-Litter is coming from LPJmL in carbon units - transformation with 0.44 is done twice reverting the effect of the transformation
-
 \hypertarget{append:TableclossAWMS}{%
 \subsection{closs in AWMS - Table}\label{append:TableclossAWMS}}
-
-\hypertarget{append:climatemap}{%
-\subsection{map on climate zone used for SCF}\label{append:climatemap}}
 
@@
 
-\hypertarget{append:initcompare}{%
-\subsection{global SOC gap in GtC for different initialization choices}\label{append:initcompare}}
+with
+Canada, Australia and New Zealand: CAZ;
+China: CHA;
+European Union: EUR;
+India: IND;
+Japan: JPN;
+Latin America: LAM;
+Middle East and north Africa: MEA;
+non-EU member states: NEU;
+other Asia: OAS; reforming countries: REF; Sub-Saharan Africa:
+SSA; United States: USA.
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{../ResultNotebooks/Output/Images/scenario_lu} \caption{Global SOC gap in GtC for the first 30 cm of soil for different initialization choices.}\label{fig:initSOC}
-\end{figure}
+\hypertarget{append:senslitterpnv}{%
+\subsection{global SOC debt in GtC for various litter parameterizations}\label{append:senslitterpnv}}
 
-\hypertarget{more-short-comings}{%
-\subsection{more short comings}\label{more-short-comings}}
+\newpage
 
-\begin{itemize}
-\item
-  Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs.
-Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer amounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.
-\item
-  Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC).
-(Note that also manure excreted to as well as tillage on pastures are neglected within this analysis, since we focus purely on cropland dynamics.)
-\item
-  Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods).
-Crop specific growing periods might be possible using LPJmL data.
-\item
-  Flooded rice area are not represented correctly as parametrization does not hold true for flooded conditions.
-\item
-  Non-net/Gross land use transitions are not tracked in this study.
-\item
-  Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types.
-Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.
-\item
-  The disaggregation of manure to build-up areas (in the case of extensive monogastrics) is leading to a lot of displaced manure (?) that is cut off
-\item
-  It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.
-  \newpage
-\end{itemize}
-
-\codedataavailability{Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/.
-Data used for the output can be found under .} %% use this section when having data sets and software code available
+\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions), github.com/pik-piam/mrsoil (soil dynamic related functions) and github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing.
+Model results are accessable under (zenode\_link).
+Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.} %% use this section when having data sets and software code available
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Karstens wrote code and paper build on work of Bodirsky (and ).
-Bodirsky, and Popp revised paper.} %% optional section
+\authorcontribution{KK wrote code and paper build on work of BLB, IW.
+JPD revised and improved the model code.
+BB, CM, MK, IW, JH, SR, MD, SM, HLC, AP revised paper.} %% optional section
 
@@
 \begin{acknowledgements}
-Thanks to the rticles contributors!
+Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors.
 \end{acknowledgements}
@@
 \bibliographystyle{copernicus}
-\bibliography{SOCbudget2.bib}
+\bibliography{SOCbudget.bib}
 %%

### 2020-12-11 SOCbudgetPaper/discussion.Rmd
compliling version

--- 2020-12-11 SOCbudgetPaper/discussion.Rmd
@@
 This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.
-LPJmL4 "
+In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010."
 ```
 
-```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
+```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="h"}
 # All defaults

### 2020-12-11 SOCbudgetPaper/methods.Rmd
compliling version

@@
 Therefore, we take the following three steps for each year of our simulation period: 
-(1) We calculate annual land-use type-specific steady states and decay rates for $SOC$ stocks (Sect. \@ref(#sec:steadystates)); 
-(2) we account for land conversion by transferring $SOC$ from and natural vegetation (Sect. \@ref(#sec:carbontransfer)),
-(3) we estimate $SOC$ stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \@ref(#sec:totalsoc)). 
-To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \@ref(#sec.initsoc)).
+(1) We calculate annual land-use type-specific steady states and decay rates for $SOC$ stocks (Sect. \@ref(sec:steadystates)); 
+(2) we account for land conversion by transferring $SOC$ from and natural vegetation (Sect. \@ref(sec:carbontransfer)),
+(3) we estimate $SOC$ stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \@ref(sec:totalsoc)). 
+To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \@ref(sec:initsoc)).
 
@@
 The remaining $AGR$ as well as all $BGR$ are expected to be left on the field.
-We limit high returning rates at $10\unit{tC}\unit{ha}^{-1}}$ to correct for outliers.
+We limit high returning rates at $10\unit{tC\,ha}^{-1}$ to correct for outliers.
 
@@
 \@ref(append:TableclossAWMS) for more details).
-We limit high application shares at $10\unit{tC}\unit{ha}^{-1}}$ to correct for outliers.
+We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.
 
@@
 
-### Parameterization of natural litter fall {#sec:scen_litterpnv}
+### Parameterization of natural litter fall {#sec:scenlitterpnv}
 
@@
 We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table @ref(tab:scen_litterpnv).
+We test however a range of different plant types as specified in Table @ref(tab:scenlitterpnv).
 

### 2020-12-11 SOCbudgetPaper/results.Rmd
compliling version

@@
 
-```{r caption_SOCmaps, echo=FALSE}
-caption_SOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas.
+```{r captionSOCmaps, echo=FALSE}
+captionSOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas.
 (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics.
-Whereas absolute losses ($\DeltaSOC$) are often highest in temperate dry regions, relative losses ($F^\mathrm{SCF}$) are often larger in tropical moist areas.
+Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
 (d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
@@
 
-```{r SOCmaps, out.width = "18cm", fig.cap = caption_SOCmaps, echo=FALSE, fig.pos="h"}
+```{r SOCmaps, out.width = "18cm", fig.cap = captionSOCmaps, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 
-```{r caption_SOCscen, echo=FALSE}
-caption_SOCscen <- "(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+```{r captionSOCscen, echo=FALSE}
+captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
 ConstManagement combines all three modifications.
-Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
+Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
 Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
@@
 
-```{r SOCscen, out.width = "18cm", fig.cap = caption_SOCscen, echo=FALSE, fig.pos="h"}
+```{r SOCscen, out.width = "18cm", fig.cap = captionSOCscen, echo=FALSE, fig.pos="h"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_horiz.png")
 ```
@@
 
-```{r caption_SOCinit, echo=FALSE}
-caption_SOCinit <- "Global $\Delta SOC$ for different SOC initialization choices in the start year 1901."
+```{r captionSOCinit, echo=FALSE}
+captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901."
 ```
 
-```{r SOCinit, out.width = "18cm", fig.cap = caption_SOCinit, echo=FALSE, fig.pos="H"}
+```{r SOCinit, out.width = "18cm", fig.cap = captionSOCinit, echo=FALSE, fig.pos="H"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scenario.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_init.png")
 ```
@@
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/OuFlowFig.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/CarbonBudget.png")
 ```

### 2020-12-13 ResultNotebooks/results_SOCcompare.Rmd
update results

--- 2020-12-10 ResultNotebooks/results_SOCcompare.Rmd
@@
 ```{r load}
-library(mrcommons)
-library(mrsoil)
-library(magpiesets)
+library(magclass)
+library(madrat)
 
-rev       <- "rev0.11"
 outputdir <- "Output/"
-
-setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-#untar(paste0(getConfig("outputfolder"),"/",rev,"histManagement_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
 
@@
 
-SOCStock     <- dimSums(dimSums(SoilCarbon[,,"actualstate"], dim=3.3)* Landuse, dim=3)
-glo_SOCStock <- dimSums(SOCStock, dim=1) 
-cell2iso     <- toolGetMapping("CountryToCellMapping.csv", type = "cell")
-iso_SOCStock <- toolCountryFill(toolAggregate(SOCStock, rel=cell2iso, from="celliso", to="iso", dim=1), fill=0)
-iso2reg      <- read.csv2("Data/histManagement/regionmappingH12.csv")
-reg_SOCStock <- toolAggregate(iso_SOCStock, rel=iso2reg, from="CountryCode", to="RegionCode")
-```
-
-Load validation data:
-
-```{r Validation Stocks}
-
-library(mrvalidation)
-
-  LPJ_IPCC   <- calcOutput("ValidCarbonStocks", datasource = "LPJ_IPCC2006", aggregate="REG+GLO", try=TRUE)
-  WISE       <- calcOutput("ValidCarbonStocks", datasource = "WISE", aggregate="REG+GLO", try=TRUE)
-  GSOC       <- calcOutput("ValidCarbonStocks", datasource = "GSOC", aggregate="REG+GLO", try=TRUE)
-  SoilGrids  <- calcOutput("ValidCarbonStocks", datasource = "SoilGrids", aggregate="REG+GLO", try=TRUE)
-  LPJmL4     <- calcOutput("ValidCarbonStocks", datasource = "LPJmL4Paper", aggregate="REG+GLO", try=TRUE)
-
+SoilCarbon[,,c("naturalstate.crop")] <- SoilCarbon[,,c("naturalstate.natveg")]
+SOCStock     <- dimSums(dimSums(SoilCarbon[,,c("naturalstate","actualstate")], dim=3.3)* Landuse, dim=3.2)
+cell2reg     <- read.csv("Data/histManagement/ClimateZoneMapping.csv")
+reg_SOCStock <- toolAggregate(SOCStock, rel=cell2reg, from="cell", to="IPCC")
 ```
@@
 
+  WISE       <- readRDS("Data/histManagement/WISE.rds")
+  GSOC       <- readRDS("Data/histManagement/GSOC.rds")
+  SoilGrids  <- readRDS("Data/histManagement/SoilGrids.rds")
+  LPJmL4     <- readRDS("Data/histManagement/LPJmL4.rds")
+
 library(knitr)
 library(stats)
-library(luplot)
 library(reshape2)
 
-SOCStock <- mbind(glo_SOCStock, reg_SOCStock)
-glo_SOC  <- mbind(setNames(SOCStock[,"y2010",],"SOC Budget"), 
+SOCStock <- reg_SOCStock
+reg_SOC  <- mbind(setNames(SOCStock[,"y2010","actualstate"],"This Study"),
+                  setNames(SOCStock[,"y2010","naturalstate"],"This Study (PNV)"),
                   setNames(LPJmL4[,"y2010",], "LPJmL4Paper"),
@@
                   setYears(setNames(GSOC[,1,],"GSOC"),"y2010"))
+glo_SOC <- mbind(setCells(dimSums(reg_SOC, dim=1),"GLO"), reg_SOC)
 
@@
 
-saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"gloSOC_comparison.rds"))
+comparison_table$Year <- "2010"
+
+saveRDS(object = comparison_table, paste0(outputdir,"gloSOC_comparison.rds"))
 
@@
 ```{r kable2image}
-library(kableExtra)
-
-table <- readRDS("Output/gloSOC_comparison.rds")
-colnames(table[[1]])[1] <- "Model/Database"
-colnames(table[[1]])[5] <- "Global SOC in GtC"
-save_kable(kable_styling(kable(table$comparison_table[c(1,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
-         file="Output/Images/TableSOC_comparison.png")
-
-save_kable(kable_styling(kable(table$comparison_table[setdiff(1:13,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
-         file="Output/Images/TableRegSOC_comparison.png")
+# library(kableExtra)
+# 
+# table <- readRDS("Output/gloSOC_comparison.rds")
+# colnames(table)[1] <- "Model/Database"
+# colnames(table)[6] <- "Global SOC in GtC"
+# save_kable(kable_styling(kable(table$comparison_table[c(1,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
+#          file="Output/Images/TableSOC_comparison.png")
+# 
+# save_kable(kable_styling(kable(table$comparison_table[setdiff(1:13,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
+#          file="Output/Images/TableRegSOC_comparison.png")
 
@@
 
-supp.labs <- getNames(glo_SOC["GLO",,invert=TRUE])
-names(supp.labs) <- c("OJ", "VC")
+table    <- readRDS("Output/gloSOC_comparison.rds")
+leg_name <- ""
 
-leg_name <- "Data Source"
+names(table) <- c("Data1","Boreal Dry", "Boreal Moist", "C-Temp Dry", "C-Temp Moist","GLO","Polar Dry", "Polar Moist",
+                  "Tropical Dry", "Tropical Moist", "Tropical Montane", "Tropical Wet", "W-Temp Dry", "W-Temp Moist", "Year")
+table        <- table[,c(1,15,6,7,8,2,3,4,5,13,14,9,10,11,12)]
 
-glob <- ggplot(data = as.data.frame(glo_SOC["GLO",,][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
+cols   <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")[c(7,3,1,4,6,8)]
+
+glob <- ggplot(data = melt(subset(table, select =c(Data1, GLO, Year))), aes(x=Year,y=value, group=Data1)) + 
   geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
-  scale_size_manual(values=c(4,2,2,2,2)) + 
-  scale_shape_manual(values=c(19,6,2,5,0)) +
+  scale_size_manual(values=c(4,2,2,2,2,2)) + 
+  scale_shape_manual(values=c(19,16,6,2,5,0)) +
+  scale_color_manual(values = cols) +
   labs(color  = leg_name, size = leg_name, shape = leg_name) +
-  xlab("") + ylab("SOC (0-30cm) in GtC")
+  xlab("") + ylab("SOC (0-30cm) in GtC") +
+  theme(axis.ticks.x=element_blank(),
+        axis.text.x=element_blank(), 
+        legend.position="right") + guides(fill=guide_legend(ncol=1,byrow=TRUE))
+        #legend.position="bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))
 
-comp <- ggplot(data = as.data.frame(glo_SOC["GLO",,invert=TRUE][,"y2010",]/1000), aes(Year,Value, group=Data1)) + 
+comp <- ggplot(data = melt(subset(table, select=-GLO)), aes(x=Year,y=value, group=Data1)) + 
   geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
-  facet_wrap(~ Region, scales="free", drop=TRUE) + 
-  scale_size_manual(values=c(4,2,2,2,2)) + 
-  scale_shape_manual(values=c(19,6,2,5,0)) +
+  facet_wrap(~ variable, scales="free", drop=TRUE, labeller = label_wrap_gen(width=8)) + 
+  scale_size_manual(values=c(4,2,2,2,2,2)) + 
+  scale_shape_manual(values=c(19,16,6,2,5,0)) +
+  scale_color_manual(values = cols) +
   labs(color  = leg_name, size = leg_name, shape = leg_name) +
-  xlab("") + ylab("SOC (0-30cm) in GtC")
+  xlab("") + ylab("SOC (0-30cm) in GtC") +
+  theme(axis.ticks.x=element_blank(),
+        axis.text.x=element_blank())
 
@@
 
-ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob, height=6, width=3, units="cm")
-ggsave(paste0(outputdir,"/reg_comparisonfigure.png"), plot=comp)
+ggsave(paste0(outputdir,"/Images/glo_comparisonfigure.png"), plot=glob, height=8, width=8, units="cm", dpi = 300)
+ggsave(paste0(outputdir,"/Images/reg_comparisonfigure.png"), plot=comp, height=10, width=14, units="cm", dpi = 300)
 

### 2020-12-13 ResultNotebooks/results_SOCscenario.Rmd
update results

@@
 
-rev <- "rev0.11"
-
-setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-
-filesNeeded <- c("SoilCarbon.rds","Landuse.rds")
-
+rev         <- "rev16"
 scenarios   <- c("histManagement","constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
@@
 
+# filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
+# setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
+# 
 # untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
@@
 ```
-
-Calc global numbers
 

### 2020-12-13 ResultNotebooks/results_SOCtable.Rmd
update results

@@
 
-rev       <- "rev0.11"
+rev       <- "rev16"
 outputdir <- "Output/"
 
-setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
 #untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
 
-SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))[,1975:2010,]
-Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))[,1975:2010,]
+SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
 
@@
 
-years             <- intersect(findset("past"), getYears(SoilCarbon))
-
-### calculating climate mask
-koeppen           <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )[,years,]
-koepp2IPCC        <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
-getNames(koeppen)         <- tolower(getNames(koeppen))
-koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
-climateIPCC       <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
-rm(koeppen)
+years             <- findset("past")
+climateIPCC       <- readRDS("Data/histManagement/IPCC_reduced.rds")
 
@@
 
-knitr::kable(SCF_table, "latex",digits=2, label="naturalstate_table")
-
 ```
@@
 
+climateIPCCr       <- readRDS("Data/histManagement/IPCC_reduced.rds")
+climateIPCC        <- readRDS("Data/histManagement/IPCC.rds")
+
+getNames(climateIPCC) <- c( "W-Temp Moist","W-Temp Dry","C-Temp Moist","C-Temp Dry","Polar Moist","Polar Dry","Boreal Moist",
+                            "Boreal Dry","Tropical Montane","Tropical Wet","Tropical Moist", "Tropical Dry")
+
+climateIPCC <- climateIPCC[,,c(5,7,3,1,6,8,4,2,9,11,10,12)]
+
 magpie_coord <- toolGetMapping("magpie_coord.rda", type="cell")
-climate2plot <- dimSums(climateIPCC*new.magpie("GLO",NULL,getNames(climateIPCC),c(1:4)), dim=3)
+climate2plot <- dimSums(climateIPCC*new.magpie("GLO",NULL,getNames(climateIPCC),c(1:12)), dim=3)
 climate2plot <- merge(as.data.frame(climate2plot), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
@@
 
-cols          <- colorRampPalette(brewer.pal(4, "RdYlBu"))(4)
+blucols          <- brewer.pal(9, "Blues")[2:5]
+viocols          <- brewer.pal(9, "Purples")[6:9]
+redcols          <- brewer.pal(9, "Reds")[6:8]
+orgcols          <- brewer.pal(9, "Oranges")[3]
+cols <- c(blucols,viocols,redcols,orgcols)
 
 ClimateZones <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(climate2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
-    scale_fill_manual(name = expression("climate zones"), 
+    geom_tile(data = subset(climate2plot, select=c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
+    scale_fill_manual(name = "", 
                       values = cols, na.value = "white",
-                      breaks= c(1,2,3,4),
-                      labels = gsub("_", " ", getNames(climateIPCC)),
+                      breaks= c(1,2,3,4,5,6,7,8,9,10,11,12),
+                      labels = getNames(climateIPCC),
                       guide = guide_legend(
@@
                       )) +
+   scale_color_manual(values=cols) +
     theme(legend.position="bottom",
@@
           plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
+          ) + guides(fill=guide_legend(nrow=2,byrow=FALSE))
 ClimateZones 
-ggsave(paste0(outputdir,"/climatezones.png"), plot=ClimateZones , width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir,"/Images/climatezones.png"), plot=ClimateZones , width = 20, height = 14, dpi = 300, units = "cm")
 ```
@@
 MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
-                                                                          unique(koepp2IPCC$ipcc_reduced), paste, sep="."))
+                                                                          getNames(climateIPCC), paste, sep="."))
 

### 2020-12-14 SOCbudgetPaper/methods.Rmd
addon

@@
 4 ([@calvo_buendia_ipcc_2019]).
-We additionally describe how we configured and extended the steady-state method (for model code see @mrsoil_2020).
+We additionally describe how we configured and extended the steady-state method (for model code see [@karstens_mrsoil_2020]).
 In Sect.
 \@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
-Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see @mrcommons_2020).
+Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see [@bodirsky_mrcommons_2020]).
 In Sect.
@@
 (1) We calculate annual land-use type-specific steady states and decay rates for $SOC$ stocks (Sect. \@ref(sec:steadystates)); 
-(2) we account for land conversion by transferring $SOC$ from and natural vegetation (Sect. \@ref(sec:carbontransfer)),
+(2) we account for land conversion by transferring $SOC$ from and to natural vegetation (Sect. \@ref(sec:carbontransfer)),
 (3) we estimate $SOC$ stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \@ref(sec:totalsoc)). 
@@
 This equation is valid for all grid cells $i$ and all years $t$.
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [(@parton_analysis_1987]).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987]).
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
@@
 
-Carbon inputs for cropland are below and above ground residues left or returned to the field (see Sect.
+Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
 \@ref(sec:residues)) and manure inputs (see Sect.
@@
     \middlehline
-    \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+    \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
                                                                         \cite{faostat_faostat_2016}, \\
@@
                                                         \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
-                                                                        lg/c generic values according to Table 5.5B\\
+                                                                        LG:C generic values according to Table 5.5B\\
                                                                         in \cite{calvo_buendia_ipcc_2019},\\
-                                                                        crop-type specific n/c-ratios from
-                                                                        \cite{dietrich_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
-                              & below ground residues &  &  \\ \cline{2-2}
+                                                                        crop-type specific N:C-ratios from
+                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
+                              & below-ground residues &  &  \\ \cline{2-2}
                               & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
@@
 Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_hs_ipcc_2006].
-Note that the SOC debt is equivalent to the negated cumulative SOC emission ([@pugh_simulated_2015]) of human land use.
+The SOC debt is equivalent to the negated cumulative SOC component of  of human land use emission ([@pugh_simulated_2015]).
 
@@
 To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
-While land-use, climate and litter input information are available till 1901, agricultural management data on residue and manure inputs are hold constant at the level of 1965.
-After a spin-up period of 64 years with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975.
-Note that irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
+While land-use, climate and litter input information are available from 1901, agricultural management data on residue and manure inputs are held constant at the level of 1965.
+After a spin-up period of 64 years (1901-1965) with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975 onwards.
+Irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
 
@@
 We compile country-specific FAO production and cropland statistics [@faostat_faostat_2016] to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which, together with the spin-up phase, restricts our analysis to the time span from 1975 to 2010.
+The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
 For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
@@
 We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production [@faostat_faostat_2016].
-Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallow land.
+By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
 
@@
 Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_hs_ipcc_2006].
-Note that deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+Deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
 We assume that all $BGR$ are left in the soil, whereas $AGR$ can be burned or harvested for other purposes such as feeding animals [@weindl_livestock_2017], fuel or for material use.
 
-A fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
+A country-specific fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
 Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
@@
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in [@mrcommons]).
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in [@bodirsky_mrcommons_2020]).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets (see [@weindl_livestock_2017]), that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
@@
 The remaining $AGR$ as well as all $BGR$ are expected to be left on the field.
-We limit high returning rates at $10\unit{tC\,ha}^{-1}$ to correct for outliers.
+We limit high residue return rates to at most $10\unit{tC\,ha}^{-1}$ in order to correct for outliers.
 
@@
 For low-income countries, we assume only extensive production systems.
-We locate them according to the share of built-up areas based on the asumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
+We locate them according to the share of built-up areas based on the assumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
 Intensive production associated with high-income countries, is distributed within a country using the share of primary-crop production, assuming that feed availability is the most determining factor for livestock location.
@@
 Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed c/n ratios (given by [@calvo_buendia_ipcc_2019]).
+Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see [@calvo_buendia_ipcc_2019]).
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 As a result $wat$ is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a (partly) irrigated grid cells, since we do not have crop group specific growing periods at hand.
-In addition, we expect, either water limitations to be no problem during the off-season (causing our assumption to be a good approximation), or irrigated growing periods to span over the whole year in water-restricted areas.
+Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
+This will lead to an overestimation of the irrigation effect.
+We expect, however, water limitations to be a minor problem during the off-season in temperature limited cropping regions, causing our assumption to not dramatically overestimate the moisture effects.
+In tropical, water-limited cropping areas, irrigated growing periods might even span over the whole year.
 
@@
 
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set); (2) the \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for these three drivers: In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set).
+The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
 manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards.
-Within the \textit{constResidue} scenario at least three effects overlay each other: (1) yields and with them residue biomass increase due to productivity gains, (2) rates of residue left or returned to fields are raising, and (3) shifts of crop groups change the amount of residue biomass due to crop-group specific harvest index values.
-The \textit{constManagement} combines all three scenarios.
+Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
+The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
 
@@
 
-Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short lg/c) and nitrogen-to-carbon (short n/c) ratios for all land under natural vegetation.
+Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short LG:C) and nitrogen-to-carbon (short N:C) ratios for all land under natural vegetation.
 This is however a strong simplification.
@@
 Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines ([@calvo_buendia_ipcc_2019]), but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
-Whereas lg/c ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, n/c ratios range between $0.001-0.05$ with an average of $0.015$ (default value).
+Whereas LG:C ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, N:C ratios range between $0.001-0.05$ with an average of $0.015$ (default value).
 We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
@@
  \tophline
-  Scenario name & plant type & n/c & lg/c \\
+  Scenario name & plant type & N:C & LG:C \\
   \middlehline

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
Update abstract 2

--- 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
@@
     affiliation: 1
+    email: bodirsky@pik-potsdam.de
+    orcid: https://orcid.org/0000-0002-8242-6712
   - given_name: Jan Philipp
@@
     affiliation: 1 
+    orcid: https://orcid.org/0000-0002-4309-6431
   - given_name: Marta
@@
     affiliation: 1 
+    orcid: https://orcid.org/0000-0002-9491-3550
   - given_name: Susanne
@@
     affiliation: 1
+    orcid: https://orcid.org/0000-0002-7651-6930    
   - given_name: Hermann
@@
 We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have around 26 GtC relative to a counterfactual natural state in 1975.
+We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
 Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
@@
 
-bibliography: SOCbudget2.bib
+bibliography: SOCbudget.bib
 running:
@@
   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions),  github.com/pik-piam/mrsoil (soil dynamic related functions) and  github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing.
-Model results are accessable under (zenode_link).
-Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
+Model results are accessible under (zenode_link).
+Software code for paper and result preparation can be found under www.github.com/k4rst3ns/historicalsocmanegement (zenodo_link).
 authorcontribution: |
-  KK wrote code and paper build on work of BLB, IW. 
+  KK, BLB and AP designed the study and the model idea.
+KK wrote the code build on work of BLB, IW.
 JPD revised and improved the model code.
-  BB, CM, MK, IW, JH, SR, MD, SM, HLC, AP revised paper.
-disclaimer: |
-  We like Copernicus.
+CM, JH and SR provided the LPJmL simulation data.
+KK wrote the paper with important contributions of BLB and CM, and  MK, JS, SR and IW.
+All authors discussed the results and commented on the manuscript.
+
 acknowledgements: |

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
update discussion 2

@@
 
-## Agricultural management matters
-
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to constant management assumptions.
-This result remains robust to variations of central model parameters and initialization stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to net reduction of the SOC debt.
-
-The evaluation of SOC stock gains and losses, however, is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 the baseline and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink according to our study.
-Annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to the often used constant management assumptions.
+This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
+While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
 
@@
 
-According to [@sanderman_soil_2017] SOC debt since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+The evaluation of SOC stock gains and losses, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink since 1975 according to our study.
+However, annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+
+According to [@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
 Our estimate of 22 GtC in 2010 for cropland debt is higher as [@sanderman_soil_2017] estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and the [@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results, suggesting that our results are in an acceptable range.
+However, there are large uncertainties in modeling SOC at the global scale, and [@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.
+This suggests that our results are within a plausible range.
 
 Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-[@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decline in SOC debt (less than 1% change).
+[@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account.
+Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
+Historic productivity increases were, however, notably larger.
+Despite large spatial heterogeneity aggregate yield increase rates are often estimate well above 50% ([@pellegrini_crop_2018], [@ray_recent_2012], [@rudel_agricultural_2009]).
 
@@
 
-Moreover, our sensitivity analysis indicates that (1) higher returned residue biomass driver by yield increases play the most important role, followed by (2) greater manure applications due to an increasing livestock sector as well as improved animal waste management systems, and (3) the adoption of no tillage practices.
-Unfortunately, disentangling drivers of higher residue and manure inputs remains unsolved within this study.
+Looking on historical cropland intensification rates of above 50% ([@rudel_agricultural_2009]), a large fraction of increased C input from residue biomass, is attributed to productivity improvements.
 
@@
 
-Modeling management effects at the global scale, however, comes with parametric and structural uncertainties.
+Modeling management effects at the global scale comes with parametric and structural uncertainties.
 High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.
-the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs.
+the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
 As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
@@
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
-However, with an evaluation of management effects (see Sect.
-\@ref(sec:ipcccompare)), there is no indication for a systematic overrating of residue biomass.
-Rather, it is also likely that we are still missing carbon inputs e.g.
-from cover- and inter-cropping to the soil.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for belowground biomass.
+However, looking on the evaluation of management effects (see Sect.
+\@ref(sec:ipcccompare)), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
+Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.
+cover- and inter-cropping practices.
 
-Another aspect concerns the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
+Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
 As shown in Fig.
-\@ref(fig:SOCinit) the SOC debt estimation halves (from ~26 GtC to ~14 GtC) as well as the decrease in SOC debt (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
+\@ref(fig:SOCinit) the SOC debt estimate almost halves (from ~26 GtC to ~14 GtC), and the SOC debt reduction is strongly reduced (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
 [@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results show that these legacy fluxes are compensated by agricultural management changes even under the strong assumption of the Initial-natveg scenario.
-However, it remains open, if our simple, depth-restricted SOC model itself can represent these fluxes correctly.
+Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes, that most of the SOC dynamic happens in the topsoil.
-In this regard several aspects are over simplified within our approach.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil.
+In this regard several aspects are strongly simplified within our approach.
 Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies ([@powlson_limited_2014]) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.
+Secondly changes to the subsoil due to tillage are neglected.
+As [@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
+SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
+Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
 
@@
 However, as pointed out by [@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions ([@wang_human-induced_2017]).
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
 Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
 Human interventions have led to large changes in SOC stocks there as well ([@sanderman_soil_2017],[@friedlingstein_global_2019]).
-This study does not try to be an comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
+This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
 
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) which are based on measurement data for croplands (see Table \@ref(tab:SCFtable)).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_hs_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]).
+Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFtable)).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
@@
 
-Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines ([@calvo_buendia_ipcc_2019]).
-Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), leaving our tropical results in between.
-Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially.
-Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
-Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\@ref(fig:SOCmaps)(c)) might be not surprising.
+Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than [@calvo_buendia_ipcc_2019].
+The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
+Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
+Our results still show  especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
+Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
+\@ref(fig:SOCmaps)(c)) may be expected.
 
@@
 
-The global estimates of SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+SoilGrids ([@hengl_soilgrids250m_2017]) especially stands out for their high estimation, whereas all other sources (including out study) are comparabely similar.
 Looking at regional results in Fig.
-\@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
-To avoid that this bias influences our results, our study focuses on SOC changes on cropland.
-Pristine natural vegetated areas without human land management cancel out in calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.
+\@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
+To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.
 
-Additionally, our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural vegetation, which are, however, only estimated in a basic manner.
-For example, we do not differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This leaves carbon inputs and decay behavior for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be used with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land may arise especially in less forested areas.
+Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
+For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
+This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
+The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
+As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
 
@@
 \@ref(sec:scen_litterpnv)).
-This shows that the general trend of decreasing SOC debt of ~4 GtC is not altered under various estimates for natural SOC stocks.
+This shows that the general trend of decreasing SOC debt of ~4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
update paper part

--- 2020-12-11 SOCbudgetPaper/SOCbudget.Rmd
@@
   codedata: |
-   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions), github.com/pik-piam/mrsoil (soil dynamic related functions) and  github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing.
+   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data, which are both based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
 Model results are accessable under (zenode_link).
@@
   Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors. 
+appendix: |
+
+  \section{Figures and tables in appendices}
+
+  \subsection{Methods}
+
+    ```{=latex}
+  \begin{figure}
+    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
+    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
+Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
+    \label{fig:CLIMzone}
+  \end{figure}
+  ```
+
+  ```{=latex}
+  \begin{table*}[h]
+  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
+Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
+  }
+  \begin{tabular}{lllllllllllll}
+  \tophline
+
+           &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
+  Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
+   \middlehline
+  tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
+  maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
+  trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
+  rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
+  soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
+  rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
+  groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
+  sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
+  oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
+  puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
+  potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
+  cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
+  sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
+  sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
+  others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
+  foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
+  cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
+  \bottomhline
+           &                 & \multicolumn{8}{l}{
+                                             \begin{minipage}[t]{0.50\columnwidth}
+                                                        nr/dm -- nitrogen to dry matter ratio\\
+                                                        wm/dm -- wet matter to dry matter ratio\\
+                                                        c/dm  -- carbon to dry matter ratio\\
+                                                        \end{minipage}} &
+                                                                          \multicolumn{3}{l}{
+                                                                              \begin{minipage}[t]{0.50\columnwidth}
+                                                                                          HI\textsuperscript{area} -- harvest index per area\\
+                                                                                          HI\textsuperscript{prod} -- harvest index per production\\
+                                                                                          RS -- root:shoot ratio \end{minipage}} 
+  \end{tabular}
+  (\#tab:c2dm)
+  \end{table*}
+  ```
+
+  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->
+
+  \subsection{Discussion}
+
+    ```{=latex}
+  \begin{figure}
+    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
+    \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.
+PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
+We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
+    \label{fig:SOCreg}
+  \end{figure}
+  ```
+
+    ```{=latex}
+  \begin{figure}
+    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
+    \caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
+\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
+Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
+    \label{fig:SOClitter}
+  \end{figure}
+  ```
+
 output:
@@
 \newpage
-
-```{r child = 'appendix.Rmd'}
-```
-\newpage

### 2020-12-14 SOCbudgetPaper/appendix.Rmd
update paper part

--- 2020-12-11 SOCbudgetPaper/appendix.Rmd
@@
 
-## map on climate zone used for SCF {#append:climatemap}
+\section{Figures and tables in appendices}
 
-## carbon 2 dry matter {#append:Tablec2dm}
+\subsection{Methods}
 
-## closs in AWMS - Table {#append:TableclossAWMS}
+```{r captionCLIMzone, echo=FALSE}
+captionCLIMzone <- "Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines ([@eggleston_ipcc_2006]) and has been reimplented by [@carre_background_2010], which is the source of this data.
+Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange."
+```
 
-## regional SOC stock in GtC from different sources {#append:regcompare}
+```{r CLIMzone, out.width = "16cm", fig.cap = captionCLIMzone, echo=FALSE}
+knitr::include_graphics("../ResultNotebooks/Output/Images/climatezones.png")
+```
+
+```{=latex}
+\begin{table*}[h]
+\caption{bla.
+}
+\begin{tabular}{lllllllllllll}
+ \tophline
+
+           &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
+Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
+ \middlehline
+tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
+maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
+trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
+rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
+soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
+rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
+groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
+sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
+oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
+puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
+potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
+cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
+sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
+sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
+others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
+foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
+cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
+\bottomhline
+           &                 & \multicolumn{8}{l}{
+                                             \begin{minipage}[t]{0.50\columnwidth}
+                                                        nr/dm -- nitrogen to dry matter ratio\\
+                                                        wm/dm -- wet matter to dry matter ratio\\
+                                                        c/dm  -- carbon to dry matter ratio\\
+                                                        \end{minipage}} &
+                                                                          \multicolumn{3}{l}{
+                                                                              \begin{minipage}[t]{0.50\columnwidth}
+                                                                                          HI\textsuperscript{area} -- harvest index per area\\
+                                                                                          HI\textsuperscript{prod} -- harvest index per production\\
+                                                                                          RS -- root:shoot ratio \end{minipage}} 
+\end{tabular}
+(\#tab:c2dm)
+\end{table*}
+```
+
+<!-- closs in AWMS - Table {#append:TableclossAWMS} -->
+
+\subsection{Discussion}
 
 ```{r captionregSOCtable, echo=FALSE}
-captionregSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
+captionregSOCtable <- "Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.
+PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
+We use the climate zone specification of the IPCC [eggleston_ipcc_2006]."
 ```
 
-```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
+```{r regSOCtable, out.width = "16cm", fig.cap = captionregSOCtable, echo=FALSE}
 # All defaults
@@
 
-with 
-Canada, Australia and New Zealand: CAZ; 
-China: CHA; 
-European Union: EUR; 
-India: IND; 
-Japan: JPN; 
-Latin America: LAM;
-Middle East and north Africa: MEA; 
-non-EU member states: NEU;
-other Asia: OAS; reforming countries: REF; Sub-Saharan Africa:
-SSA; United States: USA.
+```{r captionSOClitter, echo=FALSE}
+captionSOClitter <- "Global $\\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines ([@calvo_buendia_ipcc_2019]), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in [@century_model_2000]).
+\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
+Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates."
+```
 
-## global SOC debt in GtC for various litter parameterizations {#append:senslitterpnv}
+```{r SOClitter, out.width = "16cm", fig.cap = captionSOClitter, echo=FALSE}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_litter.png")
+```

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
update paper part

--- 2020-12-11 SOCbudgetPaper/discussion.Rmd
@@
 [@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decline in SOC debt (less than 1% change).
+The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decline in SOC debt (less than 1\% change).
 
@@
 the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs.
-As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
+As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study).
 [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
@@
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes, that most of the SOC dynamic happens in the topsoil.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes, that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are over simplified within our approach.
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) which are based on measurement data for croplands (see Table \@ref(tab:SCFtable)).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) which are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\@ref(append:climatemap)).
+\@ref(fig:CLIMzone)).
 
-```{r captionSCFtable, echo=FALSE}
-captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation.
-We compare average values for four IPCC climate zone classifications.
-IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in [@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019] are compared to 1990 and 2010 results for historic input levels of this study."
-```
-
-```{r SCFtable, out.width = "16cm", fig.cap = captionSCFtable, echo=FALSE}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
+```{r SCFglo, echo=FALSE, results = 'asis'}
+table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
+library(xtable)
+print(xtable(table$comparison_table, digit=2, comment=FALSE, 
+              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent.
+Also modelled $F^{SOC}$ have increased for all climates over time."))
 ```
@@
 Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines ([@calvo_buendia_ipcc_2019]).
-Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), leaving our tropical results in between.
+Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 ([@eggleston_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), leaving our tropical results in between.
 Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially.
@@
 
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(tab:SOCtable)).
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(fig:SOCglo)).
 
-```{r captionSOCtable, echo=FALSE}
-captionSOCtable <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+```{r captionSOCglo, echo=FALSE}
+captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
 Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
@@
 
-```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="h"}
+```{r SOCglo, out.width = "8cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 Looking at regional results in Fig.
-\@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
+\@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
 To avoid that this bias influences our results, our study focuses on SOC changes on cropland.
@@
 We conducted a sensitivity analysis (Fig.
-\@ref(append:senslitterpnv) based on various plant parameterizations from the Century model (see Sect.
-\@ref(sec:scen_litterpnv)).
+\@ref(fig:SOClitter)) based on various plant parameterizations from the Century model (see Sect.
+\@ref(sec:scenlitterpnv)).
 This shows that the general trend of decreasing SOC debt of ~4 GtC is not altered under various estimates for natural SOC stocks.

### 2020-12-14 SOCbudgetPaper/methods.Rmd
update paper part

--- 2020-12-11 SOCbudgetPaper/methods.Rmd
@@
 4 ([@calvo_buendia_ipcc_2019]).
-We additionally describe how we configured and extended the steady-state method (for model code see @mrsoil_2020).
+We additionally describe how we configured and extended the steady-state method (for model code see [@karstens_mrsoil_2020]).
 In Sect.
-\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@ipcc_eggleston_2006], [@calvo_buendia_ipcc_2019]).
 Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see @mrcommons_2020).
@@
     \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.25\textwidth}}
+    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.40\textwidth}}
     \tophline
@@
     \middlehline
-    \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+    \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                         \cite{faostat_faostat_2016}, \\
@@
                                                                         \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                        \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
-                                                                        lg/c generic values according to Table 5.5B\\
+                                                        \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+                                                                        lg/c generic values based on Table 5.5B, 5.5C \\
                                                                         in \cite{calvo_buendia_ipcc_2019},\\
                                                                         crop-type specific n/c-ratios from
-                                                                        \cite{dietrich_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
-                              & below ground residues &  &  \\ \cline{2-2}
-                              & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
+                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
+                              & below ground residues &  &  \\ 
+                              & manure                &  &  \\ 
                               \hline
@@
  \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
-                                     & \cite{hengl_soilgrids250m:_2017} \\
+                                     & \cite{hengl_soilgrids250m_2017} \\
                       \cline{2-4}
@@
 
-Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_hs_ipcc_2006].
-Note that the SOC debt is equivalent to the negated cumulative SOC emission ([@pugh_simulated_2015]) of human land use.
+Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_ipcc_2006].
+Moreover, $\Delta SOC$ is equivalent to the negated cumulative SOC component of human land use change emission ([@pugh_simulated_2015]).
 
@@
 
-Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]).
 Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
@@
 Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig.
-\@ref(append:climatemap)).
+\@ref(fig:CLIMzone)).
 The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}}$ by
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of [@eggleston_hs_ipcc_2006] and [@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of [@eggleston_ipcc_2006] and [@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see FAO2LUH2MAG_croptypes.csv in [@mrcommons]).
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see FAO2LUH2MAG_croptypes.csv in [@bodirsky_mrcommons_2020]).
 By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 
-Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_hs_ipcc_2006].
+Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_ipcc_2006].
 Note that deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
@@
 A fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
-Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_hs_ipcc_2006].
+Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15\% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_ipcc_2006].
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in [@mrcommons]).
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in [@bodirsky_mrcommons_2020]).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets (see [@weindl_livestock_2017]), that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
-We estimate a material-use share for the straw-residue group of 5% and a fuel-share of 10% for all used residue groups in low-income countries.
+We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
 For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above.
@@
 
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(append:Tablec2dm)).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(tab:c2dm)).
 
@@
 All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
+Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
 For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing. 
@@
 To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see [@mrsoil_karstens_2020]).
+In the following sections we outline name and idea of these scenarios (for technical implementation see [@karstens_mrsoil_2020]).
 
@@
 As outlined in Sect.
-@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+\@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
 As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
@@
 We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table @ref(tab:scenlitterpnv).
+We test however a range of different plant types as specified in Table \@ref(tab:scenlitterpnv).
 

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
current status local

@@
 \ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}).
-We additionally describe how we configured and extended the steady-state method (for model code see \citet{mrsoil_2020}).
+We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}).
 In Sect.
-\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
+\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{ipcc_eggleston_2006}, \citep{calvo_buendia_ipcc_2019}).
 Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citet{mrcommons_2020}).
@@
   \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.25\textwidth}}
+  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.40\textwidth}}
   \tophline
@@
   \middlehline
-  \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+  \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                       \cite{faostat_faostat_2016}, \\
@@
                                                                       \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                      \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
-                                                                      lg/c generic values according to Table 5.5B\\
+                                                      \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+                                                                      lg/c generic values based on Table 5.5B, 5.5C \\
                                                                       in \cite{calvo_buendia_ipcc_2019},\\
                                                                       crop-type specific n/c-ratios from
-                                                                      \cite{dietrich_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
-                            & below ground residues &  &  \\ \cline{2-2}
-                            & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
+                                                                      \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
+                            & below ground residues &  &  \\ 
+                            & manure                &  &  \\ 
                             \hline
@@
 
-Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_hs_ipcc_2006}.
-Note that the SOC debt is equivalent to the negated cumulative SOC emission (\citep{pugh_simulated_2015}) of human land use.
+Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
+Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land use change emission (\citep{pugh_simulated_2015}).
 
@@
 
-Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
+Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
 Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
@@
 Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig.
-\ref{append:climatemap}).
+\ref{fig:CLIMzone}).
 The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by
@@
 with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_hs_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{mrcommons}).
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{bodirsky_mrcommons_2020}).
 By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 
-Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_hs_ipcc_2006}.
+Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
 Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
@@
 Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_hs_ipcc_2006}.
+Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{mrcommons}).
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{bodirsky_mrcommons_2020}).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
@@
 
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{append:Tablec2dm}).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{tab:c2dm}).
 
@@
 To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{mrsoil_karstens_2020}).
+In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{karstens_mrsoil_2020}).
 
@@
 As outlined in Sect.
-\citet{ref}(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
 As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
@@
 We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \citet{ref}(tab:scenlitterpnv).
+We test however a range of different plant types as specified in Table \ref{tab:scenlitterpnv}.
 
@@
 Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
 }\label{fig:SOCmaps}
@@
 Focusing of the temporal dimension of SOC stock changes (see Fig.
-\ref{fig:SOCmaps}(c)), global SOC debt has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\).
+\ref{fig:SOCmaps}(c)), global \(\Delta SOC\) has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\).
 This leads to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\) and considering our estimate of the global SOC stock of \(675\unit{GtC}\) in 2010 to \(0.2\) per mille per year.
 
-\hypertarget{agricultural-management-effects-on-soc-debt}{%
-\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
+\hypertarget{agricultural-management-effects-on-delta-soc}{%
+\subsection{\texorpdfstring{Agricultural management effects on \(\Delta SOC\)}{Agricultural management effects on \textbackslash Delta SOC}}\label{agricultural-management-effects-on-delta-soc}}
 
 Figure \ref{fig:SOCscen}(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices.
-Without changes in management regimes, especially in C inflows from crop residues to the soil, global SOC debt on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\).
+Without changes in management regimes, especially in C inflows from crop residues to the soil, global \(\Delta SOC\) on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\).
 The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of \(0.06\unit{GtC yr^{-1}}\).
@@
 
-\begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+\begin{figure}
+\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
 ConstManagement combines all three modifications.
@@
 The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig.
-\ref{fig:SOcinit}.
+\ref{fig:SOCinit}.
 The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
+The \(\Delta SOC\) estimation almost halves (from \(~26\unit{GtC}\) to \(~14\unit{GtC}\)) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \(~4\unit{GtC}\) to \(~2.5\unit{GtC}\)) for the Inital-natveg scenario.
 
-\begin{figure}[H]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
+\begin{figure}
+\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
 \end{figure}
-
-Figure \ref{fig:}
 
@@
 
-\begin{figure}[h]
+\begin{figure}
 \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.}\label{fig:FlowFig}
@@
 \citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a~substantial decline in SOC debt (less than 1\% change).
+The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a substantial decline in SOC debt (less than 1\% change).
 
@@
 High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs.
-As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
+As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
 \citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
@@
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes, that most of the SOC dynamic happens in the topsoil.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes, that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are over simplified within our approach.
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) which are based on measurement data for croplands (see Table \ref{tab:SCFtable}).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) which are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\ref{append:climatemap}).
+\ref{fig:CLIMzone}).
 
-\textbackslash begin\{figure\}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/TableSCF_comparison} \textbackslash caption\{This table shows different estimates of stock change factors \(F_{SCF}\), defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation.
-We compare average values for four IPCC climate zone classifications.
-IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in \citep{eggleston_hs_ipcc_2006}, \citep{calvo_buendia_ipcc_2019} are compared to 1990 and 2010 results for historic input levels of this study.\}\label{fig:SCFtable}
-\textbackslash end\{figure\}
+\% latex table generated in R 3.6.3 by xtable 1.8-4 package
+\% Sun Dec 13 16:18:11 2020
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlllllll}
+  \hline
+ & Source & Input & Year & tropical moist & tropical dry & temperate dry & temperate moist \\ 
+  \hline
+1 & IPCC2006 & low & invariant & 0.44 & 0.55--0.61 & 0.74 & 0.66 \\ 
+  2 & IPCC2006 & medium & invariant & 0.48 & 0.58--0.64 & 0.80 & 0.69 \\ 
+  3 & IPCC2006 & high & invariant & 0.53 & 0.60--0.67 & 0.83 & 0.77 \\ 
+  4 & IPCC2019 & low & invariant & 0.76 & 0.87 & 0.70--0.71 & 0.66--0.67 \\ 
+  5 & IPCC2019 & medium & invariant & 0.83 & 0.92 & 0.76--0.77 & 0.69--0.70 \\ 
+  6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
+  7 & This Study & hist & 1990 & 0.57 & 0.61 & 0.78 & 0.76 \\ 
+  8 & This Study & hist & 2010 & 0.64 & 0.68 & 0.83 & 0.83 \\ 
+   \hline
+\end{tabular}
+\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
+Also modelled $F^{\mathrm{SCF}}$ have increased for all climates over time.}
+\end{table}
 
 Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}).
-Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_hs_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), leaving our tropical results in between.
+Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), leaving our tropical results in between.
 Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially.
@@
 
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}, \citep{schaphoff_lpjml4_2018-1}) in Table \ref{tab:SOCtable}).
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}, \citep{schaphoff_lpjml4_2018-1}) in Table \ref{fig:SOCglo}).
 
@@
 This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCtable}
+In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
 \end{figure}
@@
 Looking at regional results in Fig.
-\ref{append:regcompare}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
+\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
 To avoid that this bias influences our results, our study focuses on SOC changes on cropland.
@@
 We conducted a sensitivity analysis (Fig.
-\ref{append:senslitterpnv} based on various plant parameterizations from the Century model (see Sect.
-@ref(sec:scen\_litterpnv)).
+\ref{fig:SOClitter}) based on various plant parameterizations from the Century model (see Sect.
+\ref{sec:scenlitterpnv}).
 This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC is not altered under various estimates for natural SOC stocks.
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{append:climatemap}{%
-\subsection{map on climate zone used for SCF}\label{append:climatemap}}
-
-\hypertarget{append:Tablec2dm}{%
-\subsection{carbon 2 dry matter}\label{append:Tablec2dm}}
-
-\hypertarget{append:TableclossAWMS}{%
-\subsection{closs in AWMS - Table}\label{append:TableclossAWMS}}
-
-\hypertarget{append:regcompare}{%
-\subsection{regional SOC stock in GtC from different sources}\label{append:regcompare}}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{../ResultNotebooks/Output/Images/reg_comparisonfigure} \caption{Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.}\label{fig:regSOCtable}
-\end{figure}
-
-with
-Canada, Australia and New Zealand: CAZ;
-China: CHA;
-European Union: EUR;
-India: IND;
-Japan: JPN;
-Latin America: LAM;
-Middle East and north Africa: MEA;
-non-EU member states: NEU;
-other Asia: OAS; reforming countries: REF; Sub-Saharan Africa:
-SSA; United States: USA.
-
-\hypertarget{append:senslitterpnv}{%
-\subsection{global SOC debt in GtC for various litter parameterizations}\label{append:senslitterpnv}}
-
-\newpage
-
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions), github.com/pik-piam/mrsoil (soil dynamic related functions) and github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing.
+\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data, which are both based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
 Model results are accessable under (zenode\_link).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+\section{Figures and tables in appendices}
+
+\subsection{Methods}
+
+\begin{figure}
+\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
+\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
+Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
+\label{fig:CLIMzone}
+\end{figure}
+
+\begin{table*}[h]
+\caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
+Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
+}
+\begin{tabular}{lllllllllllll}
+\tophline
+
+         &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
+Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
+ \middlehline
+tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
+maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
+trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
+rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
+soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
+rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
+groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
+sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
+oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
+puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
+potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
+cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
+sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
+sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
+others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
+foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
+cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
+\bottomhline
+         &                 & \multicolumn{8}{l}{
+                                           \begin{minipage}[t]{0.50\columnwidth}
+                                                      nr/dm -- nitrogen to dry matter ratio\\
+                                                      wm/dm -- wet matter to dry matter ratio\\
+                                                      c/dm  -- carbon to dry matter ratio\\
+                                                      \end{minipage}} &
+                                                                        \multicolumn{3}{l}{
+                                                                            \begin{minipage}[t]{0.50\columnwidth}
+                                                                                        HI\textsuperscript{area} -- harvest index per area\\
+                                                                                        HI\textsuperscript{prod} -- harvest index per production\\
+                                                                                        RS -- root:shoot ratio \end{minipage}} 
+\end{tabular}
+\label{tab:c2dm}
+\end{table*}
+
+\subsection{Discussion}
+
+\begin{figure}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
+\caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
+In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.
+PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
+We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
+\label{fig:SOCreg}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
+\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
+\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
+Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
+\label{fig:SOClitter}
+\end{figure}
+\noappendix
 

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd

@@
     affiliation: 1
+    orcid: https://orcid.org/0000-0002-7651-6930
   - given_name: Hermann
@@
 We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have around 26 GtC relative to a counterfactual natural state in 1975.
+We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
 Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
 We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
+
 bibliography: SOCbudget.bib
@@
   codedata: |
-   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data, which are both based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
+   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
+All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
 Model results are accessable under (zenode_link).
@@
 authorcontribution: |
-  KK wrote code and paper build on work of BLB, IW. 
+  KK, BLB and AP designed the study and the model idea.
+KK wrote the code build on work of BLB, IW.
 JPD revised and improved the model code.
-  BB, CM, MK, IW, JH, SR, MD, SM, HLC, AP revised paper.
-disclaimer: |
-  We like Copernicus.
+CM, JH and SR provided the LPJmL simulation data.
+KK wrote the paper with important contributions of BLB and CM.
+MK, JS, SR and IW provided extensive feedback.
+All authors discussed the results and commented on the manuscript.
+
 acknowledgements: |

### 2020-12-14 SOCbudgetPaper/discussion.Rmd

--- 2020-12-14 SOCbudgetPaper/discussion.Rmd
@@
 
+<!--Needed???
 Looking on historical cropland intensification rates of above 50% ([@rudel_agricultural_2009]), a large fraction of increased C input from residue biomass, is attributed to productivity improvements.
+-->
 
@@
 the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
+As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study).
 [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for belowground biomass.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
 However, looking on the evaluation of management effects (see Sect.
@@
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_hs_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]).
-Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFtable)).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]).
+Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\@ref(append:climatemap)).
+\@ref(fig:CLIMzone)).
 
-```{r captionSCFtable, echo=FALSE}
-captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation.
-We compare average values for four IPCC climate zone classifications.
-IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in [@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019] are compared to 1990 and 2010 results for historic input levels of this study."
-```
-
-```{r SCFtable, out.width = "16cm", fig.cap = captionSCFtable, echo=FALSE}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
+```{r SCFglo, echo=FALSE, results = 'asis'}
+table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
+library(xtable)
+print(xtable(table$comparison_table, digit=2, comment=FALSE, 
+              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent.
+Also modelled $F^{SOC}$ have increased for all climates over time."))
 ```
@@
 Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than [@calvo_buendia_ipcc_2019].
-The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
+The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
 Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
@@
 The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids ([@hengl_soilgrids250m_2017]) especially stands out for their high estimation, whereas all other sources (including out study) are comparabely similar.
+SoilGrids ([@hengl_soilgrids250m_2017]) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
 Looking at regional results in Fig.
-\@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
+\@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
 To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
@@
 We conducted a sensitivity analysis (Fig.
-\@ref(append:senslitterpnv) based on various plant parameterizations from the Century model (see Sect.
-\@ref(sec:scen_litterpnv)).
+\@ref(fig:SOClitter)) based on various plant parameterizations from the Century model (see Sect.
+\@ref(sec:scenlitterpnv)).
 This shows that the general trend of decreasing SOC debt of ~4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.

### 2020-12-14 SOCbudgetPaper/introduction.Rmd

@@
 Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest ([@strassmann_simulating_2008], [@betts_climate_2015]).
-BKMs are designed to estimate LUC realted emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management ([@friedlingstein_global_2019], [@houghton_carbon_2012], [@hansis_relevance_2015]).
+BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management ([@friedlingstein_global_2019], [@houghton_carbon_2012], [@hansis_relevance_2015]).
 
@@
 
-<!--Recent studies on soil sequestration for net greenhouse gas removal ([@paustian_quantifying_2019], [@smith_how_2020]) therefore called for 'well-integrated data-model frameworks' including high quality management data to quantify and monitor SOC stock changes.-->
-
 Field-scale models (@del_grosso_simulated_2001], [@coleman_simulating_1997],[@smith_estimating_2010], [@taghizadeh-toosi_c-tool_2014]) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
@@
 The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historical agricultural management.
-To achieve this objective we create and provide a comprehensive dataset of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
 We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
 Using a scenario analysis, we decompose the contribution of different management activities.
-Our analysis highlights the importance of considering agricultural management for improving estimates of SOC stocks.
-We conclude, that historic agricultural management let to increasing SOC stocks in the most recent decades.

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
working pdf complilation

@@
   codedata: |
-   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
+   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
 All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
working pdf complilation

@@
 Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest (\citep{strassmann_simulating_2008}, \citep{betts_climate_2015}).
-BKMs are designed to estimate LUC realted emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management (\citep{friedlingstein_global_2019}, \citep{houghton_carbon_2012}, \citep{hansis_relevance_2015}).
+BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management (\citep{friedlingstein_global_2019}, \citep{houghton_carbon_2012}, \citep{hansis_relevance_2015}).
 
@@
 The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historical agricultural management.
-To achieve this objective we create and provide a comprehensive dataset of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
 We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
 Using a scenario analysis, we decompose the contribution of different management activities.
-Our analysis highlights the importance of considering agricultural management for improving estimates of SOC stocks.
-We conclude, that historic agricultural management let to increasing SOC stocks in the most recent decades.
 \newpage
@@
 In Sect.
-\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{ipcc_eggleston_2006}, \citep{calvo_buendia_ipcc_2019}).
-Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citet{mrcommons_2020}).
+\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
+Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citep{bodirsky_mrcommons_2020}).
 In Sect.
@@
 (1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
-(2) we account for land conversion by transferring \(SOC\) from and natural vegetation (Sect. \ref{sec:carbontransfer}),
+(2) we account for land conversion by transferring \(SOC\) from and to natural vegetation (Sect. \ref{sec:carbontransfer}),
 (3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
@@
 This equation is valid for all grid cells \(i\) and all years \(t\).
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep[(][]{parton_analysis_1987}).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model (\citep{parton_analysis_1987}).
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
@@
 
-Carbon inputs for cropland are below and above ground residues left or returned to the field (see Sect.
+Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
 \ref{sec:residues}) and manure inputs (see Sect.
@@
   \middlehline
-  \multirow{3}{*}{Cropland} & above ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+  \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                       \cite{faostat_faostat_2016}, \\
@@
                                                       \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
-                                                                      lg/c generic values based on Table 5.5B, 5.5C \\
+                                                                      LG:C generic values according to Table 5.5B\\
                                                                       in \cite{calvo_buendia_ipcc_2019},\\
-                                                                      crop-type specific n/c-ratios from
+                                                                      crop-type specific N:C-ratios from
                                                                       \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
-                            & below ground residues &  &  \\ 
+                            & below-ground residues &  &  \\
                             & manure                &  &  \\ 
@@
 Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
-Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land use change emission (\citep{pugh_simulated_2015}).
+Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land-use change emissions (\citep{pugh_simulated_2015}).
 
@@
 We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which, together with the spin-up phase, restricts our analysis to the time span from 1975 to 2010.
+The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
 For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
@@
 We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
-Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallow land.
+By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
 
@@
 Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
-Note that deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+Deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
 We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
 
-A fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
+A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
 Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
@@
 The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field.
-We limit high returning rates at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
+We limit high residue return rates to at most \(10\unit{tC\,ha}^{-1}\) in order to correct for outliers.
 
@@
 Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed c/n ratios (given by \citep{calvo_buendia_ipcc_2019}).
+Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see \citep{calvo_buendia_ipcc_2019}).
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 As a result \(wat\) is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a (partly) irrigated grid cells, since we do not have crop group specific growing periods at hand.
-In addition, we expect, either water limitations to be no problem during the off-season (causing our assumption to be a good approximation), or irrigated growing periods to span over the whole year in water-restricted areas.
+Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
+This will lead to an overestimation of the irrigation effect.
+We expect, however, water limitations to be a minor problem during the off-season in temperature limited cropping regions, causing our assumption to not dramatically overestimate the moisture effects.
+In tropical, water-limited cropping areas, irrigated growing periods might even span over the whole year.
 
@@
 
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for the three drivers: (1) In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set); (2) the \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
+To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for these three drivers: In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set).
+The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
 manure (in \(\tfrac{t}{ha}\)) at the level of 1975 onwards.
-Within the \textit{constResidue} scenario at least three effects overlay each other: (1) yields and with them residue biomass increase due to productivity gains, (2) rates of residue left or returned to fields are raising, and (3) shifts of crop groups change the amount of residue biomass due to crop-group specific harvest index values.
-The \textit{constManagement} combines all three scenarios.
+Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
+The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
 
@@
 Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
-Whereas lg/c ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, n/c ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
+Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
 We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
@@
 
-\hypertarget{agricultural-management-matters}{%
-\subsection{Agricultural management matters}\label{agricultural-management-matters}}
-
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to constant management assumptions.
-This result remains robust to variations of central model parameters and initialization stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to net reduction of the SOC debt.
-
-The evaluation of SOC stock gains and losses, however, is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 the baseline and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink according to our study.
-Annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to the often used constant management assumptions.
+This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
+While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
 
@@
 
-According to \citep{sanderman_soil_2017} SOC debt since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing.
+The evaluation of SOC stock gains and losses, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study.
+However, annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+
+According to \citep{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
 Our estimate of 22 GtC in 2010 for cropland debt is higher as \citep{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and the \citep{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results, suggesting that our results are in an acceptable range.
+However, there are large uncertainties in modeling SOC at the global scale, and \citep{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
+This suggests that our results are within a plausible range.
 
 Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account.
-The study moreover concludes that crop productivity gains (increasing yield levels by 18\%) do not lead to a substantial decline in SOC debt (less than 1\% change).
+\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account.
+Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
+Historic productivity increases were, however, notably larger.
+Despite large spatial heterogeneity aggregate yield increase rates are often estimate well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).
 
@@
 
-Moreover, our sensitivity analysis indicates that (1) higher returned residue biomass driver by yield increases play the most important role, followed by (2) greater manure applications due to an increasing livestock sector as well as improved animal waste management systems, and (3) the adoption of no tillage practices.
-Unfortunately, disentangling drivers of higher residue and manure inputs remains unsolved within this study.
-
 \hypertarget{limitations-and-uncertainties}{%
@@
 
-Modeling management effects at the global scale, however, comes with parametric and structural uncertainties.
-High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs.
-As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
+Modeling management effects at the global scale comes with parametric and structural uncertainties.
+High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
+As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
 \citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding.
-However, with an evaluation of management effects (see Sect.
-\ref{sec:ipcccompare}), there is no indication for a systematic overrating of residue biomass.
-Rather, it is also likely that we are still missing carbon inputs e.g.~from cover- and inter-cropping to the soil.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
+However, looking on the evaluation of management effects (see Sect.
+\ref{sec:ipcccompare}), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
+Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.~cover- and inter-cropping practices.
 
-Another aspect concerns the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
+Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
 As shown in Fig.
-\ref{fig:SOCinit} the SOC debt estimation halves (from \textasciitilde26 GtC to \textasciitilde14 GtC) as well as the decrease in SOC debt (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
+\ref{fig:SOCinit} the SOC debt estimate almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), and the SOC debt reduction is strongly reduced (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
 \citep{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results show that these legacy fluxes are compensated by agricultural management changes even under the strong assumption of the Initial-natveg scenario.
-However, it remains open, if our simple, depth-restricted SOC model itself can represent these fluxes correctly.
+Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes, that most of the SOC dynamic happens in the topsoil.
-In this regard several aspects are over simplified within our approach.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes that most of the SOC dynamic happens in the topsoil.
+In this regard several aspects are strongly simplified within our approach.
 Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (\citep{powlson_limited_2014}) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.
+Secondly changes to the subsoil due to tillage are neglected.
+As \citep{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
+SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
+Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
 
@@
 However, as pointed out by \citep{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions (\citep{wang_human-induced_2017}).
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC stayed on croplands.
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
 Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
 Human interventions have led to large changes in SOC stocks there as well (\citep{sanderman_soil_2017},\citep{friedlingstein_global_2019}).
-This study does not try to be an comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
+This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
 
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) which are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 (\citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}).
+Both estimates are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
@@
 \% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Sun Dec 13 16:18:11 2020
+\% Mon Dec 14 03:22:10 2020
 
@@
 Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{\mathrm{SCF}}$ have increased for all climates over time.}
+Also modelled $F^{SOC}$ have increased for all climates over time.}
 \end{table}
 
-Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}).
-Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), leaving our tropical results in between.
-Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially.
-Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
-Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\ref{fig:SOCmaps}(c)) might be not surprising.
+Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than \citep{calvo_buendia_ipcc_2019}.
+The tropical factors differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), with our estimates within this range.
+Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
+Our results still show especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
+Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
+\ref{fig:SOCmaps}(c)) may be expected.
 
@@
 
-The global estimates of SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+SoilGrids (\citep{hengl_soilgrids250m_2017}) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
 Looking at regional results in Fig.
-\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions.
-To avoid that this bias influences our results, our study focuses on SOC changes on cropland.
-Pristine natural vegetated areas without human land management cancel out in calculation of \(\Delta SOC\) and \(F^\mathrm{SCF}\).
+\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
+To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of \(\Delta SOC\) and \(F^\mathrm{SCF}\).
 
-Additionally, our estimates for total SOC stocks of the world, as well as our SOC initialization, are dominated by the representation of natural vegetation, which are, however, only estimated in a basic manner.
-For example, we do not differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This leaves carbon inputs and decay behavior for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be used with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land may arise especially in less forested areas.
+Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
+For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
+This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
+The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
+As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
 
@@
 \ref{sec:scenlitterpnv}).
-This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC is not altered under various estimates for natural SOC stocks.
+This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.
 \newpage
@@
 
-Outlook including perspective on mitigation (and soc enhancement in the future).
-\newpage
+Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
+It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historical agricultural management on global SOC stocks.
 
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data, which are both based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
+With the extraction of C at harvest and little efforts to increase cropland SOC, these stocks are depleted in comparison to natural vegetation SOC stocks in most regions.
+The reduction of further deforestation for cropland expansion thus need to be stopped quickly.
+
+However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
+Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
+The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source.
+Under the assumption that cropland management is static in time, as typically assumed in other studies on cropland SOC stocks, we also find that cropland soils are a source of CO\textsubscript{2}, building trust in our method.
+
+The estimated increase in cropland SOC is caused by changing agricultural management.
+The largest contribution to these dynamics comes from increased carbon inputs to soils from crop residues.
+Crop residue management can still be improved from the SOC perspective, e.g.~by reducing shares of residue burning or further improvements in agricultural productivity.
+Similarly, other carbon-accumulating practices, such as the application of cover crops need to be included in studies of cropland SOC dynamics, if data on these practices become available.
+
+Our findings on sequestration rates are, however, still more than an order of magnitude lower than promoted by the 4 per mille initiative.
+We conclude, that while crop productivity gains, lead to co-benefits for SOC stocks, they are not enough to help meeting ambitious climate mitigation targets.
+Soil sequestration techniques in areas with low stock change factors still might focus on closing yield gaps, whereas other SOC conserving technologies (e.g.~conservation agriculture, biochar application and agroforestry) and the avoidance of further deforestation need to be considered elsewhere.
+
+Reduced or avoided deforestation would not only reduce SOC losses on newly converted land, it would likely also require further productivity improvements in existing croplands to meet -- and thereby usually further reduce the SOC debt on croplands.
+
+Our results clearly demonstrate that agricultural management aspects need to be explicitly considered in global carbon assessments and models.
+That also implies, that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
+Our data-set and the MADRaT tool constitute a starting point for building comprehensive data sets on agricultural management aspects.
+
+\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
+All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
 Model results are accessable under (zenode\_link).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{KK wrote code and paper build on work of BLB, IW.
+\authorcontribution{KK, BLB and AP designed the study and the model idea.
+KK wrote the code build on work of BLB, IW.
 JPD revised and improved the model code.
-BB, CM, MK, IW, JH, SR, MD, SM, HLC, AP revised paper.} %% optional section
+CM, JH and SR provided the LPJmL simulation data.
+KK wrote the paper with important contributions of BLB and CM.
+MK, JS, SR and IW provided extensive feedback.
+All authors discussed the results and commented on the manuscript.} %% optional section
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\disclaimer{We like Copernicus.} %% optional section
 

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
working pdf complilation

@@
 library(xtable)
-print(xtable(table$comparison_table, digit=2, comment=FALSE, 
+print(xtable(table$comparison_table, digit=2, comment=FALSE, label="#tab:SCFglo", 
               caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.

### 2020-12-14 ResultNotebooks/results_SOCmaps.Rmd
nce pdf, update bib + type in maps

@@
     scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression(paste('croplands \u0394SOC in tons of carbon per hectare')), 
+                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare')), 
                       values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
nce pdf, update bib + type in maps

@@
 \% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 03:22:10 2020
+\% Mon Dec 14 03:39:00 2020
 
@@
 Also modelled $F^{SOC}$ have increased for all climates over time.}
+\label{tab:SCFglo}
 \end{table}
@@
 
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
+\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
 All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
nce pdf, update bib + type in maps

--- 2020-12-14 SOCbudgetPaper/discussion.Rmd
@@
 library(xtable)
-print(xtable(table$comparison_table, digit=2, comment=FALSE, 
+print(xtable(table$comparison_table, digit=2, comment=FALSE, label="tab:SCFglo", 
               caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.

### 2020-12-14 ResultNotebooks/results_SOCcompare.Rmd

@@
 
-```{r kable2image}
-# library(kableExtra)
-# 
-# table <- readRDS("Output/gloSOC_comparison.rds")
-# colnames(table)[1] <- "Model/Database"
-# colnames(table)[6] <- "Global SOC in GtC"
-# save_kable(kable_styling(kable(table$comparison_table[c(1,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
-#          file="Output/Images/TableSOC_comparison.png")
-# 
-# save_kable(kable_styling(kable(table$comparison_table[setdiff(1:13,5)], "latex", digit=2), latex_options = c("striped", "scale_down")), 
-#          file="Output/Images/TableRegSOC_comparison.png")
-
-```
-
-Plot regional stocks
-
 ```{r point plot reg stock, out.width=800, out.height=600}
 library(ggplot2)
+library(knitr)
+library(stats)
+library(reshape2)
 
@@
 
-glob <- ggplot(data = melt(subset(table, select =c(Data1, GLO, Year))), aes(x=Year,y=value, group=Data1)) + 
+glob <- ggplot(data = melt(subset(table, select =c(Data1, GLO, Year))), aes(y=Year,x=value, group=Data1)) + 
   geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
@@
   labs(color  = leg_name, size = leg_name, shape = leg_name) +
-  xlab("") + ylab("SOC (0-30cm) in GtC") +
-  theme(axis.ticks.x=element_blank(),
-        axis.text.x=element_blank(), 
-        legend.position="right") + guides(fill=guide_legend(ncol=1,byrow=TRUE))
-        #legend.position="bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))
+  ylab("") + xlab("SOC (0-30cm) in GtC") +
+  theme(axis.ticks.y =element_blank(),
+        axis.text.y=element_blank(), 
+        legend.position="bottom") + 
+  guides(fill=guide_legend(nrow=2,byrow=TRUE))
+        #legend.position="right") + guides(fill=guide_legend(ncol=1,byrow=TRUE))
 
-comp <- ggplot(data = melt(subset(table, select=-GLO)), aes(x=Year,y=value, group=Data1)) + 
+comp <- ggplot(data = melt(subset(table, select=-GLO)), aes(y=Year,x=value, group=Data1)) + 
   geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
-  facet_wrap(~ variable, scales="free", drop=TRUE, labeller = label_wrap_gen(width=8)) + 
+  facet_wrap(~ variable, scales="free", drop=TRUE, labeller = label_wrap_gen(width=18)) + 
   scale_size_manual(values=c(4,2,2,2,2,2)) + 
@@
   labs(color  = leg_name, size = leg_name, shape = leg_name) +
-  xlab("") + ylab("SOC (0-30cm) in GtC") +
-  theme(axis.ticks.x=element_blank(),
-        axis.text.x=element_blank())
+  ylab("") + xlab("SOC (0-30cm) in GtC") +
+  theme(axis.ticks.y=element_blank(),
+        axis.text.y=element_blank(),
+        legend.position = "bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))
 
@@
 
-ggsave(paste0(outputdir,"/Images/glo_comparisonfigure.png"), plot=glob, height=8, width=8, units="cm", dpi = 300)
-ggsave(paste0(outputdir,"/Images/reg_comparisonfigure.png"), plot=comp, height=10, width=14, units="cm", dpi = 300)
+ggsave(paste0(outputdir,"/Images/glo_comparisonfigure.png"), plot=glob, height=4, width=10, units="cm", dpi = 300)
+ggsave(paste0(outputdir,"/Images/reg_comparisonfigure.png"), plot=comp, height=9, width=15, units="cm", dpi = 300)
 

### 2020-12-14 ResultNotebooks/results_SOCflows.Rmd

@@
 ```{r load}
-library(mrcommons)
-library(mrvalidation)
-library(mrsoil)
+library(magclass)
+library(madrat)
 library(magpiesets)
 
-rev <- "rev0.78baseline"
-
 outputdir <- "Output/Images"
-
-setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/Baseline", extras="--keep-newer-files ")
-
-revfolder <- "Data/Baseline/"
+revfolder <- "Data/histManagement/"
 
@@
 
-files <- list.files("Data")
+year  <- "y2010"
+years <- c(paste0("y",as.numeric(substring("y2010",2))-1), year)
 
-CarbonInput      <- readRDS(paste0(revfolder,"CarbonInput.rds"))
-CarbonLitter     <- readRDS(paste0(revfolder,"CarbonLitter.rds"))
-CarbonManure     <- readRDS(paste0(revfolder,"CarbonManure.rds"))          
-CarbonResidues   <- readRDS(paste0(revfolder,"CarbonResidues.rds"))
-CarbonHarvest    <- readRDS(paste0(revfolder,"Crop_Harvest.rds"))
-SoilCarbonOut    <- readRDS(paste0(revfolder,"SoilCarbon.rds"))
+CarbonInput      <- readRDS(paste0(revfolder,"CarbonInput.rds"))[,year,]
+CarbonLitter     <- readRDS(paste0(revfolder,"CarbonLitter.rds"))[,year,]
+CarbonManure     <- readRDS(paste0(revfolder,"CarbonManure.rds"))[,year,]          
+CarbonResidues   <- readRDS(paste0(revfolder,"CarbonResidues.rds"))[,year,]
+CarbonHarvest    <- readRDS(paste0(revfolder,"Crop_Harvest.rds"))[,year,]
+CarbonVegetation <- readRDS(paste0(revfolder,"CarbonVegetation.rds"))[,years,]
+SoilCarbonOut    <- readRDS(paste0(revfolder,"SoilCarbon.rds"))[,years,]
+CarbonInflow      <- readRDS(paste0(revfolder,"CarbonInflow.rds"))[,years,]
 
-Landuse          <- readRDS(paste0(revfolder,"Landuse.rds"))
-LanduseChange    <- readRDS(paste0(revfolder,"LanduseChange.rds"))
+Landuse          <- readRDS(paste0(revfolder,"Landuse.rds"))[,years,]
+LanduseChange    <- readRDS(paste0(revfolder,"LanduseChange.rds"))[,years,]
 
-FAOmassbalance   <- readRDS(paste0(revfolder,"FAOmassbalance_ISO.rds"))
-ManureExcreted   <- readRDS(paste0(revfolder,"Manure_excreted.rds"))
-ManureRecycled   <- readRDS(paste0(revfolder,"Manure_recycled.rds"))
-ResidueBiomass   <- readRDS(paste0(revfolder,"Residue_Biomass.rds"))
-ResiduesAgDemand <- readRDS(paste0(revfolder,"ResiduesAg_FieldBalance.rds"))
+FAOmassbalance   <- readRDS(paste0(revfolder,"FAOmassbalance_ISO.rds"))[,year,"c"]
+ManureExcreted   <- readRDS(paste0(revfolder,"Manure_excreted.rds"))[,year,"c"]
+ManureRecycled   <- readRDS(paste0(revfolder,"Manure_recycled.rds"))[,year,"c"]
+ResidueBiomass   <- readRDS(paste0(revfolder,"Residue_Biomass.rds"))[,year,"c"]
+ResiduesAgDemand <- readRDS(paste0(revfolder,"ResiduesAg_FieldBalance.rds"))[,year,"c"]
 
-```
+grepSpaceTimeC    <- function(x){return(collapseNames(dimSums(x[,year,"c"],dim=1)))}
+grepSpaceTime     <- function(x){return(collapseNames(dimSums(x[,year,],dim=1)))}
 
-```{r globals}
-grepSpaceTime     <- function(x){return(dimSums(x[,"y2010","c"],dim=1))}
+TotalSOCBudget    <- NULL
 
@@
 
-CropHANPP         <- dimSums(grepSpaceTime(FAOmassbalance[,,kcr][,,c("food","waste","other_util","bioenergy","feed",
+CropHANPP         <- dimSums(grepSpaceTimeC(FAOmassbalance[,,kcr][,,c("food","waste","other_util","bioenergy","feed",
                                                                      "milling","refining","extracting","fermentation","distilling")]), dim=3)
-CropFEED          <- grepSpaceTime(dimSums(FAOmassbalance[,,kExResPast][,,"feed"], dim=3.1))
+
+CropFEED          <- grepSpaceTimeC(dimSums(FAOmassbalance[,,kExResPast][,,"feed"], dim=3.1))
 CropPRIM          <- CropHANPP-CropFEED
-ResFEED           <- grepSpaceTime(dimSums(FAOmassbalance[,,kres][,,"feed"], dim=3.1))
-ResPRIM           <- dimSums(grepSpaceTime(FAOmassbalance[,,kres][,,c("other_util","bioenergy")]), dim=3)
 
-FEED              <- grepSpaceTime(dimSums(FAOmassbalance[,,kPrim][,,"feed"], dim=3.1))
-LivstHANPP        <- dimSums(grepSpaceTime(FAOmassbalance[,,kli][,,c("food","waste","other_util")]), dim=3)
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(CropFEED, "crop_feedintake"),
+                           setNames(CropPRIM, "crop_humandemand"))
+
+ResFEED           <- grepSpaceTimeC(dimSums(FAOmassbalance[,,kres][,,"feed"], dim=3.1))
+ResPRIM           <- dimSums(grepSpaceTimeC(FAOmassbalance[,,kres][,,c("other_util","bioenergy")]), dim=3)
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(ResFEED, "ag_feedintake"),
+                           setNames(ResPRIM, "ag_humandemand"))
+
+FEED              <- grepSpaceTimeC(dimSums(FAOmassbalance[,,kPrim][,,"feed"], dim=3.1))
+LivstHANPP        <- dimSums(grepSpaceTimeC(FAOmassbalance[,,kli][,,c("food","waste","other_util")]), dim=3)
 LivstCropHANPP    <- (CropFEED+ResFEED)/FEED * LivstHANPP
 
-ResidueBiomassTot <- grepSpaceTime(dimSums(ResidueBiomass, dim=3.2))
-ResidueRecycledTot<- grepSpaceTime(dimSums(ResiduesAgDemand[,,"recycle"], dim=3.2))
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(LivstCropHANPP, "livst_humandemand"))
 
-ManureExcretedTot <- grepSpaceTime(dimSums(ManureExcreted[,,c("stubble_grazing","confinement")], dim=c(3.1,3.2)))
-ManureRecycledTot <- grepSpaceTime(dimSums(ManureRecycled, dim=3.1))
+ResidueBiomassTot <- grepSpaceTimeC(dimSums(ResidueBiomass, dim=3.2))
+ResidueRecycledTot<- grepSpaceTimeC(dimSums(ResiduesAgDemand[,,"recycle"], dim=3.2))
+ResidueLossTot    <- ResidueBiomassTot[,,"ag"] - ResidueRecycledTot - ResFEED - ResPRIM
 
-#Soil
+ResidueInput      <- grepSpaceTimeC(Landuse[,2,"crop"]* CarbonResidues)
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(ResidueBiomassTot[,,c("ag","bg")], c("ag_prod","bg_prod")),
+                           #setNames(ResidueRecycledTot, "ag_recycle"),
+                           #setNames(ResidueBiomassTot[,,"bg"], "bg_recycle"),
+                           setNames(ResidueInput[,,c("ag","bg")], c("ag_recycle","bg_recycle")),
+                           setNames(ResidueLossTot, "ag_loss"))
+
+ManureExcretedTot <- grepSpaceTimeC(dimSums(ManureExcreted[,,c("stubble_grazing","confinement")], dim=c(3.1,3.2)))
+#ManureRecycledTot <- grepSpaceTimeC(dimSums(ManureRecycled, dim=3.1))
+ManureInput       <- dimSums(grepSpaceTimeC(Landuse[,2,"crop"]* CarbonManure), dim=3.1)
+ManureLossTot     <- ManureExcretedTot - ManureInput
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(ManureExcretedTot, "manure_prod"),
+                           setNames(ManureInput, "manure_recycle"),
+                           setNames(ManureLossTot    , "manure_loss"))
+
+LitterRecycleTot  <- grepSpaceTimeC(CarbonLitter)
+
+CropBiomassTot    <- dimSums(ResidueBiomassTot, dim=3) + dimSums(grepSpaceTimeC(FAOmassbalance[,,kcr][,,"production"]),dim=3)
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(CropBiomassTot, "crop_prod"))
+
+LivstCropLoss     <- CropFEED - ManureExcretedTot - LivstCropHANPP 
+CropPRIMLoss      <- CropPRIM + ResPRIM + LivstCropHANPP - dimSums(grepSpaceTimeC(FAOmassbalance[,,c(kcr,kli,kres)][,,"other_util"]), dim=3) # only material use is not re-emitted
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(LivstCropLoss, "livst_loss"),
+                           setNames(CropPRIMLoss,  "humandemand_loss"))
 
 ### Carbon To
-CarbonTo            <- dimSums(toolConditionalReplace(SoilCarbonOut[,"y2010","carbontransfer"], "<0", 0),dim=c(1,3.3))
-CarbonToEff         <- dimSums(SoilCarbonOut[,"y2010","carbontransfer"], dim=c(1,3.3))
-CarbonToPools       <- dimSums(toolConditionalReplace(SoilCarbonOut[,"y2010","carbontransfer"], "<0", 0),dim=c(1))
+CarbonTo            <- dimSums(grepSpaceTime(toolConditionalReplace(SoilCarbonOut[,,"carbontransfer"], "<0", 0)), dim=c(3.2))
+CarbonToEff         <- dimSums(grepSpaceTime(SoilCarbonOut[,,"carbontransfer"]), dim=c(3.2))
 
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(CarbonTo[,,"natveg"], "to_natveg"),
+                           setNames(CarbonTo[,,"crop"],   "from_natveg"))
+
+SOCsub            <- grepSpaceTime(SoilCarbonOut[,,"actualstate"]*Landuse)
+SOCtot            <- dimSums(SOCsub, dim=3.2)
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(SOCsub[,,"natveg"], gsub("\\.","_",getNames(SOCsub))),
+                           SOCtot)
+
+CarbonIn          <- grepSpaceTime(SteadyState[,,"alpha_in"]*Landuse)
+
+```
+
+```{r}
 ####### delete when new output on decay is ready
@@
 ```
-
-```{r residues}
-
-### Residues
-  # originally from moinput
-  ResidueDemand       <- dimSums(collapseNames(calcOutput("ResFieldBalancePast", cellular=TRUE, products = "kres", aggregate = FALSE)[,,"c"]), dim=1)
-  ResidueRecyclingAg  <- ResidueDemand[,,"recycle"]
-  ResidueRecyclingBg  <- dimSums(calcOutput("ResBiomass", cellular=TRUE, aggregate = FALSE)[,,"bg"][,,c("c")], dim=c(1,3))
-  ResiduesMoinput     <- dimSums(ResidueRecyclingAg, dim=3) + ResidueRecyclingBg
-
-  # Residues with cut of overflow
-  ResiduesDensities   <- calcOutput("CarbonResidues", aggregate=FALSE)[,,"c"]
-  Landuse             <- calcOutput("Landuse", aggregate=FALSE)
-  ResiduesSOCbudget   <- dimSums(Landuse[,getYears(ResiduesDensities),"crop"]*ResiduesDensities, dim=c(1,3))
-
-  print(mbind(ResiduesMoinput, ResiduesSOCbudget, ResiduesSOCbudget/ResiduesMoinput))
-
-  # Corrected Residues shares for different plant parts  
-  RatioPlantParts     <- mbind(ResidueRecyclingAg, setNames(ResidueRecyclingBg, "res_roots"))/(dimSums(ResidueRecyclingAg, dim=3)+ResidueRecyclingBg)
-
-  ResiduesCorrected   <- collapseNames(RatioPlantParts * ResiduesSOCbudget)
-
-  # total residues and residue demands
-
-  ResidueBg     <- ResidueRecyclingBg
-  ResidueAg     <- dimSums(calcOutput("ResBiomass", cellular=TRUE, aggregate = FALSE)[,,"ag"][,,c("c")], dim=c(1,3))
-  ResidueTotal  <- ResidueAg + ResidueBg
-
-```
-
-Quite a lot of residues get lost, due to cutting extreme values in the later years.
-Cutting algorithm, cuts everything bigger the 95%-percentile.
-How we deal with that? 
-* Should we do a yearly extrem value cutting?
-* Should we investigate that deeper?
-* Should we leave it like this?
-
-```{r manure}
-
-### Manure
-  # originally from moinput
-  ManureApplication   <- calcOutput("ManureRecyclingCroplandPast", products = "kli", cellular = TRUE, aggregate = FALSE)[,,"c"]
-  ManureGrazing       <- calcOutput("Excretion", cellular = TRUE, attributes = "npkc", aggregate = FALSE)[,,"stubble_grazing"][,,"c"]
-  ManureMoinput       <- collapseNames(dimSums(ManureApplication, dim=1) + dimSums(ManureGrazing, dim=1))
-
-  # Manure with cut of overflow
-  ManureDensities     <- calcOutput("CarbonManure", aggregate=FALSE)[,,"c"]
-  Landuse             <- calcOutput("Landuse", aggregate=FALSE)
-  ManureSOCbudget     <- dimSums(Landuse[,getYears(ManureDensities),"crop"]*ManureDensities, dim=1)
-
-  print(mbind(dimSums(ManureMoinput, dim=3), dimSums(ManureSOCbudget, dim=3), dimSums(ManureSOCbudget, dim=3)/dimSums(ManureMoinput, dim=3)))
-
-  ManureCorrected     <- ManureSOCbudget
-
-  ManureExcreted      <- dimSums(calcOutput("Excretion", cellular = TRUE, attributes = "npkc", aggregate = FALSE)[,,"c"][,,c("confinement","stubble_grazing")], dim=c(1,3.3))
-
-  print(mbind(dimSums(ManureExcreted, dim=3), dimSums(ManureMoinput, dim=3), dimSums(ManureCorrected, dim=3)))
-  print(quantile(ManureApplication+ManureGrazing, probs=seq(0,1,0.02)))
-  range(ManureApplication+ManureGrazing)
-```
-
-Same problem here. Manure share is going down.
-
-```{r production}
-
-#Production     <- collapseNames(calcOutput("Production", products="kcr", cellular=TRUE, attributes="dm", irrigation=FALSE, aggregate = FALSE))
-Balance        <- calcOutput("FAOmassbalance", aggregate=FALSE)
-
-CropProduction <- dimSums(Balance[,,findset("kcr")][,,"production"][,,"c"], dim=c(1,3))
-FeedProduction <- dimSums(Balance[,,c("pasture",findset("kap"),findset("kres"),"scp"), invert=TRUE][,,"feed"][,,"c"], dim=c(1,3))
-ResidueFeed    <- dimSums(Balance[,,findset("kres")][,,"feed"][,,"c"], dim=c(1,3))
-
-print(mbind(CropProduction,FeedProduction,ResidueFeed))
-```

### 2020-12-14 ResultNotebooks/results_SOCscenario.Rmd

@@
 Landuse      <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
-glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))
+glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))
 
@@
   labs( x = "Year", 
-        y = "Global SOC input\n in GtC",
+        y = "SOC input in tC per ha\n (global average)",
         fill="") +
@@
   labs( x = "Year", 
-        y = "Global SOC inputs in GtC",
+        y = "SOC input in tC per ha\n (global average)",
         fill="") +

### 2020-12-14 SOCbudgetPaper/conclusion.Rmd
update written part to first official version (woohoo)

@@
 
-The conclusion goes here.
-You can modify the section name with  `\conclusions[modified heading if necessary]`.
+Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
+It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historical agricultural management on global SOC stocks.
+
+With the extraction of C at harvest and little efforts to increase cropland SOC, these stocks are depleted in comparison to natural vegetation SOC stocks in most regions.
+The reduction of further deforestation for cropland expansion thus need to be stopped quickly.
+
+However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
+Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
+The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO~2~ source.
+Under the assumption that cropland management is static in time, as typically assumed in other studies on cropland SOC stocks, we also find that cropland soils are a source of CO~2~, building trust in our method.
+
+The estimated increase in cropland SOC is caused by changing agricultural management.
+The largest contribution to these dynamics comes from increased carbon inputs to soils from crop residues.
+Crop residue management can still be improved from the SOC perspective, e.g.
+by reducing shares of residue burning or further improvements in agricultural productivity.
+Similarly, other carbon-accumulating practices, such as the application of cover crops need to be included in studies of cropland SOC dynamics, if data on these practices become available.
+
+Our findings on sequestration rates are, however, still more than an order of magnitude lower than promoted by the 4 per mille initiative.
+We conclude, that while crop productivity gains, lead to co-benefits for SOC stocks, they are not enough to help meeting ambitious climate mitigation targets.
+Soil sequestration techniques in areas with low stock change factors still might focus on closing yield gaps, whereas other SOC conserving technologies (e.g.
+conservation agriculture, biochar application and agroforestry) and the avoidance of further deforestation need to be considered elsewhere.
+
+Reduced or avoided deforestation would not only reduce SOC losses on newly converted land, it would likely also require further productivity improvements in existing croplands to meet -- and thereby usually further reduce the SOC debt on croplands.
+
+Our results clearly demonstrate that agricultural management aspects need to be explicitly considered in global carbon assessments and models.
+That also implies, that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
+Our data-set and the MADRaT tool constitute a starting point for building comprehensive data sets on agricultural management aspects.

### 2020-12-14 SOCbudgetPaper/methods.Rmd

--- 2020-12-14 SOCbudgetPaper/methods.Rmd
@@
 In Sect.
-\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]).
 Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see [@bodirsky_mrcommons_2020]).
@@
 This equation is valid for all grid cells $i$ and all years $t$.
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987]).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model ([@parton_analysis_1987]).
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
@@
     \middlehline
-    \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+    \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                         \cite{faostat_faostat_2016}, \\
@@
                                                                         \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                        \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
+                                                        \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
                                                                         LG:C generic values according to Table 5.5B\\
@@
                                                                         crop-type specific N:C-ratios from
-                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
-                              & below-ground residues &  &  \\ \cline{2-2}
-                              & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
+                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
+                              & below-ground residues &  &  \\
+                              & manure                &  &  \\ 
                               \hline
@@
 
-Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_hs_ipcc_2006].
-The SOC debt is equivalent to the negated cumulative SOC component of  of human land use emission ([@pugh_simulated_2015]).
+Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_ipcc_2006].
+Moreover, $\Delta SOC$ is equivalent to the negated cumulative SOC component of human land-use change emissions ([@pugh_simulated_2015]).
 
@@
 
-Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_hs_ipcc_2006].
+Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_ipcc_2006].
 Deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
@@
 Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_hs_ipcc_2006].
+Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_ipcc_2006].
 

### 2020-12-14 SOCbudgetPaper/results.Rmd

@@
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
 "
@@
 Focusing of the temporal dimension of SOC stock changes (see Fig.
-\@ref(fig:SOCmaps)(c)), global SOC debt has decreased around 15% over the last couple of decades to $22\unit{GtC}$.
+\@ref(fig:SOCmaps)(c)), global $\Delta SOC$  has decreased around 15% over the last couple of decades to $22\unit{GtC}$.
 This leads to a sequestration rate of $0.11\unit{GtC yr^{-1}}$ and considering our estimate of the global SOC stock of $675\unit{GtC}$ in 2010 to $0.2$ per mille per year.
 
-## Agricultural management effects on SOC debt
+## Agricultural management effects on $\Delta SOC$
 
 Figure \@ref(fig:SOCscen)(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices.
-Without changes in management regimes, especially in C inflows from crop residues to the soil, global SOC debt on cropland would be increase at a rate of $0.1\unit{GtC yr^{-1}}$.
+Without changes in management regimes, especially in C inflows from crop residues to the soil, global $\Delta SOC$  on cropland would be increase at a rate of $0.1\unit{GtC yr^{-1}}$.
 The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of $0.06\unit{GtC yr^{-1}}$.
@@
 
-```{r SOCscen, out.width = "18cm", fig.cap = captionSOCscen, echo=FALSE, fig.pos="h"}
+```{r SOCscen, out.width="14cm", fig.cap = captionSOCscen, echo=FALSE}
 # All defaults
@@
 The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig.
-\@ref(fig:SOcinit).
+\@ref(fig:SOCinit).
 The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC) as well as the decrease in of $\Delta SOC$ for over the last 35 years (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.
+The $\Delta SOC$ estimation almost halves (from $~26\unit{GtC}$ to $~14\unit{GtC}$) as well as the decrease in of $\Delta SOC$ for over the last 35 years (from $~4\unit{GtC}$ to $~2.5\unit{GtC}$) for the Inital-natveg scenario.
 
@@
 
-```{r SOCinit, out.width = "18cm", fig.cap = captionSOCinit, echo=FALSE, fig.pos="H"}
+```{r SOCinit, out.width="10cm", fig.cap = captionSOCinit, echo=FALSE}
 # All defaults
@@
 ```
- Figure \@ref(fig:)
 
@@
 
-```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
+```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE}
 # All defaults
@@
 ```
-
-<!-- ```{r table2, echo = FALSE}
-# table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
-# knitr::kable(table$comparison_table, "latex", digit=2, 
-#              caption="$F^{SOC}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{SOC}$ by additional 5-8 percent.
-Also: modelled $F^{SOC}$ have increased for all climates over time .")
-# ```
-# ```{r table, echo=FALSE}
-# table <- readRDS("../ResultNotebooks/Output/SFC_climatezone.rds")
-# knitr::kable(table$naturalstate_table[3:10,], "latex", digit=2, 
-#              caption="S)
-# ``` -->

### 2020-12-14 ResultNotebooks/results_SOCflows.Rmd
Flow figure update

@@
 
-outputdir <- "Output/Images"
+outputdir <- "Output/"
 revfolder <- "Data/histManagement/"
@@
 CarbonInflow     <- readRDS(paste0(revfolder,"CarbonInflow.rds"))[,years,]
+CarbonDecay      <- readRDS(paste0(revfolder,"Decay.rds"))[,year,]
 
@@
 TotalSOCBudget    <- mbind(TotalSOCBudget,
-                           setNames(SOCsub[,,"natveg"], gsub("\\.","_",getNames(SOCsub))),
-                           SOCtot)
+                           setNames(SOCsub, gsub("\\.","_", getNames(SOCsub))),
+                           setNames(SOCtot, paste0("soc_", getNames(SOCtot))))
 
-CarbonIn          <- grepSpaceTime(SteadyState[,,"alpha_in"]*Landuse)
+CarbonIn          <- grepSpaceTime(CarbonInflow[,,"crop"]*Landuse[,,"crop"])
+CarbonIn          <- mbind(CarbonIn[,,"ag"],CarbonIn[,,"bg"],
+                           add_dimension(dimSums(
+                             CarbonIn[,,c("livst_chick","livst_egg","livst_milk","livst_pig","livst_rum")], dim=3.1),
+                              add="subpool", nm="manure"))
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(CarbonIn, gsub("\\.","_", getNames(CarbonIn))))
+CarbonInTot       <- dimSums(CarbonIn, dim=3.2)
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(ManureInput - CarbonInTot[,,"manure"], "manure_decay"),
+                           setNames(ResidueInput[,,"ag"] - CarbonInTot[,,"ag"], "ag_decay"),
+                           setNames(ResidueInput[,,"bg"] - CarbonInTot[,,"bg"], "bg_decay"))
+
+SOCIn             <- dimSums(CarbonInTot, dim=3)
+SOCtransfer       <- CarbonToEff[,,"crop"] 
+SOCPrev           <- dimSums(grepSpaceTime(setYears(
+                       SoilCarbonOut[,1,"actualstate"][,,"crop"]*Landuse[,1,"crop"], year)), dim=3)
+SOCLoss           <- SOCPrev - SOCtot[,,"crop"] + SOCtransfer + SOCIn
+
+TotalSOCBudget    <- mbind(TotalSOCBudget,
+                           setNames(SOCLoss, "soc_decay"))
+
+write.magpie(TotalSOCBudget, paste0(outputdir,"TotalSOCBudget.cs3"))
 
@@
 ```{r}
-####### delete when new output on decay is ready
-# Decay
-
-ActiveDecay        <- mbind(setNames(calcOutput("ActiveDecay",  landtype="crop",   aggregate = FALSE),"crop"),
-                              setNames(calcOutput("ActiveDecay",  landtype="natveg", aggregate = FALSE),"natveg"))
-SlowDecay          <- mbind(setNames(calcOutput("SlowDecay",  landtype="crop",   aggregate = FALSE),"crop"),
-                              setNames(calcOutput("SlowDecay",  landtype="natveg", aggregate = FALSE),"natveg"))
-PassiveDecay       <- mbind(setNames(calcOutput("PassiveDecay", landtype="crop",   aggregate = FALSE),"crop"),
-                              setNames(calcOutput("PassiveDecay", landtype="natveg", aggregate = FALSE),"natveg"))
-decay <- calcOutput("Decay", aggregate = FALSE)
-subpools     <- c("active","slow","passive")
-names        <- as.vector(outer(getNames(Landuse), subpools, paste, sep="."))
-Decay <- new.magpie(getCells(Landuse), getYears(Landuse), names, fill = 0)
-
-Decay[,,"active"]                  <- ActiveDecay
-Decay[,,"slow"]                    <- SlowDecay
-Decay[,,"passive"]                 <- PassiveDecay
-
-rm(ActiveDecay, SlowDecay, PassiveDecay)
-
-#
-####### end delete
 

### 2020-12-14 ResultNotebooks/results_SOCmaps.Rmd
Flow figure update

@@
 ```{r load}
-library(mrcommons)
-library(mrvalidation)
-library(mrsoil)
+library(magclass)
+library(madrat)
 library(magpiesets)
 
-rev <- "rev0.11"
-
-setConfig(cachefolder  = paste0("/home/kristine/mnt/rd3mod/inputdata/cache/", rev), forcecache = TRUE)
-setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
+rev <- "rev16"
 
@@
 
-library(ggplot2)
+library(reshape2)
+
+years <- seq(1975,2010,1)
 
 SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
-SOCStock        <- dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2)
+SOCStock        <- dimSums(dimSums(SoilCarbon[,years,][,,"state",pmatch=TRUE], dim=3.3)* Landuse[,years,], dim=3.2)
 glo_SOCStock    <- dimSums(SOCStock, dim=1) 
 glo_SOCEmission <- (glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
-
-SoilCarbon <- SoilCarbon[,findset("past"),]
-Landuse    <- Landuse[,findset("past"),]
-
-```
-
-Plot cellular maps
-
-```{r plot density maps}
 
@@
 density2plot  <- merge(as.data.frame(cropland_dens), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-
-# png(paste0(outputdir,"CDens_2010.png"), width = 1280, height = 720, res = 108)
-# luplot::plotmap2(cropland_dens[,"y2010",], legend_range = c(0,300), lowcol = cols[1], midcol = cols[4], highcol = cols[7], ncol=7, legend_discrete = TRUE)
-# dev.off()
-# 
-# # png(paste0(outputdir,"CShare_1965.png"),width = 1280, height = 720, res = 108)
-# # luplot::plotmap(cropland_share[,"y1965",], legend_range = c(0,2), legend_colours = cols, main="C_share on Cropland 1965")
-# # dev.off()
-# 
-# png(paste0(outputdir,"CShare_2010.png"), width = 1280, height = 720, res = 108)
-# luplot::plotmap(cropland_share[,"y2010",], legend_range = c(0,2), legend_colours = cols, main="Stock Change Factor on Cropland 2010")
-# dev.off()
-# 
-# # png(paste0(outputdir,"CIncrease_1965.png"),width = 1280, height = 720, res = 108)
-# # luplot::plotmap(cropland_increase[,"y1965",], legend_range = c(-80,80), legend_colours = cols, main="Absoulte C change on Cropland 1965")
-# # dev.off()
-# 
-# png(paste0(outputdir,"CIncrease_2010.png"),width = 1280, height = 720, res = 108)
-# luplot::plotmap(cropland_increase[,"y2010",], legend_range = c(-80,80), legend_colours = cols, legend_title = "tC/ha", main="Absoulte C change on Cropland 2010")
-# dev.off()
 
@@
 
-SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission[,seq(1975,2010,5),])[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
+SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission)[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
   labs( x = "Year", 

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
Flow figure update

@@
 \begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t-1,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
+SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t^*,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
 \label{eq:SOCstate}
@@
 \begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} - \underbrace{SOC_{i,t-1,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
+SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} - \underbrace{SOC_{i,t^*,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
 \label{eq:steadystate2budget}
@@
 \% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 03:39:00 2020
+\% Mon Dec 14 05:34:22 2020
 

### 2020-12-14 SOCbudgetPaper/methods.Rmd
Flow figure update

@@
 \begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t-1,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
+SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t^*,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
 (\#eq:SOCstate)
@@
 \begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t-1,sub,lu} - \underbrace{SOC_{i,t-1,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
+SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} - \underbrace{SOC_{i,t^*,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
 (\#eq:steadystate2budget)

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
update again table

@@
   \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.40\textwidth}}
+  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
   \tophline
@@
   \middlehline
-  \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                       \cite{faostat_faostat_2016}, \\
-                                                                      \cite{schaphoff_lpjml4_2018} (LPJmL),\\
+                                                                      \cite{schaphoff_lpjml4_2018},\\
                                                                       \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                      \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
-                                                                      LG:C generic values according to Table 5.5B\\
-                                                                      in \cite{calvo_buendia_ipcc_2019},\\
-                                                                      crop-type specific N:C-ratios from
-                                                                      \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
-                            & below-ground residues &  &  \\
+                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
+                                                                      LG:C generic values according to Table 5.5B,\\
+                                                                      5.5C in \cite{calvo_buendia_ipcc_2019},\\
+                                                                      crop-specific N:C from \cite{bodirsky_n2o_2012}
+                                                                      \end{minipage}} \\
+                            & below-ground residues, &  &  \\
                             & manure                &  &  \\ 
                             \hline
-  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018} (LPJmL) 
-                            & \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut 
+  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
+                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                 \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
@@
 The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in \(\tfrac{t}{ha}\)) at the level of 1975 onwards.
+manure (in \(\unit{t ha^{-1}}\)) at the level of 1975 onwards.
 Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
@@
 As outlined in Sect.
-\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
 As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
 Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the (\textit{constManagement} scenario.
+We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
 
@@
 Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\citep{calvo_buendia_ipcc_2019}), but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
+Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
 Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
@@
 
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to the often used constant management assumptions.
+This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt (\(\Delta SOC\)) considerably compared to the often used constant management assumptions.
 This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
@@
 \citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
+Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a substantial decline in SOC debt (less than 1\% change).
 Historic productivity increases were, however, notably larger.
@@
 \% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 05:34:22 2020
+\% Mon Dec 14 06:14:28 2020
 

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
update again table

@@
 
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to the often used constant management assumptions.
+This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt ($\Delta SOC$) considerably compared to the often used constant management assumptions.
 This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.

### 2020-12-14 SOCbudgetPaper/methods.Rmd
update again table

@@
     \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.40\textwidth}}
+    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
     \tophline
@@
     \middlehline
-    \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+    \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                         \cite{faostat_faostat_2016}, \\
-                                                                        \cite{schaphoff_lpjml4_2018} (LPJmL),\\
+                                                                        \cite{schaphoff_lpjml4_2018},\\
                                                                         \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                        \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
-                                                                        LG:C generic values according to Table 5.5B\\
-                                                                        in \cite{calvo_buendia_ipcc_2019},\\
-                                                                        crop-type specific N:C-ratios from
-                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
-                              & below-ground residues &  &  \\
+                                                        \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
+                                                                        LG:C generic values according to Table 5.5B,\\
+                                                                        5.5C in \cite{calvo_buendia_ipcc_2019},\\
+                                                                        crop-specific N:C from \cite{bodirsky_n2o_2012}
+                                                                        \end{minipage}} \\
+                              & below-ground residues, &  &  \\
                               & manure                &  &  \\ 
                               \hline
-    Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018} (LPJmL) 
-                              & \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut 
+    Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
+                              & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                   \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
@@
 The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in $\tfrac{t}{ha}$) at the level of 1975 onwards.
+manure (in $\unit{t ha^{-1}}$) at the level of 1975 onwards.
 Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
@@
 As outlined in Sect.
-\@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+\@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
 As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
 Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the (\textit{constManagement} scenario.
+We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
 
@@
 Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines ([@calvo_buendia_ipcc_2019]), but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
+Perennial grasses are given by the IPCC guidelines [@calvo_buendia_ipcc_2019], but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
 Whereas LG:C ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, N:C ratios range between $0.001-0.05$ with an average of $0.015$ (default value).

### 2020-12-14 ResultNotebooks/results_SOCflows.Rmd
update fow fig

@@
 LivstCropLoss     <- CropFEED - ManureExcretedTot - LivstCropHANPP 
-CropPRIMLoss      <- CropPRIM + ResPRIM + LivstCropHANPP - dimSums(grepSpaceTimeC(FAOmassbalance[,,c(kcr,kli,kres)][,,"other_util"]), dim=3) # only material use is not re-emitted
+CropPRIMLoss      <- CropPRIM + ResPRIM + LivstCropHANPP # assume verything gets re-emitted at the end
 
@@
 
-write.magpie(TotalSOCBudget, paste0(outputdir,"TotalSOCBudget.cs3"))
+write.csv(as.data.frame(TotalSOCBudget), paste0(outputdir,"TotalSOCBudget.csv"))
 

### 2020-12-14 SOCbudgetPaper/introduction.Rmd
update intro

@@
 
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historical agricultural management.
+The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
 To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
update frame

@@
     email: kristine.karstens@pik-potsdam.de
+    orcid: https://orcid.org/0000-0002-5860-9789
     corresponding: true
@@
     address: Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany
+
 abstract: |
   Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-It is three times larger than the vegetation carbon pool and more than twice as big as the atmospheric carbon pool, when looking into the first meter of the Earth's soil profile.
+The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
 SOC has been depleted by anthropogenic land cover change and agricultural management.
@@
  codedata: |
-   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
-All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results are accessable under (zenode_link).
-Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
+  We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions),  github.com/pik-piam/mrsoil (soil dynamic related functions) and  github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing.
+Model results are accessible under (zenode_link).
+Software code for paper and result preparation can be found under www.github.com/k4rst3ns/historicalsocmanegement (zenodo_link).
 authorcontribution: |
@@
 CM, JH and SR provided the LPJmL simulation data.
-KK wrote the paper with important contributions of BLB and CM.
-MK, JS, SR and IW provided extensive feedback.
+KK wrote the paper with important contributions of BLB and CM, and  MK, JS, SR and IW.
 All authors discussed the results and commented on the manuscript.
@@
 appendix: |
-
-  \section{Figures and tables in appendices}
-
-  \subsection{Methods}
-
-    ```{=latex}
-  \begin{figure}
-    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
-    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
-Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
-    \label{fig:CLIMzone}
-  \end{figure}
-  ```
-
-  ```{=latex}
-  \begin{table*}[h]
-  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
-Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
-  }
-  \begin{tabular}{lllllllllllll}
-  \tophline
-
-           &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
-  Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
-   \middlehline
-  tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
-  maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
-  trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
-  rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
-  soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
-  rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
-  groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
-  sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
-  oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
-  puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
-  potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
-  cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
-  sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
-  sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
-  others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
-  foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
-  cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
-  \bottomhline
-           &                 & \multicolumn{8}{l}{
-                                             \begin{minipage}[t]{0.50\columnwidth}
-                                                        nr/dm -- nitrogen to dry matter ratio\\
-                                                        wm/dm -- wet matter to dry matter ratio\\
-                                                        c/dm  -- carbon to dry matter ratio\\
-                                                        \end{minipage}} &
-                                                                          \multicolumn{3}{l}{
-                                                                              \begin{minipage}[t]{0.50\columnwidth}
-                                                                                          HI\textsuperscript{area} -- harvest index per area\\
-                                                                                          HI\textsuperscript{prod} -- harvest index per production\\
-                                                                                          RS -- root:shoot ratio \end{minipage}} 
-  \end{tabular}
-  (\#tab:c2dm)
-  \end{table*}
-  ```
-
-  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->
-
-  \subsection{Discussion}
-
-    ```{=latex}
-  \begin{figure}
-    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
-    \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.
-PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
-We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
-    \label{fig:SOCreg}
-  \end{figure}
-  ```
-
-    ```{=latex}
-  \begin{figure}
-    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-    \caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
-\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
-Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
-    \label{fig:SOClitter}
-  \end{figure}
-  ```
 
@@
 ```
+\newpage
+
+```{r child = 'appendix.Rmd'}
+```
+\newpage

### 2020-12-14 SOCbudgetPaper/methods.Rmd
update methods

--- 2020-12-14 SOCbudgetPaper/methods.Rmd
@@
 In Sect.
-\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
 Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see [@bodirsky_mrcommons_2020]).
@@
 This equation is valid for all grid cells $i$ and all years $t$.
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model ([@parton_analysis_1987]).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987]).
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
@@
     \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.40\textwidth}}
+    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.25\textwidth} p{0.25\textwidth}}
     \tophline
@@
     \middlehline
-    \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+    \multirow{3}{*}{Cropland} & above-ground residues & \multirow{3}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
                                                                         \cite{faostat_faostat_2016}, \\
@@
                                                                         \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                        \multirow{3}{*}{\begin{minipage}[t]{1\columnwidth}\raggedright
+                                                        \multirow{2}{*}{\begin{minipage}[t]{0.28\columnwidth}\raggedright
                                                                         LG:C generic values according to Table 5.5B\\
@@
                                                                         crop-type specific N:C-ratios from
-                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\
-                              & below-ground residues &  &  \\
-                              & manure                &  &  \\ 
+                                                                        \cite{bodirsky_mrcommons_2020}\end{minipage}} \\ \cline{2-2}
+                              & below-ground residues &  &  \\ \cline{2-2}
+                              & manure                &  & Table 5.5C in \cite{calvo_buendia_ipcc_2019} \\ 
                               \hline
@@
  \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
-                                     & \cite{hengl_soilgrids250m_2017} \\
+                                     & \cite{hengl_soilgrids250m:_2017} \\
                       \cline{2-4}
@@
 
-Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_ipcc_2006].
-Moreover, $\Delta SOC$ is equivalent to the negated cumulative SOC component of human land-use change emissions ([@pugh_simulated_2015]).
+Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_hs_ipcc_2006].
+The SOC debt is equivalent to the negated cumulative SOC component of  of human land use emission ([@pugh_simulated_2015]).
 
@@
 
-Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]).
 Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
@@
 Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig.
-\@ref(fig:CLIMzone)).
+\@ref(append:climatemap)).
 The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}}$ by
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of [@eggleston_ipcc_2006] and [@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of [@eggleston_hs_ipcc_2006] and [@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see FAO2LUH2MAG_croptypes.csv in [@bodirsky_mrcommons_2020]).
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see FAO2LUH2MAG_croptypes.csv in [@mrcommons]).
 By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 \@ref(sec:landuse)).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production [@faostat_faostat_2016].
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production [@faostat_faostat_2016].
 By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
@@
 
-Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_ipcc_2006].
+Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_hs_ipcc_2006].
 Deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
@@
 Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_ipcc_2006].
+Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_hs_ipcc_2006].
 
@@
 
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(tab:c2dm)).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(append:Tablec2dm)).
 
@@
 To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see [@karstens_mrsoil_2020]).
+In the following sections we outline name and idea of these scenarios (for technical implementation see [@mrsoil_karstens_2020]).
 
@@
 As outlined in Sect.
-\@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
+@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years ((\textit{Initial-lu})).
 As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
@@
 We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \@ref(tab:scenlitterpnv).
+We test however a range of different plant types as specified in Table @ref(tab:scenlitterpnv).
 

### 2020-12-14 SOCbudgetPaper/results.Rmd
update results

@@
 # Results
+
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in [@karstens_socbudgetdata_2020].
+In the following, the most important results (see [@karstens_socbudgetpost_2020] for post-processing script) are summarized.
 
@@
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+(d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
 "
@@
 \@ref(fig:SOCmaps)(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over $100\unit{t ha^{-1}}$ in northern temperate croplands, to less than $5\unit{t ha^{-1}}$ for arid and semiarid croplands.
-Our spatially explicit results show hot spots of SOC losses and gains compared to SOC under potential natural vegetation in two different ways: 
+Values ranging between well over $100\unit{t ha^{-1}}$ in northern temperate croplands to less than $5\unit{t ha^{-1}}$ for arid and semiarid croplands.
+Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways: 
 1.
@@
 \@ref(fig:SOCmaps)(b)) indicate areas with high importance for the global SOC losses.
-The might be driven by large relative changes (e.g.
-in Central Africa) or a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.
-North-east Asia).
-In contrast, large net gains of SOC happen primarily in developed countries of the North according to our results.
+They might be driven by large relative changes (e.g.
+in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.
+North-East Asia).
+In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
 2.
 Relative SOC changes measured as stock changes factors $F^{SCF}$ (see Fig.
-\@ref(fig:SOCmaps)(c)) are a helpful tool, to measure the impact of human cropping activities.
-They indicate areas with large difference in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
+\@ref(fig:SOCmaps)(c)) are a helpful metric to analyze the impact of human cropping activities.
+They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
 Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have improved, but presumably irrigated croplands at the border to dry, unsuitable areas worldwide as well.
+Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
 
-Focusing of the temporal dimension of SOC stock changes (see Fig.
-\@ref(fig:SOCmaps)(c)), global $\Delta SOC$  has decreased around 15% over the last couple of decades to $22\unit{GtC}$.
-This leads to a sequestration rate of $0.11\unit{GtC yr^{-1}}$ and considering our estimate of the global SOC stock of $675\unit{GtC}$ in 2010 to $0.2$ per mille per year.
+The global SOC debt has decreased by about 15% in the period between 1975 and 2010 to $22\unit{GtC}$ (Fig.
+\@ref(fig:SOCmaps)(d)).
+This corresponds to a sequestration rate of $0.11\unit{GtC yr^{-1}}$.
+Considering our estimate of the global SOC stock of around $660\unit{GtC}$ in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
 
-## Agricultural management effects on $\Delta SOC$
+## Carbon flows in the agricultural system
 
-Figure \@ref(fig:SOCscen)(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices.
-Without changes in management regimes, especially in C inflows from crop residues to the soil, global $\Delta SOC$  on cropland would be increase at a rate of $0.1\unit{GtC yr^{-1}}$.
-The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of $0.06\unit{GtC yr^{-1}}$.
-Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of $0.09\unit{GtC yr^{-1}}$ for both), the effect of no-tillage gets stronger till 2010.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \@ref(fig:SOCscen)(b) as well.
+C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
+Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
+We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.
+like food, bioenergy and material) into a human demand category.
+Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
+Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.
+compost, night soil and sewage) is not recycled to soils.
+Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
+However, around 60% of C decomposes before its integrated into soils at the the litter-soil barrier.
+Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
+According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
+4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
+Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
+
+```{r caption_FlowFig, echo=FALSE}
+caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
+After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
+Large parts of C, however, get mineralized on field before entering the soil.
+Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
+Finally, SOC gets mineralized and flows back into the atmosphere."
+```
+
+```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/CarbonBudget.png")
+```
+
+## Agricultural management effects on SOC debt
+
+We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
+Without changes in management regimes, the global SOC debt on cropland would be increasing at a rate of $0.1\unit{GtC yr^{-1}}$.
+As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
+Without the historic increase in residue returning, the global SOC debt would still increase at a rate of $0.06\unit{GtC yr^{-1}}$.
+Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of $0.09\unit{GtC yr^{-1}}$ for both).
+The effect of no-tillage has been particuarly strong since the 1990s.
+The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \@ref(fig:SOCscen)(b).
 
 ```{r captionSOCscen, echo=FALSE}
-captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different scenarios deviating from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
 ConstManagement combines all three modifications.
@@
 
-```{r SOCscen, out.width="14cm", fig.cap = captionSOCscen, echo=FALSE}
+```{r SOCscen, out.width = "18cm", fig.cap = captionSOCscen, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 
-The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig.
-\@ref(fig:SOCinit).
+Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
+\@ref(fig:SOcinit)).
 The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The $\Delta SOC$ estimation almost halves (from $~26\unit{GtC}$ to $~14\unit{GtC}$) as well as the decrease in of $\Delta SOC$ for over the last 35 years (from $~4\unit{GtC}$ to $~2.5\unit{GtC}$) for the Inital-natveg scenario.
+The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC), as does the decrease in $\Delta SOC$ for the period of 1975--2010 (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.
 
 ```{r captionSOCinit, echo=FALSE}
-captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901."
+captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use."
 ```
 
-```{r SOCinit, out.width="10cm", fig.cap = captionSOCinit, echo=FALSE}
+```{r SOCinit, out.width = "18cm", fig.cap = captionSOCinit, echo=FALSE, fig.pos="H"}
 # All defaults
@@
 
-## Carbon flows in the agricultural system
-
-Figure \@ref(fig:FlowFig) illustrates the carbon flows within the agricultural production system on cropland for the year 2010.
-
-```{r caption_FlowFig, echo=FALSE}
-caption_FlowFig <- "Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues."
-```
-
-```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/CarbonBudget.png")
-```
+Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas $\Delta SOC$ vary  19-24 GtC for the year 2010 (see Fig.
+\@ref(fig:SOClitter).
+The increase of $\Delta SOC$ during the period of 1975--2010 is independent of the parameterization choice at ~4 GtC.

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
update discussion 2

--- 2020-12-14 SOCbudgetPaper/discussion.Rmd
@@
 
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt considerably compared to the often used constant management assumptions.
+This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
 This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
@@
 
-The evaluation of SOC stock gains and losses, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
 If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink since 1975 according to our study.
@@
 
-Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-[@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account.
+Furthermore, [@sanderman_soil_2017] modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+[@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
 Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
 Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimate well above 50% ([@pellegrini_crop_2018], [@ray_recent_2012], [@rudel_agricultural_2009]).
+Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50% ([@pellegrini_crop_2018], [@ray_recent_2012], [@rudel_agricultural_2009]).
 
@@
 
-<!--Needed???
-Looking on historical cropland intensification rates of above 50% ([@rudel_agricultural_2009]), a large fraction of increased C input from residue biomass, is attributed to productivity improvements.
--->
+Looking on historic cropland intensification rates of above 50% ([@rudel_agricultural_2009]), a large fraction of increased C input from residue biomass is attributed to productivity improvements.
 
@@
 the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study).
+As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities.
 [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for belowground biomass.
 However, looking on the evaluation of management effects (see Sect.
@@
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]).
-Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_hs_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]).
+Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFtable)).
 To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\@ref(fig:CLIMzone)).
+\@ref(append:climatemap)).
 
-```{r SCFglo, echo=FALSE, results = 'asis'}
-table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
-library(xtable)
-print(xtable(table$comparison_table, digit=2, comment=FALSE, label="tab:SCFglo", 
-              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time."))
+```{r captionSCFtable, echo=FALSE}
+captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation.
+We compare average values for four IPCC climate zone classifications.
+IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in [@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019] are compared to 1990 and 2010 results for historic input levels of this study."
+```
+
+```{r SCFtable, out.width = "16cm", fig.cap = captionSCFtable, echo=FALSE}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
 ```
@@
 Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than [@calvo_buendia_ipcc_2019].
-The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
+The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
 Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
@@
 
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of $F_{SCF}$ as also indicated by the time trend of the SOC debt.
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
 
@@
 
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(fig:SOCglo)).
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(tab:SOCtable)).
 
-```{r captionSOCglo, echo=FALSE}
-captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+```{r captionSOCtable, echo=FALSE}
+captionSOCtable <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
 Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
@@
 
-```{r SOCglo, out.width = "8cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
+```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 Looking at regional results in Fig.
-\@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
+\@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
 To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
 
@@
 We conducted a sensitivity analysis (Fig.
-\@ref(fig:SOClitter)) based on various plant parameterizations from the Century model (see Sect.
-\@ref(sec:scenlitterpnv)).
+\@ref(append:senslitterpnv) based on various plant parameterizations from the Century model (see Sect.
+\@ref(sec:scen_litterpnv)).
 This shows that the general trend of decreasing SOC debt of ~4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
bugfix mille

@@
 If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 

### 2020-12-14 SOCbudgetPaper/conclusion.Rmd
update conclusion

--- 2020-01-21 SOCbudgetPaper/conclusion.Rmd
@@
 
-The conclusion goes here.
-You can modify the section name with  `\conclusions[modified heading if necessary]`.
+Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
+It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
+
+Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models.
+That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
+Our data-set and the MADRaT package ([@dietrich_madrat_2020]) constitute a starting point for building comprehensive data sets on agricultural management aspects.
+
+Our study again highlights that the expansion of croplands is still a major source of CO~2~ emissions -- not only by the removal of vegetation, but also by a slow depletion of soils.
+Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
+
+However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
+Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
+The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO~2~ source.
+Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO~2~.
+The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.
+
+Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
+Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand ([@popp_land-use_2014-1]) -- and thereby usually further reduce the SOC debt on croplands.
+
+While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets.
+Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
+
+Yet, there is ample potential for further improved SOC management.
+As shown in Fig.
+\@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
+Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC).
+Improved management could include, firstly, a circular flow from the food supply chain back to soils.
+Waste composting or excreta recycling could represent a major additional N input to cropland soils ([@brenzinger_organic_2018]).
+Secondly, soil carbon sequestration techniques ([@smith_soil_2016]), deep ploughing ([@alcantara_deep_2016]) or the transformation of c inputs to more recalcitrant biochar ([@woolf_sustainable_2010]) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops ([@poeplau_carbon_2015]) and agroforestry ([@lorenz_soil_2014]) could increase total C sequestration on croplands.

### 2020-12-14 ResultNotebooks/results_SOCscenario.Rmd
update figures

@@
 
-ci2plot <- c('Above Ground Residues'="ag.crop",          
-             'Below Ground Residues'="bg.crop",          
-             'Manure from Monogastrics'="livst_chick.crop",
-             'Manure from Monogastrics'="livst_egg.crop",
-             'Manure from Monogastrics'="livst_pig.crop",
-             'Manure from Ruminants'="livst_milk.crop",  
-             'Manure from Ruminants'="livst_rum.crop",   
-             'Natural Litterfall and Fine Roots'="litfall.natveg")
+ci2plot <- c('above-ground residues'="ag.crop",          
+             'below-ground residues'="bg.crop",          
+             'manure from monogastrics'="livst_chick.crop",
+             'manure from monogastrics'="livst_egg.crop",
+             'manure from monogastrics'="livst_pig.crop",
+             'manure from ruminants'="livst_milk.crop",  
+             'manure from ruminants'="livst_rum.crop",   
+             'natural litterfall and fine roots'="litfall.natveg")
 
@@
         y = expression(paste('Global \u0394SOC in GtC')),
-        color    ="Initalitzation",
+        color    ="Initialization",
         linetype ="Management") +

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
update figures

--- 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
@@
 acknowledgements: |
-  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors. 
+  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
+Thanks to the rticles contributors as well.
+The authors thank for the data provided by FAOSTAT and LUH2v2.
+The work of Kristine Karstens was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no.
+ED78/3-2).
+The research leading to these results has received funding from the European Union’s Horizon 2020 research and innovation program under grant agreement no.
+776479 (COACCH) and .
+This work was also supported by    funded by the German Federal Ministry of Education and Research (BMBF).
+
 appendix: |

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
update figures

@@
 \% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 06:14:28 2020
+\% Mon Dec 14 21:01:55 2020
 
@@
 \begin{acknowledgements}
-Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors.
+Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
+Thanks to the rticles contributors as well.
+The authors thank for the data provided by FAOSTAT and LUH2v2.
+The work of Kristine Karstens was funded by the DFG Priority Program ``Climate Engineering: Risks, Challenges, Opportunities?'' (SPP 1689) and specifically the CEMICS2 project (grant no.
+ED78/3-2).
+The research leading to these results has received funding from the European Union's Horizon 2020 research and innovation program under grant agreement no.
+776479 (COACCH) and .
+This work was also supported by funded by the German Federal Ministry of Education and Research (BMBF).
 \end{acknowledgements}

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd

@@
   Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-It is three times larger than the vegetation carbon pool and more than twice as big as the atmospheric carbon pool, when looking into the first meter of the Earth's soil profile.
+The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
 SOC has been depleted by anthropogenic land cover change and agricultural management.

### 2020-12-14 SOCbudgetPaper/discussion.Rmd

--- 2020-12-14 SOCbudgetPaper/discussion.Rmd
@@
 
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt ($\Delta SOC$) considerably compared to the often used constant management assumptions.
+This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
 This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
@@
 
-The evaluation of SOC stock gains and losses, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
 If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 
@@
 To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
 

### 2020-12-14 SOCbudgetPaper/results.Rmd

@@
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): The SOC debt between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
+(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
 "
@@
 We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
-Without changes in management regimes, the global SOC debt on cropland would be increasing at a rate of $0.1\unit{GtC yr^{-1}}$.
+Without changes in management regimes, the global $\Delta SOC$ on cropland would be increasing at a rate of $0.1\unit{GtC yr^{-1}}$.
 As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global SOC debt would still increase at a rate of $0.06\unit{GtC yr^{-1}}$.
+Without the historic increase in residue returning, the global $\Delta SOC$ would still increase at a rate of $0.06\unit{GtC yr^{-1}}$.
 Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of $0.09\unit{GtC yr^{-1}}$ for both).

### 2020-12-14 ResultNotebooks/results_SOCscenario.Rmd
final update of figure and text

@@
 litter_scen <-  c(`LitterPNV-Mean` = "histManagement",
-                  `LitterPNV-grass`= "LitterPNV-PerennialGrasses", 
+                  `LitterPNV-Grass`= "LitterPNV-PerennialGrasses", 
                   `LitterPNV-AASS` = "LitterPNV-AASS", 

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
final update of figure and text

@@
     surname: Karstens
-    affiliation: 1
+    affiliation: {1,3}
     email: kristine.karstens@pik-potsdam.de
@@
     affiliation: 2
+    orcid: https://orcid.org/0000-0002-3784-1124
   - given_name: Isabelle
@@
 All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results are accessable under (zenode_link).
+Model results including C input data are accessable under https://doi.org/10.5281/zenodo.4320663.
 Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
@@
 KK wrote the paper with important contributions of BLB and CM.
-MK, JS, SR and IW provided extensive feedback.
+MK, JS, SR and IW provided extensive feedback to outline of the study.
 All authors discussed the results and commented on the manuscript.
@@
   Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
-Thanks to the rticles contributors as well.
+Additional thanks to the rticles contributors for providing a R Markdown template.
 The authors thank for the data provided by FAOSTAT and LUH2v2.
-The work of Kristine Karstens was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no.
+The work of KK was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no.
 ED78/3-2).
-The research leading to these results has received funding from the European Union’s Horizon 2020 research and innovation program under grant agreement no.
-776479 (COACCH) and .
-This work was also supported by    funded by the German Federal Ministry of Education and Research (BMBF).
+The research leading to these results has received funding for BLB from the European Union’s Horizon 2020 research and innovation program under grant agreement no.
+776479 (COACCH) and  no.
+821010 (CASCADES).
+The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF).
+The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no.
+774378).
 

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
final update of figure and text

--- 2020-12-14 SOCbudgetPaper/SOCbudget.tex
@@
 
-\Author[1]{Kristine}{Karstens}
+\Author[true]{Kristine}{Karstens}
 \Author[1]{Benjamin Leon}{Bodirsky}
@@
 Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-It is three times larger than the vegetation carbon pool and more than twice as big as the atmospheric carbon pool, when looking into the first meter of the Earth's soil profile.
+The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
 SOC has been depleted by anthropogenic land cover change and agricultural management.
@@
 
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons ({[}portmann\_mirca2000global\_2010{]}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons (\citep{portmann_mirca2000global_2010}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).
 
@@
                                                                       \cite{faostat_faostat_2016}, \\
-                                                                      \cite{schaphoff_lpjml4_2018},\\
+                                                                      LPJmL4 \cite{schaphoff_lpjml4_2018},\\
                                                                       \cite{weindl_livestock_2017} \end{minipage}} & 
@@
                                                                       LG:C generic values according to Table 5.5B,\\
-                                                                      5.5C in \cite{calvo_buendia_ipcc_2019},\\
+                                                                      5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
                                                                       crop-specific N:C from \cite{bodirsky_n2o_2012}
@@
                             & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
+                                IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
+                                CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
  \bottomhline
@@
 \ref{sec:landuse}).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production \citep{faostat_faostat_2016}.
 By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
@@
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see Sect.
-\ref{append:TableclossAWMS} for more details).
+For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in \citep{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}.
 We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
@@
 
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in \citep{karstens_model_2020}.
+In the following, the most important results (see \citep{karstens_result_2020} for post-processing script) are summarized.
+
 \hypertarget{soc-distribution-and-depletion}{%
@@
 \begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics.
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
+(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
 Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades.
+(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
 }\label{fig:SOCmaps}
@@
 \ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands, to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
-Our spatially explicit results show hot spots of SOC losses and gains compared to SOC under potential natural vegetation in two different ways:
+Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
+Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways:
 1.
@@
 \ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses.
-The might be driven by large relative changes (e.g.~in Central Africa) or a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-east Asia).
-In contrast, large net gains of SOC happen primarily in developed countries of the North according to our results.
+They might be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
+In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
 2.
 Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig.
-\ref{fig:SOCmaps}(c)) are a helpful tool, to measure the impact of human cropping activities.
-They indicate areas with large difference in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
+\ref{fig:SOCmaps}(c)) are a helpful metric to analyze the impact of human cropping activities.
+They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
 Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have improved, but presumably irrigated croplands at the border to dry, unsuitable areas worldwide as well.
+Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
 
-Focusing of the temporal dimension of SOC stock changes (see Fig.
-\ref{fig:SOCmaps}(c)), global \(\Delta SOC\) has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\).
-This leads to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\) and considering our estimate of the global SOC stock of \(675\unit{GtC}\) in 2010 to \(0.2\) per mille per year.
+The global SOC debt has decreased by about 15\% in the period between 1975 and 2010 to \(22\unit{GtC}\) (Fig.
+\ref{fig:SOCmaps}(d)).
+This corresponds to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\).
+Considering our estimate of the global SOC stock of around \(660\unit{GtC}\) in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
 
-\hypertarget{agricultural-management-effects-on-delta-soc}{%
-\subsection{\texorpdfstring{Agricultural management effects on \(\Delta SOC\)}{Agricultural management effects on \textbackslash Delta SOC}}\label{agricultural-management-effects-on-delta-soc}}
+\hypertarget{carbon-flows-in-the-agricultural-system}{%
+\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
 
-Figure \ref{fig:SOCscen}(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices.
-Without changes in management regimes, especially in C inflows from crop residues to the soil, global \(\Delta SOC\) on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\).
-The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of \(0.06\unit{GtC yr^{-1}}\).
-Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both), the effect of no-tillage gets stronger till 2010.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \ref{fig:SOCscen}(b) as well.
+C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
+Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
+We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.~like food, bioenergy and material) into a human demand category.
+Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
+Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not recycled to soils.
+Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
+However, around 60\% of C decomposes before its integrated into soils at the the litter-soil barrier.
+Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
+According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
+4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
+Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
+
+\begin{figure}[h]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
+After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
+Large parts of C, however, get mineralized on field before entering the soil.
+Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
+Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
+\end{figure}
+
+\hypertarget{agricultural-management-effects-on-soc-debt}{%
+\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
+
+We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)).
+Without changes in management regimes, the global \(\Delta SOC\) on cropland would be increasing at a rate of \(0.1\unit{GtC yr^{-1}}\).
+As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
+Without the historic increase in residue returning, the global \(\Delta SOC\) would still increase at a rate of \(0.06\unit{GtC yr^{-1}}\).
+Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both).
+The effect of no-tillage has been particuarly strong since the 1990s.
+The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \ref{fig:SOCscen}(b).
 
 \begin{figure}
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
 ConstManagement combines all three modifications.
@@
 
-The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig.
-\ref{fig:SOCinit}.
+Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
+\ref{fig:SOCinit}).
 The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The \(\Delta SOC\) estimation almost halves (from \(~26\unit{GtC}\) to \(~14\unit{GtC}\)) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \(~4\unit{GtC}\) to \(~2.5\unit{GtC}\)) for the Inital-natveg scenario.
+The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
 
 \begin{figure}
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
+\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
 \end{figure}
 
-\hypertarget{carbon-flows-in-the-agricultural-system}{%
-\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
-
-Figure \ref{fig:FlowFig} illustrates the carbon flows within the agricultural production system on cropland for the year 2010.
-
-\begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.}\label{fig:FlowFig}
-\end{figure}
+Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas \(\Delta SOC\) vary 19-24 GtC for the year 2010 (see Fig.
+\ref{fig:SOClitter}.
+The increase of \(\Delta SOC\) during the period of 1975--2010 is independent of the parameterization choice at \textasciitilde4 GtC.
 
@@
 
-This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt (\(\Delta SOC\)) considerably compared to the often used constant management assumptions.
+This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
 This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
@@
 
-The evaluation of SOC stock gains and losses, is complex and has several dimensions as climatic and anthropogenic effects overlap.
+The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
 If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 
@@
 
-Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a substantial decline in SOC debt (less than 1\% change).
+Furthermore, \citep{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
+Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
 Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimate well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).
+Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).
 
@@
 This is in contrast to \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
+
+Looking on historic cropland intensification rates of above 50\% (\citep{rudel_agricultural_2009}), a large fraction of increased C input from residue biomass is attributed to productivity improvements.
 
@@
 \% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 06:14:28 2020
+\% Mon Dec 14 23:33:09 2020
 
@@
 
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of \(F_{SCF}\) as also indicated by the time trend of the SOC debt.
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
 
@@
 \begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
 This makes it hard to pinpoint a specific year for these SOC estimations.
@@
 To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of \(\Delta SOC\) and \(F^\mathrm{SCF}\).
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
 
@@
 Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
-It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historical agricultural management on global SOC stocks.
+It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
 
-With the extraction of C at harvest and little efforts to increase cropland SOC, these stocks are depleted in comparison to natural vegetation SOC stocks in most regions.
-The reduction of further deforestation for cropland expansion thus need to be stopped quickly.
+Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models.
+That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
+Our data-set and the MADRaT package (\citep{dietrich_madrat_2020}) constitute a starting point for building comprehensive data sets on agricultural management aspects.
+
+Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions -- not only by the removal of vegetation, but also by a slow depletion of soils.
+Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
 
@@
 The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source.
-Under the assumption that cropland management is static in time, as typically assumed in other studies on cropland SOC stocks, we also find that cropland soils are a source of CO\textsubscript{2}, building trust in our method.
+Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO\textsubscript{2}.
+The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.
 
-The estimated increase in cropland SOC is caused by changing agricultural management.
-The largest contribution to these dynamics comes from increased carbon inputs to soils from crop residues.
-Crop residue management can still be improved from the SOC perspective, e.g.~by reducing shares of residue burning or further improvements in agricultural productivity.
-Similarly, other carbon-accumulating practices, such as the application of cover crops need to be included in studies of cropland SOC dynamics, if data on these practices become available.
+Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
+Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand (\citep{popp_land-use_2014-1}) -- and thereby usually further reduce the SOC debt on croplands.
 
-Our findings on sequestration rates are, however, still more than an order of magnitude lower than promoted by the 4 per mille initiative.
-We conclude, that while crop productivity gains, lead to co-benefits for SOC stocks, they are not enough to help meeting ambitious climate mitigation targets.
-Soil sequestration techniques in areas with low stock change factors still might focus on closing yield gaps, whereas other SOC conserving technologies (e.g.~conservation agriculture, biochar application and agroforestry) and the avoidance of further deforestation need to be considered elsewhere.
+While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets.
+Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
 
-Reduced or avoided deforestation would not only reduce SOC losses on newly converted land, it would likely also require further productivity improvements in existing croplands to meet -- and thereby usually further reduce the SOC debt on croplands.
-
-Our results clearly demonstrate that agricultural management aspects need to be explicitly considered in global carbon assessments and models.
-That also implies, that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT tool constitute a starting point for building comprehensive data sets on agricultural management aspects.
+Yet, there is ample potential for further improved SOC management.
+As shown in Fig.
+\ref{fig:FlowFig}, approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
+Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC).
+Improved management could include, firstly, a circular flow from the food supply chain back to soils.
+Waste composting or excreta recycling could represent a major additional N input to cropland soils (\citep{brenzinger_organic_2018}).
+Secondly, soil carbon sequestration techniques (\citep{smith_soil_2016}), deep ploughing (\citep{alcantara_deep_2016}) or the transformation of c inputs to more recalcitrant biochar (\citep{woolf_sustainable_2010}) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops (\citep{poeplau_carbon_2015}) and agroforestry (\citep{lorenz_soil_2014}) could increase total C sequestration on croplands.
 
@@
 All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results are accessable under (zenode\_link).
+Model results including C input data are accessable under \url{https://doi.org/10.5281/zenodo.4320663}.
 Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.} %% use this section when having data sets and software code available
@@
 KK wrote the paper with important contributions of BLB and CM.
-MK, JS, SR and IW provided extensive feedback.
+MK, JS, SR and IW provided extensive feedback to outline of the study.
 All authors discussed the results and commented on the manuscript.} %% optional section
@@
 \begin{acknowledgements}
-Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work. Thanks to the rticles contributors.
+Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
+Additional thanks to the rticles contributors for providing a R Markdown template.
+The authors thank for the data provided by FAOSTAT and LUH2v2.
+The work of KK was funded by the DFG Priority Program ``Climate Engineering: Risks, Challenges, Opportunities?'' (SPP 1689) and specifically the CEMICS2 project (grant no.
+ED78/3-2).
+The research leading to these results has received funding for BLB from the European Union's Horizon 2020 research and innovation program under grant agreement no.
+776479 (COACCH) and no.
+821010 (CASCADES).
+The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF).
+The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no.
+774378).
 \end{acknowledgements}

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
final update of figure and text

@@
 ```{r captionSOCglo, echo=FALSE}
-captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
-Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
+captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
 This makes it hard to pinpoint a specific year for these SOC estimations.

### 2020-12-14 SOCbudgetPaper/introduction.Rmd
final update of figure and text

@@
 In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.
-growing seasons ([portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g.
+growing seasons ([@portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g.
 [@pugh_simulated_2015], [@lutz_simulating_2019]).

### 2020-12-14 SOCbudgetPaper/methods.Rmd
final update of figure and text

@@
                                                                         \cite{faostat_faostat_2016}, \\
-                                                                        \cite{schaphoff_lpjml4_2018},\\
+                                                                        LPJmL4 \cite{schaphoff_lpjml4_2018},\\
                                                                         \cite{weindl_livestock_2017} \end{minipage}} & 
@@
                                                                         LG:C generic values according to Table 5.5B,\\
-                                                                        5.5C in \cite{calvo_buendia_ipcc_2019},\\
+                                                                        5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
                                                                         crop-specific N:C from \cite{bodirsky_n2o_2012}
@@
                               & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                  \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
+                                  IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
+                                  CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
    \bottomhline
@@
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in [@bodirsky_n2o_2012] (see Sect.
-\@ref(append:TableclossAWMS) for more details).
+For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in [@bodirsky_mrcommons_2020] for more details) depending on the different systems and the associated nitrogen loss rates as specified in [@bodirsky_n2o_2012].
 We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.

### 2020-12-14 SOCbudgetPaper/results.Rmd
final update of figure and text

@@
 
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in [@karstens_socbudgetdata_2020].
-In the following, the most important results (see [@karstens_socbudgetpost_2020] for post-processing script) are summarized.
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in [@karstens_model_2020].
+In the following, the most important results (see [@karstens_result_2020] for post-processing script) are summarized.
 
@@
 ```{r captionSOCmaps, echo=FALSE}
-captionSOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics.
+captionSOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
+(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
@@
 ```{r captionSOCscen, echo=FALSE}
-captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different scenarios deviating from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
 ConstManagement combines all three modifications.
@@
 Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\@ref(fig:SOcinit)).
+\@ref(fig:SOCinit)).
 The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.

### 2020-12-15 FlowFigure/manipulate_sankey.R
delete old stuff

@@ -1,256 +1,0 @@
-
-nitrogen_sankey<-function(x,svg_in,svg_out="sankey.svg",scaling=1){
-  require("xml2")
-  svg<-read_xml(svg_in)
-  svg2<-xml_find_all(svg, "//svg:g")
-  paths<-xml_find_all(svg2, ".//svg:path")
-  texts<-xml_find_all(svg2, ".//svg:text")
-  
-  mapping<-toolMappingFile(type = "sectoral",name = "svg_nitrogen.csv",readcsv = TRUE)
-  
-
-  
-  settext<-function(id,value){
-    text=texts[which(xml_attr(texts,attr="id")==id)]
-    if(length(text)!=1){warning(paste0("text id ",id," does not exist or exists multiple"))}
-    xml_text(text)<-value
-  }
-  
-  setstyle<-function(id,value,unit="px",style="stroke-width"){
-    path=paths[which(xml_attr(paths,attr="id")==id)]
-    if(length(path)!=1){stop(paste0("id ",id," does not exist or exists multiple"))}
-    text = xml_attr(path,attr="style")
-    text=strsplit(text,split = ";")[[1]]
-    pos<-which(substring(text,first = 0,last=nchar(style))==style)
-    text[pos]<-paste0(style,":",value,unit)
-    text<-paste(text,collapse=";")
-    xml_attr(path,"style")<-text
-  }
-  
-  numformat <- function(val) {
-    if((val<0.95)&(val>0.05)) {
-      return(sub("^(-?)0.", "\\1.", sprintf("%.1f", val)) )
-    } else if(val<=0.05) {
-      return(" ")
-    } else{
-      return(sprintf("%.0f", val) )  
-    }
-  }
-  
-  
-  
-  ## add year and region to header
-  
-  region<-getRegions(x)
-  if (length(region)==249){
-    region="World"
-    settext("region",paste0("of the world"))
-  } else if (length(region)==1){
-    mapping2<-toolMappingFile(type = "regional",name = "regionmappingMAgPIE.csv",readcsv = TRUE)
-    region<-mapping2[which(mapping2[,2]==region),1]
-    settext("region",paste0("of ",region))
-  } else {region <- "aggregate world region"}
-  
-
[…]

### 2020-12-15 FlowFigure/toolFlowFigure.R

@@ -1,106 +1,0 @@
-#' toolFlowFigure
-#'
-#' Create a flow figure out of data, mapping and proxy pattern
-#'
-#' @param data
-#' @param proxy
-#' @param mapping
-#' @param scaling
-#'
-#' @return file path to svg
-#' @author Benjamin Leon Bodirsky, Kristine Karstens
-#'
-#' @importFrom xml2 read_xml xml_find_all
-#' @export
-
-toolFlowFigure <- function(data, proxy, mapping, scaling=1){
-
-  proxy <- "/home/kristine/WORK/Paper/historicalsocbudget/FlowFigure/sankey_nitrogen_CHN_2010.svg"
-
-  svg      <- read_xml(proxy)
-  groups   <- xml_find_all(svg, "//svg:g")
-  paths    <- xml_find_all(groups, ".//svg:path")
-  texts    <- xml_find_all(groups, ".//svg:text")
-
-  mapping <- toolGetMapping(where = "socbudet", name = mapping)
-
-
-
-  settext<-function(id,value){
-    text=texts[which(xml_attr(texts,attr="id")==id)]
-    if(length(text)!=1){warning(paste0("text id ",id," does not exist or exists multiple"))}
-    xml_text(text)<-value
-  }
-
-  setstyle<-function(id,value,unit="px",style="stroke-width"){
-    path=paths[which(xml_attr(paths,attr="id")==id)]
-    if(length(path)!=1){stop(paste0("id ",id," does not exist or exists multiple"))}
-    text = xml_attr(path,attr="style")
-    text=strsplit(text,split = ";")[[1]]
-    pos<-which(substring(text,first = 0,last=nchar(style))==style)
-    text[pos]<-paste0(style,":",value,unit)
-    text<-paste(text,collapse=";")
-    xml_attr(path,"style")<-text
-  }
-
-  numformat <- function(val) {
-    if((val<0.95)&(val>0.05)) {
-      return(sub("^(-?)0.", "\\1.", sprintf("%.1f", val)) )
-    } else if(val<=0.05) {
-      return(" ")
-    } else{
-      return(sprintf("%.0f", val) )
-    }
-  }
-
-
-
-  ## add year and region to header
-
-  region<-getRegions(x)
-  if (length(region)==249){
-    region="World"
-    settext("region",paste0("of the world"))
-  } else if (length(region)==1){
-    mapping2<-toolMappingFile(type = "regional",name = "regionmappingMAgPIE.csv",readcsv = TRUE)
[…]

### 2020-12-17 ResultNotebooks/results_SOCcompare.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

@@
                   setNames(SOCStock[,"y2010","naturalstate"],"This Study (PNV)"),
-                  setNames(LPJmL4[,"y2010",], "LPJmL4Paper"),
+                  setNames(LPJmL4[,"y2010",], "LPJmL4"),
                   setYears(setNames(WISE[,1,],"WISE"), "y2010"),

### 2020-12-17 SOCbudgetPaper/SOCbudget.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

@@
     surname: Karstens
-    affiliation: {1,3}
+    affiliation: "1, 3"
     email: kristine.karstens@pik-potsdam.de
@@
     surname: Lotze-Campen
-    affiliation: 1
+    affiliation: "1, 3"
   - given_name: Alexander
@@
   codedata: |
-   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
-All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results including C input data are accessable under https://doi.org/10.5281/zenodo.4320663.
+   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \citep{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \citep{karstens_mrsoil_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \citep{bodirsky_mrvalidation_2020} for validation data.
+All libraries are based on the MADRaT package at github.com/pik-piam/madrat \citep{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
+Model results including C input data are accessable under https://doi.org/10.5281/zenodo.4320663 \citep{karstens_model_2020}.
 Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
@@
     ```{=latex}
-  \begin{figure}
+  \begin{figure}[h]
     \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
-    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
+    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data.
 Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
@@
   ```{=latex}
-  \begin{table*}[h]
-  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
-Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
+  \begin{sidewaystable*}[htbp]
+  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \citet{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
+Allometric coefficients are used as descriped in \citet{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
   }
@@
   \bottomhline
-           &                 & \multicolumn{8}{l}{
-                                             \begin{minipage}[t]{0.50\columnwidth}
+           &                 & \multicolumn{6}{l}{
+                                             \begin{minipage}[t]{0.25\columnwidth}
                                                         nr/dm -- nitrogen to dry matter ratio\\
@@
                                                         \end{minipage}} &
-                                                                          \multicolumn{3}{l}{
-                                                                              \begin{minipage}[t]{0.50\columnwidth}
+                                                                          \multicolumn{5}{l}{
+                                                                              \begin{minipage}[t]{0.25\columnwidth}
                                                                                           HI\textsuperscript{area} -- harvest index per area\\
@@
   (\#tab:c2dm)
-  \end{table*}
+  \end{sidewaystable*}
   ```
@@
     ```{=latex}
-  \begin{figure}
+  \begin{figure}[H]
     \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
@@
 PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
-We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
+We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
     \label{fig:SOCreg}
@@
     ```{=latex}
-  \begin{figure}
+  \begin{figure}[H]
     \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-    \caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
-\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
+    \caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \citealt{century_model_2000}).
+\textit{LitterPNV-CenturyAverage} forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
 Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}

### 2020-12-17 SOCbudgetPaper/SOCbudget.tex
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

@@
 
-\Author[true]{Kristine}{Karstens}
+\Author[1, 3]{Kristine}{Karstens}
 \Author[1]{Benjamin Leon}{Bodirsky}
@@
 \Author[1]{Isabelle}{Weindl}
-\Author[1]{Hermann}{Lotze-Campen}
+\Author[1, 3]{Hermann}{Lotze-Campen}
 \Author[1]{Alexander}{Popp}
@@
 Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times (\citep{batjes_total_1996}).
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere (\citep{friedlingstein_global_2019}, \citep{minasny_soil_2017}).
-The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (\citep{batjes_total_1996}, \citep{sanderman_soil_2017}).
+It exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}.
+Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{friedlingstein_global_2019, minasny_soil_2017}.
+The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile \citep{batjes_total_1996, sanderman_soil_2017}.
 
 Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}).
+Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018}.
 However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
@@
 The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
-\citep{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter (\citep{lal_world_2001}).
+Sanderman et al.~\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.
 
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions (\citep{friedlingstein_global_2019}).
+Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions \citep{friedlingstein_global_2019}.
 While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest (\citep{strassmann_simulating_2008}, \citep{betts_climate_2015}).
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management (\citep{friedlingstein_global_2019}, \citep{houghton_carbon_2012}, \citep{hansis_relevance_2015}).
+Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}.
+BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management \citep{friedlingstein_global_2019, houghton_carbon_2012, hansis_relevance_2015}.
 
-Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g.~\citep{bondeau_modelling_2007}, \citep{lindeskog_implications_2013}).
-\citep{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
+Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.~][]{bondeau_modelling_2007, lindeskog_implications_2013}.
+Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
 They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
 
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons (\citep{portmann_mirca2000global_2010}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010} multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][\citet{lutz_simulating_2019}]{pugh_simulated_2015}.
 
-More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
 
-Field-scale models (\citet{del_grosso_simulated_2001}{]}, \citep{coleman_simulating_1997},\citep{smith_estimating_2010}, \citep{taghizadeh-toosi_c-tool_2014}) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
+Field-scale models \citep[\citet{coleman_simulating_1997}, \citet{smith_estimating_2010}, \citet{taghizadeh-toosi_c-tool_2014}]{del_grosso_simulated_2001} are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
 However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
@@
 In Sect.
-\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}).
-We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}).
+\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}.
+We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}.
 In Sect.
-\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
-Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citep{bodirsky_mrcommons_2020}).
+\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
+Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices \citep[for model code see][]{bodirsky_mrcommons_2020}.
 In Sect.
@@
 
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
 We restrict our analysis to the first 0-30 cm of the soil profile.
@@
 This equation is valid for all grid cells \(i\) and all years \(t\).
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model (\citep{parton_analysis_1987}).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep{parton_analysis_1987}.
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
@@
 \ref{sec:residues}) and manure inputs (see Sect.
-\ref{sec:livstmanure}); for natural vegetation litterfall (including fine root turnover) (\citep{schaphoff_lpjml4_2018}) is the only source of carbon inflow to the soil.
+\ref{sec:livstmanure}); for natural vegetation litterfall including fine root turnover \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
 Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
@@
   \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      \cite{faostat_faostat_2016}, \\
-                                                                      LPJmL4 \cite{schaphoff_lpjml4_2018},\\
-                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
+                                                                      \citet{faostat_faostat_2016}, \\
+                                                                      \citet{schaphoff_lpjml4_2018},\\
+                                                                      \citet{weindl_livestock_2017} \end{minipage}} & 
                                                       \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                       LG:C generic values according to Table 5.5B,\\
-                                                                      5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
-                                                                      crop-specific N:C from \cite{bodirsky_n2o_2012}
+                                                                      5.5C from IPCC \citep{calvo_buendia_ipcc_2019},\\
+                                                                      crop-specific N:C from \citet{bodirsky_n2o_2012}
                                                                       \end{minipage}} \\
@@
                             \hline
-  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
+  Natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
                             & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
-                                CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
+                                IPCC \citep{calvo_buendia_ipcc_2019} and\\ 
+                                CENTURY (\citep{century_model_2000}) \strut \end{minipage}\tabularnewline
  \bottomhline
@@
 We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in \citep{calvo_buendia_ipcc_2019}.
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
 
@@
  \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
-                                     & \cite{hengl_soilgrids250m_2017} \\
+                                     & \citet{hengl_soilgrids250m_2017} \\
                       \cline{2-4}
 
-                      & Mircobial activity & air temperature & \cite{harris_version_2020} \\
+                      & Mircobial activity & air temperature & \citet{harris_version_2020} \\
                       \cline{2-4}
 
-                      & Water restriction & precipitation \& potential evapotranspiration & \cite{harris_version_2020} \\
+                      & Water restriction & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
                       \cline{1-4}
@@
 
-According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years (see Eq.
-5.0A in \citep{calvo_buendia_ipcc_2019}).
+According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years \citep[see Eq.
+5.0A in][]{calvo_buendia_ipcc_2019}.
 This, however, will also include naturally occurring changes due to climatic variation over time.
@@
 
-Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
-Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land-use change emissions (\citep{pugh_simulated_2015}).
+Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt \citep{sanderman_soil_2017} due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
+Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land-use change emissions \citep{pugh_simulated_2015}.
 
@@
 
-Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
+Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
 Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
@@
 with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of Eggleston et al.~\citeyearpar{eggleston_ipcc_2006} and Calvo Buendia et al.~\citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{bodirsky_mrcommons_2020}).
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups \citep[see FAO2LUH2MAG\_croptypes.csv in][]{bodirsky_mrcommons_2020}, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016}.
 By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 
-Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
+Crop residue production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
 Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
@@
 Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
-Deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+Deviating from Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
 We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
@@
 A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
-Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following Smil \citeyearpar{smil_nitrogen_1999} we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
 Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{bodirsky_mrcommons_2020}).
-Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020}.
+Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017}, that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
 We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
@@
 
-To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from \citep{weindl_livestock_2017}.
+To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of Robinson et al.~\citeyearpar{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from Weindl et al.~\citeyearpar{weindl_livestock_2017}.
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
@@
 
-Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
+Manure production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
 Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see \citep{calvo_buendia_ipcc_2019}).
+Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in \citep{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}.
+For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012}.
 We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
@@
 
-In order to derive a spatial distribution of the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from \citep{porwollik_generating_2018}.
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from \citep{fao_aquastat_2016} and half-degree resolution physical crop areas from \citep{hurtt_harmonization_2020}, applying the methodology of \citep{porwollik_generating_2018} to identify potential no-tillage grid cells.
+In order to derive a spatial distribution of the three different tillage types specified by the IPCC --- full tillage, reduced tillage and no tillage ---, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from Porwollik et al.~\citeyearpar{porwollik_generating_2018}.
+This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO \citeyearpar{fao_aquastat_2016} and half-degree resolution physical crop areas from Hurtt et al.~\citeyearpar{hurtt_harmonization_2020}, applying the methodology of Prowollik et al.~\citeyearpar{porwollik_generating_2018} to identify potential no-tillage grid cells.
 
@@
 To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{karstens_mrsoil_2020}).
+In the following sections we outline name and idea of these scenarios \citep[for technical implementation see][]{karstens_mrsoil_2020}.
 
@@
 Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
+Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file \citep{century_model_2000} for trees as parameterization were not given.
 Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
@@
 
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in \citep{karstens_model_2020}.
-In the following, the most important results (see \citep{karstens_result_2020} for post-processing script) are summarized.
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens \citeyearpar{karstens_model_2020}.
+In the following, the most important results \citep[see][]{karstens_result_2020} for post-processing script) are summarized.
 
@@
 
+\begin{figure}[h]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
+After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
+Large parts of C, however, get mineralized on field before entering the soil.
+Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
+Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
+\end{figure}
+
 C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
@@
 
-\begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
-\end{figure}
-
 \hypertarget{agricultural-management-effects-on-soc-debt}{%
 \subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
+
+\begin{figure}[h]
+\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+ConstManagement combines all three modifications.
+Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
+Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
+\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
+\begin{figure}[h]
+\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
 \end{figure}
@@
 The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
-
-\begin{figure}
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
-\end{figure}
 
@@
 The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions \citep[see][]{pugh_simulated_2015}, global croplands have acted as a carbon sink since 1975 according to our study.
+However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 \citep{minasny_soil_2017}, indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 
-According to \citep{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as \citep{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and \citep{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
+According to Sanderman et al.~\citeyearpar{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
+Our estimate of 22 GtC in 2010 for cropland debt is higher as Sanderman et al.~\citeyearpar{sanderman_soil_2017} estimations.
+However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.~\citeyearpar{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
 This suggests that our results are within a plausible range.
 
-Furthermore, \citep{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
+Furthermore, Sanderman et al.~\citeyearpar{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+Pugh et al.~\citeyearpar{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
 Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
 Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).
+Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% \citep{pellegrini_crop_2018, ray_recent_2012, rudel_agricultural_2009}.
 
@@
 We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
+This is in contrast to Pugh et al.~\citeyearpar{pugh_simulated_2015} findings of only small effects due to improved practices.
 
-Looking on historic cropland intensification rates of above 50\% (\citep{rudel_agricultural_2009}), a large fraction of increased C input from residue biomass is attributed to productivity improvements.
+Looking on historic cropland intensification rates of above 50\% \citep{rudel_agricultural_2009}, a large fraction of increased C input from residue biomass is attributed to productivity improvements.
 
@@
 High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
-\citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+As pointed out by Keel et al.~\citeyearpar{keel_large_2017} and Smith et al.~\citeyearpar{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
+Keel et al.~\citeyearpar{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
@@
 \ref{fig:SOCinit} the SOC debt estimate almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), and the SOC debt reduction is strongly reduced (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-\citep{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
+Pugh et al.~\citeyearpar{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
 Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes that most of the SOC dynamic happens in the topsoil.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and assumes that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 Secondly changes to the subsoil due to tillage are neglected.
-As \citep{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+As Powlson et al.~\citeyearpar{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
 No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
@@
 This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by \citep{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions (\citep{wang_human-induced_2017}).
+However, as pointed out by Doetterl et al.~\citeyearpar{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions \citep{wang_human-induced_2017}.
 Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
 Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well (\citep{sanderman_soil_2017},\citep{friedlingstein_global_2019}).
+Human interventions have led to large changes in SOC stocks there as well \citep{sanderman_soil_2017, friedlingstein_global_2019}.
 This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 (\citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 \citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}.
 Both estimates are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
@@
 
-\% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 23:33:09 2020
-
 \begin{table}[ht]
 \centering
+\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 \citep{eggleston_ipcc_2006} to the update in 2019 \citep{calvo_buendia_ipcc_2019}.
+leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
+Also modelled $F^{SOC}$ have increased for all climates over time.}
+\label{tab:SCFglo}
 \begin{tabular}{rlllllll}
@@
 \end{tabular}
-\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time.}
-\label{tab:SCFglo}
 \end{table}
 
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than \citep{calvo_buendia_ipcc_2019}.
-The tropical factors differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), with our estimates within this range.
+Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
+The tropical factors differ substantially between the IPCC guidelines in 2006 \citep{eggleston_ipcc_2006} and their refinement in 2019 \citep{calvo_buendia_ipcc_2019}, with our estimates within this range.
 Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
+Our results still show especially in temperate dry regions --- the smallest region area-wise --- small deviations from natural SOC stocks.
 Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
@@
 
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}, \citep{schaphoff_lpjml4_2018-1}) in Table \ref{fig:SOCglo}).
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018, schaphoff_lpjml4_2018-1} in Table \ref{fig:SOCglo}).
 
@@
 The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids (\citep{hengl_soilgrids250m_2017}) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
+SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
 Looking at regional results in Fig.
@@
 Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
-It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
+It is --- to our knowledge --- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
 
@@
 That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT package (\citep{dietrich_madrat_2020}) constitute a starting point for building comprehensive data sets on agricultural management aspects.
+Our data-set and the MADRaT package \citep{dietrich_madrat_2020} constitute a starting point for building comprehensive data sets on agricultural management aspects.
 
-Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions -- not only by the removal of vegetation, but also by a slow depletion of soils.
+Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions --- not only by the removal of vegetation, but also by a slow depletion of soils.
 Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
@@
 Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
-Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand (\citep{popp_land-use_2014-1}) -- and thereby usually further reduce the SOC debt on croplands.
+Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand \citep{popp_land-use_2014-1} --- and thereby usually further reduce the SOC debt on croplands.
 
@@
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils (\citep{brenzinger_organic_2018}).
-Secondly, soil carbon sequestration techniques (\citep{smith_soil_2016}), deep ploughing (\citep{alcantara_deep_2016}) or the transformation of c inputs to more recalcitrant biochar (\citep{woolf_sustainable_2010}) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Waste composting or excreta recycling could represent a major additional N input to cropland soils \citep{brenzinger_organic_2018}.
+Secondly, soil carbon sequestration techniques \citep{smith_soil_2016}, deep ploughing \citep{alcantara_deep_2016} or the transformation of c inputs to more recalcitrant biochar \citep{woolf_sustainable_2010} may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
 Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
-Finally, other carbon-accumulating practices, such as the cultivation of cover crops (\citep{poeplau_carbon_2015}) and agroforestry (\citep{lorenz_soil_2014}) could increase total C sequestration on croplands.
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops \citep{poeplau_carbon_2015} and agroforestry \citep{lorenz_soil_2014} could increase total C sequestration on croplands.
 
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
-All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results including C input data are accessable under \url{https://doi.org/10.5281/zenodo.4320663}.
+\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \citep{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \citep{karstens_mrsoil_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \citep{bodirsky_mrvalidation_2020} for validation data.
+All libraries are based on the MADRaT package at github.com/pik-piam/madrat \citep{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
+Model results including C input data are accessable under \url{https://doi.org/10.5281/zenodo.4320663} \citep{karstens_model_2020}.
 Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.} %% use this section when having data sets and software code available
@@
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
-\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
+\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data.
 Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
@@
 
-\begin{table*}[h]
-\caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
-Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
+\begin{sidewaystable*}[htbp]
+\caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \citet{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
+Allometric coefficients are used as descriped in \citet{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
 }
@@
 \bottomhline
-         &                 & \multicolumn{8}{l}{
-                                           \begin{minipage}[t]{0.50\columnwidth}
+         &                 & \multicolumn{6}{l}{
+                                           \begin{minipage}[t]{0.25\columnwidth}
                                                       nr/dm -- nitrogen to dry matter ratio\\
@@
                                                       \end{minipage}} &
-                                                                        \multicolumn{3}{l}{
-                                                                            \begin{minipage}[t]{0.50\columnwidth}
+                                                                        \multicolumn{5}{l}{
+                                                                            \begin{minipage}[t]{0.25\columnwidth}
                                                                                         HI\textsuperscript{area} -- harvest index per area\\
@@
 \label{tab:c2dm}
-\end{table*}
+\end{sidewaystable*}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
@@
 PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
-We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
+We use the climate zone specification of the IPCC \citep{eggleston_ipcc_2006}.}
 \label{fig:SOCreg}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
-\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
+\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \citealt{century_model_2000}).
+\textit{LitterPNV-CenturyAverage} forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
 Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}

### 2020-12-17 SOCbudgetPaper/conclusion.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

@@
 That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT package ([@dietrich_madrat_2020]) constitute a starting point for building comprehensive data sets on agricultural management aspects.
+Our data-set and the MADRaT package [@dietrich_madrat_2020] constitute a starting point for building comprehensive data sets on agricultural management aspects.
 
-Our study again highlights that the expansion of croplands is still a major source of CO~2~ emissions -- not only by the removal of vegetation, but also by a slow depletion of soils.
+Our study again highlights that the expansion of croplands is still a major source of CO~2~ emissions --- not only by the removal of vegetation, but also by a slow depletion of soils.
 Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
@@
 Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
-Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand ([@popp_land-use_2014-1]) -- and thereby usually further reduce the SOC debt on croplands.
+Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand [@popp_land-use_2014-1] --- and thereby usually further reduce the SOC debt on croplands.
 
@@
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils ([@brenzinger_organic_2018]).
-Secondly, soil carbon sequestration techniques ([@smith_soil_2016]), deep ploughing ([@alcantara_deep_2016]) or the transformation of c inputs to more recalcitrant biochar ([@woolf_sustainable_2010]) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Waste composting or excreta recycling could represent a major additional N input to cropland soils [@brenzinger_organic_2018].
+Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of c inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
 Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
-Finally, other carbon-accumulating practices, such as the cultivation of cover crops ([@poeplau_carbon_2015]) and agroforestry ([@lorenz_soil_2014]) could increase total C sequestration on croplands.
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops [@poeplau_carbon_2015] and agroforestry [@lorenz_soil_2014] could increase total C sequestration on croplands.

### 2020-12-17 SOCbudgetPaper/discussion.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

@@
 The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], global croplands have acted as a carbon sink since 1975 according to our study.
+However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 [@minasny_soil_2017], indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
 
-According to [@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as [@sanderman_soil_2017] estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and [@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.
+According to Sanderman et al.
+[-@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
+Our estimate of 22 GtC in 2010 for cropland debt is higher as Sanderman et al.
+[-@sanderman_soil_2017] estimations.
+However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.
+[-@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.
 This suggests that our results are within a plausible range.
 
-Furthermore, [@sanderman_soil_2017] modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-[@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
+Furthermore, Sanderman et al.
+[-@sanderman_soil_2017] modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+Pugh et al.
+[-@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
 Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
 Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50% ([@pellegrini_crop_2018], [@ray_recent_2012], [@rudel_agricultural_2009]).
+Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
 
@@
 We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to [@pugh_simulated_2015] findings of only small effects due to improved practices.
+This is in contrast to Pugh et al.
+[-@pugh_simulated_2015] findings of only small effects due to improved practices.
 
-Looking on historic cropland intensification rates of above 50% ([@rudel_agricultural_2009]), a large fraction of increased C input from residue biomass is attributed to productivity improvements.
+Looking on historic cropland intensification rates of above 50% [@rudel_agricultural_2009], a large fraction of increased C input from residue biomass is attributed to productivity improvements.
 
@@
 the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study).
-[@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+As pointed out by Keel et al.
+[-@keel_large_2017] and Smith et al.
+[-@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study).
+Keel et al.
+[-@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
@@
 \@ref(fig:SOCinit) the SOC debt estimate almost halves (from ~26 GtC to ~14 GtC), and the SOC debt reduction is strongly reduced (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-[@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
+Pugh et al.
+[-@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
 Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil.
+Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 Secondly changes to the subsoil due to tillage are neglected.
-As [@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
+As Powlson et al.
+[-@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
 No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
@@
 This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by [@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions ([@wang_human-induced_2017]).
+However, as pointed out by Doetterl et al.
+[-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017].
 Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
 Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well ([@sanderman_soil_2017],[@friedlingstein_global_2019]).
+Human interventions have led to large changes in SOC stocks there as well [@sanderman_soil_2017; @friedlingstein_global_2019].
 This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]).
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 [@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019].
 Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
@@
 library(xtable)
-print(xtable(table$comparison_table, digit=2, comment=FALSE, label="tab:SCFglo", 
+print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
               caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 \\citep{eggleston_ipcc_2006} to the update in 2019 \\citep{calvo_buendia_ipcc_2019}.
 leaving our results in very good agreement with the old default assumptions.
 Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time."))
+Also modelled $F^{SOC}$ have increased for all climates over time."), comment=FALSE, caption.placement = "top")
 ```
 
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than [@calvo_buendia_ipcc_2019].
-The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
+Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than Calvo Buendia et al.
+[-@calvo_buendia_ipcc_2019].
+The tropical factors differ substantially between the IPCC guidelines in 2006 [@eggleston_ipcc_2006] and their refinement in 2019 [@calvo_buendia_ipcc_2019], with our estimates within this range.
 Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show  especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
+Our results still show  especially in temperate dry regions --- the smallest region area-wise --- small deviations from natural SOC stocks.
 Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
@@
 
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(fig:SOCglo)).
+The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1] in Table \@ref(fig:SOCglo)).
 
@@
 The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids ([@hengl_soilgrids250m_2017]) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
+SoilGrids [@hengl_soilgrids250m_2017] especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
 Looking at regional results in Fig.

### 2020-12-17 SOCbudgetPaper/introduction.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

--- 2020-12-14 SOCbudgetPaper/introduction.Rmd
@@
 Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth’s soil, is the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times ([@batjes_total_1996]).
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere ([@friedlingstein_global_2019], [@minasny_soil_2017]).
-The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile ([@batjes_total_1996], [@sanderman_soil_2017]).
+It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996].
+Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@friedlingstein_global_2019; @minasny_soil_2017].
+The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile [@batjes_total_1996; @sanderman_soil_2017].
 
 Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018]).
+Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018].
 However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
@@
 The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
-[@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter ([@lal_world_2001]).
+Sanderman et al.
+[-@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter [@lal_world_2001].
 
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions ([@friedlingstein_global_2019]).
+Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@friedlingstein_global_2019].
 While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest ([@strassmann_simulating_2008], [@betts_climate_2015]).
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management ([@friedlingstein_global_2019], [@houghton_carbon_2012], [@hansis_relevance_2015]).
+Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015].
+BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management [@friedlingstein_global_2019; @houghton_carbon_2012; @hansis_relevance_2015].
 
-Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g.
-[@bondeau_modelling_2007], [@lindeskog_implications_2013]).
-[@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
+Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance [e.g.
+@bondeau_modelling_2007; @lindeskog_implications_2013].
+Pugh et al.
+[-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
 They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
@@
 In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.
-growing seasons ([portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g.
-[@pugh_simulated_2015], [@lutz_simulating_2019]).
+growing seasons [@portmann_mirca2000global_2010] multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g.
+@pugh_simulated_2015, @lutz_simulating_2019].
 
@@
 on tillage systems, see [@porwollik_generating_2018], [@prestele_spatially_2018]) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.
-[@iizumi_modeling_2019], [@minoli_modelling_2019]), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+[@iizumi_modeling_2019], [@minoli_modelling_2019]), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
 
-Field-scale models (@del_grosso_simulated_2001], [@coleman_simulating_1997],[@smith_estimating_2010], [@taghizadeh-toosi_c-tool_2014]) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
+Field-scale models [@del_grosso_simulated_2001, @coleman_simulating_1997, @smith_estimating_2010, @taghizadeh-toosi_c-tool_2014] are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
 However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.

### 2020-12-17 SOCbudgetPaper/methods.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

--- 2020-12-14 SOCbudgetPaper/methods.Rmd
@@
 \@ref(sec:carbonbudget) we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.
-4 ([@calvo_buendia_ipcc_2019]).
-We additionally describe how we configured and extended the steady-state method (for model code see [@karstens_mrsoil_2020]).
+4 [@calvo_buendia_ipcc_2019].
+We additionally describe how we configured and extended the steady-state method (for model code see [@karstens_mrsoil_2020].
 In Sect.
-\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]).
-Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see [@bodirsky_mrcommons_2020]).
+\@ref(sec:tier1) we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019].
+Section \@ref(sec:agrimanagement) provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices [for model code see @bodirsky_mrcommons_2020].
 In Sect.
@@
 Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.
-4 ([@calvo_buendia_ipcc_2019]; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon ($SOC$) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+4 [@calvo_buendia_ipcc_2019]; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon ($SOC$) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
 We restrict our analysis to the first 0-30 cm of the soil profile.
@@
 This equation is valid for all grid cells $i$ and all years $t$.
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model ([@parton_analysis_1987]).
+We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987].
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
@@
 \@ref(sec:residues)) and manure inputs (see Sect.
-\@ref(sec:livstmanure)); for natural vegetation litterfall (including fine root turnover) ([@schaphoff_lpjml4_2018]) is the only source of carbon inflow to the soil.
+\@ref(sec:livstmanure)); for natural vegetation litterfall including fine root turnover [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
 Following the IPCC guidelines [@calvo_buendia_ipcc_2019], carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
@@
     \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                        \cite{faostat_faostat_2016}, \\
-                                                                        \cite{schaphoff_lpjml4_2018},\\
-                                                                        \cite{weindl_livestock_2017} \end{minipage}} & 
+                                                                        \citet{faostat_faostat_2016}, \\
+                                                                        \citet{schaphoff_lpjml4_2018},\\
+                                                                        \citet{weindl_livestock_2017} \end{minipage}} & 
                                                         \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                         LG:C generic values according to Table 5.5B,\\
-                                                                        5.5C in \cite{calvo_buendia_ipcc_2019},\\
-                                                                        crop-specific N:C from \cite{bodirsky_n2o_2012}
+                                                                        5.5C from IPCC \citep{calvo_buendia_ipcc_2019},\\
+                                                                        crop-specific N:C from \citet{bodirsky_n2o_2012}
                                                                         \end{minipage}} \\
@@
                               \hline
-    Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
+    Natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
                               & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                  \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
+                                  IPCC \citep{calvo_buendia_ipcc_2019} and\\ 
+                                  CENTURY (\citep{century_model_2000}) \strut \end{minipage}\tabularnewline
    \bottomhline
@@
 We calculated area weighted means for $till$ and $wat$ on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \@ref(tab:datasourcedecay); equations are displayed by equation 5.0B--5.0F in [@calvo_buendia_ipcc_2019].
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \@ref(tab:datasourcedecay); equations are displayed by equation 5.0B--5.0F in Calvo Buendia et al.
+[-@calvo_buendia_ipcc_2019].
 
@@
 
-According to the IPCC guidelines $SOC$ changes can be expressed as the difference of two consecutive years (see Eq.
-5.0A in [@calvo_buendia_ipcc_2019]).
+According to the IPCC guidelines $SOC$ changes can be expressed as the difference of two consecutive years [see Eq.
+5.0A in @calvo_buendia_ipcc_2019].
 This, however, will also include naturally occurring changes due to climatic variation over time.
@@
 
-Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt ([@sanderman_soil_2017]) due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_ipcc_2006].
-Moreover, $\Delta SOC$ is equivalent to the negated cumulative SOC component of human land-use change emissions ([@pugh_simulated_2015]).
+Note that the absolute changes $\Delta SOC$ can be also interpreted as the SOC debt [@sanderman_soil_2017] due to human cropping activities; whereas relative changes $F^{\mathrm{SCF}}$ can be considered stock change factors as defined within the IPCC guidelines of 2006 [@eggleston_ipcc_2006].
+Moreover, $\Delta SOC$ is equivalent to the negated cumulative SOC component of human land-use change emissions [@pugh_simulated_2015].
 
@@
 
-Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]).
+Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019].
 Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
@@
 with $T_{c,i}$ being the translation matrix for grid cells $i$ into corresponding climate zones $c$.
-For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of [@eggleston_ipcc_2006] and [@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default $F^{\mathrm{SCF}}$ from the Tier 1 method of Eggleston et al.
+[-@eggleston_ipcc_2006] and Calvo Buendia et al.
+[@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016] (see FAO2LUH2MAG_croptypes.csv in [@bodirsky_mrcommons_2020]).
+We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups [see FAO2LUH2MAG_croptypes.csv in @bodirsky_mrcommons_2020], applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016].
 By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 
-Crop residue production and management is based on a revised methodology of [@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modeling.
+Crop residue production and management is based on a revised methodology of Bodirsky et al.
+[-@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modeling.
 Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest index values ($HI$) and root:shoot ratios ($RS$) as follows
@@
 Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_ipcc_2006].
-Deviating from [@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+Deviating from Bodirsky et al.
+[-@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
 We assume that all $BGR$ are left in the soil, whereas $AGR$ can be burned or harvested for other purposes such as feeding animals [@weindl_livestock_2017], fuel or for material use.
@@
 A country-specific fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
-Following [@smil_nitrogen_1999]) we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following Smil [-@smil_nitrogen_1999] we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
 Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_ipcc_2006].
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in [@bodirsky_mrcommons_2020]).
-Residue feed demand for five different livestock groups is based on country-specific feed baskets (see [@weindl_livestock_2017]), that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.
+[-@bodirsky_mrcommons_2020].
+Residue feed demand for five different livestock groups is based on country-specific feed baskets [see @weindl_livestock_2017], that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
 We estimate a material-use share for the straw-residue group of 5% and a fuel-share of 10% for all used residue groups in low-income countries.
@@
 
-To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of [@robinson_mapping_2014] and applying feed basket assumptions based on a revised methodology from [@weindl_livestock_2017].
+To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of Robinson et al.
+[-@robinson_mapping_2014] and applying feed basket assumptions based on a revised methodology from Weindl et al.
+[-@weindl_livestock_2017].
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
@@
 
-Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modeling.
+Manure production and management is based on a revised methodology of Bodirsky et al.
+[-@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modeling.
 Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see [@calvo_buendia_ipcc_2019]).
+Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.
+[-@calvo_buendia_ipcc_2019].
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in [@bodirsky_n2o_2012] (see Sect.
-\@ref(append:TableclossAWMS) for more details).
+For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.
+[-@bodirsky_mrcommons_2020] for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.
+[-@bodirsky_n2o_2012].
 We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.
@@
 
-In order to derive a spatial distribution of the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from  [@porwollik_generating_2018].
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from [@fao_aquastat_2016] and half-degree resolution physical crop areas from [@hurtt_harmonization_2020], applying the methodology of [@porwollik_generating_2018] to identify potential no-tillage grid cells.
+In order to derive a spatial distribution of the three different tillage types specified by the IPCC --- full tillage, reduced tillage and no tillage ---, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from Porwollik et al.
+[-@porwollik_generating_2018].
+This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO [-@fao_aquastat_2016] and half-degree resolution physical crop areas from Hurtt et al.
+[-@hurtt_harmonization_2020], applying the methodology of Prowollik et al.
+[-@porwollik_generating_2018] to identify potential no-tillage grid cells.
 
@@
 To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see [@karstens_mrsoil_2020]).
+In the following sections we outline name and idea of these scenarios [for technical implementation see @karstens_mrsoil_2020].
 
@@
 Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines [@calvo_buendia_ipcc_2019], but for woody biomass we used the CENTURY configuration file ([@century_model_2000]) for trees as parameterization were not given.
+Perennial grasses are given by the IPCC guidelines [@calvo_buendia_ipcc_2019], but for woody biomass we used the CENTURY configuration file [@century_model_2000] for trees as parameterization were not given.
 Whereas LG:C ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, N:C ratios range between $0.001-0.05$ with an average of $0.015$ (default value).

### 2020-12-17 SOCbudgetPaper/results.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

@@
 
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in [@karstens_model_2020].
-In the following, the most important results (see [@karstens_result_2020] for post-processing script) are summarized.
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens [-@karstens_model_2020].
+In the following, the most important results [see @karstens_result_2020] for post-processing script) are summarized.
 
@@
 
+```{r caption_FlowFig, echo=FALSE}
+caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
+After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
+Large parts of C, however, get mineralized on field before entering the soil.
+Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
+Finally, SOC gets mineralized and flows back into the atmosphere."
+```
+
+```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/CarbonBudget.png")
+```
+
 C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
@@
 
-```{r caption_FlowFig, echo=FALSE}
-caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere."
+## Agricultural management effects on SOC debt
+
+```{r captionSOCscen, echo=FALSE}
+captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+ConstManagement combines all three modifications.
+Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
+Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
 ```
 
-```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
+```{r SOCscen, out.width="14cm", fig.cap = captionSOCscen, echo=FALSE, fig.pos="h"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/CarbonBudget.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_horiz.png")
 ```
-
-## Agricultural management effects on SOC debt
 
@@
 
-```{r captionSOCscen, echo=FALSE}
-captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
+```{r captionSOCinit, echo=FALSE}
+captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use."
 ```
 
-```{r SOCscen, out.width="14cm", fig.cap = captionSOCscen, echo=FALSE}
+```{r SOCinit, out.width="10cm", fig.cap = captionSOCinit, echo=FALSE, fig.pos="h"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_horiz.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_init.png")
 ```
@@
 
-```{r captionSOCinit, echo=FALSE}
-captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use."
-```
-
-```{r SOCinit, out.width="10cm", fig.cap = captionSOCinit, echo=FALSE}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_init.png")
-```
-
 Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas $\Delta SOC$ vary  19-24 GtC for the year 2010 (see Fig.

### 2021-04-01 ResultNotebooks/results_SOCflows.Rmd
add revision point by point replies

@@
 
-LivstCropLoss     <- CropFEED - ManureExcretedTot - LivstCropHANPP 
+LivstCropLoss     <- CropFEED + ResFEED - ManureExcretedTot - LivstCropHANPP 
 CropPRIMLoss      <- CropPRIM + ResPRIM + LivstCropHANPP # assume verything gets re-emitted at the end

### 2021-04-01 ResultNotebooks/results_SOCmaps.Rmd
add revision point by point replies

@@
 
-rev <- "rev16"
+rev <- "rev19"
+setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
 
@@
 
-scenarios    <- c("histManagement","constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
-                  "constManagement2-1975", "constResrate-1975", "constYield-1975",
-                 "Initial-natveg", "Initial-natveg_constManangement-1975", "Initial-natveg_constManangement2-1975", 
-                 "LitterPNV-PerennialGrasses", "LitterPNV-AASS", "LitterPNV-BRCF", 
-                 "LitterPNV-MTCF", "LitterPNV-BLEAF",  "LitterPNV-TREF")
+scenarios    <- c("histManagement")#,"constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
+                  #"constManagement2-1975", "constResrate-1975", "constYield-1975",
+                 #"Initial-natveg", "Initial-natveg_constManangement-1975", "Initial-natveg_constManangement2-1975", 
+                 #"LitterPNV-PerennialGrasses", "LitterPNV-AASS", "LitterPNV-BRCF", 
+                 #"LitterPNV-MTCF", "LitterPNV-BLEAF",  "LitterPNV-TREF")
 
-# for(item in scenarios){
-#    dir.create(path=paste0("Data/",paste0(rev,item)))
-#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-#    files=filesNeeded, exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ")
-# }
+for(item in scenarios){
+   dir.create(path=paste0("Data/",paste0(rev,item)))
+   untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+   files=filesNeeded, exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ")
+}
 
@@
 
-SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))
-Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
+SoilCarbon <- read.magpie(paste0("Data/",rev,scenarios[1],"/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/",rev,scenarios[1],"/Landuse.rds"))
 
@@
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(trans = squish_trans(-18,0,10),
-                     limits = c(-28,0),
-                     breaks = c(0, -20, -22, -24, -26, -28)) +
+  scale_y_continuous(trans = squish_trans(-42,0,10),
+                     limits = c(-46,0),
+                     breaks = c(0, -42, -43, -44, -45, -46),
+                     minor_breaks = seq(-2,-42,-2)) +
   theme(rect=element_blank(),
@@
 ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid, width = 20, height = 14, dpi = 300, units = "cm")
+ggsave(paste0(outputdir,"/SOCshare.png"),    plot=SOCshare,    width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir,"/SOCincrease.png"), plot=SOCincrease, width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir,"/SOCdensity.png"),  plot=SOCdensity,  width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir,"/SOCemis.png"),     plot=SOCemis,     width = 10, height = 7, dpi = 300, units = "cm")
 

### 2021-04-01 ResultNotebooks/results_SOCscenario.Rmd
add revision point by point replies

@@
   html_document:  
+    pdf_document: default
 ---
@@
 
-rev         <- "rev16"
+rev         <- "rev19"
 scenarios   <- c("histManagement","constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
@@
 
-# filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
-# setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-# 
-# untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
-#       exdir="Data/histManagement", extras="--keep-newer-files ")
-# 
-# for(item in scenarios){
-#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-#    files=filesNeeded, exdir=paste0("Data/", item), extras="--keep-newer-files ")
-# }
+filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
+setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
+
+untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
+      exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")
+
+for(item in scenarios){
+   untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+   files=filesNeeded, exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
+}
 
@@
 for(scen in scenarios){
-  SoilCarbon <- read.magpie(paste0("Data/", scen, "/SoilCarbon.rds"))
-  Landuse    <- read.magpie(paste0("Data/", scen, "/Landuse.rds"))
+  SoilCarbon <- read.magpie(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
+  Landuse    <- read.magpie(paste0("Data/", rev, scen, "/Landuse.rds"))
   SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
@@
 
-limits <- c(-30,-21)
-breaks <- c(-22, -24,- 26, -28, -30)
+limits <- c(-48,-42)
+breaks <- c(-42, -44,- 46, -48)
 
@@
 
-limits <- c(-30,-10)
-breaks <- c(-10, -15,- 20, -25, -30)
+limits <- c(-50,-20)
+breaks <- c(-20, -30,- 40, -50)
 
@@
 
-limits <- c(-30,-18)
-breaks <- c(-18, -22, -26, -30)
+limits <- c(-48,-40)
+breaks <- c(-40, -42, -44, -46, -48)
 

### 2021-04-01 ResultNotebooks/results_SOCtable.Rmd
add revision point by point replies

@@
 
-rev       <- "rev16"
+scenarios <- c("rev19histManagement","rev18histManagement_ResAllom_2021_02_01")
 outputdir <- "Output/"
@@
 
-SoilCarbon <- read.magpie(paste0("Data/histManagement/SoilCarbon.rds"))
-Landuse    <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
-
+SoilCarbon <- read.magpie(paste0("Data/",scenarios[2],"/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/",scenarios[2],"/Landuse.rds"))
 ```
@@
 MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
-                                                                          getNames(climateIPCC), paste, sep="."))
+                                                                          getNames(climateIPCCr), paste, sep="."))
 

### 2021-06-07 Revisions/Point by point replies - Jonathan Sanderman.md
update outouts needed for final response

# Point by point replies - Jonathan Sanderman

0. This MS presents the findings of a spatially explicit implementation of a new IPCC TierII approach to soil carbon stock changes. The authors use this model to calculate how much SOC has been lost due to agriculture but then go on to run the model annually for the period 1975 – 2010 to produce a dynamic picture of SOC recovery over this modern era of farming.   The major takeaway message is that while agriculture has incurred a large SOC debt, recent agronomic improvements have led to 4 Pg C of SOC sequestration over this period of time. This detailed picture of SOC in croplands over the past several decades is of incredible importance to policy makers and as such I believe that paper can be an important contribution to the literature; however, I do have several major concerns that may or may not be addressable with revisions.

Answer to 0.:
Dear Dr. Sanderman, thank you for the thorough and helpful review of our manuscript. While checking and revising our data processing in response to your and the other reviewer's feedback, we discovered a bug in the very core of the soil model, leading to an overestimation of the transfer of carbon from active to the slow pool *just* for cropland. Additionally, we found unreasonable high forage crop production values (specifically for pumpkins used as fodder) in our input data, which are taken from FAO statistics. This made the overall intensification trend in agriculture lead to increasing carbon stocks in cropland. After correcting the bugs, this is no longer the case. Whereas this implies major revisions of our discussion and interpretation of results, we find that the essence of the paper, which is a) the availability of a soil carbon model that can account for changes in agricultural management and can be applied within integrated assessment frameworks and b) an analysis of the importance of accounting for management dynamics in SOC assessments, remains unchanged, albeit modified. We provide an assessment of how results changed after correcting the bugs as a supplement to this authors comment. Here, we respond point-by-point to the reviewers' comments. Where appropriate, this will also address the implications of the bugfix. We look forward to your response.

1. I had to read the methods section twice and spend an hour with Calvo Buendia et al., 2019 to fully understand what the authors have done. I’m still not 100% confident that I fully understand the methodology. I suggest adding an illustrative example or two graphically demonstrating how the process works. Perhaps starting with a simple case of one lu transition and then showing a more complex case of multiple lu transitions within a pixel.

Answer to 1.:
We will add more detail and graphical representation for demonstrating especially the land-transition accounting.

2. I have not been convinced that this sort of “dynamic” implementation of a steady state modeling approach is appropriate. I understand that the method was developed by the IPCC as a way to add more nuance into the Tier I emission factor approach but I don’t think the method was intended to be applied annually.  Why not go all the way to TierIII approach using the process-based dynamics that are embedded in this simplified model?   It  appears  you  have  all  the  data  assembled  to  do  this.   My  main  concern with applying a steady-state model to annual changes is we know that the recent past trajectory of SOC (particularly in the slow and passive pools) will greatly influence the short-term model response to improved management – i.e.  the model will take years to decades before SOC stocks start to rebuild if the trajectory was negative prior to the change – but this will be completely missed with the steady state application (stocks will start increasing immediately upon change).

Answer to 2.:
This has already been answered in the earlier reply ('Reply on RC1', Kristine Karstens, 26 Jan 2021).

3. What time frame are the monthly climate data averaged over to get the rate modifiers for a steady state solution? Given the passive pool has an intrinsic decay rate equivalent to >100 year turnover time, it seems that you need to have a 100+ year average climate to come up with the proper rate modifiers.

Answer to 3.:
As we do not use a steady state model, we believe the comment might be based on a misunderstanding and no averaging of climate data is needed.

4. Transfer between lu types is not clear.  I do not understand how a “respective share of the SOC is reallocated.”  My concern is that the per area SOC stock for long-term cultivated land will be much different than the per area SOC stock for newly converted cropland, so I don’t see how you can suddenly bin these separate areas into one model component.  Perhaps my request for a visual guide will help me (and other readers)understand that I have misunderstood this part of the methods.

Answer to 4.:
See above: We will add more detail and graphical representation here. This will illustrate the linearity of the problem: Accounting separately for freshly converted cropland and old cropland is in our equation system essentially the same than taking the area weighted mean.

5. Why was 1901 chosen to spin up to steady state?  We know that this was a time of rapid agricultural expansion in several major regions of the world and thus a time of rapid soil carbon loss.

Answer to 5.:
We extended the spin-up phase for a much longer period, starting in 1510 (default spin-up start in LPJmL) to capture the land-use effects. Our initial choice of starting in 1901 was driven by the climate data availability.  To extend our spin-up into the past, we will recycle climate data from 1901 to 1930, as this is done in DGVM simulations as well (Schaphoff et al. 2018a/b, von Bloh et al. 2018). The management data will be held constant on the first known level.

6. LULUC data – are these data all provided as area within each grid cell (or percent of a grid)? I think so, but please indicate. A supplemental table with the cross-walk between the LUH2 and the 17 crop groups used in this study should be included.

Answer to 6.:
Yes, all area data are in hectares and meant to represent the area within each grid cell. We will indicate that more clearly. Additionally, we will add a table to the supplemental material.

7. Please include units for eq 9 – I don’t follow the AGR calculation – it sounds like you are adding biomass and area together. Additionally, HI is usually defined as CP divided by total above ground biomass, so CP x HI is a meaningless number.L

Answer to 7.:
Our harvest index for the yield (t/ha) is calculated based on a linear function with positive intercept (ha) and a slope dependent on the yield (t/ha). This allometric function accounts for the fact that higher-yielding crops often have a lower harvest index than low-yielding crops. In our revised manuscript, we will rewrite the equations to make the functional form more visible. Instead of 
AGR_i,t,cg = CP_i,t,cg ·HI_prod,cg +CA_i,t,cg ·HI_area
we now write
AGR_i,t,cg = CA_i,t,cg (Y_i,t,cg ·HI_prod,cg + HI_area)
We will also add more explanations to the text.

7. Lack of validation. There appears to be no attempt to validate the model or the input data used to drive the model. In general, there is a lack of quantitative evaluation throughout.  There are just two qualitative quality assessments – a table comparing calculated stock change factors to Tier I estimates and a discussion on how the map looks similar to other SOC maps.

Answer to 7.:
To address the raised issue, we will add to our analysis:
* A grid level comparison of SOC stock results to SoilGrids 2.0 and LPJmL to improve spatial evaluation of our results.
* More literature comparison of our derived agricultural land emissions as they are next to the SOC stocks the most central outcome of our analysis.
* An extended qualitative discussion of uncertainty and where it might be the largest.

8. The  model  itself  was  developed  recently  as  part  of  the  2019  Refinements  to  IPCC Guidelines and those updated guidelines discuss how the model was calibrated to a set of long-term trial sites but do not report any model performance metrics. As pointed out by the authors some areas of central EU and the UK more than double SOC undercurrent agriculture than under native vegetation.  This is certainly indicative that there should be some checks against real data (see detailed elaboration on this point further down in this review).  It could be argued that point-based validation for a model run at0.5 degree resolution is meaningless but it would be an interesting exercise to see how the model reproduces trends with known SOC histories.

Answer to 8.:
Our updated model does not show the superising behaviour of strongly increasing SOC stocks in UK and central EU anymore due to the bugfix within the model and the corrected fodder production data. We will add a more detailed spatial comparison to SoilGrids 2.0 (SOC stock estimates based on point data measurements) and other model based SOC estimates (LPJmL) to account for the missing spatial dimension of model output evaluation,

9. Issues with residue C return. Given that the major takeaway from this paper is that the SOC is being sequestered due to improved yield leading to increased residue return, is there any empirical evidence that C inputs to the mineral soil have nearly doubled (Fig 3)? I think the method for calculating residue return to the soil is potentially flawed leading to this large apparent increase. The authors have assumed that both harvest index (HI) and root-to-shoot(RS) ratios have been constant through time.  However, yield improvements over the last century, and in particular the last 50 years, are a result of improvements in genetics and nutrition.  Breeding has resulted in the ability to plant most crops at much higher densities and selection towards more photosynthate being allocated to harvestable organs.  Both of these improvements have altered HI and RS ratios.  Additionally, there are strong interactions between N fertilization rate and root density.  There is a huge literature on crop breeding that support the nonstationarity of these important parameters.

Answer to 9.:
This is a very valid point that we tried to take up in the discussion part as well. We add more detail to the discussion and would like to point out the following thing:
The IPCC methodology, which we use here, tries to capture the effect of a shifting harvest index by making the harvest index a linear function of yield with a positive intercept. The parametrization of this dynamic harvest index was not possible for all crops due to a lack of literature estimates, but the most important crops like cereals or soybeans are captured.

10. Units – Gt and Mt are not SI units, please use Pg and Tg

Answer to 10.:
As In the integrated assessment community tons is most often used for reporting emissions, tons seemed the natural choice for reporting anthropogenic emissions. 
We now came aware that in natural science the Si unit system is the standard unit system. We will change all units to SI and add non-SI units in brackets were needed to bridge to IAM modelling community.

11. L13 (and elsewhere) – “we also find that SOC is very sensitive...” – this is in reference to an unvalidated model result. I’d suggest rewording these sorts of phrases to, “Our model results suggest that SOC is very sensitive...

Answer to 11.:
We will do a careful revision of naming and framing of the evaluation related statements.

12. L279 – “we provide the first world map” – no, you did not. All of the global maps that have been developed using a statistical environmental-covariate modeling approach(i.e. soilgrids and similar) implicitly include all historic land management.

Answer to 12.:
This is correct. We will rewrite the paragraph.

13. It is great that all the data are provided but I found the Karstens 2020a repository to be confusing. Can you have a description for each file in the repository? The naming convention is not clear. I did not want to download 9 Gb of data to figure it out.

Answer to 13.:
We will follow this suggestion and will add more detail to the README file including a description for each file.

14. Fig 1 – perhaps it is just the spatial scale of these small maps (and I haven’t downloaded the results to explore in more detail) but it looks like there is zero intact forest in the Congo Basin and very little intact forest in the Amazon. Also, I would have likedto see a map showing the trend in SOC spatially – how are the 4 Pg C that has been sequestered been spread across the globe? Is it all in Central EU and UK?

Answer to 14.:
Fig 1.(a) is showing cropland SOC,for every grid cell containing cropland without giving an information on the extend of cropland. We will mask out cells with very low cropland area shares, as they might let to the impression of greater cropland extent within large parts of intact forest area. We will also provide a total SOC map in the appendix.

15. Fig 2 – this is a really interesting way of summarizing the model results
Answer to 15.:
Thanks.

16. Fig 4 – the finding presented here is very counterintuitive to me. Why is the SOC debt halved when the model is initialized with natural vegetation? Shouldn’t the 1975 SOCdebt be much greater if the 1901 starting point was natural vegetation instead of actual land use? Perhaps I am just misunderstanding this sensitivity analysis.

Answer to 16.:
The initialization analysis was meant to help the reader understanding the potential maximal error of underestimating on-going emissions on croplands, that where converted to croplands before 1901. We extended the spin-up phase till 1510 now. We will move the initialization analysis to the appendix and improve the caption of the figure to make its interpretation clearer there..

17. Discussion section – in general, there is very little discussion of how these results fit into the large literature on SOC. There are many places were a reference or two would great increase the credibility of the statements that are being made.

Answer to 17.:
We will include more connections to existing literature here.
Section 4.2 – I think this section should come right before the conclusions especially as you refer to analysis that is only presented for the first time in section 4.3
Section 4.3 will be moved to the end of the result as evaulation results subsection.

18. L358-360 – the finding that northern temperate zones (particularly in EU and UK) now have SOC levels up to twice that of native state yet tropical soils have lost 40-70% of their SOC is problematic and, as the authors point out in relation to the EU example,likely points to issues with getting C input to soil correct. The EU has the perfect data set to test this model finding – the EU JRC LUCAS survey was a balanced sampling design between forested and agricultural land uses. In the tropics, it has been fairly well documented that already infertile tropical soils do not lose nearly as much SOC as their fertile temperate zone soil counterparts. While there are issues and large scale generalizations in the IPCC Tier I default factors, they do represent the consensus literature on the topic. The updated meta-analysis between the 2006 and 2019 IPCCC5guidelines when this emission factor for the tropics changed dramatically (see Table 4in this MS) points to this new knowledge.

Answer to 18.:
After our bugfix SOC stocks for the EU do not gain SOC compared to natural vegetation anymore. In general all carbon stocks are much smaller and show much higher losses compared to the Tier 1 approach for both temperate and tropical soils. This might very well still point to missing accounting of carbon inputs to the soil and will be discussed in more detail in the discussion part.
The comparison to point data is however challenging, also as point measurements do not well capture the landscape average and will likely show a very high variance. On the one hand the LUCAS database is given as soil carbon density and would need consistent bulk density data to be comparable to our results. A full comparison to point measurements may also exceed the scope of this first article. We hope with our additional comparison to SoilGrids 2.0 might help fill the evaluation gap here as SoilGrids 2.0 is based on point measurements.

19. L396 – how is this validation? It is just a comparison.

Answer to 19.:
Naming and framing of the evaluation of our results will be improved.

20. L400-406 – there is a large literature that can be drawn upon to support some of the claims made in this section.

Answer to 20.:
We will add more detail and references here

21. Section 4.4 – I do not think this is a valid comparison because SoilGrids explicitly tried to capture high carbon density soils well while your model explicitly excludes organic soils. I suggest applying an agriculture mask to all of these data sets and then redo the analysis. Additionally, ISRIC released an update to SoilGrids >6 months ago that focuses primarily on mineral soil carbon stocks. This update is probably a better comparison.

Answer to 21.:
We included the recommended SoilGrids 2.0 data for a more spatially detailed evaluation of our results especially for cropland soils.

22. L425-428 – this seems out of place.

Answer to 22.:
We will include sensitivity discussions in our extended discussion on evaluation and uncertainty of our results.

23. L453 – comparison to 4p1000 is not really fair because your model is really just the business-as-usual scenario with SOC gains simply because yields are improving glob-ally. 4p1000 is about intentional management shifts to increase SOC.

Answer to 23.:
We agree with the reviewer, that the comparison is a bit misleading. We still think, it is fruitful to compare the observed rates with ambitious targets and will reframe the discussion on this point.

### 2021-06-07 Revisions/Point by point replies - second reviewer.md
update outouts needed for final response

# Point by point replies - second reviewer

0. The authors have conducted a study evaluating the influence of management on soil organic carbon in global croplands. This is an important topic for consideration of greenhouse gas mitigation with natural solutions for climate change policy and pro-grams such as the 4 per mille initiative. As the authors mention, there are few studies that have evaluated cropland management effects on soil organic carbon, and possibly none that have addressed the influence at a global scale. The result that increased residue return to soils is the leading driver of carbon changes over the past few decades in croplands is an important finding. As the authors note, the 4 Gt C increase in carbon is less than the goals of the 4 per mille initiative, which some have argued is not realistic. I have a few concerns about the study after review of the IPCC documentation on the method that the authors selected for this analysis. I would suggest that the authors make revisions before the manuscript is accepted.

Answer to 0.:
Dear reviewer, thank you for the thorough and helpful review of our manuscript. While checking and revising our processing in response to your and the other reviewer's feedback, we discovered a bug in the very core of the soil model, leading to an overestimation of the transfer of carbon from active to the slow pool *just* for cropland. Additionally, we found unreasonable high forage crop production values (specifically for pumpkins used as fodder) in our input data, which were taken from FAO statistics. 
This made the overall intensification trend in agriculture lead to increasing carbon stocks in cropland. After correcting the bugs, this is no longer the case. 
Whereas this implies major revisions of our discussion and interpretation of results, we find that the essence of the paper, which is a) the availability of a soil carbon model that can account for changes in agricultural management and can be applied within integrated assessment frameworks and b) an analysis of the importance of accounting for management dynamics in SOC assessments, remains unchanged, albeit modified. We provide an assessment of how results changed after correcting the bugs as a supplement to this authors comment. Here, we respond point-by-point to the reviewers' comments. Where appropriate, this will also address the implications of the bugfix. We look forward to your response.

1. The Tier 2 method is in a croplands chapter of the IPCC report. The documentation in the report states that the model would need to be parameterized for other land uses.Did the authors parameterize the model for other land uses that would be considered natural vegetation? If not, the estimation of soil organic carbon for natural vegetation may not be valid. The authors seem to suggest that this is a possibility in Section 4.4 when stating the soil organic carbon and debt from land use change have to be interpreted with caution. If the model has not been parameterized for natural vegetation, I would suggest that the authors focus on cropland model results, and remove the carbon debt results. The results for the cropland alone are important, and deserve publication even if the natural vegetation estimates are not valid with this model.

Answer to 1.:
This is an important point. While our analysis focuses on croplands, estimating the natural soil carbon stocks is a necessity to account for the C entering the cropland budget via land conversion. Apart from this process, we only use natural vegetation SOC to make our results comparable to other global estimates for validation purposes.
To improve the parametrization of natural soil carbon, we will include the following model updates:
* We will improve the litterfall parameterization in natural vegetation.
* We will compare our results on soil stock under natural vegetation with the results of a model parametrized for natural vegetation (LPJmL).

2. Is it possible to estimate uncertainty with this method? IPCC methods often have large uncertainty but does this method have less uncertainty because it is a Tier 2method. If it is not possible to estimate uncertainty could the authors speculate on the level of uncertainty in the predictions. Knowing something about uncertainty would be helpful in comparisons to the modeled results from other studies that are shown in the manuscript.

Answer to 2.:
The quantitative assessment of the uncertainty of our projections unfortunately exceeds the scope of this article and would likely require a study on itself. The model includes a high number of parameters, and for most of these the uncertainty distributions have not been quantified so far. Moreover, we think that beyond parameter uncertainty, the structural uncertainty from the model design is also very high.

To still address the uncertainty within this manuscript, we will therefore discuss the uncertainty qualitatively. Most of the uncertainty in our view is included in the management data, especially in the manure numbers, as they are calculated from feed data. Additionally residue and litter estimations are based on the set of global parameters, that contains large uncertainty and at the same time have a great impact on the overall size of carbon inputs to the soil. Both effects will be discussed in more detail.

3. The authors state that a sensitivity analysis presented in Figure 4 shows that management impact is robust to the initialization of the soil organic carbon stocks at the beginning of the spin-up phase. But, the stocks and change in stocks almost halves the values if the initialization is done with natural vegetation. The initialization does make a difference, and needs further explanation.

Answer to 3.:
It is correct that the SOC stocks are highly dependent on the legacy of management. Figure 4 made however clear, that the SOC gap - so the difference between a baseline scenario and a counterfactual scenario with constant management - was not strongly dependent on the initialization. We thereby showed that the initialization does not affect our central finding.

Still, in order to improve our estimates also for the absolute SOC stock, we now extended the length of the spin-up phase, starting in 1510 (default spin-up start for introducing land use in simulations with LPJmL, see e.g. Schaphoff et al. 2018a/b, von Bloh et al. 2018 ). We will analysis and discuss our results with respect to the initialization in more detail.

4. Good to see that the authors have made a comparison to another approach to confirm the Tier 2 results. The Tier 1 method provided by the IPCC has been used for this purpose. In section 2.2, the authors present a method estimating stock change factors instead of soil organic carbon changes. But, the results in Table 4 for the stock change factors are not convincing that the methods are consistent, and the text seems unclear with discussion about larger differences with the IPCC 2019 values, which were updated by the authors and should be more accurate – I would think. Why not estimate the change in soil organic carbon for a direct comparison with the Tier 2 method instead of the stock change factors? Also, the placement of these results after the discussion seems odd, and conventionally would be presented in the results section before discussion.

Answer to 4.:
Thank you for the recommendation. We will place the comparison within the result section. For a Tier 1 approach the change in soil organic carbon has to be calculated based on a reference stock. The default stocks given by the IPCC have very low spatial resolution (42 coarse climate zone and soil type specific values), which leads to additional uncertainty when comparing directly changes in soil organic carbon.
We will add more detail to the discussion about the strong deviation of 2006 and 2019 default stock changes factors.

5. Figure 3 shows results from making certain practices constant from 1975 to 2010.The authors state around line 315 that the effect of no-till has been strong since 1990,but the effect seems minor and may not differ statistically from the histManagement with uncertainty. The conclusion about the importance of residue seems most important here.

Answer to 5.:
Due to the overall changes to the results, we will rewrite major part of the result and discussion. We will take more thoroughly into account effects of uncertainty.

6. The authors suggest that there needs to be a circular flow with food supply chain back to soils. They assumed that none of the waste from supply chains are returned to soils (near line 300) but this seems incorrect. Municipal waste and materials are amended to soils in many regions of the world although maybe there are no data on these amendments? If this is the issue, the authors could mention that they are making a conservative assumption due to lack of data.

Answer to 6.:
Indeed, further soil inputs include the application of human excreta and sewage sludge, as well as the application of processing wastes, forestry residues or biochar. Unfortunately, data on the quantity of these inputs is very scarce and often does not exist with global coverage.
In the draft of the International Nitrogen Assessment (will be provided confidentially to the reviewer), the available literature estimates for these flows (in regard to nitrogen) were reviewed. So far no reliable estimates exist, but the existing estimates indicate that these inputs must be by an order of magnitude lower than those by crop residues and manure.

7. The authors evaluate the sensitivity of the Tier 2 model for tree litter with methods in Section 2.4.3. The Tier 2 model divides litter into metabolic and structural components,and the authors have averaged lignin to nitrogen across tree components as input to the model. But forest also include deadwood and should be separated from other forest litter to model decomposition. Did the authors add a deadwood pool? I question if this model is appropriate for forest if deadwood is not modeled separately.

Answer to 7.:
We follow this suggestion and will add more detail on the litterfall of the natural vegetation. Using additional information from LPJmL, we will split up the litterfall into a wood- and soft-tissue fraction and add different parameterizations for these. 
We will not be able to add a deadwood pool, since that would require additional parameterization of turnover dynamics for this new pool. Deadwood pools are also not treated explicitly in many DGVMs but are considered part of litter pools, which distinguish woody from non-wood litter pools. The separation of litterfall into wood and soft tissue fluxes will thus add similar stock detail as in DGVMs.

8. Recommend that the authors provide more explanation for Equation 9, which determines the residue amount of C, and is a key driver of the carbon change. Harvest index is the proportion of plant biomass that is harvested, but the authors are multiplying the harvested crop product by the harvest index. But the conventional approach is ’harvested crop production divided by the harvest index’ to determine the total biomass and then subtract the harvested amount to estimate the residue. The authors are accounting for double harvesting and fallow in this calculation, which I agree is important, but some further explanation is needed about the calculation to understand how residue carbon is estimate from crop production, harvest index and area.

Answet to 8.:
Our harvest index for the yield (t/ha) is calculated based on a linear function with positive intercept (ha) and a slope dependent on the yield (t/ha). This allometric function accounts for the fact that higher-yielding crops often have a lower harvest index than low-yielding crops. In our revised manuscript, we will rewrite the equations to make the functional form more visible. Instead of 
AGR_i,t,cg = CP_i,t,cg ·HI_prod,cg +CA_i,t,cg ·HI_area
we now write
AGR_i,t,cg = CA_i,t,cg (Y_i,t,cg ·HI_prod,cg + HI_area)
We will also add more explanations to the text.

9. For the Tier 1 method, IPCC divides the reference carbon stocks by climate and soil types. Did the authors also divide the grid cells by climate and soil because only climate is mentioned in the text? And, I found a diagram in Figure 5.1 in the IPCC report that divides low, medium and high input categories. Did the authors use this diagram to classify the input? It is not clear if the authors use the diagram or developed their own. If they developed their own, is it consistent with the IPCC factors?

Answer to 9.:
We do not calculate carbon stocks and stock changes based on the Tier 1 method. Instead we calculate stock change factors from our Tier 2 approach to compare only these factors to the Tier 1 factors as given within the guidelines. We will point this out more clearly in the method and discussion part. In addition, we note here:
* To our knowledge, the IPCC method splits up soil types into mineral and organic. This analysis focuses exclusively on mineral soils, which will be pointed out clearler  throughout the paper. 
* To aggregate our Tier 2 results we assigned only one climate type for each grid cell (the most dominant one). 
* The Tier 2 method does not take into account different soil classes directly, but rather takes the sand fraction as a proxy for soil properties (like water holding capacities). .
* For the Tier 1 method no detailed analysis has been conducted . All factors shown in the manuscript are default factors without any spatial disaggregation of effects. We did not classify the inputs into Tier 1 categories (as there was no need to do so for Tier 2 and no data available to do so at the global scale).

10. What is ‘resp’ is ‘area reduction resp’ on line 110? This sentence should be revised to improve readability. I also found other sentences that were difficult to read or missing words in some cases, but did not make a list during my review. Suggest a careful review before final publication.

Answer to 10.:
We will do so.

11. I found the Tier 2 method in Chapter 5 of Volume 4 of the 2019 IPCC report, and would suggest that the authors cite this chapter rather than the entire 2019 IPCC report,which has 5 volumes. This would make it easier for others interested in the study to find the method in the IPCC report.

Answer to 11.:
We will change the citation from full report to the Chapter 5 were needed.

### 2021-06-07 ResultNotebooks/investigate_SOCbugfixDiff.Rmd
update outouts needed for final response

---
title: "Update after major correction to *Management induced changes of soil organic carbon on global croplands*"
author: "Kristine Karstens"
date: "June 8, 2021"
output: pdf_document
rmarkdown::pdf_document:
  fig_caption: yes
  includes:
    in_header: myHeader.tex
---

In the following we will provide an update on some of our main results in comparison to the version of our initial draft to help the reader understand the implication of our changes. The following major corrections and improvements have been made, that let to substantial changes (next to smaller improvements):

* We discovered a bug in the very core of the soil model, leading to an overestimation of the transfer of carbon from active to the slow pool *just* for cropland. In regions with high carbon inputs (e.g. UK or Central EU) this let to unreasonable high SOC stocks.
* Additionally, we found unreasonable high forage crop production values (especially for pumpkins used as fodder) in our input data, we used from FAO. This let to an additional decrease of residue inputs to the soil for several countries (including, Australia, Belgium and Germany). 
* We moreover improved the spin-up of our model by accounting for land-use change since 1510.

The SOC debt is not decreasing anymore, but still continuing to increase, which is visible in the following figure.

```{r captionSOCdebt, echo=FALSE}
captionSOCdebt <- "NEW!!! $\\Delta SOC$ is the difference between SOC under historic land use and potential natural vegetation and is also called the SOC debt. Within the period 1975 -- 2010 the SOC debt is continuing to increase."
```

```{r image_grobs, fig.show='hold', out.width = "70%", fig.align = "default", echo=FALSE, fig.cap = captionSOCdebt,}
library(cowplot)
library(ggplot2)

ggdraw() + draw_image("../ResultNotebooks/Output/Images/SOCemis.png")

```

\pagebreak

## SOC distribution and depletion

```{r captionSOCmaps, echo=FALSE}
captionSOCmaps <- "NEW!!! (a): Distribution of total global SOC stocks for the first 30 cm on cropland: Carbon stocks are large in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics. Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas. (d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is stopping to decrease vs. actally decrease over the the period 1975--2010. "
```

```{r SOCmaps, out.width = "16cm", fig.cap = captionSOCmaps, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
```
```{r SOCmaps2, out.width = "16cm", fig.cap = "OLD!!!", echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../../oldPaperVersion/ResultNotebooks/Output/Images/4panelfigure.png")
```
\pagebreak

With the figures above we provide a world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010. Values ranging between well over $100\mathrm{t ha^{-1}}$ in northern temperate croplands to less than $5\mathrm{t ha^{-1}}$ for arid and semiarid croplands.

\pagebreak

## Agricultural management effects on SOC debt

```{r captionSOCscen, echo=FALSE}
captionSOCscen <- "NEW!!! (a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues, manure constant or neglecting adoption of no-tillage practices. ConstManagement combines all three modifications. Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure."
```

```{r SOCscen, out.width="10cm", fig.cap = captionSOCscen, echo=FALSE}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_horiz.png")
```
```{r SOCscen2, out.width="10cm", fig.cap = "OLD!!!", echo=FALSE}
# All defaults
knitr::include_graphics("../../oldPaperVersion/ResultNotebooks/Output/Images/scenario_horiz.png")
```

Most notably difference is the trend of the historical reference line as well as the split between stylized constManagement and the baseline histManagement. Before we calculated a split of $8\mathrm{GtC}$, that now decreased to around $2\mathrm{GtC}$, which is still a third of the SOC loss of the period from 1975 -- 2010.

\pagebreak

##  Modeled management effect in comparison with default IPCC Tier 1 factors

Stock change factors $F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Our updated results are much smaller and still too low compared with the default values of the IPCC in general.

```{r SCFglo, echo=FALSE,  results = 'asis'}
table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
library(xtable)
print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
             caption="NEW!!!"),
               comment=FALSE)
```

```{r SCFglo2, echo=FALSE, results = 'asis'}
table <- readRDS("../../oldPaperVersion/ResultNotebooks/Output/SFC_comparison.rds")
library(xtable)
print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
              caption="OLD!!!"), comment=FALSE, caption.placement = "top")
```

### 2021-07-24 Revisions/Point by point replies - Jonathan Sanderman.md
update tracking of Revisions

@@
 
-0.
-This MS presents the findings of a spatially explicit implementation of a new IPCC TierII approach to soil carbon stock changes.
+\textbf{Preface: This MS presents the findings of a spatially explicit implementation of a new IPCC Tier II approach to soil carbon stock changes.
 The authors use this model to calculate how much SOC has been lost due to agriculture but then go on to run the model annually for the period 1975 – 2010 to produce a dynamic picture of SOC recovery over this modern era of farming.
 The major takeaway message is that while agriculture has incurred a large SOC debt, recent agronomic improvements have led to 4 Pg C of SOC sequestration over this period of time.
-This detailed picture of SOC in croplands over the past several decades is of incredible importance to policy makers and as such I believe that paper can be an important contribution to the literature; however, I do have several major concerns that may or may not be addressable with revisions.
+This detailed picture of SOC in croplands over the past several decades is of incredible importance to policy makers and as such I believe that paper can be an important contribution to the literature; however, I do have several major concerns that may or may not be addressable with revisions.}
 
-Answer to 0.:
+Answer to preface:
 Dear Dr.
 Sanderman, thank you for the thorough and helpful review of our manuscript.
-While checking and revising our data processing in response to your and the other reviewer's feedback, we discovered a bug in the very core of the soil model, leading to an overestimation of the transfer of carbon from active to the slow pool *just* for cropland.
+While checking and revising our data processing in response to your and the other reviewer's feedback, we discovered a bug in the soil model, leading to an overestimation of the transfer of carbon from the active to the slow pool exclusively for cropland.
 Additionally, we found unreasonable high forage crop production values (specifically for pumpkins used as fodder) in our input data, which are taken from FAO statistics.
@@
 After correcting the bugs, this is no longer the case.
-Whereas this implies major revisions of our discussion and interpretation of results, we find that the essence of the paper, which is a) the availability of a soil carbon model that can account for changes in agricultural management and can be applied within integrated assessment frameworks and b) an analysis of the importance of accounting for management dynamics in SOC assessments, remains unchanged, albeit modified.
-We provide an assessment of how results changed after correcting the bugs as a supplement to this authors comment.
+Whereas this implies major revisions of our discussion and interpretation of results, we argue that the essence of the paper remains unchanged, albeit modified.
+We suggest that our key findings are a) we introduce a soil carbon model that can account for changes in agricultural management and can be applied within integrated assessment frameworks for the first time and b) we show that it is critical to account for management dynamics in SOC assessments.
+We provide an assessment of how results changed after correcting the bugs as a supplement to this author's comment.
 Here, we respond point-by-point to the reviewers' comments.
-Where appropriate, this will also address the implications of the bugfix.
+Where appropriate, this will also address the implications of the bug fix.
 We look forward to your response.
 
-1.
+\textbf{1.
 I had to read the methods section twice and spend an hour with Calvo Buendia et al., 2019 to fully understand what the authors have done.
@@
 I suggest adding an illustrative example or two graphically demonstrating how the process works.
-Perhaps starting with a simple case of one lu transition and then showing a more complex case of multiple lu transitions within a pixel.
+Perhaps starting with a simple case of one lu transition and then showing a more complex case of multiple lu transitions within a pixel.}
 
@@
 
-2.
+\textbf{2.
 I have not been convinced that this sort of “dynamic” implementation of a steady state modeling approach is appropriate.
 I understand that the method was developed by the IPCC as a way to add more nuance into the Tier I emission factor approach but I don’t think the method was intended to be applied annually.
-Why not go all the way to TierIII approach using the process-based dynamics that are embedded in this simplified model?
+Why not go all the way to Tier III approach using the process-based dynamics that are embedded in this simplified model?
 It appears you have all the data assembled to do this.
 My  main  concern with applying a steady-state model to annual changes is we know that the recent past trajectory of SOC (particularly in the slow and passive pools) will greatly influence the short-term model response to improved management – i.e.
-the model will take years to decades before SOC stocks start to rebuild if the trajectory was negative prior to the change – but this will be completely missed with the steady state application (stocks will start increasing immediately upon change).
+the model will take years to decades before SOC stocks start to rebuild if the trajectory was negative prior to the change – but this will be completely missed with the steady state application (stocks will start increasing immediately upon change).}
 
@@
 
-3.
+\textbf{3.
 What time frame are the monthly climate data averaged over to get the rate modifiers for a steady state solution?
-Given the passive pool has an intrinsic decay rate equivalent to >100 year turnover time, it seems that you need to have a 100+ year average climate to come up with the proper rate modifiers.
+Given the passive pool has an intrinsic decay rate equivalent to >100 year turnover time, it seems that you need to have a 100+ year average climate to come up with the proper rate modifiers.}
 
@@
 
-4.
+\textbf{4.
 Transfer between lu types is not clear.
 I do not understand how a “respective share of the SOC is reallocated.”  My concern is that the per area SOC stock for long-term cultivated land will be much different than the per area SOC stock for newly converted cropland, so I don’t see how you can suddenly bin these separate areas into one model component.
-Perhaps my request for a visual guide will help me (and other readers)understand that I have misunderstood this part of the methods.
+Perhaps my request for a visual guide will help me (and other readers)understand that I have misunderstood this part of the methods.}
 
@@
 See above: We will add more detail and graphical representation here.
-This will illustrate the linearity of the problem: Accounting separately for freshly converted cropland and old cropland is in our equation system essentially the same than taking the area weighted mean.
+This will illustrate the linearity of the problem; accounting separately for freshly converted cropland and old cropland is - in our equation system - essentially the same as taking the area-weighted mean.
 
-5. Why was 1901 chosen to spin up to steady state?  We know that this was a time of rapid agricultural expansion in several major regions of the world and thus a time of rapid soil carbon loss.
+\textbf{5.
+Why was 1901 chosen to spin up to steady state?
+We know that this was a time of rapid agricultural expansion in several major regions of the world and thus a time of rapid soil carbon loss.}
 
 Answer to 5.:
-We extended the spin-up phase for a much longer period, starting in 1510 (default spin-up start in LPJmL) to capture the land-use effects.
-Our initial choice of starting in 1901 was driven by the climate data availability.
+We extended the spin-up phase for a much longer period, starting in 1510 (the default spin-up start in LPJmL) to capture the land-use effects.
+Our initial choice of starting in 1901 was driven by the availability of climate data.
 To extend our spin-up into the past, we will recycle climate data from 1901 to 1930, as this is done in DGVM simulations as well (Schaphoff et al.
@@
 
-6.
+(References: 
+von Bloh, W. et al. 2018: Implementing the nitrogen cycle into the dynamic global vegetation, hydrology, and crop growth model LPJmL (version 5.0). Geoscientific Model Development 11, 2789–2812.
+Schaphoff, S. et al. 2018a: LPJmL4 – a dynamic global vegetation model with managed land – Part 1: Model description. Geoscientific Model Development 11, 1343–1375.
+Schaphoff, S. et al. 2018b: LPJmL4 – a dynamic global vegetation model with managed land – Part 2: Model evaluation. Geoscientific Model Development 11, 1377–1403.)
+
+\textbf{6.
 LULUC data – are these data all provided as area within each grid cell (or percent of a grid)?
 I think so, but please indicate.
-A supplemental table with the cross-walk between the LUH2 and the 17 crop groups used in this study should be included.
+A supplemental table with the cross-walk between the LUH2 and the 17 crop groups used in this study should be included.}
 
@@
 
-7.
+\textbf{7.
 Please include units for eq 9 – I don’t follow the AGR calculation – it sounds like you are adding biomass and area together.
-Additionally, HI is usually defined as CP divided by total above ground biomass, so CP x HI is a meaningless number.L
+Additionally, HI is usually defined as CP divided by total above ground biomass, so CP x HI is a meaningless number.}
 
@@
 Instead of
-AGR_i,t,cg = CP_i,t,cg ·HI_prod,cg +CA_i,t,cg ·HI_area
+$AGR_{i,t,cg} = CP_{i,t,cg} \cdot HI_{prod,cg} + CA_{i,t,cg} \cdot HI_{area}$
 we now write
-AGR_i,t,cg = CA_i,t,cg (Y_i,t,cg ·HI_prod,cg + HI_area)
+$AGR_{i,t,cg} = CA_{i,t,cg} \cdot (Y_{i,t,cg} \cdot HI_{prod,cg} + HI_{area})$
 We will also add more explanations to the text.
 
-7.
+\textbf{7.
 Lack of validation.
@@
 In general, there is a lack of quantitative evaluation throughout.
-There are just two qualitative quality assessments – a table comparing calculated stock change factors to Tier I estimates and a discussion on how the map looks similar to other SOC maps.
+There are just two qualitative quality assessments – a table comparing calculated stock change factors to Tier I estimates and a discussion on how the map looks similar to other SOC maps.}
 
@@
 
-8.
+\textbf{8.
 The  model  itself  was  developed  recently  as  part  of  the  2019  Refinements  to  IPCC Guidelines and those updated guidelines discuss how the model was calibrated to a set of long-term trial sites but do not report any model performance metrics.
@@
 This is certainly indicative that there should be some checks against real data (see detailed elaboration on this point further down in this review).
-It could be argued that point-based validation for a model run at0.5 degree resolution is meaningless but it would be an interesting exercise to see how the model reproduces trends with known SOC histories.
+It could be argued that point-based validation for a model run at0.5 degree resolution is meaningless but it would be an interesting exercise to see how the model reproduces trends with known SOC histories.}
 
@@
 Our updated model does not show the superising behaviour of strongly increasing SOC stocks in UK and central EU anymore due to the bugfix within the model and the corrected fodder production data.
-We will add a more detailed spatial comparison to SoilGrids 2.0 (SOC stock estimates based on point data measurements) and other model based SOC estimates (LPJmL) to account for the missing spatial dimension of model output evaluation,
+We will add a more detailed spatial comparison to SoilGrids 2.0 (SOC stock estimates based on point data measurements) and other model based SOC estimates (LPJmL) to account for the missing spatial dimension of model output evaluation.
 
-9.
+\textbf{9.
 Issues with residue C return.
@@
 Additionally, there are strong interactions between N fertilization rate and root density.
-There is a huge literature on crop breeding that support the nonstationarity of these important parameters.
+There is a huge literature on crop breeding that support the nonstationarity of these important parameters.}
 
 Answer to 9.:
-This is a very valid point that we tried to take up in the discussion part as well. We add more detail to the discussion and would like to point out the following thing:
+This is an excellent point that we took up in the discussion. In particular, we would like to highlight the following:
 The IPCC methodology, which we use here, tries to capture the effect of a shifting harvest index by making the harvest index a linear function of yield with a positive intercept.
@@
 
-10. Units – Gt and Mt are not SI units, please use Pg and Tg
+\textbf{10. Units – Gt and Mt are not SI units, please use Pg and Tg}
 
 Answer to 10.:
-As In the integrated assessment community tons is most often used for reporting emissions, tons seemed the natural choice for reporting anthropogenic emissions. 
-We now came aware that in natural science the Si unit system is the standard unit system.
 We will change all units to SI and add non-SI units in brackets were needed to bridge to IAM modelling community.
 
-11.
+\textbf{11.
 L13 (and elsewhere) – “we also find that SOC is very sensitive...” – this is in reference to an unvalidated model result.
-I’d suggest rewording these sorts of phrases to, “Our model results suggest that SOC is very sensitive...
+I’d suggest rewording these sorts of phrases to, “Our model results suggest that SOC is very sensitive...}
 
@@
 
-12.
+\textbf{12.
 L279 – “we provide the first world map” – no, you did not.
 All of the global maps that have been developed using a statistical environmental-covariate modeling approach(i.e.
-soilgrids and similar) implicitly include all historic land management.
+soilgrids and similar) implicitly include all historic land management.}
 
@@
 
-13.
+\textbf{13.
 It is great that all the data are provided but I found the Karstens 2020a repository to be confusing.
@@
 The naming convention is not clear.
-I did not want to download 9 Gb of data to figure it out.
+I did not want to download 9 Gb of data to figure it out.}
 
@@
 
-14.
+\textbf{14.
 Fig 1 – perhaps it is just the spatial scale of these small maps (and I haven’t downloaded the results to explore in more detail) but it looks like there is zero intact forest in the Congo Basin and very little intact forest in the Amazon.
 Also, I would have likedto see a map showing the trend in SOC spatially – how are the 4 Pg C that has been sequestered been spread across the globe?
-Is it all in Central EU and UK?
+Is it all in Central EU and UK?}
 
 Answer to 14.:
-Fig 1.(a) is showing cropland SOC,for every grid cell containing cropland without giving an information on the extend of cropland.
-We will mask out cells with very low cropland area shares, as they might let to the impression of greater cropland extent within large parts of intact forest area.
+Fig 1.(a) is showing cropland SOC for every grid cell that contains cropland, without giving an information on the extent of cropland.
+We will mask out cells with very low cropland area shares, as they might give to the impression of greater cropland extent within large parts of intact forested area.
 We will also provide a total SOC map in the appendix.
 
-15. Fig 2 – this is a really interesting way of summarizing the model results
+\textbf{15. Fig 2 – this is a really interesting way of summarizing the model results}
+
 Answer to 15.:
-Thanks.
+Thank you!
 
-16.
+\textbf{16.
 Fig 4 – the finding presented here is very counterintuitive to me.
 Why is the SOC debt halved when the model is initialized with natural vegetation?
-Shouldn’t the 1975 SOCdebt be much greater if the 1901 starting point was natural vegetation instead of actual land use?
-Perhaps I am just misunderstanding this sensitivity analysis.
+Shouldn’t the 1975 SOC debt be much greater if the 1901 starting point was natural vegetation instead of actual land use?
+Perhaps I am just misunderstanding this sensitivity analysis.}
 
 Answer to 16.:
-The initialization analysis was meant to help the reader understanding the potential maximal error of underestimating on-going emissions on croplands, that where converted to croplands before 1901.
-We extended the spin-up phase till 1510 now.
-We will move the initialization analysis to the appendix and improve the caption of the figure to make its interpretation clearer there..
+The initialization analysis was meant to help the reader understand the potential maximal error of underestimating on-going emissions in croplands, that where converted to croplands before 1901.
+We have now extended the spin-up phase from 1510 to now.
+We will move the initialization analysis to the appendix and improve the caption of the figure to make its interpretation more clear.
 
-17.
+\textbf{17.
 Discussion section – in general, there is very little discussion of how these results fit into the large literature on SOC.
-There are many places were a reference or two would great increase the credibility of the statements that are being made.
+There are many places were a reference or two would great increase the credibility of the statements that are being made.}
 
@@
 
-18.
+\textbf{18.
 L358-360 – the finding that northern temperate zones (particularly in EU and UK) now have SOC levels up to twice that of native state yet tropical soils have lost 40-70% of their SOC is problematic and, as the authors point out in relation to the EU example,likely points to issues with getting C input to soil correct.
@@
 While there are issues and large scale generalizations in the IPCC Tier I default factors, they do represent the consensus literature on the topic.
-The updated meta-analysis between the 2006 and 2019 IPCCC5guidelines when this emission factor for the tropics changed dramatically (see Table 4in this MS) points to this new knowledge.
+The updated meta-analysis between the 2006 and 2019 IPCCC guidelines when this emission factor for the tropics changed dramatically (see Table 4in this MS) points to this new knowledge.}
 
 Answer to 18.:
-After our bugfix SOC stocks for the EU do not gain SOC compared to natural vegetation anymore.
-In general all carbon stocks are much smaller and show much higher losses compared to the Tier 1 approach for both temperate and tropical soils.
-This might very well still point to missing accounting of carbon inputs to the soil and will be discussed in more detail in the discussion part.
+After our bugfix, SOC stocks for the EU nore longer gain SOC compared to natural vegetation.
+In general, all carbon stocks are much smaller and show much higher losses compared to the Tier 1 approach for both temperate and tropical soils.
+This may be indicative of gaps in the accounting of carbon inputs to the soil and will be discussed in more detail within the discussion.
 The comparison to point data is however challenging, also as point measurements do not well capture the landscape average and will likely show a very high variance.
@@
 A full comparison to point measurements may also exceed the scope of this first article.
-We hope with our additional comparison to SoilGrids 2.0 might help fill the evaluation gap here as SoilGrids 2.0 is based on point measurements.
+We hope that our additional comparison to SoilGrids 2.0 might help fill the evaluation gap here as SoilGrids 2.0 is based on point measurements.
 
-19. L396 – how is this validation? It is just a comparison.
+\textbf{19. L396 – how is this validation? It is just a comparison.}
 
 Answer to 19.:
-Naming and framing of the evaluation of our results will be improved.
+The naming and framing of the evaluation of our results will be improved.
 
-20. L400-406 – there is a large literature that can be drawn upon to support some of the claims made in this section.
+\textbf{20. L400-406 – there is a large literature that can be drawn upon to support some of the claims made in this section.}
 
 Answer to 20.:
-We will add more detail and references here
+We will add more detail and references here.
 
-21.
+\textbf{21.
 Section 4.4 – I do not think this is a valid comparison because SoilGrids explicitly tried to capture high carbon density soils well while your model explicitly excludes organic soils.
@@
 Additionally, ISRIC released an update to SoilGrids >6 months ago that focuses primarily on mineral soil carbon stocks.
-This update is probably a better comparison.
+This update is probably a better comparison.}
 
@@
 
-22. L425-428 – this seems out of place.
+\textbf{22. L425-428 – this seems out of place.}
 
 Answer to 22.:
-We will include sensitivity discussions in our extended discussion on evaluation and uncertainty of our results.
+We will include a discussion of the sensitivity in our extended discussion on evaluation and uncertainty of our results.
 
-23.
+\textbf{23.
 L453 – comparison to 4p1000 is not really fair because your model is really just the business-as-usual scenario with SOC gains simply because yields are improving glob-ally.
-4p1000 is about intentional management shifts to increase SOC.
+4p1000 is about intentional management shifts to increase SOC.}
 
 Answer to 23.:
-We agree with the reviewer, that the comparison is a bit misleading. We still think, it is fruitful to compare the observed rates with ambitious targets and will reframe the discussion on this point.
+We agree with the reviewer, that the comparison is a bit misleading.
+We still think that it is fruitful to compare the observed rates with ambitious targets and will reframe the discussion on this point.

### 2021-07-24 Revisions/Point by point replies - second reviewer.md
update tracking of Revisions

@@
 Dear reviewer, thank you for the thorough and helpful review of our manuscript.
-While checking and revising our processing in response to your and the other reviewer's feedback, we discovered a bug in the very core of the soil model, leading to an overestimation of the transfer of carbon from active to the slow pool *just* for cropland.
+While checking and revising our processing in response to your and the other reviewer's feedback, we discovered a bug in the soil model, leading to an overestimation of the transfer of carbon from active to the slow pool exclusively for cropland.
 Additionally, we found unreasonable high forage crop production values (specifically for pumpkins used as fodder) in our input data, which were taken from FAO statistics.
-This made the overall intensification trend in agriculture lead to increasing carbon stocks in cropland. After correcting the bugs, this is no longer the case. 
-Whereas this implies major revisions of our discussion and interpretation of results, we find that the essence of the paper, which is a) the availability of a soil carbon model that can account for changes in agricultural management and can be applied within integrated assessment frameworks and b) an analysis of the importance of accounting for management dynamics in SOC assessments, remains unchanged, albeit modified.
-We provide an assessment of how results changed after correcting the bugs as a supplement to this authors comment.
+This made the overall intensification trend in agriculture lead to increasing carbon stocks in cropland.
+After correcting the bugs, this is no longer the case.
+Whereas this implies major revisions of our discussion and interpretation of results, we argue that the essence of the paper remains unchanged, albeit modified.
+We suggest that our key findings are a) we introduce a soil carbon model that can account for changes in agricultural management and can be applied within integrated assessment frameworks for the first time and b) we show that it is critical to account for management dynamics in SOC assessments.
+We provide an assessment of how results changed after correcting the bugs as a supplement to this author's comment.
 Here, we respond point-by-point to the reviewers' comments.
-Where appropriate, this will also address the implications of the bugfix.
+Where appropriate, this will also address the implications of the bug fix.
 We look forward to your response.
@@
 This is an important point.
-While our analysis focuses on croplands, estimating the natural soil carbon stocks is a necessity to account for the C entering the cropland budget via land conversion.
+While our analysis focuses on croplands, estimating the natural soil carbon stocks is necessary to account for the C entering the cropland budget via land conversion.
 Apart from this process, we only use natural vegetation SOC to make our results comparable to other global estimates for validation purposes.
@@
 Answer to 2.:
-The quantitative assessment of the uncertainty of our projections unfortunately exceeds the scope of this article and would likely require a study on itself.
+The quantitative assessment of the uncertainty of our projections unfortunately exceeds the scope of this article and would likely require a study in itself.
 The model includes a high number of parameters, and for most of these the uncertainty distributions have not been quantified so far.
@@
 
-To still address the uncertainty within this manuscript, we will therefore discuss the uncertainty qualitatively.
-Most of the uncertainty in our view is included in the management data, especially in the manure numbers, as they are calculated from feed data.
-Additionally residue and litter estimations are based on the set of global parameters, that contains large uncertainty and at the same time have a great impact on the overall size of carbon inputs to the soil.
+To address the uncertainty within this manuscript, we will therefore discuss the uncertainty qualitatively and add a short statement here: Most of the uncertainty in our view is included in the management data itself, and especially in the residue and manure numbers, as they are only indirectly calculated from production, feed and area data.
+The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
+As this is one of the first attempts to compile a comprehensive, global data set on these management data, it is complex to evaluate it against literature.
+Furthermore, it is still to be evaluated, if our data set covers most of the carbon inputs to the soil, as it not includes cover crops, green manure and weed biomass.
+Additionally, litter estimations are based on the set of global parameters, that contain large uncertainty and at the same time have a great impact on the overall size of SOC stocks.
 Both effects will be discussed in more detail.
@@
 It is correct that the SOC stocks are highly dependent on the legacy of management.
-Figure 4 made however clear, that the SOC gap - so the difference between a baseline scenario and a counterfactual scenario with constant management - was not strongly dependent on the initialization.
-We thereby showed that the initialization does not affect our central finding.
+However, Figure 4 shows that the SOC gap - the difference between a baseline scenario and a counterfactual scenario with constant management - was not strongly dependent on the initialization.
+We therefore conclusde that the initialization does not affect our central finding.
 
@@
 2018a/b, von Bloh et al.
-2018 ).
-We will analysis and discuss our results with respect to the initialization in more detail.
+2018).
+We will analyse and discuss our results with respect to the initialization in more detail.
+
+(von Bloh, W. et al. 2018: Implementing the nitrogen cycle into the dynamic global vegetation, hydrology, and crop growth model LPJmL (version 5.0). Geoscientific Model Development 11, 2789–2812.
+Schaphoff, S. et al. 2018a: LPJmL4 – a dynamic global vegetation model with managed land – Part 1: Model description. Geoscientific Model Development 11, 1343–1375.
+Schaphoff, S. et al. 2018b: LPJmL4 – a dynamic global vegetation model with managed land – Part 2: Model evaluation. Geoscientific Model Development 11, 1377–1403.)
 
@@
 We will place the comparison within the result section.
-For a Tier 1 approach the change in soil organic carbon has to be calculated based on a reference stock.
-The default stocks given by the IPCC have very low spatial resolution (42 coarse climate zone and soil type specific values), which leads to additional uncertainty when comparing directly changes in soil organic carbon.
-We will add more detail to the discussion about the strong deviation of 2006 and 2019 default stock changes factors.
+For a Tier 1 approach the change in soil organic carbon must be calculated based on a reference stock.
+The default stocks given by the IPCC have a very low spatial resolution (42 coarse climate zones and soil type specific values), which lead to additional uncertainty when directly comparing changes in soil organic carbon.
+We will add more detail to the discussion of the strong deviation between the 2006 and 2019 default stock change factors.
 
@@
 Answer to 5.:
-Due to the overall changes to the results, we will rewrite major part of the result and discussion. We will take more thoroughly into account effects of uncertainty.
+Due to the overall changes in the results, we will rewrite major part of the result and discussion. We will take more thoroughly into account the effects of uncertainty.
 
@@
 In the draft of the International Nitrogen Assessment (will be provided confidentially to the reviewer), the available literature estimates for these flows (in regard to nitrogen) were reviewed.
-So far no reliable estimates exist, but the existing estimates indicate that these inputs must be by an order of magnitude lower than those by crop residues and manure.
+So far no reliable estimates exist, but the existing estimates indicate that these inputs must be by an order of magnitude lower than those of crop residues and manure.
 
@@
 Instead of
-AGR_i,t,cg = CP_i,t,cg ·HI_prod,cg +CA_i,t,cg ·HI_area
+$AGR_{i,t,cg} = CP_{i,t,cg} \cdot HI_{prod,cg} + CA_{i,t,cg} \cdot HI_{area}$
 we now write
-AGR_i,t,cg = CA_i,t,cg (Y_i,t,cg ·HI_prod,cg + HI_area)
+$AGR_{i,t,cg} = CA_{i,t,cg} \cdot (Y_{i,t,cg} \cdot HI_{prod,cg} + HI_{area})$
 We will also add more explanations to the text.
@@
 In addition, we note here:
-* To our knowledge, the IPCC method splits up soil types into mineral and organic. This analysis focuses exclusively on mineral soils, which will be pointed out clearler  throughout the paper. 
+* To our knowledge, the IPCC method splits up soil types into mineral and organic. This analysis focuses exclusively on mineral soils, which will be pointed out more clearly throughout the paper. 
 * To aggregate our Tier 2 results we assigned only one climate type for each grid cell (the most dominant one).
-* The Tier 2 method does not take into account different soil classes directly, but rather takes the sand fraction as a proxy for soil properties (like water holding capacities). .
-* For the Tier 1 method no detailed analysis has been conducted .
+* The Tier 2 method does not take into account different soil classes directly, but rather takes the sand fraction as a proxy for soil properties (like water holding capacities).
+* For the Tier 1 method no detailed analysis has been conducted.
 All factors shown in the manuscript are default factors without any spatial disaggregation of effects.
@@
 Answer to 11.:
-We will change the citation from full report to the Chapter 5 were needed.
+We will change the citation from full report to the Chapter 5 where needed.

### 2021-07-24 ResultNotebooks/create_LUtransition.Rmd
update results scripts

---
  title: "LU transitions overview figs"
output:
  md_document:
  variant: markdown_github
---

```{r arrange plot, out.height=1000 , out.width=2000}
library(cowplot)
library(ggplot2)

start     <- "../FlowFigure/LUC_start.pdf"
trans     <- "../FlowFigure/LUC_transition.pdf"
hetero    <- "../FlowFigure/LUC_hetero.pdf"
homo      <- "../FlowFigure/LUC_homo.pdf"
text      <- "../FlowFigure/LUC_text.pdf"
outputdir <- "Output/Images"

p1 <- ggdraw() + draw_image(start)
p2 <- ggdraw() + draw_image(trans)
p3 <- ggdraw() + draw_image(hetero)
p4 <- ggdraw() + draw_image(homo)
pt <- ggdraw() + draw_image(text)

grid_1 <- plot_grid(p1, p2, p3, ncol=3)
grid_2 <- plot_grid (pt, p4, ncol=2, rel_widths = c(2, 1))
grid   <- plot_grid(grid_1, grid_2, nrow=2)
grid
ggsave(paste0(outputdir,"/LUtransition.png"), plot=grid, width = 20, height = 10, dpi = 400, units = "cm", bg="white")

```

### 2021-07-24 ResultNotebooks/investigate_SOCbugfixDiff.Rmd
update results scripts

@@
 UK or Central EU) this let to unreasonable high SOC stocks.
-* Additionally, we found unreasonable high forage crop production values (especially for pumpkins used as fodder) in our input data, we used from FAO.
-This let to an additional decrease of residue inputs to the soil for several countries (including, Australia, Belgium and Germany).
+* Additionally, we found unreasonable high forage crop production values for pumpkins used as fodder in our input data, we used from FAO.
+This led to an additional decrease of residue inputs to the soil for several countries (including, Australia, Belgium and Germany).
+The values were excluded and replaced by zeros.
 * We moreover improved the spin-up of our model by accounting for land-use change since 1510.
 
-The SOC debt is not decreasing anymore, but still continuing to increase, which is visible in the following figure.
+## Summary of main changes
 
-```{r captionSOCdebt, echo=FALSE}
-captionSOCdebt <- "NEW!!!
-$\\Delta SOC$ is the difference between SOC under historic land use and potential natural vegetation and is also called the SOC debt.
-Within the period 1975 -- 2010 the SOC debt is continuing to increase."
-```
+The SOC debt is not decreasing anymore, but still continuously increases.
+Unreasonable high numbers for SOC stocks and stock change factors compared to natural vegetation stocks have vanished.
+The impact of management is lower but still considerably.
+Stock change factors for cropland SOC are now globally lower than default factor from the IPCC for all four climate zones.
 
-```{r image_grobs, fig.show='hold', out.width = "70%", fig.align = "default", echo=FALSE, fig.cap = captionSOCdebt,}
-library(cowplot)
-library(ggplot2)
+In the following, we provide the comparison of updated (first) and original (second) figures for three of our main results:
 
-ggdraw() + draw_image("../ResultNotebooks/Output/Images/SOCemis.png")
-
-```
+* SOC distribution and depletion (Fig. 1 in the manuscript)
+* Agricultural management effects on SOC debt (Fig. 3 in the manuscript)
+* Modeled management effect in comparison with default IPCC Tier 1 factors (Tab. 4 in the manuscript)
 
@@
 ```{r captionSOCmaps, echo=FALSE}
-captionSOCmaps <- "NEW!!!
+captionSOCmaps <- "Updated figure!
 (a): Distribution of total global SOC stocks for the first 30 cm on cropland: Carbon stocks are large in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
+(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hot spots of SOC dynamics.
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is stopping to decrease vs.
-actally decrease over the the period 1975--2010.
-"
+(d): SOC debt is the difference between SOC under historic land use and potential natural vegetation.
+Within the period 1975 -- 2010 the SOC debt is continuing to increase."
 ```
@@
 ```
-```{r SOCmaps2, out.width = "16cm", fig.cap = "OLD!!!", echo=FALSE, fig.pos="h"}
+```{r SOCmaps2, out.width = "16cm", fig.cap = "Original figure of the manuscript", echo=FALSE, fig.pos="h"}
 # All defaults
@@
 Values ranging between well over $100\mathrm{t ha^{-1}}$ in northern temperate croplands to less than $5\mathrm{t ha^{-1}}$ for arid and semiarid croplands.
+The correction of to high SOC stock values is visible for all four figure parts, as it corrects the unreasonable high values in UK and Central Europe (see old figure).
 
@@
 ```{r captionSOCscen, echo=FALSE}
-captionSOCscen <- "NEW!!!
-(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues, manure constant or neglecting adoption of no-tillage practices.
+captionSOCscen <- "Updated figure!
+(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios deviate from historic agricultural management by holding effects of carbon inflows from residues or manure constant, or neglecting adoption of no-tillage practices.
 ConstManagement combines all three modifications.
@@
 ```
-```{r SOCscen2, out.width="10cm", fig.cap = "OLD!!!", echo=FALSE}
+```{r SOCscen2, out.width="10cm", fig.cap = "Original figure of the manuscript", echo=FALSE}
 # All defaults
@@
 
-Stock change factors $F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Our updated results are much smaller and still too low compared with the default values of the IPCC in general.
+Stock change factors $F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Our updated results are much smaller and still considerably lower compared to the default values of the IPCC for all four climate zones.
 
@@
 print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
-             caption="NEW!!!"),
+             caption="Updated table."),
                comment=FALSE)
@@
 print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
-              caption="OLD!!!"), comment=FALSE, caption.placement = "top")
+              caption="Original table of the manuscript."), comment=FALSE)
 ```

### 2021-07-24 ResultNotebooks/investigate_SOCflows.Rmd
update results scripts

--- 2021-04-01 ResultNotebooks/results_SOCflows.Rmd
@@
 
-year  <- "y2010"
-years <- c(paste0("y",as.numeric(substring("y2010",2))-1), year)
+year  <- paste0("y",seq(1975,2010,5))
+years <- sort(c(paste0("y",as.numeric(substring(year,2))-1), year))
 
@@
 
-grepSpaceTimeC    <- function(x){return(collapseNames(dimSums(x[,year,"c"],dim=1)))}
-grepSpaceTime     <- function(x){return(collapseNames(dimSums(x[,year,],dim=1)))}
+grepSpaceTimeC    <- function(x){return(collapseNames(dimSums(x["DEU",year,"c"],dim=1)))}
+grepSpaceTime     <- function(x){return(collapseNames(dimSums(x["DEU",year,],dim=1)))}
 
@@
 
-ResidueInput      <- grepSpaceTimeC(Landuse[,2,"crop"]* CarbonResidues)
+ResidueInput      <- grepSpaceTimeC(Landuse[,year,"crop"]* CarbonResidues)
 
@@
 #ManureRecycledTot <- grepSpaceTimeC(dimSums(ManureRecycled, dim=3.1))
-ManureInput       <- dimSums(grepSpaceTimeC(Landuse[,2,"crop"]* CarbonManure), dim=3.1)
+ManureInput       <- dimSums(grepSpaceTimeC(Landuse[,year,"crop"]* CarbonManure), dim=3.1)
 ManureLossTot     <- ManureExcretedTot - ManureInput
@@
 SOCPrev           <- dimSums(grepSpaceTime(setYears(
-                       SoilCarbonOut[,1,"actualstate"][,,"crop"]*Landuse[,1,"crop"], year)), dim=3)
+                       SoilCarbonOut[,setdiff(years,year),"actualstate"][,,"crop"]*
+                         Landuse[,setdiff(years,year),"crop"], year)), dim=3)
 SOCLoss           <- SOCPrev - SOCtot[,,"crop"] + SOCtransfer + SOCIn
@@
 
-write.csv(as.data.frame(TotalSOCBudget), paste0(outputdir,"TotalSOCBudget.csv"))
-
+write.csv(as.data.frame(TotalSOCBudget), paste0(outputdir,"DEUSOCBudget.csv"))
+write.magpie(TotalSOCBudget, paste0(outputdir,"DEUSOCBudget.cs3"))
 ```
@@
 ```{r}
+# investigate high yields
 
-DecayCut          <- Decay
-DecayCut[Decay>1] <- 1
+temp_crops <- c("maiz","tece","potato","sugr_beet","rapeseed","others","foddr")
 
-### Carbon In
-f1        <- param[,,"f1"]
-f2_crop   <- param[,,"f2_ft"]
-f2_natveg <- param[,,"f2_nt"]
-f3        <- param[,,"f3"]
+library(mrcommons)
+setConfig(cachefolder  = "/home/kristine/mnt/rd3mod/inputdata/cache/rev16", forcecache = TRUE)
+setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
+setConfig(mappingfolder = "/home/kristine/mnt/rd3mod/inputdata/mappings/")
 
-# f2 has to be updated with new test runs
+croparea <- calcOutput("Croparea", aggregate = FALSE)
 
-MineralizeFracInput <- setCells(mbind(add_dimension(collapseNames(CarbonInput[,,"totalC",invert=TRUE][1,,"res"]), dim=3.1, add = "lu", nm = "crop"),
-                                      add_dimension(collapseNames(CarbonInput[,,"totalC",invert=TRUE][1,,"res"]), dim=3.1, add = "lu", nm = "natveg")), "GLO")
-MineralizeFracInput[,,"metabDOC"]        <- f1
-MineralizeFracInput[,,"crop.strucDOC"]   <- f2_crop
-MineralizeFracInput[,,"natveg.strucDOC"] <- f2_natveg
-MineralizeFracInput[,,"ligninC"]         <- f3
-
-input2lu <- as.data.frame(list(c("res", findset("kli"), "litfall"),
-                            c(rep("crop",6), "natveg")), col.names=c("input","lu"))
-
-CarbonInEff       <- collapseNames(dimSums(toolAggregate(CarbonInput[,"y2010",][,,"totalC",invert=TRUE], rel=input2lu, dim=3.1) *
-                                            MineralizeFracInput[,"y2010",] * Landuse[,"y2010",], dim=c(1,3.2)))
-
-CarbonInSource   <- collapseNames(dimSums(CarbonInput[,"y2010",][,,"totalC",invert=TRUE]  * 
-                                            toolAggregate(MineralizeFracInput[,"y2010",] * Landuse[,"y2010",], rel=input2lu, from="lu", to="input", dim=3.1), dim=c(1,3.2)))
-
-CarbonInRaw       <- collapseNames(dimSums(SoilCarbonOut[,"y2010","steadystate"]*DecayCut[,"y2010",]*Landuse[,"y2010",], dim=c(1,3.3)))
-
-### Carbon Out
-
-CarbonOutRaw      <- collapseNames(dimSums(SoilCarbonOut[,"y2010","interstate"]*DecayCut[,"y2010",]*Landuse[,"y2010",], dim=c(1,3.3)))
-CarbonOutEff      <- CarbonOutRaw + CarbonInEff - CarbonInRaw
-
-### SOC State
-
-SOCstate <- dimSums(SoilCarbonOut[,c("y2009","y2010"),c("actualstate","interstate")]*Landuse[,c("y2009","y2010"),], dim=c(1,3.3))
+print((grepSpaceTimeC(FAOmassbalance[,,temp_crops][,,"production"])))
+print(grepSpaceTime(croparea[,,temp_crops]))
+print((grepSpaceTimeC(FAOmassbalance[,,temp_crops][,,"production"]))/grepSpaceTime(croparea[,,temp_crops]))
 

### 2021-07-24 ResultNotebooks/investigate_SOCvalidate.Rmd
update results scripts

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
  bookdown::pdf_document2: default
   keep_md: true
---

```{r load}
library(magclass)
library(madrat)

outputdir <- "Output/Images/"

SoilCarbon <- read.magpie("Data/rev26histManagement/SoilCarbon.rds")
Landuse    <- read.magpie("Data/rev26histManagement/Landuse.rds")

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

cropshare       <- toolConditionalReplace(Landuse[,,"crop"]/dimSums(Landuse, dim=3), "is.na()", 0)
SOCStock_natveg <- dimSums(SoilCarbon[,,c("naturalstate.natveg")], dim=3)
SOCStock_lu     <- dimSums(SoilCarbon[,,c("actualstate")]*Landuse, dim=3)/dimSums(Landuse, dim=3)

SOCStock_natveg[cropshare<=0.05] <- NA
SOCStock_lu[    cropshare<=0.05] <- NA

```

```{r load soc validation stocks}

  library(mrvalidation)

  setConfig(cachefolder  = "/home/kristine/mnt/rd3mod/inputdata/cache/rev26/", forcecache = TRUE)
  setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
  setConfig(mappingfolder = "/home/kristine/mnt/rd3mod/inputdata/mappings/")

  SoilGrids2   <- readRDS("Data/additional/SoilGrids2.rds")
  LPJmL4_CRU   <- toolCoord2Isocell(readRDS("Data/additional/LPJmL4_CRU.rds"))
  LPJmL4_soilc <- readSource("LPJmL_new", subtype="LPJmL4_for_MAgPIE_84a69edd:CRU4:historical:soilc_layer", convert="onlycorrect")*0.01
  LPJmL4_soilc <- toolCoord2Isocell(LPJmL4_soilc)
  LPJmL4_soilc <- setNames(LPJmL4_soilc[,,"layer1"] + 1/3*LPJmL4_soilc[,,"layer2"], "layer_0-30")
  LPJmL4_soilc <- LPJmL4_soilc[,getYears(Landuse),]
  LPJmL4_CRU   <- LPJmL4_CRU[,getYears(Landuse),]

  cells      <- getCells(LPJmL4_soilc)

  LPJmL4_soilc[   cropshare     <=0.05] <- NA
  LPJmL4_CRU[     cropshare     <=0.05] <- NA
  SoilGrids[cropshare[,"y2010",]<=0.05] <- NA

```

```{r correation plot soilc, fig.height=7.5, fig.width=6}

regToCell <- toolGetMapping("clustermap_rev4.51+mrmagpie10_c200_h12.rds",type="regional")
regions   <- levels(regToCell$region)
climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
climate$N <- as.integer(substring(climate$cell,5))
climate   <- climate[order(climate$N),]

# climate
library(RColorBrewer)
library(scales)
blucols          <- brewer.pal(9, "Blues")[2:5]
viocols          <- brewer.pal(9, "Purples")[6:9]
redcols          <- brewer.pal(9, "Reds")[6:8]
orgcols          <- brewer.pal(9, "Oranges")[3]
cols <- c(blucols[4], viocols[1], redcols[1], orgcols[1])
paired <- brewer.pal(12, "Paired")

years <- seq(1975, 2010, 5)

climates <- unique(climate$IPCC_reduced)

layout(matrix(c(1,2,3,4,5,5), ncol=2, byrow=TRUE), 
       heights = c(0.38, 0.42, 0.2), widths =  c(0.52,0.48))

xlabel <- c("", "", "LPJmL4 (t/ha)", "LPJmL4 (t/ha)")
ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")

layout.show(n=5)

mar <- list(c(2,4,1,1),
            c(2,2,1,1),
            c(4,4,1,1),
            c(4,2,1,1))

for(i in 1:length(climates)){

  icells <- as.array(climate$cell[climate$IPCC_reduced==climates[i]])
  par(mar = mar[[i]])
  lim <- quantile(c(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,]), na.rm = TRUE, probs = 0.999)
  #lim <- max(LPJmL4[icells,years,],SOCStock_natveg[icells,years,], na.rm = TRUE)

  plot(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,], asp = 1, 
       xlab = xlabel[i], ylab = ylabel[i],
       col = "white", pch = ".",
       xlim = c(0, lim), ylim = c(0, lim))

  for(r in 1:length(regions)){

    rcells <- intersect(icells,as.array(regToCell$cell[regToCell$region==rev(regions)[r]]))

    points(LPJmL4_soilc[rcells,years,],SOCStock_natveg[rcells,years,],
       col = alpha(paired[r], 0.2), pch = ".")

  }

  lines(c(0,lim), c(0,lim), col="grey")
  lines(c(0,lim/2), c(0,lim), col="grey", lty = 2)
  text(0, lim, paste0("cor=",round(cor(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,], 
                                 use = "complete.obs"),2)), adj = c(0,1))
  text(lim/3, lim, paste0("r²=",round(cor(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,], 
                                       use = "complete.obs")^2, 2)), adj = c(0,1))
  text(lim*2/3, lim, paste0("EF=",round(qualV::EF(as.array(LPJmL4_soilc[icells,years,][!is.na(LPJmL4_soilc[icells,years,])]),
                                             as.array(SOCStock_natveg[icells,years,][!is.na(SOCStock_natveg[icells,years,])])),
                                             2)), adj = c(0,1))
  text(lim/2, 0, paste0("Climate zone: ", gsub("_", " ", climates[i])))
}

plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
legend("top", horiz = TRUE, legend = rev(regions), col = paired, pch=16, bty ="n")
dev.copy(png, paste0(outputdir, "scatterCompareLPJmL_pnv.png"), width=1200, height=1500, res=250)
dev.off()

library(ggplot2)
theme_set(theme_bw())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce) 
library(scales)

legend_size <- 10
year        <- 2010

map.world     <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
cropland_diff <- collapseNames(SOCStock_natveg[,year,] - LPJmL4_soilc[,year,])
diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]

SOCdiff <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(diff2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
                      name = expression(paste('natural \u0394SOC in tons of carbon per hectare (this study - LPJmL)')), 
                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
                      breaks=c(-50,-20,-10,10,20,50,100),
                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
                      guide = guide_legend(
                        title.theme=element_text(size=legend_size),
                        direction = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = F,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
    theme(legend.position="bottom",
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          legend.text=element_text(size=legend_size),
          rect=element_blank(),
          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

SOCdiff
```
```{r correation plot CRU, fig.height=7.5, fig.width=6}

regToCell <- toolGetMapping("clustermap_rev4.51+mrmagpie10_c200_h12.rds",type="regional")
regions   <- levels(regToCell$region)
climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
climate$N <- as.integer(substring(climate$cell,5))
climate   <- climate[order(climate$N),]

# climate
library(RColorBrewer)
library(scales)
blucols          <- brewer.pal(9, "Blues")[2:5]
viocols          <- brewer.pal(9, "Purples")[6:9]
redcols          <- brewer.pal(9, "Reds")[6:8]
orgcols          <- brewer.pal(9, "Oranges")[3]
cols <- c(blucols[4], viocols[1], redcols[1], orgcols[1])
paired <- brewer.pal(12, "Paired")

years <- seq(1975, 2010, 5)

climates <- unique(climate$IPCC_reduced)

layout(matrix(c(1,2,3,4,5,5), ncol=2, byrow=TRUE), 
       heights = c(0.38, 0.42, 0.2), widths =  c(0.52,0.48))

xlabel <- c("", "", "LPJmL4 (t/ha)", "LPJmL4 (t/ha)")
ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")

layout.show(n=5)

mar <- list(c(2,4,1,1),
            c(2,2,1,1),
            c(4,4,1,1),
            c(4,2,1,1))

for(i in 1:length(climates)){

  icells <- as.array(climate$cell[climate$IPCC_reduced==climates[i]])
  par(mar = mar[[i]])
  lim <- quantile(c(LPJmL4_CRU[icells,years,],SOCStock_natveg[icells,years,]), na.rm = TRUE, probs = 0.999)
  #lim <- max(LPJmL4[icells,years,],SOCStock_natveg[icells,years,], na.rm = TRUE)

  plot(LPJmL4_CRU[icells,years,],SOCStock_natveg[icells,years,], asp = 1, 
       xlab = xlabel[i], ylab = ylabel[i],
       col = "white", pch = ".",
       xlim = c(0, lim), ylim = c(0, lim))

  for(r in 1:length(regions)){

    rcells <- intersect(icells,as.array(regToCell$cell[regToCell$region==rev(regions)[r]]))

    points(LPJmL4_CRU[rcells,years,],SOCStock_natveg[rcells,years,],
       col = alpha(paired[r], 0.2), pch = ".")

  }

  lines(c(0,lim), c(0,lim), col="grey")
  lines(c(0,lim/2), c(0,lim), col="grey", lty = 2)
  text(0, lim, paste0("cor=",round(cor(LPJmL4_CRU[icells,years,],SOCStock_natveg[icells,years,], 
                                 use = "complete.obs"),2)), adj = c(0,1))
  text(lim/2, 0, paste0("Climate zone: ", gsub("_", " ", climates[i])))
}

plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
legend("top", horiz = TRUE, legend = rev(regions), col = paired, pch=16, bty ="n")
dev.copy(png, paste0(outputdir, "scatterCompareLPJmLCRU_pnv.png"), width=1200, height=1500, res=250)
dev.off()

legend_size <- 10
year        <- 2010

map.world     <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
cropland_diff <- collapseNames(SOCStock_natveg[,year,] - LPJmL4_CRU[,year,])
diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]

SOCdiff <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(diff2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
                      name = expression(paste('natural \u0394SOC in tons of carbon per hectare (this study - LPJmL)')), 
                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
                      breaks=c(-50,-20,-10,10,20,50,100),
                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
                      guide = guide_legend(
                        title.theme=element_text(size=legend_size),
                        direction = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = F,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
    theme(legend.position="bottom",
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          legend.text=element_text(size=legend_size),
          rect=element_blank(),
          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

SOCdiff

```
```{r corr measures}

# corr <- new.magpie(getCells(SOCStock_natveg), 1, names = c("r", "r²", "EF", "GRI", "RMSE", "meanbias","sdratio"), fill = NA)
# 
# for (j in getCells(SOCStock_natveg))
#   {
#   
#   corr[j,,"r"]    <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])
#   corr[j,,"r²"]   <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])^2
#   corr[j,,"EF"]   <- EF(LPJmL4_CRU[j,years,], SOCStock_natveg[j,years,])
#   corr[j,,"GRI"]  <- GRI(LPJmL4_CRU[j,years,] ,SOCStock_natveg[j,years,])
#   corr[j,,"RMSE"] <- RMSE(LPJmL4_CRU[j,years,], SOCStock_natveg[j,years,])
#   print(j)
# }
# 
#   corr[,,"meanbias"] <- apply(SOCStock_natveg[,years,], 1, mean) - apply(LPJmL4_CRU[,years,], 1, mean)
#   corr[,,"sdratio"]  <-  apply(SOCStock_natveg[,years,], 1, sd)  /apply(LPJmL4_CRU[,years,], 1, sd)
# 
# write.magpie(corr, "corr.rds")

luplot::plotmap(corr[,,"meanbias"], legend_range = c(-30,30))
luplot::plotmap(corr[,,"sdratio"], legend_range = c(0,2))
luplot::plotmap(corr[,,"r"], legend_range = c(-1,1))
```

```{r correation plot soilgrids, fig.height=7.5, fig.width=6}

regToCell <- toolGetMapping("clustermap_rev4.51+mrmagpie10_c200_h12.rds",type="regional")
regions   <- levels(regToCell$region)
climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
climate$N <- as.integer(substring(climate$cell,5))
climate   <- climate[order(climate$N),]

# climate
library(RColorBrewer)
library(scales)
blucols          <- brewer.pal(9, "Blues")[2:5]
viocols          <- brewer.pal(9, "Purples")[6:9]
redcols          <- brewer.pal(9, "Reds")[6:8]
orgcols          <- brewer.pal(9, "Oranges")[3]
cols <- c(blucols[4], viocols[1], redcols[1], orgcols[1])
paired <- brewer.pal(12, "Paired")

years <- seq(1975, 2010, 5)

climates <- unique(climate$IPCC_reduced)

layout(matrix(c(1,2,3,4,5,5), ncol=2, byrow=TRUE), 
       heights = c(0.38, 0.42, 0.2), widths =  c(0.52,0.48))

xlabel <- c("", "", "SoilGrids (t/ha) (timeless)", "SoilGrids (t/ha) (timeless)")
ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")

layout.show(n=5)

mar <- list(c(2,4,1,1),
            c(2,2,1,1),
            c(4,4,1,1),
            c(4,2,1,1))

for(i in 1:length(climates)){

  icells <- as.array(climate$cell[climate$IPCC_reduced==climates[i]])
  par(mar = mar[[i]])
  lim <- quantile(c(SoilGrids[icells,,],SOCStock_natveg[icells,years,]), na.rm = TRUE, probs = 0.999)
  #lim <- max(SoilGrids[icells,,],SOCStock_natveg[icells,years,], na.rm = TRUE)

  plot(rep(SoilGrids[icells,,],8),SOCStock_lu[icells,years,], asp = 1, 
       xlab = xlabel[i], ylab = ylabel[i],
       col = "white", pch = ".",
       xlim = c(0, lim), ylim = c(0, lim))

  for(r in 1:length(regions)){

    rcells <- intersect(icells,as.array(regToCell$cell[regToCell$region==rev(regions)[r]]))

    points(rep(SoilGrids[rcells,,],8),SOCStock_lu[rcells,years,],
       col = alpha(paired[r], 0.2), pch = ".")

  }

  lines(c(0,lim), c(0,lim), col="grey")
  text(0, lim, paste0("cor=",round(cor(rep(SoilGrids[icells,,],8),SOCStock_lu[icells,years,], 
                                 use = "complete.obs"),2)), adj = c(0,1))
  text(lim/2, 0, paste0("Climate zone: ", gsub("_", " ", climates[i])))
}

plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
legend("top", horiz = TRUE, legend = rev(regions), col = paired, pch=16, bty ="n")

dev.copy(png, paste0(outputdir, "scatterCompareSoilGrids_crop.png"), width=1200, height=1500, res=250)
dev.off()

```

```{r map2010}

legend_size <- 10
year        <- 2010

map.world     <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

#magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
load("/home/kristine/WORK/R libraries/mrcommons/R/sysdata.rda")
cropland_diff <- collapseNames(SOCStock_natveg[,year,] - setYears(SoilGrids2[,1,], NULL))
diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]

SOCdiff2010 <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(diff2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare  (this study - SoilGrid2)')), 
                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
                      breaks=c(-50,-20,-10,10,20,50,100),
                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
                      guide = guide_legend(
                        title.theme=element_text(size=legend_size),
                        direction = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = F,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
    theme(legend.position="bottom",
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          legend.text=element_text(size=legend_size),
          rect=element_blank(),
          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

SOCdiff2010

```

```{r mapmean}
legend_size <- 10
year        <- 2010

map.world     <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

#magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
load("/home/kristine/WORK/R libraries/mrcommons/R/sysdata.rda")
cropland_diff <- collapseNames(magpply(SOCStock_natveg[,years,], mean, 1) - setYears(SoilGrids2[,1,], NULL))
diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]

SOCdiffmean <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(diff2plot, Year=="0", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare  (this study - SoilGrid2)')), 
                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
                      breaks=c(-50,-20,-10,10,20,50,100),
                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
                      guide = guide_legend(
                        title.theme=element_text(size=legend_size),
                        direction = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = F,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
    theme(legend.position="bottom",
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          legend.text=element_text(size=legend_size),
          rect=element_blank(),
          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

SOCdiffmean
SOCdiff2010
SOCdiff_LPJvsSG2
```
```{r mapLPJsoilGrids}
legend_size <- 10
year        <- 2010

map.world     <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

#magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
load("/home/kristine/WORK/R libraries/mrcommons/R/sysdata.rda")
cropland_diff <- collapseNames(magpply(LPJmL4_CRU[,years,], mean, 1) - setYears(SoilGrids2[,1,], NULL))
diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]

SOCdiff_LPJvsSG2 <- ggplot(data = world) + geom_sf() +
    geom_tile(data = subset(diff2plot, Year=="0", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare  (this study - SoilGrid2)')), 
                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
                      breaks=c(-50,-20,-10,10,20,50,100),
                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
                      guide = guide_legend(
                        title.theme=element_text(size=legend_size),
                        direction = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust = 0.5,
                        label.hjust = 1,
                        nrow = 1,
                        byrow = F,
                        label.position = "bottom",
                        keyheight = 0.5,
                        keywidth  = 2
                      )) +
    theme(legend.position="bottom",
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          legend.text=element_text(size=legend_size),
          rect=element_blank(),
          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

SOCdiff_LPJvsSG2
```

### 2021-07-24 ResultNotebooks/results_NC.Rmd
update results scripts

---
  title: "Global N release binding"
output:
  md_document:
  variant: markdown_github
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev       <- "rev19"
outputdir <- "Output/"

#filesNeeded  <- c("SoilCarbon.rds","Landuse.rds","LanduseChange.rds")
#untar(paste0(getConfig("outputfolder"),"/",rev, "histManagement_ResAllom_2021_01_12_h12_carbonbudget.tgz"), files=filesNeeded, exdir="Data/histManagement_ResAllom", extras="--keep-newer-files ")

# SoilCarbon    <- read.magpie(paste0("Data/histManagement_ResAllom/SoilCarbon.rds"))
# Landuse       <- read.magpie(paste0("Data/histManagement_ResAllom/Landuse.rds"))
# LanduseChange <- read.magpie(paste0("Data/histManagement_ResAllom/LanduseChange.rds"))

SoilCarbon    <- read.magpie(paste0("Data/rev19histManagement/SoilCarbon.rds"))
Landuse       <- read.magpie(paste0("Data/rev19histManagement/Landuse.rds"))
LanduseChange <- read.magpie(paste0("Data/rev19histManagement/LanduseChange.rds"))

```

Calc global numbers

```{r Carbon Emissions (glo)}

SoilCarbon        <- dimSums(SoilCarbon, dim="subpool")

# Carbon lost at Cropland staying Cropland

years     <- getYears(SoilCarbon, as.integer = TRUE)
#years     <- years[-1]

StayingCropland <- (SoilCarbon[,tail(years,-1),"actualstate.crop"]-setYears(SoilCarbon[,head(years,-1),"actualstate.crop"], head(years,-1)+1))*(Landuse[,tail(years,-1),"crop"]-LanduseChange[,tail(years,-1),"expansion.crop"])

# Carbon lost at newly comverted Cropland

NewlyConverted <- (SoilCarbon[,tail(years,-1),"actualstate.crop"]-setYears(SoilCarbon[,head(years,-1),"actualstate.natveg"], head(years,-1)+1))*(LanduseChange[,tail(years,-1),"expansion.crop"])

StayingCropland_gain <- StayingCropland_loss <- StayingCropland
StayingCropland_gain[StayingCropland<0] <- 0
StayingCropland_loss[StayingCropland>0] <- 0

NewlyConverted_gain <- NewlyConverted_loss <- NewlyConverted
NewlyConverted_gain[NewlyConverted<0] <- 0
NewlyConverted_loss[NewlyConverted>0] <- 0

NetChange <- NetChange_gain <- NetChange_loss <- NewlyConverted + StayingCropland
NetChange_gain[NetChange<0] <- 0
NetChange_loss[NetChange>0] <- 0

diffs <- mbind(setNames(NetChange,      "tot.net"),
               setNames(NetChange_gain, "tot.gain"),
               setNames(NetChange_loss, "tot.loss"),
               setNames(StayingCropland,      "stay.net"),
               setNames(StayingCropland_gain, "stay.gain"),
               setNames(StayingCropland_loss, "stay.loss"),
               setNames(NewlyConverted,       "conv.net"),
               setNames(NewlyConverted_gain,  "conv.gain"),
               setNames(NewlyConverted_loss,  "conv.loss")
               )

iso2cell <- toolGetMapping("CountryToCellMapping.csv", type="cell")
diffsIso <- toolAggregate(diffs, rel=iso2cell, from="celliso", to="iso", dim=1)

NChange <- dimSums(diffs, dim=1)/15
NChangeIso <- diffsIso/15

print(NChange)
print(NChangeIso[,"y2010",])
write.magpie(NChangeIso,"Output/NChange.csv")
write.magpie(NChange,"Output/NChangeGlo.csv")

```

```{r som}

SOM  <- read.magpie("Data/additional/SOM.nc")
SOM2 <- read.magpie("Data/additional/SOM2.nc")

years    <- getYears(SOM, as.integer = TRUE)
getCells(SOM) <- getCells(SOM2) <- getCells(Landuse)
getNames(SOM, dim=1) <- getNames(SOM2, dim=1) <- c("crop", "natveg")

SOC  <- mbind(add_dimension(collapseNames(toolConditionalReplace(SOM[,, "soilc"]/Landuse,"is.na()", 0)), dim=3.1, add="model", nm="som"), 
              add_dimension(collapseNames(toolConditionalReplace(SOM2[,,"soilc"]/Landuse,"is.na()", 0)), dim=3.1, add="model", nm="som2"))

StayingCroplandSOM <- collapseNames((SOC[,tail(years,-1),"crop"]-setYears(SOC[,head(years,-1),"crop"],head(years,-1)+1))*
  (Landuse[,tail(years,-1),"crop"]-LanduseChange[,tail(years,-1),"expansion.crop"]))

# Carbon lost at newly converted Cropland

NewlyConvertedSOM <- collapseNames((SOC[,tail(years,-1),"crop"]-setYears(SOC[,head(years,-1),"natveg"],head(years,-1)+1))*
  (LanduseChange[,tail(years,-1),"expansion.crop"]))

StayingCroplandSOM_gain <- StayingCroplandSOM_loss <- StayingCroplandSOM
StayingCroplandSOM_gain[StayingCroplandSOM<0] <- 0
StayingCroplandSOM_loss[StayingCroplandSOM>0] <- 0

NewlyConvertedSOM_gain <- NewlyConvertedSOM_loss <- NewlyConvertedSOM
NewlyConvertedSOM_gain[NewlyConvertedSOM<0] <- 0
NewlyConvertedSOM_loss[NewlyConvertedSOM>0] <- 0

NetChangeSOM <- NetChangeSOM_gain <- NetChangeSOM_loss <- NewlyConvertedSOM + StayingCroplandSOM
NetChangeSOM_gain[NetChangeSOM<0] <- 0
NetChangeSOM_loss[NetChangeSOM>0] <- 0

diffsom <- mbind(add_dimension(NetChangeSOM, dim=3.2, add="land.mode", nm="tot.net"),
                 add_dimension(NetChangeSOM_gain, dim=3.2, add="land.mode", nm="tot.gain"),
                 add_dimension(NetChangeSOM_loss, dim=3.2, add="land.mode", nm="tot.loss"),
                 add_dimension(StayingCroplandSOM,      dim=3.2, add="land.mode", nm="stay.net"),
                 add_dimension(StayingCroplandSOM_gain, dim=3.2, add="land.mode", nm="stay.gain"),
                 add_dimension(StayingCroplandSOM_loss, dim=3.2, add="land.mode", nm="stay.loss"),
                 add_dimension(NewlyConvertedSOM,       dim=3.2, add="land.mode", nm="conv.net"),
                 add_dimension(NewlyConvertedSOM_gain,  dim=3.2, add="land.mode", nm="conv.gain"),
                 add_dimension(NewlyConvertedSOM_loss,  dim=3.2, add="land.mode", nm="conv.loss")
)

iso2cell <- toolGetMapping("CountryToCellMapping.csv", type="cell")
diffsomIso <- toolAggregate(diffsom, rel=iso2cell, from="celliso", to="iso", dim=1)

NChangeSOM <- dimSums(diffsom, dim=1)/15
NChangeSOMIso <- diffsomIso/15

print(NChangeSOM)
print(NChangeSOMIso[,"y2010",])

NChangeALL    <- mbind(NChangeSOM, add_dimension(NChange, dim=3.1, add="model", nm="mrsoil"))
NChangeIsoALL <- mbind(NChangeSOMIso, add_dimension(NChangeIso, dim=3.1, add="model", nm="mrsoil"))

write.magpie(NChangeSOMIso,"Output/NChangeALL.csv")
write.magpie(NChangeSOM,"Output/NChangeALLGlo.csv")

SOMloss <- dimSums(-SOM[,,"delta_soilc"][,,"crop"]/15, dim=1)

```

### 2021-07-24 ResultNotebooks/results_SOCmaps.Rmd
update results scripts

@@
 
-rev <- "rev19"
+rev <- "rev26"
 setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
@@
 
-for(item in scenarios){
-   dir.create(path=paste0("Data/",paste0(rev,item)))
-   untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-   files=filesNeeded, exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ")
-}
+# for(item in scenarios){
+#    dir.create(path=paste0("Data/",paste0(rev,item)))
+#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+#    files=filesNeeded, exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ")
+# }
 
@@
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(trans = squish_trans(-42,0,10),
-                     limits = c(-46,0),
-                     breaks = c(0, -42, -43, -44, -45, -46),
-                     minor_breaks = seq(-2,-42,-2)) +
+  scale_y_continuous(trans = squish_trans(-32,0,10),
+                     limits = c(-38,0),
+                     breaks = c(0, -32, -33, -34, -35, -36, -37, -38),
+                     minor_breaks = seq(-1,-29,-2)) +
   theme(rect=element_blank(),
@@
   annotate(geom="text", label="SOC debt", x=factor(1995), y=-20, size = 3) +
-  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -24, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
+  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -34, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
 

### 2021-07-24 ResultNotebooks/results_SOCscenario.Rmd
update results scripts

--- 2020-12-14 ResultNotebooks/results_SOCscenario.Rmd
@@
 
-rev         <- "rev16"
-scenarios   <- c("histManagement","constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
-                  "constManagement2-1975", "constResrate-1975", "constYield-1975",
-                 "Initial-natveg", "Initial-natveg_constManagement-1975", "Initial-natveg_constManagement2-1975", 
-                 "LitterPNV-PerennialGrasses", "LitterPNV-AASS", "LitterPNV-BRCF", 
-                 "LitterPNV-MTCF", "LitterPNV-BLEAF",  "LitterPNV-TREF")
+rev         <- "rev26"
+scenarios   <- c("rev26histManagement", "rev26constManagement", "rev26constManagement2", "rev26constManagement_constClimate", "rev26constManagement2_constClimate", "rev26constManure", "rev26constResidues", "rev26constResrate", "rev26constTillage", "rev26constYield", "rev26Initial-1901_constManagement", "rev26Initial-1901_constManagement2", "rev26Initial-1901", "rev26Initial-0_constManagement", "rev26Initial-0_constManagement2", "rev26Initial-0")
 
@@
 # untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
-#       exdir="Data/histManagement", extras="--keep-newer-files ")
-# 
+#       exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")
+
 # for(item in scenarios){
 #    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-#    files=filesNeeded, exdir=paste0("Data/", item), extras="--keep-newer-files ")
+#    files=filesNeeded, exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
 # }
@@
 
-CarbonInput  <- read.magpie(paste0("Data/histManagement/CarbonInput.rds"))
-Landuse      <- read.magpie(paste0("Data/histManagement/Landuse.rds"))
+CarbonInput  <- read.magpie(paste0("Data/rev26histManagement/CarbonInput.rds"))
+Landuse      <- read.magpie(paste0("Data/rev26histManagement/Landuse.rds"))
 glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))
@@
              'manure from ruminants'="livst_rum.crop",   
-             'natural litterfall and fine roots'="litfall.natveg")
+             'natural litterfall wood fraction'="wood.natveg",
+             'natural litterfall leaf fraction'="leaf.natveg")
 
@@
 
-manage_scen <-  c(histManagement   = "histManagement",
-                  constTillage     = "constTillage-mixedtill", 
-                  constManure      = "constManure-1975",
-                  constResidues    = "constResidues-1975", 
-                  constManagement  = "constManagement-1975")
+manage_scen <-  c(histManagement   = "rev26histManagement",
+                  constTillage     = "rev26constTillage", 
+                  constManure      = "rev26constManure",
+                  constResidues    = "rev26constResidues", 
+                  constManagement  = "rev26constManagement")#,
+                  #constManagementNoCC = "rev26constManagement_constClimate")
 
@@
 
-limits <- c(-30,-21)
-breaks <- c(-22, -24,- 26, -28, -30)
+limits <- c(-40,-32)
+breaks <- c(-32, -34,- 36, -38, -40)
 
@@
 
-init_scen   <-  c(`histManagement.Initial-natveg`  = "Initial-natveg", 
-                  `constManagement.Initial-natveg` = "Initial-natveg_constManagement-1975",
-                  `histManagement.Initial-lu`      = "histManagement", 
-                  `constManagement.Initial-lu`     = "constManagement-1975")
+init_scen   <-  c(`histManagement.Initial-natveg`          = "rev26Initial-1901", 
+                  `constManagement.Initial-natveg`         = "rev26Initial-1901_constManagement",
+                  `histManagement.Initial-eq`              = "rev26Initial-0", 
+                  `constManagement.Initial-eq`             = "rev26Initial-0_constManagement",
+                  `histManagement.Initial-spinup1530`      = "rev26histManagement", 
+                  `constManagement.Initial-spinup1530`     = "rev26constManagement")
 
@@
 legend_size <- 10
-scen_cols   <- c("#785EF0","#FE6100","#FFB000")
+scen_cols   <- c("#648FFF","#785EF0","#DC267F")
 input_cols  <- scen_cols[1:4]
 
-limits <- c(-30,-10)
-breaks <- c(-10, -15,- 20, -25, -30)
+limits <- c(-55,-25)
+breaks <- c( -30,-40,-50 )
 
@@
   scale_y_continuous(limits=limits, breaks=breaks) +
-  scale_color_manual(values=c(scen_cols)[1:2], labels = unique(gsub("(.*)\\.(.*)","\\2", names(init_scen)))) +
+  scale_color_manual(values=c(scen_cols)[1:3], labels = unique(gsub("(.*)\\.(.*)","\\2", names(init_scen)))) +
   scale_linetype_manual(values=c(1,2), labels=unique(gsub("(.*)\\.(.*)","\\1", names(init_scen)))) +

### 2021-07-24 ResultNotebooks/results_SOCtable.Rmd
update results scripts

@@
 
-scenarios <- c("rev19histManagement","rev18histManagement_ResAllom_2021_02_01")
+scenarios <- c("rev26histManagement") #,"rev18histManagement_ResAllom_2021_02_01")
 outputdir <- "Output/"
@@
 
-SoilCarbon <- read.magpie(paste0("Data/",scenarios[2],"/SoilCarbon.rds"))
-Landuse    <- read.magpie(paste0("Data/",scenarios[2],"/Landuse.rds"))
+SoilCarbon <- read.magpie(paste0("Data/",scenarios[1],"/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/",scenarios[1],"/Landuse.rds"))
 ```
@@
 years             <- findset("past")
-climateIPCC       <- readRDS("Data/histManagement/IPCC_reduced.rds")
+climateIPCCr      <- readRDS("Data/rev26histManagement/IPCC_reduced.rds")
 
-### calculating stock change factors
+### calculating stock change factors cropand weighted
 SOCDens           <- dimSums(SoilCarbon[,years,c("actualstate","naturalstate")], dim=3.3) 
@@
 
-CShare_ipcczone   <- dimSums(climateIPCC * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
-                     dimSums(climateIPCC * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)
+CShare_ipcczone   <- dimSums(climateIPCCr * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
+                     dimSums(climateIPCCr * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)
+
+# CShare_ipcczone   <- dimSums(climateIPCCr * toolConditionalReplace(SOCDens[,,"crop"][,,"actualstate"]/SOCDens[,,"natveg"][,,"naturalstate"],
+#                                                                  c("is.na()", "is.nan()", "is.infinite()",">3"), c(1,1,1,5)) * Landuse[,years,"crop"], dim=1)/
+#                     dimSums(climateIPCCr * Landuse[,years,"crop"], dim=1)
 
@@
 
-climateIPCCr       <- readRDS("Data/histManagement/IPCC_reduced.rds")
-climateIPCC        <- readRDS("Data/histManagement/IPCC.rds")
+climateIPCCr       <- readRDS("Data/rev26histManagement/IPCC_reduced.rds")
+climateIPCC        <- readRDS("Data/rev26histManagement/IPCC.rds")
 
@@
 save_kable(save, file="Output/Images/TableSCF_comparison.png")
+table
 ```

### 2021-08-09 SOCbudgetPaper/SOCbudget.Rmd
remove copyright statement fix typo in reference
udate paper draft - intro, abstract, small changes in methods

@@
 Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
-Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
-We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
-
+We combine it with as reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
+We estimate that due to arable farming, soils have lost around 32 GtC relative to a counterfactual natural state in 1975.
+Within the period 1975--2010 this SOC debt continued to increasing by 5 GtC ($0.14\unit{GtC yr^{-1}}$) to around 37 GtC.
+However, accounting for historical management rather than assuming constant management conditions led to 2 GtC less $0.06\unit{GtC yr^{-1}}$  emissions.
+We also find that SOC is most sensitive to management decisions concern soil inputs such as residue returning, indicating that biomass increasing SOC sequestration techniques might be (the more) promising negative emission techniques (NET).
+Using the here proposed reduced-complexity SOC model it might be possible to track these management-driven SOC gains within integrated (land-use) assessment modeling accounting dynamically for their socio-economic NET potential.
 bibliography: SOCbudget.bib
@@
 # See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
-copyrightstatement: |
-  The author's copyright for this publication is transferred to institution/company.
+# copyrightstatement: |
+#   The author's copyright for this publication is transferred to Potsdam Institute for Climate Impact Research (PIK).
 availability:

### 2021-08-09 SOCbudgetPaper/introduction.Rmd
udate paper draft - intro, abstract, small changes in methods

@@
 However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
-Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
+Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land (cite more book keeping models??).
 
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions (find corresponding definition of LULUC emission in Pongratz, 2014), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
 Sanderman et al.
@@
 Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015].
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management [@friedlingstein_global_2019; @houghton_carbon_2012; @hansis_relevance_2015].
+BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO~2~ fertilization, N deposition.
+Whereas BKM largely improved in estimating additional emission from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management [@friedlingstein_global_2019; @houghton_carbon_2012; @hansis_relevance_2015] (other citations).
 
@@
 Field-scale models [@del_grosso_simulated_2001, @coleman_simulating_1997, @smith_estimating_2010, @taghizadeh-toosi_c-tool_2014] are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge [@morais_detailed_2019].
 
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
+However, managed soil of the world have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (very long list of citations).
+These soil carbon sequestration (SCS) techniques have been proposed together with other nature-based solution as a mitigation and negative emission technique (NET) to combat climate change.
+Technical SCS potentials have been estimated by (1) up-scaling measured SCS potentials from field-scale experiments (cite), or (2) modeling SCS potentials using stylized future scenarios (cite).
+They range between 0.1 -- 3 GtC per year (cite) with best estimates at 0.5 -- 1.4 GtC per year.
+Assuming SCS application from 2030 till 2100, this accumulates to 35 -- 98 GtC sequestered again, which is just slightly less than the estimated for total, global SOC debt.
+Assessing a more realistic potential considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as also integrated assessment models (cite) have not been able to track management impacts on SOC stocks so far.
+
+The objective of our study is to provide global as well as spatially explicit SOC and SOC debt estimates that considers spatially explicit and time-variant agricultural management.
 To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
-Using a scenario analysis, we decompose the contribution of different management activities.
+We simulate SOC stocks, dynamics and the SOC debt for 1975--2010 using a yearly, spatial explicit, reduced-complexity SOC model.
+We decompose the contribution of different management activities through a scenario analysis, identifying the most impacting management decisions for SOC development.
+Moreover, we compare our model performance against other SOC stock and SOC emission estimates, to evaluate suitability of this approach for further research.

### 2021-08-09 SOCbudgetPaper/methods.Rmd
udate paper draft - intro, abstract, small changes in methods

@@
 We calculate $SOC$ stocks based on the area shares of land-use types ($lu$) within our grid cells ($i$).
-If land is converted from one land-use type $lu=\{crop,natveg\}$ into the other $!lu=\{natveg,crop\}$, a respective share of the $SOC$ is reallocated.
+If land is converted from one land-use type $lu=\{crop,natveg\}$ into the other $!lu=\{natveg,crop\}$, a respective share of the $SOC$ is reallocated within our budget.
+As we do not distinguish between freshly converted and old cropland (see \@ref(fig:lutransitions)), we update the average SOC stock per area within this modeling step.
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{t^*}$ via
@@
 
-To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
-While land-use, climate and litter input information are available from 1901, agricultural management data on residue and manure inputs are held constant at the level of 1965.
-After a spin-up period of 64 years (1901-1965) with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975 onwards.
-Irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
+We initialize our SOC sub-pools using a three-step approach, since input data availability is limited for climate and litter (from 1901) as well as for agricultural management data (from 1965):
+
+Firstly, to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we account for land-use change from 1510.
+We repeat climate and subsequently also litter input data from 1901-1930 to simulate a constant climate for this first initializing period.
+Agricultural management data is hold constant at the level of 1965.
+In 1510 we assume all SOC pools to be in natural steady-state, implicating that all land-use change emissions prior to that time, are redistributed to 1510.
+Due to this long first spin-up period we suppose that the effect of this late accounting and appertaining delayed legacy fluxes can be neglected in 1901.
+This approaches follows others studies looking on effects of land-use change and management (cite LPJmL (which?)).
+
+Secondly, to account for the impact of a changing climate, we account not only for land-use change, but also for historical climate and consequently natural litter inputs to the soil, from 1901 to 1965 still considering constant agricultural input data.
+
+Thirdly, we run the model for additional 10 years with historic dynamic input data and start analyzing results from 1975 onward.
+This is in line with IPCC method suggestion to have a 5-20 year spin-up period.
+
+Irrigation areas are part of the water effect associated with the climate data set and therefore dynamic from 1901 on, whereas data on no-tillage areas is only available after 1974.
 

### 2021-08-09 ResultNotebooks/investigate_SOCbugfixDiff.Rmd
update results

@@
   fig_caption: yes
-  includes:
-    in_header: myHeader.tex
 ---

### 2021-08-17 ResultNotebooks/create_LUtransition.Rmd
update LUC figure for methods

@@
 
-```{r arrange plot, out.height=1000 , out.width=2000}
+```{r arrange plot, out.height=3000 , out.width=3300}
 library(cowplot)
@@
 
-start     <- "../FlowFigure/LUC_start.pdf"
-trans     <- "../FlowFigure/LUC_transition.pdf"
-hetero    <- "../FlowFigure/LUC_hetero.pdf"
-homo      <- "../FlowFigure/LUC_homo.pdf"
-text      <- "../FlowFigure/LUC_text.pdf"
+start     <- "../FlowFigure/LUC_start_rotated.pdf"
+trans     <- "../FlowFigure/LUC_transition_rotated.pdf"
+hetero    <- "../FlowFigure/LUC_hetero_rotated.pdf"
+homo      <- "../FlowFigure/LUC_homo_rotated.pdf"
+text      <- "../FlowFigure/LUC_text_rotated.pdf"
 outputdir <- "Output/Images"
@@
 grid
-ggsave(paste0(outputdir,"/LUtransition.png"), plot=grid, width = 20, height = 10, dpi = 400, units = "cm", bg="white")
+ggsave(paste0(outputdir,"/LUtransition_rotated.png"), plot=grid, width = 33, height = 30, dpi = 400, units = "cm", bg="white")
 

### 2021-08-17 SOCbudgetPaper/methods.Rmd
update method part and pdf

@@
+---
+output:
+  pdf_document: default
+  html_document: default
+---
 # Methods
@@
 
+```{r captionLUtrans, echo=FALSE}
+captionLUtrans <- "Scheme of land-use transition representation for a simplified case.."
+```
+
+```{r LUtrans, out.width = "16cm", fig.cap = captionLUtrans, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images//LUtransition.png")
+```
+
 We calculate $SOC$ stocks based on the area shares of land-use types ($lu$) within our grid cells ($i$).
 If land is converted from one land-use type $lu=\{crop,natveg\}$ into the other $!lu=\{natveg,crop\}$, a respective share of the $SOC$ is reallocated within our budget.
-As we do not distinguish between freshly converted and old cropland (see \@ref(fig:lutransitions)), we update the average SOC stock per area within this modeling step.
+As we do not distinguish between newly converted and old cropland (see \@ref(fig:lutransitions)), we update the average SOC stock per area within this modeling step.
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{t^*}$ via
@@
 
-with $A_{lu}$ being the land-use type specific areas, $AR_{lu}$ and $AE_{lu}$ the area reduction resp.
-area expansion of the two land-use types.
+with $A_{lu}$ being the land-use type specific areas, $AR_{lu}$ the area reduction and $AE_{lu}$ the area expansion of the two land-use types.
 Data sources and methodology on land-use states and changes are described in Sect.
@@
 [-@eggleston_ipcc_2006] and Calvo Buendia et al.
-[@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+[-@calvo_buendia_ipcc_2019] as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 [-@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modeling.
-Starting from harvested crop production ($CP$) estimates and their respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific harvest index values ($HI$) and root:shoot ratios ($RS$) as follows
+Starting from crop yield estimates ($Y$) and respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific ratios for above-ground residues to harvested biomass ($r^{\mathrm{ag,prod}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{tC}}$), above-ground residues to harvested area ($r^{\mathrm{ag,area}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{ha}}$) and below-ground residues to above-ground biomass ($r^{\mathrm{bg}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{tC}}$) as follows
 
@@
 \begin{aligned}
-AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
+AGR_{i,t,cg} & = CA_{i,t,cg} \cdot \left( Y_{i,t,cg} \cdot  r^{\mathrm{ag,prod}}_{cg}+  r^{\mathrm{ag,area}}_{cg} \right)
 \qquad & \textrm{and} \\
-BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
+BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot r^{\mathrm{bg}}_{cg} \qquad                                            & \forall\quad cg, i, t.
 (\#eq:resbiomass)
@@
 
-Following the IPCC guidelines, we split the harvest index into a production ($HI^{\mathrm{prod}}$) and an area dependent ($HI^{\mathrm{area}}$) fraction [@eggleston_ipcc_2006].
+Following the IPCC guidelines, we split the above-ground residue calculations into a production ($r^{\mathrm{ag,prod}}$) and an area dependent ($r^{\mathrm{ag,area}}$) fraction [@hergoualch_n2o_in_ipcc_2019].
 Deviating from Bodirsky et al.
@@
 The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in $\unit{t ha^{-1}}$) at the level of 1975 onwards.
+manure (in $\unit{t ha^{-1}}$) at the level of 1975 onward.
 Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
 The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
-
-### Initialization choice {#sec:scen_initial}
-
-As outlined in Sect.
-\@ref(sec:initsoc) we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
-As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
-Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
-
-### Parameterization of natural litter fall {#sec:scenlitterpnv}
-
-Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short LG:C) and nitrogen-to-carbon (short N:C) ratios for all land under natural vegetation.
-This is however a strong simplification.
-Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines [@calvo_buendia_ipcc_2019], but for woody biomass we used the CENTURY configuration file [@century_model_2000] for trees as parameterization were not given.
-Whereas LG:C ratios ranges between $0.09-0.35$ with an average of $0.24$ (default value) for different tree compartments and different tree types, N:C ratios range between $0.001-0.05$ with an average of $0.015$ (default value).
-We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \@ref(tab:scenlitterpnv).
-
-```{=latex}
-
- \begin{table*}[h]
- \caption{Scenarios for sensitivity analysis on litter parameterization}
- \begin{tabular}{l l l l}
- \tophline
-  Scenario name & plant type & N:C & LG:C \\
-  \middlehline
-  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
-  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
-  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
-  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
-  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
-  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
-  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- (\#tab:scenlitterpnv)
- \end{table*}
-```

### 2022-02-24 SOCbudgetPaper/results.Rmd
remove copyright statement fix typo in reference
update of result part

@@
 In the following, the most important results [see @karstens_result_2020] for post-processing script) are summarized.
+Additionally, we compared our model results to other literature estimates within the last section (see Sect.
+\@ref(sec:validate)).
 
@@
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
-"
+(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is increasing over time, meaning net SOC losses on global croplands over the period 1975--2010.
+The speed of SOC losses is however decreasing."
 ```
@@
 In Fig.
-\@ref(fig:SOCmaps)(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
+\@ref(fig:SOCmaps)(a) we provide a world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
 Values ranging between well over $100\unit{t ha^{-1}}$ in northern temperate croplands to less than $5\unit{t ha^{-1}}$ for arid and semiarid croplands.
@@
 North-East Asia).
-In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
 2.
@@
 Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
+Conversely, irrigated croplands at the border to dry, unsuitable areas worldwide have high stock change factors.
 
-The global SOC debt has decreased by about 15% in the period between 1975 and 2010 to $22\unit{GtC}$ (Fig.
+The global SOC debt has increased by about 14% in the period between 1975 and 2010 to $39.6\unit{GtC}$ (Fig.
 \@ref(fig:SOCmaps)(d)).
-This corresponds to a sequestration rate of $0.11\unit{GtC yr^{-1}}$.
-Considering our estimate of the global SOC stock of around $660\unit{GtC}$ in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
+This corresponds to a loss rate of $0.14\unit{GtC yr^{-1}}$.
+Considering our estimate of the global SOC stock of around $705\unit{GtC}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
 
@@
 We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
-Without changes in management regimes, the global $\Delta SOC$ on cropland would be increasing at a rate of $0.1\unit{GtC yr^{-1}}$.
+Without changes in management regimes, the global $\Delta SOC$ on cropland would be decreasing at a rate of $0.14\unit{GtC yr^{-1}}$.
 As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global $\Delta SOC$ would still increase at a rate of $0.06\unit{GtC yr^{-1}}$.
-Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of $0.09\unit{GtC yr^{-1}}$ for both).
-The effect of no-tillage has been particuarly strong since the 1990s.
+Without the historic increase in residue returning, the global $\Delta SOC$ would still increase at a rate of $0.19\unit{GtC yr^{-1}}$.
+Both the constManure and constTillage scenarios show only small deviations from the historic values (loss rate of $0.15\unit{GtC yr^{-1}}$ for both).
+The effect of no-tillage only became visible from 2000 onwards.
 The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \@ref(fig:SOCscen)(b).
 
-```{r captionSOCinit, echo=FALSE}
-captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use."
+## Model validation {#sec:validate}
+
+To evaluate our model results we take five steps: (1) we compare our stock change factors (see Sect.
+\@ref(#sec:tier1)) to IPCC default assumptions ([@lasco_cropland_in_ipcc_2006], [@ogle_cropland_in_ipcc_2019]); (2) we compare our global (and climate-zone specific) total SOC stocks to other literature estimates; (3) we contrast our results to point measurements; (4) to evaluate the representation of our natural SOC stocks we correlated LPJmL4 SOC stocks for potential natural vegetation with our natural state SOC results on grid level; and (5) we do a similar correlation analysis for our modeled actual SOC stocks in comparison to the results of SoilGrids 2.0 [@poggio_soilgrids_2021].
+
+### Stock change factors compared to IPCC assumptions {#sec:ipcccompare}
+
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors $F^{\\mathrm{SCF}}$ of 2006 [@lasco_cropland_in_ipcc_2006] and their refinements in 2019 [@ogle_cropland_in_ipcc_2019].
+Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
+To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
+\@ref(fig:CLIMzone)) by weighting grid-level $F^{\\mathrm{SCF}}$ with cropland area.
+
+```{r SCFglo, echo=FALSE, results = 'asis'}
+table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
+library(xtable)
+print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
+              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC.
+For the tropical regions the assumptions changed notablly from the guidelines in 2006 \\citep{lasco_cropland_in_ipcc_2006} to the update in 2019 \\citep{ogle_cropland_in_ipcc_2019}, leaving our results in very good agreement with the old default assumptions.
+Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent.
+Modelled $F^{SCF}$ have increased or stayed constant for all climates over time."), comment=FALSE, caption.placement = "top")
 ```
 
-```{r SOCinit, out.width="10cm", fig.cap = captionSOCinit, echo=FALSE, fig.pos="h"}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_init.png")
+### Global SOC stocks comparison
+
+We contrast our global SOC stocks with a wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1; @poggio_soilgrids_2021; @sanderman_soil_2017] in Fig.
+\@ref(fig:SOCglo).
+
+```{r captionSOCglo, echo=FALSE}
+captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, SoilGrids 2.0, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
+This makes it hard to pinpoint a specific year for these SOC estimations.
+In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation, estimates from \\citep{@sanderman_soil_2017} and this study for the year 2010."
 ```
 
-Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\@ref(fig:SOCinit)).
-The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC), as does the decrease in $\Delta SOC$ for the period of 1975--2010 (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.
+```{r SOCglo, out.width = "8cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
+```
 
-Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas $\Delta SOC$ vary  19-24 GtC for the year 2010 (see Fig.
-\@ref(fig:SOClitter).
-The increase of $\Delta SOC$ during the period of 1975--2010 is independent of the parameterization choice at ~4 GtC.
+The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
+SoilGrids [@hengl_soilgrids250m_2017] especially stands out for their high estimation, whereas all other sources (including our study) are comparably similar.
+Looking at regional results in Fig.
+\@ref(fig:SOCreg), our estimates turn out to be in within the range of other estimates for most of the world, with the largest deviations for polar dry, tropical dry and tropical wet areas.
+
+### SOC point-based validation
+
+We correlate our SOC results for native vegetated and cropped areas in 2010 with literature values from point measurements (for data base see appendix of [@sanderman_soil_2017]).
+
+```{r captionSOCpoint, echo=FALSE}
+captionSOCpoint <- "Correlation between modeled and measured SOC stocks: Given the wide span between minimun and maximum measured SOC stocks within in a given cell, we correlated median values with our modeled results.
+Both cropland and areas with natural vegetation (here called Native) are generally lower within our results compared to the measured data."
+```
+
+```{r SOCpoint, out.width = "10cm", fig.cap = captionSOCpoint, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scatterComparePointData_2010.png")
+```
+
+### Natural SOC stock comparison with LPJmL4
+
+As described in Sect.
+\@ref(#sec:steadystates) the natural litterfall of LPJmL4 is the carbon input source for the areas with natural vegetation.
+Here we compare SOC values from a LPJmL4 run for potential natural vegetation with our estimates SOC stocks for a potential natural state.
+Note that both models see the same climate conditions and the same natural litterfall and just differ within their SOC dynamics including the processing of the litterfall.
+We restrict the validation data to the year 2010 and only account for grid cells with more than $1000\unit{ha}$ of cropland.
+
+```{r captionSOCLPJmL4, echo=FALSE}
+captionSOCLPJmL4 <- "Correlation between modeled SOC stocks of LPJmL4 and this study for a potential natural state: Spatial correlation of natural SOC stock values are good especially for dry climate zones.
+For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results except some outliers."
+```
+
+```{r SOCLPJmL4, out.width = "13cm", fig.cap = captionSOCLPJmL4, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png")
+```
+
+### Actual SOC stock comparison with SoilGrids 2.0.
+
+SoilGrids 2.0 [@poggio_soilgrids_2021] is a digital soil mapping approaches that uses over 240 000 soil profile observations to produce high resolution soil maps including SOC stocks and their uncertainties.
+To evaluate the performance of our model on the global scale we correlate SoilGrids 2.0 SOC stock predictions, that were aggregated to 0.5 degree resolution, to our estimates from 2010 in Fig.
+\@ref(fig:SOCSoilGrids2).
+To focus our comparison on cropland areas, we mask out grid cells with less than $1000\unit{ha}$.
+Additionally, we use the uncertainty estimates from SoilGrids 2.0 (fifth and 95th quantile) in Fig.
+\@ref(fig:SOCSoilGrids2Q) to highlight areas, where our modeled SOC stocks have very low agreement with SoilGrids 2.0 data.
+
+```{r captionSOCSoilGrids2, echo=FALSE}
+captionSOCSoilGrids2 <- "Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0: Spatial correlation is fair for tropical climate zones, whereas it is poor for especially temperate moist areas, where our modeled SOC stocks are considerable higher than SoilGrids 2.0 estimates.
+However, for low SOC stocks thoughout the different climate zones our model results seem to be lower than predicted values by Soil Grids 2.0."
+```
+
+```{r SOCSoilGrids2, out.width = "13cm", fig.cap = captionSOCSoilGrids2, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareSoilGrids2_2010_lu.png")
+```
+
+```{r captionSOCSoilGrids2Q, echo=FALSE}
+captionSOCSoilGrids2Q <- "Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0: For the vast majorty of grid cells our model results are between the fifth and 95th quantile of SoilGrids 2.0 estimates.
+We underestimate SOC stocks especially close to dry areas (e.g.
+close to the Sahara) and overestimates stocks for some areas within the center of North America, Middle East Asia and Western South America."
+```
+
+```{r SOCSoilGrids2Q, out.width = "14cm", fig.cap = captionSOCSoilGrids2Q, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/SOCquantilesSoilGrids2.png")

### 2022-02-26 SOCbudgetPaper/results.Rmd
update results after bennis comments inclusion

@@
 # Results
-
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens [-@karstens_model_2020].
-In the following, the most important results [see @karstens_result_2020] for post-processing script) are summarized.
-Additionally, we compared our model results to other literature estimates within the last section (see Sect.
-\@ref(sec:validate)).
 
@@
 Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is increasing over time, meaning net SOC losses on global croplands over the period 1975--2010.
+(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is increasing over time, meaning net SOC losses on global cropland over the period 1975--2010.
 The speed of SOC losses is however decreasing."
@@
 
+The global SOC debt has increased by about 14% in the period between 1975 and 2010 to $39.6\unit{GtC}$ (Fig.
+\@ref(fig:SOCmaps)(d)).
+This corresponds to a loss rate of $0.14\unit{GtC\,yr^{-1}}$.
+Considering our estimate of the global SOC stock of around $705\unit{GtC}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+
 In Fig.
-\@ref(fig:SOCmaps)(a) we provide a world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over $100\unit{t ha^{-1}}$ in northern temperate croplands to less than $5\unit{t ha^{-1}}$ for arid and semiarid croplands.
+\@ref(fig:SOCmaps)(a) we provide a world map of SOC stocks estimates for the first 30 cm on cropland considering historic management data at the global scale for the year 2010.
+Values ranging between over $100\unit{t\,ha^{-1}}$ in northern temperate cropland to less than $5\unit{t\,ha^{-1}}$ for arid and semiarid cropland.
 Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways: 
@@
 They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
-Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, irrigated croplands at the border to dry, unsuitable areas worldwide have high stock change factors.
-
-The global SOC debt has increased by about 14% in the period between 1975 and 2010 to $39.6\unit{GtC}$ (Fig.
-\@ref(fig:SOCmaps)(d)).
-This corresponds to a loss rate of $0.14\unit{GtC yr^{-1}}$.
-Considering our estimate of the global SOC stock of around $705\unit{GtC}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+Large parts of tropical cropland seems to suffer from strongly reduced relative stocks indicating SOC degradation.
+Conversely, irrigated cropland at the border to dry, unsuitable areas worldwide shows a strong relative increase in stocks.
 
@@
 caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
+After accounting for losses within the agricultural system three major C inputs are applied to cropland: manure, above- and below-ground residues.
 Large parts of C, however, get mineralized on field before entering the soil.
@@
 ```{r captionSOCscen, echo=FALSE}
-captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
+captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management (histManagement) by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
 ConstManagement combines all three modifications.
@@
 
-We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
-Without changes in management regimes, the global $\Delta SOC$ on cropland would be decreasing at a rate of $0.14\unit{GtC yr^{-1}}$.
-As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global $\Delta SOC$ would still increase at a rate of $0.19\unit{GtC yr^{-1}}$.
-Both the constManure and constTillage scenarios show only small deviations from the historic values (loss rate of $0.15\unit{GtC yr^{-1}}$ for both).
-The effect of no-tillage only became visible from 2000 onwards.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \@ref(fig:SOCscen)(b).
+We analyze the relative impact of different management practices by comparing the actual historic management scenario with counterfactual scenarios, where individual management practices are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
+As shown by the constResidues scenario, changes in residue return rates dominate the management effects.
+Without the historic increase in residue returning, the global $\Delta SOC$ would still decrease at a rate of $0.19\unit{GtC\,yr^{-1}}$ compared to $0.14\unit{GtC\,yr^{-1}}$ for the histManagement estimates.
+Both the constManure and constTillage scenarios show with $0.15\unit{GtC\,yr^{-1}}$ only small deviations from the historic argicultural management values.
+The effect of no-tillage only becomes visible from 2000 onwards.
+The large contribution of residues relative to manure also becomes visible when considering the yearly C inputs of residues and manure to soils over a period of 35 years (Fig. \@ref(fig:SOCscen)(b)).
 
@@
 To evaluate our model results we take five steps: (1) we compare our stock change factors (see Sect.
-\@ref(#sec:tier1)) to IPCC default assumptions ([@lasco_cropland_in_ipcc_2006], [@ogle_cropland_in_ipcc_2019]); (2) we compare our global (and climate-zone specific) total SOC stocks to other literature estimates; (3) we contrast our results to point measurements; (4) to evaluate the representation of our natural SOC stocks we correlated LPJmL4 SOC stocks for potential natural vegetation with our natural state SOC results on grid level; and (5) we do a similar correlation analysis for our modeled actual SOC stocks in comparison to the results of SoilGrids 2.0 [@poggio_soilgrids_2021].
+\@ref(#sec:tier1)) to IPCC default assumptions ([@lasco_cropland_in_ipcc_2006], [@ogle_cropland_in_ipcc_2019]); (2) we compare our global (and climate-zone specific) total SOC stocks to other literature estimates; (3) we compare our results to point measurements.
+To evaluate the representation of our natural SOC stocks (4) we correlated LPJmL4 SOC stocks for potential natural vegetation with our natural state SOC results on grid level; and (5) we do a similar correlation analysis for our modeled actual SOC stocks in comparison to the results of SoilGrids 2.0 [@poggio_soilgrids_2021].
+Here we use our SOC stocks estimates under actual land-use patterns as SoilGrids 2.0 takes land use and cover information into account for their high resolution global soil map.
 
@@
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors $F^{\\mathrm{SCF}}$ of 2006 [@lasco_cropland_in_ipcc_2006] and their refinements in 2019 [@ogle_cropland_in_ipcc_2019].
-Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\@ref(fig:CLIMzone)) by weighting grid-level $F^{\\mathrm{SCF}}$ with cropland area.
+To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock change factors $F^{\mathrm{SCF}}$ of 2006 [@lasco_cropland_in_ipcc_2006] and their refinements in 2019 [@ogle_cropland_in_ipcc_2019].
+Both estimates are based on measurement data for cropland (see Table \@ref(tab:SCFglo)).
+To allow for comparison, we aggregate our stock change factors by weighting grid-level $F^{\mathrm{SCF}}$ with cropland area to derive factors for the four IPCC climate zones (Fig.
+\@ref(fig:CLIMzone)).
 
@@
 SoilGrids [@hengl_soilgrids250m_2017] especially stands out for their high estimation, whereas all other sources (including our study) are comparably similar.
-Looking at regional results in Fig.
-\@ref(fig:SOCreg), our estimates turn out to be in within the range of other estimates for most of the world, with the largest deviations for polar dry, tropical dry and tropical wet areas.
+Regional results (Fig.
+\@ref(fig:SOCreg)) show that our estimates are within the range of other estimates for most of the world, with the largest deviations for polar dry, tropical dry and tropical wet areas.
 
@@
 
-We correlate our SOC results for native vegetated and cropped areas in 2010 with literature values from point measurements (for data base see appendix of [@sanderman_soil_2017]).
+We correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements (for data base see appendix of [@sanderman_soil_2017]).
 
@@
 captionSOCpoint <- "Correlation between modeled and measured SOC stocks: Given the wide span between minimun and maximum measured SOC stocks within in a given cell, we correlated median values with our modeled results.
-Both cropland and areas with natural vegetation (here called Native) are generally lower within our results compared to the measured data."
+Both cropland and areas with natural vegetation are by tendency lower within our results than within the point measurements."
 ```
@@
 
-As described in Sect.
-\@ref(#sec:steadystates) the natural litterfall of LPJmL4 is the carbon input source for the areas with natural vegetation.
-Here we compare SOC values from a LPJmL4 run for potential natural vegetation with our estimates SOC stocks for a potential natural state.
-Note that both models see the same climate conditions and the same natural litterfall and just differ within their SOC dynamics including the processing of the litterfall.
-We restrict the validation data to the year 2010 and only account for grid cells with more than $1000\unit{ha}$ of cropland.
+Estimates of SOC stocks under natural vegetation influence our modeled results for croplands when the croplands have a history of land conversion.
+We therefore also validate our model simulation for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4.
+Both models see the same climate conditions and the same natural litterfall and just differ within their SOC dynamics including the processing of the litterfall.
+For our purpose only areas which are used as croplands are of relevance; we therefore restrict the validation cells with more than $1000\unit{ha}$ of cropland (capturing 99.9% of global cropland area).
 
 ```{r captionSOCLPJmL4, echo=FALSE}
-captionSOCLPJmL4 <- "Correlation between modeled SOC stocks of LPJmL4 and this study for a potential natural state: Spatial correlation of natural SOC stock values are good especially for dry climate zones.
+captionSOCLPJmL4 <- "Correlation between modeled SOC stocks of LPJmL4 and this study for a potential natural state for the year 2010: Spatial correlation of natural SOC stock values are good especially for dry climate zones.
 For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results except some outliers."
@@
 SoilGrids 2.0 [@poggio_soilgrids_2021] is a digital soil mapping approaches that uses over 240 000 soil profile observations to produce high resolution soil maps including SOC stocks and their uncertainties.
-To evaluate the performance of our model on the global scale we correlate SoilGrids 2.0 SOC stock predictions, that were aggregated to 0.5 degree resolution, to our estimates from 2010 in Fig.
+To evaluate the performance of our model on the global scale we correlate SoilGrids 2.0 SOC stock predictions, that were aggregated to 0.5 degree resolution, to our estimates for the year 2010 in Fig.
 \@ref(fig:SOCSoilGrids2).
 To focus our comparison on cropland areas, we mask out grid cells with less than $1000\unit{ha}$.
-Additionally, we use the uncertainty estimates from SoilGrids 2.0 (fifth and 95th quantile) in Fig.
-\@ref(fig:SOCSoilGrids2Q) to highlight areas, where our modeled SOC stocks have very low agreement with SoilGrids 2.0 data.
+Additionally, we use the uncertainty estimates from SoilGrids 2.0 in Fig.
+\@ref(fig:SOCSoilGrids2Q) to highlight areas, where our modeled SOC stocks that are below the 5th or above the 95th quantile of SoilGrids 2.0 data.
 
@@
 ```{r captionSOCSoilGrids2Q, echo=FALSE}
-captionSOCSoilGrids2Q <- "Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0: For the vast majorty of grid cells our model results are between the fifth and 95th quantile of SoilGrids 2.0 estimates.
+captionSOCSoilGrids2Q <- "Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0: For the vast majorty of grid cells our model results are between the 5th and 95th quantile of SoilGrids 2.0 estimates.
 We underestimate SOC stocks especially close to dry areas (e.g.

### 2022-02-28 SOCbudgetPaper/discussion.Rmd
major update of discussion

@@
 
-This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
-This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
+We have (1) developed a reduced-complexity model and (2) compiled a spatially explicit time series data set of agricultural management data that helped us to (3) analyze the role of agricultural management in historical cropland SOC dynamics.
+Our study shows that agricultural management data alter estimates of the SOC stock size and slows down the increase of SOC debt compared to the often used constant management assumptions.
 
-## SOC debt and SOC drivers in literature
+It is important to evaluate the validity of our results as modeling management effects on SOC dynamics at the global scale comes with large uncertainties.
+The quantitative assessment of the uncertainty of our projections unfortunately exceeds the scope of this article and would likely require a study in itself.
+The model includes a large number of parameters, and for most of these the uncertainty distributions have not been quantified so far.
+Moreover, we think that beyond parameter uncertainty, the structural uncertainty from the model design is high.
+The management data itself is prone to uncertainties as well, as most of it is only indirectly calculated from reported data.
 
-The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 [@minasny_soil_2017], indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
+In the following, we give a qualitative assessment of the uncertainties and limitations of this study as well as discuss our three study objectives and results against existing literature.
 
-According to Sanderman et al.
-[-@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as Sanderman et al.
-[-@sanderman_soil_2017] estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.
-[-@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.
-This suggests that our results are within a plausible range.
+## Comprehensive historical agricultural management data set
 
-Furthermore, Sanderman et al.
-[-@sanderman_soil_2017] modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-Pugh et al.
-[-@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
-Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
+Our spatially explicit time series data set of agricultural management
+is based on country-specific FAO production and cropland statistics [@faostat_faostat_2016] as well as 0.5 degree land-use data from LUH2 [@hurtt_harmonization_2020].
+Starting from these two source, we compile a harmonized and consistent data set for the major flows within the agricultural system (\@ref(fig:FlowFig)) using rule-based assumptions (e.g.
+from the IPCC guidelines Vol.
+4 [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019]) and only a few auxiliary data sets (e.g.
+[@porwollik_generating_2018]).
 
-Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to Pugh et al.
-[-@pugh_simulated_2015] findings of only small effects due to improved practices.
+For some of the aspects covered in our data set, for example livestock distribution [@robinson_mapping_2014] or manure production and application [@zhang_global_2017] very well compiled data sets in high resolution exists, that capture real world conditions much better than our estimates.
+However, they often come with the caveats of either being static in time, demanding large sets of auxiliary data or being inconsistent to each other.
+Within IAM modeling, not only the SOC model itself, but also the input data that drives the model, has to be computed endogenously and dynamically in time, leaving these data sets, that can not be easily projected into the future unsuitable.
 
-Looking on historic cropland intensification rates of above 50% [@rudel_agricultural_2009], a large fraction of increased C input from residue biomass is attributed to productivity improvements.
+For most of our parameters used in our rule-based calculation no uncertainty estimates exist.
+This is why, in our view, most of the uncertainty is included in the management data itself, and especially in the residue and manure numbers, as they are only indirectly calculated from production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017] using large sets of parameter assumptions.
+The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
+As this is one of the first attempts to compile a comprehensive, global data set on these management data, it is complex to evaluate it against literature.
+Furthermore, it is still to be evaluated, if our data set covers most of the carbon inputs to the soil, as it not includes cover crops, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
+Unfortunately, data on the quantity of these inputs is very scarce and often does not exist with global coverage.
 
-## Limitations and uncertainties
+> (aus den point-by-point replies): In the draft of the International Nitrogen Assessment (will be provided confidentially to the reviewer), the available literature estimates for these flows (in regard to nitrogen) were reviewed.
+So far no reliable estimates exist, but the existing estimates indicate that these inputs must be by an order of magnitude lower than those of crop residues and manure.
 
-Modeling management effects at the global scale comes with parametric and structural uncertainties.
-High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.
-the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
+As SOC inputs from above-ground residues had the strongest effect on SOC debt dynamics (see Fig.
+\@ref(fig:SOCscen)).
 As pointed out by Keel et al.
@@
 While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
-However, looking on the evaluation of management effects (see Sect.
-\@ref(sec:ipcccompare)), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
-Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.
-cover- and inter-cropping practices.
+Looking on the evaluation of management effects (see Sect.
+\@ref(sec:ipcccompare)), there is no general indication of overestimating SCFs compared to IPCC default assumptions.
+In contrast, it is even more likely that we are still missing carbon inputs to the soil from e.g.
+cover- and inter-cropping practices as SCF are lower than predicted by the IPCC.
 
-Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
-As shown in Fig.
-\@ref(fig:SOCinit) the SOC debt estimate almost halves (from ~26 GtC to ~14 GtC), and the SOC debt reduction is strongly reduced (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-Pugh et al.
-[-@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
+## Reduced complexity SOC model
 
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamic happens in the topsoil.
+The newly developed reduced-complexity SOC model is based on a Tier 2 modeling approach.
+This reduces the computational and data demand of the model in comparison to DGVMs.
+Still, the effects of water supply from spatially heterogenic rainfall and irrigation as well as tillage systems can be accounted for in the computation of SOC decay rates.
+Similarly, C inputs into the soils reflect differences in productivity in space and over time as well as C losses in the removal rates and processing steps of residues and manure production and application rates to cropland.
+
+The substantial impact of changing management practices in time is indicated by the development of our estimated stock change factors (see Table \@ref(tab:gloSCF)) as well as by the time trend of the SOC debt (see Fig.
+\@ref(fig:SOCmaps)(a)).
+Residue management has changed over the last decades, especially with fading out of residue burning practices in several regions and increased general productivity, this shows a clear impact on the SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
+
+The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly made for agricultural soils, whereas we apply it also to soils under PNV.
+This is necessary in order to represent SOC losses under land-use change in a dynamics way, which is a main driver of SOC dynamics (cite).
+Also, the comparison against simulated PNV data with LPJmL4 shows substantial skill in reproducing PNV SOC stocks (Fig.
+\@ref(fig:SOCLPJmL4)).
+
+Using litterfall estimates from LPJmL4, we have been able to estimate the total SOC stocks of the world, which is dominated by SOC under natural vegetation.
+However, as the world's SOC stock is highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1; @poggio_soilgrids_2021; @sanderman_soil_2017] in Fig.
+\@ref(fig:SOCglo), the only conclusion we can draw from this is, that our result is within a plausible range.
+
+To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management drop out in our calculation of SOC debt and stock change factors.
+
+Focusing on cropland SOC, some differences between our simulated SOC patterns and reference data show specific shortcomings that need to be considered in the application of our model and in the interpretation of results.
+In some mountainous regions, we strongly overestimate SOC stocks compared to SoilGrids 2.0 [@poggio_soilgrids_2021].
+This could indicate that we are not capable to capture specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils [@borrelli_assessment_2017]) or limited accuracy in the SoilGrids 2.0 data for these regions.
+
+In comparison with aggregated results from the IPCC guidelines, our model shows too strong SOC losses for tropical regions (Table \@ref(tab:SCFglo)).
+While tropical soils might suffer from low C input rates due to large yield gaps (cite global yield gap atlas?) and high shares of residue removal (MADRAT?
+FAO for residue burning???), initial SOC stocks under PNV are also comparatively low and thus have little potential for large absolute SOC declines.
+The revised IPCC factors also indicate that also relative SOC differences are much smaller in the tropics, which cannot be reproduced well by our model.
+With the production-increasing impact of irrigation and fertilization on carbon-poor dryland soils, SOC under cropland can also be higher than under PNV with stock change factors above 1 (see Fig.
+\@ref(fig:SOCmaps)(d)), but areas are much smaller than areas where the stock change factors are well below unity.
+
+Generally, limiting the analysis to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamic happens in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 [-@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
+No-tillage effects may seem larger than they actually are, if only focusing on the topsoil.
 SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
@@
 Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
-Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well [@sanderman_soil_2017; @friedlingstein_global_2019].
-This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
 
-## Modeled management effect in line with default IPCC assumptions {#sec:ipcccompare}
+> where to put this?
+With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors (see Table \@ref(tab:gloSCF)) as also indicated by the time trend of the SOC debt.
+Residue management has changed over the last decades, especially with fading out of residue burning practices in several regions and increased general productivity, this shows a clear impact on the SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 [@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019].
-Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\@ref(fig:CLIMzone)).
+## SOC debt and SOC drivers
 
-```{r SCFglo, echo=FALSE, results = 'asis'}
-table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
-library(xtable)
-print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
-              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 \\citep{eggleston_ipcc_2006} to the update in 2019 \\citep{calvo_buendia_ipcc_2019}.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time."), comment=FALSE, caption.placement = "top")
-```
+The analysis of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
+There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period [e.g.
+@friedlingstein_global_2019].
+There is uncertainty with respect to the overall size of these emissions from different methods and reference points and with respect to the contribution of cropland and agricultural management to these emissions.
+In order to mitigate climate change, there is hope that cropland management can be adapted to sequester atmospheric C in cropland soils [@minasny_soil_2017].
+Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], we find that global cropland has acted as a carbon source since 1975.
+Annual C loss rates of 0.2 per 1000 still have the opposite trend as the promoted 4 per 1000 C sequestration rate target [@minasny_soil_2017].
+Dedicated efforts to increase cropland SOC are thus necessary, as management improvements at historical rates are not enough to counteract ongoing SOC degradation on cropland.
 
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than Calvo Buendia et al.
-[-@calvo_buendia_ipcc_2019].
-The tropical factors differ substantially between the IPCC guidelines in 2006 [@eggleston_ipcc_2006] and their refinement in 2019 [@calvo_buendia_ipcc_2019], with our estimates within this range.
-Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show  especially in temperate dry regions --- the smallest region area-wise --- small deviations from natural SOC stocks.
-Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\@ref(fig:SOCmaps)(c)) may be expected.
+According to Sanderman et al.
+[-@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
+Our estimate of 39.6 GtC in 2010 for cropland debt is thus twice as high as their estimate.
+However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.
+[-@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.
 
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
+Furthermore, Sanderman et al.
+[-@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+Pugh et al.
+[-@pugh_simulated_2015] considered management effects like tillage and residue returning in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
+Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
+Historical productivity increases were, however, notably larger.
+Despite large spatial heterogeneity, aggregate yield increase rates are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
 
-## SOC stocks in line with literature
+Our study uses a dynamic management data set as driver for SOC dynamics, including historical yield trends and associated increases in residue inputs.
+[@herzfeld_soc_2021] also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
+Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
+This is consistent with our finding that increasing SOC inputs from above-ground residues had the strongest effect on the slowing-down of the SOC debt increase (Fig.
+@ref(fig:SOCscen)).
 
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1] in Table \@ref(fig:SOCglo)).
+While we find substantially larger effects of productivity gains than the 1% reported by Pugh et al.
+[-@pugh_simulated_2015], this is not sufficient to compensate SOC losses from moderate global cropland expansion of around 11% between 1974 and 2010.
 
-```{r captionSOCglo, echo=FALSE}
-captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
-This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010."
-```
+The effects of agricultural productivity on cropland SOC dynamics can be directly accounted for in our modeling approach, whereas process-based studies [@pugh_simulated_2015; @herzfeld_soc_2021] often lack data on relevant management aspects that drive production increases.
+Elliott et al.
+[-@elliott_management_trends_2018] show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, genetic traits, such as kernel number and radiation use efficiency.
+As such, it will remain challenging for process-based DGVMs to capture the trend of agricultural productivity on cropland SOC dynamics.
 
-```{r SOCglo, out.width = "8cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
-```
+Our study emphasizes again that the expansion of croplands is still a major source of CO~2~ emissions --- not only by the removal of vegetation, but also by a slow depletion of C stocks in soils.
+Our estimates indicate a SOC debt of 39.6 GtC in 2010, and every additional deforestated hectare adds to this debt.
+Avoided deforestation and other environmental regulation leads to intensification on existing cropland [@humpenoder_large-scale_2018] and our results show that such intensification could lead to increased cropland SOC, if residues are returned to the soil --- amplifying the C sequestration potential of avoided deforestation.
 
-The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids [@hengl_soilgrids250m_2017] especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
-Looking at regional results in Fig.
-\@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
-To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
-
-Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
-For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
-
-We conducted a sensitivity analysis (Fig.
-\@ref(fig:SOClitter)) based on various plant parameterizations from the Century model (see Sect.
-\@ref(sec:scenlitterpnv)).
-This shows that the general trend of decreasing SOC debt of ~4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.
+Yet, there is ample potential for further improved SOC management.
+As shown in Fig.
+\@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
+Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the soil surface (1.4 GtC).
+Improved management could include, firstly, a circular flow from the food supply chain back to soils.
+Waste composting or excreta recycling could represent a major additional N input to cropland soils [@brenzinger_organic_2018].
+Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of c inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops [@poeplau_carbon_2015; @porwollik_covercrops_2022] and agroforestry [@lorenz_soil_2014] could increase total C sequestration on cropland.
+However, data on adoption rates and patterns are missing to help analyzing their impact on historical SOC trends and the future potential of cover crops and agroforestry for increasing SOC stocks has not been explored yet.

### 2022-02-28 ResultNotebooks/create_LUtransition.Rmd
Major update of finalized Paper - Draft 2

--- 2021-07-24 ResultNotebooks/create_LUtransition.Rmd
@@
 
-```{r arrange plot, out.height=1000 , out.width=2000}
+```{r arrange plot, out.height=3000 , out.width=3300}
 library(cowplot)
 library(ggplot2)
+library(magick)
+library(grImport)
 
-start     <- "../FlowFigure/LUC_start.pdf"
-trans     <- "../FlowFigure/LUC_transition.pdf"
-hetero    <- "../FlowFigure/LUC_hetero.pdf"
-homo      <- "../FlowFigure/LUC_homo.pdf"
-text      <- "../FlowFigure/LUC_text.pdf"
-outputdir <- "Output/Images"
+path <- "../FlowFigure/"
 
-p1 <- ggdraw() + draw_image(start)
-p2 <- ggdraw() + draw_image(trans)
-p3 <- ggdraw() + draw_image(hetero)
-p4 <- ggdraw() + draw_image(homo)
-pt <- ggdraw() + draw_image(text)
+start     <- "LUC_start.svg"
+trans     <- "LUC_trans.svg"
+hetero    <- "LUC_hetero.svg"
+homo      <- "LUC_homo.svg"
+text      <- "LUC_text.svg"
 
-grid_1 <- plot_grid(p1, p2, p3, ncol=3)
-grid_2 <- plot_grid (pt, p4, ncol=2, rel_widths = c(2, 1))
-grid   <- plot_grid(grid_1, grid_2, nrow=2)
+#for(tmp in c(start, trans, hetero, homo, text)) PostScriptTrace(paste0(path, tmp), paste0("Output/", tmp, ".xml"))
+for(tmp in c(start, trans, hetero, homo, text)) grConvert::convertPicture(paste0(path, tmp), paste0("Output/", gsub("\\.(.*)", "\\.svg", tmp)))
+
+start     <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", start))))
+trans     <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", trans))))
+hetero    <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", hetero))))
+homo      <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", homo))))
+text      <- grImport2::pictureGrob(grImport2::readPicture(paste0("Output/", gsub("\\.(.*)", "\\.svg", text))))
+
+# start     <- pictureGrob(readPicture(paste0("Output/", start, ".xml")))
+# trans     <- pictureGrob(readPicture(paste0("Output/", trans, ".xml")))
+# hetero    <- pictureGrob(readPicture(paste0("Output/", hetero, ".xml")))
+# homo      <- pictureGrob(readPicture(paste0("Output/", homo, ".xml")))
+# text      <- pictureGrob(readPicture(paste0("Output/", text, ".xml")))
+
+# gridExtra::grid.arrange(g[1], g[2], g[3], g[4], g[5], nrow=2)
+
+#PostScriptTrace("../FlowFigure/LUC_start.svg")
+
+# # pt <- ggdraw() + draw_image(text)
+# 
+# #grid_1 <- plot_grid(p1, p2, p3, ncol=3)
+# #grid_2 <- plot_grid (pt, p4, ncol=2, rel_widths = c(2, 1))
+grid   <- plot_grid(start, trans, hetero, homo, text, nrow=2)
 grid
-ggsave(paste0(outputdir,"/LUtransition.png"), plot=grid, width = 20, height = 10, dpi = 400, units = "cm", bg="white")
+ggsave(paste0(outputdir,"/LUtransition_rotated.pdf"), plot=grid, width = 22, height = 30, units = "cm", bg="white")
 

### 2022-02-28 ResultNotebooks/investigate_SOCvalidate.Rmd
Major update of finalized Paper - Draft 2

@@
 library(madrat)
+library(mrcommons)
 
@@
 
-SoilCarbon <- read.magpie("Data/rev26histManagement/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/rev26histManagement/Landuse.rds")
+rev         <- "rev29"
+
+SoilCarbon <- readRDS(paste0("Data/", rev, "histManagement/SoilCarbon.rds"))
+Landuse    <- readRDS(paste0("Data/", rev, "histManagement/Landuse.rds"))
 
@@
 
-cropshare       <- toolConditionalReplace(Landuse[,,"crop"]/dimSums(Landuse, dim=3), "is.na()", 0)
+cropland        <- Landuse[,,"crop"]
 SOCStock_natveg <- dimSums(SoilCarbon[,,c("naturalstate.natveg")], dim=3)
@@
 
-SOCStock_natveg[cropshare<=0.05] <- NA
-SOCStock_lu[    cropshare<=0.05] <- NA
+SOCStock_natveg[cropland <= 0.001] <- NA
+SOCStock_lu[    cropland <= 0.001] <- NA
 
@@
 
-  library(mrvalidation)
+  # SoilGrids2
+  SoilGrids2       <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_grid.rds"))
+  SoilGrids2_full  <- mbind(setNames(SoilGrids2,     "SOC"),
+                            setNames(readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_Q0p95.rds")), "SOC_Q95"),
+                            setNames(readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_Q0p05.rds")), "SOC_Q05"))
 
-  setConfig(cachefolder  = "/home/kristine/mnt/rd3mod/inputdata/cache/rev26/", forcecache = TRUE)
-  setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
-  setConfig(mappingfolder = "/home/kristine/mnt/rd3mod/inputdata/mappings/")
-
-  SoilGrids2   <- readRDS("Data/additional/SoilGrids2.rds")
-  LPJmL4_CRU   <- toolCoord2Isocell(readRDS("Data/additional/LPJmL4_CRU.rds"))
-  LPJmL4_soilc <- readSource("LPJmL_new", subtype="LPJmL4_for_MAgPIE_84a69edd:CRU4:historical:soilc_layer", convert="onlycorrect")*0.01
-  LPJmL4_soilc <- toolCoord2Isocell(LPJmL4_soilc)
-  LPJmL4_soilc <- setNames(LPJmL4_soilc[,,"layer1"] + 1/3*LPJmL4_soilc[,,"layer2"], "layer_0-30")
-  LPJmL4_soilc <- LPJmL4_soilc[,getYears(Landuse),]
+  #LPJmL4
+  LPJmL4_CRU <- toolCoord2Isocell(readRDS(paste0("Data/", rev, "histManagement/LPJmL4_CRU.rds")))
   LPJmL4_CRU <- LPJmL4_CRU[,getYears(Landuse),]
 
-  cells      <- getCells(LPJmL4_soilc)
-
-  LPJmL4_soilc[   cropshare     <=0.05] <- NA
-  LPJmL4_CRU[     cropshare     <=0.05] <- NA
-  SoilGrids[cropshare[,"y2010",]<=0.05] <- NA
+  LPJmL4_CRU[cropland            <= 0.001] <- NA
+  SoilGrids2[cropland[,"y2010",] <= 0.001] <- NA
+  SoilGrids2_full[cropland[,"y2010",] <= 0.001] <- NA
 
@@
 
-```{r correation plot soilc, fig.height=7.5, fig.width=6}
+```{r corr LPJmL4 measures}
 
-regToCell <- toolGetMapping("clustermap_rev4.51+mrmagpie10_c200_h12.rds",type="regional")
-regions   <- levels(regToCell$region)
-climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
-climate$N <- as.integer(substring(climate$cell,5))
-climate   <- climate[order(climate$N),]
-
-# climate
-library(RColorBrewer)
-library(scales)
-blucols          <- brewer.pal(9, "Blues")[2:5]
-viocols          <- brewer.pal(9, "Purples")[6:9]
-redcols          <- brewer.pal(9, "Reds")[6:8]
-orgcols          <- brewer.pal(9, "Oranges")[3]
-cols <- c(blucols[4], viocols[1], redcols[1], orgcols[1])
-paired <- brewer.pal(12, "Paired")
+library(stats)
+library(weights)
+library(MetricsWeighted)
 
@@
 
-climates <- unique(climate$IPCC_reduced)
+validData <- new.magpie(getCells(SOCStock_natveg), 1, 
+                        names = c("LPJmL_CRU", "SOCbudget", "pearson", "meanbias","sdratio", "rmse"), fill = NA)
 
-layout(matrix(c(1,2,3,4,5,5), ncol=2, byrow=TRUE), 
-       heights = c(0.38, 0.42, 0.2), widths =  c(0.52,0.48))
+# for (j in getCells(SOCStock_natveg)) {
+#   
+#   validData[j,,"pearson"] <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])
+#   # validData[j,,"rmse"]    <- MetricsWeighted::rmse(predicted    = SOCStock_natveg[j,years,],
+#   #                                                   actual      = LPJmL4_CRU[j,years,],
+#   #                                           na.rm = TRUE)
+# }
 
-xlabel <- c("", "", "LPJmL4 (t/ha)", "LPJmL4 (t/ha)")
-ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")
+validData[,,"SOCbudget"] <- apply(SOCStock_natveg_raw[,years,], 1, mean)
+validData[,,"LPJmL_CRU"] <- apply(LPJmL4_CRU_raw[,years,], 1, mean)
+validData[,,"meanbias"]  <- validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"]
+validData[,,"sdratio"]   <- apply(SOCStock_natveg_raw[,years,], 1, sd)  / apply(LPJmL4_CRU_raw[,years,], 1, sd)
 
-layout.show(n=5)
+gloValidData <- new.magpie("GLO", 1, 
+                        names = c("socbudget", "lpjml_cru", 
+                                  "pearson", "AE", "RMSE", "sdratio", "meanbias"), fill = NA)
 
-mar <- list(c(2,4,1,1),
-            c(2,2,1,1),
-            c(4,4,1,1),
-            c(4,2,1,1))
+gloValidData[,, "socbudget"]     <- dimSums(validData[,,"SOCbudget"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
+                                      dim = 1, na.rm = TRUE)
 
-for(i in 1:length(climates)){
+gloValidData[,, "lpjml_cru"]     <- dimSums(validData[,,"LPJmL_CRU"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
+                                      dim = 1, na.rm = TRUE)
 
-  icells <- as.array(climate$cell[climate$IPCC_reduced==climates[i]])
-  par(mar = mar[[i]])
-  lim <- quantile(c(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,]), na.rm = TRUE, probs = 0.999)
-  #lim <- max(LPJmL4[icells,years,],SOCStock_natveg[icells,years,], na.rm = TRUE)
+gloValidData[,, "AE"]      <- dimSums(validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"], 
+                                      dim = 1, na.rm = TRUE)
 
-  plot(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,], asp = 1, 
-       xlab = xlabel[i], ylab = ylabel[i],
-       col = "white", pch = ".",
-       xlim = c(0, lim), ylim = c(0, lim))
+gloValidData[,, "RMSE"]    <- MetricsWeighted::rmse(predicted   = as.vector(validData[,,"SOCbudget"]),
+                                                    actual      = as.vector(validData[,,"LPJmL_CRU"]),
+                                            na.rm = TRUE, w = as.vector(cropshare[,"y2010",]))
 
-  for(r in 1:length(regions)){
+gloValidData[,, "meanbias"] <- stats::weighted.mean(validData[, , "meanbias"],
+                                                    w = cropshare[, "y2010", ], na.rm = TRUE)
+gloValidData[,, "sdratio"]  <- stats::sd(validData[, , "SOCbudget"], na.rm = TRUE) /
+                               stats::sd(validData[, , "LPJmL_CRU"], na.rm = TRUE)
+gloValidData[,, "pearson"]  <- weights::wtd.cors(x = validData[,,"SOCbudget"], 
+                                                 y = validData[,,"LPJmL_CRU"], 
+                                                 weight = cropshare[,"y2010",])
 
-    rcells <- intersect(icells,as.array(regToCell$cell[regToCell$region==rev(regions)[r]]))
+print(gloValidData)
 
-    points(LPJmL4_soilc[rcells,years,],SOCStock_natveg[rcells,years,],
-       col = alpha(paired[r], 0.2), pch = ".")
+write.magpie(validData,    "validData.rds")
+write.magpie(gloValidData, "gloValidData.rds")
 
-  }
+#  without time dim
+# gewichtet meanbias -> mean value ->map, global -> gewichtet  
+# -> RMSE - 
+# -> MAE 
+# -> Pearson 
+# -> sdratio (gewichtet???)
 
-  lines(c(0,lim), c(0,lim), col="grey")
-  lines(c(0,lim/2), c(0,lim), col="grey", lty = 2)
-  text(0, lim, paste0("cor=",round(cor(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,], 
-                                 use = "complete.obs"),2)), adj = c(0,1))
-  text(lim/3, lim, paste0("r²=",round(cor(LPJmL4_soilc[icells,years,],SOCStock_natveg[icells,years,], 
-                                       use = "complete.obs")^2, 2)), adj = c(0,1))
-  text(lim*2/3, lim, paste0("EF=",round(qualV::EF(as.array(LPJmL4_soilc[icells,years,][!is.na(LPJmL4_soilc[icells,years,])]),
-                                             as.array(SOCStock_natveg[icells,years,][!is.na(SOCStock_natveg[icells,years,])])),
-                                             2)), adj = c(0,1))
-  text(lim/2, 0, paste0("Climate zone: ", gsub("_", " ", climates[i])))
-}
-
-plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
-legend("top", horiz = TRUE, legend = rev(regions), col = paired, pch=16, bty ="n")
-dev.copy(png, paste0(outputdir, "scatterCompareLPJmL_pnv.png"), width=1200, height=1500, res=250)
-dev.off()
-
-library(ggplot2)
-theme_set(theme_bw())
-library(sf)
-library(rnaturalearth)
-library(rnaturalearthdata)
-library(RColorBrewer)
-library(sp)
-library(rworldmap)
-library(oce) 
-library(scales)
-
-legend_size <- 10
-year        <- 2010
-
-map.world     <- map_data("world")
-world         <- ne_countries(scale = "medium", returnclass = "sf")
-data(coastlineWorld)  
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
-
-magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
-cropland_diff <- collapseNames(SOCStock_natveg[,year,] - LPJmL4_soilc[,year,])
-diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]
-
-SOCdiff <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(diff2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression(paste('natural \u0394SOC in tons of carbon per hectare (this study - LPJmL)')), 
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
-
-SOCdiff
-```
-```{r correation plot CRU, fig.height=7.5, fig.width=6}
-
-regToCell <- toolGetMapping("clustermap_rev4.51+mrmagpie10_c200_h12.rds",type="regional")
-regions   <- levels(regToCell$region)
-climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
-climate$N <- as.integer(substring(climate$cell,5))
-climate   <- climate[order(climate$N),]
-
-# climate
-library(RColorBrewer)
-library(scales)
-blucols          <- brewer.pal(9, "Blues")[2:5]
-viocols          <- brewer.pal(9, "Purples")[6:9]
-redcols          <- brewer.pal(9, "Reds")[6:8]
-orgcols          <- brewer.pal(9, "Oranges")[3]
-cols <- c(blucols[4], viocols[1], redcols[1], orgcols[1])
-paired <- brewer.pal(12, "Paired")
-
-years <- seq(1975, 2010, 5)
-
-climates <- unique(climate$IPCC_reduced)
-
-layout(matrix(c(1,2,3,4,5,5), ncol=2, byrow=TRUE), 
-       heights = c(0.38, 0.42, 0.2), widths =  c(0.52,0.48))
-
-xlabel <- c("", "", "LPJmL4 (t/ha)", "LPJmL4 (t/ha)")
-ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")
-
-layout.show(n=5)
-
-mar <- list(c(2,4,1,1),
-            c(2,2,1,1),
-            c(4,4,1,1),
-            c(4,2,1,1))
-
-for(i in 1:length(climates)){
-
-  icells <- as.array(climate$cell[climate$IPCC_reduced==climates[i]])
-  par(mar = mar[[i]])
-  lim <- quantile(c(LPJmL4_CRU[icells,years,],SOCStock_natveg[icells,years,]), na.rm = TRUE, probs = 0.999)
-  #lim <- max(LPJmL4[icells,years,],SOCStock_natveg[icells,years,], na.rm = TRUE)
-
-  plot(LPJmL4_CRU[icells,years,],SOCStock_natveg[icells,years,], asp = 1, 
-       xlab = xlabel[i], ylab = ylabel[i],
-       col = "white", pch = ".",
-       xlim = c(0, lim), ylim = c(0, lim))
-
-  for(r in 1:length(regions)){
-
-    rcells <- intersect(icells,as.array(regToCell$cell[regToCell$region==rev(regions)[r]]))
-
-    points(LPJmL4_CRU[rcells,years,],SOCStock_natveg[rcells,years,],
-       col = alpha(paired[r], 0.2), pch = ".")
-
-  }
-
-  lines(c(0,lim), c(0,lim), col="grey")
-  lines(c(0,lim/2), c(0,lim), col="grey", lty = 2)
-  text(0, lim, paste0("cor=",round(cor(LPJmL4_CRU[icells,years,],SOCStock_natveg[icells,years,], 
-                                 use = "complete.obs"),2)), adj = c(0,1))
-  text(lim/2, 0, paste0("Climate zone: ", gsub("_", " ", climates[i])))
-}
-
-plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
-legend("top", horiz = TRUE, legend = rev(regions), col = paired, pch=16, bty ="n")
-dev.copy(png, paste0(outputdir, "scatterCompareLPJmLCRU_pnv.png"), width=1200, height=1500, res=250)
-dev.off()
-
-legend_size <- 10
-year        <- 2010
-
-map.world     <- map_data("world")
-world         <- ne_countries(scale = "medium", returnclass = "sf")
-data(coastlineWorld)  
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
-
-magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
-cropland_diff <- collapseNames(SOCStock_natveg[,year,] - LPJmL4_CRU[,year,])
-diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]
-
-SOCdiff <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(diff2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression(paste('natural \u0394SOC in tons of carbon per hectare (this study - LPJmL)')), 
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
-
-SOCdiff
-
-```
-```{r corr measures}
-
-# corr <- new.magpie(getCells(SOCStock_natveg), 1, names = c("r", "r²", "EF", "GRI", "RMSE", "meanbias","sdratio"), fill = NA)
-# 
-# for (j in getCells(SOCStock_natveg))
-#   {
-#   
-#   corr[j,,"r"]    <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])
-#   corr[j,,"r²"]   <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])^2
-#   corr[j,,"EF"]   <- EF(LPJmL4_CRU[j,years,], SOCStock_natveg[j,years,])
-#   corr[j,,"GRI"]  <- GRI(LPJmL4_CRU[j,years,] ,SOCStock_natveg[j,years,])
-#   corr[j,,"RMSE"] <- RMSE(LPJmL4_CRU[j,years,], SOCStock_natveg[j,years,])
-#   print(j)
-# }
-# 
-#   corr[,,"meanbias"] <- apply(SOCStock_natveg[,years,], 1, mean) - apply(LPJmL4_CRU[,years,], 1, mean)
-#   corr[,,"sdratio"]  <-  apply(SOCStock_natveg[,years,], 1, sd)  /apply(LPJmL4_CRU[,years,], 1, sd)
-# 
-# write.magpie(corr, "corr.rds")
-
-luplot::plotmap(corr[,,"meanbias"], legend_range = c(-30,30))
-luplot::plotmap(corr[,,"sdratio"], legend_range = c(0,2))
-luplot::plotmap(corr[,,"r"], legend_range = c(-1,1))
-```
-
-```{r correation plot soilgrids, fig.height=7.5, fig.width=6}
-
-regToCell <- toolGetMapping("clustermap_rev4.51+mrmagpie10_c200_h12.rds",type="regional")
-regions   <- levels(regToCell$region)
-climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
-climate$N <- as.integer(substring(climate$cell,5))
-climate   <- climate[order(climate$N),]
-
-# climate
-library(RColorBrewer)
-library(scales)
-blucols          <- brewer.pal(9, "Blues")[2:5]
-viocols          <- brewer.pal(9, "Purples")[6:9]
-redcols          <- brewer.pal(9, "Reds")[6:8]
-orgcols          <- brewer.pal(9, "Oranges")[3]
-cols <- c(blucols[4], viocols[1], redcols[1], orgcols[1])
-paired <- brewer.pal(12, "Paired")
-
-years <- seq(1975, 2010, 5)
-
-climates <- unique(climate$IPCC_reduced)
-
-layout(matrix(c(1,2,3,4,5,5), ncol=2, byrow=TRUE), 
-       heights = c(0.38, 0.42, 0.2), widths =  c(0.52,0.48))
-
-xlabel <- c("", "", "SoilGrids (t/ha) (timeless)", "SoilGrids (t/ha) (timeless)")
-ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")
-
-layout.show(n=5)
-
-mar <- list(c(2,4,1,1),
-            c(2,2,1,1),
-            c(4,4,1,1),
-            c(4,2,1,1))
-
-for(i in 1:length(climates)){
-
-  icells <- as.array(climate$cell[climate$IPCC_reduced==climates[i]])
-  par(mar = mar[[i]])
-  lim <- quantile(c(SoilGrids[icells,,],SOCStock_natveg[icells,years,]), na.rm = TRUE, probs = 0.999)
-  #lim <- max(SoilGrids[icells,,],SOCStock_natveg[icells,years,], na.rm = TRUE)
-
-  plot(rep(SoilGrids[icells,,],8),SOCStock_lu[icells,years,], asp = 1, 
-       xlab = xlabel[i], ylab = ylabel[i],
-       col = "white", pch = ".",
-       xlim = c(0, lim), ylim = c(0, lim))
-
-  for(r in 1:length(regions)){
-
-    rcells <- intersect(icells,as.array(regToCell$cell[regToCell$region==rev(regions)[r]]))
-
-    points(rep(SoilGrids[rcells,,],8),SOCStock_lu[rcells,years,],
-       col = alpha(paired[r], 0.2), pch = ".")
-
-  }
-
-  lines(c(0,lim), c(0,lim), col="grey")
-  text(0, lim, paste0("cor=",round(cor(rep(SoilGrids[icells,,],8),SOCStock_lu[icells,years,], 
-                                 use = "complete.obs"),2)), adj = c(0,1))
-  text(lim/2, 0, paste0("Climate zone: ", gsub("_", " ", climates[i])))
-}
-
-plot(1, type = "n", axes=F, xlab="", ylab="") # Create empty plot
-legend("top", horiz = TRUE, legend = rev(regions), col = paired, pch=16, bty ="n")
-
-dev.copy(png, paste0(outputdir, "scatterCompareSoilGrids_crop.png"), width=1200, height=1500, res=250)
-dev.off()
+# zeitlich dynamic -> Pearson
+# + sdratio -> SI
 
@@
 
-```{r map2010}
+### LPJmL4 - CRU data
+```{r LPJmL4_CRU plots, fig.height=6, fig.width=12}
 
-legend_size <- 10
-year        <- 2010
+validData_raw <- validData <- validData2    <- read.magpie("validData.rds")
+gloValidData <- read.magpie("gloValidData.rds")
 
-map.world     <- map_data("world")
-world         <- ne_countries(scale = "medium", returnclass = "sf")
-data(coastlineWorld)  
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+brbg_hcl <- colorspace::diverging_hcl(10, h = c(180, 50), c = 80, l = c(20, 95), power = c(0.7, 1.3))
 
-#magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
-load("/home/kristine/WORK/R libraries/mrcommons/R/sysdata.rda")
-cropland_diff <- collapseNames(SOCStock_natveg[,year,] - setYears(SoilGrids2[,1,], NULL))
-diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+validData[cropshare[,"y2010",]  <= 0.05] <- NA
+validData2[cropland[,"y2010",]  <= 0.001] <- NA
 
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]
-
-SOCdiff2010 <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(diff2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare  (this study - SoilGrid2)')), 
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
-
-SOCdiff2010
+luplot::plotmap(validData[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
+luplot::plotmap(validData[,,"meanbias"]/ validData[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
+luplot::plotmap(validData2[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
+luplot::plotmap(validData2[,,"meanbias"]/ validData2[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
+luplot::plotmap(validData_raw[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
+luplot::plotmap(validData_raw[,,"meanbias"]/ validData_raw[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
 
@@
 
-```{r mapmean}
-legend_size <- 10
-year        <- 2010
+```{r quantileSOILGRIDS2, fig.height=6, fig.width=12}
 
-map.world     <- map_data("world")
-world         <- ne_countries(scale = "medium", returnclass = "sf")
-data(coastlineWorld)  
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+InOrOut <- SOCStock_lu[,"y2010",]
+InOrOut[] <- 0
+InOrOut[cropshare[,"y2010",] > 0.05] <- 0.5
+InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q05"] & 
+           SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q95"]] <- 1
+InOrOut[SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q05"]] <- 0.3
+InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q95"]] <- 0.6
+luplot::plotmap(InOrOut)
 
-#magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
-load("/home/kristine/WORK/R libraries/mrcommons/R/sysdata.rda")
-cropland_diff <- collapseNames(magpply(SOCStock_natveg[,years,], mean, 1) - setYears(SoilGrids2[,1,], NULL))
-diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+```
 
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]
+```{r corr soilgrids measures}
 
-SOCdiffmean <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(diff2plot, Year=="0", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare  (this study - SoilGrid2)')), 
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
+library(stats)
+library(weights)
+library(MetricsWeighted)
 
-SOCdiffmean
-SOCdiff2010
-SOCdiff_LPJvsSG2
+years <- 2010
+
+validData <- new.magpie(getCells(SOCStock_natveg), 1, 
+                        names = c("LPJmL_CRU", "SOCbudget", "pearson", "meanbias","sdratio", "rmse"), fill = NA)
+
+# for (j in getCells(SOCStock_natveg)) {
+#   
+#   validData[j,,"pearson"] <- cor(LPJmL4_CRU[j,years,],SOCStock_natveg[j,years,])
+#   # validData[j,,"rmse"]    <- MetricsWeighted::rmse(predicted    = SOCStock_natveg[j,years,],
+#   #                                                   actual      = LPJmL4_CRU[j,years,],
+#   #                                           na.rm = TRUE)
+# }
+
+validData[,,"SOCbudget"] <- apply(SOCStock_lu_raw[,years,], 1, mean)
+validData[,,"LPJmL_CRU"] <- apply(SoilGrids2_raw[,years,], 1, mean)
+validData[,,"meanbias"]  <- validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"]
+validData[,,"sdratio"]   <- apply(SOCStock_lu_raw[,years,], 1, sd)  / apply(SoilGrids2_raw[,years,], 1, sd)
+
+gloValidData <- new.magpie("GLO", 1, 
+                        names = c("socbudget", "lpjml_cru", 
+                                  "pearson", "AE", "RMSE", "sdratio", "meanbias"), fill = NA)
+
+gloValidData[,, "socbudget"]     <- dimSums(validData[,,"SOCbudget"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
+                                      dim = 1, na.rm = TRUE)
+
+gloValidData[,, "lpjml_cru"]     <- dimSums(validData[,,"LPJmL_CRU"] * dimSums(Landuse[,"y2010",], dim = c(2,3)), 
+                                      dim = 1, na.rm = TRUE)
+
+gloValidData[,, "AE"]      <- dimSums(validData[,,"SOCbudget"] - validData[,,"LPJmL_CRU"], 
+                                      dim = 1, na.rm = TRUE)
+
+gloValidData[,, "RMSE"]    <- MetricsWeighted::rmse(predicted   = as.vector(validData[,,"SOCbudget"]),
+                                                    actual      = as.vector(validData[,,"LPJmL_CRU"]),
+                                            na.rm = TRUE, w = as.vector(cropshare[,"y2010",]))
+
+gloValidData[,, "meanbias"] <- stats::weighted.mean(validData[, , "meanbias"],
+                                                    w = cropshare[, "y2010", ], na.rm = TRUE)
+gloValidData[,, "sdratio"]  <- stats::sd(validData[, , "SOCbudget"], na.rm = TRUE) /
+                               stats::sd(validData[, , "LPJmL_CRU"], na.rm = TRUE)
+gloValidData[,, "pearson"]  <- weights::wtd.cors(x = validData[,,"SOCbudget"], 
+                                                 y = validData[,,"LPJmL_CRU"], 
+                                                 weight = cropshare[,"y2010",])
+
+print(gloValidData)
+
+write.magpie(validData,    "validData_soilgrids.rds")
+write.magpie(gloValidData, "gloValidData_soilgrids.rds")
+
+#  without time dim
+# gewichtet meanbias -> mean value ->map, global -> gewichtet  
+# -> RMSE - 
+# -> MAE 
+# -> Pearson 
+# -> sdratio (gewichtet???)
+
+# zeitlich dynamic -> Pearson
+# + sdratio -> SI
+
 ```
-```{r mapLPJsoilGrids}
-legend_size <- 10
-year        <- 2010
 
-map.world     <- map_data("world")
-world         <- ne_countries(scale = "medium", returnclass = "sf")
-data(coastlineWorld)  
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+### LPJmL4 - CRU data
+```{r LPJmL4_CRU plots, fig.height=6, fig.width=12}
 
-#magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
-load("/home/kristine/WORK/R libraries/mrcommons/R/sysdata.rda")
-cropland_diff <- collapseNames(magpply(LPJmL4_CRU[,years,], mean, 1) - setYears(SoilGrids2[,1,], NULL))
-diff2plot     <- merge(as.data.frame(cropland_diff), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+validData_raw <- validData <- validData2    <- read.magpie("validData_soilgrids.rds")
+gloValidData <- read.magpie("gloValidData_soilgrids.rds")
 
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)[7:1]
+brbg_hcl <- colorspace::diverging_hcl(10, h = c(180, 50), c = 80, l = c(20, 95), power = c(0.7, 1.3))
 
-SOCdiff_LPJvsSG2 <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(diff2plot, Year=="0", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
-                      name = expression(paste('cropland \u0394SOC in tons of carbon per hectare  (this study - SoilGrid2)')), 
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
+validData[cropshare[,"y2010",]  <= 0.05] <- NA
+validData2[cropland[,"y2010",]  <= 0.001] <- NA
 
-SOCdiff_LPJvsSG2
+luplot::plotmap(validData[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
+luplot::plotmap(validData[,,"meanbias"]/ validData[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
+luplot::plotmap(validData2[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
+luplot::plotmap(validData2[,,"meanbias"]/ validData2[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
+luplot::plotmap(validData_raw[,,"meanbias"], legend_range = c(-100,100), legend_colours=brbg_hcl)
+luplot::plotmap(validData_raw[,,"meanbias"]/ validData_raw[,,"SOCbudget"], legend_range = c(-0.75,0.75), legend_colours=brbg_hcl)
+
 ```

### 2022-02-28 ResultNotebooks/results_NC.Rmd
Major update of finalized Paper - Draft 2

@@
 
-rev       <- "rev19"
+scenarios <- c("rev27histManagement") #,"rev18histManagement_ResAllom_2021_02_01")
 outputdir <- "Output/"
 
-#filesNeeded  <- c("SoilCarbon.rds","Landuse.rds","LanduseChange.rds")
-#untar(paste0(getConfig("outputfolder"),"/",rev, "histManagement_ResAllom_2021_01_12_h12_carbonbudget.tgz"), files=filesNeeded, exdir="Data/histManagement_ResAllom", extras="--keep-newer-files ")
+#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
 
-# SoilCarbon    <- read.magpie(paste0("Data/histManagement_ResAllom/SoilCarbon.rds"))
-# Landuse       <- read.magpie(paste0("Data/histManagement_ResAllom/Landuse.rds"))
-# LanduseChange <- read.magpie(paste0("Data/histManagement_ResAllom/LanduseChange.rds"))
-
-SoilCarbon    <- read.magpie(paste0("Data/rev19histManagement/SoilCarbon.rds"))
-Landuse       <- read.magpie(paste0("Data/rev19histManagement/Landuse.rds"))
-LanduseChange <- read.magpie(paste0("Data/rev19histManagement/LanduseChange.rds"))
+SoilCarbon    <- read.magpie(paste0("Data/",scenarios[1],"/SoilCarbon.rds"))
+Landuse       <- read.magpie(paste0("Data/",scenarios[1],"/Landuse.rds"))
+LanduseChange <- read.magpie(paste0("Data/",scenarios[1],"/LanduseChange.rds"))
 

### 2022-02-28 ResultNotebooks/results_SOCcompare.Rmd
add revision point by point replies
Major update of finalized Paper - Draft 2

@@
 
+rev        <- "rev30"
+
 outputdir  <- "Output/"
 
-SoilCarbon <- read.magpie("Data/histManagement/SoilCarbon.rds")
-Landuse    <- read.magpie("Data/histManagement/Landuse.rds")
+SoilCarbon <- read.magpie(paste0("Data/", rev, "histManagement/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/", rev, "histManagement/Landuse.rds"))
 
@@
 SOCStock     <- dimSums(dimSums(SoilCarbon[,,c("naturalstate","actualstate")], dim=3.3)* Landuse, dim=3.2)
-cell2reg     <- read.csv("Data/histManagement/ClimateZoneMapping.csv")
+cell2reg     <- read.csv("Data/additional/ClimateZoneMapping.csv")
 reg_SOCStock <- toolAggregate(SOCStock, rel=cell2reg, from="cell", to="IPCC")
@@
 
-  WISE       <- readRDS("Data/histManagement/WISE.rds")
-  GSOC       <- readRDS("Data/histManagement/GSOC.rds")
-  SoilGrids  <- readRDS("Data/histManagement/SoilGrids.rds")
-  LPJmL4     <- readRDS("Data/histManagement/LPJmL4.rds")
+  WISE         <- readRDS(paste0("Data/", rev, "histManagement/WISE.rds"))
+  GSOC         <- readRDS(paste0("Data/", rev, "histManagement/GSOC.rds"))
+  SoilGrids    <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids.rds"))
+  SoilGrids2   <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids2.rds"))
+  LPJmL4       <- readRDS(paste0("Data/", rev, "histManagement/LPJmL4.rds"))
+  LPJmL4_CRU   <- readRDS(paste0("Data/", rev, "histManagement/LPJmL4_CRU.rds"))
+  SOCDebtPaper <- readRDS(paste0("Data/", rev, "histManagement/SOCDebtPaperC.rds"))
 
@@
                   setYears(setNames(SoilGrids[,1,],"SoilGrids"),"y2010"),
-                  setYears(setNames(GSOC[,1,],"GSOC"),"y2010"))
+                  setYears(setNames(GSOC[,1,],"GSOC"),"y2010"),
+                  setYears(setNames(SoilGrids2[,1,],"SoilGrids2"),"y2010"),
+                  setYears(setNames(SOCDebtPaper[,"y2010",],"SOCDebtPaper"),"y2010"))
 glo_SOC <- mbind(setCells(dimSums(reg_SOC, dim=1),"GLO"), reg_SOC)
@@
 
-cols   <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")[c(7,3,1,4,6,8)]
+cols   <-c("black",   "#DF536B", "#61D04F", "#2297E6",
+           "#28E2E5", "#CD0BBC", "#F5C710", "gray62")[c(2, 7, 1, 3, 4, 5, 6, 8)]
 
@@
   geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
-  scale_size_manual(values=c(4,2,2,2,2,2)) + 
-  scale_shape_manual(values=c(19,16,6,2,5,0)) +
+  scale_size_manual( values = c( 4,  2, 2, 2, 2, 2, 2,  2)) + 
+  scale_shape_manual(values = c(19, 16, 6, 2, 5, 0, 8, 10)) +
   scale_color_manual(values = cols) +
@@
         legend.position="bottom") + 
-  guides(fill=guide_legend(nrow=2,byrow=TRUE))
+  guides(fill=guide_legend(nrow=2, byrow=TRUE))
         #legend.position="right") + guides(fill=guide_legend(ncol=1,byrow=TRUE))
@@
   facet_wrap(~ variable, scales="free", drop=TRUE, labeller = label_wrap_gen(width=18)) + 
-  scale_size_manual(values=c(4,2,2,2,2,2)) + 
-  scale_shape_manual(values=c(19,16,6,2,5,0)) +
+  scale_size_manual( values = c( 4,  2, 2, 2, 2, 2, 2,  2)) + 
+  scale_shape_manual(values = c(19, 16, 6, 2, 5, 0, 8, 10)) +
   scale_color_manual(values = cols) +

### 2022-02-28 ResultNotebooks/results_SOCflows.Rmd
Major update of finalized Paper - Draft 2

@@
 outputdir <- "Output/"
-revfolder <- "Data/histManagement/"
+revfolder <- "Data/rev30histManagement/"
 
@@
                              CarbonIn[,,c("livst_chick","livst_egg","livst_milk","livst_pig","livst_rum")], dim=3.1),
-                              add="subpool", nm="manure"))
+                              add="inputs", nm="manure"))
 

### 2022-02-28 ResultNotebooks/results_SOCmaps.Rmd
Major update of finalized Paper - Draft 2

--- 2021-04-01 ResultNotebooks/results_SOCmaps.Rmd
@@
 
-rev <- "rev19"
-setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
+rev <- "rev30"
 
@@
 
-for(item in scenarios){
-   dir.create(path=paste0("Data/",paste0(rev,item)))
-   untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-   files=filesNeeded, exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ")
-}
+# for(item in scenarios){
+#    dir.create(path=paste0("Data/",paste0(rev,item)))
+#    untar(paste0("Data/", paste0(rev,item), "_h12_carbonbudget.tgz"),
+#    exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ") # files=filesNeeded,
+# }
 
@@
 cropland_dens    <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)
-cropland_dens[Landuse[,,"crop"]==0] <- NA
+cropland_dens[Landuse[,,"crop"] <= 0.001] <- NA
 
-cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/dimSums(SoilCarbon[,,"naturalstate"][,,"natveg"], dim=3)
-cropland_share[Landuse[,,"crop"]==0] <- NA
+cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/(dimSums(SoilCarbon[,,"naturalstate"][,,"natveg"], dim=3) + 1e-8)
+cropland_share[Landuse[,,"crop"] <= 0.001] <- NA
 
 cropland_increase <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)-dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3) 
-cropland_increase[Landuse[,,"crop"]==0] <- NA
+cropland_increase[Landuse[,,"crop"] <= 0.001] <- NA
 
@@
   scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(trans = squish_trans(-42,0,10),
-                     limits = c(-46,0),
-                     breaks = c(0, -42, -43, -44, -45, -46),
-                     minor_breaks = seq(-2,-42,-2)) +
+  scale_y_continuous(trans = squish_trans(-34,0,10),
+                     limits = c(-40,0),
+                     breaks = c(0, -34, -35, -36, -37, -38, -39, -40),
+                     minor_breaks = seq(-1,-29,-2)) +
   theme(rect=element_blank(),
@@
   annotate(geom="text", label="SOC debt", x=factor(1995), y=-20, size = 3) +
-  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -24, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
+  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -36, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
 
-cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(7)
+SOCplotmap   <- function(data, name, cols, limits, values = NULL, breaks, labels) {
 
-SOCshare <- ggplot(data = world) + geom_sf() + 
-    geom_tile(data = subset(share2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(0,2), colours = cols, 
-                      name = expression("cropland SOC stock change factor (unitless)"), 
-                      breaks=c(0.3,0.6,0.9,1.1,1.4,1.7,2),
-                      labels=c("0     0.3","  0.6", "  0.9","  1.1","  1.4","  1.7","> 2"), na.value = "white",
+  out <- ggplot(data = world) + geom_sf() + 
+    geom_tile(data = subset(data, Year == "2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    scale_fill_stepsn(limits   = limits, 
+                      colours  = cols, 
+                      name     = name,
+                      values   = values,
+                      breaks   = breaks,
+                      labels   = labels, 
+                      na.value = "white",
                       guide    = guide_legend(
-                        title.theme=element_text(size=legend_size),
+                        title.theme    = element_text(size=legend_size),
                         direction      = "horizontal",
@@
                       )) +
-    theme(legend.position="bottom",
+    theme(legend.position = "bottom",
           axis.title.x     = element_blank(), 
           axis.title.y     = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
+          legend.text      = element_text(size = legend_size),
+          rect             = element_blank(),
+          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7,0.7)),
           plot.margin      = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
@@
 
-#geom_map(dat=map.world, map = map.world, aes(map_id=region), fill="NA", color="black") +
+    return(out)
+}
 
-SOCincrease <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(increase2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(-100,100), colours = cols, 
+cols        <- colorRampPalette(brewer.pal(11, "RdYlBu"))(11)[c(1:6,7,9,11)]
+SOCshare    <- SOCplotmap( data     = share2plot,
+                            limits   = c(0,2),
+                            cols     = cols,
+                            name     = expression("cropland SOC stock change factor (unitless)"),
+                            values   = c(0,0.2,0.4,0.6,0.8,1.0,1.2,2)/2,
+                            breaks   = c(0.2,0.4,0.6,0.8,1.0,1.2,2),
+                            labels   = c("< 0.2","0.4", "0.6","0.8","1.0","1.2","> 1.2"))
+
+SOCincrease <- SOCplotmap( data     = increase2plot,
+                            limits   = c(-100, 100),
+                            cols     = cols,
                             name     = expression(paste('cropland \u0394SOC in tons of carbon per hectare')),
-                      values=(c(-100,-50,-20,-10,10,20,50,100)+100)/200,
-                      breaks=c(-50,-20,-10,10,20,50,100),
-                      labels=c("< -100    -50", "  -20","  -10","  10","  20","  50"," > 100"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
+                            values   = (c(-100, -40, -30, -20, -10, 0, 10, 100)+100)/200,
+                            breaks   =  c(-40, -30, -20, -10, 0, 10, 100),
+                            labels   = c("< -40", "        -30","        -20","      -10 ","     0","    10","> 10"))
 
 cols <- colorRampPalette(brewer.pal(9, "YlGn"))(8)[2:8]
-
-SOCdensity <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(density2plot, Year=="2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
-    scale_fill_stepsn(limits=c(0,150), colours = cols, 
+SOCdensity  <- SOCplotmap( data     = density2plot,
+                            limits   = c(0, 150), 
+                            cols     = cols, 
                             name     = expression("cropland SOC in tons of carbon per hectare"), 
-                      values=c(0,5,10,20,40,70,100,150)/150,
-                      breaks=c(5,10,20,40,70,100,150),
-                      labels=c("0     5","  10","  20","  40","  70","  100"," >150"), na.value = "white",
-                      guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
-                        direction = "horizontal",
-                        title.position = 'top',
-                        # I shift the labels around, the should be placed 
-                        # exactly at the right end of each legend key
-                        title.hjust = 0.5,
-                        label.hjust = 1,
-                        nrow = 1,
-                        byrow = F,
-                        label.position = "bottom",
-                        keyheight = 0.5,
-                        keywidth  = 2
-                      )) +
-    theme(legend.position="bottom",
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
-          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          )
+                            values   = c(0, 5, 10, 20, 40, 70, 100, 150) / 150,
+                            breaks   = c(5, 10, 20, 40, 70, 100, 150),
+                            labels   = c("0     5","  10","  20","  40","  70","  100"," >150"))
 
@@
 library(cowplot)
+SOCemis
+SOCincrease
 SOCdensity
-SOCincrease
-SOCemis
 SOCshare
 
-grid <- plot_grid(SOCdensity,SOCincrease, SOCshare, SOCemis,  ncol=2, labels=c("(a)","(b)", "(c)", "(d)"),
+grid <- plot_grid(SOCemis,SOCdensity,SOCincrease, SOCshare,  ncol=2, labels=c("(a)","(b)", "(c)", "(d)"),
                   label_fontface = "plain", label_size = 10)

### 2022-02-28 ResultNotebooks/results_SOCscattervalidate.Rmd
Major update of finalized Paper - Draft 2

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
  bookdown::pdf_document2: default
   keep_md: true
---

```{r load}
library(magclass)
library(madrat)
library(mrcommons)

outputdir <- "Output/Images/"

rev         <- "rev30"

SoilCarbon <- readRDS(paste0("Data/", rev, "histManagement/SoilCarbon.rds"))
Landuse    <- readRDS(paste0("Data/", rev, "histManagement/Landuse.rds"))

```

Calculate global carbon stocks

```{r Carbon Stock (cell iso reg glo)}

cropland        <- Landuse[,,"crop"]
SOCStock_natveg <- dimSums(SoilCarbon[,,c("naturalstate.natveg")], dim=3)
SOCStock_lu     <- dimSums(SoilCarbon[,,c("actualstate")]*Landuse, dim=3)/dimSums(Landuse, dim=3)

SOCStock_natveg[cropland <= 0.001] <- NA
SOCStock_lu[    cropland <= 0.001] <- NA

```

```{r load soc validation stocks}

  # SoilGrids2
  SoilGrids2       <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_grid.rds"))
  SoilGrids2_full  <- mbind(setNames(SoilGrids2,     "SOC"),
                            setNames(readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_Q0p95.rds")), "SOC_Q95"),
                            setNames(readRDS(paste0("Data/", rev, "histManagement/SoilGrids2_Q0p05.rds")), "SOC_Q05"))

  #LPJmL4
  LPJmL4_CRU <- toolCoord2Isocell(readRDS(paste0("Data/", rev, "histManagement/LPJmL4_CRU.rds")))
  LPJmL4_CRU <- LPJmL4_CRU[,getYears(Landuse),]

  LPJmL4_CRU[cropland            <= 0.001] <- NA
  SoilGrids2[cropland[,"y2010",] <= 0.001] <- NA
  SoilGrids2_full[cropland[,"y2010",] <= 0.001] <- NA

```

```{r correation plot CRU, fig.height=6, fig.width=6}

climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
climate$N <- as.integer(substring(climate$cell, 5))
climate   <- climate[order(climate$N), ]
climates  <- unique(climate$IPCC_reduced)

layout(matrix(c(1, 2, 3, 4), ncol = 2, byrow = TRUE), 
       heights = c(0.48, 0.52), widths =  c(0.52, 0.48))

xlabel <- c("", "", "LPJmL4 (t/ha) in 2010", "LPJmL4 (t/ha) in 2010")
ylabel <- c("This Study - PNV (t/ha) in 2010", "", "This Study - PNV (t/ha) in 2010", "")

layout.show(n = 4)

mar <- list(c(2, 4, 1, 1),
            c(2, 2, 1, 1),
            c(4, 4, 1, 1),
            c(4, 2, 1, 1))

years <- "y2010"

for (i in 1:length(climates)) {

  icells <- as.array(climate$cell[climate$IPCC_reduced == climates[i]])
  par(mar = mar[[i]])
  max <- max(    LPJmL4_CRU[icells, years, ], SOCStock_natveg[icells, years, ], na.rm = TRUE)
  min <- c(10, 20, 10, 20)

  plot(LPJmL4_CRU[icells, years, ], SOCStock_natveg[icells, years, ], asp = 1, 
       xlab = xlabel[i], ylab = ylabel[i],
       col = "white", pch = ".", log = "xy",
       xlim = c(min[i], max), ylim = c(min[i], max))

  points(LPJmL4_CRU[icells ,years ,],SOCStock_natveg[icells, years, ], pch = ".")

  lines(c(min[i], max), c(min[i], max), col = "grey")

  text(min[i], max, paste0("R2=", round(cor(LPJmL4_CRU[icells, years, ],SOCStock_natveg[icells, years, ], 
                                 use = "complete.obs"), 2)), adj = c(0,1))
  text(max, min[i], paste0("Climate zone: ", gsub("_", " ", climates[i])), adj = c(1, 0))
}

dev.copy(png, paste0(outputdir, "scatterCompareLPJmL42010_pnv.png"), width = 1500, height = 1500, res = 250)
dev.off()

print(round(cor(LPJmL4_CRU[, years, ],SOCStock_natveg[, years, ], use = "complete.obs"), 2))

```

```{r correation plot SoilGrids2, fig.height=6, fig.width=6}

climate   <- read.csv("Data/additional/ClimateZoneMapping.csv")
climate$N <- as.integer(substring(climate$cell, 5))
climate   <- climate[order(climate$N),]
climates  <- unique(climate$IPCC_reduced)

layout(matrix(c(1, 2, 3, 4), ncol = 2, byrow = TRUE), 
       heights = c(0.48, 0.52), widths =  c(0.52, 0.48))

xlabel <- c("", "", "SoilGrids2 (t/ha)", "SoilGrids2 (t/ha)")
ylabel <- c("This Study (t/ha)", "", "This Study (t/ha)", "")

layout.show(n = 4)

mar <- list(c(2, 4, 1, 1),
            c(2, 2, 1, 1),
            c(4, 4, 1, 1),
            c(4, 2, 1, 1))

years <- "y2010"

for (i in 1:length(climates)) {

  icells <- as.array(climate$cell[climate$IPCC_reduced == climates[i]])
  par(mar = mar[[i]])
  max <- max(    SoilGrids2[icells ,years, ], SOCStock_lu[icells, years, ], na.rm = TRUE)
  min <- c(5, 10, 5, 10)

  plot(SoilGrids2[icells, years, ],SOCStock_lu[icells, years, ], asp = 1, 
       xlab = xlabel[i], ylab = ylabel[i],
       col = "white", pch = ".", log = "xy",
       xlim = c(min[i], max), ylim = c(min[i], max))

  points(SoilGrids2[icells, years, ],SOCStock_lu[icells, years, ], pch = ".")

  lines(c(min[i], max), c(min[i], max), col = "grey")

  text(min[i], max, paste0("R2=", round(cor(SoilGrids2[icells, years, ], SOCStock_lu[icells, years, ], 
                                 use = "complete.obs"), 2)), adj = c(0,1))
  text(max, min[i], paste0("Climate zone: ", gsub("_", " ", climates[i])), adj = c(1, 0))
}

dev.copy(png, paste0(outputdir, "scatterCompareSoilGrids2_2010_lu.png"), width = 1500, height = 1500, res = 250)
dev.off()

print(round(cor(SoilGrids2[, years, ],SOCStock_lu[, years, ], use = "complete.obs"), 2))

```

```{r quantileSOILGRIDS2, fig.height=6, fig.width=12}

InOrOut <- SOCStock_lu[,"y2010",]
InOrOut[] <- NA
InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q05"] & 
        SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q95"]] <- 1
InOrOut[SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q05"]] <- 0.5
InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q95"]] <- 1.5

magpie_coord      <- toolGetMapping("magpie_coord.rda", type="cell")
soilgrid22plot    <- merge(as.data.frame(InOrOut), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")

```

```{r plot quantileSOILGRIDS2, fig.height=6, fig.width=12}

library(ggplot2)
theme_set(theme_bw())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce) 
library(scales)

legend_size <- 10

map.world <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)  
coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])

SOCplotmap   <- function(data, name, cols, limits, values = NULL, breaks, labels) {

  out <- ggplot(data = world) + geom_sf() + 
    geom_tile(data = subset(data, Year == "2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
    scale_fill_manual(name     = name,
                      values   = cols,
                      labels   = labels, 
                      na.value = "white",
                      guide    = guide_legend(
                        title.theme    = element_text(size=legend_size),
                        direction      = "horizontal",
                        title.position = 'top',
                        # I shift the labels around, the should be placed 
                        # exactly at the right end of each legend key
                        title.hjust    = 0.5,
                        label.hjust    = 1,
                        nrow           = 1,
                        byrow          = F,
                        label.position = "bottom",
                        keyheight      = 0.5,
                        keywidth       = 2
                      )) +
    theme(legend.position = "bottom",
          axis.title.x     = element_blank(), 
          axis.title.y     = element_blank(),
          legend.text      = element_text(size = legend_size),
          rect             = element_blank(),
          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7,0.7)),
          plot.margin      = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

    return(out)
}

cols              <- c("#648FFF", "#DC267F", "#FFB000")[c(1, 3, 2)]
SoilGrids2InMap   <- SOCplotmap( data     =  soilgrid22plot,
                                 limits   = c(0.5, 1.5),
                                 cols     = cols,
                                 name     = expression("Modeled SOC stock in relation to uncertainty range of SoilGrids 2.0"),
                            breaks   = c(0.5, 1.0, 1.5),
                            labels   = c("lower than Q0.05", "inside [Q0.05, Q0.95] ", "higher than Q0.95", "No Cropland"))

SoilGrids2InMap
ggsave(paste0(outputdir, "/SOCquantilesSoilGrids2.png"),    plot=SoilGrids2InMap,    width = 15, height = 10, dpi = 300, units = "cm")

```

### 2022-02-28 SOCbudgetPaper/SOCbudget.tex
remove copyright statement fix typo in reference
Major update of finalized Paper - Draft 2

@@
 
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
+%% Journal abbreviations (please use the same for preprints and final revised papers)
 
@@
 % Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
 % Atmospheric Chemistry and Physics (acp)
@@
 % E&G Quaternary Science Journal (egqsj)
+% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
 % European Journal of Mineralogy (ejm)
@@
 % Ocean Science (os)
+% Polarforschung - Journal of the German Society for Polar Research (polf)
 % Primate Biology (pb)
 % Proceedings of the International Association of Hydrological Sciences (piahs)
+% Safety of Nuclear Waste Disposal (sand)
 % Scientific Drilling (sd)
@@
 
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
-
 % Pandoc citation processing
@@
 % The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
+% % % From pandoc table feature
+% \usepackage{longtable,booktabs,array}
+% % \usepackage{calc} % for calculating minipage widths
+% % Correct order of tables after \paragraph or \subparagraph
+% \usepackage{etoolbox}
+% \makeatletter
+% \patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+% \makeatother
+% % Allow footnotes in longtable head/foot
+% \IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+% \makesavenoteenv{longtable}
 % 
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+%
 \begin{document}
@@
 Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land cover change and agricultural management.
+The first meter of the Earth's soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
+SOC has been depleted by anthropogenic land-cover change and agricultural management.
 However, the latter has so far not been well represented in global carbon stock assessments.
 While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
-Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
-Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
-We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
+Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
+We estimate that due to arable farming, soils have lost around 34.6 GtC relative to a counterfactual hypothetical natural state in 1975.
+Within the period 1975--2010 this SOC debt continued to expand by 5 GtC (\(0.14\unit{GtC yr^{-1}}\)) to around 39.6 GtC.
+However, accounting for historical management led to 2.1 GtC less (\(0.06\unit{GtC yr^{-1}}\)) emissions than under the assumption of constant management.
+We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique.
+The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling.
 \end{abstract}
 
-\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
-
 \newpage
@@
 
-Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool.
+Soil organic carbon (SOC), the amount of organic carbon stored in the Earth's soil, constitutes the largest terrestrial organic carbon pool.
 It exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}.
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{friedlingstein_global_2019, minasny_soil_2017}.
+Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{frielingstein_glob_carbon_2020, minasny_soil_2017}.
 The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile \citep{batjes_total_1996, sanderman_soil_2017}.
@@
 However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
-Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
-
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
+Compared to climatic and geological driving forces, human interventions alter terrestrial carbon pools over much shorter time scales and are currently one of the most dominant drivers of SOC changes on managed land \citep{hansis_relevance_2015, bastos_comparison_2021}.
+
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al.~\citeyearpar{pongratz_terminolgy_2014}), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state.
 Sanderman et al.~\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.
 
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions \citep{friedlingstein_global_2019}.
-While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}.
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management \citep{friedlingstein_global_2019, houghton_carbon_2012, hansis_relevance_2015}.
+Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs), Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions \citep{frielingstein_glob_carbon_2020}.
+While providing estimates of the magnitude of SOC losses due to land-cover change, most models lack a detailed representation of agricultural management.
+Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}.
+BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization and N deposition.
+Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management \citep{frielingstein_glob_carbon_2020, houghton_carbon_2012, hansis_relevance_2015, bastos_comparison_2021}.
 
 Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.~][]{bondeau_modelling_2007, lindeskog_implications_2013}.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
-They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
-
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010} multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][\citet{lutz_simulating_2019}]{pugh_simulated_2015}.
-
-More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
-
-Field-scale models \citep[\citet{coleman_simulating_1997}, \citet{smith_estimating_2010}, \citet{taghizadeh-toosi_c-tool_2014}]{del_grosso_simulated_2001} are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
-
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
-To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
-Using a scenario analysis, we decompose the contribution of different management activities.
+Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data.
+They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.
+
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010}, multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][]{pugh_simulated_2015, lutz_simulating_2019}.
+Herzfeld et al.~\citeyearpar{herzfeld_soc_2021} account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC.
+Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.
+
+More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g.~on tillage systems, see \citep{porwollik_generating_2019, prestele_spatially_2018}) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019, minoli_modelling_2019}), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+
+Field-scale models \citep{del_grosso_simulated_2001, coleman_simulating_1997, smith_estimating_2010, taghizadeh-toosi_c-tool_2014} are able to better account for historical agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge \citep{morais_detailed_2019}.
+
+Managed soils have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (soil carbon sequestration (SCS) techniques).
+However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) \citep{popp_land-use_2016, rogelj_scenarios_2018} have not integrated soil management into their mitigation pathways.
+More detailed process-based models are typically computationally too demanding to be integrated into optimization-based IAMs.
+Better accounting for soil carbon management in IAMs thus requires a light-weight model suitable for iterative modeling with detailed options to represent agricultural soil management.
+
+The objectives of our study are (1) to develop a reduced-complexity SOC model able to account for SCS in IAM frameworks; (2) to create a comprehensive data set of the global gridded management time series, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices; and (3) to provide global as well as spatially explicit SOC and SOC debt estimates that consider spatially explicit and time-variant agricultural management.
+We decompose the contribution of different management activities through a scenario analysis, identifying the most impacting management decisions for SOC development.
+Moreover, we compare our model performance against other SOC stock and SOC emission estimates, to evaluate the suitability of this reduced-complexity approach for integration into IAM modeling.
 \newpage
@@
 In Sect.
-\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}.
-We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}.
+\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail in the refinement of the IPCC guidelines vol.~4 Chapter 5 on ``Cropland'' \citep{ogle_cropland_in_ipcc_2019}.
+We additionally describe how we configured and extended the Tier 2 modeling approach \citep[for model code see][]{karstens_mrsoil_2020}.
 In Sect.
@@
 In Sect.
-\ref{sec:scenarios} we define the scenarios used to complement our historic model results.
+\ref{sec:scenarios} we define the management scenarios used to analyze the role of different management aspects in historical cropland SOC dynamics.
 
 \hypertarget{sec:carbonbudget}{%
-\subsection{SOC stocks and stock changes following the steady-state method}\label{sec:carbonbudget}}
-
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+\subsection{SOC stocks and stock changes following the Tier 2 modeling approach}\label{sec:carbonbudget}}
+
+Following the Tier 2 modeling approach of the refinement of the IPCC guidelines vol.~4 Chapter 5 on ``Cropland'' \citep{ogle_cropland_in_ipcc_2019}; referred to as \textit{Tier 2 modeling approach} in the following), we estimate SOC stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
 We restrict our analysis to the first 0-30 cm of the soil profile.
-Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
+Moreover, we assume the current SOC state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
 Therefore, we take the following three steps for each year of our simulation period:
@@
 This equation is valid for all grid cells \(i\) and all years \(t\).
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep{parton_analysis_1987}.
+We use the Tier 2 modeling approach for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep{parton_analysis_1987}.
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
@@
 
-Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
+Carbon inputs for cropland are below- and above-ground crop residues left or returned to the field (see Sect.
 \ref{sec:residues}) and manure inputs (see Sect.
-\ref{sec:livstmanure}); for natural vegetation litterfall including fine root turnover \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
-Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
+\ref{sec:livstmanure}); for natural vegetation, litterfall including fine root turnover \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
+Following the IPCC guidelines \citep{ogle_cropland_in_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
 For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients.
 This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \ref{tab:datasourceinputs}.
+
+Whereas residue and manure default lignin and nitrogen fractions are given by the IPCC guidelines \citep{ogle_cropland_in_ipcc_2019}, we use plant-functional type and plant-organ specific parameterization for the natural litterfall.
+Global distribution of plant-functional types is given by \citep{schaphoff_lpjml4_2018} as well as separation of litter into leaf, fine root and wood litter compartments excluding litter biomass burnt in wild fires.
+Leaf litter parameters are given by Brovkin et al.~\citeyearpar{brovkin_plant-driven_2012}, fine root to leaf litter lignin ratio by Guo et al.~\citeyearpar{guo_leaf_root_decompostion_2021}, lignin content of wood litter by Rahman et al.~\citeyearpar{rahman_lignin_2013} and nitrogen content scaling factors for leaf to fine roots and leaf to wood litter by von Bloh et al.~\citeyearpar{vBloh_lpjml5_2018}.
+Data sources for all considered carbon inputs as well as for lignin and nitrogen parameterization are listed in Table \ref{tab:datasourceinputs}.
 
   \begin{table*}[h]
-  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
+  \caption{type and data sources for carbon inputs and parameterization to different land-use types }
   \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
@@
   \middlehline
-  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
+  \multirow{3}{*}{cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                       \citet{faostat_faostat_2016}, \\
@@
                          LG:C generic values according to Table 5.5B,\\
-                                                                      5.5C from IPCC \citep{calvo_buendia_ipcc_2019},\\
+                         5.5C from IPCC \citep{ogle_cropland_in_ipcc_2019},\\
                          crop-specific N:C from \citet{bodirsky_n2o_2012}
@@
                             \hline
-  Natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
+  natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
                       & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                IPCC \citep{calvo_buendia_ipcc_2019} and\\ 
-                                CENTURY (\citep{century_model_2000}) \strut \end{minipage}\tabularnewline
+                          leaf N and LG concentration from \citet{brovkin_plant-driven_2012}, 
+                          root to leaf litter LG ratio \citet{guo_leaf_root_decompostion_2021},  
+                          lignin content of wood litter \citet{rahman_lignin_2013} and 
+                          nitrogen scaling factors for leaf to root and wood litter 
+                          from \citet{vBloh_lpjml5_2018} \strut \end{minipage}\tabularnewline
  \bottomhline
@@
 
-The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space (\(i\)) and time (\(t\)).
-Following the steady-state method \citep{calvo_buendia_ipcc_2019}, we consider temperature (\(temp\)), water (\(wat\)), sand fraction (\(sf\)) and tillage (\(till\)) effects to account for spatial variation of decay rates.
-Thus, \(k_{sub}\) are given by
+The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space \(i\) and time \(t\).
+Following the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019}, we consider temperature \(temp\), water \(wat\), sand-fraction \(sf\), and tillage \(till\) effects to account for spatial and temporal variation of decay rates.
+Thus, \(k_{sub}\) rates are given by
 
@@
 \ref{sec:irrigation}) practices on decay rates.
-We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
+We calculate area-weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation practices.
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equations 5.0B--5.0F in Ogle et al.~\citeyearpar{ogle_cropland_in_ipcc_2019}.
 
  \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
+ \caption{type and data sources for carbon inputs to different land-use types}
  \begin{tabular}{l l l l}
@@
  \middlehline
- \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
+ \multirow{2}{*}{all} & soil quality & sand fraction of the first 0-30 cm 
                                      & \citet{hengl_soilgrids250m_2017} \\
@@
 
-                      & Mircobial activity & air temperature & \citet{harris_version_2020} \\
+                      & mircobial activity & air temperature & \citet{harris_version_2020} \\
                       \cline{2-4}
 
-                      & Water restriction & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
+                      & soil moisture & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
                       \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
+\multirow{2}{*}{\begin{minipage}[t]{0.15\columnwidth}\raggedright\strut cropland\\(additionally)\strut\end{minipage}} & soil moisture*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
                       \cline{2-4}
 
-                      & Soil disturbance & tillage & Sect. \ref{sec:irrigation} \\
+                      & soil disturbance & tillage & Sect. \ref{sec:tillage} \\
  \bottomhline
@@
 
-We calculate \(SOC\) stocks based on the area shares of land-use types (\(lu\)) within our grid cells (\(i\)).
-If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the \(SOC\) is reallocated.
+\begin{figure}[h]
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/LUtransitions} \caption{Scheme of land-use transition representation. 
+Given an initial land-use pattern (as in this example 2 ha land under natural vegetation and 1 ha of cropland), there are separate $SOC$ stocks for natural vegetation and cropland.
+While in this example we assume $SOC$ under natural vegetation to be in steady state, the cropland $SOC$ stock approaches its steady state without having reached it yet (a).
+Upon cropland expansion (in this example half of the natural vegetation is cleared to be used as cropland), $SOC$ stocks on cropland increase due to a transfer of land from natural vegetation (b).
+Separately accounting for SOC dynamics on newly converted and existing cropland (c) leads to the same overall SOC dynamics as if SOC stocks are averaged (d), due to the linearity of Eq.
+\ref{eq:SOCstate} and the cropland-age independend decay rates (see Eq.
+\ref{eq:decayrates})}\label{fig:LUtrans}
+\end{figure}
+
+We calculate SOC stocks based on the area shares of land-use types \(lu\) within our grid cells \(i\).
+If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the SOC is reallocated within our budget.
+We do not distinguish between newly converted and existing cropland, but can work with the average carbon content as the relative decay of SOC is proportional to the SOC stock (see \ref{fig:LUtrans}).
 We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{t^*}\) via
@@
 
-with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) and \(AE_{lu}\) the area reduction resp.
-area expansion of the two land-use types.
+with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) the area reduction and \(AE_{lu}\) the area expansion of the two land-use types.
 Data sources and methodology on land-use states and changes are described in Sect.
@@
 
-\(SOC\) converge towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
+SOC converges towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
 
@@
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
+SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific $SOC$ stock within cell}}}_{\text{$SOC_{i,t}$ -- total $SOC$ stock within cell}}.
 \label{eq:totalstock}
@@
 According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years \citep[see Eq.
-5.0A in][]{calvo_buendia_ipcc_2019}.
+5.0A in][]{ogle_cropland_in_ipcc_2019}.
 This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
+For our study, we define the absolute and relative \(SOC\) changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
 The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by
@@
 
-To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
-While land-use, climate and litter input information are available from 1901, agricultural management data on residue and manure inputs are held constant at the level of 1965.
-After a spin-up period of 64 years (1901-1965) with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975 onwards.
-Irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
+The initialization of \(SOC\) pools is very important and has to include the proper accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historical land-use change (LUC) and the slow decline of \(SOC\) under crop cultivation, before it reaches a new equilibrium.
+We initialize our \(SOC\) sub-pools using a three-step approach, since input data availability is limited for climate and litter estimates (starting only in 1901) as well as for agricultural management data (starting only in 1965):
+
+Firstly, in order to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we consider land-use change from 1510 onwards.
+In 1510, we assume all \(SOC\) pools to be in natural steady-state, implying that all land-use change prior to that time occurs in 1510.
+We assume that by 1901, all cropland converted in 1510 has reached its new steady state, so that it is not necessary to explicitly account for even older land conversion.
+Model inputs for 1901--1930 for climate and natural vegetation litterfall are repeated for 1510--1900 to mimic constant climate conditions for this first initializing period.
+Similarly, agricultural management data are held constant at the level of 1965 until 1965.
+This approach follows others studies looking on effects of land-use change and management \citep[e.g.~][]{schaphoff_lpjml4_2018-1, herzfeld_soc_2021}.
+
+Secondly, with the availability of transient climate data after 1901, we account not only for land-use change, but also for historical climate change and consequently natural litter inputs to the soil from 1901 to 1965 still considering constant agricultural input data, which are not available prior to 1965.
+
+Thirdly, we run the model for 10 years from 1965 to 1975 with historical dynamic data on agricultural management and start analyzing results from 1975 onward.
+This is in line with the IPCC guidelines vol.~4 method suggestion to have a 5-20 year spin-up period \citep{ogle_cropland_in_ipcc_2019}.
+
+With transient climate considered after 1901, decay rates \(k_{sub}\) become dynamic in time.
+As the decay rates are also affected by irrigation and tillage (see Sect.
+\ref{sec:steadystates}), we also account for transient changes in irrigated areas after 1901.
+Data on no-tillage practices are only available after 1974 and we assume conventional tillage on all cropland prior to 1975.
 
@@
 
-Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
+Additionally to the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019} and the detailed analysis of management data coming with it, \(SOC\) changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
 Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Estimates of \(F^{\mathrm{SCF}}\) are differentiated by crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows.
 Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig.
@@
 with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of Eggleston et al.~\citeyearpar{eggleston_ipcc_2006} and Calvo Buendia et al.~\citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of Eggleston et al.~\citeyearpar{eggleston_ipcc_2006} and Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
+The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1510--1974).
 For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
@@
 
-Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups \citep[see FAO2LUH2MAG\_croptypes.csv in][]{bodirsky_mrcommons_2020}, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016}.
-By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
+Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set (short LUH2), which we sum up from quarter-degree to half-degree resolution.
+We disaggregate the physical area (in Mha) of the five different cropland subcategories (c3ann: C3 annual crops, c3per: C3 perennial crops, c4ann: C4 annual crops, c4per: C4 perennial crops, c3nfx: C3 nitrogen-fixing crops) of LUH2 into our 17 crop groups (see Table \ref{tab:fao2luh2mag}), applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016}.
+By calculating country-specific multicropping factors \(MCF\) using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
 Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
@@
 \ref{sec:landuse}).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production \citep{faostat_faostat_2016}.
-By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
-
-Crop residue production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
-Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}.
+Note
+
+Crop residue production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and key aspects are explained as they play a central role in soil carbon modeling.
+Starting from crop yield estimates \(Y\) and respective physical crop area \(CA\), we estimate total above-ground \(AGR\) and below-ground \(BGR\) residue biomass (in tonnes) using crop group \(cg\) specific ratios for above-ground residues to harvested biomass \(r^{\mathrm{ag,prod}}_{cg}\) in \(\unit{tC\,\unit{tC}^{-1}}\), above-ground residues to harvested area \(r^{\mathrm{ag,area}}_{cg}\) in \(\unit{tC\,\unit{ha}^{-1}}\) and below-ground residues to above-ground biomass \(r^{\mathrm{bg}}_{cg}\) in \(\unit{tC\,\unit{tC}^{-1}}\) as follows
 
@@
 \begin{aligned}
-AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
+AGR_{i,t,cg} & = CA_{i,t,cg} \cdot \left( Y_{i,t,cg} \cdot  r^{\mathrm{ag,prod}}_{cg} + MCF_{i,t} \cdot r^{\mathrm{ag,area}}_{cg} \right)
 \qquad & \textrm{and} \\
-BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
+BGR_{i,t,cg} & = (CA_{i,t,cg} \cdot Y_{i,t,cg} + AGR_{i,t,cg}) \cdot r^{\mathrm{bg}}_{cg} \quad.
 \label{eq:resbiomass}
@@
 
-Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
-Deviating from Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+Following the IPCC guidelines, we split the above-ground residue calculations into a yield-dependent slope (\(r^{\mathrm{ag,prod}}\)) and a positive intercept (\(r^{\mathrm{ag,area}}\)) fraction \citep{hergoualch_n2o_in_ipcc_2019}.
+Residues biomass therefore increases under-proportionally with rising yields, reflecting a shifting harvest index of higher-yielding breeds.
+Deviating from Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} we use harvested instead of physical crop area (denoted in Eq.
+\eqref{eq:resbiomass} by MCF described in Sect.
+\ref{sec:landuse}) to account for increased residue biomass due to multiple cropping (multiple harvests with each lower yields) and decreased residue amounts due to fallow land.
 We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
@@
 A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
-Following Smil \citeyearpar{smil_nitrogen_1999} we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following Smil \citeyearpar{smil_nitrogen_1999} we assume a burn share of 25\% for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income for the periods before 1995.
+After 1995 we estimate a linear decline of burn shares to 10\% for low-income countries and 0\% for high-income countries till 2025 to account for recent increases in air pollution regulation.
+The estimated trends show good agreement with fire-satellite-image derived estimates by the Global Fire Database \citep{vanderWerf_fire_2017}.
 Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.
@@
 
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{tab:c2dm}).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon and nitrogen to dry matter ratios (see Table \ref{tab:c2dm}).
 
@@
 
-Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems at the present day a lot of the manure has to be stored and can be applied on croplands.
-We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
+Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems a lot of the manure has to be stored and can be applied on cropland.
+We assume that manure is applied in close proximity to its excretion, so that the distribution of livestock is the driving factor for the spatial pattern of manure application.
 
@@
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-Due to great feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
+Due to high feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
 We distinguish between grazed pasture, which is converted into livestock products in extensive systems, and primary-crop feed stuff, which we consider to be consumed in intensive systems.
@@
 Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
+Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.~\citeyearpar{ogle_cropland_in_ipcc_2019}.
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012}.
-We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
+We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
 
@@
 
-The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for their cropland subcategories.
+The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for its cropland subcategories.
 We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient \(wat\) on decay rates using these shares to compute the weighted mean between rainfed and irrigated \(wat\) factors.
 As a result \(wat\) is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
+Furthermore, we suppose the irrigation effect to be present for all 12 months of a year, since we do not have consistent crop group specific growing periods available.
 This will lead to an overestimation of the irrigation effect.
@@
 In order to derive a spatial distribution of the three different tillage types specified by the IPCC --- full tillage, reduced tillage and no tillage ---, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from Porwollik et al.~\citeyearpar{porwollik_generating_2018}.
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO \citeyearpar{fao_aquastat_2016} and half-degree resolution physical crop areas from Hurtt et al.~\citeyearpar{hurtt_harmonization_2020}, applying the methodology of Prowollik et al.~\citeyearpar{porwollik_generating_2018} to identify potential no-tillage grid cells.
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historical global gridded tillage data set from Porwollik et al.~\citeyearpar{porwollik_generating_2019}.
+This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO \citeyearpar{fao_aquastat_2016} and half-degree resolution physical crop areas from Hurtt et al.~\citeyearpar{hurtt_harmonization_2020}, applying the methodology of Prowollik et al.~\citeyearpar{porwollik_generating_2019} to identify potential no-tillage grid cells.
 
 \hypertarget{sec:scenarios}{%
-\subsection{Scenario definitions}\label{sec:scenarios}}
-
-To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios \citep[for technical implementation see][]{karstens_mrsoil_2020}.
-
-\hypertarget{sec:scen_management}{%
-\subsubsection{Management scenarios}\label{sec:scen_management}}
+\subsection{Management scenario definitions}\label{sec:scenarios}}
 
@@
 The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in \(\unit{t ha^{-1}}\)) at the level of 1975 onwards.
+manure (in \(\unit{t ha^{-1}}\)) at the level of 1975 onward.
 Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
 The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
-
-\hypertarget{sec:scen_initial}{%
-\subsubsection{Initialization choice}\label{sec:scen_initial}}
-
-As outlined in Sect.
-\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
-As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
-Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
-
-\hypertarget{sec:scenlitterpnv}{%
-\subsubsection{Parameterization of natural litter fall}\label{sec:scenlitterpnv}}
-
-Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short LG:C) and nitrogen-to-carbon (short N:C) ratios for all land under natural vegetation.
-This is however a strong simplification.
-Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file \citep{century_model_2000} for trees as parameterization were not given.
-Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
-We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \ref{tab:scenlitterpnv}.
-
- \begin{table*}[h]
- \caption{Scenarios for sensitivity analysis on litter parameterization}
- \begin{tabular}{l l l l}
- \tophline
-  Scenario name & plant type & N:C & LG:C \\
-  \middlehline
-  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
-  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
-  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
-  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
-  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
-  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
-  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:scenlitterpnv}
- \end{table*}
-
 \newpage
@@
 \begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
-Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
-}\label{fig:SOCmaps}
+\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data.
+Panel (a) shows global $\Delta SOC$ between historical land use and potential natural vegetation (PNV).
+The distribution of total global SOC stocks for the first 30 cm on cropland for the year 2010 is depcited in panel (b).
+Absolute (c) and relative (cd SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains.}\label{fig:SOCmaps}
 \end{figure}
 
+The global SOC debt has increased by about 14\% in the period between 1975 and 2010 from \(34.6\) to \(39.6\unit{GtC}\) (Fig.
+\ref{fig:SOCmaps}(a)).
+This corresponds to an average loss rate of \(0.14\unit{GtC\,yr^{-1}}\) in comparison to a hypothetical potential natural vegetation (PNV) state.
+Considering our estimate of the global SOC stock of around \(705\unit{GtC}\) in the upper \(30\unit{cm}\) in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+The speed of this SOC loss has decreased towards the end of the modeling period.
+
 In Fig.
-\ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
-Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways:
+\ref{fig:SOCmaps}(b) we provide a world map of SOC stocks estimates for the first \(30\unit{cm}\) on cropland considering historical management data for the year 2010.
+Values range between over \(100\unit{t\,ha^{-1}}\) in northern temperate cropland to less than \(5\unit{t\,ha^{-1}}\) for arid and semiarid cropland.
+Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under PNV in two complementary ways:
 1.
-Absolute SOC changes \(\Delta SOC\) (see Fig.
-\ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses.
-They might be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
-In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
+Absolute SOC changes \(\Delta SOC\) (Fig.
+\ref{fig:SOCmaps}(c)) indicate areas with high importance for the global SOC loss.
+They can be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
 2.
-Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig.
-\ref{fig:SOCmaps}(c)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
-Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
-
-The global SOC debt has decreased by about 15\% in the period between 1975 and 2010 to \(22\unit{GtC}\) (Fig.
-\ref{fig:SOCmaps}(d)).
-This corresponds to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\).
-Considering our estimate of the global SOC stock of around \(660\unit{GtC}\) in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
+Relative SOC changes measured as stock changes factors \(F^{SCF}\) (Fig.
+\ref{fig:SOCmaps}(d)) are a helpful metric to analyze the impact of human cropping activities.
+They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome by improved agricultural practices.
+Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation.
+Conversely, irrigated cropland at the border to dry, unsuitable areas worldwide shows a strong relative increase in SOC stocks.
 
@@
 \begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC).
+Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand.
+After accounting for losses within the agricultural system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues.
+Large parts of C, however, are mineralized on the field before entering the soil.
+Additionally, C is transferred to and from the global agricultural soil stock via land-use change between cropland and natural vegetation.
+Finally, SOC is mineralized and flows back to the atmosphere.}\label{fig:FlowFig}
 \end{figure}
 
-C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
-Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
-We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.~like food, bioenergy and material) into a human demand category.
-Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
-Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not recycled to soils.
-Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
-However, around 60\% of C decomposes before its integrated into soils at the the litter-soil barrier.
-Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
-According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
-4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
-Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
+C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvest organ, above- and below-ground residues).
+Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (\(729\unit{MtC}\) in 2010) are directly returned to the field.
+We divide crop biomass usage into feed usage and aggregate all other usage types (e.g.~food, bioenergy and material) into a human demand category.
+Livestock feed demand for crop organ harvest and above-ground residues of \(1136\unit{MtC}\) is roughly equal to the human demand of \(1129\unit{MtC}\).
+Whereas large parts of feed intake are returned to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not returned to soils.
+Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with \(1350\unit{MtC}\) returned to the fields in 2010.
+However, around 60\% of this organic C decomposes before it is integrated into soils.
+Due to the different composition of organic C, proportionally more C enters the slow pool from manure than from crop residue.
+According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of \(257\unit{MtC}\) in 2010.
+The agricultural system receives \(4585\unit{MtC}\) assimilated by crop plants and releases \(3554\unit{MtC}\) mostly through respiration.
+Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around \(33\unit{MtC}\) for the year 2010.
 
@@
 \begin{figure}[h]
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
+\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios.
+The stylized scenarios deviate from historical argicultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage).
+The scenario constManagement combines all three modifications.
+Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a hypothetical natural vegetation state.
+Panel (b) shows the carbon inflows from crop residue and manure.}\label{fig:SOCscen}
 \end{figure}
 
-We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)).
-Without changes in management regimes, the global \(\Delta SOC\) on cropland would be increasing at a rate of \(0.1\unit{GtC yr^{-1}}\).
-As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global \(\Delta SOC\) would still increase at a rate of \(0.06\unit{GtC yr^{-1}}\).
-Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both).
-The effect of no-tillage has been particuarly strong since the 1990s.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \ref{fig:SOCscen}(b).
-
-\begin{figure}[h]
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
-\end{figure}
-
-Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\ref{fig:SOCinit}).
-The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
-
-Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas \(\Delta SOC\) vary 19-24 GtC for the year 2010 (see Fig.
-\ref{fig:SOClitter}.
-The increase of \(\Delta SOC\) during the period of 1975--2010 is independent of the parameterization choice at \textasciitilde4 GtC.
-
-\newpage
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
-This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
-
-\hypertarget{soc-debt-and-soc-drivers-in-literature}{%
-\subsection{SOC debt and SOC drivers in literature}\label{soc-debt-and-soc-drivers-in-literature}}
-
-The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions \citep[see][]{pugh_simulated_2015}, global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 \citep{minasny_soil_2017}, indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
-
-According to Sanderman et al.~\citeyearpar{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as Sanderman et al.~\citeyearpar{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.~\citeyearpar{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
-This suggests that our results are within a plausible range.
-
-Furthermore, Sanderman et al.~\citeyearpar{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
-Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% \citep{pellegrini_crop_2018, ray_recent_2012, rudel_agricultural_2009}.
-
-Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to Pugh et al.~\citeyearpar{pugh_simulated_2015} findings of only small effects due to improved practices.
-
-Looking on historic cropland intensification rates of above 50\% \citep{rudel_agricultural_2009}, a large fraction of increased C input from residue biomass is attributed to productivity improvements.
-
-\hypertarget{limitations-and-uncertainties}{%
-\subsection{Limitations and uncertainties}\label{limitations-and-uncertainties}}
-
-Modeling management effects at the global scale comes with parametric and structural uncertainties.
-High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by Keel et al.~\citeyearpar{keel_large_2017} and Smith et al.~\citeyearpar{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
-Keel et al.~\citeyearpar{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
-However, looking on the evaluation of management effects (see Sect.
-\ref{sec:ipcccompare}), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
-Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.~cover- and inter-cropping practices.
-
-Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
-As shown in Fig.
-\ref{fig:SOCinit} the SOC debt estimate almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), and the SOC debt reduction is strongly reduced (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
-
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and assumes that most of the SOC dynamic happens in the topsoil.
-In this regard several aspects are strongly simplified within our approach.
-Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
-This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected.
-As Powlson et al.~\citeyearpar{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
-SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
-Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
-
-This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by Doetterl et al.~\citeyearpar{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions \citep{wang_human-induced_2017}.
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
-Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well \citep{sanderman_soil_2017, friedlingstein_global_2019}.
-This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
+We analyze the relative impact of different management practices by comparing the actual historical management scenario with counterfactual scenarios, where individual management practices (residues in constResidues, manure in constManure, tillage practices in constTillage, all three in constManagement) are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)).
+As shown by the difference between the constResidues scenario and the other counterfactuals, changes in residue return rates dominate the management effects.
+Without the historical increase in C inputs from residues to agricultural soils, the global \(\Delta SOC\) would decrease to \(41.7\unit{GtC}\) at a rate of \(0.20\unit{GtC\,yr^{-1}}\) --- a 35\% increase compared to \(0.14\unit{GtC\,yr^{-1}}\) for the histManagement estimates.
+Both the constManure and constTillage scenarios show only small deviations from the historical argicultural management values with \(0.15\unit{GtC\,yr^{-1}}\).
+The effect of no-tillage only becomes discernible from 2000 onwards.
+The large contribution of residues relative to manure also becomes visible when considering the annual C inputs of residues and manure to soils over a period of 35 years (Fig. \ref{fig:SOCscen}(b)).
+
+\hypertarget{sec:validate}{%
+\subsection{Model evaluation}\label{sec:validate}}
+
+To evaluate our model results against reference data in five steps: (1) we compare our stock change factors (see Sect.
+\ref{sec:tier1}) to IPCC default assumptions \citep{lasco_cropland_in_ipcc_2006, ogle_cropland_in_ipcc_2019}; (2) we compare our global (and climate-zone specific) total SOC stocks to other literature estimates; (3) we compare our results to point measurements.
+To evaluate the representation of our natural SOC stocks (4) we correlated LPJmL4 SOC stocks for PNV with our natural state SOC results on grid level; and (5) we do a similar correlation analysis for our modeled actual SOC stocks in comparison to the results of SoilGrids 2.0 \citep{poggio_soilgrids_2021}, which accounts for actual land use too.
 
 \hypertarget{sec:ipcccompare}{%
-\subsection{Modeled management effect in line with default IPCC assumptions}\label{sec:ipcccompare}}
-
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 \citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}.
-Both estimates are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
+\subsubsection{Stock change factors compared to IPCC assumptions}\label{sec:ipcccompare}}
+
+To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors \(F^{\mathrm{SCF}}\) of 2006 \citep{lasco_cropland_in_ipcc_2006} and their refinements in 2019 \citep{ogle_cropland_in_ipcc_2019}.
+Both estimates are based on measurement data for cropland (see Table \ref{tab:SCFglo}).
+To allow for comparison, we aggregate our stock change factors weighted by grid-level \(F^{\mathrm{SCF}}\) cropland area to derive average factors for the four IPCC climate zones (Fig.
 \ref{fig:CLIMzone}).
@@
 \centering
-\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 \citep{eggleston_ipcc_2006} to the update in 2019 \citep{calvo_buendia_ipcc_2019}.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time.}
+\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.} 
 \label{tab:SCFglo}
@@
   6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
-  7 & This Study & hist & 1990 & 0.57 & 0.61 & 0.78 & 0.76 \\ 
-  8 & This Study & hist & 2010 & 0.64 & 0.68 & 0.83 & 0.83 \\ 
+  7 & This Study & hist & 1975 & 0.48 & 0.59 & 0.64 & 0.59 \\ 
+  8 & This Study & hist & 2010 & 0.5 & 0.64 & 0.64 & 0.59 \\ 
    \hline
@@
 
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
-The tropical factors differ substantially between the IPCC guidelines in 2006 \citep{eggleston_ipcc_2006} and their refinement in 2019 \citep{calvo_buendia_ipcc_2019}, with our estimates within this range.
-Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show especially in temperate dry regions --- the smallest region area-wise --- small deviations from natural SOC stocks.
-Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\ref{fig:SOCmaps}(c)) may be expected.
-
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
-
-\hypertarget{soc-stocks-in-line-with-literature}{%
-\subsection{SOC stocks in line with literature}\label{soc-stocks-in-line-with-literature}}
-
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018, schaphoff_lpjml4_2018-1} in Table \ref{fig:SOCglo}).
+Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC.
+For the tropical regions the IPCC factors changed notably from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} to the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
+Our results are in good agreement with the 2006 IPCC factors.
+Modelled \(F^{SCF}\) have increased or stayed constant for all climate zones over time (1975-2010).
+
+\hypertarget{global-soc-stocks-comparison}{%
+\subsubsection{Global SOC stocks comparison}\label{global-soc-stocks-comparison}}
+
+We compare our global SOC stocks with a wide range of global SOC stock estimates for the first 30 cm of the soil profile, using data from WISE \citep{batjes_harmonized_2016}, SoilGrids \citep{hengl_soilgrids250m_2017}, GSOC \citep{fao_global_2018}, LPJmL4 \citep{schaphoff_lpjml4_2018-1}, SoilGrids2 \citep{poggio_soilgrids_2021}, and SOCDebtPaper \citep{sanderman_soil_2017} in Fig.
+\ref{fig:SOCglo}.
 
 \begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
-This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
+\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \citep{schaphoff_lpjml4_2018-1} and estimates from \citep{sanderman_soil_2017}.
+We show values of this study for the year 2010 accounting for the historical land-use dynamics as well as for an hypothetical PNV.}\label{fig:SOCglo}
 \end{figure}
 
-The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
-Looking at regional results in Fig.
-\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
-To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
-
-Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
-For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
-
-We conducted a sensitivity analysis (Fig.
-\ref{fig:SOClitter}) based on various plant parameterizations from the Century model (see Sect.
-\ref{sec:scenlitterpnv}).
-This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.
+The global estimates of the total SOC stock of the upper \(30\,cm\) from this study are in the middle of the wide range of other modeled or observation-based estimates.
+SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, whereas SoilGrid2.0 \citep{poggio_soilgrids_2021} marks the lower end.
+Regional results (Fig.
+\ref{fig:SOCreg}) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.
+
+\hypertarget{point-based-evaluation}{%
+\subsubsection{Point-based evaluation}\label{point-based-evaluation}}
+
+We correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements (for data base see appendix of \citep{sanderman_soil_2017}).
+
+\begin{figure}[h]
+\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010} \caption{Correlation between modeled and measured SOC stocks.
+Given the wide span between minimun and maximum measured SOC stocks within in a given cell, we correlated median values with our modeled results.
+Both cropland and areas with natural vegetation tend to be lower in our results than in the point measurements.}\label{fig:SOCpoint}
+\end{figure}
+
+\hypertarget{natural-soc-stock-comparison-with-lpjml4}{%
+\subsubsection{Natural SOC stock comparison with LPJmL4}\label{natural-soc-stock-comparison-with-lpjml4}}
+
+Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time.
+We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation.
+Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.
+With our focus on cropland SOC dynamics, we compare only cells with more than \(1000\unit{ha}\) of cropland (capturing 99.9\% of global cropland area).
+
+\begin{figure}[h]
+\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv} \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010.
+The grey lines indicate the 1:1 line.}\label{fig:SOCLPJmL4}
+\end{figure}
+
+Spatial correlations of PNV SOC stock values are high (global \(R2=0.81\)), especially for dry climate zones (Fig.
+\ref{fig:SOCLPJmL4}).
+For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results.
+
+\hypertarget{actual-soc-stock-comparison-with-soilgrids-2.0}{%
+\subsubsection{Actual SOC stock comparison with SoilGrids 2.0}\label{actual-soc-stock-comparison-with-soilgrids-2.0}}
+
+SoilGrids 2.0 \citep{poggio_soilgrids_2021} is a digital soil mapping approach that uses over 240 000 soil profile observations to produce high resolution soil maps including SOC stocks and estimates of their uncertainties.
+To evaluate the performance of our model at the global scale, we correlate SoilGrids 2.0 SOC stock values, which were aggregated to 0.5 degree resolution, to our estimates for the year 2010 in Fig.
+\ref{fig:SOCSoilGrids2}.
+To focus our comparison on cropland areas, we mask out grid cells with less than \(1000\unit{ha}\) of cropland.
+Spatial correlation is moderate for tropical climate zones, whereas it is low for temperate moist areas.
+In tropical dry and temperate dry areas, we simulate also very low SOC values (below \$10\unit{tC\,ha^{-1}}), which is not found in SoilGrids 2.0 whereas our modeled SOC stocks can be substantially higher in temperate moist areas than reported by SoilGrids 2.0.
+Additionally, we use the uncertainty estimates from SoilGrids 2.0 in Fig.
+\ref{fig:SOCSoilGrids2Q} to identify areas, where our modeled SOC stocks that are below the 5th or above the 95th percentile of the SoilGrids 2.0 data.
+For the vast majority of grid cells our model results are between the 5th and 95th percentile of SoilGrids 2.0 estimates.
+We underestimate SOC stocks especially in dry areas (e.g.~close to the Sahara).
+Overestimated stocks are often situated in mountainous regions.
+
+\begin{figure}[h]
+\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareSoilGrids2_2010_lu} \caption{Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0.}\label{fig:SOCSoilGrids2}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/SOCquantilesSoilGrids2} \caption{Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0.}\label{fig:SOCSoilGrids2Q}
+\end{figure}
 \newpage
 
-\conclusions
-
-Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
-It is --- to our knowledge --- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
-
-Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models.
-That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT package \citep{dietrich_madrat_2020} constitute a starting point for building comprehensive data sets on agricultural management aspects.
-
-Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions --- not only by the removal of vegetation, but also by a slow depletion of soils.
-Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
-
-However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
-Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
-The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source.
-Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO\textsubscript{2}.
-The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.
-
-Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
-Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand \citep{popp_land-use_2014-1} --- and thereby usually further reduce the SOC debt on croplands.
-
-While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets.
-Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
-
-Yet, there is ample potential for further improved SOC management.
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+We have (1) developed a reduced-complexity model and (2) compiled a spatially explicit time series data set of agricultural management data in order to (3) analyze the role of agricultural management in historical cropland SOC dynamics.
+Our study shows that information on agricultural management alters estimates of the SOC debt and slows down loss of SOC compared to the often used constant management assumptions.
+
+It is important to evaluate the validity of our results as modeling management effects on SOC dynamics at the global scale comes with large uncertainties.
+The model includes a large number of parameters, and for most of these the uncertainty distributions have not been quantified so far.
+Moreover, we think that beyond parameter uncertainty, the structural uncertainty from the model design is high.
+The management data itself is prone to uncertainties as well, as most of it is only indirectly calculated from reported data.
+
+In the following, we give a qualitative assessment of the uncertainties and limitations of this study as well as discuss our three study objectives and results against existing literature.
+
+\hypertarget{comprehensive-historical-agricultural-management-data-set}{%
+\subsection{Comprehensive historical agricultural management data set}\label{comprehensive-historical-agricultural-management-data-set}}
+
+Our spatially explicit time series dataset of agricultural management
+is based on country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} as well as 0.5 degree land-use data from LUH2 \citep{hurtt_harmonization_2020}.
+Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the agricultural system (\ref{fig:FlowFig}) using a mass balance approach from the IPCC guidelines Vol.
+4 \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and other auxiliary data sets (e.g.~\citep{porwollik_generating_2019}).
+
+For some of the aspects covered in our data set, for example livestock distribution \citep{robinson_mapping_2014} or manure production and application \citep{zhang_global_2017}, well-compiled data sets in high resolution exist that capture real world conditions much better than our estimates.
+However, they often come with the caveats of either being static in time, demanding large sets of auxiliary data or being inconsistent with each other.
+
+For most of the parameters used in our management estimates no uncertainty estimates exist.
+This is why, in our view, most of the uncertainty with respect to the impacts of SOC management is included in the management data itself, and especially in the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data \citep{faostat_faostat_2016, weindl_livestock_2017}.
+The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
+Previous modeling studies of SOC carbon on cropland often only used stylized scenarios of management practices \citep{pugh_simulated_2015, lutz_simulating_2019}, rather than trying to estimate real management.
+
+While our dataset, by including crop residues and manure, likely the largest carbon inputs to soils, it does not account for a list of minor carbon inputs from cover crops, agroforestry, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
+Including these sources would correct our estimates upwards and bring our estimates closer to the IPCC stock change factors(see Sect.
+\ref{sec:ipcccompare}).
+Unfortunately, data on the quantity of these inputs is very scarce and does not exist with global coverage.
+
+SOC inputs from above-ground residues had the strongest management effect on SOC debt dynamics on cropland (see Fig.
+\ref{fig:SOCscen}).
+As pointed out by Keel et al.~\citeyearpar{keel_large_2017} and Smith et al.~\citeyearpar{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
+Keel et al.~\citeyearpar{keel_large_2017} question whether below-ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass, in particular for below-ground biomass.
+
+\hypertarget{reduced-complexity-soc-model}{%
+\subsection{Reduced complexity SOC model}\label{reduced-complexity-soc-model}}
+
+Our reduced-complexity SOC model is based on a Tier 2 modeling approach.
+This reduces the computational and data demand of the model in comparison to DGVMs, while still allowing for the explicit representation of agricultural management practices.
+Along with the effects of various C inputs, the impacts of water supply from rainfall and irrigation as well as tillage systems can also be accounted for in the computation of SOC decay rates.
+As such, the model can reflect the spatial and temporal heterogeneity in both management and biophysical conditions.
+
+The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Table \ref{tab:SCFglo}) as well as by the time trend of the SOC debt (see Fig.
+\ref{fig:SOCmaps}(a)).
+Residue management has changed over the last decades, especially with the phasing out of residue burning practices in several regions and increased general productivity, showing a clear impact on SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
+
+The Tier 2 approach \citep{ogle_cropland_in_ipcc_2019} used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV.
+This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is a main driver of SOC dynamics (cite).
+The comparison of simulated PNV data with LPJmL4 shows the model's substantial capability in reproducing PNV SOC stocks (Fig.
+\ref{fig:SOCLPJmL4}).
+
+Using litterfall estimates from LPJmL4, we have been able to estimate the total SOC stocks of the world, which is dominated by SOC under natural vegetation.
+However, as the world's SOC stock is highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018, schaphoff_lpjml4_2018-1, poggio_soilgrids_2021, sanderman_soil_2017} in Fig.
+\ref{fig:SOCglo}, the only conclusion we can draw from this is that our result is within a plausible range.
+To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management drop out in our calculation of SOC debt and stock change factors.
+Natural SOC estimates only influence results when natural land is converted to cropland.
+
+Comparing the geographic SOC patterns to Soil Grids 2.0 {[}\citet{poggio_soilgrids_2021}; see Fig.
+\ref{fig:SOCSoilGrids2Q}{]}, we find that our model estimates values of SOC stocks greater than the estimated confidence intervals in Soil Grid 2.0s for some mountainous regions across the globe.
+This could indicate that we are not capable in capturing specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils \citep{borrelli_assessment_2017}).
+A large swathe of eastern North America was heavily affected by the dust bowl event, with wind erosion removing large parts of the topsoils, a process not considered in our model.
+Similarly, we likely overestimate SOC stocks for the loess soils in northern China and the Altiplano in Latin America; in both cases erosion is a likely reason.
+In contrast, we estimate lower SOC stocks at the edges of the Sahara, where uncertain local water availability and artifical irrigation may dominate spatial SOC patterns.
+
+In our model, erosion should however only affect the spatial pattern but not the aggregate SOC pool.
+As pointed out by Doetterl et al.~\citeyearpar{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions \citep{wang_human-induced_2017}.
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focused especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
+Erosion and degradation impacts on yields and therefore on soil C inputs are captured by our method as we base them on FAO statistics of actual production.
+Yet the distribution of production below the country level - which we allocate proportional to LPJmL production potentials that do not reflect erosion feedback to yields - will overestimate yields and therefore biomass inputs to eroded soils.
+
+In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Table \ref{tab:SCFglo}) for almost all regions.
+
+Tropical soils might suffer from low C input rates due to large yield gaps \citep{global_yieldgap_2022} and high shares of residue removal and burning in lower-income countries {[}\citet{smil_crop_1999}; williams\_influence\_1997; jain\_emission\_2014{]}.
+Yet, even when comparing our etimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values, while it shows very good agreement with the older default values from IPCC \citeyearpar{lasco_cropland_in_ipcc_2006}.
+Don et al.~\citeyearpar{don_impact_2011} estimated SOC losses for tropical soils of around 25\% on average corresponding to a stock change factor of 0.75, but also reported a wide range of measured SOC changes from -80\% to +58\%.
+Fujisaki et al.~\citeyearpar{fujisaki_forest_2015} however found much lower loss rates of around 9\%, attributing the difference to the different time period length since the conversion to cropland.
+As our results do not specifically account for cropland age and most of the cropland is older than 20 years (as assumed for the default IPCC Tier 1 stock change factors) our stock change factors have to be lower by definition following the steady-state assumption that cropland will continue to approach a new equilibrium.
+For the same reason, our estimates for temperate regions might be lower than both IPCC \citeyearpar{lasco_cropland_in_ipcc_2006} and IPCC \citeyearpar{ogle_cropland_in_ipcc_2019} default values.
+With the production-increasing impact of irrigation and fertilization on carbon-poor dryland soils, SOC under cropland can also be higher than under PNV with stock change factors above 1 (see Fig.
+\ref{fig:SOCmaps}(d)), but these areas are much smaller than where the stock change factors are well below unity.
+
+Generally, limiting the analysis to the first 30 cm of the soil profile follows the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and assumes that most of the SOC dynamics happen in the topsoil.
+In this regard several aspects are strongly simplified within our approach.
+Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
+This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
+Second, changes to the subsoil due to tillage are neglected.
+As Powlson et al.~\citeyearpar{powlson_limited_2014} have shown, the subsoil can be make a large difference in evaluating total SOC losses or gains for no-tillage systems.
+No-tillage effects may seem larger than they actually are if only topsoil is considered.
+SOC transfers to deeper soil layers under tillage might enhance subsoil SOC compared to no-till practices.
+Finally, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
+
+\hypertarget{soc-debt-and-soc-drivers}{%
+\subsection{SOC debt and SOC drivers}\label{soc-debt-and-soc-drivers}}
+
+The analysis of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
+There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period \citep[e.g.~][]{frielingstein_glob_carbon_2020}.
+There is uncertainty with respect to the overall size of these emissions from different methods and reference points and with respect to the contribution of cropland and agricultural management to these emissions.
+In order to mitigate greenhouse gas emissions, it is essential to stop the decline of SOC stocks or even transform cropland management to sequester atmospheric C in cropland soils \citep{minasny_soil_2017}.
+Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions \citep[see][]{pugh_simulated_2015}, we find that global cropland has acted as a emissions source since 1975.
+Annual C loss rates of 0.2 per 1000 C still have the opposite trend as the promoted 4 per 1000 C sequestration rate target \citep{minasny_soil_2017}.
+Dedicated efforts to increase cropland SOC are thus necessary, as management improvements at historical rates are not enough to counteract ongoing SOC degradation on cropland.
+Yet our study also shows the substantial impact of changing management on the development of SOC debt (Fig.
+\ref{fig:SOCscen}).
+
+According to Sanderman et al.~\citeyearpar{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
+Our estimate of 39.6 GtC in 2010 for cropland debt is thus twice as high as their estimate.
+However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.~\citeyearpar{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
+
+Furthermore, Sanderman et al.~\citeyearpar{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+Pugh et al.~\citeyearpar{pugh_simulated_2015} considered management effects like tillage and incorporation of residues in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
+Their study moreover concludes that yield gains (by 18\% in their simulations) do not lead to a substantial decline in SOC debt (less than 1\% change).
+Historical yield increases, however, are often estimated to be well above 50\% \citep{pellegrini_crop_2018, ray_recent_2012, rudel_agricultural_2009}.
+While we find substantially larger effects of productivity gains than the 1\% reported by Pugh et al.~\citeyearpar{pugh_simulated_2015}, this is not sufficient to compensate SOC losses from moderate global cropland expansion of around 11\% between 1974 and 2010.
+
+The effects of agricultural productivity on cropland SOC dynamics, including historical yield trends and associated increases in residue inputs, can be directly accounted for in our modeling approach.
+In contrast, process-based studies \citep{pugh_simulated_2015, herzfeld_soc_2021} often lack data on relevant management aspects that drive production increases.
+\citep{herzfeld_soc_2021} also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
+Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
+This is consistent with our finding that increasing SOC inputs from above-ground residues had the strongest effect on the slowing-down of the SOC debt increase (Fig.
+\ref{fig:SOCscen}).
+
+Elliott et al.~\citeyearpar{elliott_management_trends_2018} show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, and genetic traits such as kernel number and radiation use efficiency.
+As such, it will remain challenging for process-based DGVMs to capture the trend of agricultural productivity on cropland SOC dynamics.
+
+Our study emphasizes again that the expansion of cropland is still a major source of CO\textsubscript{2} emissions --- not only through the removal of vegetation, but also by a slow depletion of C stocks in soils.
+Our estimates indicate a SOC debt of 39.6 GtC in 2010, and every additional deforestated hectare adds to this debt.
+Avoided deforestation and other environmental regulation leads to intensification on existing cropland \citep{humpenoder_large-scale_2018} and our results show that such intensification could lead to increased cropland SOC, if residues are returned to the soil --- amplifying the C sequestration potential of avoided deforestation.
+
+There is also ample potential for further improved SOC management.
 As shown in Fig.
 \ref{fig:FlowFig}, approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC).
+However, even larger losses occur at the end of the food supply chain (1.2 GtC year), at the soil surface (1.4 GtC), during residue burning (0.3 Gt C) and with manure management (0.2).
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils \citep{brenzinger_organic_2018}.
-Secondly, soil carbon sequestration techniques \citep{smith_soil_2016}, deep ploughing \citep{alcantara_deep_2016} or the transformation of c inputs to more recalcitrant biochar \citep{woolf_sustainable_2010} may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Waste composting or excreta recycling could represent a major additional C input to cropland soils \citep{brenzinger_organic_2018}.
+Secondly, soil carbon sequestration techniques \citep{smith_soil_2016}, deep ploughing \citep{alcantara_deep_2016} or the transformation of C inputs to more recalcitrant biochar \citep{woolf_sustainable_2010} may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
 Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
-Finally, other carbon-accumulating practices, such as the cultivation of cover crops \citep{poeplau_carbon_2015} and agroforestry \citep{lorenz_soil_2014} could increase total C sequestration on croplands.
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops \citep{poeplau_carbon_2015, porwollik_covercrops_2022} and agroforestry \citep{lorenz_soil_2014} could increase total C sequestration on cropland.
+\newpage
+
+\conclusions
+
+We have compiled a spatially explicit and time-variant data set on agricultural management aspects relevant for cropland SOC dynamics.
+We have also developed a reduced-complexity SOC model that is able to be applied in optimization-based IAM frameworks, for which detailed process-based models are computationally too expensive.
+Making use of these data and model, we are able to estimate spatially explicit SOC stocks, SOC debts, and stock change factors considering agricultural management.
+It is --- to our knowledge --- the first study that analyzes the role of time-variant and spatially explicit historical agricultural management for global SOC dynamics.
+
+Our results demonstrate that historical changes in agricultural management have shaped the SOC debt on cropland.
+It is thus necessary to explicitly consider agricultural management in a dynamic manner in global carbon assessments and models, especially when exploring climate mitigation pathways with so-called land-based solutions \citep[e.g.~][]{popp_land-use_2016}.
+That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
+Our open-source model \citep{karstens_mrsoil_2020}, published data-set \citep{karstens_model_2020} and the flexible data processing with the MADRaT package \citep{dietrich_madrat_2020} constitute a starting point for building comprehensive data sets on agricultural management aspects.
+
+With the reduced-complexity SOC model we are able to account for agricultural management effects on cropland SOC dynamics within optimization-based IAM frameworks.
+Reduced input data requirements such as accounting for changes in productivity rather than reproducing the processes that lead to such changes in productivity \citep{elliott_management_trends_2018} will help to explore the role of agricultural management in sustainable development pathway analyses \citep{sorgel_sustainable_2021}.
+However, we clearly see that increases in agricultural productivity are not sufficient to create positive net SOC sequestration in cropland soils.
+More management options that explicitly target the sequestration of C in cropland soils need to be considered.
+Our open-source model can be expanded to account for additional management options for carbon farming, such as cover crops, agroforestry, or biochar applications.
 
@@
 
+\begin{table}[htbp]
+  \centering
+  \caption{Mapping of FAO crop production items \citep{faostat_faostat_2016} to LUH2 categories \citep{hurtt_harmonization_2020} and to crop groups used in this study.}
+  \begin{tabular}{lllll}
+  \tophline
+      LUH short name & FAO Production Item & Crop group – madrat & Short name – madrat \\ 
+  \middlehline    
+      c3ann & 122|Sweet potatoes & Tropical roots & cassav\_sp \\ 
+      c3ann & 125|Cassava & Tropical roots & cassav\_sp \\ 
+      c3ann & 135|Yautia (cocoyam) & Tropical roots & cassav\_sp \\ 
+      c3ann & 136|Taro (cocoyam) & Tropical roots & cassav\_sp \\ 
+      c3ann & 137|Yams & Tropical roots & cassav\_sp \\ 
+      c3ann & "149|Roots and tubers, nes" & Tropical roots & cassav\_sp \\ 
+      c3per & 486|Bananas & Tropical roots & cassav\_sp \\ 
+      c3per & 489|Plantains & Tropical roots & cassav\_sp \\ 
+      c3ann & 328|Seed cotton & Cotton seed & cottn\_pro \\ 
+      c3ann & "638|Forage and silage, rye grass" & Forage & foddr \\ 
+      c3ann & "639|Forage and silage, grasses nes" & Forage & foddr \\ 
+      c3ann & "642|Forage and silage, green oilseeds" & Forage & foddr \\ 
+      c3ann & 644|Cabbage for fodder & Forage & foddr \\ 
+      c3ann & 645|Pumpkins for Fodder & Forage & foddr \\ 
+      c3ann & 646|Turnips for fodder & Forage & foddr \\ 
+      c3ann & 647|Beets for fodder & Forage & foddr \\ 
+      c3ann & 648|Carrots for fodder & Forage & foddr \\ 
+      c3ann & 649|Swedes for fodder & Forage & foddr \\ 
+      c3ann & 651|Forage products & Forage & foddr \\ 
+      c3ann & 655|Vegetables and roots fodder & Forage & foddr \\ 
+      c4ann & "636|Forage and silage, maize" & Forage & foddr \\ 
+      c4ann & "637|Forage and silage, sorghum" & Forage & foddr \\ 
+      c3nfx & "640|Forage and silage, clover" & Forage & foddr \\ 
+      c3nfx & "641|Forage and silage, alfalfa" & Forage & foddr \\ 
+      c3nfx & "643|Forage and silage, legumes" & Forage & foddr \\ 
+      c3nfx & "242|Groundnuts, with shell" & Groundnuts & groundnut \\ 
+      c4ann & 56|Maize & Maize & maiz \\ 
+      c3per & "254|Oil, palm fruit" & Oilpalms & oilpalm \\ 
+      c3per & 254|Oil palm fruit & Oilpalms & oilpalm \\ 
+      c3ann & 358|Cabbages and other brassicas & Fruits Vegetables Nuts & others \\ 
+      c3ann & 366|Artichokes & Fruits Vegetables Nuts & others \\ 
+      c3ann & 367|Asparagus & Fruits Vegetables Nuts & others \\ 
+      c3ann & 372|Lettuce and chicory & Fruits Vegetables Nuts & others \\ 
+      c3ann & 373|Spinach & Fruits Vegetables Nuts & others \\ 
+      c3ann & 378|Cassava leaves & Fruits Vegetables Nuts & others \\ 
+      c3ann & 388|Tomatoes & Fruits Vegetables Nuts & others \\ 
+      c3ann & 393|Cauliflowers and broccoli & Fruits Vegetables Nuts & others \\ 
+      c3ann & "394|Pumpkins, squash and gourds" & Fruits Vegetables Nuts & others \\ 
+      c3ann & 397|Cucumbers and gherkins & Fruits Vegetables Nuts & others \\ 
+      c3ann & 399|Eggplants (aubergines) & Fruits Vegetables Nuts & others \\ 
+      c3ann & "401|Chillies and peppers, green" & Fruits Vegetables Nuts & others \\ 
+      c3ann & "402|Onions, shallots, green" & Fruits Vegetables Nuts & others \\ 
+      c3ann & "403|Onions, dry" & Fruits Vegetables Nuts & others \\ 
+      c3ann & 406|Garlic & Fruits Vegetables Nuts & others \\ 
+      c3ann & "407|Leeks, other alliaceous vegetables" & Fruits Vegetables Nuts & others \\ 
+      c3ann & 426|Carrots and turnips & Fruits Vegetables Nuts & others \\ 
+      c3ann & 430|Okra & Fruits Vegetables Nuts & others \\ 
+      c3ann & 449|Mushrooms and truffles & Fruits Vegetables Nuts & others \\ 
+      c3ann & 459|Chicory roots & Fruits Vegetables Nuts & others \\ 
+      c3ann & "463|Vegetables, fresh nes" & Fruits Vegetables Nuts & others \\ 
+      c3ann & 544|Strawberries & Fruits Vegetables Nuts & others \\ 
+      c3ann & 567|Watermelons & Fruits Vegetables Nuts & others \\ 
+      c3ann & "568|Melons, other (inccantaloupes)" & Fruits Vegetables Nuts & others \\ 
+      c3ann & "689|Chillies and peppers, dry" & Fruits Vegetables Nuts & others \\ 
+      c3ann & "711|Anise, badian, fennel, coriander" & Fruits Vegetables Nuts & others \\ 
+      c3ann & 720|Ginger & Fruits Vegetables Nuts & others \\ 
+      c3ann & "723|Spices, nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & "216|Brazil nuts, with shell" & Fruits Vegetables Nuts & others \\ 
+      c3per & "217|Cashew nuts, with shell" & Fruits Vegetables Nuts & others \\ 
+      c3per & 220|Chestnut & Fruits Vegetables Nuts & others \\ 
+      c3per & "221|Almonds, with shell" & Fruits Vegetables Nuts & others \\ 
+      c3per & "222|Walnuts, with shell" & Fruits Vegetables Nuts & others \\ 
+      c3per & 223|Pistachios & Fruits Vegetables Nuts & others \\ 
+      c3per & 224|Kola nuts & Fruits Vegetables Nuts & others \\ 
+      c3per & "225|Hazelnuts, with shell" & Fruits Vegetables Nuts & others \\ 
+      c3per & 226|Areca nuts & Fruits Vegetables Nuts & others \\ 
+      c3per & "234|Nuts, nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & 461|Carobs & Fruits Vegetables Nuts & others \\ 
+      c3per & 490|Oranges & Fruits Vegetables Nuts & others \\ 
+      c3per & "495|Tangerines, mandarins, clementines, satsumas" & Fruits Vegetables Nuts & others \\ 
+      c3per & 497|Lemons and limes & Fruits Vegetables Nuts & others \\ 
+      c3per & 507|Grapefruit (inc pomelos) & Fruits Vegetables Nuts & others \\ 
+      c3per & "512|Fruit, citrus nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & 515|Apples & Fruits Vegetables Nuts & others \\ 
+      c3per & 521|Pears & Fruits Vegetables Nuts & others \\ 
+      c3per & 523|Quinces & Fruits Vegetables Nuts & others \\ 
+      c3per & 526|Apricots & Fruits Vegetables Nuts & others \\ 
+      c3per & "530|Cherries, sour" & Fruits Vegetables Nuts & others \\ 
+      c3per & 531|Cherries & Fruits Vegetables Nuts & others \\ 
+      c3per & 534|Peaches and nectarines & Fruits Vegetables Nuts & others \\ 
+      c3per & 536|Plums and sloes & Fruits Vegetables Nuts & others \\ 
+      c3per & "541|Fruit, stone nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & "542|Fruit, pome nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & 547|Raspberries & Fruits Vegetables Nuts & others \\ 
+      c3per & 549|Gooseberries & Fruits Vegetables Nuts & others \\ 
+      c3per & 550|Currants & Fruits Vegetables Nuts & others \\ 
+      c3per & 552|Blueberries & Fruits Vegetables Nuts & others \\ 
+      c3per & 554|Cranberries & Fruits Vegetables Nuts & others \\ 
+      c3per & 558|Berries nes & Fruits Vegetables Nuts & others \\ 
+      c3per & 560|Grapes & Fruits Vegetables Nuts & others \\ 
+      c3per & 569|Figs & Fruits Vegetables Nuts & others \\ 
+      c3per & "571|Mangoes, mangosteens, guavas" & Fruits Vegetables Nuts & others \\ 
+      c3per & 572|Avocados & Fruits Vegetables Nuts & others \\ 
+      c3per & 574|Pineapples & Fruits Vegetables Nuts & others \\ 
+      c3per & 577|Dates & Fruits Vegetables Nuts & others \\ 
+      c3per & 587|Persimmons & Fruits Vegetables Nuts & others \\ 
+      c3per & 591|Cashewapple & Fruits Vegetables Nuts & others \\ 
+      c3per & 592|Kiwi fruit & Fruits Vegetables Nuts & others \\ 
+      c3per & 600|Papayas & Fruits Vegetables Nuts & others \\ 
+      c3per & "603|Fruit, tropical fresh nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & "619|Fruit, fresh nes" & Fruits Vegetables Nuts & others \\ 
+      c3per & "656|Coffee, green" & Fruits Vegetables Nuts & others \\ 
+      c3per & "661|Cocoa, beans" & Fruits Vegetables Nuts & others \\ 
+      c3per & 667|Tea & Fruits Vegetables Nuts & others \\ 
+      c3per & 671|Mate & Fruits Vegetables Nuts & others \\ 
+      c3per & 687|Pepper (piper spp) & Fruits Vegetables Nuts & others \\ 
+      c3per & 692|Vanilla & Fruits Vegetables Nuts & others \\ 
+      c3per & 693|Cinnamon (canella) & Fruits Vegetables Nuts & others \\ 
+      c3per & 698|Cloves & Fruits Vegetables Nuts & others \\ 
+      c3per & "702|Nutmeg, mace and cardamoms" & Fruits Vegetables Nuts & others \\ 
+      c4ann & "446|Maize, green" & Fruits Vegetables Nuts & others \\ 
+      c3nfx & "414|Beans, green" & Fruits Vegetables Nuts & others \\ 
+      c3nfx & "417|Peas, green" & Fruits Vegetables Nuts & others \\ 
+      c3nfx & "420|Vegetables, leguminous nes" & Fruits Vegetables Nuts & others \\ 
+      c3nfx & 423|String beans & Fruits Vegetables Nuts & others \\ 
+      c3ann & 677|Hops & Fruits Vegetables Nuts & others \\ 
+      c3ann & 748|Peppermint & Fruits Vegetables Nuts & others \\ 
+      c3ann & 116|Potatoes & Potatoes & potato \\ 
+      c3nfx & "176|Beans, dry" & Pulses & puls\_pro \\ 
+      c3nfx & "181|Broad beans, horse beans, dry" & Pulses & puls\_pro \\ 
+      c3nfx & "187|Peas, dry" & Pulses & puls\_pro \\ 
+      c3nfx & 191|Chick peas & Pulses & puls\_pro \\ 
+      c3nfx & "195|Cow peas, dry" & Pulses & puls\_pro \\ 
+      c3nfx & 197|Pigeon peas & Pulses & puls\_pro \\ 
+      c3nfx & 201|Lentils & Pulses & puls\_pro \\ 
+      c3nfx & 203|Bambara beans & Pulses & puls\_pro \\ 
+      c3nfx & 205|Vetches & Pulses & puls\_pro \\ 
+      c3nfx & 210|Lupins & Pulses & puls\_pro \\ 
+      c3nfx & "211|Pulses, nes" & Pulses & puls\_pro \\ 
+      c3ann & 265|Castor oil seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 270|Rapeseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 280|Safflower seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 289|Sesame seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 292|Mustard seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 296|Poppy seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 299|Melonseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 333|Linseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & 336|Hempseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 249|Coconuts & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 260|Olives & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 263|Karite nuts (sheanuts) & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 275|Tung nuts & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 277|Jojoba seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 305|Tallowtree seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 310|Kapok fruit & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 311|Kapokseed in shell & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3per & 339|Oilseeds nes & Other oil crops (incl rapeseed) & rapeseed \\ 
+      c3ann & "27|Rice, paddy" & Rice & rice\_pro \\ 
+      c3nfx & 236|Soybeans & Soybean & soybean \\ 
+      c3ann & 157|Sugar beet & Sugar beet & sugr\_beet \\ 
+      c4per & 156|Sugar cane & Sugar cane & sugr\_cane \\ 
+      c3ann & 267|Sunflower seed & Sunflower & sunflower \\ 
+      c3ann & 15|Wheat & Temperate cereals & tece \\ 
+      c3ann & 44|Barley & Temperate cereals & tece \\ 
+      c3ann & 71|Rye & Temperate cereals & tece \\ 
+      c3ann & 75|Oats & Temperate cereals & tece \\ 
+      c3ann & 89|Buckwheat & Temperate cereals & tece \\ 
+      c3ann & 92|Quinoa & Temperate cereals & tece \\ 
+      c3ann & 94|Fonio & Temperate cereals & tece \\ 
+      c3ann & 97|Triticale & Temperate cereals & tece \\ 
+      c3ann & 101|Canary seed & Temperate cereals & tece \\ 
+      c3ann & "103|Grain, mixed" & Temperate cereals & tece \\ 
+      c3ann & "108|Cereals, nes" & Temperate cereals & tece \\ 
+      c4ann & 68|Popcorn & Temperate cereals & tece \\ 
+      c4ann & 79|Millet & Tropical cereals & trce \\ 
+      c4ann & 83|Sorghum & Tropical cereals & trce \\ 
+      \bottomhline
+  \end{tabular}
+  \label{tab:fao2luh2mag}
+\end{table}
+
 \subsection{Discussion}
@@
 \label{fig:SOCreg}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \citealt{century_model_2000}).
-\textit{LitterPNV-CenturyAverage} forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
-Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
-\label{fig:SOClitter}
 \end{figure}

### 2022-02-28 SOCbudgetPaper2/SOCbudget.Rmd
Major update of finalized Paper - Draft 2

--- 2020-12-17 SOCbudgetPaper/SOCbudget.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
-
-%% copernicus_rticles_template (flag for rticles template detection - do not remove!)
-
-%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
-
-%% 2-column papers and discussion papers
-\documentclass[gc, manuscript]{copernicus}
-
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
-
-% Advances in Geosciences (adgeo)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
-% Atmospheric Chemistry and Physics (acp)
-% Atmospheric Measurement Techniques (amt)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% DEUQUA Special Publications (deuquasp)
-% Drinking Water Engineering and Science (dwes)
-% Earth Surface Dynamics (esurf)
-% Earth System Dynamics (esd)
-% Earth System Science Data (essd)
-% E&G Quaternary Science Journal (egqsj)
-% European Journal of Mineralogy (ejm)
-% Fossil Record (fr)
-% Geochronology (gchron)
-% Geographica Helvetica (gh)
-% Geoscience Communication (gc)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% History of Geo- and Space Sciences (hgss)
-% Hydrology and Earth System Sciences (hess)
-% Journal of Bone and Joint Infection (jbji)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Magnetic Resonance (mr)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Primate Biology (pb)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Weather and Climate Dynamics (wcd)
-% Web Ecology (we)
-% Wind Energy Science (wes)
-
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
-
-% Pandoc citation processing
-
-% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
-%
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-\begin{document}
-
-\title{Management induced changes of soil organic carbon on global croplands}
-
-\Author[1, 3]{Kristine}{Karstens}
-\Author[1]{Benjamin Leon}{Bodirsky}
-\Author[1]{Jan Philipp}{Dietrich}
-\Author[2]{Marta}{Dondini}
-\Author[1]{Jens}{Heinke}
-\Author[2]{Matthias}{Kuhnert}
-\Author[1]{Christoph}{Müller}
-\Author[1]{Susanne}{Rolinski}
-\Author[2]{Pete}{Smith}
-\Author[1]{Isabelle}{Weindl}
-\Author[1, 3]{Hermann}{Lotze-Campen}
-\Author[1]{Alexander}{Popp}
-
-\affil[1]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany}
-\affil[2]{Institute of Biological \& Environmental Sciences, University of Aberdeen, Aberdeen, UK}
-\affil[3]{Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany}
-
-\runningtitle{R Markdown Template for Copernicus}
-
-\runningauthor{Nüst et al.}
-
-\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
-
-\maketitle
-
-\begin{abstract}
+---
+title: Management induced changes of soil organic carbon on global croplands 
+journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
+author:
+  - given_name: Kristine
+    surname: Karstens
+    affiliation: "1, 3"
+    email: kristine.karstens@pik-potsdam.de
+    orcid: https://orcid.org/0000-0002-5860-9789
+    corresponding: true
+  - given_name: Benjamin Leon
+    surname: Bodirsky
+    affiliation: 1
+    email: bodirsky@pik-potsdam.de
+    orcid: https://orcid.org/0000-0002-8242-6712
+  - given_name: Jan Philipp
+    surname: Dietrich
+    affiliation: 1 
+    orcid: https://orcid.org/0000-0002-4309-6431
+  - given_name: Marta
+    surname: Dondini
+    affiliation: 2
+  - given_name: Jens
+    surname: Heinke
+    affiliation: 1
+  - given_name: Matthias
+    surname: Kuhnert
+    affiliation: 2
+  - given_name: Christoph
+    surname: Müller
+    affiliation: 1 
+    orcid: https://orcid.org/0000-0002-9491-3550
+  - given_name: Susanne
+    surname: Rolinski
+    affiliation: 1
+  - given_name: Pete
+    surname: Smith
+    affiliation: 2
+    orcid: https://orcid.org/0000-0002-3784-1124
+  - given_name: Isabelle
+    surname: Weindl
+    affiliation: 1
+    orcid: https://orcid.org/0000-0002-7651-6930
+  - given_name: Hermann
+    surname: Lotze-Campen
+    affiliation: "1, 3"
+  - given_name: Alexander
+    surname: Popp
+    affiliation: 1
+affiliation:
+  - code: 1
+    address: Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany
+  - code: 2
+    address: Institute of Biological & Environmental Sciences, University of Aberdeen, Aberdeen, UK
+  - code: 3
+    address: Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany
+abstract: |
   Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land cover change and agricultural management.
+The first meter of the Earth's soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
+SOC has been depleted by anthropogenic land-cover change and agricultural management.
 However, the latter has so far not been well represented in global carbon stock assessments.
 While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
-Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
-Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
-We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
-\end{abstract}
+Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
+We estimate that due to arable farming, soils have lost around 34.6 GtC relative to a counterfactual hypothetical natural state in 1975.
+Within the period 1975--2010 this SOC debt continued to expand by 5 GtC ($0.14\unit{GtC yr^{-1}}$) to around 39.6 GtC.
+However, accounting for historical management led to 2.1 GtC less ($0.06\unit{GtC yr^{-1}}$) emissions than under the assumption of constant management.
+We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique.
+The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling.
+bibliography: SOCbudget.bib
+running:
+  title: R Markdown Template for Copernicus
+  author: Nüst et al.
+competinginterests: |
+  The authors declare no competing interests.
+# OPTIONAL:
+algorithms: true
+# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
+# copyrightstatement: |
+#   The author's copyright for this publication is transferred to Potsdam Institute for Climate Impact Research (PIK).
+availability:
+  codedata: |
+   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \citep{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \citep{karstens_mrsoil_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \citep{bodirsky_mrvalidation_2020} for validation data.
+All libraries are based on the MADRaT package at github.com/pik-piam/madrat \citep{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
+Model results including C input data are accessable under https://doi.org/10.5281/zenodo.4320663 \citep{karstens_model_2020}.
+Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.
+authorcontribution: |
+  KK, BLB and AP designed the study and the model idea.
+KK wrote the code build on work of BLB, IW.
+JPD revised and improved the model code.
+CM, JH and SR provided the LPJmL simulation data.
+KK wrote the paper with important contributions of BLB and CM.
+MK, JS, SR and IW provided extensive feedback to outline of the study.
+All authors discussed the results and commented on the manuscript.
 
-\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
+acknowledgements: |
+  Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
+Additional thanks to the rticles contributors for providing a R Markdown template.
+The authors thank for the data provided by FAOSTAT and LUH2v2.
+The work of KK was funded by the DFG Priority Program “Climate Engineering: Risks, Challenges, Opportunities?” (SPP 1689) and specifically the CEMICS2 project (grant no.
+ED78/3-2).
+The research leading to these results has received funding for BLB from the European Union’s Horizon 2020 research and innovation program under grant agreement no.
+776479 (COACCH) and  no.
+821010 (CASCADES).
+The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF).
+The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no.
+774378).
 
-\newpage
+appendix: |
 
-\introduction
-
-Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}.
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{friedlingstein_global_2019, minasny_soil_2017}.
-The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile \citep{batjes_total_1996, sanderman_soil_2017}.
-
-Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018}.
-However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
-Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
-
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
-Sanderman et al.~\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.
-
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions \citep{friedlingstein_global_2019}.
-While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}.
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management \citep{friedlingstein_global_2019, houghton_carbon_2012, hansis_relevance_2015}.
-
-Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.~][]{bondeau_modelling_2007, lindeskog_implications_2013}.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
-They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
-
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010} multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][\citet{lutz_simulating_2019}]{pugh_simulated_2015}.
-
-More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
-
-Field-scale models \citep[\citet{coleman_simulating_1997}, \citet{smith_estimating_2010}, \citet{taghizadeh-toosi_c-tool_2014}]{del_grosso_simulated_2001} are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
-
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
-To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
-Using a scenario analysis, we decompose the contribution of different management activities.
-\newpage
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-In Sect.
-\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}.
-We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}.
-In Sect.
-\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
-Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices \citep[for model code see][]{bodirsky_mrcommons_2020}.
-In Sect.
-\ref{sec:scenarios} we define the scenarios used to complement our historic model results.
-
-\hypertarget{sec:carbonbudget}{%
-\subsection{SOC stocks and stock changes following the steady-state method}\label{sec:carbonbudget}}
-
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 \citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
-We restrict our analysis to the first 0-30 cm of the soil profile.
-Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
-Therefore, we take the following three steps for each year of our simulation period:
-(1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
-(2) we account for land conversion by transferring \(SOC\) from and to natural vegetation (Sect. \ref{sec:carbontransfer}),
-(3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
-To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \ref{sec:initsoc}).
-
-\hypertarget{sec:steadystates}{%
-\subsubsection{Steady-state SOC stocks and decay rates}\label{sec:steadystates}}
-
-In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{\mathrm{eq}}\) are given by
-\begin{equation}
-SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
-\label{eq:inoutflow}
-\end{equation}
-with \(C^{\textrm{in}}\) being the carbon inputs to the soil, \(k\) denotes the soil organic carbon decay rate.
-This equation is valid for all grid cells \(i\) and all years \(t\).
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model \citep{parton_analysis_1987}.
-Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
-We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
-
-Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
-\ref{sec:residues}) and manure inputs (see Sect.
-\ref{sec:livstmanure}); for natural vegetation litterfall including fine root turnover \citep{schaphoff_lpjml4_2018} is the only source of carbon inflow to the soil.
-Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
-For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients.
-This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \ref{tab:datasourceinputs}.
-
-  \begin{table*}[h]
-  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
-  \tophline
-  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
-  \middlehline
-  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      \citet{faostat_faostat_2016}, \\
-                                                                      \citet{schaphoff_lpjml4_2018},\\
-                                                                      \citet{weindl_livestock_2017} \end{minipage}} & 
-                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      LG:C generic values according to Table 5.5B,\\
-                                                                      5.5C from IPCC \citep{calvo_buendia_ipcc_2019},\\
-                                                                      crop-specific N:C from \citet{bodirsky_n2o_2012}
-                                                                      \end{minipage}} \\
-                            & below-ground residues, &  &  \\
-                            & manure                &  &  \\ 
-                            \hline
-  Natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
-                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                IPCC \citep{calvo_buendia_ipcc_2019} and\\ 
-                                CENTURY (\citep{century_model_2000}) \strut \end{minipage}\tabularnewline
- \bottomhline
- \end{tabular}
- \label{tab:datasourceinputs}
- \belowtable{}
- \end{table*}
-
-The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space (\(i\)) and time (\(t\)).
-Following the steady-state method \citep{calvo_buendia_ipcc_2019}, we consider temperature (\(temp\)), water (\(wat\)), sand fraction (\(sf\)) and tillage (\(till\)) effects to account for spatial variation of decay rates.
-Thus, \(k_{sub}\) are given by
-
-\begin{equation}
-\begin{aligned}
-& k_{i,t,\mathrm{active},lu}  & = &~ k_{\mathrm{active}}  ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ & \cdot~ sf_{i}\\
-& k_{i,t,\mathrm{slow},lu}    & = &~ k_{\mathrm{slow}}    ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ &\\
-& k_{i,t,\mathrm{passive},lu} & = &~ k_{\mathrm{passive}} ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ & ~ & 
-\label{eq:decayrates}
-\end{aligned}.
-\end{equation}
-
-For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect.
-\ref{sec:tillage}) and irrigation (see Sect.
-\ref{sec:irrigation}) practices on decay rates.
-We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
-
- \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
- \begin{tabular}{l l l l}
- \tophline
-  land-use types   & type of decay driver & parameter use to represent driver & data source \\
- \middlehline
- \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
-                                     & \citet{hengl_soilgrids250m_2017} \\
-                      \cline{2-4}
-
-                      & Mircobial activity & air temperature & \citet{harris_version_2020} \\
-                      \cline{2-4}
-
-                      & Water restriction & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
-                      \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
-                      \cline{2-4}
-
-                      & Soil disturbance & tillage & Sect. \ref{sec:irrigation} \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:datasourcedecay}
- \end{table*}
-
-\hypertarget{sec:carbontransfer}{%
-\subsubsection{SOC transfer between land-use types}\label{sec:carbontransfer}}
-
-We calculate \(SOC\) stocks based on the area shares of land-use types (\(lu\)) within our grid cells (\(i\)).
-If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the \(SOC\) is reallocated.
-We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{t^*}\) via
-
-\begin{equation}
-SOC_{i,t^*,sub,lu} = SOC_{i,t-1,sub,lu} - \frac{SOC_{i,t-1,sub,lu}}{A_{i,t-1,lu}} \cdot  AR_{i,t,lu} + \frac{SOC_{i,t-1,sub,!lu}}{A_{i,t-1,!lu}} \cdot  AE_{i,t,lu}
-\label{eq:ctransfer}
-\end{equation}
-
-with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) and \(AE_{lu}\) the area reduction resp.
-area expansion of the two land-use types.
-Data sources and methodology on land-use states and changes are described in Sect.
-\ref{sec:landuse}.
-
-\hypertarget{sec:totalsoc}{%
-\subsubsection{Total SOC stocks and stock changes}\label{sec:totalsoc}}
-
-\(SOC\) converge towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
-
-\begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t^*,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
-\label{eq:SOCstate}
-\end{equation}
-
-Note that the decay rates have to be multiplied by one year (\(1a\)) to form a dimensionless factor.
-Reformulating this equation, we obtain a mass balance equation as follows
-
-\begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} - \underbrace{SOC_{i,t^*,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
-\label{eq:steadystate2budget}
-\end{equation}
-
-The global \(SOC\) stock for each time step \(t\) can then be calculated via
-
-\begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
-\label{eq:totalstock}
-\end{equation}
-
-According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years \citep[see Eq.
-5.0A in][]{calvo_buendia_ipcc_2019}.
-This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
-The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by
-
-\begin{equation}
-\Delta SOC_{i,t} = SOC_{i,t} - SOC^{\mathrm{pnv}}_{i,t}\qquad \text{and} \qquad  F^{\mathrm{SCF}}_{i,t} = \frac{SOC_{i,t}}{SOC^{\mathrm{pnv}}_{i,t}} .
-\label{eq:stockdiff}
-\end{equation}
-
-Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt \citep{sanderman_soil_2017} due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
-Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land-use change emissions \citep{pugh_simulated_2015}.
-
-\hypertarget{sec:initsoc}{%
-\subsubsection{Initialization of SOC pools}\label{sec:initsoc}}
-
-To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
-While land-use, climate and litter input information are available from 1901, agricultural management data on residue and manure inputs are held constant at the level of 1965.
-After a spin-up period of 64 years (1901-1965) with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975 onwards.
-Irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
-
-\hypertarget{sec:tier1}{%
-\subsection{SOC stocks and stock changes following Tier 1}\label{sec:tier1}}
-
-Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019}.
-Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig.
-\ref{fig:CLIMzone}).
-The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by
-
-\begin{equation}
-SOC_{i,t} = \sum_{c,m} T_{c,i} \cdot SOC^{\mathrm{ref}}_{i,t} \cdot F^{\mathrm{SCF}}_{c,m}
-\label{eq:tier1}
-\end{equation}
-
-with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of Eggleston et al.~\citeyearpar{eggleston_ipcc_2006} and Calvo Buendia et al.~\citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
-
-\hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data at 0.5 degree resolution}\label{sec:agrimanagement}}
-
-We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
-
-\hypertarget{sec:landuse}{%
-\subsubsection{Land use and land-use change}\label{sec:landuse}}
-
-Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups \citep[see FAO2LUH2MAG\_croptypes.csv in][]{bodirsky_mrcommons_2020}, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016}.
-By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
-Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
-
-\hypertarget{sec:residues}{%
-\subsubsection{Crop and crop residues production}\label{sec:residues}}
-
-Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see Sect.
-\ref{sec:landuse}).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production \citep{faostat_faostat_2016}.
-By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
-
-Crop residue production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
-Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
-
-\begin{equation}
-\begin{aligned}
-AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
-\qquad & \textrm{and} \\
-BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
-\label{eq:resbiomass}
-\end{aligned}
-\end{equation}
-
-Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
-Deviating from Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
-We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
-
-A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
-Following Smil \citeyearpar{smil_nitrogen_1999} we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.
-
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020}.
-Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017}, that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
-We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above.
-The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field.
-We limit high residue return rates to at most \(10\unit{tC\,ha}^{-1}\) in order to correct for outliers.
-
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{tab:c2dm}).
-
-\hypertarget{sec:livstmanure}{%
-\subsubsection{Livestock distribution and manure excretion}\label{sec:livstmanure}}
-
-Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems at the present day a lot of the manure has to be stored and can be applied on croplands.
-We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
-
-To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of Robinson et al.~\citeyearpar{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from Weindl et al.~\citeyearpar{weindl_livestock_2017}.
-We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-Due to great feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
-We distinguish between grazed pasture, which is converted into livestock products in extensive systems, and primary-crop feed stuff, which we consider to be consumed in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to distinguish between intensive and extensive production systems.
-For low-income countries, we assume only extensive production systems.
-We locate them according to the share of built-up areas based on the assumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
-Intensive production associated with high-income countries, is distributed within a country using the share of primary-crop production, assuming that feed availability is the most determining factor for livestock location.
-For middle-income countries we split the livestock production into extensive and intensive systems based on the per-capita income.
-
-Manure production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
-Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
-Depending on the feed system we assume manure to be handled in four different ways:
-All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
-To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012}.
-We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
-
-\hypertarget{sec:irrigation}{%
-\subsubsection{Irrigation}\label{sec:irrigation}}
-
-The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for their cropland subcategories.
-We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient \(wat\) on decay rates using these shares to compute the weighted mean between rainfed and irrigated \(wat\) factors.
-As a result \(wat\) is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
-This will lead to an overestimation of the irrigation effect.
-We expect, however, water limitations to be a minor problem during the off-season in temperature limited cropping regions, causing our assumption to not dramatically overestimate the moisture effects.
-In tropical, water-limited cropping areas, irrigated growing periods might even span over the whole year.
-
-\hypertarget{sec:tillage}{%
-\subsubsection{Tillage}\label{sec:tillage}}
-
-In order to derive a spatial distribution of the three different tillage types specified by the IPCC --- full tillage, reduced tillage and no tillage ---, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from Porwollik et al.~\citeyearpar{porwollik_generating_2018}.
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO \citeyearpar{fao_aquastat_2016} and half-degree resolution physical crop areas from Hurtt et al.~\citeyearpar{hurtt_harmonization_2020}, applying the methodology of Prowollik et al.~\citeyearpar{porwollik_generating_2018} to identify potential no-tillage grid cells.
-
-\hypertarget{sec:scenarios}{%
-\subsection{Scenario definitions}\label{sec:scenarios}}
-
-To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios \citep[for technical implementation see][]{karstens_mrsoil_2020}.
-
-\hypertarget{sec:scen_management}{%
-\subsubsection{Management scenarios}\label{sec:scen_management}}
-
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for these three drivers: In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set).
-The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in \(\unit{t ha^{-1}}\)) at the level of 1975 onwards.
-Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
-The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
-
-\hypertarget{sec:scen_initial}{%
-\subsubsection{Initialization choice}\label{sec:scen_initial}}
-
-As outlined in Sect.
-\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
-As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
-Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
-
-\hypertarget{sec:scenlitterpnv}{%
-\subsubsection{Parameterization of natural litter fall}\label{sec:scenlitterpnv}}
-
-Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short LG:C) and nitrogen-to-carbon (short N:C) ratios for all land under natural vegetation.
-This is however a strong simplification.
-Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file \citep{century_model_2000} for trees as parameterization were not given.
-Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
-We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \ref{tab:scenlitterpnv}.
-
- \begin{table*}[h]
- \caption{Scenarios for sensitivity analysis on litter parameterization}
- \begin{tabular}{l l l l}
- \tophline
-  Scenario name & plant type & N:C & LG:C \\
-  \middlehline
-  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
-  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
-  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
-  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
-  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
-  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
-  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:scenlitterpnv}
- \end{table*}
-
-\newpage
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens \citeyearpar{karstens_model_2020}.
-In the following, the most important results \citep[see][]{karstens_result_2020} for post-processing script) are summarized.
-
-\hypertarget{soc-distribution-and-depletion}{%
-\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}
-
-\begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
-Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
-}\label{fig:SOCmaps}
-\end{figure}
-
-In Fig.
-\ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
-Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways:
-1.
-Absolute SOC changes \(\Delta SOC\) (see Fig.
-\ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses.
-They might be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
-In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
-2.
-Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig.
-\ref{fig:SOCmaps}(c)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
-Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
-
-The global SOC debt has decreased by about 15\% in the period between 1975 and 2010 to \(22\unit{GtC}\) (Fig.
-\ref{fig:SOCmaps}(d)).
-This corresponds to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\).
-Considering our estimate of the global SOC stock of around \(660\unit{GtC}\) in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
-
-\hypertarget{carbon-flows-in-the-agricultural-system}{%
-\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
-
-\begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
-\end{figure}
-
-C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
-Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
-We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.~like food, bioenergy and material) into a human demand category.
-Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
-Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not recycled to soils.
-Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
-However, around 60\% of C decomposes before its integrated into soils at the the litter-soil barrier.
-Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
-According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
-4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
-Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
-
-\hypertarget{agricultural-management-effects-on-soc-debt}{%
-\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
-
-\begin{figure}[h]
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
-\end{figure}
-
-We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)).
-Without changes in management regimes, the global \(\Delta SOC\) on cropland would be increasing at a rate of \(0.1\unit{GtC yr^{-1}}\).
-As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global \(\Delta SOC\) would still increase at a rate of \(0.06\unit{GtC yr^{-1}}\).
-Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both).
-The effect of no-tillage has been particuarly strong since the 1990s.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \ref{fig:SOCscen}(b).
-
-\begin{figure}[h]
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
-\end{figure}
-
-Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\ref{fig:SOCinit}).
-The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
-
-Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas \(\Delta SOC\) vary 19-24 GtC for the year 2010 (see Fig.
-\ref{fig:SOClitter}.
-The increase of \(\Delta SOC\) during the period of 1975--2010 is independent of the parameterization choice at \textasciitilde4 GtC.
-
-\newpage
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
-This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
-
-\hypertarget{soc-debt-and-soc-drivers-in-literature}{%
-\subsection{SOC debt and SOC drivers in literature}\label{soc-debt-and-soc-drivers-in-literature}}
-
-The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions \citep[see][]{pugh_simulated_2015}, global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 \citep{minasny_soil_2017}, indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
-
-According to Sanderman et al.~\citeyearpar{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as Sanderman et al.~\citeyearpar{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.~\citeyearpar{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
-This suggests that our results are within a plausible range.
-
-Furthermore, Sanderman et al.~\citeyearpar{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
-Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% \citep{pellegrini_crop_2018, ray_recent_2012, rudel_agricultural_2009}.
-
-Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to Pugh et al.~\citeyearpar{pugh_simulated_2015} findings of only small effects due to improved practices.
-
-Looking on historic cropland intensification rates of above 50\% \citep{rudel_agricultural_2009}, a large fraction of increased C input from residue biomass is attributed to productivity improvements.
-
-\hypertarget{limitations-and-uncertainties}{%
-\subsection{Limitations and uncertainties}\label{limitations-and-uncertainties}}
-
-Modeling management effects at the global scale comes with parametric and structural uncertainties.
-High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by Keel et al.~\citeyearpar{keel_large_2017} and Smith et al.~\citeyearpar{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
-Keel et al.~\citeyearpar{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
-However, looking on the evaluation of management effects (see Sect.
-\ref{sec:ipcccompare}), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
-Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.~cover- and inter-cropping practices.
-
-Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
-As shown in Fig.
-\ref{fig:SOCinit} the SOC debt estimate almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), and the SOC debt reduction is strongly reduced (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
-
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and assumes that most of the SOC dynamic happens in the topsoil.
-In this regard several aspects are strongly simplified within our approach.
-Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
-This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected.
-As Powlson et al.~\citeyearpar{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
-SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
-Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
-
-This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by Doetterl et al.~\citeyearpar{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions \citep{wang_human-induced_2017}.
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
-Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well \citep{sanderman_soil_2017, friedlingstein_global_2019}.
-This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
-
-\hypertarget{sec:ipcccompare}{%
-\subsection{Modeled management effect in line with default IPCC assumptions}\label{sec:ipcccompare}}
-
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 \citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}.
-Both estimates are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\ref{fig:CLIMzone}).
-
-\begin{table}[ht]
-\centering
-\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 \citep{eggleston_ipcc_2006} to the update in 2019 \citep{calvo_buendia_ipcc_2019}.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time.}
-\label{tab:SCFglo}
-\begin{tabular}{rlllllll}
-  \hline
- & Source & Input & Year & tropical moist & tropical dry & temperate dry & temperate moist \\ 
-  \hline
-1 & IPCC2006 & low & invariant & 0.44 & 0.55--0.61 & 0.74 & 0.66 \\ 
-  2 & IPCC2006 & medium & invariant & 0.48 & 0.58--0.64 & 0.80 & 0.69 \\ 
-  3 & IPCC2006 & high & invariant & 0.53 & 0.60--0.67 & 0.83 & 0.77 \\ 
-  4 & IPCC2019 & low & invariant & 0.76 & 0.87 & 0.70--0.71 & 0.66--0.67 \\ 
-  5 & IPCC2019 & medium & invariant & 0.83 & 0.92 & 0.76--0.77 & 0.69--0.70 \\ 
-  6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
-  7 & This Study & hist & 1990 & 0.57 & 0.61 & 0.78 & 0.76 \\ 
-  8 & This Study & hist & 2010 & 0.64 & 0.68 & 0.83 & 0.83 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019}.
-The tropical factors differ substantially between the IPCC guidelines in 2006 \citep{eggleston_ipcc_2006} and their refinement in 2019 \citep{calvo_buendia_ipcc_2019}, with our estimates within this range.
-Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show especially in temperate dry regions --- the smallest region area-wise --- small deviations from natural SOC stocks.
-Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\ref{fig:SOCmaps}(c)) may be expected.
-
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
-
-\hypertarget{soc-stocks-in-line-with-literature}{%
-\subsection{SOC stocks in line with literature}\label{soc-stocks-in-line-with-literature}}
-
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile \citep{batjes_harmonized_2016, hengl_soilgrids250m_2017, fao_global_2018, schaphoff_lpjml4_2018-1} in Table \ref{fig:SOCglo}).
-
-\begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
-This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
-\end{figure}
-
-The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
-Looking at regional results in Fig.
-\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
-To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
-
-Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
-For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
-
-We conducted a sensitivity analysis (Fig.
-\ref{fig:SOClitter}) based on various plant parameterizations from the Century model (see Sect.
-\ref{sec:scenlitterpnv}).
-This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.
-\newpage
-
-\conclusions
-
-Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
-It is --- to our knowledge --- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
-
-Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models.
-That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT package \citep{dietrich_madrat_2020} constitute a starting point for building comprehensive data sets on agricultural management aspects.
-
-Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions --- not only by the removal of vegetation, but also by a slow depletion of soils.
-Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
-
-However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
-Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
-The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source.
-Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO\textsubscript{2}.
-The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.
-
-Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
-Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand \citep{popp_land-use_2014-1} --- and thereby usually further reduce the SOC debt on croplands.
-
-While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets.
-Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
-
-Yet, there is ample potential for further improved SOC management.
-As shown in Fig.
-\ref{fig:FlowFig}, approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC).
-Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils \citep{brenzinger_organic_2018}.
-Secondly, soil carbon sequestration techniques \citep{smith_soil_2016}, deep ploughing \citep{alcantara_deep_2016} or the transformation of c inputs to more recalcitrant biochar \citep{woolf_sustainable_2010} may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
-Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
-Finally, other carbon-accumulating practices, such as the cultivation of cover crops \citep{poeplau_carbon_2015} and agroforestry \citep{lorenz_soil_2014} could increase total C sequestration on croplands.
-
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \citep{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \citep{karstens_mrsoil_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \citep{bodirsky_mrvalidation_2020} for validation data.
-All libraries are based on the MADRaT package at github.com/pik-piam/madrat \citep{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results including C input data are accessable under \url{https://doi.org/10.5281/zenodo.4320663} \citep{karstens_model_2020}.
-Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.} %% use this section when having data sets and software code available
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
   \section{Figures and tables in appendices}
@@
 
+    ```{=latex}
   \begin{figure}[h]
@@
   \end{figure}
+  ```
 
+  ```{=latex}
   \begin{sidewaystable*}[htbp]
@@
   \end{tabular}
-\label{tab:c2dm}
+  (\#tab:c2dm)
   \end{sidewaystable*}
+  ```
+
+  ```{=latex}
+
+  \begin{table}[htbp]
+    \centering
+    \caption{Mapping of FAO crop production items \citep{faostat_faostat_2016} to LUH2 categories \citep{hurtt_harmonization_2020} and to crop groups used in this study.}
+    \begin{tabular}{lllll}
+    \tophline
+        LUH short name & FAO Production Item & Crop group – madrat & Short name – madrat \\ 
+    \middlehline    
+        c3ann & 122|Sweet potatoes & Tropical roots & cassav\_sp \\ 
+        c3ann & 125|Cassava & Tropical roots & cassav\_sp \\ 
+        c3ann & 135|Yautia (cocoyam) & Tropical roots & cassav\_sp \\ 
+        c3ann & 136|Taro (cocoyam) & Tropical roots & cassav\_sp \\ 
+        c3ann & 137|Yams & Tropical roots & cassav\_sp \\ 
+        c3ann & "149|Roots and tubers, nes" & Tropical roots & cassav\_sp \\ 
+        c3per & 486|Bananas & Tropical roots & cassav\_sp \\ 
+        c3per & 489|Plantains & Tropical roots & cassav\_sp \\ 
+        c3ann & 328|Seed cotton & Cotton seed & cottn\_pro \\ 
+        c3ann & "638|Forage and silage, rye grass" & Forage & foddr \\ 
+        c3ann & "639|Forage and silage, grasses nes" & Forage & foddr \\ 
+        c3ann & "642|Forage and silage, green oilseeds" & Forage & foddr \\ 
+        c3ann & 644|Cabbage for fodder & Forage & foddr \\ 
+        c3ann & 645|Pumpkins for Fodder & Forage & foddr \\ 
+        c3ann & 646|Turnips for fodder & Forage & foddr \\ 
+        c3ann & 647|Beets for fodder & Forage & foddr \\ 
+        c3ann & 648|Carrots for fodder & Forage & foddr \\ 
+        c3ann & 649|Swedes for fodder & Forage & foddr \\ 
+        c3ann & 651|Forage products & Forage & foddr \\ 
+        c3ann & 655|Vegetables and roots fodder & Forage & foddr \\ 
+        c4ann & "636|Forage and silage, maize" & Forage & foddr \\ 
+        c4ann & "637|Forage and silage, sorghum" & Forage & foddr \\ 
+        c3nfx & "640|Forage and silage, clover" & Forage & foddr \\ 
+        c3nfx & "641|Forage and silage, alfalfa" & Forage & foddr \\ 
+        c3nfx & "643|Forage and silage, legumes" & Forage & foddr \\ 
+        c3nfx & "242|Groundnuts, with shell" & Groundnuts & groundnut \\ 
+        c4ann & 56|Maize & Maize & maiz \\ 
+        c3per & "254|Oil, palm fruit" & Oilpalms & oilpalm \\ 
+        c3per & 254|Oil palm fruit & Oilpalms & oilpalm \\ 
+        c3ann & 358|Cabbages and other brassicas & Fruits Vegetables Nuts & others \\ 
+        c3ann & 366|Artichokes & Fruits Vegetables Nuts & others \\ 
+        c3ann & 367|Asparagus & Fruits Vegetables Nuts & others \\ 
+        c3ann & 372|Lettuce and chicory & Fruits Vegetables Nuts & others \\ 
+        c3ann & 373|Spinach & Fruits Vegetables Nuts & others \\ 
+        c3ann & 378|Cassava leaves & Fruits Vegetables Nuts & others \\ 
+        c3ann & 388|Tomatoes & Fruits Vegetables Nuts & others \\ 
+        c3ann & 393|Cauliflowers and broccoli & Fruits Vegetables Nuts & others \\ 
+        c3ann & "394|Pumpkins, squash and gourds" & Fruits Vegetables Nuts & others \\ 
+        c3ann & 397|Cucumbers and gherkins & Fruits Vegetables Nuts & others \\ 
+        c3ann & 399|Eggplants (aubergines) & Fruits Vegetables Nuts & others \\ 
+        c3ann & "401|Chillies and peppers, green" & Fruits Vegetables Nuts & others \\ 
+        c3ann & "402|Onions, shallots, green" & Fruits Vegetables Nuts & others \\ 
+        c3ann & "403|Onions, dry" & Fruits Vegetables Nuts & others \\ 
+        c3ann & 406|Garlic & Fruits Vegetables Nuts & others \\ 
+        c3ann & "407|Leeks, other alliaceous vegetables" & Fruits Vegetables Nuts & others \\ 
+        c3ann & 426|Carrots and turnips & Fruits Vegetables Nuts & others \\ 
+        c3ann & 430|Okra & Fruits Vegetables Nuts & others \\ 
+        c3ann & 449|Mushrooms and truffles & Fruits Vegetables Nuts & others \\ 
+        c3ann & 459|Chicory roots & Fruits Vegetables Nuts & others \\ 
+        c3ann & "463|Vegetables, fresh nes" & Fruits Vegetables Nuts & others \\ 
+        c3ann & 544|Strawberries & Fruits Vegetables Nuts & others \\ 
+        c3ann & 567|Watermelons & Fruits Vegetables Nuts & others \\ 
+        c3ann & "568|Melons, other (inccantaloupes)" & Fruits Vegetables Nuts & others \\ 
+        c3ann & "689|Chillies and peppers, dry" & Fruits Vegetables Nuts & others \\ 
+        c3ann & "711|Anise, badian, fennel, coriander" & Fruits Vegetables Nuts & others \\ 
+        c3ann & 720|Ginger & Fruits Vegetables Nuts & others \\ 
+        c3ann & "723|Spices, nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & "216|Brazil nuts, with shell" & Fruits Vegetables Nuts & others \\ 
+        c3per & "217|Cashew nuts, with shell" & Fruits Vegetables Nuts & others \\ 
+        c3per & 220|Chestnut & Fruits Vegetables Nuts & others \\ 
+        c3per & "221|Almonds, with shell" & Fruits Vegetables Nuts & others \\ 
+        c3per & "222|Walnuts, with shell" & Fruits Vegetables Nuts & others \\ 
+        c3per & 223|Pistachios & Fruits Vegetables Nuts & others \\ 
+        c3per & 224|Kola nuts & Fruits Vegetables Nuts & others \\ 
+        c3per & "225|Hazelnuts, with shell" & Fruits Vegetables Nuts & others \\ 
+        c3per & 226|Areca nuts & Fruits Vegetables Nuts & others \\ 
+        c3per & "234|Nuts, nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & 461|Carobs & Fruits Vegetables Nuts & others \\ 
+        c3per & 490|Oranges & Fruits Vegetables Nuts & others \\ 
+        c3per & "495|Tangerines, mandarins, clementines, satsumas" & Fruits Vegetables Nuts & others \\ 
+        c3per & 497|Lemons and limes & Fruits Vegetables Nuts & others \\ 
+        c3per & 507|Grapefruit (inc pomelos) & Fruits Vegetables Nuts & others \\ 
+        c3per & "512|Fruit, citrus nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & 515|Apples & Fruits Vegetables Nuts & others \\ 
+        c3per & 521|Pears & Fruits Vegetables Nuts & others \\ 
+        c3per & 523|Quinces & Fruits Vegetables Nuts & others \\ 
+        c3per & 526|Apricots & Fruits Vegetables Nuts & others \\ 
+        c3per & "530|Cherries, sour" & Fruits Vegetables Nuts & others \\ 
+        c3per & 531|Cherries & Fruits Vegetables Nuts & others \\ 
+        c3per & 534|Peaches and nectarines & Fruits Vegetables Nuts & others \\ 
+        c3per & 536|Plums and sloes & Fruits Vegetables Nuts & others \\ 
+        c3per & "541|Fruit, stone nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & "542|Fruit, pome nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & 547|Raspberries & Fruits Vegetables Nuts & others \\ 
+        c3per & 549|Gooseberries & Fruits Vegetables Nuts & others \\ 
+        c3per & 550|Currants & Fruits Vegetables Nuts & others \\ 
+        c3per & 552|Blueberries & Fruits Vegetables Nuts & others \\ 
+        c3per & 554|Cranberries & Fruits Vegetables Nuts & others \\ 
+        c3per & 558|Berries nes & Fruits Vegetables Nuts & others \\ 
+        c3per & 560|Grapes & Fruits Vegetables Nuts & others \\ 
+        c3per & 569|Figs & Fruits Vegetables Nuts & others \\ 
+        c3per & "571|Mangoes, mangosteens, guavas" & Fruits Vegetables Nuts & others \\ 
+        c3per & 572|Avocados & Fruits Vegetables Nuts & others \\ 
+        c3per & 574|Pineapples & Fruits Vegetables Nuts & others \\ 
+        c3per & 577|Dates & Fruits Vegetables Nuts & others \\ 
+        c3per & 587|Persimmons & Fruits Vegetables Nuts & others \\ 
+        c3per & 591|Cashewapple & Fruits Vegetables Nuts & others \\ 
+        c3per & 592|Kiwi fruit & Fruits Vegetables Nuts & others \\ 
+        c3per & 600|Papayas & Fruits Vegetables Nuts & others \\ 
+        c3per & "603|Fruit, tropical fresh nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & "619|Fruit, fresh nes" & Fruits Vegetables Nuts & others \\ 
+        c3per & "656|Coffee, green" & Fruits Vegetables Nuts & others \\ 
+        c3per & "661|Cocoa, beans" & Fruits Vegetables Nuts & others \\ 
+        c3per & 667|Tea & Fruits Vegetables Nuts & others \\ 
+        c3per & 671|Mate & Fruits Vegetables Nuts & others \\ 
+        c3per & 687|Pepper (piper spp) & Fruits Vegetables Nuts & others \\ 
+        c3per & 692|Vanilla & Fruits Vegetables Nuts & others \\ 
+        c3per & 693|Cinnamon (canella) & Fruits Vegetables Nuts & others \\ 
+        c3per & 698|Cloves & Fruits Vegetables Nuts & others \\ 
+        c3per & "702|Nutmeg, mace and cardamoms" & Fruits Vegetables Nuts & others \\ 
+        c4ann & "446|Maize, green" & Fruits Vegetables Nuts & others \\ 
+        c3nfx & "414|Beans, green" & Fruits Vegetables Nuts & others \\ 
+        c3nfx & "417|Peas, green" & Fruits Vegetables Nuts & others \\ 
+        c3nfx & "420|Vegetables, leguminous nes" & Fruits Vegetables Nuts & others \\ 
+        c3nfx & 423|String beans & Fruits Vegetables Nuts & others \\ 
+        c3ann & 677|Hops & Fruits Vegetables Nuts & others \\ 
+        c3ann & 748|Peppermint & Fruits Vegetables Nuts & others \\ 
+        c3ann & 116|Potatoes & Potatoes & potato \\ 
+        c3nfx & "176|Beans, dry" & Pulses & puls\_pro \\ 
+        c3nfx & "181|Broad beans, horse beans, dry" & Pulses & puls\_pro \\ 
+        c3nfx & "187|Peas, dry" & Pulses & puls\_pro \\ 
+        c3nfx & 191|Chick peas & Pulses & puls\_pro \\ 
+        c3nfx & "195|Cow peas, dry" & Pulses & puls\_pro \\ 
+        c3nfx & 197|Pigeon peas & Pulses & puls\_pro \\ 
+        c3nfx & 201|Lentils & Pulses & puls\_pro \\ 
+        c3nfx & 203|Bambara beans & Pulses & puls\_pro \\ 
+        c3nfx & 205|Vetches & Pulses & puls\_pro \\ 
+        c3nfx & 210|Lupins & Pulses & puls\_pro \\ 
+        c3nfx & "211|Pulses, nes" & Pulses & puls\_pro \\ 
+        c3ann & 265|Castor oil seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 270|Rapeseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 280|Safflower seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 289|Sesame seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 292|Mustard seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 296|Poppy seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 299|Melonseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 333|Linseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & 336|Hempseed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 249|Coconuts & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 260|Olives & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 263|Karite nuts (sheanuts) & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 275|Tung nuts & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 277|Jojoba seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 305|Tallowtree seed & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 310|Kapok fruit & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 311|Kapokseed in shell & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3per & 339|Oilseeds nes & Other oil crops (incl rapeseed) & rapeseed \\ 
+        c3ann & "27|Rice, paddy" & Rice & rice\_pro \\ 
+        c3nfx & 236|Soybeans & Soybean & soybean \\ 
+        c3ann & 157|Sugar beet & Sugar beet & sugr\_beet \\ 
+        c4per & 156|Sugar cane & Sugar cane & sugr\_cane \\ 
+        c3ann & 267|Sunflower seed & Sunflower & sunflower \\ 
+        c3ann & 15|Wheat & Temperate cereals & tece \\ 
+        c3ann & 44|Barley & Temperate cereals & tece \\ 
+        c3ann & 71|Rye & Temperate cereals & tece \\ 
+        c3ann & 75|Oats & Temperate cereals & tece \\ 
+        c3ann & 89|Buckwheat & Temperate cereals & tece \\ 
+        c3ann & 92|Quinoa & Temperate cereals & tece \\ 
+        c3ann & 94|Fonio & Temperate cereals & tece \\ 
+        c3ann & 97|Triticale & Temperate cereals & tece \\ 
+        c3ann & 101|Canary seed & Temperate cereals & tece \\ 
+        c3ann & "103|Grain, mixed" & Temperate cereals & tece \\ 
+        c3ann & "108|Cereals, nes" & Temperate cereals & tece \\ 
+        c4ann & 68|Popcorn & Temperate cereals & tece \\ 
+        c4ann & 79|Millet & Tropical cereals & trce \\ 
+        c4ann & 83|Sorghum & Tropical cereals & trce \\ 
+        \bottomhline
+    \end{tabular}
+    (\#tab:fao2luh2mag)
+  \end{table}
+  ```
+
+  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->
 
@@
 
+    ```{=latex}
   \begin{figure}[H]
@@
   \end{figure}
+  ```
 
-\begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \citealt{century_model_2000}).
-\textit{LitterPNV-CenturyAverage} forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
-Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
-\label{fig:SOClitter}
-\end{figure}
-\noappendix
+output:
+  bookdown::pdf_book:
+    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
+    md_extensions: +raw_attribute
+  md_document:
+    variant: markdown_github
+  bookdown::word_document2: default
+  rticles::copernicus_article: default
+  bookdown::pdf_document2: default
+---
+\newpage
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{KK, BLB and AP designed the study and the model idea.
-KK wrote the code build on work of BLB, IW.
-JPD revised and improved the model code.
-CM, JH and SR provided the LPJmL simulation data.
-KK wrote the paper with important contributions of BLB and CM.
-MK, JS, SR and IW provided extensive feedback to outline of the study.
-All authors discussed the results and commented on the manuscript.} %% optional section
+```{r child = 'introduction.Rmd'}
+```
+\newpage
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
+```{r child = 'methods.Rmd'}
+```
+\newpage
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+```{r child = 'results.Rmd'}
+```
+\newpage
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgements}
-Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
-Additional thanks to the rticles contributors for providing a R Markdown template.
-The authors thank for the data provided by FAOSTAT and LUH2v2.
-The work of KK was funded by the DFG Priority Program ``Climate Engineering: Risks, Challenges, Opportunities?'' (SPP 1689) and specifically the CEMICS2 project (grant no.
-ED78/3-2).
-The research leading to these results has received funding for BLB from the European Union's Horizon 2020 research and innovation program under grant agreement no.
-776479 (COACCH) and no.
-821010 (CASCADES).
-The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF).
-The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no.
-774378).
-\end{acknowledgements}
+```{r child = 'discussion.Rmd'}
+```
+\newpage
 
-%% REFERENCES
-%% DN: pre-configured to BibTeX for rticles
-
-%% The reference list is compiled as follows:
-%%
-%% \begin{thebibliography}{}
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL1}
-%% REFERENCE 1
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL2}
-%% REFERENCE 2
-%%
-%% \end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{SOCbudget.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-\end{document}
+```{r child = 'conclusion.Rmd'}
+```

### 2022-02-28 SOCbudgetPaper2/conclusion.Rmd
Major update of finalized Paper - Draft 2

--- 2020-12-14 SOCbudgetPaper/SOCbudget.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
-
-%% copernicus_rticles_template (flag for rticles template detection - do not remove!)
-
-%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
-
-%% 2-column papers and discussion papers
-\documentclass[gc, manuscript]{copernicus}
-
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
-
-% Advances in Geosciences (adgeo)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
-% Atmospheric Chemistry and Physics (acp)
-% Atmospheric Measurement Techniques (amt)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% DEUQUA Special Publications (deuquasp)
-% Drinking Water Engineering and Science (dwes)
-% Earth Surface Dynamics (esurf)
-% Earth System Dynamics (esd)
-% Earth System Science Data (essd)
-% E&G Quaternary Science Journal (egqsj)
-% European Journal of Mineralogy (ejm)
-% Fossil Record (fr)
-% Geochronology (gchron)
-% Geographica Helvetica (gh)
-% Geoscience Communication (gc)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% History of Geo- and Space Sciences (hgss)
-% Hydrology and Earth System Sciences (hess)
-% Journal of Bone and Joint Infection (jbji)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Magnetic Resonance (mr)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Primate Biology (pb)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Weather and Climate Dynamics (wcd)
-% Web Ecology (we)
-% Wind Energy Science (wes)
-
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
-
-% Pandoc citation processing
-
-% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
-%
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-\begin{document}
-
-\title{Management induced changes of soil organic carbon on global croplands}
-
-\Author[true]{Kristine}{Karstens}
-\Author[1]{Benjamin Leon}{Bodirsky}
-\Author[1]{Jan Philipp}{Dietrich}
-\Author[2]{Marta}{Dondini}
-\Author[1]{Jens}{Heinke}
-\Author[2]{Matthias}{Kuhnert}
-\Author[1]{Christoph}{Müller}
-\Author[1]{Susanne}{Rolinski}
-\Author[2]{Pete}{Smith}
-\Author[1]{Isabelle}{Weindl}
-\Author[1]{Hermann}{Lotze-Campen}
-\Author[1]{Alexander}{Popp}
-
-\affil[1]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany}
-\affil[2]{Institute of Biological \& Environmental Sciences, University of Aberdeen, Aberdeen, UK}
-\affil[3]{Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany}
-
-\runningtitle{R Markdown Template for Copernicus}
-
-\runningauthor{Nüst et al.}
-
-\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
-
-\maketitle
-
-\begin{abstract}
-Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land cover change and agricultural management.
-However, the latter has so far not been well represented in global carbon stock assessments.
-While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
-Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
-Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
-We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
-\end{abstract}
-
-\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
-
-\newpage
-
-\introduction
-
-Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times (\citep{batjes_total_1996}).
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere (\citep{friedlingstein_global_2019}, \citep{minasny_soil_2017}).
-The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (\citep{batjes_total_1996}, \citep{sanderman_soil_2017}).
-
-Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}).
-However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
-Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
-
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
-\citep{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter (\citep{lal_world_2001}).
-
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions (\citep{friedlingstein_global_2019}).
-While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest (\citep{strassmann_simulating_2008}, \citep{betts_climate_2015}).
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management (\citep{friedlingstein_global_2019}, \citep{houghton_carbon_2012}, \citep{hansis_relevance_2015}).
-
-Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g.~\citep{bondeau_modelling_2007}, \citep{lindeskog_implications_2013}).
-\citep{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
-They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
-
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons (\citep{portmann_mirca2000global_2010}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).
-
-More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
-
-Field-scale models (\citet{del_grosso_simulated_2001}{]}, \citep{coleman_simulating_1997},\citep{smith_estimating_2010}, \citep{taghizadeh-toosi_c-tool_2014}) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
-
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
-To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
-Using a scenario analysis, we decompose the contribution of different management activities.
-\newpage
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-In Sect.
-\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}).
-We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}).
-In Sect.
-\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
-Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citep{bodirsky_mrcommons_2020}).
-In Sect.
-\ref{sec:scenarios} we define the scenarios used to complement our historic model results.
-
-\hypertarget{sec:carbonbudget}{%
-\subsection{SOC stocks and stock changes following the steady-state method}\label{sec:carbonbudget}}
-
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
-We restrict our analysis to the first 0-30 cm of the soil profile.
-Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
-Therefore, we take the following three steps for each year of our simulation period:
-(1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
-(2) we account for land conversion by transferring \(SOC\) from and to natural vegetation (Sect. \ref{sec:carbontransfer}),
-(3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
-To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \ref{sec:initsoc}).
-
-\hypertarget{sec:steadystates}{%
-\subsubsection{Steady-state SOC stocks and decay rates}\label{sec:steadystates}}
-
-In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{\mathrm{eq}}\) are given by
-\begin{equation}
-SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
-\label{eq:inoutflow}
-\end{equation}
-with \(C^{\textrm{in}}\) being the carbon inputs to the soil, \(k\) denotes the soil organic carbon decay rate.
-This equation is valid for all grid cells \(i\) and all years \(t\).
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model (\citep{parton_analysis_1987}).
-Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
-We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
-
-Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
-\ref{sec:residues}) and manure inputs (see Sect.
-\ref{sec:livstmanure}); for natural vegetation litterfall (including fine root turnover) (\citep{schaphoff_lpjml4_2018}) is the only source of carbon inflow to the soil.
-Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
-For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients.
-This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \ref{tab:datasourceinputs}.
-
-  \begin{table*}[h]
-  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
-  \tophline
-  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
-  \middlehline
-  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      \cite{faostat_faostat_2016}, \\
-                                                                      LPJmL4 \cite{schaphoff_lpjml4_2018},\\
-                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      LG:C generic values according to Table 5.5B,\\
-                                                                      5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
-                                                                      crop-specific N:C from \cite{bodirsky_n2o_2012}
-                                                                      \end{minipage}} \\
-                            & below-ground residues, &  &  \\
-                            & manure                &  &  \\ 
-                            \hline
-  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
-                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
-                                CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
- \bottomhline
- \end{tabular}
- \label{tab:datasourceinputs}
- \belowtable{}
- \end{table*}
-
-The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space (\(i\)) and time (\(t\)).
-Following the steady-state method \citep{calvo_buendia_ipcc_2019}, we consider temperature (\(temp\)), water (\(wat\)), sand fraction (\(sf\)) and tillage (\(till\)) effects to account for spatial variation of decay rates.
-Thus, \(k_{sub}\) are given by
-
-\begin{equation}
-\begin{aligned}
-& k_{i,t,\mathrm{active},lu}  & = &~ k_{\mathrm{active}}  ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ & \cdot~ sf_{i}\\
-& k_{i,t,\mathrm{slow},lu}    & = &~ k_{\mathrm{slow}}    ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ &\\
-& k_{i,t,\mathrm{passive},lu} & = &~ k_{\mathrm{passive}} ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ & ~ & 
-\label{eq:decayrates}
-\end{aligned}.
-\end{equation}
-
-For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect.
-\ref{sec:tillage}) and irrigation (see Sect.
-\ref{sec:irrigation}) practices on decay rates.
-We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in \citep{calvo_buendia_ipcc_2019}.
-
- \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
- \begin{tabular}{l l l l}
- \tophline
-  land-use types   & type of decay driver & parameter use to represent driver & data source \\
- \middlehline
- \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
-                                     & \cite{hengl_soilgrids250m_2017} \\
-                      \cline{2-4}
-
-                      & Mircobial activity & air temperature & \cite{harris_version_2020} \\
-                      \cline{2-4}
-
-                      & Water restriction & precipitation \& potential evapotranspiration & \cite{harris_version_2020} \\
-                      \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
-                      \cline{2-4}
-
-                      & Soil disturbance & tillage & Sect. \ref{sec:irrigation} \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:datasourcedecay}
- \end{table*}
-
-\hypertarget{sec:carbontransfer}{%
-\subsubsection{SOC transfer between land-use types}\label{sec:carbontransfer}}
-
-We calculate \(SOC\) stocks based on the area shares of land-use types (\(lu\)) within our grid cells (\(i\)).
-If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the \(SOC\) is reallocated.
-We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{t^*}\) via
-
-\begin{equation}
-SOC_{i,t^*,sub,lu} = SOC_{i,t-1,sub,lu} - \frac{SOC_{i,t-1,sub,lu}}{A_{i,t-1,lu}} \cdot  AR_{i,t,lu} + \frac{SOC_{i,t-1,sub,!lu}}{A_{i,t-1,!lu}} \cdot  AE_{i,t,lu}
-\label{eq:ctransfer}
-\end{equation}
-
-with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) and \(AE_{lu}\) the area reduction resp.
-area expansion of the two land-use types.
-Data sources and methodology on land-use states and changes are described in Sect.
-\ref{sec:landuse}.
-
-\hypertarget{sec:totalsoc}{%
-\subsubsection{Total SOC stocks and stock changes}\label{sec:totalsoc}}
-
-\(SOC\) converge towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
-
-\begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t^*,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
-\label{eq:SOCstate}
-\end{equation}
-
-Note that the decay rates have to be multiplied by one year (\(1a\)) to form a dimensionless factor.
-Reformulating this equation, we obtain a mass balance equation as follows
-
-\begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} - \underbrace{SOC_{i,t^*,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
-\label{eq:steadystate2budget}
-\end{equation}
-
-The global \(SOC\) stock for each time step \(t\) can then be calculated via
-
-\begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
-\label{eq:totalstock}
-\end{equation}
-
-According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years (see Eq.
-5.0A in \citep{calvo_buendia_ipcc_2019}).
-This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
-The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by
-
-\begin{equation}
-\Delta SOC_{i,t} = SOC_{i,t} - SOC^{\mathrm{pnv}}_{i,t}\qquad \text{and} \qquad  F^{\mathrm{SCF}}_{i,t} = \frac{SOC_{i,t}}{SOC^{\mathrm{pnv}}_{i,t}} .
-\label{eq:stockdiff}
-\end{equation}
-
-Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
-Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land-use change emissions (\citep{pugh_simulated_2015}).
-
-\hypertarget{sec:initsoc}{%
-\subsubsection{Initialization of SOC pools}\label{sec:initsoc}}
-
-To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
-While land-use, climate and litter input information are available from 1901, agricultural management data on residue and manure inputs are held constant at the level of 1965.
-After a spin-up period of 64 years (1901-1965) with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975 onwards.
-Irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
-
-\hypertarget{sec:tier1}{%
-\subsection{SOC stocks and stock changes following Tier 1}\label{sec:tier1}}
-
-Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
-Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig.
-\ref{fig:CLIMzone}).
-The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by
-
-\begin{equation}
-SOC_{i,t} = \sum_{c,m} T_{c,i} \cdot SOC^{\mathrm{ref}}_{i,t} \cdot F^{\mathrm{SCF}}_{c,m}
-\label{eq:tier1}
-\end{equation}
-
-with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
-
-\hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data at 0.5 degree resolution}\label{sec:agrimanagement}}
-
-We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
-
-\hypertarget{sec:landuse}{%
-\subsubsection{Land use and land-use change}\label{sec:landuse}}
-
-Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{bodirsky_mrcommons_2020}).
-By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
-Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
-
-\hypertarget{sec:residues}{%
-\subsubsection{Crop and crop residues production}\label{sec:residues}}
-
-Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see Sect.
-\ref{sec:landuse}).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production \citep{faostat_faostat_2016}.
-By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
-
-Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
-Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
-
-\begin{equation}
-\begin{aligned}
-AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
-\qquad & \textrm{and} \\
-BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
-\label{eq:resbiomass}
-\end{aligned}
-\end{equation}
-
-Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
-Deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
-We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
-
-A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
-Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.
-
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{bodirsky_mrcommons_2020}).
-Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
-We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above.
-The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field.
-We limit high residue return rates to at most \(10\unit{tC\,ha}^{-1}\) in order to correct for outliers.
-
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{tab:c2dm}).
-
-\hypertarget{sec:livstmanure}{%
-\subsubsection{Livestock distribution and manure excretion}\label{sec:livstmanure}}
-
-Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems at the present day a lot of the manure has to be stored and can be applied on croplands.
-We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
-
-To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from \citep{weindl_livestock_2017}.
-We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-Due to great feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
-We distinguish between grazed pasture, which is converted into livestock products in extensive systems, and primary-crop feed stuff, which we consider to be consumed in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to distinguish between intensive and extensive production systems.
-For low-income countries, we assume only extensive production systems.
-We locate them according to the share of built-up areas based on the assumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
-Intensive production associated with high-income countries, is distributed within a country using the share of primary-crop production, assuming that feed availability is the most determining factor for livestock location.
-For middle-income countries we split the livestock production into extensive and intensive systems based on the per-capita income.
-
-Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
-Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see \citep{calvo_buendia_ipcc_2019}).
-Depending on the feed system we assume manure to be handled in four different ways:
-All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
-To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in \citep{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}.
-We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
-
-\hypertarget{sec:irrigation}{%
-\subsubsection{Irrigation}\label{sec:irrigation}}
-
-The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for their cropland subcategories.
-We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient \(wat\) on decay rates using these shares to compute the weighted mean between rainfed and irrigated \(wat\) factors.
-As a result \(wat\) is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
-This will lead to an overestimation of the irrigation effect.
-We expect, however, water limitations to be a minor problem during the off-season in temperature limited cropping regions, causing our assumption to not dramatically overestimate the moisture effects.
-In tropical, water-limited cropping areas, irrigated growing periods might even span over the whole year.
-
-\hypertarget{sec:tillage}{%
-\subsubsection{Tillage}\label{sec:tillage}}
-
-In order to derive a spatial distribution of the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from \citep{porwollik_generating_2018}.
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from \citep{fao_aquastat_2016} and half-degree resolution physical crop areas from \citep{hurtt_harmonization_2020}, applying the methodology of \citep{porwollik_generating_2018} to identify potential no-tillage grid cells.
-
-\hypertarget{sec:scenarios}{%
-\subsection{Scenario definitions}\label{sec:scenarios}}
-
-To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{karstens_mrsoil_2020}).
-
-\hypertarget{sec:scen_management}{%
-\subsubsection{Management scenarios}\label{sec:scen_management}}
-
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for these three drivers: In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set).
-The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in \(\unit{t ha^{-1}}\)) at the level of 1975 onwards.
-Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
-The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
-
-\hypertarget{sec:scen_initial}{%
-\subsubsection{Initialization choice}\label{sec:scen_initial}}
-
-As outlined in Sect.
-\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
-As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
-Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
-
-\hypertarget{sec:scenlitterpnv}{%
-\subsubsection{Parameterization of natural litter fall}\label{sec:scenlitterpnv}}
-
-Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short LG:C) and nitrogen-to-carbon (short N:C) ratios for all land under natural vegetation.
-This is however a strong simplification.
-Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
-Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
-We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \ref{tab:scenlitterpnv}.
-
- \begin{table*}[h]
- \caption{Scenarios for sensitivity analysis on litter parameterization}
- \begin{tabular}{l l l l}
- \tophline
-  Scenario name & plant type & N:C & LG:C \\
-  \middlehline
-  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
-  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
-  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
-  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
-  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
-  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
-  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:scenlitterpnv}
- \end{table*}
-
-\newpage
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in \citep{karstens_model_2020}.
-In the following, the most important results (see \citep{karstens_result_2020} for post-processing script) are summarized.
-
-\hypertarget{soc-distribution-and-depletion}{%
-\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}
-
-\begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
-Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
-}\label{fig:SOCmaps}
-\end{figure}
-
-In Fig.
-\ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
-Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways:
-1.
-Absolute SOC changes \(\Delta SOC\) (see Fig.
-\ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses.
-They might be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
-In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
-2.
-Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig.
-\ref{fig:SOCmaps}(c)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
-Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
-
-The global SOC debt has decreased by about 15\% in the period between 1975 and 2010 to \(22\unit{GtC}\) (Fig.
-\ref{fig:SOCmaps}(d)).
-This corresponds to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\).
-Considering our estimate of the global SOC stock of around \(660\unit{GtC}\) in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
-
-\hypertarget{carbon-flows-in-the-agricultural-system}{%
-\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
-
-C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
-Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
-We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.~like food, bioenergy and material) into a human demand category.
-Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
-Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not recycled to soils.
-Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
-However, around 60\% of C decomposes before its integrated into soils at the the litter-soil barrier.
-Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
-According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
-4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
-Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
-
-\begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
-\end{figure}
-
-\hypertarget{agricultural-management-effects-on-soc-debt}{%
-\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
-
-We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)).
-Without changes in management regimes, the global \(\Delta SOC\) on cropland would be increasing at a rate of \(0.1\unit{GtC yr^{-1}}\).
-As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global \(\Delta SOC\) would still increase at a rate of \(0.06\unit{GtC yr^{-1}}\).
-Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both).
-The effect of no-tillage has been particuarly strong since the 1990s.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \ref{fig:SOCscen}(b).
-
-\begin{figure}
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
-\end{figure}
-
-Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\ref{fig:SOCinit}).
-The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
-
-\begin{figure}
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
-\end{figure}
-
-Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas \(\Delta SOC\) vary 19-24 GtC for the year 2010 (see Fig.
-\ref{fig:SOClitter}.
-The increase of \(\Delta SOC\) during the period of 1975--2010 is independent of the parameterization choice at \textasciitilde4 GtC.
-
-\newpage
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
-This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
-
-\hypertarget{soc-debt-and-soc-drivers-in-literature}{%
-\subsection{SOC debt and SOC drivers in literature}\label{soc-debt-and-soc-drivers-in-literature}}
-
-The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
-
-According to \citep{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as \citep{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and \citep{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
-This suggests that our results are within a plausible range.
-
-Furthermore, \citep{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
-Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).
-
-Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
-
-Looking on historic cropland intensification rates of above 50\% (\citep{rudel_agricultural_2009}), a large fraction of increased C input from residue biomass is attributed to productivity improvements.
-
-\hypertarget{limitations-and-uncertainties}{%
-\subsection{Limitations and uncertainties}\label{limitations-and-uncertainties}}
-
-Modeling management effects at the global scale comes with parametric and structural uncertainties.
-High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
-\citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
-However, looking on the evaluation of management effects (see Sect.
-\ref{sec:ipcccompare}), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
-Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.~cover- and inter-cropping practices.
-
-Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
-As shown in Fig.
-\ref{fig:SOCinit} the SOC debt estimate almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), and the SOC debt reduction is strongly reduced (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-\citep{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
-
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes that most of the SOC dynamic happens in the topsoil.
-In this regard several aspects are strongly simplified within our approach.
-Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil.
-This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected.
-As \citep{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
-SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
-Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
-
-This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by \citep{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions (\citep{wang_human-induced_2017}).
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
-Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well (\citep{sanderman_soil_2017},\citep{friedlingstein_global_2019}).
-This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
-
-\hypertarget{sec:ipcccompare}{%
-\subsection{Modeled management effect in line with default IPCC assumptions}\label{sec:ipcccompare}}
-
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 (\citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}).
-Both estimates are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\ref{fig:CLIMzone}).
-
-\% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 23:33:09 2020
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlllllll}
-  \hline
- & Source & Input & Year & tropical moist & tropical dry & temperate dry & temperate moist \\ 
-  \hline
-1 & IPCC2006 & low & invariant & 0.44 & 0.55--0.61 & 0.74 & 0.66 \\ 
-  2 & IPCC2006 & medium & invariant & 0.48 & 0.58--0.64 & 0.80 & 0.69 \\ 
-  3 & IPCC2006 & high & invariant & 0.53 & 0.60--0.67 & 0.83 & 0.77 \\ 
-  4 & IPCC2019 & low & invariant & 0.76 & 0.87 & 0.70--0.71 & 0.66--0.67 \\ 
-  5 & IPCC2019 & medium & invariant & 0.83 & 0.92 & 0.76--0.77 & 0.69--0.70 \\ 
-  6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
-  7 & This Study & hist & 1990 & 0.57 & 0.61 & 0.78 & 0.76 \\ 
-  8 & This Study & hist & 2010 & 0.64 & 0.68 & 0.83 & 0.83 \\ 
-   \hline
-\end{tabular}
-\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time.}
-\label{tab:SCFglo}
-\end{table}
-
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than \citep{calvo_buendia_ipcc_2019}.
-The tropical factors differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), with our estimates within this range.
-Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
-Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\ref{fig:SOCmaps}(c)) may be expected.
-
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
-
-\hypertarget{soc-stocks-in-line-with-literature}{%
-\subsection{SOC stocks in line with literature}\label{soc-stocks-in-line-with-literature}}
-
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}, \citep{schaphoff_lpjml4_2018-1}) in Table \ref{fig:SOCglo}).
-
-\begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
-This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
-\end{figure}
-
-The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids (\citep{hengl_soilgrids250m_2017}) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
-Looking at regional results in Fig.
-\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
-To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
-
-Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
-For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
-
-We conducted a sensitivity analysis (Fig.
-\ref{fig:SOClitter}) based on various plant parameterizations from the Century model (see Sect.
-\ref{sec:scenlitterpnv}).
-This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.
-\newpage
-
 \conclusions
 
-Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
-It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
+We have compiled a spatially explicit and time-variant data set on agricultural management aspects relevant for cropland SOC dynamics.
+We have also developed a reduced-complexity SOC model that is able to be applied in optimization-based IAM frameworks, for which detailed process-based models are computationally too expensive.
+Making use of these data and model, we are able to estimate spatially explicit SOC stocks, SOC debts, and stock change factors considering agricultural management.
+It is --- to our knowledge --- the first study that analyzes the role of time-variant and spatially explicit historical agricultural management for global SOC dynamics.
 
-Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models.
+Our results demonstrate that historical changes in agricultural management have shaped the SOC debt on cropland.
+It is thus necessary to explicitly consider agricultural management in a dynamic manner in global carbon assessments and models, especially when exploring climate mitigation pathways with so-called land-based solutions [e.g.
+@popp_land-use_2016].
 That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT package (\citep{dietrich_madrat_2020}) constitute a starting point for building comprehensive data sets on agricultural management aspects.
+Our open-source model [@karstens_mrsoil_2020], published data-set [@karstens_model_2020] and the flexible data processing with the MADRaT package [@dietrich_madrat_2020] constitute a starting point for building comprehensive data sets on agricultural management aspects.
 
-Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions -- not only by the removal of vegetation, but also by a slow depletion of soils.
-Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
-
-However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
-Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
-The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source.
-Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO\textsubscript{2}.
-The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.
-
-Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
-Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand (\citep{popp_land-use_2014-1}) -- and thereby usually further reduce the SOC debt on croplands.
-
-While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets.
-Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
-
-Yet, there is ample potential for further improved SOC management.
-As shown in Fig.
-\ref{fig:FlowFig}, approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC).
-Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils (\citep{brenzinger_organic_2018}).
-Secondly, soil carbon sequestration techniques (\citep{smith_soil_2016}), deep ploughing (\citep{alcantara_deep_2016}) or the transformation of c inputs to more recalcitrant biochar (\citep{woolf_sustainable_2010}) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
-Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
-Finally, other carbon-accumulating practices, such as the cultivation of cover crops (\citep{poeplau_carbon_2015}) and agroforestry (\citep{lorenz_soil_2014}) could increase total C sequestration on croplands.
-
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
-All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results including C input data are accessable under \url{https://doi.org/10.5281/zenodo.4320663}.
-Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.} %% use this section when having data sets and software code available
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Figures and tables in appendices}
-
-\subsection{Methods}
-
-\begin{figure}
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
-\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
-Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
-\label{fig:CLIMzone}
-\end{figure}
-
-\begin{table*}[h]
-\caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
-Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
-}
-\begin{tabular}{lllllllllllll}
-\tophline
-
-         &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
-Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
- \middlehline
-tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
-maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
-trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
-rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
-soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
-rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
-groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
-sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
-oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
-puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
-potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
-cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
-sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
-sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
-others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
-foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
-cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
-\bottomhline
-         &                 & \multicolumn{8}{l}{
-                                           \begin{minipage}[t]{0.50\columnwidth}
-                                                      nr/dm -- nitrogen to dry matter ratio\\
-                                                      wm/dm -- wet matter to dry matter ratio\\
-                                                      c/dm  -- carbon to dry matter ratio\\
-                                                      \end{minipage}} &
-                                                                        \multicolumn{3}{l}{
-                                                                            \begin{minipage}[t]{0.50\columnwidth}
-                                                                                        HI\textsuperscript{area} -- harvest index per area\\
-                                                                                        HI\textsuperscript{prod} -- harvest index per production\\
-                                                                                        RS -- root:shoot ratio \end{minipage}} 
-\end{tabular}
-\label{tab:c2dm}
-\end{table*}
-
-\subsection{Discussion}
-
-\begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
-\caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.
-PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
-We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
-\label{fig:SOCreg}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
-\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
-Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
-\label{fig:SOClitter}
-\end{figure}
-\noappendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{KK, BLB and AP designed the study and the model idea.
-KK wrote the code build on work of BLB, IW.
-JPD revised and improved the model code.
-CM, JH and SR provided the LPJmL simulation data.
-KK wrote the paper with important contributions of BLB and CM.
-MK, JS, SR and IW provided extensive feedback to outline of the study.
-All authors discussed the results and commented on the manuscript.} %% optional section
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgements}
-Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
-Additional thanks to the rticles contributors for providing a R Markdown template.
-The authors thank for the data provided by FAOSTAT and LUH2v2.
-The work of KK was funded by the DFG Priority Program ``Climate Engineering: Risks, Challenges, Opportunities?'' (SPP 1689) and specifically the CEMICS2 project (grant no.
-ED78/3-2).
-The research leading to these results has received funding for BLB from the European Union's Horizon 2020 research and innovation program under grant agreement no.
-776479 (COACCH) and no.
-821010 (CASCADES).
-The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF).
-The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no.
-774378).
-\end{acknowledgements}
-
-%% REFERENCES
-%% DN: pre-configured to BibTeX for rticles
-
-%% The reference list is compiled as follows:
-%%
-%% \begin{thebibliography}{}
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL1}
-%% REFERENCE 1
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL2}
-%% REFERENCE 2
-%%
-%% \end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{SOCbudget.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-\end{document}
+With the reduced-complexity SOC model we are able to account for agricultural management effects on cropland SOC dynamics within optimization-based IAM frameworks.
+Reduced input data requirements such as accounting for changes in productivity rather than reproducing the processes that lead to such changes in productivity [@elliott_management_trends_2018] will help to explore the role of agricultural management in sustainable development pathway analyses [@sorgel_sustainable_2021].
+However, we clearly see that increases in agricultural productivity are not sufficient to create positive net SOC sequestration in cropland soils.
+More management options that explicitly target the sequestration of C in cropland soils need to be considered.
+Our open-source model can be expanded to account for additional management options for carbon farming, such as cover crops, agroforestry, or biochar applications.

### 2022-02-28 SOCbudgetPaper2/discussion.Rmd
Major update of finalized Paper - Draft 2

--- 2020-12-14 SOCbudgetPaper/SOCbudget.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
+# Discussion
 
-%% copernicus_rticles_template (flag for rticles template detection - do not remove!)
+We have (1) developed a reduced-complexity model and (2) compiled a spatially explicit time series data set of agricultural management data in order to (3) analyze the role of agricultural management in historical cropland SOC dynamics.
+Our study shows that information on agricultural management alters estimates of the SOC debt and slows down loss of SOC compared to the often used constant management assumptions.
 
-%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
+It is important to evaluate the validity of our results as modeling management effects on SOC dynamics at the global scale comes with large uncertainties.
+The model includes a large number of parameters, and for most of these the uncertainty distributions have not been quantified so far.
+Moreover, we think that beyond parameter uncertainty, the structural uncertainty from the model design is high.
+The management data itself is prone to uncertainties as well, as most of it is only indirectly calculated from reported data.
 
-%% 2-column papers and discussion papers
-\documentclass[gc, manuscript]{copernicus}
+In the following, we give a qualitative assessment of the uncertainties and limitations of this study as well as discuss our three study objectives and results against existing literature.
 
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
+## Comprehensive historical agricultural management data set
 
-% Advances in Geosciences (adgeo)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
-% Atmospheric Chemistry and Physics (acp)
-% Atmospheric Measurement Techniques (amt)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% DEUQUA Special Publications (deuquasp)
-% Drinking Water Engineering and Science (dwes)
-% Earth Surface Dynamics (esurf)
-% Earth System Dynamics (esd)
-% Earth System Science Data (essd)
-% E&G Quaternary Science Journal (egqsj)
-% European Journal of Mineralogy (ejm)
-% Fossil Record (fr)
-% Geochronology (gchron)
-% Geographica Helvetica (gh)
-% Geoscience Communication (gc)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% History of Geo- and Space Sciences (hgss)
-% Hydrology and Earth System Sciences (hess)
-% Journal of Bone and Joint Infection (jbji)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Magnetic Resonance (mr)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Primate Biology (pb)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Weather and Climate Dynamics (wcd)
-% Web Ecology (we)
-% Wind Energy Science (wes)
+Our spatially explicit time series dataset of agricultural management
+is based on country-specific FAO production and cropland statistics [@faostat_faostat_2016] as well as 0.5 degree land-use data from LUH2 [@hurtt_harmonization_2020].
+Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the agricultural system (\@ref(fig:FlowFig)) using a mass balance approach from the IPCC guidelines Vol.
+4 [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and other auxiliary data sets (e.g.
+[@porwollik_generating_2019]).
 
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
+For some of the aspects covered in our data set, for example livestock distribution [@robinson_mapping_2014] or manure production and application [@zhang_global_2017], well-compiled data sets in high resolution exist that capture real world conditions much better than our estimates.
+However, they often come with the caveats of either being static in time, demanding large sets of auxiliary data or being inconsistent with each other.
 
-% Pandoc citation processing
+For most of the parameters used in our management estimates no uncertainty estimates exist.
+This is why, in our view, most of the uncertainty with respect to the impacts of SOC management is included in the management data itself, and especially in the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017].
+The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
+Previous modeling studies of SOC carbon on cropland often only used stylized scenarios of management practices [@pugh_simulated_2015; @lutz_simulating_2019], rather than trying to estimate real management.
 
-% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
-%
-\usepackage{algorithmic}
-\usepackage{algorithm}
+While our dataset, by including crop residues and manure, likely the largest carbon inputs to soils, it does not account for a list of minor carbon inputs from cover crops, agroforestry, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
+Including these sources would correct our estimates upwards and bring our estimates closer to the IPCC stock change factors(see Sect.
+\@ref(sec:ipcccompare)).
+Unfortunately, data on the quantity of these inputs is very scarce and does not exist with global coverage.
 
-\begin{document}
+SOC inputs from above-ground residues had the strongest management effect on SOC debt dynamics on cropland (see Fig.
+\@ref(fig:SOCscen)).
+As pointed out by Keel et al.
+[-@keel_large_2017] and Smith et al.
+[-@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study).
+Keel et al.
+[-@keel_large_2017] question whether below-ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass, in particular for below-ground biomass.
 
-\title{Management induced changes of soil organic carbon on global croplands}
+## Reduced complexity SOC model
 
-\Author[true]{Kristine}{Karstens}
-\Author[1]{Benjamin Leon}{Bodirsky}
-\Author[1]{Jan Philipp}{Dietrich}
-\Author[2]{Marta}{Dondini}
-\Author[1]{Jens}{Heinke}
-\Author[2]{Matthias}{Kuhnert}
-\Author[1]{Christoph}{Müller}
-\Author[1]{Susanne}{Rolinski}
-\Author[2]{Pete}{Smith}
-\Author[1]{Isabelle}{Weindl}
-\Author[1]{Hermann}{Lotze-Campen}
-\Author[1]{Alexander}{Popp}
+Our reduced-complexity SOC model is based on a Tier 2 modeling approach.
+This reduces the computational and data demand of the model in comparison to DGVMs, while still allowing for the explicit representation of agricultural management practices.
+Along with the effects of various C inputs, the impacts of water supply from rainfall and irrigation as well as tillage systems can also be accounted for in the computation of SOC decay rates.
+As such, the model can reflect the spatial and temporal heterogeneity in both management and biophysical conditions.
 
-\affil[1]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany}
-\affil[2]{Institute of Biological \& Environmental Sciences, University of Aberdeen, Aberdeen, UK}
-\affil[3]{Humboldt-Universität zu Berlin, Department of Agricultural Economics, Unter den Linden 6, 10099 Berlin, Germany}
+The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Table \@ref(tab:SCFglo)) as well as by the time trend of the SOC debt (see Fig.
+\@ref(fig:SOCmaps)(a)).
+Residue management has changed over the last decades, especially with the phasing out of residue burning practices in several regions and increased general productivity, showing a clear impact on SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
 
-\runningtitle{R Markdown Template for Copernicus}
+The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV.
+This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is a main driver of SOC dynamics (cite).
+The comparison of simulated PNV data with LPJmL4 shows the model's substantial capability in reproducing PNV SOC stocks (Fig.
+\@ref(fig:SOCLPJmL4)).
 
-\runningauthor{Nüst et al.}
+Using litterfall estimates from LPJmL4, we have been able to estimate the total SOC stocks of the world, which is dominated by SOC under natural vegetation.
+However, as the world's SOC stock is highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1; @poggio_soilgrids_2021; @sanderman_soil_2017] in Fig.
+\@ref(fig:SOCglo), the only conclusion we can draw from this is that our result is within a plausible range.
+To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland.
+Pristine natural vegetated areas (like permafrost and rain forests) without human land management drop out in our calculation of SOC debt and stock change factors.
+Natural SOC estimates only influence results when natural land is converted to cropland.
 
-\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}
+Comparing the geographic SOC patterns to Soil Grids 2.0 [@poggio_soilgrids_2021; see Fig.
+\@ref(fig:SOCSoilGrids2Q)], we find that our model estimates values of SOC stocks greater than the estimated confidence intervals in Soil Grid 2.0s for some mountainous regions across the globe.
+This could indicate that we are not capable in capturing specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils [@borrelli_assessment_2017]).
+A large swathe of eastern North America was heavily affected by the dust bowl event, with wind erosion removing large parts of the topsoils, a process not considered in our model.
+Similarly, we likely overestimate SOC stocks for the loess soils in northern China and the Altiplano in Latin America; in both cases erosion is a likely reason.
+In contrast, we estimate lower SOC stocks at the edges of the Sahara, where uncertain local water availability and artifical irrigation may dominate spatial SOC patterns.
 
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
+In our model, erosion should however only affect the spatial pattern but not the aggregate SOC pool.
+As pointed out by Doetterl et al.
+[-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017].
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focused especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
+Erosion and degradation impacts on yields and therefore on soil C inputs are captured by our method as we base them on FAO statistics of actual production.
+Yet the distribution of production below the country level - which we allocate proportional to LPJmL production potentials that do not reflect erosion feedback to yields - will overestimate yields and therefore biomass inputs to eroded soils.
 
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
+In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Table \@ref(tab:SCFglo)) for almost all regions.
 
-\firstpage{1}
+Tropical soils might suffer from low C input rates due to large yield gaps [@global_yieldgap_2022] and high shares of residue removal and burning in lower-income countries [@smil_crop_1999; williams_influence_1997; jain_emission_2014].
+Yet, even when comparing our etimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values, while it shows very good agreement with the older default values from IPCC [-@lasco_cropland_in_ipcc_2006].
+Don et al.
+[-@don_impact_2011] estimated SOC losses for tropical soils of around 25% on average corresponding to a stock change factor of 0.75, but also reported a wide range of measured SOC changes from -80% to +58%.
+Fujisaki et al.
+[-@fujisaki_forest_2015] however found much lower loss rates of around 9%, attributing the difference to the different time period length since the conversion to cropland.
+As our results do not specifically account for cropland age and most of the cropland is older than 20 years (as assumed for the default IPCC Tier 1 stock change factors) our stock change factors have to be lower by definition following the steady-state assumption that cropland will continue to approach a new equilibrium.
+For the same reason, our estimates for temperate regions might be lower than both IPCC [-@lasco_cropland_in_ipcc_2006] and IPCC [-@ogle_cropland_in_ipcc_2019] default values.
+With the production-increasing impact of irrigation and fertilization on carbon-poor dryland soils, SOC under cropland can also be higher than under PNV with stock change factors above 1 (see Fig.
+\@ref(fig:SOCmaps)(d)), but these areas are much smaller than where the stock change factors are well below unity.
 
-\maketitle
-
-\begin{abstract}
-Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land cover change and agricultural management.
-However, the latter has so far not been well represented in global carbon stock assessments.
-While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
-Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975.
-Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity.
-We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
-\end{abstract}
-
-\copyrightstatement{The author's copyright for this publication is transferred to institution/company.}
-
-\newpage
-
-\introduction
-
-Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth's soil, is the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times (\citep{batjes_total_1996}).
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere (\citep{friedlingstein_global_2019}, \citep{minasny_soil_2017}).
-The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (\citep{batjes_total_1996}, \citep{sanderman_soil_2017}).
-
-Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
-Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}).
-However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
-Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
-
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
-\citep{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter (\citep{lal_world_2001}).
-
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions (\citep{friedlingstein_global_2019}).
-While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest (\citep{strassmann_simulating_2008}, \citep{betts_climate_2015}).
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization, and agricultural management (\citep{friedlingstein_global_2019}, \citep{houghton_carbon_2012}, \citep{hansis_relevance_2015}).
-
-Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g.~\citep{bondeau_modelling_2007}, \citep{lindeskog_implications_2013}).
-\citep{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
-They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
-
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons (\citep{portmann_mirca2000global_2010}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).
-
-More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.~on tillage systems, see \citep{porwollik_generating_2018}, \citep{prestele_spatially_2018}) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019}, \citep{minoli_modelling_2019}), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
-
-Field-scale models (\citet{del_grosso_simulated_2001}{]}, \citep{coleman_simulating_1997},\citep{smith_estimating_2010}, \citep{taghizadeh-toosi_c-tool_2014}) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
-
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
-To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
-Using a scenario analysis, we decompose the contribution of different management activities.
-\newpage
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-In Sect.
-\ref{sec:carbonbudget} we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}).
-We additionally describe how we configured and extended the steady-state method (for model code see \citep{karstens_mrsoil_2020}).
-In Sect.
-\ref{sec:tier1} we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
-Section \ref{sec:agrimanagement} provides a detailed description of the global, gridded management data used to drive the model, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices (for model code see \citep{bodirsky_mrcommons_2020}).
-In Sect.
-\ref{sec:scenarios} we define the scenarios used to complement our historic model results.
-
-\hypertarget{sec:carbonbudget}{%
-\subsection{SOC stocks and stock changes following the steady-state method}\label{sec:carbonbudget}}
-
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.~4 (\citep{calvo_buendia_ipcc_2019}; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon (\(SOC\)) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
-We restrict our analysis to the first 0-30 cm of the soil profile.
-Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
-Therefore, we take the following three steps for each year of our simulation period:
-(1) We calculate annual land-use type-specific steady states and decay rates for \(SOC\) stocks (Sect. \ref{sec:steadystates});
-(2) we account for land conversion by transferring \(SOC\) from and to natural vegetation (Sect. \ref{sec:carbontransfer}),
-(3) we estimate \(SOC\) stocks and changes based on the stocks of the previous time step, the steady state stocks and the decay rate (Sect. \ref{sec:totalsoc}).
-To initialize the first year of our simulation period we use a spin-up period of 74 years (Sect. \ref{sec:initsoc}).
-
-\hypertarget{sec:steadystates}{%
-\subsubsection{Steady-state SOC stocks and decay rates}\label{sec:steadystates}}
-
-In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{\mathrm{eq}}\) are given by
-\begin{equation}
-SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
-\label{eq:inoutflow}
-\end{equation}
-with \(C^{\textrm{in}}\) being the carbon inputs to the soil, \(k\) denotes the soil organic carbon decay rate.
-This equation is valid for all grid cells \(i\) and all years \(t\).
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools \(sub\) (active, slow and passive) and interactions between them, following the approach in the Century model (\citep{parton_analysis_1987}).
-Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
-We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
-
-Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
-\ref{sec:residues}) and manure inputs (see Sect.
-\ref{sec:livstmanure}); for natural vegetation litterfall (including fine root turnover) (\citep{schaphoff_lpjml4_2018}) is the only source of carbon inflow to the soil.
-Following the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
-For each component the sum of all carbon input sources is allocated to the respective \(SOC\) sub-pools via transfer coefficients.
-This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \ref{tab:datasourceinputs}.
-
-  \begin{table*}[h]
-  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
-  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
-  \tophline
-  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
-  \middlehline
-  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      \cite{faostat_faostat_2016}, \\
-                                                                      LPJmL4 \cite{schaphoff_lpjml4_2018},\\
-                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
-                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
-                                                                      LG:C generic values according to Table 5.5B,\\
-                                                                      5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
-                                                                      crop-specific N:C from \cite{bodirsky_n2o_2012}
-                                                                      \end{minipage}} \\
-                            & below-ground residues, &  &  \\
-                            & manure                &  &  \\ 
-                            \hline
-  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
-                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
-                                CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
- \bottomhline
- \end{tabular}
- \label{tab:datasourceinputs}
- \belowtable{}
- \end{table*}
-
-The sub-pool specific decay rates \(k_{sub}\) are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space (\(i\)) and time (\(t\)).
-Following the steady-state method \citep{calvo_buendia_ipcc_2019}, we consider temperature (\(temp\)), water (\(wat\)), sand fraction (\(sf\)) and tillage (\(till\)) effects to account for spatial variation of decay rates.
-Thus, \(k_{sub}\) are given by
-
-\begin{equation}
-\begin{aligned}
-& k_{i,t,\mathrm{active},lu}  & = &~ k_{\mathrm{active}}  ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ & \cdot~ sf_{i}\\
-& k_{i,t,\mathrm{slow},lu}    & = &~ k_{\mathrm{slow}}    ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ &\cdot~ till_{i,t,lu} ~ &\\
-& k_{i,t,\mathrm{passive},lu} & = &~ k_{\mathrm{passive}} ~ &\cdot~ temp_{i,t} ~ &\cdot~ wat_{i,t,lu} ~ & ~ & 
-\label{eq:decayrates}
-\end{aligned}.
-\end{equation}
-
-For natural vegetation, we assume rainfed and non-tilled conditions, whereas for cropland, we distinguish the effect of different tillage (see Sect.
-\ref{sec:tillage}) and irrigation (see Sect.
-\ref{sec:irrigation}) practices on decay rates.
-We calculated area weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equation 5.0B--5.0F in \citep{calvo_buendia_ipcc_2019}.
-
- \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
- \begin{tabular}{l l l l}
- \tophline
-  land-use types   & type of decay driver & parameter use to represent driver & data source \\
- \middlehline
- \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm 
-                                     & \cite{hengl_soilgrids250m_2017} \\
-                      \cline{2-4}
-
-                      & Mircobial activity & air temperature & \cite{harris_version_2020} \\
-                      \cline{2-4}
-
-                      & Water restriction & precipitation \& potential evapotranspiration & \cite{harris_version_2020} \\
-                      \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. \ref{sec:irrigation} \\ 
-                      \cline{2-4}
-
-                      & Soil disturbance & tillage & Sect. \ref{sec:irrigation} \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:datasourcedecay}
- \end{table*}
-
-\hypertarget{sec:carbontransfer}{%
-\subsubsection{SOC transfer between land-use types}\label{sec:carbontransfer}}
-
-We calculate \(SOC\) stocks based on the area shares of land-use types (\(lu\)) within our grid cells (\(i\)).
-If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the \(SOC\) is reallocated.
-We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{t^*}\) via
-
-\begin{equation}
-SOC_{i,t^*,sub,lu} = SOC_{i,t-1,sub,lu} - \frac{SOC_{i,t-1,sub,lu}}{A_{i,t-1,lu}} \cdot  AR_{i,t,lu} + \frac{SOC_{i,t-1,sub,!lu}}{A_{i,t-1,!lu}} \cdot  AE_{i,t,lu}
-\label{eq:ctransfer}
-\end{equation}
-
-with \(A_{lu}\) being the land-use type specific areas, \(AR_{lu}\) and \(AE_{lu}\) the area reduction resp.
-area expansion of the two land-use types.
-Data sources and methodology on land-use states and changes are described in Sect.
-\ref{sec:landuse}.
-
-\hypertarget{sec:totalsoc}{%
-\subsubsection{Total SOC stocks and stock changes}\label{sec:totalsoc}}
-
-\(SOC\) converge towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
-
-\begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} + (SOC^{\mathrm{eq}}_{i,t,sub,lu} - SOC_{i,t^*,sub,lu}) \cdot k_{i,t,sub,lu} \cdot 1\unit{a}.
-\label{eq:SOCstate}
-\end{equation}
-
-Note that the decay rates have to be multiplied by one year (\(1a\)) to form a dimensionless factor.
-Reformulating this equation, we obtain a mass balance equation as follows
-
-\begin{equation}
-SOC_{i,t,sub,lu} = SOC_{i,t^*,sub,lu} - \underbrace{SOC_{i,t^*,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{\mathrm{eq}}_{i,t,sub,lu} \cdot k_{i,t,sub,lu} \cdot 1\unit{a}}^{\text{input (using equation (1))}}.
-\label{eq:steadystate2budget}
-\end{equation}
-
-The global \(SOC\) stock for each time step \(t\) can then be calculated via
-
-\begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
-\label{eq:totalstock}
-\end{equation}
-
-According to the IPCC guidelines \(SOC\) changes can be expressed as the difference of two consecutive years (see Eq.
-5.0A in \citep{calvo_buendia_ipcc_2019}).
-This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we defined the absolute and relative SOC changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
-The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by
-
-\begin{equation}
-\Delta SOC_{i,t} = SOC_{i,t} - SOC^{\mathrm{pnv}}_{i,t}\qquad \text{and} \qquad  F^{\mathrm{SCF}}_{i,t} = \frac{SOC_{i,t}}{SOC^{\mathrm{pnv}}_{i,t}} .
-\label{eq:stockdiff}
-\end{equation}
-
-Note that the absolute changes \(\Delta SOC\) can be also interpreted as the SOC debt (\citep{sanderman_soil_2017}) due to human cropping activities; whereas relative changes \(F^{\mathrm{SCF}}\) can be considered stock change factors as defined within the IPCC guidelines of 2006 \citep{eggleston_ipcc_2006}.
-Moreover, \(\Delta SOC\) is equivalent to the negated cumulative SOC component of human land-use change emissions (\citep{pugh_simulated_2015}).
-
-\hypertarget{sec:initsoc}{%
-\subsubsection{Initialization of SOC pools}\label{sec:initsoc}}
-
-To initialize all SOC sub-pools we assume that cropland and natural vegetation are in a land-use type specific steady state for the initialization year 1901.
-While land-use, climate and litter input information are available from 1901, agricultural management data on residue and manure inputs are held constant at the level of 1965.
-After a spin-up period of 64 years (1901-1965) with constant agricultural input data, we run the model for additional 10 years with historic input data and start analyzing results from 1975 onwards.
-Irrigation areas are part of the land-use data set and therefore dynamic from 1901 on, whereas data on no-tillage area is only available after 1974.
-
-\hypertarget{sec:tier1}{%
-\subsection{SOC stocks and stock changes following Tier 1}\label{sec:tier1}}
-
-Additionally to the steady-state method \citep{calvo_buendia_ipcc_2019} and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}).
-Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-Estimates of \(F^{\mathrm{SCF}}\) are differentiated by different crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
-Moreover, estimates of \(F^{\mathrm{SCF}}\) vary for different climatic zones (\(c\)) specified by the IPCC (see Fig.
-\ref{fig:CLIMzone}).
-The actual \(SOC\) stocks are thus calculated based on a given reference stock \(SOC^{\mathrm{ref}}\) by
-
-\begin{equation}
-SOC_{i,t} = \sum_{c,m} T_{c,i} \cdot SOC^{\mathrm{ref}}_{i,t} \cdot F^{\mathrm{SCF}}_{c,m}
-\label{eq:tier1}
-\end{equation}
-
-with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of \citep{eggleston_ipcc_2006} and \citep{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
-
-\hypertarget{sec:agrimanagement}{%
-\subsection{Agricultural management data at 0.5 degree resolution}\label{sec:agrimanagement}}
-
-We compile country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
-For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
-
-\hypertarget{sec:landuse}{%
-\subsubsection{Land use and land-use change}\label{sec:landuse}}
-
-Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016} (see FAO2LUH2MAG\_croptypes.csv in \citep{bodirsky_mrcommons_2020}).
-By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
-Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
-
-\hypertarget{sec:residues}{%
-\subsubsection{Crop and crop residues production}\label{sec:residues}}
-
-Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see Sect.
-\ref{sec:landuse}).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production \citep{faostat_faostat_2016}.
-By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
-
-Crop residue production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and key aspects are explained here given its central role for soil carbon modeling.
-Starting from harvested crop production (\(CP\)) estimates and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest index values (\(HI\)) and root:shoot ratios (\(RS\)) as follows
-
-\begin{equation}
-\begin{aligned}
-AGR_{i,t,cg} & = CP_{i,t,cg} \cdot HI^{\mathrm{prod}}_{cg} + CA_{i,t,cg} \cdot HI^{\mathrm{area}}_{cg}
-\qquad & \textrm{and} \\
-BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot RS_{cg} \qquad                                            & \forall\quad cg, i, t.
-\label{eq:resbiomass}
-\end{aligned}
-\end{equation}
-
-Following the IPCC guidelines, we split the harvest index into a production (\(HI^{\mathrm{prod}}\)) and an area dependent (\(HI^{\mathrm{area}}\)) fraction \citep{eggleston_ipcc_2006}.
-Deviating from \citep{bodirsky_n2o_2012} we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
-We assume that all \(BGR\) are left in the soil, whereas \(AGR\) can be burned or harvested for other purposes such as feeding animals \citep{weindl_livestock_2017}, fuel or for material use.
-
-A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
-Following \citep{smil_nitrogen_1999}) we assume a burn share of \(25\%\) for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
-Depending on the crop group, 80--90\% of the carbon in the crop residues burned in the fields is lost within the combustion process \citep{eggleston_ipcc_2006}.
-
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in \citep{bodirsky_mrcommons_2020}).
-Residue feed demand for five different livestock groups is based on country-specific feed baskets (see \citep{weindl_livestock_2017}), that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
-We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
-For high-income countries, no withdrawal for material or fuel use is assumed, and use shares of middle-income countries are linearly interpolated based on per-capita income, following the same rationale as for the share of burnt residues described above.
-The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field.
-We limit high residue return rates to at most \(10\unit{tC\,ha}^{-1}\) in order to correct for outliers.
-
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \ref{tab:c2dm}).
-
-\hypertarget{sec:livstmanure}{%
-\subsubsection{Livestock distribution and manure excretion}\label{sec:livstmanure}}
-
-Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems at the present day a lot of the manure has to be stored and can be applied on croplands.
-We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
-
-To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of \citep{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from \citep{weindl_livestock_2017}.
-We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
-Due to great feed demand of ruminants, we assume that ruminant livestock is located where the production of feed occurs to minimize transport of feed.
-We distinguish between grazed pasture, which is converted into livestock products in extensive systems, and primary-crop feed stuff, which we consider to be consumed in intensive systems.
-For poultry, egg and monogastric meat production we use the per-capita income of the country to distinguish between intensive and extensive production systems.
-For low-income countries, we assume only extensive production systems.
-We locate them according to the share of built-up areas based on the assumption that these animals are held in subsistence or small-holder farming systems with a high labor-per-animal ratio.
-Intensive production associated with high-income countries, is distributed within a country using the share of primary-crop production, assuming that feed availability is the most determining factor for livestock location.
-For middle-income countries we split the livestock production into extensive and intensive systems based on the per-capita income.
-
-Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
-Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see \citep{calvo_buendia_ipcc_2019}).
-Depending on the feed system we assume manure to be handled in four different ways:
-All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
-Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
-For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
-To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems.
-Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in \citep{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}.
-We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.
-
-\hypertarget{sec:irrigation}{%
-\subsubsection{Irrigation}\label{sec:irrigation}}
-
-The LUH2v2 \citep{hurtt_harmonization_2020} data set provides irrigated fractions for their cropland subcategories.
-We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient \(wat\) on decay rates using these shares to compute the weighted mean between rainfed and irrigated \(wat\) factors.
-As a result \(wat\) is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
-This will lead to an overestimation of the irrigation effect.
-We expect, however, water limitations to be a minor problem during the off-season in temperature limited cropping regions, causing our assumption to not dramatically overestimate the moisture effects.
-In tropical, water-limited cropping areas, irrigated growing periods might even span over the whole year.
-
-\hypertarget{sec:tillage}{%
-\subsubsection{Tillage}\label{sec:tillage}}
-
-In order to derive a spatial distribution of the three different tillage types specified by the IPCC -- full tillage, reduced tillage and no tillage --, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from \citep{porwollik_generating_2018}.
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from \citep{fao_aquastat_2016} and half-degree resolution physical crop areas from \citep{hurtt_harmonization_2020}, applying the methodology of \citep{porwollik_generating_2018} to identify potential no-tillage grid cells.
-
-\hypertarget{sec:scenarios}{%
-\subsection{Scenario definitions}\label{sec:scenarios}}
-
-To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios (for technical implementation see \citep{karstens_mrsoil_2020}).
-
-\hypertarget{sec:scen_management}{%
-\subsubsection{Management scenarios}\label{sec:scen_management}}
-
-To single out the impact of tillage practices, residue and manure inputs, we defined scenario with constant values for these three drivers: In the \textit{constTillage} scenario the adoption of no-tillage practices are neglected (adoption starts in 1974 according to the available data set).
-The \textit{constResidues} and the \textit{constManure} scenario assume constant input rates from residues resp.
-manure (in \(\unit{t ha^{-1}}\)) at the level of 1975 onwards.
-Within the \textit{constResidue} scenario at different effects overlay each other: yields and with them residue biomass increase due to productivity gains; rates of residue left or returned to fields are raising; and shifts of cropping pattern change the amount of residue biomass due to crop-group specific harvest index values.
-The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
-
-\hypertarget{sec:scen_initial}{%
-\subsubsection{Initialization choice}\label{sec:scen_initial}}
-
-As outlined in Sect.
-\ref{sec:initsoc} we assume landuse-type specific steady-state SOC stocks for the start year 1901 followed by a long spin-up period of 74 years (\textit{Initial-lu}).
-As some SOC compartments decay over very long timescales, the initialisation setting might strongly effect the overall outcome of SOC stocks and changes.
-Thus we conduct counterfactual scenario \textit{Initial-natveg}, initializing the start year with the steady-state SOC under potential natural vegetation for all land-use types.
-We additionally combined the \textit{Initial-natveg} scenario with the \textit{constManagement} scenario.
-
-\hypertarget{sec:scenlitterpnv}{%
-\subsubsection{Parameterization of natural litter fall}\label{sec:scenlitterpnv}}
-
-Due to missing information about the litterfall composition, we assume woody potential natural vegetation for the lignin-to-carbon (short LG:C) and nitrogen-to-carbon (short N:C) ratios for all land under natural vegetation.
-This is however a strong simplification.
-Therefore, we test the sensitivity of the model for other natural vegetation parameterizations.
-Perennial grasses are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, but for woody biomass we used the CENTURY configuration file (\citep{century_model_2000}) for trees as parameterization were not given.
-Whereas LG:C ratios ranges between \(0.09-0.35\) with an average of \(0.24\) (default value) for different tree compartments and different tree types, N:C ratios range between \(0.001-0.05\) with an average of \(0.015\) (default value).
-We decided to average over all tree compartments and tree types equally for the default values, since we have no data available on tree type distribution and tree compartment composition.
-We test however a range of different plant types as specified in Table \ref{tab:scenlitterpnv}.
-
- \begin{table*}[h]
- \caption{Scenarios for sensitivity analysis on litter parameterization}
- \begin{tabular}{l l l l}
- \tophline
-  Scenario name & plant type & N:C & LG:C \\
-  \middlehline
-  \textit{LitterPNV-Mean} & mix over all tree types & 0.015  & 0.25 \\
-  \textit{LitterPNV-Grass} & perennial grasses     & 0.029  & 0.11 \\
-  \textit{LitterPNV-AASS}  & Arctic alpine subshrub           & 0.022  & 0.22 \\
-  \textit{LitterPNV-BLEAF} & Temperate broadleaf evergreen    & 0.013  & 0.27 \\
-  \textit{LitterPNV-BRCF}  & Boreal coniferous forest         & 0.013  & 0.17 \\
-  \textit{LitterPNV-MTCF}  & Maritime coniferous forest       & 0.0094 & 0.22 \\
-  \textit{LitterPNV-TREF}  & Tropical evergreen forest        & 0.016  & 0.30 \\
- \bottomhline
- \end{tabular}
- \belowtable{}
- \label{tab:scenlitterpnv}
- \end{table*}
-
-\newpage
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in \citep{karstens_model_2020}.
-In the following, the most important results (see \citep{karstens_result_2020} for post-processing script) are summarized.
-
-\hypertarget{soc-distribution-and-depletion}{%
-\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}
-
-\begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
-Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
-}\label{fig:SOCmaps}
-\end{figure}
-
-In Fig.
-\ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
-Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways:
-1.
-Absolute SOC changes \(\Delta SOC\) (see Fig.
-\ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses.
-They might be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
-In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
-2.
-Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig.
-\ref{fig:SOCmaps}(c)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
-Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
-
-The global SOC debt has decreased by about 15\% in the period between 1975 and 2010 to \(22\unit{GtC}\) (Fig.
-\ref{fig:SOCmaps}(d)).
-This corresponds to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\).
-Considering our estimate of the global SOC stock of around \(660\unit{GtC}\) in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
-
-\hypertarget{carbon-flows-in-the-agricultural-system}{%
-\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
-
-C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
-Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
-We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.~like food, bioenergy and material) into a human demand category.
-Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
-Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not recycled to soils.
-Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
-However, around 60\% of C decomposes before its integrated into soils at the the litter-soil barrier.
-Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
-According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
-4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
-Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
-
-\begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
-\end{figure}
-
-\hypertarget{agricultural-management-effects-on-soc-debt}{%
-\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}
-
-We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)).
-Without changes in management regimes, the global \(\Delta SOC\) on cropland would be increasing at a rate of \(0.1\unit{GtC yr^{-1}}\).
-As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global \(\Delta SOC\) would still increase at a rate of \(0.06\unit{GtC yr^{-1}}\).
-Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both).
-The effect of no-tillage has been particuarly strong since the 1990s.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \ref{fig:SOCscen}(b).
-
-\begin{figure}
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
-\end{figure}
-
-Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\ref{fig:SOCinit}).
-The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.
-
-\begin{figure}
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
-\end{figure}
-
-Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas \(\Delta SOC\) vary 19-24 GtC for the year 2010 (see Fig.
-\ref{fig:SOClitter}.
-The increase of \(\Delta SOC\) during the period of 1975--2010 is independent of the parameterization choice at \textasciitilde4 GtC.
-
-\newpage
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions.
-This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks.
-While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.
-
-\hypertarget{soc-debt-and-soc-drivers-in-literature}{%
-\subsection{SOC debt and SOC drivers in literature}\label{soc-debt-and-soc-drivers-in-literature}}
-
-The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study.
-However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.
-
-According to \citep{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
-Our estimate of 22 GtC in 2010 for cropland debt is higher as \citep{sanderman_soil_2017} estimations.
-However, there are large uncertainties in modeling SOC at the global scale, and \citep{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
-This suggests that our results are within a plausible range.
-
-Furthermore, \citep{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-\citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change).
-Historic productivity increases were, however, notably larger.
-Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).
-
-Our study for the first time uses a dynamic management dataset as driver for SOC dynamics.
-We show that the moderate global cropland expansion of around 11\% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices.
-This is in contrast to \citep{pugh_simulated_2015} findings of only small effects due to improved practices.
-
-Looking on historic cropland intensification rates of above 50\% (\citep{rudel_agricultural_2009}), a large fraction of increased C input from residue biomass is attributed to productivity improvements.
-
-\hypertarget{limitations-and-uncertainties}{%
-\subsection{Limitations and uncertainties}\label{limitations-and-uncertainties}}
-
-Modeling management effects at the global scale comes with parametric and structural uncertainties.
-High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g.~the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs.
-As pointed out by \citep{keel_large_2017} and \citep{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
-\citep{keel_large_2017} question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
-Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
-However, looking on the evaluation of management effects (see Sect.
-\ref{sec:ipcccompare}), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions.
-Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g.~cover- and inter-cropping practices.
-
-Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965.
-As shown in Fig.
-\ref{fig:SOCinit} the SOC debt estimate almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), and the SOC debt reduction is strongly reduced (from \textasciitilde4 GtC to \textasciitilde2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation.
-\citep{pugh_simulated_2015} pointed out the importance of accounting for the land-use history, as many CO\textsubscript{2} emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium.
-Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.
-
-Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines (\citep{eggleston_ipcc_2006}, \citep{calvo_buendia_ipcc_2019}) and assumes that most of the SOC dynamic happens in the topsoil.
+Generally, limiting the analysis to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamics happen in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected.
-As \citep{powlson_limited_2014} have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-No-tillage effects may seem larger than the actually are, if only focusing on the topsoil.
-SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
-Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
+Second, changes to the subsoil due to tillage are neglected.
+As Powlson et al.
+[-@powlson_limited_2014] have shown, the subsoil can be make a large difference in evaluating total SOC losses or gains for no-tillage systems.
+No-tillage effects may seem larger than they actually are if only topsoil is considered.
+SOC transfers to deeper soil layers under tillage might enhance subsoil SOC compared to no-till practices.
+Finally, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
 
-This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by \citep{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions (\citep{wang_human-induced_2017}).
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
-Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry.
-Human interventions have led to large changes in SOC stocks there as well (\citep{sanderman_soil_2017},\citep{friedlingstein_global_2019}).
-This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.
+## SOC debt and SOC drivers
 
-\hypertarget{sec:ipcccompare}{%
-\subsection{Modeled management effect in line with default IPCC assumptions}\label{sec:ipcccompare}}
+The analysis of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
+There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period [e.g.
+@frielingstein_glob_carbon_2020].
+There is uncertainty with respect to the overall size of these emissions from different methods and reference points and with respect to the contribution of cropland and agricultural management to these emissions.
+In order to mitigate greenhouse gas emissions, it is essential to stop the decline of SOC stocks or even transform cropland management to sequester atmospheric C in cropland soils [@minasny_soil_2017].
+Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], we find that global cropland has acted as a emissions source since 1975.
+Annual C loss rates of 0.2 per 1000 C still have the opposite trend as the promoted 4 per 1000 C sequestration rate target [@minasny_soil_2017].
+Dedicated efforts to increase cropland SOC are thus necessary, as management improvements at historical rates are not enough to counteract ongoing SOC degradation on cropland.
+Yet our study also shows the substantial impact of changing management on the development of SOC debt (Fig.
+\@ref(fig:SOCscen)).
 
-To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 (\citep{eggleston_ipcc_2006} and their refinements in 2019 \citep{calvo_buendia_ipcc_2019}).
-Both estimates are based on measurement data for croplands (see Table \ref{tab:SCFglo}).
-To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig.
-\ref{fig:CLIMzone}).
+According to Sanderman et al.
+[-@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
+Our estimate of 39.6 GtC in 2010 for cropland debt is thus twice as high as their estimate.
+However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.
+[-@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results.
 
-\% latex table generated in R 3.6.3 by xtable 1.8-4 package
-\% Mon Dec 14 23:33:09 2020
+Furthermore, Sanderman et al.
+[-@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+Pugh et al.
+[-@pugh_simulated_2015] considered management effects like tillage and incorporation of residues in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
+Their study moreover concludes that yield gains (by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
+Historical yield increases, however, are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
+While we find substantially larger effects of productivity gains than the 1% reported by Pugh et al.
+[-@pugh_simulated_2015], this is not sufficient to compensate SOC losses from moderate global cropland expansion of around 11% between 1974 and 2010.
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlllllll}
-  \hline
- & Source & Input & Year & tropical moist & tropical dry & temperate dry & temperate moist \\ 
-  \hline
-1 & IPCC2006 & low & invariant & 0.44 & 0.55--0.61 & 0.74 & 0.66 \\ 
-  2 & IPCC2006 & medium & invariant & 0.48 & 0.58--0.64 & 0.80 & 0.69 \\ 
-  3 & IPCC2006 & high & invariant & 0.53 & 0.60--0.67 & 0.83 & 0.77 \\ 
-  4 & IPCC2019 & low & invariant & 0.76 & 0.87 & 0.70--0.71 & 0.66--0.67 \\ 
-  5 & IPCC2019 & medium & invariant & 0.83 & 0.92 & 0.76--0.77 & 0.69--0.70 \\ 
-  6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
-  7 & This Study & hist & 1990 & 0.57 & 0.61 & 0.78 & 0.76 \\ 
-  8 & This Study & hist & 2010 & 0.64 & 0.68 & 0.83 & 0.83 \\ 
-   \hline
-\end{tabular}
-\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general.
-For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019.
-leaving our results in very good agreement with the old default assumptions.
-Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\mathrm{SCF}}$ by additional 5-8 percent.
-Also modelled $F^{SOC}$ have increased for all climates over time.}
-\label{tab:SCFglo}
-\end{table}
+The effects of agricultural productivity on cropland SOC dynamics, including historical yield trends and associated increases in residue inputs, can be directly accounted for in our modeling approach.
+In contrast, process-based studies [@pugh_simulated_2015; @herzfeld_soc_2021] often lack data on relevant management aspects that drive production increases.
+[@herzfeld_soc_2021] also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
+Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
+This is consistent with our finding that increasing SOC inputs from above-ground residues had the strongest effect on the slowing-down of the SOC debt increase (Fig.
+\@ref(fig:SOCscen)).
 
-Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than \citep{calvo_buendia_ipcc_2019}.
-The tropical factors differ substantially between the IPCC guidelines in 2006 (\citep{eggleston_ipcc_2006}) and their refinement in 2019 (\citep{calvo_buendia_ipcc_2019}), with our estimates within this range.
-Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially.
-Our results still show especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks.
-Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig.
-\ref{fig:SOCmaps}(c)) may be expected.
+Elliott et al.
+[-@elliott_management_trends_2018] show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, and genetic traits such as kernel number and radiation use efficiency.
+As such, it will remain challenging for process-based DGVMs to capture the trend of agricultural productivity on cropland SOC dynamics.
 
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.
+Our study emphasizes again that the expansion of cropland is still a major source of CO~2~ emissions --- not only through the removal of vegetation, but also by a slow depletion of C stocks in soils.
+Our estimates indicate a SOC debt of 39.6 GtC in 2010, and every additional deforestated hectare adds to this debt.
+Avoided deforestation and other environmental regulation leads to intensification on existing cropland [@humpenoder_large-scale_2018] and our results show that such intensification could lead to increased cropland SOC, if residues are returned to the soil --- amplifying the C sequestration potential of avoided deforestation.
 
-\hypertarget{soc-stocks-in-line-with-literature}{%
-\subsection{SOC stocks in line with literature}\label{soc-stocks-in-line-with-literature}}
-
-The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile (\citep{batjes_harmonized_2016}, \citep{hengl_soilgrids250m_2017}, \citep{fao_global_2018}, \citep{schaphoff_lpjml4_2018-1}) in Table \ref{fig:SOCglo}).
-
-\begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time.
-This makes it hard to pinpoint a specific year for these SOC estimations.
-In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
-\end{figure}
-
-The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates.
-SoilGrids (\citep{hengl_soilgrids250m_2017}) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar.
-Looking at regional results in Fig.
-\ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas.
-To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland.
-Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.
-
-Additionally, our estimates for total SOC stocks of the world (as well as our SOC initialization) are dominated by the representation of natural vegetation, which are only estimated in a basic manner.
-For example, we do not differentiate the parameterization of nitrogen and lignin content of litterfall for woody and grass plant types.
-This renders carbon inputs and decay dynamics for natural litterfall rather uncertain.
-The absolute values of SOC stocks and debt from land-use change have to be interpreted with caution.
-As our default litter parameterization accounts for woody plant types, larger uncertainty in natural land SOC dynamics may arise especially in less forested areas.
-
-We conducted a sensitivity analysis (Fig.
-\ref{fig:SOClitter}) based on various plant parameterizations from the Century model (see Sect.
-\ref{sec:scenlitterpnv}).
-This shows that the general trend of decreasing SOC debt of \textasciitilde4 GtC within the period of 1975--2010 is not altered under various estimates for natural SOC stocks.
-\newpage
-
-\conclusions
-
-Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management.
-It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.
-
-Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models.
-That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
-Our data-set and the MADRaT package (\citep{dietrich_madrat_2020}) constitute a starting point for building comprehensive data sets on agricultural management aspects.
-
-Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions -- not only by the removal of vegetation, but also by a slow depletion of soils.
-Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.
-
-However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975.
-Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils.
-The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source.
-Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO\textsubscript{2}.
-The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.
-
-Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks.
-Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand (\citep{popp_land-use_2014-1}) -- and thereby usually further reduce the SOC debt on croplands.
-
-While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets.
-Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
-
-Yet, there is ample potential for further improved SOC management.
+There is also ample potential for further improved SOC management.
 As shown in Fig.
-\ref{fig:FlowFig}, approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC).
+\@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
+However, even larger losses occur at the end of the food supply chain (1.2 GtC year), at the soil surface (1.4 GtC), during residue burning (0.3 Gt C) and with manure management (0.2).
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils (\citep{brenzinger_organic_2018}).
-Secondly, soil carbon sequestration techniques (\citep{smith_soil_2016}), deep ploughing (\citep{alcantara_deep_2016}) or the transformation of c inputs to more recalcitrant biochar (\citep{woolf_sustainable_2010}) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Waste composting or excreta recycling could represent a major additional C input to cropland soils [@brenzinger_organic_2018].
+Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of C inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
 Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
-Finally, other carbon-accumulating practices, such as the cultivation of cover crops (\citep{poeplau_carbon_2015}) and agroforestry (\citep{lorenz_soil_2014}) could increase total C sequestration on croplands.
-
-\codedataavailability{We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrsoil_2020} for soil dynamic related functions and github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data.
-All libraries are based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing.
-Model results including C input data are accessable under \url{https://doi.org/10.5281/zenodo.4320663}.
-Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.} %% use this section when having data sets and software code available
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Figures and tables in appendices}
-
-\subsection{Methods}
-
-\begin{figure}
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
-\caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data.
-Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
-\label{fig:CLIMzone}
-\end{figure}
-
-\begin{table*}[h]
-\caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
-Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
-}
-\begin{tabular}{lllllllllllll}
-\tophline
-
-         &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
-Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
- \middlehline
-tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
-maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
-trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
-rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
-soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
-rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
-groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
-sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
-oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
-puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
-potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
-cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
-sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
-sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
-others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
-foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
-cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
-\bottomhline
-         &                 & \multicolumn{8}{l}{
-                                           \begin{minipage}[t]{0.50\columnwidth}
-                                                      nr/dm -- nitrogen to dry matter ratio\\
-                                                      wm/dm -- wet matter to dry matter ratio\\
-                                                      c/dm  -- carbon to dry matter ratio\\
-                                                      \end{minipage}} &
-                                                                        \multicolumn{3}{l}{
-                                                                            \begin{minipage}[t]{0.50\columnwidth}
-                                                                                        HI\textsuperscript{area} -- harvest index per area\\
-                                                                                        HI\textsuperscript{prod} -- harvest index per production\\
-                                                                                        RS -- root:shoot ratio \end{minipage}} 
-\end{tabular}
-\label{tab:c2dm}
-\end{table*}
-
-\subsection{Discussion}
-
-\begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
-\caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
-In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010.
-PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model.
-We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
-\label{fig:SOCreg}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-\caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}.
-\textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
-Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
-\label{fig:SOClitter}
-\end{figure}
-\noappendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{KK, BLB and AP designed the study and the model idea.
-KK wrote the code build on work of BLB, IW.
-JPD revised and improved the model code.
-CM, JH and SR provided the LPJmL simulation data.
-KK wrote the paper with important contributions of BLB and CM.
-MK, JS, SR and IW provided extensive feedback to outline of the study.
-All authors discussed the results and commented on the manuscript.} %% optional section
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing interests.} %% this section is mandatory even if you declare that no competing interests are present
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgements}
-Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
-Additional thanks to the rticles contributors for providing a R Markdown template.
-The authors thank for the data provided by FAOSTAT and LUH2v2.
-The work of KK was funded by the DFG Priority Program ``Climate Engineering: Risks, Challenges, Opportunities?'' (SPP 1689) and specifically the CEMICS2 project (grant no.
-ED78/3-2).
-The research leading to these results has received funding for BLB from the European Union's Horizon 2020 research and innovation program under grant agreement no.
-776479 (COACCH) and no.
-821010 (CASCADES).
-The work of SR, JS and IW was also supported by CLIMASTEPPE (01DJ8012), EXIMO (01LP1903D) and FOCUS (031B0787B) funded by the German Federal Ministry of Education and Research (BMBF).
-The input of PS, MK and MD contributes to the Soils-R-GGREAT project (NE/P019455/1) and CIRCASA (EU H2020; grant agreement no.
-774378).
-\end{acknowledgements}
-
-%% REFERENCES
-%% DN: pre-configured to BibTeX for rticles
-
-%% The reference list is compiled as follows:
-%%
-%% \begin{thebibliography}{}
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL1}
-%% REFERENCE 1
-%%
-%% \bibitem[AUTHOR(YEAR)]{LABEL2}
-%% REFERENCE 2
-%%
-%% \end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{SOCbudget.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-\end{document}
+Finally, other carbon-accumulating practices, such as the cultivation of cover crops [@poeplau_carbon_2015; @porwollik_covercrops_2022] and agroforestry [@lorenz_soil_2014] could increase total C sequestration on cropland.

### 2022-02-28 SOCbudgetPaper2/introduction.Rmd
Major update of finalized Paper - Draft 2

\introduction

Soil organic carbon (SOC), the amount of organic carbon stored in the Earth’s soil, constitutes the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996]. Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@frielingstein_glob_carbon_2020; @minasny_soil_2017]. The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile [@batjes_total_1996; @sanderman_soil_2017].

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018]. However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological driving forces, human interventions alter terrestrial carbon pools over much shorter time scales and are currently one of the most dominant drivers of SOC changes on managed land [@hansis_relevance_2015; @bastos_comparison_2021].

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al. [-@pongratz_terminolgy_2014]), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state. Sanderman et al. [-@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter [@lal_world_2001].

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs), Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@frielingstein_glob_carbon_2020; ]. While providing estimates of the magnitude of SOC losses due to land-cover change, most models lack a detailed representation of agricultural management. Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015]. BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO~2~ fertilization and N deposition. Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management [@frielingstein_glob_carbon_2020; @houghton_carbon_2012; @hansis_relevance_2015; @bastos_comparison_2021].

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance [e.g. @bondeau_modelling_2007; @lindeskog_implications_2013]. Pugh et al. [-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons [@portmann_mirca2000global_2010], multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g. @pugh_simulated_2015, @lutz_simulating_2019]. Herzfeld et al. [-@herzfeld_soc_2021] account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC. Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.

More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g. on tillage systems, see [@porwollik_generating_2019], [@prestele_spatially_2018]) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g. [@iizumi_modeling_2019], [@minoli_modelling_2019]), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models [@del_grosso_simulated_2001, @coleman_simulating_1997, @smith_estimating_2010, @taghizadeh-toosi_c-tool_2014] are able to better account for historical agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge [@morais_detailed_2019].

Managed soils have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (soil carbon sequestration (SCS) techniques). However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) [@popp_land-use_2016; @rogelj_scenarios_2018] have not integrated soil management into their mitigation pathways. More detailed process-based models are typically computationally too demanding to be integrated into optimization-based IAMs. Better accounting for soil carbon management in IAMs thus requires a light-weight model suitable for iterative modeling with detailed options to represent agricultural soil management.

The objectives of our study are (1) to develop a reduced-complexity SOC model able to account for SCS in IAM frameworks; (2) to create a comprehensive data set of the global gridded management time series, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices; and (3) to provide global as well as spatially explicit SOC and SOC debt estimates that consider spatially explicit and time-variant agricultural management. 
We decompose the contribution of different management activities through a scenario analysis, identifying the most impacting management decisions for SOC development. Moreover, we compare our model performance against other SOC stock and SOC emission estimates, to evaluate the suitability of this reduced-complexity approach for integration into IAM modeling.

### 2022-02-28 SOCbudgetPaper2/methods.Rmd
Major update of finalized Paper - Draft 2

--- 2021-08-17 SOCbudgetPaper/methods.Rmd
@@
----
-output:
-  pdf_document: default
-  html_document: default
----
 # Methods
@@
 In Sect.
-\@ref(sec:carbonbudget) we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.
-4 [@calvo_buendia_ipcc_2019].
-We additionally describe how we configured and extended the steady-state method (for model code see [@karstens_mrsoil_2020].
+\@ref(sec:carbonbudget) we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail in the refinement of the IPCC guidelines vol.
+4 Chapter 5 on Cropland [@ogle_cropland_in_ipcc_2019].
+We additionally describe how we configured and extended the Tier 2 modeling approach [for model code see @karstens_mrsoil_2020].
 In Sect.
@@
 In Sect.
-\@ref(sec:scenarios) we define the scenarios used to complement our historic model results.
+\@ref(sec:scenarios) we define the management scenarios used to analyze the role of different management aspects in historical cropland SOC dynamics.
 
-## SOC stocks and stock changes following the steady-state method {#sec:carbonbudget}
+## SOC stocks and stock changes following the Tier 2 modeling approach {#sec:carbonbudget}
 
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.
-4 [@calvo_buendia_ipcc_2019]; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon ($SOC$) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+Following the Tier 2 modeling approach of the refinement of the IPCC guidelines vol.
+4 Chapter 5 on Cropland [@ogle_cropland_in_ipcc_2019]; referred to as \textit{Tier 2 modeling approach} in the following), we estimate SOC stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
 We restrict our analysis to the first 0-30 cm of the soil profile.
@@
 This equation is valid for all grid cells $i$ and all years $t$.
-We use the steady-state method for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987].
+We use the Tier 2 modeling approach for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987].
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
@@
 
-Carbon inputs for cropland are below- and above-ground residues left or returned to the field (see Sect.
+Carbon inputs for cropland are below- and above-ground crop residues left or returned to the field (see Sect.
 \@ref(sec:residues)) and manure inputs (see Sect.
-\@ref(sec:livstmanure)); for natural vegetation litterfall including fine root turnover [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
-Following the IPCC guidelines [@calvo_buendia_ipcc_2019], carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
+\@ref(sec:livstmanure)); for natural vegetation, litterfall including fine root turnover [@schaphoff_lpjml4_2018] is the only source of carbon inflow to the soil.
+Following the IPCC guidelines [@ogle_cropland_in_ipcc_2019], carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content.
 For each component the sum of all carbon input sources is allocated to the respective $SOC$ sub-pools via transfer coefficients.
 This implies that both the amount of carbon and its structural composition determine the effective inflow into the different pools.
-Data sources for all considered carbon inputs as well as for lignin and nitrogen content are listed in Table \@ref(tab:datasourceinputs).
+
+Whereas residue and manure default lignin and nitrogen fractions are given by the IPCC guidelines [@ogle_cropland_in_ipcc_2019], we use plant-functional type and plant-organ specific parameterization for the natural litterfall.
+Global distribution of plant-functional types is given by [@schaphoff_lpjml4_2018] as well as separation of litter into leaf, fine root and wood litter compartments excluding litter biomass burnt in wild fires.
+Leaf litter parameters are given by Brovkin et al.
+[-@brovkin_plant-driven_2012], fine root to leaf litter lignin ratio by Guo et al.
+[-@guo_leaf_root_decompostion_2021], lignin content of wood litter by Rahman et al.
+[-@rahman_lignin_2013] and nitrogen content scaling factors for leaf to fine roots and leaf to wood litter by von Bloh et al.
+[-@vBloh_lpjml5_2018].
+Data sources for all considered carbon inputs as well as for lignin and nitrogen parameterization are listed in Table \@ref(tab:datasourceinputs).
 
@@
     \begin{table*}[h]
-    \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
+    \caption{type and data sources for carbon inputs and parameterization to different land-use types }
     \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
@@
     \middlehline
-    \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
+    \multirow{3}{*}{cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                         \citet{faostat_faostat_2016}, \\
@@
                            LG:C generic values according to Table 5.5B,\\
-                                                                        5.5C from IPCC \citep{calvo_buendia_ipcc_2019},\\
+                           5.5C from IPCC \citep{ogle_cropland_in_ipcc_2019},\\
                            crop-specific N:C from \citet{bodirsky_n2o_2012}
@@
                               \hline
-    Natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
+    natural vegetation        & annual litterfall &  \citet{schaphoff_lpjml4_2018}
                         & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
-                                  IPCC \citep{calvo_buendia_ipcc_2019} and\\ 
-                                  CENTURY (\citep{century_model_2000}) \strut \end{minipage}\tabularnewline
+                            leaf N and LG concentration from \citet{brovkin_plant-driven_2012}, 
+                            root to leaf litter LG ratio \citet{guo_leaf_root_decompostion_2021},  
+                            lignin content of wood litter \citet{rahman_lignin_2013} and 
+                            nitrogen scaling factors for leaf to root and wood litter 
+                            from \citet{vBloh_lpjml5_2018} \strut \end{minipage}\tabularnewline
    \bottomhline
@@
 
-The sub-pool specific decay rates $k_{sub}$ are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space ($i$) and time ($t$).
-Following the steady-state method [@calvo_buendia_ipcc_2019], we consider temperature ($temp$), water ($wat$), sand fraction ($sf$) and tillage ($till$) effects to account for spatial variation of decay rates.
-Thus, $k_{sub}$ are given by
+The sub-pool specific decay rates $k_{sub}$ are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over space $i$ and time $t$.
+Following the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019], we consider temperature $temp$, water $wat$, sand-fraction $sf$, and tillage $till$ effects to account for spatial and temporal variation of decay rates.
+Thus, $k_{sub}$ rates are given by
 
@@
 \@ref(sec:irrigation)) practices on decay rates.
-We calculated area weighted means for $till$ and $wat$ on cropland for each grid cell, using area shares for the different tillage and irrigation pratices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \@ref(tab:datasourcedecay); equations are displayed by equation 5.0B--5.0F in Calvo Buendia et al.
-[-@calvo_buendia_ipcc_2019].
+We calculate area-weighted means for $till$ and $wat$ on cropland for each grid cell, using area shares for the different tillage and irrigation practices.
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \@ref(tab:datasourcedecay); equations are displayed by equations 5.0B--5.0F in Ogle et al.
+[-@ogle_cropland_in_ipcc_2019].
 
@@
  \begin{table*}[h]
- \caption{Type and data sources for carbon inputs to different land-use types}
+ \caption{type and data sources for carbon inputs to different land-use types}
  \begin{tabular}{l l l l}
@@
 
-                      & Mircobial activity & air temperature & \citet{harris_version_2020} \\
+                      & mircobial activity & air temperature & \citet{harris_version_2020} \\
                       \cline{2-4}
 
-                      & Water restriction & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
+                      & soil moisture & precipitation \& potential evapotranspiration & \citet{harris_version_2020} \\
                       \cline{1-4}
-\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & Sect. @ref(sec:irrigation) \\ 
+\multirow{2}{*}{\begin{minipage}[t]{0.15\columnwidth}\raggedright\strut cropland\\(additionally)\strut\end{minipage}} & soil moisture*  & irrigation  & Sect. @ref(sec:irrigation) \\ 
                       \cline{2-4}
 
-                      & Soil disturbance & tillage & Sect. @ref(sec:irrigation) \\
+                      & soil disturbance & tillage & Sect. @ref(sec:tillage) \\
  \bottomhline
@@
 
-```{r captionLUtrans, echo=FALSE}
-captionLUtrans <- "Scheme of land-use transition representation for a simplified case.."
+```{r captionLUtrans, echo=FALSE, include=FALSE}
+captionLUtrans <- "Scheme of land-use transition representation. 
+Given an initial land-use pattern (as in this example 2 ha land under natural vegetation and 1 ha of cropland), there are separate SOC stocks for natural vegetation and cropland.
+While in this example we assume SOC under natural vegetation to be in steady state, the cropland SOC stock approaches its steady state without having reached it yet (a).
+Upon cropland expansion (in this example half of the natural vegetation is cleared to be used as cropland), SOC stocks on cropland increase due to a transfer of land from natural vegetation (b).
+Separately accounting for SOC dynamics on newly converted and existing cropland (c) leads to the same overall SOC dynamics as if SOC stocks are averaged (d), due to the linearity of Eq.
+\ref{eq:SOCstate} and the cropland-age independend decay rates (see Eq.
+\ref{eq:decayrates})"
 ```
 
-```{r LUtrans, out.width = "16cm", fig.cap = captionLUtrans, echo=FALSE, fig.pos="h"}
+```{r LUtrans, out.width = "8cm", fig.cap = captionLUtrans, echo=FALSE, fig.pos="h"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images//LUtransition.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/LUtransitions.png")
 ```
 
-We calculate $SOC$ stocks based on the area shares of land-use types ($lu$) within our grid cells ($i$).
+We calculate $SOC$ stocks based on the area shares of land-use types $lu$ within our grid cells $i$.
 If land is converted from one land-use type $lu=\{crop,natveg\}$ into the other $!lu=\{natveg,crop\}$, a respective share of the $SOC$ is reallocated within our budget.
-As we do not distinguish between newly converted and old cropland (see \@ref(fig:lutransitions)), we update the average SOC stock per area within this modeling step.
+We do not distinguish between newly converted and existing cropland, but can work with the average carbon content as the relative decay of SOC is proportional to the SOC stock (see \@ref(fig:LUtrans)).
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{t^*}$ via
@@
 
-$SOC$ converge towards the calculated steady-state stock $SOC^{\mathrm{eq}}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like
+$SOC$ converges towards the calculated steady-state stock $SOC^{\mathrm{eq}}$ for each grid cell $i$, each annual time step $t$, each land-use type $lu$ and each sub-pool $sub$ like
 
@@
 According to the IPCC guidelines $SOC$ changes can be expressed as the difference of two consecutive years [see Eq.
-5.0A in @calvo_buendia_ipcc_2019].
+5.0A in @ogle_cropland_in_ipcc_2019].
 This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we defined the absolute and relative SOC changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
+For our study, we define the absolute and relative SOC changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
 The absolute changes $\Delta SOC$ and relative changes $F^{\mathrm{SCF}}$ are thus given by
@@
 
-We initialize our SOC sub-pools using a three-step approach, since input data availability is limited for climate and litter (from 1901) as well as for agricultural management data (from 1965):
+The initialization of SOC pools is very important and has to include the proper accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historical land-use change (LUC) and the slow decline of SOC under crop cultivation, before it reaches a new equilibrium.
+We initialize our SOC sub-pools using a three-step approach, since input data availability is limited for climate and litter estimates (starting only in 1901) as well as for agricultural management data (starting only in 1965):
 
-Firstly, to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we account for land-use change from 1510.
-We repeat climate and subsequently also litter input data from 1901-1930 to simulate a constant climate for this first initializing period.
-Agricultural management data is hold constant at the level of 1965.
-In 1510 we assume all SOC pools to be in natural steady-state, implicating that all land-use change emissions prior to that time, are redistributed to 1510.
-Due to this long first spin-up period we suppose that the effect of this late accounting and appertaining delayed legacy fluxes can be neglected in 1901.
-This approaches follows others studies looking on effects of land-use change and management (cite LPJmL (which?)).
+Firstly, in order to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we consider land-use change from 1510 onwards.
+In 1510, we assume all SOC pools to be in natural steady-state, implying that all land-use change prior to that time occurs in 1510.
+We assume that by 1901, all cropland converted in 1510 has reached its new steady state, so that it is not necessary to explicitly account for even older land conversion.
+Model inputs for 1901--1930 for climate and natural vegetation litterfall are repeated for 1510--1900 to mimic constant climate conditions for this first initializing period.
+Similarly, agricultural management data are held constant at the level of 1965 until 1965.
+This approach follows others studies looking on effects of land-use change and management [e.g.
+@schaphoff_lpjml4_2018-1; @herzfeld_soc_2021].
 
-Secondly, to account for the impact of a changing climate, we account not only for land-use change, but also for historical climate and consequently natural litter inputs to the soil, from 1901 to 1965 still considering constant agricultural input data.
+Secondly, with the availability of transient climate data after 1901, we account not only for land-use change, but also for historical climate change and consequently natural litter inputs to the soil from 1901 to 1965 still considering constant agricultural input data, which are not available prior to 1965.
 
-Thirdly, we run the model for additional 10 years with historic dynamic input data and start analyzing results from 1975 onward.
-This is in line with IPCC method suggestion to have a 5-20 year spin-up period.
+Thirdly, we run the model for 10 years from 1965 to 1975 with historical dynamic data on agricultural management and start analyzing results from 1975 onward.
+This is in line with the IPCC guidelines vol.
+4 method suggestion to have a 5-20 year spin-up period [@ogle_cropland_in_ipcc_2019].
 
-Irrigation areas are part of the water effect associated with the climate data set and therefore dynamic from 1901 on, whereas data on no-tillage areas is only available after 1974.
+With transient climate considered after 1901, decay rates $k_{sub}$ become dynamic in time.
+As the decay rates are also affected by irrigation and tillage (see Sect.
+\@ref(sec:steadystates)), we also account for transient changes in irrigated areas after 1901.
+Data on no-tillage practices are only available after 1974 and we assume conventional tillage on all cropland prior to 1975.
 
-<!-- backup
-```{=latex}
- \begin{table*}[h]
- \caption{Setup of spin-up and simulation phase}
- \begin{tabular}{l l l}
- \tophline
- \multirow{3}{*}{1901--1964} 
-   & Constant management spin-up phase with
-   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
-   & constant management data from 1965 (manure and residue inputs, tillage and irrigation areas)\\
-                      \cline{1-4}
- \multirow{3}{*}{1965--1974}
-   & Dynamic management spin-up phase with
-   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
-   & historic management data (manure and residue inputs, tillage and irrigation areas)
-                      \cline{1-4}
- \multirow{3 }{*}{1975--2010}                     
-   & Simulation phase 
-   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
-   & historic management data (manure and residue inputs, tillage and irrigation areas)              
- \bottomhline
- \end{tabular}
- \belowtable{}
- (\#tab:datasourcedecay)
- \end{table*}
-``` -->
 ## SOC stocks and stock changes following Tier 1 {#sec:tier1}
 
-Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019].
+Additionally to the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019].
 Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-Estimates of $F^{\mathrm{SCF}}$ are differentiated by different crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Estimates of $F^{\mathrm{SCF}}$ are differentiated by crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows.
 Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig.
@@
 We compile country-specific FAO production and cropland statistics [@faostat_faostat_2016] to a harmonized and consistent data set.
-The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1901-1974).
+The data is prepared in 5-year time steps from 1965 to 2010, which restricts our analysis to the time span from 1975 to 2010 (after a spin-up phase from 1510--1974).
 For all the following data, if not declared differently, we interpolate values linearly between the time steps and keep them constant before 1965.
@@
 
-Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups [see FAO2LUH2MAG_croptypes.csv in @bodirsky_mrcommons_2020], applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016].
-By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
+Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set (short LUH2), which we sum up from quarter-degree to half-degree resolution.
+We disaggregate the physical area (in Mha) of the five different cropland subcategories (c3ann: C3 annual crops, c3per: C3 perennial crops, c4ann: C4 annual crops, c4per: C4 perennial crops, c3nfx: C3 nitrogen-fixing crops) of LUH2 into our 17 crop groups [see FAO2LUH2MAG_croptypes.csv in @bodirsky_mrcommons_2020], applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016].
+By calculating country-specific multicropping factors $MCF$ using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
 Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
@@
 \@ref(sec:landuse)).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production [@faostat_faostat_2016].
-By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production [@faostat_faostat_2016].
+Note
 
 Crop residue production and management is based on a revised methodology of Bodirsky et al.
-[-@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modeling.
-Starting from crop yield estimates ($Y$) and respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific ratios for above-ground residues to harvested biomass ($r^{\mathrm{ag,prod}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{tC}}$), above-ground residues to harvested area ($r^{\mathrm{ag,area}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{ha}}$) and below-ground residues to above-ground biomass ($r^{\mathrm{bg}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{tC}}$) as follows
+[-@bodirsky_n2o_2012] and key aspects are explained as they play a central role in soil carbon modeling.
+Starting from crop yield estimates $Y$ and respective physical crop area $CA$, we estimate total above-ground $AGR$ and below-ground $BGR$ residue biomass (in tonnes) using crop group $cg$ specific ratios for above-ground residues to harvested biomass $r^{\mathrm{ag,prod}}_{cg}$ in $\unit{tC\,\unit{tC}^{-1}}$, above-ground residues to harvested area $r^{\mathrm{ag,area}}_{cg}$ in $\unit{tC\,\unit{ha}^{-1}}$ and below-ground residues to above-ground biomass $r^{\mathrm{bg}}_{cg}$ in $\unit{tC\,\unit{tC}^{-1}}$ as follows
 
@@
 \begin{aligned}
-AGR_{i,t,cg} & = CA_{i,t,cg} \cdot \left( Y_{i,t,cg} \cdot  r^{\mathrm{ag,prod}}_{cg}+  r^{\mathrm{ag,area}}_{cg} \right)
+AGR_{i,t,cg} & = CA_{i,t,cg} \cdot \left( Y_{i,t,cg} \cdot  r^{\mathrm{ag,prod}}_{cg} + MCF_{i,t} \cdot r^{\mathrm{ag,area}}_{cg} \right)
 \qquad & \textrm{and} \\
-BGR_{i,t,cg} & = (CP_{i,t,cg} + AGR_{i,t,cg}) \cdot r^{\mathrm{bg}}_{cg} \qquad                                            & \forall\quad cg, i, t.
+BGR_{i,t,cg} & = (CA_{i,t,cg} \cdot Y_{i,t,cg} + AGR_{i,t,cg}) \cdot r^{\mathrm{bg}}_{cg} \quad.
 (\#eq:resbiomass)
@@
 
-Following the IPCC guidelines, we split the above-ground residue calculations into a production ($r^{\mathrm{ag,prod}}$) and an area dependent ($r^{\mathrm{ag,area}}$) fraction [@hergoualch_n2o_in_ipcc_2019].
+Following the IPCC guidelines, we split the above-ground residue calculations into a yield-dependent slope ($r^{\mathrm{ag,prod}}$) and a positive intercept ($r^{\mathrm{ag,area}}$) fraction [@hergoualch_n2o_in_ipcc_2019].
+Residues biomass therefore increases under-proportionally with rising yields, reflecting a shifting harvest index of higher-yielding breeds.
 Deviating from Bodirsky et al.
-[-@bodirsky_n2o_2012] we use harvested instead of physical crop area to account for increased residue biomass due to multiple cropping and decreased residue amounts due to fallow land.
+[-@bodirsky_n2o_2012] we use harvested instead of physical crop area (denoted in Eq.
+\@ref(eq:resbiomass) by MCF described in Sect.
+\@ref(sec:landuse)) to account for increased residue biomass due to multiple cropping (multiple harvests with each lower yields) and decreased residue amounts due to fallow land.
 We assume that all $BGR$ are left in the soil, whereas $AGR$ can be burned or harvested for other purposes such as feeding animals [@weindl_livestock_2017], fuel or for material use.
@@
 A country-specific fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
-Following Smil [-@smil_nitrogen_1999] we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following Smil [-@smil_nitrogen_1999] we assume a burn share of 25% for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income for the periods before 1995.
+After 1995 we estimate a linear decline of burn shares to 10% for low-income countries and 0% for high-income countries till 2025 to account for recent increases in air pollution regulation.
+The estimated trends show good agreement with fire-satellite-image derived estimates by the Global Fire Database [@vanderWerf_fire_2017].
 Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_ipcc_2006].
@@
 
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(tab:c2dm)).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon and nitrogen to dry matter ratios (see \@ref(tab:c2dm)).
 
@@
 
-Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems 	at the present day a lot of the manure has to be stored and can be applied on croplands.
-We assume that manure application happens at its excretion place, so that the livestock distribution is the driving factor of the spatial pattern of manuring.
+Manure especially from ruminants is often excreted at pastures and rangelands, but due to the intensification of livestock systems a lot of the manure has to be stored and can be applied on cropland.
+We assume that manure is applied in close proximity to its excretion, so that the distribution of livestock is the driving factor for the spatial pattern of manure application.
 
@@
 Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.
-[-@calvo_buendia_ipcc_2019].
+[-@ogle_cropland_in_ipcc_2019].
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 [-@bodirsky_n2o_2012].
-We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.
+We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
 
@@
 
-The LUH2v2 [@hurtt_harmonization_2020] data set provides irrigated fractions for their cropland subcategories.
+The LUH2v2 [@hurtt_harmonization_2020] data set provides irrigated fractions for its cropland subcategories.
 We sum up irrigation area shares for all crop groups within a grid cell, and calculate the water effect coefficient $wat$ on decay rates using these shares to compute the weighted mean between rainfed and irrigated $wat$ factors.
 As a result $wat$ is the same for all crop groups within a grid cell.
-Furthermore, we suppose the irrigation effect to be present for all 12 months of a year in a grid cell including irrigated areas, since we do not have consistent crop group specific growing periods available.
+Furthermore, we suppose the irrigation effect to be present for all 12 months of a year, since we do not have consistent crop group specific growing periods available.
 This will lead to an overestimation of the irrigation effect.
@@
 In order to derive a spatial distribution of the three different tillage types specified by the IPCC --- full tillage, reduced tillage and no tillage ---, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historic global gridded tillage data set from Porwollik et al.
-[-@porwollik_generating_2018].
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historical global gridded tillage data set from Porwollik et al.
+[-@porwollik_generating_2019].
 This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO [-@fao_aquastat_2016] and half-degree resolution physical crop areas from Hurtt et al.
 [-@hurtt_harmonization_2020], applying the methodology of Prowollik et al.
-[-@porwollik_generating_2018] to identify potential no-tillage grid cells.
+[-@porwollik_generating_2019] to identify potential no-tillage grid cells.
 
-## Scenario definitions {#sec:scenarios}
-
-To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization and parameter choices, we perform a set of scenario runs.
-In the following sections we outline name and idea of these scenarios [for technical implementation see @karstens_mrsoil_2020].
-
-### Management scenarios {#sec:scen_management}
+## Management scenario definitions {#sec:scenarios}
 

### 2022-02-28 SOCbudgetPaper2/results.Rmd
Major update of finalized Paper - Draft 2

# Results

Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens [-@karstens_model_2020]. In the following, the most important results [see @karstens_result_2020] for post-processing script) are summarized.

## SOC distribution and depletion

```{r captionSOCmaps, echo=FALSE}
captionSOCmaps <- "Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data. Panel (a) shows global $\\Delta SOC$ between historical land use and potential natural vegetation (PNV). The distribution of total global SOC stocks for the first 30 cm on cropland for the year 2010 is depcited in panel (b). Absolute (c) and relative (cd SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains."
```

```{r SOCmaps, out.width = "18cm", fig.cap = captionSOCmaps, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/4panelfigure.png")
```

The global SOC debt has increased by about 14% in the period between 1975 and 2010 from $34.6$ to $39.6\unit{GtC}$ (Fig. \@ref(fig:SOCmaps)(a)). This corresponds to an average loss rate of $0.14\unit{GtC\,yr^{-1}}$ in comparison to a hypothetical potential natural vegetation (PNV) state. Considering our estimate of the global SOC stock of around $705\unit{GtC}$ in the upper $30\unit{cm}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010. The speed of this SOC loss has decreased towards the end of the modeling period.

In Fig. \@ref(fig:SOCmaps)(b) we provide a world map of SOC stocks estimates for the first $30\unit{cm}$ on cropland considering historical management data for the year 2010. Values range between over $100\unit{t\,ha^{-1}}$ in northern temperate cropland to less than $5\unit{t\,ha^{-1}}$ for arid and semiarid cropland.
Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under PNV in two complementary ways: 
1. Absolute SOC changes $\Delta SOC$ (Fig. \@ref(fig:SOCmaps)(c)) indicate areas with high importance for the global SOC loss. They can be driven by large relative changes (e.g. in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g. North-East Asia). 
2. Relative SOC changes measured as stock changes factors $F^{SCF}$ (Fig. \@ref(fig:SOCmaps)(d)) are a helpful metric to analyze the impact of human cropping activities. They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome by improved agricultural practices. Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation. Conversely, irrigated cropland at the border to dry, unsuitable areas worldwide shows a strong relative increase in SOC stocks.

## Carbon flows in the agricultural system

```{r caption_FlowFig, echo=FALSE}
caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in MtC). Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand. After accounting for losses within the agricultural system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues. Large parts of C, however, are mineralized on the field before entering the soil. Additionally, C is transferred to and from the global agricultural soil stock via land-use change between cropland and natural vegetation. Finally, SOC is mineralized and flows back to the atmosphere."
```

```{r FlowFig, out.width = "16cm", fig.cap = caption_FlowFig, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/CarbonBudget.png")
```

C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvest organ, above- and below-ground residues). Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues ($729\unit{MtC}$ in 2010) are directly returned to the field. We divide crop biomass usage into feed usage and aggregate all other usage types (e.g. food, bioenergy and material) into a human demand category. Livestock feed demand for crop organ harvest and above-ground residues of $1136\unit{MtC}$ is roughly equal to the human demand of $1129\unit{MtC}$. Whereas large parts of feed intake are returned to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g. compost, night soil and sewage) is not returned to soils. Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with $1350\unit{MtC}$ returned to the fields in 2010. However, around 60% of this organic C decomposes before it is integrated into soils. Due to the different composition of organic C, proportionally more C enters the slow pool from manure than from crop residue. 
According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of $257\unit{MtC}$ in 2010. The agricultural system receives $4585\unit{MtC}$ assimilated by crop plants and releases $3554\unit{MtC}$ mostly through respiration. Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around $33\unit{MtC}$ for the year 2010.

## Agricultural management effects on SOC debt

```{r captionSOCscen, echo=FALSE}
captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios. The stylized scenarios deviate from historical argicultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage). The scenario constManagement combines all three modifications. Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a hypothetical natural vegetation state. Panel (b) shows the carbon inflows from crop residue and manure."
```

```{r SOCscen, out.width="14cm", fig.cap = captionSOCscen, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_horiz.png")
```

We analyze the relative impact of different management practices by comparing the actual historical management scenario with counterfactual scenarios, where individual management practices (residues in constResidues, manure in constManure, tillage practices in constTillage, all three in constManagement) are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)). As shown by the difference between the constResidues scenario and the other counterfactuals, changes in residue return rates dominate the management effects. Without the historical increase in C inputs from residues to agricultural soils, the global $\Delta SOC$ would decrease to $41.7\unit{GtC}$ at a rate of $0.20\unit{GtC\,yr^{-1}}$ --- a 35% increase compared to $0.14\unit{GtC\,yr^{-1}}$ for the histManagement estimates. Both the constManure and constTillage scenarios show only small deviations from the historical argicultural management values with $0.15\unit{GtC\,yr^{-1}}$. The effect of no-tillage only becomes discernible from 2000 onwards. 
The large contribution of residues relative to manure also becomes visible when considering the annual C inputs of residues and manure to soils over a period of 35 years (Fig. \@ref(fig:SOCscen)(b)).

## Model evaluation {#sec:validate}

To evaluate our model results against reference data in five steps: (1) we compare our stock change factors (see Sect. \@ref(#sec:tier1)) to IPCC default assumptions [@lasco_cropland_in_ipcc_2006; @ogle_cropland_in_ipcc_2019]; (2) we compare our global (and climate-zone specific) total SOC stocks to other literature estimates; (3) we compare our results to point measurements. To evaluate the representation of our natural SOC stocks (4) we correlated LPJmL4 SOC stocks for PNV with our natural state SOC results on grid level; and (5) we do a similar correlation analysis for our modeled actual SOC stocks in comparison to the results of SoilGrids 2.0 [@poggio_soilgrids_2021], which accounts for actual land use too.

### Stock change factors compared to IPCC assumptions {#sec:ipcccompare}

To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors $F^{\mathrm{SCF}}$ of 2006 [@lasco_cropland_in_ipcc_2006] and their refinements in 2019 [@ogle_cropland_in_ipcc_2019]. Both estimates are based on measurement data for cropland (see Table \@ref(tab:SCFglo)). To allow for comparison, we aggregate our stock change factors weighted by grid-level $F^{\mathrm{SCF}}$ cropland area to derive average factors for the four IPCC climate zones (Fig. \@ref(fig:CLIMzone)).

```{r SCFglo, echo=FALSE, results = 'asis'}
table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
library(xtable)
print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \\citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \\citep{ogle_cropland_in_ipcc_2019}."), comment=FALSE, caption.placement = "top")
```
Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC. For the tropical regions the IPCC factors changed notably from the guidelines in 2006 [@lasco_cropland_in_ipcc_2006] to the update in 2019 [@ogle_cropland_in_ipcc_2019]. Our results are in good agreement with the 2006 IPCC factors. Modelled $F^{SCF}$ have increased or stayed constant for all climate zones over time (1975-2010).

### Global SOC stocks comparison

We compare our global SOC stocks with a wide range of global SOC stock estimates for the first 30 cm of the soil profile, using data from WISE [@batjes_harmonized_2016], SoilGrids [@hengl_soilgrids250m_2017], GSOC [@fao_global_2018], LPJmL4 [@schaphoff_lpjml4_2018-1], SoilGrids2 [@poggio_soilgrids_2021], and SOCDebtPaper [@sanderman_soil_2017] in Fig. \@ref(fig:SOCglo).

```{r captionSOCglo, echo=FALSE}
captionSOCglo <- "Modeled as well as observation-based estimates for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \\citep{schaphoff_lpjml4_2018-1} and estimates from \\citep{sanderman_soil_2017}. We show values of this study for the year 2010 accounting for the historical land-use dynamics as well as for an hypothetical PNV."
```

```{r SOCglo, out.width = "8cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
```

The global estimates of the total SOC stock of the upper $30\,cm$ from this study are in the middle of the wide range of other modeled or observation-based estimates. SoilGrids [@hengl_soilgrids250m_2017] especially stands out with its high estimate, whereas SoilGrid2.0 [@poggio_soilgrids_2021] marks the lower end. Regional results (Fig. \@ref(fig:SOCreg)) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.

### Point-based evaluation

We correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements (for data base see appendix of [@sanderman_soil_2017]).

```{r captionSOCpoint, echo=FALSE}
captionSOCpoint <- "Correlation between modeled and measured SOC stocks. Given the wide span between minimun and maximum measured SOC stocks within in a given cell, we correlated median values with our modeled results. Both cropland and areas with natural vegetation tend to be lower in our results than in the point measurements."
```

```{r SOCpoint, out.width = "10cm", fig.cap = captionSOCpoint, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scatterComparePointData_2010.png")
```

### Natural SOC stock comparison with LPJmL4

Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time. We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation. Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics. With our focus on cropland SOC dynamics, we compare only cells with more than $1000\unit{ha}$ of cropland (capturing 99.9% of global cropland area).

```{r captionSOCLPJmL4, echo=FALSE}
captionSOCLPJmL4 <- "Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line."
```

```{r SOCLPJmL4, out.width = "13cm", fig.cap = captionSOCLPJmL4, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png")
```
Spatial correlations of PNV SOC stock values are high (global $R2=0.81$), especially for dry climate zones (Fig. \@ref(fig:SOCLPJmL4)). For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results.

### Actual SOC stock comparison with SoilGrids 2.0

SoilGrids 2.0 [@poggio_soilgrids_2021] is a digital soil mapping approach that uses over 240 000 soil profile observations to produce high resolution soil maps including SOC stocks and estimates of their uncertainties. To evaluate the performance of our model at the global scale, we correlate SoilGrids 2.0 SOC stock values, which were aggregated to 0.5 degree resolution, to our estimates for the year 2010 in Fig. \@ref(fig:SOCSoilGrids2). To focus our comparison on cropland areas, we mask out grid cells with less than $1000\unit{ha}$ of cropland. Spatial correlation is moderate for tropical climate zones, whereas it is low for temperate moist areas. In tropical dry and temperate dry areas, we simulate also very low SOC values (below $10\unit{tC\,ha^{-1}}), which is not found in SoilGrids 2.0 whereas our modeled SOC stocks can be substantially higher in temperate moist areas than reported by SoilGrids 2.0. 
Additionally, we use the uncertainty estimates from SoilGrids 2.0 in Fig. \@ref(fig:SOCSoilGrids2Q) to identify areas, where our modeled SOC stocks that are below the 5th or above the 95th percentile of the SoilGrids 2.0 data. 
For the vast majority of grid cells our model results are between the 5th and 95th percentile of SoilGrids 2.0 estimates. We underestimate SOC stocks especially in dry areas (e.g. close to the Sahara). Overestimated stocks are often situated in mountainous regions.

```{r captionSOCSoilGrids2, echo=FALSE}
captionSOCSoilGrids2 <- "Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0."
```

```{r SOCSoilGrids2, out.width = "13cm", fig.cap = captionSOCSoilGrids2, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareSoilGrids2_2010_lu.png")
```

```{r captionSOCSoilGrids2Q, echo=FALSE}
captionSOCSoilGrids2Q <- "Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0."
```

```{r SOCSoilGrids2Q, out.width = "14cm", fig.cap = captionSOCSoilGrids2Q, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/SOCquantilesSoilGrids2.png")
```

### 2022-05-25 SOCbudgetPaper/SOCbudget.Rmd
updated Paper version

@@
 abstract: |
-  Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land cover change and agricultural management.
-However, the latter has so far not been well represented in global carbon stock assessments.
+  Soil organic carbon (SOC) is one of the largest terrestrial carbon (C) stocks on Earth.
+The first meter of the Earth's soils profile stores three times as much C as the vegetation and twice the amount of C in the atmosphere.
+SOC has been depleted by anthropogenic land-cover change and agricultural management.
+However, the latter has so far not been well represented in global C stock assessments.
 While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
-Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We combine it with as reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 32 GtC relative to a counterfactual natural state in 1975.
-Within the period 1975--2010 this SOC debt continued to increasing by 5 GtC ($0.14\unit{GtC yr^{-1}}$) to around 37 GtC.
-However, accounting for historical management rather than assuming constant management conditions led to 2 GtC less $0.06\unit{GtC yr^{-1}}$  emissions.
-We also find that SOC is most sensitive to management decisions concern soil inputs such as residue returning, indicating that biomass increasing SOC sequestration techniques might be (the more) promising negative emission techniques (NET).
-Using the here proposed reduced-complexity SOC model it might be possible to track these management-driven SOC gains within integrated (land-use) assessment modeling accounting dynamically for their socio-economic NET potential.
+Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
+We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
+We estimate that due to arable farming, soils have lost around $34.6\,\unit{GtC}$ relative to a counterfactual hypothetical natural state in 1975.
+Within the period 1975--2010 this SOC debt continued to expand by $5\,\unit{GtC}$ ($0.14\,\unit{GtC\,yr^{-1}}$) to around 39.6 GtC.
+However, accounting for historical management led to $2.1\,\unit{GtC}$ less ($0.06\,\unit{GtC\,yr^{-1}}$) emissions than under the assumption of constant management.
+We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique.
+The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling.
 bibliography: SOCbudget.bib
@@
 # copyrightstatement: |
-#   The author's copyright for this publication is transferred to Potsdam Institute for Climate Impact Research (PIK).
 availability:
@@
 acknowledgements: |
+
   Thanks to Vera Porwollik for contributing the time resolved tillage data set based on her previous work.
@@
 774378).
-
 appendix: |
@@
   \section{Figures and tables in appendices}
-
   \subsection{Methods}
-
       ```{=latex}
@@
 
-  \subsection{Discussion}
+  \subsection{Results}
 
@@
     \label{fig:SOCreg}
-  \end{figure}
-  ```
-
-    ```{=latex}
-  \begin{figure}[H]
-    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
-    \caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines \citep{calvo_buendia_ipcc_2019}, we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \citealt{century_model_2000}).
-\textit{LitterPNV-CenturyAverage} forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally.
-Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
-    \label{fig:SOClitter}
   \end{figure}

### 2022-05-25 SOCbudgetPaper/SOCbudget.tex
updated Paper version

@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth.
-The first meter of the Earth's soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere.
+Soil organic carbon (SOC) is one of the largest terrestrial carbon (C) stocks on Earth.
+The first meter of the Earth's soils profile stores three times as much C as the vegetation and twice the amount of C in the atmosphere.
 SOC has been depleted by anthropogenic land-cover change and agricultural management.
-However, the latter has so far not been well represented in global carbon stock assessments.
+However, the latter has so far not been well represented in global C stock assessments.
 While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale.
@@
 We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils.
-We estimate that due to arable farming, soils have lost around 34.6 GtC relative to a counterfactual hypothetical natural state in 1975.
-Within the period 1975--2010 this SOC debt continued to expand by 5 GtC (\(0.14\unit{GtC yr^{-1}}\)) to around 39.6 GtC.
-However, accounting for historical management led to 2.1 GtC less (\(0.06\unit{GtC yr^{-1}}\)) emissions than under the assumption of constant management.
+We estimate that due to arable farming, soils have lost around \(34.6\,\unit{GtC}\) relative to a counterfactual hypothetical natural state in 1975.
+Within the period 1975--2010 this SOC debt continued to expand by \(5\,\unit{GtC}\) (\(0.14\,\unit{GtC\,yr^{-1}}\)) to around 39.6 GtC.
+However, accounting for historical management led to \(2.1\,\unit{GtC}\) less (\(0.06\,\unit{GtC\,yr^{-1}}\)) emissions than under the assumption of constant management.
 We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique.
@@
 Managed soils have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (soil carbon sequestration (SCS) techniques).
-However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) \citep{popp_land-use_2016, rogelj_scenarios_2018} have not integrated soil management into their mitigation pathways.
+However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) \citep{popp_land-use_2016, rogelj_scenarios_2018, forster_sr1p5_2018} have not integrated soil management into their mitigation pathways.
 More detailed process-based models are typically computationally too demanding to be integrated into optimization-based IAMs.
@@
 
-Following the Tier 2 modeling approach of the refinement of the IPCC guidelines vol.~4 Chapter 5 on ``Cropland'' \citep{ogle_cropland_in_ipcc_2019}; referred to as \textit{Tier 2 modeling approach} in the following), we estimate SOC stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+Following the Tier 2 modeling approach of the refinement of the IPCC guidelines vol.~4 Chapter 5 on ``Cropland'' \citep{ogle_cropland_in_ipcc_2019}; referred to as \textit{Tier 2 modeling approach} in the following), we estimate \(SOC\) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
 We restrict our analysis to the first 0-30 cm of the soil profile.
-Moreover, we assume the current SOC state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
+Moreover, we assume the current \(SOC\) state converges towards a steady state, which itself depends on biophysical, climatic and agronomic conditions.
 Therefore, we take the following three steps for each year of our simulation period:
@@
 Upon cropland expansion (in this example half of the natural vegetation is cleared to be used as cropland), $SOC$ stocks on cropland increase due to a transfer of land from natural vegetation (b).
-Separately accounting for SOC dynamics on newly converted and existing cropland (c) leads to the same overall SOC dynamics as if SOC stocks are averaged (d), due to the linearity of Eq.
-\ref{eq:SOCstate} and the cropland-age independend decay rates (see Eq.
-\ref{eq:decayrates})}\label{fig:LUtrans}
-\end{figure}
-
-We calculate SOC stocks based on the area shares of land-use types \(lu\) within our grid cells \(i\).
-If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the SOC is reallocated within our budget.
-We do not distinguish between newly converted and existing cropland, but can work with the average carbon content as the relative decay of SOC is proportional to the SOC stock (see \ref{fig:LUtrans}).
+Explicitly representing newly converted cropland and existing cropland to account for SOC dynamics (c) leads to the same weighted mean value as averaging $SOC$ stocks (d), due to the linearity of Eq.
+\ref{eq:SOCstate} and cropland-age independent decay rates (see Eq.
+\ref{eq:decayrates}).}\label{fig:LUtrans}
+\end{figure}
+
+We calculate \(SOC\) stocks based on the area shares of land-use types \(lu\) within our grid cells \(i\).
+If land is converted from one land-use type \(lu=\{crop,natveg\}\) into the other \(!lu=\{natveg,crop\}\), a respective share of the \(SOC\) is reallocated within our budget.
+We do not distinguish between newly converted and existing cropland, but can work with the average carbon content as the relative decay of \(SOC\) is proportional to the \(SOC\) stock (see \ref{fig:LUtrans}).
 We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{t^*}\) via
@@
 
-SOC converges towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
+\(SOC\) converges towards the calculated steady-state stock \(SOC^{\mathrm{eq}}\) for each grid cell \(i\), each annual time step \(t\), each land-use type \(lu\) and each sub-pool \(sub\) like
 
@@
 Land-use patterns are based on the Land-Use Harmonization 2 \citep{hurtt_harmonization_2020} data set (short LUH2), which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area (in Mha) of the five different cropland subcategories (c3ann: C3 annual crops, c3per: C3 perennial crops, c4ann: C4 annual crops, c4per: C4 perennial crops, c3nfx: C3 nitrogen-fixing crops) of LUH2 into our 17 crop groups (see Table \ref{tab:fao2luh2mag}), applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016}.
+We disaggregate the physical area (given as total land area in million ha) of the five different cropland subcategories (c3ann: C3 annual crops, c3per: C3 perennial crops, c4ann: C4 annual crops, c4per: C4 perennial crops, c3nfx: C3 nitrogen-fixing crops) of LUH2 into our 17 crop groups \citep[see Table ``FAO2LUH2MAG\_croptypes.csv'' in][]{karstens_model_2020}, applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data \citep{faostat_faostat_2016}.
 By calculating country-specific multicropping factors \(MCF\) using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
@@
 Crop residue production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and key aspects are explained as they play a central role in soil carbon modeling.
-Starting from crop yield estimates \(Y\) and respective physical crop area \(CA\), we estimate total above-ground \(AGR\) and below-ground \(BGR\) residue biomass (in tonnes) using crop group \(cg\) specific ratios for above-ground residues to harvested biomass \(r^{\mathrm{ag,prod}}_{cg}\) in \(\unit{tC\,\unit{tC}^{-1}}\), above-ground residues to harvested area \(r^{\mathrm{ag,area}}_{cg}\) in \(\unit{tC\,\unit{ha}^{-1}}\) and below-ground residues to above-ground biomass \(r^{\mathrm{bg}}_{cg}\) in \(\unit{tC\,\unit{tC}^{-1}}\) as follows
+Starting from crop yield estimates \(Y\) and respective physical crop area \(CA\), we estimate total above-ground \(AGR\) and below-ground \(BGR\) residue biomass (in tonnes) using crop group \(cg\) specific ratios for above-ground residues to harvested biomass \(r^{\mathrm{ag,prod}}_{cg}\) in \((tDM\,ha^{-1})(tDM\,ha^{-1})^{-1}\) above-ground residues to harvested area \(r^{\mathrm{ag,area}}_{cg}\) in \(\unit{tDM\,\unit{ha}^{-1}}\) and below-ground residues to above-ground biomass \(r^{\mathrm{bg}}_{cg}\) in \(\unit{tDM\,\unit{tDM}^{-1}}\) as follows
 
@@
 A country-specific fixed share of the \(AGR\) is assumed to be burned on field depending on the per-capita income of the country.
-Following Smil \citeyearpar{smil_nitrogen_1999} we assume a burn share of 25\% for low-income countries according to World Bank definitions (\(<\,1000\,\tfrac{USD}{yr\,cap}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr\,cap}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income for the periods before 1995.
+Following Smil \citeyearpar{smil_nitrogen_1999} we assume a burn share of 25\% for low-income countries according to World Bank definitions (\(<\,10000\,\unit{USD}\,\unit{yr}^{-1}\,\unit{cap}^{-1}\)), 15\% for high-income (\(>\,10000\,\unit{USD}\,\unit{yr}^{-1}\,\unit{cap}^{-1}\)) and linearly interpolate shares for all middle-income countries depending on their per-capita income for the periods before 1995.
 After 1995 we estimate a linear decline of burn shares to 10\% for low-income countries and 0\% for high-income countries till 2025 to account for recent increases in air pollution regulation.
@@
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020}.
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020}).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017}, that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
@@
 The remaining \(AGR\) as well as all \(BGR\) are expected to be left on the field.
-We limit high residue return rates to at most \(10\unit{tC\,ha}^{-1}\) in order to correct for outliers.
+We limit high residue return rates to at most \(10\,\unit{tC\,ha}^{-1}\) in order to correct for outliers.
 
@@
 For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012}.
-We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
+We limit high application shares at \(10\,\unit{tC\,ha}^{-1}\) to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
 
@@
 Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens \citeyearpar{karstens_model_2020}.
-In the following, the most important results \citep[see][]{karstens_result_2020} for post-processing script) are summarized.
+In the following, the most important results \citep[see][ for post-processing script]{karstens_result_2020} are summarized.
 
@@
 The distribution of total global SOC stocks for the first 30 cm on cropland for the year 2010 is depcited in panel (b).
-Absolute (c) and relative (cd SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains.}\label{fig:SOCmaps}
-\end{figure}
-
-The global SOC debt has increased by about 14\% in the period between 1975 and 2010 from \(34.6\) to \(39.6\unit{GtC}\) (Fig.
+Absolute (c) and relative (d) SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains.}\label{fig:SOCmaps}
+\end{figure}
+
+The global SOC debt has increased by about 14\% in the period between 1975 and 2010 from \(34.6\) to \(39.6\,\unit{GtC}\) (Fig.
 \ref{fig:SOCmaps}(a)).
-This corresponds to an average loss rate of \(0.14\unit{GtC\,yr^{-1}}\) in comparison to a hypothetical potential natural vegetation (PNV) state.
-Considering our estimate of the global SOC stock of around \(705\unit{GtC}\) in the upper \(30\unit{cm}\) in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+This corresponds to an average loss rate of \(0.14\,\unit{GtC\,yr^{-1}}\) in comparison to a hypothetical potential natural vegetation (PNV) state.
+Considering our estimate of the global SOC stock of around \(705\,\unit{GtC}\) in the upper \(30\,\unit{cm}\) in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
 The speed of this SOC loss has decreased towards the end of the modeling period.
@@
 In Fig.
-\ref{fig:SOCmaps}(b) we provide a world map of SOC stocks estimates for the first \(30\unit{cm}\) on cropland considering historical management data for the year 2010.
-Values range between over \(100\unit{t\,ha^{-1}}\) in northern temperate cropland to less than \(5\unit{t\,ha^{-1}}\) for arid and semiarid cropland.
+\ref{fig:SOCmaps}(b) we provide a world map of SOC stocks estimates for the first \(30\,\unit{cm}\) on cropland considering historical management data for the year 2010.
+Values range between over \(100\,\unit{t\,ha^{-1}}\) in northern temperate cropland to less than \(5\,\unit{t\,ha^{-1}}\) for arid and semiarid cropland.
 Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under PNV in two complementary ways:
@@
 Absolute SOC changes \(\Delta SOC\) (Fig.
-\ref{fig:SOCmaps}(c)) indicate areas with high importance for the global SOC loss.
+\ref{fig:SOCmaps}(c)) indicate areas with high importance for the global \(SOC\) loss.
 They can be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia).
@@
 
+\begin{figure}[h]
+\includegraphics[width=9cm]{../ResultNotebooks/Output/Images/2panelfigure} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile.
+Panel (a) shows global $\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types.
+Computing the difference between the $\Delta SOC$ estimate for 2010 and for 1975 (b) depicts areas of soil depletion (SOC debt increase, red) and net-sequestration (SOC debt decline, blue).}\label{fig:SOCdebt}
+\end{figure}
+
+The spatial distribution of the total \(\Delta SOC\) summed over all land-use types (Fig.
+\ref{fig:SOCdebt}(a)) and its change from 1975 to 2010 (Fig.
+\ref{fig:SOCdebt}(b)) reveals areas of SOC debt decline and increase.
+Regions with large cropland expansion (e.g.~Brazil, Southeast Asia, Canada) continue to lose SOC, whereas regions with cropland reduction (and thus SOC restoration) or with accumulating cropland SOC can be found e.g.~in highly productive areas of Europe and Central USA.
+
 \hypertarget{carbon-flows-in-the-agricultural-system}{%
@@
 \begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC).
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in $\unit{MtC}$).
 Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand.
@@
 
-C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvest organ, above- and below-ground residues).
-Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (\(729\unit{MtC}\) in 2010) are directly returned to the field.
+C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvested organ, above- and below-ground residues).
+Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (\(729\,\unit{MtC}\) in 2010) are directly returned to the field.
 We divide crop biomass usage into feed usage and aggregate all other usage types (e.g.~food, bioenergy and material) into a human demand category.
-Livestock feed demand for crop organ harvest and above-ground residues of \(1136\unit{MtC}\) is roughly equal to the human demand of \(1129\unit{MtC}\).
-Whereas large parts of feed intake are returned to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not returned to soils.
-Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with \(1350\unit{MtC}\) returned to the fields in 2010.
+Livestock feed demand for crop organ harvest and above-ground residues of \(1136\,\unit{MtC}\) is roughly equal to the human demand of \(1129\,\unit{MtC}\).
+Whereas large parts of feed intake are returned to the soils via manure (C input from manure at \(384\,\unit{MtC}\)), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not returned to soils.
+Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with \(1350\,\unit{MtC}\) returned to the fields in the year 2010.
 However, around 60\% of this organic C decomposes before it is integrated into soils.
 Due to the different composition of organic C, proportionally more C enters the slow pool from manure than from crop residue.
-According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of \(257\unit{MtC}\) in 2010.
-The agricultural system receives \(4585\unit{MtC}\) assimilated by crop plants and releases \(3554\unit{MtC}\) mostly through respiration.
-Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around \(33\unit{MtC}\) for the year 2010.
+According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of \(257\,\unit{MtC}\) in 2010.
+The agricultural system receives \(4585\,\unit{MtC}\) assimilated by crop plants and releases \(3554\,\unit{MtC}\) mostly through respiration.
+Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around \(33\,\unit{MtC}\) for the year 2010.
 
@@
 \begin{figure}[h]
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios.
-The stylized scenarios deviate from historical argicultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage).
+\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in $\unit{GtC}$ for different management scenarios.
+The stylized scenarios deviate from historical agricultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage).
 The scenario constManagement combines all three modifications.
@@
 As shown by the difference between the constResidues scenario and the other counterfactuals, changes in residue return rates dominate the management effects.
-Without the historical increase in C inputs from residues to agricultural soils, the global \(\Delta SOC\) would decrease to \(41.7\unit{GtC}\) at a rate of \(0.20\unit{GtC\,yr^{-1}}\) --- a 35\% increase compared to \(0.14\unit{GtC\,yr^{-1}}\) for the histManagement estimates.
-Both the constManure and constTillage scenarios show only small deviations from the historical argicultural management values with \(0.15\unit{GtC\,yr^{-1}}\).
+Without the historical increase in C inputs from residues to agricultural soils, the global \(\Delta SOC\) would decrease to \(41.7\,\unit{GtC}\) at a rate of \(0.20\,\unit{GtC\,yr^{-1}}\) --- a 35\% increase compared to \(0.14\,\unit{GtC\,yr^{-1}}\) for the histManagement estimates.
+Both the constManure and constTillage scenarios show only small deviations from the historical agricultural management values with \(0.15\,\unit{GtC\,yr^{-1}}\).
 The effect of no-tillage only becomes discernible from 2000 onwards.
@@
 Our results are in good agreement with the 2006 IPCC factors.
-Modelled \(F^{SCF}\) have increased or stayed constant for all climate zones over time (1975-2010).
+Modeled \(F^{SCF}\) have increased or stayed constant for all climate zones over time (1975-2010).
 
@@
 
-We compare our global SOC stocks with a wide range of global SOC stock estimates for the first 30 cm of the soil profile, using data from WISE \citep{batjes_harmonized_2016}, SoilGrids \citep{hengl_soilgrids250m_2017}, GSOC \citep{fao_global_2018}, LPJmL4 \citep{schaphoff_lpjml4_2018-1}, SoilGrids2 \citep{poggio_soilgrids_2021}, and SOCDebtPaper \citep{sanderman_soil_2017} in Fig.
+We compare our global SOC stocks with a wide range of global SOC stock estimates for the first 30 cm of the soil profile, using data from WISE \citep{batjes_harmonized_2016}, SoilGrids \citep{hengl_soilgrids250m_2017}, GSOC \citep{fao_global_2018}, LPJmL4 \citep{schaphoff_lpjml4_2018-1}, SoilGrids 2.0 \citep{poggio_soilgrids_2021}, and SOCDebtPaper \citep{sanderman_soil_2017} in Fig.
 \ref{fig:SOCglo}.
@@
 \begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area.
+\includegraphics[width=11cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area.
 The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \citep{schaphoff_lpjml4_2018-1} and estimates from \citep{sanderman_soil_2017}.
@@
 
-The global estimates of the total SOC stock of the upper \(30\,cm\) from this study are in the middle of the wide range of other modeled or observation-based estimates.
-SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, whereas SoilGrid2.0 \citep{poggio_soilgrids_2021} marks the lower end.
+The global estimates of the total SOC stock of the upper 30 cm from this study are in the middle of the wide range of other modeled or observation-based estimates.
+SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, whereas SoilGrids 2.0 \citep{poggio_soilgrids_2021} marks the lower end.
 Regional results (Fig.
@@
 \begin{figure}[h]
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010} \caption{Correlation between modeled and measured SOC stocks.
-Given the wide span between minimun and maximum measured SOC stocks within in a given cell, we correlated median values with our modeled results.
+\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010} \caption{Correlation between modeled and measured $SOC$ stocks.
+Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results.
 Both cropland and areas with natural vegetation tend to be lower in our results than in the point measurements.}\label{fig:SOCpoint}
@@
 Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.
-With our focus on cropland SOC dynamics, we compare only cells with more than \(1000\unit{ha}\) of cropland (capturing 99.9\% of global cropland area).
+With our focus on cropland SOC dynamics, we compare only cells with more than \(1000\,\unit{ha}\) of cropland (capturing 99.9\% of global cropland area).
 
@@
 \ref{fig:SOCSoilGrids2}.
-To focus our comparison on cropland areas, we mask out grid cells with less than \(1000\unit{ha}\) of cropland.
+To focus our comparison on cropland areas, we mask out grid cells with less than \(1000\,\unit{ha}\) of cropland.
 Spatial correlation is moderate for tropical climate zones, whereas it is low for temperate moist areas.
-In tropical dry and temperate dry areas, we simulate also very low SOC values (below \$10\unit{tC\,ha^{-1}}), which is not found in SoilGrids 2.0 whereas our modeled SOC stocks can be substantially higher in temperate moist areas than reported by SoilGrids 2.0.
+In tropical dry and temperate dry areas, we simulate also very low SOC values (below \(10\,\unit{tC\,ha^{-1}}\)), which is not found in SoilGrids 2.0 whereas our modeled SOC stocks can be substantially higher in temperate moist areas than reported by SoilGrids 2.0.
 Additionally, we use the uncertainty estimates from SoilGrids 2.0 in Fig.
@@
 
-Our spatially explicit time series dataset of agricultural management
-is based on country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} as well as 0.5 degree land-use data from LUH2 \citep{hurtt_harmonization_2020}.
+Our spatially explicit time series data set of agricultural management is based on country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} as well as 0.5 degree land-use data from LUH2 \citep{hurtt_harmonization_2020}.
 Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the agricultural system (\ref{fig:FlowFig}) using a mass balance approach from the IPCC guidelines Vol.
-4 \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and other auxiliary data sets (e.g.~\citep{porwollik_generating_2019}).
+4 \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and other auxiliary data sets \citep[e.g.~][]{porwollik_generating_2019}.
 
@@
 
-While our dataset, by including crop residues and manure, likely the largest carbon inputs to soils, it does not account for a list of minor carbon inputs from cover crops, agroforestry, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
-Including these sources would correct our estimates upwards and bring our estimates closer to the IPCC stock change factors(see Sect.
+While our data set, by including crop residues and manure, likely the largest carbon inputs to soils, it does not account for a list of minor carbon inputs from cover crops, agroforestry, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
+Including these sources would correct our estimates upwards and bring our estimates closer to the IPCC stock change factors (see Sect.
 \ref{sec:ipcccompare}).
@@
 The Tier 2 approach \citep{ogle_cropland_in_ipcc_2019} used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV.
-This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is a main driver of SOC dynamics (cite).
+This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics.
 The comparison of simulated PNV data with LPJmL4 shows the model's substantial capability in reproducing PNV SOC stocks (Fig.
@@
 
-Comparing the geographic SOC patterns to Soil Grids 2.0 {[}\citet{poggio_soilgrids_2021}; see Fig.
-\ref{fig:SOCSoilGrids2Q}{]}, we find that our model estimates values of SOC stocks greater than the estimated confidence intervals in Soil Grid 2.0s for some mountainous regions across the globe.
+Comparing the geographic SOC patterns to Soil Grids 2.0 \citep{poggio_soilgrids_2021} (see Fig.
+\ref{fig:SOCSoilGrids2Q}), we find that our model estimates values of SOC stocks greater than the estimated confidence intervals in Soil Grid 2.0 for some mountainous regions across the globe.
 This could indicate that we are not capable in capturing specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils \citep{borrelli_assessment_2017}).
@@
 In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Table \ref{tab:SCFglo}) for almost all regions.
-
-Tropical soils might suffer from low C input rates due to large yield gaps \citep{global_yieldgap_2022} and high shares of residue removal and burning in lower-income countries {[}\citet{smil_crop_1999}; williams\_influence\_1997; jain\_emission\_2014{]}.
+Tropical soils might suffer from low C input rates due to large yield gaps (Global Yield Gap and Water Productivity Atlas.
+Available URL: www.yieldgap.org (accessed on: 03/01/2022)) and high shares of residue removal and burning in lower-income countries \citep{smil_crop_1999, williams_influence_1997, jain_emission_2014}.
 Yet, even when comparing our etimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values, while it shows very good agreement with the older default values from IPCC \citeyearpar{lasco_cropland_in_ipcc_2006}.
@@
 In contrast, process-based studies \citep{pugh_simulated_2015, herzfeld_soc_2021} often lack data on relevant management aspects that drive production increases.
-\citep{herzfeld_soc_2021} also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
+Herzfeld et al.~\citeyearpar{herzfeld_soc_2021} also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
 Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
@@
 \section{Figures and tables in appendices}
-
 \subsection{Methods}
@@
 
-\begin{table}[htbp]
-  \centering
-  \caption{Mapping of FAO crop production items \citep{faostat_faostat_2016} to LUH2 categories \citep{hurtt_harmonization_2020} and to crop groups used in this study.}
-  \begin{tabular}{lllll}
-  \tophline
-      LUH short name & FAO Production Item & Crop group – madrat & Short name – madrat \\ 
-  \middlehline    
-      c3ann & 122|Sweet potatoes & Tropical roots & cassav\_sp \\ 
-      c3ann & 125|Cassava & Tropical roots & cassav\_sp \\ 
-      c3ann & 135|Yautia (cocoyam) & Tropical roots & cassav\_sp \\ 
-      c3ann & 136|Taro (cocoyam) & Tropical roots & cassav\_sp \\ 
-      c3ann & 137|Yams & Tropical roots & cassav\_sp \\ 
-      c3ann & "149|Roots and tubers, nes" & Tropical roots & cassav\_sp \\ 
-      c3per & 486|Bananas & Tropical roots & cassav\_sp \\ 
-      c3per & 489|Plantains & Tropical roots & cassav\_sp \\ 
-      c3ann & 328|Seed cotton & Cotton seed & cottn\_pro \\ 
-      c3ann & "638|Forage and silage, rye grass" & Forage & foddr \\ 
-      c3ann & "639|Forage and silage, grasses nes" & Forage & foddr \\ 
-      c3ann & "642|Forage and silage, green oilseeds" & Forage & foddr \\ 
-      c3ann & 644|Cabbage for fodder & Forage & foddr \\ 
-      c3ann & 645|Pumpkins for Fodder & Forage & foddr \\ 
-      c3ann & 646|Turnips for fodder & Forage & foddr \\ 
-      c3ann & 647|Beets for fodder & Forage & foddr \\ 
-      c3ann & 648|Carrots for fodder & Forage & foddr \\ 
-      c3ann & 649|Swedes for fodder & Forage & foddr \\ 
-      c3ann & 651|Forage products & Forage & foddr \\ 
-      c3ann & 655|Vegetables and roots fodder & Forage & foddr \\ 
-      c4ann & "636|Forage and silage, maize" & Forage & foddr \\ 
-      c4ann & "637|Forage and silage, sorghum" & Forage & foddr \\ 
-      c3nfx & "640|Forage and silage, clover" & Forage & foddr \\ 
-      c3nfx & "641|Forage and silage, alfalfa" & Forage & foddr \\ 
-      c3nfx & "643|Forage and silage, legumes" & Forage & foddr \\ 
-      c3nfx & "242|Groundnuts, with shell" & Groundnuts & groundnut \\ 
-      c4ann & 56|Maize & Maize & maiz \\ 
-      c3per & "254|Oil, palm fruit" & Oilpalms & oilpalm \\ 
-      c3per & 254|Oil palm fruit & Oilpalms & oilpalm \\ 
-      c3ann & 358|Cabbages and other brassicas & Fruits Vegetables Nuts & others \\ 
-      c3ann & 366|Artichokes & Fruits Vegetables Nuts & others \\ 
-      c3ann & 367|Asparagus & Fruits Vegetables Nuts & others \\ 
-      c3ann & 372|Lettuce and chicory & Fruits Vegetables Nuts & others \\ 
-      c3ann & 373|Spinach & Fruits Vegetables Nuts & others \\ 
-      c3ann & 378|Cassava leaves & Fruits Vegetables Nuts & others \\ 
-      c3ann & 388|Tomatoes & Fruits Vegetables Nuts & others \\ 
-      c3ann & 393|Cauliflowers and broccoli & Fruits Vegetables Nuts & others \\ 
-      c3ann & "394|Pumpkins, squash and gourds" & Fruits Vegetables Nuts & others \\ 
-      c3ann & 397|Cucumbers and gherkins & Fruits Vegetables Nuts & others \\ 
-      c3ann & 399|Eggplants (aubergines) & Fruits Vegetables Nuts & others \\ 
-      c3ann & "401|Chillies and peppers, green" & Fruits Vegetables Nuts & others \\ 
-      c3ann & "402|Onions, shallots, green" & Fruits Vegetables Nuts & others \\ 
-      c3ann & "403|Onions, dry" & Fruits Vegetables Nuts & others \\ 
-      c3ann & 406|Garlic & Fruits Vegetables Nuts & others \\ 
-      c3ann & "407|Leeks, other alliaceous vegetables" & Fruits Vegetables Nuts & others \\ 
-      c3ann & 426|Carrots and turnips & Fruits Vegetables Nuts & others \\ 
-      c3ann & 430|Okra & Fruits Vegetables Nuts & others \\ 
-      c3ann & 449|Mushrooms and truffles & Fruits Vegetables Nuts & others \\ 
-      c3ann & 459|Chicory roots & Fruits Vegetables Nuts & others \\ 
-      c3ann & "463|Vegetables, fresh nes" & Fruits Vegetables Nuts & others \\ 
-      c3ann & 544|Strawberries & Fruits Vegetables Nuts & others \\ 
-      c3ann & 567|Watermelons & Fruits Vegetables Nuts & others \\ 
-      c3ann & "568|Melons, other (inccantaloupes)" & Fruits Vegetables Nuts & others \\ 
-      c3ann & "689|Chillies and peppers, dry" & Fruits Vegetables Nuts & others \\ 
-      c3ann & "711|Anise, badian, fennel, coriander" & Fruits Vegetables Nuts & others \\ 
-      c3ann & 720|Ginger & Fruits Vegetables Nuts & others \\ 
-      c3ann & "723|Spices, nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & "216|Brazil nuts, with shell" & Fruits Vegetables Nuts & others \\ 
-      c3per & "217|Cashew nuts, with shell" & Fruits Vegetables Nuts & others \\ 
-      c3per & 220|Chestnut & Fruits Vegetables Nuts & others \\ 
-      c3per & "221|Almonds, with shell" & Fruits Vegetables Nuts & others \\ 
-      c3per & "222|Walnuts, with shell" & Fruits Vegetables Nuts & others \\ 
-      c3per & 223|Pistachios & Fruits Vegetables Nuts & others \\ 
-      c3per & 224|Kola nuts & Fruits Vegetables Nuts & others \\ 
-      c3per & "225|Hazelnuts, with shell" & Fruits Vegetables Nuts & others \\ 
-      c3per & 226|Areca nuts & Fruits Vegetables Nuts & others \\ 
-      c3per & "234|Nuts, nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & 461|Carobs & Fruits Vegetables Nuts & others \\ 
-      c3per & 490|Oranges & Fruits Vegetables Nuts & others \\ 
-      c3per & "495|Tangerines, mandarins, clementines, satsumas" & Fruits Vegetables Nuts & others \\ 
-      c3per & 497|Lemons and limes & Fruits Vegetables Nuts & others \\ 
-      c3per & 507|Grapefruit (inc pomelos) & Fruits Vegetables Nuts & others \\ 
-      c3per & "512|Fruit, citrus nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & 515|Apples & Fruits Vegetables Nuts & others \\ 
-      c3per & 521|Pears & Fruits Vegetables Nuts & others \\ 
-      c3per & 523|Quinces & Fruits Vegetables Nuts & others \\ 
-      c3per & 526|Apricots & Fruits Vegetables Nuts & others \\ 
-      c3per & "530|Cherries, sour" & Fruits Vegetables Nuts & others \\ 
-      c3per & 531|Cherries & Fruits Vegetables Nuts & others \\ 
-      c3per & 534|Peaches and nectarines & Fruits Vegetables Nuts & others \\ 
-      c3per & 536|Plums and sloes & Fruits Vegetables Nuts & others \\ 
-      c3per & "541|Fruit, stone nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & "542|Fruit, pome nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & 547|Raspberries & Fruits Vegetables Nuts & others \\ 
-      c3per & 549|Gooseberries & Fruits Vegetables Nuts & others \\ 
-      c3per & 550|Currants & Fruits Vegetables Nuts & others \\ 
-      c3per & 552|Blueberries & Fruits Vegetables Nuts & others \\ 
-      c3per & 554|Cranberries & Fruits Vegetables Nuts & others \\ 
-      c3per & 558|Berries nes & Fruits Vegetables Nuts & others \\ 
-      c3per & 560|Grapes & Fruits Vegetables Nuts & others \\ 
-      c3per & 569|Figs & Fruits Vegetables Nuts & others \\ 
-      c3per & "571|Mangoes, mangosteens, guavas" & Fruits Vegetables Nuts & others \\ 
-      c3per & 572|Avocados & Fruits Vegetables Nuts & others \\ 
-      c3per & 574|Pineapples & Fruits Vegetables Nuts & others \\ 
-      c3per & 577|Dates & Fruits Vegetables Nuts & others \\ 
-      c3per & 587|Persimmons & Fruits Vegetables Nuts & others \\ 
-      c3per & 591|Cashewapple & Fruits Vegetables Nuts & others \\ 
-      c3per & 592|Kiwi fruit & Fruits Vegetables Nuts & others \\ 
-      c3per & 600|Papayas & Fruits Vegetables Nuts & others \\ 
-      c3per & "603|Fruit, tropical fresh nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & "619|Fruit, fresh nes" & Fruits Vegetables Nuts & others \\ 
-      c3per & "656|Coffee, green" & Fruits Vegetables Nuts & others \\ 
-      c3per & "661|Cocoa, beans" & Fruits Vegetables Nuts & others \\ 
-      c3per & 667|Tea & Fruits Vegetables Nuts & others \\ 
-      c3per & 671|Mate & Fruits Vegetables Nuts & others \\ 
-      c3per & 687|Pepper (piper spp) & Fruits Vegetables Nuts & others \\ 
-      c3per & 692|Vanilla & Fruits Vegetables Nuts & others \\ 
-      c3per & 693|Cinnamon (canella) & Fruits Vegetables Nuts & others \\ 
-      c3per & 698|Cloves & Fruits Vegetables Nuts & others \\ 
-      c3per & "702|Nutmeg, mace and cardamoms" & Fruits Vegetables Nuts & others \\ 
-      c4ann & "446|Maize, green" & Fruits Vegetables Nuts & others \\ 
-      c3nfx & "414|Beans, green" & Fruits Vegetables Nuts & others \\ 
-      c3nfx & "417|Peas, green" & Fruits Vegetables Nuts & others \\ 
-      c3nfx & "420|Vegetables, leguminous nes" & Fruits Vegetables Nuts & others \\ 
-      c3nfx & 423|String beans & Fruits Vegetables Nuts & others \\ 
-      c3ann & 677|Hops & Fruits Vegetables Nuts & others \\ 
-      c3ann & 748|Peppermint & Fruits Vegetables Nuts & others \\ 
-      c3ann & 116|Potatoes & Potatoes & potato \\ 
-      c3nfx & "176|Beans, dry" & Pulses & puls\_pro \\ 
-      c3nfx & "181|Broad beans, horse beans, dry" & Pulses & puls\_pro \\ 
-      c3nfx & "187|Peas, dry" & Pulses & puls\_pro \\ 
-      c3nfx & 191|Chick peas & Pulses & puls\_pro \\ 
-      c3nfx & "195|Cow peas, dry" & Pulses & puls\_pro \\ 
-      c3nfx & 197|Pigeon peas & Pulses & puls\_pro \\ 
-      c3nfx & 201|Lentils & Pulses & puls\_pro \\ 
-      c3nfx & 203|Bambara beans & Pulses & puls\_pro \\ 
-      c3nfx & 205|Vetches & Pulses & puls\_pro \\ 
-      c3nfx & 210|Lupins & Pulses & puls\_pro \\ 
-      c3nfx & "211|Pulses, nes" & Pulses & puls\_pro \\ 
-      c3ann & 265|Castor oil seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 270|Rapeseed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 280|Safflower seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 289|Sesame seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 292|Mustard seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 296|Poppy seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 299|Melonseed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 333|Linseed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & 336|Hempseed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 249|Coconuts & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 260|Olives & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 263|Karite nuts (sheanuts) & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 275|Tung nuts & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 277|Jojoba seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 305|Tallowtree seed & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 310|Kapok fruit & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 311|Kapokseed in shell & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3per & 339|Oilseeds nes & Other oil crops (incl rapeseed) & rapeseed \\ 
-      c3ann & "27|Rice, paddy" & Rice & rice\_pro \\ 
-      c3nfx & 236|Soybeans & Soybean & soybean \\ 
-      c3ann & 157|Sugar beet & Sugar beet & sugr\_beet \\ 
-      c4per & 156|Sugar cane & Sugar cane & sugr\_cane \\ 
-      c3ann & 267|Sunflower seed & Sunflower & sunflower \\ 
-      c3ann & 15|Wheat & Temperate cereals & tece \\ 
-      c3ann & 44|Barley & Temperate cereals & tece \\ 
-      c3ann & 71|Rye & Temperate cereals & tece \\ 
-      c3ann & 75|Oats & Temperate cereals & tece \\ 
-      c3ann & 89|Buckwheat & Temperate cereals & tece \\ 
-      c3ann & 92|Quinoa & Temperate cereals & tece \\ 
-      c3ann & 94|Fonio & Temperate cereals & tece \\ 
-      c3ann & 97|Triticale & Temperate cereals & tece \\ 
-      c3ann & 101|Canary seed & Temperate cereals & tece \\ 
-      c3ann & "103|Grain, mixed" & Temperate cereals & tece \\ 
-      c3ann & "108|Cereals, nes" & Temperate cereals & tece \\ 
-      c4ann & 68|Popcorn & Temperate cereals & tece \\ 
-      c4ann & 79|Millet & Tropical cereals & trce \\ 
-      c4ann & 83|Sorghum & Tropical cereals & trce \\ 
-      \bottomhline
-  \end{tabular}
-  \label{tab:fao2luh2mag}
-\end{table}
-
-\subsection{Discussion}
+\subsection{Results}
 

### 2022-05-25 SOCbudgetPaper/discussion.Rmd
updated Paper version

@@
 
-We have (1) developed a reduced-complexity model and (2) compiled a spatially explicit time series data set of agricultural management data that helped us to (3) analyze the role of agricultural management in historical cropland SOC dynamics.
-Our study shows that agricultural management data alter estimates of the SOC stock size and slows down the increase of SOC debt compared to the often used constant management assumptions.
+We have (1) developed a reduced-complexity model and (2) compiled a spatially explicit time series data set of agricultural management data in order to (3) analyze the role of agricultural management in historical cropland SOC dynamics.
+Our study shows that information on agricultural management alters estimates of the SOC debt and slows down loss of SOC compared to the often used constant management assumptions.
 
 It is important to evaluate the validity of our results as modeling management effects on SOC dynamics at the global scale comes with large uncertainties.
-The quantitative assessment of the uncertainty of our projections unfortunately exceeds the scope of this article and would likely require a study in itself.
 The model includes a large number of parameters, and for most of these the uncertainty distributions have not been quantified so far.
@@
 
-Our spatially explicit time series data set of agricultural management
-is based on country-specific FAO production and cropland statistics [@faostat_faostat_2016] as well as 0.5 degree land-use data from LUH2 [@hurtt_harmonization_2020].
-Starting from these two source, we compile a harmonized and consistent data set for the major flows within the agricultural system (\@ref(fig:FlowFig)) using rule-based assumptions (e.g.
-from the IPCC guidelines Vol.
-4 [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019]) and only a few auxiliary data sets (e.g.
-[@porwollik_generating_2018]).
+Our spatially explicit time series data set of agricultural management is based on country-specific FAO production and cropland statistics [@faostat_faostat_2016] as well as 0.5 degree land-use data from LUH2 [@hurtt_harmonization_2020].
+Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the agricultural system (\@ref(fig:FlowFig)) using a mass balance approach from the IPCC guidelines Vol.
+4 [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and other auxiliary data sets [e.g.
+@porwollik_generating_2019].
 
-For some of the aspects covered in our data set, for example livestock distribution [@robinson_mapping_2014] or manure production and application [@zhang_global_2017] very well compiled data sets in high resolution exists, that capture real world conditions much better than our estimates.
-However, they often come with the caveats of either being static in time, demanding large sets of auxiliary data or being inconsistent to each other.
-Within IAM modeling, not only the SOC model itself, but also the input data that drives the model, has to be computed endogenously and dynamically in time, leaving these data sets, that can not be easily projected into the future unsuitable.
+For some of the aspects covered in our data set, for example livestock distribution [@robinson_mapping_2014] or manure production and application [@zhang_global_2017], well-compiled data sets in high resolution exist that capture real world conditions much better than our estimates.
+However, they often come with the caveats of either being static in time, demanding large sets of auxiliary data or being inconsistent with each other.
 
-For most of our parameters used in our rule-based calculation no uncertainty estimates exist.
-This is why, in our view, most of the uncertainty is included in the management data itself, and especially in the residue and manure numbers, as they are only indirectly calculated from production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017] using large sets of parameter assumptions.
+For most of the parameters used in our management estimates no uncertainty estimates exist.
+This is why, in our view, most of the uncertainty with respect to the impacts of SOC management is included in the management data itself, and especially in the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017].
 The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
-As this is one of the first attempts to compile a comprehensive, global data set on these management data, it is complex to evaluate it against literature.
-Furthermore, it is still to be evaluated, if our data set covers most of the carbon inputs to the soil, as it not includes cover crops, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
-Unfortunately, data on the quantity of these inputs is very scarce and often does not exist with global coverage.
+Previous modeling studies of SOC carbon on cropland often only used stylized scenarios of management practices [@pugh_simulated_2015; @lutz_simulating_2019], rather than trying to estimate real management.
 
-> (aus den point-by-point replies): In the draft of the International Nitrogen Assessment (will be provided confidentially to the reviewer), the available literature estimates for these flows (in regard to nitrogen) were reviewed.
-So far no reliable estimates exist, but the existing estimates indicate that these inputs must be by an order of magnitude lower than those of crop residues and manure.
+While our data set, by including crop residues and manure, likely the largest carbon inputs to soils, it does not account for a list of minor carbon inputs from cover crops, agroforestry, green manure, weed biomass as well as application of human excreta, sewage sludge, processing wastes, forestry residues or biochar.
+Including these sources would correct our estimates upwards and bring our estimates closer to the IPCC stock change factors (see Sect.
+\@ref(sec:ipcccompare)).
+Unfortunately, data on the quantity of these inputs is very scarce and does not exist with global coverage.
 
-As SOC inputs from above-ground residues had the strongest effect on SOC debt dynamics (see Fig.
+SOC inputs from above-ground residues had the strongest management effect on SOC debt dynamics on cropland (see Fig.
 \@ref(fig:SOCscen)).
@@
 Keel et al.
-[-@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+[-@keel_large_2017] question whether below-ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
-While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass.
-Looking on the evaluation of management effects (see Sect.
-\@ref(sec:ipcccompare)), there is no general indication of overestimating SCFs compared to IPCC default assumptions.
-In contrast, it is even more likely that we are still missing carbon inputs to the soil from e.g.
-cover- and inter-cropping practices as SCF are lower than predicted by the IPCC.
+While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass, in particular for below-ground biomass.
 
@@
 
-The newly developed reduced-complexity SOC model is based on a Tier 2 modeling approach.
-This reduces the computational and data demand of the model in comparison to DGVMs.
-Still, the effects of water supply from spatially heterogenic rainfall and irrigation as well as tillage systems can be accounted for in the computation of SOC decay rates.
-Similarly, C inputs into the soils reflect differences in productivity in space and over time as well as C losses in the removal rates and processing steps of residues and manure production and application rates to cropland.
+Our reduced-complexity SOC model is based on a Tier 2 modeling approach.
+This reduces the computational and data demand of the model in comparison to DGVMs, while still allowing for the explicit representation of agricultural management practices.
+Along with the effects of various C inputs, the impacts of water supply from rainfall and irrigation as well as tillage systems can also be accounted for in the computation of SOC decay rates.
+As such, the model can reflect the spatial and temporal heterogeneity in both management and biophysical conditions.
 
-The substantial impact of changing management practices in time is indicated by the development of our estimated stock change factors (see Table \@ref(tab:gloSCF)) as well as by the time trend of the SOC debt (see Fig.
+The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Table \@ref(tab:SCFglo)) as well as by the time trend of the SOC debt (see Fig.
 \@ref(fig:SOCmaps)(a)).
-Residue management has changed over the last decades, especially with fading out of residue burning practices in several regions and increased general productivity, this shows a clear impact on the SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
+Residue management has changed over the last decades, especially with the phasing out of residue burning practices in several regions and increased general productivity, showing a clear impact on SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
 
-The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly made for agricultural soils, whereas we apply it also to soils under PNV.
-This is necessary in order to represent SOC losses under land-use change in a dynamics way, which is a main driver of SOC dynamics (cite).
-Also, the comparison against simulated PNV data with LPJmL4 shows substantial skill in reproducing PNV SOC stocks (Fig.
+The Tier 2 approach [@ogle_cropland_in_ipcc_2019] used here is explicitly designed for agricultural soils, whereas we apply it also to soils under PNV.
+This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics.
+The comparison of simulated PNV data with LPJmL4 shows the model's substantial capability in reproducing PNV SOC stocks (Fig.
 \@ref(fig:SOCLPJmL4)).
@@
 However, as the world's SOC stock is highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1; @poggio_soilgrids_2021; @sanderman_soil_2017] in Fig.
-\@ref(fig:SOCglo), the only conclusion we can draw from this is, that our result is within a plausible range.
-
+\@ref(fig:SOCglo), the only conclusion we can draw from this is that our result is within a plausible range.
 To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland.
 Pristine natural vegetated areas (like permafrost and rain forests) without human land management drop out in our calculation of SOC debt and stock change factors.
+Natural SOC estimates only influence results when natural land is converted to cropland.
 
-Focusing on cropland SOC, some differences between our simulated SOC patterns and reference data show specific shortcomings that need to be considered in the application of our model and in the interpretation of results.
-In some mountainous regions, we strongly overestimate SOC stocks compared to SoilGrids 2.0 [@poggio_soilgrids_2021].
-This could indicate that we are not capable to capture specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils [@borrelli_assessment_2017]) or limited accuracy in the SoilGrids 2.0 data for these regions.
+Comparing the geographic SOC patterns to Soil Grids 2.0 [@poggio_soilgrids_2021] (see Fig.
+\@ref(fig:SOCSoilGrids2Q)), we find that our model estimates values of SOC stocks greater than the estimated confidence intervals in Soil Grid 2.0 for some mountainous regions across the globe.
+This could indicate that we are not capable in capturing specific processes that would reduce the vegetation's productivity (such as erosion on steep slopes or shallow soils [@borrelli_assessment_2017]).
+A large swathe of eastern North America was heavily affected by the dust bowl event, with wind erosion removing large parts of the topsoils, a process not considered in our model.
+Similarly, we likely overestimate SOC stocks for the loess soils in northern China and the Altiplano in Latin America; in both cases erosion is a likely reason.
+In contrast, we estimate lower SOC stocks at the edges of the Sahara, where uncertain local water availability and artifical irrigation may dominate spatial SOC patterns.
 
-In comparison with aggregated results from the IPCC guidelines, our model shows too strong SOC losses for tropical regions (Table \@ref(tab:SCFglo)).
-While tropical soils might suffer from low C input rates due to large yield gaps (cite global yield gap atlas?) and high shares of residue removal (MADRAT?
-FAO for residue burning???), initial SOC stocks under PNV are also comparatively low and thus have little potential for large absolute SOC declines.
-The revised IPCC factors also indicate that also relative SOC differences are much smaller in the tropics, which cannot be reproduced well by our model.
+In our model, erosion should however only affect the spatial pattern but not the aggregate SOC pool.
+As pointed out by Doetterl et al.
+[-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017].
+Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focused especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
+Erosion and degradation impacts on yields and therefore on soil C inputs are captured by our method as we base them on FAO statistics of actual production.
+Yet the distribution of production below the country level - which we allocate proportional to LPJmL production potentials that do not reflect erosion feedback to yields - will overestimate yields and therefore biomass inputs to eroded soils.
+
+In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Table \@ref(tab:SCFglo)) for almost all regions.
+Tropical soils might suffer from low C input rates due to large yield gaps (Global Yield Gap and Water Productivity Atlas.
+Available URL: www.yieldgap.org (accessed on: 03/01/2022)) and high shares of residue removal and burning in lower-income countries [@smil_crop_1999; @williams_influence_1997; @jain_emission_2014].
+Yet, even when comparing our etimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values, while it shows very good agreement with the older default values from IPCC [-@lasco_cropland_in_ipcc_2006].
+Don et al.
+[-@don_impact_2011] estimated SOC losses for tropical soils of around 25% on average corresponding to a stock change factor of 0.75, but also reported a wide range of measured SOC changes from -80% to +58%.
+Fujisaki et al.
+[-@fujisaki_forest_2015] however found much lower loss rates of around 9%, attributing the difference to the different time period length since the conversion to cropland.
+As our results do not specifically account for cropland age and most of the cropland is older than 20 years (as assumed for the default IPCC Tier 1 stock change factors) our stock change factors have to be lower by definition following the steady-state assumption that cropland will continue to approach a new equilibrium.
+For the same reason, our estimates for temperate regions might be lower than both IPCC [-@lasco_cropland_in_ipcc_2006] and IPCC [-@ogle_cropland_in_ipcc_2019] default values.
 With the production-increasing impact of irrigation and fertilization on carbon-poor dryland soils, SOC under cropland can also be higher than under PNV with stock change factors above 1 (see Fig.
-\@ref(fig:SOCmaps)(d)), but areas are much smaller than areas where the stock change factors are well below unity.
+\@ref(fig:SOCmaps)(d)), but these areas are much smaller than where the stock change factors are well below unity.
 
-Generally, limiting the analysis to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamic happens in the topsoil.
+Generally, limiting the analysis to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamics happen in the topsoil.
 In this regard several aspects are strongly simplified within our approach.
@@
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Secondly changes to the subsoil due to tillage are neglected.
+Second, changes to the subsoil due to tillage are neglected.
 As Powlson et al.
-[-@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems.
-No-tillage effects may seem larger than they actually are, if only focusing on the topsoil.
-SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices.
-Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
-
-This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion.
-However, as pointed out by Doetterl et al.
-[-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017].
-Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
-
-> where to put this?
-With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors (see Table \@ref(tab:gloSCF)) as also indicated by the time trend of the SOC debt.
-Residue management has changed over the last decades, especially with fading out of residue burning practices in several regions and increased general productivity, this shows a clear impact on the SOC dynamics --- underlining the importance to account for these effects in soil carbon modeling.
+[-@powlson_limited_2014] have shown, the subsoil can be make a large difference in evaluating total SOC losses or gains for no-tillage systems.
+No-tillage effects may seem larger than they actually are if only topsoil is considered.
+SOC transfers to deeper soil layers under tillage might enhance subsoil SOC compared to no-till practices.
+Finally, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
 
@@
 There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period [e.g.
-@friedlingstein_global_2019].
+@frielingstein_glob_carbon_2020].
 There is uncertainty with respect to the overall size of these emissions from different methods and reference points and with respect to the contribution of cropland and agricultural management to these emissions.
-In order to mitigate climate change, there is hope that cropland management can be adapted to sequester atmospheric C in cropland soils [@minasny_soil_2017].
-Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], we find that global cropland has acted as a carbon source since 1975.
-Annual C loss rates of 0.2 per 1000 still have the opposite trend as the promoted 4 per 1000 C sequestration rate target [@minasny_soil_2017].
+In order to mitigate greenhouse gas emissions, it is essential to stop the decline of SOC stocks or even transform cropland management to sequester atmospheric C in cropland soils [@minasny_soil_2017].
+Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], we find that global cropland has acted as a emissions source since 1975.
+Annual C loss rates of 0.2 per 1000 C still have the opposite trend as the promoted 4 per 1000 C sequestration rate target [@minasny_soil_2017].
 Dedicated efforts to increase cropland SOC are thus necessary, as management improvements at historical rates are not enough to counteract ongoing SOC degradation on cropland.
+Yet our study also shows the substantial impact of changing management on the development of SOC debt (Fig.
+\@ref(fig:SOCscen)).
 
@@
 Pugh et al.
-[-@pugh_simulated_2015] considered management effects like tillage and residue returning in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
-Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
-Historical productivity increases were, however, notably larger.
-Despite large spatial heterogeneity, aggregate yield increase rates are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
-
-Our study uses a dynamic management data set as driver for SOC dynamics, including historical yield trends and associated increases in residue inputs.
-[@herzfeld_soc_2021] also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
-Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
-This is consistent with our finding that increasing SOC inputs from above-ground residues had the strongest effect on the slowing-down of the SOC debt increase (Fig.
-@ref(fig:SOCscen)).
-
+[-@pugh_simulated_2015] considered management effects like tillage and incorporation of residues in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
+Their study moreover concludes that yield gains (by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
+Historical yield increases, however, are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
 While we find substantially larger effects of productivity gains than the 1% reported by Pugh et al.
@@
 
-The effects of agricultural productivity on cropland SOC dynamics can be directly accounted for in our modeling approach, whereas process-based studies [@pugh_simulated_2015; @herzfeld_soc_2021] often lack data on relevant management aspects that drive production increases.
+The effects of agricultural productivity on cropland SOC dynamics, including historical yield trends and associated increases in residue inputs, can be directly accounted for in our modeling approach.
+In contrast, process-based studies [@pugh_simulated_2015; @herzfeld_soc_2021] often lack data on relevant management aspects that drive production increases.
+Herzfeld et al.
+[-@herzfeld_soc_2021] also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
+Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
+This is consistent with our finding that increasing SOC inputs from above-ground residues had the strongest effect on the slowing-down of the SOC debt increase (Fig.
+\@ref(fig:SOCscen)).
+
 Elliott et al.
-[-@elliott_management_trends_2018] show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, genetic traits, such as kernel number and radiation use efficiency.
+[-@elliott_management_trends_2018] show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, and genetic traits such as kernel number and radiation use efficiency.
 As such, it will remain challenging for process-based DGVMs to capture the trend of agricultural productivity on cropland SOC dynamics.
 
-Our study emphasizes again that the expansion of croplands is still a major source of CO~2~ emissions --- not only by the removal of vegetation, but also by a slow depletion of C stocks in soils.
+Our study emphasizes again that the expansion of cropland is still a major source of CO~2~ emissions --- not only through the removal of vegetation, but also by a slow depletion of C stocks in soils.
 Our estimates indicate a SOC debt of 39.6 GtC in 2010, and every additional deforestated hectare adds to this debt.
@@
 
-Yet, there is ample potential for further improved SOC management.
+There is also ample potential for further improved SOC management.
 As shown in Fig.
 \@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the soil surface (1.4 GtC).
+However, even larger losses occur at the end of the food supply chain (1.2 GtC year), at the soil surface (1.4 GtC), during residue burning (0.3 Gt C) and with manure management (0.2).
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.
-Waste composting or excreta recycling could represent a major additional N input to cropland soils [@brenzinger_organic_2018].
-Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of c inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
+Waste composting or excreta recycling could represent a major additional C input to cropland soils [@brenzinger_organic_2018].
+Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of C inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools.
 Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs.
 Finally, other carbon-accumulating practices, such as the cultivation of cover crops [@poeplau_carbon_2015; @porwollik_covercrops_2022] and agroforestry [@lorenz_soil_2014] could increase total C sequestration on cropland.
-However, data on adoption rates and patterns are missing to help analyzing their impact on historical SOC trends and the future potential of cover crops and agroforestry for increasing SOC stocks has not been explored yet.

### 2022-05-25 SOCbudgetPaper/methods.Rmd
updated Paper version

@@
 In Sect.
-\@ref(sec:carbonbudget) we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail within the refinement of the IPCC guidelines vol.
-4 [@calvo_buendia_ipcc_2019].
-We additionally describe how we configured and extended the steady-state method (for model code see [@karstens_mrsoil_2020].
+\@ref(sec:carbonbudget) we introduce the basic concept of SOC dynamics as applied in this study and explained in more detail in the refinement of the IPCC guidelines vol.
+4 Chapter 5 on "Cropland" [@ogle_cropland_in_ipcc_2019].
+We additionally describe how we configured and extended the Tier 2 modeling approach [for model code see @karstens_mrsoil_2020].
 In Sect.
@@
 In Sect.
-\@ref(sec:scenarios) we define the scenarios used to complement our historic model results.
+\@ref(sec:scenarios) we define the management scenarios used to analyze the role of different management aspects in historical cropland SOC dynamics.
 
-## SOC stocks and stock changes following the steady-state method {#sec:carbonbudget}
+## SOC stocks and stock changes following the Tier 2 modeling approach {#sec:carbonbudget}
 
-Following the Tier 2 steady-state approach of the refinement of the IPCC guidelines vol.
-4 [@calvo_buendia_ipcc_2019]; referred to as \textit{steady-state method} in the following), we estimate soil organic carbon ($SOC$) stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
+Following the Tier 2 modeling approach of the refinement of the IPCC guidelines vol.
+4 Chapter 5 on "Cropland" [@ogle_cropland_in_ipcc_2019]; referred to as \textit{Tier 2 modeling approach} in the following), we estimate $SOC$ stocks and their change over time for cropland at half-degree resolution from 1975 to 2010.
 We restrict our analysis to the first 0-30 cm of the soil profile.
@@
 
-```{r captionLUtrans, echo=FALSE}
-captionLUtrans <- "Scheme of land-use transition representation for a simplified case.."
+```{r captionLUtrans, echo=FALSE, include=FALSE}
+captionLUtrans <- "Scheme of land-use transition representation.
+Given an initial land-use pattern (as in this example 2 ha land under natural vegetation and 1 ha of cropland), there are separate $SOC$ stocks for natural vegetation and cropland.
+While in this example we assume $SOC$ under natural vegetation to be in steady state, the cropland $SOC$ stock approaches its steady state without having reached it yet (a).
+Upon cropland expansion (in this example half of the natural vegetation is cleared to be used as cropland), $SOC$ stocks on cropland increase due to a transfer of land from natural vegetation (b).
+Explicitly representing newly converted cropland and existing cropland to account for SOC dynamics (c) leads to the same weighted mean value as averaging $SOC$ stocks (d), due to the linearity of Eq.
+\\ref{eq:SOCstate} and cropland-age independent decay rates (see Eq.
+\\ref{eq:decayrates})."
 ```
 
-```{r LUtrans, out.width = "16cm", fig.cap = captionLUtrans, echo=FALSE, fig.pos="h"}
+```{r LUtrans, out.width = "8cm", fig.cap = captionLUtrans, echo=FALSE, fig.pos="h"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images//LUtransition.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/LUtransitions.png")
 ```
 
-We calculate $SOC$ stocks based on the area shares of land-use types ($lu$) within our grid cells ($i$).
+We calculate $SOC$ stocks based on the area shares of land-use types $lu$ within our grid cells $i$.
 If land is converted from one land-use type $lu=\{crop,natveg\}$ into the other $!lu=\{natveg,crop\}$, a respective share of the $SOC$ is reallocated within our budget.
-As we do not distinguish between newly converted and old cropland (see \@ref(fig:lutransitions)), we update the average SOC stock per area within this modeling step.
+We do not distinguish between newly converted and existing cropland, but can work with the average carbon content as the relative decay of $SOC$ is proportional to the $SOC$ stock (see \@ref(fig:LUtrans)).
 We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{t^*}$ via
@@
 \begin{equation}
-SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific SOC stock within cell}}}_{\text{$SOC_{i,t}$ -- total SOC stock within cell}}.
+SOC_{t} = \sum_{i} \underbrace{\sum_{lu} \overbrace{\sum_{sub} SOC_{i,t,sub,lu}}^{\text{$SOC_{i,t,lu}$ -- land-use type specific $SOC$ stock within cell}}}_{\text{$SOC_{i,t}$ -- total $SOC$ stock within cell}}.
 (\#eq:totalstock)
@@
 According to the IPCC guidelines $SOC$ changes can be expressed as the difference of two consecutive years [see Eq.
-5.0A in @calvo_buendia_ipcc_2019].
+5.0A in @ogle_cropland_in_ipcc_2019].
 This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we defined the absolute and relative SOC changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
+For our study, we define the absolute and relative $SOC$ changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
 The absolute changes $\Delta SOC$ and relative changes $F^{\mathrm{SCF}}$ are thus given by
@@
 
-We initialize our SOC sub-pools using a three-step approach, since input data availability is limited for climate and litter (from 1901) as well as for agricultural management data (from 1965):
+The initialization of $SOC$ pools is very important and has to include the proper accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historical land-use change (LUC) and the slow decline of $SOC$ under crop cultivation, before it reaches a new equilibrium.
+We initialize our $SOC$ sub-pools using a three-step approach, since input data availability is limited for climate and litter estimates (starting only in 1901) as well as for agricultural management data (starting only in 1965):
 
-Firstly, to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we account for land-use change from 1510.
-We repeat climate and subsequently also litter input data from 1901-1930 to simulate a constant climate for this first initializing period.
-Agricultural management data is hold constant at the level of 1965.
-In 1510 we assume all SOC pools to be in natural steady-state, implicating that all land-use change emissions prior to that time, are redistributed to 1510.
-Due to this long first spin-up period we suppose that the effect of this late accounting and appertaining delayed legacy fluxes can be neglected in 1901.
-This approaches follows others studies looking on effects of land-use change and management (cite LPJmL (which?)).
+Firstly, in order to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we consider land-use change from 1510 onwards.
+In 1510, we assume all $SOC$ pools to be in natural steady-state, implying that all land-use change prior to that time occurs in 1510.
+We assume that by 1901, all cropland converted in 1510 has reached its new steady state, so that it is not necessary to explicitly account for even older land conversion.
+Model inputs for 1901--1930 for climate and natural vegetation litterfall are repeated for 1510--1900 to mimic constant climate conditions for this first initializing period.
+Similarly, agricultural management data are held constant at the level of 1965 until 1965.
+This approach follows others studies looking on effects of land-use change and management [e.g.
+@schaphoff_lpjml4_2018-1; @herzfeld_soc_2021].
 
-Secondly, to account for the impact of a changing climate, we account not only for land-use change, but also for historical climate and consequently natural litter inputs to the soil, from 1901 to 1965 still considering constant agricultural input data.
+Secondly, with the availability of transient climate data after 1901, we account not only for land-use change, but also for historical climate change and consequently natural litter inputs to the soil from 1901 to 1965 still considering constant agricultural input data, which are not available prior to 1965.
 
-Thirdly, we run the model for additional 10 years with historic dynamic input data and start analyzing results from 1975 onward.
-This is in line with IPCC method suggestion to have a 5-20 year spin-up period.
+Thirdly, we run the model for 10 years from 1965 to 1975 with historical dynamic data on agricultural management and start analyzing results from 1975 onward.
+This is in line with the IPCC guidelines vol.
+4 method suggestion to have a 5-20 year spin-up period [@ogle_cropland_in_ipcc_2019].
 
-Irrigation areas are part of the water effect associated with the climate data set and therefore dynamic from 1901 on, whereas data on no-tillage areas is only available after 1974.
+With transient climate considered after 1901, decay rates $k_{sub}$ become dynamic in time.
+As the decay rates are also affected by irrigation and tillage (see Sect.
+\@ref(sec:steadystates)), we also account for transient changes in irrigated areas after 1901.
+Data on no-tillage practices are only available after 1974 and we assume conventional tillage on all cropland prior to 1975.
 
-<!-- backup
-```{=latex}
- \begin{table*}[h]
- \caption{Setup of spin-up and simulation phase}
- \begin{tabular}{l l l}
- \tophline
- \multirow{3}{*}{1901--1964} 
-   & Constant management spin-up phase with
-   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
-   & constant management data from 1965 (manure and residue inputs, tillage and irrigation areas)\\
-                      \cline{1-4}
- \multirow{3}{*}{1965--1974}
-   & Dynamic management spin-up phase with
-   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
-   & historic management data (manure and residue inputs, tillage and irrigation areas)
-                      \cline{1-4}
- \multirow{3 }{*}{1975--2010}                     
-   & Simulation phase 
-   & historic land-use change (LUH2v2), climate (CRU4p2) and litter inputs (LPJmL4)
-   & historic management data (manure and residue inputs, tillage and irrigation areas)              
- \bottomhline
- \end{tabular}
- \belowtable{}
- (\#tab:datasourcedecay)
- \end{table*}
-``` -->
 ## SOC stocks and stock changes following Tier 1 {#sec:tier1}
 
-Additionally to the steady-state method [@calvo_buendia_ipcc_2019] and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019].
+Additionally to the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019] and the detailed analysis of management data coming with it, $SOC$ changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019].
 Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
-Estimates of $F^{\mathrm{SCF}}$ are differentiated by different crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these.
+Estimates of $F^{\mathrm{SCF}}$ are differentiated by crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows.
 Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig.
@@
 
-Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set, which we sum up from quarter-degree to half-degree resolution.
-We disaggregate the physical area of the five different cropland subcategories (c3ann, c3per, c4ann, c4per, c3nfx) of LUH2 into our 17 crop groups [see FAO2LUH2MAG_croptypes.csv in @bodirsky_mrcommons_2020], applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016].
-By calculating country-specific multicropping factors using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
+Land-use patterns are based on the Land-Use Harmonization 2 [@hurtt_harmonization_2020] data set (short LUH2), which we sum up from quarter-degree to half-degree resolution.
+We disaggregate the physical area (given as total land area in million ha) of the five different cropland subcategories (c3ann: C3 annual crops, c3per: C3 perennial crops, c4ann: C4 annual crops, c4per: C4 perennial crops, c3nfx: C3 nitrogen-fixing crops) of LUH2 into our 17 crop groups [see Table "FAO2LUH2MAG_croptypes.csv" in @karstens_model_2020], applying the relative shares for each grid cell based on the country- and year-specific area harvested shares of FAOSTAT data [@faostat_faostat_2016].
+By calculating country-specific multicropping factors $MCF$ using FAOSTAT data, we are able to compute crop-group specific area harvested on grid cell level.
 Land-use transitions are calculated as net area differences of the land-use data at half-degree resolution, considering no split up into crop-group specific areas but only total cropland and natural vegetation areas.
@@
 \@ref(sec:landuse)).
-We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production [@faostat_faostat_2016].
-By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.
+We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production [@faostat_faostat_2016].
+Note
 
 Crop residue production and management is based on a revised methodology of Bodirsky et al.
-[-@bodirsky_n2o_2012] and key aspects are explained here given its central role for soil carbon modeling.
-Starting from crop yield estimates ($Y$) and respective harvested crop area ($CA$), we estimate above-ground ($AGR$) and below-ground ($BGR$) residual biomass using crop group ($cg$) specific ratios for above-ground residues to harvested biomass ($r^{\mathrm{ag,prod}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{tC}}$), above-ground residues to harvested area ($r^{\mathrm{ag,area}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{ha}}$) and below-ground residues to above-ground biomass ($r^{\mathrm{bg}}_{cg}$ in $\tfrac{\unit{tC}}{\unit{tC}}$) as follows
+[-@bodirsky_n2o_2012] and key aspects are explained as they play a central role in soil carbon modeling.
+Starting from crop yield estimates $Y$ and respective physical crop area $CA$, we estimate total above-ground $AGR$ and below-ground $BGR$ residue biomass (in tonnes) using crop group $cg$ specific ratios for above-ground residues to harvested biomass $r^{\mathrm{ag,prod}}_{cg}$ in $(tDM\,ha^{-1})(tDM\,ha^{-1})^{-1}$ above-ground residues to harvested area $r^{\mathrm{ag,area}}_{cg}$ in $\unit{tDM\,\unit{ha}^{-1}}$ and below-ground residues to above-ground biomass $r^{\mathrm{bg}}_{cg}$ in $\unit{tDM\,\unit{tDM}^{-1}}$ as follows
 
@@
 A country-specific fixed share of the $AGR$ is assumed to be burned on field depending on the per-capita income of the country.
-Following Smil [-@smil_nitrogen_1999] we assume a burn share of $25\%$ for low-income countries according to World Bank definitions ($<\,1000\,\tfrac{USD}{yr\,cap}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr\,cap}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income.
+Following Smil [-@smil_nitrogen_1999] we assume a burn share of 25% for low-income countries according to World Bank definitions ($<\,10000\,\unit{USD}\,\unit{yr}^{-1}\,\unit{cap}^{-1}$), 15% for high-income  ($>\,10000\,\unit{USD}\,\unit{yr}^{-1}\,\unit{cap}^{-1}$) and linearly interpolate shares for all middle-income countries depending on their per-capita income for the periods before 1995.
+After 1995 we estimate a linear decline of burn shares to 10% for low-income countries and 0% for high-income countries till 2025 to account for recent increases in air pollution regulation.
+The estimated trends show good agreement with fire-satellite-image derived estimates by the Global Fire Database [@vanderWerf_fire_2017].
 Depending on the crop group, 80–90% of the carbon in the crop residues burned in the fields is lost within the combustion process [@eggleston_ipcc_2006].
@@
 From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.
-[-@bodirsky_mrcommons_2020].
+[-@bodirsky_mrcommons_2020]).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets [see @weindl_livestock_2017], that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
@@
 The remaining $AGR$ as well as all $BGR$ are expected to be left on the field.
-We limit high residue return rates to at most $10\unit{tC\,ha}^{-1}$ in order to correct for outliers.
+We limit high residue return rates to at most $10\,\unit{tC\,ha}^{-1}$ in order to correct for outliers.
 
-To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon resp. nitrogen to dry matter (c/dm, n/dm) ratios (see \@ref(tab:c2dm)).
+To transform dry matter estimates into carbon and nitrogen, we compiled crop-group and plant-part specific carbon and nitrogen to dry matter ratios (see Table \@ref(tab:c2dm)).
 
@@
 [-@bodirsky_n2o_2012].
-We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.
+We limit high application shares at $10\,\unit{tC\,ha}^{-1}$ to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
 

### 2022-05-25 SOCbudgetPaper/results.Rmd
updated Paper version

--- 2020-12-17 SOCbudgetPaper/results.Rmd
@@
 Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens [-@karstens_model_2020].
-In the following, the most important results [see @karstens_result_2020] for post-processing script) are summarized.
+In the following, the most important results [see @karstens_result_2020 for post-processing script] are summarized.
 
@@
 ```{r captionSOCmaps, echo=FALSE}
-captionSOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas.
-(b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics.
-Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas.
-(d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010.
-"
+captionSOCmaps <- "Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data.
+Panel (a) shows global $\\Delta SOC$ between historical land use and potential natural vegetation (PNV).
+The distribution of total global SOC stocks for the first 30 cm on cropland for the year 2010 is depcited in panel (b).
+Absolute (c) and relative (d) SOC stocks changes for the year 2010 are compared to a potential natural state identify different hotspots of SOC losses and gains."
 ```
@@
 
+The global SOC debt has increased by about 14% in the period between 1975 and 2010 from $34.6$ to $39.6\,\unit{GtC}$ (Fig.
+\@ref(fig:SOCmaps)(a)).
+This corresponds to an average loss rate of $0.14\,\unit{GtC\,yr^{-1}}$ in comparison to a hypothetical potential natural vegetation (PNV) state.
+Considering our estimate of the global SOC stock of around $705\,\unit{GtC}$ in the upper $30\,\unit{cm}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+The speed of this SOC loss has decreased towards the end of the modeling period.
+
 In Fig.
-\@ref(fig:SOCmaps)(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010.
-Values ranging between well over $100\unit{t ha^{-1}}$ in northern temperate croplands to less than $5\unit{t ha^{-1}}$ for arid and semiarid croplands.
-Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways: 
+\@ref(fig:SOCmaps)(b) we provide a world map of SOC stocks estimates for the first $30\,\unit{cm}$ on cropland considering historical management data for the year 2010.
+Values range between over $100\,\unit{t\,ha^{-1}}$ in northern temperate cropland to less than $5\,\unit{t\,ha^{-1}}$ for arid and semiarid cropland.
+Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under PNV in two complementary ways: 
 1.
-Absolute SOC changes $\Delta SOC$ (see Fig.
-\@ref(fig:SOCmaps)(b)) indicate areas with high importance for the global SOC losses.
-They might be driven by large relative changes (e.g.
+Absolute SOC changes $\Delta SOC$ (Fig.
+\@ref(fig:SOCmaps)(c)) indicate areas with high importance for the global $SOC$ loss.
+They can be driven by large relative changes (e.g.
 in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.
 North-East Asia).
-In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
 2.
-Relative SOC changes measured as stock changes factors $F^{SCF}$ (see Fig.
-\@ref(fig:SOCmaps)(c)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices.
-Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation.
-Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.
+Relative SOC changes measured as stock changes factors $F^{SCF}$ (Fig.
+\@ref(fig:SOCmaps)(d)) are a helpful metric to analyze the impact of human cropping activities.
+They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome by improved agricultural practices.
+Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation.
+Conversely, irrigated cropland at the border to dry, unsuitable areas worldwide shows a strong relative increase in SOC stocks.
 
-The global SOC debt has decreased by about 15% in the period between 1975 and 2010 to $22\unit{GtC}$ (Fig.
-\@ref(fig:SOCmaps)(d)).
-This corresponds to a sequestration rate of $0.11\unit{GtC yr^{-1}}$.
-Considering our estimate of the global SOC stock of around $660\unit{GtC}$ in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.
+```{r captionSOCdebt, echo=FALSE}
+captionSOCdebt <- "Global total $\\Delta SOC$ and $\\Delta SOC$ change for the first 30 cm of the soil profile.
+Panel (a) shows global $\\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types.
+Computing the difference between the $\\Delta SOC$ estimate for 2010 and for 1975 (b) depicts areas of soil depletion (SOC debt increase, red) and net-sequestration (SOC debt decline, blue)."
+```
+
+```{r SOCdebt, out.width = "9cm", fig.cap = captionSOCdebt, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/2panelfigure.png")
+```
+The spatial distribution of the total $\Delta SOC$ summed over all land-use types (Fig.
+\@ref(fig:SOCdebt)(a))  and its change from 1975 to 2010 (Fig.
+\@ref(fig:SOCdebt)(b)) reveals areas of SOC debt decline and increase.
+Regions with large cropland expansion (e.g.
+Brazil, Southeast Asia, Canada) continue to lose SOC, whereas regions with cropland reduction (and thus SOC restoration) or with accumulating cropland SOC can be found e.g.
+in highly productive areas of Europe and Central USA.
 
@@
 ```{r caption_FlowFig, echo=FALSE}
-caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues.
-Large parts of C, however, get mineralized on field before entering the soil.
-Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation.
-Finally, SOC gets mineralized and flows back into the atmosphere."
+caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in $\\unit{MtC}$).
+Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand.
+After accounting for losses within the agricultural system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues.
+Large parts of C, however, are mineralized on the field before entering the soil.
+Additionally, C is transferred to and from the global agricultural soil stock via land-use change between cropland and natural vegetation.
+Finally, SOC is mineralized and flows back to the atmosphere."
 ```
@@
 
-C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues).
-Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field.
-We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.
-like food, bioenergy and material) into a human demand category.
-Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC.
-Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.
-compost, night soil and sewage) is not recycled to soils.
-Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010.
-However, around 60% of C decomposes before its integrated into soils at the the litter-soil barrier.
-Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue.
-According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010.
-4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system.
-Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.
+C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvested organ, above- and below-ground residues).
+Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues ($729\,\unit{MtC}$ in 2010) are directly returned to the field.
+We divide crop biomass usage into feed usage and aggregate all other usage types (e.g.
+food, bioenergy and material) into a human demand category.
+Livestock feed demand for crop organ harvest and above-ground residues of $1136\,\unit{MtC}$ is roughly equal to the human demand of $1129\,\unit{MtC}$.
+Whereas large parts of feed intake are returned to the soils via manure (C input from manure at $384\,\unit{MtC}$), we assume the carbon demanded from humans (ending up as e.g.
+compost, night soil and sewage) is not returned to soils.
+Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with $1350\,\unit{MtC}$ returned to the fields in the year 2010.
+However, around 60% of this organic C decomposes before it is integrated into soils.
+Due to the different composition of organic C, proportionally more C enters the slow pool from manure than from crop residue.
+According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of $257\,\unit{MtC}$ in 2010.
+The agricultural system receives $4585\,\unit{MtC}$ assimilated by crop plants and releases $3554\,\unit{MtC}$ mostly through respiration.
+Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around $33\,\unit{MtC}$ for the year 2010.
 
@@
 ```{r captionSOCscen, echo=FALSE}
-captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage).
-ConstManagement combines all three modifications.
-Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state.
-Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
+captionSOCscen <- "(a) Global $\\Delta SOC$ in $\\unit{GtC}$ for different management scenarios.
+The stylized scenarios deviate from historical agricultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage).
+The scenario constManagement combines all three modifications.
+Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a hypothetical natural vegetation state.
+Panel (b) shows the carbon inflows from crop residue and manure."
 ```
@@
 
-We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
-Without changes in management regimes, the global $\Delta SOC$ on cropland would be increasing at a rate of $0.1\unit{GtC yr^{-1}}$.
-As shown by the constResidue scenario, changes in residue return rates dominate the management effects.
-Without the historic increase in residue returning, the global $\Delta SOC$ would still increase at a rate of $0.06\unit{GtC yr^{-1}}$.
-Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of $0.09\unit{GtC yr^{-1}}$ for both).
-The effect of no-tillage has been particuarly strong since the 1990s.
-The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \@ref(fig:SOCscen)(b).
+We analyze the relative impact of different management practices by comparing the actual historical management scenario with counterfactual scenarios, where individual management practices (residues in constResidues, manure in constManure, tillage practices in constTillage, all three in constManagement) are kept static at the 1975 values (Figure \@ref(fig:SOCscen)(a)).
+As shown by the difference between the constResidues scenario and the other counterfactuals, changes in residue return rates dominate the management effects.
+Without the historical increase in C inputs from residues to agricultural soils, the global $\Delta SOC$ would decrease to $41.7\,\unit{GtC}$ at a rate of $0.20\,\unit{GtC\,yr^{-1}}$ --- a 35% increase compared to $0.14\,\unit{GtC\,yr^{-1}}$ for the histManagement estimates.
+Both the constManure and constTillage scenarios show only small deviations from the historical agricultural management values with $0.15\,\unit{GtC\,yr^{-1}}$.
+The effect of no-tillage only becomes discernible from 2000 onwards.
+The large contribution of residues relative to manure also becomes visible when considering the annual C inputs of residues and manure to soils over a period of 35 years (Fig. \@ref(fig:SOCscen)(b)).
 
-```{r captionSOCinit, echo=FALSE}
-captionSOCinit <- "Global $\\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use."
+## Model evaluation {#sec:validate}
+
+To evaluate our model results against reference data in five steps: (1) we compare our stock change factors (see Sect.
+\@ref(sec:tier1)) to IPCC default assumptions [@lasco_cropland_in_ipcc_2006; @ogle_cropland_in_ipcc_2019]; (2) we compare our global (and climate-zone specific) total SOC stocks to other literature estimates; (3) we compare our results to point measurements.
+To evaluate the representation of our natural SOC stocks (4) we correlated LPJmL4 SOC stocks for PNV with our natural state SOC results on grid level; and (5) we do a similar correlation analysis for our modeled actual SOC stocks in comparison to the results of SoilGrids 2.0 [@poggio_soilgrids_2021], which accounts for actual land use too.
+
+### Stock change factors compared to IPCC assumptions {#sec:ipcccompare}
+
+To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors $F^{\mathrm{SCF}}$ of 2006 [@lasco_cropland_in_ipcc_2006] and their refinements in 2019 [@ogle_cropland_in_ipcc_2019].
+Both estimates are based on measurement data for cropland (see Table \@ref(tab:SCFglo)).
+To allow for comparison, we aggregate our stock change factors weighted by grid-level $F^{\mathrm{SCF}}$ cropland area to derive average factors for the four IPCC climate zones (Fig.
+\@ref(fig:CLIMzone)).
+
+```{r SCFglo, echo=FALSE, results = 'asis'}
+table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
+library(xtable)
+print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
+              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \\citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \\citep{ogle_cropland_in_ipcc_2019}."), comment=FALSE, caption.placement = "top")
+```
+Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC.
+For the tropical regions the IPCC factors changed notably from the guidelines in 2006 [@lasco_cropland_in_ipcc_2006] to the update in 2019 [@ogle_cropland_in_ipcc_2019].
+Our results are in good agreement with the 2006 IPCC factors.
+Modeled $F^{SCF}$ have increased or stayed constant for all climate zones over time (1975-2010).
+
+### Global SOC stocks comparison
+
+We compare our global SOC stocks with a wide range of global SOC stock estimates for the first 30 cm of the soil profile, using data from WISE [@batjes_harmonized_2016], SoilGrids [@hengl_soilgrids250m_2017], GSOC [@fao_global_2018], LPJmL4 [@schaphoff_lpjml4_2018-1], SoilGrids 2.0 [@poggio_soilgrids_2021], and SOCDebtPaper [@sanderman_soil_2017] in Fig.
+\@ref(fig:SOCglo).
+
+```{r captionSOCglo, echo=FALSE}
+captionSOCglo <- "Modeled as well as observation-based estimates for global SOC stock in $\\unit{GtC}$ for the first 30 cm of soil aggregated over all land area.
+The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \\citep{schaphoff_lpjml4_2018-1} and estimates from \\citep{sanderman_soil_2017}.
+We show values of this study for the year 2010 accounting for the historical land-use dynamics as well as for an hypothetical PNV."
 ```
 
-```{r SOCinit, out.width="10cm", fig.cap = captionSOCinit, echo=FALSE, fig.pos="h"}
+```{r SOCglo, out.width = "11cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_init.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
 ```
 
-Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig.
-\@ref(fig:SOCinit)).
-The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901.
-The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC), as does the decrease in $\Delta SOC$ for the period of 1975--2010 (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.
+The global estimates of the total SOC stock of the upper 30 cm from this study are in the middle of the wide range of other modeled or observation-based estimates.
+SoilGrids [@hengl_soilgrids250m_2017] especially stands out with its high estimate, whereas SoilGrids 2.0 [@poggio_soilgrids_2021] marks the lower end.
+Regional results (Fig.
+\@ref(fig:SOCreg)) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.
 
-Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas $\Delta SOC$ vary  19-24 GtC for the year 2010 (see Fig.
-\@ref(fig:SOClitter).
-The increase of $\Delta SOC$ during the period of 1975--2010 is independent of the parameterization choice at ~4 GtC.
+### Point-based evaluation
+
+We correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements (for data base see appendix of [@sanderman_soil_2017]).
+
+```{r captionSOCpoint, echo=FALSE}
+captionSOCpoint <- "Correlation between modeled and measured $SOC$ stocks.
+Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results.
+Both cropland and areas with natural vegetation tend to be lower in our results than in the point measurements."
+```
+
+```{r SOCpoint, out.width = "10cm", fig.cap = captionSOCpoint, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scatterComparePointData_2010.png")
+```
+
+### Natural SOC stock comparison with LPJmL4
+
+Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time.
+We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation.
+Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.
+With our focus on cropland SOC dynamics, we compare only cells with more than $1000\,\unit{ha}$ of cropland (capturing 99.9% of global cropland area).
+
+```{r captionSOCLPJmL4, echo=FALSE}
+captionSOCLPJmL4 <- "Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line."
+```
+
+```{r SOCLPJmL4, out.width = "13cm", fig.cap = captionSOCLPJmL4, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png")
+```
+Spatial correlations of PNV SOC stock values are high (global $R2=0.81$), especially for dry climate zones (Fig.
+\@ref(fig:SOCLPJmL4)).
+For temperate and tropical moist areas estimates of this study tend to be a bit lower compared to LPJmL4 results.
+
+### Actual SOC stock comparison with SoilGrids 2.0
+
+SoilGrids 2.0 [@poggio_soilgrids_2021] is a digital soil mapping approach that uses over 240 000 soil profile observations to produce high resolution soil maps including SOC stocks and estimates of their uncertainties.
+To evaluate the performance of our model at the global scale, we correlate SoilGrids 2.0 SOC stock values, which were aggregated to 0.5 degree resolution, to our estimates for the year 2010 in Fig.
+\@ref(fig:SOCSoilGrids2).
+To focus our comparison on cropland areas, we mask out grid cells with less than $1000\,\unit{ha}$ of cropland.
+Spatial correlation is moderate for tropical climate zones, whereas it is low for temperate moist areas.
+In tropical dry and temperate dry areas, we simulate also very low SOC values (below $10\,\unit{tC\,ha^{-1}}$), which is not found in SoilGrids 2.0 whereas our modeled SOC stocks can be substantially higher in temperate moist areas than reported by SoilGrids 2.0.
+Additionally, we use the uncertainty estimates from SoilGrids 2.0 in Fig.
+\@ref(fig:SOCSoilGrids2Q) to identify areas, where our modeled SOC stocks that are below the 5th or above the 95th percentile of the SoilGrids 2.0 data.
+For the vast majority of grid cells our model results are between the 5th and 95th percentile of SoilGrids 2.0 estimates.
+We underestimate SOC stocks especially in dry areas (e.g.
+close to the Sahara).
+Overestimated stocks are often situated in mountainous regions.
+
+```{r captionSOCSoilGrids2, echo=FALSE}
+captionSOCSoilGrids2 <- "Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0."
+```
+
+```{r SOCSoilGrids2, out.width = "13cm", fig.cap = captionSOCSoilGrids2, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareSoilGrids2_2010_lu.png")
+```
+
+```{r captionSOCSoilGrids2Q, echo=FALSE}
+captionSOCSoilGrids2Q <- "Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0."
+```
+
+```{r SOCSoilGrids2Q, out.width = "14cm", fig.cap = captionSOCSoilGrids2Q, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/SOCquantilesSoilGrids2.png")
+```

### 2022-06-14 ResultNotebooks/results_SOCpointvalidate.Rmd

---
title: "SOC budget analysis in comparsion"
output:
  html_document:
    df_print: paged
  pdf_document: default
  bookdown::pdf_document2: default
   keep_md: true
---

```{r load}
library(magclass)
library(madrat)
library(mrcommons)

outputdir <- "Output/Images/"

rev         <- "rev30"
scen        <- "histManagement"

SoilCarbon <- readRDS(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
Landuse    <- readRDS(paste0("Data/", rev, scen, "/Landuse.rds"))

```

```{r extract carbon stock for validation year}

year            <- "y2010"
SOCStock_natveg <- dimSums(SoilCarbon[, year, c("actualstate.natveg")], dim = 3)
SOCStock_crop   <- dimSums(SoilCarbon[, year, c("actualstate.crop")],   dim = 3)

SOCcrop_table      <- as.data.frame(setCells(SOCStock_crop, gsub("\\.", "_", getCells(SOCStock_crop))))
SOCcrop_table$Cell <- gsub("_", "\\.", SOCcrop_table$Region)
SOCcrop_table$Native  <- 0
SOCcrop_table         <- SOCcrop_table[-c(2:4)] 
names(SOCcrop_table)[names(SOCcrop_table) == "Value"] <- "thisStudy"

SOCnatveg_table      <- as.data.frame(setCells(SOCStock_natveg, gsub("\\.", "_", getCells(SOCStock_natveg))))
SOCnatveg_table$Cell <- gsub("_", "\\.", SOCnatveg_table$Region)
SOCnatveg_table$Native  <- 1
SOCnatveg_table         <- SOCnatveg_table[-c(2:4)] 
names(SOCnatveg_table)[names(SOCnatveg_table) == "Value"] <- "thisStudy"
```

### Sanderman validation data (point data)

```{r reformat point validation data for plotting, warning=FALSE}

  library(stringr)

  sand  <- readRDS(paste0("Data/", rev, "histManagement/calcSOCPointData.rds"))
  getCells(sand)  <- gsub("\\.", "_", getCells(sand)) 
  sand_table      <- as.data.frame(sand)
  sand_table      <- reshape2::dcast(sand_table, Region ~ Data1, value.var = "Value")
  sand_table$ID   <- str_extract(sand_table$Region, "[^_]*$")
  sand_table$Cell <- gsub("_", "\\.", str_extract(sand_table$Region, "[^_]*_[^_]*"))
  sand_table      <- sand_table[names(sand_table) %in% c("Cell", "Native", "SOC", "ID")]
  names(sand_table)[names(sand_table) == "SOC"] <- "ValidationData"

  getCells(sand) <- str_extract(getCells(sand), "[^_]*_[^_]*")
  unique         <- unique(getCells(sand))
  names          <- as.vector(outer(c("natveg", "crop"), 
                                    c("median", "mean", "max", "min", "q25", "q75"), 
                                    paste, sep = "."))
  sand_reduced   <- new.magpie(unique, NULL, names, fill = NA)

  for(u in unique) {
    sand_reduced[u, , "natveg.mean"]   <- mean(    as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
    sand_reduced[u, , "crop.mean"]     <- mean(    as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
    sand_reduced[u, , "natveg.median"] <- median(  as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
    sand_reduced[u, , "crop.median"]   <- median(  as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
    sand_reduced[u, , "natveg.max"]    <- max(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
    sand_reduced[u, , "crop.max"]      <- max(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
    sand_reduced[u, , "natveg.min"]    <- min(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
    sand_reduced[u, , "crop.min"]      <- min(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
    sand_reduced[u, , "natveg.q25"]    <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1], probs = 0.25)
    sand_reduced[u, , "crop.q25"]      <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0], probs = 0.25)
    sand_reduced[u, , "natveg.q75"]    <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1], probs = 0.75)
    sand_reduced[u, , "crop.q75"]      <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0], probs = 0.75)
  }

  sand_crop_table <- reshape2::dcast(as.data.frame(sand_reduced[, , "crop"]), 
                                        Region ~ Data2, value.var = "Value")
  sand_crop_table$Cell   <- gsub("_", "\\.", sand_crop_table$Region)
  sand_crop_table$Native <- 0
  sand_crop_table        <- sand_crop_table[complete.cases(sand_crop_table), ][-1]

  sand_natveg_table <- reshape2::dcast(as.data.frame(sand_reduced[, , "natveg"]), 
                                        Region ~ Data2, value.var = "Value")
  sand_natveg_table$Cell   <- gsub("_", "\\.", sand_natveg_table$Region)
  sand_natveg_table$Native <- 1
  sand_natveg_table        <- sand_natveg_table[complete.cases(sand_natveg_table), ][-1]

```

```{r merge validation and output data}

allPoints   <- rbind(merge(sand_table, SOCcrop_table, by = c("Cell", "Native")),
                     merge(sand_table, SOCnatveg_table, by = c("Cell", "Native")))

statsPoints <- rbind(merge(sand_crop_table,   SOCcrop_table,   by = c("Cell", "Native")),
                     merge(sand_natveg_table, SOCnatveg_table, by = c("Cell", "Native")))

correlations <- new.magpie("GLO", NULL, names = unlist(outer(c("median", "mean", "allP"), 
                                                            c("crop", "natveg", "allLU"),
                                                            paste, sep = ".")))

correlations[, , "median.allLU"]  <- cor.test(statsPoints$median, statsPoints$thisStudy)$estimate^2
correlations[, , "mean.allLU"]    <- cor.test(statsPoints$mean,   statsPoints$thisStudy)$estimate^2

correlations[, , "median.natveg"] <- cor.test(statsPoints$median[statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
correlations[, , "mean.natveg"]   <- cor.test(statsPoints$mean[  statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
correlations[, , "median.crop"]   <- cor.test(statsPoints$median[statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2
correlations[, , "mean.crop"]     <- cor.test(statsPoints$mean[  statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2

correlations[, , "allP.allLU"]    <- cor.test(allPoints$ValidationData, allPoints$thisStudy)$estimate^2
correlations[, , "allP.natveg"]   <- cor.test(allPoints$ValidationData[allPoints$Native == 0], allPoints$thisStudy[allPoints$Native == 0])$estimate^2
correlations[, , "allP.crop"]     <- cor.test(allPoints$ValidationData[allPoints$Native == 1], allPoints$thisStudy[allPoints$Native == 1])$estimate^2

```

```{r sanderman2, fig.height=8, fig.width=8}

library(ggplot2)
theme_set(theme_bw())
legend_size <- 14
scen_cols   <- c("#785EF0", "#FE6100")
input_cols  <- scen_cols[1:4]

SOCPointValid  <- ggplot() + 
                  geom_point(data = statsPoints, 
                            aes(alpha = str_wrap("Median (R2=0.13)", 7), x = median, y = thisStudy, colour = factor(Native)),
                            shape = 3, size = 2) +
                  geom_segment(data = statsPoints,
                               aes(alpha = "Min-Max", x = min, y = thisStudy, xend = max, yend = thisStudy, 
                                   colour = factor(Native)), 
                               size = 0.5) +
                  geom_point(data = allPoints, 
                             aes(alpha = "Point Data", x = ValidationData, y = thisStudy, colour = factor(Native)),
                             shape = 1, size = 0.5) +
                  labs( x = "Point measurement data (tC/ha) ", y = "Predicted value by this study (tC/ha)") +
                  scale_x_continuous(trans = 'log10', limits = c(7, 470)) + 
                  scale_y_continuous(trans = 'log10', limits = c(7, 470)) +
                  coord_fixed(1) + theme(aspect.ratio=1) +
                  geom_abline(slope=1, intercept=0, color = "grey", size = 0.5, linetype = "dashed") + 
                  theme(rect=element_blank(),
                        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
                        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
                        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
                        axis.text.y = element_text(size = legend_size),
                        legend.position= c(0.87, 0.2), legend.direction="vertical",
                        legend.title = element_text(size = legend_size), 
                        legend.text  = element_text(size = legend_size),
                        legend.spacing.y = unit(0.1, "cm")) +
                  scale_color_manual(values = scen_cols, labels = c("Cropland", str_wrap("Natural vegetation", 7)), name = "") +
                  scale_alpha_manual(name = NULL,
                                     values = c(1, 1, 1),
                                     breaks = c(str_wrap("Median (R2=0.13)", 7), "Min-Max", "Point Data"),
                                     guide = guide_legend(override.aes = list(linetype = c(0, 1, 0),
                                                                                 shape = c(3, NA, 1),
                                                                                  size = c(2, 0.5, 0.5), 
                                                                                 color = "black")))

SOCPointValid

ggsave(paste0(outputdir,"scatterComparePointData_2010.png"), plot=SOCPointValid, height=16, width=16, units="cm", dpi = 300)

```

```{r sanderman4, fig.height=7.5, fig.width=7.5}

year         <- "y2010"

sand_data    <- readRDS("Data/additional/dataframe.rds")

SOCcrop_data <- as.data.frame(setCells(SOCStock_crop_raw[, year, ], 
                                       gsub("\\.", "_", getCells(SOCStock_crop_raw))))

SOCcrop_data$celliso <- gsub("_", "\\.", SOCcrop_data$Region)
SOCcrop_data$native  <- 0
SOCcrop_data <- SOCcrop_data[-c(1:4)]

SOCnatveg_data <- as.data.frame(setCells(SOCStock_natveg_raw[, year, ], 
                                       gsub("\\.", "_", getCells(SOCStock_crop_raw))))

SOCnatveg_data$celliso <- gsub("_", "\\.", SOCnatveg_data$Region)
SOCnatveg_data$native  <- 1
SOCnatveg_data <- SOCnatveg_data[-c(1:4)]

test <- rbind(merge(sand_data, SOCcrop_data, by = c("celliso", "native"))[-c(3, 4)],
              merge(sand_data, SOCnatveg_data, by = c("celliso", "native"))[-c(3, 4)])

testcrop <- merge(sand_data, SOCcrop_data, by = c("celliso", "native"))[-c(3, 4)] 
testcrop$Value <- as.factor(testcrop$Value)

p <- ggplot(testcrop, aes(x=Value, y=soc)) + geom_boxplot() + coord_flip()

  ### RAW DATA

  points(sand[,,"soc"][sand[,,"native"] == 0],SOCStock_crop_raw[getCells(sand),year,][sand[,,"native"] == 0], col ="red")
  lines(c(-1000, 1000), c(-1000, 1000), col = "grey")

  ### MEAN DATA

  plot(sand_reduced[,,"soc_nat"][                        !is.na(sand_reduced[,,"soc_nat"])],
       SOCStock_natveg_raw[getCells(sand_reduced),year,][!is.na(sand_reduced[,,"soc_nat"])], 
       xlab = "Sandermann", ylab = "This Study",
       xlim = c(0, unname(lim)), ylim = c(0, unname(lim)))

  points(sand_reduced[,,"soc_crop"][                     !is.na(sand_reduced[,,"soc_crop"])],
         SOCStock_crop_raw[getCells(sand_reduced),year,][!is.na(sand_reduced[,,"soc_crop"])], 
         col ="red")
  lines(c(-1000, 1000), c(-1000, 1000), col = "grey")

  for (u in unique) {

    lines(c(sand_reduced[u,,"soc_crop"] - sand_reduced[u,,"sd_crop"], 
            sand_reduced[u,,"soc_crop"] + sand_reduced[u,,"sd_crop"]), 
          c(SOCStock_crop_raw[u,year,],
            SOCStock_crop_raw[u,year,]), col = "pink")

    lines(c(sand_reduced[u,,"soc_nat"] - sand_reduced[u,,"sd_nat"], 
            sand_reduced[u,,"soc_nat"] + sand_reduced[u,,"sd_nat"]), 
          c(SOCStock_natveg_raw[u,year,],
            SOCStock_natveg_raw[u,year,]), col = "grey")

  }

```

```{r sandermann3}
  x <- c(sand[, , "soc"][sand[, , "native"] == 0],
         sand[, , "soc"][sand[, , "native"] == 1])

  y <- c(SOCStock_crop_raw[    getCells(sand), year, ][sand[, , "native"] == 0], 
         SOCStock_natveg_raw[getCells(sand), year, ][sand[, , "native"] == 1])#

  x_clean <- c(sand[, , "soc"][sand[, , "native"] == 0],
               sand[, , "soc"]["CHN.9964",,,invert=TRUE][sand[, , "native"] == 1])

  y_clean <- c(SOCStock_crop_raw[  getCells(sand), year, ][sand[, , "native"] == 0], 
               SOCStock_natveg_raw[getCells(sand), year, ]["CHN.9964",,,invert=TRUE][sand[, , "native"] == 1])

  x_crop <- c(sand[, , "soc"][sand[, , "native"] == 0])

  y_crop <- c(SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0])

  x_red  <- c(sand_reduced[, , "soc_nat" ][!is.na(sand_reduced[, , "soc_nat" ])],
              sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])])

  y_red  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_nat" ])],
              SOCStock_crop_raw[  getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])

  x_red2  <- c(sand_reduced[, , "soc_nat" ]["CHN.9964",,,invert=TRUE][!is.na(sand_reduced[, , "soc_nat" ])],
              sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])])

  y_red2  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ]["CHN.9964",,,invert=TRUE][!is.na(sand_reduced[, , "soc_nat" ])],
              SOCStock_crop_raw[  getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])

  cor(x, y)^2
  cor.test(x, y)$estimate^2
  cor.test(x_clean, y_clean)$estimate^2
  cor.test(x_crop, y_crop)$estimate^2
  cor.test(x_red, y_red)$estimate^2
  cor.test(log(x_red2), log(y_red2))$estimate^2

```

### 2022-06-14 ResultNotebooks/results_SOCscenario.Rmd

@@
 
-rev         <- "rev26"
-scenarios   <- c("rev26histManagement", "rev26constManagement", "rev26constManagement2", "rev26constManagement_constClimate", "rev26constManagement2_constClimate", "rev26constManure", "rev26constResidues", "rev26constResrate", "rev26constTillage", "rev26constYield", "rev26Initial-1901_constManagement", "rev26Initial-1901_constManagement2", "rev26Initial-1901", "rev26Initial-0_constManagement", "rev26Initial-0_constManagement2", "rev26Initial-0")
+rev         <- "rev30"
+scenarios   <- c("histManagement", "constClimate-1975", "constManagement-1975", "constManagement2-1975", "constManagement-1975_constClimate-1975", "constManagement2-1975_constClimate-1975", "constManure-1975", "constResidues-1975", "constResrate-1975", "constTillage-mixedtill", "constYield-1975")
 
-# filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
-# setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
-# 
+ filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
+# setConfig(outputfolder = "/home/kristine/WORK/Paper/historicalsocbudget/ResultNotebooks/Data/")
+
 #  untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
 #        exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")
-
+# 
 # for(item in scenarios){
 #     untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-#    files=filesNeeded, exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
+#     exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
 #  }
@@
 for(scen in scenarios){
-  SoilCarbon <- read.magpie(paste0("Data/", scen, "/SoilCarbon.rds"))
-  Landuse    <- read.magpie(paste0("Data/", scen, "/Landuse.rds"))
+  print(scen)
+  SoilCarbon <- read.magpie(paste0("Data/", paste0(rev,scen), "/SoilCarbon.rds"))
+  Landuse    <- read.magpie(paste0("Data/", paste0(rev,scen), "/Landuse.rds"))
   SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
@@
 
-CarbonInput  <- read.magpie(paste0("Data/rev26histManagement/CarbonInput.rds"))
-Landuse      <- read.magpie(paste0("Data/rev26histManagement/Landuse.rds"))
+CarbonInput  <- read.magpie(paste0("Data/", rev, "histManagement/CarbonInput.rds"))
+Landuse      <- read.magpie(paste0("Data/", rev, "histManagement/Landuse.rds"))
 glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))
@@
 
-manage_scen <-  c(histManagement   = "rev26histManagement",
-                  constTillage     = "rev26constTillage", 
-                  constManure      = "rev26constManure",
-                  constResidues    = "rev26constResidues", 
-                  constManagement  = "rev26constManagement")#,
+manage_scen <-  c(histManagement   = "histManagement",
+                  constTillage     = "constTillage-mixedtill", 
+                  constManure      = "constManure-1975",
+                  constResidues    = "constResidues-1975", 
+                  constManagement  = "constManagement-1975")#,
                   #constManagementNoCC = "rev26constManagement_constClimate")
@@
 
-limits <- c(-40,-32)
-breaks <- c(-32, -34,- 36, -38, -40)
+limits <- c(-42,-34)
+breaks <- c(-34, -36,- 38, -40, -42)
 

### 2022-06-14 ResultNotebooks/results_SOCtable.Rmd

@@
 
-scenarios <- c("rev26histManagement") #,"rev18histManagement_ResAllom_2021_02_01")
+scenarios <- c("rev30histManagement") #,"rev18histManagement_ResAllom_2021_02_01")
 outputdir <- "Output/"
@@
 years             <- findset("past")
-climateIPCCr      <- readRDS("Data/rev26histManagement/IPCC_reduced.rds")
+climateIPCCr      <- readRDS("Data/rev30histManagement/IPCC_reduced.rds")
 
@@
 
-CShare_ipcczone   <- dimSums(climateIPCCr * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
-                     dimSums(climateIPCCr * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)
+# CShare_ipcczone   <- dimSums(climateIPCCr * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
+#                     dimSums(climateIPCCr * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)
 
-# CShare_ipcczone   <- dimSums(climateIPCCr * toolConditionalReplace(SOCDens[,,"crop"][,,"actualstate"]/SOCDens[,,"natveg"][,,"naturalstate"],
-#                                                                  c("is.na()", "is.nan()", "is.infinite()",">3"), c(1,1,1,5)) * Landuse[,years,"crop"], dim=1)/
-#                     dimSums(climateIPCCr * Landuse[,years,"crop"], dim=1)
+CShare_ipcczone   <- dimSums(climateIPCCr * toolConditionalReplace(SOCDens[,,"crop"][,,"actualstate"]/SOCDens[,,"natveg"][,,"naturalstate"],
+                                                                  c("is.na()", "is.nan()", "is.infinite()",">3"), c(1,1,1,5)) * Landuse[,years,"crop"], dim=1) /
+                     dimSums(climateIPCCr * Landuse[,years,"crop"], dim=1)
 
 getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))
+getCells(CShare_ipcczone) <- "GLO"
 
@@
 
-climateIPCCr       <- readRDS("Data/rev26histManagement/IPCC_reduced.rds")
-climateIPCC        <- readRDS("Data/rev26histManagement/IPCC.rds")
+climateIPCCr       <- readRDS("Data/rev30histManagement/IPCC_reduced.rds")
+climateIPCC        <- readRDS("Data/rev30histManagement/IPCC.rds")
 
@@
 
-MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
+MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL, outer(c("IPCC2006","IPCC2019"),
                                                                           getNames(climateIPCCr), paste, sep="."))
@@
 comparison <- mbind(CShare_ipcc, 
-                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),NULL),
-                                  nm="This Study.hist.1990",add="model.input.year"),2),
+                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1975","naturalstate"]),NULL),
+                                  nm="This Study.hist.1975",add="model.input.year"),2),
                     round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),NULL),
@@
 
-save_kable(save, file="Output/Images/TableSCF_comparison.png")
+save_kable(save, file="Output/Images/TableSCF_comparison_new.png")
 table

### 2022-06-14 SOCbudgetPaper/introduction.Rmd

--- 2020-12-17 SOCbudgetPaper/introduction.Rmd
@@
 
-Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth’s soil, is the largest terrestrial organic carbon pool.
+Soil organic carbon (SOC), the amount of organic carbon stored in the Earth’s soil, constitutes the largest terrestrial organic carbon pool.
 It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996].
-Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@friedlingstein_global_2019; @minasny_soil_2017].
+Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@frielingstein_glob_carbon_2020; @minasny_soil_2017].
 The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile [@batjes_total_1996; @sanderman_soil_2017].
@@
 However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis.
-Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.
+Compared to climatic and geological driving forces, human interventions alter terrestrial carbon pools over much shorter time scales and are currently one of the most dominant drivers of SOC changes on managed land [@hansis_relevance_2015; @bastos_comparison_2021].
 
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state.
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al.
+[-@pongratz_terminolgy_2014]), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state.
 Sanderman et al.
@@
 
-Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@friedlingstein_global_2019].
-While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management.
-Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015].
-BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management [@friedlingstein_global_2019; @houghton_carbon_2012; @hansis_relevance_2015].
+Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs), Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@frielingstein_glob_carbon_2020].
+While providing estimates of the magnitude of SOC losses due to land-cover change, most models lack a detailed representation of agricultural management.
+Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015].
+BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO~2~ fertilization and N deposition.
+Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management [@frielingstein_glob_carbon_2020; @houghton_carbon_2012; @hansis_relevance_2015; @bastos_comparison_2021].
 
@@
 Pugh et al.
-[-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data.
-They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.
+[-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data.
+They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.
 
 In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.
-growing seasons [@portmann_mirca2000global_2010] multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g.
-@pugh_simulated_2015, @lutz_simulating_2019].
+growing seasons [@portmann_mirca2000global_2010], multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g.
+@pugh_simulated_2015; @lutz_simulating_2019].
+Herzfeld et al.
+[-@herzfeld_soc_2021] account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC.
+Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.
 
-More data sets on spatially explicit agricultural management time series with global coverage become available (e.g.
-on tillage systems, see [@porwollik_generating_2018], [@prestele_spatially_2018]) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g.
-[@iizumi_modeling_2019], [@minoli_modelling_2019]), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g.
+on tillage systems, see [@porwollik_generating_2019; @prestele_spatially_2018]) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g.
+[@iizumi_modeling_2019; @minoli_modelling_2019]), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
 
-Field-scale models [@del_grosso_simulated_2001, @coleman_simulating_1997, @smith_estimating_2010, @taghizadeh-toosi_c-tool_2014] are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
-However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
+Field-scale models [@del_grosso_simulated_2001; @coleman_simulating_1997; @smith_estimating_2010; @taghizadeh-toosi_c-tool_2014] are able to better account for historical agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites.
+However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge [@morais_detailed_2019].
 
-The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management.
-To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices.
-We simulate SOC stocks, dynamics and the SOC debt for 1975--2010.
-Using a scenario analysis, we decompose the contribution of different management activities.
+Managed soils have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (soil carbon sequestration (SCS) techniques).
+However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) [@popp_land-use_2016; @rogelj_scenarios_2018; @forster_sr1p5_2018] have not integrated soil management into their mitigation pathways.
+More detailed process-based models are typically computationally too demanding to be integrated into optimization-based IAMs.
+Better accounting for soil carbon management in IAMs thus requires a light-weight model suitable for iterative modeling with detailed options to represent agricultural soil management.
+
+The objectives of our study are (1) to develop a reduced-complexity SOC model able to account for SCS in IAM frameworks; (2) to create a comprehensive data set of the global gridded management time series, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices; and (3) to provide global as well as spatially explicit SOC and SOC debt estimates that consider spatially explicit and time-variant agricultural management.
+We decompose the contribution of different management activities through a scenario analysis, identifying the most impacting management decisions for SOC development.
+Moreover, we compare our model performance against other SOC stock and SOC emission estimates, to evaluate the suitability of this reduced-complexity approach for integration into IAM modeling.

### 2022-07-10 SOCbudgetPaper/SOCbudget.Rmd
update

@@
 abstract: |
-  Soil organic carbon (SOC) is one of the largest terrestrial carbon (C) stocks on Earth.
-The first meter of the Earth's soils profile stores three times as much C as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land-cover change and agricultural management.
+  Soil organic carbon (SOC), one of the largest terrestrial carbon (C) stocks on Earth, has been depleted by anthropogenic land-cover change and agricultural management.
 However, the latter has so far not been well represented in global C stock assessments.
@@
 
+    ```{=latex}
+  \begin{figure}[H]
+    \includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
+    \caption{Correlation between modeled and measured $SOC$ stocks.
+Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results.
+Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
+    \label{fig:SOCpoint}
+  \end{figure}
+   ```
+
+     ```{=latex}
+  \begin{figure}[H]
+    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
+  \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
+    \label{fig:SOCLPJmL4}
+  \end{figure}
+   ```
+
 output:

### 2022-07-12 SOCbudgetPaper/results.Rmd
update

@@
 
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens [-@karstens_model_2020].
-In the following, the most important results [see @karstens_result_2020 for post-processing script] are summarized.
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in the data repository from Karstens [-@karstens_model_2020].
+In the following, the most important results [see @karstens_result_2020 for post-processing scripts] are summarized.
 
@@
 This corresponds to an average loss rate of $0.14\,\unit{GtC\,yr^{-1}}$ in comparison to a hypothetical potential natural vegetation (PNV) state.
-Considering our estimate of the global SOC stock of around $705\,\unit{GtC}$ in the upper $30\,\unit{cm}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+Considering our estimate of the global SOC stock of around $705\,\unit{GtC}$ in the upper $30\,\unit{cm}$ in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975 and 2010 in comparison to the PNV state.
 The speed of this SOC loss has decreased towards the end of the modeling period.
+Note that the SOC stock itself --- without comparing it to a a PNV state --- increases during the period between 1975 and 2010 from $705\,\unit{GtC}$ to $712\,\unit{GtC}$, which corresponds to overall SOC stock increase of $0.2\,\unit{GtC\,yr^{-1}}$.
 
@@
 Both estimates are based on measurement data for cropland (see Table \@ref(tab:SCFglo)).
-To allow for comparison, we aggregate our stock change factors weighted by grid-level $F^{\mathrm{SCF}}$ cropland area to derive average factors for the four IPCC climate zones (Fig.
+To allow for comparison, we aggregate our stock change factors weighted by grid-level cropland area to derive average factors for the four IPCC climate zones (Fig.
 \@ref(fig:CLIMzone)).
+Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking the time scale of the stock change.
 
@@
 Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC.
-For the tropical regions the IPCC factors changed notably from the guidelines in 2006 [@lasco_cropland_in_ipcc_2006] to the update in 2019 [@ogle_cropland_in_ipcc_2019].
+For the tropical regions the IPCC factors increased notably from the guidelines in 2006 [@lasco_cropland_in_ipcc_2006] to the update in 2019 [@ogle_cropland_in_ipcc_2019] due to the inclusion of more data points.
 Our results are in good agreement with the 2006 IPCC factors.
@@
 The global estimates of the total SOC stock of the upper 30 cm from this study are in the middle of the wide range of other modeled or observation-based estimates.
-SoilGrids [@hengl_soilgrids250m_2017] especially stands out with its high estimate, whereas SoilGrids 2.0 [@poggio_soilgrids_2021] marks the lower end.
 Regional results (Fig.
 \@ref(fig:SOCreg)) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.
+SoilGrids [@hengl_soilgrids250m_2017] especially stands out with its high estimate, since they include the litter horizon on top of the soil, that might dominate especially polar and boreal soils.
+SoilGrids 2.0 [@poggio_soilgrids_2021] however, excludes litter C and thus marks the lower end for particularly northern regions.
+For the same reason it is also more comparable to our results, which do not account for litter C as well.
 
@@
 
-We correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements (for data base see appendix of [@sanderman_soil_2017]).
-
-```{r captionSOCpoint, echo=FALSE}
-captionSOCpoint <- "Correlation between modeled and measured $SOC$ stocks.
-Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results.
-Both cropland and areas with natural vegetation tend to be lower in our results than in the point measurements."
-```
-
-```{r SOCpoint, out.width = "10cm", fig.cap = captionSOCpoint, echo=FALSE, fig.pos="h"}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scatterComparePointData_2010.png")
-```
+In Fig.
+\@ref(fig:SOCpoint) we correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements [for data base see appendix of @sanderman_soil_2017].
+The goodness of the fit is very low with an $R^2$ of $0.13$.
+Individually the correlations are even lower with a $R^2$ of $0.09$ for cropland and $0.08$ for areas of natural vegetation.
+This point to the fact that differences between land-use type SOC stocks could be better matched than the spatial pattern of the rather small point measurement data base.
+Due to the low number of small-scale measurements, statistical properties of the point data variability are not derived and thus, could not be used to improve the point-to-grid-cell comparison [see @rammig_generic_2018].
 
@@
 Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time.
-We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation.
+As the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019] is not specifically parameterized for natural vegetation it is important to evaluate its suitability to produce reasonable results in that domain as well at least comparable to other modeling approaches.
+We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.\@ref(fig:SOCLPJmL4)).
 Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.
 With our focus on cropland SOC dynamics, we compare only cells with more than $1000\,\unit{ha}$ of cropland (capturing 99.9% of global cropland area).
-
-```{r captionSOCLPJmL4, echo=FALSE}
-captionSOCLPJmL4 <- "Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line."
-```
-
-```{r SOCLPJmL4, out.width = "13cm", fig.cap = captionSOCLPJmL4, echo=FALSE, fig.pos="h"}
-# All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png")
-```
 Spatial correlations of PNV SOC stock values are high (global $R2=0.81$), especially for dry climate zones (Fig.

### 2022-07-12 SOCbudgetPaper/methods.Rmd
update

@@
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific.
-We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
+We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and grasslands (referred to as natural vegetation in the following).
+Note that purpose grown annual forage crops are counted as cropland, whereas grazed pastures and mowed meadows (perennials) are assigned to natural vegetation.
+Carbon flows regarding livestock are split in the sense that only carbon originated from cropland feed sources are considered in this study.
 
@@
 Similarly, agricultural management data are held constant at the level of 1965 until 1965.
-This approach follows others studies looking on effects of land-use change and management [e.g.
-@schaphoff_lpjml4_2018-1; @herzfeld_soc_2021].
+We acknowledge that this introduces a bias as agricultural management has changed prior to 1965, but this approach follows others studies looking on effects of land-use change and management [e.g.
+@schaphoff_lpjml4_2018-1; @herzfeld_soc_2021] and is limited by data availability on harvest statistics (and other management effects).
 
@@
 Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+Note that IPCC factors are derived under the assumption that there is a linear change between steady states over 20 years.
 Estimates of $F^{\mathrm{SCF}}$ are differentiated by crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows.

### 2022-07-12 SOCbudgetPaper/discussion.Rmd
update

@@
 For most of the parameters used in our management estimates no uncertainty estimates exist.
-This is why, in our view, most of the uncertainty with respect to the impacts of SOC management is included in the management data itself, and especially in the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017].
+This is why, in our view, a large part of the uncertainty with respect to the impacts of SOC management --- next to the parametric and structural uncertainty of the soil model --- is included in the management data itself.
+This is especially the case for the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data [@faostat_faostat_2016; @weindl_livestock_2017].
 The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
@@
 This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics.
-The comparison of simulated PNV data with LPJmL4 shows the model's substantial capability in reproducing PNV SOC stocks (Fig.
+The comparison of simulated PNV data with LPJmL4 shows the model's capability in reproducing PNV SOC stocks (Fig.
 \@ref(fig:SOCLPJmL4)).
+Concurrently, the point-data comparison (see Fig.
+\@ref(fig:SOCpoint)) shows low correlation for PNV, however also for cropland sites.
+This might also point to the fact that SOC stocks can vary at field and local scale considerable and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f
 
@@
 \@ref(fig:SOCglo), the only conclusion we can draw from this is that our result is within a plausible range.
+
+Additionally, we find our $0.2\,\unit{GtC\,yr^{-1}}$ SOC stock change within the period between 1975 and 2010 for the first 30 cm of the soil profile at the upper end of estimates comparing it to estimates of Ito et al.
+[-@ito_soil_2020] of $0.18\,\pm\,0.41\,\unit{GtC\,yr^{-1}}$ within the period between 1850 and 2014 for whole soil profile.
+This might be not surprising as the CO~2~ effect is most likely stronger and land-use change effects weaker within our later and shorter modeling period compared to a mean value of the period between 1850 and 2014.
+The large uncertainty within estimates of SOC stock and its changes [@ito_soil_2020] again points to the large structural uncertainty within SOC modeling.
+
 To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland.
@@
 Natural SOC estimates only influence results when natural land is converted to cropland.
+Moreover, the temporal dynamic of the SOC debt and stock change factors is not (or only to a small degree) altered by climatic or atmospheric effects on SOC stocks, as they cancel out by taking the difference (for the SOC debt) and ratio (for the stock change factors) of cropland SOC and SOC under hypothetical PNV conditions.
 
@@
 
-In our model, erosion should however only affect the spatial pattern but not the aggregate SOC pool.
+In our model, erosion might however only affect the spatial pattern but not the aggregate SOC pool.
 As pointed out by Doetterl et al.
 [-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017].
+Concurrently, other studies have claimed erosion moves SOC into aquatic reservoirs [@zhang_simulating_2020] and thus changing total global terrestrial SOC stock.
 Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focused especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
@@
 Available URL: www.yieldgap.org (accessed on: 03/01/2022)) and high shares of residue removal and burning in lower-income countries [@smil_crop_1999; @williams_influence_1997; @jain_emission_2014].
-Yet, even when comparing our etimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values, while it shows very good agreement with the older default values from IPCC [-@lasco_cropland_in_ipcc_2006].
-Don et al.
+Yet, even when comparing our estimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values [-@ogle_cropland_in_ipcc_2019], while it shows good agreement with the older default values from IPCC [-@lasco_cropland_in_ipcc_2006].
+However, the revised estimates of the IPCC included much more and more recent data points, calling for a closer look on causes for the large deviations between our results and the refined Tier 1 factors.
+On the one hand, our approach does not account for unharvested carbon inputs from weeds or biomass cover on short-term fallows.
+Shifting agriculture with fallow periods might be prominent in tropical regions.
+While long-term fallow land (>4 years) is excluded from FAOSTAT as cropland, short-term fallow is not.
+Thus, our carbon inputs for these areas might be underestimated, leading to too low stock changes factors.
+On the other hand, older studies by Don et al.
 [-@don_impact_2011] estimated SOC losses for tropical soils of around 25% on average corresponding to a stock change factor of 0.75, but also reported a wide range of measured SOC changes from -80% to +58%.
@@
 \@ref(fig:SOCscen)).
+In line with this, Dangal et al.
+[-@dangal_improving_2022] finds that no-tillage has only minor impacts on SOC dynamics across parts of the US.
 

### 2022-07-20 SOCbudgetPaper/SOCbudget.Rmd
updated manuscript

@@
 However, accounting for historical management led to $2.1\,\unit{GtC}$ less ($0.06\,\unit{GtC\,yr^{-1}}$) emissions than under the assumption of constant management.
-We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique.
-The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling.
+We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that SOC enhancement by biomass retention may be a promising negative emissions technique.
+The reduced-complexity SOC model may allow to simulate management-induced SOC enhancement also within computationally demanding integrated (land-use) assessment modeling.
 bibliography: SOCbudget.bib
@@
   \begin{figure}[H]
-    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
+    \includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
+    \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510).
+On the other hand, assuming all SOC to be in land-use specific steady-state already in 1901 (Initial-eq) leads to the opposite effect of an already higher $\Delta SOC$ in 1975 and a less steeper increase till 2010.
+However, the difference between the dynamic historical management assumption (histManagement) compared to constant management assumption from 1975 on (constManagement) are of a similar size for all initialization choices.}
+    \label{fig:SOCinit}
+  \end{figure}
+  ```
+
+    ```{=latex}
+  \begin{figure}[H]
+    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
     \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.

### 2022-07-20 SOCbudgetPaper/discussion.Rmd
updated manuscript

@@
 Our spatially explicit time series data set of agricultural management is based on country-specific FAO production and cropland statistics [@faostat_faostat_2016] as well as 0.5 degree land-use data from LUH2 [@hurtt_harmonization_2020].
-Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the agricultural system (\@ref(fig:FlowFig)) using a mass balance approach from the IPCC guidelines Vol.
+Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the cropping system (\@ref(fig:FlowFig)) using a mass balance approach from the IPCC guidelines Vol.
 4 [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and other auxiliary data sets [e.g.
@@
 
+The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes effect these values strongly (see Fig.
+\@ref(fig:SOCinit)).
+According to our sensitivity analysis the SOC debt depending on the initialization choice has been between $33.3\,\unit{GtC}$ and $50.7\,\unit{GtC}$, with our best guess at $39.6\,\unit{GtC}$.
+Concurrently, our results indicate, that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
+
 Comparing the geographic SOC patterns to Soil Grids 2.0 [@poggio_soilgrids_2021] (see Fig.
@@
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Second, changes to the subsoil due to tillage are neglected.
+Second, changes to the subsoil are neglected, which is most important for tillage effects.
+Other management practices might be more equally effecting top- and subsoil as they do not directly the relocates carbon vertically.
 As Powlson et al.
@@
 SOC transfers to deeper soil layers under tillage might enhance subsoil SOC compared to no-till practices.
+However, the effect of no-till to the subsoil as well as its overall importance as a SCS measure is still debated [@ogle_climate_2019].
 Finally, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
@@
 Yet our study also shows the substantial impact of changing management on the development of SOC debt (Fig.
+\@ref(fig:SOCdebt) and Fig.
 \@ref(fig:SOCscen)).
@@
 [-@pugh_simulated_2015] considered management effects like tillage and incorporation of residues in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
-Their study moreover concludes that yield gains (by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change).
-Historical yield increases, however, are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].
-While we find substantially larger effects of productivity gains than the 1% reported by Pugh et al.
-[-@pugh_simulated_2015], this is not sufficient to compensate SOC losses from moderate global cropland expansion of around 11% between 1974 and 2010.
+Their study moreover concludes that yield gains (by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% SOC increase).
+Historical yield increases, however, are estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009] and often lead to an increase in below- and above-ground residue biomass inputs to the soil.
+While we find substantially larger SOC increase in response to productivity gains than the 1% reported by Pugh et al.
+[-@pugh_simulated_2015], this is not sufficient to compensate SOC losses from global cropland expansion of around 11% between 1974 and 2010.
 
@@
 As shown in Fig.
-\@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-However, even larger losses occur at the end of the food supply chain (1.2 GtC year), at the soil surface (1.4 GtC), during residue burning (0.3 Gt C) and with manure management (0.2).
+\@ref(fig:FlowFig), the annual SOC respiration (1.3 GtC per year) is slightly above one quarter of total annual net C uptake by crops (4.6 GtC per year).
+C compounds have to be respired by soil organisms to maintain basic soil functions and regulate the nutrient cycle, which often leaves limited options to decrease C losses via SOC respiration [@janzen_soil_2006].
+However, similar C losses occur at the end of the food supply chain (1.2 GtC per year), at the soil surface (1.5 GtC per year), and smaller but still considerable during residue burning (0.2 GtC per year) and within animal waste management systems (0.2 GtC per year).
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.

### 2022-07-20 SOCbudgetPaper/introduction.Rmd

@@
 
-Soil organic carbon (SOC), the amount of organic carbon stored in the Earth’s soil, constitutes the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996].
+Soil organic carbon (SOC), the amount of organic carbon stored in the Earth’s soil, exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996].
 Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@frielingstein_glob_carbon_2020; @minasny_soil_2017].
@@
 Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015].
-BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO~2~ fertilization and N deposition.
-Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management [@frielingstein_glob_carbon_2020; @houghton_carbon_2012; @hansis_relevance_2015; @bastos_comparison_2021].
+BKMs are designed to estimate LUC related emissions and have largely improved in estimating additional emissions from wood harvest and shifting cultivation.
+However, state of the art models do not consider impacts of varying agricultural management [@frielingstein_glob_carbon_2020; @houghton_carbon_2012; @hansis_relevance_2015; @bastos_comparison_2021].
 
@@
 Pugh et al.
-[-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data.
+[-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with uniform scenario assumptions on management rather than with historical management data.
 They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.
 
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.
-growing seasons [@portmann_mirca2000global_2010], multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g.
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time [e.g.
+for growing seasons in @portmann_mirca2000global_2010; multiple cropping systems in @waha_multiple_2020; irrigation systems in @jagermeyr_water_2015] or as stylized (in the sense of uniform management assumptions) scenarios [e.g.
 @pugh_simulated_2015; @lutz_simulating_2019].
@@
 [-@herzfeld_soc_2021] account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC.
-Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.
+Within their stylized future management scenarios under future climate change they find that none of the management aspects considered (residue management, no-tillage) can create a positive SOC stock change on current cropland areas that counteracts the still continuing legacy flux from the initial land-use change.
 

### 2022-07-20 SOCbudgetPaper/methods.Rmd

@@
 We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and grasslands (referred to as natural vegetation in the following).
-Note that purpose grown annual forage crops are counted as cropland, whereas grazed pastures and mowed meadows (perennials) are assigned to natural vegetation.
-Carbon flows regarding livestock are split in the sense that only carbon originated from cropland feed sources are considered in this study.
+Forage crops are included within cropland, whereas pastures (including mowed meadows (perennials) and rangelands) are assigned to natural vegetation.
+Carbon flows connected to livestock are only considered in this study when they originate from cropland feed sources, while the manure originating from pasture biomass is disregarded, implicitly assuming that this manure is excreted or applied to pastures.
 
@@
 
-## Management scenario definitions {#sec:scenarios}
+## Scenario definitions {#sec:scenarios}
+
+To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization choices, we perform a set of scenario runs.
+In the following section we outline name and idea of these scenarios [for technical implementation see @karstens_mrsoil_2020].
 
@@
 The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
+
+As outlined in Sect.
+\@ref(sec:initsoc) we assume to start in a steady-state SOC stocks for the start year of the spin-up in 1510 followed by a long spin-up period of (\textit{Initial-spinup1510}).
+As some SOC compartments decay over very long timescales, the initialization setting might strongly effect the overall outcome of SOC stocks and changes.
+Thus we conduct two counterfactual scenarios \textit{Initial-natveg} and \textit{Initial-eq}.
+Whereas in \textit{Initial-natveg} we assume SOC stocks to be in a steady-state SOC under potential natural vegetation for all land-use types in 1901, SOC stocks start in their land-use type specific steady state in 1901 for the \textit{Initial-eq} scenario.
+In that way, the two scenarios mark the two extreme cases: In \textit{Initial-eq} all legacy fluxes already appeared in 1901, whereas in \textit{Initial-natveg} all legacy fluxes before 1901 still have to appear.
+We additionally combined the counterfactual scenarios with the \textit{constManagement} scenario.

### 2022-07-20 SOCbudgetPaper/results.Rmd

@@
 
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in the data repository from Karstens [-@karstens_model_2020].
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data from manure and residues as well as SOC stock results for all scenario runs can be found in the data repository from Karstens [-@karstens_model_2020].
 In the following, the most important results [see @karstens_result_2020 for post-processing scripts] are summarized.
@@
 Panel (a) shows global $\\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types.
-Computing the difference between the $\\Delta SOC$ estimate for 2010 and for 1975 (b) depicts areas of soil depletion (SOC debt increase, red) and net-sequestration (SOC debt decline, blue)."
+Computing the difference between the $\\Delta SOC$ estimate for 2010 and for 1975 (b1) depicts areas of SOC depletion (SOC debt increase, red) and SOC accumulation (SOC debt decline, blue).
+Panel (b2) shows the of land-use change (LUC) induced change of ∆SOC between 1975 and 2010, whereas panel (b3) depicts the change due to changing agricultural management (MAN)."
 ```
@@
 # All defaults
-knitr::include_graphics("../ResultNotebooks/Output/Images/2panelfigure.png")
+knitr::include_graphics("../ResultNotebooks/Output/Images/SpatialDebtFigure.png")
 ```
@@
 \@ref(fig:SOCdebt)(a))  and its change from 1975 to 2010 (Fig.
-\@ref(fig:SOCdebt)(b)) reveals areas of SOC debt decline and increase.
+\@ref(fig:SOCdebt)(b1)) reveals areas of SOC debt decline and increase.
 Regions with large cropland expansion (e.g.
@@
 
-## Carbon flows in the agricultural system
+## Carbon flows in the cropland system
 
 ```{r caption_FlowFig, echo=FALSE}
-caption_FlowFig <- "Global carbon flows within the agricultural system for the year 2010 (in $\\unit{MtC}$).
+caption_FlowFig <- "Global carbon flows within the cropland system for the year 2010 (in $\\unit{MtC}$).
 Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues.
+After accounting for losses within the cropland system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues.
 Large parts of C, however, are mineralized on the field before entering the soil.
-Additionally, C is transferred to and from the global agricultural soil stock via land-use change between cropland and natural vegetation.
+Additionally, C is transferred to and from global cropland soil stock via land-use change between cropland and natural vegetation.
 Finally, SOC is mineralized and flows back to the atmosphere."
@@
 
-C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvested organ, above- and below-ground residues).
+C is removed from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvested organ, above- and below-ground residues).
 Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues ($729\,\unit{MtC}$ in 2010) are directly returned to the field.
@@
 According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of $257\,\unit{MtC}$ in 2010.
-The agricultural system receives $4585\,\unit{MtC}$ assimilated by crop plants and releases $3554\,\unit{MtC}$ mostly through respiration.
+The cropland system receives $4585\,\unit{MtC}$ assimilated by crop plants and releases $3554\,\unit{MtC}$ mostly through respiration.
 Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around $33\,\unit{MtC}$ for the year 2010.
@@
 The large contribution of residues relative to manure also becomes visible when considering the annual C inputs of residues and manure to soils over a period of 35 years (Fig. \@ref(fig:SOCscen)(b)).
+
+Using the constManagement results, that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results.
+The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state while calculating $\Delta SOC$.
+The increasing SOC debt on global cropland are primarily caused by LUC (red areas in Fig.
+\@ref(fig:SOCdebt)(b2)).
+Deteriorated management also contributed to increasing SOC debt in parts of Sub-saharan Africa and Central Asia.
+In contrast, agricultural management has led to an decrease in SOC debt in the USA, Europe, as well as in parts of China and India (blue areas in Fig.
+\@ref(fig:SOCdebt)(b3)), which is not visible in the total $\Delta SOC$ change as LUC was happening at the same time.
+
+Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks (see Fig.
+\@ref(fig:SOCinit) and Sect.
+\@ref(sec:scenarios)) with around $2.15\,\unit{GtC}$ difference in SOC debt between the histManagement and the constManagement scenario.
+However, the SOC debt and SOC debt change varies with the different initialization choices.
+Whereas the default assumption (Initial-spinup1510) shows a $\Delta SOC$ of $39.6\,\unit{GtC}$ for the year 2010, the Initial-natveg scenario with high legacy fluxes to come only has a $\Delta SOC$ of $33.3\,\unit{GtC}$ and the Initial-eq scenarios with only little legacy fluxes left already a $\Delta SOC$ of $50.7\,\unit{GtC}$.
 

### 2022-07-20 ResultNotebooks/create_ZenodoOutput.Rmd
update of output scripts and figure

---
  title: "SOC budget aóutput creator"
output:
  html_document:
  df_print: paged
pdf_document: default
bookdown::pdf_document2: default
keep_md: true
---

Load all folder and create NetCDF files:

```{r load files}

rev        <- "Data/rev30"
scenarios   <- c("histManagement", "constManagement-1975", "constManure-1975", "constResidues-1975", "constTillage-mixedtill")

for (s in scenarios[1]) {

   path  <- paste0(rev, s)
   files <- list.files(path, pattern = "*rds")
   # dir.create(s)

   for (f in files) {

     tmp <- readRDS(paste0(path, "/", f))
     f   <- gsub(".rds", "", f)
     items <- getItems(tmp)

     if (length(unlist(items[1])) == 59199) output <- ".nc" else output <- ".csv"
     print(paste(s, f))
     print(unlist(items[2])[c(1, length(unlist(items[2])))])
     print(unname(unlist(items[3])))

     # write.magpie(tmp, file_name = paste0(f, output), file_folder = paste0(outputdir, s))
   }

   # tar(paste0(s, ".tgz"), files = s, compression = "gzip")

}
```

### 2022-07-20 ResultNotebooks/results_SOCattribution.Rmd
update of output scripts and figure

---
  title: "Global SOC scenarios"
output:
  md_document:
  variant: markdown_github
html_document:  
  pdf_document: default
---

```{r load}
library(mrcommons)
library(mrvalidation)
library(mrsoil)
library(magpiesets)

rev         <- "rev30"
scenarios   <- c("histManagement", "constManagement-1975")
filesNeeded <- c("SoilCarbon.rds", "Landuse.rds")
outputdir   <- "Output/"
```

```{r Load data and calc globals}

SOCscen <- NULL

for (scen in scenarios) {
  print(scen)
  SoilCarbon <- read.magpie(
                  paste0("Data/", paste0(rev, scen), "/SoilCarbon.rds"))
  Landuse    <- read.magpie(
                  paste0("Data/", paste0(rev, scen), "/Landuse.rds"))
  SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
  SOCStock   <- toolConditionalReplace(dimSums(dimSums(
    SoilCarbon[, , c("actualstate", "naturalstate")], dim = 3.3) *
      Landuse, dim = 3.2), "is.nan()", 0)
  SOCscen    <- mbind(SOCscen, 
                      add_dimension(SOCStock,
                                    dim = 3.1, nm = scen, add = "scenario"))
}

SOCDebt       <- collapseNames(SOCscen[, , "actualstate"] -
                                 SOCscen[, , "naturalstate"])
SOCDebtChange <- collapseDim(SOCDebt[, c(1975:2010), ] -
                               SOCDebt[, "y1975", ])

landMask   <- collapseDim(Landuse[, 1, 1])
landMask[] <- NA
for (y in c(1975:2010)) landMask[Landuse[, y, "crop"] > 0.001] <- 1

LUCattr <- toolConditionalReplace(
             SOCDebtChange[, "y2010", "constManagement-1975"],
             "is.na()", 0) * landMask

MANattr <- toolConditionalReplace(
             (SOCDebtChange[, "y2010", "histManagement"] -
              SOCDebtChange[, "y2010", "constManagement-1975"]),
             "is.na()", 0) * landMask

saveRDS(
  mbind(
    add_dimension(collapseNames(SOCDebt[, "y2010", "histManagement"]) *
                    landMask, dim = 3.1, add = "type", nm = "SOCDebt"),
    add_dimension(collapseNames(SOCDebtChange[, "y2010", "histManagement"]) *
                    landMask, dim = 3.1, add = "type", nm = "SOCDebtChange"),
    add_dimension(collapseNames(LUCattr),
                  dim = 3.1, add = "type", nm = "LUCdiff"),
    add_dimension(collapseNames(MANattr),
                  dim = 3.1, add = "type", nm = "MANdiff")),
  paste0(outputdir, "SOCDebtSpatial.rds"))

# LUCattr <- toolConditionalReplace(
#              SOCDebtChange[,"y2010", "constManagement-1975"] /
#              SOCDebtChange[,"y2010","histManagement"],
#              "is.na()", 0) * landMask
#
# MANattr <- toolConditionalReplace(
#              (SOCDebtChange[,"y2010","histManagement"] -
#               SOCDebtChange[,"y2010", "constManagement-1975"]) /
#               SOCDebtChange[,"y2010","histManagement"],
#              "is.na()", 0) * landMask
#
# signLUC   <- landMask
# signLUC[] <- 1
# signLUC[SOCDebtChange[,"y2010","constManagement-1975"] < 0] <- (-1)
#
# signMAN   <- landMask
# signMAN[] <- 1
# signMAN[(SOCDebtChange[,"y2010","histManagement"] -
#            SOCDebtChange[,"y2010","constManagement-1975"]) < 0] <- (-1)
#
# signLUCattr <- collapseNames(abs(LUCattr) * signLUC)
# signMANattr <- collapseNames(abs(MANattr) * signMAN)
#
# saveRDS(mbind(add_dimension(signLUCattr, dim = 3.1,
#                             add = "type", nm = "LUCattr"),
#               add_dimension(signMANattr, dim = 3.1,
#                             add = "type", nm = "MANattr")),
#         paste0(outputdir, "SOCattribution.rds"))
```

```{r grid plot, out.height=800 , out.width=1200}

library(reshape2)

SOCDebtSpatial <- readRDS(paste0(outputdir, "SOCDebtSpatial.rds"))
magpie_coord   <- toolGetMapping("magpie_coord.rda",
                                type = "cell", where = "mappingfolder")

SOCDebt2plot       <- 
  merge(as.data.frame(SOCDebtSpatial[, , "SOCDebt"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
SOCDebtChange2plot <- 
  merge(as.data.frame(SOCDebtSpatial[, , "SOCDebtChange"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
LUCattri2plot      <- 
  merge(as.data.frame(SOCDebtSpatial[, , "LUCdiff"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
MANattri2plot      <- 
  merge(as.data.frame(SOCDebtSpatial[, , "MANdiff"]),
        data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")

library(ggplot2)
theme_set(theme_bw())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(sp)
library(rworldmap)
library(oce)
library(scales)
library(cowplot)

legend_size <- 10

map.world <- map_data("world")
world         <- ne_countries(scale = "medium", returnclass = "sf")
data(coastlineWorld)
coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]],
                              latitude = coastlineWorld[["latitude"]])

SOCplotmap   <- function(data, name, cols, limits,
                         values = NULL, breaks, labels, legend_width) {

out <- ggplot(data = world) + geom_sf() +
              geom_tile(data = subset(data, Year == "2010", 
                                      c("lon", "lat", "Value")),
                        aes(x = lon, y = lat, fill = Value)) +
        scale_fill_stepsn(limits   = limits,
                      colours  = cols,
                      name     = name,
                      values   = values,
                      breaks   = breaks,
                      labels   = labels,
                      na.value = "white",
                      guide    = guide_legend(
                        title.theme    = element_text(size = legend_size),
                        direction      = "horizontal",
                        title.position = "top",
                        # I shift the labels around, the should be placed
                        # exactly at the right end of each legend key
                        title.hjust    = 0.5,
                        label.hjust    = 1,
                        nrow           = 1,
                        byrow          = FALSE,
                        label.position = "bottom",
                        keyheight      = 0.5,
                        keywidth       = legend_width
                      )) +
    theme(legend.position = "bottom",
          axis.title.x     = element_blank(),
          axis.title.y     = element_blank(),
          legend.text      = element_text(size = legend_size),
          rect             = element_blank(),
          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7, 0.7)),
          plot.margin      = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
          )

    return(out)
}

cols     <- colorRampPalette(brewer.pal(11, "RdYlBu"))(11)[c(1:6, 7, 9, 11)]
lim_debt <- 10
val_debt <- c(-10, -6, -4, -2, -1, 0, 1, 10)
debt          <- 
  SOCplotmap(data     = SOCDebt2plot,
             limits   = c(-lim_debt, lim_debt),
             cols     = cols,
             name     = expression("\u0394SOC in 2010 in million tons of carbon"),
             values   = (val_debt + lim_debt) / (2 * lim_debt),
             breaks   = val_debt[-1],
             labels   =  c("< -6", "    -4", "    -2", "    -1 ", "   0", "  1", "> 1"),
             legend_width = 2.3)

cols          <- colorRampPalette(brewer.pal(11, "RdYlBu"))(10)
lim_abs       <- 1.5
val_abs       <- seq(-lim_abs, lim_abs, length.out = 11)
debttrend_abs <-
  SOCplotmap(data     = SOCDebtChange2plot,
             limits   = c(-lim_abs, lim_abs),
             cols     = cols,
             name     = expression("\u0394SOC change from 1975 to 2010 in million tons of carbon"),
             values   = (val_abs + lim_abs) / (2 * lim_abs),
             breaks   = val_abs[-1],
             labels   = c(paste0("< ", val_abs[2]),
                          as.character(val_abs[c(-1, -2, -11)]),
                          paste0("> ", val_abs[10])),
             legend_width = 1.5)

debttrend_luc <- 
  SOCplotmap(data     = LUCattri2plot,
             limits   = c(-lim_abs, lim_abs),
             cols     = cols,
             name     = expression("\u0394SOC change from 1975 to 2010 in million tons of carbon"),
             values   = (val_abs + lim_abs) / (2 * lim_abs),
             breaks   = val_abs[-1],
             labels   = c(paste0("< ", val_abs[2]),
                          as.character(val_abs[c(-1, -2, -11)]),
                          paste0("> ", val_abs[10])),
             legend_width = 1.5)

debttrend_man <- 
  SOCplotmap(data     = MANattri2plot,
             limits   = c(-lim_abs, lim_abs),
             cols     = cols,
             name     = expression("\u0394SOC change from 1975 to 2010 in million tons of carbon"),
             values   = (val_abs + lim_abs) / (2 * lim_abs),
             breaks   = val_abs[-1],
             labels   = c(paste0("< ", val_abs[2]),
                          as.character(val_abs[c(-1, -2, -11)]),
                          paste0("> ", val_abs[10])),
             legend_width = 1.5)

debt
debttrend_abs
debttrend_luc
debttrend_man

library(gridExtra)
library(cowplot)

grid  <- plot_grid(debt,
                   debttrend_abs + guides(fill = "none"),
                   debttrend_luc + guides(fill = "none"),
                   debttrend_man, ncol = 1,
                   labels = c("(a)  - \u0394SOC ",
                              "(b1) - total \u0394SOC change",
                              "(b2) - \u0394SOC change due to LUC",
                              "(b3) - \u0394SOC change due to MAN"),
                   label_fontface = "plain", label_size = 10,
                   align = "v",  rel_heights = c(1.40, 1, 1, 1.40),
                   label_x = 0.01, hjust = 0, vjust = 1.1
                   )

grid

ggsave(paste0(outputdir, "Images/SpatialDebtFigure.png"), plot = grid,
       width = 10, height = 26, dpi = 300, units = "cm")
```

### 2022-07-20 ResultNotebooks/results_SOCcompare.Rmd
update of output scripts and figure

@@
 ---
-title: "SOC budget analysis in comparsion"
+title: "SOC budget analysis in comparsion to other global SOC estimates"
 output:
@@
 rev        <- "rev30"
-
 outputdir  <- "Output/"
@@
 Landuse    <- read.magpie(paste0("Data/", rev, "histManagement/Landuse.rds"))
-
 ```
@@
 
-SoilCarbon[,,c("naturalstate.crop")] <- SoilCarbon[,,c("naturalstate.natveg")]
-SOCStock     <- dimSums(dimSums(SoilCarbon[,,c("naturalstate","actualstate")], dim=3.3)* Landuse, dim=3.2)
+SoilCarbon[, , c("naturalstate.crop")] <- SoilCarbon[, , c("naturalstate.natveg")]
+SOCStock     <- dimSums(dimSums(SoilCarbon[, , c("naturalstate", "actualstate")],
+                                dim = 3.3) * Landuse, dim = 3.2)
 cell2reg     <- read.csv("Data/additional/ClimateZoneMapping.csv")
-reg_SOCStock <- toolAggregate(SOCStock, rel=cell2reg, from="cell", to="IPCC")
+reg_SOCStock <- toolAggregate(SOCStock, rel = cell2reg, from = "cell", to = "IPCC")
 ```
@@
 
-  WISE         <- readRDS(paste0("Data/", rev, "histManagement/WISE.rds"))
-  GSOC         <- readRDS(paste0("Data/", rev, "histManagement/GSOC.rds"))
-  SoilGrids    <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids.rds"))
-  SoilGrids2   <- readRDS(paste0("Data/", rev, "histManagement/SoilGrids2.rds"))
-  LPJmL4       <- readRDS(paste0("Data/", rev, "histManagement/LPJmL4.rds"))
-  LPJmL4_CRU   <- readRDS(paste0("Data/", rev, "histManagement/LPJmL4_CRU.rds"))
-  SOCDebtPaper <- readRDS(paste0("Data/", rev, "histManagement/SOCDebtPaperC.rds"))
+WISE         <- readRDS(paste0("Data/",
+                                 rev, "histManagement/WISE.rds"))
+  GSOC         <- readRDS(paste0("Data/",
+                                 rev, "histManagement/GSOC.rds"))
+  SoilGrids    <- readRDS(paste0("Data/",
+                                 rev, "histManagement/SoilGrids.rds"))
+  SoilGrids2   <- readRDS(paste0("Data/",
+                                 rev, "histManagement/SoilGrids2.rds"))
+  LPJmL4       <- readRDS(paste0("Data/",
+                                 rev, "histManagement/LPJmL4.rds"))
+  LPJmL4_CRU   <- readRDS(paste0("Data/",
+                                 rev, "histManagement/LPJmL4_CRU.rds"))
+  SOCDebtPaper <- readRDS(paste0("Data/",
+                                 rev, "histManagement/SOCDebtPaperC.rds"))
 
@@
 SOCStock <- reg_SOCStock
-reg_SOC  <- mbind(setNames(SOCStock[,"y2010","actualstate"],"This Study"),
-                  setNames(SOCStock[,"y2010","naturalstate"],"This Study (PNV)"),
-                  setNames(LPJmL4[,"y2010",], "LPJmL4"),
-                  setYears(setNames(WISE[,1,],"WISE"), "y2010"),
-                  setYears(setNames(SoilGrids[,1,],"SoilGrids"),"y2010"),
-                  setYears(setNames(GSOC[,1,],"GSOC"),"y2010"),
-                  setYears(setNames(SoilGrids2[,1,],"SoilGrids2"),"y2010"),
-                  setYears(setNames(SOCDebtPaper[,"y2010",],"SOCDebtPaper"),"y2010"))
-glo_SOC <- mbind(setCells(dimSums(reg_SOC, dim=1),"GLO"), reg_SOC)
+reg_SOC  <-
+  mbind(setNames(SOCStock[, "y2010", "actualstate"], "This Study"),
+        setNames(SOCStock[, "y2010", "naturalstate"], "This Study (PNV)"),
+        setNames(LPJmL4[, "y2010", ], "LPJmL4"),
+        setYears(setNames(WISE[, 1, ], "WISE"), "y2010"),
+        setYears(setNames(SoilGrids[, 1, ], "SoilGrids"), "y2010"),
+        setYears(setNames(GSOC[, 1, ], "GSOC"), "y2010"),
+        setYears(setNames(SoilGrids2[, 1, ], "SoilGrids 2.0"), "y2010"),
+        setYears(setNames(SOCDebtPaper[, "y2010", ], "SOCDebtPaper"), "y2010"))
+glo_SOC <- mbind(setCells(dimSums(reg_SOC, dim = 1), "GLO"), reg_SOC)
 
-#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+# print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
 
-comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(glo_SOC/1000)), 
-                           Data1 ~ Region, value.var="Value", 
-                            row.names=levels(as.data.frame(glo_SOC)$Data1)))
+comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(glo_SOC / 1000)),
+                           Data1 ~ Region, value.var = "Value",
+                            row.names = levels(as.data.frame(glo_SOC)$Data1)))
 
@@
 
-saveRDS(object = comparison_table, paste0(outputdir,"gloSOC_comparison.rds"))
-
+saveRDS(object = comparison_table, paste0(outputdir, "gloSOC_comparison.rds"))
 ```
@@
 
-names(table) <- c("Data1","Boreal Dry", "Boreal Moist", "C-Temp Dry", "C-Temp Moist","GLO","Polar Dry", "Polar Moist",
-                  "Tropical Dry", "Tropical Moist", "Tropical Montane", "Tropical Wet", "W-Temp Dry", "W-Temp Moist", "Year")
-table        <- table[,c(1,15,6,7,8,2,3,4,5,13,14,9,10,11,12)]
+names(table) <- c("Data1", "Boreal Dry", "Boreal Moist", "C-Temp Dry",
+                  "C-Temp Moist", "GLO", "Polar Dry", "Polar Moist",
+                  "Tropical Dry", "Tropical Moist", "Tropical Montane",
+                  "Tropical Wet", "W-Temp Dry", "W-Temp Moist", "Year")
+table        <- table[, c(1, 15, 6, 7, 8, 2, 3, 4, 5, 13, 14, 9, 10, 11, 12)]
 
-cols   <-c("black",   "#DF536B", "#61D04F", "#2297E6",
-           "#28E2E5", "#CD0BBC", "#F5C710", "gray62")[c(2, 7, 1, 3, 4, 5, 6, 8)]
+cols   <- c("black",   "#DF536B", "#61D04F",
+           "#2297E6", "#28E2E5", "#CD0BBC",
+           "#F5C710", "gray62")[c(2, 7, 1, 3, 4, 5, 6, 8)]
 
-glob <- ggplot(data = melt(subset(table, select =c(Data1, GLO, Year))), aes(y=Year,x=value, group=Data1)) + 
-  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
-  scale_size_manual( values = c( 4,  2, 2, 2, 2, 2, 2,  2)) + 
+glob <- ggplot(data = melt(subset(table, select = c(Data1, GLO, Year))),
+               aes(x = Year, y = value, group = Data1)) +
+  geom_point(aes(colour = Data1, size = Data1, shape = Data1)) +
+  scale_size_manual(values = c(4,  2, 2, 2, 2, 2, 2,  2)) +
   scale_shape_manual(values = c(19, 16, 6, 2, 5, 0, 8, 10)) +
@@
   labs(color  = leg_name, size = leg_name, shape = leg_name) +
-  ylab("") + xlab("SOC (0-30cm) in GtC") +
-  theme(axis.ticks.y =element_blank(),
-        axis.text.y=element_blank(), 
-        legend.position="bottom") + 
-  guides(fill=guide_legend(nrow=2, byrow=TRUE))
-        #legend.position="right") + guides(fill=guide_legend(ncol=1,byrow=TRUE))
+  xlab("") + ylab("SOC (0-30cm) in GtC") +
+  theme(axis.ticks.x = element_blank(),
+        axis.text.x  = element_blank(),
+        legend.position = "right")
 
-comp <- ggplot(data = melt(subset(table, select=-GLO)), aes(y=Year,x=value, group=Data1)) + 
-  geom_point(aes(colour=Data1, size=Data1, shape=Data1)) + 
-  facet_wrap(~ variable, scales="free", drop=TRUE, labeller = label_wrap_gen(width=18)) + 
-  scale_size_manual( values = c( 4,  2, 2, 2, 2, 2, 2,  2)) + 
+comp <- ggplot(data = melt(subset(table, select = -GLO)),
+               aes(x = Year, y = value, group = Data1)) +
+  geom_point(aes(colour = Data1, size = Data1, shape = Data1)) +
+  facet_wrap(~variable, scales = "free", drop = TRUE,
+             labeller = label_wrap_gen(width = 18)) +
+  scale_size_manual(values = c(4,  2, 2, 2, 2, 2, 2,  2)) +
   scale_shape_manual(values = c(19, 16, 6, 2, 5, 0, 8, 10)) +
@@
   labs(color  = leg_name, size = leg_name, shape = leg_name) +
-  ylab("") + xlab("SOC (0-30cm) in GtC") +
-  theme(axis.ticks.y=element_blank(),
-        axis.text.y=element_blank(),
-        legend.position = "bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))
+  xlab("") + ylab("SOC (0-30cm) in GtC") +
+  theme(axis.ticks.x = element_blank(),
+        axis.text.x  = element_blank(),
+        legend.position = "bottom") +
+  guides(fill = guide_legend(nrow = 3, byrow = TRUE))
 
@@
 
-ggsave(paste0(outputdir,"/Images/glo_comparisonfigure.png"), plot=glob, height=4, width=10, units="cm", dpi = 300)
-ggsave(paste0(outputdir,"/Images/reg_comparisonfigure.png"), plot=comp, height=9, width=15, units="cm", dpi = 300)
-
+ggsave(paste0(outputdir, "/Images/glo_comparisonfigure.png"),
+       plot = glob, height =  7, width =  7, units = "cm", dpi = 300)
+ggsave(paste0(outputdir, "/Images/reg_comparisonfigure.png"),
+       plot = comp, height = 15, width = 14, units = "cm", dpi = 300)
 ```
@@
 glob
-ggsave(paste0(outputdir,"/glo_comparisonfigure.png"), plot=glob)
+ggsave(paste0(outputdir, "/glo_comparisonfigure.png"), plot = glob)
 ```

### 2022-07-20 ResultNotebooks/results_SOCmaps.Rmd
update of output scripts and figure

@@
 rev         <- "rev30"
-
-filesNeeded <- c("SoilCarbon.rds","Landuse.rds")
-
-scenarios    <- c("histManagement")#,"constManagement-1975", "constTillage-mixedtill", "constResidues-1975", "constManure-1975",
-                  #"constManagement2-1975", "constResrate-1975", "constYield-1975",
-                 #"Initial-natveg", "Initial-natveg_constManangement-1975", "Initial-natveg_constManangement2-1975", 
-                 #"LitterPNV-PerennialGrasses", "LitterPNV-AASS", "LitterPNV-BRCF", 
-                 #"LitterPNV-MTCF", "LitterPNV-BLEAF",  "LitterPNV-TREF")
-
-# for(item in scenarios){
-#    dir.create(path=paste0("Data/",paste0(rev,item)))
-#    untar(paste0("Data/", paste0(rev,item), "_h12_carbonbudget.tgz"),
-#    exdir=paste0("Data/", paste0(rev,item)), extras="--keep-newer-files ") # files=filesNeeded,
-# }
-
+filesNeeded <- c("SoilCarbon.rds", "Landuse.rds")
+scenarios   <- c("histManagement")
 outputdir   <- "Output/Images"
 
-SoilCarbon <- read.magpie(paste0("Data/",rev,scenarios[1],"/SoilCarbon.rds"))
-Landuse    <- read.magpie(paste0("Data/",rev,scenarios[1],"/Landuse.rds"))
-
+SoilCarbon <- read.magpie(paste0("Data/", rev, scenarios[1], "/SoilCarbon.rds"))
+Landuse    <- read.magpie(paste0("Data/", rev, scenarios[1], "/Landuse.rds"))
 ```
@@
 
-years <- seq(1975,2010,1)
+years <- seq(1975, 2010, 1)
 
-SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
-SOCStock        <- dimSums(dimSums(SoilCarbon[,years,][,,"state",pmatch=TRUE], dim=3.3)* Landuse[,years,], dim=3.2)
-glo_SOCStock    <- dimSums(SOCStock, dim=1) 
-glo_SOCEmission <- (glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
+SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
+SOCStock        <- dimSums(
+                     dimSums(
+                       SoilCarbon[, years, ][, , "state", pmatch = TRUE],
+                       dim = 3.3) *
+                     Landuse[, years, ],
+                     dim = 3.2)
+glo_SOCStock    <- dimSums(SOCStock, dim = 1)
+glo_SOCEmission <- (glo_SOCStock[, , "naturalstate"] -
+                      glo_SOCStock[, , "actualstate"]) / 1000
 
-cropland_dens    <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)
-cropland_dens[Landuse[,,"crop"] <= 0.001] <- NA
+cropland_dens   <- dimSums(SoilCarbon[, , "actualstate"][, , "crop"], dim = 3)
+cropland_dens[Landuse[, , "crop"] <= 0.001] <- NA
 
-cropland_share <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)/(dimSums(SoilCarbon[,,"naturalstate"][,,"natveg"], dim=3) + 1e-8)
-cropland_share[Landuse[,,"crop"] <= 0.001] <- NA
+cropland_share <- dimSums(SoilCarbon[, , "actualstate"][, , "crop"],
+                          dim = 3) /
+                 (dimSums(SoilCarbon[, , "naturalstate"][, , "natveg"],
+                          dim = 3) + 1e-8)
+cropland_share[Landuse[, , "crop"] <= 0.001] <- NA
 
-cropland_increase <- dimSums(SoilCarbon[,,"actualstate"][,,"crop"], dim=3)-dimSums(SoilCarbon[,,"actualstate"][,,"natveg"], dim=3) 
-cropland_increase[Landuse[,,"crop"] <= 0.001] <- NA
+cropland_increase <- dimSums(SoilCarbon[, , "actualstate"][, , "crop"],
+                             dim = 3) -
+                     dimSums(SoilCarbon[, , "naturalstate"][, , "natveg"],
+                             dim = 3)
+cropland_increase[Landuse[, , "crop"] <= 0.001] <- NA
 
-magpie_coord  <- toolGetMapping("magpie_coord.rda", type="cell")
-share2plot    <- merge(as.data.frame(cropland_share), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-increase2plot <- merge(as.data.frame(cropland_increase), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-density2plot  <- merge(as.data.frame(cropland_dens), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-
+magpie_coord  <- toolGetMapping("magpie_coord.rda", type = "cell")
+share2plot    <- merge(as.data.frame(cropland_share),
+                       data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
+increase2plot <- merge(as.data.frame(cropland_increase),
+                       data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
+density2plot  <- merge(as.data.frame(cropland_dens),
+                       data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
 ```
@@
 data(coastlineWorld)
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]],
+                              latitude  = coastlineWorld[["latitude"]])
 
@@
     # apply transformation
-    x[isq] <- from + (x[isq] - from)/factor
-    x[ito] <- from + (to - from)/factor + (x[ito] - to)
+    x[isq] <- from + (x[isq] - from) / factor
+    x[ito] <- from + (to - from) / factor + (x[ito] - to)
     return(x)
@@
     # get indices for the relevant regions
-    isq <- x > from & x < from + (to - from)/factor
-    ito <- x >= from + (to - from)/factor
+    isq <- x > from & x < from + (to - from) / factor
+    ito <- x >= from + (to - from) / factor
     # apply transformation
     x[isq] <- from + (x[isq] - from) * factor
-    x[ito] <- to + (x[ito] - (from + (to - from)/factor))
+    x[ito] <- to + (x[ito] - (from + (to - from) / factor))
     return(x)
@@
 
-SOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCEmission)[,c(3,6)],  aes(x = Year, y = Value, group=1)) +
-  labs( x = "Year", 
-        y = expression(paste('Global \u0394SOC in GtC'))) +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(trans = squish_trans(-34,0,10),
-                     limits = c(-40,0),
+SOCemis <- ggplot() +
+  geom_line(data = as.data.frame(-glo_SOCEmission)[, c(3, 6)],
+            aes(x = Year, y = Value, group = 1)) +
+  labs(x = "Year",
+        y = expression(paste("Global \u0394SOC in GtC"))) +
+  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
+  scale_y_continuous(trans = squish_trans(-34, 0, 10),
+                     limits = c(-40, 0),
                      breaks = c(0, -34, -35, -36, -37, -38, -39, -40),
-                     minor_breaks = seq(-1,-29,-2)) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size),
-        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 10, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size)#,
-        #plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
+                     minor_breaks = seq(-1, -29, -2)) +
+  theme(rect = element_blank(),
+        axis.title.y = element_text(size = legend_size),
+        axis.title.x = element_text(size = legend_size,
+                                    margin = margin(t = 10,
+                                                    r = 0,
+                                                    b = 10,
+                                                    l = 0,
+                                                   unit = "pt")),
+        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8,
+                                   size = legend_size)
         ) +
-  annotate(geom="text", label="SOC debt", x=factor(1995), y=-20, size = 3) +
-  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -36, arrow=arrow(end="both", type="closed", length=unit(0.1, "inches")))
+  annotate(geom = "text", label = "SOC debt",
+           x = factor(1995), y = -20, size = 3) +
+  annotate("segment", x = factor(1990), xend = factor(1990), y = 0, yend = -36,
+           arrow = arrow(end    = "both",
+                         type   = "closed",
+                         length = unit(0.1, "inches")))
 
-SOCplotmap   <- function(data, name, cols, limits, values = NULL, breaks, labels) {
+SOCplotmap   <- function(data, name, cols, limits,
+                         values = NULL, breaks, labels) {
 
   out <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(data, Year == "2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = Value)) +
+    geom_tile(data = subset(data, Year == "2010", c("lon", "lat", "Value")),
+              aes(x = lon, y = lat, fill = Value)) +
     scale_fill_stepsn(limits   = limits,
@@
                       guide    = guide_legend(
-                        title.theme    = element_text(size=legend_size),
+                        title.theme    = element_text(size = legend_size),
                         direction      = "horizontal",
-                        title.position = 'top',
+                        title.position = "top",
                         # I shift the labels around, the should be placed
@@
                         nrow           = 1,
-                        byrow          = F,
+                        byrow          = FALSE,
                         label.position = "bottom",
@@
           rect             = element_blank(),
-          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7,0.7)),
+          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7, 0.7)),
           plot.margin      = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
@@
 
-cols        <- colorRampPalette(brewer.pal(11, "RdYlBu"))(11)[c(1:6,7,9,11)]
-SOCshare    <- SOCplotmap( data     = share2plot,
-                            limits   = c(0,2),
+cols        <- colorRampPalette(brewer.pal(11, "RdYlBu"))(11)[c(1:6, 7, 9, 11)]
+SOCshare    <-
+  SOCplotmap(
+    data     = share2plot,
+    limits   = c(0, 2),
     cols     = cols,
     name     = expression("cropland SOC stock change factor (unitless)"),
-                            values   = c(0,0.2,0.4,0.6,0.8,1.0,1.2,2)/2,
-                            breaks   = c(0.2,0.4,0.6,0.8,1.0,1.2,2),
-                            labels   = c("< 0.2","0.4", "0.6","0.8","1.0","1.2","> 1.2"))
+    values   = c(0, 0.2, 0.4, 0.6, 0.8, 1.0, 1.2, 2) / 2,
+    breaks   = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2, 2),
+    labels   = c("< 0.2", "0.4", "0.6", "0.8", "1.0", "1.2", "> 1.2"))
 
-SOCincrease <- SOCplotmap( data     = increase2plot,
+SOCincrease <-
+  SOCplotmap(
+    data     = increase2plot,
     limits   = c(-100, 100),
     cols     = cols,
-                            name     = expression(paste('cropland \u0394SOC in tons of carbon per hectare')),
-                            values   = (c(-100, -40, -30, -20, -10, 0, 10, 100)+100)/200,
+    name     = expression(
+                 paste("cropland \u0394SOC in tons of carbon per hectare")),
+    values   = (c(-100, -40, -30, -20, -10, 0, 10, 100) + 100) / 200,
     breaks   =  c(-40, -30, -20, -10, 0, 10, 100),
-                            labels   = c("< -40", "        -30","        -20","      -10 ","     0","    10","> 10"))
+    labels   =  c("< -40", "        -30", "        -20",
+                  "      -10 ", "     0", "    10", "> 10"))
 
-cols <- colorRampPalette(brewer.pal(9, "YlGn"))(8)[2:8]
-SOCdensity  <- SOCplotmap( data     = density2plot,
+cols <-
+  colorRampPalette(
+    c("white",
+      brewer.pal(9, "YlGn")[c(1:2)],
+      "yellow",
+      brewer.pal(9, "YlGn")[c(4:7)],
+      brewer.pal(9, "YlGnBu")[c(6:9)],
+      "black"))(21)[c(1, 5, 7, 10, 13, 16, 19) + 1]
+
+SOCdensity  <-
+  SOCplotmap(
+    data     = density2plot,
     limits   = c(0, 150),
@@
     breaks   = c(5, 10, 20, 40, 70, 100, 150),
-                            labels   = c("0     5","  10","  20","  40","  70","  100"," >150"))
+    labels   = c("0     5", "  10", "  20", "  40", "  70", "  100", " >150"))
 
@@
 library(cowplot)
+
 SOCemis
@@
 
-grid <- plot_grid(SOCemis,SOCdensity,SOCincrease, SOCshare,  ncol=2, labels=c("(a)","(b)", "(c)", "(d)"),
+grid <- plot_grid(SOCemis, SOCdensity, SOCincrease, SOCshare,
+                  ncol = 2, labels = c("(a)", "(b)", "(c)", "(d)"),
                   label_fontface = "plain", label_size = 10)
+
 grid
-ggsave(paste0(outputdir,"/4panelfigure.png"), plot=grid, width = 20, height = 14, dpi = 300, units = "cm")
-ggsave(paste0(outputdir,"/SOCshare.png"),    plot=SOCshare,    width = 10, height = 7, dpi = 300, units = "cm")
-ggsave(paste0(outputdir,"/SOCincrease.png"), plot=SOCincrease, width = 10, height = 7, dpi = 300, units = "cm")
-ggsave(paste0(outputdir,"/SOCdensity.png"),  plot=SOCdensity,  width = 10, height = 7, dpi = 300, units = "cm")
-ggsave(paste0(outputdir,"/SOCemis.png"),     plot=SOCemis,     width = 10, height = 7, dpi = 300, units = "cm")
-
+ggsave(paste0(outputdir, "/4panelfigure.png"),
+       plot = grid, width = 20, height = 14, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "/SOCshare.png"),
+       plot = SOCshare,    width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "/SOCincrease.png"),
+       plot = SOCincrease, width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "/SOCdensity.png"),
+       plot = SOCdensity,  width = 10, height = 7, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "/SOCemis.png"),
+       plot = SOCemis,     width = 10, height = 7, dpi = 300, units = "cm")
 ```

### 2022-07-20 ResultNotebooks/results_SOCpointvalidate.Rmd
update of output scripts and figure

@@
 ---
-title: "SOC budget analysis in comparsion"
+title: "SOC budget analysis in comparsionto point-based measurements"
 output:
@@
 Landuse    <- readRDS(paste0("Data/", rev, scen, "/Landuse.rds"))
-
 ```
@@
 
-  for(u in unique) {
-    sand_reduced[u, , "natveg.mean"]   <- mean(    as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
-    sand_reduced[u, , "crop.mean"]     <- mean(    as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
-    sand_reduced[u, , "natveg.median"] <- median(  as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
-    sand_reduced[u, , "crop.median"]   <- median(  as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
-    sand_reduced[u, , "natveg.max"]    <- max(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
-    sand_reduced[u, , "crop.max"]      <- max(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
-    sand_reduced[u, , "natveg.min"]    <- min(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1])
-    sand_reduced[u, , "crop.min"]      <- min(     as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0])
-    sand_reduced[u, , "natveg.q25"]    <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1], probs = 0.25)
-    sand_reduced[u, , "crop.q25"]      <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0], probs = 0.25)
-    sand_reduced[u, , "natveg.q75"]    <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==1], probs = 0.75)
-    sand_reduced[u, , "crop.q75"]      <- quantile(as.array(sand[u, ,"SOC"])[sand[u,,"Native"]==0], probs = 0.75)
+  for (u in unique) {
+    sand_reduced[u, , "natveg.mean"]   <- mean(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
+    sand_reduced[u, , "crop.mean"]     <- mean(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
+    sand_reduced[u, , "natveg.median"] <- median(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
+    sand_reduced[u, , "crop.median"]   <- median(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
+    sand_reduced[u, , "natveg.max"]    <- max(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
+    sand_reduced[u, , "crop.max"]      <- max(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
+    sand_reduced[u, , "natveg.min"]    <- min(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1])
+    sand_reduced[u, , "crop.min"]      <- min(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0])
+    sand_reduced[u, , "natveg.q25"]    <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1], probs = 0.25)
+    sand_reduced[u, , "crop.q25"]      <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0], probs = 0.25)
+    sand_reduced[u, , "natveg.q75"]    <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 1], probs = 0.75)
+    sand_reduced[u, , "crop.q75"]      <- quantile(as.array(sand[u, , "SOC"])[sand[u, , "Native"] == 0], probs = 0.75)
   }
@@
   sand_natveg_table        <- sand_natveg_table[complete.cases(sand_natveg_table), ][-1]
-
 ```
@@
 correlations[, , "median.natveg"] <- cor.test(statsPoints$median[statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
-correlations[, , "mean.natveg"]   <- cor.test(statsPoints$mean[  statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
+correlations[, , "mean.natveg"]   <- cor.test(statsPoints$mean[statsPoints$Native == 1], statsPoints$thisStudy[statsPoints$Native == 1])$estimate^2
 correlations[, , "median.crop"]   <- cor.test(statsPoints$median[statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2
-correlations[, , "mean.crop"]     <- cor.test(statsPoints$mean[  statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2
+correlations[, , "mean.crop"]     <- cor.test(statsPoints$mean[statsPoints$Native == 0], statsPoints$thisStudy[statsPoints$Native == 0])$estimate^2
 
@@
 correlations[, , "allP.crop"]     <- cor.test(allPoints$ValidationData[allPoints$Native == 1], allPoints$thisStudy[allPoints$Native == 1])$estimate^2
-
 ```
@@
                   geom_point(data = statsPoints,
-                            aes(alpha = str_wrap("Median (R2=0.13)", 7), x = median, y = thisStudy, colour = factor(Native)),
+                            aes(alpha = str_wrap("Median (R2=0.13)", 7),
+                                x = median, y = thisStudy, colour = factor(Native)),
                             shape = 3, size = 2) +
                   geom_segment(data = statsPoints,
-                               aes(alpha = "Min-Max", x = min, y = thisStudy, xend = max, yend = thisStudy, 
+                               aes(alpha = "Min-Max",
+                                   x = min, y = thisStudy, xend = max, yend = thisStudy,
                                    colour = factor(Native)),
@@
                   geom_point(data = allPoints,
-                             aes(alpha = "Point Data", x = ValidationData, y = thisStudy, colour = factor(Native)),
+                             aes(alpha = "Point Data",
+                                 x = ValidationData, y = thisStudy, colour = factor(Native)),
                              shape = 1, size = 0.5) +
-                  labs( x = "Point measurement data (tC/ha) ", y = "Predicted value by this study (tC/ha)") +
-                  scale_x_continuous(trans = 'log10', limits = c(7, 470)) + 
-                  scale_y_continuous(trans = 'log10', limits = c(7, 470)) +
-                  coord_fixed(1) + theme(aspect.ratio=1) +
-                  geom_abline(slope=1, intercept=0, color = "grey", size = 0.5, linetype = "dashed") + 
-                  theme(rect=element_blank(),
-                        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
-                        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
-                        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
+                  labs(x = "Point measurement data (tC/ha) ", y = "Predicted value by this study (tC/ha)") +
+                  scale_x_continuous(trans = "log10", limits = c(7, 470)) +
+                  scale_y_continuous(trans = "log10", limits = c(7, 470)) +
+                  coord_fixed(1) + theme(aspect.ratio = 1) +
+                  geom_abline(slope = 1, intercept = 0, color = "grey", size = 0.5, linetype = "dashed") +
+                  theme(rect = element_blank(),
+                        axis.title.y = element_text(size = legend_size,
+                                                   margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
+                        axis.title.x = element_text(size = legend_size,
+                                                   margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
+                        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, size = legend_size),
                         axis.text.y = element_text(size = legend_size),
-                        legend.position= c(0.87, 0.2), legend.direction="vertical",
+                        legend.position = c(0.87, 0.2), legend.direction = "vertical",
                         legend.title = element_text(size = legend_size),
@@
                         legend.spacing.y = unit(0.1, "cm")) +
-                  scale_color_manual(values = scen_cols, labels = c("Cropland", str_wrap("Natural vegetation", 7)), name = "") +
+                  scale_color_manual(values = scen_cols,
+                                     labels = c("Cropland", str_wrap("Natural vegetation", 7)), name = "") +
                   scale_alpha_manual(name = NULL,
@@
 
-ggsave(paste0(outputdir,"scatterComparePointData_2010.png"), plot=SOCPointValid, height=16, width=16, units="cm", dpi = 300)
-
+ggsave(paste0(outputdir, "scatterComparePointData_2010.png"), 
+       plot = SOCPointValid, height = 16, width = 16, units = "cm", dpi = 300)
 ```
@@
 
-p <- ggplot(testcrop, aes(x=Value, y=soc)) + geom_boxplot() + coord_flip()
+p <- ggplot(testcrop, aes(x = Value, y = soc)) + geom_boxplot() + coord_flip()
 
@@
 
-  points(sand[,,"soc"][sand[,,"native"] == 0],SOCStock_crop_raw[getCells(sand),year,][sand[,,"native"] == 0], col ="red")
+  points(sand[, , "soc"][sand[, , "native"] == 0], SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0], col = "red")
   lines(c(-1000, 1000), c(-1000, 1000), col = "grey")
@@
 
-  plot(sand_reduced[,,"soc_nat"][                        !is.na(sand_reduced[,,"soc_nat"])],
-       SOCStock_natveg_raw[getCells(sand_reduced),year,][!is.na(sand_reduced[,,"soc_nat"])], 
+  plot(sand_reduced[, , "soc_nat"][!is.na(sand_reduced[, , "soc_nat"])],
+       SOCStock_natveg_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_nat"])],
        xlab = "Sandermann", ylab = "This Study",
@@
 
-  points(sand_reduced[,,"soc_crop"][                     !is.na(sand_reduced[,,"soc_crop"])],
-         SOCStock_crop_raw[getCells(sand_reduced),year,][!is.na(sand_reduced[,,"soc_crop"])], 
-         col ="red")
+  points(sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])],
+         SOCStock_crop_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])],
+         col = "red")
   lines(c(-1000, 1000), c(-1000, 1000), col = "grey")
@@
 
-    lines(c(sand_reduced[u,,"soc_crop"] - sand_reduced[u,,"sd_crop"], 
-            sand_reduced[u,,"soc_crop"] + sand_reduced[u,,"sd_crop"]), 
-          c(SOCStock_crop_raw[u,year,],
-            SOCStock_crop_raw[u,year,]), col = "pink")
+    lines(c(sand_reduced[u, , "soc_crop"] - sand_reduced[u, , "sd_crop"],
+            sand_reduced[u, , "soc_crop"] + sand_reduced[u, , "sd_crop"]),
+          c(SOCStock_crop_raw[u, year, ],
+            SOCStock_crop_raw[u, year, ]), col = "pink")
 
-    lines(c(sand_reduced[u,,"soc_nat"] - sand_reduced[u,,"sd_nat"], 
-            sand_reduced[u,,"soc_nat"] + sand_reduced[u,,"sd_nat"]), 
-          c(SOCStock_natveg_raw[u,year,],
-            SOCStock_natveg_raw[u,year,]), col = "grey")
+    lines(c(sand_reduced[u, , "soc_nat"] - sand_reduced[u, , "sd_nat"],
+            sand_reduced[u, , "soc_nat"] + sand_reduced[u, , "sd_nat"]),
+          c(SOCStock_natveg_raw[u, year, ],
+            SOCStock_natveg_raw[u, year, ]), col = "grey")
 
   }
-
 ```
@@
 
-  y <- c(SOCStock_crop_raw[    getCells(sand), year, ][sand[, , "native"] == 0], 
-         SOCStock_natveg_raw[getCells(sand), year, ][sand[, , "native"] == 1])#
+  y <- c(SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0],
+         SOCStock_natveg_raw[getCells(sand), year, ][sand[, , "native"] == 1]) #
 
   x_clean <- c(sand[, , "soc"][sand[, , "native"] == 0],
-               sand[, , "soc"]["CHN.9964",,,invert=TRUE][sand[, , "native"] == 1])
+               sand[, , "soc"]["CHN.9964", , , invert = TRUE][sand[, , "native"] == 1])
 
-  y_clean <- c(SOCStock_crop_raw[  getCells(sand), year, ][sand[, , "native"] == 0], 
-               SOCStock_natveg_raw[getCells(sand), year, ]["CHN.9964",,,invert=TRUE][sand[, , "native"] == 1])
+  y_clean <- c(SOCStock_crop_raw[getCells(sand), year, ][sand[, , "native"] == 0],
+               SOCStock_natveg_raw[getCells(sand), year, ]["CHN.9964", , , invert = TRUE][sand[, , "native"] == 1])
 
@@
 
-  x_red  <- c(sand_reduced[, , "soc_nat" ][!is.na(sand_reduced[, , "soc_nat" ])],
+  x_red  <- c(sand_reduced[, , "soc_nat"][!is.na(sand_reduced[, , "soc_nat"])],
               sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])])
 
-  y_red  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_nat" ])],
-              SOCStock_crop_raw[  getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])
+  y_red  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_nat"])],
+              SOCStock_crop_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])
 
-  x_red2  <- c(sand_reduced[, , "soc_nat" ]["CHN.9964",,,invert=TRUE][!is.na(sand_reduced[, , "soc_nat" ])],
+  x_red2  <- c(sand_reduced[, , "soc_nat"]["CHN.9964", , , invert = TRUE][!is.na(sand_reduced[, , "soc_nat"])],
               sand_reduced[, , "soc_crop"][!is.na(sand_reduced[, , "soc_crop"])])
 
-  y_red2  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ]["CHN.9964",,,invert=TRUE][!is.na(sand_reduced[, , "soc_nat" ])],
-              SOCStock_crop_raw[  getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])
+  y_red2  <- c(SOCStock_natveg_raw[getCells(sand_reduced), year, ]["CHN.9964", , , invert = TRUE][!is.na(sand_reduced[, , "soc_nat"])],
+              SOCStock_crop_raw[getCells(sand_reduced), year, ][!is.na(sand_reduced[, , "soc_crop"])])
 
@@
   cor.test(log(x_red2), log(y_red2))$estimate^2
-
 ```

### 2022-07-20 ResultNotebooks/results_SOCscattervalidate.Rmd
update of output scripts and figure

@@
 ---
-title: "SOC budget analysis in comparsion"
+title: "SOC budget analysis in comparsion to LPJmL4 ad SoilGrids 2.0"
 output:
@@
 Landuse    <- readRDS(paste0("Data/", rev, "histManagement/Landuse.rds"))
-
 ```
@@
 
-cropland        <- Landuse[,,"crop"]
-SOCStock_natveg <- dimSums(SoilCarbon[,,c("naturalstate.natveg")], dim=3)
-SOCStock_lu     <- dimSums(SoilCarbon[,,c("actualstate")]*Landuse, dim=3)/dimSums(Landuse, dim=3)
+cropland        <- Landuse[, , "crop"]
+SOCStock_natveg <- dimSums(SoilCarbon[, , c("naturalstate.natveg")], dim = 3)
+SOCStock_lu     <- dimSums(SoilCarbon[, , c("actualstate")] * Landuse, dim = 3) / dimSums(Landuse, dim = 3)
 
 SOCStock_natveg[cropland <= 0.001] <- NA
-SOCStock_lu[    cropland <= 0.001] <- NA
-
+SOCStock_lu[cropland <= 0.001] <- NA
 ```
@@
 
-  #LPJmL4
+  # LPJmL4
   LPJmL4_CRU <- toolCoord2Isocell(readRDS(paste0("Data/", rev, "histManagement/LPJmL4_CRU.rds")))
-  LPJmL4_CRU <- LPJmL4_CRU[,getYears(Landuse),]
+  LPJmL4_CRU <- LPJmL4_CRU[, getYears(Landuse), ]
 
   LPJmL4_CRU[cropland            <= 0.001] <- NA
-  SoilGrids2[cropland[,"y2010",] <= 0.001] <- NA
-  SoilGrids2_full[cropland[,"y2010",] <= 0.001] <- NA
-
+  SoilGrids2[cropland[, "y2010", ] <= 0.001] <- NA
+  SoilGrids2_full[cropland[, "y2010", ] <= 0.001] <- NA
 ```
@@
   par(mar = mar[[i]])
-  max <- max(    LPJmL4_CRU[icells, years, ], SOCStock_natveg[icells, years, ], na.rm = TRUE)
+  max <- max(LPJmL4_CRU[icells, years, ], SOCStock_natveg[icells, years, ], na.rm = TRUE)
   min <- c(10, 20, 10, 20)
@@
 
-  points(LPJmL4_CRU[icells ,years ,],SOCStock_natveg[icells, years, ], pch = ".")
+  points(LPJmL4_CRU[icells, years, ], SOCStock_natveg[icells, years, ], pch = ".")
 
@@
 
-  text(min[i], max, paste0("R2=", round(cor(LPJmL4_CRU[icells, years, ],SOCStock_natveg[icells, years, ], 
-                                 use = "complete.obs"), 2)), adj = c(0,1))
+  text(min[i], max, paste0("R2=", round(cor(LPJmL4_CRU[icells, years, ], SOCStock_natveg[icells, years, ],
+                                 use = "complete.obs"), 2)), adj = c(0, 1))
   text(max, min[i], paste0("Climate zone: ", gsub("_", " ", climates[i])), adj = c(1, 0))
@@
 
-print(round(cor(LPJmL4_CRU[, years, ],SOCStock_natveg[, years, ], use = "complete.obs"), 2))
-
+print(round(cor(LPJmL4_CRU[, years, ], SOCStock_natveg[, years, ], use = "complete.obs"), 2))
 ```
@@
 climate$N <- as.integer(substring(climate$cell, 5))
-climate   <- climate[order(climate$N),]
+climate   <- climate[order(climate$N), ]
 climates  <- unique(climate$IPCC_reduced)
@@
   par(mar = mar[[i]])
-  max <- max(    SoilGrids2[icells ,years, ], SOCStock_lu[icells, years, ], na.rm = TRUE)
+  max <- max(SoilGrids2[icells, years, ], SOCStock_lu[icells, years, ], na.rm = TRUE)
   min <- c(5, 10, 5, 10)
 
-  plot(SoilGrids2[icells, years, ],SOCStock_lu[icells, years, ], asp = 1, 
+  plot(SoilGrids2[icells, years, ], SOCStock_lu[icells, years, ], asp = 1,
        xlab = xlabel[i], ylab = ylabel[i],
@@
 
-  points(SoilGrids2[icells, years, ],SOCStock_lu[icells, years, ], pch = ".")
+  points(SoilGrids2[icells, years, ], SOCStock_lu[icells, years, ], pch = ".")
 
@@
   text(min[i], max, paste0("R2=", round(cor(SoilGrids2[icells, years, ], SOCStock_lu[icells, years, ],
-                                 use = "complete.obs"), 2)), adj = c(0,1))
+                                 use = "complete.obs"), 2)), adj = c(0, 1))
   text(max, min[i], paste0("Climate zone: ", gsub("_", " ", climates[i])), adj = c(1, 0))
@@
 
-print(round(cor(SoilGrids2[, years, ],SOCStock_lu[, years, ], use = "complete.obs"), 2))
-
+print(round(cor(SoilGrids2[, years, ], SOCStock_lu[, years, ], use = "complete.obs"), 2))
 ```
@@
 
-InOrOut <- SOCStock_lu[,"y2010",]
+InOrOut <- SOCStock_lu[, "y2010", ]
 InOrOut[] <- NA
-InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q05"] & 
-        SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q95"]] <- 1
-InOrOut[SOCStock_lu[,"y2010",] < SoilGrids2_full[,"y2010","SOC_Q05"]] <- 0.5
-InOrOut[SOCStock_lu[,"y2010",] > SoilGrids2_full[,"y2010","SOC_Q95"]] <- 1.5
+InOrOut[SOCStock_lu[, "y2010", ] > SoilGrids2_full[, "y2010", "SOC_Q05"] &
+        SOCStock_lu[, "y2010", ] < SoilGrids2_full[, "y2010", "SOC_Q95"]] <- 1
+InOrOut[SOCStock_lu[, "y2010", ] < SoilGrids2_full[, "y2010", "SOC_Q05"]] <- 0.5
+InOrOut[SOCStock_lu[, "y2010", ] > SoilGrids2_full[, "y2010", "SOC_Q95"]] <- 1.5
 
-magpie_coord      <- toolGetMapping("magpie_coord.rda", type="cell")
-soilgrid22plot    <- merge(as.data.frame(InOrOut), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
-
+magpie_coord      <- toolGetMapping("magpie_coord.rda", type = "cell")
+soilgrid22plot    <- merge(as.data.frame(InOrOut), data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
 ```
@@
 data(coastlineWorld)
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]], latitude = coastlineWorld[["latitude"]])
 
@@
   out <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(data, Year == "2010", c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
+    geom_tile(data = subset(data, Year == "2010", c("lon", "lat", "Value")), aes(x = lon, y = lat, fill = factor(Value))) +
     scale_fill_manual(name     = name,
@@
 cols              <- c("#648FFF", "#DC267F", "#FFB000")[c(1, 3, 2)]
-SoilGrids2InMap   <- SOCplotmap( data     =  soilgrid22plot,
+SoilGrids2InMap   <- SOCplotmap(data     =  soilgrid22plot,
                                  limits   = c(0.5, 1.5),
@@
 SoilGrids2InMap
-ggsave(paste0(outputdir, "/SOCquantilesSoilGrids2.png"),    plot=SoilGrids2InMap,    width = 15, height = 10, dpi = 300, units = "cm")
-
+ggsave(paste0(outputdir, "/SOCquantilesSoilGrids2.png"),    plot = SoilGrids2InMap,    width = 15, height = 10, dpi = 300, units = "cm")
 ```

### 2022-07-20 ResultNotebooks/results_SOCscenario.Rmd
update of output scripts and figure

@@
 rev         <- "rev30"
-scenarios   <- c("histManagement", "constClimate-1975", "constManagement-1975", "constManagement2-1975", "constManagement-1975_constClimate-1975", "constManagement2-1975_constClimate-1975", "constManure-1975", "constResidues-1975", "constResrate-1975", "constTillage-mixedtill", "constYield-1975")
+scenarios   <- c("histManagement", 
+                 "constManagement-1975", 
+                 "constManure-1975", 
+                 "constResidues-1975", 
+                 "constTillage-mixedtill")
 
- filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
+ filesNeeded  <- c("SoilCarbon.rds", "Landuse.rds")
 # setConfig(outputfolder = "/home/kristine/WORK/Paper/historicalsocbudget/ResultNotebooks/Data/")
@@
 outputdir <- "Output/"
-
 ```
@@
 
-for(scen in scenarios){
+for (scen in scenarios) {
+
   print(scen)
-  SoilCarbon <- read.magpie(paste0("Data/", paste0(rev,scen), "/SoilCarbon.rds"))
-  Landuse    <- read.magpie(paste0("Data/", paste0(rev,scen), "/Landuse.rds"))
-  SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
-  SOCStock        <- toolConditionalReplace(dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2), "is.nan()",0)
-  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), 
-                                                       dim=3.1, nm=scen, add="scenario")) 
+  SoilCarbon <- read.magpie(paste0("Data/", paste0(rev, scen), "/SoilCarbon.rds"))
+  Landuse    <- read.magpie(paste0("Data/", paste0(rev, scen), "/Landuse.rds"))
+  SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
+  SOCStock  <- toolConditionalReplace(
+                 dimSums(dimSums(
+                   SoilCarbon[, , "state", pmatch = TRUE], dim = 3.3) * 
+                     Landuse, dim = 3.2), "is.nan()", 0)
+  glo_SOCStock <- mbind(glo_SOCStock, 
+                        add_dimension(dimSums(SOCStock, dim = 1),
+                                      dim = 3.1, nm = scen, 
+                                      add = "scenario"))
 }
 
-glo_SOCDebt  <- collapseNames(glo_SOCStock[,,"naturalstate"] - glo_SOCStock[,,"actualstate"])/1000
-glo_SOCShare <- collapseNames(toolConditionalReplace(glo_SOCStock[,,"actualstate"]/glo_SOCStock[,,"naturalstate"], "is.na()",0) )
+glo_SOCDebt  <- collapseNames(glo_SOCStock[, , "naturalstate"] - 
+                                glo_SOCStock[, , "actualstate"]) / 1000
+glo_SOCShare <- collapseNames(
+                  toolConditionalReplace(glo_SOCStock[, , "actualstate"] /
+                                           glo_SOCStock[, , "naturalstate"], 
+                                         "is.na()", 0))
 
-saveRDS(mbind(add_dimension(collapseNames(glo_SOCStock[,,"actualstate"]), dim=3.1, add="type", nm="stock"),
-              add_dimension(              glo_SOCDebt,                    dim=3.1, add="type", nm="debt"),
-              add_dimension(              glo_SOCShare,                   dim=3.1, add="type", nm="share")), 
+saveRDS(mbind(add_dimension(
+                collapseNames(glo_SOCStock[, , "actualstate"]), 
+                dim = 3.1, add = "type", nm = "stock"),
+              add_dimension(glo_SOCDebt,                    
+                dim = 3.1, add = "type", nm = "debt"),
+              add_dimension(glo_SOCShare,                   
+                dim = 3.1, add = "type", nm = "share")),
         paste0(outputdir, "glo_SOCscenarions.rds"))
@@
 
-CarbonInput  <- read.magpie(paste0("Data/", rev, "histManagement/CarbonInput.rds"))
-Landuse      <- read.magpie(paste0("Data/", rev, "histManagement/Landuse.rds"))
-glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))
+CarbonInput  <- read.magpie(paste0("Data/", rev, 
+                                   "histManagement/CarbonInput.rds"))
+Landuse      <- read.magpie(paste0("Data/", rev, 
+                                   "histManagement/Landuse.rds"))
+glo_SOCInput <- collapseNames(dimSums(CarbonInput[, , "totalC"] * 
+                                        Landuse, dim = c(1))) / collapseNames(dimSums(Landuse, dim = c(1)))
 
-ci2plot <- c('above-ground residues'="ag.crop",          
-             'below-ground residues'="bg.crop",          
-             'manure from monogastrics'="livst_chick.crop",
-             'manure from monogastrics'="livst_egg.crop",
-             'manure from monogastrics'="livst_pig.crop",
-             'manure from ruminants'="livst_milk.crop",  
-             'manure from ruminants'="livst_rum.crop",   
-             'natural litterfall wood fraction'="wood.natveg",
-             'natural litterfall leaf fraction'="leaf.natveg")
+ci2plot <- c("above-ground residues" = "ag.crop",
+             "below-ground residues" = "bg.crop",
+             "manure from monogastrics" = "livst_chick.crop",
+             "manure from monogastrics" = "livst_egg.crop",
+             "manure from monogastrics" = "livst_pig.crop",
+             "manure from ruminants" = "livst_milk.crop",
+             "manure from ruminants" = "livst_rum.crop",
+             "natural litterfall wood fraction" = "wood.natveg",
+             "natural litterfall leaf fraction" = "leaf.natveg")
 
-glo_SOCInput <- toolAggregate(glo_SOCInput, rel=data.frame(from=ci2plot,to=names(ci2plot)), from="from", to="to", dim=3)
-
+glo_SOCInput <- toolAggregate(glo_SOCInput, 
+                  rel = data.frame(from = ci2plot, to = names(ci2plot)), 
+                  from = "from", to = "to", dim = 3)
 ```
@@
 
-glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
+glo_SOCDebt <- collapseNames(readRDS(
+                 paste0(outputdir, "glo_SOCscenarions.rds"))[, , "debt"])
 
@@
                   constResidues    = "constResidues-1975",
-                  constManagement  = "constManagement-1975")#,
-                  #constManagementNoCC = "rev26constManagement_constClimate")
+                  constManagement  = "constManagement-1975")
 
@@
 legend_size <- 10
-scen_cols   <- c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000")[5:1]
+scen_cols   <- c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000")[5:1]
 input_cols  <- scen_cols[1:4]
 
-limits <- c(-42,-34)
-breaks <- c(-34, -36,- 38, -40, -42)
+limits <- c(-42, -34)
+breaks <- c(-34, -36, -38, -40, -42)
+years  <- seq(1975, 2010, 1)
 
-hSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),manage_scen]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
-  labs( x = "Year", 
-        y = expression(paste('Global \u0394SOC in GtC')),
-        color="") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits=limits, breaks=breaks) +
-  scale_color_manual(values=scen_cols, labels = names(manage_scen)) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
-        legend.position="bottom", legend.direction="vertical",
-        legend.margin=margin(-10,-10,0,0),
-        legend.box.margin=margin(-10,-10,0,0))
+hSOCemis <- 
+  ggplot() + 
+  geom_line(data = as.data.frame(-glo_SOCDebt[, years, manage_scen]),  
+            aes(x = Year, y = Value, group = Data1, color = Data1)) +
+  labs(x = "Year", y = expression(paste("Global \u0394SOC in GtC")), 
+       color = "") +
+  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
+  scale_y_continuous(limits = limits, breaks = breaks) +
+  scale_color_manual(values = scen_cols, labels = names(manage_scen)) +
+  theme(rect = element_blank(),
+        axis.title.y = element_text(size = legend_size, 
+                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.title.x = element_text(size = legend_size, 
+                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
+                                   size = legend_size),
+        legend.position = "bottom", legend.direction = "vertical",
+        legend.margin = margin(-10, -10, 0, 0),
+        legend.box.margin = margin(-10, -10, 0, 0))
 
-hSOCinput <- ggplot(data= as.data.frame(glo_SOCInput[,seq(1975,2010,1),c(3:4,2,1)]),  aes(x = Year, y = Value, group=Data1, fill=Data1)) + geom_area() +
-  labs( x = "Year", 
-        y = "SOC input in tC per ha\n (global average)",
-        fill="") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
+hSOCinput <- 
+  ggplot(data = as.data.frame(glo_SOCInput[, years, c(3:4, 2, 1)]),  
+         aes(x = Year, y = Value, group = Data1, fill = Data1)) + geom_area() +
+  labs(x = "Year", y = "SOC input in tC per ha\n (global average)", fill = "") +
+  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
   scale_y_continuous() +
-  scale_fill_manual(values=input_cols) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
-        legend.position="bottom", 
-        legend.direction="vertical",
-        legend.margin=margin(-10,-10,0,0),
-        legend.box.margin=margin(-10,-10,0,0))
+  scale_fill_manual(values = input_cols) +
+  theme(rect = element_blank(),
+        axis.title.y = element_text(size = legend_size, 
+                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.title.x = element_text(size = legend_size, 
+                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
+                                   size = legend_size),
+        legend.position = "bottom",
+        legend.direction = "vertical",
+        legend.margin = margin(-10, -10, 0, 0),
+        legend.box.margin = margin(-10, -10, 0, 0))
 
-vSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1974,2010,1),manage_scen]),  aes(x = Year, y = Value, group=Data1, color=Data1)) +
-  labs( x = "Year", 
-        y = expression(paste('Global \u0394SOC in GtC')),
-        color="") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
+vSOCemis <- 
+  ggplot() + geom_line(data = as.data.frame(-glo_SOCDebt[, years, manage_scen]),  
+                       aes(x = Year, y = Value, group = Data1, color = Data1)) +
+  labs(x = "Year", y = expression(paste("Global \u0394SOC in GtC")),
+       color = "") +
+  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
   scale_y_continuous(limits = limits,
                      breaks = breaks) +
-  scale_color_manual(values=scen_cols, labels = names(manage_scen)) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
-        legend.position="right", legend.direction="vertical")
+  scale_color_manual(values = scen_cols, labels = names(manage_scen)) +
+  theme(rect = element_blank(),
+        axis.title.y = element_text(size = legend_size, 
+                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.title.x = element_text(size = legend_size, 
+                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
+                                   size = legend_size),
+        legend.position = "right", legend.direction = "vertical")
 
-vSOCinput <- ggplot(data= as.data.frame(glo_SOCInput[,seq(1975,2010,1),c(3:4,2,1)]),  aes(x = Year, y = Value, group=Data1, fill=Data1)) + geom_area() +
-  labs( x = "Year", 
-        y = "SOC input in tC per ha\n (global average)",
-        fill="") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
+vSOCinput <- 
+  ggplot(data = as.data.frame(glo_SOCInput[, years, c(3:4, 2, 1)]),  
+         aes(x = Year, y = Value, group = Data1, fill = Data1)) + geom_area() +
+  labs(x = "Year", y = "SOC input in tC per ha\n (global average)", fill = "") +
+  scale_x_discrete(breaks = seq(1975, 2010, 5)) +
   scale_y_continuous() +
-  scale_fill_manual(values=input_cols) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t = 10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size), legend.direction="vertical")
+  scale_fill_manual(values = input_cols) +
+  theme(rect = element_blank(),
+        axis.title.y = element_text(size = legend_size, 
+                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.title.x = element_text(size = legend_size, 
+                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
+                                   size = legend_size), 
+        legend.direction = "vertical")
 
@@
 
-hgrid <- plot_grid(hSOCemis, hSOCinput, ncol=2, labels=c("(a)","(b)"),
-                  label_fontface = "plain", label_size = 10, align="hv")
+hgrid <- plot_grid(hSOCemis, hSOCinput, ncol = 2, labels = c("(a)", "(b)"),
+                  label_fontface = "plain", label_size = 10, align = "hv")
 
-vgrid <- plot_grid(plot_grid(vSOCemis  + theme(legend.position = "none") , 
-                             vSOCinput + theme(legend.position = "none"), ncol=1, labels=c("(a)","(b)"),
-                             label_fontface = "plain", label_size = 10, align="v", axis="tblr"),
-                   plot_grid(get_legend(vSOCemis), get_legend(vSOCinput), ncol=1,
-                             label_fontface = "plain", label_size = 10, align="v"),
-                   align="h")
+vgrid <- plot_grid(plot_grid(vSOCemis  + theme(legend.position = "none"),
+                             vSOCinput + theme(legend.position = "none"), 
+                             ncol = 1, labels = c("(a)", "(b)"),
+                             label_fontface = "plain", label_size = 10, 
+                             align = "v", axis = "tblr"),
+                   plot_grid(get_legend(vSOCemis), get_legend(vSOCinput), 
+                             ncol = 1, label_fontface = "plain", 
+                             label_size = 10, align = "v"),
+                   align = "h")
 
@@
 
-ggsave(paste0(outputdir,"Images/scenario_horiz.png"), plot=hgrid, width = 15, height = 10, dpi = 300, units = "cm")
-ggsave(paste0(outputdir,"Images/scenario_verti.png"), plot=vgrid, width = 12, height = 15, dpi = 300, units = "cm")
-
+ggsave(paste0(outputdir, "Images/scenario_horiz.png"), 
+       plot = hgrid, width = 15, height = 10, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "Images/scenario_verti.png"), 
+       plot = vgrid, width = 12, height = 15, dpi = 300, units = "cm")
 ```
@@
 
-glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
+glo_SOCStock_init <- NULL
+init_scenarios    <- c("Initial-0_constManagement-1975",
+                       "Initial-1901_constManagement-1975")
 
-init_scen   <-  c(`histManagement.Initial-natveg`          = "rev26Initial-1901", 
-                  `constManagement.Initial-natveg`         = "rev26Initial-1901_constManagement",
-                  `histManagement.Initial-eq`              = "rev26Initial-0", 
-                  `constManagement.Initial-eq`             = "rev26Initial-0_constManagement",
-                  `histManagement.Initial-spinup1530`      = "rev26histManagement", 
-                  `constManagement.Initial-spinup1530`     = "rev26constManagement")
+for (scen in init_scenarios) {
+  print(scen)
+  SoilCarbon <- read.magpie(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
+  Landuse    <- read.magpie(paste0("Data/", rev, scen, "/Landuse.rds"))
+  SoilCarbon[, , "naturalstate.crop"] <- SoilCarbon[, , "naturalstate.natveg"]
+  SOCStock   <- toolConditionalReplace(
+                  dimSums(dimSums(
+                    SoilCarbon[, , c("naturalstate", 
+                                     "actualstate")], dim = 3.3) * 
+                      Landuse, dim = 3.2), "is.nan()", 0)
+
+  glo_SOCStock_init <- mbind(glo_SOCStock_init, 
+                            add_dimension(dimSums(SOCStock, dim = 1),
+                                          dim = 3.1, 
+                                          nm = scen, 
+                                          add = "scenario"))
+}
+
+glo_SOCDebt_init  <- collapseNames(glo_SOCStock_init[, , "naturalstate"] - 
+                                glo_SOCStock_init[, , "actualstate"]) / 1000
+
+saveRDS(add_dimension(glo_SOCDebt_init,                    
+                      dim = 3.1, add = "type", nm = "debt_init"),
+        paste0(outputdir, "glo_SOCscenarions_initConst.rds"))
+
+glo_SOCDebt <-
+  mbind(collapseNames(readRDS(
+          paste0(outputdir,
+                 "glo_SOCscenarions.rds"))[, , "debt"]),
+        collapseNames(readRDS(
+          paste0(outputdir,
+                 "glo_SOCscenarions_init.rds"))[, , "debt_init"]),
+        collapseNames(readRDS(
+          paste0(outputdir, 
+                 "glo_SOCscenarions_initConst.rds"))[, , "debt_init"]))
+
+init_scen   <-  
+  c( `histManagement.Initial-natveg`         = "Initial-1901",
+    `constManagement.Initial-natveg`         = "Initial-1901_constManagement-1975",
+     `histManagement.Initial-eq`             = "Initial-0",
+    `constManagement.Initial-eq`             = "Initial-0_constManagement-1975",
+     `histManagement.Initial-spinup1510`     = "histManagement",
+    `constManagement.Initial-spinup1510`     = "constManagement-1975")
 
@@
 legend_size <- 10
-scen_cols   <- c("#648FFF","#785EF0","#DC267F")
+scen_cols   <- c("#648FFF", "#785EF0", "#DC267F")
+
 input_cols  <- scen_cols[1:4]
 
-limits <- c(-55,-25)
-breaks <- c( -30,-40,-50 )
+limits <- c(-55, -25)
+breaks <- c(-30, -40, -50)
 
-initSOCemis <- ggplot() + geom_path(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),]), 
-                                    aes(x = Year, y = Value, group=interaction(Data1,Data2), linetype=Data1, color=Data2)) +
-  labs( x = "Year", 
-        y = expression(paste('Global \u0394SOC in GtC')),
-        color    ="Initialization",
-        linetype ="Management") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits=limits, breaks=breaks) +
-  scale_color_manual(values=c(scen_cols)[1:3], labels = unique(gsub("(.*)\\.(.*)","\\2", names(init_scen)))) +
-  scale_linetype_manual(values=c(1,2), labels=unique(gsub("(.*)\\.(.*)","\\1", names(init_scen)))) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
-        legend.position="right", legend.direction="vertical")
+startY <- 1965
+
+initSOCemis <- 
+  ggplot() + 
+  geom_path(data = as.data.frame(-glo_SOCDebt[, seq(startY, 2010, 1), ]),
+            aes(x = Year, y = Value, group = interaction(Data1, Data2), 
+                linetype = Data1, color = Data2)) +
+  labs(x = "Year",
+        y = expression(paste("Global \u0394SOC in GtC")),
+        color    = "Initialization",
+        linetype = "Management") +
+  scale_x_discrete(breaks = seq(startY, 2010, 5)) +
+  scale_y_continuous(limits = limits, breaks = breaks) +
+  scale_color_manual(values = c(scen_cols), 
+                     labels = unique(gsub("(.*)\\.(.*)", "\\2", 
+                                          names(init_scen)))) +
+  scale_linetype_manual(values = c(1, 2), 
+                        labels = unique(gsub("(.*)\\.(.*)", "\\1", 
+                                             names(init_scen)))) +
+  theme(rect = element_blank(),
+        axis.title.y = element_text(size = legend_size, 
+                                    margin = margin(t = 0, r = 10, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.title.x = element_text(size = legend_size, 
+                                    margin = margin(t = 10, r = 0, b = 0, l = 0, 
+                                                    unit = "pt")),
+        axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, 
+                                   size = legend_size),
+        legend.position = "right", legend.direction = "vertical")
 
@@
 
-ggsave(paste0(outputdir,"Images/scenario_init.png"), plot= initSOCemis, width = 10, height = 6, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "Images/scenario_init.png"), 
+       plot = initSOCemis, width = 12, height = 8, dpi = 300, units = "cm")
 ```
-
-```{r Plot Litter param Scenarios, out.height="6cm", out.width="18cm"}
-
-glo_SOCDebt <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"debt"])
-glo_SOCStock <- collapseNames(readRDS(paste0(outputdir, "glo_SOCscenarions.rds"))[,,"stock"])
-
-litter_scen <-  c(`LitterPNV-Mean` = "histManagement",
-                  `LitterPNV-Grass`= "LitterPNV-PerennialGrasses", 
-                  `LitterPNV-AASS` = "LitterPNV-AASS", 
-                  `LitterPNV-BRCF` = "LitterPNV-BRCF", 
-                  `LitterPNV-MTCF` = "LitterPNV-MTCF", 
-                  `LitterPNV-BLEAF`= "LitterPNV-BLEAF",  
-                  `LitterPNV-TREF` = "LitterPNV-TREF")
-
-library(ggplot2)
-library(scales)
-library(gridExtra)
-library(cowplot)
-
-theme_set(theme_bw())
-legend_size <- 10
-scen_cols   <- c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000")[5:1]
-input_cols  <- scen_cols[1:4]
-
-limits <- c(-48,-40)
-breaks <- c(-40, -42, -44, -46, -48)
-
-litterSOCemis <- ggplot() + geom_line(data= as.data.frame(-glo_SOCDebt[,seq(1975,2010,1),litter_scen]), 
-                                    aes(x = Year, y = Value, group=Data1, color=Data1, size=Data1)) +
-  labs( x = "Year", 
-        y = expression(paste('Global \u0394SOC in GtC')),
-        color="") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits=limits, breaks=breaks) +
-  scale_color_manual(values=c("red", scen_cols, "blue"), labels = names(litter_scen)) +
-  scale_size_manual(values=c(1,rep(0.5, length(litter_scen)-1)), labels = names(litter_scen)) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
-        legend.position = "none")
-
-litterSOCemis
-
-limits <- c(625,705)
-breaks <- c(635, 655, 675, 695)
-
-litterSOCstock <- ggplot() + geom_line(data= as.data.frame(glo_SOCStock[,seq(1975,2010,1),litter_scen]/1000), 
-                                    aes(x = Year, y = Value, group=Data1, color=Data1, size=Data1)) +
-  labs( x = "Year", 
-        y = "Global SOC stock\n in GtC",
-        color="Litter parameterization",
-        size="Litter parameterization") +
-  scale_x_discrete(breaks=seq(1975,2010,5)) +
-  scale_y_continuous(limits=limits, breaks=breaks) +
-  scale_color_manual(values=c("red", scen_cols, "blue"), labels = names(litter_scen)) +
-  scale_size_manual(values=c(1,rep(0.5, length(litter_scen)-1)), labels = names(litter_scen)) +
-  theme(rect=element_blank(),
-        axis.title.y= element_text(size=legend_size, margin = margin(t = 0, r =10, b = 0, l = 0, unit = "pt")),
-        axis.title.x= element_text(size=legend_size, margin = margin(t =10, r = 0, b = 0, l = 0, unit = "pt")),
-        axis.text.x = element_text(angle=60, hjust=0.8, vjust=0.8, size=legend_size),
-        legend.direction="vertical",
-        legend.margin=margin(-10,-10,0,0),
-        legend.box.margin=margin(-10,-10,0,0)) +
-  guides(color=guide_legend(), size = guide_legend())
-
-litterSOCstock
-
-littergrid <- plot_grid(litterSOCemis, litterSOCstock + theme(legend.position = "none"), get_legend(litterSOCstock),
-                                  ncol=3, labels=c("(a)","(b)"), axis = "l",
-                                  label_fontface = "plain", label_size = 10, align="v")
-littergrid
-
-ggsave(paste0(outputdir,"Images/scenario_litter.png"), plot=littergrid, width = 18, height = 6, dpi = 300, units = "cm")
-
-```

### 2022-07-20 ResultNotebooks/results_SOCtable.Rmd
update of output scripts and figure

--- 2021-04-01 ResultNotebooks/results_SOCtable.Rmd
@@
 library(mrvalidation)
-library(mrsoil)
+library(mrcommons)
 library(magpiesets)
 
-scenarios <- c("rev19histManagement","rev18histManagement_ResAllom_2021_02_01")
 outputdir <- "Output/"
 
-#untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"), exdir="Data/histManagement", extras="--keep-newer-files ")
+rev         <- "rev30"
+scen        <- "histManagement"
 
-SoilCarbon <- read.magpie(paste0("Data/",scenarios[2],"/SoilCarbon.rds"))
-Landuse    <- read.magpie(paste0("Data/",scenarios[2],"/Landuse.rds"))
+# untar(paste0(getConfig("outputfolder"),"/",rev, "_h12_carbonbudget.tgz"),
+#       exdir="Data/histManagement", extras="--keep-newer-files ")
+
+SoilCarbon <- readRDS(paste0("Data/", rev, scen, "/SoilCarbon.rds"))
+Landuse    <- readRDS(paste0("Data/", rev, scen, "/Landuse.rds"))
+
 ```
@@
 library(reshape2)
+library(Hmisc)
 
-years             <- findset("past")
-climateIPCC       <- readRDS("Data/histManagement/IPCC_reduced.rds")
+years             <- seq(1975, 2010, 5)
+climateIPCCr      <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))
 
-### calculating stock change factors
-SOCDens           <- dimSums(SoilCarbon[,years,c("actualstate","naturalstate")], dim=3.3) 
-SOCDens[,,"naturalstate.crop"] <- SOCDens[,,"actualstate.crop"]
+### calculating stock change factors cropand weighted
+SOCDens <- dimSums(SoilCarbon[, years, c("actualstate", "naturalstate")], 
+                   dim = 3.3)
+SOCDens[, , "naturalstate.crop"] <- SOCDens[, , "actualstate.crop"]
 
-CShare_ipcczone   <- dimSums(climateIPCC * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
-                     dimSums(climateIPCC * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)
+Cshare_cell       <- collapseNames(climateIPCCr * 
+                       toolConditionalReplace(
+                         SOCDens[, , "crop"][, , "actualstate"] /
+                         SOCDens[, , "natveg"][, , "naturalstate"],
+                         c("is.na()", "is.nan()", "is.infinite()"), 
+                         c(NA, NA, NA)))
 
-getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))
+CShare_study      <- new.magpie("GLO", years,  
+                          outer(getNames(climateIPCCr),
+                                c("Q0p05", "Q0p5", "Q0p95"), 
+                                paste, sep = "."))
 
-SCF_table  <- data.frame(dcast(as.data.frame(collapseNames(CShare_ipcczone[,years,"naturalstate"])), 
-                             Year ~ Data1, value.var="Value", row.names=years), row.names = "Year")
-saveRDS(list(SCF_table = SCF_table), paste0(outputdir, "SFC_climatezone.rds"))
+cropMask  <- (Landuse[, years, "crop"] > 0.001)
 
+for (zone in getNames(climateIPCCr)) {
+
+  tempMMask <- setYears(
+                 climateIPCCr[, rep(1, length(years)) , zone] == 1,
+                 years) & cropMask
+
+  for(y in years) {
+
+    cropTempM <- Landuse[, y , "crop"][tempMMask[, y, ]]
+    CShare_study[, y , zone] <- 
+    Hmisc::wtd.quantile(Cshare_cell[, y, zone][tempMMask[, y, ]], 
+                        cropTempM, 
+                        probs = c(0.05, 0.5, 0.95))  
+  }
+}
+
+getNames(CShare_study) <- gsub("_", " ", getNames(CShare_study))
+getCells(CShare_study) <- "GLO"
+
+SCF_table  <- as.data.frame(CShare_study)[-1]
+
+saveRDS(SCF_table, paste0(outputdir, "SFC_climatezone+range.rds"))
 ```
@@
 
-climateIPCCr       <- readRDS("Data/histManagement/IPCC_reduced.rds")
-climateIPCC        <- readRDS("Data/histManagement/IPCC.rds")
+climateIPCCr <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))
+climateIPCC  <- readRDS(paste0("Data/", re, scen, "/IPCC.rds"))
 
-getNames(climateIPCC) <- c( "W-Temp Moist","W-Temp Dry","C-Temp Moist","C-Temp Dry","Polar Moist","Polar Dry","Boreal Moist",
-                            "Boreal Dry","Tropical Montane","Tropical Wet","Tropical Moist", "Tropical Dry")
+getNames(climateIPCC) <- c("W-Temp Moist", "W-Temp Dry", "C-Temp Moist", 
+                           "C-Temp Dry", "Polar Moist", "Polar Dry", 
+                           "Boreal Moist",  "Boreal Dry", "Tropical Montane", 
+                           "Tropical Wet", "Tropical Moist", "Tropical Dry")
 
-climateIPCC <- climateIPCC[,,c(5,7,3,1,6,8,4,2,9,11,10,12)]
+climateIPCC <- climateIPCC[, , c(5, 7, 3, 1, 6, 8, 4, 2, 9, 11, 10, 12)]
 
-magpie_coord <- toolGetMapping("magpie_coord.rda", type="cell")
-climate2plot <- dimSums(climateIPCC*new.magpie("GLO",NULL,getNames(climateIPCC),c(1:12)), dim=3)
-climate2plot <- merge(as.data.frame(climate2plot), data.frame(magpie_coord, Cell=c(1:59199)), by="Cell")
+magpie_coord <- toolGetMapping("magpie_coord.rda", type = "cell")
+climate2plot <- dimSums(climateIPCC * 
+                          new.magpie("GLO", NULL, 
+                                     getNames(climateIPCC), 
+                                     c(1:12)), 
+                        dim = 3)
+climate2plot <- merge(as.data.frame(climate2plot), 
+                      data.frame(magpie_coord, Cell = c(1:59199)), by = "Cell")
 
@@
 data(coastlineWorld)
-coastline.world <- data.frame(longitude=coastlineWorld[["longitude"]],latitude=coastlineWorld[["latitude"]])
+coastline.world <- data.frame(longitude = coastlineWorld[["longitude"]], 
+                              latitude = coastlineWorld[["latitude"]])
 
@@
 orgcols          <- brewer.pal(9, "Oranges")[3]
-cols <- c(blucols,viocols,redcols,orgcols)
+cols <- c(blucols, viocols, redcols, orgcols)
 
 ClimateZones <- ggplot(data = world) + geom_sf() +
-    geom_tile(data = subset(climate2plot, select=c("lon","lat","Value")), aes(x = lon, y = lat, fill = factor(Value))) +
+    geom_tile(data = subset(climate2plot, select = c("lon", "lat", "Value")), 
+              aes(x = lon, y = lat, fill = factor(Value))) +
     scale_fill_manual(name = "",
                       values = cols, na.value = "white",
-                      breaks= c(1,2,3,4,5,6,7,8,9,10,11,12),
+                      breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12),
                       labels = getNames(climateIPCC),
                       guide = guide_legend(
-                        title.theme=element_text(size=legend_size),
+                        title.theme = element_text(size = legend_size),
                         direction = "horizontal",
-                        title.position = 'top',
+                        title.position = "top",
                         # I shift the labels around, the should be placed
@@
                         nrow = 1,
-                        byrow = F,
+                        byrow = FALSE,
                         label.position = "bottom",
@@
                       )) +
-   scale_color_manual(values=cols) +
-    theme(legend.position="bottom",
+   scale_color_manual(values = cols) +
+    theme(legend.position = "bottom",
           axis.title.x = element_blank(),
           axis.title.y = element_blank(),
-          legend.text=element_text(size=legend_size),
-          rect=element_blank(),
-          panel.background=element_rect(fill="#F8F8FF", size=c(0.7,0.7)),
+          legend.text = element_text(size = legend_size),
+          rect = element_blank(),
+          panel.background = element_rect(fill = "#F8F8FF", size = c(0.7, 0.7)),
           plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
-          ) + guides(fill=guide_legend(nrow=2,byrow=FALSE))
+          ) + guides(fill = guide_legend(nrow = 2, byrow = FALSE))
 ClimateZones
-ggsave(paste0(outputdir,"/Images/climatezones.png"), plot=ClimateZones , width = 20, height = 14, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "/Images/climatezones.png"), 
+       plot = ClimateZones, width = 20, height = 14, dpi = 300, units = "cm")
 ```
@@
 
-MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
-                                                                          getNames(climateIPCCr), paste, sep="."))
+climateIPCCr <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))
 
-MShare_ipcc <- add_dimension(MShare_ipcc, dim=3.2, nm="medium.invariant", add="input.year")
-LShare_ipcc <- add_dimension(LShare_ipcc, dim=3.2, nm="low.invariant",    add="input.year")
-HShare_ipcc <- add_dimension(HShare_ipcc, dim=3.2, nm="high.invariant",   add="input.year")
+categories   <- c("IPCC2006.low", "IPCC2006.medium", "IPCC2006.high",
+                  "IPCC2019.low", "IPCC2019.medium", "IPCC2019.high" )
 
-MShare_ipcc[,,"IPCC2006.medium"][,,"tropical_moist"]  <- "0.48"
-MShare_ipcc[,,"IPCC2006.medium"][,,"tropical_dry"]    <- "0.58--0.64" 
-MShare_ipcc[,,"IPCC2006.medium"][,,"temperate_moist"] <- "0.69"
-MShare_ipcc[,,"IPCC2006.medium"][,,"temperate_dry"]   <- "0.80"
+CShare_ipcc <- new.magpie("GLO", NULL,  
+                          outer(categories, 
+                                getNames(climateIPCCr),
+                                paste, sep = "."))
 
-MShare_ipcc[,,"IPCC2019.medium"][,,"tropical_moist"]  <- "0.83"
-MShare_ipcc[,,"IPCC2019.medium"][,,"tropical_dry"]    <- "0.92" 
-MShare_ipcc[,,"IPCC2019.medium"][,,"temperate_moist"] <- "0.69--0.70"
-MShare_ipcc[,,"IPCC2019.medium"][,,"temperate_dry"]   <- "0.76--0.77"
+y2006m <- "IPCC2006.medium" 
+y2006l <- "IPCC2006.low" 
+y2006h <- "IPCC2006.high" 
+y2019m <- "IPCC2019.medium" 
+y2019l <- "IPCC2019.low" 
+y2019h <- "IPCC2019.high"
 
-#Low input regime factor
-LShare_ipcc[,,"IPCC2006.low"][,,"tropical_moist"]  <- "0.44"
-LShare_ipcc[,,"IPCC2006.low"][,,"tropical_dry"]    <- "0.55--0.61" 
-LShare_ipcc[,,"IPCC2006.low"][,,"temperate_moist"] <- "0.66"
-LShare_ipcc[,,"IPCC2006.low"][,,"temperate_dry"]   <- "0.74"
+# Medium input regime factor
+CShare_ipcc[, , y2006m][, , "tropical_moist"]  <- 0.48
+CShare_ipcc[, , y2006m][, , "tropical_dry"]    <- 0.58
+CShare_ipcc[, , y2006m][, , "temperate_moist"] <- 0.69
+CShare_ipcc[, , y2006m][, , "temperate_dry"]   <- 0.80
 
-LShare_ipcc[,,"IPCC2019.low"][,,"tropical_moist"]  <- "0.76"
-LShare_ipcc[,,"IPCC2019.low"][,,"tropical_dry"]    <- "0.87" 
-LShare_ipcc[,,"IPCC2019.low"][,,"temperate_moist"] <- "0.66--0.67"
-LShare_ipcc[,,"IPCC2019.low"][,,"temperate_dry"]   <- "0.70--0.71"
+CShare_ipcc[, , y2019m][, , "tropical_moist"]   <- 0.83
+CShare_ipcc[, , y2019m][, , "tropical_dry"]     <- 0.92
+CShare_ipcc[, , y2019m][, , "temperate_moist"]  <- 0.70
+CShare_ipcc[, , y2019m][, , "temperate_dry"]    <- 0.77
 
-#High input regime factor
-HShare_ipcc[,,"IPCC2006.high"][,,"tropical_moist"]  <- "0.53"
-HShare_ipcc[,,"IPCC2006.high"][,,"tropical_dry"]    <- "0.60--0.67" 
-HShare_ipcc[,,"IPCC2006.high"][,,"temperate_moist"] <- "0.77"
-HShare_ipcc[,,"IPCC2006.high"][,,"temperate_dry"]   <- "0.83"
+# Low input regime factor
+CShare_ipcc[, , y2006l][, , "tropical_moist"]  <- round(0.48 * 0.92, 2)
+CShare_ipcc[, , y2006l][, , "tropical_dry"]    <- round(0.58 * 0.95, 2)
+CShare_ipcc[, , y2006l][, , "temperate_moist"] <- round(0.69 * 0.92, 2)
+CShare_ipcc[, , y2006l][, , "temperate_dry"]   <- round(0.80 * 0.95, 2)
 
-HShare_ipcc[,,"IPCC2019.high"][,,"tropical_moist"]  <- "0.92"
-HShare_ipcc[,,"IPCC2019.high"][,,"tropical_dry"]    <- "0.96" 
-HShare_ipcc[,,"IPCC2019.high"][,,"temperate_moist"] <- "0.77--0.78"
-HShare_ipcc[,,"IPCC2019.high"][,,"temperate_dry"]   <- "0.79--0.80"
+CShare_ipcc[, , y2019l][, , "tropical_moist"]   <- round(0.83 * 0.92, 2)
+CShare_ipcc[, , y2019l][, , "tropical_dry"]     <- round(0.92 * 0.95, 2)
+CShare_ipcc[, , y2019l][, , "temperate_moist"]  <- round(0.70 * 0.92, 2)
+CShare_ipcc[, , y2019l][, , "temperate_dry"]    <- round(0.77 * 0.95, 2)
 
-CShare_ipcc <- mbind(LShare_ipcc,MShare_ipcc,HShare_ipcc)
+# High input regime factor
+CShare_ipcc[, , y2006h][, , "tropical_moist"]  <- round(0.48 * 1.11, 2)
+CShare_ipcc[, , y2006h][, , "tropical_dry"]    <- round(0.58 * 1.04, 2)
+CShare_ipcc[, , y2006h][, , "temperate_moist"] <- round(0.69 * 1.11, 2)
+CShare_ipcc[, , y2006h][, , "temperate_dry"]   <- round(0.80 * 1.04, 2)
+
+CShare_ipcc[, , y2019h][, , "tropical_moist"]   <- round(0.83 * 1.11, 2)
+CShare_ipcc[, , y2019h][, , "tropical_dry"]     <- round(0.92 * 1.04, 2)
+CShare_ipcc[, , y2019h][, , "temperate_moist"]  <- round(0.70 * 1.11, 2)
+CShare_ipcc[, , y2019h][, , "temperate_dry"]    <- round(0.77 * 1.04, 2)
+
 getNames(CShare_ipcc) <- gsub("_", " ", getNames(CShare_ipcc))
 
+CShare_study  <- as.magpie(readRDS(
+                   paste0(outputdir, "SFC_climatezone+range.rds")))
+
+CShare_median   <- collapseNames(CShare_ipcc[, , y2006m])
+CShare_range    <- CShare_study[, "y2010", c("Q0p05", "Q0p5", "Q0p95")] 
+CShare_median[] <- CShare_study[, "y2010", "Q0p5"] 
+CShare_range    <- round(
+                     add_dimension(
+                       setYears(
+                         CShare_range, 
+                         NULL),
+                       nm = "This Study (5th - 95th percentile).historical 2010 (5th - 95th percentile)", 
+                       add = "model"), 
+                     2)
+CShare_median   <- round(
+                     add_dimension(
+                       setYears(
+                         CShare_median, 
+                         NULL),
+                       nm = "This Study (5th - 95th percentile).historical 2010 (median)", 
+                       add = "model"), 
+                     2)
+
 comparison <- mbind(CShare_ipcc,
-                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),NULL),
-                                  nm="This Study.hist.1990",add="model.input.year"),2),
-                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),NULL),
-                                  nm="This Study.hist.2010",add="model.input.year"),2))
+                    CShare_median)
 
-#print(dimOrder(comparison[,,sort(getNames(comparison,dim=1))], c(2,1)))
+comparison_table  <- as.data.frame(collapseNames(comparison))[c(-1, -2, -3)]
 
-comparison_table  <- data.frame(dcast(as.data.frame(collapseNames(comparison)), 
-                           Data1 + Data2 + Data3 ~ Data4, value.var="Value", 
-                            row.names=levels(as.data.frame(comparison)$Data1)))
+range_table       <- as.data.frame(CShare_range)[c(-1, -2, -3)]                          
+colnames(comparison_table) <- gsub("\\.", " ", colnames(comparison_table))
+colnames(comparison_table)[1:4] <- c("Source", "Management", "Region", "Data")
+colnames(range_table)[1:4]      <- c("Source", "Management", "Region", "Data")
 
-colnames(comparison_table) <- gsub("\\.", " ", colnames(comparison_table))
-colnames(comparison_table)[1:3] <- c("Source", "Input", "Year")
-
-saveRDS(list(comparison_table = comparison_table), paste0(outputdir,"SFC_comparison.rds"))
-
+saveRDS(list(comparison_table = comparison_table), 
+        paste0(outputdir, "SFC_comparison.rds"))
 ```
@@
 getwd()
-table <- readRDS(paste0(outputdir,"/SFC_comparison.rds"))
+table <- readRDS(paste0(outputdir, "/SFC_comparison.rds"))
 
-save  <- kbl(table$comparison_table, latex = T, digit=2)%>%
-            kable_styling(latex_options = c("scale_down","striped"), full_width = F)%>%
-              row_spec(7:8, bold=TRUE)
+save  <- kbl(table$comparison_table, latex = TRUE, digit = 2) %>%
+            kable_styling(latex_options = c("scale_down", "striped"), full_width = FALSE) %>%
+              row_spec(7:8, bold = TRUE)
 
-save_kable(save, file="Output/Images/TableSCF_comparison.png")
+save_kable(save, file = "Output/Images/TableSCF_comparison_new.png")
+table
 ```
+
+```{r table2boxplot, fig.height=3, fig.width=6}
+
+#comparison_table
+range_table <- dcast(range_table, Source + Management + Region ~ Data)
+
+cols <- c("#648FFF", "#DC267F", "#FFB000")[3:1]
+
+SCFTable <- 
+  ggplot() + 
+    geom_linerange(data = range_table,
+                   aes(x = Region, ymin = Q0p05, ymax = Q0p95, color = Source),
+                   position = position_nudge(x = 0.165), size = 0.6) +
+    geom_point(data = comparison_table, 
+               aes(y = Data, x = Region, group = Source, color = Source, shape = Management), 
+               position = position_dodge(width = 0.5), 
+               size = 2) +
+    labs(y = "Stock Change Factor",
+         x = "Climate Zone") +
+    scale_color_manual(values = cols) +
+    guides(shape = guide_legend(order = 2),col = guide_legend(order = 1)) +
+    theme(rect = element_blank(),
+                        axis.title.y = element_text(size = legend_size,
+                                                   margin = margin(t = 0, r = 10, b = 0, l = 0, 
+                                                                   unit = "pt")),
+                        axis.title.x = element_text(size = legend_size,
+                                                   margin = margin(t = 10, r = 0, b = 0, l = 0, 
+                                                                   unit = "pt")))
+
+ggsave(paste0(outputdir, "Images/SCF_comparison.png"), 
+       plot = SCFTable, height = 8, width = 18, units = "cm", dpi = 300)
+```

### 2022-07-20 ResultNotebooks/results_SOCtable.Rmd
update to version before bug check

@@
 cols        <- c("#648FFF", "#DC267F", "#FFB000")[3:1]
+legend_size <- 12
 
@@
     scale_color_manual(values = cols) +
+    scale_x_discrete(labels = c("tropical\nmoist", "tropical\ndry", 
+                                "temperate\ndry",  "temperate\nmoist")) +
     guides(shape = guide_legend(order = 2),col = guide_legend(order = 1)) +
@@
                                       margin = margin(t = 10, r = 0, b = 0, l = 0, 
-                                                                   unit = "pt")))
+                                                                   unit = "pt")),
+          axis.text.x  = element_text(size = legend_size),
+          axis.text.y  = element_text(size = legend_size),
+          legend.title = element_text(size = legend_size),
+          legend.text  = element_text(size = legend_size),
+          legend.spacing.y = unit(0.1, "cm"))
+
+SCFTable
 

### 2022-07-20 SOCbudgetPaper/SOCbudget.tex
update to version before bug check

@@
 \begin{abstract}
-Soil organic carbon (SOC) is one of the largest terrestrial carbon (C) stocks on Earth.
-The first meter of the Earth's soils profile stores three times as much C as the vegetation and twice the amount of C in the atmosphere.
-SOC has been depleted by anthropogenic land-cover change and agricultural management.
+Soil organic carbon (SOC), one of the largest terrestrial carbon (C) stocks on Earth, has been depleted by anthropogenic land-cover change and agricultural management.
 However, the latter has so far not been well represented in global C stock assessments.
@@
 However, accounting for historical management led to \(2.1\,\unit{GtC}\) less (\(0.06\,\unit{GtC\,yr^{-1}}\)) emissions than under the assumption of constant management.
-We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique.
-The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling.
+We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that SOC enhancement by biomass retention may be a promising negative emissions technique.
+The reduced-complexity SOC model may allow to simulate management-induced SOC enhancement also within computationally demanding integrated (land-use) assessment modeling.
 \end{abstract}
@@
 
-Soil organic carbon (SOC), the amount of organic carbon stored in the Earth's soil, constitutes the largest terrestrial organic carbon pool.
-It exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}.
+Soil organic carbon (SOC), the amount of organic carbon stored in the Earth's soil, exceeds the carbon in the atmospheric and vegetation pools multiple times \citep{batjes_total_1996}.
 Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO\textsubscript{2} in the atmosphere \citep{frielingstein_glob_carbon_2020, minasny_soil_2017}.
@@
 
-The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al.~\citeyearpar{pongratz_terminolgy_2014}), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state.
-Sanderman et al.~\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.
+The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al.
+\citeyearpar{pongratz_terminolgy_2014}), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state.
+Sanderman et al.
+\citeyearpar{sanderman_soil_2017} identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter \citep{lal_world_2001}.
 
@@
 Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest \citep{strassmann_simulating_2008, betts_climate_2015}.
-BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO\textsubscript{2} fertilization and N deposition.
-Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management \citep{frielingstein_glob_carbon_2020, houghton_carbon_2012, hansis_relevance_2015, bastos_comparison_2021}.
-
-Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.~][]{bondeau_modelling_2007, lindeskog_implications_2013}.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data.
+BKMs are designed to estimate LUC related emissions and have largely improved in estimating additional emissions from wood harvest and shifting cultivation.
+However, state of the art models do not consider impacts of varying agricultural management \citep{frielingstein_glob_carbon_2020, houghton_carbon_2012, hansis_relevance_2015, bastos_comparison_2021}.
+
+Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance \citep[e.g.][]{bondeau_modelling_2007, lindeskog_implications_2013}.
+Pugh et al.
+\citeyearpar{pugh_simulated_2015} explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with uniform scenario assumptions on management rather than with historical management data.
 They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.
 
-In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons \citep{portmann_mirca2000global_2010}, multiple cropping systems \citep{waha_multiple_2020}, irrigation systems \citep{jagermeyr_water_2015}) or as stylized scenarios \citep[e.g.~][]{pugh_simulated_2015, lutz_simulating_2019}.
-Herzfeld et al.~\citeyearpar{herzfeld_soc_2021} account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC.
-Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.
-
-More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g.~on tillage systems, see \citep{porwollik_generating_2019, prestele_spatially_2018}) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g.~\citep{iizumi_modeling_2019, minoli_modelling_2019}), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
+In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time \citetext{\citealp[e.g.~for growing seasons in][]{portmann_mirca2000global_2010}; \citealp[multiple cropping systems in][]{waha_multiple_2020}; \citealp[irrigation systems in][]{jagermeyr_water_2015}} or as stylized (in the sense of uniform management assumptions) scenarios \citep[e.g.][]{pugh_simulated_2015, lutz_simulating_2019}.
+Herzfeld et al.
+\citeyearpar{herzfeld_soc_2021} account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC.
+Within their stylized future management scenarios under future climate change they find that none of the management aspects considered (residue management, no-tillage) can create a positive SOC stock change on current cropland areas that counteracts the still continuing legacy flux from the initial land-use change.
+
+More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g.~on tillage systems, see \citep{porwollik_generating_2019, prestele_spatially_2018}) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g.
+\citep{iizumi_modeling_2019, minoli_modelling_2019}), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.
 
@@
 Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type \(lu\) specific.
-We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).
+We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and grasslands (referred to as natural vegetation in the following).
+Forage crops are included within cropland, whereas pastures (including mowed meadows (perennials) and rangelands) are assigned to natural vegetation.
+Carbon flows connected to livestock are only considered in this study when they originate from cropland feed sources, while the manure originating from pasture biomass is disregarded, implicitly assuming that this manure is excreted or applied to pastures.
 
@@
 Global distribution of plant-functional types is given by \citep{schaphoff_lpjml4_2018} as well as separation of litter into leaf, fine root and wood litter compartments excluding litter biomass burnt in wild fires.
-Leaf litter parameters are given by Brovkin et al.~\citeyearpar{brovkin_plant-driven_2012}, fine root to leaf litter lignin ratio by Guo et al.~\citeyearpar{guo_leaf_root_decompostion_2021}, lignin content of wood litter by Rahman et al.~\citeyearpar{rahman_lignin_2013} and nitrogen content scaling factors for leaf to fine roots and leaf to wood litter by von Bloh et al.~\citeyearpar{vBloh_lpjml5_2018}.
+Leaf litter parameters are given by Brovkin et al.
+\citeyearpar{brovkin_plant-driven_2012}, fine root to leaf litter lignin ratio by Guo et al.
+\citeyearpar{guo_leaf_root_decompostion_2021}, lignin content of wood litter by Rahman et al.
+\citeyearpar{rahman_lignin_2013} and nitrogen content scaling factors for leaf to fine roots and leaf to wood litter by von Bloh et al.
+\citeyearpar{vBloh_lpjml5_2018}.
 Data sources for all considered carbon inputs as well as for lignin and nitrogen parameterization are listed in Table \ref{tab:datasourceinputs}.
@@
 We calculate area-weighted means for \(till\) and \(wat\) on cropland for each grid cell, using area shares for the different tillage and irrigation practices.
-Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equations 5.0B--5.0F in Ogle et al.~\citeyearpar{ogle_cropland_in_ipcc_2019}.
+Data sources as well as used parameters for the different decay drivers for all land-use types are listed in Table \ref{tab:datasourcedecay}; equations are displayed by equations 5.0B--5.0F in Ogle et al.
+\citeyearpar{ogle_cropland_in_ipcc_2019}.
 
@@
 Similarly, agricultural management data are held constant at the level of 1965 until 1965.
-This approach follows others studies looking on effects of land-use change and management \citep[e.g.~][]{schaphoff_lpjml4_2018-1, herzfeld_soc_2021}.
+We acknowledge that this introduces a bias as agricultural management has changed prior to 1965, but this approach follows others studies on effects of land-use change and management \citep[e.g.][]{schaphoff_lpjml4_2018-1, herzfeld_soc_2021} and is limited by data availability on harvest statistics (and other management effects).
 
@@
 Here, stocks are calculated via stock change factors (\(F^{\mathrm{SCF}}\)) given by the IPCC for the topsoil (0-30 cm) and based on observational data.
+Note that IPCC factors are derived under the assumption that there is a linear change between steady states over 20 years.
 Estimates of \(F^{\mathrm{SCF}}\) are differentiated by crops, management and input systems (here summarized under \(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows.
@@
 with \(T_{c,i}\) being the translation matrix for grid cells \(i\) into corresponding climate zones \(c\).
-For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of Eggleston et al.~\citeyearpar{eggleston_ipcc_2006} and Calvo Buendia et al.~\citeyearpar{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
+For this analysis, we use the default \(F^{\mathrm{SCF}}\) from the Tier 1 method of Eggleston et al.
+\citeyearpar{eggleston_ipcc_2006} and Calvo Buendia et al.
+\citeyearpar{calvo_buendia_ipcc_2019} as a comparison and consistency check for our more detailed Tier 2 steady-state approach.
 
@@
 
-Crop residue production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and key aspects are explained as they play a central role in soil carbon modeling.
+Crop residue production and management is based on a revised methodology of Bodirsky et al.
+\citeyearpar{bodirsky_n2o_2012} and key aspects are explained as they play a central role in soil carbon modeling.
 Starting from crop yield estimates \(Y\) and respective physical crop area \(CA\), we estimate total above-ground \(AGR\) and below-ground \(BGR\) residue biomass (in tonnes) using crop group \(cg\) specific ratios for above-ground residues to harvested biomass \(r^{\mathrm{ag,prod}}_{cg}\) in \((tDM\,ha^{-1})(tDM\,ha^{-1})^{-1}\) above-ground residues to harvested area \(r^{\mathrm{ag,area}}_{cg}\) in \(\unit{tDM\,\unit{ha}^{-1}}\) and below-ground residues to above-ground biomass \(r^{\mathrm{bg}}_{cg}\) in \(\unit{tDM\,\unit{tDM}^{-1}}\) as follows
@@
 Residues biomass therefore increases under-proportionally with rising yields, reflecting a shifting harvest index of higher-yielding breeds.
-Deviating from Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} we use harvested instead of physical crop area (denoted in Eq.
+Deviating from Bodirsky et al.
+\citeyearpar{bodirsky_n2o_2012} we use harvested instead of physical crop area (denoted in Eq.
 \eqref{eq:resbiomass} by MCF described in Sect.
@@
 
-From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020}).
+From our 17 crop groups, we compile four residue groups (straw, high- and low-lignin residues, residues without dual use), of which the first three are taken away from the field for other purposes (see mappingCrop2Residue.csv in Bodirsky et al.
+\citeyearpar{bodirsky_mrcommons_2020}).
 Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017}, that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
@@
 
-To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of Robinson et al.~\citeyearpar{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from Weindl et al.~\citeyearpar{weindl_livestock_2017}.
+To disaggregate country level FAOSTAT livestock production data to half-degree resolution, we use the following rule-based assumptions, drawing from the approach of Robinson et al.
+\citeyearpar{robinson_mapping_2014} and applying feed basket assumptions based on a revised methodology from Weindl et al.
+\citeyearpar{weindl_livestock_2017}.
 We differentiate between ruminant and monogastric systems, as well as extensive and intensive systems.
@@
 
-Manure production and management is based on a revised methodology of Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
+Manure production and management is based on a revised methodology of Bodirsky et al.
+\citeyearpar{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling.
 Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted.
-Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.~\citeyearpar{ogle_cropland_in_ipcc_2019}.
+Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see Calvo Buendia et al.
+\citeyearpar{ogle_cropland_in_ipcc_2019}.
 Depending on the feed system we assume manure to be handled in four different ways:
@@
 Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields.
-For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.~\citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.~\citeyearpar{bodirsky_n2o_2012}.
+For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in Bodirsky et al.
+\citeyearpar{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in Bodirsky et al.
+\citeyearpar{bodirsky_n2o_2012}.
 We limit high application shares at \(10\,\unit{tC\,ha}^{-1}\) to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
@@
 In order to derive a spatial distribution of the three different tillage types specified by the IPCC --- full tillage, reduced tillage and no tillage ---, we assume that all natural land and pastures are not tilled, whereas annual crops are under full and perennials under reduced tillage per default.
-Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historical global gridded tillage data set from Porwollik et al.~\citeyearpar{porwollik_generating_2019}.
-This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO \citeyearpar{fao_aquastat_2016} and half-degree resolution physical crop areas from Hurtt et al.~\citeyearpar{hurtt_harmonization_2020}, applying the methodology of Prowollik et al.~\citeyearpar{porwollik_generating_2019} to identify potential no-tillage grid cells.
+Furthermore, we assume no tillage in cropland cells specified as no tillage cell based on the historical global gridded tillage data set from Porwollik et al.
+\citeyearpar{porwollik_generating_2019}.
+This data set is extended to the period of 1975--2010 by combining country-level data on areas under conservation agriculture from FAO \citeyearpar{fao_aquastat_2016} and half-degree resolution physical crop areas from Hurtt et al.
+\citeyearpar{hurtt_harmonization_2020}, applying the methodology of Prowollik et al.
+\citeyearpar{porwollik_generating_2019} to identify potential no-tillage grid cells.
 
 \hypertarget{sec:scenarios}{%
-\subsection{Management scenario definitions}\label{sec:scenarios}}
+\subsection{Scenario definitions}\label{sec:scenarios}}
+
+To highlight the impact of changing management effects and to assess the sensitivity of the model towards different initialization choices, we perform a set of scenario runs.
+In the following section we outline name and idea of these scenarios \citep[for technical implementation see][]{karstens_mrsoil_2020}.
 
@@
 The \textit{constManagement} combines all three scenarios \textit{constTillage}, \textit{constResidues} and \textit{constManure}.
+
+As outlined in Sect.
+\ref{sec:initsoc} we assume to start in a steady-state SOC stocks for the start year of the spin-up in 1510 followed by a long spin-up period of (\textit{Initial-spinup1510}).
+As some SOC compartments decay over very long timescales, the initialization setting might strongly effect the overall outcome of SOC stocks and changes.
+Thus we conduct two counterfactual scenarios \textit{Initial-natveg} and \textit{Initial-eq}.
+Whereas in \textit{Initial-natveg} we assume SOC stocks to be in a steady-state SOC under potential natural vegetation for all land-use types in 1901, SOC stocks start in their land-use type specific steady state in 1901 for the \textit{Initial-eq} scenario.
+In that way, the two scenarios mark the two extreme cases: In \textit{Initial-eq} all legacy fluxes already appeared in 1901, whereas in \textit{Initial-natveg} all legacy fluxes before 1901 still have to appear.
+We additionally combined the counterfactual scenarios with the \textit{constManagement} scenario.
 \newpage
@@
 
-Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in Karstens \citeyearpar{karstens_model_2020}.
-In the following, the most important results \citep[see][ for post-processing script]{karstens_result_2020} are summarized.
+Detailed results for the spatially explicit global SOC budget including intermediate results on input data from manure and residues as well as SOC stock results for all scenario runs can be found in the data repository from Karstens \citeyearpar{karstens_model_2020}.
+In the following, the most important results \citep[see][ for post-processing scripts]{karstens_result_2020} are summarized.
 
@@
 This corresponds to an average loss rate of \(0.14\,\unit{GtC\,yr^{-1}}\) in comparison to a hypothetical potential natural vegetation (PNV) state.
-Considering our estimate of the global SOC stock of around \(705\,\unit{GtC}\) in the upper \(30\,\unit{cm}\) in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975--2010.
+Considering our estimate of the global SOC stock of around \(705\,\unit{GtC}\) in the upper \(30\,\unit{cm}\) in 1975, global SOC decreased by 0.2 per 1000 per year for the period between 1975 and 2010 in comparison to the PNV state.
 The speed of this SOC loss has decreased towards the end of the modeling period.
+Note that the SOC stock itself --- without comparing it to a a PNV state --- increases during the period between 1975 and 2010 from \(705\,\unit{GtC}\) to \(712\,\unit{GtC}\), which corresponds to overall SOC stock increase of \(0.2\,\unit{GtC\,yr^{-1}}\).
 
@@
 \begin{figure}[h]
-\includegraphics[width=9cm]{../ResultNotebooks/Output/Images/2panelfigure} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile.
+\includegraphics[width=6.3cm]{../ResultNotebooks/Output/Images/SpatialDebtFigure} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile.
 Panel (a) shows global $\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types.
-Computing the difference between the $\Delta SOC$ estimate for 2010 and for 1975 (b) depicts areas of soil depletion (SOC debt increase, red) and net-sequestration (SOC debt decline, blue).}\label{fig:SOCdebt}
+Computing the difference between the $\Delta SOC$ estimate for 2010 and for 1975 (b1) depicts areas of SOC depletion (SOC debt increase, red) and SOC accumulation (SOC debt decline, blue).
+Panel (b2) shows the of land-use change (LUC) induced change of $\Delta SOC$ between 1975 and 2010, whereas panel (b3) depicts the change due to changing agricultural management (MAN).}\label{fig:SOCdebt}
 \end{figure}
@@
 \ref{fig:SOCdebt}(a)) and its change from 1975 to 2010 (Fig.
-\ref{fig:SOCdebt}(b)) reveals areas of SOC debt decline and increase.
+\ref{fig:SOCdebt}(b1)) reveals areas of SOC debt decline and increase.
 Regions with large cropland expansion (e.g.~Brazil, Southeast Asia, Canada) continue to lose SOC, whereas regions with cropland reduction (and thus SOC restoration) or with accumulating cropland SOC can be found e.g.~in highly productive areas of Europe and Central USA.
 
-\hypertarget{carbon-flows-in-the-agricultural-system}{%
-\subsection{Carbon flows in the agricultural system}\label{carbon-flows-in-the-agricultural-system}}
-
-\begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in $\unit{MtC}$).
+\hypertarget{carbon-flows-in-the-cropland-system}{%
+\subsection{Carbon flows in the cropland system}\label{carbon-flows-in-the-cropland-system}}
+
+\begin{figure}[h]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the cropland system for the year 2010 (in $\unit{MtC}$).
 Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand.
-After accounting for losses within the agricultural system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues.
+After accounting for losses within the cropland system, there are three major C inputs to cropland SOC: manure, above- and below-ground residues.
 Large parts of C, however, are mineralized on the field before entering the soil.
-Additionally, C is transferred to and from the global agricultural soil stock via land-use change between cropland and natural vegetation.
+Additionally, C is transferred to and from global cropland soil stock via land-use change between cropland and natural vegetation.
 Finally, SOC is mineralized and flows back to the atmosphere.}\label{fig:FlowFig}
@@
 
-C is sequestered from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvested organ, above- and below-ground residues).
+C is removed from the atmosphere via plant growth and allocated to different plant parts, which we aggregate to three pools (harvested organ, above- and below-ground residues).
 Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (\(729\,\unit{MtC}\) in 2010) are directly returned to the field.
@@
 According to our model results, land-use change dynamics led to a C transfer from natural vegetation to cropland of \(257\,\unit{MtC}\) in 2010.
-The agricultural system receives \(4585\,\unit{MtC}\) assimilated by crop plants and releases \(3554\,\unit{MtC}\) mostly through respiration.
+The cropland system receives \(4585\,\unit{MtC}\) assimilated by crop plants and releases \(3554\,\unit{MtC}\) mostly through respiration.
 Accounting for SOC transfer and decomposition, the net SOC decrease of global cropland is around \(33\,\unit{MtC}\) for the year 2010.
@@
 
+Using the constManagement results, that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results.
+The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state while calculating \(\Delta SOC\).
+The increasing SOC debt on global cropland are primarily caused by LUC (red areas in Fig.
+\ref{fig:SOCdebt}(b2)).
+Deteriorated management also contributed to increasing SOC debt in parts of Sub-saharan Africa and Central Asia.
+In contrast, agricultural management has led to an decrease in SOC debt in the USA, Europe, as well as in parts of China and India (blue areas in Fig.
+\ref{fig:SOCdebt}(b3)), which is not visible in the total \(\Delta SOC\) change as LUC was happening at the same time.
+
+Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks (see Fig.
+\ref{fig:SOCinit} and Sect.
+\ref{sec:scenarios}) with around \(2.15\,\unit{GtC}\) difference in SOC debt between the histManagement and the constManagement scenario.
+However, the SOC debt and SOC debt change varies with the different initialization choices.
+Whereas the default assumption (Initial-spinup1510) shows a \(\Delta SOC\) of \(39.6\,\unit{GtC}\) for the year 2010, the Initial-natveg scenario with high legacy fluxes to come only has a \(\Delta SOC\) of \(33.3\,\unit{GtC}\) and the Initial-eq scenarios with only little legacy fluxes left already a \(\Delta SOC\) of \(50.7\,\unit{GtC}\).
+
 \hypertarget{sec:validate}{%
@@
 To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors \(F^{\mathrm{SCF}}\) of 2006 \citep{lasco_cropland_in_ipcc_2006} and their refinements in 2019 \citep{ogle_cropland_in_ipcc_2019}.
-Both estimates are based on measurement data for cropland (see Table \ref{tab:SCFglo}).
-To allow for comparison, we aggregate our stock change factors weighted by grid-level \(F^{\mathrm{SCF}}\) cropland area to derive average factors for the four IPCC climate zones (Fig.
+Both estimates are based on measurement data for cropland (see Fig.
+\ref{fig:SCFglo}).
+To allow for comparison, we aggregate our stock change factors weighted by grid-level cropland area to derive median factors for the four IPCC climate zones (Fig.
 \ref{fig:CLIMzone}).
-
-\begin{table}[ht]
-\centering
-\caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.} 
-\label{tab:SCFglo}
-\begin{tabular}{rlllllll}
-  \hline
- & Source & Input & Year & tropical moist & tropical dry & temperate dry & temperate moist \\ 
-  \hline
-1 & IPCC2006 & low & invariant & 0.44 & 0.55--0.61 & 0.74 & 0.66 \\ 
-  2 & IPCC2006 & medium & invariant & 0.48 & 0.58--0.64 & 0.80 & 0.69 \\ 
-  3 & IPCC2006 & high & invariant & 0.53 & 0.60--0.67 & 0.83 & 0.77 \\ 
-  4 & IPCC2019 & low & invariant & 0.76 & 0.87 & 0.70--0.71 & 0.66--0.67 \\ 
-  5 & IPCC2019 & medium & invariant & 0.83 & 0.92 & 0.76--0.77 & 0.69--0.70 \\ 
-  6 & IPCC2019 & high & invariant & 0.92 & 0.96 & 0.79--0.80 & 0.77--0.78 \\ 
-  7 & This Study & hist & 1975 & 0.48 & 0.59 & 0.64 & 0.59 \\ 
-  8 & This Study & hist & 2010 & 0.5 & 0.64 & 0.64 & 0.59 \\ 
-   \hline
-\end{tabular}
-\end{table}
+Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking the time scale of the stock change.
+
+\begin{figure}[h]
+\includegraphics[width=15cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
+IPCC factors vary a bit with different types of management (low, medium, high).
+The results of the study span over a wide range -- indicated by the span between the 5th and 95th percentile -- but not overlapping with the IPCCs 2019 factors for tropical moist.
+The median for most regions is considerably lower than the IPCC factors.}\label{fig:SCFglo}
+\end{figure}
 
 Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC.
-For the tropical regions the IPCC factors changed notably from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} to the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
-Our results are in good agreement with the 2006 IPCC factors.
-Modeled \(F^{SCF}\) have increased or stayed constant for all climate zones over time (1975-2010).
+For the tropical regions the IPCC factors increased notably from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} to the update in 2019 \citep{ogle_cropland_in_ipcc_2019} due to the inclusion of more data points.
+Our results span over a broad range due to the different ages of the cropland but also due to different agricultural management practices within climate regions.
 
@@
 \begin{figure}[h]
-\includegraphics[width=11cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area.
+\includegraphics[width=7cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area.
 The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \citep{schaphoff_lpjml4_2018-1} and estimates from \citep{sanderman_soil_2017}.
@@
 The global estimates of the total SOC stock of the upper 30 cm from this study are in the middle of the wide range of other modeled or observation-based estimates.
-SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, whereas SoilGrids 2.0 \citep{poggio_soilgrids_2021} marks the lower end.
 Regional results (Fig.
 \ref{fig:SOCreg}) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.
+SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, since they include the litter horizon on top of the soil, that might dominate especially polar and boreal soils.
+SoilGrids 2.0 \citep{poggio_soilgrids_2021} however, excludes litter C and thus marks the lower end for particularly northern regions.
+For the same reason it is also more comparable to our results, which do not account for litter C as well.
 
@@
 
-We correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements (for data base see appendix of \citep{sanderman_soil_2017}).
-
-\begin{figure}[h]
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010} \caption{Correlation between modeled and measured $SOC$ stocks.
-Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results.
-Both cropland and areas with natural vegetation tend to be lower in our results than in the point measurements.}\label{fig:SOCpoint}
-\end{figure}
+In Fig.
+\ref{fig:SOCpoint} we correlate our SOC results for natural vegetation and cropland in 2010 with literature values from point measurements \citep[for data base see appendix of][]{sanderman_soil_2017}.
+The goodness of the fit is very low with an \(R^2\) of \(0.13\).
+Individually the correlations are even lower with a \(R^2\) of \(0.09\) for cropland and \(0.08\) for areas of natural vegetation.
+This point to the fact that differences between land-use type SOC stocks could be better matched than the spatial pattern of the rather small point measurement data base.
+Due to the low number of small-scale measurements, statistical properties of the point data variability are not derived and thus, could not be used to improve the point-to-grid-cell comparison \citep[see][]{rammig_generic_2018}.
 
@@
 Estimates of SOC stocks under natural vegetation influence our modeled results for cropland, which has been converted from natural vegetation at some point in time.
-We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation.
+As the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019} is not specifically parameterized for natural vegetation it is important to evaluate its suitability to produce reasonable results in that domain as well at least comparable to other modeling approaches.
+We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.\ref{fig:SOCLPJmL4}).
 Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.
 With our focus on cropland SOC dynamics, we compare only cells with more than \(1000\,\unit{ha}\) of cropland (capturing 99.9\% of global cropland area).
-
-\begin{figure}[h]
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv} \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010.
-The grey lines indicate the 1:1 line.}\label{fig:SOCLPJmL4}
-\end{figure}
-
 Spatial correlations of PNV SOC stock values are high (global \(R2=0.81\)), especially for dry climate zones (Fig.
@@
 \end{figure}
+
 \newpage
@@
 Our spatially explicit time series data set of agricultural management is based on country-specific FAO production and cropland statistics \citep{faostat_faostat_2016} as well as 0.5 degree land-use data from LUH2 \citep{hurtt_harmonization_2020}.
-Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the agricultural system (\ref{fig:FlowFig}) using a mass balance approach from the IPCC guidelines Vol.
-4 \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and other auxiliary data sets \citep[e.g.~][]{porwollik_generating_2019}.
+Starting from these two sources, we derive a harmonized and consistent data set for the major C flows within the cropping system (\ref{fig:FlowFig}) using a mass balance approach from the IPCC guidelines Vol.
+4 \citep{eggleston_ipcc_2006, calvo_buendia_ipcc_2019} and other auxiliary data sets \citep[e.g.][]{porwollik_generating_2019}.
 
@@
 For most of the parameters used in our management estimates no uncertainty estimates exist.
-This is why, in our view, most of the uncertainty with respect to the impacts of SOC management is included in the management data itself, and especially in the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data \citep{faostat_faostat_2016, weindl_livestock_2017}.
+This is why, in our view, a large part of the uncertainty with respect to the impacts of SOC management --- next to the parametric and structural uncertainty of the soil model --- is included in the management data itself.
+This is especially the case for the residue and manure production and application numbers, as these are only indirectly derived from crop and livestock production, feed and area data \citep{faostat_faostat_2016, weindl_livestock_2017}.
 The uncertainty of recycling shares adds on top of the uncertain total numbers of manure and residue biomass.
@@
 \ref{fig:SOCscen}).
-As pointed out by Keel et al.~\citeyearpar{keel_large_2017} and Smith et al.~\citeyearpar{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
-Keel et al.~\citeyearpar{keel_large_2017} question whether below-ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
+As pointed out by Keel et al.
+\citeyearpar{keel_large_2017} and Smith et al.
+\citeyearpar{smith_how_2020}, carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \ref{tab:c2dm} for coefficients used in this study).
+Keel et al.
+\citeyearpar{keel_large_2017} question whether below-ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains.
 Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources.
@@
 
-The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Table \ref{tab:SCFglo}) as well as by the time trend of the SOC debt (see Fig.
+The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Fig.
+\ref{fig:SCFglo}) as well as by the time trend of the SOC debt (see Fig.
 \ref{fig:SOCmaps}(a)).
@@
 This is necessary in order to represent SOC losses under land-use change in a dynamic way, as this is an important driver of SOC dynamics.
-The comparison of simulated PNV data with LPJmL4 shows the model's substantial capability in reproducing PNV SOC stocks (Fig.
+The comparison of simulated PNV data with LPJmL4 shows the model's capability in reproducing PNV SOC stocks (Fig.
 \ref{fig:SOCLPJmL4}).
+Concurrently, the point-data comparison (see Fig.
+\ref{fig:SOCpoint}) shows low correlation for PNV, however also for cropland sites.
+This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison \citep[see][]{rammig_generic_2018}.f
 
@@
 \ref{fig:SOCglo}, the only conclusion we can draw from this is that our result is within a plausible range.
+
+Additionally, we find our \(0.2\,\unit{GtC\,yr^{-1}}\) SOC stock change within the period between 1975 and 2010 for the first 30 cm of the soil profile at the upper end of estimates comparing it to estimates of Ito et al.
+\citeyearpar{ito_soil_2020} of \(0.18\,\pm\,0.41\,\unit{GtC\,yr^{-1}}\) within the period between 1850 and 2014 for whole soil profile.
+This might be not surprising as the CO\textsubscript{2} effect is most likely stronger and land-use change effects weaker within our later and shorter modeling period compared to a mean value of the period between 1850 and 2014.
+The large uncertainty within estimates of SOC stock and its changes \citep{ito_soil_2020} again points to the large structural uncertainty within SOC modeling.
+
 To avoid a strong impact of natural land representation and its uncertainties on our results, we focus on SOC changes on cropland.
@@
 Natural SOC estimates only influence results when natural land is converted to cropland.
+Moreover, the temporal dynamic of the SOC debt and stock change factors is not (or only to a small degree) altered by climatic or atmospheric effects on SOC stocks, as they cancel out by taking the difference (for the SOC debt) and ratio (for the stock change factors) of cropland SOC and SOC under hypothetical PNV conditions.
+
+The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes effect these values strongly (see Fig.
+\ref{fig:SOCinit}).
+According to our sensitivity analysis the SOC debt depending on the initialization choice has been between \(33.3\,\unit{GtC}\) and \(50.7\,\unit{GtC}\), with our best guess at \(39.6\,\unit{GtC}\).
+Concurrently, our results indicate, that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
 
@@
 
-In our model, erosion should however only affect the spatial pattern but not the aggregate SOC pool.
-As pointed out by Doetterl et al.~\citeyearpar{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions \citep{wang_human-induced_2017}.
+In our model, erosion might however only affect the spatial pattern but not the aggregate SOC pool.
+As pointed out by Doetterl et al.
+\citeyearpar{doetterl_erosion_2016}, the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions \citep{wang_human-induced_2017}.
+Concurrently, other studies have claimed erosion moves SOC into aquatic reservoirs \citep{zhang_simulating_2020} and thus changing total global terrestrial SOC stock.
 Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focused especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
@@
 
-In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Table \ref{tab:SCFglo}) for almost all regions.
+In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Fig.
+\ref{fig:SCFglo}) for almost all regions.
 Tropical soils might suffer from low C input rates due to large yield gaps (Global Yield Gap and Water Productivity Atlas.
 Available URL: www.yieldgap.org (accessed on: 03/01/2022)) and high shares of residue removal and burning in lower-income countries \citep{smil_crop_1999, williams_influence_1997, jain_emission_2014}.
-Yet, even when comparing our etimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values, while it shows very good agreement with the older default values from IPCC \citeyearpar{lasco_cropland_in_ipcc_2006}.
-Don et al.~\citeyearpar{don_impact_2011} estimated SOC losses for tropical soils of around 25\% on average corresponding to a stock change factor of 0.75, but also reported a wide range of measured SOC changes from -80\% to +58\%.
-Fujisaki et al.~\citeyearpar{fujisaki_forest_2015} however found much lower loss rates of around 9\%, attributing the difference to the different time period length since the conversion to cropland.
+Yet, even when comparing our estimates to the low-input stock change factors of the IPCC, our SOC loss is roughly twice as large as the revised 2019 IPCC default values \citeyearpar{ogle_cropland_in_ipcc_2019}, while it shows good agreement with the older default values from IPCC \citeyearpar{lasco_cropland_in_ipcc_2006}.
+However, the revised estimates of the IPCC included much more and more recent data points, calling for a closer look on causes for the large deviations between our results and the refined Tier 1 factors.
+On the one hand, our approach does not account for unharvested carbon inputs from weeds or biomass cover on short-term fallows.
+Shifting agriculture with fallow periods might be prominent in tropical regions.
+While long-term fallow land (\textgreater4 years) is excluded from FAOSTAT as cropland, short-term fallow is not.
+Thus, our carbon inputs for these areas might be underestimated, leading to too low stock change factors.
+On the other hand, older studies by Don et al.
+\citeyearpar{don_impact_2011} estimated SOC losses for tropical soils of around 25\% on average corresponding to a stock change factor of 0.75, but also reported a wide range of measured SOC changes from -80\% to +58\%.
+Fujisaki et al.
+\citeyearpar{fujisaki_forest_2015} however found much lower loss rates of around 9\%, attributing the difference to the different time period lengths since the conversion to cropland.
 As our results do not specifically account for cropland age and most of the cropland is older than 20 years (as assumed for the default IPCC Tier 1 stock change factors) our stock change factors have to be lower by definition following the steady-state assumption that cropland will continue to approach a new equilibrium.
@@
 This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas.
-Second, changes to the subsoil due to tillage are neglected.
-As Powlson et al.~\citeyearpar{powlson_limited_2014} have shown, the subsoil can be make a large difference in evaluating total SOC losses or gains for no-tillage systems.
+Second, changes to the subsoil are neglected, which is most important for tillage effects.
+Other management practices might be more equally effecting top- and subsoil as they do not directly the relocates carbon vertically.
+As Powlson et al.
+\citeyearpar{powlson_limited_2014} have shown, the subsoil can be make a large difference in evaluating total SOC losses or gains for no-tillage systems.
 No-tillage effects may seem larger than they actually are if only topsoil is considered.
 SOC transfers to deeper soil layers under tillage might enhance subsoil SOC compared to no-till practices.
+However, the effect of no-till to the subsoil as well as its overall importance as a SCS measure is still debated \citep{ogle_climate_2019}.
 Finally, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.
@@
 The analysis of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap.
-There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period \citep[e.g.~][]{frielingstein_glob_carbon_2020}.
+There is broad consensus that land conversion to cropland has caused substantial C emissions over the historical period \citep[e.g.][]{frielingstein_glob_carbon_2020}.
 There is uncertainty with respect to the overall size of these emissions from different methods and reference points and with respect to the contribution of cropland and agricultural management to these emissions.
@@
 Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions \citep[see][]{pugh_simulated_2015}, we find that global cropland has acted as a emissions source since 1975.
+Comparing our SOC loss rate (the change of SOC debt) of \(0.14\,\unit{GtC\,yr^{-1}}\) to estimates on land-use change induced emissions of \(2.0\,\pm\,1.0\,\unit{GtC\,yr^{-1}}\) (sum of `Bookkeeping LULCC emissions' and `Loss of additional sink capacity' for the years 2009--2018 in Gasser et al.
+\citeyearpar{gasser_historical_2020}), we find SOC emissions of the first 30 cm of the soil profile to be a minor contributor to overall land-use change induced emissions.
 Annual C loss rates of 0.2 per 1000 C still have the opposite trend as the promoted 4 per 1000 C sequestration rate target \citep{minasny_soil_2017}.
@@
 Yet our study also shows the substantial impact of changing management on the development of SOC debt (Fig.
+\ref{fig:SOCdebt} and Fig.
 \ref{fig:SOCscen}).
 
-According to Sanderman et al.~\citeyearpar{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
+According to Sanderman et al.
+\citeyearpar{sanderman_soil_2017}, the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands.
 Our estimate of 39.6 GtC in 2010 for cropland debt is thus twice as high as their estimate.
-However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.~\citeyearpar{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
-
-Furthermore, Sanderman et al.~\citeyearpar{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
-Pugh et al.~\citeyearpar{pugh_simulated_2015} considered management effects like tillage and incorporation of residues in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
-Their study moreover concludes that yield gains (by 18\% in their simulations) do not lead to a substantial decline in SOC debt (less than 1\% change).
-Historical yield increases, however, are often estimated to be well above 50\% \citep{pellegrini_crop_2018, ray_recent_2012, rudel_agricultural_2009}.
-While we find substantially larger effects of productivity gains than the 1\% reported by Pugh et al.~\citeyearpar{pugh_simulated_2015}, this is not sufficient to compensate SOC losses from moderate global cropland expansion of around 11\% between 1974 and 2010.
+However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al.
+\citeyearpar{sanderman_soil_2017} pointed out that their results might be conservatively low compared to experimental results.
+
+Furthermore, Sanderman et al.
+\citeyearpar{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management.
+Pugh et al.
+\citeyearpar{pugh_simulated_2015} considered management effects like tillage and incorporation of residues in stylized and static scenarios only, so that they could not account for historical management effects on SOC dynamics.
+Their study moreover concludes that yield gains (by 18\% in their simulations) do not lead to a substantial decline in SOC debt (less than 1\% SOC increase).
+Historical yield increases, however, are estimated to be well above 50\% \citep{pellegrini_crop_2018, ray_recent_2012, rudel_agricultural_2009} and often lead to an increase in below- and above-ground residue biomass inputs to the soil.
+While we find substantially larger SOC increase in response to productivity gains than the 1\% reported by Pugh et al.
+\citeyearpar{pugh_simulated_2015}, this is not sufficient to compensate SOC losses from global cropland expansion of around 11\% between 1974 and 2010.
 
@@
 In contrast, process-based studies \citep{pugh_simulated_2015, herzfeld_soc_2021} often lack data on relevant management aspects that drive production increases.
-Herzfeld et al.~\citeyearpar{herzfeld_soc_2021} also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
+Herzfeld et al.
+\citeyearpar{herzfeld_soc_2021} also consider historical management trends for fertilizer and manure inputs as well as on residue removal rates and tillage systems, but cannot reproduce the substantial increase in agricultural productivity over the last decades.
 Still, they find that compared to no-tillage systems, residue management has much larger potential to affect the strength of their projected future global cropland SOC decline.
@@
 \ref{fig:SOCscen}).
-
-Elliott et al.~\citeyearpar{elliott_management_trends_2018} show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, and genetic traits such as kernel number and radiation use efficiency.
+In line with this, Dangal et al.
+\citeyearpar{dangal_improving_2022} finds that no-tillage has only minor impacts on SOC dynamics across parts of the US.
+
+Elliott et al.
+\citeyearpar{elliott_management_trends_2018} show that yield trends in the USA can be reproduced by models, but require information on inputs that are not available at the global scale, such as annual data on sowing dates, planting densities, and genetic traits such as kernel number and radiation use efficiency.
 As such, it will remain challenging for process-based DGVMs to capture the trend of agricultural productivity on cropland SOC dynamics.
@@
 As shown in Fig.
-\ref{fig:FlowFig}, approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year).
-However, even larger losses occur at the end of the food supply chain (1.2 GtC year), at the soil surface (1.4 GtC), during residue burning (0.3 Gt C) and with manure management (0.2).
+\ref{fig:FlowFig}, the annual SOC respiration (1.3 GtC per year) is slightly above one quarter of total annual net C uptake by crops (4.6 GtC per year).
+C compounds have to be respired by soil organisms to maintain basic soil functions and regulate the nutrient cycle, which often leaves limited options to decrease C losses via SOC respiration \citep{janzen_soil_2006}.
+However, similar C losses occur at the end of the food supply chain (1.2 GtC per year), at the soil surface (1.5 GtC per year), and smaller but still considerable during residue burning (0.2 GtC per year) and within animal waste management systems (0.2 GtC per year).
 Improved management could include, firstly, a circular flow from the food supply chain back to soils.
@@
 Our results demonstrate that historical changes in agricultural management have shaped the SOC debt on cropland.
-It is thus necessary to explicitly consider agricultural management in a dynamic manner in global carbon assessments and models, especially when exploring climate mitigation pathways with so-called land-based solutions \citep[e.g.~][]{popp_land-use_2016}.
+It is thus necessary to explicitly consider agricultural management in a dynamic manner in global carbon assessments and models, especially when exploring climate mitigation pathways with so-called land-based solutions \citep[e.g.][]{popp_land-use_2016}.
 That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data.
@@
 \section{Figures and tables in appendices}
+
 \subsection{Methods}
@@
 \begin{figure}[h]
-  \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
+\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones.png} 
 \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data.
@@
 \tophline
-
          &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
@@
 
-\begin{figure}[H]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
+\begin{figure}[h]
+\includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
+\caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510).
+On the other hand, assuming all SOC to be in land-use specific steady-state already in 1901 (Initial-eq) leads to the opposite effect of an already higher $\Delta SOC$ in 1975 and a less steeper increase till 2010.
+However, the difference between the dynamic historical management assumption (histManagement) compared to constant management assumption from 1975 on (constManagement) are of a similar size for all initialization choices.}
+\label{fig:SOCinit}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
 \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
@@
 \label{fig:SOCreg}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
+\caption{Correlation between modeled and measured $SOC$ stocks.
+Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results.
+Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
+\label{fig:SOCpoint}
+\end{figure}
+
+ \begin{figure}[h]
+\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
+\caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
+\label{fig:SOCLPJmL4}
 \end{figure}

### 2022-07-20 SOCbudgetPaper/discussion.Rmd
update to version before bug check

@@
 
-The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Table \@ref(tab:SCFglo)) as well as by the time trend of the SOC debt (see Fig.
+The substantial impact of changing management practices through time is indicated by the development of our estimated stock change factors (see Fig.
+\@ref(fig:SCFglo)) as well as by the time trend of the SOC debt (see Fig.
 \@ref(fig:SOCmaps)(a)).
@@
 \@ref(fig:SOCpoint)) shows low correlation for PNV, however also for cropland sites.
-This might also point to the fact that SOC stocks can vary at field and local scale considerable and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f
+This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f
 
@@
 
-In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Table \@ref(tab:SCFglo)) for almost all regions.
+In comparison with default stock change factors of the IPCC guidelines, our model estimates a stronger decline of SOC stocks (Fig.
+\@ref(fig:SCFglo)) for almost all regions.
 Tropical soils might suffer from low C input rates due to large yield gaps (Global Yield Gap and Water Productivity Atlas.
@@
 While long-term fallow land (>4 years) is excluded from FAOSTAT as cropland, short-term fallow is not.
-Thus, our carbon inputs for these areas might be underestimated, leading to too low stock changes factors.
+Thus, our carbon inputs for these areas might be underestimated, leading to too low stock change factors.
 On the other hand, older studies by Don et al.
@@
 Fujisaki et al.
-[-@fujisaki_forest_2015] however found much lower loss rates of around 9%, attributing the difference to the different time period length since the conversion to cropland.
+[-@fujisaki_forest_2015] however found much lower loss rates of around 9%, attributing the difference to the different time period lengths since the conversion to cropland.
 As our results do not specifically account for cropland age and most of the cropland is older than 20 years (as assumed for the default IPCC Tier 1 stock change factors) our stock change factors have to be lower by definition following the steady-state assumption that cropland will continue to approach a new equilibrium.
@@
 Defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on cropland as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], we find that global cropland has acted as a emissions source since 1975.
+Comparing our SOC loss rate (the change of SOC debt) of $0.14\,\unit{GtC\,yr^{-1}}$ to estimates on land-use change induced emissions of $2.0\,\pm\,1.0\,\unit{GtC\,yr^{-1}}$ (sum of 'Bookkeeping LULCC emissions' and 'Loss of additional sink capacity' for the years 2009--2018 in Gasser et al.
+[-@gasser_historical_2020]), we find SOC emissions of the first 30 cm of the soil profile to be a minor contributor to overall land-use change induced emissions.
 Annual C loss rates of 0.2 per 1000 C still have the opposite trend as the promoted 4 per 1000 C sequestration rate target [@minasny_soil_2017].

### 2022-07-20 SOCbudgetPaper/methods.Rmd
update to version before bug check

@@
 Similarly, agricultural management data are held constant at the level of 1965 until 1965.
-We acknowledge that this introduces a bias as agricultural management has changed prior to 1965, but this approach follows others studies looking on effects of land-use change and management [e.g.
+We acknowledge that this introduces a bias as agricultural management has changed prior to 1965, but this approach follows others studies on effects of land-use change and management [e.g.
 @schaphoff_lpjml4_2018-1; @herzfeld_soc_2021] and is limited by data availability on harvest statistics (and other management effects).

### 2022-07-20 SOCbudgetPaper/results.Rmd
update to version before bug check

@@
 Computing the difference between the $\\Delta SOC$ estimate for 2010 and for 1975 (b1) depicts areas of SOC depletion (SOC debt increase, red) and SOC accumulation (SOC debt decline, blue).
-Panel (b2) shows the of land-use change (LUC) induced change of ∆SOC between 1975 and 2010, whereas panel (b3) depicts the change due to changing agricultural management (MAN)."
+Panel (b2) shows the of land-use change (LUC) induced change of $\\Delta SOC$ between 1975 and 2010, whereas panel (b3) depicts the change due to changing agricultural management (MAN)."
 ```
 
-```{r SOCdebt, out.width = "9cm", fig.cap = captionSOCdebt, echo=FALSE, fig.pos="h"}
+```{r SOCdebt, out.width = "6.3cm", fig.cap = captionSOCdebt, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 To evaluate our modeled SOC stocks and stock changes under agricultural management, we compare our results to the default IPCC stock change factors $F^{\mathrm{SCF}}$ of 2006 [@lasco_cropland_in_ipcc_2006] and their refinements in 2019 [@ogle_cropland_in_ipcc_2019].
-Both estimates are based on measurement data for cropland (see Table \@ref(tab:SCFglo)).
-To allow for comparison, we aggregate our stock change factors weighted by grid-level cropland area to derive average factors for the four IPCC climate zones (Fig.
+Both estimates are based on measurement data for cropland (see Fig.
+\@ref(fig:SCFglo)).
+To allow for comparison, we aggregate our stock change factors weighted by grid-level cropland area to derive median factors for the four IPCC climate zones (Fig.
 \@ref(fig:CLIMzone)).
@@
 
-```{r SCFglo, echo=FALSE, results = 'asis'}
-table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
-library(xtable)
-print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
-              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \\citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \\citep{ogle_cropland_in_ipcc_2019}."), comment=FALSE, caption.placement = "top")
+```{r captionSCFglo, echo=FALSE}
+captionSCFglo <- "$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \\citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \\citep{ogle_cropland_in_ipcc_2019}.
+IPCC factors vary a bit with different types of management (low, medium, high).
+The results of the study span over a wide range -- indicated by the span between the 5th and 95th percentile -- but not overlapping with the IPCCs 2019 factors for tropical moist.
+The median for most regions is considerably lower than the IPCC factors."
 ```
+
+```{r SCFglo, out.width = "15cm", fig.cap = captionSCFglo, echo=FALSE, fig.pos="h"}
+# All defaults
+knitr::include_graphics("../ResultNotebooks/Output/Images/SCF_comparison.png")
+```
+
 Stock change factors for temperate climate zones of this study are lower than the default values of the IPCC.
 For the tropical regions the IPCC factors increased notably from the guidelines in 2006 [@lasco_cropland_in_ipcc_2006] to the update in 2019 [@ogle_cropland_in_ipcc_2019] due to the inclusion of more data points.
-Our results are in good agreement with the 2006 IPCC factors.
-Modeled $F^{SCF}$ have increased or stayed constant for all climate zones over time (1975-2010).
+Our results span over a broad range due to the different ages of the cropland but also due to different agricultural management practices within climate regions.
 
@@
 
-```{r SOCglo, out.width = "11cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
+```{r SOCglo, out.width = "7cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
 # All defaults

### 2022-07-20 ResultNotebooks/results_SOCscenario.Rmd
update to version before bug check
update to final version - 2. resubmit

@@
 theme_set(theme_bw())
-legend_size <- 10
+legend_size <- 12
 scen_cols   <- c("#648FFF", "#785EF0", "#DC267F")
@@
 
-startY <- 1965
+startY <- 1975
 
@@
                                    size = legend_size),
+        axis.text.y  = element_text(size = legend_size),
+        legend.title = element_text(size = legend_size),
+        legend.text  = element_text(size = legend_size),
         legend.position = "right", legend.direction = "vertical")
@@
 ggsave(paste0(outputdir, "Images/scenario_init.png"), 
-       plot = initSOCemis, width = 12, height = 8, dpi = 300, units = "cm")
+       plot = initSOCemis, width = 15, height = 10, dpi = 300, units = "cm")
 ```

### 2022-07-20 SOCbudgetPaper/SOCbudget.Rmd
update to final version - 2. resubmit

@@
 appendix: |
+  \section{Figures and tables in appendices}
 
-  \section{Figures and tables in appendices}
   \subsection{Methods}
+
     ```{=latex}
     \begin{figure}[h]
-    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
+    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones.png} 
     \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data.
@@
   \tophline
-
            &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
@@
 
-  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->
-
   \subsection{Results}
@@
     ```{=latex}
-  \begin{figure}[H]
+    \begin{figure}[h]
     \includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
@@
     ```{=latex}
-  \begin{figure}[H]
-    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
+    \begin{figure}[h]
+    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
     \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
@@
     ```{=latex}
-  \begin{figure}[H]
+    \begin{figure}[h]
     \includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
@@
     ```{=latex}
-  \begin{figure}[H]
+     \begin{figure}[h]
     \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}

### 2022-07-20 SOCbudgetPaper/SOCbudget.tex
update to final version - 2. resubmit

@@
 This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we define the absolute and relative \(SOC\) changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\), that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
+For our study, we define the absolute and relative \(SOC\) changes in relation to a potential natural state \(SOC^{\mathrm{pnv}}\) under the same climatic conditions in grid cell \(i\) at time \(t\) that is based on the natural vegetation \(SOC\) calculations as defined above without accounting for land conversion from cropland at any time.
 The absolute changes \(\Delta SOC\) and relative changes \(F^{\mathrm{SCF}}\) are thus given by
@@
 \citeyearpar{bodirsky_mrcommons_2020}).
-Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017}, that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
+Residue feed demand for five different livestock groups is based on country-specific feed baskets \citep[see][]{weindl_livestock_2017} that differentiate between the residue groups and take available \(AGR\) biomass as well as livestock productivity into account.
 We estimate a material-use share for the straw-residue group of 5\% and a fuel-share of 10\% for all used residue groups in low-income countries.
@@
 \citeyearpar{bodirsky_n2o_2012}.
-We limit high application shares at \(10\,\unit{tC\,ha}^{-1}\) to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
+We limit high application shares at \(10\,\unit{tC\,ha}^{-1}\) to correct for outliers that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
 
@@
 \ref{fig:SOCmaps}(d)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome by improved agricultural practices.
+They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation that may hold potential to be overcome by improved agricultural practices.
 Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation.
@@
 
-Using the constManagement results, that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results.
-The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state while calculating \(\Delta SOC\).
+Using the constManagement results that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results.
+The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state when calculating \(\Delta SOC\).
 The increasing SOC debt on global cropland are primarily caused by LUC (red areas in Fig.
@@
 \ref{fig:CLIMzone}).
-Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking the time scale of the stock change.
-
-\begin{figure}[h]
-\includegraphics[width=15cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
+Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking age of the cropland.
+
+\begin{figure}[h]
+\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
 IPCC factors vary a bit with different types of management (low, medium, high).
@@
 \ref{fig:SOCreg}) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.
-SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, since they include the litter horizon on top of the soil, that might dominate especially polar and boreal soils.
+SoilGrids \citep{hengl_soilgrids250m_2017} especially stands out with its high estimate, since they include the litter horizon on top of the soil that might dominate especially polar and boreal soils.
 SoilGrids 2.0 \citep{poggio_soilgrids_2021} however, excludes litter C and thus marks the lower end for particularly northern regions.
@@
 As the Tier 2 modeling approach \citep{ogle_cropland_in_ipcc_2019} is not specifically parameterized for natural vegetation it is important to evaluate its suitability to produce reasonable results in that domain as well at least comparable to other modeling approaches.
-We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.\ref{fig:SOCLPJmL4}).
+We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.
+\ref{fig:SOCLPJmL4}).
 Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.
@@
 \ref{fig:SOCpoint}) shows low correlation for PNV, however also for cropland sites.
-This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison \citep[see][]{rammig_generic_2018}.f
+This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties that could improve the point-to-grid-cell comparison \citep[see][]{rammig_generic_2018}.f
 
@@
 
-The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes effect these values strongly (see Fig.
+The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes affect these values strongly (see Fig.
 \ref{fig:SOCinit}).
-According to our sensitivity analysis the SOC debt depending on the initialization choice has been between \(33.3\,\unit{GtC}\) and \(50.7\,\unit{GtC}\), with our best guess at \(39.6\,\unit{GtC}\).
-Concurrently, our results indicate, that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
+According to our sensitivity analysis the SOC debt varies between \(33.3\,\unit{GtC}\) and \(50.7\,\unit{GtC}\) depending on the initialization choice, with our best guess at \(39.6\,\unit{GtC}\).
+Concurrently, our results indicate that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
 

### 2022-07-20 SOCbudgetPaper/discussion.Rmd
update to final version - 2. resubmit

@@
 \@ref(fig:SOCpoint)) shows low correlation for PNV, however also for cropland sites.
-This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties, that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f
+This might also point to the fact that SOC stocks can vary at field and local scale considerably and thus a very high number of point data is needed to derive statistical properties that could improve the point-to-grid-cell comparison [see @rammig_generic_2018].f
 
@@
 
-The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes effect these values strongly (see Fig.
+The initialization of SOC stocks, however, is important for the size of the SOC debt and its change over time, since the presence and size of legacy fluxes affect these values strongly (see Fig.
 \@ref(fig:SOCinit)).
-According to our sensitivity analysis the SOC debt depending on the initialization choice has been between $33.3\,\unit{GtC}$ and $50.7\,\unit{GtC}$, with our best guess at $39.6\,\unit{GtC}$.
-Concurrently, our results indicate, that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
+According to our sensitivity analysis the SOC debt varies between $33.3\,\unit{GtC}$ and $50.7\,\unit{GtC}$ depending on the initialization choice, with our best guess at $39.6\,\unit{GtC}$.
+Concurrently, our results indicate that the impact of the dynamic agricultural management is robust to the initialization of SOC stocks.
 

### 2022-07-20 SOCbudgetPaper/methods.Rmd
update to final version - 2. resubmit

@@
 This, however, will also include naturally occurring changes due to climatic variation over time.
-For our study, we define the absolute and relative $SOC$ changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$, that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
+For our study, we define the absolute and relative $SOC$ changes in relation to a potential natural state $SOC^{\mathrm{pnv}}$ under the same climatic conditions in grid cell $i$ at time $t$ that is based on the natural vegetation $SOC$ calculations as defined above without accounting for land conversion from cropland at any time.
 The absolute changes $\Delta SOC$ and relative changes $F^{\mathrm{SCF}}$ are thus given by
@@
 [-@bodirsky_mrcommons_2020]).
-Residue feed demand for five different livestock groups is based on country-specific feed baskets [see @weindl_livestock_2017], that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
+Residue feed demand for five different livestock groups is based on country-specific feed baskets [see @weindl_livestock_2017] that differentiate between the residue groups and take available $AGR$ biomass as well as livestock productivity into account.
 We estimate a material-use share for the straw-residue group of 5% and a fuel-share of 10% for all used residue groups in low-income countries.
@@
 [-@bodirsky_n2o_2012].
-We limit high application shares at $10\,\unit{tC\,ha}^{-1}$ to correct for outliers, that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
+We limit high application shares at $10\,\unit{tC\,ha}^{-1}$ to correct for outliers that can occur due to inconsistencies between FAO production and 0.5 degree land-use data.
 

### 2022-07-20 SOCbudgetPaper/results.Rmd
update to final version - 2. resubmit

@@
 \@ref(fig:SOCmaps)(d)) are a helpful metric to analyze the impact of human cropping activities.
-They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome by improved agricultural practices.
+They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation that may hold potential to be overcome by improved agricultural practices.
 Large parts of tropical cropland seem to suffer from strongly reduced relative stocks, indicating SOC degradation.
@@
 
-Using the constManagement results, that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results.
-The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state while calculating $\Delta SOC$.
+Using the constManagement results that only include land-use change (LUC) related changes of the SOC debt between 1975 and 2010, we are able to subtract the LUC effect from the overall SOC debt change within the histManagement results.
+The remaining effect can be attributed to the changing agricultural management (MAN) as other drivers such as climatic effects have been already canceled out by taking the difference to a PNV reference state when calculating $\Delta SOC$.
 The increasing SOC debt on global cropland are primarily caused by LUC (red areas in Fig.
@@
 \@ref(fig:CLIMzone)).
-Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking the time scale of the stock change.
+Note that IPCC Tier 1 factors are derived under the assumption that there is a linear change between steady states over 20 years, whereas our aggregated factors just reflect the relative change compared to a given potential natural vegetation reference stock without specifically tracking age of the cropland.
 
@@
 
-```{r SCFglo, out.width = "15cm", fig.cap = captionSCFglo, echo=FALSE, fig.pos="h"}
+```{r SCFglo, out.width = "16cm", fig.cap = captionSCFglo, echo=FALSE, fig.pos="h"}
 # All defaults
@@
 \@ref(fig:SOCreg)) show that our estimates are well within the range of other estimates for most regions, but at the lower end for tropical moist and tropical wet areas.
-SoilGrids [@hengl_soilgrids250m_2017] especially stands out with its high estimate, since they include the litter horizon on top of the soil, that might dominate especially polar and boreal soils.
+SoilGrids [@hengl_soilgrids250m_2017] especially stands out with its high estimate, since they include the litter horizon on top of the soil that might dominate especially polar and boreal soils.
 SoilGrids 2.0 [@poggio_soilgrids_2021] however, excludes litter C and thus marks the lower end for particularly northern regions.
@@
 As the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019] is not specifically parameterized for natural vegetation it is important to evaluate its suitability to produce reasonable results in that domain as well at least comparable to other modeling approaches.
-We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.\@ref(fig:SOCLPJmL4)).
+We therefore also compare our modeled results for SOC under natural vegetation (derived using litterfall of LPJmL4) against estimates of SOC by LPJmL4 for a PNV simulation (see Fig.
+\@ref(fig:SOCLPJmL4)).
 Both models are driven by the same climate conditions and the same natural litterfall and just differ in the representation of SOC and litter dynamics.

### 2022-08-19 SOCbudgetPaper/SOCbudget.tex
updated paper version (with hand edited tex file for picture placement issue)
update paper files that where submitted to bg

@@
 \begin{figure}[h]
-\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/LUtransitions} \caption{Scheme of land-use transition representation.
+\includegraphics[width=8cm]{LUtransitions.png} \caption{Scheme of land-use transition representation.
 Given an initial land-use pattern (as in this example 2 ha land under natural vegetation and 1 ha of cropland), there are separate $SOC$ stocks for natural vegetation and cropland.
@@
 \begin{figure}[h]
-\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data.
+\includegraphics[width=18cm]{4PanelFigure.png} \caption{Global SOC stocks and SOC stock changes on cropland for the first 30 cm of the soil profile considering historical management data.
 Panel (a) shows global $\Delta SOC$ between historical land use and potential natural vegetation (PNV).
@@
 \begin{figure}[h]
-\includegraphics[width=6.3cm]{../ResultNotebooks/Output/Images/SpatialDebtFigure} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile.
+\includegraphics[width=6.3cm]{SpatialDebtFigure.png} \caption{Global total $\Delta SOC$ and $\Delta SOC$ change for the first 30 cm of the soil profile.
 Panel (a) shows global $\Delta SOC$ as the difference between $SOC$ under historical land use and $SOC$ under potential natural vegetation (PNV) in the year 2010 summed over all land-use types.
@@
 \begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the cropland system for the year 2010 (in $\unit{MtC}$).
+\includegraphics[width=16cm]{CarbonBudget.png} \caption{Global carbon flows within the cropland system for the year 2010 (in $\unit{MtC}$).
 Carbon is first photosynthesised by crop plants and then used for livestock feed and various other usages subsumed under human demand.
@@
 \begin{figure}[h]
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in $\unit{GtC}$ for different management scenarios.
+\includegraphics[width=14cm]{ScenarioManageHoriz.png} \caption{(a) Global $\Delta SOC$ in $\unit{GtC}$ for different management scenarios.
 The stylized scenarios deviate from historical agricultural management patterns (histManagement) by holding effects of carbon inflows from residues (constResidues) or manure (constManure) constant at the 1975 level, or neglecting adoption of no-tillage practices over time (constTillage).
@@
 \begin{figure}[h]
-\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/SCF_comparison} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
+\includegraphics[width=16cm]{SCFcomparison.png} \caption{$F^{\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors from the guidelines in 2006 \citep{lasco_cropland_in_ipcc_2006} and the update in 2019 \citep{ogle_cropland_in_ipcc_2019}.
 IPCC factors vary a bit with different types of management (low, medium, high).
@@
 \begin{figure}[h]
-\includegraphics[width=7cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area.
+\includegraphics[width=7cm]{gloComparisonFigure.png} \caption{Modeled as well as observation-based estimates for global SOC stock in $\unit{GtC}$ for the first 30 cm of soil aggregated over all land area.
 The comparison against observation-based data (SoilGrids, SoilGrids 2.0, GSOC and WISE) is supplemented by modeled data from LPJmL4 \citep{schaphoff_lpjml4_2018-1} and estimates from \citep{sanderman_soil_2017}.
@@
 \begin{figure}[h]
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareSoilGrids2_2010_lu} \caption{Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0.}\label{fig:SOCSoilGrids2}
+\includegraphics[width=13cm]{scatterCompareSoilGrids2_2010_lu.png} \caption{Correlation between modeled SOC stocks of this study and projected values from SoilGrids 2.0.}\label{fig:SOCSoilGrids2}
 \end{figure}
@@
 \begin{figure}[h]
-\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/SOCquantilesSoilGrids2} \caption{Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0.}\label{fig:SOCSoilGrids2Q}
+\includegraphics[width=14cm]{SOCquantilesSoilGrids2.png} \caption{Global map on SOC results compared to uncertainty estimates from SoilGrids 2.0.}\label{fig:SOCSoilGrids2Q}
 \end{figure}
@@
 Our open-source model can be expanded to account for additional management options for carbon farming, such as cover crops, agroforestry, or biochar applications.
+\newpage
 
@@
 
-\begin{figure}[h]
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones.png} 
+\begin{figure}[!hp]
+\includegraphics[width=13cm]{ClimateZones.png} 
 \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \citep{eggleston_ipcc_2006} and has been reimplented by \citet{carre_background_2010}, which is the source of this data.
@@
 
-\begin{sidewaystable*}[htbp]
+\begin{table*}[!htbp]
 \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \citet{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions.
@@
 \tophline
-         &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
+         &                                 & 
+         \multicolumn{3}{l}{\shortstack[l]{Harvested\\Organs}} & 
+         \multicolumn{3}{l}{\shortstack[l]{Above-ground\\Residues}} & 
+         \multicolumn{2}{l}{\shortstack[l]{Below-ground\\Residues}} & 
+         \multicolumn{3}{l}{\shortstack[l]{Allometric\\coefficients}}                \\
 Crop code  & Crop Type          & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
@@
 soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
-rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
+rapeseed   & \begin{tabular}{@{}l@{}}Other oil crops\\(incl rapeseed)\end{tabular} & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
 groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
@@
 sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
-others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
+others     & \begin{tabular}{@{}l@{}}Fruits, Vegetables,\\Nuts\end{tabular}          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
 foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
@@
 \bottomhline
-         &                 & \multicolumn{6}{l}{
-                                           \begin{minipage}[t]{0.25\columnwidth}
+         &                 & \multicolumn{5}{l}{
+                                           \begin{minipage}[t]{0.3\columnwidth}
                                                       nr/dm -- nitrogen to dry matter ratio\\
@@
                                                       \end{minipage}} &
-                                                                        \multicolumn{5}{l}{
-                                                                            \begin{minipage}[t]{0.25\columnwidth}
+                                                                        \multicolumn{6}{l}{
+                                                                            \begin{minipage}[t]{0.3\columnwidth}
                                                                                         HI\textsuperscript{area} -- harvest index per area\\
@@
   \label{tab:c2dm}
-\end{sidewaystable*}
+\end{table*}
+
+\newpage
 
@@
 
-\begin{figure}[h]
-\includegraphics[width=12cm]{../ResultNotebooks/Output/Images/scenario_init.png} 
+\begin{figure}[!hp]
+\includegraphics[width=12cm]{ScenarioInit.png} 
 \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting in 1901 with steady-state SOC under vegetation for all land-use types without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use (Initial-spinup1510).
@@
 
-\begin{figure}[h]
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure.png} 
+\begin{figure}[!hp]
+\includegraphics[width=13cm]{regComparisonFigure.png} 
 \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations.
@@
 
-\begin{figure}[h]
-\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
+\newpage
+
+\begin{figure}[ht]
+\includegraphics[width=10cm]{scatterComparePointData_2010.png} 
 \caption{Correlation between modeled and measured $SOC$ stocks.
@@
 
- \begin{figure}[h]
-\includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
+\newpage
+
+ \begin{figure}[ht]
+\includegraphics[width=13cm]{scatterCompareLPJmL42010_pnv.png}
 \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
@@
 \noappendix
+
+\newpage
 

### 2022-10-21 ResultNotebooks/results_SOCscattervalidate.Rmd
update Images

@@
 
-dev.copy(png, paste0(outputdir, "scatterCompareSoilGrids2_2010_lu.png"), width = 1500, height = 1500, res = 250)
+dev.copy(png, paste0(outputdir, "scatterCompareSoilGrids2_2010_lu.png"), width = 1300, height = 1300, res = 250)
 dev.off()
@@
                         # exactly at the right end of each legend key
-                        title.hjust    = 0.5,
-                        label.hjust    = 1,
+                        title.hjust    = 0,
+                        label.hjust    = 0,
                         nrow           = 1,
@@
     theme(legend.position = "bottom",
+          legend.direction = "horizontal",
+          legend.margin    = margin(t = 0, r = 0, b = 0, l = 1, unit = "cm"),
           axis.title.x     = element_blank(),
@@
                                 cols     = cols,
-                                 name     = expression("Modeled SOC stock in relation to uncertainty range of SoilGrids 2.0"),
+                                name     = NULL,
                                 breaks   = c(0.5, 1.0, 1.5),
-                            labels   = c("lower than Q0.05", "inside [Q0.05, Q0.95] ", "higher than Q0.95", "No Cropland"))
+                                labels   = c("< Q0.05", "inside [Q0.05, Q0.95] ", "> Q0.95", ""))
 
 SoilGrids2InMap
-ggsave(paste0(outputdir, "/SOCquantilesSoilGrids2.png"),    plot = SoilGrids2InMap,    width = 15, height = 10, dpi = 300, units = "cm")
+ggsave(paste0(outputdir, "/SOCquantilesSoilGrids2.png"),    plot = SoilGrids2InMap,    width = 10, height = 7, dpi = 400, units = "cm")
 ```

### 2022-10-21 ResultNotebooks/results_SOCscenario.Rmd
update Images

@@
                                    size = legend_size),
-        legend.position = "right", legend.direction = "vertical")
+        legend.position = "right", legend.direction = "vertical",
+        legend.margin = margin(0, 0, -1, 0),
+        legend.box.margin = margin(0, 0, 0, -20))
 
@@
                                    size = legend_size), 
-        legend.direction = "vertical")
+        legend.direction = "vertical", 
+        legend.margin = margin(-140, 0, 0, 0),
+        legend.box.margin = margin(0, 0, 0, -20))
 
@@
                              label_size = 10, align = "v"),
-                   align = "h")
+                   align = "h", rel_widths = c(1.4, 1))
 
@@
 ggsave(paste0(outputdir, "Images/scenario_verti.png"), 
-       plot = vgrid, width = 12, height = 15, dpi = 300, units = "cm")
+       plot = vgrid, width = 13, height = 15, dpi = 300, units = "cm")
 ```

### 2022-10-21 ResultNotebooks/results_SOCtable.Rmd
update Images

--- 2022-07-20 ResultNotebooks/results_SOCtable.Rmd
@@
 climateIPCCr <- readRDS(paste0("Data/", rev, scen, "/IPCC_reduced.rds"))
-climateIPCC  <- readRDS(paste0("Data/", re, scen, "/IPCC.rds"))
+climateIPCC  <- readRDS(paste0("Data/", rev, scen, "/IPCC.rds"))
 
@@
 
-magpie_coord <- toolGetMapping("magpie_coord.rda", type = "cell")
+magpie_coord <- toolGetMapping("magpie_coord.rda", type = "cell", where="mappingfolder")
 climate2plot <- dimSums(climateIPCC * 
@@
 ```{r table2boxplot, fig.height=3, fig.width=6}
+library(ggplot2)
 
@@
                                       margin = margin(t = 10, r = 0, b = 0, l = 0, 
-                                                                   unit = "pt")))
+                                                                   unit = "pt")),
+          axis.text.x  = element_text(size = legend_size),
+          axis.text.y  = element_text(size = legend_size),
+          legend.title = element_text(size = legend_size),
+          legend.text  = element_text(size = legend_size),
+          legend.spacing.y = unit(0.1, "cm"),
+          legend.position="bottom",
+          legend.direction = "vertical",
+          legend.box = "horizontal"
+          )
+
+SCFTable
 
 ggsave(paste0(outputdir, "Images/SCF_comparison.png"), 
-       plot = SCFTable, height = 8, width = 18, units = "cm", dpi = 300)
+       plot = SCFTable, height = 11, width = 17, units = "cm", dpi = 300)
 ```
